### 2014-11-28 linear_iso.tex
initilize

\section{Vertical shear instability of locally
  isothermal perturbations}
The linear problem simplifies considerably for the case
$\gamma=1$. That is, for locally isothermal perturbations where
\begin{align}
  W=Q,
\end{align}
so that
\begin{align}
  \delta v_z = \frac{1}{\ii\sigma}\frac{dQ}{dz}. 
\end{align}
Inserting this into the expression for $\delta v_x$ and the continuity
equation, we arrive at a single equation for the linear stability
problem,
\begin{align}\label{iso_ode}
  \frac{d^2Q}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dQ}{dz} +
  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)=0, 
\end{align} 
subject to appropriate boundary conditions.

\subsection{Approximations}
Eq. \ref{iso_ode} is still complicated because the coefficients depend
on $z$ in a non-trivial manner. Furthermore, the eigenfrequency
$\sigma^2$ appears in $D$. In order to make progress, in this section
we make the following approximations.

\emph{Thin disk.} We consider thin disks with $h\ll1$, so we may expand the
background density and vertical shear in powers of $z/r$. To lowest
order we obtain
\begin{align}
  &\rho(z) \simeq \rho_{00} \exp{\left(-\frac{z^2}{2H^2}\right)},\\
  &\frac{d\Omega^2}{dz} \simeq \frac{q\Omega_k^2z}{r^2}. 
\end{align}

\emph{Low frequency.} We consider eigenvalues such that
$|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
\kappa^2$. The problem simplifies because the eigenvalue now only
appears in the last term in Eq. \ref{iso_ode}.

% \emph{Nearly constant $\kappa$}. We are mainly interested in the
% effect of vertical shear, i.e. $q\neq 0$. This is already captured
% explicitly in Eq. \ref{iso_ode}. We therefore simplify further by
% ignoring the vertical dependence of $\kappa^2$. In a thin disk, this
% quantity is nearly Keplerian.

With the above approximations and introducing the dimensionless
vertical co-ordinate $\hat{z} \equiv z/H$, the governing equation becomes
\begin{align}\label{iso_ode2}
  &\frac{d^2Q}{d\hat{z}^2} - \left[ 1 + \ii qh
    \left(k_xH\right)\frac{\Omega_k^2}{\kappa^2}\right]\hat{z}\frac{dQ}{d\hat{z}}\notag\\
  &+ \sigma^2H^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\kappa^2}\right)Q =
  0. 
\end{align}

We are mainly interested in the effect of vertical shear, i.e. $q\neq
0$. This is already captured explicitly in Eq. \ref{iso_ode2}. We 
therefore simplify further by ignoring the vertical dependence of
$\kappa^2$ and replace it by $\Omega_k^2$. This is expected to be
valid in a thin disk. Our final equation is then

\begin{align}\label{iso_ode3}
  \frac{d^2Q}{d\hat{z}^2} - \left(1 + \ii qh
    \hat{k}\right)\hat{z}\frac{dQ}{d\hat{z}} +
  \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)Q = 
  0, 
\end{align}
where we have introduced the dimensionless wavenumber $\hat{k}
\equiv k_x H$ for convenience.

\subsection{Neccessity of vertical shear for instability}
Writing
\begin{align}
  V \equiv Q\exp{\left[-\frac{\hat{z}^2}{2}(\ii q h
      \hat{k})\right]},
\end{align}
we have an alternative form of the governing equation 
\begin{align}\label{iso_alt}
  \frac{d}{d\hat{z}}\left[w(\hat{z})\left(\frac{dV}{d\hat{z}} + \ii q h
    \hat{k}\hat{z}V\right)\right] + n^2V w(\hat{z}) =0, 
\end{align}
where 
\begin{align}
  w(\hat{z}) \equiv \exp{\left(-\frac{\hat{z}^2}{2}\right)},  
\end{align}
and we have defined a new eigenvalue
\begin{align}
  n^2 \equiv \left(\frac{\sigma}{\Omega_k}\right)^2(1+\hat{k}^2). 
\end{align}

Next, we multiply Eq. \ref{iso_alt} by $V^{*}$, integrate
vertically and assume boundary terms vanish to obtain
\begin{align}\label{integral_relation}
  n^2\int_{-\infty}^{\infty}w|V|^2d\hat{z} 
  =&\int_{-\infty}^{\infty}w\left|\frac{dV}{d\hat{z}}\right|^2d\hat{z}\notag\\
  +&\ii q h \hat{k}\int_{-\infty}^{\infty}w\hat{z}V\frac{dV^*}{d\hat{z}}d\hat{z}.   
\end{align}
It follows that for instability ($\imag\sigma>0$ and therefore $\imag
n^2 \neq 0$), it is neccessary to have $q\neq 0$, i.e. vertical
shear. We will use this integral relation to estimate growth
rates.

\subsection{Stability in the absence of vertical shear}
When $q=0$, the governing equation is
\begin{align}
  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dV}{d\hat{z}}\right] + n^2V
  w(\hat{z}) =0, 
\end{align}
i.e. Hermite's differential equation. For bounded solutions we have 
\begin{align}
  V\propto \He_n(\hat{z}),
\end{align}
and $n$ is a non-negative integer. This translates to
\begin{align}
  \left|\frac{\sigma}{\Omega_k}\right| = n
  \left(1+\hat{k}^2\right)^{-1/2}. 
\end{align}
Since we have assumed $|\sigma^2|\ll \kappa^2$, our analysis is only
valid for large wavenumbers such that $\hat{k}^2\gg 
n^2-1$. Physically, this corresponds to radial length-scales much
smaller than the local disk scale height. From

\subsection{Instability in the presence of weak vertical shear}
To investigate the destabilizing effect of vertical shear, we first
consider a system with $q\equiv0$, for which the eigenfunctions are
Hermite polynomials and the real eigenfrequencies are known. We then
perturb this system by introducing a weak vertical shear ($|q|\ll1$)
and linearize the governing equation. This procedure is
\begin{align}   
  &q \to 0 + \delta q\\
  &V \to \He_n + \delta V \\
  &\sigma \to \sigma + \delta\sigma, 
\end{align}
where for this exercise we restore $\sigma$ as the eigenfrequency.

Linearizing the integral Eq. \ref{integral_relation} and taking the
imaginary part, we have

### 2014-11-28 paper.tex
initilize

%\documentclass[12pt]{aastex}
\documentclass[iop]{emulateapj}
%\documentclass{article}
%\usepackage{draftwatermark}
\usepackage{paralist}
\shorttitle{Vertical shear instability}
\shortauthors{M-K. Lin, A. Youdin}

\usepackage{amsmath}
%\usepackage{url}
\usepackage{bm}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\newcommand{\p}{\partial}
\newcommand{\lmax}{l_\mathrm{max}}
\newcommand{\sbar}{\bar{\sigma}} 
\newcommand{\He}{\operatorname{He}}
\newcommand{\imgi}{\mathrm{i}} 
\newcommand{\ii}{\mathrm{i}}
\newcommand{\C}{\mathcal{C}^\lambda}
\newcommand{\wop}[1]{\mathcal{V}^{(#1)}}
\newcommand{\avg}[1]{\langle{#1}\rangle}
%\newcommand{\max}[1]{\mathrm{max}({#1})}
%\newcommand{\min}[1]{\mathrm{min}({#1})}
\newcommand{\zhat}{\hat{z}}
\newcommand{\rhat}{\hat{r}}
\newcommand{\he}{\mathcal{H}}
\newcommand{\amp}{\mathcal{A}}
\newcommand{\dd}{\delta}
\newcommand{\real}{\operatorname{Re}}
\newcommand{\imag}{\operatorname{Im}}
\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\zmax}{Z_s}
\newcommand{\wbar}{\tilde{W}}
\newcommand{\qbar}{\tilde{Q}}
\newcommand{\dbigH}{\frac{H^\prime}{H}}
\newcommand{\Rp}{\frac{\rho_0^\prime}{\rho_0}}
\newcommand{\Fp}{\frac{g^\prime}{g}}
\newcommand{\Rpp}{\frac{\rho_0^{\prime\prime}}{\rho_0}}
\newcommand{\Fpp}{\frac{g^{\prime\prime}}{g}}
\newcommand{\ils}{L_s^{-1}}
\newcommand{\ilp}{L_p^{-1}}
\newcommand{\ihp}{H_p^{-1}}
\newcommand{\ihs}{H_s^{-1}}
\newcommand{\ciso}{c_\mathrm{iso}}
\newcommand{\csmid}{c_{s0}} 
\newcommand{\hiso}{h_\mathrm{iso}}
\newcommand{\Hiso}{H_\mathrm{iso}}
\newcommand{\zeus}{\texttt{ZEUS-MP }}
\newcommand{\rsph}{r_\mathrm{sph}}
\newcommand{\dvx}{\dd v_x}
\newcommand{\dvy}{\dd v_y}
\newcommand{\dvz}{\dd v_z}
\newcommand{\dbx}{\dd B_x}
\newcommand{\dby}{\dd B_y}
\newcommand{\dbz}{\dd B_z}
\newcommand{\w}{ \widetilde{W}}
%\newcommand{\sbar}{\bar{\sigma}}
\newcommand{\dphi}{\dd \Phi}
\newcommand{\actaa}{Acta Astron.}
\defcitealias{lin12}{L12}

\begin{document}

\title{Linear vertical shear instability in astrophysical disks: a parameter study}

\author{Min-Kai Lin, Andrew Youdin}
\affil{ $^1$Department of Astronomy and Steward Observatory, University of
  Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
\email{minkailin@email.arizona.edu}

\begin{abstract}

\end{abstract}

\section{Introduction}

\input{setup}
\input{linear_iso}
%\input{results}
%\input{results_gi}
%\input{results_azi} 
%\input{summary}

%\acknowledgements
%\appendix
%\input{appendix}

\bibliographystyle{apj}
\bibliography{ref}

\end{document}

### 2014-11-28 setup.tex
initilize

 \section{Baroclinic disk equilibra}
 We consider power-law disks with midplane density
 \begin{align}
   \rho_0(r) = \rho_{00}\left(\frac{r}{r_0}\right)^p,
 \end{align}
 where $\rho_{00}$ is the midplane density at the fiducial radius
 $r=r_0$, and midplane sound-speed 
 \begin{align}
   c_{s0}^2(r)=c_{s00}^2\left(\frac{r}{r_0}\right)^q, 
 \end{align}
 where $c_{s00}$ is the midplane sound-speed at $r=r_0$. 
 The three-dimensional equilibrium pressure $P$ and density $p$ are
 related by 
 \begin{align}
   P(r,z) = 
   \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z), 
 \end{align}
 where we refer to $\Gamma$ as the polytropic index.

 \subsection{Locally isothermal disks}
 For the special case $\Gamma=1$ the disk is isothermal in the
 vertical direction. The equilibrium density $\rho(r,z)$ and rotation
 $\Omega(r,z)$ is then 
 \begin{align}
   \rho(r,z) &=
   \rho_0(r)\exp{\left[\frac{v_k^2}{c_{s0}^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\   
   \Omega^2(r,z)& = \Omega_k^2(r) \left[(p+q)h^2(r) + (1 + q) -
     \frac{1}{\sqrt{1+z^2/r^2}}\right],
 \end{align}
 where $\Omega_k\equiv\sqrt{GM/r^3}=v_k/r$ is the Keplerian rotation 
 and 
 \begin{align}
   h(r) = \frac{c_{s0}(r)}{v_k(r)} 
 \end{align}
 is the local disk aspect ratio. The disk scale height is $H=hr$.

 The radial shear is 
 \begin{align}
   S(r,z) \equiv r\frac{\p\Omega}{\p r},  
 \end{align}
 and we note that the square of the epicyclic frequency $\kappa$ is
 given by
 \begin{align}
   \kappa^2(r,z) \equiv 2\Omega(2\Omega + S). 
 \end{align}
 The vertical shear is 
 \begin{align}
   \frac{\p\Omega^2}{\p z} =
   \frac{q\Omega_k^2z}{r^2\left(1+z^2/r^2\right)^{3/2}}. 
 \end{align}

 \section{The vertically global shearing box} 
% We are interested in the stability problem on radial length-scales
% much smaller than the local disk scale-height $H$. 
 We work within the framework of the \emph{vertically global shearing
   box} (VGSB) developed by \cite{mcnally14}. 
 In this approximation the background density, pressure, radial shear
 and vertical shear are all assumed to only 
 depend on height. That is, the expressions for these quantities are
 evaluated at $r=r_0$.

 The axisymmetric VGSB equations read
 \begin{align}
   & \frac{\p\rho}{\p t} + \frac{\p}{\p x}\left(\rho v_x\right) +
   \frac{\p}{\p z}\left(\rho v_z\right)=0,\\
   & \frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
   v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}} = - 
   2\Omega_0\hat{\bm{z}}\times\bm{v} - S_0v_x\hat{\bm{y}} \notag\\ 
   &\phantom{\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
   v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}}=}-\frac{1}{\rho}\nabla P - \frac{\p\Phi_0}{\p z}\hat{\bm{z}},\\
   & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P = -\gamma P
   \nabla\cdot\bm{v} -
   \frac{\rho}{t_c}\left(\frac{P}{\rho}-\left.\frac{P}{\rho}\right|_{t=0}\right), 
 \end{align}
 where $\bm{v}$ is the velocity field relative to the bulk flow
 (i.e. $\bm{v}=\bm{0}$ in equilibrium),  
 subscript $0$ indicates evaluation at $r=r_0$. In the energy equation
 the ratio of specific heats $\gamma$ is assumed constant, 
 and we have included a thermal relaxation term with timescale $t_c$.

 \subsection{Linearization}
 We perturb the system such that
 \begin{align}
   \rho \to \rho(z) + \delta\rho(z)\exp{\left(\ii k_x x - \ii\sigma
       t\right)},    
 \end{align}
 and similarly for other fluid variables, where $k_x$ is a horizontal
 wavenumber and $\sigma$ is a (generally) complex frequency. For
 clarity we drop the subscript $0$ on the frequencies and shear rates
 below.

 {\bf below: at the moment only valid for locally isothermal
   equilibrium ($\Gamma=1$)} 
 Writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, 
 the linearized system of equations is 
 \begin{align}
   -&\ii\sigma \frac{Q}{c_s^2} + \ii k_x \delta v_x + \frac{d\dd
     v_z}{dz} + \frac{d\ln{\rho}}{dz}\delta v_z = 0,\\
   &\ii\sigma \delta v_x + 2\Omega\delta v_y = \ii k_x W,\\
   -&\frac{\kappa^2}{2\Omega}\delta v_x + \ii \sigma\delta v_y =
   r\frac{d\Omega}{dz}\delta v_z, \\
   & \ii\sigma\delta v_z = \frac{dW}{dz} +
   \frac{d\ln{\rho}}{dz}\left(W-Q\right), \\
   -&\ii\sigma W + \frac{1}{t_c}\left(W-Q\right) = -\ii\sigma\gamma Q +
   c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma-1\right)\delta v_z.
 \end{align}

 It is convenient to express $\delta v_x$ as a function of $W$ and
 $\delta v_z$,
 \begin{align}
   \delta v_x = -\frac{1}{D}\left(\sigma k_x W +
     r\frac{d\Omega^2}{dz}\delta v_z\right),
 \end{align}
 where
 \begin{align}
   D \equiv \kappa^2 - \sigma^2.
 \end{align}

### 2014-11-28 linear_iso.tex
minor update

@@
   \frac{d}{d\hat{z}}\left[w(\hat{z})\left(\frac{dV}{d\hat{z}} + \ii q h
-    \hat{k}\hat{z}V\right)\right] + n^2V w(\hat{z}) =0, 
+    \hat{k}\hat{z}V\right)\right] + nV w(\hat{z}) =0, 
 \end{align}
@@
 \begin{align}
-  n^2 \equiv \left(\frac{\sigma}{\Omega_k}\right)^2(1+\hat{k}^2). 
+  n \equiv \left(\frac{\sigma}{\Omega_k}\right)^2(1+\hat{k}^2). 
 \end{align}
@@
 \begin{align}\label{integral_relation}
-  n^2\int_{-\infty}^{\infty}w|V|^2d\hat{z} 
+  n \int_{-\infty}^{\infty}w|V|^2d\hat{z} 
   =&\int_{-\infty}^{\infty}w\left|\frac{dV}{d\hat{z}}\right|^2d\hat{z}\notag\\
@@
 It follows that for instability ($\imag\sigma>0$ and therefore $\imag
-n^2 \neq 0$), it is neccessary to have $q\neq 0$, i.e. vertical
+n \neq 0$), it is neccessary to have $q\neq 0$, i.e. vertical
 shear. We will use this integral relation to estimate growth
 rates.
+
+We also recognize that in Eq. \ref{integral_relation}, 
+the factor $q\hat{z}$ represents vertical shear and $w(\hat{z})$ is
+the background density profile. Assuming $V$ is either odd or even in
+$\hat{z}$,  then for the second integral on the right
+hand side to be non-zero, it is neccessary to have a vertical shear
+profile that has odd symmetry in the vertical co-ordinate.
 
@@
 \begin{align}
-  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dV}{d\hat{z}}\right] + n^2V
+  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dV}{d\hat{z}}\right] + nV
   w(\hat{z}) =0, 
@@
 \begin{align}
-  \left|\frac{\sigma}{\Omega_k}\right| = n
+  \left|\frac{\sigma}{\Omega_k}\right| = \sqrt{n}
   \left(1+\hat{k}^2\right)^{-1/2}. 
@@
 valid for large wavenumbers such that $\hat{k}^2\gg 
-n^2-1$. Physically, this corresponds to radial length-scales much
+n-1$. Physically, this corresponds to radial length-scales much
 smaller than the local disk scale height. From
@@
 imaginary part, we have
+\begin{align}
+  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
+  \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
+  \He_n^2(\hat{z}) d\hat{z} \notag\\
+  = \delta q h \hat{k} 
+  \int_{-\infty}^{\infty}
+  w(\zhat)\hat{z}\He_n(\hat{z})\frac{d\He_n}{d\zhat} d\hat{z}
+\end{align}
+Integrating the right hand side by parts and recognizing $\He_2 =
+\hat{z}^2 -1$, we have
+\begin{align}
+  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
+  \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
+  \He_n^2(\hat{z}) d\hat{z} \notag\\
+  = \frac{1}{2}\delta q h \hat{k} 
+  \int_{-\infty}^{\infty}
+  w(\zhat)\He_2(\hat{z})\He_n^2(\hat{z})d\hat{z}. 
+\end{align}
+Finally, we evaluate the integrals using following properties of Hermite polynomials
+\begin{align}
+  \int_{-\infty}^{\infty}
+  w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl}, 
+\end{align}
+and
+\begin{align}
+  \int_{-\infty}^{\infty}
+  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi =
+  \frac{\sqrt{2\pi}m!k!l!}{(s-m)!(s-k)!(s-l)!}, 
+\end{align}
+where $s = (m+k+l)/2$. We then obtain
+\begin{align}
+ \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
+ \left(1+\hat{k}^2\right) = \delta q h \hat{k} \frac{n!}{(n-1)!}.
+\end{align}
+Inserting the original eigenfrequency $\sigma = \pm
+\sqrt{n}\Omega_k(1+\hat{k}^2)^{-1/2}$ gives
+\begin{align}
+  \frac{\imag(\delta\sigma)}{\Omega_k} = \pm \frac{1}{2}\delta q h
+  \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
+\end{align}
+ For $\hat{k}\gg 1$ we have
+\begin{align}
+  \frac{\imag(\delta\sigma)}{\Omega_k} \simeq \pm \frac{1}{2}\delta q h
+  \sqrt{n} \sgn{\hat{k}}. 
+\end{align}

### 2014-11-28 linear_iso.tex
minor update

@@
 \end{align}
-i.e. Hermite's differential equation. For bounded solutions we have 
+i.e. Hermite's differential equation. If we impose that the kinetic
+energy density remain bounded at infinity, then  
 \begin{align}
@@
 \end{align}
-and $n$ is a non-negative integer. This translates to
+and $n$ is a non-negative integer. This translates to a real
+eigenfrequency $\sigma$ such that
 \begin{align}
@@
 
+\subsection{Simple solutions with vertical shear}
+The differential equation, Eq. \ref{iso_ode3}, can be solved by
+\begin{align}
+  Q\propto \hat{z}. 
+\end{align} 
+%Although $Q$ diverges at infinity, this solution is
+%acceptable because the implied kinetic energy densities remain
+%bounded, since the density decays rapidly. 
+In this case the eigenfrequency must satisfy 
+\begin{align}
+  \left(\frac{\sigma}{\Omega_k}\right)^2 = \frac{1+\ii q h
+    \hat{k}}{1+\hat{k}^2}, 
+\end{align}
+and $\sigma$ is generally complex\footnote{
+We note that an even simpler solution to Eq. \ref{iso_ode3} is
+$Q=\mathrm{constant}$, but this implies $\sigma^2=0$ and therefore
+stability, so this case is of no interest.}, with  
+%If we write $\sigma =
+%(\hat{\omega} + \ii \hat{\nu})\Omega_k$ with $\hat{\omega},\hat{\nu}$
+%real, we have
+\begin{align}
+  \left[\frac{\imag(\delta\sigma)}{\Omega_k}\right]^2 =
+  \frac{\sqrt{1+q^2h^2\hat{k}^2} - 1}{2\left(1+\hat{k}^2\right)}. 
+\end{align}
+This expression can be simplified in the limit $|q|\ll 1$. Then we have
+\begin{align}\label{simple_growth}
+  \frac{\imag(\sigma)}{\Omega_k} = \pm
+  \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1).  
+\end{align}
+
 \subsection{Instability in the presence of weak vertical shear}
-To investigate the destabilizing effect of vertical shear, we first
+We can also investigate the destabilizing effect of vertical shear
+without explicitly solving the full governing equation,
+Eq. \ref{iso_ode3}. We first
 consider a system with $q\equiv0$, for which the eigenfunctions are
@@
 \end{align}
-Integrating the right hand side by parts and recognizing $\He_2 =
-\hat{z}^2 -1$, we have
+Recognizing $\hat{z} = \He_1$ and using $d\He_n/d\hat{z} = n
+\He_{n-1}$ we have
  \begin{align}
@@
    \He_n^2(\hat{z}) d\hat{z} \notag\\
-  = \frac{1}{2}\delta q h \hat{k} 
+   =\delta q h \hat{k} 
    \int_{-\infty}^{\infty}
-  w(\zhat)\He_2(\hat{z})\He_n^2(\hat{z})d\hat{z}. 
+   w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
  \end{align}
+% Integrating the right hand side by parts and recognizing $\He_2 =
+% \hat{z}^2 -1$, we have
+% \begin{align}
+%   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
+%   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
+%   \He_n^2(\hat{z}) d\hat{z} \notag\\
+%   = \frac{1}{2}\delta q h \hat{k} 
+%   \int_{-\infty}^{\infty}
+%   w(\zhat)\He_2(\hat{z})\He_n^2(\hat{z})d\hat{z}. 
+% \end{align}
 Finally, we evaluate the integrals using following properties of Hermite polynomials
@@
 where $s = (m+k+l)/2$. We then obtain
+% \begin{align}
+%  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
+%  \left(1+\hat{k}^2\right) = \delta q h \hat{k} \frac{n!}{(n-1)!}.
+% \end{align
  \begin{align}
   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
- \left(1+\hat{k}^2\right) = \delta q h \hat{k} \frac{n!}{(n-1)!}.
+  \left(1+\hat{k}^2\right) = \delta q h \hat{k} n.
  \end{align}
@@
 \end{align}
- For $\hat{k}\gg 1$ we have
+This result agrees with Eq. \ref{simple_growth} for $n=1$, since the
+function $\He_1$ solves the governing equation exactly. For $\hat{k}\gg 1$ we have
 \begin{align}

### 2014-11-29 linear_iso.tex
minor update
explicit solution to vert iso ode added

@@
       r}{D}\frac{d\Omega^2}{dz}\right) \frac{dQ}{dz} +
-  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)=0, 
+  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)Q=0, 
 \end{align} 
@@
 
-\subsection{Simple solutions with vertical shear}
-The differential equation, Eq. \ref{iso_ode3}, can be solved by
+\subsection{Polynomial solutions with vertical shear}
+We can seek power-series solution to Eq. \ref{iso_ode3},
 \begin{align}
-  Q\propto \hat{z}. 
+  Q(\zhat) = \sum_{m=0}^\infty a_m\zhat^m. 
 \end{align}
-%Although $Q$ diverges at infinity, this solution is
-%acceptable because the implied kinetic energy densities remain
-%bounded, since the density decays rapidly. 
-In this case the eigenfrequency must satisfy 
+Then the coefficients must satisfy the recurrence relation
 \begin{align}
-  \left(\frac{\sigma}{\Omega_k}\right)^2 = \frac{1+\ii q h
-    \hat{k}}{1+\hat{k}^2}, 
+  (m+2)(m+1)a_{m+2} +
+  \left[n - m\left(1+\ii q h \hat{k}\right)\right] a_m = 0, 
 \end{align}
-and $\sigma$ is generally complex\footnote{
-We note that an even simpler solution to Eq. \ref{iso_ode3} is
-$Q=\mathrm{constant}$, but this implies $\sigma^2=0$ and therefore
-stability, so this case is of no interest.}, with  
-%If we write $\sigma =
-%(\hat{\omega} + \ii \hat{\nu})\Omega_k$ with $\hat{\omega},\hat{\nu}$
-%real, we have
+where $n$ is not neccessarily an integer. Indeed, if we demand
+a polynomial of degree $M$ as the solution, then the eigenfrequency
+must satisfy
 \begin{align}
-  \left[\frac{\imag(\delta\sigma)}{\Omega_k}\right]^2 =
-  \frac{\sqrt{1+q^2h^2\hat{k}^2} - 1}{2\left(1+\hat{k}^2\right)}. 
+\left(\frac{\sigma}{\Omega_k}\right)^2 = M\left(\frac{1+\ii q h
+    \hat{k}}{1+\hat{k}^2}\right).
 \end{align}
-This expression can be simplified in the limit $|q|\ll 1$. Then we have
+The eigenfrequency $\sigma$ is therefore complex. We are interested in
+unstable modes for which $\imag{\sigma}$ is the growth rate and is
+given by 
+\begin{align}
+%   &\left[\frac{\real(\sigma)}{\Omega_k}\right]^2 =
+%  \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} + 1\right)\\
+   \left[\frac{\imag(\sigma)}{\Omega_k}\right]^2 =
+   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} - 
+    1\right). 
+\end{align}
+This expression can be simplified in the limit $|q|\ll 1$, 
  \begin{align}\label{simple_growth}
-  \frac{\imag(\sigma)}{\Omega_k} = \pm
-  \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1).  
+   \frac{\imag(\sigma)}{\Omega_k} = \pm \sqrt{M} 
+   \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
  \end{align}
+ with $M$ being an integer.
+
+% The differential equation, Eq. \ref{iso_ode3}, can be solved by
+% \begin{align}
+%   Q\propto \hat{z}. 
+% \end{align} 
+% %Although $Q$ diverges at infinity, this solution is
+% %acceptable because the implied kinetic energy densities remain
+% %bounded, since the density decays rapidly. 
+% In this case the eigenfrequency must satisfy 
+% \begin{align}
+%   \left(\frac{\sigma}{\Omega_k}\right)^2 = \frac{1+\ii q h
+%     \hat{k}}{1+\hat{k}^2}, 
+% \end{align}
+% and $\sigma$ is generally complex\footnote{
+% We note that an even simpler solution to Eq. \ref{iso_ode3} is
+% $Q=\mathrm{constant}$, but this implies $\sigma^2=0$ and therefore
+% stability, so this case is of no interest.}, with  
+% %If we write $\sigma =
+% %(\hat{\omega} + \ii \hat{\nu})\Omega_k$ with $\hat{\omega},\hat{\nu}$
+% %real, we have
+% \begin{align}
+%   \left[\frac{\imag(\delta\sigma)}{\Omega_k}\right]^2 =
+%   \frac{\sqrt{1+q^2h^2\hat{k}^2} - 1}{2\left(1+\hat{k}^2\right)}. 
+% \end{align}
+% This expression can be simplified in the limit $|q|\ll 1$. Then we have
+% \begin{align}\label{simple_growth}
+%   \frac{\imag(\sigma)}{\Omega_k} = \pm
+%   \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1).  
+%\end{align}
 
@@
 \end{align}
-This result agrees with Eq. \ref{simple_growth} for $n=1$, since the
-function $\He_1$ solves the governing equation exactly. For $\hat{k}\gg 1$ we have
+This result agrees with Eq. \ref{simple_growth}.  
+%for $n=1$, since the
+%function $\He_1$ solves the governing equation exactly. 
+For $\hat{k}\gg 1$ we have
 \begin{align}

### 2014-11-30 linear_iso.tex
extensions

@@
   isothermal perturbations}
+{\bf need to update notation, use $\epsilon$ instead of $h$}
 The linear problem simplifies considerably for the case
-$\gamma=1$. That is, for locally isothermal perturbations where
+$\Gamma=\gamma=1$. Then 
 \begin{align}
-  W=Q,
+  W=Q. 
 \end{align}
-so that
-\begin{align}
-  \delta v_z = \frac{1}{\ii\sigma}\frac{dQ}{dz}. 
-\end{align}
-Inserting this into the expression for $\delta v_x$ and the continuity
-equation, we arrive at a single equation for the linear stability
-problem,
+%so that
+%\begin{align}
+ % \delta v_z = \frac{1}{\ii\sigma}\frac{dQ}{dz}. 
+%\end{align}
+%Inserting this into the expression for $\delta v_x$ and the continuity
+%equation, 
+Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
+problem for isothermal perturbations in a vertically isothermal disk, 
 \begin{align}\label{iso_ode}
@@
 subject to appropriate boundary conditions.
+
+We remark that Eq .\ref{iso_ode} is also applicable for
+$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$, since in that case
+$W\simeq Q$ as well. The following discussion therefore also holds for
+vertically non-isothermal, adiabatic disks.
 
@@
 \begin{align}
-  \int_{-\infty}^{\infty}
-  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi =
-  \frac{\sqrt{2\pi}m!k!l!}{(s-m)!(s-k)!(s-l)!}, 
+  &\int_{-\infty}^{\infty}
+  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi\notag\\& =
+  \frac{\sqrt{2\pi}m!k!l!}{(j-m)!(j-k)!(j-l)!}, 
 \end{align}
-where $s = (m+k+l)/2$. We then obtain
+where $j = (m+k+l)/2$. We then obtain
 % \begin{align}

### 2014-11-30 setup.tex
extensions

@@
  \end{align}
- where $c_{s00}$ is the midplane sound-speed at $r=r_0$. 
+ where $c_{s00}$ is the midplane sound-speed at $r=r_0$. We also
+ define
+\begin{align}
+  \epsilon(r) \equiv \frac{c_{s0}(r)}{v_k(r)}
+\end{align}
+to parametrize the magnitude of the sound speed, where
+$v_k=r\Omega_k=\sqrt{GM/r}$ is the Keplerian azimuthal velocity.
+
 The three-dimensional equilibrium pressure $P$ and density $p$ are
@@
 \end{align}
- where we refer to $\Gamma$ as the polytropic index.
+where we refer to $\Gamma$ as the polytropic index. We define the
+equilibrium sound-speed
+\begin{align}
+  c_s^2\equiv \frac{dP}{d\rho} =
+  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1}.  
+\end{align}
 
- \subsection{Locally isothermal disks}
+The equilibrium density is given by vertical hydrostatic balance. This
+gives, for $\Gamma\neq1$,  
+\begin{align}\label{eqm_dens}
+  &\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} = 1 +
+  \frac{\left(\Gamma-1\right)}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right). 
+\end{align}
+Note that for $\Gamma\neq1$ the disk has a surface $z=H(r)$ such that
+$\rho(r,H)=0$. If we parametrize
+\begin{align}
+  H(r)\equiv h r,
+\end{align}
+then the imposed sound-speed and disk thickness are related by 
+\begin{align}
+  h^2(r) = \left[1-\frac{\epsilon^2(r)}{\Gamma-1}\right]^{-2}-1. 
+\end{align}
+
+The equilibrium rotation profile is obtained by balancing radial
+pressure force, centrifugal force and radial gravity from the central
+star. This gives 
+\begin{align}\label{eqm_rot}
+  \frac{\Omega^2(r,z)}{\Omega_k^2(r)}=1 +
+  \left(p+\frac{s}{\Gamma}\right)\epsilon^2(r) 
+  +\frac{s}{\Gamma} \left(1-\frac{1}{\sqrt{1+z^2/r^2}}\right), 
+\end{align}
+where $s\equiv q+p(1-\Gamma)$. We also define the radial shear rate  
+\begin{align}
+  S(r,z) \equiv r\frac{\p\Omega}{\p r},  
+\end{align}
+and the square of the epicyclic frequency $\kappa$ 
+\begin{align}
+  \kappa^2(r,z) \equiv 2\Omega(2\Omega + S). 
+\end{align}
+We note that rotation profile generally depends on both $r$ and $z$,
+unless the equilibrium is barotropic with $s=0$, in which case
+$\Omega$ only depends on $r$.
+
+\subsection{Vertically isothermal disks}
 For the special case $\Gamma=1$ the disk is isothermal in the
@@
  \begin{align}
-   \rho(r,z) &=
-   \rho_0(r)\exp{\left[\frac{v_k^2}{c_{s0}^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\   
-   \Omega^2(r,z)& = \Omega_k^2(r) \left[(p+q)h^2(r) + (1 + q) -
-     \frac{1}{\sqrt{1+z^2/r^2}}\right],
+   \frac{\rho(r,z)}{\rho_0(r)} &= \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\   
+   \frac{\Omega^2(r,z)}{\Omega_k^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
+     \frac{1}{\sqrt{1+z^2/r^2}}\right),
  \end{align}
- where $\Omega_k\equiv\sqrt{GM/r^3}=v_k/r$ is the Keplerian rotation 
- and 
+which may be obtained from Eq. \ref{eqm_dens} and Eq. \ref{eqm_rot} by
+taking the limit $\Gamma\to 1$.
+
+The vertically isothermal disk has no surface, but we can define a
+characteristic scale-height
 \begin{align}
-   h(r) = \frac{c_{s0}(r)}{v_k(r)} 
+  H \equiv \epsilon r \quad\quad(\Gamma \equiv 1),   
 \end{align} 
- is the local disk aspect ratio. The disk scale height is $H=hr$.
+and in this case $h = \epsilon$.
 
- The radial shear is 
- \begin{align}
-   S(r,z) \equiv r\frac{\p\Omega}{\p r},  
- \end{align}
- and we note that the square of the epicyclic frequency $\kappa$ is
- given by
- \begin{align}
-   \kappa^2(r,z) \equiv 2\Omega(2\Omega + S). 
- \end{align}
- The vertical shear is 
- \begin{align}
-   \frac{\p\Omega^2}{\p z} =
-   \frac{q\Omega_k^2z}{r^2\left(1+z^2/r^2\right)^{3/2}}. 
- \end{align}
+% where $\Omega_k\equiv\sqrt{GM/r^3}=v_k/r$ is the Keplerian rotation 
+ % and 
+ % \begin{align}
+ %   h(r) = \frac{c_{s0}(r)}{v_k(r)} 
+ % \end{align}
+ % is the local disk aspect ratio. 
+%The disk scale height is $H=hr$.
+
+ % The vertical shear is 
+ % \begin{align}
+ %   \frac{\p\Omega^2}{\p z} =
+ %   \frac{q\Omega_k^2z}{r^2\left(1+z^2/r^2\right)^{3/2}}. 
+ % \end{align}
 
@@
  We work within the framework of the \emph{vertically global shearing
-   box} (VGSB) developed by \cite{mcnally14}. 
- In this approximation the background density, pressure, radial shear
- and vertical shear are all assumed to only 
- depend on height. That is, the expressions for these quantities are
- evaluated at $r=r_0$.
+   box} (VGSB) developed by \cite{mcnally14}. In this approximation
+ one sets up a local Cartesian co-ordinate  system $(x,y,z)$ centered
+ on a fiducial point $(r,\phi,z)=(r_0,\phi_0,0)$ in the global disk, with $\phi_0(t) =
+ \Omega(r_0,0)t$. The local $x,y,z$ axis correspond to the
+ global radial, azimuthal and vertical directions.
 
+ In this shearing box the equilibrium density, pressure and 
+ temperatures are assumed to depend only on $z$, by evaluating these
+ quantities at $r=r_0$. The background rotation is approximated by a
+ linear shear in $x$, but left unaltered in $z$. 
+% This corresponds to
+%  setting 
+% \begin{align}
+%   \rho(r,z)\to\rho(r_0,z),\,P \to P(r_0,z), c_s(r,z)^2 \to
+%   c_s^2(r_0,z),  \\
+% \end{align}
  The axisymmetric VGSB equations read
@@
 
- {\bf below: at the moment only valid for locally isothermal
-   equilibrium ($\Gamma=1$)} 
+ % {\bf below: at the moment only valid for locally isothermal
+ %  equilibrium ($\Gamma=1$)} 
  Writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, 
@@
    \frac{d\ln{\rho}}{dz}\left(W-Q\right), \\
-   -&\ii\sigma W + \frac{1}{t_c}\left(W-Q\right) = -\ii\sigma\gamma Q +
-   c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma-1\right)\delta v_z.
+   -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
+   c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.
  \end{align}
+It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa$
+are evaluated at $r=r_0$ using their global expressions above.
 
- It is convenient to express $\delta v_x$ as a function of $W$ and
- $\delta v_z$,
+We can eliminate the perturbed velocities between the linearized
+equations to obtain a pair or ordinary differential equations
+for $W,Q$ as
 \begin{align}
-   \delta v_x = -\frac{1}{D}\left(\sigma k_x W +
-     r\frac{d\Omega^2}{dz}\delta v_z\right),
+  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
+  &=\left(\frac{\ii
+    k_x r}{D}\frac{d\Omega^2}{dz} -
+  \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
+  \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
+&- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
+  \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
+&\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
+\frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
+&=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
+\left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
 \end{align}
+ % It is convenient to express $\delta v_x$ as a function of $W$ and
+ % $\delta v_z$,
+ % \begin{align}
+ %   \delta v_x = -\frac{1}{D}\left(\sigma k_x W +
+ %     r\frac{d\Omega^2}{dz}\delta v_z\right),
+ % \end{align}
 where

### 2014-12-01 linear_iso.tex
major restructuring

@@
-\section{Vertical shear instability of locally
+\section{Vertically isothermal disks with 
   isothermal perturbations}
-{\bf need to update notation, use $\epsilon$ instead of $h$}
 The linear problem simplifies considerably for the case
@@
 subject to appropriate boundary conditions. 
-
 We remark that Eq .\ref{iso_ode} is also applicable for
 $\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$, since in that case
-$W\simeq Q$ as well. The following discussion therefore also holds for
-vertically non-isothermal, adiabatic disks.
+$W\simeq Q$ as well. 
+%The following discussion therefore also holds for
+%vertically non-isothermal, adiabatic disks.
 
-\subsection{Approximations}
-Eq. \ref{iso_ode} is still complicated because the coefficients depend
-on $z$ in a non-trivial manner. Furthermore, the eigenfrequency
-$\sigma^2$ appears in $D$. In order to make progress, in this section
-we make the following approximations.
+\subsection{Neccessity of vertical shear for instability}\label{integral_relation}
+We can establish a neccessary condition for instability with the
+following assumptions:
 
-\emph{Thin disk.} We consider thin disks with $h\ll1$, so we may expand the
-background density and vertical shear in powers of $z/r$. To lowest
-order we obtain
-\begin{align}
-  &\rho(z) \simeq \rho_{00} \exp{\left(-\frac{z^2}{2H^2}\right)},\\
-  &\frac{d\Omega^2}{dz} \simeq \frac{q\Omega_k^2z}{r^2}. 
-\end{align}
-
-\emph{Low frequency.} We consider eigenvalues such that
+\begin{enumerate}
+\item We consider eigenvalues such that
   $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
@@
   appears in the last term in Eq. \ref{iso_ode}. 
+\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
+0$. This is already captured explicitly in Eq. \ref{iso_ode} through
+the factor $d\Omega^2/dz$. We 
+therefore simplify further by ignoring the vertical dependence of
+$\kappa^2$. For clarity, we will replace  $\kappa^2$ with $\Omega_k^2$.  
+\end{enumerate}
+
+With these approximations we can re-write the governing equation as
+\begin{align}\label{iso_ode1}
+&\frac{d}{dz}\left(\rho e^{-\ii k_x r
+    \Omega^2/\Omega_k^2}\frac{dQ}{dz}\right) +
+\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho Q e^{-\ii k_x r
+  \Omega^2/\Omega_k^2} \notag\\
+&=0.
+\end{align}
+% Writing
+% \begin{align}
+%   V \equiv Q\exp{\left[-\frac{\hat{z}^2}{2}(\ii q \epsilon
+%       \hat{k})\right]},
+% \end{align}
+% we have an alternative form of the governing equation 
+% \begin{align}\label{iso_alt}
+%   \frac{d}{d\hat{z}}\left[w(\hat{z})\left(\frac{dV}{d\hat{z}} + \ii q \epsilon
+%     \hat{k}\hat{z}V\right)\right] + nV w(\hat{z}) =0, 
+% \end{align}
+% where 
+% \begin{align}
+%   w(\hat{z}) \equiv \exp{\left(-\frac{\hat{z}^2}{2}\right)},  
+% \end{align}
+% and we have defined a new eigenvalue
+% \begin{align}
+%   n \equiv \left(\frac{\sigma}{\Omega_k}\right)^2(1+\hat{k}^2). 
+% \end{align}
+
+Next, we multiply Eq. \ref{iso_ode1} by $Q^{*}\exp{\left(\ii k_x r
+    \Omega^2/\Omega_k\right)}$, integrate 
+vertically and assume boundary terms vanish to obtain
+\begin{align}\label{neccessary_cond}
+   &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\infty}^{\infty}\rho|Q|^2dz \notag\\
+  &=\int_{-\infty}^{\infty}\rho\left|\frac{dQ}{dz}\right|^2dz 
+  +\frac{\ii k_x r}{\Omega_k^2}\int_{-\infty}^{\infty}\rho\frac{d\Omega^2}{dz}Q^*\frac{dQ}{dz}dz. 
+\end{align}
+%\begin{align}\label{integral_relation}
+ % n \int_{-\infty}^{\infty}w|V|^2d\hat{z} 
+ % =&\int_{-\infty}^{\infty}w\left|\frac{dV}{d\hat{z}}\right|^2d\hat{z}\notag\\
+ % +&\ii q \epsilon \hat{k}\int_{-\infty}^{\infty}w\hat{z}V\frac{dV^*}{d\hat{z}}d\hat{z}.   
+%\end{align}
+
+It follows that for instability ($\imag\sigma>0$), it is neccessary to
+have $d\Omega^2/dz\neq 0$, i.e. vertical 
+shear.
+
+% We recognize that in Eq. \ref{integral_relation}, $w(\hat{z})$
+% represents the background density profile and
+% the factor $\epsilon q \hat{z}$ represents vertical shear. 
+% Making the replacement $w\to \rho$ and 
+% $\epsilon q \hat{z}\to (r/\Omega_k^2)d\Omega^2/dz$ in
+% Eq. \ref{integral_relation}  gives the corresponding integral relation
+% without the thin-disk approximation for the the equilibrium density
+% and vertical shear profile (Eq. \ref{thin_vshear} ---
+% \ref{thin_dens}).
+
+%Assuming $V$ is either odd or even in
+%$\hat{z}$,  then for the second integral on the right
+%hand side to be non-zero, it is neccessary to have a vertical shear
+%profile that has odd symmetry in the vertical co-ordinate.
+
+\subsection{Thin-disk limit}
+The full expressions for the equilibrium density and vertical shear
+depends on $z$ in a non-trivial manner. Thus, to simplify the problem 
+we consider thin-disks ($\epsilon\ll1$), 
+% Furthermore, the eigenfrequency
+% $\sigma^2$ appears in $D$. In order to make progress, in this section
+% we make the following approximations. 
+so we may expand the background density and vertical shear in powers
+of $z/r$. To lowest order we obtain
+\begin{align}
+  &\rho(z) \simeq \rho_{00} \exp{\left(-\frac{z^2}{2\Hiso^2}\right)},\label{thin_dens}\\
+  &\frac{d\Omega^2}{dz} \simeq \frac{q\Omega_k^2z}{r^2}. \label{thin_vshear}
+\end{align}
 
@@
 
-With the above approximations and introducing the dimensionless
-vertical co-ordinate $\hat{z} \equiv z/H$, the governing equation becomes
-\begin{align}\label{iso_ode2}
-  &\frac{d^2Q}{d\hat{z}^2} - \left[ 1 + \ii qh
-    \left(k_xH\right)\frac{\Omega_k^2}{\kappa^2}\right]\hat{z}\frac{dQ}{d\hat{z}}\notag\\
-  &+ \sigma^2H^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\kappa^2}\right)Q =
-  0. 
-\end{align}
-
-We are mainly interested in the effect of vertical shear, i.e. $q\neq
-0$. This is already captured explicitly in Eq. \ref{iso_ode2}. We 
-therefore simplify further by ignoring the vertical dependence of
-$\kappa^2$ and replace it by $\Omega_k^2$. This is expected to be
-valid in a thin disk. Our final equation is then
-
+With this approximation and those in \S\ref{integral_relation}, the
+governing equation becomes
 \begin{align}\label{iso_ode3}
-  \frac{d^2Q}{d\hat{z}^2} - \left(1 + \ii qh
+  \frac{d^2Q}{d\hat{z}^2} - \left(1 + \ii q\epsilon
     \hat{k}\right)\hat{z}\frac{dQ}{d\hat{z}} +
@@
 \end{align}
-where we have introduced the dimensionless wavenumber $\hat{k}
-\equiv k_x H$ for convenience.
+where we have introduced the dimensionless vertical co-ordinate
+$\hat{z} \equiv z/\Hiso$ and wavenumber $\hat{k} \equiv k_x \Hiso$.
 
-\subsection{Neccessity of vertical shear for instability}
-Writing
-\begin{align}
-  V \equiv Q\exp{\left[-\frac{\hat{z}^2}{2}(\ii q h
-      \hat{k})\right]},
-\end{align}
-we have an alternative form of the governing equation 
-\begin{align}\label{iso_alt}
-  \frac{d}{d\hat{z}}\left[w(\hat{z})\left(\frac{dV}{d\hat{z}} + \ii q h
-    \hat{k}\hat{z}V\right)\right] + nV w(\hat{z}) =0, 
+In the thin-disk approximation the integral relation
+Eq. \ref{neccessary_cond} becomes
+\begin{align}\label{integral_relation_thin}
+ & \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
+  \int_{-\infty}^{\infty}w|Q|^2d\hat{z}  \notag\\ 
+  & =\int_{-\infty}^{\infty}w\left|\frac{dQ}{d\hat{z}}\right|^2d\hat{z}
+  +\ii q \epsilon
+  \hat{k}\int_{-\infty}^{\infty}w\hat{z}Q^*\frac{dQ}{d\hat{z}}d\hat{z}.    
 \end{align}
@@
 \begin{align}
-  w(\hat{z}) \equiv \exp{\left(-\frac{\hat{z}^2}{2}\right)},  
+  w(\hat{z}) \equiv \exp{\left(-\frac{\hat{z}^2}{2}\right)}.  
 \end{align}
-and we have defined a new eigenvalue
+
+% the governing equation becomes
+% \begin{align}\label{iso_ode2}
+%   &\frac{d^2Q}{d\hat{z}^2} - \left[ 1 + \ii q\epsilon
+%     \left(k_x\Hiso\right)\frac{\Omega_k^2}{\kappa^2}\right]\hat{z}\frac{dQ}{d\hat{z}}\notag\\
+%   &+ \sigma^2\Hiso^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\kappa^2}\right)Q =
+%   0. 
+% \end{align}
+
+\subsection{Stability in the absence of vertical shear}
+When $q=0$, Eq. \ref{iso_ode3} is
+\begin{align}\label{hermite_ode}
+  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dQ}{d\hat{z}}\right] + nQ
+  w(\hat{z}) =0, 
+\end{align}
+where we have defined a new eigenvalue
 \begin{align}
@@
 \end{align} 
-
-Next, we multiply Eq. \ref{iso_alt} by $V^{*}$, integrate
-vertically and assume boundary terms vanish to obtain
-\begin{align}\label{integral_relation}
-  n \int_{-\infty}^{\infty}w|V|^2d\hat{z} 
-  =&\int_{-\infty}^{\infty}w\left|\frac{dV}{d\hat{z}}\right|^2d\hat{z}\notag\\
-  +&\ii q h \hat{k}\int_{-\infty}^{\infty}w\hat{z}V\frac{dV^*}{d\hat{z}}d\hat{z}.   
-\end{align}
-It follows that for instability ($\imag\sigma>0$ and therefore $\imag
-n \neq 0$), it is neccessary to have $q\neq 0$, i.e. vertical
-shear. We will use this integral relation to estimate growth
-rates.
-
-We also recognize that in Eq. \ref{integral_relation}, 
-the factor $q\hat{z}$ represents vertical shear and $w(\hat{z})$ is
-the background density profile. Assuming $V$ is either odd or even in
-$\hat{z}$,  then for the second integral on the right
-hand side to be non-zero, it is neccessary to have a vertical shear
-profile that has odd symmetry in the vertical co-ordinate.
-
-\subsection{Stability in the absence of vertical shear}
-When $q=0$, the governing equation is
+Eq. \ref{iso_ode3} is Hermite's differential equation. If we impose
+that the kinetic energy density remain bounded at infinity, then  
 \begin{align}
-  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dV}{d\hat{z}}\right] + nV
-  w(\hat{z}) =0, 
-\end{align}
-i.e. Hermite's differential equation. If we impose that the kinetic
-energy density remain bounded at infinity, then  
-\begin{align}
-  V\propto \He_n(\hat{z}),
+  Q \propto \He_n(\hat{z}),
 \end{align}
@@
 n-1$. Physically, this corresponds to radial length-scales much
-smaller than the local disk scale height. From
+smaller than the local disk scale height.
 
@@
 \begin{align}
-\left(\frac{\sigma}{\Omega_k}\right)^2 = M\left(\frac{1+\ii q h
+\left(\frac{\sigma}{\Omega_k}\right)^2 = M\left(\frac{1+\ii q \epsilon
     \hat{k}}{1+\hat{k}^2}\right).
@@
    \left[\frac{\imag(\sigma)}{\Omega_k}\right]^2 =
-   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} - 
+   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
     1\right). 
@@
    \frac{\imag(\sigma)}{\Omega_k} = \pm \sqrt{M} 
-   \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
+   \frac{q\epsilon\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
  \end{align}
@@
   &q \to 0 + \delta q\\
-  &V \to \He_n + \delta V \\
+  &Q \to \He_n + \delta V \\
   &\sigma \to \sigma + \delta\sigma, 
@@
 
-Linearizing the integral Eq. \ref{integral_relation} and taking the
+Linearizing the integral relation in the thin-disk limit, Eq. \ref{integral_relation_thin}, and taking the
 imaginary part, we have
@@
   \He_n^2(\hat{z}) d\hat{z} \notag\\
-  = \delta q h \hat{k} 
+  = \delta q \epsilon \hat{k} 
   \int_{-\infty}^{\infty}
@@
    \He_n^2(\hat{z}) d\hat{z} \notag\\
-   =\delta q h \hat{k} 
+   =\delta q \epsilon \hat{k} 
    \int_{-\infty}^{\infty}
@@
   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
-  \left(1+\hat{k}^2\right) = \delta q h \hat{k} n.
+  \left(1+\hat{k}^2\right) = \delta q \epsilon \hat{k} n.
  \end{align}
@@
 \begin{align}
-  \frac{\imag(\delta\sigma)}{\Omega_k} = \pm \frac{1}{2}\delta q h
+  \frac{\imag(\delta\sigma)}{\Omega_k} = \pm \frac{1}{2}\delta q \epsilon
   \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
@@
 \begin{align}
-  \frac{\imag(\delta\sigma)}{\Omega_k} \simeq \pm \frac{1}{2}\delta q h
+  \frac{\imag(\delta\sigma)}{\Omega_k} \simeq \pm \frac{1}{2}\delta q \epsilon
   \sqrt{n} \sgn{\hat{k}}. 

### 2014-12-01 setup.tex
major restructuring

@@
 \begin{align}
-  H \equiv \epsilon r \quad\quad(\Gamma \equiv 1),   
+  \Hiso \equiv \epsilon r \quad\quad(\Gamma \equiv 1),   
 \end{align} 
-and in this case $h = \epsilon$.
+and in this case the characteristic aspect-ratio $h = \epsilon$.
 
@@
 \end{align}
+
+\subsection{Boundary conditions}
+In numerical calculations a finite vertical domain $z\in[-1,1]\zmax$
+is adopted. For $\Gamma\neq1$ we take $\zmax$ to be slightly less than
+$H$ to avoid the zero-density surface. At $z=\pm\zmax$ we impose
+\begin{align}
+  \delta v_z = 0,
+\end{align}
+for a solid boundary, or 
+\begin{align}
+  \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0,  
+\end{align}
+for a free surface, where $\bm{\xi}$ is the Lagrangian
+displacement.
+
+%These boundary conditions are also numerically
+%convenient because they can be recast to contain the eigenfrequency
+%$\sigma$. This means that the the full set of ODEs, including boundary
+%conditions can be written in the form $L(\bm{U}) = \sigma R (\bm{U})$

### 2014-12-01 linear_iso.tex
backup

--- 2014-11-29 linear_iso.tex
@@
 
-\subsection{Polynomial solutions with vertical shear}
-We can seek power-series solution to Eq. \ref{iso_ode3},
+\subsection{Simple solutions with vertical shear}
+The differential equation, Eq. \ref{iso_ode3}, can be solved by
 \begin{align}
-  Q(\zhat) = \sum_{m=0}^\infty a_m\zhat^m. 
+  Q\propto \hat{z}. 
 \end{align} 
-Then the coefficients must satisfy the recurrence relation
+%Although $Q$ diverges at infinity, this solution is
+%acceptable because the implied kinetic energy densities remain
+%bounded, since the density decays rapidly. 
+In this case the eigenfrequency must satisfy 
 \begin{align}
-  (m+2)(m+1)a_{m+2} +
-  \left[n - m\left(1+\ii q h \hat{k}\right)\right] a_m = 0, 
+  \left(\frac{\sigma}{\Omega_k}\right)^2 = \frac{1+\ii q h
+    \hat{k}}{1+\hat{k}^2}, 
 \end{align}
-where $n$ is not neccessarily an integer. Indeed, if we demand
-a polynomial of degree $M$ as the solution, then the eigenfrequency
-must satisfy
+and $\sigma$ is generally complex\footnote{
+We note that an even simpler solution to Eq. \ref{iso_ode3} is
+$Q=\mathrm{constant}$, but this implies $\sigma^2=0$ and therefore
+stability, so this case is of no interest.}, with  
+%If we write $\sigma =
+%(\hat{\omega} + \ii \hat{\nu})\Omega_k$ with $\hat{\omega},\hat{\nu}$
+%real, we have
 \begin{align}
-\left(\frac{\sigma}{\Omega_k}\right)^2 = M\left(\frac{1+\ii q h
-    \hat{k}}{1+\hat{k}^2}\right).
+  \left[\frac{\imag(\sigma)}{\Omega_k}\right]^2 =
+  \frac{\sqrt{1+q^2h^2\hat{k}^2} - 1}{2\left(1+\hat{k}^2\right)}. 
 \end{align}
-The eigenfrequency $\sigma$ is therefore complex. We are interested in
-unstable modes for which $\imag{\sigma}$ is the growth rate and is
-given by 
-\begin{align}
-%   &\left[\frac{\real(\sigma)}{\Omega_k}\right]^2 =
-%  \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} + 1\right)\\
-   \left[\frac{\imag(\sigma)}{\Omega_k}\right]^2 =
-   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} - 
-    1\right). 
+This expression can be simplified in the limit $|q|\ll 1$. Then we have
+\begin{align}\label{simple_growth}
+  \frac{\imag(\sigma)}{\Omega_k} = \pm
+  \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1).  
 \end{align}
-This expression can be simplified in the limit $|q|\ll 1$, 
- \begin{align}\label{simple_growth}
-   \frac{\imag(\sigma)}{\Omega_k} = \pm \sqrt{M} 
-   \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
- \end{align}
- with $M$ being an integer.
-
-% The differential equation, Eq. \ref{iso_ode3}, can be solved by
-% \begin{align}
-%   Q\propto \hat{z}. 
-% \end{align} 
-% %Although $Q$ diverges at infinity, this solution is
-% %acceptable because the implied kinetic energy densities remain
-% %bounded, since the density decays rapidly. 
-% In this case the eigenfrequency must satisfy 
-% \begin{align}
-%   \left(\frac{\sigma}{\Omega_k}\right)^2 = \frac{1+\ii q h
-%     \hat{k}}{1+\hat{k}^2}, 
-% \end{align}
-% and $\sigma$ is generally complex\footnote{
-% We note that an even simpler solution to Eq. \ref{iso_ode3} is
-% $Q=\mathrm{constant}$, but this implies $\sigma^2=0$ and therefore
-% stability, so this case is of no interest.}, with  
-% %If we write $\sigma =
-% %(\hat{\omega} + \ii \hat{\nu})\Omega_k$ with $\hat{\omega},\hat{\nu}$
-% %real, we have
-% \begin{align}
-%   \left[\frac{\imag(\delta\sigma)}{\Omega_k}\right]^2 =
-%   \frac{\sqrt{1+q^2h^2\hat{k}^2} - 1}{2\left(1+\hat{k}^2\right)}. 
-% \end{align}
-% This expression can be simplified in the limit $|q|\ll 1$. Then we have
-% \begin{align}\label{simple_growth}
-%   \frac{\imag(\sigma)}{\Omega_k} = \pm
-%   \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1).  
-%\end{align}
 
@@
   \int_{-\infty}^{\infty}
-  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi =
+  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi \notag\\=
   \frac{\sqrt{2\pi}m!k!l!}{(s-m)!(s-k)!(s-l)!}, 
@@
 \end{align}
-This result agrees with Eq. \ref{simple_growth}.  
-%for $n=1$, since the
-%function $\He_1$ solves the governing equation exactly. 
-For $\hat{k}\gg 1$ we have
+This result agrees with Eq. \ref{simple_growth} for $n=1$, since the
+function $\He_1$ solves the governing equation exactly. For $\hat{k}\gg 1$ we have
 \begin{align}

### 2014-12-03 linear_iso.tex
starting to add numerical results

--- 2014-12-01 linear_iso.tex
@@
 \begin{align}
-  W=Q. 
+  Q=W. 
 \end{align}
-%so that
-%\begin{align}
- % \delta v_z = \frac{1}{\ii\sigma}\frac{dQ}{dz}. 
-%\end{align}
-%Inserting this into the expression for $\delta v_x$ and the continuity
-%equation, 
+
 Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
@@
 \begin{align}\label{iso_ode}
-  \frac{d^2Q}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
-      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dQ}{dz} +
-  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)Q=0, 
+  \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
+      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
+  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
 \end{align} 
-subject to appropriate boundary conditions. 
-We remark that Eq .\ref{iso_ode} is also applicable for
-$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$, since in that case
-$W\simeq Q$ as well. 
-%The following discussion therefore also holds for
-%vertically non-isothermal, adiabatic disks.
+subject to appropriate boundary conditions. We remark that Eq
+.\ref{iso_ode} is also applicable for    
+$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. adiabatic
+disks with adiabatic vertical stratification), since in that case  
+$W\simeq W$ as well.
 
@@
   &\frac{d}{dz}\left(\rho e^{-\ii k_x r
-    \Omega^2/\Omega_k^2}\frac{dQ}{dz}\right) +
-\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho Q e^{-\ii k_x r
+      \Omega^2/\Omega_k^2}\frac{dW}{dz}\right) +
+  \sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho W e^{-\ii k_x r
     \Omega^2/\Omega_k^2} \notag\\
@@
 \end{align}
-% Writing
-% \begin{align}
-%   V \equiv Q\exp{\left[-\frac{\hat{z}^2}{2}(\ii q \epsilon
-%       \hat{k})\right]},
-% \end{align}
-% we have an alternative form of the governing equation 
-% \begin{align}\label{iso_alt}
-%   \frac{d}{d\hat{z}}\left[w(\hat{z})\left(\frac{dV}{d\hat{z}} + \ii q \epsilon
-%     \hat{k}\hat{z}V\right)\right] + nV w(\hat{z}) =0, 
-% \end{align}
-% where 
-% \begin{align}
-%   w(\hat{z}) \equiv \exp{\left(-\frac{\hat{z}^2}{2}\right)},  
-% \end{align}
-% and we have defined a new eigenvalue
-% \begin{align}
-%   n \equiv \left(\frac{\sigma}{\Omega_k}\right)^2(1+\hat{k}^2). 
-% \end{align}
-
-Next, we multiply Eq. \ref{iso_ode1} by $Q^{*}\exp{\left(\ii k_x r
+Next, we multiply Eq. \ref{iso_ode1} by $W^{*}\exp{\left(\ii k_x r
     \Omega^2/\Omega_k\right)}$, integrate 
@@
 \begin{align}\label{neccessary_cond}
-   &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\infty}^{\infty}\rho|Q|^2dz \notag\\
-  &=\int_{-\infty}^{\infty}\rho\left|\frac{dQ}{dz}\right|^2dz 
-  +\frac{\ii k_x r}{\Omega_k^2}\int_{-\infty}^{\infty}\rho\frac{d\Omega^2}{dz}Q^*\frac{dQ}{dz}dz. 
+   &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\infty}^{\infty}\rho|W|^2dz \notag\\
+  &=\int_{-\infty}^{\infty}\rho\left|\frac{dW}{dz}\right|^2dz 
+  +\frac{\ii k_x r}{\Omega_k^2}\int_{-\infty}^{\infty}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
 \end{align}
-%\begin{align}\label{integral_relation}
- % n \int_{-\infty}^{\infty}w|V|^2d\hat{z} 
- % =&\int_{-\infty}^{\infty}w\left|\frac{dV}{d\hat{z}}\right|^2d\hat{z}\notag\\
- % +&\ii q \epsilon \hat{k}\int_{-\infty}^{\infty}w\hat{z}V\frac{dV^*}{d\hat{z}}d\hat{z}.   
-%\end{align}
-
 It follows that for instability ($\imag\sigma>0$), it is neccessary to
-have $d\Omega^2/dz\neq 0$, i.e. vertical 
-shear.
-
-% We recognize that in Eq. \ref{integral_relation}, $w(\hat{z})$
-% represents the background density profile and
-% the factor $\epsilon q \hat{z}$ represents vertical shear. 
-% Making the replacement $w\to \rho$ and 
-% $\epsilon q \hat{z}\to (r/\Omega_k^2)d\Omega^2/dz$ in
-% Eq. \ref{integral_relation}  gives the corresponding integral relation
-% without the thin-disk approximation for the the equilibrium density
-% and vertical shear profile (Eq. \ref{thin_vshear} ---
-% \ref{thin_dens}).
-
-%Assuming $V$ is either odd or even in
-%$\hat{z}$,  then for the second integral on the right
-%hand side to be non-zero, it is neccessary to have a vertical shear
-%profile that has odd symmetry in the vertical co-ordinate.
+have $d\Omega^2/dz\neq 0$, i.e. vertical shear.
 
@@
 we consider thin-disks ($\epsilon\ll1$), 
-% Furthermore, the eigenfrequency
-% $\sigma^2$ appears in $D$. In order to make progress, in this section
-% we make the following approximations. 
 so we may expand the background density and vertical shear in powers
@@
 
-% \emph{Nearly constant $\kappa$}. We are mainly interested in the
-% effect of vertical shear, i.e. $q\neq 0$. This is already captured
-% explicitly in Eq. \ref{iso_ode}. We therefore simplify further by
-% ignoring the vertical dependence of $\kappa^2$. In a thin disk, this
-% quantity is nearly Keplerian.
-
 With this approximation and those in \S\ref{integral_relation}, the
@@
 \begin{align}\label{iso_ode3}
-  \frac{d^2Q}{d\hat{z}^2} - \left(1 + \ii q\epsilon
-    \hat{k}\right)\hat{z}\frac{dQ}{d\hat{z}} +
-  \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)Q = 
+  \frac{d^2W}{d\hat{z}^2} - \left(1 + \ii q\epsilon
+    \hat{k}\right)\hat{z}\frac{dW}{d\hat{z}} +
+  \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)W = 
   0, 
@@
 where we have introduced the dimensionless vertical co-ordinate
-$\hat{z} \equiv z/\Hiso$ and wavenumber $\hat{k} \equiv k_x \Hiso$.
+$\hat{z} \equiv z/\Hiso$ and wavenumber $\hat{k} \equiv k_x \Hiso$. We 
+remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
+that derived by \cite{nelson13}, although we have taken a different
+route.
 
@@
  & \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
-  \int_{-\infty}^{\infty}w|Q|^2d\hat{z}  \notag\\ 
-  & =\int_{-\infty}^{\infty}w\left|\frac{dQ}{d\hat{z}}\right|^2d\hat{z}
+  \int_{-\infty}^{\infty}w|W|^2d\hat{z}  \notag\\ 
+  & =\int_{-\infty}^{\infty}w\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}
   +\ii q \epsilon
-  \hat{k}\int_{-\infty}^{\infty}w\hat{z}Q^*\frac{dQ}{d\hat{z}}d\hat{z}.    
+  \hat{k}\int_{-\infty}^{\infty}w\hat{z}W^*\frac{dW}{d\hat{z}}d\hat{z}.    
 \end{align}
@@
 
-% the governing equation becomes
-% \begin{align}\label{iso_ode2}
-%   &\frac{d^2Q}{d\hat{z}^2} - \left[ 1 + \ii q\epsilon
-%     \left(k_x\Hiso\right)\frac{\Omega_k^2}{\kappa^2}\right]\hat{z}\frac{dQ}{d\hat{z}}\notag\\
-%   &+ \sigma^2\Hiso^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\kappa^2}\right)Q =
-%   0. 
-% \end{align}
-
 \subsection{Stability in the absence of vertical shear}
@@
 \begin{align}\label{hermite_ode}
-  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dQ}{d\hat{z}}\right] + nQ
+  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dW}{d\hat{z}}\right] + nW
   w(\hat{z}) =0, 
@@
 \begin{align}
-  Q \propto \He_n(\hat{z}),
+  W \propto \He_n(\hat{z}),
 \end{align}
@@
 \begin{align}
-  Q(\zhat) = \sum_{m=0}^\infty a_m\zhat^m. 
+  W(\zhat) = \sum_{m=0}^\infty a_m\zhat^m. 
 \end{align}
@@
 given by 
+\begin{align}\label{simple_growth}
+   \frac{\imag(\sigma)}{\Omega_k} =\sqrt{
+   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
+    1\right)}. 
+\end{align}
+For fixed $q$ and $\epsilon$, the growth rate vanishes for both
+$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. The maximum growth rate
+occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
 \begin{align}
-%   &\left[\frac{\real(\sigma)}{\Omega_k}\right]^2 =
-%  \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} + 1\right)\\
-   \left[\frac{\imag(\sigma)}{\Omega_k}\right]^2 =
-   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
-    1\right). 
+  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
 \end{align}
-This expression can be simplified in the limit $|q|\ll 1$, 
- \begin{align}\label{simple_growth}
-   \frac{\imag(\sigma)}{\Omega_k} = \pm \sqrt{M} 
-   \frac{q\epsilon\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
+and
+\begin{align}
+\mathrm{max}\left[\imag(\sigma)\right] =\frac{\sqrt{M}|\epsilon
+  q|}{2\sqrt{1+|\epsilon q|}}\Omega_k. 
 \end{align}
- with $M$ being an integer.
 
-% The differential equation, Eq. \ref{iso_ode3}, can be solved by
-% \begin{align}
-%   Q\propto \hat{z}. 
+% This expression can be simplified in the limit $|q|\ll 1$, 
+%  \begin{align}\label{simple_growth}
+%    \frac{\imag(\sigma)}{\Omega_k} = \pm \sqrt{M} 
+%    \frac{q\epsilon\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
 %  \end{align}
-% %Although $Q$ diverges at infinity, this solution is
-% %acceptable because the implied kinetic energy densities remain
-% %bounded, since the density decays rapidly. 
-% In this case the eigenfrequency must satisfy 
-% \begin{align}
-%   \left(\frac{\sigma}{\Omega_k}\right)^2 = \frac{1+\ii q h
-%     \hat{k}}{1+\hat{k}^2}, 
-% \end{align}
-% and $\sigma$ is generally complex\footnote{
-% We note that an even simpler solution to Eq. \ref{iso_ode3} is
-% $Q=\mathrm{constant}$, but this implies $\sigma^2=0$ and therefore
-% stability, so this case is of no interest.}, with  
-% %If we write $\sigma =
-% %(\hat{\omega} + \ii \hat{\nu})\Omega_k$ with $\hat{\omega},\hat{\nu}$
-% %real, we have
-% \begin{align}
-%   \left[\frac{\imag(\delta\sigma)}{\Omega_k}\right]^2 =
-%   \frac{\sqrt{1+q^2h^2\hat{k}^2} - 1}{2\left(1+\hat{k}^2\right)}. 
-% \end{align}
-% This expression can be simplified in the limit $|q|\ll 1$. Then we have
-% \begin{align}\label{simple_growth}
-%   \frac{\imag(\sigma)}{\Omega_k} = \pm
-%   \frac{qh\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1).  
-%\end{align}
+%  with $M$ being an integer.
 
@@
   &q \to 0 + \delta q\\
-  &Q \to \He_n + \delta V \\
+  &W \to \He_n + \delta V \\
   &\sigma \to \sigma + \delta\sigma, 
@@
  \end{align}
-% Integrating the right hand side by parts and recognizing $\He_2 =
-% \hat{z}^2 -1$, we have
-% \begin{align}
-%   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
-%   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-%   \He_n^2(\hat{z}) d\hat{z} \notag\\
-%   = \frac{1}{2}\delta q h \hat{k} 
-%   \int_{-\infty}^{\infty}
-%   w(\zhat)\He_2(\hat{z})\He_n^2(\hat{z})d\hat{z}. 
-% \end{align}
-Finally, we evaluate the integrals using following properties of Hermite polynomials
+
+Finally, we evaluate the integrals using the results
 \begin{align}
@@
 where $j = (m+k+l)/2$. We then obtain
-% \begin{align}
-%  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
-%  \left(1+\hat{k}^2\right) = \delta q h \hat{k} \frac{n!}{(n-1)!}.
-% \end{align
+
  \begin{align}
@@
 \end{align}
-This result agrees with Eq. \ref{simple_growth}.  
+This result agrees with Eq. \ref{simple_growth} in the limit
+$|q|\to0$. 
 %for $n=1$, since the
@@
 \end{align}
+
+\subsection{Numerical solutions}
+In this section we numerically solve the linear stability problem
+corresponding to Eq. \ref{iso_ode1}. We thus relax the thin-disk
+assumption, using the full expressions for $\rho$, $\Omega$ and
+$\kappa$, so that we may consider vertical domains up to $z\sim r$.
+However, we retain the low-frequency approximation
+$|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue problem. This
+level of approximation lies in between the full problem 
+Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and the thin-disk
+limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.
+
+We solve Eq. \ref{iso_ode1} using a pseudo-spectral method by
+expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$ and
+discretizing the equation on a grid with
+$\zmax=10H_\mathrm{iso}$. This procedure converts the linear problem
+to a standard matrix eigenvalue problem. %In the low-frequency
+%approximation the eigenvalue is $\sigma^2$.
+
+\subsubsection{Fiducial calculation}
+We first validate the low-frequency approximation by solving a 
+fiducial problem with $q=-1$, $p=-1.5$, $\epsilon=0.1$ and $k_x =
+200\pi/r$ with solid boundaries. This is the setup considered by
+\cite{mcnally14} who solved the full problem.
+
+Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
+\ii\nu$. This plot is effectively identical to that in
+\cite{mcnally14}. The low-frequency approximation reproduces the
+three solution branches defined by \citep{nelson13}: corrugation modes
+with odd symmetry, breathing modes with even symmetry, and surface
+modes which correspond to the nearly-vertical locus in
+Fig. \ref{lowfreq_eigen}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
+  \caption{Eigenvalues in the low-frequency approximation for the
+    vertical shear instability in a vertically isothermal disk evolved
+    isothermally ($\gamma=\Gamma=1$). The disk parameters are $q=-1$,
+    $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
+    wavenumber is $k_x=200\pi/r$. This is the set up considered in
+    \cite{mcnally14}. \label{lowfreq_eigen}
+  }
+\end{figure}
+
+We plot in Fig. \ref{lowfreq_eigenfunc} the structurally simplest
+eigenfunction --- the fundamental corrugation mode corresponding to
+the eigenvalue with $\mathrm{min}|\sigma|$. The perturbation $W$ is
+roughly proportional to $z$, except close the boundaries where it
+flattens because $\delta v_z=dW/dz=0$ is imposed there.
+
+For $k_x = 200\pi/r$ and $\epsilon = 0.1$, the dimensionless
+wavenumber $\hat{k} = k_xH_\mathrm{iso}=20\pi$. Then the expected
+growth rate according to Eq. \ref{simple_growth} with $M=1$ is
+\begin{align*}
+  \nu = 0.2606\epsilon\Omega_k,
+\end{align*}
+as obtained numerically (Fig. \ref{lowfreq_eigen}). This agreement is
+surpsingly good, given that Eq. \ref{simple_growth} assumes the
+thin-disk approximation and imposes a different boundary condition to
+that in the numerical calculations.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/eigenvector_iso}
+  \caption{Eigenfunction of the fundamental corrugation mode,
+    corresponding to the 
+    bottom-left eigenvalue displayed in Fig. \ref{lowfreq_eigen}. The
+    solid and dashed lines are $\real W$ and $\imag W$, respectively. 
+    \label{lowfreq_eigenfunc}
+  }
+\end{figure}
+
+\subsubsection{Influence of vertical boundaries}

### 2014-12-04 linear_iso.tex
starting to add numerical results
added proof on maximum growth rate in limit of weak shear

@@
 disks with adiabatic vertical stratification), since in that case  
-$W\simeq W$ as well.
+$Q\simeq W$ as well.
 
@@
 have $d\Omega^2/dz\neq 0$, i.e. vertical shear.
+
+\subsection{Maximum growth rate in the limit of weak shear}   
+In the above discussions we have assumed a vertical domain   
+$\hat{z}\in[-\infty,\infty]$ which permits analytical expressions for
+the growth rates. Here we consider a finite vertical domain
+$\hat{z}\in[-\hat{z}_\mathrm{max},\hat{z}_\mathrm{max}]$ and show that
+the growth rate is limited by the maximum vertical shear in
+the domain. Since we do not persue explicit solutions to the linear
+problem, we relax the thin-disk approximation and restore the full
+expressions for $\rho$ and $d\Omega^2/dz$.
+
+We begin with the integral relation Eq. \ref{neccessary_cond} with a
+finite vertical domain but still assuming boundary terms vanish,
+\begin{align}
+  &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z}  
+  \notag\\ 
+  &=\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}      
+  +\ii
+  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z},    
+\end{align}   
+where $f(\hat{z})$ is defined such that
+\begin{align}
+  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
+\end{align}
+In the thin-disk limit, $f(\hat{z})=\hat{z}$.
+
+We now take the same approach as in the previous section. We first
+assume an neutral solution corresponding to $q\equiv0$, for which
+$\sigma$ is real and we can take $W$ to be real as well. We then
+set $q\to\delta q$ and linearize. This procedure means
+\begin{align}
+  &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}  
+  =\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z},\\ 
+  &\frac{2\sigma\imag{\delta\sigma}}{\Omega_k^2}\left(1+\hat{k}^2\right)
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}\notag\\
+  &=\epsilon\hat{k}\delta q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}. 
+\end{align}
+Eliminating $\sigma^2$ between the above equations, we obtain
+\begin{align}\label{sigma_finite_domain}
+  &4
+  \left[\frac{\imag{\left(\delta\sigma\right)}}{\Omega_k}\right]^2\left(1+\hat{k}^2\right) 
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  W^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z}\notag\\
+  &=\left(\epsilon\hat{k}\delta q\right)^2\left(\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}\right)^2. 
+\end{align}
+On the left hand side of this equation, we apply the Cauchy-Schwarz
+inequality to obtain
+\begin{align}
+  &\left( \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+    |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z}\right)^2\notag\\&\leq
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho 
+  W^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z}.
+\end{align}
+On the right hand side of Eq. \ref{sigma_finite_domain} we have
+\begin{align}
+  \left|\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+    f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}\right|\leq \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  |f(\hat{z})| |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z} \notag\\
+  \leq
+  \mathrm{max}\left(|f|\right)\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z},
+\end{align}
+where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
+$z\in[-\zmax,\zmax]$. Inserting these inequalities into
+Eq. \ref{sigma_finite_domain} gives
+\begin{align}
+  \left|\frac{\imag{(\delta\sigma)}}{\Omega_k}\right| \leq
+  \frac{\epsilon |\hat{k}\delta
+    q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
+\end{align}
 

### 2014-12-04 linear_iso.tex
new arguments on growth rate limit

--- 2014-12-03 linear_iso.tex
@@
 Next, we multiply Eq. \ref{iso_ode1} by $W^{*}\exp{\left(\ii k_x r
-    \Omega^2/\Omega_k\right)}$, integrate 
-vertically and assume boundary terms vanish to obtain
+    \Omega^2/\Omega_k\right)}$ and integrate 
+vertically from $z=-\zmax$ to $z=\zmax$. We assume either $W$ or
+$dW/dz$ vanishes at $z=\pm \zmax$, or that $\zmax$ is sufficiently
+large so that the boundary terms are negligible because of the
+decaying background density with increasing height. Then,
 \begin{align}\label{neccessary_cond}
-   &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\infty}^{\infty}\rho|W|^2dz \notag\\
-  &=\int_{-\infty}^{\infty}\rho\left|\frac{dW}{dz}\right|^2dz 
-  +\frac{\ii k_x r}{\Omega_k^2}\int_{-\infty}^{\infty}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
+   &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\zmax}^{\zmax}\rho|W|^2dz \notag\\
+  &=\int_{-\zmax}^{\zmax}\rho\left|\frac{dW}{dz}\right|^2dz 
+  +\frac{\ii k_x r}{\Omega_k^2}\int_{-\zmax}^{\zmax}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
 \end{align}
@@
 
+\subsection{Maximum growth rate}   
+Here we show that the growth rate is limited by the maximum vertical
+shear in the domain. It is convenient to introduce the dimensionless
+vertical co-ordinate $\hat{z} = z/\Hiso$ and wavenumber
+$\hat{k}=k_x\Hiso$. Then 
+the integral relation Eq. \ref{neccessary_cond} becomes 
+\begin{align}\label{integral_relation1}
+  &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z}  
+  \notag\\ 
+  &=\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}      
+  +\ii
+  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z},    
+\end{align}   
+where $f(\hat{z})$ is defined such that
+\begin{align}\label{fz_shear}
+  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
+\end{align}
+
+For convenience, we write $\hat{\sigma}\equiv\sigma/\Omega_k = \hat{\omega} +
+\ii\hat{\nu}$ with $\hat{\omega},\hat{\nu}$ being the dimensionless
+mode frequency and growth rates. Then the real and imaginary parts of
+Eq. \ref{integral_relation1} are
+\begin{align}
+  &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
+   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z} -
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}
+  \notag\\
+  &=\real\left[\ii
+  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+ f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right], \notag\\
+& 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
+   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z}\notag\\
+   &=\imag\left[\ii
+  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+ f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right].
+\end{align}
+Adding the square of these equations give
+\begin{align}
+&\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
+   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z} -
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\right]^2\notag\\
+&+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+   |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\notag\\
+   &=\left|\ii
+  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+ f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|^2.
+\end{align}
+It is clear that
+\begin{align}\label{sigma_finite_domain} 
+&4\hat{\nu}^2\left(1+\hat{k}^2\right) 
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+   |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\notag\\
+   &\leq\left|
+  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+ f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|^2.
+\end{align}
+% We now take the same approach as in the previous section. We first
+% assume an neutral solution corresponding to $q\equiv0$, for which
+% $\sigma$ is real and we can take $W$ to be real as well. We then
+% set $q\to\delta q$ and linearize. This procedure means
+% \begin{align}
+%   &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
+%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}  
+%   =\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z},\\ 
+%   &\frac{2\sigma\imag{\delta\sigma}}{\Omega_k^2}\left(1+\hat{k}^2\right)
+%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}\notag\\
+%   &=\epsilon\hat{k}\delta q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+%   f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}. 
+% \end{align}
+% Eliminating $\sigma^2$ between the above equations, we obtain
+% \begin{align}\label{sigma_finite_domain}
+%   &4
+%   \left[\frac{\imag{\left(\delta\sigma\right)}}{\Omega_k}\right]^2\left(1+\hat{k}^2\right) 
+%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+%   W^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z}\notag\\
+%   &=\left(\epsilon\hat{k}\delta q\right)^2\left(\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+%   f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}\right)^2. 
+% \end{align}
+On the left hand side of this inequality, we apply the Cauchy-Schwarz
+inequality to obtain
+\begin{align}
+  &\left( \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+    |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z}\right)^2\notag\\&\leq
+  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho 
+  |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}.
+\end{align}
+On the right hand side of Eq. \ref{sigma_finite_domain} we have
+\begin{align}
+  \left|\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+    f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|\leq \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  \left|f(\hat{z})W^*\frac{dW}{d\hat{z}}\right|d\hat{z} \notag\\
+  \leq
+  \mathrm{max}\left(|f|\right)\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
+  |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z},
+\end{align}
+where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
+$z\in[-\zmax,\zmax]$. Inserting these inequalities into
+Eq. \ref{sigma_finite_domain} gives
+\begin{align}\label{max_growth}
+  \hat{\nu}\equiv\left|\frac{\imag{(\sigma)}}{\Omega_k}\right| \leq
+  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
+\end{align}
+Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
+follows that the maximum possible growth rate of unstable modes,
+satisfying the above boundary conditions, is limited by the maximum
+value of the vertical shear in the domain considered.
+
+For the vertical shear profile of the vertically isothermal
+disk, $f(\hat{z}) =
+\hat{z}\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$. If we consider
+$\zmax\to\infty$,  then the growth rate is limited by the global
+maximum of the vertical shear, which occurs at $\zhat
+=1/\sqrt{2}\epsilon$ for the vertically isothermal disk.  
+For vertical domains smaller than this height, the growth rate is
+limited by $|f|$ at the vertical boundaries considered. For 
+$|\epsilon\hat{z}|\ll1$ we have $f\simeq \hat{z}$, in which case we
+expect growth rates to increase linearly with the vertical domain
+size.
+
 \subsection{Thin-disk limit}
-The full expressions for the equilibrium density and vertical shear
-depends on $z$ in a non-trivial manner. Thus, to simplify the problem 
-we consider thin-disks ($\epsilon\ll1$), 
+We can make further progress for thin-disks ($\epsilon\ll1$), 
 so we may expand the background density and vertical shear in powers
@@
 \begin{align}
-  &\rho(z) \simeq \rho_{00} \exp{\left(-\frac{z^2}{2\Hiso^2}\right)},\label{thin_dens}\\
-  &\frac{d\Omega^2}{dz} \simeq \frac{q\Omega_k^2z}{r^2}. \label{thin_vshear}
+  &\rho(\hat{z}) \simeq \rho_{00}
+  \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv \rho_{00}w(\hat{z}),\label{thin_dens}\\
+  &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}. \label{thin_vshear}
 \end{align}
-
 With this approximation and those in \S\ref{integral_relation}, the
@@
   \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)W = 
-  0, 
+  0.
 \end{align}
-where we have introduced the dimensionless vertical co-ordinate
-$\hat{z} \equiv z/\Hiso$ and wavenumber $\hat{k} \equiv k_x \Hiso$. We 
-remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
+We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
 that derived by \cite{nelson13}, although we have taken a different
@@
 
-In the thin-disk approximation the integral relation
-Eq. \ref{neccessary_cond} becomes
-\begin{align}\label{integral_relation_thin}
- & \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
-  \int_{-\infty}^{\infty}w|W|^2d\hat{z}  \notag\\ 
-  & =\int_{-\infty}^{\infty}w\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}
-  +\ii q \epsilon
-  \hat{k}\int_{-\infty}^{\infty}w\hat{z}W^*\frac{dW}{d\hat{z}}d\hat{z}.    
-\end{align}
-where 
-\begin{align}
-  w(\hat{z}) \equiv \exp{\left(-\frac{\hat{z}^2}{2}\right)}.  
-\end{align}
+To complete the problem we must specify boundary conditions. The
+maximum vertical domain size we should consider is limited by the thin disk
+approximation (i.e. $z\lesssim r$). It is most common to take $\zmax\to\infty$
+and impose that the kinetic  energy remain bounded at infinity.  This
+is appropriate for the density because both the true density profile
+and its thin-disk approximation decay rapidly away from the midplane.
+
+However, the thin-disk representation of vertical shear diverges with 
+$\hat{z}$, unlike the true vertical shear profile which eventually
+decays.  Thus taking $\zmax\to$ will permit unphysically large growth
+rates (as demonstrated below). Nevertheless, the infinite disk is
+useful to consider since it permits an analytical discussion.
 
@@
 
-Linearizing the integral relation in the thin-disk limit, Eq. \ref{integral_relation_thin}, and taking the
+Linearizing the integral relation Eq. \ref{integral_relation1} in the
+thin-disk limit, and taking the
 imaginary part, we have

### 2014-12-04 linear_iso.tex
new arguments on growth rate limit

@@
 maximum of the vertical shear, which occurs at $\zhat
-=1/\sqrt{2}\epsilon$ for the vertically isothermal disk.  
+=1/\sqrt{2}\epsilon$.  
 For vertical domains smaller than this height, the growth rate is

### 2014-12-14 linear_iso.tex
minor update with notes

@@
 \end{enumerate}
+
+{\bf implicit limit on domain size. cannot be too large, otherwise omega reaches small values. should remain of order
+omegak}
 
@@
 size.
+
+{\bf get very similar result for if we replaced $\kappa^2(z)$ by $\Omega^2(z)$ instead. then the $(1+\hat{k}^2)$ factor should be 
+inside integral with extra factor of $\Omega_k^2/\Omega^2$ multiplying the $\hat{k}^2$. but for subkeplerian disks $\Omega^2<\Omega_k^2$ so can replace
+$\Omega^2$ by $\Omega^2_k$ without affecting the inequality on the lhs. on rhs need to replace $f$ by $\tilde{f}\equiv f\Omega_k^2/\Omega^2$. inequality is then not very 
+useful if zmax is very large since $\Omega^2$ may reach small values or even zero at finite height , so $\mathrm{max}|\tilde{f}|$ is large or infinite. 
+but still useful for zmax not too large such that $\Omega^2$ is similar to $\Omega^2$.}
+
+{\bf zmax is limited by: $\Omega^2$ and $\kappa^2$ do not become small, since low frequency assumption will fail if domain reaches heights such that $\Omega\sim0$. also can't assume
+$\Omega$ or $\kappa^2$ is roughly constant in coefficient of ode. (vertical derivatives always kept explicitly) 	
+}
 

### 2014-12-14 linear_adia.tex
notation change, begun adiabatic notes

\section{Adiabatic disks} 
We now examine the VSI in adiabatic disks ($t_c\to\infty$). For 
analytical discussion we consider the initially isothermal case with 
$\Gamma=1$. We eliminate variables in favor of $\delta v_z$ and
$\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
\begin{align}
&\Delta\left(1 + \frac{\gamma c_s^2 k_x^2}{D}\right) = \frac{d\delta
  v_z}{dz} + \frac{\ii k_x^2}{D}\left(\ii c_s^2 \frac{d\ln\rho}{dz}
  - \frac{r}{k_x}\frac{d\Omega^2}{dz}\right)\delta v_z,\label{adia_iso1}\\
& -\sigma^2\delta v_z = \gamma c_s^2 \frac{d\Delta}{dz} +
 c_s^2\frac{d}{dz}\left(\frac{d\ln\rho}{dz}\delta v_z\right) +
 \left(\gamma-1\right)\frac{d\ln\rho}{dz} \Delta.\label{adia_iso2}
\end{align}

We again make the low-frequency approximation
and ignore the vertical dependence of $\Omega$ and $\kappa$ except
where their vertical derivatives appear explicitly in the above
equations. We thus make the replacement $D\to \Omega_k^2$, as before
for isothermal perturbations. Here, we also use
the fact that in the global disk,
\begin{align}
  r\frac{\p \Omega^2}{\p z} = - \frac{\p\ln\rho}{\p z}\frac{\p
    c_s^2}{\p r} = -\frac{\p \ln\rho}{\p z} \frac{qc_s^2}{r},
\end{align}
where the second equality follows from the power-law prescription of
the sound-speed for $\Gamma=1$.

Combining Eq. \ref{adia_iso1} and Eq. \ref{adia_iso2} with the above
substitutions we obtain, in terms of dimensionless variables defined previously, an equation for $\delta v_z$,
\begin{align}
  0 =& \dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime\notag\\
  &+\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right)\right.\notag\\
  &\phantom{0=}\left.-\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon
          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
\end{align}
We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
integrate vertically, assume boundary terms vanish when integrating by
parts, to obtain
\begin{align}
  &\hat{\sigma}^2\left(\frac{1}{\gamma} +
    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
  v_z|^2 d\zhat \notag\\
  &=  \left(\frac{\gamma-1}{\gamma}\right)
  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
  v_z)^\prime|^2 d\zhat\notag\\
  &\phantom{=} +
  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
    2}|\delta v_z|^2 d\zhat\notag\\
  &\phantom{=} + \ii\epsilon q \hat{k}
  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
  \delta v_z d\zhat.
\end{align}

\subsection{Explicit solutions in the thin disk limit}
To simplify further, we consider the problem in the thin-disk limit so
that $\ln \rho \simeq -\zhat^2/2$. Eq. \ref{adia_iso3} becomes 
\begin{align}
  0 = \delta v_z^{\prime\prime} - \zhat A \delta v_z^\prime + \left(B
    - C \zhat^2\right)\delta v_z,
\end{align}
with
\begin{align}
  &A \equiv 1 + \ii \epsilon q \hat{k},\\
  &B \equiv \hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right) -
  \left(\frac{1}{\gamma} + \ii \hat{k} \epsilon q\right)\\
  &C \equiv \frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
      \epsilon q}{\hat{k}}\right).
\end{align}

% \begin{align}
%   0 =& \frac{d^2\delta v_z}{d\zhat^2} - \hat{z}\left(1 + \ii \hat{k}
%     \epsilon q \right)\frac{d\delta v_z}{d\hat{z}} \notag\\ 
% &+ \left[\hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right) -
%   \left(\frac{1}{\gamma} + \ii \hat{k} \epsilon q\right)\right.\notag\\
%   &\phantom{0=}-\left.\frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
%       \epsilon q}{\hat{k}}\right)\zhat^2\right]\delta v_z.\label{adia_iso_final}
% \end{align}
% Eq. \ref{adia_iso_final} will be the basis for discussion below.

### 2014-12-14 linear_iso.tex
notation change, begun adiabatic notes

@@
 
-With these approximations we can re-write the governing equation as
-\begin{align}\label{iso_ode1}
-  &\frac{d}{dz}\left(\rho e^{-\ii k_x r
-      \Omega^2/\Omega_k^2}\frac{dW}{dz}\right) +
-  \sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho W e^{-\ii k_x r
-    \Omega^2/\Omega_k^2} \notag\\
-  &=0.
+With these approximations and introducing the dimensionless variables
+\begin{align}
+  \hat{z} = z/\Hiso,\quad \hat{k}=k_x\Hiso, \quad \hat{\sigma} = \sigma/\Omega_k,
 \end{align}
-Next, we multiply Eq. \ref{iso_ode1} by $W^{*}\exp{\left(\ii k_x r
-    \Omega^2/\Omega_k\right)}$ and integrate 
-vertically from $z=-\zmax$ to $z=\zmax$. We assume either $W$ or
-$dW/dz$ vanishes at $z=\pm \zmax$, or that $\zmax$ is sufficiently
-large so that the boundary terms are negligible because of the
-decaying background density with increasing height. Then,
-\begin{align}\label{neccessary_cond}
-   &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\zmax}^{\zmax}\rho|W|^2dz \notag\\
-  &=\int_{-\zmax}^{\zmax}\rho\left|\frac{dW}{dz}\right|^2dz 
-  +\frac{\ii k_x r}{\Omega_k^2}\int_{-\zmax}^{\zmax}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
+the governing equation can be re-written as
+\begin{align}
+  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
+    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
 \end{align}
-It follows that for instability ($\imag\sigma>0$), it is neccessary to
-have $d\Omega^2/dz\neq 0$, i.e. vertical shear.
-
-\subsection{Maximum growth rate}   
-Here we show that the growth rate is limited by the maximum vertical
-shear in the domain. It is convenient to introduce the dimensionless
-vertical co-ordinate $\hat{z} = z/\Hiso$ and wavenumber
-$\hat{k}=k_x\Hiso$. Then 
-the integral relation Eq. \ref{neccessary_cond} becomes 
-\begin{align}\label{integral_relation1}
-  &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
-  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z}  
-  \notag\\ 
-  &=\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}      
-  +\ii
-  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-  f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z},    
-\end{align}   
-where $f(\hat{z})$ is defined such that
+where the primes denote $d/d\zhat$ and $f(\zhat)$ is defined such that
 \begin{align}\label{fz_shear}
@@
 
-For convenience, we write $\hat{\sigma}\equiv\sigma/\Omega_k = \hat{\omega} +
-\ii\hat{\nu}$ with $\hat{\omega},\hat{\nu}$ being the dimensionless
+Next, we multiply Eq. \ref{iso_ode1} by $\rho W^*$ and integrate 
+vertically from $\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary
+terms when integrating by parts, by assuming $W$ or
+$W^\prime$ vanishes at the boundaries, or that the boundaries are 
+sufficiently far away so that the boundary terms are negligible because of the
+decaying background density with increasing height. Then,
+\begin{align}
+  &\hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat \notag\\
+  &=\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
+  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
+\end{align}
+It follows that for instability ($\imag\hat{\sigma}>0$), it is neccessary to
+have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical shear.
+
+% With these approximations we can re-write the governing equation as
+% \begin{align}\label{iso_ode1}
+%   &\frac{d}{dz}\left(\rho e^{-\ii k_x r
+%       \Omega^2/\Omega_k^2}\frac{dW}{dz}\right) +
+%   \sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho W e^{-\ii k_x r
+%     \Omega^2/\Omega_k^2} \notag\\
+%   &=0.
+% \end{align}
+% Next, we multiply Eq. \ref{iso_ode1} by $W^{*}\exp{\left(\ii k_x r
+%     \Omega^2/\Omega_k\right)}$ and integrate 
+% vertically from $z=-\zmax$ to $z=\zmax$. We assume either $W$ or
+% $dW/dz$ vanishes at $z=\pm \zmax$, or that $\zmax$ is sufficiently
+% large so that the boundary terms are negligible because of the
+% decaying background density with increasing height. Then,
+% \begin{align}\label{neccessary_cond}
+%    &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\zmax}^{\zmax}\rho|W|^2dz \notag\\
+%   &=\int_{-\zmax}^{\zmax}\rho\left|\frac{dW}{dz}\right|^2dz 
+%   +\frac{\ii k_x r}{\Omega_k^2}\int_{-\zmax}^{\zmax}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
+% \end{align}
+% It follows that for instability ($\imag\sigma>0$), it is neccessary to
+% have $d\Omega^2/dz\neq 0$, i.e. vertical shear.
+
+\subsection{Maximum growth rate}   
+Here we show that the growth rate is limited by the maximum vertical
+shear in the domain. For convenience, we write $\hat{\sigma} = \hat{\omega} +
+\ii\hat{\nu}$ with $\hat{\omega},\hat{\nu}$ being the real dimensionless
 mode frequency and growth rates. Then the real and imaginary parts of
@@
   &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
-   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z} -
-  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}
+  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
+  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
   \notag\\
   &=\real\left[\ii
-  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
- f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right], \notag\\
+    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
+    f(\hat{z}) W^*W^\prime d\zhat\right], \notag\\
 & 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
-   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z}\notag\\
+   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}\notag\\
    &=\imag\left[\ii
-  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
- f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right].
+     \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
+     f(\hat{z}) W^*W^\prime d\hat{z}\right].
 \end{align}
@@
 &\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
-   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho|W|^2d\hat{z} -
-  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\right]^2\notag\\
+   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
+  \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2\notag\\
 &+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
-  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-   |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\notag\\
+  \int_{\zhat_1}^{\zhat_2}\rho
+   |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
    &=\left|\ii
-  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
- f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|^2.
+  \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
+ f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
 \end{align}
@@
 &4\hat{\nu}^2\left(1+\hat{k}^2\right) 
-  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-   |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}\notag\\
+  \int_{\zhat_1}^{\zhat_2}\rho
+   |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
    &\leq\left|
-  \epsilon\hat{k}q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
- f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|^2.
+  \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
+ f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
 \end{align}
-% We now take the same approach as in the previous section. We first
-% assume an neutral solution corresponding to $q\equiv0$, for which
-% $\sigma$ is real and we can take $W$ to be real as well. We then
-% set $q\to\delta q$ and linearize. This procedure means
-% \begin{align}
-%   &\left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)
-%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}  
-%   =\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z},\\ 
-%   &\frac{2\sigma\imag{\delta\sigma}}{\Omega_k^2}\left(1+\hat{k}^2\right)
-%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho W^2d\hat{z}\notag\\
-%   &=\epsilon\hat{k}\delta q\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-%   f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}. 
-% \end{align}
-% Eliminating $\sigma^2$ between the above equations, we obtain
-% \begin{align}\label{sigma_finite_domain}
-%   &4
-%   \left[\frac{\imag{\left(\delta\sigma\right)}}{\Omega_k}\right]^2\left(1+\hat{k}^2\right) 
-%   \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-%   W^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left(\frac{dW}{d\hat{z}}\right)^2d\hat{z}\notag\\
-%   &=\left(\epsilon\hat{k}\delta q\right)^2\left(\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-%   f(\hat{z}) W\frac{dW}{d\hat{z}}d\hat{z}\right)^2. 
-% \end{align}
 On the left hand side of this inequality, we apply the Cauchy-Schwarz
@@
 \begin{align}
-  &\left( \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-    |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z}\right)^2\notag\\&\leq
-  \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho 
-  |W|^2d\hat{z}\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho\left|\frac{dW}{d\hat{z}}\right|^2d\hat{z}.
+  &\left( \int_{\zhat_1}^{\zhat_2}\rho
+    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
+  \int_{\zhat_1}^{\zhat_2}\rho 
+  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
 \end{align}
@@
 \begin{align}
-  \left|\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-    f(\hat{z}) W^*\frac{dW}{d\hat{z}}d\hat{z}\right|\leq \int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-  \left|f(\hat{z})W^*\frac{dW}{d\hat{z}}\right|d\hat{z} \notag\\
+  \left|\int_{\zhat_1}^{\zhat_2}\rho
+    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
+  \left|f(\hat{z})W^*W^\prime \right|d\hat{z} \notag\\
   \leq
-  \mathrm{max}\left(|f|\right)\int_{-z_\mathrm{max}}^{z_\mathrm{max}}\rho
-  |W|\left|\frac{dW}{d\hat{z}}\right|d\hat{z},
+  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
+  |W|\left|W^\prime \right|d\hat{z},
 \end{align}
 where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
-$z\in[-\zmax,\zmax]$. Inserting these inequalities into
+$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
 Eq. \ref{sigma_finite_domain} gives
 \begin{align}\label{max_growth}
-  \hat{\nu}\equiv\left|\frac{\imag{(\sigma)}}{\Omega_k}\right| \leq
+  |\hat{\nu}|\leq
   \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
@@
 \begin{align}\label{iso_ode3}
-  \frac{d^2W}{d\hat{z}^2} - \left(1 + \ii q\epsilon
-    \hat{k}\right)\hat{z}\frac{dW}{d\hat{z}} +
-  \left(\frac{\sigma}{\Omega_k}\right)^2\left(1+\hat{k}^2\right)W = 
+  W^{\prime\prime} - \left(1 + \ii q\epsilon
+    \hat{k}\right)\hat{z}W^\prime  +
+  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
   0.
@@
 maximum vertical domain size we should consider is limited by the thin disk
-approximation (i.e. $z\lesssim r$). It is most common to take $\zmax\to\infty$
+approximation (i.e. $z\lesssim r$). It is common to take $\mathrm{max}|z|\to\infty$
 and impose that the kinetic  energy remain bounded at infinity.  This
@@
 $\hat{z}$, unlike the true vertical shear profile which eventually
-decays.  Thus taking $\zmax\to$ will permit unphysically large growth
+decays.  Thus, taking $\zhat_1\to -\infty$ and $\zhat_2\to\infty$ will permit unphysically large growth
 rates (as demonstrated below). Nevertheless, the infinite disk is
@@
 \begin{align}\label{hermite_ode}
-  \frac{d}{d\hat{z}}\left[w(\hat{z})\frac{dW}{d\hat{z}}\right] + nW
+  \frac{d}{d\hat{z}}\left[w(\hat{z})W^\prime \right] + nW
   w(\hat{z}) =0, 
@@
 \begin{align}
-  n \equiv \left(\frac{\sigma}{\Omega_k}\right)^2(1+\hat{k}^2). 
+  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
 \end{align} 
@@
 \begin{align}
-  \left|\frac{\sigma}{\Omega_k}\right| = \sqrt{n}
+  \left|\hat{\sigma}\right| = \sqrt{n}
   \left(1+\hat{k}^2\right)^{-1/2}. 
@@
   (m+2)(m+1)a_{m+2} +
-  \left[n - m\left(1+\ii q h \hat{k}\right)\right] a_m = 0, 
+  \left[n - m\left(1+\ii \epsilon q  \hat{k}\right)\right] a_m = 0, 
 \end{align}
@@
 \begin{align}
-\left(\frac{\sigma}{\Omega_k}\right)^2 = M\left(\frac{1+\ii q \epsilon
+\hat{\sigma}^2 = M\left(\frac{1+\ii q \epsilon
     \hat{k}}{1+\hat{k}^2}\right).
@@
 The eigenfrequency $\sigma$ is therefore complex. We are interested in
-unstable modes for which $\imag{\sigma}$ is the growth rate and is
+unstable modes for which $\imag{\hat{\sigma}}>0$ is the growth rate and is
 given by 
 \begin{align}\label{simple_growth}
-   \frac{\imag(\sigma)}{\Omega_k} =\sqrt{
+  \hat{\nu} =\sqrt{
    \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
@@
 \begin{align}
-\mathrm{max}\left[\imag(\sigma)\right] =\frac{\sqrt{M}|\epsilon
-  q|}{2\sqrt{1+|\epsilon q|}}\Omega_k. 
+\mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{M}|\epsilon
+  q|}{2\sqrt{1+|\epsilon q|}}. 
 \end{align}
@@
   &q \to 0 + \delta q\\
-  &W \to \He_n + \delta V \\
-  &\sigma \to \sigma + \delta\sigma, 
+  &W \to \He_n + \delta W \\
+  &\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}. 
 \end{align}
-where for this exercise we restore $\sigma$ as the eigenfrequency.
+%where for this exercise we restore $\sigma$ as the eigenfrequency.
 
@@
 \begin{align}
-  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
+  2\hat{\sigma}\imag(\delta\hat{\sigma})
   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
@@
   \int_{-\infty}^{\infty}
-  w(\zhat)\hat{z}\He_n(\hat{z})\frac{d\He_n}{d\zhat} d\hat{z}
+  w(\zhat)\hat{z}\He_n(\hat{z})\He_n^\prime(\zhat) d\hat{z}
 \end{align}
-Recognizing $\hat{z} = \He_1$ and using $d\He_n/d\hat{z} = n
+Recognizing $\hat{z} = \He_1(\hat{z})$ and using $\He_n^\prime = n
 \He_{n-1}$ we have
  \begin{align}
-   \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
+   2\hat{\sigma}\imag(\delta\hat{\sigma})
    \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
@@
  \begin{align}
-  \frac{2\sigma\imag(\delta\sigma)}{\Omega_k^2}
+   2\hat{\sigma}\imag(\delta\hat{\sigma})
   \left(1+\hat{k}^2\right) = \delta q \epsilon \hat{k} n.
  \end{align}
-Inserting the original eigenfrequency $\sigma = \pm
-\sqrt{n}\Omega_k(1+\hat{k}^2)^{-1/2}$ gives
+Inserting the original eigenfrequency $\hat{\sigma} = \pm
+\sqrt{n}(1+\hat{k}^2)^{-1/2}$ gives
 \begin{align}
-  \frac{\imag(\delta\sigma)}{\Omega_k} = \pm \frac{1}{2}\delta q \epsilon
+  \imag(\delta\hat{\sigma})= \pm \frac{1}{2}\delta q \epsilon
   \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
@@
 \begin{align}
-  \frac{\imag(\delta\sigma)}{\Omega_k} \simeq \pm \frac{1}{2}\delta q \epsilon
+  \imag(\delta\hat{\sigma})\simeq \pm \frac{1}{2}\delta q \epsilon
   \sqrt{n} \sgn{\hat{k}}. 

### 2014-12-15 linear_adia.tex
minor update

@@
 \end{align}
+{\bf note: incompressible mode, set delta to zero. only works for
+  constant gravity.}
 
@@
   \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
-  \delta v_z d\zhat.
+  \delta v_z d\zhat.\label{adia_integral}
 \end{align}
+When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
+$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
+expected, a sub-adiabatically stratified disk is stable in the absense
+of vertical shear. In the presence of vertical shear $q\neq0$ and
+$\hat{\sigma}$ is generally complex, which implies the possibility of
+instability.
 
-\subsection{Explicit solutions in the thin disk limit}
+Note that the term with $\ln\rho^{\prime 2}$ in the integrand (third
+term on the RHS) represents stabilization by the vertical entropy
+gradient when there is no vertical shear. It gives an imaginary
+contribution to $\hat{\sigma}$ when $q\neq0$, but this is a factor
+$\hat{k}$ smaller than the real (stabilizing) contribution, since we
+consider $\hat{k}\gg1$. Thus, instability is expected to be due to the
+last term on the RHS of Eq. \ref{adia_integral}.
+
+\subsection{Explicit solutions in the thin-disk limit}
 To simplify further, we consider the problem in the thin-disk limit so
@@
   &C \equiv \frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
-      \epsilon q}{\hat{k}}\right).
+      \epsilon q}{\hat{k}}\right).\label{adia_thin}
+\end{align}
+Let
+\begin{align}
+  \delta v_z(\zhat) =
+  g(\zhat)\exp\left(\frac{\alpha\zhat^2}{2}\right), \label{adia_ansatz}
+\end{align}
+where $\alpha$ is a constant to be chosen for convenience. We require
+the vertical kinetic energy density $\rho|\delta v_z|^2$ to remain
+finite.  Then assuming
+$g(\zhat)$ is a polynomial, we require 
+\begin{align}
+  \real\alpha < \frac{1}{2}. 
 \end{align}
 
+Inserting the ansatz Eq. \ref{adia_ansatz} into Eq. \ref{adia_thin}
+gives
+\begin{align}
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
+      \alpha\right)g
+    +\underbrace{\left(\alpha^2 - \alpha A - C\right)}_\text{set to zero}\zhat^2 g.
+\end{align}
+We choose $\alpha$ to make the coefficient of $\zhat^2g$ vanish. That
+is,
+\begin{align}
+  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right).
+\end{align} 
+We now have an equation in the same form as Eq. \ref{iso_ode3}:
+\begin{align}
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime +
+  \left(B + \alpha\right)g.
+\end{align}
+We seek polynomial solutions 
+\begin{align}
+  g(\zhat) = \sum_{m=0}^M b_m \zhat^m,
+\end{align}
+% We again seek power-series solutions 
 % \begin{align}
-%   0 =& \frac{d^2\delta v_z}{d\zhat^2} - \hat{z}\left(1 + \ii \hat{k}
-%     \epsilon q \right)\frac{d\delta v_z}{d\hat{z}} \notag\\ 
-% &+ \left[\hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right) -
-%   \left(\frac{1}{\gamma} + \ii \hat{k} \epsilon q\right)\right.\notag\\
-%   &\phantom{0=}-\left.\frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
-%       \epsilon q}{\hat{k}}\right)\zhat^2\right]\delta v_z.\label{adia_iso_final}
+%   g(\zhat) = \sum_{m=0}^\infty b_m \zhat^m,
 % \end{align}
-% Eq. \ref{adia_iso_final} will be the basis for discussion below.
+% which requires
+% \begin{align}
+%   (m+2)(m+1)b_{m+2} + \left[\left(2\alpha - A\right)m + B + \alpha\right]=0.
+% \end{align}
+which requires
+\begin{align}
+  B + \alpha = M\left(A-2\alpha\right).\label{adia_disp}
+\end{align}
+Inserting the definition of $B$ we get
+\begin{align}
+  &\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) \notag\\ &=
+  \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
+  \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}. 
+\end{align}
+%Recalling that $B$ depends $\hat{\sigma}$, Eq. \ref{adia_disp} gives
+%the eigenfrequency as a function of disk paramters.
+
+% \subsection{Special case of a thin-disk with $\gamma=2$} 
+% In principle, we can solve Eq. \ref{adia_disp} for the mode
+% frequencies and growth rates. This is complicated in the
+% general case, but simplifies considerably for $\gamma=2$.

### 2014-12-16 linear_adia.tex
more adiabatic discussion, plus a new fig of growth rates

@@
-\section{Adiabatic disks} 
-We now examine the VSI in adiabatic disks ($t_c\to\infty$). For 
-analytical discussion we consider the initially isothermal case with 
-$\Gamma=1$. We eliminate variables in favor of $\delta v_z$ and
+\subsection{Adiabatic perturbations} 
+We now extend the above discussion to vertically isothermal disks
+($\Gamma=1$) with $\gamma\neq1$ and $t_c\to\infty$, i.e. adiabatic
+evolution. We eliminate variables in favor of $\delta v_z$ and
 $\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
@@
 contribution to $\hat{\sigma}$ when $q\neq0$, but this is a factor
-$\hat{k}$ smaller than the real (stabilizing) contribution, since we
-consider $\hat{k}\gg1$. Thus, instability is expected to be due to the
+$|\epsilon/\hat{k}|$ smaller than the real (stabilizing) contribution,
+since $\epsilon\ll 1$ for a thin disk and we (usually) consider
+$\hat{k}\gg1$. Thus, instability is expected to be due to the 
 last term on the RHS of Eq. \ref{adia_integral}.
 
-\subsection{Explicit solutions in the thin-disk limit}
+\subsubsection{Explicit solutions in the thin-disk limit}
 To simplify further, we consider the problem in the thin-disk limit so
@@
 \begin{align}
-  B + \alpha = M\left(A-2\alpha\right).\label{adia_disp}
+  B + \alpha = M\left(A-2\alpha\right).
 \end{align}
@@
   \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
-  \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}. 
+  \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}\label{adia_disp}. 
 \end{align}
@@
 
-% \subsection{Special case of a thin-disk with $\gamma=2$} 
-% In principle, we can solve Eq. \ref{adia_disp} for the mode
-% frequencies and growth rates. This is complicated in the
-% general case, but simplifies considerably for $\gamma=2$.
+\subsubsection{The case of $\gamma=2$}
+In principle, we can solve Eq. \ref{adia_disp} for the mode
+frequencies and growth rates. This is complicated in 
+general, but simplifies for $\gamma=2$, for which the quantity
+$A^2+4C$ is real. In this case we find
+\begin{align}
+  \hat{\nu}^2 =
+  \frac{1}{2}\frac{(2M+1)}{\sqrt{1+2\hat{k}^2}}&\left[
+    \sqrt{1-\frac{4M(M+1)\hat{k}^2\epsilon^2q^2}{(2M+1)^2(1+2\hat{k}^2)}}\right.\notag\\
+  &\phantom{=}
+  \left.- \sqrt{1 - \frac{\hat{k}^2\epsilon^2q^2}{(1+2\hat{k}^2)}}\right].
+\end{align}
+In the limit of a thin disk $\epsilon\to0$ or weak shear $|q|\to0$, we
+have
+\begin{align}
+  \hat{\nu}^2 = \frac{\epsilon^2
+    q^2}{4(2M+1)}\frac{\hat{k}^2}{(1+2\hat{k}^2)^{3/2}},\label{gam2_growth}
+\end{align}
+and the maximum growth rate occurs at $\hat{k}=\hat{k}_\mathrm{opt}=1$,  
+\begin{align}
+  \mathrm{max}(\hat{\nu}) &= \frac{|\epsilon
+    q|}{2(3^{3/4}\sqrt{2M+1})}\label{gam2_max_growth}.
+  % & \leq \frac{|\epsilon
+  % q|}{2\times3^{3/4}}\notag.
+\end{align}
+Eq. \ref{gam2_max_growth} show that growth rates decrease with $M$,
+and there exists a maximum possible growth rate at $M=0$. This is
+despite the fact that in the thin-disk approximation, for which
+$d\Omega^2/dz\propto z$, there is infinite vertical shear
+available.
+
+\subsubsection{Comparison with $\gamma=1$}
+We can compare growth rates obtained for $\gamma=2$ to those for
+$\gamma=1$ in the previous section. If we compare growth rates at their 
+respective optimum radial wavelengths, we find
+that for $\gamma=2$ the \emph{maximum} possible growth rate, occuring
+at $M=0$, is still $\sim2$ times smaller than the \emph{minimum} growth
+rate for $\gamma=1$. That is,
+\begin{align}
+  &\mathrm{max}(\hat{\nu}) \geq \frac{|\epsilon q|}{2\sqrt{1+|\epsilon
+      q|}} & \text{at }\hat{k} = \hat{k}_\mathrm{opt} \text{ for }   \gamma=1,  \\
+  &\mathrm{max}(\hat{\nu}) \leq \frac{|\epsilon q|}{2\times 3^{3/4}}
+           & \text{at }\hat{k} = \hat{k}_\mathrm{opt} \text{ for } \gamma=2.
+\end{align}
+
+We can also compare growth rates in the limit $\hat{k}\gg1$ using
+Eq. \ref{simple_growth} and Eq. \ref{gam2_growth},
+\begin{align}
+  &\hat{\nu} \geq \sqrt{\frac{|\epsilon q|}{2\hat{k}}}
+  &  \text{as }  \hat{k}\to\infty  \text{ for }  \gamma=1, \\
+  &\hat{\nu} \leq \frac{|\epsilon q|}{2^{7/4}\sqrt{\hat{k}}}
+  &\text{as }  \hat{k}\to\infty  \text{ for }  \gamma=2.
+\end{align}
+Then the growth rate of perturbations with large $|\hat{k}|$ in a
+$\gamma=2$ disk is at most $2^{-5/4}\sqrt{|\epsilon q|}$ times that in a
+$\gamma=1$ disk. For $\epsilon = 0.1$ and $|q|=1$, this factor is
+$\sim 0.1$.
+
+In Fig. \ref{adia_growth} we plot the growth rate $|\hat{\nu}|$ as a
+function of the radial wavenumber $\hat{k}$, for a range of adiabatic
+indices, using Eq. \ref{adia_disp}. The presence of a positive
+vertical entropy gradient is strongly stabilizing. This is because the
+vertical shear grows as $z$ away from the midplane (and has a maximum
+when the thin-disk approximation is relaxed), but the
+square of the bouyancy frequency, which is stabilizing, grows as $z^2$
+away from the midplane. The entropy $P/\rho^\gamma$ formally diverges
+for $z\to\infty$ even without the thin-disk approximation.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/rate_theory_gmma}
+  \caption{Growth rate of the fundamental VSI mode in vertically
+    isothermal disks ($\Gamma=1$) subject to adiabatic perturbations,
+    calculated in thin disk approximation from Eq. \ref{adia_disp}
+    with $M=0$, $\epsilon=0.1$ and  
+    $|q|=1$. Growth rates are shown as a function of the radial
+    wavenumber for different values of the adiabatic index
+    $\gamma$. For $\gamma\equiv 1$ the perturbations are isothermal
+    and there is no vertical entropy gradient.\label{adia_growth}}  
+\end{figure}

### 2014-12-16 linear_iso.tex
more adiabatic discussion, plus a new fig of growth rates

@@
-\section{Vertically isothermal disks with 
-  isothermal perturbations}
+\section{Analytical discussion of vertically isothermal disks}
+\subsection{Isothermal perturbations}
 The linear problem simplifies considerably for the case
@@
 
-\subsection{Neccessity of vertical shear for instability}\label{integral_relation}
+\subsubsection{Neccessity of vertical shear for instability}\label{integral_relation}
 We can establish a neccessary condition for instability with the
@@
 
-\subsection{Maximum growth rate}   
+\subsubsection{Maximum growth rate}   
 Here we show that the growth rate is limited by the maximum vertical
@@
 
-\subsection{Thin-disk limit}
+\subsubsection{Thin-disk limit}
 We can make further progress for thin-disks ($\epsilon\ll1$), 
@@
 
-\subsection{Stability in the absence of vertical shear}
+\subsubsection{Stability in the absence of vertical shear}
 When $q=0$, Eq. \ref{iso_ode3} is
@@
 
-\subsection{Polynomial solutions with vertical shear}
+\subsubsection{Polynomial solutions with vertical shear}
 We can seek power-series solution to Eq. \ref{iso_ode3},
@@
 \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{M}|\epsilon
-  q|}{2\sqrt{1+|\epsilon q|}}. 
+  q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
 \end{align}
@@
 
-\subsection{Instability in the presence of weak vertical shear}
+\subsubsection{Instability in the presence of weak vertical shear}
 We can also investigate the destabilizing effect of vertical shear
@@
 \end{align}
-
-\subsection{Numerical solutions}
-In this section we numerically solve the linear stability problem
-corresponding to Eq. \ref{iso_ode1}. We thus relax the thin-disk
-assumption, using the full expressions for $\rho$, $\Omega$ and
-$\kappa$, so that we may consider vertical domains up to $z\sim r$.
-However, we retain the low-frequency approximation
-$|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue problem. This
-level of approximation lies in between the full problem 
-Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and the thin-disk
-limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.
-
-We solve Eq. \ref{iso_ode1} using a pseudo-spectral method by
-expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$ and
-discretizing the equation on a grid with
-$\zmax=10H_\mathrm{iso}$. This procedure converts the linear problem
-to a standard matrix eigenvalue problem. %In the low-frequency
-%approximation the eigenvalue is $\sigma^2$.
-
-\subsubsection{Fiducial calculation}
-We first validate the low-frequency approximation by solving a 
-fiducial problem with $q=-1$, $p=-1.5$, $\epsilon=0.1$ and $k_x =
-200\pi/r$ with solid boundaries. This is the setup considered by
-\cite{mcnally14} who solved the full problem.
-
-Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
-\ii\nu$. This plot is effectively identical to that in
-\cite{mcnally14}. The low-frequency approximation reproduces the
-three solution branches defined by \citep{nelson13}: corrugation modes
-with odd symmetry, breathing modes with even symmetry, and surface
-modes which correspond to the nearly-vertical locus in
-Fig. \ref{lowfreq_eigen}.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
-  \caption{Eigenvalues in the low-frequency approximation for the
-    vertical shear instability in a vertically isothermal disk evolved
-    isothermally ($\gamma=\Gamma=1$). The disk parameters are $q=-1$,
-    $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
-    wavenumber is $k_x=200\pi/r$. This is the set up considered in
-    \cite{mcnally14}. \label{lowfreq_eigen}
-  }
-\end{figure}
-
-We plot in Fig. \ref{lowfreq_eigenfunc} the structurally simplest
-eigenfunction --- the fundamental corrugation mode corresponding to
-the eigenvalue with $\mathrm{min}|\sigma|$. The perturbation $W$ is
-roughly proportional to $z$, except close the boundaries where it
-flattens because $\delta v_z=dW/dz=0$ is imposed there.
-
-For $k_x = 200\pi/r$ and $\epsilon = 0.1$, the dimensionless
-wavenumber $\hat{k} = k_xH_\mathrm{iso}=20\pi$. Then the expected
-growth rate according to Eq. \ref{simple_growth} with $M=1$ is
-\begin{align*}
-  \nu = 0.2606\epsilon\Omega_k,
-\end{align*}
-as obtained numerically (Fig. \ref{lowfreq_eigen}). This agreement is
-surpsingly good, given that Eq. \ref{simple_growth} assumes the
-thin-disk approximation and imposes a different boundary condition to
-that in the numerical calculations.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/eigenvector_iso}
-  \caption{Eigenfunction of the fundamental corrugation mode,
-    corresponding to the 
-    bottom-left eigenvalue displayed in Fig. \ref{lowfreq_eigen}. The
-    solid and dashed lines are $\real W$ and $\imag W$, respectively. 
-    \label{lowfreq_eigenfunc}
-  }
-\end{figure}
-
-\subsubsection{Influence of vertical boundaries}

### 2014-12-16 numerical.tex
more adiabatic discussion, plus a new fig of growth rates

\section{Numerical solutions}
In this section we numerically solve the linear stability problem
corresponding to Eq. \ref{iso_ode1}. We thus relax the thin-disk
assumption, using the full expressions for $\rho$, $\Omega$ and
$\kappa$, so that we may consider vertical domains up to $z\sim r$.
However, we retain the low-frequency approximation
$|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue problem. This
level of approximation lies in between the full problem 
Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and the thin-disk
limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.

We solve Eq. \ref{iso_ode1} using a pseudo-spectral method by
expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$ and
discretizing the equation on a grid with
$\zmax=10H_\mathrm{iso}$. This procedure converts the linear problem
to a standard matrix eigenvalue problem. %In the low-frequency
%approximation the eigenvalue is $\sigma^2$.

\subsection{Fiducial calculation}
We first validate the low-frequency approximation by solving a 
fiducial problem with $q=-1$, $p=-1.5$, $\epsilon=0.1$ and $k_x =
200\pi/r$ with solid boundaries. This is the setup considered by
\cite{mcnally14} who solved the full problem.

Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
\ii\nu$. This plot is effectively identical to that in
\cite{mcnally14}. The low-frequency approximation reproduces the
three solution branches defined by \citep{nelson13}: corrugation modes
with odd symmetry, breathing modes with even symmetry, and surface
modes which correspond to the nearly-vertical locus in
Fig. \ref{lowfreq_eigen}.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
  \caption{Eigenvalues in the low-frequency approximation for the
    vertical shear instability in a vertically isothermal disk evolved
    isothermally ($\gamma=\Gamma=1$). The disk parameters are $q=-1$,
    $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
    wavenumber is $k_x=200\pi/r$. This is the set up considered in
    \cite{mcnally14}. \label{lowfreq_eigen}
  }
\end{figure}

We plot in Fig. \ref{lowfreq_eigenfunc} the structurally simplest
eigenfunction --- the fundamental corrugation mode corresponding to
the eigenvalue with $\mathrm{min}|\sigma|$. The perturbation $W$ is
roughly proportional to $z$, except close the boundaries where it
flattens because $\delta v_z=dW/dz=0$ is imposed there.

For $k_x = 200\pi/r$ and $\epsilon = 0.1$, the dimensionless
wavenumber $\hat{k} = k_xH_\mathrm{iso}=20\pi$. Then the expected
growth rate according to Eq. \ref{simple_growth} with $M=1$ is
\begin{align*}
  \nu = 0.2606\epsilon\Omega_k,
\end{align*}
as obtained numerically (Fig. \ref{lowfreq_eigen}). This agreement is
surpsingly good, given that Eq. \ref{simple_growth} assumes the
thin-disk approximation and imposes a different boundary condition to
that in the numerical calculations.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvector_iso}
  \caption{Eigenfunction of the fundamental corrugation mode,
    corresponding to the 
    bottom-left eigenvalue displayed in Fig. \ref{lowfreq_eigen}. The
    solid and dashed lines are $\real W$ and $\imag W$, respectively. 
    \label{lowfreq_eigenfunc}
  }
\end{figure}

### 2014-12-16 linear_adia.tex
update adia fig

@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/rate_theory_gmma}
-  \caption{Growth rate of the fundamental VSI mode in vertically
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/rate_theory_om}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_nu}
+  \caption{Real frequency (top) and growth rate (bottom) of the fundamental VSI mode in vertically 
     isothermal disks ($\Gamma=1$) subject to adiabatic perturbations,
@@
     with $M=0$, $\epsilon=0.1$ and  
-    $|q|=1$. Growth rates are shown as a function of the radial
+    $|q|=1$. Frequencies are shown as a function of the radial
     wavenumber for different values of the adiabatic index

### 2014-12-17 appendix.tex
write up edits

--- 2014-12-16 linear_iso.tex
@@
-\section{Analytical discussion of vertically isothermal disks}
-\subsection{Isothermal perturbations}
-The linear problem simplifies considerably for the case
-$\Gamma=\gamma=1$. Then 
-\begin{align}
-  Q=W. 
-\end{align}
-
-Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
-problem for isothermal perturbations in a vertically isothermal disk, 
-\begin{align}\label{iso_ode}
-  \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
-      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
-  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
-\end{align} 
-subject to appropriate boundary conditions. We remark that Eq
-.\ref{iso_ode} is also applicable for    
-$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. adiabatic
-disks with adiabatic vertical stratification), since in that case  
-$Q\simeq W$ as well.
-
-\subsubsection{Neccessity of vertical shear for instability}\label{integral_relation}
-We can establish a neccessary condition for instability with the
-following assumptions:
-
-\begin{enumerate}
-\item We consider eigenvalues such that
-  $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
-  \kappa^2$. The problem simplifies because the eigenvalue now only
-  appears in the last term in Eq. \ref{iso_ode}. 
-\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
-  0$. This is already captured explicitly in Eq. \ref{iso_ode} through
-  the factor $d\Omega^2/dz$. We 
-  therefore simplify further by ignoring the vertical dependence of
-  $\kappa^2$. For clarity, we will replace  $\kappa^2$ with $\Omega_k^2$.  
-\end{enumerate}
-
-{\bf implicit limit on domain size. cannot be too large, otherwise omega reaches small values. should remain of order
-  omegak}
-
-With these approximations and introducing the dimensionless variables
-\begin{align}
-  \hat{z} = z/\Hiso,\quad \hat{k}=k_x\Hiso, \quad \hat{\sigma} = \sigma/\Omega_k,
-\end{align}
-the governing equation can be re-written as
-\begin{align}
-  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
-    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
-\end{align}
-where the primes denote $d/d\zhat$ and $f(\zhat)$ is defined such that
-\begin{align}\label{fz_shear}
-  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
-\end{align}
-
-Next, we multiply Eq. \ref{iso_ode1} by $\rho W^*$ and integrate 
-vertically from $\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary
-terms when integrating by parts, by assuming $W$ or
-$W^\prime$ vanishes at the boundaries, or that the boundaries are 
-sufficiently far away so that the boundary terms are negligible because of the
-decaying background density with increasing height. Then,
-\begin{align}
-  &\hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat \notag\\
-  &=\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
-  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
-\end{align}
-It follows that for instability ($\imag\hat{\sigma}>0$), it is neccessary to
-have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical shear.
-
-% With these approximations we can re-write the governing equation as
-% \begin{align}\label{iso_ode1}
-%   &\frac{d}{dz}\left(\rho e^{-\ii k_x r
-%       \Omega^2/\Omega_k^2}\frac{dW}{dz}\right) +
-%   \sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho W e^{-\ii k_x r
-%     \Omega^2/\Omega_k^2} \notag\\
-%   &=0.
-% \end{align}
-% Next, we multiply Eq. \ref{iso_ode1} by $W^{*}\exp{\left(\ii k_x r
-%     \Omega^2/\Omega_k\right)}$ and integrate 
-% vertically from $z=-\zmax$ to $z=\zmax$. We assume either $W$ or
-% $dW/dz$ vanishes at $z=\pm \zmax$, or that $\zmax$ is sufficiently
-% large so that the boundary terms are negligible because of the
-% decaying background density with increasing height. Then,
-% \begin{align}\label{neccessary_cond}
-%    &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\zmax}^{\zmax}\rho|W|^2dz \notag\\
-%   &=\int_{-\zmax}^{\zmax}\rho\left|\frac{dW}{dz}\right|^2dz 
-%   +\frac{\ii k_x r}{\Omega_k^2}\int_{-\zmax}^{\zmax}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
-% \end{align}
-% It follows that for instability ($\imag\sigma>0$), it is neccessary to
-% have $d\Omega^2/dz\neq 0$, i.e. vertical shear.
-
-\subsubsection{Maximum growth rate}   
-Here we show that the growth rate is limited by the maximum vertical
-shear in the domain. For convenience, we write $\hat{\sigma} = \hat{\omega} +
-\ii\hat{\nu}$ with $\hat{\omega},\hat{\nu}$ being the real dimensionless
-mode frequency and growth rates. Then the real and imaginary parts of
-Eq. \ref{integral_relation1} are
-\begin{align}
-  &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
-  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
-  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
-  \notag\\
-  &=\real\left[\ii
-    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\zhat\right], \notag\\
-& 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
-   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}\notag\\
-   &=\imag\left[\ii
-     \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
-     f(\hat{z}) W^*W^\prime d\hat{z}\right].
-\end{align}
-Adding the square of these equations give
-\begin{align}
-&\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
-   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
-  \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2\notag\\
-&+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
-  \int_{\zhat_1}^{\zhat_2}\rho
-   |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
-   &=\left|\ii
-  \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
- f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
-\end{align}
-It is clear that
-\begin{align}\label{sigma_finite_domain} 
-&4\hat{\nu}^2\left(1+\hat{k}^2\right) 
-  \int_{\zhat_1}^{\zhat_2}\rho
-   |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
-   &\leq\left|
-  \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
- f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
-\end{align}
-On the left hand side of this inequality, we apply the Cauchy-Schwarz
-inequality to obtain
-\begin{align}
-  &\left( \int_{\zhat_1}^{\zhat_2}\rho
-    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
-  \int_{\zhat_1}^{\zhat_2}\rho 
-  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
-\end{align}
-On the right hand side of Eq. \ref{sigma_finite_domain} we have
-\begin{align}
-  \left|\int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
-  \left|f(\hat{z})W^*W^\prime \right|d\hat{z} \notag\\
-  \leq
-  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
-  |W|\left|W^\prime \right|d\hat{z},
-\end{align}
-where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
-$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
-Eq. \ref{sigma_finite_domain} gives
-\begin{align}\label{max_growth}
-  |\hat{\nu}|\leq
-  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
-\end{align}
-Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
-follows that the maximum possible growth rate of unstable modes,
-satisfying the above boundary conditions, is limited by the maximum
-value of the vertical shear in the domain considered.
-
-For the vertical shear profile of the vertically isothermal
-disk, $f(\hat{z}) =
-\hat{z}\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$. If we consider
-$\zmax\to\infty$,  then the growth rate is limited by the global
-maximum of the vertical shear, which occurs at $\zhat
-=1/\sqrt{2}\epsilon$.  
-For vertical domains smaller than this height, the growth rate is
-limited by $|f|$ at the vertical boundaries considered. For 
-$|\epsilon\hat{z}|\ll1$ we have $f\simeq \hat{z}$, in which case we
-expect growth rates to increase linearly with the vertical domain
-size.
-
-{\bf get very similar result for if we replaced $\kappa^2(z)$ by $\Omega^2(z)$ instead. then the $(1+\hat{k}^2)$ factor should be 
-inside integral with extra factor of $\Omega_k^2/\Omega^2$ multiplying the $\hat{k}^2$. but for subkeplerian disks $\Omega^2<\Omega_k^2$ so can replace
-$\Omega^2$ by $\Omega^2_k$ without affecting the inequality on the lhs. on rhs need to replace $f$ by $\tilde{f}\equiv f\Omega_k^2/\Omega^2$. inequality is then not very 
-useful if zmax is very large since $\Omega^2$ may reach small values or even zero at finite height , so $\mathrm{max}|\tilde{f}|$ is large or infinite. 
-but still useful for zmax not too large such that $\Omega^2$ is similar to $\Omega^2$.}
-
-{\bf zmax is limited by: $\Omega^2$ and $\kappa^2$ do not become small, since low frequency assumption will fail if domain reaches heights such that $\Omega\sim0$. also can't assume
-$\Omega$ or $\kappa^2$ is roughly constant in coefficient of ode. (vertical derivatives always kept explicitly) 	
-}
-
-\subsubsection{Thin-disk limit}
-We can make further progress for thin-disks ($\epsilon\ll1$), 
-so we may expand the background density and vertical shear in powers
-of $z/r$. To lowest order we obtain 
-\begin{align}
-  &\rho(\hat{z}) \simeq \rho_{00}
-  \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv \rho_{00}w(\hat{z}),\label{thin_dens}\\
-  &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}. \label{thin_vshear}
-\end{align}
-With this approximation and those in \S\ref{integral_relation}, the
-governing equation becomes 
-\begin{align}\label{iso_ode3}
-  W^{\prime\prime} - \left(1 + \ii q\epsilon
-    \hat{k}\right)\hat{z}W^\prime  +
-  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
-  0.
-\end{align}
-We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
-that derived by \cite{nelson13}, although we have taken a different
-route.
-
-To complete the problem we must specify boundary conditions. The
-maximum vertical domain size we should consider is limited by the thin disk
-approximation (i.e. $z\lesssim r$). It is common to take $\mathrm{max}|z|\to\infty$
-and impose that the kinetic  energy remain bounded at infinity.  This
-is appropriate for the density because both the true density profile
-and its thin-disk approximation decay rapidly away from the midplane.
-
-However, the thin-disk representation of vertical shear diverges with 
-$\hat{z}$, unlike the true vertical shear profile which eventually
-decays.  Thus, taking $\zhat_1\to -\infty$ and $\zhat_2\to\infty$ will permit unphysically large growth
-rates (as demonstrated below). Nevertheless, the infinite disk is
-useful to consider since it permits an analytical discussion.
-
-\subsubsection{Stability in the absence of vertical shear}
-When $q=0$, Eq. \ref{iso_ode3} is
-\begin{align}\label{hermite_ode}
-  \frac{d}{d\hat{z}}\left[w(\hat{z})W^\prime \right] + nW
-  w(\hat{z}) =0, 
-\end{align}
-where we have defined a new eigenvalue
-\begin{align}
-  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
-\end{align} 
-Eq. \ref{iso_ode3} is Hermite's differential equation. If we impose
-that the kinetic energy density remain bounded at infinity, then  
-\begin{align}
-  W \propto \He_n(\hat{z}),
-\end{align}
-and $n$ is a non-negative integer. This translates to a real
-eigenfrequency $\sigma$ such that
-\begin{align}
-  \left|\hat{\sigma}\right| = \sqrt{n}
-  \left(1+\hat{k}^2\right)^{-1/2}. 
-\end{align}
-Since we have assumed $|\sigma^2|\ll \kappa^2$, our analysis is only
-valid for large wavenumbers such that $\hat{k}^2\gg 
-n-1$. Physically, this corresponds to radial length-scales much
-smaller than the local disk scale height.
-
-\subsubsection{Polynomial solutions with vertical shear}
-We can seek power-series solution to Eq. \ref{iso_ode3},
-\begin{align}
-  W(\zhat) = \sum_{m=0}^\infty a_m\zhat^m. 
-\end{align}
-Then the coefficients must satisfy the recurrence relation
-\begin{align}
-  (m+2)(m+1)a_{m+2} +
-  \left[n - m\left(1+\ii \epsilon q  \hat{k}\right)\right] a_m = 0, 
-\end{align}
-where $n$ is not neccessarily an integer. Indeed, if we demand
-a polynomial of degree $M$ as the solution, then the eigenfrequency
-must satisfy
-\begin{align}
-\hat{\sigma}^2 = M\left(\frac{1+\ii q \epsilon
-    \hat{k}}{1+\hat{k}^2}\right).
-\end{align}
-The eigenfrequency $\sigma$ is therefore complex. We are interested in
-unstable modes for which $\imag{\hat{\sigma}}>0$ is the growth rate and is
-given by 
-\begin{align}\label{simple_growth}
-  \hat{\nu} =\sqrt{
-   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
-    1\right)}. 
-\end{align}
-For fixed $q$ and $\epsilon$, the growth rate vanishes for both
-$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. The maximum growth rate
-occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
-\begin{align}
-  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
-\end{align}
-and
-\begin{align}
-\mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{M}|\epsilon
-  q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
-\end{align}
-
-% This expression can be simplified in the limit $|q|\ll 1$, 
-%  \begin{align}\label{simple_growth}
-%    \frac{\imag(\sigma)}{\Omega_k} = \pm \sqrt{M} 
-%    \frac{q\epsilon\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
-%  \end{align}
-%  with $M$ being an integer.
-
-\subsubsection{Instability in the presence of weak vertical shear}
+\section{Alternative inference of instability in the presence of weak
+  vertical shear}\label{pert_theory}
 We can also investigate the destabilizing effect of vertical shear
@@
 \begin{align}   
-  &q \to 0 + \delta q\\
-  &W \to \He_n + \delta W \\
-  &\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}. 
+  q \to 0 + \delta q,\quad
+  W \to \He_n + \delta W,\quad
+  \hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}. 
 \end{align}
@@
   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-  \He_n^2(\hat{z}) d\hat{z} \notag\\
+  \He_n^2(\hat{z}) d\hat{z}
   = \delta q \epsilon \hat{k} 
@@
    \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-   \He_n^2(\hat{z}) d\hat{z} \notag\\
+   \He_n^2(\hat{z}) d\hat{z}
    =\delta q \epsilon \hat{k} 
@@
   \int_{-\infty}^{\infty}
-  w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl}, 
-\end{align}
-and
-\begin{align}
-  &\int_{-\infty}^{\infty}
-  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi\notag\\& =
+  w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl} \quad
+% \end{align}
+\text{and} \quad
+% \begin{align}
+  \int_{-\infty}^{\infty}
+  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi =
   \frac{\sqrt{2\pi}m!k!l!}{(j-m)!(j-k)!(j-l)!}, 
@@
 where $j = (m+k+l)/2$. We then obtain
-
  \begin{align}

### 2014-12-17 linear_adia.tex
write up edits

@@
 \subsection{Adiabatic perturbations} 
-We now extend the above discussion to vertically isothermal disks
-($\Gamma=1$) with $\gamma\neq1$ and $t_c\to\infty$, i.e. adiabatic
-evolution. We eliminate variables in favor of $\delta v_z$ and
-$\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
+We now extend the above discussion to general $\gamma$, but in the
+adiabatic limit ($t_c\to\infty$). We eliminate variables from 
+Eq. \ref{lin_mass}---\ref{lin_energy} in favor of $\delta v_z$ 
+and $\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
 \begin{align}
@@
 equations. We thus make the replacement $D\to \Omega_k^2$, as before
-for isothermal perturbations. Here, we also use
-the fact that in the global disk,
+for isothermal perturbations. Here, we also use the fact that in the
+global disk, 
 \begin{align}
-  r\frac{\p \Omega^2}{\p z} = - \frac{\p\ln\rho}{\p z}\frac{\p
-    c_s^2}{\p r} = -\frac{\p \ln\rho}{\p z} \frac{qc_s^2}{r},
+  r\frac{\p \Omega^2}{\p z} = -\frac{\p \ln\rho}{\p z} \frac{qc_s^2}{r},
 \end{align}
-where the second equality follows from the power-law prescription of
-the sound-speed for $\Gamma=1$.
+for the power-law prescription of the midplane sound-speed for
+$\Gamma=1$.
 
 Combining Eq. \ref{adia_iso1} and Eq. \ref{adia_iso2} with the above
-substitutions we obtain, in terms of dimensionless variables defined previously, an equation for $\delta v_z$,
+substitutions we obtain, in terms of dimensionless variables defined
+previously, an equation for $\delta v_z$, 
 \begin{align}
@@
 since $\epsilon\ll 1$ for a thin disk and we (usually) consider
-$\hat{k}\gg1$. Thus, instability is expected to be due to the 
+$\hat{k}\gg1$. Thus, instability is expected to associated with the 
 last term on the RHS of Eq. \ref{adia_integral}.
@@
 \subsubsection{Explicit solutions in the thin-disk limit}
-To simplify further, we consider the problem in the thin-disk limit so
-that $\ln \rho \simeq -\zhat^2/2$. Eq. \ref{adia_iso3} becomes 
+We can construct  solutions in the thin-disk limit, in which 
+$\ln(\rho/\rho_0) \simeq -\zhat^2/2$. Eq. \ref{adia_iso3} becomes   
 \begin{align}
@@
 \end{align}
-where $\alpha$ is a constant to be chosen for convenience. We require
+where $\alpha$ is a constant to be chosen for convenience. Inserting
+Eq. \ref{adia_ansatz} into Eq. \ref{adia_thin} gives
+\begin{align}
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
+    \alpha\right)g
+  +\left(\alpha^2 - \alpha A - C\right)\zhat^2 g.
+\end{align}
+We choose $\alpha$ to make the coefficient of $\zhat^2g$
+vanish. Furthermore, we require 
 the vertical kinetic energy density $\rho|\delta v_z|^2$ to remain
-finite.  Then assuming
-$g(\zhat)$ is a polynomial, we require 
+finite. Then assuming $g(\zhat)$ is a polynomial, we require 
 \begin{align}
@@
 \end{align}
-
-Inserting the ansatz Eq. \ref{adia_ansatz} into Eq. \ref{adia_thin}
-gives
-\begin{align}
-  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
-      \alpha\right)g
-    +\underbrace{\left(\alpha^2 - \alpha A - C\right)}_\text{set to zero}\zhat^2 g.
-\end{align}
-We choose $\alpha$ to make the coefficient of $\zhat^2g$ vanish. That
-is,
+We therefore choose 
 \begin{align}
@@
 \end{align}
+
 We now have an equation in the same form as Eq. \ref{iso_ode3}:
@@
 \end{align}
-Eq. \ref{gam2_max_growth} show that growth rates decrease with $M$,
-and there exists a maximum possible growth rate at $M=0$. This is
-despite the fact that in the thin-disk approximation, for which
-$d\Omega^2/dz\propto z$, there is infinite vertical shear
-available.
+Unlike for isothermal perturbations, Eq. \ref{gam2_max_growth} show
+that growth rates decrease with $M$, and there exists a maximum
+possible growth rate at $M=0$. This is despite the fact that in the
+thin-disk approximation, for which $d\Omega^2/dz\propto z$, there is
+infinite vertical shear available.
+
+The acceptable solution for $\alpha$ is
+\begin{align}
+  \alpha = \frac{1}{2}\left[\left(1+\ii\epsilon q \hat{k}\right) -
+    \sqrt{1 + \hat{k}^2\left(2-\epsilon^2 q^2\right)}\right]. 
+\end{align}
+Since $\epsilon\ll 1$, $\real\alpha<0$, so the vertical velocity
+eventually decays for large heights. This is unlike for isothermal
+perturbations, in which perturbation magnitudes increase with $|z|$.
 
@@
 respective optimum radial wavelengths, we find
-that for $\gamma=2$ the \emph{maximum} possible growth rate, occuring
+that for $\gamma=2$ the \emph{maximum} growth rate, occuring
 at $M=0$, is still $\sim2$ times smaller than the \emph{minimum} growth

### 2014-12-17 linear_iso.tex
minor notation change
write up edits

@@
 \section{Analytical discussion of vertically isothermal disks} 
+We begin with a discussion of vertically isothermal disks
+where $\Gamma=1$, then $c_s$ is a constant in the shearing box. We
+examine separately the limiting case of isothermal perturbations
+($\gamma\equiv1$) and adiabatic perturbations ($\gamma\neq 1$ and
+$t_c\to\infty$).   
 \subsection{Isothermal perturbations}
@@
 \end{align}
-
 Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
@@
 .\ref{iso_ode} is also applicable for    
-$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. adiabatic
-disks with adiabatic vertical stratification), since in that case  
-$Q\simeq W$ as well.
+$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. neutrally
+stratified adiabatic disks), since in that case  
+$Q\simeq W$ as well. However, the following discussion is only
+valid for $\gamma=\Gamma=1$ because $c_s$ is taken to be a constant.
 
-\subsubsection{Neccessity of vertical shear for instability}\label{integral_relation}
-We can establish a neccessary condition for instability with the
-following assumptions:
+\subsubsection{Approximations}
+To facilitate an analytical discussion, we simplify the governing
+equation Eq. \ref{iso_ode} with the following approximations:
 
@@
   0$. This is already captured explicitly in Eq. \ref{iso_ode} through
-  the factor $d\Omega^2/dz$. We 
-  therefore simplify further by ignoring the vertical dependence of
-  $\kappa^2$. For clarity, we will replace  $\kappa^2$ with $\Omega_k^2$.  
+  the factor $d\Omega^2/dz$. We therefore simplify further by ignoring
+  the vertical dependence of $\kappa^2$ elsewhere. For clarity, we
+  will replace $\kappa^2$ with $\Omega_k^2$.%, although the discussion
+  % is unchanged if we replaced 
+  % $\kappa^2$ with $\kappa^2(0)$, for example.    
 \end{enumerate}
+These assumptions amount to making the replacement $D\to\Omega_k^2$.  
+We will consider sound-speed profiles with $q<0$ in the global 
+disk. Then $\Omega^2(z)$ decreases away from the midplane,
+and may reach small values for sufficiently large $|z|$, which
+invalidates the low-frequency approximation. In principle, then, the
+above approximations implicitly limits us to consider thin disks ($\epsilon
+\ll 1$) with vertical domain sizes not too large, so that the rotation
+is nearly-Keplerian.
 
-{\bf implicit limit on domain size. cannot be too large, otherwise omega reaches small values. should remain of order
-  omegak}
-
-With these approximations and introducing the dimensionless variables
+\subsubsection{Neccessity of vertical shear for
+  instability}\label{integral_relation} 
+With the above approximations and introducing the dimensionless variables
 \begin{align}
@@
 
-Next, we multiply Eq. \ref{iso_ode1} by $\rho W^*$ and integrate 
-vertically from $\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary
+We can establish a necessary condition for instability by multipling
+Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
+$\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
 terms when integrating by parts, by assuming $W$ or
@@
 
-% With these approximations we can re-write the governing equation as
-% \begin{align}\label{iso_ode1}
-%   &\frac{d}{dz}\left(\rho e^{-\ii k_x r
-%       \Omega^2/\Omega_k^2}\frac{dW}{dz}\right) +
-%   \sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right) \rho W e^{-\ii k_x r
-%     \Omega^2/\Omega_k^2} \notag\\
-%   &=0.
-% \end{align}
-% Next, we multiply Eq. \ref{iso_ode1} by $W^{*}\exp{\left(\ii k_x r
-%     \Omega^2/\Omega_k\right)}$ and integrate 
-% vertically from $z=-\zmax$ to $z=\zmax$. We assume either $W$ or
-% $dW/dz$ vanishes at $z=\pm \zmax$, or that $\zmax$ is sufficiently
-% large so that the boundary terms are negligible because of the
-% decaying background density with increasing height. Then,
-% \begin{align}\label{neccessary_cond}
-%    &\sigma^2\left(\frac{1}{c_s^2}+\frac{k_x^2}{\Omega_k^2}\right)\int_{-\zmax}^{\zmax}\rho|W|^2dz \notag\\
-%   &=\int_{-\zmax}^{\zmax}\rho\left|\frac{dW}{dz}\right|^2dz 
-%   +\frac{\ii k_x r}{\Omega_k^2}\int_{-\zmax}^{\zmax}\rho\frac{d\Omega^2}{dz}W^*\frac{dW}{dz}dz. 
-% \end{align}
-% It follows that for instability ($\imag\sigma>0$), it is neccessary to
-% have $d\Omega^2/dz\neq 0$, i.e. vertical shear.
-
 \subsubsection{Maximum growth rate}   
 Here we show that the growth rate is limited by the maximum vertical
-shear in the domain. For convenience, we write $\hat{\sigma} = \hat{\omega} +
-\ii\hat{\nu}$ with $\hat{\omega},\hat{\nu}$ being the real dimensionless
-mode frequency and growth rates. Then the real and imaginary parts of
+shear in the domain. Writing $\hat{\sigma} = \hat{\omega} +
+\ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, the real and
+imaginary parts of 
 Eq. \ref{integral_relation1} are
@@
 
-For the vertical shear profile of the vertically isothermal
-disk, $f(\hat{z}) =
-\hat{z}\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$. If we consider
-$\zmax\to\infty$,  then the growth rate is limited by the global
-maximum of the vertical shear, which occurs at $\zhat
-=1/\sqrt{2}\epsilon$.  
-For vertical domains smaller than this height, the growth rate is
-limited by $|f|$ at the vertical boundaries considered. For 
-$|\epsilon\hat{z}|\ll1$ we have $f\simeq \hat{z}$, in which case we
-expect growth rates to increase linearly with the vertical domain
-size.
-
-{\bf get very similar result for if we replaced $\kappa^2(z)$ by $\Omega^2(z)$ instead. then the $(1+\hat{k}^2)$ factor should be 
-inside integral with extra factor of $\Omega_k^2/\Omega^2$ multiplying the $\hat{k}^2$. but for subkeplerian disks $\Omega^2<\Omega_k^2$ so can replace
-$\Omega^2$ by $\Omega^2_k$ without affecting the inequality on the lhs. on rhs need to replace $f$ by $\tilde{f}\equiv f\Omega_k^2/\Omega^2$. inequality is then not very 
-useful if zmax is very large since $\Omega^2$ may reach small values or even zero at finite height , so $\mathrm{max}|\tilde{f}|$ is large or infinite. 
-but still useful for zmax not too large such that $\Omega^2$ is similar to $\Omega^2$.}
-
-{\bf zmax is limited by: $\Omega^2$ and $\kappa^2$ do not become small, since low frequency assumption will fail if domain reaches heights such that $\Omega\sim0$. also can't assume
-$\Omega$ or $\kappa^2$ is roughly constant in coefficient of ode. (vertical derivatives always kept explicitly) 	
-}
+For the vertically isothermal disk we have $|f(\hat{z})| =
+|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, which maximises
+at $|\zhat|=1/\sqrt{2}\epsilon$. For vertical domains smaller 
+than this height, the growth rate is limited by $|f|$ at the vertical
+boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
+\hat{z}$, in which case we can expect growth rates to increase linearly
+with the vertical domain size.   %no faster than linear
 
 \subsubsection{Thin-disk limit}
-We can make further progress for thin-disks ($\epsilon\ll1$), 
-so we may expand the background density and vertical shear in powers
+We can make further progress for thin disks ($\epsilon\ll1$), 
+by expanding the background density and vertical shear profile in powers
 of $z/r$. To lowest order we obtain 
 \begin{align}
-  &\rho(\hat{z}) \simeq \rho_{00}
-  \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv \rho_{00}w(\hat{z}),\label{thin_dens}\\
-  &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}. \label{thin_vshear}
+  &\frac{\rho}{\rho_0} \simeq 
+  \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv w(\hat{z}),\label{thin_dens}\\
+  &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}, \label{thin_vshear}
 \end{align} 
-With this approximation and those in \S\ref{integral_relation}, the
-governing equation becomes 
+%With this approximation and those in \S\ref{integral_relation}, 
+and the governing equation becomes 
 \begin{align}\label{iso_ode3}
@@
 To complete the problem we must specify boundary conditions. The 
-maximum vertical domain size we should consider is limited by the thin disk
-approximation (i.e. $z\lesssim r$). It is common to take $\mathrm{max}|z|\to\infty$
-and impose that the kinetic  energy remain bounded at infinity.  This
-is appropriate for the density because both the true density profile
+maximum vertical domain size should be limited by the thin
+disk and low-frequency approximations (i.e. $z\lesssim r$).     
+It is, however, common to take $\mathrm{max}|z|\to\infty$ and impose
+that the kinetic  energy remain bounded at infinity. This 
+is appropriate for the density field because both the true density profile 
 and its thin-disk approximation decay rapidly away from the midplane.
@@
 $\hat{z}$, unlike the true vertical shear profile which eventually
-decays.  Thus, taking $\zhat_1\to -\infty$ and $\zhat_2\to\infty$ will permit unphysically large growth
+decays.  Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit unphysically large growth
 rates (as demonstrated below). Nevertheless, the infinite disk is
@@
 \begin{align}\label{hermite_ode}
-  \frac{d}{d\hat{z}}\left[w(\hat{z})W^\prime \right] + nW
+  \left[w(\hat{z})W^\prime \right]^\prime + nW
   w(\hat{z}) =0, 
@@
 \end{align}
-Since we have assumed $|\sigma^2|\ll \kappa^2$, our analysis is only
-valid for large wavenumbers such that $\hat{k}^2\gg 
+Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
+analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
 n-1$. Physically, this corresponds to radial length-scales much
@@
 \begin{align}
-  W(\zhat) = \sum_{m=0}^\infty a_m\zhat^m. 
+  W(\zhat) = \sum_{l=0}^\infty a_l\zhat^l. 
 \end{align}
@@
 \begin{align}
-  (m+2)(m+1)a_{m+2} +
-  \left[n - m\left(1+\ii \epsilon q  \hat{k}\right)\right] a_m = 0, 
+  (l+2)(l+1)a_{l+2} +
+  \left[n - l\left(1+\ii \epsilon q  \hat{k}\right)\right] a_l = 0, 
 \end{align}
 where $n$ is not neccessarily an integer. Indeed, if we demand
-a polynomial of degree $M$ as the solution, then the eigenfrequency
+a polynomial of degree $L$ as the solution, then the eigenfrequency
 must satisfy
 \begin{align}
-\hat{\sigma}^2 = M\left(\frac{1+\ii q \epsilon
+\hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
     \hat{k}}{1+\hat{k}^2}\right).
 \end{align}
-The eigenfrequency $\sigma$ is therefore complex. We are interested in
+The eigenfrequency $\hat{\sigma}$ is therefore complex for
+$L\geq1$. We are interested in 
 unstable modes for which $\imag{\hat{\sigma}}>0$ is the growth rate and is
@@
   \hat{\nu} =\sqrt{
-   \frac{M}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
+   \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
     1\right)}. 
@@
 \begin{align}
-\mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{M}|\epsilon
+  \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
     q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
 \end{align}
+Since $L\geq1$, there is no limit to the maximum growth rate. This is
+an artifact of the thin-disk approximation because the vertical 
+shear $d\Omega^2/dz\propto z$ increases indefinitely with
+height. However, we find Eq. \ref{iso_max_growth} compares well with
+numerical solutions of the governing equation for small $L$.
 
@@
 
-\subsubsection{Instability in the presence of weak vertical shear}
-We can also investigate the destabilizing effect of vertical shear
-without explicitly solving the full governing equation,
-Eq. \ref{iso_ode3}. We first
-consider a system with $q\equiv0$, for which the eigenfunctions are
-Hermite polynomials and the real eigenfrequencies are known. We then
-perturb this system by introducing a weak vertical shear ($|q|\ll1$)
-and linearize the governing equation. This procedure is
-\begin{align}   
-  &q \to 0 + \delta q\\
-  &W \to \He_n + \delta W \\
-  &\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}. 
-\end{align}
-%where for this exercise we restore $\sigma$ as the eigenfrequency.
-
-Linearizing the integral relation Eq. \ref{integral_relation1} in the
-thin-disk limit, and taking the
-imaginary part, we have
-\begin{align}
-  2\hat{\sigma}\imag(\delta\hat{\sigma})
-  \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-  \He_n^2(\hat{z}) d\hat{z} \notag\\
-  = \delta q \epsilon \hat{k} 
-  \int_{-\infty}^{\infty}
-  w(\zhat)\hat{z}\He_n(\hat{z})\He_n^\prime(\zhat) d\hat{z}
-\end{align}
-Recognizing $\hat{z} = \He_1(\hat{z})$ and using $\He_n^\prime = n
-\He_{n-1}$ we have
- \begin{align}
-   2\hat{\sigma}\imag(\delta\hat{\sigma})
-   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-   \He_n^2(\hat{z}) d\hat{z} \notag\\
-   =\delta q \epsilon \hat{k} 
-   \int_{-\infty}^{\infty}
-   w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
- \end{align}
-
-Finally, we evaluate the integrals using the results
-\begin{align}
-  \int_{-\infty}^{\infty}
-  w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl}, 
-\end{align}
-and
-\begin{align}
-  &\int_{-\infty}^{\infty}
-  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi\notag\\& =
-  \frac{\sqrt{2\pi}m!k!l!}{(j-m)!(j-k)!(j-l)!}, 
-\end{align}
-where $j = (m+k+l)/2$. We then obtain
-
- \begin{align}
-   2\hat{\sigma}\imag(\delta\hat{\sigma})
-  \left(1+\hat{k}^2\right) = \delta q \epsilon \hat{k} n.
- \end{align}
-Inserting the original eigenfrequency $\hat{\sigma} = \pm
-\sqrt{n}(1+\hat{k}^2)^{-1/2}$ gives
-\begin{align}
-  \imag(\delta\hat{\sigma})= \pm \frac{1}{2}\delta q \epsilon
-  \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
-\end{align}
-This result agrees with Eq. \ref{simple_growth} in the limit
-$|q|\to0$. 
-%for $n=1$, since the
-%function $\He_1$ solves the governing equation exactly. 
-For $\hat{k}\gg 1$ we have
-\begin{align}
-  \imag(\delta\hat{\sigma})\simeq \pm \frac{1}{2}\delta q \epsilon
-  \sqrt{n} \sgn{\hat{k}}. 
-\end{align}
+In Appendix \ref{pert_theory} we give an alternative method to
+infer instability in the presence of vertical shear without explicitly
+solving the governing equation.

### 2014-12-17 setup.tex
minor update with notes
write up edits

@@
 \section{Baroclinic disk equilibra}
- We consider power-law disks with midplane density
+We consider  an inviscid, non-self-gravitating three-dimensional (3D)
+disk orbiting a central star of mass $M_*$. 
+We adopt cylindrical
+co-ordinates $(r,\phi, z)$ centered on the star to describe the global
+disk. The equilibrium disk is steady and axisymmetric with density,
+pressure and rotation profiles $\rho(r,z)$, $P(r,z)$ and
+$\Omega(r,z)$, respectively.
+
+We examine disks with power-law profiles for the midplane density
+$\rho_0(r)\equiv\rho(r,0)$: 
 \begin{align}
@@
 where $\rho_{00}$ is the midplane density at the fiducial radius
- $r=r_0$, and midplane sound-speed 
+$r=r_0$ and $p$ is the power-law index. The equilibrium pressure $P$
+and density $\rho$ are related by 
+\begin{align}
+  P(r,z) &= 
+  \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z),\notag\\
+  & \equiv K(r)\rho^{\Gamma}(r,z). \label{eqm_press}
+\end{align}
+where the polytropic index $\Gamma$ is a parameter and 
+the midplane sound-speed $c_{s0}(r)$ is taken to be another power-law: 
 \begin{align}
@@
 \end{align}
- where $c_{s00}$ is the midplane sound-speed at $r=r_0$. We also
- define
+where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the power-law index for
+the midplane temperature. We define
 \begin{align}
@@
 \end{align}
-to parametrize the magnitude of the sound speed, where
-$v_k=r\Omega_k=\sqrt{GM/r}$ is the Keplerian azimuthal velocity.
-
-The three-dimensional equilibrium pressure $P$ and density $p$ are
-related by 
-\begin{align}
-  P(r,z) = 
-  \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z), 
-\end{align}
-where we refer to $\Gamma$ as the polytropic index. We define the
-equilibrium sound-speed
+to parametrize the magnitude of the sound-speed, where
+$v_k=r\Omega_k=\sqrt{GM_*/r}$ is the Keplerian azimuthal velocity and
+$G$ is the gravitational constant. 
+The global sound-speed $c_s(r,z)$ is given by
 \begin{align}
@@
 
-The equilibrium density is given by vertical hydrostatic balance. This
-gives, for $\Gamma\neq1$,  
+The equilibrium disk satisfies
+\begin{align}
+  0 &= \frac{1}{\rho}\frac{\p P}{\p z} + \frac{\p \Phi_*}{\p z},\label{vert_equilibrium}\\
+  r\Omega^2&= \frac{1}{\rho}\frac{\p P}{\p r} + \frac{\p \Phi_*}{\p
+    r},\label{hori_equilibrium} 
+\end{align} 
+where $\Phi_* = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational
+potential of the central star. Note that the disk is in general
+baroclinic with $\Omega = \Omega(r,z)$, since
+Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} imply, with
+Eq. \ref{eqm_press},  
+\begin{align}
+  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p z}\frac{dK}{dr}.
+\end{align}
+
+The explicit expression for the density field for
+$\Gamma\neq1$ is given  by
 \begin{align}\label{eqm_dens}
   &\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} = 1 +
-  \frac{\left(\Gamma-1\right)}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right). 
+  \frac{\left(\Gamma-1\right)}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right),
 \end{align}
@@
 
-The equilibrium rotation profile is obtained by balancing radial
-pressure force, centrifugal force and radial gravity from the central
-star. This gives 
+Similarly, the equilibrium rotation is 
 \begin{align}\label{eqm_rot}
@@
 \end{align}
-We note that rotation profile generally depends on both $r$ and $z$,
-unless the equilibrium is barotropic with $s=0$, in which case
-$\Omega$ only depends on $r$.
+
+% {\bf need to mention: generally consider sub-keplerian disks -> caveat is that it's entirely possble for angular 
+% velocity to reach zero at finite height. give example later for isothermal disk. this limits domain size, but this critical height is usually
+% very large for thin disks}
 
@@
 \begin{align}
-   \frac{\rho(r,z)}{\rho_0(r)} &= \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\   
+  \frac{\rho(r,z)}{\rho_0(r)} &=
+  \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\    
   \frac{\Omega^2(r,z)}{\Omega_k^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
@@
 which may be obtained from Eq. \ref{eqm_dens} and Eq. \ref{eqm_rot} by
-taking the limit $\Gamma\to 1$.
-
-The vertically isothermal disk has no surface, but we can define a
-characteristic scale-height
+taking the limit $\Gamma\to 1$. The vertically isothermal disk has no
+surface, but we can define a characteristic scale-height 
 \begin{align}
-  \Hiso \equiv \epsilon r \quad\quad(\Gamma \equiv 1),   
+  \Hiso \equiv \epsilon r.% \quad\quad(\Gamma \equiv 1),   
 \end{align} 
@@
 
-% where $\Omega_k\equiv\sqrt{GM/r^3}=v_k/r$ is the Keplerian rotation 
- % and 
- % \begin{align}
- %   h(r) = \frac{c_{s0}(r)}{v_k(r)} 
- % \end{align}
- % is the local disk aspect ratio. 
-%The disk scale height is $H=hr$.
-
- % The vertical shear is 
- % \begin{align}
- %   \frac{\p\Omega^2}{\p z} =
- %   \frac{q\Omega_k^2z}{r^2\left(1+z^2/r^2\right)^{3/2}}. 
- % \end{align}
-
 \section{The vertically global shearing box} 
-% We are interested in the stability problem on radial length-scales
-% much smaller than the local disk scale-height $H$. 
- We work within the framework of the \emph{vertically global shearing
+We are interested in the axisymmetric stability of the above system,
+on radial length-scales much smaller than the local radius $r$, but
+may be global in the vertical direction. As such, we work within the
+framework of the \emph{vertically global shearing 
   box} (VGSB) developed by \cite{mcnally14}. In this approximation
@@
 
- In this shearing box the equilibrium density, pressure and 
- temperatures are assumed to depend only on $z$, by evaluating these
- quantities at $r=r_0$. The background rotation is approximated by a
- linear shear in $x$, but left unaltered in $z$. 
-% This corresponds to
-%  setting 
-% \begin{align}
-%   \rho(r,z)\to\rho(r_0,z),\,P \to P(r_0,z), c_s(r,z)^2 \to
-%   c_s^2(r_0,z),  \\
-% \end{align}
+In the this shearing box the equilibrium density, pressure and
+sound-speed are assumed to depend only on $z$, by evaluating these
+quantities at $r=r_0$ in the global disk. The background rotation is
+approximated by a linear shear in $x$, but left unaltered in $z$.
+
 The axisymmetric VGSB equations read 
@@
 and similarly for other fluid variables, where $k_x$ is a horizontal
- wavenumber and $\sigma$ is a (generally) complex frequency. For
- clarity we drop the subscript $0$ on the frequencies and shear rates
- below.
-
- % {\bf below: at the moment only valid for locally isothermal
- %  equilibrium ($\Gamma=1$)} 
- Writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, 
- the linearized system of equations is 
+wavenumber and $\sigma = \omega + \ii \nu$ is a (generally) complex
+frequency, with $\omega$ amd $\nu$ being the real frequency and growth
+rate, respectively. For clarity we drop the subscript $0$ on the
+frequencies and shear rates below. Then writing $\delta P /\rho \equiv W$
+and $c_s^2\delta\rho/\rho\equiv Q$, the linearized system of equations
+is   
  \begin{align}
    -&\ii\sigma \frac{Q}{c_s^2} + \ii k_x \delta v_x + \frac{d\dd
-     v_z}{dz} + \frac{d\ln{\rho}}{dz}\delta v_z = 0,\\
-   &\ii\sigma \delta v_x + 2\Omega\delta v_y = \ii k_x W,\\
+     v_z}{dz} + \frac{d\ln{\rho}}{dz}\delta v_z = 0,\label{lin_mass}\\
+   &\ii\sigma \delta v_x + 2\Omega\delta v_y = \ii k_x W,\label{lin_vx}\\
    -&\frac{\kappa^2}{2\Omega}\delta v_x + \ii \sigma\delta v_y =
-   r\frac{d\Omega}{dz}\delta v_z, \\
+   r\frac{d\Omega}{dz}\delta v_z, \label{lin_vy}\\
    & \ii\sigma\delta v_z = \frac{dW}{dz} +
-   \frac{d\ln{\rho}}{dz}\left(W-Q\right), \\
+   \frac{d\ln{\rho}}{dz}\left(W-Q\right), \label{lin_vz}\\
    -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
-   c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.
+   c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
  \end{align}
@@
 \end{align}
+Eq. \ref{ode_w}---\ref{ode_Q} are convenient for numerical work, 
+but alternative forms are more suitable for analytical discussion, and 
+will be derived later.
 

### 2014-12-17 linear_adia.tex
minor updates

@@
     - \frac{r}{k_x}\frac{d\Omega^2}{dz}\right)\delta v_z,\label{adia_iso1}\\
-  & -\sigma^2\delta v_z = \gamma c_s^2 \frac{d\Delta}{dz} +
-  c_s^2\frac{d}{dz}\left(\frac{d\ln\rho}{dz}\delta v_z\right) +
+  & -\frac{\sigma^2}{c_s^2}\delta v_z = \gamma \frac{d\Delta}{dz} +
+  \frac{d}{dz}\left(\frac{d\ln\rho}{dz}\delta v_z\right) +
   \left(\gamma-1\right)\frac{d\ln\rho}{dz} \Delta.\label{adia_iso2}
 \end{align}
-{\bf note: incompressible mode, set delta to zero. only works for
-  constant gravity.}
+%{\bf note: incompressible mode, set delta to zero. only works for
+% constant gravity.}
 
-We again make the low-frequency approximation
-and ignore the vertical dependence of $\Omega$ and $\kappa$ except
-where their vertical derivatives appear explicitly in the above
-equations. We thus make the replacement $D\to \Omega_k^2$, as before
-for isothermal perturbations. Here, we also use the fact that in the
+We make the low-frequency approximation and the replacement
+$D\to \Omega_k^2$ in Eq. \ref{adia_iso1}---\ref{adia_iso2}. For the
+analysis to be carried out, the latter is a more limiting
+approximation than for isothermal perturbations because we make the
+replacement before combining the pair of equations above.
+
+Here, we also use the fact that in the
 global disk, 
@@
 
-Combining Eq. \ref{adia_iso1} and Eq. \ref{adia_iso2} with the above
-substitutions we obtain, in terms of dimensionless variables defined
+Making the above substitutions and combining Eq. \ref{adia_iso1}---
+Eq. \ref{adia_iso2} we obtain, in terms of dimensionless variables defined
 previously, an equation for $\delta v_z$, 
@@
 \end{align}
+
 We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and

### 2014-12-18 appendix.tex
added appendix b, edited first numerical stuff

@@
 \end{align}
+
+\section{Improving the nearly-Keplerian approximation for adiabatic
+  disks}\label{adia_improve}
+In \S\ref{analytic_adia} we made the replacement $D\to\Omega_k^2$
+before combining Eq. \ref{adia_iso1}---\ref{adia_iso2} to derive
+Eq. \ref{adia_iso3}. This procedure ignores the vertical dependence of
+$D$. We show here that this has no significant consequence for thin
+disks.
+
+Differentiating Eq. \ref{adia_iso1} properly gives
+\begin{align}
+  \left(D + \gamma k_x^2 c_s^2\right)\frac{d\Delta}{dz}=& D
+  \frac{d^2\delta v_z}{dz^2} + k_x^2c_s^2\left[\frac{\gamma}{\left(D + \gamma k_x^2 c_s^2\right)}\frac{dD}{dz} + \ii
+    \frac{d\ln\rho}{dz}\left(\ii +
+      \frac{q}{k_xr}\right)\right]\frac{d\delta v_z}{dz}\notag\\
+  &+ \ii k_x^2 c_s^2 \left(\ii +
+      \frac{q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} -
+      \frac{1}{\left(D + \gamma
+          k_x^2c_s^2\right)}\frac{dD}{dz}\frac{d\ln\rho}{dz}\right]\delta
+    v_z. \label{adia_diso1}
+\end{align}
+The terms that were ignored in deriving Eq. \ref{adia_iso3} are those
+proportional to $dD/dz$ in Eq. \ref{adia_diso1}. We estimate the
+importance of these terms by noting that 
+\begin{align}
+  \frac{dD}{dz}\equiv \frac{d\kappa^2}{dz} \simeq \frac{d\Omega^2}{dz}
+  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2},
+\end{align}
+for thin disks. Then Eq. \ref{adia_diso1} becomes
+\begin{align}
+\left(D + \gamma k_x^2 c_s^2\right)\frac{d\Delta}{dz}=& D
+  \frac{d^2\delta v_z}{dz^2} - k_x^2c_s^2\frac{d\ln\rho}{dz}\left[ 1 - 
+      \frac{\ii q}{k_xr}  +  \frac{\gamma q c_s^2}{r^2\left(D +
+        \gamma k_x^2 c_s^2\right)}\right]\frac{d\delta
+    v_z}{dz}\notag\\ 
+  &- k_x^2 c_s^2 \left(1 - 
+      \frac{\ii q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} + 
+      \frac{qc_s^2}{r^2\left(D + \gamma
+          k_x^2c_s^2\right)}\left(\frac{d\ln\rho}{dz}\right)^2\right]\delta
+    v_z. \label{adia_diso2}
+\end{align}
+Since $D\sim \Omega_k^2$ in the low-frequency limit, we 
+see from Eq. \ref{adia_diso2} that the neglected terms are
+$O(\epsilon^2)$. We can make this explicit by combining
+Eq. \ref{adia_diso1}---\ref{adia_diso2} and Eq. \ref{adia_iso2}, then
+set $D\to\Omega_k^2$. In terms of non-dimensional variables, the result
+is  
+\begin{align}
+   &\delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon q\hat{k} -
+    \frac{\gamma q \epsilon^2\hat{k}^2}{1+\gamma
+      \hat{k}^2}\right)\ln\rho^{\prime}\delta v_z^\prime +
+  \left[\left(\frac{1}{\gamma} + \ii \epsilon q
+      \hat{k}\right)\ln\rho^{\prime\prime} - \hat{k}^2\left(1 -
+      \frac{\ii\epsilon
+        q}{\hat{k}}\right)\left(\frac{\gamma-1}{\gamma} +
+      \frac{q\epsilon^2}{1+\gamma\hat{k}^2}\right)\ln\rho^{\prime
+      2}\right]\delta v_z \notag\\&=
+  -\hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right)\delta v_z\label{adia_diso3}.
+\end{align}   
+Eq. \ref{adia_diso3} differs from Eq. \ref{adia_iso3} by terms
+proportional to $\epsilon^2$. For a thin disk, $\epsilon\ll1$, so
+neglecting these terms has no qualitative effect on the discussion in
+\S\ref{analytic_adia}.

### 2014-12-18 linear_adia.tex
added appendix b, edited first numerical stuff

@@
-\subsection{Adiabatic perturbations} 
+\subsection{Adiabatic perturbations}\label{analytic_adia}
 We now extend the above discussion to general $\gamma$, but in the
@@
 We make the low-frequency approximation and the replacement 
-$D\to \Omega_k^2$ in Eq. \ref{adia_iso1}---\ref{adia_iso2}. For the
-analysis to be carried out, the latter is a more limiting
-approximation than for isothermal perturbations because we make the
-replacement before combining the pair of equations above.
-
-Here, we also use the fact that in the
+$D\to \Omega_k^2$ in Eq. \ref{adia_iso1}---\ref{adia_iso2}. For
+simplicity we do this before combining these equations later. We  
+show in Appendix \ref{adia_improve} that making these approximations
+after combining  Eq. \ref{adia_iso1}---\ref{adia_iso2} adds unecessary
+complexity for thin disks. Here, we also use the fact that in the
 global disk, 
@@
 
-Making the above substitutions and combining Eq. \ref{adia_iso1}---
+Making the above substitutions and eliminating $\Delta$ between Eq. \ref{adia_iso1}---
 Eq. \ref{adia_iso2} we obtain, in terms of dimensionless variables defined
@@
 
-Note that the term with $\ln\rho^{\prime 2}$ in the integrand (third
-term on the RHS) represents stabilization by the vertical entropy
-gradient when there is no vertical shear. It gives an imaginary
-contribution to $\hat{\sigma}$ when $q\neq0$, but this is a factor
-$|\epsilon/\hat{k}|$ smaller than the real (stabilizing) contribution,
+Note that when $\gamma>1$ and $q=0$, the term with $\ln\rho^{\prime
+  2}$ in the integrand (third term on the RHS) represents
+stabilization by the vertical entropy gradient (real bouyancy frequency). 
+It gives an imaginary contribution to $\hat{\sigma}$ when $q\neq0$,
+but this is a factor $|\epsilon/\hat{k}|$ smaller than the real
+(stabilizing) contribution, 
 since $\epsilon\ll 1$ for a thin disk and we (usually) consider
-$\hat{k}\gg1$. Thus, instability is expected to associated with the 
+$\hat{k}\gg1$. Thus, instability is expected to be associated with the 
 last term on the RHS of Eq. \ref{adia_integral}.
@@
 We choose $\alpha$ to make the coefficient of $\zhat^2g$
-vanish. Furthermore, we require 
-the vertical kinetic energy density $\rho|\delta v_z|^2$ to remain
-finite. Then assuming $g(\zhat)$ is a polynomial, we require 
+vanish, and impose the vertical kinetic energy density
+$\rho|\delta v_z|^2$ to remain finite. Then assuming $g(\zhat)$ is a
+polynomial, we require  
 \begin{align}
@@
 \end{align}
+This gives the eigenfrequency $\hat{sigma}$ and completes the solution
+description.
+
 %Recalling that $B$ depends $\hat{\sigma}$, Eq. \ref{adia_disp} gives
@@
 In principle, we can solve Eq. \ref{adia_disp} for the mode
-frequencies and growth rates. This is complicated in 
+frequencies and growth rates as a function of the disk parameters and
+the radial wavenumber. This is complicated in 
 general, but simplifies for $\gamma=2$, for which the quantity
@@
 \end{align}
-Unlike for isothermal perturbations, Eq. \ref{gam2_max_growth} show
-that growth rates decrease with $M$, and there exists a maximum
-possible growth rate at $M=0$. This is despite the fact that in the
-thin-disk approximation, for which $d\Omega^2/dz\propto z$, there is
-infinite vertical shear available.
+Unlike isothermal perturbations, Eq. \ref{gam2_max_growth} shows that
+there is a maximum possible growth rate for $M=0$, and growth rates decrease with increasing $M$. This
+is despite the fact that in the thin-disk approximation 
+$d\Omega^2/dz\propto z$, so there is infinite vertical shear
+available.
 
-The acceptable solution for $\alpha$ is
+The acceptable solution for $\alpha$ when $\gamma=2$ is
 \begin{align}
@@
 Since $\epsilon\ll 1$, $\real\alpha<0$, so the vertical velocity
-eventually decays for large heights. This is unlike for isothermal
-perturbations, in which perturbation magnitudes increase with $|z|$.
+eventually decays for large heights. For $\hat{k}\gg 1$ the
+perturbation decays rapidly away from the midplane. This is again
+unlike isothermal perturbations, whose magnitude increase with $|z|$.
+
+%gauss decay means boundary conditions have no influence, all perts go
+%to zero
 
@@
 
-In Fig. \ref{adia_growth} we plot the growth rate $|\hat{\nu}|$ as a
+In Fig. \ref{adia_growth} we plot the eigenfrequencies as a
 function of the radial wavenumber $\hat{k}$, for a range of adiabatic
-indices, using Eq. \ref{adia_disp}. The presence of a positive
-vertical entropy gradient is strongly stabilizing. This is because the
-vertical shear grows as $z$ away from the midplane (and has a maximum
-when the thin-disk approximation is relaxed), but the
-square of the bouyancy frequency, which is stabilizing, grows as $z^2$
-away from the midplane. The entropy $P/\rho^\gamma$ formally diverges
-for $z\to\infty$ even without the thin-disk approximation.
+indices, using Eq. \ref{adia_disp}. The low-frequency approximation
+is questionable for $\hat{k}\lesssim4$ since
+$|\hat{\sigma}|^2\gtrsim0.2$. For large $\hat{k}$ the presence of a
+positive vertical entropy gradient is strongly stabilizing. This is
+because the vertical shear grows as $z$ away from the midplane (and
+has a maximum when the thin-disk approximation is relaxed), but square
+of the bouyancy frequency, which is stabilizing, grows as $z^2$ away
+from the midplane. %The entropy $P/\rho^\gamma$ formally diverges   
+%for $z\to\infty$ even without the thin-disk approximation.
 

### 2014-12-18 numerical.tex
minor updates
added appendix b, edited first numerical stuff

@@
-\section{Numerical solutions}
-In this section we numerically solve the linear stability problem
-corresponding to Eq. \ref{iso_ode1}. We thus relax the thin-disk
-assumption, using the full expressions for $\rho$, $\Omega$ and
-$\kappa$, so that we may consider vertical domains up to $z\sim r$.
-However, we retain the low-frequency approximation
-$|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue problem. This
-level of approximation lies in between the full problem 
-Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and the thin-disk
-limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.
+\section{Numerical solutions in the low frequency limit}
+In this section we numerically solve the linear stability problem.  
+We relax the thin-disk assumption, using the full, non-Keplerian
+expressions of $\rho$, $\Omega$ and $\kappa$, so that we may consider
+vertical domains up to $z\sim r$. However, we retain the low-frequency
+approximation $|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue
+problem, which limits us to consider large radial wavenumbers
+$k_x$. We will relax this in the next section when considering
+moderate values of $k_x$. % This level of approximation lies in between
+% the full problem Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and
+% the thin-disk limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.
 
-We solve Eq. \ref{iso_ode1} using a pseudo-spectral method by
-expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$ and
-discretizing the equation on a grid with
+%We will solve different versions of the linearized equations depending
+%on the disk properties, but in all cases we use
+
+\subsection{Vertically isothermal disks with isothermal perturbations}   
+We first validate the low-frequency approximation by re-producing
+results from \cite{mcnally14} who solved the full problem for
+vertically isothermal disks ($\Gamma=1$) subject to isothermal
+perturbations ($\gamma=1$). The parameters are $q=-1$,
+$p=-1.5$, $\epsilon=0.1$ and $k_x = 200\pi/r$ (or $\hat{k} = 20\pi$),
+with solid vertical boundaries.
+
+For this problem we solve Eq. \ref{iso_ode1}, using a pseudo-spectral
+method by expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$
+and discretizing the equation on a grid with
 $\zmax=10H_\mathrm{iso}$. This procedure converts the linear problem
-to a standard matrix eigenvalue problem. %In the low-frequency
-%approximation the eigenvalue is $\sigma^2$.
-
-\subsection{Fiducial calculation}
-We first validate the low-frequency approximation by solving a 
-fiducial problem with $q=-1$, $p=-1.5$, $\epsilon=0.1$ and $k_x =
-200\pi/r$ with solid boundaries. This is the setup considered by
-\cite{mcnally14} who solved the full problem.
+to a standard matrix eigenvalue problem.
 
@@
 \ii\nu$. This plot is effectively identical to that in
-\cite{mcnally14}. The low-frequency approximation reproduces the
-three solution branches defined by \citep{nelson13}: corrugation modes
-with odd symmetry, breathing modes with even symmetry, and surface
-modes which correspond to the nearly-vertical locus in
-Fig. \ref{lowfreq_eigen}.
+\cite{mcnally14}. The fundamental VSI mode, with one node in $W$, has
+the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}.  
+The expected growth rate from Eq. \ref{simple_growth} with 
+$L=1$ is  
+\begin{align*}
+  \nu = 0.2606\epsilon\Omega_k,
+\end{align*}
+as obtained numerically. This agreement is surpsingly good, given that
+Eq. \ref{simple_growth} assumes the thin-disk approximation and
+imposes a different boundary condition to that in the numerical
+calculation. The eigenfunction $W$ is qualitatively proportional to
+$z$, as expected for the fundamental mode according to explicit
+solutions discussed in \S\ref{iso_poly}.
 
@@
 
-We plot in Fig. \ref{lowfreq_eigenfunc} the structurally simplest
-eigenfunction --- the fundamental corrugation mode corresponding to
-the eigenvalue with $\mathrm{min}|\sigma|$. The perturbation $W$ is
-roughly proportional to $z$, except close the boundaries where it
-flattens because $\delta v_z=dW/dz=0$ is imposed there.
-
-For $k_x = 200\pi/r$ and $\epsilon = 0.1$, the dimensionless
-wavenumber $\hat{k} = k_xH_\mathrm{iso}=20\pi$. Then the expected
-growth rate according to Eq. \ref{simple_growth} with $M=1$ is
-\begin{align*}
-  \nu = 0.2606\epsilon\Omega_k,
-\end{align*}
-as obtained numerically (Fig. \ref{lowfreq_eigen}). This agreement is
-surpsingly good, given that Eq. \ref{simple_growth} assumes the
-thin-disk approximation and imposes a different boundary condition to
-that in the numerical calculations.
-
 \begin{figure}
   \includegraphics[width=\linewidth]{figures/eigenvector_iso}
-  \caption{Eigenfunction of the fundamental corrugation mode,
-    corresponding to the 
-    bottom-left eigenvalue displayed in Fig. \ref{lowfreq_eigen}. The
+  \caption{Eigenfunction of the fundamental VSI,
+    corresponding to the bottom-left eigenvalue displayed in
+    Fig. \ref{lowfreq_eigen} (smallest $|\sigma|$). The
     solid and dashed lines are $\real W$ and $\imag W$, respectively. 
@@
 \end{figure}
+
+\subsection{Stabilization of the VSI by a positive vertical entropy
+  gradient}

### 2014-12-18 appendix.tex
improved appendix b

@@
 The terms that were ignored in deriving Eq. \ref{adia_iso3} are those
-proportional to $dD/dz$ in Eq. \ref{adia_diso1}. We estimate the
-importance of these terms by noting that 
+proportional to $dD/dz$ in Eq. \ref{adia_diso1}. Now, in the global
+disk we have
+\begin{align}\label{dkappa2}
+  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
+    r}\frac{\p\Omega^2}{\p z} = -
+  \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(2 + q +
+    \frac{z^2/r^2-2}{z^2/r^2+1}\right). 
+\end{align}
+For the local problem we may then write
 \begin{align}
-  \frac{dD}{dz}\equiv \frac{d\kappa^2}{dz} \simeq \frac{d\Omega^2}{dz}
-  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2},
+  \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z;q),
 \end{align}
-for thin disks. Then Eq. \ref{adia_diso1} becomes
+where the function $F$ corresponds to the bracket in
+Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
+3+q$ as $|z|\to\infty$, i.e. its magnitude is of order unity. 
+% We estimate the
+% importance of these terms by noting that 
+% \begin{align}
+%   \frac{dD}{dz}\equiv \frac{d\kappa^2}{dz} \simeq \frac{d\Omega^2}{dz}
+%   = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2},
+% \end{align}
+% for thin disks. 
+Eq. \ref{adia_diso1} becomes
 \begin{align}
@@
   \frac{d^2\delta v_z}{dz^2} - k_x^2c_s^2\frac{d\ln\rho}{dz}\left[ 1 - 
-      \frac{\ii q}{k_xr}  +  \frac{\gamma q c_s^2}{r^2\left(D +
+      \frac{\ii q}{k_xr}  +  \frac{\gamma q c_s^2F}{r^2\left(D +
         \gamma k_x^2 c_s^2\right)}\right]\frac{d\delta
@@
       \frac{\ii q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} + 
-      \frac{qc_s^2}{r^2\left(D + \gamma
+      \frac{qc_s^2F}{r^2\left(D + \gamma
           k_x^2c_s^2\right)}\left(\frac{d\ln\rho}{dz}\right)^2\right]\delta
@@
    &\delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon q\hat{k} -
-    \frac{\gamma q \epsilon^2\hat{k}^2}{1+\gamma
+    \frac{\gamma q \epsilon^2\hat{k}^2F}{1+\gamma
       \hat{k}^2}\right)\ln\rho^{\prime}\delta v_z^\prime +
@@
         q}{\hat{k}}\right)\left(\frac{\gamma-1}{\gamma} +
-      \frac{q\epsilon^2}{1+\gamma\hat{k}^2}\right)\ln\rho^{\prime
+      \frac{q\epsilon^2F}{1+\gamma\hat{k}^2}\right)\ln\rho^{\prime
       2}\right]\delta v_z \notag\\&=

### 2014-12-18 linear_iso.tex
added appendix b, edited first numerical stuff
improved appendix b

@@
   \kappa^2$. The problem simplifies because the eigenvalue now only
-  appears in the last term in Eq. \ref{iso_ode}. 
+  appears in the last term in Eq. \ref{iso_ode}. We call this the
+  low-frequency approximation. 
 \item We are mainly interested in the effect of vertical shear, i.e. $q\neq
@@
   the vertical dependence of $\kappa^2$ elsewhere. For clarity, we
-  will replace $\kappa^2$ with $\Omega_k^2$.%, although the discussion
+  will replace $\kappa^2$ with $\Omega_k^2$. 
+  % , although the discussion
   % is unchanged if we replaced 
@@
 \end{enumerate}
-These assumptions amount to making the replacement $D\to\Omega_k^2$.  
-We will consider sound-speed profiles with $q<0$ in the global 
+Making both of these approximations is equivalent to the replacement
+$D\to\Omega_k^2$.  We will consider sound-speed profiles with $q<0$ in the global 
 disk. Then $\Omega^2(z)$ decreases away from the midplane,
@@
 
-\subsubsection{Polynomial solutions with vertical shear}
+\subsubsection{Polynomial solutions with vertical shear}\label{iso_poly}
 We can seek power-series solution to Eq. \ref{iso_ode3},

### 2014-12-18 numerical.tex
improved appendix b

@@
 
-For this problem we solve Eq. \ref{iso_ode1}, using a pseudo-spectral
+For this problem we solve Eq. \ref{iso_ode}, with $D$ replaced by
+$\kappa^2$, using a pseudo-spectral
 method by expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$

### 2014-12-19 linear_adia.tex
regular updates

@@
 \end{align}
-This gives the eigenfrequency $\hat{sigma}$ and completes the solution
+This gives the eigenfrequency $\hat{\sigma}$ and completes the solution
 description.
@@
   \alpha = \frac{1}{2}\left[\left(1+\ii\epsilon q \hat{k}\right) -
-    \sqrt{1 + \hat{k}^2\left(2-\epsilon^2 q^2\right)}\right]. 
+    \sqrt{1 + \hat{k}^2\left(2-\epsilon^2 q^2\right)}\right].\label{gam2_alpha}
 \end{align}
@@
   &\hat{\nu} \leq \frac{|\epsilon q|}{2^{7/4}\sqrt{\hat{k}}}
-  &\text{as }  \hat{k}\to\infty  \text{ for }  \gamma=2.
+  &\text{as }  \hat{k}\to\infty  \text{ for }  \gamma=2.\label{gam2_growth_rate}
 \end{align}
@@
 \end{figure}
+
+\subsubsection{Qualitative expectations in the presence of thermal
+  relaxation}

### 2014-12-19 numerical.tex
regular updates

--- 2014-12-18 numerical.tex
@@
 
-\subsection{Vertically isothermal disks with isothermal perturbations}   
+\subsection{Vertically isothermal disks with isothermal
+  perturbations}\label{vertiso_pertiso} 
 We first validate the low-frequency approximation by re-producing
@@
 
-For this problem we solve Eq. \ref{iso_ode1}, using a pseudo-spectral
+For this problem we solve Eq. \ref{iso_ode}, with $D$ replaced by
+$\kappa^2$, using a pseudo-spectral
 method by expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$
 and discretizing the equation on a grid with
-$\zmax=10H_\mathrm{iso}$. This procedure converts the linear problem
+$\zmax=10\epsilon r$. This procedure converts the linear problem
 to a standard matrix eigenvalue problem.
@@
   gradient}
+We demonstrate the (strong) stabilizing effect of a  
+positive vertical entropy gradient. We solve the more general form of
+the linear problem, Eq. \ref{ode_w}---\ref{ode_Q}, again with the
+replacement $D\to\kappa^2$. We consider an effectively vertically
+isothermal disk by setting $\Gamma=1.011$ with adiabatic index
+$\gamma=2$. We also set $\beta=10^9$ so the cooling term is
+effectively switched off. Other disk and perturbation parameters are
+the same as in \S\ref{vertiso_pertiso}.
+
+We use the same pseudo-spectral method as before. Here we set
+$\zmax = 2\epsilon r$ and impose a free surface. This boundary
+condition was found to be most convenient for numerical
+implementation. Since we expect the perturbations to decay rapidly
+away from the midplane (see \S\ref{analytic_adia}), we do not expect 
+boundary conditions to play an important role, provided that $\zmax$ is
+sufficiently large.
+
+Fig. \ref{lowfreq_eigen_adia} show the eigenvalues for this
+problem. The growth rates are about an order of magnitude smaller than
+the previous case. The fundamental VSI mode is that with
+$\mathrm{max}|\nu|$. Its expected and numerically-calculated growth
+rates compares well, with 
+\begin{align*}
+  \nu = 0.03751 \epsilon \Omega_k &\quad \text{(from
+    Eq. \ref{gam2_growth_rate})},\\
+  \nu = 0.03682 \epsilon \Omega_k &\quad \text{(numerical)}.
+\end{align*}
+The fundamental mode is plotted in Fig. \ref{lowfreq_eigenfunc_adia},
+and clearly show that perturbations are rapidly stabilized away from
+the midplane. For the adopted disk and perturbation parameters 
+($|q|=1$ and $\hat{k} = 20\pi$), Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$,
+corresponding to a characteristic decay lengthscale of $\simeq
+0.15\epsilon r$, which is consistent with that observed in 
+Fig. \ref{lowfreq_eigenfunc_adia}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/eigenvalues_adia}
+  \caption{Eigenvalues in the low-frequency approximation for the
+    vertical shear instability in a nearly-vertically isothermal disk
+    ($\Gamma=1.011$) evolved adiabatically with $\gamma=2$ and
+    $\beta=10^9$. The disk
+    parameters are $q=-1$, 
+    $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
+    wavenumber is $k_x=200\pi/r$. \label{lowfreq_eigen_adia}
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/eigenvectorvz_adia}
+  \caption{Fundamental VSI mode in a nearly-vertically
+    isothermal disk ($\Gamma=1.011$) evolved adiabatically
+    ($\gamma=2$, $\beta=10^9$). This eigenfunction 
+    corresponds to the top-left eigenvalue displayed in 
+    Fig. \ref{lowfreq_eigen_adia} (largest $|\nu|$).  
+    The solid and dashed lines are $\real \delta v_z$ and $\imag
+    \delta v_z$, respectively.  
+    \label{lowfreq_eigenfunc_adia}
+  }
+\end{figure}
+
+\subsection{Effect of cooling}

### 2014-12-19 setup.tex
regular updates

@@
   c_s^2\equiv \frac{dP}{d\rho} =
-  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1}. 
+  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} = \frac{\Gamma P}{\rho}. 
 \end{align}
+For discussion purposes it is useful to define the vertical buoyancy frequency $N_z$ via
+\begin{align}
+  N_z^2 = c_s^2 \left(\frac{\gamma -
+      \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   
+\end{align}
+and we are interested in neutrally or sub-adiabatically stratified
+disks with $\Gamma\leq \gamma$ so that $N_z^2\geq0$.
 
@@
 the ratio of specific heats $\gamma$ is assumed constant, 
-and we have included a thermal relaxation term with timescale $t_c$.
+and we have included a thermal relaxation term with timescale
+$t_c=\beta\Omega_k^{-1}$ with $\beta$  being a dimensionless input
+parameter.
 

### 2014-12-19 linear_adia.tex
added beta cooling requirements

@@
   relaxation}
+We have shown that when $\gamma>1$ so that there is a positive
+vertical entropy gradient and perturbations evolve
+adiabatically ($t_c\to\infty$) that the VSI is strongly
+surpressed. This is the limit where the term proportional to
+$t_c^{-1}$ in the linearized energy equation (Eq. \ref{lin_energy})
+becomes negligible compared to the term proportional to
+$\gamma/\Gamma-1$. The latter term is associated with the vertical
+entropy gradient, or buoyancy frequency $N_z$.
+
+Reversing the argument, we expect to recover the VSI when the bouyancy
+frequency becomes negligible compared to the cooling rate, or 
+\begin{align}
+  t_c \ll \frac{1}{N_z}. 
+\end{align} 
+Parametrizing $t_c = \beta\Omega_k^{-1}$ and inserting the expression
+for $N_z$ appropriate for vertically isothermal disks, we expect
+\begin{align}
+  \beta \ll
+  \frac{1}{H_\mathrm{iso}\sqrt{\frac{\gamma-1}{\gamma}}\left|\frac{\p\ln\rho}{\p
+          z}\right|} 
+\end{align}
+is required to enable the VSI. This inequality is satisfied at the
+midplane and as $|z|\to\infty$. It can be satisifed at all heights if
+it is satisfied at the height of $\mathrm{max}|\p\ln\rho/\p z|$. 
+Using the expression for the density field for vertically isothermal
+disks (Eq. \ref{vertiso_eqm}), we expect 
+\begin{align}
+  \beta \ll \epsilon\frac{3^{3/2}}{2}\sqrt{\frac{\gamma}{\gamma-1}},
+\end{align}
+i.e. $\beta\ll O(\epsilon)$, is a sufficient condition for
+bouyancy to be negligible.

### 2014-12-20 numerical.tex
replace numerical section with that of full equations

@@
-\section{Numerical solutions in the low frequency limit}
-In this section we numerically solve the linear stability problem.  
-We relax the thin-disk assumption, using the full, non-Keplerian
-expressions of $\rho$, $\Omega$ and $\kappa$, so that we may consider
-vertical domains up to $z\sim r$. However, we retain the low-frequency
-approximation $|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue
-problem, which limits us to consider large radial wavenumbers
-$k_x$. We will relax this in the next section when considering
-moderate values of $k_x$. % This level of approximation lies in between
-% the full problem Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and
-% the thin-disk limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.
+\section{Numerical solutions}
+In this section we numerically solve the linear stability problem
+corresponding to Eq. \ref{lin_mass}---\ref{lin_energy}. 
+We relax the thin-disk, low-frequency and nearly-Keplerian
+approximations used in the discussion above, and 
+use the full, non-Keplerian expressions of $\rho$, $\Omega$ and
+$\kappa$, so that we may consider vertical domains up to $z\sim r$.
 
-%We will solve different versions of the linearized equations depending
-%on the disk properties, but in all cases we use
+% \subsection{Boundary conditions}
+% In numerical calculations a finite vertical domain $z\in[-1,1]\zmax$
+% is adopted. For $\Gamma\neq1$ we take $\zmax$ to be slightly less than
+% $H$ to avoid the zero-density surface. At $z=\pm\zmax$ we impose
+% \begin{align}
+%   \delta v_z = 0,
+% \end{align}
+% for a solid boundary, or 
+% \begin{align}
+%   \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0,  
+% \end{align}
+% for a free surface, where $\bm{\xi}$ is the Lagrangian
+% displacement.
 
-\subsection{Vertically isothermal disks with isothermal
-  perturbations}\label{vertiso_pertiso} 
-We first validate the low-frequency approximation by re-producing
-results from \cite{mcnally14} who solved the full problem for
-vertically isothermal disks ($\Gamma=1$) subject to isothermal
-perturbations ($\gamma=1$). The parameters are $q=-1$,
-$p=-1.5$, $\epsilon=0.1$ and $k_x = 200\pi/r$ (or $\hat{k} = 20\pi$),
-with solid vertical boundaries.
+We solve the linear problem by expanding the
+perturbations in Chebyshev polynomials $T_l$ up to $l=512$
+and discretizing the equations on a grid with
+$z\in[-\zmax,\zmax]$. At the vertical boundaries we impose a free
+surface, 
+\begin{align}
+  \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0 \quad \text{at } z=\pm\zmax,
+\end{align}
+where $\bm{\xi}$ is the Lagrangian displacement. This boundary
+condition was found to be most convenient for numerical
+implementation. The above discretization procedure converts the linear  
+system of differential equations to a standard matrix eigenvalue
+problem.
 
-For this problem we solve Eq. \ref{iso_ode}, with $D$ replaced by
-$\kappa^2$, using a pseudo-spectral
-method by expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$
-and discretizing the equation on a grid with
-$\zmax=10\epsilon r$. This procedure converts the linear problem
-to a standard matrix eigenvalue problem.
+Unless otherwise stated, our fiducial disk model is a
+nearly-vertically isothermal disk with $\Gamma=1.011$, $p=-1.5$,
+$q=-1$, $\epsilon=0.1$, and the perturbation wavenumber $k_x = 200\pi$
+(or $\hat{k}=20\pi$).
+
+\subsection{VSI in a neutrally stratified disk}\label{vertiso_pertiso} 
+We first test our numerical code, as well as the discussion in
+\S\ref{iso_discuss}, by considering nearly-isothermal
+perturbations. We set $\gamma=1.011=\Gamma$, and effectively switch
+off thermal relaxation by choosing $\beta=10^9$. The vertical domain is
+$\zmax = 10\epsilon r$. This setup is very similar to that adopted in
+\cite{mcnally14}, who calculated the VSI in vertically isothermal
+disks subject to isothermal perturbations ($\Gamma=\gamma=1$).
 
@@
 the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}.  
-The expected growth rate from Eq. \ref{simple_growth} with 
-$L=1$ is  
+Its expected growth rate, calculated from Eq. \ref{simple_growth} with  
+$L=1$, and the numerically-obtained value are 
 \begin{align*}
-  \nu = 0.2606\epsilon\Omega_k,
+  \nu = 0.2606\epsilon\Omega_k &\quad \text{(from
+    Eq. \ref{simple_growth})},\\
+  \nu = 0.2577 \epsilon \Omega_k &\quad \text{(numerical)}.
 \end{align*}
-as obtained numerically. This agreement is surpsingly good, given that
-Eq. \ref{simple_growth} assumes the thin-disk approximation and
+This agreement is surpsingly good, given that
+Eq. \ref{simple_growth} assumes a thin, nearly-Keplerian disk in the 
+low-frequency approximation, and
 imposes a different boundary condition to that in the numerical
-calculation. The eigenfunction $W$ is qualitatively proportional to
+calculation. The eigenfrequency is not very sensitive to the vertical boundary
+condition because the expression for $\sigma$, given by
+Eq. \ref{integral_relation1}, involve integrands with $\rho$ as the 
+weight function, which rapidly decays away from the midplane. The
+eigenfunction $W$ is qualitatively proportional to 
 $z$, as expected for the fundamental mode according to explicit
@@
   \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
-  \caption{Eigenvalues in the low-frequency approximation for the
-    vertical shear instability in a vertically isothermal disk evolved
-    isothermally ($\gamma=\Gamma=1$). The disk parameters are $q=-1$,
+  \caption{Eigenvalues of the
+    vertical shear instability in a neutrally-stratified,
+    nearly-vertically isothermal disk  ($\gamma=\Gamma=1.011$). The
+    disk parameters are $q=-1$,  
     $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
@@
 
+We have also solved this problem with in the low-frequency 
+approximation by setting $D\to\kappa^2$ in Eq. \ref{iso_ode}. We
+obtained very similar results, as expected for modes with
+$|\sigma/\Omega_k|\ll1$.
+
 \subsection{Stabilization of the VSI by a positive vertical entropy
@@
 We demonstrate the (strong) stabilizing effect of a  
-positive vertical entropy gradient. We solve the more general form of
-the linear problem, Eq. \ref{ode_w}---\ref{ode_Q}, again with the
-replacement $D\to\kappa^2$. We consider an effectively vertically
-isothermal disk by setting $\Gamma=1.011$ with adiabatic index
-$\gamma=2$. We also set $\beta=10^9$ so the cooling term is
-effectively switched off. Other disk and perturbation parameters are
-the same as in \S\ref{vertiso_pertiso}.
+positive vertical entropy gradient by setting $\gamma=2$.  
+Since we expect the perturbations to decay rapidly
+away from the midplane (see \S\ref{analytic_adia}), we consider a
+smaller domain with $\zmax = 2\epsilon r$. 
+Other disk and perturbation parameters are the same as in
+\S\ref{vertiso_pertiso}.
 
-We use the same pseudo-spectral method as before. Here we set
-$\zmax = 2\epsilon r$ and impose a free surface. This boundary
-condition was found to be most convenient for numerical
-implementation. Since we expect the perturbations to decay rapidly
-away from the midplane (see \S\ref{analytic_adia}), we do not expect 
-boundary conditions to play an important role, provided that $\zmax$ is
-sufficiently large.
+% We use the same pseudo-spectral method as before. Here we set
+% $\zmax = 2\epsilon r$ and impose a free surface. This boundary
+% condition was found to be most convenient for numerical
+% implementation. Since we expect the perturbations to decay rapidly
+% away from the midplane (see \S\ref{analytic_adia}), we do not expect 
+% boundary conditions to play an important role, provided that $\zmax$ is
+% sufficiently large.
 
@@
     Eq. \ref{gam2_growth_rate})},\\
-  \nu = 0.03682 \epsilon \Omega_k &\quad \text{(numerical)}.
+  \nu = 0.03672 \epsilon \Omega_k &\quad \text{(numerical)}.
 \end{align*}
@@
   \includegraphics[width=\linewidth]{figures/eigenvalues_adia}
-  \caption{Eigenvalues in the low-frequency approximation for the
+  \caption{Eigenvalues in for the
     vertical shear instability in a nearly-vertically isothermal disk
@@
 
-\subsection{Effect of cooling}
+We have also solved this problem in the low-frequency approximation
+using Eq. \ref{ode_w}---\ref{ode_Q} with the replacement
+$D\to\kappa^2$. We again obtained very similar results.
+
+\subsection{Effect of thermal relaxation}
+{\bf vertical isothermal disk with tcool}
+
+\subsection{Vertically non-isothermal disks}
+{\bf big gamma > 1, with tcool}

### 2014-12-20 codes/cool_full/vsi.f90
added code backup

@@ -1,0 +1,431 @@
+module global
+  character*6 :: vbc
+  logical :: gterms
+  integer :: nz, nzeff, bignz
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls  
+  real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:)
+  complex*16, parameter :: ii = (0d0,1d0)
+  complex*16, allocatable :: bigW(:), dbigW(:), bigQ(:), vx(:), vy(:), vz(:), dvz(:) 
+end module global
+
+program vsi
+  use global
+  implicit none
+  character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
+  integer, allocatable :: ipiv(:), eigen_test(:)
+  integer :: i,ip1, j,k, loc(1), lmax, nzmid, lwork, info
+  real*8 :: T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2, m, zbar,lmode
+  real*8, allocatable :: rwork(:)
+  complex*16, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)
+  complex*16, allocatable :: L1(:,:), L2(:,:), L3(:,:), L4(:,:), &
+       L5(:,:), L6(:,:), L7(:,:), L8(:,:), &
+       L9(:,:), L10(:,:), L11(:,:), &
+       L1bar(:,:), L2bar(:,:), L3bar(:,:), L4bar(:,:), L5bar(:,:), &
+       L6bar(:,:)
+  complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
+       vr(:,:), rhs(:,:)
+  real*8, external :: dlogrho_dz, domega2_dz, omega2_z, logrho_z, kappa2_z, d2logrho_dz2, csq_z 
+  namelist /params/ smallq, smallp, eps, gmma, bgmma, bcool, kx, zmax, vbc, gterms
+  namelist /grid/ nz
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=grid)
+  close(7)
[…]

### 2014-12-20 codes/cool_iso/vsi.f90

@@ -1,0 +1,381 @@
+module global
+  character*6 :: vbc
+  integer :: nz, nzeff, bignz
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, smallh, kx, zmax, gmma, bcool  
+  real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), freq(:), growth(:)
+  complex*16, parameter :: ii = (0d0,1d0)
+  complex*16, allocatable :: bigW(:), bigQ(:)  
+end module global
+
+program vsi
+  use global
+  implicit none
+  character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
+  integer, allocatable :: ipiv(:), eigen_test(:)
+  integer :: i,ip1, j,k, loc(1), lmax, nzmid, lwork, info
+  real*8 :: T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2, m, zbar,lmode
+  real*8, allocatable :: rwork(:)
+  complex*16, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)
+  complex*16, allocatable :: L1(:,:), L2(:,:), L3(:,:), L4(:,:), &
+       L1bar(:,:), L2bar(:,:), L3bar(:,:), L4bar(:,:), L5bar(:,:), &
+       L6bar(:,:)
+  complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
+       vr(:,:), rhs(:,:)
+  real*8, external :: dlogrho_dz, domega2_dz, shear_z, omega2_z, logrho_z, kappa2_z, d2logrho_dz2
+  namelist /params/ smallq, smallp, smallh, gmma, bcool, kx, zmax, vbc
+  namelist /grid/ nz
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=grid)
+  close(7)
+
+  if(mod(nz,2).eq.0) then
+     print*, 'Nz needs to be odd but Nz=', nz
+     stop
+  else
+     Nzmid = (Nz+1)/2
+  endif
+
+  if(vbc.ne.'nolagp') then
+     nzeff = nz-2
+  else
+     nzeff = nz
+  endif
+
[…]

### 2014-12-20 codes/cool_lowfreq/vsi.f90

@@ -1,0 +1,411 @@
+module global
+  character*6 :: vbc
+  integer :: nz, nzeff, bignz
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls  
+  real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:)
+  complex*16, parameter :: ii = (0d0,1d0)
+  complex*16, allocatable :: bigW(:), dbigW(:), bigQ(:), dbigQ(:), d2bigW(:), vz(:), dvz(:)
+end module global
+
+program vsi
+  use global
+  implicit none
+  character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
+  integer, allocatable :: ipiv(:), eigen_test(:)
+  integer :: i,ip1, j,k, loc(1), lmax, nzmid, lwork, info
+  real*8 :: T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2, m, zbar,lmode
+  real*8, allocatable :: rwork(:)
+  complex*16, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)
+  complex*16, allocatable :: L1(:,:), L2(:,:), L3(:,:), L4(:,:), &
+       L1bar(:,:), L2bar(:,:), L3bar(:,:), L4bar(:,:), L5bar(:,:), &
+       L6bar(:,:)
+  complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
+       vr(:,:), rhs(:,:)
+  real*8, external :: dlogrho_dz, domega2_dz, omega2_z, logrho_z, kappa2_z, d2logrho_dz2, csq_z 
+  namelist /params/ smallq, smallp, eps, gmma, bgmma, bcool, kx, zmax, vbc
+  namelist /grid/ nz
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=grid)
+  close(7)
+
+  if(mod(nz,2).eq.0) then
+     print*, 'Nz needs to be odd but Nz=', nz
+     stop
[…]

### 2014-12-20 codes/global/vsi.f90

@@ -1,0 +1,345 @@
+module global
+  character*6 :: vbc
+  integer :: nz, nzeff, bignz
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, smallh, kx, zmax 
+  real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), freq(:), growth(:)
+  complex*16, parameter :: ii = (0d0,1d0)
+  complex*16, allocatable :: bigW(:) 
+end module global
+
+program vsi
+  use global
+  implicit none
+  character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
+  integer, allocatable :: ipiv(:), eigen_test(:)
+  integer :: i,ip1, j,k, loc(1), lmax, nzmid, lwork, info
+  real*8 :: T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2, m, zbar,lmode
+  real*8, allocatable :: rwork(:)
+  complex*16, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)
+  complex*16, allocatable :: L1(:,:), L2(:,:), L3(:,:)
+  complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
+       vr(:,:), rhs(:,:)
+  real*8, external :: dlogrho_dz, domega2_dz, shear_z, omega2_z, logrho_z, kappa2_z, d2logrho_dz2
+  namelist /params/ smallq, smallp, smallh, kx, zmax, vbc
+  namelist /grid/ nz
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=grid)
+  close(7)
+
+  if(mod(nz,2).eq.0) then
+     print*, 'Nz needs to be odd but Nz=', nz
+     stop
+  else
+     Nzmid = (Nz+1)/2
+  endif
+
+  if(vbc.ne.'nolagp') then
+  nzeff = nz-2
+  else
+  nzeff = nz
+  endif 
+
+  bignz = 2*nzeff
+
+  !allocate grids
+  allocate(zaxis(nz))
+  allocate(logrho(nz))
+  allocate(dlogrho(nz))
+  allocate(d2logrho(nz))
+  allocate(omega2(nz))
+  allocate(domega2(nz))
[…]

### 2014-12-20 codes/global/vsi_lowfreq.f90

@@ -1,0 +1,332 @@
+module global
+  character*6 :: vbc
+  integer :: nz, lwork, info, nzmid, nzeff
+  integer, allocatable :: ipiv(:), eigen_test(:)
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, smallh, kx, zmax, zbar,lmode, T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2
+  real*8 :: dz
+  real*8, allocatable :: zaxis(:), rwork(:), freq(:), growth(:)
+  real*8, allocatable :: logrho(:), dlogrho(:), omega2(:), domega2(:), kappa2(:)
+  complex*16, parameter :: ii = (0d0,1d0)
+  complex*16 :: sig2, sig
+  complex*16, allocatable :: bigpi(:),  T(:,:), Tp(:,:), Tpp(:,:), matrix(:,:), w(:), vl(:,:), vr(:,:), rhs(:,:)
+  complex*16, allocatable :: work(:), alpha(:), beta(:)
+end module global
+
+program vsi
+  use global
+  implicit none
+  character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
+  integer :: i,j,k, loc(1), lmax, ip1
+  real*8, external :: dlogrho_dz, domega2_dz, omega2_z, shear_z, logrho_z
+  real*8 :: m
+  namelist /params/ smallq, smallp, smallh, kx, zmax, vbc
+  namelist /grid/ nz
+  
+  !read input parameters.
+  open(7, file="input")
+  read(7, nml=params)
+  read(7, nml=grid)
+  close(7)
+  
+  if(mod(nz,2).eq.0) then
+     print*, 'Nz needs to be odd but Nz=', nz
+     stop
+  else
+     Nzmid = (Nz+1)/2
+  endif
+  
+  !allocate grids
+  allocate(zaxis(nz))
+  allocate(logrho(nz))
+  allocate(dlogrho(nz))
+  allocate(omega2(nz))
+  allocate(domega2(nz))
+  allocate(kappa2(nz))
+  allocate(bigpi(nz))
+  
+  allocate(T(nz,nz))
+  allocate(Tp(nz,nz))
+  allocate(Tpp(nz,nz))
+  
+  
+  if(vbc.ne.'nolagp') then
+     nzeff = nz-2
+  else
+     nzeff = nz
[…]

### 2014-12-20 codes/cool_full/var_kx/vsi.f90
added var_kx code

@@ -1,0 +1,456 @@
+module global
+  character*6 :: vbc
+  logical :: gterms
+  integer :: nz, nzeff, bignz, nkx
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls
+  real*8 :: kxmin, kxmax, dlogkx
+  real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:)
+  real*8, allocatable :: kaxis(:)
+  complex*16, parameter :: ii = (0d0,1d0)
+  complex*16, allocatable :: bigW(:), dbigW(:), bigQ(:), vx(:), vy(:), vz(:), dvz(:) 
+end module global
+
+program vsi
+  use global
+  implicit none
+  character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
+  integer, allocatable :: ipiv(:), eigen_test(:)
+  integer :: i,ip1, j,k, loc(1), lmax, nzmid, lwork, info
+  real*8 :: T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2, m, zbar,lmode
+  real*8, allocatable :: rwork(:)
+  complex*16, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)
+  complex*16, allocatable :: L1(:,:), L2(:,:), L3(:,:), L4(:,:), &
+       L5(:,:), L6(:,:), L7(:,:), L8(:,:), &
+       L9(:,:), L10(:,:), L11(:,:), &
+       L1bar(:,:), L2bar(:,:), L3bar(:,:), L4bar(:,:), L5bar(:,:), &
+       L6bar(:,:)
+  complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
+       vr(:,:), rhs(:,:)
+  real*8, external :: dlogrho_dz, domega2_dz, omega2_z, logrho_z, kappa2_z, d2logrho_dz2, csq_z 
+  namelist /params/ smallq, smallp, eps, gmma, bgmma, bcool, zmax, vbc, gterms
+  namelist /loop/ kxmin, kxmax, nkx 
+  namelist /grid/ nz
+  
+  !read input parameters.
+  open(7, file="input")
[…]

### 2014-12-20 linear_adia.tex
corrected beta vs nz estimate, updated numerical adiabatic section

@@
 $\gamma/\Gamma-1$. The latter term is associated with the vertical
-entropy gradient, or buoyancy frequency $N_z$.
+entropy gradient, or buoyancy frequency squared, $N_z^2$.
 
@@
 \begin{align}
-  t_c \ll \frac{1}{N_z}. 
+  t_c \ll \frac{\Omega}{\gamma N_z^2}. 
 \end{align} 
 Parametrizing $t_c = \beta\Omega_k^{-1}$ and inserting the expression
-for $N_z$ appropriate for vertically isothermal disks, we expect
+for $N_z^2$ appropriate for vertically isothermal disks, we expect
 \begin{align}
   \beta \ll
-  \frac{1}{H_\mathrm{iso}\sqrt{\frac{\gamma-1}{\gamma}}\left|\frac{\p\ln\rho}{\p
-          z}\right|} 
+  \frac{1}{H^2_\mathrm{iso}\left(\gamma-1\right)\left(\frac{\p\ln\rho}{\p
+          z}\right)^2} 
 \end{align}
-is required to enable the VSI. This inequality is satisfied at the
+is required to enable the VSI, where we used $\Omega\sim\Omega_k$.  
+This inequality is satisfied at the
 midplane and as $|z|\to\infty$. It can be satisifed at all heights if
 it is satisfied at the height of $\mathrm{max}|\p\ln\rho/\p z|$. 
-Using the expression for the density field for vertically isothermal
+Using the full expression for the density field for vertically isothermal
 disks (Eq. \ref{vertiso_eqm}), we expect 
 \begin{align}
-  \beta \ll \epsilon\frac{3^{3/2}}{2}\sqrt{\frac{\gamma}{\gamma-1}},
+  \beta \ll \frac{27}{4(\gamma-1)}\epsilon^2,
 \end{align}
-i.e. $\beta\ll O(\epsilon)$, is a sufficient condition for
+i.e. $\beta\ll O(\epsilon^2)$, is a sufficient condition for
 bouyancy to be negligible.

### 2014-12-20 numerical.tex
corrected beta vs nz estimate, updated numerical adiabatic section

@@
 nearly-vertically isothermal disk with $\Gamma=1.011$, $p=-1.5$,
-$q=-1$, $\epsilon=0.1$, and the perturbation wavenumber $k_x = 200\pi$
+$q=-1$, $\epsilon=0.1$, and the perturbation wavenumber $k_x = 200\pi/r$
 (or $\hat{k}=20\pi$).
@@
 ($|q|=1$ and $\hat{k} = 20\pi$), Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$,
-corresponding to a characteristic decay lengthscale of $\simeq
+corresponding to a characteristic decay length scale of $\simeq
 0.15\epsilon r$, which is consistent with that observed in 
 Fig. \ref{lowfreq_eigenfunc_adia}.
+
+Note that the decay lengthscale is rougly the height within which
+vertical shear is larger than the buoyancy frequency squared. That is,
+\begin{align*}
+\left|r\frac{d\Omega^2}{dz}\right|\gtrsim N_z^2
+\end{align*}
+for
+\begin{align*}
+  \left|\frac{z}{\epsilon r}\right| \lesssim \frac{\epsilon|q|\gamma}{\gamma-1},
+\end{align*}
+in a thin, vertically isothermal disk. This evaluates to $z\sim
+0.2\epsilon r$ for the current disk model, consistent with
+Fig. \ref{lowfreq_eigenfunc_adia}.
+
+Furthermore, the growth rate to should be no larger than the
+maximum vertical shear available within this decay lengthscale. Using 
+Eq. \ref{max_growth} as a rough guide, 
+\begin{align}
+  \frac{\nu}{\epsilon\Omega_k}<\frac{\epsilon |q|^2}\gamma{2(\gamma-1)},
+\end{align}
+or $\nu< 0.1\epsilon\Omega_k$ is expected in our disk model,
+consistent with numerical results.
 

### 2014-12-21 codes/cool_full/var_kx/vsi.f90
updated variable var_kx code

@@ -1,12 +1,12 @@
 module global
   character*6 :: vbc
-  logical :: gterms
-  integer :: nz, nzeff, bignz, nkx
+  logical :: gterms, eigen_trial, eigenv_out
+  integer :: nz, nzeff, bignz, nkx, nb
   real*8, parameter :: pi = 2d0*acos(0d0)
   real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls
-  real*8 :: kxmin, kxmax, dlogkx
+  real*8 :: kxmin, kxmax, dlogkx, bmin, bmax, db
   real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:)
-  real*8, allocatable :: kaxis(:)
+  real*8, allocatable :: kaxis(:), baxis(:)
   complex*16, parameter :: ii = (0d0,1d0)
   complex*16, allocatable :: bigW(:), dbigW(:), bigQ(:), vx(:), vy(:), vz(:), dvz(:) 
 end module global
@@ -16,9 +16,11 @@ program vsi
   implicit none
   character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
   integer, allocatable :: ipiv(:), eigen_test(:)
-  integer :: i,ip1, j,k, loc(1), lmax, nzmid, lwork, info
+  integer :: i,ip1, j,k, n, loc(1), lmax, nzmid, lwork, info
   real*8 :: T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2, m, zbar,lmode
+  real*8 :: w_re, w_im, dummy
   real*8, allocatable :: rwork(:)
+  complex*16 :: wold
   complex*16, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)
   complex*16, allocatable :: L1(:,:), L2(:,:), L3(:,:), L4(:,:), &
        L5(:,:), L6(:,:), L7(:,:), L8(:,:), &
@@ -26,10 +28,10 @@ program vsi
        L1bar(:,:), L2bar(:,:), L3bar(:,:), L4bar(:,:), L5bar(:,:), &
        L6bar(:,:)
   complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
-       vr(:,:), rhs(:,:)
+       vr(:,:), rhs(:,:), wtrial(:)
[…]

### 2014-12-22 codes/cool_full/bloop/vsi.f90
updated codes, added gloop and renamed var_kx to bloop. these codes loop over gmma or bcool

@@ -131,9 +131,9 @@ program vsi
   allocate(ipiv(bignz))
   
   !setup kaxis. input is kxHiso. then output
-  dlogkx = (kxmax-kxmin)/(nkx-1d0)
+  dlogkx = log10(kxmax/kxmin)/(nkx-1d0)
   do i=1, nkx
-  kaxis(i) = kxmin + dlogkx*(i-1d0)
+  kaxis(i) = 10d0**(log10(kxmin) + dlogkx*(i-1d0))
   enddo
   open(10,file='kaxis.dat')
   do i=1, nkx
@@ -357,7 +357,7 @@ program vsi
      endif
      wtrial(k) = w(i)
  
-     if( (mod(n,10).eq.0).or.(n.eq.1)) then !output
+     if(mod(n-1,10).eq.0) then !output
      write(20,fmt='(3(e22.15,x))'), freq(i), growth(i), kaxis(k) 
      if(n.eq.1) print*, 'kloop, kxHiso=', k, kaxis(k), growth(i)
 
@@ -378,10 +378,10 @@ program vsi
      endif
      endif
   enddo!kx loop
-  if( (mod(n,10).eq.0).or.(n.eq.1)) write(10,fmt='(e22.15,x)'), baxis(n)
   if(n.eq.1) print*, 'bcool, kxHsio, max growth='
   loc = maxloc(dimag(wtrial))  
   print*,   baxis(n), kaxis(loc(1)), dimag(wtrial(loc(1))) 
+  if(mod(n-1,10).eq.0) write(10,fmt='(4(e22.15,x))'), baxis(n), kaxis(loc(1)), dble(wtrial(loc(1))), dimag(wtrial(loc(1)))
   enddo!bcool loop
  
   close(10)  

### 2014-12-22 codes/cool_full/gloop/vsi.f90

@@ -1,0 +1,510 @@
+module global
+  character*6 :: vbc
+  logical :: gterms, eigen_trial, eigenv_out
+  integer :: nz, nzeff, bignz, nkx, ng
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls
+  real*8 :: kxmin, kxmax, dlogkx, gmin, gmax, dg
+  real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:)
+  real*8, allocatable :: kaxis(:), gaxis(:)
+  complex*16, parameter :: ii = (0d0,1d0)
+  complex*16, allocatable :: bigW(:), dbigW(:), bigQ(:), vx(:), vy(:), vz(:), dvz(:) 
+end module global
+
+program vsi
+  use global
+  implicit none
+  character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
+  integer, allocatable :: ipiv(:), eigen_test(:)
+  integer :: i,ip1, j,k, n, loc(1), lmax, nzmid, lwork, info
+  real*8 :: T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2, m, zbar,lmode
+  real*8 :: w_re, w_im, dummy
+  real*8, allocatable :: rwork(:)
+  complex*16 :: wold
+  complex*16, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)
+  complex*16, allocatable :: L1(:,:), L2(:,:), L3(:,:), L4(:,:), &
+       L5(:,:), L6(:,:), L7(:,:), L8(:,:), &
+       L9(:,:), L10(:,:), L11(:,:), &
+       L1bar(:,:), L2bar(:,:), L3bar(:,:), L4bar(:,:), L5bar(:,:), &
+       L6bar(:,:)
+  complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
+       vr(:,:), rhs(:,:), wtrial(:)
+  real*8, external :: dlogrho_dz, domega2_dz, omega2_z, logrho_z, kappa2_z, d2logrho_dz2, csq_z 
+  namelist /params/ smallq, smallp, eps, bcool, bgmma, zmax, vbc, gterms
[…]

### 2014-12-22 numerical.tex
corrected beta vs nz estimate, updated numerical adiabatic section
updated codes, added gloop and renamed var_kx to bloop. these codes loop over gmma or bcool
minor paper udpate

@@
 
-Note that the decay lengthscale is rougly the height within which
+Coincidentally, this decay lengthscale is rougly the height within which
 vertical shear is larger than the buoyancy frequency squared. That is,
@@
 \begin{align}
-  \frac{\nu}{\epsilon\Omega_k}<\frac{\epsilon |q|^2}\gamma{2(\gamma-1)},
+  \frac{\nu}{\epsilon\Omega_k}<\frac{\epsilon |q|^2\gamma}{2(\gamma-1)},
 \end{align}
@@
 
+\subsubsection{Other adiabatic indices}
+
 \subsection{Effect of thermal relaxation}
@@
 \subsection{Vertically non-isothermal disks}
-{\bf big gamma > 1, with tcool}
+{\bf big gamma larger than 1 but equal to small gamma, with tcool}
+
+\subsection{Non-uniform thermal relaxation} 
+{\bf variable tcool, large at midplane, small in atmosphere. simplest
+  model is to multiply tcool by background density}

### 2014-12-23 linear_iso.tex
replace numerical section with that of full equations
added nearly isothermal discussion, only consider special case of fundamental mode for now. should generalize

@@
 examine separately the limiting case of isothermal perturbations
-($\gamma\equiv1$) and adiabatic perturbations ($\gamma\neq 1$ and
-$t_c\to\infty$).   
-\subsection{Isothermal perturbations}
+($\gamma\equiv1$ and any value of $t_c$), nearly-isothermal
+perturbations ($\gamma\geq 1$ and $t_c\to0$), and 
+adiabatic perturbations ($\gamma\neq 1$ and $t_c\to\infty$).   
+\subsection{Isothermal perturbations}\label{iso_discuss}
 The linear problem simplifies considerably for the case
@@
 \end{align} 
-subject to appropriate boundary conditions. We remark that Eq
-.\ref{iso_ode} is also applicable for    
+subject to appropriate boundary conditions. Eq. \ref{iso_ode} also
+applies to $\gamma>1$ for $t_c\to0$.
+
+We remark that Eq.\ref{iso_ode} is also applicable for    
 $\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. neutrally
@@
 must satisfy
-\begin{align}
+\begin{align}\label{sig2_iso}
 \hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
@@
 solving the governing equation.
+
+\subsection{Nearly-isothermal perturbations}
+We can use the explicit solution developed above to examine the effect
+of cooling $t_c$ on perturbations with $\gamma>1$. We shall consider
+the limit $t_c\to0$, for which perturbations can be considered
+nearly-isothermal.
+
+We begin by writing the pair of ODEs for $W$ and $Q$
+(Eq. \ref{ode_w}---\ref{ode_Q}) in the thin-disk, low-frequency, 
+nearly-Keplerian approximations as used previously, as well as
+adopting dimensionless co-ordinates and variables, 
+\begin{align}
+  &\hat{\sigma}^2\left(Q +  \hat{k}^2W\right) \notag\\ 
+  &= \hat{z} \left(1 +
+    \ii\epsilon q \hat{k} \right)\left[W^\prime - \zhat \left(W -
+      Q\right)\right] - \left[W^\prime - \zhat \left(W -
+      Q\right)\right]^\prime\label{nearly_iso1},\\
+&\beta \hat{\sigma}^2\left(W - \gamma Q\right) + \ii\hat{\sigma}
+    \left(W-Q\right)\notag\\
+   & = -\beta\zhat\left(\gamma-1\right) \left[W^\prime - \zhat \left(W -
+      Q\right)\right]\label{nearly_iso2},
+\end{align}
+where we recall $\beta = t_c\Omega_k$ is the dimensionless cooling
+time. It is evident that for $\beta\equiv 0$, we recover
+Eq. \ref{iso_ode3}.
+
+Since we have developed solutions for $\beta\equiv 0$, we may ask how
+do the eigenfrequencies and eigenfunctions change when we change
+$\beta$ from zero to a small but finite value? For sufficiently small
+$\beta$, we expect the new solutions to only differ slighly from that
+above. Then we may linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2} to
+determine the effect of adding small finite cooling.
+
+Let 
+\begin{align}\label{nearly_iso_pert}
+  \beta \to 0 + \delta\beta,\, W\to W+\delta W,\, Q \to Q+\delta
+  Q,\,\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}, 
+\end{align} 
+where $W=Q$ is the polynomial solution described in \S\ref{iso_poly}
+and $\hat{\sigma}^2$ is given by Eq. \ref{sig2_iso}. Inserting
+Eq. \ref{nearly_iso_pert} into
+Eq. \ref{nearly_iso1}---\ref{nearly_iso2} and keeping only first order
+terms, we obtain
+\begin{align}
+  &2\hat{\sigma}\delta\hat{\sigma}\left(1+\hat{k}^2\right)W +
+  \hat{\sigma}^2 \left(\delta Q + \hat{k}^2\delta W\right)\notag\\
+  &= \zhat \left(1+\ii\epsilon q \hat{k}\right)\left[\delta W^\prime -
+    \zhat \left(\delta W - 
+      \delta Q\right)\right] - \delta W^{\prime\prime}\notag\\
+  &\phantom{=} + \left[\zhat \left(\delta W -
+     \delta Q\right)\right]^\prime,\label{nearly_iso3}\\
+&\ii\hat{\sigma} \left(\delta W - \delta Q\right) =
+\delta\beta\left(\gamma-1\right)\left(\hat{\sigma}^2W - \hat{z}W^\prime\right).\label{nearly_iso4}
+\end{align}
+
+We could proceed further by considering the set of polynomial
+solutions for $W$ described in \S\ref{iso_poly} altogether, but it is
+simplest to look at the fundamental mode. Thus
+we consider the `basic state'
+\begin{align*}
+  W = \hat{z},\quad \hat{\sigma}^2\left(1+\hat{k}^2\right) = 1 +
+  \ii\epsilon q \hat{k},
+\end{align*}
+then we can assume
+\begin{align}
+  \delta W = b \zhat^3,
+\end{align}
+where $b$ is a constant to be determined. Eliminating $\delta Q$ from
+Eq. \ref{nearly_iso3}---\ref{nearly_iso4}, inserting the above
+expressions for $W$ and $\delta W$ and balancing coefficients of
+$\zhat$ and $\zhat^3$, we find
+\begin{align}
+ & 2\left(1+\ii\epsilon q \hat{k}\right) \delta\hat{\sigma} \notag\\& = -6b\hat{\sigma} -
+2\ii \delta \beta\left(\gamma-1\right)\left(\hat{\sigma}^2-1\right) - \ii
+  \hat{\sigma}^2\delta\beta\left(\gamma-1\right),\\
+& 0=  2b\hat{\sigma}+\ii\delta\beta
+\left(\gamma-1\right)\left(\hat{\sigma}^2 - 1\right),
+\end{align}
+which gives
+\begin{align}\label{nearly_iso_dsig}
+  \delta \hat{\sigma} =
+  -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii\epsilon
+    q - \hat{k}\right)^2}{2\left(1+\ii\epsilon q \hat{k}\right)\left(1+\hat{k}^2\right)^2}\delta\beta.
+\end{align}
+The imaginary part of Eq. \ref{nearly_iso_dsig} is
+\begin{align}
+   \delta\hat{\nu} =
+  -\frac{\left(\gamma-1\right)\hat{k}^2 \left(\hat{k}^2 -
+      2\epsilon^2q^2\hat{k}^2 - \epsilon^2q^2\right)}{2\left(1+\epsilon^2 q^2
+        \hat{k}^2\right)\left(1+\hat{k}^2\right)^2}\delta\beta.  
+\end{align}
+Since $\epsilon \ll 1$, introducing finite cooling $\delta\beta>0$
+implies $\delta\nu < 0$, i.e. stabilization. Interestingly, for both
+$\hat{k}\to0$ and $\hat{k}\to\infty$ we find $\delta\hat{\nu}\to0$, so
+stabilization by finite cooling is ineffective at very large or very small scales.

### 2014-12-23 linear_adia.tex
minor updates

@@
 \subsection{Adiabatic perturbations}\label{analytic_adia}
-We now extend the above discussion to general $\gamma$, but in the
-adiabatic limit ($t_c\to\infty$). We eliminate variables from 
+We now consider general $\gamma$ in the
+adiabatic limit ($t_c\to\infty$). This is the opposite regime to
+isothermal perturbations. We eliminate variables from 
 Eq. \ref{lin_mass}---\ref{lin_energy} in favor of $\delta v_z$ 

### 2014-12-23 linear_adia.tex
added numerical results for variable gmma

@@
 
-\subsubsection{Comparison with $\gamma=1$}
+\subsubsection{Comparison with $\gamma=1$}\label{adia_compare_gam1}
 We can compare growth rates obtained for $\gamma=2$ to those for
@@
 
-\subsubsection{Qualitative expectations in the presence of thermal
-  relaxation}
+\subsubsection{Qualitative condition for buoyancy to be negligible in
+  the presence of thermal relaxation}
 We have shown that when $\gamma>1$ so that there is a positive

### 2014-12-23 linear_iso.tex
added numerical results for variable gmma

@@
 \subsection{Nearly-isothermal perturbations}
-We can use the explicit solution developed above to examine the effect
-of cooling $t_c$ on perturbations with $\gamma>1$. We shall consider
-the limit $t_c\to0$, for which perturbations can be considered
-nearly-isothermal.
+The explicit solution developed in \S\ref{iso_poly} is valid for
+non-isothermal perturbations $\gamma>1$ if the thermal relaxation
+(cooling) timescale $t_c\equiv0$. We may ask how
+do the eigenfrequencies and eigenfunctions change when we change
+$t_c$ from zero to a small but finite value? For sufficiently small
+$t_c$, for which the perturbations may be considered
+nearly-isothermal, we expect the new solutions to only differ slighly
+from that above.
+
+% Then we may linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2} to
+% determine the effect of adding small finite cooling.
 
@@
 
-Since we have developed solutions for $\beta\equiv 0$, we may ask how
-do the eigenfrequencies and eigenfunctions change when we change
-$\beta$ from zero to a small but finite value? For sufficiently small
-$\beta$, we expect the new solutions to only differ slighly from that
-above. Then we may linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2} to
-determine the effect of adding small finite cooling.
-
-Let 
+We linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2} about a known
+solution for $\beta\equiv0$. Let 
 \begin{align}\label{nearly_iso_pert}
@@
 where $W=Q$ is the polynomial solution described in \S\ref{iso_poly}
-and $\hat{\sigma}^2$ is given by Eq. \ref{sig2_iso}. Inserting
+with $\hat{\sigma}^2$ given by Eq. \ref{sig2_iso}. Inserting
 Eq. \ref{nearly_iso_pert} into
@@
 \end{align}
+This is a pair of inhomogeneous ODEs with a forcing by $W$. The
+homogeneous problem is, of course, just the ODE for $\beta=0$ solved
+in \S\ref{iso_poly}. Hence we only seek the particular solution.
 
-We could proceed further by considering the set of polynomial
-solutions for $W$ described in \S\ref{iso_poly} altogether, but it is
-simplest to look at the fundamental mode. Thus
-we consider the `basic state'
+We could proceed by considering the entire set of polynomial
+solutions for $W$, but it is simplest to look at the fundamental
+mode. Thus we consider the `basic state' 
 \begin{align*}
@@
 \end{align*}
-then we can assume
+then
 \begin{align}
@@
 \end{align}
-which gives
+where the expression for $\hat{\sigma}^2$ was used. Solving for
+$\delta\hat{\sigma}$ gives 
 \begin{align}\label{nearly_iso_dsig}
@@
 Since $\epsilon \ll 1$, introducing finite cooling $\delta\beta>0$
-implies $\delta\nu < 0$, i.e. stabilization. Interestingly, for both
+implies $\delta\hat{\nu} < 0$, i.e. stabilization. Interestingly, for both
 $\hat{k}\to0$ and $\hat{k}\to\infty$ we find $\delta\hat{\nu}\to0$, so
-stabilization by finite cooling is ineffective at very large or very small scales.
+stabilization by finite cooling is ineffective at very large or very
+small scales. We plot Eq. \ref{nearly_iso_dsig} in Fig. \ref{domegadbeta} with
+$\epsilon=0.05$, $q=-1.0$ and $\gamma=1.4$. For this disk model the
+optimum wavenumber for stabilization is $\hat{k}\sim 6$.
+
+% This contrasts to the case without vertical shear
+% ($q=0$), in which $\p\hat{\sigma}/\p\beta\to\mathrm{constant}$ as 
+% $\hat{k}\to\infty$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/domegadbeta}
+  \caption{Rate of change of the fundamental VSI eigenfrequency
+    $\sigma$ with respect to the introduction of a small but finite
+    cooling time. The plotted quantity is $\p\hat{\sigma}/\p\beta$ at $\beta=0$, evaluated using
+    Eq. \ref{nearly_iso_dsig} with 
+    disk parameters $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The real
+    (imaginary) part of $\p\hat{\sigma}/\p\beta$ is shown as the solid
+    (dotted) lines. 
+    \label{domegadbeta}}  
+\end{figure}
+
+One effect of introducing finite thermal relaxation is buoyancy. We
+show in the next section that buoyancy strongly stabilizes the
+VSI. The solution $\delta W\propto\zhat^3$ then makes sense because it
+is most significant for large $\zhat$, i.e. away from the midplane
+where the effect of bouyancy first appears as $t_c$ is increased.

### 2014-12-23 numerical.tex
added numerical results for variable gmma

@@
 \subsubsection{Other adiabatic indices}
+In Fig. \ref{adia_growth_num} we plot the eigenfrequencies of the
+fundamental VSI mode for a range of $\gamma$ values. This plot is to
+be compared with Fig. \ref{adia_growth}.
+
+% The numerically-determined eigenfrequencies are qualitatively
+% consistent with the discussion in \S\ref{adia_compare_gam1}.
+There is good agreement between Fig. \ref{adia_growth_num} and
+Fig. \ref{adia_growth} for $k_xH_\mathrm{iso}\gtrsim10$. There is a
+mistach for smaller wavenumbers, especially in the growth
+rates. Actual growth rates decrease slower as $k_xH_\mathrm{iso}\to0$
+than that in Fig. \ref{adia_growth}. Disagreement at small
+$k_xH_\mathrm{iso}$ is not surprising because the low-frequency
+approximation used to obtain Fig. \ref{adia_growth} becomes invalid since
+$|\omega|\sim\Omega_k$ for small $k_x$.
+
+Given the number of approximations made in \S\ref{adia_compare_gam1},
+Fig. \ref{adia_growth_num} compares well with
+Fig. \ref{adia_growth}. The main qualitative features of the full
+numerical solution in Fig. \ref{adia_growth_num} are captured by
+Fig. \ref{adia_growth}: stabilization for
+$k_xH_\mathrm{iso}\to0$ and $k_xH_\mathrm{iso}\to\infty$,
+stabilization for increasing $\gamma$ (but ineffective at small $k_x$), 
+and the shift of the most unstable wavenumber to smaller values as
+$\gamma$ is increased.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_real1}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag1}
+  \caption{Real frequency (top) and growth rate 
+    (bottom) of the fundamental VSI mode in nearly-vertically   
+    isothermal disks ($\Gamma=1.011$) with $\epsilon=0.1$ and  
+    $q=-1$, subject to adiabatic perturbations
+    ($\beta=10^9$). The eigenfrequencies are calculated by numerically
+    solving the full linear eigenvalue problem. \label{adia_growth_num}}   
+\end{figure}
 
 \subsection{Effect of thermal relaxation}
-{\bf vertical isothermal disk with tcool}
+We now introduce thermal relaxation. Here we consider a slightly
+different disk model with $\epsilon=0.05$ and
+$\gamma=1.4$. This model was used by \citep{nelson13}, where the
+authors find a cooling timescale of $\beta\geq 0.1$ was sufficient to
+stabilize the VSI in non-linear hydrodynamic simulations.
 
-\subsection{Vertically non-isothermal disks}
-{\bf big gamma larger than 1 but equal to small gamma, with tcool}
+% \subsection{Vertically non-isothermal disks}
+% {\bf big gamma larger than 1 but equal to small gamma, with tcool}
 
-\subsection{Non-uniform thermal relaxation} 
-{\bf variable tcool, large at midplane, small in atmosphere. simplest
-  model is to multiply tcool by background density}
+% \subsection{Non-uniform thermal relaxation} 
+% {\bf variable tcool, large at midplane, small in atmosphere. simplest
+%   model is to multiply tcool by background density}

### 2014-12-27 linear_adia.tex
extended analytic discussion to thermal relaxation cases

@@
   0 = \delta v_z^{\prime\prime} - \zhat A \delta v_z^\prime + \left(B
-    - C \zhat^2\right)\delta v_z,
+    - C \zhat^2\right)\delta v_z, \label{adia_iso4} 
 \end{align}
@@
 \begin{align}
-  B + \alpha = M\left(A-2\alpha\right).
+  B + \alpha = M\left(A-2\alpha\right).\label{adia_disp0}
 \end{align}
@@
 bouyancy to be negligible.
+
+\subsection{Perturbations with thermal relaxation}
+Having examined the extreme limits of isothermal ($t_c\to0$) and
+adiabatic ($t_c\to\infty$) perturbations, we now generalize further to consider
+perturbations with intermediate thermodynamic responses with general
+values of $t_c$.
+
+In this section we will pursue explicit solutions in the low-frequency,
+thin-disk, nearly-Keplerian approximations as used in the development
+for adiabatic solutions. The pair of ODEs for $W$ and $Q$
+(Eq. \ref{ode_w}---\ref{ode_Q}) with these approximations and 
+written with dimensionless variables, are
+\begin{align}
+  &\hat{\sigma}^2\left(Q +  \hat{k}^2W\right) \notag\\ 
+  &= \hat{z} \left(1 +
+    \ii\epsilon q \hat{k} \right)\left[W^\prime - \zhat \left(W -
+      Q\right)\right] - \left[W^\prime - \zhat \left(W -
+      Q\right)\right]^\prime\label{nearly_iso1},\\
+&\beta \hat{\sigma}^2\left(W - \gamma Q\right) + \ii\hat{\sigma}
+    \left(W-Q\right)\notag\\
+   & = -\beta\zhat\left(\gamma-1\right) \left[W^\prime - \zhat \left(W -
+      Q\right)\right]\label{nearly_iso2},
+\end{align}
+where we recall $\beta = t_c\Omega_k$ is the dimensionless cooling
+time. It is evident that for $\beta\equiv 0$, we recover
+Eq. \ref{iso_ode3}. In addition, from the expression of the vertical
+velocity perturbation (Eq. \ref{lin_vz}) we have
+\begin{align}
+  W^\prime - \zhat \left(W -
+      Q\right) = \ii\hat{\sigma}c_s\delta v_z. \label{nearly_vz}
+\end{align}
+
+\subsubsection{Perturbative effect of introducing a small but finite
+  thermal relaxation time}
+
+It is instructive to first ask: how
+do the eigenfrequencies and eigenfunctions change when we change
+$t_c$ from zero to a small but finite value? For sufficiently small
+$t_c$, for which the perturbations may be considered
+nearly-isothermal, we expect the new solutions to only differ slighly
+from the explicit isothermal solutions developed in \S\ref{iso_poly}.
+
+%is valid for
+%non-isothermal perturbations $\gamma>1$ if the thermal relaxation
+%(cooling) timescale $t_c\equiv0$.
+
+% Then we may linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2}
+
+To determine the (assumed) small effect of introducing a short thermal
+relaxation, we linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2} about a known
+solution for $\beta\equiv0$. Let 
+\begin{align}\label{nearly_iso_pert}
+  \beta \to 0 + \delta\beta,\, W\to W+\delta W,\, Q \to Q+\delta
+  Q,\,\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}, 
+\end{align} 
+where $W=Q$ is the polynomial solution described in \S\ref{iso_poly}
+with $\hat{\sigma}^2$ given by Eq. \ref{sig2_iso}. Inserting
+Eq. \ref{nearly_iso_pert} into
+Eq. \ref{nearly_iso1}---\ref{nearly_iso2} and keeping only first order
+terms, we obtain
+\begin{align}
+  &2\hat{\sigma}\delta\hat{\sigma}\left(1+\hat{k}^2\right)W +
+  \hat{\sigma}^2 \left(\delta Q + \hat{k}^2\delta W\right)\notag\\
+  &= \zhat \left(1+\ii\epsilon q \hat{k}\right)\left[\delta W^\prime -
+    \zhat \left(\delta W - 
+      \delta Q\right)\right] - \delta W^{\prime\prime}\notag\\
+  &\phantom{=} + \left[\zhat \left(\delta W -
+     \delta Q\right)\right]^\prime,\label{nearly_iso3}\\
+&\ii\hat{\sigma} \left(\delta W - \delta Q\right) =
+\delta\beta\left(\gamma-1\right)\left(\hat{\sigma}^2W - \hat{z}W^\prime\right).\label{nearly_iso4}
+\end{align}
+This is a pair of inhomogeneous ODEs with a forcing by $W$. The
+homogeneous problem is, of course, just the ODE for $\beta=0$ solved
+in \S\ref{iso_poly}. Hence we only seek the particular solution.
+
+We could proceed by considering the entire set of polynomial
+solutions for $W$, but it is simplest to look at the fundamental
+mode. Thus we consider the `basic state' 
+\begin{align*}
+  W = \hat{z},\quad \hat{\sigma}^2\left(1+\hat{k}^2\right) = 1 +
+  \ii\epsilon q \hat{k},
+\end{align*}
+then
+\begin{align}
+  \delta W = b \zhat^3,
+\end{align}
+where $b$ is a constant to be determined. Eliminating $\delta Q$ from
+Eq. \ref{nearly_iso3}---\ref{nearly_iso4}, inserting the above
+expressions for $W$ and $\delta W$ and balancing coefficients of
+$\zhat$ and $\zhat^3$, we find
+\begin{align}
+ & 2\left(1+\ii\epsilon q \hat{k}\right) \delta\hat{\sigma} \notag\\& = -6b\hat{\sigma} -
+2\ii \delta \beta\left(\gamma-1\right)\left(\hat{\sigma}^2-1\right) - \ii
+  \hat{\sigma}^2\delta\beta\left(\gamma-1\right),\\
+& 0=  2b\hat{\sigma}+\ii\delta\beta
+\left(\gamma-1\right)\left(\hat{\sigma}^2 - 1\right),
+\end{align}
+where the expression for $\hat{\sigma}^2$ was used. Solving for
+$\delta\hat{\sigma}$ gives 
+\begin{align}\label{nearly_iso_dsig}
+  \delta \hat{\sigma} =
+  -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii\epsilon
+    q - \hat{k}\right)^2}{2\left(1+\ii\epsilon q \hat{k}\right)\left(1+\hat{k}^2\right)^2}\delta\beta.
+\end{align}
+The imaginary part of Eq. \ref{nearly_iso_dsig} is
+\begin{align}
+   \delta\hat{\nu} =
+  -\frac{\left(\gamma-1\right)\hat{k}^2 \left(\hat{k}^2 -
+      2\epsilon^2q^2\hat{k}^2 - \epsilon^2q^2\right)}{2\left(1+\epsilon^2 q^2
+        \hat{k}^2\right)\left(1+\hat{k}^2\right)^2}\delta\beta.  
+\end{align}
+Since $\epsilon \ll 1$, introducing finite cooling $\delta\beta>0$
+implies $\delta\hat{\nu} < 0$, i.e. stabilization. Interestingly, for both
+$\hat{k}\to0$ and $\hat{k}\to\infty$ we find $\delta\hat{\nu}\to0$, so
+stabilization by finite cooling is ineffective at very large or very
+small scales. We plot Eq. \ref{nearly_iso_dsig} in Fig. \ref{domegadbeta} with
+$\epsilon=0.05$, $q=-1.0$ and $\gamma=1.4$. For this disk model the
+optimum wavenumber for stabilization is $\hat{k}\sim 5$---$6$.
+
+% This contrasts to the case without vertical shear
+% ($q=0$), in which $\p\hat{\sigma}/\p\beta\to\mathrm{constant}$ as 
+% $\hat{k}\to\infty$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/domegadbeta}
+  \caption{Rate of change of the fundamental VSI eigenfrequency
+    $\sigma$ with respect to the introduction of a small but finite
+    cooling time. The plotted quantity is $\p\hat{\sigma}/\p\beta$ at $\beta=0$, evaluated using
+    Eq. \ref{nearly_iso_dsig} with 
+    disk parameters $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The real
+    (imaginary) part of $\p\hat{\sigma}/\p\beta$ is shown as the solid
+    (dotted) lines. 
+    \label{domegadbeta}}  
+\end{figure}
+
+One effect of introducing finite thermal relaxation is buoyancy. % We
+% show in the next section that buoyancy strongly stabilizes the
+% VSI.
+The solution $\delta W\propto\zhat^3$ then makes sense because it
+is most significant for large $\zhat$, i.e. away from the midplane
+where the effect of bouyancy first appears as $t_c$ is increased.
+
+\subsubsection{General explicit solutions with thermal relaxation}
+To develop explicit solutions of
+Eq. \ref{nearly_iso1}---\ref{nearly_vz}, we eliminate variables in
+favor of $\delta v_z$ to obtain
+\begin{align}
+  0 = \delta v_z^{\prime\prime} - z \widetilde{A}\delta v_z^\prime +
+  (\widetilde{B} - \widetilde{C}\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
+\end{align}
+with
+\begin{align}
+  &\widetilde{A} \equiv 1 + \ii \epsilon q \hat{k},\\
+  &\widetilde{B} \equiv \hat{\sigma}^2\left(\chi + \hat{k}^2\right) -
+  \left(\chi + \ii \epsilon q \hat{k}\right),\\
+  &\widetilde{C} \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
+      \epsilon q\hat{k}\right),
+\end{align}
+where
+\begin{align}
+  \chi \equiv \frac{1-\ii\hat{\sigma}\beta}{1-\ii\hat{\sigma}\beta\gamma}.
+\end{align}
+
+The governing equation Eq. \ref{nearly_iso_explicit} has the same form
+as in the adiabatic case (Eq. \ref{adia_iso4}), so we can apply the
+same solution method. However, the coefficients
+$\widetilde{B}$ and $\widetilde{C}$ now have a more complicated
+dependence on $\hat{\sigma}$.  The adiabatic dispersion relation,
+Eq. \ref{adia_disp0}, generalizes to a $6^\mathrm{th}$ order
+polynomial in $\hat{\sigma}$:
+\begin{align}
+  0 = \sum_{l=0}^{6}c_l\hat{\sigma}^l,\label{relax_disp}
+\end{align}
+where the coefficients $c_l$ are given in Appendix \ref{relax_coeff}.
+Eq. \ref{relax_disp} contrasts to Eq. \ref{adia_disp}, which is
+quadratic in $\hat{\sigma}$ for adiabatic perturbations.
+
+It is simplest to solve Eq. \ref{relax_disp} numerically. In
+Fig. \ref{relax_disp_fig} we plot the growth rate of the fundamental
+($M=0$) VSI mode as a function of $k_xH_\mathrm{iso}$ for a range of
+dimensionless thermal relaxation timescales $\beta$, in a disk with
+$\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/rate_theory_grow_relax}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_grow_relax2}
+  \caption{Growth rate of the fundamental VSI mode in a disk with
+    dimensionless thermal relaxation timescales
+    $\beta\in[10^{-3},1]$ (top) and $\beta\in[1,10^3]$ (bottom).
+Other disk disk parameters are
+    $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The growth rates are
+    calculated from the dispersion relation Eq. \ref{relax_disp}. 
+    \label{relax_disp_fig}}  
+\end{figure}
+
+%ate_theory, eps=0.05, smallq=-1d0, gmma=1.4, bmax=1.0, ncases=11, mmode=0.0, root=1, legend=[0.102,0.14,5.4,0.355], yrange=[0,6]

### 2014-12-27 linear_iso.tex
updated figures log axis
extended analytic discussion to thermal relaxation cases

--- 2014-12-18 linear_iso.tex
@@
 examine separately the limiting case of isothermal perturbations
-($\gamma\equiv1$) and adiabatic perturbations ($\gamma\neq 1$ and
-$t_c\to\infty$).   
-\subsection{Isothermal perturbations}
+($\gamma\equiv1$ and any value of $t_c$, or any value of $\gamma$ with
+$t_c = 0$) and adiabatic perturbations ($\gamma\neq 1$ and $t_c\to\infty$).   
+\subsection{Isothermal perturbations}\label{iso_discuss}
 The linear problem simplifies considerably for the case

### 2014-12-29 appendix.tex
replace numerical section with that of full equations
extended analytic discussion to thermal relaxation cases
restructing analytic discussion

--- 2014-12-17 appendix.tex
@@
+\section{Full governing equation for vertically isothermal disks}\label{adia_improve}
+% \section{Improving the nearly-Keplerian approximation for adiabatic
+%   disks}\label{adia_improve}
+ In \S\ref{analytic_adia} we made the replacement $D\to\Omega_k^2$
+ before combining Eq. \ref{adia_iso1}---\ref{adia_iso2} to derive
+ Eq. \ref{adia_iso3}.
+
+This procedure ignores the vertical dependence of
+ $D=\kappa^2(z) - \sigma^2$.
+
+We show here that this has no significant
+ consequence for thin disks.
+
+In the global disk with $\Gamma=1$ we have
+\begin{align}\label{dkappa2}
+  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
+    r}\frac{\p\Omega^2}{\p z} = -
+  \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(2 + q +
+    \frac{z^2/r^2-2}{z^2/r^2+1}\right). 
+\end{align}
+For the local problem we may then write
+\begin{align}
+  \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z;q),
+\end{align}
+where the function $F$ corresponds to the bracket in
+Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
+3+q$ as $|z|\to\infty$, i.e. its magnitude is of order unity.
+
+ Eliminating $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
+ Eq. \ref{lin_vz} for vertically isothermal disks, retaining the
+ vertical dependence of $D$ and including thermal relaxation, we
+ obtain 
+\begin{align}
+  0 =& \frac{d^2\delta v_z}{dz^2} + \left[1 + \frac{\ii k_x c_s^2
+      q}{Dr} - \frac{k_x^2c_s^2}{\left(k_x^2c_s^2 + \chi
+        D\right)}\frac{qc_s^2F}{Dr^2}\right]\frac{d\ln\rho}{dz}\frac{d\delta
+    v_z}{dz} \notag\\
+  &+ \left\{\sigma^2\left(\frac{k_x^2}{D} +
+      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
+        q}{Dr}\right)\frac{d^2\ln\rho}{dz^2} -
+    \frac{c_s^2}{D}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
+      \frac{\ii k_x q}{r}\right)
+   \left[\left(1-\chi\right) +
+     \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
+   \right\}\delta v_z,
+\end{align}
+where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
+t_c \gamma\right)$. Making the low-frequency approximation with the
+replacement $D\to \Omega_k^2$ gives, in terms of dimensionless
+variables,
+\begin{align}
+   &\delta v_z ^{\prime\prime} + \left[1 + \ii\epsilon q\hat{k} -
+    \frac{ \hat{k}^2}{
+      \left(\hat{k}^2+\chi\right)}q \epsilon^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
+  \left\{\left(\chi + \ii \epsilon q
+      \hat{k}\right)\ln\rho^{\prime\prime} - \ln\rho^{\prime
+      2}\left(\khat^2 -
+      \ii\epsilon
+      q\hat{k}\right)\left[1 - \chi +
+      \frac{\chi}{\left(\hat{k}^2+\chi\right)}q\epsilon^2F\right]\right\}\delta v_z \notag\\&=
+  -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
+\end{align}
+
 \section{Alternative inference of instability in the presence of weak
@@
 \end{align}
+
+% Differentiating Eq. \ref{adia_iso1} properly gives
+% \begin{align}
+%   \left(D + \gamma k_x^2 c_s^2\right)\frac{d\Delta}{dz}=& D
+%   \frac{d^2\delta v_z}{dz^2} + k_x^2c_s^2\left[\frac{\gamma}{\left(D + \gamma k_x^2 c_s^2\right)}\frac{dD}{dz} + \ii
+%     \frac{d\ln\rho}{dz}\left(\ii +
+%       \frac{q}{k_xr}\right)\right]\frac{d\delta v_z}{dz}\notag\\
+%   &+ \ii k_x^2 c_s^2 \left(\ii +
+%       \frac{q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} -
+%       \frac{1}{\left(D + \gamma
+%           k_x^2c_s^2\right)}\frac{dD}{dz}\frac{d\ln\rho}{dz}\right]\delta
+%     v_z. \label{adia_diso1}
+% \end{align}
+% The terms that were ignored in deriving Eq. \ref{adia_iso3} are those
+% proportional to $dD/dz$ in Eq. \ref{adia_diso1}. Now, in the global
+% disk with $\Gamma=1$ we have
+% \begin{align}\label{dkappa2}
+%   \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
+%     r}\frac{\p\Omega^2}{\p z} = -
+%   \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(2 + q +
+%     \frac{z^2/r^2-2}{z^2/r^2+1}\right). 
+% \end{align}
+% For the local problem we may then write
+% \begin{align}
+%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z;q),
+% \end{align}
+% where the function $F$ corresponds to the bracket in
+% Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
+% 3+q$ as $|z|\to\infty$, i.e. its magnitude is of order unity. 
+% % We estimate the
+% % importance of these terms by noting that 
+% % \begin{align}
+% %   \frac{dD}{dz}\equiv \frac{d\kappa^2}{dz} \simeq \frac{d\Omega^2}{dz}
+% %   = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2},
+% % \end{align}
+% % for thin disks. 
+% Eq. \ref{adia_diso1} becomes
+% \begin{align}
+% \left(D + \gamma k_x^2 c_s^2\right)\frac{d\Delta}{dz}=& D
+%   \frac{d^2\delta v_z}{dz^2} - k_x^2c_s^2\frac{d\ln\rho}{dz}\left[ 1 - 
+%       \frac{\ii q}{k_xr}  +  \frac{\gamma q c_s^2F}{r^2\left(D +
+%         \gamma k_x^2 c_s^2\right)}\right]\frac{d\delta
+%     v_z}{dz}\notag\\ 
+%   &- k_x^2 c_s^2 \left(1 - 
+%       \frac{\ii q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} + 
+%       \frac{qc_s^2F}{r^2\left(D + \gamma
+%           k_x^2c_s^2\right)}\left(\frac{d\ln\rho}{dz}\right)^2\right]\delta
+%     v_z. \label{adia_diso2}
+% \end{align}
+% Since $D\sim \Omega_k^2$ in the low-frequency limit, we 
+% see from Eq. \ref{adia_diso2} that the neglected terms are
+% $O(\epsilon^2)$. We can make this explicit by combining
+% Eq. \ref{adia_diso1}---\ref{adia_diso2} and Eq. \ref{adia_iso2}, then
+% set $D\to\Omega_k^2$. In terms of non-dimensional variables, the result
+% is  
+% \begin{align}
+%    &\delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon q\hat{k} -
+%     \frac{\gamma q \epsilon^2\hat{k}^2F}{1+\gamma
+%       \hat{k}^2}\right)\ln\rho^{\prime}\delta v_z^\prime +
+%   \left[\left(\frac{1}{\gamma} + \ii \epsilon q
+%       \hat{k}\right)\ln\rho^{\prime\prime} - \hat{k}^2\left(1 -
+%       \frac{\ii\epsilon
+%         q}{\hat{k}}\right)\left(\frac{\gamma-1}{\gamma} +
+%       \frac{q\epsilon^2F}{1+\gamma\hat{k}^2}\right)\ln\rho^{\prime
+%       2}\right]\delta v_z \notag\\&=
+%   -\hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right)\delta v_z\label{adia_diso3}.
+% \end{align}   
+% Eq. \ref{adia_diso3} differs from Eq. \ref{adia_iso3} by terms
+% proportional to $\epsilon^2$. For a thin disk, $\epsilon\ll1$, so
+% neglecting these terms has no qualitative effect on the discussion in
+% \S\ref{analytic_adia}.
+
+\section{Coefficients for the dispersion relation for perturbations
+with thermal relaxation}\label{relax_coeff}
+The coefficients of the dispersion relation Eq. \ref{relax_disp} is
+given by:
+\begin{align}
+  &c_0 = M(M+1)\widetilde{A}^2,\\
+  &c_1 = \ii\beta\left\{\left(1-\gamma\right)\left[1 +
+      \khat^2\left(1+2M\right)^2 - 4 \ii\epsilon q\khat M (M+1)\right] 
+    - 2\widetilde{A}^2\gamma M (M+1)\right\},\\
+  &c_2 = \left(\khat^2 + 1\right)\widetilde{A} + \beta^2\left\{(1-\gamma)\left[1
+      + \gamma \khat^2(1+2M)^2 - 4\ii\epsilon q \khat \gamma M(M+1)
+    \right]
+    -\gamma^2 \widetilde{A}^2 M(M+1)
+  \right\},\\
+  &c_3 = \beta\left\{\epsilon q \khat + \gamma \left[\ii + \epsilon q
+      k \left(1+2\khat^2\right)\right] - 3\ii - 2\ii
+    \khat^2\right\},\\
+  &c_4 =
+  \beta^2\left(1+\gamma\khat^2\right)\left[\gamma\left(1-\ii\epsilon q
+    \khat\right)-2\right] - \left(1+\khat^2\right)^2,\\
+&c_5 = 2\ii\beta\left(1+\khat^2\right)\left(1+\gamma\khat^2\right),\\
+&c_6 = \beta^2\left(1+\gamma\khat^2\right)^2.
+\end{align}

### 2014-12-29 linear_adia.tex
restructing analytic discussion

@@
 \subsection{Adiabatic perturbations}\label{analytic_adia}
-We now consider general $\gamma$ in the
-adiabatic limit ($t_c\to\infty$). This is the opposite regime to
-isothermal perturbations. We eliminate variables from 
-Eq. \ref{lin_mass}---\ref{lin_energy} in favor of $\delta v_z$ 
-and $\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
-\begin{align}
-  &\Delta\left(1 + \frac{\gamma c_s^2 k_x^2}{D}\right) = \frac{d\delta
-    v_z}{dz} + \frac{\ii k_x^2}{D}\left(\ii c_s^2 \frac{d\ln\rho}{dz}
-    - \frac{r}{k_x}\frac{d\Omega^2}{dz}\right)\delta v_z,\label{adia_iso1}\\
-  & -\frac{\sigma^2}{c_s^2}\delta v_z = \gamma \frac{d\Delta}{dz} +
-  \frac{d}{dz}\left(\frac{d\ln\rho}{dz}\delta v_z\right) +
-  \left(\gamma-1\right)\frac{d\ln\rho}{dz} \Delta.\label{adia_iso2}
-\end{align}
-%{\bf note: incompressible mode, set delta to zero. only works for
-% constant gravity.}
+We now consider % general $\gamma$
+perturbations in the adiabatic limit ($t_c,\beta\to\infty$). This is the opposite regime to
+isothermal perturbations. In this case the parameter $\chi = 1/\gamma$
+and Eq. \ref{vertiso_gov_nondim} can be written as 
+% We eliminate variables from 
+% Eq. \ref{lin_mass}---\ref{lin_energy} in favor of $\delta v_z$ 
+% and $\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
+% \begin{align}
+%   &\Delta\left(1 + \frac{\gamma c_s^2 k_x^2}{D}\right) = \frac{d\delta
+%     v_z}{dz} + \frac{\ii k_x^2}{D}\left(\ii c_s^2 \frac{d\ln\rho}{dz}
+%     - \frac{r}{k_x}\frac{d\Omega^2}{dz}\right)\delta v_z,\label{adia_iso1}\\
+%   & -\frac{\sigma^2}{c_s^2}\delta v_z = \gamma \frac{d\Delta}{dz} +
+%   \frac{d}{dz}\left(\frac{d\ln\rho}{dz}\delta v_z\right) +
+%   \left(\gamma-1\right)\frac{d\ln\rho}{dz} \Delta.\label{adia_iso2}
+% \end{align}
+% %{\bf note: incompressible mode, set delta to zero. only works for
+% % constant gravity.}
 
-We make the low-frequency approximation and the replacement 
-$D\to \Omega_k^2$ in Eq. \ref{adia_iso1}---\ref{adia_iso2}. For
-simplicity we do this before combining these equations later. We  
-show in Appendix \ref{adia_improve} that making these approximations
-after combining  Eq. \ref{adia_iso1}---\ref{adia_iso2} adds unecessary
-complexity for thin disks. Here, we also use the fact that in the
-global disk, 
-\begin{align}
-  r\frac{\p \Omega^2}{\p z} = -\frac{\p \ln\rho}{\p z} \frac{qc_s^2}{r},
-\end{align}
-for the power-law prescription of the midplane sound-speed for
-$\Gamma=1$.
+% We make the low-frequency approximation and the replacement 
+% $D\to \Omega_k^2$ in Eq. \ref{adia_iso1}---\ref{adia_iso2}. For
+% simplicity we do this before combining these equations later. We  
+% show in Appendix \ref{adia_improve} that making these approximations
+% after combining  Eq. \ref{adia_iso1}---\ref{adia_iso2} adds unecessary
+% complexity for thin disks. Here, we also use the fact that in the
+% global disk, 
+% \begin{align}
+%   r\frac{\p \Omega^2}{\p z} = -\frac{\p \ln\rho}{\p z} \frac{qc_s^2}{r},
+% \end{align}
+% for the power-law prescription of the midplane sound-speed for
+% $\Gamma=1$.
 
-Making the above substitutions and eliminating $\Delta$ between Eq. \ref{adia_iso1}---
-Eq. \ref{adia_iso2} we obtain, in terms of dimensionless variables defined
-previously, an equation for $\delta v_z$, 
+% Making the above substitutions and eliminating $\Delta$ between Eq. \ref{adia_iso1}---
+% Eq. \ref{adia_iso2} we obtain, in terms of dimensionless variables defined
+% previously, an equation for $\delta v_z$,
+
 \begin{align}
@@
 
-It is instructive to first ask: how
+It is instructive to first ask the more analytically tractable question: how
 do the eigenfrequencies and eigenfunctions change when we change

### 2014-12-29 linear_iso.tex
restructing analytic discussion

@@
 ($\gamma\equiv1$ and any value of $t_c$, or any value of $\gamma$ with
-$t_c = 0$) and adiabatic perturbations ($\gamma\neq 1$ and $t_c\to\infty$).   
-\subsection{Isothermal perturbations}\label{iso_discuss}
-The linear problem simplifies considerably for the case
-$\Gamma=\gamma=1$.  Then 
-\begin{align}
-  Q=W. 
-\end{align}
-Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
-problem for isothermal perturbations in a vertically isothermal disk, 
-\begin{align}\label{iso_ode}
-  \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
-      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
-  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
-\end{align} 
-subject to appropriate boundary conditions. Eq. \ref{iso_ode} also
-applies to $\gamma>1$ for $t_c\to0$.
+$t_c = 0$) and adiabatic perturbations ($\gamma\neq 1$ and
+$t_c\to\infty$), before considering the effect of thermal relaxation.
 
-We remark that Eq.\ref{iso_ode} is also applicable for    
-$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. neutrally
-stratified adiabatic disks), since in that case  
-$Q\simeq W$ as well. However, the following discussion is only
-valid for $\gamma=\Gamma=1$ because $c_s$ is taken to be a constant.
-
-\subsubsection{Approximations}
-To facilitate an analytical discussion, we simplify the governing
-equation Eq. \ref{iso_ode} with the following approximations:
+\subsection{Approximate governing equation for general perturbations}\label{approx_gov}
+To facilitate a simple discussion, we derive a governing
+equation with the following approximations:
 
@@
   $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
-  \kappa^2$. The problem simplifies because the eigenvalue now only
-  appears in the last term in Eq. \ref{iso_ode}. We call this the
+  \kappa^2$. % The problem simplifies because the eigenvalue now only
+  % appears in the last term in Eq. \ref{iso_ode}.
+  We call this the
   low-frequency approximation. 
 \item We are mainly interested in the effect of vertical shear, i.e. $q\neq
-  0$. This is already captured explicitly in Eq. \ref{iso_ode} through
-  the factor $d\Omega^2/dz$. We therefore simplify further by ignoring
-  the vertical dependence of $\kappa^2$ elsewhere. For clarity, we
-  will replace $\kappa^2$ with $\Omega_k^2$. 
+  0$. This is already captured explicitly in the linearized equations through
+  the factor $d\Omega/dz$ in Eq. \ref{lin_vy}. We therefore simplify
+  further by ignoring the vertical dependence of $\Omega$ and $\kappa$
+  elsewhere. For clarity, we replace $\kappa^2$ with $\Omega_k^2$. 
   % , although the discussion
@@
 Making both of these approximations is equivalent to the replacement
-$D\to\Omega_k^2$.  We will consider sound-speed profiles with $q<0$ in the global 
+$D\to\Omega_k^2$.  After making this replacement, we eliminate $W$ and
+$Q$ between  Eq. \ref{lin_vz} and Eq. \ref{ode_w}---\ref{ode_Q},
+making use of Eq. \ref{vertical_shear}, to
+obtain an equation for $\delta v_z$ when $\Gamma=1$,
+\begin{align}
+  0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
+      q}{\Omega_k^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
+    v_z}{dz} \notag\\
+  &+ \left[\sigma^2\left(\frac{k_x^2}{\Omega_k^2} +
+      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
+        q}{\Omega_k^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
+    &\phantom{+=}\left.-
+    \frac{c_s^2}{\Omega_k^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
+      \frac{\ii k_x q}{r}\right)
+   \left(1-\chi\right) 
+   \right]\delta v_z,\label{vertiso_gov}
+\end{align}
+where
+\begin{align}
+  \chi \equiv \frac{1-\ii\sigma t_c}{1-\ii\sigma t_c\gamma}.
+\end{align}
+
+It is convenient to introduce the dimensionless variables
+\begin{align}
+  \hat{z} = z/\Hiso,\quad \hat{k}=k_x\Hiso, \quad \hat{\sigma} = \sigma/\Omega_k,
+\end{align}
+then Eq. \ref{vertiso_gov} becomes
+\begin{align}
+  0=& \delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon
+    q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime \notag\\
+  &+
+  \left[\hat{\sigma}^2\left(\hat{k}^2+\chi\right) +
+    \left(  \chi + \ii \epsilon q\hat{k}\right)\ln\rho^{\prime\prime}\right.
+    \notag\\
+    &\left.- \ln\rho^{\prime
+      2}\left(\khat^2 -
+      \ii\epsilon
+      q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
+\end{align}
+where $^\prime$ denotes $d/d\zhat$. Eq. \ref{vertiso_gov_nondim} will
+be the basis for our discussion of vertically isothermal disks,
+although further simplifications or maniputations will be performed
+depending the type of perturbations being considered.
+
+In deriving Eq. \ref{vertiso_gov}, the vertical dependence of
+$D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
+$D\to\Omega_k^2$ was made before eliminating variables in favor of
+$\delta v_z$. In Appendix \ref{adia_improve} we show that making this
+replacement after eliminating variables, which requires the vertical
+derivative of $D$, adds unecessary complexity for thin disks.
+
+We will consider sound-speed profiles with $q<0$ in the global 
 disk. Then $\Omega^2(z)$ decreases away from the midplane,
@@
 
-\subsubsection{Neccessity of vertical shear for
-  instability}\label{integral_relation} 
-With the above approximations and introducing the dimensionless variables
+\subsection{Isothermal perturbations}\label{iso_discuss}
+Isothermal perturbations corresponds to $\gamma=1$ for any thermal
+relaxation timescale $\beta$, or for any adiabatc index $\gamma$ when
+$\beta=0$. In this case Eq. \ref{ode_Q} implies 
 \begin{align}
-  \hat{z} = z/\Hiso,\quad \hat{k}=k_x\Hiso, \quad \hat{\sigma} = \sigma/\Omega_k,
+  Q = W, 
 \end{align}
-the governing equation can be re-written as
+ and so 
+ \begin{align}
+   \ii\hat{\sigma}c_s\delta v_z = W^\prime
+   % \quad   \text{(isothermal perturbations)},
+ \end{align}
+from Eq. \ref{lin_vz}. For isothermal perturbations 
+the parameter $\chi=1$ so it is easiest to work with
+$W$ by integrating Eq. \ref{vertiso_gov_nondim}. For later discussion
+it is convenient to write the result as
+
 \begin{align}
@@
 \end{align}
-where the primes denote $d/d\zhat$ and $f(\zhat)$ is defined such that
+where $f(\zhat)$ is defined such that
 \begin{align}\label{fz_shear}
@@
 
+% The linear problem simplifies considerably for the case
+% $\Gamma=\gamma=1$.  Then 
+% \begin{align}
+%   Q=W. 
+% \end{align}
+% Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
+% problem for isothermal perturbations in a vertically isothermal disk, 
+% \begin{align}\label{iso_ode}
+%   \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
+%       r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
+%   \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
+% \end{align} 
+% subject to appropriate boundary conditions. Eq. \ref{iso_ode} also
+% applies to $\gamma>1$ for $t_c\to0$.
+
+Note that for isothermal perturbations the issue of ignoring the
+vertical dependence of $D$ described in \S\ref{approx_gov} and Appendix
+\ref{adia_improve} is irrelevant, since
+Eq. \ref{iso_ode1} may be derived directly by substituting $Q=W$ in
+Eq. \ref{ode_w}, i.e. no derivatives of $D$ needs to be taken.
+
+We remark that Eq. \ref{iso_ode1} is also applicable for  $\gamma=\Gamma\neq 1$ in the
+limit $t_c\to\infty$ (i.e. neutrally stratified adiabatic disks),
+since in that case $Q\simeq W$ as well. However, the following
+discussion is only  valid for $\gamma=\Gamma=1$ because $c_s$ is taken
+to be a constant.
+
+\subsubsection{Neccessity of vertical shear for
+  instability}\label{integral_relation} 
 We can establish a necessary condition for instability by multipling

### 2014-12-29 setup.tex
added beta cooling requirements
replace numerical section with that of full equations
restructing analytic discussion

@@
 \begin{align}
-  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p z}\frac{dK}{dr}.
+  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p z}\frac{dK}{dr}.\label{vertical_shear}
 \end{align}
@@
   \frac{\Omega^2(r,z)}{\Omega_k^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
-    \frac{1}{\sqrt{1+z^2/r^2}}\right),
+    \frac{1}{\sqrt{1+z^2/r^2}}\right),\label{vertiso_eqm}
 \end{align}
@@
 \end{align}
-Eq. \ref{ode_w}---\ref{ode_Q} are convenient for numerical work, 
-but alternative forms are more suitable for analytical discussion, and 
-will be derived later.
-
-\subsection{Boundary conditions}
-In numerical calculations a finite vertical domain $z\in[-1,1]\zmax$
-is adopted. For $\Gamma\neq1$ we take $\zmax$ to be slightly less than
-$H$ to avoid the zero-density surface. At $z=\pm\zmax$ we impose
-\begin{align}
-  \delta v_z = 0,
-\end{align}
-for a solid boundary, or 
-\begin{align}
-  \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0,  
-\end{align}
-for a free surface, where $\bm{\xi}$ is the Lagrangian
-displacement.
+% Eq. \ref{ode_w}---\ref{ode_Q} are convenient for numerical work, 
+% but alternative forms are more suitable for analytical discussion, and 
+% will be derived later.
 

### 2014-12-29 linear_adia.tex
restructing analytic discussion

@@
 
-\subsubsection{Explicit solutions in the thin-disk limit}
+\subsubsection{Explicit solutions in the thin-disk limit}\label{adia_explicit}
 We can construct solutions in the thin-disk limit, in which 
@@
 perturbations with intermediate thermodynamic responses with general
-values of $t_c$.
+values of $t_c$ or $\beta$.
 
-In this section we will pursue explicit solutions in the low-frequency,
-thin-disk, nearly-Keplerian approximations as used in the development
-for adiabatic solutions. The pair of ODEs for $W$ and $Q$
-(Eq. \ref{ode_w}---\ref{ode_Q}) with these approximations and 
-written with dimensionless variables, are
+In this section we will pursue explicit solutions in the thin-disk
+approximation. The governing equation, Eq. \ref{vertiso_gov_nondim}, is
+then 
 \begin{align}
-  &\hat{\sigma}^2\left(Q +  \hat{k}^2W\right) \notag\\ 
-  &= \hat{z} \left(1 +
-    \ii\epsilon q \hat{k} \right)\left[W^\prime - \zhat \left(W -
-      Q\right)\right] - \left[W^\prime - \zhat \left(W -
-      Q\right)\right]^\prime\label{nearly_iso1},\\
-&\beta \hat{\sigma}^2\left(W - \gamma Q\right) + \ii\hat{\sigma}
-    \left(W-Q\right)\notag\\
-   & = -\beta\zhat\left(\gamma-1\right) \left[W^\prime - \zhat \left(W -
-      Q\right)\right]\label{nearly_iso2},
+  0 = \delta v_z^{\prime\prime} - z \widetilde{A}\delta v_z^\prime +
+  (\widetilde{B} - \widetilde{C}\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
 \end{align}
-where we recall $\beta = t_c\Omega_k$ is the dimensionless cooling
-time. It is evident that for $\beta\equiv 0$, we recover
-Eq. \ref{iso_ode3}. In addition, from the expression of the vertical
-velocity perturbation (Eq. \ref{lin_vz}) we have
+with
 \begin{align}
-  W^\prime - \zhat \left(W -
-      Q\right) = \ii\hat{\sigma}c_s\delta v_z. \label{nearly_vz}
+  &\widetilde{A} \equiv 1 + \ii \epsilon q \hat{k},\\
+  &\widetilde{B} \equiv \hat{\sigma}^2\left(\chi + \hat{k}^2\right) -
+  \left(\chi + \ii \epsilon q \hat{k}\right),\\
+  &\widetilde{C} \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
+      \epsilon q\hat{k}\right), 
+\end{align}
+and $\chi =
+\left(1-\ii\hat{\sigma}\beta\right)/\left(1-\ii\hat{\sigma}\beta\gamma\right)
+$ in terms of dimensionless variables.
+
+\subsubsection{The effect of introducing a small but finite
+  thermal relaxation time} 
+It is instructive to first ask  the more analytically tractable
+question: how 
+do the eigenfrequencies and eigenfunctions change when we change
+$t_c$ from zero to a small but finite value? For sufficiently small
+$\beta$, for which the perturbations may be considered 
+nearly-isothermal, we expect the new solutions to only differ slightly
+from the explicit isothermal solutions developed in \S\ref{iso_poly}.
+
+To determine the (assumed) small effect of introducing a short thermal
+relaxation, we linearize Eq. \ref{nearly_iso_explicit} about a known 
+solution for $\beta\equiv0$  ($\chi\equiv 1$). 
+Let 
+\begin{align}\label{nearly_iso_pert}
+  \beta \to 0 + \delta\beta,\, \delta v_z\to \delta v_z+\delta
+  v_{z1},\,\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}, 
+\end{align}
+which implies 
+\begin{align}
+  \chi \to 1 + \delta\chi = 1 + \ii \hat{\sigma}\left(\gamma-1\right)\delta\beta.
+\end{align}
+Here $\delta v_z$ corresponds to the explicit solution developed in
+\S\ref{iso_poly} with  $\hat{\sigma}^2$ given by Eq. \ref{sig2_iso}. Inserting
+Eq. \ref{nearly_iso_pert} into Eq. \ref{nearly_iso_explicit}  and keeping
+only first order terms, we obtain 
+\begin{align}
+ 0 =& \delta v_{z1}^{\prime\prime} - z\left(1 + \ii\epsilon q
+   k\right)\delta v_{z1}^\prime\notag\\
+ &+ \left[2\hat{\sigma}\delta\hat{\sigma}\left(1 + \khat^2\right) +
+   \left(\hat{\sigma}^2 - 1\right)\delta\chi \right.\notag\\
+&\phantom{++} \left. + \zhat^2\left(\khat^2 -
+     \ii\epsilon q \khat\right)\delta\chi\right]\delta v_z\notag\\
+ &+\left[\hat{\sigma}^2\left(1 + \khat^2\right) - \left(1+\ii\epsilon
+     q \khat\right)\right]\delta v_{z1}.\label{nearly_iso4}
 \end{align}
 
-\subsubsection{Perturbative effect of introducing a small but finite
-  thermal relaxation time}
-
-It is instructive to first ask the more analytically tractable question: how
-do the eigenfrequencies and eigenfunctions change when we change
-$t_c$ from zero to a small but finite value? For sufficiently small
-$t_c$, for which the perturbations may be considered
-nearly-isothermal, we expect the new solutions to only differ slighly
-from the explicit isothermal solutions developed in \S\ref{iso_poly}.
-
-%is valid for
-%non-isothermal perturbations $\gamma>1$ if the thermal relaxation
-%(cooling) timescale $t_c\equiv0$.
-
-% Then we may linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2}
-
-To determine the (assumed) small effect of introducing a short thermal
-relaxation, we linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2} about a known
-solution for $\beta\equiv0$. Let 
-\begin{align}\label{nearly_iso_pert}
-  \beta \to 0 + \delta\beta,\, W\to W+\delta W,\, Q \to Q+\delta
-  Q,\,\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}, 
-\end{align} 
-where $W=Q$ is the polynomial solution described in \S\ref{iso_poly}
-with $\hat{\sigma}^2$ given by Eq. \ref{sig2_iso}. Inserting
-Eq. \ref{nearly_iso_pert} into
-Eq. \ref{nearly_iso1}---\ref{nearly_iso2} and keeping only first order
-terms, we obtain
-\begin{align}
-  &2\hat{\sigma}\delta\hat{\sigma}\left(1+\hat{k}^2\right)W +
-  \hat{\sigma}^2 \left(\delta Q + \hat{k}^2\delta W\right)\notag\\
-  &= \zhat \left(1+\ii\epsilon q \hat{k}\right)\left[\delta W^\prime -
-    \zhat \left(\delta W - 
-      \delta Q\right)\right] - \delta W^{\prime\prime}\notag\\
-  &\phantom{=} + \left[\zhat \left(\delta W -
-     \delta Q\right)\right]^\prime,\label{nearly_iso3}\\
-&\ii\hat{\sigma} \left(\delta W - \delta Q\right) =
-\delta\beta\left(\gamma-1\right)\left(\hat{\sigma}^2W - \hat{z}W^\prime\right).\label{nearly_iso4}
-\end{align}
-This is a pair of inhomogeneous ODEs with a forcing by $W$. The
-homogeneous problem is, of course, just the ODE for $\beta=0$ solved
-in \S\ref{iso_poly}. Hence we only seek the particular solution.
-
-We could proceed by considering the entire set of polynomial
-solutions for $W$, but it is simplest to look at the fundamental
-mode. Thus we consider the `basic state' 
+We could proceed by considering the entire set of polynomial solutions
+for $W$, and hence $\delta v_z$, discussed in \S\ref{iso_poly}, but it
+is simplest to look at the fundamental mode. Thus we consider the `basic state' 
  \begin{align*}
-  W = \hat{z},\quad \hat{\sigma}^2\left(1+\hat{k}^2\right) = 1 +
+   \delta v_z = 1,\quad \hat{\sigma}^2\left(1+\hat{k}^2\right) = 1 +
    \ii\epsilon q \hat{k},
  \end{align*}
-then
+ then we may seek 
  \begin{align}
-  \delta W = b \zhat^3,
+   \delta v_{z1} = a \zhat^2 + b,
  \end{align}
-where $b$ is a constant to be determined. Eliminating $\delta Q$ from
-Eq. \ref{nearly_iso3}---\ref{nearly_iso4}, inserting the above
-expressions for $W$ and $\delta W$ and balancing coefficients of
-$\zhat$ and $\zhat^3$, we find
+where $a$, $b$ are constants.
+
+Inserting the above expressions for $\delta v_z$, $\delta v_{z1}$ and
+$\hat{\sigma}^2$ into Eq. \ref{nearly_iso4}, then balancing
+coefficients of $\zhat^2$ and constants, we require
 \begin{align}
- & 2\left(1+\ii\epsilon q \hat{k}\right) \delta\hat{\sigma} \notag\\& = -6b\hat{\sigma} -
-2\ii \delta \beta\left(\gamma-1\right)\left(\hat{\sigma}^2-1\right) - \ii
-  \hat{\sigma}^2\delta\beta\left(\gamma-1\right),\\
-& 0=  2b\hat{\sigma}+\ii\delta\beta
-\left(\gamma-1\right)\left(\hat{\sigma}^2 - 1\right),
+0 &= 2a + 2\hat{\sigma}\left(1+\khat^2\right)\delta\hat{\sigma} +
+\left(\hat{\sigma}^2 -1\right)\delta\chi,\notag\\
+0&= 2a\left(1+\ii\epsilon q \khat\right) - \left(\khat^2 - \ii\epsilon
+q \khat\right)\delta\chi.
 \end{align}
-where the expression for $\hat{\sigma}^2$ was used. Solving for
-$\delta\hat{\sigma}$ gives 
+
+Using the expressions for $\delta\chi$ and $\hat{\sigma}$ we can solve
+for $\delta\hat{\sigma}$. This gives 
 \begin{align}\label{nearly_iso_dsig}
@@
 % VSI.
-The solution $\delta W\propto\zhat^3$ then makes sense because it
+The $\zhat^2$ depedence in $\delta v_{z1}$ then makes sense because it
 is most significant for large $\zhat$, i.e. away from the midplane
@@
 
-\subsubsection{General explicit solutions with thermal relaxation}
-To develop explicit solutions of
-Eq. \ref{nearly_iso1}---\ref{nearly_vz}, we eliminate variables in
-favor of $\delta v_z$ to obtain
-\begin{align}
-  0 = \delta v_z^{\prime\prime} - z \widetilde{A}\delta v_z^\prime +
-  (\widetilde{B} - \widetilde{C}\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
-\end{align}
-with
-\begin{align}
-  &\widetilde{A} \equiv 1 + \ii \epsilon q \hat{k},\\
-  &\widetilde{B} \equiv \hat{\sigma}^2\left(\chi + \hat{k}^2\right) -
-  \left(\chi + \ii \epsilon q \hat{k}\right),\\
-  &\widetilde{C} \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
-      \epsilon q\hat{k}\right),
-\end{align}
-where
-\begin{align}
-  \chi \equiv \frac{1-\ii\hat{\sigma}\beta}{1-\ii\hat{\sigma}\beta\gamma}.
-\end{align}
-
-The governing equation Eq. \ref{nearly_iso_explicit} has the same form
-as in the adiabatic case (Eq. \ref{adia_iso4}), so we can apply the
-same solution method. However, the coefficients
-$\widetilde{B}$ and $\widetilde{C}$ now have a more complicated
-dependence on $\hat{\sigma}$.  The adiabatic dispersion relation,
+\subsubsection{General dispersion relation with thermal relaxation}\label{disp_relax}
+Eq. \ref{nearly_iso_explicit} is same form
+as the governing equation for the adiabatic case
+(Eq. \ref{adia_iso4}). However, the coefficients $\widetilde{B}$ and $\widetilde{C}$
+have a complicated dependence on $\hat{\sigma}$. 
+Applying the same solution method as in \S\ref{adia_explicit}, we 
+% Thus calculating the
+% eigenfrequency $\hat{\sigma}$ generally requires a numerical
+% approach, which we do  in \S\ref{disp_relax}. However, analytical
+% results may be obtained in the limit $\beta\to0$. 
+find the adiabatic dispersion relation,
 Eq. \ref{adia_disp0}, generalizes to a $6^\mathrm{th}$ order
@@
 
-%ate_theory, eps=0.05, smallq=-1d0, gmma=1.4, bmax=1.0, ncases=11, mmode=0.0, root=1, legend=[0.102,0.14,5.4,0.355], yrange=[0,6]
+%ate_theory, eps=0.05, smallq=-1d0, gmma=1.4, bmax=1.0, ncases=11,
+%mmode=0.0, root=1, legend=[0.102,0.14,5.4,0.355], yrange=[0,6]
+
+% The pair of ODEs for $W$ and $Q$
+% (Eq. \ref{ode_w}---\ref{ode_Q}) with these approximations and 
+% written with dimensionless variables, are
+% \begin{align}
+%   &\hat{\sigma}^2\left(Q +  \hat{k}^2W\right) \notag\\ 
+%   &= \hat{z} \left(1 +
+%     \ii\epsilon q \hat{k} \right)\left[W^\prime - \zhat \left(W -
+%       Q\right)\right] - \left[W^\prime - \zhat \left(W -
+%       Q\right)\right]^\prime\label{nearly_iso1},\\
+% &\beta \hat{\sigma}^2\left(W - \gamma Q\right) + \ii\hat{\sigma}
+%     \left(W-Q\right)\notag\\
+%    & = -\beta\zhat\left(\gamma-1\right) \left[W^\prime - \zhat \left(W -
+%       Q\right)\right]\label{nearly_iso2},
+% \end{align}
+% where we recall $\beta = t_c\Omega_k$ is the dimensionless cooling
+% time. It is evident that for $\beta\equiv 0$, we recover
+% Eq. \ref{iso_ode3}. In addition, from the expression of the vertical
+% velocity perturbation (Eq. \ref{lin_vz}) we have
+% \begin{align}
+%   W^\prime - \zhat \left(W -
+%       Q\right) = \ii\hat{\sigma}c_s\delta v_z. \label{nearly_vz}
+% \end{align}
+
+% %is valid for
+% %non-isothermal perturbations $\gamma>1$ if the thermal relaxation
+% %(cooling) timescale $t_c\equiv0$.
+
+% % Then we may linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2}
+
+% To determine the (assumed) small effect of introducing a short thermal
+% relaxation, we linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2} about a known
+% solution for $\beta\equiv0$. Let 
+% \begin{align}\label{nearly_iso_pert}
+%   \beta \to 0 + \delta\beta,\, W\to W+\delta W,\, Q \to Q+\delta
+%   Q,\,\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}, 
+% \end{align} 
+% where $W=Q$ is the polynomial solution described in \S\ref{iso_poly}
+% with $\hat{\sigma}^2$ given by Eq. \ref{sig2_iso}. Inserting
+% Eq. \ref{nearly_iso_pert} into
+% Eq. \ref{nearly_iso1}---\ref{nearly_iso2} and keeping only first order
+% terms, we obtain
+% \begin{align}
+%   &2\hat{\sigma}\delta\hat{\sigma}\left(1+\hat{k}^2\right)W +
+%   \hat{\sigma}^2 \left(\delta Q + \hat{k}^2\delta W\right)\notag\\
+%   &= \zhat \left(1+\ii\epsilon q \hat{k}\right)\left[\delta W^\prime -
+%     \zhat \left(\delta W - 
+%       \delta Q\right)\right] - \delta W^{\prime\prime}\notag\\
+%   &\phantom{=} + \left[\zhat \left(\delta W -
+%      \delta Q\right)\right]^\prime,\label{nearly_iso3}\\
+% &\ii\hat{\sigma} \left(\delta W - \delta Q\right) =
+% \delta\beta\left(\gamma-1\right)\left(\hat{\sigma}^2W - \hat{z}W^\prime\right).\label{nearly_iso4}
+% \end{align}
+% This is a pair of inhomogeneous ODEs with a forcing by $W$. The
+% homogeneous problem is, of course, just the ODE for $\beta=0$ solved
+% in \S\ref{iso_poly}. Hence we only seek the particular solution.
+
+% We could proceed by considering the entire set of polynomial
+% solutions for $W$, but it is simplest to look at the fundamental
+% mode. Thus we consider the `basic state' 
+% \begin{align*}
+%   W = \hat{z},\quad \hat{\sigma}^2\left(1+\hat{k}^2\right) = 1 +
+%   \ii\epsilon q \hat{k},
+% \end{align*}
+% then
+% \begin{align}
+%   \delta W = b \zhat^3,
+% \end{align}
+% where $b$ is a constant to be determined. Eliminating $\delta Q$ from
+% Eq. \ref{nearly_iso3}---\ref{nearly_iso4}, inserting the above
+% expressions for $W$ and $\delta W$ and balancing coefficients of
+% $\zhat$ and $\zhat^3$, we find
+% \begin{align}
+%  & 2\left(1+\ii\epsilon q \hat{k}\right) \delta\hat{\sigma} \notag\\& = -6b\hat{\sigma} -
+% 2\ii \delta \beta\left(\gamma-1\right)\left(\hat{\sigma}^2-1\right) - \ii
+%   \hat{\sigma}^2\delta\beta\left(\gamma-1\right),\\
+% & 0=  2b\hat{\sigma}+\ii\delta\beta
+% \left(\gamma-1\right)\left(\hat{\sigma}^2 - 1\right),
+% \end{align}
+% where the expression for $\hat{\sigma}^2$ was used.

### 2014-12-29 appendix.tex
restructing analytic discussion

@@
 %   disks}\label{adia_improve}
- In \S\ref{analytic_adia} we made the replacement $D\to\Omega_k^2$
- before combining Eq. \ref{adia_iso1}---\ref{adia_iso2} to derive
- Eq. \ref{adia_iso3}.
-
-This procedure ignores the vertical dependence of
- $D=\kappa^2(z) - \sigma^2$.
-
-We show here that this has no significant
+In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
+eliminating variables to obtain a single equation for $\delta v_z$,
+Eq. \ref{vertiso_gov}.  This procedure ignores the vertical dependence of
+$D=\kappa^2(z) - \sigma^2$. We show here that this has no significant
 consequence for thin disks.
@@
 \end{align}   
+Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
+proportional to $\epsilon^2$. For a thin disk, $\epsilon\ll1$, so
+neglecting these terms has no qualitative effect on the discussion in
+\S\ref{analytic_adia} and \S\ref{analytic_relax}.  This issue does not
+arise for isothermal perturbations discussed in \S\ref{iso_discuss}
+since in that case we work with a governing equation for $W$ instead.
 

### 2014-12-29 linear_adia.tex
restructing analytic discussion

@@
 
-\subsection{Perturbations with thermal relaxation}
+\subsection{Perturbations with thermal relaxation}\label{analytic_relax}
 Having examined the extreme limits of isothermal ($t_c\to0$) and
@@
 We could proceed by considering the entire set of polynomial solutions
-for $W$, and hence $\delta v_z$, discussed in \S\ref{iso_poly}, but it
+for $W$, and hence for $\delta v_z$, as derived in \S\ref{iso_poly}, but it
 is simplest to look at the fundamental mode. Thus we consider the `basic state' 
@@
 
-\subsubsection{General dispersion relation with thermal relaxation}\label{disp_relax}
+\subsubsection{Dispersion relation with thermal relaxation}\label{disp_relax}
 Eq. \ref{nearly_iso_explicit} is same form
@@
 Eq. \ref{adia_disp0}, generalizes to a $6^\mathrm{th}$ order
-polynomial in $\hat{\sigma}$:
+polynomial in $\hat{\sigma}$ with finite thermal relaxation:
 \begin{align}
@@
 dimensionless thermal relaxation timescales $\beta$, in a disk with
-$\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.
+$\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. Consistent with the previous
+discussion, for $\beta\to0$ thermal relaxation has no effect as
+$k_x\to0$ or $k_x\to\infty$.
+
+Fig. \ref{relax_disp_fig} shows that a growing fundamental VSI mode
+exists both in the isothermal ($\beta\to0$) and adiabatic
+($\beta\to\infty$) limits, although the latter growth rates are
+significantly smaller and perturbations are restricted to the
+midplane. However, for $0.1 \lesssim \beta\lesssim
+\beta 30$, this mode is stabilized for $k_xH_\mathrm{iso}\gtrsim
+10$. Although results for small $k_x$ violates the low-frequency
+approximation, Fig. \ref{relax_disp_fig} suggests that thermal
+relaxation needs to be significantly shorter than the dynamical time
+in order for the VSI with appreciable growth rates to develop.
 

### 2014-12-29 linear_adia.tex
new fig for analytic discussion with therm relax

@@
   2}$ in the integrand (third term on the RHS) represents
-stabilization by the vertical entropy gradient (real bouyancy frequency). 
+stabilization by the vertical entropy gradient (real buoyancy frequency). 
 It gives an imaginary contribution to $\hat{\sigma}$ when $q\neq0$,
@@
 Unlike isothermal perturbations, Eq. \ref{gam2_max_growth} shows that
-there is a maximum possible growth rate for $M=0$, and growth rates decrease with increasing $M$. This
+maximum growth rate occurs at $M=0$, and growth rates decrease with increasing $M$. This
 is despite the fact that in the thin-disk approximation 
@@
 i.e. $\beta\ll O(\epsilon^2)$, is a sufficient condition for
-bouyancy to be negligible.
+buoyancy to be negligible.
 
@@
 is most significant for large $\zhat$, i.e. away from the midplane
-where the effect of bouyancy first appears as $t_c$ is increased.
+where the effect of buoyancy first appears as $t_c$ is increased.
 
@@
 dimensionless thermal relaxation timescales $\beta$, in a disk with
-$\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. Consistent with the previous
-discussion, for $\beta\to0$ thermal relaxation has no effect as
-$k_x\to0$ or $k_x\to\infty$.
+$\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.  Similarly, 
+in Fig. \ref{relax_bcrit} we solve the dispersion relation for the
+critical $\beta$ such that the growth rate of the fundamental VSI mode
+is zero.
 
-Fig. \ref{relax_disp_fig} shows that a growing fundamental VSI mode
-exists both in the isothermal ($\beta\to0$) and adiabatic
+Fig. \ref{relax_disp_fig}---\ref{relax_bcrit} shows that a growing fundamental
+VSI mode exists both in the isothermal ($\beta\to0$) and adiabatic
 ($\beta\to\infty$) limits, although the latter growth rates are
-significantly smaller and perturbations are restricted to the
-midplane. However, for $0.1 \lesssim \beta\lesssim
-\beta 30$, this mode is stabilized for $k_xH_\mathrm{iso}\gtrsim
-10$. Although results for small $k_x$ violates the low-frequency
-approximation, Fig. \ref{relax_disp_fig} suggests that thermal
-relaxation needs to be significantly shorter than the dynamical time
-in order for the VSI with appreciable growth rates to develop.
+smaller by at least a factor of two. Furthermore, the adiabatic VSI is 
+is limited to a thin layer about the disk midplane so it may be
+considered stable. Then Fig. \ref{relax_bcrit} shows that 
+$\beta\lesssim 0.1$ is required for instability in this disk model. This critical
+timescale is indepedent of $k_x$ for
+$k_xH_\mathrm{iso}\gtrsim10$. Fig. \ref{relax_disp_fig} is also consistent with previous
+discussions which show that thermal relaxation has no effect for
+$k_x\to0$ or $k_x\to\infty$ as $\beta\to0$,
 
@@
 %mmode=0.0, root=1, legend=[0.102,0.14,5.4,0.355], yrange=[0,6]
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/bcrit_theory}
+  \caption{Condition on the thermal relaxation timescale $\beta$ for the existence of unstable fundamental VSI
+    modes in a vertically isothermal disk with $\epsilon=0.05$, $q=-1$
+    and $\gamma=1.4$. For sufficiently small $\beta$ the isothermal VSI opertates,
+    while for sufficiently large $\beta$ the adiabatic VSI
+    operates. For intermediate values of $\beta$ the VSI growth rate
+    is negative (i.e. stabilized). 
+    \label{relax_bcrit}}  
+\end{figure}
 

### 2014-12-29 linear_iso.tex
new fig for analytic discussion with therm relax

@@
   0$. This is already captured explicitly in the linearized equations through
-  the factor $d\Omega/dz$ in Eq. \ref{lin_vy}. We therefore simplify
+  the term $d\Omega/dz$. We therefore simplify
   further by ignoring the vertical dependence of $\Omega$ and $\kappa$
@@
 boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
-\hat{z}$, in which case we can expect growth rates to increase linearly
-with the vertical domain size.   %no faster than linear
+\hat{z}$, in which case we can expect growth rates to increase at most linearly
+with respect to the vertical domain size.   %no faster than linear
 

### 2014-12-29 linear_adia.tex
updated analytical discussion

@@
 timescale is indepedent of $k_x$ for
-$k_xH_\mathrm{iso}\gtrsim10$. Fig. \ref{relax_disp_fig} is also consistent with previous
-discussions which show that thermal relaxation has no effect for
-$k_x\to0$ or $k_x\to\infty$ as $\beta\to0$,
+$k_xH_\mathrm{iso}\gtrsim10$. The top panel of Fig. \ref{relax_disp_fig} is also
+consistent with the previous
+discussion that show thermal relaxation has no effect for
+$k_x\to0$ or $k_x\to\infty$ as $\beta\to0$.
 

### 2014-12-30 linear_adia.tex
added analytical derivation of critical beta critical (initial order of mag est may not be quite right, but no contradictions)

@@
   the presence of thermal relaxation}
+{\bf this section might be wrong, but keep for reference for now}
 We have shown that when $\gamma>1$ so that there is a positive
@@
 
+\subsubsection{Upper limit on the thermal relaxation timescale for the
+  fundamental isothermal VSI mode}
+Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
+fundamental isothermal VSI tends to a constant for large
+$\khat$. Using this clue, we can calculate this critical $\beta$ by
+considering $\khat\to\infty$. We proceed as follows.
+
+At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
+is real. The dispersion relation for the fundamental mode is then
+\begin{align}\label{relax_disp_fund}
+  0 = \sum_{l=1}^{6}c_l\hat{\omega}^l \quad \text{for $M=\hat{\nu}=0$}.  
+\end{align} 
+Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} in
+the limit $\khat\to\infty$ we find
+\begin{align}
+  0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
+  \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
+  &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
+  0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
+  - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3.\label{relax_cond2}
+\end{align}
+We seek solutions with
+\begin{align}
+  \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \khat\hat{\omega} \text{ finite}.
+\end{align}
+Then
+\begin{align}
+  \khat\hat{\omega} = \frac{(\gamma-1)\beta}{\epsilon q}
+\end{align}
+from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
+that
+\begin{align}
+  \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
+        q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2}. 
+\end{align}
+For a thin disk $\epsilon\ll 1$, we infer 
+\begin{align}
+  \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
+  \beta_\mathrm{crit} 
+\end{align}
+is required for the fundamental isothermal VSI to operate.  For
+$\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$ we find
+$\beta_\mathrm{crit}=0.125$, in agreement with Fig. \ref{relax_bcrit}
+at large $\khat$.
+
 % The pair of ODEs for $W$ and $Q$

### 2015-01-21 linear_adia.tex
regular update, added analytical calc of critical timescale for adia vsi

@@
 
-\subsubsection{Qualitative condition for buoyancy to be negligible in
-  the presence of thermal relaxation}
-{\bf this section might be wrong, but keep for reference for now}
-We have shown that when $\gamma>1$ so that there is a positive
-vertical entropy gradient and perturbations evolve
-adiabatically ($t_c\to\infty$) that the VSI is strongly
-surpressed. This is the limit where the term proportional to
-$t_c^{-1}$ in the linearized energy equation (Eq. \ref{lin_energy})
-becomes negligible compared to the term proportional to
-$\gamma/\Gamma-1$. The latter term is associated with the vertical
-entropy gradient, or buoyancy frequency squared, $N_z^2$.
+% \subsubsection{Qualitative condition for buoyancy to be negligible in
+%   the presence of thermal relaxation}
+% {\bf this section might be wrong, but keep for reference for now}
+% We have shown that when $\gamma>1$ so that there is a positive
+% vertical entropy gradient and perturbations evolve
+% adiabatically ($t_c\to\infty$) that the VSI is strongly
+% surpressed. This is the limit where the term proportional to
+% $t_c^{-1}$ in the linearized energy equation (Eq. \ref{lin_energy})
+% becomes negligible compared to the term proportional to
+% $\gamma/\Gamma-1$. The latter term is associated with the vertical
+% entropy gradient, or buoyancy frequency squared, $N_z^2$.
 
-Reversing the argument, we expect to recover the VSI when the bouyancy
-frequency becomes negligible compared to the cooling rate, or 
-\begin{align}
-  t_c \ll \frac{\Omega}{\gamma N_z^2}. 
-\end{align} 
-Parametrizing $t_c = \beta\Omega_k^{-1}$ and inserting the expression
-for $N_z^2$ appropriate for vertically isothermal disks, we expect
-\begin{align}
-  \beta \ll
-  \frac{1}{H^2_\mathrm{iso}\left(\gamma-1\right)\left(\frac{\p\ln\rho}{\p
-          z}\right)^2} 
-\end{align}
-is required to enable the VSI, where we used $\Omega\sim\Omega_k$.  
-This inequality is satisfied at the
-midplane and as $|z|\to\infty$. It can be satisifed at all heights if
-it is satisfied at the height of $\mathrm{max}|\p\ln\rho/\p z|$. 
-Using the full expression for the density field for vertically isothermal
-disks (Eq. \ref{vertiso_eqm}), we expect 
-\begin{align}
-  \beta \ll \frac{27}{4(\gamma-1)}\epsilon^2,
-\end{align}
-i.e. $\beta\ll O(\epsilon^2)$, is a sufficient condition for
-buoyancy to be negligible.
+% Reversing the argument, we expect to recover the VSI when the bouyancy
+% frequency becomes negligible compared to the cooling rate, or 
+% \begin{align}
+%   t_c \ll \frac{\Omega}{\gamma N_z^2}. 
+% \end{align} 
+% Parametrizing $t_c = \beta\Omega_k^{-1}$ and inserting the expression
+% for $N_z^2$ appropriate for vertically isothermal disks, we expect
+% \begin{align}
+%   \beta \ll
+%   \frac{1}{H^2_\mathrm{iso}\left(\gamma-1\right)\left(\frac{\p\ln\rho}{\p
+%           z}\right)^2} 
+% \end{align}
+% is required to enable the VSI, where we used $\Omega\sim\Omega_k$.  
+% This inequality is satisfied at the
+% midplane and as $|z|\to\infty$. It can be satisifed at all heights if
+% it is satisfied at the height of $\mathrm{max}|\p\ln\rho/\p z|$. 
+% Using the full expression for the density field for vertically isothermal
+% disks (Eq. \ref{vertiso_eqm}), we expect 
+% \begin{align}
+%   \beta \ll \frac{27}{4(\gamma-1)}\epsilon^2,
+% \end{align}
+% i.e. $\beta\ll O(\epsilon^2)$, is a sufficient condition for
+% buoyancy to be negligible.
 
@@
     $\beta\in[10^{-3},1]$ (top) and $\beta\in[1,10^3]$ (bottom).
-Other disk disk parameters are
+Other disk parameters are
     $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The growth rates are
@@
 \subsubsection{Upper limit on the thermal relaxation timescale for the
-  fundamental isothermal VSI mode}
+  fundamental isothermal VSI mode}\label{iso_vsi_beta_crit}
 Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
@@
 at large $\khat$.
+
+\subsubsection{Lower limit on the thermal relaxation timescale for the
+  fundamental adiabatic VSI mode}
+{\bf Note: for reference only. Not to be included in paper.} 
+Fig. \ref{relax_bcrit} indicates there exists a lower limit on the
+thermal relaxation time for the system to behave adiabatically and
+support the fundamental adiabatic VSI. This critical
+timescale satisfies Eq. \ref{relax_cond1}---\ref{relax_cond2} with 
+\begin{align}
+  \beta \sim \beta_0\khat^{1/2},\quad \hat{\omega}  \sim
+  \hat{\omega}_0\khat^{-1/2} \text{ as } \khat\to\infty.
+\end{align}
+The constants $\beta_0$ and $\hat{\omega}_0$ are determined by the
+simultaneous equations
+\begin{align}
+  & 0 = \gamma\beta_0^2(1-\gamma) + 2\epsilon q \gamma
+  \beta_0\hat{\omega}_0 - \hat{\omega}_0^2 +
+  \beta_0^2\gamma^2\hat{\omega}_0^4.\\
+  & 0= \beta_0(1-\gamma) + \epsilon q \hat{\omega}_0 - \epsilon q
+  \gamma^2 \beta_0^2\hat{\omega}_0^3. 
+\end{align}
+For $\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$, the appropriate
+solution is $\beta_0 = 10.33$ and $\hat{\omega}_0 = 0.74$.
 

### 2015-01-21 numerical.tex
restructing analytic discussion
regular update, added analytical calc of critical timescale for adia vsi

@@
 nearly-vertically isothermal disk with $\Gamma=1.011$, $p=-1.5$,
-$q=-1$, $\epsilon=0.1$, and the perturbation wavenumber $k_x = 200\pi/r$
-(or $\hat{k}=20\pi$).
+$q=-1$, $\epsilon=0.1$, $\zmax = 10\epsilon r$, and the perturbation
+wavenumber $k_x = 200\pi/r$ (or $\hat{k}=20\pi$).
 
@@
 We have also solved this problem with in the low-frequency 
-approximation by setting $D\to\kappa^2$ in Eq. \ref{iso_ode}. We
+approximation by setting $D\to\kappa^2$ in Eq. \ref{ode_w} (with $Q=W$). We
 obtained very similar results, as expected for modes with
@@
 
-\subsubsection{Other adiabatic indices}
+\subsection{Other adiabatic indices}
 In Fig. \ref{adia_growth_num} we plot the eigenfrequencies of the
-fundamental VSI mode for a range of $\gamma$ values. This plot is to
-be compared with Fig. \ref{adia_growth}.
+fundamental VSI mode for $\gamma\in[1,2]$. To select the appropriate
+eigenvalues we vary the parameters slowly away from a reference solution.
 
@@
     ($\beta=10^9$). The eigenfrequencies are calculated by numerically
-    solving the full linear eigenvalue problem. \label{adia_growth_num}}   
+    solving the full linear eigenvalue problem. This plot is to
+    be compared with Fig. \ref{adia_growth}. \label{adia_growth_num}}   
 \end{figure}
@@
 We now introduce thermal relaxation. Here we consider a slightly
-different disk model with $\epsilon=0.05$ and
-$\gamma=1.4$. This model was used by \citep{nelson13}, where the
-authors find a cooling timescale of $\beta\geq 0.1$ was sufficient to
-stabilize the VSI in non-linear hydrodynamic simulations.
+different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as used by
+\cite{nelson13}. Fig. \ref{relax_growth_num} plots the growth rates
+of the fundamental mode function of $k_x$. %corrugation mode
+% We first obtain the eigenvalues for the
+% fundamental mode for $k_xH_\mathrm{iso}=0$, then vary parameters
+% slowly to obtain eigenvalues for other values of $k_x$ or $\beta$.
 
-% \subsection{Vertically non-isothermal disks}
-% {\bf big gamma larger than 1 but equal to small gamma, with tcool}
+As above, Fig. \ref{relax_growth_num} compares well with
+Fig. \ref{relax_disp_fig} except at $k_xH_\mathrm{iso}\lesssim 1$
+where the low-frequency approximation fails. The non-smoothness in the
+curve for $\beta=0.063$ at $k_xH_\mathrm{iso}\gtrsim50$ was found to be
+associated with perturbations near the vertical boundaries. This is
+likely attributable to the free surface boundary condition 
+(cf. regularity condition at infinity in \S\ref{disp_relax}).
 
-% \subsection{Non-uniform thermal relaxation} 
-% {\bf variable tcool, large at midplane, small in atmosphere. simplest
-%   model is to multiply tcool by background density}
+Fig. \ref{relax_growth_num} confirms our analytical discussion that
+introducing a small but finite cooling time stabilizes the fundamental
+VSI, but there is a preferable wavenumber around $k_xH_\mathrm{iso} =
+O(10)$ that maximizes this effect.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag_bloop2}
+  \caption{Growth rates of the fundamental VSI mode in
+    nearly-vertically isothermal disks ($\Gamma=1.011$) with
+    $\epsilon=0.05$ and $q=-1$, with thermal relaxation timescale
+    $\beta$. The eigenfrequencies are calculated by numerically
+    solving the full linear eigenvalue problem. This plot is to
+    be compared with Fig. \ref{relax_disp_fig}. \label{relax_growth_num}}   
+\end{figure}
+
+We visiualize the effect of thermal relaxation in
+Fig. \ref{relax_eigenW_num} by comparing the eigenfunction $W$ for
+$\beta=0.01$ and $\beta=0.125$, the latter is close to the critical
+value beyond which the isothermal fundamental VSI is stabilized
+(\S\ref{iso_vsi_beta_crit}).
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/eigenvectorW_beta0d01}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0cm]{figures/eigenvectorW_beta0d125}
+  \caption{Numerically-calculated eigenfunction of the fundamental VSI
+    mode in a disk with dimensionless thermal relaxation timescale
+    $\beta \equiv t_c\Omega_k = 0.01$ (top) and $0.125$ (bottom) for
+    perturbations with a radial wavenumber $k_xH_\mathrm{iso}=10$. The
+    real (imaginary) part of $W$ is plotted as the solid (dotted)
+    line. Other disk parameters are
+    $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.
+    \label{relax_eigenW_num}}  
+\end{figure}
+
+\citeauthor{nelson13} find a thermal relaxation timescale of
+$\beta\gtrsim 0.6$ stabilized the VSI for this disk model in non-linear
+hydrodynamic simulations. This is consistent with our numerical
+results.

### 2015-01-21 numerical.tex
regular update

@@
 We now introduce thermal relaxation. Here we consider a slightly
-different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as used by
-\cite{nelson13}. Fig. \ref{relax_growth_num} plots the growth rates
-of the fundamental mode function of $k_x$. %corrugation mode
+different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as
+adopted in \cite{nelson13}. Fig. \ref{relax_growth_num} plots the
+growth rates of the fundamental mode function of $k_x$. %corrugation
+                                %mode 
 % We first obtain the eigenvalues for the
@@
 
-As above, Fig. \ref{relax_growth_num} compares well with
-Fig. \ref{relax_disp_fig} except at $k_xH_\mathrm{iso}\lesssim 1$
+As above, Fig. \ref{relax_growth_num} is largely consistent with 
+Fig. \ref{relax_disp_fig} except for $k_xH_\mathrm{iso}\lesssim 1$
 where the low-frequency approximation fails. The non-smoothness in the
-curve for $\beta=0.063$ at $k_xH_\mathrm{iso}\gtrsim50$ was found to be
-associated with perturbations near the vertical boundaries. This is
-likely attributable to the free surface boundary condition 
-(cf. regularity condition at infinity in \S\ref{disp_relax}).
+curve for $\beta=0.063$ at $k_xH_\mathrm{iso}\sim50$ was found to be
+associated with large perturbations near the vertical boundaries. This is
+likely attributable to the free surface boundary condition imposed in
+the numerical calculation (cf. regularity condition at infinity in
+\S\ref{disp_relax}).
 
@@
 \begin{figure}
+   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_imag_bloop}
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag_bloop2}
@@
     $\beta$. The eigenfrequencies are calculated by numerically
-    solving the full linear eigenvalue problem. This plot is to
-    be compared with Fig. \ref{relax_disp_fig}. \label{relax_growth_num}}   
+    solving the full linear eigenvalue problem. \label{relax_growth_num}}   
 \end{figure}
 
-We visiualize the effect of thermal relaxation in
+We visualize the effect of thermal relaxation in 
 Fig. \ref{relax_eigenW_num} by comparing the eigenfunction $W$ for
-$\beta=0.01$ and $\beta=0.125$, the latter is close to the critical
-value beyond which the isothermal fundamental VSI is stabilized
-(\S\ref{iso_vsi_beta_crit}).
+$\beta=0.01$ and $\beta=0.125$. The latter value is close to the critical
+value beyond which the fundamental isothermal VSI is stabilized
+(\S\ref{iso_vsi_beta_crit}).  Increasing $\beta$ restricts 
+the region in which $W\sim z$ closer to the midplane, with oscillatory
+behavior emerging at the boundaries, where buoyancy first becomes
+important.
 
@@
 
-\citeauthor{nelson13} find a thermal relaxation timescale of
-$\beta\gtrsim 0.6$ stabilized the VSI for this disk model in non-linear
-hydrodynamic simulations. This is consistent with our numerical
-results.
+%\citeauthor{nelson13} find a thermal relaxation timescale of
+% $\beta\gtrsim 0.6$ stabilized the VSI for this disk model in non-linear
+% hydrodynamic simulations. This is consistent with our numerical
+% results.
+
+\subsection{Vertically non-isothermal disks}
+%gamma = Gamma = 1.4, vary bcool
+
+%gamma = 2.0, Gamma = 1.4, vary bcool

### 2015-01-22 application.tex
adding applications

\section{Application to protoplanetary disks}
As an application of our results, we estimate where in a 
protoplanetary disk the VSI may operate.

We consider 
the Minimum Mass Solar Nebula (MMSN) described in \cite{chiang10}

### 2015-01-22 numerical.tex
adding applications

@@
 
-We have also solved this problem with in the low-frequency 
-approximation by setting $D\to\kappa^2$ in Eq. \ref{ode_w} (with $Q=W$). We
-obtained very similar results, as expected for modes with
-$|\sigma/\Omega_k|\ll1$.
+% We have also solved this problem with in the low-frequency 
+% approximation by setting $D\to\kappa^2$ in Eq. \ref{ode_w} (with $Q=W$). We
+% obtained very similar results, as expected for modes with
+% $|\sigma/\Omega_k|\ll1$.
 
@@
 
-We have also solved this problem in the low-frequency approximation
-using Eq. \ref{ode_w}---\ref{ode_Q} with the replacement
-$D\to\kappa^2$. We again obtained very similar results.
+% We have also solved this problem in the low-frequency approximation
+% using Eq. \ref{ode_w}---\ref{ode_Q} with the replacement
+% $D\to\kappa^2$. We again obtained very similar results.
 

### 2015-01-22 application.tex
adding youdin notes

@@
 As an application of our results, we estimate where in a
-protoplanetary disk the VSI may operate.
+protoplanetary disk do the thermodynamic conditions allow the VSI to 
+operate. Specifically, we consider the fundamental isothermal VSI, 
+for which we can use the simple criterion Eq. \ref{iso_vsi_cond}. The
+condition derived below may be considered as a sufficient condition for the VSI.
 
-We consider 
+In our disk models we have adopted a simple thermal relaxation
+function in the energy equation. To connect our model with a realistic
+disk, we consider the more physically-motivated form of the energy
+equation with radiative cooling 
+\begin{align}\label{real_energy}
+\rho T \frac{DS}{Dt} = - \left(\nabla\cdot\bm{F} - Q_+\right), 
+\end{align}
+where the temperature $T$ is defined such that $P=\mathcal{R}\rho
+T/\mu$, where $\mathcal{R}$ is the ideal gas constant and $\mu$ is
+the mean molecular weight. In Eq. \ref{real_energy} $S$ is the
+specific entropy,
+\begin{align}
+  S \equiv c_v\ln{\left(\frac{p}{\rho^{\gamma}}\right)} 
+\end{align}
+where $c_v \equiv \left(\gamma-1\right)^{-1}\mathcal{R}/\mu$ is the specific heat capcity at
+constant volume, $D/Dt$ is the convective derivative, $\bm{F}$ is the
+radiative flux, and $Q_+$ is a heating term such that an equilibrium
+solution exists.
+
+\subsection{Newtonian cooling}
+Our aim is to extract from Eq. \ref{real_energy} the dimensionless
+cooling time $\beta$ as used in our linear calculations. To do so, we
+consider perturbations with small radial lengthscales $L\equiv
+2\pi/k_x<l_\mathrm{rad}$ where 
+\begin{align}
+l_\mathrm{rad} \equiv \frac{1}{\kappa_d\rho} 
+\end{align} 
+is the photon mean-free-path and $\kappa_d$ is the (dust) opacity. For
+$L< l_\mathrm{rad}$ the radiative cooling is given by {\bf (reference?)}
+\begin{align}
+\nabla\cdot\bm{F} = 4 \rho \kappa_d \sigma_s T^4,
+\end{align}   
+where $\sigma_s$ is the Stefan-Boltzmann constant. We choose the
+heating term 
+\begin{align}
+  Q_+ = 4\rho\kappa_d\sigma_s T^{4}_{t=0}. 
+\end{align}
+Then Eq. \ref{real_energy} becomes
+\begin{align}
+  \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} -
+  \frac{4P\kappa_d\sigma_s}{Tc_v}\left(T^4 - T_{t=0}^4\right). 
+\end{align}
+When linearized, the last term on the right hand side becomes
+\begin{align}
+  -\frac{16\sigma_s}{c_v}\kappa_d T^3\left(\delta P -
+    \frac{P}{\rho}\delta\rho\right). 
+\end{align}
+By comparing this expression with the linearized version of the energy
+equation used in our calculations (Eq. \ref{energy_eq}), we identify 
+\begin{align}
+  t_c = \frac{c_v}{16\sigma_s\kappa_dT^3}.
+\end{align}
+For the fundamental isothermal VSI to operate, we need $\beta =
+t_c\Omega_k< \beta_\mathrm{crit}$ as given by
+Eq. \ref{iso_vsi_cond}. This translates to
+\begin{align} 
+  \frac{\mathcal{R}\Omega_k}{16\mu\sigma_s\kappa_dT^3} < \epsilon|q|. 
+\end{align}
+
 the Minimum Mass Solar Nebula (MMSN) described in \cite{chiang10}

### 2015-01-23 appendix.tex
adding applications section

@@
 \end{align}
+
+\section{Fiducial model for a protoplanetary disk}\label{mmsn}
+For results application in \S\ref{application} we use the disk model
+of \cite{chiang10}. This disk model orbits a Solar-mass star and 
+has a surface density distribution
+\begin{align}\label{mmsn_sigma}
+  \Sigma = 2200
+  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-3/2}\mathrm{g}\,\mathrm{cm}^{-2},  
+\end{align}
+a mid-plane density distribution
+\begin{align}
+  \rho_0 = 2.7\times10^{-9}
+  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
+\end{align}
+which implies $p=-39/14$, 
+a temperature profile
+\begin{align}\label{mmsn_temp}
+  T = 120\hat{T}\left(\frac{r}{\mathrm{AU}}\right)^{-3/7} \mathrm{K},
+\end{align}
+which implies $q=-3/7$, 
+and the fiducial disk aspect-ratio 
+\begin{align}
+  \epsilon = 0.022\left(\frac{r}{\mathrm{AU}}\right)^{2/7}.
+\end{align}
+In the above expressions, $\hat{\Sigma}$ and
+$\hat{T}$ are dimensionless coefficients used to scale the model
+relative to the MMSN. Their nomial values are unity. In addition, we
+use the opacity model {\bf(reference?)}
+ \begin{align}
+   \kappa_d &= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2
+   \mathrm{cm}^2\,\mathrm{g}
+    =
+   2.88\hat{\kappa}_d\hat{T}^2\left(\frac{r}{\mathrm{AU}}\right)^{-6/7}\mathrm{cm}^2\,\mathrm{g}^{-1},   
+ \end{align}
+where the second equality follows from the model temperature profile
+above, and $\hat{\kappa}_d$ has similar meaning as $\hat{\Sigma}$ and
+$\hat{T}$.

### 2015-01-23 application.tex
adding applications section

@@
-\section{Application to protoplanetary disks}
+\section{Application to protoplanetary disks}\label{application}
+
+\subsection{Numerical example}
+
+\subsection{Connection with radiative cooling}
 As an application of our results, we estimate where in a 
@@
 
-\subsection{Newtonian cooling}
+\subsubsection{Newtonian cooling}
 Our aim is to extract from Eq. \ref{real_energy} the dimensionless
@@
 Eq. \ref{iso_vsi_cond}. This translates to
-\begin{align} 
+\begin{align}\label{newton_vsi_app} 
   \frac{\mathcal{R}\Omega_k}{16\mu\sigma_s\kappa_dT^3} < \epsilon|q|. 
@@
 
-the Minimum Mass Solar Nebula (MMSN) described in \cite{chiang10}
+We evaluate Eq. \ref{newton_vsi_app} for the Minimum Mass Solar Nebula
+(MMSN) described in \cite{chiang10} and summarized in Appendix
+\ref{mmsn} to find
+\begin{align}
+  \frac{r}{\mathrm{AU}}\lesssim 14.24\left(\mu\hat{\kappa}_d
+  \right)^{14/5}\hat{T}^{14}.
+\end{align}
+For the MMSN with $\mu=2$ this is
+\begin{align*}
+r\lesssim 100\mathrm{AU}. 
+\end{align*}
+
+On the other hand, we have assumed the perturbation lengthscale
+$L<l_\mathrm{rad}$, or
+\begin{align} 
+  \frac{2\pi}{\khat} < \frac{l_\mathrm{rad}}{H_\mathrm{iso}} =
+  \frac{1}{\kappa_d\rho H_\mathrm{iso}} \sim
+  \frac{1}{\kappa_d\Sigma}. 
+\end{align}
+For the fiducial disk model, this becomes 
+% \begin{align}
+%   \khat \gtrsim 3.98\times10^4
+%   \hat{\kappa_d}\hat{\Sigma}\hat{T}^2\left(\frac{r}{\mathrm{AU}}\right)^{-33/14}. 
+% \end{align}
+\begin{align}
+  \left(\frac{r}{\mathrm{AU}}\right) \gtrsim 
+  89.4\left(\hat{\kappa}_d\hat{\Sigma}\right)^{14/33}\hat{T}^{28/33}\khat^{-14/33}.  
+\end{align}
+For the MMSN and $\khat=10$, say, this is
+\begin{align*}
+  r \gtrsim 34 \mathrm{AU}. 
+\end{align*}
+% These rough estimates suggest that the VSI can operate in the outer
+% parts of protoplanetary disks.
+
+\subsubsection{Radiative diffusion}
+For perturbation lengthscales $L>l_\mathrm{rad}$,  
+let us assume that temperature fluctuations are smoothed out by
+radiative diffusion. To model this in linear theory we can redefine
+the cooling time as the operator 
+\begin{align}
+  t_c^{-1} \to -\eta \nabla^2, 
+\end{align}
+where $\eta$ is the diffusion coefficient. This, of course,
+fundamentally changes the linear problem by introducing higher order
+derivatives. We defer a proper analysis of this problem to a follow-up
+study. Here, we are interested in order-of-magnitude estimates. We
+thus proceed by assuming vertical derivatives can be neglected in the
+Laplacian. Then
+\begin{align}
+  t_c^{-1} \sim \eta k_x^2 = \hat{\eta} \Omega_k \khat^2,
+\end{align}
+where $\hat{\eta} = \eta/H_\mathrm{iso}^2\Omega_k$ is the
+dimensionless diffusion coefficient, so that
+\begin{align}
+  \beta = \frac{1}{\hat{\eta}\khat^2}. 
+\end{align}
+
+Our linearized equations remain valid for thermal relaxation by
+radiative diffusion with the above approximation, provided we replace
+the dimensionless cooling timescale $\beta$ by
+$1/\hat{\eta}\khat^2$. Then the criterion for the fundamental
+isothermal VSI, Eq. \ref{iso_vsi_cond} becomes
+\begin{align}\label{iso_vsi_cond_diff}
+  \frac{(\gamma-1)}{\epsilon |q|\khat^2}< \hat{\eta}. 
+\end{align}
+
+Now, the dimensional diffusion cofficient is {\bf(reference needed?)}
+\begin{align}
+  \eta = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 c_v}, 
+\end{align}
+so that
+\begin{align}
+  \hat{\eta} \simeq  \frac{16\sigma_s T^3}{3\kappa_d\Sigma^2
+    c_v\Omega_k}. 
+\end{align}
+The criterion Eq. \ref{iso_vsi_cond_diff} becomes
+\begin{align}
+\frac{3\kappa_d\Sigma^2\mathcal{R}
+  \Omega_k}{16\mu\sigma_s T^3} < \epsilon |q| \khat^2. 
+\end{align}
+For the fiducial disk model this translates to
+\begin{align}
+  \left(\frac{r}{\mathrm{AU}}\right) \gtrsim
+  76.5\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)^{14/57}\mu^{-14/57}\khat^{-28/57}. 
+\end{align} 
+Inserting $\mu=2$ and $\khat=10$ we find, for the MMSN, that
+\begin{align*}
+  r\gtrsim 21 \mathrm{AU}. 
+\end{align*}
+
+We have, however, assumed that perturbation lengthscales
+$L>l_\mathrm{rad}$, which is the opposite limit to the previous
+section. Thus, for the MMSN and $\khat=10$ the above estimate is
+restricted to $r\lesssim 34\mathrm{AU}$. This suggests that the VSI
+can operate between $\sim 20$---$30\mathrm{AU}$ in the MMSN.
+
+Finally, we need to verify the radiative diffusion approximation. This
+requires an optical depth greater than unity, or
+\begin{align}
+  l_\mathrm{rad}< H_\mathrm{iso}. 
+\end{align}
+For the fiducial disk model this is
+\begin{align}
+  \left(\frac{r}{\mathrm{AU}}\right) \lesssim 41
+  \left(\hat{\kappa}_d\hat{T}^2\hat{\Sigma}\right)^{14/33}, 
+\end{align}
+so the above estimate is valid for the MMSN.

### 2015-01-23 application.tex
rough version of applications section done

@@
 \section{Application to protoplanetary disks}\label{application} 
-
-\subsection{Numerical example}
-
-\subsection{Connection with radiative cooling}
 As an application of our results, we estimate where in a 
@@
 
-\subsubsection{Newtonian cooling}
+\subsection{Newtonian cooling}
 Our aim is to extract from Eq. \ref{real_energy} the dimensionless
@@
 equation used in our calculations (Eq. \ref{energy_eq}), we identify 
-\begin{align}
+\begin{align}\label{tc_newton_cool}
   t_c = \frac{c_v}{16\sigma_s\kappa_dT^3}.
@@
 
-\subsubsection{Radiative diffusion}
+\subsubsection{Numerical example 1}
+Consider the MMSN at $r=50\mathrm{AU}$, with $\gamma=1.4$ and
+$\mu=2$. Then we find
+the dimensionless thermal relaxation time, using
+Eq. \ref{tc_newton_cool}, is
+\begin{align*}
+  \beta \simeq 0.057. 
+\end{align*}
+At $r=50\mathrm{AU}$, the disk aspect ratio is $\epsilon \simeq
+0.067$ which implies, using Eq. \ref{iso_vsi_cond}, a critical 
+relaxation timescale of  
+\begin{align*}
+  \beta_\mathrm{crit} \simeq 0.072. 
+\end{align*}
+Thus, the disk supports the fundamental isothermal VSI, but only
+marginally. Using the above value of $\beta$ and the disk parameters
+for the MMSN at $50\mathrm{AU}$ in our linear code we find, for a
+perturbation wavenumber $\khat=10$, a growth rate of
+\begin{align*}
+  \nu = 2.6\times10^{-3}\Omega_k.
+\end{align*}
+This corresponds to a characteristic growth time of $t_g\simeq 60$
+orbits. 
+%only find one unstable mode -> close to marginal
+
+\subsection{Radiative diffusion}
 For perturbation lengthscales $L>l_\mathrm{rad}$,  
@@
 so that
-\begin{align}
+\begin{align}\label{diff_coeff_dimensionless}
   \hat{\eta} \simeq  \frac{16\sigma_s T^3}{3\kappa_d\Sigma^2
@@
 so the above estimate is valid for the MMSN.
+
+\subsubsection{Numerical example 2}
+Consider the MMSN at $r=25\mathrm{AU}$, with $\gamma=1.4$ and $\mu=2$
+as before. Evaluating Eq. \ref{diff_coeff_dimensionless} at
+$25\mathrm{AU}$ for the MMSN, we find
+\begin{align*}
+  \frac{1}{\hat{\eta}} \simeq 1.12. 
+\end{align*}
+At $r=25\mathrm{AU}$ the disk aspect ratio is
+$\epsilon\simeq0.055$. Then Eq. \ref{iso_vsi_cond_diff} becomes
+\begin{align*}
+  \khat^2\gtrsim 15. 
+\end{align*}
+
+Using the above disk parameters with $\beta =
+1/\hat{\eta}\khat^2\simeq0.011$ in our linear code, we find, for
+$\khat=10$ a growth rate of 
+\begin{align*}
+\nu = 8.5\times10^{-3}\Omega_k. 
+\end{align*} 
+This corresponds to a characteristic growth time of $t_g\simeq19$
+orbits. 
+%other modes exist but fundamental isothermal is among the fastest
+%(second harmonic only slightly faster)

### 2015-01-24 application.tex
re-written applications section

--- 2015-01-22 application.tex
@@
-\section{Application to protoplanetary disks}
+\section{Application to protoplanetary disks}\label{application} 
 As an application of our results, we estimate where in a 
@@
 
-\subsection{Newtonian cooling}
-Our aim is to extract from Eq. \ref{real_energy} the dimensionless
-cooling time $\beta$ as used in our linear calculations. To do so, we
-consider perturbations with small radial lengthscales $L\equiv
-2\pi/k_x<l_\mathrm{rad}$ where 
-\begin{align}
+Our aim is to relate the right hand side (RHS) of Eq. \ref{real_energy} to
+the thermal relaxation time $t_c$ as used in our linear
+calculations. To do so, we first consider separately perturbations
+with radial lengthscales $L\equiv 2\pi/k_x\ll l_\mathrm{rad}$ and 
+$L\gg l_\mathrm{rad}$, where      
+\begin{align}\label{lrad}
   l_\mathrm{rad} \equiv \frac{1}{\kappa_d\rho} 
 \end{align} 
-is the photon mean-free-path and $\kappa_d$ is the (dust) opacity. For
-$L< l_\mathrm{rad}$ the radiative cooling is given by {\bf (reference?)}
+is the photon mean-free-path and $\kappa_d$ is the (dust) opacity.
+
+\subsection{Newtonian cooling}
+For $L\ll l_\mathrm{rad}$ the radiative cooling is given by {\bf (reference?)}
 \begin{align}
@@
 \end{align}
-When linearized, the last term on the right hand side becomes
+When linearized, the last term on the RHS becomes
 \begin{align}
@@
 equation used in our calculations (Eq. \ref{energy_eq}), we identify 
+\begin{align}\label{tc_newton_cool} 
+  t_\mathrm{cool} = \frac{c_v}{16\sigma_s\kappa_dT^3}
+\end{align}
+as the thermal relaxation timescale for perturbation lengthscales
+$L\ll l_\mathrm{rad}$.
+
+\subsection{Radiative diffusion}
+For perturbation lengthscales $L\gg l_\mathrm{rad}$, let us assume that
+temperature fluctuations are smoothed out by radiative diffusion. To
+model this in linear theory, we choose the heating term such that
 \begin{align}
-  t_c = \frac{c_v}{16\sigma_s\kappa_dT^3}.
+  Q_+ - \nabla\cdot\bm{F} \equiv \nabla\cdot\left[k_\mathrm{rad}\nabla
+    \left(T-T_{t=0}\right)\right], 
 \end{align} 
-For the fundamental isothermal VSI to operate, we need $\beta =
-t_c\Omega_k< \beta_\mathrm{crit}$ as given by
-Eq. \ref{iso_vsi_cond}. This translates to
+where $k_\mathrm{rad}$ is the conduction coefficient to be defined
+implicitly. Eq. \ref{real_energy} becomes
 \begin{align}
-  \frac{\mathcal{R}\Omega_k}{16\mu\sigma_s\kappa_dT^3} < \epsilon|q|. 
+  \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} + \frac{P}{\rho T
+    c_v}\nabla\cdot\left[k_\mathrm{rad}\nabla\left(T-T_{t=0}\right)\right].     
+\end{align}
+When linearized, the last term on the RHS becomes
+\begin{align}\label{diff_cool_proper}
+  \frac{P}{\rho T c_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
+    T\right). 
+\end{align}
+This, of course, fundamentally changes the linear problem by
+introducing higher order derivatives. We defer a proper analysis of
+this problem to a follow-up study. Here, we are interested in
+order-of-magnitude estimates. We thus proceed by assuming vertical
+derivatives can be neglected in Eq. \ref{diff_cool_proper} and write
+\begin{align}\label{diff_cool_approx}
+  \frac{P}{\rho T c_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
+    T\right) &\to -\frac{k_\mathrm{rad}}{\rho
+    c_v}k_x^2P\left(\frac{\delta T}{T}\right)\notag\\
+  &\equiv -\eta k_x^2 \left(\delta P - \frac{P}{\rho}\delta\rho\right), 
+\end{align}
+%we are interested in radially-thin perturbations  
+where $\eta=k_\mathrm{rad}/\rho c_v$ is the diffusion coefficient and
+is given in terms of physical disk parameters as 
+\begin{align}\label{eta_def}
+  \eta = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 c_v}. 
+\end{align}
+From Eq. \ref{diff_cool_approx} we identify the thermal relaxation
+time for diffusion as 
+\begin{align}\label{tc_diff_cool} 
+  t_\mathrm{diff} = \frac{1}{\eta k_x^2}\equiv \frac{\Omega_k^{-1}}{\hat{\eta}\khat^2}, 
+\end{align}
+where $\hat{\eta} = \eta/H_\mathrm{iso}^2\Omega_k$ is the 
+dimensionless diffusion coefficient defined for convenience, and is
+approximately 
+\begin{align}\label{diff_coeff_dimensionless}
+  \hat{\eta} \simeq  \frac{16\sigma_s T^3}{3\kappa_d\Sigma^2
+    c_v\Omega_k}. 
 \end{align}
 
-the Minimum Mass Solar Nebula (MMSN) described in \cite{chiang10}
+\subsection{Toy model for realistic thermal relaxation}
+To unify $t_\mathrm{cool}$ and $t_\mathrm{diff}$, we define the
+effective thermal relaxation timescale in linear theory as
+\begin{align}\label{tc_def}
+  t_c &\equiv t_\mathrm{cool} + t _\mathrm{diff} =
+  \frac{l_\mathrm{rad}^2}{3\eta} + \frac{1}{\eta k_x^2},  
+\end{align}
+where the relation between $t_\mathrm{cool}$, $l_\mathrm{rad}$ and
+$\eta$ can be inferred from Eq. \ref{lrad}, Eq. \ref{tc_newton_cool}
+and Eq. \ref{eta_def}. Eq. \ref{tc_def} is a simple prescription so
+that for small scales (large $k_x$), $t_c\to t_\mathrm{cool}$, while
+for large scales (small $k_x$), $t_c\to t_\mathrm{diff}$.
+
+The dimensionless thermal relaxation time $\beta$ is then
+\begin{align}\label{real_beta}
+  \beta \equiv t_c\Omega_k =
+  \frac{1}{\hat{\eta}}\left(\frac{l_\mathrm{rad}^2}{3H_\mathrm{iso}^2}
+    + \frac{1}{\khat^2}\right)  \simeq
+  \frac{1}{\hat{\eta}}\left(\frac{1}{3\kappa_d^2\Sigma^2} 
+    + \frac{1}{\khat^2}\right).
+\end{align}
+
+\subsection{Evaluation for a fiducial disk model}
+We evaluate Eq. \ref{real_beta} for the protoplanetary disk model
+described in \cite{chiang10} and summarized in Appendix \ref{mmsn}
+(where an opacity model is also chosen). Defining $\zeta\equiv
+r/\mathrm{AU}$, we find Eq. \ref{real_beta} becomes
+\begin{align}\label{beta_mmsn}
+\beta =
+\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)\zeta^{-57/14}\left(\frac{8.3\times10^{-9}}{\hat{\kappa}_d^2\hat{T}^4\hat{\Sigma}^2}\zeta^{33/7}+\frac{1}{\khat^2}\right).
+\end{align}
+We also evaluate the criterion Eq. \ref{iso_vsi_cond} for the fiducial
+disk model, 
+\begin{align}\label{bcrit_mmsn}
+  \beta_\mathrm{crit} = \frac{9.4\times10^{-3}}{(\gamma
+    -1)}\zeta^{2/7},
+\end{align}
+and we recall $\beta < \beta_\mathrm{crit}$ permits the fundamental 
+isothermal VSI to operate.
+
+In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn} and
+Eq. \ref{bcrit_mmsn} for the MMSN with $\mu = 2$ and
+$\gamma=1.4$. Although it is possible to satisfy
+$\beta<\beta_\mathrm{crit}$ at small radii by choosing a sufficiently
+large $\khat$, our linear calculations suggest growth rates eventually
+decay as $\khat\to\infty$. We thus restrict our attention to
+$\khat\lesssim 10^2$.
+
+Fig. \ref{mmsn_bcrit_bcool} suggests that the fundamental isothermal
+VSI can operate in the outer parts of the MMSN at radii on the order
+of tens of AU. However, the thermal relaxation time $\beta$ is not
+much shorter than $\beta_\mathrm{crit}$, suggesting the VSI may be
+weak.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/bcrit_mmsn}  
+  \caption{Dimensionless thermal relaxation timescales $\beta$ in the MMSN
+    using Eq. \ref{beta_mmsn} for three values of the 
+    perturbation radial wavenumber: $\khat=1$ (dotted), $\khat=10$
+    (dashed) and $\khat=100$ (dashed-dot). The relaxation timescale
+    $\beta_\mathrm{crit}$, below which the fundamental isothermal VSI
+    operates, is shown as the solid line. 
+    \label{mmsn_bcrit_bcool}}   
+\end{figure}
+
+\subsection{Numerical example}
+Consider perturbations with $\khat=10$ in the 
+the MMSN at $r=30\mathrm{AU}$, with $\gamma=1.4$ and
+$\mu=2$. Then we find
+the dimensionless thermal relaxation time, using
+Eq. \ref{tc_newton_cool}, is
+\begin{align*}
+  \beta \simeq 0.046. 
+\end{align*}
+At $r=30\mathrm{AU}$, the disk aspect ratio is $\epsilon \simeq
+0.058$ which implies, using Eq. \ref{iso_vsi_cond}, a critical 
+relaxation timescale of  
+\begin{align*}
+  \beta_\mathrm{crit} \simeq 0.062. 
+\end{align*}
+Thus, the disk supports the fundamental isothermal VSI, but only
+marginally. Using the above value of $\beta$, MMSN disk parameters
+and $\khat=10$ in our linear code, we find a growth rate of
+\begin{align*}
+  \nu = 2.7\times10^{-3}\Omega_k.
+\end{align*}
+This corresponds to a characteristic growth time of $t_g\simeq 60$
+orbits.
+
+ %only find one unstable mode -> close to marginal
+
+ %larger kx

### 2015-01-24 application.tex
regular update, code updates

@@
 of tens of AU. However, the thermal relaxation time $\beta$ is not
-much shorter than $\beta_\mathrm{crit}$, suggesting the VSI may be
+much shorter than $\beta_\mathrm{crit}$, so the VSI may be
 weak.
@@
 \subsection{Numerical example}
-Consider perturbations with $\khat=10$ in the 
-the MMSN at $r=30\mathrm{AU}$, with $\gamma=1.4$ and
-$\mu=2$. Then we find
-the dimensionless thermal relaxation time, using
-Eq. \ref{tc_newton_cool}, is
-\begin{align*}
-  \beta \simeq 0.046. 
-\end{align*}
-At $r=30\mathrm{AU}$, the disk aspect ratio is $\epsilon \simeq
-0.058$ which implies, using Eq. \ref{iso_vsi_cond}, a critical 
+Consider  the MMSN at $r=30\mathrm{AU}$, with $\gamma=1.4$ and
+$\mu=2$. Eq. \ref{mmsn_epsilon} gives $\epsilon \simeq
+0.058$ which implies, using Eq. \ref{bcrit_mmsn}, a critical 
 relaxation timescale of  
@@
 \end{align*} 
-Thus, the disk supports the fundamental isothermal VSI, but only
-marginally. Using the above value of $\beta$, MMSN disk parameters
-and $\khat=10$ in our linear code, we find a growth rate of
+To estimate a thermal relaxation timescale, we set $\khat=20$ 
+in Eq. \ref{beta_mmsn} to obtain 
 \begin{align*}
-  \nu = 2.7\times10^{-3}\Omega_k.
+  \beta \simeq 0.042
 \end{align*}
-This corresponds to a characteristic growth time of $t_g\simeq 60$
-orbits.
+at $r=30\mathrm{AU}$. This indicates the disk supports the fundamental
+isothermal VSI, but only marginally. Using the above value of $\beta$,
+MMSN disk parameters and $\khat=20$ in our linear code, we find a
+growth rate of 
+\begin{align*}
+  \nu = 3.3\times10^{-3}\Omega_k.
+\end{align*}
+This corresponds to a characteristic growth time of $t_g\sim 50$
+orbits. In fact, this mode was found to be the most unstable for
+$\khat=20$. 
+% and
+% physically acceptable -> other modes
+% are crazy
+
+We also solved the linear problem with varying $\khat$  
+and $\beta$ determined by Eq. \ref{beta_mmsn}. We find the most
+unstable wavenumber occurs at $\khat\simeq60$ with $\nu\simeq
+3.6\times10^{-3}\Omega_k$, i.e. only slightly faster than that above. 
+However, the eigenfunctions for large $\khat$ were found to have large
+amplitude oscillations near the vertical boundaries.
 
  %only find one unstable mode -> close to marginal 
-
  %larger kx

### 2015-01-24 codes/cool_full/bloop/vsi.f90

@@ -1,7 +1,7 @@
 module global
   character*6 :: vbc
-  logical :: gterms, eigen_trial, eigenv_out
-  integer :: nz, nzeff, bignz, nkx, nb
+  logical :: gterms, eigen_trial, eigenv_out, eigen_refine
+  integer :: nz, nzeff, bignz, nkx, nb, ntrials, output_freq
   real*8, parameter :: pi = 2d0*acos(0d0)
   real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls
   real*8 :: kxmin, kxmax, dlogkx, bmin, bmax, db
@@ -28,10 +28,10 @@ program vsi
        L1bar(:,:), L2bar(:,:), L3bar(:,:), L4bar(:,:), L5bar(:,:), &
        L6bar(:,:)
   complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
-       vr(:,:), rhs(:,:), wtrial(:)
+       vr(:,:), rhs(:,:), wtrial(:), wtrial_tot(:)
   real*8, external :: dlogrho_dz, domega2_dz, omega2_z, logrho_z, kappa2_z, d2logrho_dz2, csq_z 
   namelist /params/ smallq, smallp, eps, gmma, bgmma, zmax, vbc, gterms
-  namelist /loop/ kxmin, kxmax, nkx, bmin, bmax, nb, eigen_trial, eigenv_out
+  namelist /loop/ kxmin, kxmax, nkx, bmin, bmax, nb, eigen_trial, eigenv_out, eigen_refine
   namelist /grid/ nz
   
   !read input parameters.
@@ -69,11 +69,22 @@ program vsi
   nzeff = nz
   bignz = 5*nzeff
 
+  if(eigen_refine.eq..false.) then
+  ntrials=nkx
+  output_freq = 1
+  else
+  ntrials     = nkx*nb
+  output_freq = 1
+  endif
+
+  
+
   !allocate grids
   allocate(kaxis(nkx))
   allocate(baxis(nb))
   allocate(wtrial(nkx))
-
+  allocate(wtrial_tot(ntrials))
+  
   allocate(zaxis(nz))
   allocate(logrho(nz))
   allocate(dlogrho(nz))
@@ -132,8 +143,10 @@ program vsi
   
   !setup kaxis. input is kxHiso. then output
[…]

### 2015-01-24 codes/cool_full/bloop_mmsn/vsi.f90

@@ -1,12 +1,12 @@
 module global
   character*6 :: vbc
   logical :: gterms, eigen_trial, eigenv_out, eigen_refine
-  integer :: nz, nzeff, bignz, nkx, ng, ntrials, output_freq
-  real*8, parameter :: pi = 2d0*acos(0d0)
-  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls
-  real*8 :: kxmin, kxmax, dlogkx, gmin, gmax, dg
+  integer :: nz, nzeff, bignz, nkx, nb, ntrials, output_freq
+  real*8, parameter :: pi = 2d0*acos(0d0), mu=2d0 !fix mean molecular weight 
+  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls, dist
+  real*8 :: kxmin, kxmax, dlogkx, bmin, bmax, db
   real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:)
-  real*8, allocatable :: kaxis(:), gaxis(:)
+  real*8, allocatable :: kaxis(:), baxis(:)
   complex*16, parameter :: ii = (0d0,1d0)
   complex*16, allocatable :: bigW(:), dbigW(:), bigQ(:), vx(:), vy(:), vz(:), dvz(:) 
 end module global
@@ -30,8 +30,8 @@ program vsi
   complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
        vr(:,:), rhs(:,:), wtrial(:), wtrial_tot(:)
   real*8, external :: dlogrho_dz, domega2_dz, omega2_z, logrho_z, kappa2_z, d2logrho_dz2, csq_z 
-  namelist /params/ smallq, smallp, eps, bcool, bgmma, zmax, vbc, gterms
-  namelist /loop/ kxmin, kxmax, nkx, gmax, ng, eigen_trial, eigenv_out, eigen_refine
+  namelist /params/ smallq, smallp, dist, gmma, bgmma, zmax, vbc, gterms
+  namelist /loop/ kxmin, kxmax, nkx, bmin, bmax, nb, eigen_trial, eigenv_out, eigen_refine
   namelist /grid/ nz
   
   !read input parameters.
@@ -40,6 +40,8 @@ program vsi
   read(7, nml=loop)
   read(7, nml=grid)
   close(7)
+
[…]

### 2015-01-24 codes/cool_full/vsi.f90

@@ -283,7 +283,8 @@ program vsi
   do i=1, bignz
      !discard modes with eigenvalues too large (approx eqns assume |sigma^2|<<kappa2
      !discard modes that grow too slowly
-     if((abs(w(i)).le.1d0/eps).and.(growth(i).ge.eps**2d0)) then
+!     if((abs(w(i)).le.1d0/eps).and.(growth(i).ge.eps**3d0)) then
+      if(growth(i).gt.0d0) then
         eigen_test(i) = 1 
         write(20,fmt='(2(e22.15,x))'), freq(i), growth(i)
         bigW = matmul(T,vr(1:nzeff,i))

### 2015-01-25 application.tex
almost done with numerical results section

@@
 \begin{align*}
-  \beta \simeq 0.042
+  \beta \simeq 0.042.
 \end{align*}
-at $r=30\mathrm{AU}$. This indicates the disk supports the fundamental
-isothermal VSI, but only marginally. Using the above value of $\beta$,
-MMSN disk parameters and $\khat=20$ in our linear code, we find a
+(In fact, Fig. \ref{mmsn_bcrit_bcool} shows that $\beta$  is not
+sensitive to $\khat$ at $r=30\mathrm{AU}$, provided that $\khat\gg1$.)
+The disk supports the fundamental isothermal VSI, but only
+marginally. Using the above value of $\beta$, MMSN disk parameters and
+$\khat=20$ in our linear code, we find a  
 growth rate of 
@@
 This corresponds to a characteristic growth time of $t_g\sim 50$
-orbits. In fact, this mode was found to be the most unstable for
+orbits. This mode was found to be the most unstable for
 $\khat=20$. 
@@
 unstable wavenumber occurs at $\khat\simeq60$ with $\nu\simeq
-3.6\times10^{-3}\Omega_k$, i.e. only slightly faster than that above. 
+3.6\times10^{-3}\Omega_k$, i.e. only slightly faster than $\khat=20$.   
 However, the eigenfunctions for large $\khat$ were found to have large
-amplitude oscillations near the vertical boundaries.
+amplitude oscillations near the vertical boundaries. This is shown in
+Fig. \ref{mmsn_eigenW}. These modes may not be important since they
+mostly affect the disk atmosphere where there is little mass.
 
- %only find one unstable mode -> close to marginal 
- %larger kx
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/eigenvectorW_mmsnkx20.ps}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0cm]{figures/eigenvectorW_mmsnkx60.ps}
+  \caption{Most unstable VSI modes in the MMSN at
+    $r=30\mathrm{AU}$. The eigenfunction $W$ for $\khat=20$ (top) and
+    $\khat=60$ (bottom) are shown. The real (imaginary) part of $W$
+    is plotted as the solid (dotted) lines.\label{mmsn_eigenW}}  
+\end{figure}
+
+% only find one unstable mode -> close to marginal 
+% larger kx

### 2015-01-25 numerical.tex
almost done with numerical results section

@@
 \subsection{Vertically non-isothermal disks}
-%gamma = Gamma = 1.4, vary bcool
+We briefly consider vertically non-isothermal disks with
+$\Gamma=1.4$. Other disk parameters are $p=0$, $q=-1.0$ and $\epsilon
+= 0.05$. These are again chosen for comparison with \cite{nelson13}. 
+In this case we set $\zmax\simeq2.2\epsilon r$, which is close to the
+zero-density surface.
+
+\subsubsection{$N_z^2\equiv 0$}
+
+Consider first the case $\gamma=1.4$ so that there is no vertical
+entropy gradient. We plot in Fig. the growth rates for the
+fundamental VSI for $\beta\in[10^{-2},10^{2}]$. We find growth rates
+do not reach zero, implying the VSI operates regardless of the thermal
+relaxation timescale if $\gamma=\Gamma$. This result is consistent
+with \cite{nelson13}.
+
+We also note that growth rates for $\beta = 10^{-2}$ and
+$\beta=10^{2}$ behave similarly. This is not surprising because
+the linearized energy equation (Eq. \ref{lin_energy}) implies that 
+\begin{align}
+  W = \frac{\left(\Gamma^{-1} -
+      \ii\hat{\sigma}\beta\right)}{\left(1-\ii\hat{\sigma}\beta\right)}
+  Q \text{ for } \gamma\equiv \Gamma.   
+\end{align}
+Thus, $W \to Q/\Gamma$ as $\beta\to0$ and $W\to Q$ as
+$\beta\to\infty$. Since $\Gamma=O(1)$, these two limits are expected
+to produce qualitatively similar results.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/growth_vnoniso1} 
+  \caption{Numerically-determined growth rates of the fundamental VSI
+    mode in neutrally-buoyant, vertically non-isothermal disks
+    ($\Gamma=\gamma=1.4$), with $p=0$, $q=-1$ and  $\epsilon=0.05$, for
+    a range of thermal relaxation timescales
+    $\beta$. \label{growth_vnoniso1}}     
+\end{figure}
+
+Interestingly, Fig. \ref{growth_vnoniso1} show that growth rates
+maximize for $\beta=O(1)$. However, all growth rates are similar to
+order-of-magnitude for $0.1\lesssim\khat\lesssim 100$.
+
+\subsubsection{$N_z^2\geq0$}
 
 %gamma = 2.0, Gamma = 1.4, vary bcool
+
+Finally, we introduce buoyancy in a vertically non-isothermal disk by
+setting $\gamma=2$. This situation was not simulated in 
+\cite{nelson13}, but is similar in spirit with vertically isothermal 
+disks evolved with $\gamma>1$, as considered in the previous section.
+
+Fig. plots the growth rates for this case. The trend is
+qualitatively similar to that for vertically isothermal disks with
+$N_z^2\geq 0$ (Fig. \ref{relax_growth_num}), namely growth rates reach
+zero at finite $\khat$.

### 2015-01-25 numerical.tex
going through numerical section, up to sec 5.3 now

@@
 $\kappa$, so that we may consider vertical domains up to $z\sim r$.
-
-% \subsection{Boundary conditions}
-% In numerical calculations a finite vertical domain $z\in[-1,1]\zmax$
-% is adopted. For $\Gamma\neq1$ we take $\zmax$ to be slightly less than
-% $H$ to avoid the zero-density surface. At $z=\pm\zmax$ we impose
-% \begin{align}
-%   \delta v_z = 0,
-% \end{align}
-% for a solid boundary, or 
-% \begin{align}
-%   \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0,  
-% \end{align}
-% for a free surface, where $\bm{\xi}$ is the Lagrangian
-% displacement.
 
@@
 
-Unless otherwise stated, our fiducial disk model is a
-nearly-vertically isothermal disk with $\Gamma=1.011$, $p=-1.5$,
-$q=-1$, $\epsilon=0.1$, $\zmax = 10\epsilon r$, and the perturbation
-wavenumber $k_x = 200\pi/r$ (or $\hat{k}=20\pi$).
+Unless otherwise stated, our disk model is a nearly-vertically
+isothermal disk with $\Gamma=1.011$, $p=-1.5$, $q=-1$,
+$\epsilon=0.1$.
+
+%mention mode selection -> focus on fundamental corrugation ->
+%simulations eventually dominated by them 
+%choice of kx is O(10), since nelson observe kxH \sim 30
 
 \subsection{VSI in a neutrally stratified disk}\label{vertiso_pertiso} 
-We first test our numerical code, as well as the discussion in
-\S\ref{iso_discuss}, by considering nearly-isothermal
-perturbations. We set $\gamma=1.011=\Gamma$, and effectively switch
-off thermal relaxation by choosing $\beta=10^9$. The vertical domain is
-$\zmax = 10\epsilon r$. This setup is very similar to that adopted in
-\cite{mcnally14}, who calculated the VSI in vertically isothermal
-disks subject to isothermal perturbations ($\Gamma=\gamma=1$).
+We first calculate the VSI in a neutrally straitified disk ($N_z^2\equiv0$) by setting
+$\gamma=\Gamma$, and effectively switch off thermal relaxation by
+choosing $\beta=10^9$. We set the vertical domain to $\zmax =
+10\epsilon r$.  Here, we consider a perturbation wavenumber
+$\khat=20\pi$ or $k_x = 200\pi/r$. This setup is very similar to that
+adopted in \cite{mcnally14}, who calculated the VSI in vertically
+isothermal disks subject to isothermal perturbations
+($\Gamma=\gamma=1$).  This allows us to test our numerical code.
 
 Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
-\ii\nu$. This plot is effectively identical to that in
-\cite{mcnally14}. The fundamental VSI mode, with one node in $W$, has
+\ii\nu$. The fundamental VSI mode, with one node in $W$, has
 the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}. 
-Its expected growth rate, calculated from Eq. \ref{simple_growth} with  
-$L=1$, and the numerically-obtained value are 
+The eigenfunction $W$ is qualitatively proportional to 
+$z$ (especially for $z\to0$),  as expected for the fundamental mode
+according to explicit solutions developed in \S\ref{iso_poly}.    
+These plots are consistent with those in \cite{mcnally14}.
+
+The expected growth rate of the fundamental VSI, calculated from
+Eq. \ref{simple_growth} with $L=1$, and the numerically-obtained value
+are  
 \begin{align*}
@@
 Eq. \ref{integral_relation1}, involve integrands with $\rho$ as the 
-weight function, which rapidly decays away from the midplane. The
-eigenfunction $W$ is qualitatively proportional to 
-$z$, as expected for the fundamental mode according to explicit
-solutions discussed in \S\ref{iso_poly}.
+weight function, which rapidly decays away from the midplane.
 
@@
   \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
-  \caption{Eigenvalues of the
-    vertical shear instability in a neutrally-stratified,
+  \caption{Eigenvalues of the VSI in a neutrally-stratified,
     nearly-vertically isothermal disk  ($\gamma=\Gamma=1.011$). The
-    disk parameters are $q=-1$,  
-    $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
-    wavenumber is $k_x=200\pi/r$. This is the set up considered in
-    \cite{mcnally14}. \label{lowfreq_eigen}
+    disk parameters are $q=-1$, $p=-1.5$, $\epsilon=0.1$, and the perturbation radial
+    wavenumber is $\khat=20\pi$. \label{lowfreq_eigen}
   }
@@
     corresponding to the bottom-left eigenvalue displayed in
-    Fig. \ref{lowfreq_eigen} (smallest $|\sigma|$). The
-    solid and dashed lines are $\real W$ and $\imag W$, respectively. 
+    Fig. \ref{lowfreq_eigen} (smallest $|\sigma|$). The real
+    (imaginary) parts of $W$ are plotted as solid (dotted) lines. 
     \label{lowfreq_eigenfunc}
@@
 
-% We have also solved this problem with in the low-frequency 
-% approximation by setting $D\to\kappa^2$ in Eq. \ref{ode_w} (with $Q=W$). We
-% obtained very similar results, as expected for modes with
-% $|\sigma/\Omega_k|\ll1$.
-
-\subsection{Stabilization of the VSI by a positive vertical entropy
-  gradient}
+\subsection{VSI in a stably stratified disk}
 We demonstrate the (strong) stabilizing effect of a  
-positive vertical entropy gradient by setting $\gamma=2$.  
+positive vertical entropy gradient ($N_z^2\geq0$) by setting $\gamma=2$.  
 Since we expect the perturbations to decay rapidly
@@
 
-% We use the same pseudo-spectral method as before. Here we set
-% $\zmax = 2\epsilon r$ and impose a free surface. This boundary
-% condition was found to be most convenient for numerical
-% implementation. Since we expect the perturbations to decay rapidly
-% away from the midplane (see \S\ref{analytic_adia}), we do not expect 
-% boundary conditions to play an important role, provided that $\zmax$ is
-% sufficiently large.
-
 Fig. \ref{lowfreq_eigen_adia} show the eigenvalues for this
-problem. The growth rates are about an order of magnitude smaller than
-the previous case. The fundamental VSI mode is that with
-$\mathrm{max}|\nu|$. Its expected and numerically-calculated growth
+problem. The growth rates are much smaller than
+those in the previous section. The fundamental VSI mode is that with
+$\mathrm{max}\nu$. Its expected and numerically-calculated growth
 rates compares well, with 
@@
 and clearly show that perturbations are rapidly stabilized away from
-the midplane. For the adopted disk and perturbation parameters 
-($|q|=1$ and $\hat{k} = 20\pi$), Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$,
-corresponding to a characteristic decay length scale of $\simeq
-0.15\epsilon r$, which is consistent with that observed in 
-Fig. \ref{lowfreq_eigenfunc_adia}.
+the midplane. For the adopted disk and perturbation parameters,
+Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$, corresponding
+to a characteristic decay length scale of $\sqrt{2/|\alpha|}\epsilon r\simeq 0.2\epsilon r$,
+which is consistent with that observed in  Fig. \ref{lowfreq_eigenfunc_adia}.
 
-Coincidentally, this decay lengthscale is rougly the height within which
+This decay lengthscale is rougly the height within which
 vertical shear is larger than the buoyancy frequency squared. That is,
@@
 \end{align*}
-in a thin, vertically isothermal disk. This evaluates to $z\sim
+in a thin, vertically isothermal disk. This evaluates to $|z|\lesssim
 0.2\epsilon r$ for the current disk model, consistent with
@@
 \end{align}
-or $\nu< 0.1\epsilon\Omega_k$ is expected in our disk model,
+or $\nu< 0.1\epsilon\Omega_k$ in our disk model,
 consistent with numerical results.
@@
   \caption{Eigenvalues in for the
-    vertical shear instability in a nearly-vertically isothermal disk
-    ($\Gamma=1.011$) evolved adiabatically with $\gamma=2$ and
-    $\beta=10^9$. The disk
+    VSI in a nearly vertically isothermal disk
+    ($\Gamma=1.011$) evolved adiabatically with $\gamma=2$. The disk
     parameters are $q=-1$, 
-    $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
-    wavenumber is $k_x=200\pi/r$. \label{lowfreq_eigen_adia}
+    $p=-1.5$, $\epsilon=0.1$, and the perturbation radial
+    wavenumber is $\khat=20\pi$. \label{lowfreq_eigen_adia}
   }
@@
   \includegraphics[width=\linewidth]{figures/eigenvectorvz_adia}
-  \caption{Fundamental VSI mode in a nearly-vertically
+  \caption{Fundamental VSI mode in a nearly vertically
     isothermal disk ($\Gamma=1.011$) evolved adiabatically
-    ($\gamma=2$, $\beta=10^9$). This eigenfunction 
+    with $\gamma=2$. This eigenfunction 
     corresponds to the top-left eigenvalue displayed in 
     Fig. \ref{lowfreq_eigen_adia} (largest $|\nu|$).  
-    The solid and dashed lines are $\real \delta v_z$ and $\imag
-    \delta v_z$, respectively.  
+    The  real (imaginary) part of $\delta v_z$ are shown as solid
+    (dotted) lines. 
     \label{lowfreq_eigenfunc_adia}
@@
 
-% We have also solved this problem in the low-frequency approximation
-% using Eq. \ref{ode_w}---\ref{ode_Q} with the replacement
-% $D\to\kappa^2$. We again obtained very similar results.
+Although a positive vertical entropy gradient is stabilizing, we note
+that $N_z^2\propto z^2$  while vertical shear $rd\Omega^2/dz\propto
+z$, so that there is always a region sufficiently close to the
+midplane in which vertical shear dominates over bouyancy and the VSI
+can operate. However, the growth rates are small and the instability
+only affects the regions close the midplane. These modes may not be
+important in practice.
 

### 2015-01-25 setup.tex
adding youdin notes
almost done with numerical results section
going through numerical section, up to sec 5.3 now

@@
 pressure and rotation profiles $\rho(r,z)$, $P(r,z)$ and
-$\Omega(r,z)$, respectively.
+$\Omega(r,z)$, respectively. The equilibrium disk satisfies
+\begin{align}
+  0 &= \frac{1}{\rho}\frac{\p P}{\p z} + \frac{\p \Phi_*}{\p z},\label{vert_equilibrium}\\
+  r\Omega^2&= \frac{1}{\rho}\frac{\p P}{\p r} + \frac{\p \Phi_*}{\p
+    r},\label{hori_equilibrium} 
+\end{align} 
+where $\Phi_* = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational
+potential of the central star.
 
@@
 \begin{align}
-  \epsilon(r) \equiv \frac{c_{s0}(r)}{v_k(r)}
+  \epsilon(r) \equiv \frac{c_{s0}(r)}{v_k(r)} \equiv
+  \frac{H_\mathrm{iso}}{r} 
 \end{align}
@@
 $v_k=r\Omega_k=\sqrt{GM_*/r}$ is the Keplerian azimuthal velocity and
-$G$ is the gravitational constant. 
+$G$ is the gravitational constant, and $H_\mathrm{iso}$ is a
+characteristic scale-height.  Thus, $\epsilon$ is also the
+characteristic disk aspect-ratio.  
+% From $\epsilon$ we can define a
+% characteristic scale-height  
+% \begin{align}
+%   \Hiso \equiv \epsilon r.% \quad\quad(\Gamma \equiv 1),   
+% \end{align} 
 The global sound-speed $c_s(r,z)$ is given by
@@
 disks with $\Gamma\leq \gamma$ so that $N_z^2\geq0$.
-
-The equilibrium disk satisfies
-\begin{align}
-  0 &= \frac{1}{\rho}\frac{\p P}{\p z} + \frac{\p \Phi_*}{\p z},\label{vert_equilibrium}\\
-  r\Omega^2&= \frac{1}{\rho}\frac{\p P}{\p r} + \frac{\p \Phi_*}{\p
-    r},\label{hori_equilibrium} 
-\end{align} 
-where $\Phi_* = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational
-potential of the central star. Note that the disk is in general
-baroclinic with $\Omega = \Omega(r,z)$, since
-Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} imply, with
-Eq. \ref{eqm_press},  
-\begin{align}
-  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p z}\frac{dK}{dr}.\label{vertical_shear}
-\end{align}
 
@@
 \end{align}
-
 Similarly, the equilibrium rotation is 
@@
   \kappa^2(r,z) \equiv 2\Omega(2\Omega + S). 
+\end{align}
+
+Note that the disk is in general baroclinic with $\Omega =
+\Omega(r,z)$. This is explicit from Eq. \ref{eqm_rot}, but is also
+implied by Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} and 
+Eq. \ref{eqm_press},  
+\begin{align}
+  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
+    z}\frac{dK}{dr}.\label{vertical_shear}
 \end{align}
@@
 which may be obtained from Eq. \ref{eqm_dens} and Eq. \ref{eqm_rot} by
-taking the limit $\Gamma\to 1$. The vertically isothermal disk has no
-surface, but we can define a characteristic scale-height 
-\begin{align}
-  \Hiso \equiv \epsilon r.% \quad\quad(\Gamma \equiv 1),   
-\end{align} 
-and in this case the characteristic aspect-ratio $h = \epsilon$.
+taking the limit $\Gamma\to 1$. We will, in fact, primarily focus on
+vertically isothermal disks for simplicitly.
 
@@
   \nabla\cdot\bm{v} -
-  \frac{\rho}{t_c}\left(\frac{P}{\rho}-\left.\frac{P}{\rho}\right|_{t=0}\right), 
+  \frac{\rho}{t_c}\left(\frac{P}{\rho}-\left.\frac{P}{\rho}\right|_{t=0}\right), \label{energy_eq}
 \end{align}

### 2015-01-26 appendix.tex
regular update, code updates
editing numerical results, redid mmsn application for 50au and corrected espilon to include mu dependence

@@
 \end{align}
-which implies $p=-39/14$, 
+which implies $p=-39/14$, {\bf ASK: should there be $\hat{T}$ and $\mu$ dependence for midplane density?}
 a temperature profile
@@
 and the fiducial disk aspect-ratio 
-\begin{align}
-  \epsilon = 0.022\left(\frac{r}{\mathrm{AU}}\right)^{2/7}.
+\begin{align}\label{mmsn_epsilon}
+  \epsilon = 3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}.
 \end{align}
+%0.033576258
 In the above expressions, $\hat{\Sigma}$ and

### 2015-01-26 application.tex
editing numerical results, redid mmsn application for 50au and corrected espilon to include mu dependence

--- 2015-01-24 application.tex
@@
 \begin{align}\label{bcrit_mmsn}
-  \beta_\mathrm{crit} = \frac{9.4\times10^{-3}}{(\gamma
-    -1)}\zeta^{2/7},
+  \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
+    -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\zeta^{2/7},
 \end{align}
@@
 In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn} and
-Eq. \ref{bcrit_mmsn} for the MMSN with $\mu = 2$ and
+Eq. \ref{bcrit_mmsn} for the MMSN with $\mu = 2.33$ and
 $\gamma=1.4$. Although it is possible to satisfy
@@
 \subsection{Numerical example}
-Consider perturbations with $\khat=10$ in the 
-the MMSN at $r=30\mathrm{AU}$, with $\gamma=1.4$ and
-$\mu=2$. Then we find
-the dimensionless thermal relaxation time, using
-Eq. \ref{tc_newton_cool}, is
-\begin{align*}
-  \beta \simeq 0.046. 
-\end{align*}
-At $r=30\mathrm{AU}$, the disk aspect ratio is $\epsilon \simeq
-0.058$ which implies, using Eq. \ref{iso_vsi_cond}, a critical 
+Consider  the MMSN at $r=50\mathrm{AU}$, with $\gamma=1.4$ and
+$\mu=2.33$. Eq. \ref{mmsn_epsilon} gives $\epsilon \simeq
+0.067$ which implies, using Eq. \ref{bcrit_mmsn}, a critical 
 relaxation timescale of  
 \begin{align*}
-  \beta_\mathrm{crit} \simeq 0.062. 
+  \beta_\mathrm{crit} \simeq 0.072. 
 \end{align*} 
-Thus, the disk supports the fundamental isothermal VSI, but only
-marginally. Using the above value of $\beta$, MMSN disk parameters
-and $\khat=10$ in our linear code, we find a growth rate of
+To estimate a thermal relaxation timescale, we set $\khat=10$ 
+in Eq. \ref{beta_mmsn} to obtain 
 \begin{align*}
-  \nu = 2.7\times10^{-3}\Omega_k.
+  \beta \simeq 0.049.
 \end{align*}
-This corresponds to a characteristic growth time of $t_g\simeq 60$
+(In fact, Fig. \ref{mmsn_bcrit_bcool} shows that $\beta$  is not
+sensitive to $\khat$ at $r=50\mathrm{AU}$ for $\khat\gg1$.)
+The disk supports the fundamental isothermal VSI, but only
+marginally. Using the above value of $\beta$, MMSN disk parameters and
+$\khat=10$ in our linear code, we find a 
+growth rate of 
+\begin{align*}
+  \nu = 4.0\times10^{-3}\Omega_k.
+\end{align*}
+This corresponds to a characteristic growth time $\sim 40$
 orbits. 
+%This mode was found to be the most unstable for
+%$\khat=10$.
 
- %only find one unstable mode -> close to marginal
+% and
+% physically acceptable -> other modes
+% are crazy
 
- %larger kx
+%We also solved the linear problem with varying $\khat$  
+%and $\beta$ determined by Eq. \ref{beta_mmsn}. We find the most
+%unstable wavenumber occurs at $\khat\simeq60$ with $\nu\simeq
+%3.6\times10^{-3}\Omega_k$, i.e. only slightly faster than $\khat=20$.   
+%However, the eigenfunctions for large $\khat$ were found to have large
+%amplitude oscillations near the vertical boundaries. This is shown in
+%Fig. \ref{mmsn_eigenW}. These modes may not be important since they
+%mostly affect the disk atmosphere where there is little mass.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/eigenvectorW_mmsnkx10.ps}
+%  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0cm]{figures/eigenvectorW_mmsnkx60.ps}
+  \caption{Most unstable VSI modes in the MMSN at
+    $r=30\mathrm{AU}$. The eigenfunction $W$ for $\khat=20$ (top) and
+    $\khat=60$ (bottom) are shown. The real (imaginary) part of $W$
+    is plotted as the solid (dotted) lines.\label{mmsn_eigenW}}  
+\end{figure}
+
+% only find one unstable mode -> close to marginal 
+% larger kx

### 2015-01-26 linear_adia.tex
adding youdin notes
almost done with numerical results section
editing numerical results, redid mmsn application for 50au and corrected espilon to include mu dependence

@@
 
-% \subsubsection{Qualitative condition for buoyancy to be negligible in
-%   the presence of thermal relaxation}
-% {\bf this section might be wrong, but keep for reference for now}
-% We have shown that when $\gamma>1$ so that there is a positive
-% vertical entropy gradient and perturbations evolve
-% adiabatically ($t_c\to\infty$) that the VSI is strongly
-% surpressed. This is the limit where the term proportional to
-% $t_c^{-1}$ in the linearized energy equation (Eq. \ref{lin_energy})
-% becomes negligible compared to the term proportional to
-% $\gamma/\Gamma-1$. The latter term is associated with the vertical
-% entropy gradient, or buoyancy frequency squared, $N_z^2$.
-
-% Reversing the argument, we expect to recover the VSI when the bouyancy
-% frequency becomes negligible compared to the cooling rate, or 
-% \begin{align}
-%   t_c \ll \frac{\Omega}{\gamma N_z^2}. 
-% \end{align} 
-% Parametrizing $t_c = \beta\Omega_k^{-1}$ and inserting the expression
-% for $N_z^2$ appropriate for vertically isothermal disks, we expect
-% \begin{align}
-%   \beta \ll
-%   \frac{1}{H^2_\mathrm{iso}\left(\gamma-1\right)\left(\frac{\p\ln\rho}{\p
-%           z}\right)^2} 
-% \end{align}
-% is required to enable the VSI, where we used $\Omega\sim\Omega_k$.  
-% This inequality is satisfied at the
-% midplane and as $|z|\to\infty$. It can be satisifed at all heights if
-% it is satisfied at the height of $\mathrm{max}|\p\ln\rho/\p z|$. 
-% Using the full expression for the density field for vertically isothermal
-% disks (Eq. \ref{vertiso_eqm}), we expect 
-% \begin{align}
-%   \beta \ll \frac{27}{4(\gamma-1)}\epsilon^2,
-% \end{align}
-% i.e. $\beta\ll O(\epsilon^2)$, is a sufficient condition for
-% buoyancy to be negligible.
-
 \subsection{Perturbations with thermal relaxation}\label{analytic_relax}
@@
 \subsubsection{The effect of introducing a small but finite
-  thermal relaxation time} 
+  thermal relaxation time}\label{relax_pert}
 It is instructive to first ask  the more analytically tractable
@@
 For a thin disk $\epsilon\ll 1$, we infer 
-\begin{align}
+\begin{align}\label{iso_vsi_cond}
   \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
@@
 solution is $\beta_0 = 10.33$ and $\hat{\omega}_0 = 0.74$.
-
-% The pair of ODEs for $W$ and $Q$
-% (Eq. \ref{ode_w}---\ref{ode_Q}) with these approximations and 
-% written with dimensionless variables, are
-% \begin{align}
-%   &\hat{\sigma}^2\left(Q +  \hat{k}^2W\right) \notag\\ 
-%   &= \hat{z} \left(1 +
-%     \ii\epsilon q \hat{k} \right)\left[W^\prime - \zhat \left(W -
-%       Q\right)\right] - \left[W^\prime - \zhat \left(W -
-%       Q\right)\right]^\prime\label{nearly_iso1},\\
-% &\beta \hat{\sigma}^2\left(W - \gamma Q\right) + \ii\hat{\sigma}
-%     \left(W-Q\right)\notag\\
-%    & = -\beta\zhat\left(\gamma-1\right) \left[W^\prime - \zhat \left(W -
-%       Q\right)\right]\label{nearly_iso2},
-% \end{align}
-% where we recall $\beta = t_c\Omega_k$ is the dimensionless cooling
-% time. It is evident that for $\beta\equiv 0$, we recover
-% Eq. \ref{iso_ode3}. In addition, from the expression of the vertical
-% velocity perturbation (Eq. \ref{lin_vz}) we have
-% \begin{align}
-%   W^\prime - \zhat \left(W -
-%       Q\right) = \ii\hat{\sigma}c_s\delta v_z. \label{nearly_vz}
-% \end{align}
-
-% %is valid for
-% %non-isothermal perturbations $\gamma>1$ if the thermal relaxation
-% %(cooling) timescale $t_c\equiv0$.
-
-% % Then we may linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2}
-
-% To determine the (assumed) small effect of introducing a short thermal
-% relaxation, we linearize Eq. \ref{nearly_iso1}---\ref{nearly_iso2} about a known
-% solution for $\beta\equiv0$. Let 
-% \begin{align}\label{nearly_iso_pert}
-%   \beta \to 0 + \delta\beta,\, W\to W+\delta W,\, Q \to Q+\delta
-%   Q,\,\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}, 
-% \end{align} 
-% where $W=Q$ is the polynomial solution described in \S\ref{iso_poly}
-% with $\hat{\sigma}^2$ given by Eq. \ref{sig2_iso}. Inserting
-% Eq. \ref{nearly_iso_pert} into
-% Eq. \ref{nearly_iso1}---\ref{nearly_iso2} and keeping only first order
-% terms, we obtain
-% \begin{align}
-%   &2\hat{\sigma}\delta\hat{\sigma}\left(1+\hat{k}^2\right)W +
-%   \hat{\sigma}^2 \left(\delta Q + \hat{k}^2\delta W\right)\notag\\
-%   &= \zhat \left(1+\ii\epsilon q \hat{k}\right)\left[\delta W^\prime -
-%     \zhat \left(\delta W - 
-%       \delta Q\right)\right] - \delta W^{\prime\prime}\notag\\
-%   &\phantom{=} + \left[\zhat \left(\delta W -
-%      \delta Q\right)\right]^\prime,\label{nearly_iso3}\\
-% &\ii\hat{\sigma} \left(\delta W - \delta Q\right) =
-% \delta\beta\left(\gamma-1\right)\left(\hat{\sigma}^2W - \hat{z}W^\prime\right).\label{nearly_iso4}
-% \end{align}
-% This is a pair of inhomogeneous ODEs with a forcing by $W$. The
-% homogeneous problem is, of course, just the ODE for $\beta=0$ solved
-% in \S\ref{iso_poly}. Hence we only seek the particular solution.
-
-% We could proceed by considering the entire set of polynomial
-% solutions for $W$, but it is simplest to look at the fundamental
-% mode. Thus we consider the `basic state' 
-% \begin{align*}
-%   W = \hat{z},\quad \hat{\sigma}^2\left(1+\hat{k}^2\right) = 1 +
-%   \ii\epsilon q \hat{k},
-% \end{align*}
-% then
-% \begin{align}
-%   \delta W = b \zhat^3,
-% \end{align}
-% where $b$ is a constant to be determined. Eliminating $\delta Q$ from
-% Eq. \ref{nearly_iso3}---\ref{nearly_iso4}, inserting the above
-% expressions for $W$ and $\delta W$ and balancing coefficients of
-% $\zhat$ and $\zhat^3$, we find
-% \begin{align}
-%  & 2\left(1+\ii\epsilon q \hat{k}\right) \delta\hat{\sigma} \notag\\& = -6b\hat{\sigma} -
-% 2\ii \delta \beta\left(\gamma-1\right)\left(\hat{\sigma}^2-1\right) - \ii
-%   \hat{\sigma}^2\delta\beta\left(\gamma-1\right),\\
-% & 0=  2b\hat{\sigma}+\ii\delta\beta
-% \left(\gamma-1\right)\left(\hat{\sigma}^2 - 1\right),
-% \end{align}
-% where the expression for $\hat{\sigma}^2$ was used.

### 2015-01-26 numerical.tex
editing numerical results, redid mmsn application for 50au and corrected espilon to include mu dependence

@@
   \caption{Eigenvalues of the VSI in a neutrally-stratified,
-    nearly-vertically isothermal disk  ($\gamma=\Gamma=1.011$). The
-    disk parameters are $q=-1$, $p=-1.5$, $\epsilon=0.1$, and the perturbation radial
+    nearly-vertically isothermal disk  ($\gamma=\Gamma=1.011$) with 
+    $(p,q,\epsilon)=(-1.5,-1,0.1)$, and the perturbation radial
     wavenumber is $\khat=20\pi$. \label{lowfreq_eigen}
@@
     VSI in a nearly vertically isothermal disk
-    ($\Gamma=1.011$) evolved adiabatically with $\gamma=2$. The disk
-    parameters are $q=-1$, 
-    $p=-1.5$, $\epsilon=0.1$, and the perturbation radial
+    ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
+    evolved adiabatically with $\gamma=2$. The perturbation radial
     wavenumber is $\khat=20\pi$. \label{lowfreq_eigen_adia}
@@
   \caption{Fundamental VSI mode in a nearly vertically
-    isothermal disk ($\Gamma=1.011$) evolved adiabatically
+    isothermal disk ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
+    evolved adiabatically
     with $\gamma=2$. This eigenfunction 
@@
 \subsection{Other adiabatic indices}
-In Fig. \ref{adia_growth_num} we plot the eigenfrequencies of the
-fundamental VSI mode for $\gamma\in[1,2]$. To select the appropriate
-eigenvalues we vary the parameters slowly away from a reference solution.
+To see the transition from neutrally-bouyant to stably stratified cases above, 
+we repeated the previous calculation with $\gamma\in[1,2]$. 
+We restore the vertical domain to $\zmax=10\epsilon r$. 
+Fig. \ref{adia_growth_num} shows the eigenfrequencies of the
+fundamental VSI mode.
 
+%To select the appropriate
+%eigenvalues we vary the parameters slowly away from a reference solution.   
 % The numerically-determined eigenfrequencies are qualitatively
@@
 There is good agreement between Fig. \ref{adia_growth_num} and
-Fig. \ref{adia_growth} for $k_xH_\mathrm{iso}\gtrsim10$. There is a
-mistach for smaller wavenumbers, especially in the growth
-rates. Actual growth rates decrease slower as $k_xH_\mathrm{iso}\to0$
+Fig. \ref{adia_growth} for $\khat=k_xH_\mathrm{iso}\gtrsim10$. There is a
+mismatch for smaller wavenumbers, especially in the growth
+rates. Actual growth rates decrease slower as $\khat\to0$
 than that in Fig. \ref{adia_growth}. Disagreement at small
-$k_xH_\mathrm{iso}$ is not surprising because the low-frequency
+$\khat$ is not unexpected because the low-frequency
 approximation used to obtain Fig. \ref{adia_growth} becomes invalid since
-$|\omega|\sim\Omega_k$ for small $k_x$.
+$|\omega|\sim\Omega_k$ for small $\khat$.
 
@@
 Fig. \ref{adia_growth}: stabilization for 
-$k_xH_\mathrm{iso}\to0$ and $k_xH_\mathrm{iso}\to\infty$,
-stabilization for increasing $\gamma$ (but ineffective at small $k_x$), 
+$\khat\to0$ and $\khat\to\infty$,
+stabilization for increasing $\gamma$ (but ineffective at small $\khat$), 
 and the shift of the most unstable wavenumber to smaller values as
@@
   \caption{Real frequency (top) and growth rate 
-    (bottom) of the fundamental VSI mode in nearly-vertically   
-    isothermal disks ($\Gamma=1.011$) with $\epsilon=0.1$ and  
-    $q=-1$, subject to adiabatic perturbations
-    ($\beta=10^9$). The eigenfrequencies are calculated by numerically
+    (bottom) of the fundamental VSI in a nearly vertically   
+    isothermal disks ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
+    subject to adiabatic perturbations with different values of $\gamma$. 
+    The eigenfrequencies are calculated by numerically
     solving the full linear eigenvalue problem. This plot is to
@@
 different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as
-adopted in \cite{nelson13}. Fig. \ref{relax_growth_num} plots the
-growth rates of the fundamental mode function of $k_x$. %corrugation
-                                %mode 
+adopted in simulations performed by \cite{nelson13}. 
+Fig. \ref{relax_growth_num} plots the
+growth rates of the fundamental mode function of $\khat$.
+
+%corrugation mode
 % We first obtain the eigenvalues for the
@@
 
-As above, Fig. \ref{relax_growth_num} is largely consistent with 
-Fig. \ref{relax_disp_fig} except for $k_xH_\mathrm{iso}\lesssim 1$
-where the low-frequency approximation fails. The non-smoothness in the
-curve for $\beta=0.063$ at $k_xH_\mathrm{iso}\sim50$ was found to be
-associated with large perturbations near the vertical boundaries. This is
-likely attributable to the free surface boundary condition imposed in
-the numerical calculation (cf. regularity condition at infinity in
-\S\ref{disp_relax}).
+Fig. \ref{relax_growth_num} is largely consistent with 
+Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
+where the low-frequency approximation fails. 
+For $\beta=0.063$ at $\khat\gtrsim50$ we found eigenfunctions had 
+large perturbations near the vertical boundaries, which may be 
+responsible for the non-smoothness in the curve at large $\khat$.
 
-Fig. \ref{relax_growth_num} confirms our analytical discussion that
+%likely attributable to the free surface boundary condition imposed in
+%the numerical calculation (cf. regularity condition at infinity in
+%\S\ref{disp_relax}).
+
+Fig. \ref{relax_growth_num} also confirms our analytical discussion that
 introducing a small but finite cooling time stabilizes the fundamental
-VSI, but there is a preferable wavenumber around $k_xH_\mathrm{iso} =
-O(10)$ that maximizes this effect.
+VSI, but there is a preferable wavenumber that maximizes this effect (\S\ref{relax_pert}).
+This occurs at $\khat= O(10)$ for the current disk model.
 
@@
     nearly-vertically isothermal disks ($\Gamma=1.011$) with
-    $\epsilon=0.05$ and $q=-1$, with thermal relaxation timescale
+    $(p,q,\epsilon)=(-1.5,-1,0.05)$ and a range of thermal relaxation timescales 
     $\beta$. The eigenfrequencies are calculated by numerically 
@@
 
-\subsubsection{$N_z^2\equiv 0$}
+\subsubsection{$\gamma=\Gamma \neq 1$, $N_z^2\equiv 0$}
 
@@
   \caption{Numerically-determined growth rates of the fundamental VSI
-    mode in neutrally-buoyant, vertically non-isothermal disks
-    ($\Gamma=\gamma=1.4$), with $p=0$, $q=-1$ and  $\epsilon=0.05$, for
+    mode in neutrally-buoyant ($N_z^2\equiv0$), vertically non-isothermal disks
+    ($\Gamma=\gamma=1.4$), with $(p,q,\epsilon)=(0,-1,0.05)$, for
     a range of thermal relaxation timescales 
@@
 
-\subsubsection{$N_z^2\geq0$}
+\subsubsection{$\gamma\neq\Gamma \neq 1$, $N_z^2\geq0$}
 
@@
 zero at finite $\khat$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/growth_vnoniso2}
+  \caption{Same as Fig. \ref{growth_vnoniso1}, but with $\gamma=2$ so that
+    $N_z^2\geq0$. \label{growth_vnoniso2}}
+\end{figure}

### 2015-01-26 appendix.tex
updating

@@
 For results application in \S\ref{application} we use the disk model
-of \cite{chiang10}. This disk model orbits a Solar-mass star and 
-has a surface density distribution
+described in \cite{chiang10}. This disk model orbits a Solar-mass star and 
+has the surface density distribution
 \begin{align}\label{mmsn_sigma}
@@
 \end{align}
-a mid-plane density distribution
-\begin{align}
-  \rho_0 = 2.7\times10^{-9}
-  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
-\end{align}
-which implies $p=-39/14$, {\bf ASK: should there be $\hat{T}$ and $\mu$ dependence for midplane density?}
-a temperature profile
+and the temperature profile
 \begin{align}\label{mmsn_temp}
@@
 \end{align}
-which implies $q=-3/7$, 
-and the fiducial disk aspect-ratio 
+which implies $q=-3/7$. In the above expressions, $\hat{\Sigma}$ and
+$\hat{T}$ are dimensionless coefficients used to scale the model
+relative to the MMSN. Their nomial values are unity. By assuming a
+vertically isothermal disk, we deduce the disk aspect-ratio 
 \begin{align}\label{mmsn_epsilon}
-  \epsilon = 3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}.
+  \epsilon =
+  3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}, 
 \end{align}
+and the mid-plane density distribution 
+\begin{align}
+%  \rho_0 = 2.7\times10^{-9}
+%  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
+\rho_0 = 1.7\times10^{-9}
+  \hat{\Sigma}\left(\frac{\hat{T}}{\mu}\right)^{-1/2}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},
+\end{align}
+which implies $p=-39/14$. 
 %0.033576258
-In the above expressions, $\hat{\Sigma}$ and
-$\hat{T}$ are dimensionless coefficients used to scale the model
-relative to the MMSN. Their nomial values are unity. In addition, we
+In addition, we
 use the opacity model {\bf(reference?)}

### 2015-01-27 application.tex
writing summary now

@@
 As an application of our results, we estimate where in a 
-protoplanetary disk do the thermodynamic conditions allow the VSI to 
+protoplanetary disk do thermodynamic conditions allow the VSI to 
 operate. Specifically, we consider the fundamental isothermal VSI, 
-for which we can use the simple criterion Eq. \ref{iso_vsi_cond}. The
-condition derived below may be considered as a sufficient condition for the VSI.
+for which we can use the criterion Eq. \ref{iso_vsi_cond}.
 
@@
 function in the energy equation. To connect our model with a realistic
-disk, we consider the more physically-motivated form of the energy
+disk, we consider a more physically-motivated form of the energy
 equation with radiative cooling 
@@
 \end{align}
-where the temperature $T$ is defined such that $P=\mathcal{R}\rho
+where the temperature $T$ is defined through $P=\mathcal{R}\rho
 T/\mu$, where $\mathcal{R}$ is the ideal gas constant and $\mu$ is
@@
 \end{align}
-where $c_v \equiv \left(\gamma-1\right)^{-1}\mathcal{R}/\mu$ is the specific heat capcity at
-constant volume, $D/Dt$ is the convective derivative, $\bm{F}$ is the
-radiative flux, and $Q_+$ is a heating term such that an equilibrium
-solution exists.
+where $c_v \equiv \left(\gamma-1\right)^{-1}\mathcal{R}/\mu$ is the
+specific heat capcity at constant volume, $D/Dt$ is the convective
+derivative, $\bm{F}$ is the radiative flux, and $Q_+$ is a heating
+term such that an equilibrium solution exists when $D/Dt=0$.
 
@@
 is the photon mean-free-path and $\kappa_d$ is the (dust) opacity. 
+%need to comment on kappa_d(rho,T)
 
@@
 this problem to a follow-up study. Here, we are interested in
-order-of-magnitude estimates. We thus proceed by assuming vertical
-derivatives can be neglected in Eq. \ref{diff_cool_proper} and write
+order-of-magnitude estimates for perturbations with radial
+lengthscales much smaller than vertical lengthscales. We thus proceed
+by assuming vertical derivatives can be neglected in
+Eq. \ref{diff_cool_proper} and write  
 \begin{align}\label{diff_cool_approx}
@@
 \end{align}
-%we are interested in radially-thin perturbations  
+% we are interested in radially-thin perturbations  
 where $\eta=k_\mathrm{rad}/\rho c_v$ is the diffusion coefficient and
@@
 where $\hat{\eta} = \eta/H_\mathrm{iso}^2\Omega_k$ is the 
-dimensionless diffusion coefficient defined for convenience, and is
-approximately 
-\begin{align}\label{diff_coeff_dimensionless}
-  \hat{\eta} \simeq  \frac{16\sigma_s T^3}{3\kappa_d\Sigma^2
-    c_v\Omega_k}. 
-\end{align}
+dimensionless diffusion coefficient.
 
@@
 
-The dimensionless thermal relaxation time $\beta$ is then
+In order to utilize our linear results, which assumes $t_c$ is a
+constant, we need to make further 
+approximations. Specifically, we consider vertically-isothermal disks
+and replace explicit $\rho$ depedences in the above expressions by $\Sigma/
+H_\mathrm{iso}$, where $\Sigma$ is the surface density at the fiducial
+radius of interest. For example, the dimensionless diffusion
+coefficient becomes 
+\begin{align}\label{diff_coeff_dimensionless}
+  \hat{\eta} \to  \frac{16\sigma_s T^3}{3\kappa_d\Sigma^2
+    c_v\Omega_k}. 
+\end{align}
+Furthermore, we will choose a temperature-dependent opacity model for
+simplicity, so that $\kappa_d=\kappa_d(T)$. The dimensionless thermal
+relaxation time $\beta$  
 \begin{align}\label{real_beta}
@@
   \frac{1}{\hat{\eta}}\left(\frac{1}{3\kappa_d^2\Sigma^2} 
-    + \frac{1}{\khat^2}\right).
+    + \frac{1}{\khat^2}\right)
 \end{align}
+is then a constant for vertically-isothermal disks, 
+and we can directly apply previous results.
 
@@
   \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
-    -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\zeta^{2/7},
+    -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\zeta^{2/7}. 
 \end{align}
-and we recall $\beta < \beta_\mathrm{crit}$ permits the fundamental 
-isothermal VSI to operate.
+We recall $\beta < \beta_\mathrm{crit}$ is a neccessary condition for the
+fundamental isothermal VSI when $\khat\gg 1$. For $\khat\to0$, this
+becomes a sufficient condition (see Fig. \ref{relax_bcrit} and
+\S\ref{iso_vsi_beta_crit}).
 
 In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn} and
-Eq. \ref{bcrit_mmsn} for the MMSN with $\mu = 2.33$ and
-$\gamma=1.4$. Although it is possible to satisfy
-$\beta<\beta_\mathrm{crit}$ at small radii by choosing a sufficiently
-large $\khat$, our linear calculations suggest growth rates eventually
-decay as $\khat\to\infty$. We thus restrict our attention to
-$\khat\lesssim 10^2$.
+Eq. \ref{bcrit_mmsn} for the Minimum Mass Solar Nebula (MMSN) by setting
+the scalings $\hat{T}=\hat{\Sigma}=\hat{\kappa}_d=1$, with $\mu =
+2.33$ and $\gamma=1.4$. The figure suggests that the fundamental
+isothermal VSI operates in the outer parts of the MMSN at radii on
+the order of tens of AU. However, the thermal relaxation time $\beta$
+is not much shorter than $\beta_\mathrm{crit}$, so the VSI may be 
+weak.
 
-Fig. \ref{mmsn_bcrit_bcool} suggests that the fundamental isothermal
-VSI can operate in the outer parts of the MMSN at radii on the order
-of tens of AU. However, the thermal relaxation time $\beta$ is not
-much shorter than $\beta_\mathrm{crit}$, so the VSI may be
-weak.
+We remark that it is possible to satisfy $\beta<\beta_\mathrm{crit}$
+at small radii by choosing a sufficiently large $\khat$. However, our
+linear calculations suggest growth rates eventually decay as 
+$\khat\to\infty$. Thus, very small-scale VSI in the inner disk is
+probably unimportant.
 
@@
     perturbation radial wavenumber: $\khat=1$ (dotted), $\khat=10$
-    (dashed) and $\khat=100$ (dashed-dot). The relaxation timescale
-    $\beta_\mathrm{crit}$, below which the fundamental isothermal VSI
-    operates, is shown as the solid line. 
+    (dashed) and $\khat=100$ (dashed-dot). The solid line is the
+    critical thermal relaxation timescale $\beta_\mathrm{crit}$, and 
+    $\beta<\beta_\mathrm{crit}$
+    is necessary for the fundamental isothermal VSI to operate  
+    for $\khat\gg1$. 
     \label{mmsn_bcrit_bcool}}   
@@
 \subsection{Numerical example}
-Consider  the MMSN at $r=50\mathrm{AU}$, with $\gamma=1.4$ and
-$\mu=2.33$. Eq. \ref{mmsn_epsilon} gives $\epsilon \simeq
+Consider  the MMSN at $r=50\mathrm{AU}$. Eq. \ref{mmsn_epsilon} gives
+$\epsilon \simeq 
 0.067$ which implies, using Eq. \ref{bcrit_mmsn}, a critical 
@@
 sensitive to $\khat$ at $r=50\mathrm{AU}$ for $\khat\gg1$.)
-The disk supports the fundamental isothermal VSI, but only
-marginally. Using the above value of $\beta$, MMSN disk parameters and
-$\khat=10$ in our linear code, we find a 
-growth rate of 
+The disk is unstable to the fundamental isothermal VSI, but not
+strongly.
+
+Using the MMSN disk parameters at $r=50\mathrm{AU}$ and the
+scale-dependent thermal relaxation $\beta(\khat)$ given by
+Eq. \ref{beta_mmsn} in our linear code, we calculate the fundamental
+VSI growth rates in Fig. \ref{mmsn_growth}. Growth rates
+are weakly depedent on $\khat$ until $\khat\gtrsim50$ beyond which the
+mode is rapidly stabilized. A representative growth rate is 
 \begin{align*}
-  \nu = 4.0\times10^{-3}\Omega_k.
+  \nu =0.059\epsilon\Omega_k = 4.0\times10^{-3}\Omega_k 
 \end{align*}
-This corresponds to a characteristic growth time $\sim 40$
-orbits. 
-%This mode was found to be the most unstable for
-%$\khat=10$.
-
-% and
-% physically acceptable -> other modes
-% are crazy
-
-%We also solved the linear problem with varying $\khat$  
-%and $\beta$ determined by Eq. \ref{beta_mmsn}. We find the most
-%unstable wavenumber occurs at $\khat\simeq60$ with $\nu\simeq
-%3.6\times10^{-3}\Omega_k$, i.e. only slightly faster than $\khat=20$.   
-%However, the eigenfunctions for large $\khat$ were found to have large
-%amplitude oscillations near the vertical boundaries. This is shown in
-%Fig. \ref{mmsn_eigenW}. These modes may not be important since they
-%mostly affect the disk atmosphere where there is little mass.
+for $\khat=10$. This corresponds to a characteristic growth time $\sim
+40$ orbits.
 
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/eigenvectorW_mmsnkx10.ps}
-%  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0cm]{figures/eigenvectorW_mmsnkx60.ps}
-  \caption{Most unstable VSI modes in the MMSN at
-    $r=30\mathrm{AU}$. The eigenfunction $W$ for $\khat=20$ (top) and
-    $\khat=60$ (bottom) are shown. The real (imaginary) part of $W$
-    is plotted as the solid (dotted) lines.\label{mmsn_eigenW}}  
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/growth_mmsn_50AU} 
+  \caption{VSI growth rates in the MMSN at
+    $r=50\mathrm{AU}$. The thermal relaxation timescale $\beta(\khat)$
+    used for this calculation is given by
+    Eq. \ref{beta_mmsn} and is scale-dependent. 
+    \label{mmsn_growth}}  
 \end{figure}
 
+Fig. \ref{mmsn_eigenW} show several eigenfunctions for the modes in
+Fig. \ref{mmsn_growth}. As $\khat$ is increased, the disturbance
+become more concentrated near the vertical boundaries. Thus, modes
+with very large $\khat$ may not be important in practice because they
+predominantly affect the disk atmosphere, which contains little mass.
+
+\begin{figure}
+   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvectorW_mmsnkx1.ps}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvectorW_mmsnkx10.ps} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/eigenvectorW_mmsnkx50.ps} 
+  \caption{Fundamental VSI modes in the MMSN at
+    $r=50\mathrm{AU}$. The eigenfunction $W$ for $\khat=1$ (top),
+    $\khat=10$ 
+    (middle) and $\khat=50$ (bottom) are shown. The real (imaginary)
+    part of $W$ is plotted as the solid (dotted)
+    lines. The thermal relaxation timescale $\beta(\khat)$ is given by
+    Eq. \ref{beta_mmsn} and is scale-dependent. \label{mmsn_eigenW}}    
+\end{figure}
 % only find one unstable mode -> close to marginal 

### 2015-01-27 linear_adia.tex
writing summary now

@@
 
-\subsubsection{Lower limit on the thermal relaxation timescale for the
-  fundamental adiabatic VSI mode}
-{\bf Note: for reference only. Not to be included in paper.} 
-Fig. \ref{relax_bcrit} indicates there exists a lower limit on the
-thermal relaxation time for the system to behave adiabatically and
-support the fundamental adiabatic VSI. This critical
-timescale satisfies Eq. \ref{relax_cond1}---\ref{relax_cond2} with 
-\begin{align}
-  \beta \sim \beta_0\khat^{1/2},\quad \hat{\omega}  \sim
-  \hat{\omega}_0\khat^{-1/2} \text{ as } \khat\to\infty.
-\end{align}
-The constants $\beta_0$ and $\hat{\omega}_0$ are determined by the
-simultaneous equations
-\begin{align}
-  & 0 = \gamma\beta_0^2(1-\gamma) + 2\epsilon q \gamma
-  \beta_0\hat{\omega}_0 - \hat{\omega}_0^2 +
-  \beta_0^2\gamma^2\hat{\omega}_0^4.\\
-  & 0= \beta_0(1-\gamma) + \epsilon q \hat{\omega}_0 - \epsilon q
-  \gamma^2 \beta_0^2\hat{\omega}_0^3. 
-\end{align}
-For $\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$, the appropriate
-solution is $\beta_0 = 10.33$ and $\hat{\omega}_0 = 0.74$.
+% \subsubsection{Lower limit on the thermal relaxation timescale for the
+%   fundamental adiabatic VSI mode}
+% {\bf Note: for reference only. Not to be included in paper.} 
+% Fig. \ref{relax_bcrit} indicates there exists a lower limit on the
+% thermal relaxation time for the system to behave adiabatically and
+% support the fundamental adiabatic VSI. This critical
+% timescale satisfies Eq. \ref{relax_cond1}---\ref{relax_cond2} with 
+% \begin{align}
+%   \beta \sim \beta_0\khat^{1/2},\quad \hat{\omega}  \sim
+%   \hat{\omega}_0\khat^{-1/2} \text{ as } \khat\to\infty.
+% \end{align}
+% The constants $\beta_0$ and $\hat{\omega}_0$ are determined by the
+% simultaneous equations
+% \begin{align}
+%   & 0 = \gamma\beta_0^2(1-\gamma) + 2\epsilon q \gamma
+%   \beta_0\hat{\omega}_0 - \hat{\omega}_0^2 +
+%   \beta_0^2\gamma^2\hat{\omega}_0^4.\\
+%   & 0= \beta_0(1-\gamma) + \epsilon q \hat{\omega}_0 - \epsilon q
+%   \gamma^2 \beta_0^2\hat{\omega}_0^3. 
+% \end{align}
+% For $\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$, the appropriate
+% solution is $\beta_0 = 10.33$ and $\hat{\omega}_0 = 0.74$.

### 2015-01-27 numerical.tex
writing summary now

@@
 \begin{figure}
-   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_imag_bloop}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag_bloop2}
+   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+   0cm]{figures/compare_eigen_imag_bloop} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/compare_eigen_imag_bloop2} 
   \caption{Growth rates of the fundamental VSI mode in 
     nearly-vertically isothermal disks ($\Gamma=1.011$) with
-    $(p,q,\epsilon)=(-1.5,-1,0.05)$ and a range of thermal relaxation timescales 
+    $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ under a
+    range of thermal relaxation timescales   
     $\beta$. The eigenfrequencies are calculated by numerically 
@@
 value beyond which the fundamental isothermal VSI is stabilized
-(\S\ref{iso_vsi_beta_crit}).  Increasing $\beta$ restricts 
+(\S\ref{iso_vsi_beta_crit}). Hence the latter growth rate is smaller
+than the former by a factor $\sim 20$. Increasing $\beta$ restricts 
 the region in which $W\sim z$ closer to the midplane, with oscillatory
 behavior emerging at the boundaries, where buoyancy first becomes
-important.
+important. % We also find this trend for fixed $\beta$ but increasing
+% $\khat$.
 
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/eigenvectorW_beta0d01}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 0cm]{figures/eigenvectorW_beta0d125}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvectorW_beta0d01} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/eigenvectorW_beta0d125} 
   \caption{Numerically-calculated eigenfunction of the fundamental VSI
-    mode in a disk with dimensionless thermal relaxation timescale
-    $\beta \equiv t_c\Omega_k = 0.01$ (top) and $0.125$ (bottom) for
-    perturbations with a radial wavenumber $k_xH_\mathrm{iso}=10$. The
+    mode in a nearly vertically isothermal disk ($\Gamma=1.011$) with
+    $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ and a dimensionless 
+    thermal relaxation timescale $\beta \equiv t_c\Omega_k = 0.01$
+    (top) and $0.125$ (bottom). The
+    perturbation radial wavenumber is $\khat_\mathrm{iso}=10$. The
     real (imaginary) part of $W$ is plotted as the solid (dotted)
-    line. Other disk parameters are
-    $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.
+    line.
     \label{relax_eigenW_num}}  
@@
 We briefly consider vertically non-isothermal disks with
-$\Gamma=1.4$. Other disk parameters are $p=0$, $q=-1.0$ and $\epsilon
-= 0.05$. These are again chosen for comparison with \cite{nelson13}. 
-In this case we set $\zmax\simeq2.2\epsilon r$, which is close to the
-zero-density surface.
+$\Gamma=1.4$ and $(p,q,\epsilon)=(0,-1,0.05)$ as simulated in
+\cite{nelson13}. In this case we set $\zmax\simeq2.2\epsilon r$, which
+is close to the zero-density surface.
 
 \subsubsection{$\gamma=\Gamma \neq 1$, $N_z^2\equiv 0$}
-
 Consider first the case $\gamma=1.4$ so that there is no vertical
@@
 $\beta\to\infty$. Since $\Gamma=O(1)$, these two limits are expected
-to produce qualitatively similar results.
+to produce qualitatively similar results. Interestingly,
+Fig. \ref{growth_vnoniso1} show that growth rates 
+maximize for $\beta=O(1)$. However, all growth rates are similar to
+order-of-magnitude for $0.1\lesssim\khat\lesssim 100$.
 
@@
 
-Interestingly, Fig. \ref{growth_vnoniso1} show that growth rates
-maximize for $\beta=O(1)$. However, all growth rates are similar to
-order-of-magnitude for $0.1\lesssim\khat\lesssim 100$.
-
 \subsubsection{$\gamma\neq\Gamma \neq 1$, $N_z^2\geq0$}
-
 %gamma = 2.0, Gamma = 1.4, vary bcool 
-
 Finally, we introduce buoyancy in a vertically non-isothermal disk by
-setting $\gamma=2$. This situation was not simulated in 
+setting $\gamma=2$. This scenario was not considered by  
 \cite{nelson13}, but is similar in spirit with vertically isothermal 
 disks evolved with $\gamma>1$, as considered in the previous section. 
-
-Fig. plots the growth rates for this case. The trend is
-qualitatively similar to that for vertically isothermal disks with
-$N_z^2\geq 0$ (Fig. \ref{relax_growth_num}), namely growth rates reach
-zero at finite $\khat$.
+Fig. \ref{growth_vnoniso2} plots the fundamental VSI growth rates for
+this case. We observe the same trend as for vertically isothermal
+disks with $N_z^2\geq 0$ (Fig. \ref{relax_growth_num}), namely growth
+rates reach zero at finite $\khat$.
 

### 2015-01-27 summary.tex
writing summary now

\section{Summary and discussion}
%what have we done, relation to previous work 
In this paper we have studied the linear axisymmetric vertical shear
instability (VSI) in astrophysical disks that arises from the
height-dependence of the equilibrium angular velocity profile
$\Omega(r,z)$ of a baroclinic  disk. The VSI was analysed and
explained earlier in the local limit by \cite{urpin98} and
\cite{urpin03}, simulated by \cite{arlt04} and more  recently by
\cite{nelson13}.

We generalized previous vertically-global linear stability calculations
of the VSI \citep{nelson13,mcnally14}, which assumed
isothermal perturbations, to include an energy equation. Our effort
may also be regarded as an extension of \cite{urpin03}, which 
included an energy equation, by relaxing the Boussinesq and  
vertically-local approximations adopted in that analysis.

In accordance with the above studies, which show that the VSI   
applies to perturbations with small radial lengthscales, we retained
the radially-local approximation by using the `vertically-global
shearing box' (VGSB) formalism developed by \cite{mcnally14}, who 
showed the linear isothermal VSI studied in \cite{nelson13}, can
be captured by the VGSB. We added 
an energy source term in the VGSB equations which relaxes the disk
temperature to its initial value on a timescale
$t_c=\beta\Omega_k^{-1}$. This form of thermal relaxation was also 
employed in the global numerical simulations of \cite{nelson13}, and allows
the thermodynamic response of the disk to be varied in a controlled 
manner.

We mostly focused on vertically isothermal disks ($\Gamma=1$) because
it permits a largely analytical discussion. Our analysis is an
extension of the linear models considered in \cite{lubow93} by
including vertical shear, but in the limit of low-frequency
perturbations in  nearly-Keplerian disks. These approximations were
made for simplicity, but also motivated by previous work
\citep{nelson13}.

For isothermal perturbations ($\gamma=1$ or $\beta=0$) we explicitly
demonstrated instability 
associated with vertical shear, and that the VSI growth rate is
limited by the maximum vertical shear within the vertical domain of
interest. In the opposite limit of adiabatic perturbations ($\gamma>1$
and $\beta\to\infty$), we find VSI growth rates are dramatically
reduced on account of a stabilizing 
vertical entropy gradient. The VSI is not completely
surpressed because vertical shear dominates over buoyancy 
sufficiently close the disk midplane. However, perturbations grow
slowly and is constrained to regions immediately about the disk
midplane. Thus, the adiabatic VSI is likely unimportant in practice.

For non-isothermal ($\gamma\neq1$) perturbations with intermediate
thermal relaxation timescales, the general dispersion relation
(Eq. \ref{relax_disp}) is complicated and requires a numerical
study. However, for the fundamental VSI, characterized by  $\delta
v_z\sim\mathrm{constant}$ near $z=0$, some simplification is
possible. In particular, we quantified the thermal relaxation
timescale $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$, below
which small-scale perturbations ($\khat\gg 1$) may be regarded as isothermal
and the standard VSI operates. We believe this explicit expression for
$\beta_\mathrm{crit}$ is the most useful result from our
investigation.

%summary of results

%application to MMSN

\subsection{Caveats and outlooks}

### 2015-01-27 application.tex
completed rough summary

@@
 
-\subsection{Newtonian cooling}
+\subsection{Newtonian cooling}\label{newton_cool}
 For $L\ll l_\mathrm{rad}$ the radiative cooling is given by {\bf (reference?)}
@@
 
-\subsection{Toy model for realistic thermal relaxation}
+\subsection{Toy model for realistic thermal relaxation}\label{toy_relax}
 To unify $t_\mathrm{cool}$ and $t_\mathrm{diff}$, we define the
@@
 
-\subsection{Numerical example}
+\subsection{Numerical example}\label{mmsn_example}
 Consider  the MMSN at $r=50\mathrm{AU}$. Eq. \ref{mmsn_epsilon} gives

### 2015-01-27 linear_iso.tex
regular update
completed rough summary

@@
-\section{Analytical discussion of vertically isothermal disks} 
+\section{Analytical discussion of vertically isothermal disks}\label{analytical}
 We begin with a discussion of vertically isothermal disks
@@
 \end{enumerate}
-Making both of these approximations is equivalent to the replacement
+Making both of these approximations is equivalent to setting 
 $D\to\Omega_k^2$.  After making this replacement, we eliminate $W$ and

### 2015-01-27 summary.tex
completed rough summary

@@
 For non-isothermal ($\gamma\neq1$) perturbations with intermediate
-thermal relaxation timescales, the general dispersion relation
+$\beta$, the general dispersion relation
 (Eq. \ref{relax_disp}) is complicated and requires a numerical
-study. However, for the fundamental VSI, characterized by  $\delta
+treatment. However, for the fundamental VSI, characterized by  $\delta
 v_z\sim\mathrm{constant}$ near $z=0$, some simplification is
@@
 
-%summary of results
+We also solved the linear stability problem numerically using a
+pseudo-spectral approach. Numerical solutions are mostly consistent
+with our analytical expectations, except for perturbations with small
+radial wavenumber ($\khat\to0$) where the low-frequency approximation
+is inappropriate.
+
+We briefly studied vertically non-isothermal disks using our linear
+code. We confirm the finding by \cite{nelson13} that in the absence
+of vertical entropy gradients, the VSI operates and is insensitive to thermal relaxation. When a
+stable stratification is introduced, the dependence of VSI growth
+rates on the thermal relaxation timescale is qualitatively similar to
+that observed for vertically isothermal disks.
 
@@
 
-\subsection{Caveats and outlooks}
+We applied our linear results to assess the relevance of the VSI in
+protoplanetary disks, specifically the Minimum Mass Solar Nebulae (MMSN),
+using disk models described in \cite{chiang10}. We find the MMSN is
+unstable to the VSI at $10$s of AU, with characteristic
+growth times of $10$s of orbits. However, we note the strong
+temperature dependence of the physical thermal relaxation timescale
+in the disk model ($\beta$ decreases as $T^{-5}$ for perturbations with small
+radial scale). Thus, increasing the disk temperature can signifcantly 
+shorten the typical growth times for the VSI. This is also attributed
+to a larger value of $\beta_\mathrm{crit}$, since increasing $T$ also
+increases the aspect-ratio $\epsilon$.
+
+\subsection{Future extensions} 
+We highlight below some improvements to the present  
+linear models that should be pursued:
+
+\emph{Realistic thermal evolution.} We have treated the thermal
+relaxation time $t_c$ as a constant for simplicity. However, a
+physical expression for $t_c$ will depend on 
+$z$. For example, the Newtonian cooling timescale
+(\S\ref{newton_cool}) will depend on $z$ for vertically non-isothermal
+disks, and through the opacity since 
+$\kappa_d=\kappa_d(\rho,T)$ in general.
+
+For vertically isothermal disks and a temperature-dependent opacity
+model, as considered in \S\ref{toy_relax} and Appendix \ref{mmsn},
+there is still a vertical dependence of the effective thermal relaxation time
+because the radiative diffusion coefficient depends on $\rho$. We
+artificially discarded this vertical dependence by 
+replacing $\rho\to\Sigma/H_\mathrm{iso}$, but can restore it
+by making the transformation
+$\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv
+\tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
+$\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example calculation
+in \S\ref{mmsn_example}  with $\beta\to\tilde{\beta}$, but found
+similar results.
+
+A proper treatment for radiative diffusion will ultimately involve
+the Laplacian operator. In this case additional boundary conditions 
+must be supplemented. We will consider this in a future work.
+
+\emph{Additional physics.} Depending on where the VSI operates in a
+disk, additional effects may become important. For example, in the
+outer parts of a protoplanetary disk, where we estimate the VSI may
+operate, disk self-gravity (SG) may play a role. Specifically, the vertical
+shear profile depends on the vertical density stratification
+(Eq. \ref{vertical_shear}), which wil be enhanced by vertical SG,
+which will enhance the VSI. Inclusion of the Poisson equation should
+therefore be considered.
+
+We have neglected viscosity. \cite{nelson13} have shown that an
+alpha viscosity of $\lesssim O(10^{-4})$ is required for the VSI to
+operate. It is of interest to quantify the effect of viscosity
+on the linear VSI. This would help identify, more accurately, the
+appropriate regions in a protoplanetary disk that are sufficiently
+laminar to develop the VSI.  This problem is, however, complicated by
+the fact that proper inclusion of viscosity introduces viscous heating
+in the energy equation.
+
+% \emph{Theoretical generalizations.} Our analytical discussion
+% (\S\ref{analytical}) can be extended in several ways. One could
+% relax the low-frequency approximation to improve the agreement between
+% numerical and analytical eigenfrequencies at small $\khat$. However,
+% such modes are radially more global, and would need a global
+% cylindrical disk model to study them. Another minor improvement is to
+
+%analytical extensions -> nonisothermal, all frequency, non-keplerian,
+%diffusion 
+%viscosity self-gravity 
+%disk models: polytropic connected with isothermal

### 2015-01-27 application.tex
drafted summary

@@
 % larger kx
+
+\subsubsection{Vertically-depedent thermal relaxation} 
+Note that for a vertically isothermal disk and opacity $\kappa_d(T)$, 
+we can restore the $z$ dependence in $\beta$ (due to the $\rho$
+dependence in  $\eta$, see Eq. \ref{eta_def}) by making the transformation
+$\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv 
+\tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
+$\hat{\rho}=\rho/\rho(z=0)$.  We have have repeated the example
+calculation above with $\beta\to\tilde{\beta}$. We find growth rates
+were only slightly reduced at low $\khat$ and eigenfunctions are
+similar.

### 2015-01-27 summary.tex
drafted summary

@@
 
-We applied our linear results to assess the relevance of the VSI in
+We applied our linear results to assess the applicability of the VSI in
 protoplanetary disks, specifically the Minimum Mass Solar Nebulae (MMSN),
@@
 $\kappa_d=\kappa_d(\rho,T)$ in general. 
-
-For vertically isothermal disks and a temperature-dependent opacity
-model, as considered in \S\ref{toy_relax} and Appendix \ref{mmsn},
-there is still a vertical dependence of the effective thermal relaxation time
-because the radiative diffusion coefficient depends on $\rho$. We
-artificially discarded this vertical dependence by 
-replacing $\rho\to\Sigma/H_\mathrm{iso}$, but can restore it
-by making the transformation
-$\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv
-\tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
-$\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example calculation
-in \S\ref{mmsn_example}  with $\beta\to\tilde{\beta}$, but found
-similar results.
-
-A proper treatment for radiative diffusion will ultimately involve
-the Laplacian operator. In this case additional boundary conditions 
-must be supplemented. We will consider this in a future work.
+A proper treatment for radiative diffusion, on the other hand, will
+ultimately involve the Laplacian operator. In this case additional
+boundary conditions must be supplemented. We will consider this in a
+future work.
 
@@
 
+\emph{Geometry.} Studying the VSI in the shearing box has an 
+inconsistency. Namely, the equilibrium temperature in the
+box only depends on height (if at all). However, it is precisely the
+radial temperature (or entropy) gradient in the global disk that
+gives rise to vertical shear. This is not a serious issue for
+perturbations with small radial wavelengths, 
+%explicitly checked by including global terms 
+but will need to be resolved by adopting global disk models  
+if one wishes to generalize the VSI to large-scale perturbations.
+
 % \emph{Theoretical generalizations.} Our analytical discussion

### 2015-01-28 paper.tex
minor update
major restructuring
notation change, begun adiabatic notes
more adiabatic discussion, plus a new fig of growth rates
write up edits
minor updates
extended analytic discussion to thermal relaxation cases
new fig for analytic discussion with therm relax
adding applications
writing summary now
completed rough summary
drafted summary
abstract added, intro left to do

@@
 \newcommand{\p}{\partial}
+\newcommand{\zmax}{z_\mathrm{max}}
 \newcommand{\lmax}{l_\mathrm{max}}
@@
 \newcommand{\zhat}{\hat{z}}
+\newcommand{\khat}{\hat{k}}
 \newcommand{\rhat}{\hat{r}}
@@
 \newcommand{\sgn}{\operatorname{sgn}}
-\newcommand{\zmax}{Z_s}
 \newcommand{\wbar}{\tilde{W}}
@@
 
-\title{Linear vertical shear instability in astrophysical disks: a parameter study}
+\title{Linear analysis of the vertical shear instability in astrophysical disks}
 
@@
   Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
-\email{minkailin@email.arizona.edu}
+\email{minkailin@email.arizona.edu, youdin@email.arizona.edu}
 
 \begin{abstract}
-
+  Baroclinic astrophysical disks with equilibrium rotation profiles
+  $\Omega$ that depends on the distance away from the disk midplane are
+  potentially unstable to the Vertical Shear Instability (VSI). It is
+  is necessary to understand the linear instability before
+  considering the VSI as a hydrodynamic route to turbulent angular
+  momentum transport in accretion disks. We study the linear,
+  axisymmetric VSI in radially-local but vertically-global models
+  of accretion disks, including an energy equation with thermal
+  relaxation on a characteristic timescale $t_c=\beta\Omega_k^{-1}$
+  where $\beta$ is a dimensionless parameter and $\Omega_k$ is the
+  Keplerian rotation frequency. We explore how the VSI depends on the
+  thermodynamic response  of the disk by varying $\beta$. We find
+  increasing $\beta$ away from zero dramatically reduces VSI growth
+  rates because this generally introduces a strongly stabilizing
+  vertical entropy gradient. Our linear calculations are consistent
+  with previous hydrodynamic simulations, which indicate that the VSI
+  requires a short thermal relaxation timscale. For vertically
+  isothermal disks, we quantify `fast'
+  thermal relaxation as $\beta < \epsilon|q|/(\gamma-1)$, where
+  $\epsilon$ is the disk aspect-ratio, $q$ is the radial power-law
+  index for the equilibrium temperature (which gives rise to vertical
+  shear) and $\gamma$ is the adiabatic index. We apply our linear
+  results to a disk model of the Minimum Mass Solar Nebulae (MMSN), and
+  find the MMSN is unstable to the VSI at tens of AU with a
+  characteristic growth times of tens of local orbital periods. 
 \end{abstract}
@@
 \section{Introduction}
+% stuff about hydrodynamic angular momentum transport 
+Understanding the mechanisms of angular mometum transport is the
+central problem in the theory of accretion disks
+\citep{lydenbell74}. The
+magneto-rotational instability in the context of astrophysical disks 
+\citep{balbus91} is now accepted as a robust pathway to
+magneto-hydrodynamic (MHD) turbulence, which transports angular momentum
+outwards and thus enables mass accretion.
+
+%stuff about vsi
+
+%recent calculations (nelson, mcnally)
+
+%why linear approach? computationally fast (need very high resolution) 
+%can be used as a first estimate
+
+%what we're doing
+
+%paper organization
 
@@
 \input{linear_iso}
-%\input{results}
-%\input{results_gi}
-%\input{results_azi} 
-%\input{summary}
+\input{linear_adia}
+\input{numerical}
+\input{application} 
+\input{summary}
 
 %\acknowledgements
-%\appendix
-%\input{appendix}
+\appendix
+\input{appendix}
 

### 2015-01-28 summary.tex
abstract added, intro left to do

@@
 of vertical entropy gradients, the VSI operates and is insensitive to thermal relaxation. When a
-stable stratification is introduced, the dependence of VSI growth
-rates on the thermal relaxation timescale is qualitatively similar to
+stable stratification is introduced, the reduction of VSI growth
+rates by a finite thermal relaxation timescale is qualitatively similar to
 that observed for vertically isothermal disks.

### 2015-01-31 paper.tex
reg update
MAJOR REVISION UNDERWAY

@@
   potentially unstable to the Vertical Shear Instability (VSI). It is
-  is necessary to understand the linear instability before
+  is important to understand the linear instability before
   considering the VSI as a hydrodynamic route to turbulent angular
@@
   axisymmetric VSI in radially-local but vertically-global models
-  of accretion disks, including an energy equation with thermal
-  relaxation on a characteristic timescale $t_c=\beta\Omega_k^{-1}$
+  of accretion disks. We include an energy equation with thermal
+  relaxation on a characteristic timescale $t_c=\beta\Omega_k^{-1}$,
   where $\beta$ is a dimensionless parameter and $\Omega_k$ is the
-  Keplerian rotation frequency. We explore how the VSI depends on the
+  Keplerian rotation frequency, and explore how the VSI depends on the
   thermodynamic response  of the disk by varying $\beta$. We find
@@
   vertical entropy gradient. Our linear calculations are consistent
-  with previous hydrodynamic simulations, which indicate that the VSI
+  with previous hydrodynamic simulations. These simulations indicate that the VSI 
   requires a short thermal relaxation timscale. For vertically 
@@
   $\epsilon$ is the disk aspect-ratio, $q$ is the radial power-law
-  index for the equilibrium temperature (which gives rise to vertical
+  index for the equilibrium temperature profile (which gives rise to vertical 
   shear) and $\gamma$ is the adiabatic index. We apply our linear
@@
 central problem in the theory of accretion disks
-\citep{lydenbell74}. The
-magneto-rotational instability in the context of astrophysical disks 
-\citep{balbus91} is now accepted as a robust pathway to
-magneto-hydrodynamic (MHD) turbulence, which transports angular momentum
-outwards and thus enables mass accretion.
+\citep{lyndenbell74}. The  
+magneto-rotational instability in the context of differentially
+rotating disks \citep[MRI,][]{balbus91} is now accepted as a robust
+pathway to magneto-hydrodynamic (MHD) turbulence, which transports
+angular momentum outwards and thus enables mass accretion.
+
+However, MHD turbulence may not be applicable for all 
+accretion disks. For example, there may be regions in protoplanetary
+disks --- called dead zones ---  which do not support the MRI 
+\citep{gammie96,landry13}. In fact, recent state-of-the-art
+
+%in this respect purely hydrodynamical turbulence should be
 

### 2015-01-31 setup.tex
MAJOR REVISION UNDERWAY

@@
 \end{align}
-and we are interested in neutrally or sub-adiabatically stratified
+where $\gamma$ is the constant adiabatic index. We are interested in
+neutrally or sub-adiabatically stratified 
 disks with $\Gamma\leq \gamma$ so that $N_z^2\geq0$.
@@
 
+\subsection{Stability condition against axisymmetric adiabatic
+  perturbations}
+A rotating flow is stable against axisymmetric adiabatic perturbations
+if both Solberg-Hoiland criteria  are satisfied:
+\begin{align}
+  \kappa^2 - \frac{1}{\rho c_p}\nabla P \cdot \nabla S &> 0,\\
+  -\frac{\p P}{\p z}\left(\frac{\p j^2}{\p r}\frac{\p S}{ \p z} -
+    \frac{\p j^2}{\p z}\frac{\p S}{\p R} \right)&>0.  
+\end{align}
+\citep{tassoul78}, where $S = c_p\ln{P^{1/\gamma}/\rho}$ is the
+specific entropy, $c_p$ is the heat capacity at constant
+pressure (assumed to be constant), and $j\equiv r^2\Omega$ is the
+specific angular momentum.
+
+For rotationally-supported, thin disks, pressure gradients are small 
+compared to rotation, so the first criterion is generally
+satisfied. Thus, to assess stability, we evaluate the second criterion
+for the above disk models in $z>0$ where $\p_zP<0$, without loss of
+generality,   
+\begin{align}
+  \gamma - \Gamma > \frac{s\epsilon^2}{\Gamma}
+  \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s
+    -\left(\gamma-\Gamma\right)\frac{\p\ln{P}}{\p\ln{r}} \right]. 
+\end{align} 
+For typical model parameters the right hand side is
+$O(\epsilon^2)$. Hence, to have instability in the adiabatic limit 
+requires $\gamma < \Gamma + O(\epsilon^2)$. Since we consider
+$\epsilon\ll1$, this implies the equilibrium disk must have a small
+entropy gradient in order to be unstable to axisymmetric adiabatic
+perturbations.
+
 \section{The vertically global shearing box} 
@@
 
-In the this shearing box the equilibrium density, pressure and
+In this shearing box the equilibrium density, pressure and
 sound-speed are assumed to depend only on $z$, by evaluating these
 quantities at $r=r_0$ in the global disk. The background rotation is
-approximated by a linear shear in $x$, but left unaltered in $z$.
+approximated by a linear shear in $x$, but left unaltered in
+$z$. However, the vertical shear $\p_z\Omega$ is also assumed to only
+depend on $z$.
 
@@
 subscript $0$ indicates evaluation at $r=r_0$. In the energy equation 
-the ratio of specific heats $\gamma$ is assumed constant, 
-and we have included a thermal relaxation term with timescale
+we have added a thermal relaxation term with timescale
 $t_c=\beta\Omega_k^{-1}$ with $\beta$  being a dimensionless input
 parameter.
+
+\subsection{Limitations}
+Before using the VGSB to study the VSI, it is very important to be
+aware of several limitations of the box model. The VGSB momentum  
+equation have the same form as that for the standard shearing box 
+\citep{goldreich65}, except for the appearence of the term 
+$v_zr_0d\Omega_0/dz$. This term appears in the global stability
+problem and is the only term associated with vertical shear that is 
+ported from the global disk into the local model.
+
+%horizontal structure not ported
+
+Note that, within the VGSB framework, the vertical shear 
+$d\Omega_0/dz$ is a free parameter (as are $\Omega_0$ and $S_0$). 
+This is different to the global problem because in that case, the
+vertical shear is a result of the disk radial structure, which is
+ignored in the VGSB equilibria. 
+%because there is only one
+%equilibrium condition, which determines the vertical density/pressure
+%structure. 
+In effect, we are studying dynamics in the standard shearing box
+subject to an additional azimuthal force $-v_zr_0 d \Omega_0/dz$. This
+force is inspired by the vertical shear in the global disk, and its
+functional form is taken from global equilibrium. 
+%but should be considered as a model parameter
+
+%horizontal structure that gives rises 
+%vertical shear is an input force
+%only porting vertical shear  
+%vertical shear doesn't depend on x 
+%not all results can be extrapolated to the global context
+
+In Appendix we make the connection between the linear problem in the
+VGSB to that in the global disk model.
 

### 2015-01-31 summary.tex
MAJOR REVISION UNDERWAY

@@
 
+%relation with kley simulations
+
 \subsection{Future extensions} 
@@
 
-\emph{Geometry.} Studying the VSI in the shearing box has an 
-inconsistency. Namely, the equilibrium temperature in the
+\emph{Geometry.} Extrapolating our results from the shearing box to a global disk  
+has an inconsistency. Namely, the equilibrium temperature in the
 box only depends on height (if at all). However, it is precisely the
 radial temperature (or entropy) gradient in the global disk that
-gives rise to vertical shear. This is not a serious issue for
+gives rise to vertical shear. We do not expect this to be a serious issue for 
 perturbations with small radial wavelengths, 

### 2015-01-31 appendix.tex
new appedix on local-global connection

@@
+\section{Local model for a global problem}
+Here, we discuss the relation between the linear problem in the
+VGSB framework and the global linear problem. We begin with the
+axisymmetric linearized fluid equations in cylindrical co-ordinates, including
+thermal relaxation, 
+\begin{align}
+\ii\sigma\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
+\left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
+    \delta v_z}{\p z}\right] + \frac{1}{t_c}\left(\delta P -
+  \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,\\
+  \ii\sigma\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
+  z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\\
+\ii\sigma\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
+\delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
+\ii\sigma\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
+r\frac{\p\Omega}{\p z}\delta v_z,\\
+\ii \sigma\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
+\frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},
+\end{align}
+where the coefficients $\hat{g}_c=1$ are inserted to label terms
+associated with the background radial disk structure.
+
+We assume perturbations have the form $W(z)\exp{\ii k_r r}$ such
+that $|k_rr|\gg1$, so that curvature terms can be ignored. For these
+small radial lengthscale perturbations we also approximate the
+background structure and radial gradients as uniform in $r$, and
+evaluate all cofficients in the above equations at the fiducial radius
+$r=r_0$. This procedure produces $\phi$ and $z$ momentum equations
+that are equilvalent to those for $y$ and $z$ momenta in the VGSB.
+
+Using the notation and disk models in the main text, the
+energy, density and $r$ momentum equations become 
+\begin{align}
+  \ii \sigma W  &=c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
+      z}\right|_0 \delta v_z + \frac{\gamma}{\Gamma} \frac{d\delta
+    v_z}{dz}\right) + \left[\ii k_rc_s^2(r_0,z)
+  \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
+      r}\right|_0\right]\delta v_r  +
+\frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\\
+\ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_r + \hat{g}_c
+  \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_r + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
+      z}\right|_0 \delta v_z + \frac{d\delta
+    v_z}{dz}\right), \\
+\ii \sigma \delta v_r & = \ii k_r W  -
+\hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
+  r}\right|_0Q - 2\Omega(r_0,z)\delta v_\phi,
+\end{align}
+where subscript $0$ denotes evaluation at $r=r_0$. To obtain the
+linear problem in the VGSB, we take the radially-local 
+approximation further and formally set $\hat{g}_c=0$, and define the
+new co-ordinate $x=r - r_0$ with $k_r\to k_x$. It is clear that
+this approximation improves with increasing $|k_x|$.
+
+We may also reverse this procedure by reinstating the `global
+correction' terms proportional to $\hat{g}_c$ in the linear VGSB
+equations. Of course, doing so implies the basic state is not uniform
+in $x$, which would be inconsistent with the full VGSB equations in
+equilibrium.
+
+%nevertheless we have calculated
+%plot growth rate v.s. bcool
+
 \section{Full governing equation for vertically isothermal disks}\label{adia_improve}

### 2015-02-02 appendix.tex
revised analysis part

@@
-\section{Local model for a global problem}
+\section{Local model for a global problem}\label{global_corr}
 Here, we discuss the relation between the linear problem in the

### 2015-02-02 linear_adia.tex
revised analysis part

@@
-\subsection{Adiabatic perturbations}\label{analytic_adia}
-We now consider % general $\gamma$
-perturbations in the adiabatic limit ($t_c,\beta\to\infty$). This is the opposite regime to
-isothermal perturbations. In this case the parameter $\chi = 1/\gamma$
-and Eq. \ref{vertiso_gov_nondim} can be written as 
-% We eliminate variables from 
-% Eq. \ref{lin_mass}---\ref{lin_energy} in favor of $\delta v_z$ 
-% and $\Delta \equiv \nabla\cdot\delta\bm{v}$ to obtain
-% \begin{align}
-%   &\Delta\left(1 + \frac{\gamma c_s^2 k_x^2}{D}\right) = \frac{d\delta
-%     v_z}{dz} + \frac{\ii k_x^2}{D}\left(\ii c_s^2 \frac{d\ln\rho}{dz}
-%     - \frac{r}{k_x}\frac{d\Omega^2}{dz}\right)\delta v_z,\label{adia_iso1}\\
-%   & -\frac{\sigma^2}{c_s^2}\delta v_z = \gamma \frac{d\Delta}{dz} +
-%   \frac{d}{dz}\left(\frac{d\ln\rho}{dz}\delta v_z\right) +
-%   \left(\gamma-1\right)\frac{d\ln\rho}{dz} \Delta.\label{adia_iso2}
-% \end{align}
-% %{\bf note: incompressible mode, set delta to zero. only works for
-% % constant gravity.}
-
-% We make the low-frequency approximation and the replacement 
-% $D\to \Omega_k^2$ in Eq. \ref{adia_iso1}---\ref{adia_iso2}. For
-% simplicity we do this before combining these equations later. We  
-% show in Appendix \ref{adia_improve} that making these approximations
-% after combining  Eq. \ref{adia_iso1}---\ref{adia_iso2} adds unecessary
-% complexity for thin disks. Here, we also use the fact that in the
-% global disk, 
-% \begin{align}
-%   r\frac{\p \Omega^2}{\p z} = -\frac{\p \ln\rho}{\p z} \frac{qc_s^2}{r},
-% \end{align}
-% for the power-law prescription of the midplane sound-speed for
-% $\Gamma=1$.
-
-% Making the above substitutions and eliminating $\Delta$ between Eq. \ref{adia_iso1}---
-% Eq. \ref{adia_iso2} we obtain, in terms of dimensionless variables defined
-% previously, an equation for $\delta v_z$,
-
+\subsection{Perturbations with thermal relaxation}\label{analytic_relax}
+We now introduce thermal relaxation ($\beta>0$). We pursue explicit
+solutions in the thin-disk approximation. The governing equation,
+Eq. \ref{vertiso_gov_nondim}, is then 
 \begin{align}
-  0 =& \dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
-    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime\notag\\
-  &+\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right)\right.\notag\\
-  &\phantom{0=}\left.-\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
-          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
-\end{align}
-
-We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
-integrate vertically, assume boundary terms vanish when integrating by
-parts, to obtain
-\begin{align}
-  &\hat{\sigma}^2\left(\frac{1}{\gamma} +
-    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
-  v_z|^2 d\zhat \notag\\
-  &=  \left(\frac{\gamma-1}{\gamma}\right)
-  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
-  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
-  v_z)^\prime|^2 d\zhat\notag\\
-  &\phantom{=} +
-  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
-      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
-    2}|\delta v_z|^2 d\zhat\notag\\
-  &\phantom{=} + \ii\epsilon q \hat{k}
-  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
-  \delta v_z d\zhat.\label{adia_integral}
-\end{align}
-When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
-$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
-expected, a sub-adiabatically stratified disk is stable in the absense
-of vertical shear. In the presence of vertical shear $q\neq0$ and
-$\hat{\sigma}$ is generally complex, which implies the possibility of
-instability.
-
-Note that when $\gamma>1$ and $q=0$, the term with $\ln\rho^{\prime
-  2}$ in the integrand (third term on the RHS) represents
-stabilization by the vertical entropy gradient (real buoyancy frequency). 
-It gives an imaginary contribution to $\hat{\sigma}$ when $q\neq0$,
-but this is a factor $|\epsilon/\hat{k}|$ smaller than the real
-(stabilizing) contribution, 
-since $\epsilon\ll 1$ for a thin disk and we (usually) consider
-$\hat{k}\gg1$. Thus, instability is expected to be associated with the 
-last term on the RHS of Eq. \ref{adia_integral}.
-
-\subsubsection{Explicit solutions in the thin-disk limit}\label{adia_explicit}
-We can construct solutions in the thin-disk limit, in which 
-$\ln(\rho/\rho_0) \simeq -\zhat^2/2$. Eq. \ref{adia_iso3} becomes   
-\begin{align}
-  0 = \delta v_z^{\prime\prime} - \zhat A \delta v_z^\prime + \left(B
-    - C \zhat^2\right)\delta v_z, \label{adia_iso4} 
+  0 = \delta v_z^{\prime\prime} - z A\delta v_z^\prime +
+  (B - C\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
 \end{align}
@@
   &A \equiv 1 + \ii \epsilon q \hat{k},\\
-  &B \equiv \hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right) -
-  \left(\frac{1}{\gamma} + \ii \hat{k} \epsilon q\right)\\
-  &C \equiv \frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
-      \epsilon q}{\hat{k}}\right).\label{adia_thin}
-\end{align}
-Let
-\begin{align}
-  \delta v_z(\zhat) =
-  g(\zhat)\exp\left(\frac{\alpha\zhat^2}{2}\right), \label{adia_ansatz}
-\end{align}
-where $\alpha$ is a constant to be chosen for convenience. Inserting
-Eq. \ref{adia_ansatz} into Eq. \ref{adia_thin} gives
-\begin{align}
-  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
-    \alpha\right)g
-  +\left(\alpha^2 - \alpha A - C\right)\zhat^2 g.
-\end{align}
-We choose $\alpha$ to make the coefficient of $\zhat^2g$
-vanish, and impose the vertical kinetic energy density
-$\rho|\delta v_z|^2$ to remain finite. Then assuming $g(\zhat)$ is a
-polynomial, we require  
-\begin{align}
-  \real\alpha < \frac{1}{2}. 
-\end{align}
-We therefore choose 
-\begin{align}
-  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right). 
-\end{align}
-
-We now have an equation in the same form as Eq. \ref{iso_ode3}:
-\begin{align}
-  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime +
-  \left(B + \alpha\right)g.
-\end{align}
-We seek polynomial solutions 
-\begin{align}
-  g(\zhat) = \sum_{m=0}^M b_m \zhat^m,
-\end{align}
-% We again seek power-series solutions 
-% \begin{align}
-%   g(\zhat) = \sum_{m=0}^\infty b_m \zhat^m,
-% \end{align}
-% which requires
-% \begin{align}
-%   (m+2)(m+1)b_{m+2} + \left[\left(2\alpha - A\right)m + B + \alpha\right]=0.
-% \end{align}
-which requires
-\begin{align}
-  B + \alpha = M\left(A-2\alpha\right).\label{adia_disp0}
-\end{align}
-Inserting the definition of $B$ we get
-\begin{align}
-  &\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) \notag\\ &=
-  \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
-  \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}\label{adia_disp}. 
-\end{align}
-This gives the eigenfrequency $\hat{\sigma}$ and completes the solution
-description.
-
-%Recalling that $B$ depends $\hat{\sigma}$, Eq. \ref{adia_disp} gives
-%the eigenfrequency as a function of disk paramters.
-
-\subsubsection{The case of $\gamma=2$}
-In principle, we can solve Eq. \ref{adia_disp} for the mode
-frequencies and growth rates as a function of the disk parameters and
-the radial wavenumber. This is complicated in 
-general, but simplifies for $\gamma=2$, for which the quantity
-$A^2+4C$ is real. In this case we find
-\begin{align}
-  \hat{\nu}^2 =
-  \frac{1}{2}\frac{(2M+1)}{\sqrt{1+2\hat{k}^2}}&\left[
-    \sqrt{1-\frac{4M(M+1)\hat{k}^2\epsilon^2q^2}{(2M+1)^2(1+2\hat{k}^2)}}\right.\notag\\
-  &\phantom{=}
-  \left.- \sqrt{1 - \frac{\hat{k}^2\epsilon^2q^2}{(1+2\hat{k}^2)}}\right].
-\end{align}
-In the limit of a thin disk $\epsilon\to0$ or weak shear $|q|\to0$, we
-have 
-\begin{align}
-  \hat{\nu}^2 = \frac{\epsilon^2
-    q^2}{4(2M+1)}\frac{\hat{k}^2}{(1+2\hat{k}^2)^{3/2}},\label{gam2_growth}
-\end{align}
-and the maximum growth rate occurs at $\hat{k}=\hat{k}_\mathrm{opt}=1$,  
-\begin{align}
-  \mathrm{max}(\hat{\nu}) &= \frac{|\epsilon
-    q|}{2(3^{3/4}\sqrt{2M+1})}\label{gam2_max_growth}.
-  % & \leq \frac{|\epsilon
-  % q|}{2\times3^{3/4}}\notag.
-\end{align}
-Unlike isothermal perturbations, Eq. \ref{gam2_max_growth} shows that
-maximum growth rate occurs at $M=0$, and growth rates decrease with increasing $M$. This
-is despite the fact that in the thin-disk approximation 
-$d\Omega^2/dz\propto z$, so there is infinite vertical shear
-available.
-
-The acceptable solution for $\alpha$ when $\gamma=2$ is
-\begin{align}
-  \alpha = \frac{1}{2}\left[\left(1+\ii\epsilon q \hat{k}\right) -
-    \sqrt{1 + \hat{k}^2\left(2-\epsilon^2 q^2\right)}\right].\label{gam2_alpha}
-\end{align}
-Since $\epsilon\ll 1$, $\real\alpha<0$, so the vertical velocity
-eventually decays for large heights. For $\hat{k}\gg 1$ the
-perturbation decays rapidly away from the midplane. This is again
-unlike isothermal perturbations, whose magnitude increase with $|z|$.
-
-%gauss decay means boundary conditions have no influence, all perts go
-%to zero
-
-\subsubsection{Comparison with $\gamma=1$}\label{adia_compare_gam1}
-We can compare growth rates obtained for $\gamma=2$ to those for
-$\gamma=1$ in the previous section. If we compare growth rates at their 
-respective optimum radial wavelengths, we find
-that for $\gamma=2$ the \emph{maximum} growth rate, occuring
-at $M=0$, is still $\sim2$ times smaller than the \emph{minimum} growth
-rate for $\gamma=1$. That is,
-\begin{align}
-  &\mathrm{max}(\hat{\nu}) \geq \frac{|\epsilon q|}{2\sqrt{1+|\epsilon
-      q|}} & \text{at }\hat{k} = \hat{k}_\mathrm{opt} \text{ for }   \gamma=1,  \\
-  &\mathrm{max}(\hat{\nu}) \leq \frac{|\epsilon q|}{2\times 3^{3/4}}
-           & \text{at }\hat{k} = \hat{k}_\mathrm{opt} \text{ for } \gamma=2.
-\end{align}
-
-We can also compare growth rates in the limit $\hat{k}\gg1$ using
-Eq. \ref{simple_growth} and Eq. \ref{gam2_growth},
-\begin{align}
-  &\hat{\nu} \geq \sqrt{\frac{|\epsilon q|}{2\hat{k}}}
-  &  \text{as }  \hat{k}\to\infty  \text{ for }  \gamma=1, \\
-  &\hat{\nu} \leq \frac{|\epsilon q|}{2^{7/4}\sqrt{\hat{k}}}
-  &\text{as }  \hat{k}\to\infty  \text{ for }  \gamma=2.\label{gam2_growth_rate}
-\end{align}
-Then the growth rate of perturbations with large $|\hat{k}|$ in a
-$\gamma=2$ disk is at most $2^{-5/4}\sqrt{|\epsilon q|}$ times that in a
-$\gamma=1$ disk. For $\epsilon = 0.1$ and $|q|=1$, this factor is
-$\sim 0.1$.
-
-In Fig. \ref{adia_growth} we plot the eigenfrequencies as a
-function of the radial wavenumber $\hat{k}$, for a range of adiabatic
-indices, using Eq. \ref{adia_disp}. The low-frequency approximation
-is questionable for $\hat{k}\lesssim4$ since
-$|\hat{\sigma}|^2\gtrsim0.2$. For large $\hat{k}$ the presence of a
-positive vertical entropy gradient is strongly stabilizing. This is
-because the vertical shear grows as $z$ away from the midplane (and
-has a maximum when the thin-disk approximation is relaxed), but square
-of the bouyancy frequency, which is stabilizing, grows as $z^2$ away
-from the midplane. %The entropy $P/\rho^\gamma$ formally diverges   
-%for $z\to\infty$ even without the thin-disk approximation.
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/rate_theory_om}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_nu}
-  \caption{Real frequency (top) and growth rate (bottom) of the fundamental VSI mode in vertically 
-    isothermal disks ($\Gamma=1$) subject to adiabatic perturbations,
-    calculated in thin disk approximation from Eq. \ref{adia_disp}
-    with $M=0$, $\epsilon=0.1$ and  
-    $|q|=1$. Frequencies are shown as a function of the radial
-    wavenumber for different values of the adiabatic index
-    $\gamma$. For $\gamma\equiv 1$ the perturbations are isothermal
-    and there is no vertical entropy gradient.\label{adia_growth}}  
-\end{figure}
-
-\subsection{Perturbations with thermal relaxation}\label{analytic_relax}
-Having examined the extreme limits of isothermal ($t_c\to0$) and
-adiabatic ($t_c\to\infty$) perturbations, we now generalize further to consider
-perturbations with intermediate thermodynamic responses with general
-values of $t_c$ or $\beta$.
-
-In this section we will pursue explicit solutions in the thin-disk
-approximation. The governing equation, Eq. \ref{vertiso_gov_nondim}, is
-then 
-\begin{align}
-  0 = \delta v_z^{\prime\prime} - z \widetilde{A}\delta v_z^\prime +
-  (\widetilde{B} - \widetilde{C}\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
-\end{align}
-with
-\begin{align}
-  &\widetilde{A} \equiv 1 + \ii \epsilon q \hat{k},\\
-  &\widetilde{B} \equiv \hat{\sigma}^2\left(\chi + \hat{k}^2\right) -
+  &B \equiv \hat{\sigma}^2\left(\chi + \hat{k}^2\right) -
   \left(\chi + \ii \epsilon q \hat{k}\right),\\
-  &\widetilde{C} \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
+  &C \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
     \epsilon q\hat{k}\right), 
@@
 $ in terms of dimensionless variables.
+
+We remark that Eq. \ref{nearly_iso_explicit} is in the same
+form as Eq. 41 in \cite{lubow93}, which describes adiabatic axisymmetric waves in
+a vertically isothermal disk without vertical shear. The two equations
+become identical after  setting $q=0$ (no vertical shear) and
+$\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
+and making the low-frequency, Keplerian disk approximation in
+\citeauthor{lubow93} after a change of variables.
 
@@
 
-One effect of introducing finite thermal relaxation is buoyancy. % We
-% show in the next section that buoyancy strongly stabilizes the
-% VSI.
+One effect of introducing finite thermal relaxation is buoyancy. 
 The $\zhat^2$ depedence in $\delta v_{z1}$ then makes sense because it
@@
 
-\subsubsection{Dispersion relation with thermal relaxation}\label{disp_relax}
-Eq. \ref{nearly_iso_explicit} is same form
-as the governing equation for the adiabatic case
-(Eq. \ref{adia_iso4}). However, the coefficients $\widetilde{B}$ and $\widetilde{C}$
-have a complicated dependence on $\hat{\sigma}$. 
-Applying the same solution method as in \S\ref{adia_explicit}, we 
-% Thus calculating the
-% eigenfrequency $\hat{\sigma}$ generally requires a numerical
-% approach, which we do  in \S\ref{disp_relax}. However, analytical
-% results may be obtained in the limit $\beta\to0$. 
-find the adiabatic dispersion relation,
-Eq. \ref{adia_disp0}, generalizes to a $6^\mathrm{th}$ order
-polynomial in $\hat{\sigma}$ with finite thermal relaxation:
+\subsubsection{Dispersion relation with thermal
+  relaxation}\label{disp_relax}
+To obtain a dispersion relation similar to Eq. \ref{sig2_iso} for
+isothermal perturbations, we first write 
+\begin{align}
+  \delta v_z(\zhat) =
+  g(\zhat)\exp\left(\frac{\alpha\zhat^2}{2}\right), \label{adia_ansatz}
+\end{align}
+where $\alpha$ is a constant to be chosen for convenience. Inserting
+Eq. \ref{adia_ansatz} into Eq. \ref{nearly_iso_explicit} gives
+\begin{align}
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
+    \alpha\right)g
+  +\left(\alpha^2 - \alpha A - C\right)\zhat^2 g.
+\end{align}
+We choose $\alpha$ to make the coefficient of $\zhat^2g$
+vanish, and impose the vertical kinetic energy density
+$\rho|\delta v_z|^2$ to remain finite. Then assuming $g(\zhat)$ is a
+polynomial, we require  
+\begin{align}
+  \real\alpha < \frac{1}{2}. 
+\end{align}
+We choose 
+\begin{align}
+  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right),  
+\end{align} 
+by considering the case $q=0$ and $\beta\to\infty$ as examined in
+\cite{lubow93}.
+
+We now have an equation in the same form as that for isothermal
+perturbations (Eq. \ref{iso_ode3}):
+\begin{align}
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime +
+  \left(B + \alpha\right)g.
+\end{align}
+We seek polynomial solutions 
+\begin{align}
+  g(\zhat) = \sum_{m=0}^M b_m \zhat^m,
+\end{align}
+which requires
+\begin{align}
+  B + \alpha = M\left(A-2\alpha\right).\label{adia_disp0}
+\end{align}
+% We remark that for $q=0$ and $\beta\to\infty$, Eq. \ref{adia_disp0} is
+% the same as Eq. 54 in \cite{lubow93} when the low-frequency, Keplerian
+% approximation is made in the latter. 
+% Eq. \ref{nearly_iso_explicit} is same form
+% as the governing equation for the adiabatic case
+% (Eq. \ref{adia_iso4}). However, the coefficients $\widetilde{B}$ and $\widetilde{C}$
+% have a complicated dependence on $\hat{\sigma}$. 
+% Applying the same solution method as in \S\ref{adia_explicit}, we 
+% find the adiabatic dispersion relation, 
+This is a $6^\mathrm{th}$ order 
+polynomial in $\hat{\sigma}$:  
 \begin{align}
@@
 where the coefficients $c_l$ are given in Appendix \ref{relax_coeff}.
-Eq. \ref{relax_disp} contrasts to Eq. \ref{adia_disp}, which is
-quadratic in $\hat{\sigma}$ for adiabatic perturbations.
+% Eq. \ref{relax_disp} contrasts to Eq. \ref{adia_disp}, which is
+% quadratic in $\hat{\sigma}$ for adiabatic perturbations.
 
@@
 Fig. \ref{relax_disp_fig} we plot the growth rate of the fundamental
-($M=0$) VSI mode as a function of $k_xH_\mathrm{iso}$ for a range of
+($M=0$) VSI mode as a function of $\khat$ for a range of
 dimensionless thermal relaxation timescales $\beta$, in a disk with
-$\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.  Similarly, 
-in Fig. \ref{relax_bcrit} we solve the dispersion relation for the
-critical $\beta$ such that the growth rate of the fundamental VSI mode
-is zero.
-
-Fig. \ref{relax_disp_fig}---\ref{relax_bcrit} shows that a growing fundamental
-VSI mode exists both in the isothermal ($\beta\to0$) and adiabatic
-($\beta\to\infty$) limits, although the latter growth rates are
-smaller by at least a factor of two. Furthermore, the adiabatic VSI is 
-is limited to a thin layer about the disk midplane so it may be
-considered stable. Then Fig. \ref{relax_bcrit} shows that 
-$\beta\lesssim 0.1$ is required for instability in this disk model. This critical
-timescale is indepedent of $k_x$ for
-$k_xH_\mathrm{iso}\gtrsim10$. The top panel of Fig. \ref{relax_disp_fig} is also
-consistent with the previous
-discussion that show thermal relaxation has no effect for
-$k_x\to0$ or $k_x\to\infty$ as $\beta\to0$.
+$\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. Increasing $\beta$ even
+slightly above zero significantly reduces the growth rates for
+$\khat\gg 1$. (There is little effect at small $\khat$ but this case is
+not of interest because the local model breaks down.) Thus, 
+a growing fundamental VSI mode exists only for nearly
+isothermal perturbations ($\beta\ll 1$).  The plot is also consistent with our previous
+discussion that there exists an intermediate wavenumber for which
+finite thermal relaxation has maximal effect (for small $\beta$). 
+%$k_x\to0$ or $k_x\to\infty$ as $\beta\to0$.
 
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/rate_theory_grow_relax}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_grow_relax2}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm 0cm]{figures/rate_theory_grow_relax}
+%  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_grow_relax2}
   \caption{Growth rate of the fundamental VSI mode in a disk with
     dimensionless thermal relaxation timescales
-    $\beta\in[10^{-3},1]$ (top) and $\beta\in[1,10^3]$ (bottom).
-Other disk parameters are
+    $\beta\in[10^{-3},1]$.  Other disk parameters are
     $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The growth rates are
@@
 
-%ate_theory, eps=0.05, smallq=-1d0, gmma=1.4, bmax=1.0, ncases=11,
+% ate_theory, eps=0.05, smallq=-1d0, gmma=1.4, bmax=1.0, ncases=11,
 %mmode=0.0, root=1, legend=[0.102,0.14,5.4,0.355], yrange=[0,6]
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/bcrit_theory}
-  \caption{Condition on the thermal relaxation timescale $\beta$ for the existence of unstable fundamental VSI
-    modes in a vertically isothermal disk with $\epsilon=0.05$, $q=-1$
-    and $\gamma=1.4$. For sufficiently small $\beta$ the isothermal VSI opertates,
-    while for sufficiently large $\beta$ the adiabatic VSI
-    operates. For intermediate values of $\beta$ the VSI growth rate
-    is negative (i.e. stabilized). 
-    \label{relax_bcrit}}  
-\end{figure}
+\subsubsection{Upper limit on the thermal relaxation timescale for the
+  fundamental VSI mode}\label{iso_vsi_beta_crit}
 
-\subsubsection{Upper limit on the thermal relaxation timescale for the
-  fundamental isothermal VSI mode}\label{iso_vsi_beta_crit}
-Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
-fundamental isothermal VSI tends to a constant for large
-$\khat$. Using this clue, we can calculate this critical $\beta$ by
-considering $\khat\to\infty$. We proceed as follows.
-
+Here we estimate the maximum thermal relaxation timescale
+$\beta$ that allows the fundamental VSI for $\khat\gg1$.   
+% Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
+% fundamental isothermal VSI tends to a constant for large
+% $\khat$. Using this clue, we can calculate this critical $\beta$ by
+% considering $\khat\to\infty$. We proceed as follows. 
 At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
@@
 \begin{align}
-  \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \khat\hat{\omega} \text{ finite}.
+  \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
 \end{align}
@@
 \begin{align}
-  \khat\hat{\omega} = \frac{(\gamma-1)\beta}{\epsilon q}
+  \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
 \end{align}
@@
   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
-        q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2}. 
+        q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
 \end{align}
+at marginal stability for perturbations with large $\khat$. 
 For a thin disk $\epsilon\ll 1$, we infer 
@@
 \end{align}
-is required for the fundamental isothermal VSI to operate.  For
-$\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$ we find
-$\beta_\mathrm{crit}=0.125$, in agreement with Fig. \ref{relax_bcrit}
-at large $\khat$.
+is required for the fundamental VSI to operate.
+
+In Fig. \ref{relax_bcrit} we plot the critical thermal relaxation
+timescale obtained from solving Eq. \ref{relax_disp_fund} numerically 
+for a disk with $\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$, for which
+we have $\beta_\mathrm{crit}=0.125$ from Eq. \ref{iso_vsi_cond}. Our
+upper estimate agrees well with the numerical solution for $\khat\gtrsim
+10$. From our discussion for isothermal perturbations
+(\S\ref{iso_discuss}), we expect $\beta<\beta_\mathrm{crit}$ to be a
+sufficient condition for the VSI.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/bcrit_theory}
+  \caption{Condition on the thermal relaxation timescale $\beta$ for
+    the existence of unstable fundamental VSI mode in a vertically
+    isothermal disk with $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.  
+    The dotted line denotes the asymptotic value
+    $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$ for large $\khat$. 
+    \label{relax_bcrit}}  
+\end{figure}
 
@@
 % solution is $\beta_0 = 10.33$ and $\hat{\omega}_0 = 0.74$.
+
+\subsection{Spurious growth of adiabatic perturbations in the local
+  model}\label{analytic_adia} 
+For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
+simply use the Soilberg-Hoiland critera to assess stability
+(\S\ref{solberg}). Nevertheless, we briefly consider this case in the
+local framework to demonstrate its limitations. For $\beta\to\infty$
+the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
+be written as
+
+\begin{align}
+  0 =& \dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
+    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime\notag\\
+  &+\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right)\right.\notag\\
+  &\phantom{0=}\left.-\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
+          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
+\end{align}
+
+We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
+integrate vertically, assume boundary terms vanish when integrating by
+parts, to obtain
+\begin{align}
+  &\hat{\sigma}^2\left(\frac{1}{\gamma} +
+    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
+  v_z|^2 d\zhat \notag\\
+  &=  \left(\frac{\gamma-1}{\gamma}\right)
+  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
+  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
+  v_z)^\prime|^2 d\zhat\notag\\
+  &\phantom{=} +
+  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
+      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
+    2}|\delta v_z|^2 d\zhat\notag\\
+  &\phantom{=} + \ii\epsilon q \hat{k}
+  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
+  \delta v_z d\zhat.\label{adia_integral}
+\end{align}
+When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
+$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
+expected, a sub-adiabatically stratified disk is stable in the absense
+of vertical shear.
+
+% If we follow through the solution method as before, we obtain a
+% dispersion relation 
+% \begin{align}
+%   &\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) \notag\\ &=
+%   \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
+%   \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}\label{adia_disp}.
+% \end{align}
+% This equation is the same as Eq. 54 in \cite{lubow93} when $q=0$ and
+% the low-frequency, Keplerian approximation is made for the latter.
+
+However, in the presence of vertical shear $q\neq0$,
+Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
+$\khat$, which implies the possibility of instability, regardless of
+how much $\gamma$ deviates from unity. This is contradicts the second 
+Solberg-Hoiland criterion (Eq. \ref{solberg2}) and is a result of the
+radially-local approximation (Appendix \ref{global_corr}). To 
+avoid this, we must restrict our attention to short thermal relaxation
+timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
+($\gamma=\Gamma$) when allowing vertical shear.
+
+% Note that when $\gamma>1$ and $q=0$, the term with $\ln\rho^{\prime
+%   2}$ in the integrand (third term on the RHS) represents
+% stabilization by the vertical entropy gradient (real buoyancy frequency). 
+% It gives an imaginary contribution to $\hat{\sigma}$ when $q\neq0$,
+% but this is a factor $|\epsilon/\hat{k}|$ smaller than the real
+% (stabilizing) contribution, 
+% since $\epsilon\ll 1$ for a thin disk and we (usually) consider
+% $\hat{k}\gg1$. Thus, instability is expected to be associated with the 
+% last term on the RHS of Eq. \ref{adia_integral}.
+
+% \subsubsection{Explicit solutions in the thin-disk limit}\label{adia_explicit}
+% We can construct solutions in the thin-disk limit, in which 
+% $\ln(\rho/\rho_0) \simeq -\zhat^2/2$. Eq. \ref{adia_iso3} becomes   
+% \begin{align}
+%   0 = \delta v_z^{\prime\prime} - \zhat A \delta v_z^\prime + \left(B
+%     - C \zhat^2\right)\delta v_z, \label{adia_iso4} 
+% \end{align}
+% with
+% \begin{align}
+%   &A \equiv 1 + \ii \epsilon q \hat{k},\\
+%   &B \equiv \hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right) -
+%   \left(\frac{1}{\gamma} + \ii \hat{k} \epsilon q\right)\\
+%   &C \equiv \frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
+%       \epsilon q}{\hat{k}}\right).\label{adia_thin}
+% \end{align}
+
+% Inserting the definition of $B$ we get
+% \begin{align}
+%   &\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) \notag\\ &=
+%   \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
+%   \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}\label{adia_disp}. 
+% \end{align}
+% This gives the eigenfrequency $\hat{\sigma}$ and completes the solution
+% description.
+
+% %Recalling that $B$ depends $\hat{\sigma}$, Eq. \ref{adia_disp} gives
+% %the eigenfrequency as a function of disk paramters.
+
+% \subsubsection{The case of $\gamma=2$}
+% In principle, we can solve Eq. \ref{adia_disp} for the mode
+% frequencies and growth rates as a function of the disk parameters and
+% the radial wavenumber. This is complicated in 
+% general, but simplifies for $\gamma=2$, for which the quantity
+% $A^2+4C$ is real. In this case we find
+% \begin{align}
+%   \hat{\nu}^2 =
+%   \frac{1}{2}\frac{(2M+1)}{\sqrt{1+2\hat{k}^2}}&\left[
+%     \sqrt{1-\frac{4M(M+1)\hat{k}^2\epsilon^2q^2}{(2M+1)^2(1+2\hat{k}^2)}}\right.\notag\\
+%   &\phantom{=}
+%   \left.- \sqrt{1 - \frac{\hat{k}^2\epsilon^2q^2}{(1+2\hat{k}^2)}}\right].
+% \end{align}
+% In the limit of a thin disk $\epsilon\to0$ or weak shear $|q|\to0$, we
+% have 
+% \begin{align}
+%   \hat{\nu}^2 = \frac{\epsilon^2
+%     q^2}{4(2M+1)}\frac{\hat{k}^2}{(1+2\hat{k}^2)^{3/2}},\label{gam2_growth}
+% \end{align}
+% and the maximum growth rate occurs at $\hat{k}=\hat{k}_\mathrm{opt}=1$,  
+% \begin{align}
+%   \mathrm{max}(\hat{\nu}) &= \frac{|\epsilon
+%     q|}{2(3^{3/4}\sqrt{2M+1})}\label{gam2_max_growth}.
+%   % & \leq \frac{|\epsilon
+%   % q|}{2\times3^{3/4}}\notag.
+% \end{align}
+% Unlike isothermal perturbations, Eq. \ref{gam2_max_growth} shows that
+% maximum growth rate occurs at $M=0$, and growth rates decrease with increasing $M$. This
+% is despite the fact that in the thin-disk approximation 
+% $d\Omega^2/dz\propto z$, so there is infinite vertical shear
+% available.
+
+% The acceptable solution for $\alpha$ when $\gamma=2$ is
+% \begin{align}
+%   \alpha = \frac{1}{2}\left[\left(1+\ii\epsilon q \hat{k}\right) -
+%     \sqrt{1 + \hat{k}^2\left(2-\epsilon^2 q^2\right)}\right].\label{gam2_alpha}
+% \end{align}
+% Since $\epsilon\ll 1$, $\real\alpha<0$, so the vertical velocity
+% eventually decays for large heights. For $\hat{k}\gg 1$ the
+% perturbation decays rapidly away from the midplane. This is again
+% unlike isothermal perturbations, whose magnitude increase with $|z|$.
+
+% %gauss decay means boundary conditions have no influence, all perts go
+% %to zero
+
+% \subsubsection{Comparison with $\gamma=1$}\label{adia_compare_gam1}
+% We can compare growth rates obtained for $\gamma=2$ to those for
+% $\gamma=1$ in the previous section. If we compare growth rates at their 
+% respective optimum radial wavelengths, we find
+% that for $\gamma=2$ the \emph{maximum} growth rate, occuring
+% at $M=0$, is still $\sim2$ times smaller than the \emph{minimum} growth
+% rate for $\gamma=1$. That is,
+% \begin{align}
+%   &\mathrm{max}(\hat{\nu}) \geq \frac{|\epsilon q|}{2\sqrt{1+|\epsilon
+%       q|}} & \text{at }\hat{k} = \hat{k}_\mathrm{opt} \text{ for }   \gamma=1,  \\
+%   &\mathrm{max}(\hat{\nu}) \leq \frac{|\epsilon q|}{2\times 3^{3/4}}
+%            & \text{at }\hat{k} = \hat{k}_\mathrm{opt} \text{ for } \gamma=2.
+% \end{align}
+
+% We can also compare growth rates in the limit $\hat{k}\gg1$ using
+% Eq. \ref{simple_growth} and Eq. \ref{gam2_growth},
+% \begin{align}
+%   &\hat{\nu} \geq \sqrt{\frac{|\epsilon q|}{2\hat{k}}}
+%   &  \text{as }  \hat{k}\to\infty  \text{ for }  \gamma=1, \\
+%   &\hat{\nu} \leq \frac{|\epsilon q|}{2^{7/4}\sqrt{\hat{k}}}
+%   &\text{as }  \hat{k}\to\infty  \text{ for }  \gamma=2.\label{gam2_growth_rate}
+% \end{align}
+% Then the growth rate of perturbations with large $|\hat{k}|$ in a
+% $\gamma=2$ disk is at most $2^{-5/4}\sqrt{|\epsilon q|}$ times that in a
+% $\gamma=1$ disk. For $\epsilon = 0.1$ and $|q|=1$, this factor is
+% $\sim 0.1$.
+
+% In Fig. \ref{adia_growth} we plot the eigenfrequencies as a
+% function of the radial wavenumber $\hat{k}$, for a range of adiabatic
+% indices, using Eq. \ref{adia_disp}. The low-frequency approximation
+% is questionable for $\hat{k}\lesssim4$ since
+% $|\hat{\sigma}|^2\gtrsim0.2$. For large $\hat{k}$ the presence of a
+% positive vertical entropy gradient is strongly stabilizing. This is
+% because the vertical shear grows as $z$ away from the midplane (and
+% has a maximum when the thin-disk approximation is relaxed), but square
+% of the bouyancy frequency, which is stabilizing, grows as $z^2$ away
+% from the midplane. %The entropy $P/\rho^\gamma$ formally diverges   
+% %for $z\to\infty$ even without the thin-disk approximation.
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/rate_theory_om}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_nu}
+%   \caption{Real frequency (top) and growth rate (bottom) of the fundamental VSI mode in vertically 
+%     isothermal disks ($\Gamma=1$) subject to adiabatic perturbations,
+%     calculated in thin disk approximation from Eq. \ref{adia_disp}
+%     with $M=0$, $\epsilon=0.1$ and  
+%     $|q|=1$. Frequencies are shown as a function of the radial
+%     wavenumber for different values of the adiabatic index
+%     $\gamma$. For $\gamma\equiv 1$ the perturbations are isothermal
+%     and there is no vertical entropy gradient.\label{adia_growth}}  
+% \end{figure}

### 2015-02-02 linear_iso.tex
revised analysis part

@@
 where $\Gamma=1$, then $c_s$ is a constant in the shearing box. We
-examine separately the limiting case of isothermal perturbations
-($\gamma\equiv1$ and any value of $t_c$, or any value of $\gamma$ with
-$t_c = 0$) and adiabatic perturbations ($\gamma\neq 1$ and
-$t_c\to\infty$), before considering the effect of thermal relaxation.
+first examine the limiting case of isothermal perturbations
+($\gamma\equiv1$ and any value of $\beta$, or any value of $\gamma$ with
+$\beta = 0$) before considering the effect of thermal relaxation. In the
+final section we consider adiabatic disks ($\gamma>1$ and
+$\beta\to\infty$) to demonstrate an important limitation of the 
+local model.
 
@@
   elsewhere. For clarity, we replace $\kappa^2$ with $\Omega_k^2$.
-  % , although the discussion
-  % is unchanged if we replaced 
-  % $\kappa^2$ with $\kappa^2(0)$, for example.    
+
 \end{enumerate}
@@
     v_z}{dz} \notag\\
-  &+ \left[\sigma^2\left(\frac{k_x^2}{\Omega_k^2} +
-      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
+  % &+ \left[\sigma^2\left(\frac{k_x^2}{\Omega_k^2} +
+  %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
+  &+ \left[\frac{\sigma^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_k^2} +
+      \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
         q}{\Omega_k^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
@@
 
-We will consider sound-speed profiles with $q<0$ in the global 
-disk. Then $\Omega^2(z)$ decreases away from the midplane,
-and may reach small values for sufficiently large $|z|$, which
-invalidates the low-frequency approximation. In principle, then, the
-above approximations implicitly limits us to consider thin disks ($\epsilon
-\ll 1$) with vertical domain sizes not too large, so that the rotation
-is nearly-Keplerian.
+% We will consider sound-speed profiles with $q<0$ in the global 
+% disk. Then $\Omega^2(z)$ decreases away from the midplane,
+% and may reach small values for sufficiently large $|z|$, which
+% invalidates the low-frequency approximation. In principle, then, the
+% above approximations implicitly limits us to consider thin disks ($\epsilon
+% \ll 1$) with vertical domain sizes not too large, so that the rotation
+% is nearly-Keplerian.
+
+Note that we can generalize Eq. \ref{vertiso_gov_nondim} to other
+(non-Keplerian) rotation profiles by redfining $\khat$
+ as $k_xc_s/\kappa$, $\hat{\sigma}$ as $\sigma/\kappa$, and 
+ $\epsilon$ as $c_s/r\kappa$.
+
+% Consider rotation profiles with radial dependence
+% $\propto r^{-m}$. Then the same dimensionless governing equation 
+% (Eq. \ref{vertiso_gov_nondim}) holds, provided we redfine $\khat$
+% as $k_xc_s/\kappa$, $\hat{\sigma}$ as $\sigma/\kappa$, and reset
+% $\epsilon$ to $\epsilon/sqrt{4-2m}$.
 
@@
 relaxation timescale $\beta$, or for any adiabatc index $\gamma$ when
-$\beta=0$. In this case Eq. \ref{ode_Q} implies 
+$\beta=0$. In either case $\chi=1$ and Eq. \ref{ode_Q} implies 
 \begin{align}

### 2015-02-02 setup.tex
revised analysis part

@@
 characteristic disk aspect-ratio.
-% From $\epsilon$ we can define a
-% characteristic scale-height  
-% \begin{align}
-%   \Hiso \equiv \epsilon r.% \quad\quad(\Gamma \equiv 1),   
-% \end{align} 
+
 The global sound-speed $c_s(r,z)$ is given by
@@
 
-% {\bf need to mention: generally consider sub-keplerian disks -> caveat is that it's entirely possble for angular  
-% velocity to reach zero at finite height. give example later for isothermal disk. this limits domain size, but this critical height is usually
-% very large for thin disks}
-
 \subsection{Vertically isothermal disks}
@@
 \subsection{Stability condition against axisymmetric adiabatic
-  perturbations}
+  perturbations}\label{solberg}
 A rotating flow is stable against axisymmetric adiabatic perturbations
@@
 generality,   
-\begin{align}
+\begin{align}\label{solberg2}
   \gamma - \Gamma > \frac{s\epsilon^2}{\Gamma}
@@
 sound-speed are assumed to depend only on $z$, by evaluating these
-quantities at $r=r_0$ in the global disk. The background rotation is
-approximated by a linear shear in $x$, but left unaltered in
-$z$. However, the vertical shear $\p_z\Omega$ is also assumed to only
-depend on $z$.
+quantities at $r=r_0$ in the global disk. The background rotation
+$\Omega$ is approximated by a linear shear in $x$, but left unaltered
+in $z$. However, the vertical shear $\p_z\Omega$ is also assumed to
+only depend on $z$.
 
@@
 parameter.
-
-\subsection{Limitations}
-Before using the VGSB to study the VSI, it is very important to be
-aware of several limitations of the box model. The VGSB momentum  
-equation have the same form as that for the standard shearing box 
-\citep{goldreich65}, except for the appearence of the term 
-$v_zr_0d\Omega_0/dz$. This term appears in the global stability
-problem and is the only term associated with vertical shear that is 
-ported from the global disk into the local model.
-
-%horizontal structure not ported
-
-Note that, within the VGSB framework, the vertical shear 
-$d\Omega_0/dz$ is a free parameter (as are $\Omega_0$ and $S_0$). 
-This is different to the global problem because in that case, the
-vertical shear is a result of the disk radial structure, which is
-ignored in the VGSB equilibria. 
-%because there is only one
-%equilibrium condition, which determines the vertical density/pressure
-%structure. 
-In effect, we are studying dynamics in the standard shearing box
-subject to an additional azimuthal force $-v_zr_0 d \Omega_0/dz$. This
-force is inspired by the vertical shear in the global disk, and its
-functional form is taken from global equilibrium. 
-%but should be considered as a model parameter
-
-%horizontal structure that gives rises 
-%vertical shear is an input force
-%only porting vertical shear  
-%vertical shear doesn't depend on x 
-%not all results can be extrapolated to the global context
-
-In Appendix we make the connection between the linear problem in the
-VGSB to that in the global disk model.
 
@@
 \end{align}
- % It is convenient to express $\delta v_x$ as a function of $W$ and
- % $\delta v_z$,
- % \begin{align}
- %   \delta v_x = -\frac{1}{D}\left(\sigma k_x W +
- %     r\frac{d\Omega^2}{dz}\delta v_z\right),
- % \end{align}
 where
@@
 \end{align}
-% Eq. \ref{ode_w}---\ref{ode_Q} are convenient for numerical work, 
-% but alternative forms are more suitable for analytical discussion, and 
-% will be derived later.
 
-%These boundary conditions are also numerically
-%convenient because they can be recast to contain the eigenfrequency
-%$\sigma$. This means that the the full set of ODEs, including boundary
-%conditions can be written in the form $L(\bm{U}) = \sigma R (\bm{U})$
+\subsection{Limitations}
+Before using the above equations to study the VSI, it is very important to be
+aware of several limitations of the VGSB model. The VGSB momentum  
+equations have the same form as that for the standard shearing box 
+\citep{goldreich65}, except for the appearence of the term 
+$v_zr_0d\Omega_0/dz$. This term appears in the global stability
+problem and is the only term associated with vertical shear that is 
+ported from the global disk into the local model.
+
+Note that, within the VGSB framework, the vertical shear 
+$d\Omega_0/dz$ is a free parameter (as are $\Omega_0$ and $S_0$). 
+This is different to the global problem because in that case, the
+vertical shear is a result of the disk radial structure, which is
+ignored in the VGSB equilibria. 
+In effect, we are studying dynamics in the standard shearing box
+subject to an additional azimuthal force $-v_zr_0 d \Omega_0/dz$. This
+force is inspired by the vertical shear in the global disk, and its
+functional form is taken from global equilibrium.
+
+%horizontal structure that gives rises 
+%vertical shear is an input force
+%only porting vertical shear  
+%vertical shear doesn't depend on x 
+%not all results can be extrapolated to the global context
+
+In Appendix \ref{global_corr} we make the connection between the
+linear problem in the VGSB to that in the global disk model. We
+emphasize that not all calculations in the local model can be 
+consistently interpreted in the global context. However, for
+perturbations with small radial lengthscales (large $|k_x|$) and fast
+thermal relaxation ($\beta\ll1$), which characterizes the VSI, the
+VGSB framework is adequate.

### 2015-02-02 linear_adia.tex
revised analysis part

@@
 \subsection{Spurious growth of adiabatic perturbations in the local
-  model}\label{analytic_adia} 
+  model with vertical shear}\label{analytic_adia} 
 For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may

### 2015-02-03 appendix.tex
updating

@@
       r}\right|_0\right]\delta v_r  +
-\frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\\
+\frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
 \ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_r + \hat{g}_c
@@
       z}\right|_0 \delta v_z + \frac{d\delta
-    v_z}{dz}\right), \\
+    v_z}{dz}\right),\label{gcorr_terms2} \\
 \ii \sigma \delta v_r & = \ii k_r W  -
 \hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
-  r}\right|_0Q - 2\Omega(r_0,z)\delta v_\phi,
+  r}\right|_0Q - 2\Omega(r_0,z)\delta v_\phi,\label{gcorr_terms3}
 \end{align}
@@
 linear problem in the VGSB, we take the radially-local 
-approximation further and formally set $\hat{g}_c=0$, and define the
-new co-ordinate $x=r - r_0$ with $k_r\to k_x$. It is clear that
-this approximation improves with increasing $|k_x|$.
+approximation further and formally set $\hat{g}_c=0$, $r\to x$ and
+$k_r\to k_x$. It is clear that this approximation improves with
+increasing $|k_x|$.
 
-We may also reverse this procedure by reinstating the `global
+\subsection{Effect of global radial gradients}
+We may also reverse the above procedure by reinstating the `global
 correction' terms proportional to $\hat{g}_c$ in the linear VGSB
@@
 in $x$, which would be inconsistent with the full VGSB equations in
-equilibrium.
+equilibrium. Nevertheless, we illustrate in Fig. \ref{gcorr_compare} the effect of
+these global radial gradients on the growth rate of the fundamental
+VSI.
+
+The `global correction' terms do not change the qualitative dependence
+of the fundamental VSI growth rates on the thermal relaxation
+timescale. The difference is negligble for $\beta\lesssim 1$ (i.e.
+fast thermal relaxation). The critical relaxation timescale for the
+fundamental VSI is unaffected by these global radients.
+
+However, in the adiabatic limit $\beta\to\infty$ the linear VGSB
+equations give a positive growth rate as discussed in
+\S\ref{analytic_adia}, which is inconsistent with the second
+Solberg-Hoiland criterion (Eq. \ref{solberg2}). The correct behaviour
+($\nu\to0$ as $\beta\to\infty$) is obtained with global correction.
+
+It is interesting to note that there is a thermal relaxation timescale
+that maximizes the mode decay rate. Here, it is $\beta\simeq2$ or
+about $0.3$ orbital periods. This is, in fact, consistent with Fig. 12
+of \cite{nelson13}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
+  0cm]{figures/gcorr_compare} 
+  \caption{Growth rate of
+    the fundamental VSI mode as a function of the thermal
+    relaxatin time $\beta$, computed from the linear VGSB equations used in
+    the main text (dotted) and that including additional global
+    radial gradient terms in  Eq. \ref{gcorr_terms1}---\ref{gcorr_terms3} (solid). 
+    The disk parameters are $\Gamma=1.011$,
+    $\gamma=1.4$ and $(p,q,\epsilon)=(-1.5,-1,0.05)$. The perturbation
+    wavenumber is $\khat=20$. 
+    {\bf waiting for
+      high resolution} 
+    \label{gcorr_compare}}  
+\end{figure}
 
@@
 
-\section{Alternative inference of instability in the presence of weak
-  vertical shear}\label{pert_theory}
-We can also investigate the destabilizing effect of vertical shear
-without explicitly solving the full governing equation,
-Eq. \ref{iso_ode3}. We first
-consider a system with $q\equiv0$, for which the eigenfunctions are
-Hermite polynomials and the real eigenfrequencies are known. We then
-perturb this system by introducing a weak vertical shear ($|q|\ll1$)
-and linearize the governing equation. This procedure is
-\begin{align}   
-  q \to 0 + \delta q,\quad
-  W \to \He_n + \delta W,\quad
-  \hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}. 
-\end{align}
-%where for this exercise we restore $\sigma$ as the eigenfrequency.
+% \section{Alternative inference of instability in the presence of weak
+%   vertical shear}\label{pert_theory}
+% We can also investigate the destabilizing effect of vertical shear
+% without explicitly solving the full governing equation,
+% Eq. \ref{iso_ode3}. We first
+% consider a system with $q\equiv0$, for which the eigenfunctions are
+% Hermite polynomials and the real eigenfrequencies are known. We then
+% perturb this system by introducing a weak vertical shear ($|q|\ll1$)
+% and linearize the governing equation. This procedure is
+% \begin{align}   
+%   q \to 0 + \delta q,\quad
+%   W \to \He_n + \delta W,\quad
+%   \hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}. 
+% \end{align}
+% %where for this exercise we restore $\sigma$ as the eigenfrequency.
 
-Linearizing the integral relation Eq. \ref{integral_relation1} in the
-thin-disk limit, and taking the
-imaginary part, we have
-\begin{align}
-  2\hat{\sigma}\imag(\delta\hat{\sigma})
-  \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-  \He_n^2(\hat{z}) d\hat{z}
-  = \delta q \epsilon \hat{k} 
-  \int_{-\infty}^{\infty}
-  w(\zhat)\hat{z}\He_n(\hat{z})\He_n^\prime(\zhat) d\hat{z}
-\end{align}
-Recognizing $\hat{z} = \He_1(\hat{z})$ and using $\He_n^\prime = n
-\He_{n-1}$ we have
- \begin{align}
-   2\hat{\sigma}\imag(\delta\hat{\sigma})
-   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-   \He_n^2(\hat{z}) d\hat{z}
-   =\delta q \epsilon \hat{k} 
-   \int_{-\infty}^{\infty}
-   w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
- \end{align}
+% Linearizing the integral relation Eq. \ref{integral_relation1} in the
+% thin-disk limit, and taking the
+% imaginary part, we have
+% \begin{align}
+%   2\hat{\sigma}\imag(\delta\hat{\sigma})
+%   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
+%   \He_n^2(\hat{z}) d\hat{z}
+%   = \delta q \epsilon \hat{k} 
+%   \int_{-\infty}^{\infty}
+%   w(\zhat)\hat{z}\He_n(\hat{z})\He_n^\prime(\zhat) d\hat{z}
+% \end{align}
+% Recognizing $\hat{z} = \He_1(\hat{z})$ and using $\He_n^\prime = n
+% \He_{n-1}$ we have
+%  \begin{align}
+%    2\hat{\sigma}\imag(\delta\hat{\sigma})
+%    \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
+%    \He_n^2(\hat{z}) d\hat{z}
+%    =\delta q \epsilon \hat{k} 
+%    \int_{-\infty}^{\infty}
+%    w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
+%  \end{align}
 
-Finally, we evaluate the integrals using the results
-\begin{align}
-  \int_{-\infty}^{\infty}
-  w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl} \quad
+% Finally, we evaluate the integrals using the results
+% \begin{align}
+%   \int_{-\infty}^{\infty}
+%   w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl} \quad
+% % \end{align}
+% \text{and} \quad
+% % \begin{align}
+%   \int_{-\infty}^{\infty}
+%   w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi =
+%   \frac{\sqrt{2\pi}m!k!l!}{(j-m)!(j-k)!(j-l)!}, 
 % \end{align}
-\text{and} \quad
+% where $j = (m+k+l)/2$. We then obtain
 %  \begin{align}
-  \int_{-\infty}^{\infty}
-  w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi =
-  \frac{\sqrt{2\pi}m!k!l!}{(j-m)!(j-k)!(j-l)!}, 
-\end{align}
-where $j = (m+k+l)/2$. We then obtain
- \begin{align}
-   2\hat{\sigma}\imag(\delta\hat{\sigma})
-  \left(1+\hat{k}^2\right) = \delta q \epsilon \hat{k} n.
- \end{align}
-Inserting the original eigenfrequency $\hat{\sigma} = \pm
-\sqrt{n}(1+\hat{k}^2)^{-1/2}$ gives
-\begin{align}
-  \imag(\delta\hat{\sigma})= \pm \frac{1}{2}\delta q \epsilon
-  \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
-\end{align}
-This result agrees with Eq. \ref{simple_growth} in the limit
-$|q|\to0$. 
-%for $n=1$, since the
-%function $\He_1$ solves the governing equation exactly. 
-For $\hat{k}\gg 1$ we have
-\begin{align}
-  \imag(\delta\hat{\sigma})\simeq \pm \frac{1}{2}\delta q \epsilon
-  \sqrt{n} \sgn{\hat{k}}. 
-\end{align}
+%    2\hat{\sigma}\imag(\delta\hat{\sigma})
+%   \left(1+\hat{k}^2\right) = \delta q \epsilon \hat{k} n.
+%  \end{align}
+% Inserting the original eigenfrequency $\hat{\sigma} = \pm
+% \sqrt{n}(1+\hat{k}^2)^{-1/2}$ gives
+% \begin{align}
+%   \imag(\delta\hat{\sigma})= \pm \frac{1}{2}\delta q \epsilon
+%   \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
+% \end{align}
+% This result agrees with Eq. \ref{simple_growth} in the limit
+% $|q|\to0$. 
+% %for $n=1$, since the
+% %function $\He_1$ solves the governing equation exactly. 
+% For $\hat{k}\gg 1$ we have
+% \begin{align}
+%   \imag(\delta\hat{\sigma})\simeq \pm \frac{1}{2}\delta q \epsilon
+%   \sqrt{n} \sgn{\hat{k}}. 
+% \end{align}
 

### 2015-02-03 linear_adia.tex
updating

@@
 timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
-($\gamma=\Gamma$) when allowing vertical shear.
+($\gamma=\Gamma$) when $q\neq0$.
 

### 2015-02-03 numerical.tex
completed rough summary
updating

@@
 use the full, non-Keplerian expressions of $\rho$, $\Omega$ and
-$\kappa$, so that we may consider vertical domains up to $z\sim r$.
+$\kappa$.
+%, so that we may consistently consider vertical domains up to
+%$z\sim r$.
 
@@
 where $\bm{\xi}$ is the Lagrangian displacement. This boundary
-condition was found to be most convenient for numerical
-implementation. The above discretization procedure converts the linear  
-system of differential equations to a standard matrix eigenvalue
-problem.
-
-Unless otherwise stated, our disk model is a nearly-vertically
-isothermal disk with $\Gamma=1.011$, $p=-1.5$, $q=-1$,
-$\epsilon=0.1$.
+condition is convenient for numerical implementation since it conforms
+to a standard eigenvalue problem. The above discretization procedure
+converts the linear system of differential equations to a set of 
+algebraic equations, for which we use matrix routines in the LAPACK
+package to solve.
 
@@
 
-\subsection{VSI in a neutrally stratified disk}\label{vertiso_pertiso} 
-We first calculate the VSI in a neutrally straitified disk ($N_z^2\equiv0$) by setting
-$\gamma=\Gamma$, and effectively switch off thermal relaxation by
-choosing $\beta=10^9$. We set the vertical domain to $\zmax =
-10\epsilon r$.  Here, we consider a perturbation wavenumber
-$\khat=20\pi$ or $k_x = 200\pi/r$. This setup is very similar to that
-adopted in \cite{mcnally14}, who calculated the VSI in vertically
-isothermal disks subject to isothermal perturbations
-($\Gamma=\gamma=1$).  This allows us to test our numerical code.
+\subsection{VSI in a locally isothermal, neutrally stratified disk}\label{vertiso_pertiso} 
+%  disk model is a nearly-vertically
+% isothermal disk with $\Gamma=1.011$, $p=-1.5$, $q=-1$,
+% $\epsilon=0.1$. 
+We first calculate the VSI in a nearly vertcally isothermal, neutrally
+straitified disk ($N_z^2\equiv0$) by setting
+$\gamma=\Gamma=1.011$ and $\beta=10^{-3}$. Other disk parameters are 
+$(p,q,\epsilon)=(-1.5,-1,0.1)$ and the vertical domain is $\zmax =
+10\epsilon r$. This setup is very similar to that adopted in
+\cite{mcnally14}, who calculated the VSI in vertically isothermal
+disks subject to isothermal perturbations ($\Gamma=\gamma=1$).  This
+allows us to test our numerical code.
 
-Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
-\ii\nu$. The fundamental VSI mode, with one node in $W$, has
+%behaviour as func of kx
+
+Fig. \ref{iso_eigen_kx} compares the numerically-obtained eigenvalues
+for the fundamental mode and that caluclated from
+Eq. \ref{simple_growth} (with $L=1$). For completeness we
+included cases with $\khat\ll 1$ even though the local
+approximation does not hold in this regime. The agreement is good for
+$\khat\gtrsim10$. The mismatch is significant for $\khat\ll1$ since
+the low-frequency approximation used to obtain Eq. \ref{simple_growth}
+is no longer valid: $|\omega|\sim\Omega_k$ as $\khat\to 0$. This means
+that our analytic results should only be applied to perturbations with
+$\khat\gg 1$, i.e. small radial-lengthscales.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/compare_eigen_imag_iso} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/compare_eigen_real_iso}
+  \caption{Growth rate (top) and real frequency (bottom) of the
+    fundamental VSI mode in a disk with $\gamma=\Gamma=1.011$,
+    $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+    $\beta=10^{-3}$. The solid (dotted) line is obtained numerically
+    (analytically).  
+    \label{iso_eigen_kx} 
+  }
+\end{figure}
+
+%particular kx 
+As an example of an unstable mode, we consider a perturbation
+wavenumber $\khat=20\pi$ \citep[or $k_x = 200\pi/r$ as used
+in][]{mcnally14}. Fig. \ref{lowfreq_eigen} shows the eigenvalues
+$\sigma = \omega + \ii\nu$. The fundamental VSI mode, with one node in $W$, has
 the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}. 
@@
 
-The expected growth rate of the fundamental VSI, calculated from
-Eq. \ref{simple_growth} with $L=1$, and the numerically-obtained value
+The growth rate of the fundamental VSI calculated from
+Eq. \ref{simple_growth} with $L=1$ and the numerically-obtained value
 are  
@@
     Eq. \ref{simple_growth})},\\
-  \nu = 0.2577 \epsilon \Omega_k &\quad \text{(numerical)}.
+  \nu = 0.2575 \epsilon \Omega_k &\quad \text{(numerical)}.
 \end{align*}
@@
 imposes a different boundary condition to that in the numerical
-calculation. The eigenfrequency is not very sensitive to the vertical boundary
-condition because the expression for $\sigma$, given by
-Eq. \ref{integral_relation1}, involve integrands with $\rho$ as the 
+calculation. These values are also close to that found by
+\cite{mcnally14}. The eigenfrequency is not sensitive to the
+vertical boundary condition because the expression for $\sigma$, given
+by Eq. \ref{integral_relation1}, involve integrands with $\rho$ as the  
 weight function, which rapidly decays away from the midplane.
@@
     nearly-vertically isothermal disk  ($\gamma=\Gamma=1.011$) with 
-    $(p,q,\epsilon)=(-1.5,-1,0.1)$, and the perturbation radial
-    wavenumber is $\khat=20\pi$. \label{lowfreq_eigen}
+    $(p,q,\epsilon)=(-1.5,-1,0.1)$ and $\beta=10^{-3}$. The
+    perturbation wavenumber is $\khat=20\pi$. \label{lowfreq_eigen} 
   }
@@
 
-\subsection{VSI in a stably stratified disk}
-We demonstrate the (strong) stabilizing effect of a  
-positive vertical entropy gradient ($N_z^2\geq0$) by setting $\gamma=2$.  
-Since we expect the perturbations to decay rapidly
-away from the midplane (see \S\ref{analytic_adia}), we consider a
-smaller domain with $\zmax = 2\epsilon r$. 
-Other disk and perturbation parameters are the same as in
-\S\ref{vertiso_pertiso}.
+\subsection{Effect of thermal relaxation}
+We now consider stably stratified disks ($\gamma > \Gamma$) with
+thermal relaxation $\beta\in[10^{-3},1]$. Here we use a slightly 
+different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as
+adopted in simulations performed by \cite{nelson13}. Other parameters
+are unchanged from the previous section.
 
-Fig. \ref{lowfreq_eigen_adia} show the eigenvalues for this
-problem. The growth rates are much smaller than
-those in the previous section. The fundamental VSI mode is that with
-$\mathrm{max}\nu$. Its expected and numerically-calculated growth
-rates compares well, with 
-\begin{align*}
-  \nu = 0.03751 \epsilon \Omega_k &\quad \text{(from
-    Eq. \ref{gam2_growth_rate})},\\
-  \nu = 0.03672 \epsilon \Omega_k &\quad \text{(numerical)}.
-\end{align*}
-The fundamental mode is plotted in Fig. \ref{lowfreq_eigenfunc_adia},
-and clearly show that perturbations are rapidly stabilized away from
-the midplane. For the adopted disk and perturbation parameters,
-Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$, corresponding
-to a characteristic decay length scale of $\sqrt{2/|\alpha|}\epsilon r\simeq 0.2\epsilon r$,
-which is consistent with that observed in  Fig. \ref{lowfreq_eigenfunc_adia}.
+Fig. \ref{relax_growth_num} shows the growth rates of the fundamental  
+mode as a function of $\khat$. This plot is largely consistent with  
+Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
+where the low-frequency approximation fails. (The local model not valid
+for $\khat\ll1$ anyway). For $\khat\gtrsim 10$ introducing thermal
+relaxation rapidly stabilizes the fundamental mode.
 
-This decay lengthscale is rougly the height within which
-vertical shear is larger than the buoyancy frequency squared. That is,
-\begin{align*}
-\left|r\frac{d\Omega^2}{dz}\right|\gtrsim N_z^2
-\end{align*}
-for
-\begin{align*}
-  \left|\frac{z}{\epsilon r}\right| \lesssim \frac{\epsilon|q|\gamma}{\gamma-1},
-\end{align*}
-in a thin, vertically isothermal disk. This evaluates to $|z|\lesssim
-0.2\epsilon r$ for the current disk model, consistent with
-Fig. \ref{lowfreq_eigenfunc_adia}.
-
-Furthermore, the growth rate to should be no larger than the
-maximum vertical shear available within this decay lengthscale. Using 
-Eq. \ref{max_growth} as a rough guide, 
-\begin{align}
-  \frac{\nu}{\epsilon\Omega_k}<\frac{\epsilon |q|^2\gamma}{2(\gamma-1)},
-\end{align}
-or $\nu< 0.1\epsilon\Omega_k$ in our disk model,
-consistent with numerical results.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/eigenvalues_adia}
-  \caption{Eigenvalues in for the
-    VSI in a nearly vertically isothermal disk
-    ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
-    evolved adiabatically with $\gamma=2$. The perturbation radial
-    wavenumber is $\khat=20\pi$. \label{lowfreq_eigen_adia}
-  }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/eigenvectorvz_adia}
-  \caption{Fundamental VSI mode in a nearly vertically
-    isothermal disk ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
-    evolved adiabatically
-    with $\gamma=2$. This eigenfunction 
-    corresponds to the top-left eigenvalue displayed in 
-    Fig. \ref{lowfreq_eigen_adia} (largest $|\nu|$).  
-    The  real (imaginary) part of $\delta v_z$ are shown as solid
-    (dotted) lines. 
-    \label{lowfreq_eigenfunc_adia}
-  }
-\end{figure}
-
-Although a positive vertical entropy gradient is stabilizing, we note
-that $N_z^2\propto z^2$  while vertical shear $rd\Omega^2/dz\propto
-z$, so that there is always a region sufficiently close to the
-midplane in which vertical shear dominates over bouyancy and the VSI
-can operate. However, the growth rates are small and the instability
-only affects the regions close the midplane. These modes may not be
-important in practice.
-
-\subsection{Other adiabatic indices}
-To see the transition from neutrally-bouyant to stably stratified cases above, 
-we repeated the previous calculation with $\gamma\in[1,2]$. 
-We restore the vertical domain to $\zmax=10\epsilon r$. 
-Fig. \ref{adia_growth_num} shows the eigenfrequencies of the
-fundamental VSI mode.
-
-%To select the appropriate
-%eigenvalues we vary the parameters slowly away from a reference solution.   
-% The numerically-determined eigenfrequencies are qualitatively
-% consistent with the discussion in \S\ref{adia_compare_gam1}.
-There is good agreement between Fig. \ref{adia_growth_num} and
-Fig. \ref{adia_growth} for $\khat=k_xH_\mathrm{iso}\gtrsim10$. There is a
-mismatch for smaller wavenumbers, especially in the growth
-rates. Actual growth rates decrease slower as $\khat\to0$
-than that in Fig. \ref{adia_growth}. Disagreement at small
-$\khat$ is not unexpected because the low-frequency
-approximation used to obtain Fig. \ref{adia_growth} becomes invalid since
-$|\omega|\sim\Omega_k$ for small $\khat$.
-
-Given the number of approximations made in \S\ref{adia_compare_gam1},
-Fig. \ref{adia_growth_num} compares well with 
-Fig. \ref{adia_growth}. The main qualitative features of the full
-numerical solution in Fig. \ref{adia_growth_num} are captured by
-Fig. \ref{adia_growth}: stabilization for 
-$\khat\to0$ and $\khat\to\infty$,
-stabilization for increasing $\gamma$ (but ineffective at small $\khat$), 
-and the shift of the most unstable wavenumber to smaller values as
-$\gamma$ is increased.
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_real1}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag1}
-  \caption{Real frequency (top) and growth rate 
-    (bottom) of the fundamental VSI in a nearly vertically   
-    isothermal disks ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
-    subject to adiabatic perturbations with different values of $\gamma$. 
-    The eigenfrequencies are calculated by numerically
-    solving the full linear eigenvalue problem. This plot is to
-    be compared with Fig. \ref{adia_growth}. \label{adia_growth_num}}   
-\end{figure}
-
-\subsection{Effect of thermal relaxation}
-We now introduce thermal relaxation. Here we consider a slightly
-different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as
-adopted in simulations performed by \cite{nelson13}. 
-Fig. \ref{relax_growth_num} plots the
-growth rates of the fundamental mode function of $\khat$.
-
+% For $\beta=0.063$ at $\khat\gtrsim50$ we found eigenfunctions had 
+% large perturbations near the vertical boundaries, which may be 
+% responsible for the non-smoothness in the curve at large $\khat$. 
 %corrugation mode
@@
 % slowly to obtain eigenvalues for other values of $k_x$ or $\beta$.  
-
-Fig. \ref{relax_growth_num} is largely consistent with 
-Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
-where the low-frequency approximation fails. 
-For $\beta=0.063$ at $\khat\gtrsim50$ we found eigenfunctions had 
-large perturbations near the vertical boundaries, which may be 
-responsible for the non-smoothness in the curve at large $\khat$.
-
 %likely attributable to the free surface boundary condition imposed in
@@
 \begin{figure}
-   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
    0cm]{figures/compare_eigen_imag_bloop} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  1cm]{figures/compare_eigen_imag_bloop2} 
+  % \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  % 1cm]{figures/compare_eigen_imag_bloop2} 
   \caption{Growth rates of the fundamental VSI mode in 
@@
 
-%\citeauthor{nelson13} find a thermal relaxation timescale of
-% $\beta\gtrsim 0.6$ stabilized the VSI for this disk model in non-linear
-% hydrodynamic simulations. This is consistent with our numerical
-% results.
+For this disk model, \citeauthor{nelson13} found a thermal relaxation
+timescale of $\beta\gtrsim 0.6$ stabilized the VSI in non-linear
+hydrodynamic simulations. This is consistent with our numerical and
+analytic results.
+
+\subsubsection{Dependence on disk parameters}
+We test our estimate for the critical relaxation time
+$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}). We use the above disk
+parameters as a fiducial case and vary $q\in[-0.2,-1.2]$,
+$\gamma\in[1.2,2.0]$ and $\epsilon\in[]$
+separately. The numerically-obtained $\beta_\mathrm{crit}$ is shown in
+Fig. \ref{bcrit_compare} in comparison with Eq. \ref{iso_vsi_cond}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
+  0cm]{figures/bcrit_compare_q.ps} 
+  \caption{Dependence of the upper limit to the thermal relaxation timescale
+    $\beta_\mathrm{crit}$ for the fundamental VSI mode on disk
+    parameters. The fiducial parameters are $\Gamma=1.011$,
+    $\gamma=1.4$ and $(p,q,\epsilon)=(-1.5,-1,0.05)$. Top: varying
+    $q\in[-1.2,-0.2]$; middle: varying $\gamma\in[1.2,2.0]$; bottom:
+    varying $\epsilon\in[]$. The perturbation wavenumber is
+    $\khat=20$. {\bf waiting for high resolution}
+    \label{bcrit_compare}}  
+\end{figure}
 
@@
 \end{figure}
+
+% \subsection{VSI in a stably stratified disk}
+% We demonstrate the (strong) stabilizing effect of a  
+% positive vertical entropy gradient ($N_z^2\geq0$) by setting $\gamma=2$.  
+% Since we expect the perturbations to decay rapidly
+% away from the midplane (see \S\ref{analytic_adia}), we consider a
+% smaller domain with $\zmax = 2\epsilon r$. 
+% Other disk and perturbation parameters are the same as in
+% \S\ref{vertiso_pertiso}.
+
+% Fig. \ref{lowfreq_eigen_adia} show the eigenvalues for this
+% problem. The growth rates are much smaller than
+% those in the previous section. The fundamental VSI mode is that with
+% $\mathrm{max}\nu$. Its expected and numerically-calculated growth
+% rates compares well, with 
+% \begin{align*}
+%   \nu = 0.03751 \epsilon \Omega_k &\quad \text{(from
+%     Eq. \ref{gam2_growth_rate})},\\
+%   \nu = 0.03672 \epsilon \Omega_k &\quad \text{(numerical)}.
+% \end{align*}
+% The fundamental mode is plotted in Fig. \ref{lowfreq_eigenfunc_adia},
+% and clearly show that perturbations are rapidly stabilized away from
+% the midplane. For the adopted disk and perturbation parameters,
+% Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$, corresponding
+% to a characteristic decay length scale of $\sqrt{2/|\alpha|}\epsilon r\simeq 0.2\epsilon r$,
+% which is consistent with that observed in  Fig. \ref{lowfreq_eigenfunc_adia}.
+
+% This decay lengthscale is rougly the height within which
+% vertical shear is larger than the buoyancy frequency squared. That is,
+% \begin{align*}
+% \left|r\frac{d\Omega^2}{dz}\right|\gtrsim N_z^2
+% \end{align*}
+% for
+% \begin{align*}
+%   \left|\frac{z}{\epsilon r}\right| \lesssim \frac{\epsilon|q|\gamma}{\gamma-1},
+% \end{align*}
+% in a thin, vertically isothermal disk. This evaluates to $|z|\lesssim
+% 0.2\epsilon r$ for the current disk model, consistent with
+% Fig. \ref{lowfreq_eigenfunc_adia}.
+
+% Furthermore, the growth rate to should be no larger than the
+% maximum vertical shear available within this decay lengthscale. Using 
+% Eq. \ref{max_growth} as a rough guide, 
+% \begin{align}
+%   \frac{\nu}{\epsilon\Omega_k}<\frac{\epsilon |q|^2\gamma}{2(\gamma-1)},
+% \end{align}
+% or $\nu< 0.1\epsilon\Omega_k$ in our disk model,
+% consistent with numerical results.
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth]{figures/eigenvalues_adia}
+%   \caption{Eigenvalues in for the
+%     VSI in a nearly vertically isothermal disk
+%     ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
+%     evolved adiabatically with $\gamma=2$. The perturbation radial
+%     wavenumber is $\khat=20\pi$. \label{lowfreq_eigen_adia}
+%   }
+% \end{figure}
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth]{figures/eigenvectorvz_adia}
+%   \caption{Fundamental VSI mode in a nearly vertically
+%     isothermal disk ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
+%     evolved adiabatically
+%     with $\gamma=2$. This eigenfunction 
+%     corresponds to the top-left eigenvalue displayed in 
+%     Fig. \ref{lowfreq_eigen_adia} (largest $|\nu|$).  
+%     The  real (imaginary) part of $\delta v_z$ are shown as solid
+%     (dotted) lines. 
+%     \label{lowfreq_eigenfunc_adia}
+%   }
+% \end{figure}
+
+% Although a positive vertical entropy gradient is stabilizing, we note
+% that $N_z^2\propto z^2$  while vertical shear $rd\Omega^2/dz\propto
+% z$, so that there is always a region sufficiently close to the
+% midplane in which vertical shear dominates over bouyancy and the VSI
+% can operate. However, the growth rates are small and the instability
+% only affects the regions close the midplane. These modes may not be
+% important in practice.
+
+% \subsection{Other adiabatic indices}
+% To see the transition from neutrally-bouyant to stably stratified cases above, 
+% we repeated the previous calculation with $\gamma\in[1,2]$. 
+% We restore the vertical domain to $\zmax=10\epsilon r$. 
+% Fig. \ref{adia_growth_num} shows the eigenfrequencies of the
+% fundamental VSI mode.
+
+% %To select the appropriate
+% %eigenvalues we vary the parameters slowly away from a reference solution.   
+% % The numerically-determined eigenfrequencies are qualitatively
+% % consistent with the discussion in \S\ref{adia_compare_gam1}.
+% There is good agreement between Fig. \ref{adia_growth_num} and
+% Fig. \ref{adia_growth} for $\khat=k_xH_\mathrm{iso}\gtrsim1$. There is a
+% mismatch for smaller wavenumbers, especially in the growth
+% rates. Actual growth rates decrease slower as $\khat\to0$
+% than that in Fig. \ref{adia_growth}. Disagreement at small
+% $\khat$ is not unexpected because the low-frequency
+% approximation used to obtain Fig. \ref{adia_growth} becomes invalid since
+% $|\omega|\sim\Omega_k$ for small $\khat$.
+
+% Given the number of approximations made in \S\ref{adia_compare_gam1},
+% Fig. \ref{adia_growth_num} compares well with 
+% Fig. \ref{adia_growth}. The main qualitative features of the full
+% numerical solution in Fig. \ref{adia_growth_num} are captured by
+% Fig. \ref{adia_growth}: stabilization for 
+% $\khat\to0$ and $\khat\to\infty$,
+% stabilization for increasing $\gamma$ (but ineffective at small $\khat$), 
+% and the shift of the most unstable wavenumber to smaller values as
+% $\gamma$ is increased.
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_real1}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag1}
+%   \caption{Real frequency (top) and growth rate 
+%     (bottom) of the fundamental VSI in a nearly vertically   
+%     isothermal disks ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
+%     subject to adiabatic perturbations with different values of $\gamma$. 
+%     The eigenfrequencies are calculated by numerically
+%     solving the full linear eigenvalue problem. This plot is to
+%     be compared with Fig. \ref{adia_growth}. \label{adia_growth_num}}   
+% \end{figure}

### 2015-02-06 appendix.tex
major revision underway, update to include global correction in numerical sims

@@
 
-The `global correction' terms do not change the qualitative dependence
+The global correction terms do not change the qualitative dependence
 of the fundamental VSI growth rates on the thermal relaxation
@@
 fast thermal relaxation). The critical relaxation timescale for the
-fundamental VSI is unaffected by these global radients.
+fundamental VSI is unaffected by these global radial radients.
 

### 2015-02-06 application.tex
abstract added, intro left to do
major revision underway, update to include global correction in numerical sims

@@
 protoplanetary disk do thermodynamic conditions allow the VSI to 
-operate. Specifically, we consider the fundamental isothermal VSI, 
+operate. Specifically, we consider the fundamental VSI, 
 for which we can use the criterion Eq. \ref{iso_vsi_cond}.
@@
 where the temperature $T$ is defined through $P=\mathcal{R}\rho
-T/\mu$, where $\mathcal{R}$ is the ideal gas constant and $\mu$ is
-the mean molecular weight. In Eq. \ref{real_energy} $S$ is the
-specific entropy,
-\begin{align}
-  S \equiv c_v\ln{\left(\frac{p}{\rho^{\gamma}}\right)} 
-\end{align}
-where $c_v \equiv \left(\gamma-1\right)^{-1}\mathcal{R}/\mu$ is the
-specific heat capcity at constant volume, $D/Dt$ is the convective
+T/\mu$, $\mathcal{R}$ is the ideal gas constant and $\mu$ is
+the mean molecular weight, % In Eq. \ref{real_energy} $S$ is the
+% specific entropy,
+% \begin{align}
+%   S \equiv c_v\ln{\left(\frac{p}{\rho^{\gamma}}\right)} 
+% \end{align}
+% where $c_v \equiv \left(\gamma-1\right)^{-1}\mathcal{R}/\mu$ is the
+% specific heat capcity at constant volume
+$D/Dt$ is the convective
 derivative, $\bm{F}$ is the radiative flux, and $Q_+$ is a heating
@@
 calculations. To do so, we first consider separately perturbations
-with radial lengthscales $L\equiv 2\pi/k_x\ll l_\mathrm{rad}$ and 
-$L\gg l_\mathrm{rad}$, where      
+with radial lengthscales $l\equiv 2\pi/k_x\ll l_\mathrm{rad}$ and 
+$l\gg l_\mathrm{rad}$, where      
 \begin{align}\label{lrad}
@@
 \subsection{Newtonian cooling}\label{newton_cool}
-For $L\ll l_\mathrm{rad}$ the radiative cooling is given by {\bf (reference?)}
+For $l\ll l_\mathrm{rad}$ the radiative cooling is given by {\bf (reference?)}
 \begin{align}
@@
 \subsection{Radiative diffusion}
-For perturbation lengthscales $L\gg l_\mathrm{rad}$, let us assume that
+For perturbation lengthscales $l\gg l_\mathrm{rad}$, we assume 
 temperature fluctuations are smoothed out by radiative diffusion. To
@@
 \end{align}
-We recall $\beta < \beta_\mathrm{crit}$ is a neccessary condition for the
-fundamental isothermal VSI when $\khat\gg 1$. For $\khat\to0$, this
+We recall $\beta < \beta_\mathrm{crit}$ is neccessary for the
+fundamental VSI when $\khat\gg 1$. For $\khat\to0$, this
 becomes a sufficient condition (see Fig. \ref{relax_bcrit} and
@@
 the scalings $\hat{T}=\hat{\Sigma}=\hat{\kappa}_d=1$, with $\mu =
-2.33$ and $\gamma=1.4$. The figure suggests that the fundamental
-isothermal VSI operates in the outer parts of the MMSN at radii on
+2.33$ and $\gamma=1.4$. The figure suggests that the fundamental VSI
+operates in the outer parts of the MMSN at radii on 
 the order of tens of AU. However, the thermal relaxation time $\beta$
@@
 sensitive to $\khat$ at $r=50\mathrm{AU}$ for $\khat\gg1$.)
-The disk is unstable to the fundamental isothermal VSI, but not
+The MMSN is unstable to the fundamental VSI, but not
 strongly.
@@
 calculation above with $\beta\to\tilde{\beta}$. We find growth rates
-were only slightly reduced at low $\khat$ and eigenfunctions are
-similar.
+were reduced, but at most by $\sim 15\%$ at $\khat=1$. Eigenfunctions
+were similar.
+
+%at most by 15%

### 2015-02-06 numerical.tex
major revision underway, update to include global correction in numerical sims

@@
 Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
-where the low-frequency approximation fails. (The local model not valid
-for $\khat\ll1$ anyway). For $\khat\gtrsim 10$ introducing thermal
+where the low-frequency approximation fails. (In any case, the local model not valid
+for $\khat\ll1$.) For $\khat\gtrsim 10$ introducing thermal
 relaxation rapidly stabilizes the fundamental mode.
@@
 
-\subsubsection{Dependence on disk parameters}
-We test our estimate for the critical relaxation time
-$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}). We use the above disk
+\subsection{Critical thermal relaxation timescale}
+We test our estimate for the maximum thermal relaxation timescale 
+$\beta_\mathrm{crit}$ for the fundamental
+VSI (Eq. \ref{iso_vsi_cond}). We use the previous disk
 parameters as a fiducial case and vary $q\in[-0.2,-1.2]$,
@@
 
+We find Fig. \ref{bcrit_compare} generally agrees with
+Eq. \ref{iso_vsi_cond}.  The match is better for smaller
+$\beta_\mathrm{crit}$. There is a noticeable difference as
+$\gamma\to1$ because the  derivation of $\beta_\mathrm{crit}$ assumed
+$\gamma>1$.
+
 \begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.cm 0cm
+  0cm]{figures/bcrit_compare_q.ps} 
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
-  0cm]{figures/bcrit_compare_q.ps} 
+  0.8cm]{figures/bcrit_compare_g.ps} 
   \caption{Dependence of the upper limit to the thermal relaxation timescale
@@
     varying $\epsilon\in[]$. The perturbation wavenumber is
-    $\khat=20$. {\bf waiting for high resolution}
+    $\khat=20$. {\bf still need variable epsilon case}
     \label{bcrit_compare}}  
@@
 fundamental VSI for $\beta\in[10^{-2},10^{2}]$. We find growth rates
-do not reach zero, implying the VSI operates regardless of the thermal
+do not reach zero, implying instability regardless of the thermal
 relaxation timescale if $\gamma=\Gamma$. This result is consistent
-with \cite{nelson13}.
+with \cite{nelson13}.  Note that for $\beta\to\infty$ (adiabatic
+perturbations) and $\gamma=\Gamma$, the disk can be unstable according
+to the Solberg-Hoiland criterion  Eq. \ref{solberg2}.
 
-We also note that growth rates for $\beta = 10^{-2}$ and
+Growth rates for $\beta = 10^{-2}$ and
 $\beta=10^{2}$ behave similarly. This is not surprising because
@@
 disks with $N_z^2\geq 0$ (Fig. \ref{relax_growth_num}), namely growth
-rates reach zero at finite $\khat$.
+rates reach zero at finite $\khat$. From Fig.\ref{growth_vnoniso2} we
+expect the VSI to be strongly stabilized even for small, finite
+thermal relaxation times.
 

### 2015-02-06 paper.tex
major revision underway, update to include global correction in numerical sims

@@
 accretion disks. For example, there may be regions in protoplanetary  
-disks --- called dead zones ---  which do not support the MRI 
-\citep{gammie96,landry13}. In fact, recent state-of-the-art
+disks which do not support the MRI \citep{gammie96,landry13}. In fact,
+state-of-the-art  numerical simulations suggest portions of PPDs
+remain laminar. %transport by wind, large-scale field 
+% cite bai, lesur,
+% gressel
 
-%in this respect purely hydrodynamical turbulence should be
+In this respect, purely hydrodynamical turbulence remains of interest
+in the context of PPDs. One possibiliy is the vertical shear
+instability (VSI), which has recently been demonstrated through
+hydrodynamical simulations \citep{nelson13} %and kley simulations
 
 %stuff about vsi 
+The VSI is an axisymmetric instability in three-dimensional (3D) disks
+for which the rotation frequency $\Omega$ depends on the height from
+the disk midplane.
 

### 2015-02-06 summary.tex
major revision underway, update to include global correction in numerical sims

@@
 shearing box' (VGSB) formalism developed by \cite{mcnally14}, who 
-showed the linear isothermal VSI studied in \cite{nelson13}, can
-be captured by the VGSB. We added 
-an energy source term in the VGSB equations which relaxes the disk
+showed the linear VSI studied by \cite{nelson13} in a locally
+isothermal disk can be captured by the VGSB.
+
+% The presence of a
+% vertical shear, but 
+% is not consistent with the
+% global
+
+We added an energy source term in the VGSB equations which relaxes the disk
 temperature to its initial value on a timescale
@@
 employed in the global numerical simulations of \cite{nelson13}, and allows
-the thermodynamic response of the disk to be varied in a controlled 
-manner.
+the thermodynamic response of the disk to be controlled.
 
@@
 extension of the linear models considered in \cite{lubow93} by
-including vertical shear, but in the limit of low-frequency
+adding vertical shear, but in the limit of low-frequency
 perturbations in nearly-Keplerian disks. These approximations were
@@
 For isothermal perturbations ($\gamma=1$ or $\beta=0$) we explicitly
-demonstrated instability 
-associated with vertical shear, and that the VSI growth rate is
-limited by the maximum vertical shear within the vertical domain of
-interest. In the opposite limit of adiabatic perturbations ($\gamma>1$
-and $\beta\to\infty$), we find VSI growth rates are dramatically
-reduced on account of a stabilizing 
-vertical entropy gradient. The VSI is not completely
-surpressed because vertical shear dominates over buoyancy 
-sufficiently close the disk midplane. However, perturbations grow
-slowly and is constrained to regions immediately about the disk
-midplane. Thus, the adiabatic VSI is likely unimportant in practice.
+demonstrated instability  associated with vertical shear, and that the
+VSI growth rate is limited by the maximum vertical shear within the
+vertical domain of interest. The fundamental VSI mode is characterized
+by one node in the density perturbation at the mid-plane
+(i.e. anti-symmetric about $z=0$), which corresponds to a vertical
+velocity perturbation  $\delta v_z\sim\mathrm{constant}$ near $z=0$.
 
-For non-isothermal ($\gamma\neq1$) perturbations with intermediate
-$\beta$, the general dispersion relation
-(Eq. \ref{relax_disp}) is complicated and requires a numerical
-treatment. However, for the fundamental VSI, characterized by  $\delta
-v_z\sim\mathrm{constant}$ near $z=0$, some simplification is
-possible. In particular, we quantified the thermal relaxation
+% In the opposite limit of adiabatic perturbations ($\gamma>1$
+% and $\beta\to\infty$), we find VSI growth rates are dramatically
+% reduced on account of a stabilizing 
+% vertical entropy gradient. The VSI is not completely
+% surpressed because vertical shear dominates over buoyancy 
+% sufficiently close the disk midplane. However, perturbations grow
+% slowly and is constrained to regions immediately about the disk
+% midplane. Thus, the adiabatic VSI is likely unimportant in practice.
+
+For non-isothermal  perturbations with thermal relaxation ($\gamma>1$,
+$\beta>0$) , the general dispersion relation (Eq. \ref{relax_disp}) is
+complicated and requires a numerical treatment. However, for the
+fundamental VSI some simplification is  possible. In particular, we quantified the critical thermal relaxation
 timescale $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$, below
 which small-scale perturbations ($\khat\gg 1$) may be regarded as isothermal
-and the standard VSI operates. We believe this explicit expression for
+and the fundamental VSI operates. We believe this explicit expression for
 $\beta_\mathrm{crit}$ is the most useful result from our
@@
 radial wavenumber ($\khat\to0$) where the low-frequency approximation
-is inappropriate.
+is inappropriate. The local model is also not valid for such modes.
 
@@
 code. We confirm the finding by \cite{nelson13} that in the absence
-of vertical entropy gradients, the VSI operates and is insensitive to thermal relaxation. When a
-stable stratification is introduced, the reduction of VSI growth
-rates by a finite thermal relaxation timescale is qualitatively similar to
-that observed for vertically isothermal disks.
+of vertical entropy gradients, the VSI operates and is insensitive to
+thermal relaxation. When a stable stratification is introduced, the
+rapid reduction in VSI growth rates by a finite thermal relaxation timescale
+is qualitatively similar to that observed for vertically isothermal
+disks.
 
@@
 using disk models described in \cite{chiang10}. We find the MMSN is
-unstable to the VSI at $10$s of AU, with characteristic
+unstable to the fundamental VSI at $10$s of AU, with characteristic
 growth times of $10$s of orbits. However, we note the strong
@@
 in the disk model ($\beta$ decreases as $T^{-5}$ for perturbations with small
-radial scale). Thus, increasing the disk temperature can signifcantly 
+radial scale). Thus, increasing the disk temperature even slightly can signifcantly 
 shorten the typical growth times for the VSI. This is also attributed
@@
 future work.
+
+%need to compare kley sims
 

### 2015-02-07 appendix.tex
undergoing major revisions

@@
-\section{Local model for a global problem}\label{global_corr}
+\section{Limitations of the vertically global shearing box}\label{global_corr}
 Here, we discuss the relation between the linear problem in the
-VGSB framework and the global linear problem. We begin with the
-axisymmetric linearized fluid equations in cylindrical co-ordinates, including
-thermal relaxation, 
-\begin{align}
-\ii\sigma\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
-\left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
-    \delta v_z}{\p z}\right] + \frac{1}{t_c}\left(\delta P -
-  \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,\\
-  \ii\sigma\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
-  z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\\
-\ii\sigma\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
-\delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
-\ii\sigma\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
-r\frac{\p\Omega}{\p z}\delta v_z,\\
-\ii \sigma\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
-\frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},
-\end{align}
-where the coefficients $\hat{g}_c=1$ are inserted to label terms
-associated with the background radial disk structure.
+VGSB framework and the global linear problem.
 
-We assume perturbations have the form $W(z)\exp{\ii k_r r}$ such
-that $|k_rr|\gg1$, so that curvature terms can be ignored. For these
-small radial lengthscale perturbations we also approximate the
-background structure and radial gradients as uniform in $r$, and
-evaluate all cofficients in the above equations at the fiducial radius
-$r=r_0$. This procedure produces $\phi$ and $z$ momentum equations
+ This procedure produces $\phi$ and $z$ momentum equations
 that are equilvalent to those for $y$ and $z$ momenta in the VGSB.
@@
 %plot growth rate v.s. bcool
+
+\subsection{Spurious growth of adiabatic perturbations in the local
+  model with vertical shear}\label{analytic_adia} 
+For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
+simply use the Soilberg-Hoiland critera to assess stability
+(\S\ref{solberg}). Nevertheless, we briefly consider this case in the
+local framework to demonstrate its limitations. For $\beta\to\infty$
+the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
+be written as
+
+\begin{align}
+  0 =& \dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
+    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime\notag\\
+  &+\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right)\right.\notag\\
+  &\phantom{0=}\left.-\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
+          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
+\end{align}
+
+We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
+integrate vertically, assume boundary terms vanish when integrating by
+parts, to obtain
+\begin{align}
+  &\hat{\sigma}^2\left(\frac{1}{\gamma} +
+    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
+  v_z|^2 d\zhat \notag\\
+  &=  \left(\frac{\gamma-1}{\gamma}\right)
+  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
+  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
+  v_z)^\prime|^2 d\zhat\notag\\
+  &\phantom{=} +
+  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
+      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
+    2}|\delta v_z|^2 d\zhat\notag\\
+  &\phantom{=} + \ii\epsilon q \hat{k}
+  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
+  \delta v_z d\zhat.\label{adia_integral}
+\end{align}
+When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
+$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
+expected, a sub-adiabatically stratified disk is stable in the absense
+of vertical shear.
+
+However, in the presence of vertical shear $q\neq0$,
+Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
+$\khat$, which implies the possibility of instability, regardless of
+how much $\gamma$ deviates from unity. This is contradicts the second 
+Solberg-Hoiland criterion (Eq. \ref{solberg2}) and is a result of the
+radially-local approximation (Appendix \ref{global_corr}). To 
+avoid this, we must restrict our attention to short thermal relaxation
+timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
+($\gamma=\Gamma$) when $q\neq0$.
 

### 2015-02-07 linear_adia.tex
undergoing major revisions

--- 2015-02-02 linear_adia.tex
@@
 $\hat{k}\to0$ and $\hat{k}\to\infty$ we find $\delta\hat{\nu}\to0$, so
-stabilization by finite cooling is ineffective at very large or very
-small scales. We plot Eq. \ref{nearly_iso_dsig} in Fig. \ref{domegadbeta} with
-$\epsilon=0.05$, $q=-1.0$ and $\gamma=1.4$. For this disk model the
-optimum wavenumber for stabilization is $\hat{k}\sim 5$---$6$.
+stabilization by small finite cooling is ineffective at very large or
+very small scales. 
+% We plot Eq. \ref{nearly_iso_dsig} in
+% Fig. \ref{domegadbeta} with 
+% $\epsilon=0.05$, $q=-1.0$ and $\gamma=1.4$. For this disk model the
+% optimum wavenumber for stabilization is $\hat{k}\sim 5$---$6$.
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/domegadbeta}
-  \caption{Rate of change of the fundamental VSI eigenfrequency
-    $\sigma$ with respect to the introduction of a small but finite
-    cooling time. The plotted quantity is $\p\hat{\sigma}/\p\beta$ at $\beta=0$, evaluated using
-    Eq. \ref{nearly_iso_dsig} with 
-    disk parameters $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The real
-    (imaginary) part of $\p\hat{\sigma}/\p\beta$ is shown as the solid
-    (dotted) lines. 
-    \label{domegadbeta}}  
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\linewidth]{figures/domegadbeta}
+%   \caption{Rate of change of the fundamental VSI eigenfrequency
+%     $\sigma$ with respect to the introduction of a small but finite
+%     cooling time. The plotted quantity is $\p\hat{\sigma}/\p\beta$ at $\beta=0$, evaluated using
+%     Eq. \ref{nearly_iso_dsig} with 
+%     disk parameters $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The real
+%     (imaginary) part of $\p\hat{\sigma}/\p\beta$ is shown as the solid
+%     (dotted) lines. 
+%     \label{domegadbeta}}  
+% \end{figure}
 
@@
   relaxation}\label{disp_relax}
-To obtain a dispersion relation similar to Eq. \ref{sig2_iso} for
-isothermal perturbations, we first write 
+To obtain a dispersion relation similar to that for 
+isothermal perturbations (Eq. \ref{sig2_iso}), we first write 
 \begin{align}
@@
 where the coefficients $c_l$ are given in Appendix \ref{relax_coeff}.
-% Eq. \ref{relax_disp} contrasts to Eq. \ref{adia_disp}, which is
-% quadratic in $\hat{\sigma}$ for adiabatic perturbations.
+This dispersion relation generally requires a numerical
+solution. However, for the fundamental mode ($M=0$) some analytical
+progress can be made, which we consider next.
 
-It is simplest to solve Eq. \ref{relax_disp} numerically. In
-Fig. \ref{relax_disp_fig} we plot the growth rate of the fundamental
-($M=0$) VSI mode as a function of $\khat$ for a range of
-dimensionless thermal relaxation timescales $\beta$, in a disk with
-$\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. Increasing $\beta$ even
-slightly above zero significantly reduces the growth rates for
-$\khat\gg 1$. (There is little effect at small $\khat$ but this case is
-not of interest because the local model breaks down.) Thus, 
-a growing fundamental VSI mode exists only for nearly
-isothermal perturbations ($\beta\ll 1$).  The plot is also consistent with our previous
-discussion that there exists an intermediate wavenumber for which
-finite thermal relaxation has maximal effect (for small $\beta$). 
-%$k_x\to0$ or $k_x\to\infty$ as $\beta\to0$.
+% It is simplest to solve Eq. \ref{relax_disp} numerically. In
+% Fig. \ref{relax_disp_fig} we plot the growth rate of the fundamental
+% ($M=0$) VSI mode as a function of $\khat$ for a range of
+% dimensionless thermal relaxation timescales $\beta$, in a disk with
+% $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. Increasing $\beta$ even
+% slightly above zero significantly reduces the growth rates for
+% $\khat\gg 1$. (There is little effect at small $\khat$ but this case is
+% not of interest because the local model breaks down.) Thus, 
+% a growing fundamental VSI mode exists only for nearly
+% isothermal perturbations ($\beta\ll 1$).  The plot is also consistent with our previous
+% discussion that there exists an intermediate wavenumber for which
+% finite thermal relaxation has maximal effect (for small $\beta$).
 
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm 0cm]{figures/rate_theory_grow_relax}
-%  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_grow_relax2}
-  \caption{Growth rate of the fundamental VSI mode in a disk with
-    dimensionless thermal relaxation timescales
-    $\beta\in[10^{-3},1]$.  Other disk parameters are
-    $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The growth rates are
-    calculated from the dispersion relation Eq. \ref{relax_disp}. 
-    \label{relax_disp_fig}}  
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm 0cm]{figures/rate_theory_grow_relax}
+% %  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_grow_relax2}
+%   \caption{Growth rate of the fundamental VSI mode in a disk with
+%     dimensionless thermal relaxation timescales
+%     $\beta\in[10^{-3},1]$.  Other disk parameters are
+%     $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The growth rates are
+%     calculated from the dispersion relation Eq. \ref{relax_disp}. 
+%     \label{relax_disp_fig}}  
+% \end{figure}
 
@@
 \end{align}
-is required for the fundamental VSI to operate.
+is required for the fundamental VSI to operate. From our discussion
+for isothermal perturbations (\S\ref{iso_discuss}), we expect
+$\beta<\beta_\mathrm{crit}$ to be a sufficient condition for the
+VSI. That is, when $\beta<\beta_\mathrm{crit}$, perturbations are
+effectively isothermal. For a thin disk 
+$\beta_\mathrm{crit}\ll 1$, which suggests that finite thermal
+relaxation rapidly stabilizes the VSI.
 
-In Fig. \ref{relax_bcrit} we plot the critical thermal relaxation
-timescale obtained from solving Eq. \ref{relax_disp_fund} numerically 
-for a disk with $\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$, for which
-we have $\beta_\mathrm{crit}=0.125$ from Eq. \ref{iso_vsi_cond}. Our
-upper estimate agrees well with the numerical solution for $\khat\gtrsim
-10$. From our discussion for isothermal perturbations
-(\S\ref{iso_discuss}), we expect $\beta<\beta_\mathrm{crit}$ to be a
-sufficient condition for the VSI.
+% In Fig. \ref{relax_bcrit} we plot the critical thermal relaxation
+% timescale obtained from solving Eq. \ref{relax_disp_fund} numerically 
+% for a disk with $\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$, for which
+% we have $\beta_\mathrm{crit}=0.125$ from Eq. \ref{iso_vsi_cond}. Our
+% upper estimate agrees well with the numerical solution for $\khat\gtrsim
+% 10$.
 
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/bcrit_theory}
-  \caption{Condition on the thermal relaxation timescale $\beta$ for
-    the existence of unstable fundamental VSI mode in a vertically
-    isothermal disk with $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.  
-    The dotted line denotes the asymptotic value
-    $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$ for large $\khat$. 
-    \label{relax_bcrit}}  
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\linewidth]{figures/bcrit_theory}
+%   \caption{Condition on the thermal relaxation timescale $\beta$ for
+%     the existence of unstable fundamental VSI mode in a vertically
+%     isothermal disk with $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.  
+%     The dotted line denotes the asymptotic value
+%     $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$ for large $\khat$. 
+%     \label{relax_bcrit}}  
+% \end{figure}
 
@@
 
-\subsection{Spurious growth of adiabatic perturbations in the local
-  model}\label{analytic_adia} 
-For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
-simply use the Soilberg-Hoiland critera to assess stability
-(\S\ref{solberg}). Nevertheless, we briefly consider this case in the
-local framework to demonstrate its limitations. For $\beta\to\infty$
-the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
-be written as
-
-\begin{align}
-  0 =& \dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
-    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime\notag\\
-  &+\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right)\right.\notag\\
-  &\phantom{0=}\left.-\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
-          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
-\end{align}
-
-We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
-integrate vertically, assume boundary terms vanish when integrating by
-parts, to obtain
-\begin{align}
-  &\hat{\sigma}^2\left(\frac{1}{\gamma} +
-    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
-  v_z|^2 d\zhat \notag\\
-  &=  \left(\frac{\gamma-1}{\gamma}\right)
-  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
-  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
-  v_z)^\prime|^2 d\zhat\notag\\
-  &\phantom{=} +
-  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
-      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
-    2}|\delta v_z|^2 d\zhat\notag\\
-  &\phantom{=} + \ii\epsilon q \hat{k}
-  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
-  \delta v_z d\zhat.\label{adia_integral}
-\end{align}
-When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
-$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
-expected, a sub-adiabatically stratified disk is stable in the absense
-of vertical shear.
-
 % If we follow through the solution method as before, we obtain a
@@
 % the low-frequency, Keplerian approximation is made for the latter.
-
-However, in the presence of vertical shear $q\neq0$,
-Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
-$\khat$, which implies the possibility of instability, regardless of
-how much $\gamma$ deviates from unity. This is contradicts the second 
-Solberg-Hoiland criterion (Eq. \ref{solberg2}) and is a result of the
-radially-local approximation (Appendix \ref{global_corr}). To 
-avoid this, we must restrict our attention to short thermal relaxation
-timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
-($\gamma=\Gamma$) when allowing vertical shear.
 

### 2015-02-07 linear_iso.tex
updating
undergoing major revisions

@@
-\section{Analytical discussion of vertically isothermal disks}\label{analytical}
+\section{Analytical discussion of vertically isothermal
+  disks in the framework of the vertically global shearing
+  box}\label{analytical}  
 We begin with a discussion of vertically isothermal disks
-where $\Gamma=1$, then $c_s$ is a constant in the shearing box. We
-first examine the limiting case of isothermal perturbations
-($\gamma\equiv1$ and any value of $\beta$, or any value of $\gamma$ with
-$\beta = 0$) before considering the effect of thermal relaxation. In the
-final section we consider adiabatic disks ($\gamma>1$ and
-$\beta\to\infty$) to demonstrate an important limitation of the 
-local model.
+where $\Gamma=1$, then $c_s$ is a constant. We  
+first examine isothermal perturbations ($\gamma\equiv1$ and any value
+of $\beta$, or any value of $\gamma$ with $\beta = 0$) before
+considering the effect of finite thermal relaxation. 
+In order to facilitate an analytical discussion, we introduce a number
+of approximations below that allows a simple governing equation to be
+derived and directly compared with previous works.
 
-\subsection{Approximate governing equation for general perturbations}\label{approx_gov}
-To facilitate a simple discussion, we derive a governing
-equation with the following approximations:
+\subsection{The linear problem in the vertically global
+  shearing box} 
+We connect the above linearized equations to the vertically global
+shearing box (VGSB) formalism developed by \cite{mcnally14}. The VGSB
+is a generalization of the standard shearing box \citep{goldreich65}
+by allowing the shear flow in the shearing box to be height-dependent.
 
+The linear problem in the VGSB framework can be derived from 
+Eq. \ref{lin_mass}---\ref{lin_vz} by setting $\hat{g}_c=0$, since 
+the VGSB, like the standard shearing box, is homogeneous in the radial 
+direction in equilibrium. The resulting linear problem is very similar
+to that for local axisymmetric waves in accretion disks already
+discussed by \cite{lubow93}, but modified by the presence of vertical
+shear and thermal relaxation. This is convenient as it allows us to
+directly apply their solution methods.
+
+% but modified by the vertical shear term
+% $r\p_z\Omega\delta \delta v_z$.
+
+After setting $\hat{g}_c=0$ in Eq. \ref{lin_mass}, \ref{lin_vx} and
+\ref{lin_energy}, we can eliminate the perturbed velocities between
+the linearized equations to obtain a pair or ordinary differential
+equations for $W,Q$ as
+\begin{align}
+  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
+  &=\left(\frac{\ii
+      k_x r}{D}\frac{d\Omega^2}{dz} -
+    \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
+    \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
+  &- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
+    \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
+  &\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
+  \frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
+  &=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
+  \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
+\end{align}
+where
+\begin{align}
+  D \equiv \kappa^2 - \sigma^2,
+\end{align}
+and we have replaced the notation $\p_z$ by $d/dz$ since we are
+considering solutions at a fixed radius.
+
+The VGSB is a self-consistent framework but extrapolating results 
+obtained from the VGSB to the global disk has limitations. This is
+because within the VGSB framework, $d\Omega/dz$ can be freely
+specified. In effect, by adopting the VGSB framework, we are studying
+dynamics in the standard shearing box subject to an additional
+azimuthal force $\delta v_zr d \Omega/dz$ with $d\Omega/dz$ as an
+input parameter. This is different to the global problem because in
+that case, the vertical shear is a result of the background radial
+disk structure, which is ignored in the VGSB.
+
+We further discuss caveats of the VGSB in Appendix \ref{global_corr}.
+However, we find that for perturbations with small radial lengthscales
+(large $|k_x|$) and fast thermal relaxation ($\beta\ll1$), which
+characterizes the VSI, the VGSB framework is adequate for the purpose
+of a simple analytical discussion.
+
+% In effect, we are studying dynamics in the standard shearing box
+% subject to an additional azimuthal force $-v_zr_0 d \Omega_0/dz$.
+
+% $d\Omega_0/dz$ is a free parameter (as are $\Omega_0$ and $S_0$). 
+% This is different to the global problem because in that case, the
+% vertical shear is a result of the disk radial structure, which is
+% ignored in the VGSB equilibria. 
+% In effect, we are studying dynamics in the standard shearing box
+% subject to an additional azimuthal force $-v_zr_0 d \Omega_0/dz$. This
+% force is inspired by the vertical shear in the global disk, and its
+% functional form is taken from global equilibrium.
+
+% \subsection{Limitations}
+% Before using the above equations to study the VSI, it is very important to be
+% aware of several limitations of the VGSB model. The VGSB momentum  
+% equations have the same form as that for the standard shearing box 
+% \citep{goldreich65}, except for the appearence of the term 
+% $v_zr_0d\Omega_0/dz$. This term appears in the global stability
+% problem and is the only term associated with vertical shear that is 
+% ported from the global disk into the local model.
+
+% In Appendix \ref{global_corr} we make the connection between the
+% linear problem in the VGSB to that in the global disk model. We
+% emphasize that not all calculations in the local model can be 
+% consistently interpreted in the global context. However, for
+% perturbations with small radial lengthscales (large $|k_x|$) and fast
+% thermal relaxation ($\beta\ll1$), which characterizes the VSI, the
+% VGSB framework is adequate.
+
+\subsection{Approximate governing equation}\label{approx_gov}
+We now derive a governing equation with further approximations:
 \begin{enumerate}
@@
 where $^\prime$ denotes $d/d\zhat$. Eq. \ref{vertiso_gov_nondim} will
-be the basis for our discussion of vertically isothermal disks,
-although further simplifications or maniputations will be performed
-depending the type of perturbations being considered.
+be the basis for our discussion of vertically isothermal disks. % ,
+% although further simplifications or maniputations will be performed
+% depending the type of perturbations being considered.
 
@@
 \end{align} 
-Eq. \ref{iso_ode3} is Hermite's differential equation. If we impose
+Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
 that the kinetic energy density remain bounded at infinity, then  
@@
 
-In Appendix \ref{pert_theory} we give an alternative method to
-infer instability in the presence of vertical shear without explicitly
-solving the governing equation.
+% In Appendix \ref{pert_theory} we give an alternative method to
+% infer instability in the presence of vertical shear without explicitly
+% solving the governing equation.

### 2015-02-07 numerical.tex
undergoing major revisions

@@
 We relax the thin-disk, low-frequency and nearly-Keplerian
-approximations used in the discussion above, and 
-use the full, non-Keplerian expressions of $\rho$, $\Omega$ and
-$\kappa$.
+approximations used in the analytical discussion above, and  
+use the full expressions of $\rho$, $\Omega$ and
+$\kappa$ given in \S\ref{eqm}. We also restore $\hat{g}_c=1$ to
+account for the background radial disk structure.
+
 %, so that we may consistently consider vertical domains up to
@@
 
-%mention mode selection -> focus on fundamental corrugation ->
-%simulations eventually dominated by them 
-%choice of kx is O(10), since nelson observe kxH \sim 30
+% mention mode selection -> focus on fundamental corrugation ->
+% simulations eventually dominated by them 
+% choice of kx is O(10), since nelson observe kxH \sim 30
 
@@
 $(p,q,\epsilon)=(-1.5,-1,0.1)$ and the vertical domain is $\zmax =
-10\epsilon r$. This setup is very similar to that adopted in
+5\epsilon r$. This setup is similar to that adopted in
 \cite{mcnally14}, who calculated the VSI in vertically isothermal
@@
 Fig. \ref{iso_eigen_kx} compares the numerically-obtained eigenvalues
-for the fundamental mode and that caluclated from
-Eq. \ref{simple_growth} (with $L=1$). For completeness we
-included cases with $\khat\ll 1$ even though the local
-approximation does not hold in this regime. The agreement is good for
-$\khat\gtrsim10$. The mismatch is significant for $\khat\ll1$ since
-the low-frequency approximation used to obtain Eq. \ref{simple_growth}
-is no longer valid: $|\omega|\sim\Omega_k$ as $\khat\to 0$. This means
-that our analytic results should only be applied to perturbations with
-$\khat\gg 1$, i.e. small radial-lengthscales.
+for the fundamental mode and that calculated from 
+Eq. \ref{sig2_iso} (with $L=1$). There is good agreement for 
+$\khat\gtrsim10$. The match worsens for smaller $\khat$ because the 
+the low-frequency approximation used to obtain Eq. \ref{sig2_iso} 
+is no longer valid, as $|\omega|$ is not small compared to $\Omega_k$
+for $\khat\lesssim 10$.
+
+% This means
+% that our analytic results should only be applied to perturbations with
+% $\khat\gg 1$, i.e. small radial-lengthscales.
 
@@
 
-%particular kx 
-As an example of an unstable mode, we consider a perturbation
-wavenumber $\khat=20\pi$ \citep[or $k_x = 200\pi/r$ as used
-in][]{mcnally14}. Fig. \ref{lowfreq_eigen} shows the eigenvalues
-$\sigma = \omega + \ii\nu$. The fundamental VSI mode, with one node in $W$, has
-the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}. 
-The eigenfunction $W$ is qualitatively proportional to 
-$z$ (especially for $z\to0$),  as expected for the fundamental mode
-according to explicit solutions developed in \S\ref{iso_poly}.    
-These plots are consistent with those in \cite{mcnally14}.
+Fig. \ref{lowfreq_eigenfunc} compares the fundamental VSI eigenfunction
+$W$ for $\khat=1$ and $\khat=20\pi$. The latter value, corresponding
+to $k_x= 200\pi/r$, was used by \cite{mcnally14}. We find $W$ is
+similar for all $\khat\gtrsim10$ and is proportional to $z$, as
+expected for the fundamental mode according to explicit solutions
+developed in \S\ref{iso_poly}.
 
-The growth rate of the fundamental VSI calculated from
-Eq. \ref{simple_growth} with $L=1$ and the numerically-obtained value
-are  
+However, for $\khat\lesssim 10$, e.g. $\khat=1$ in 
+Fig. \ref{lowfreq_eigenfunc}, the perturbation magnitude near the
+boundaries begin to dominate. For even smaller $\khat\lesssim 1$ (not
+shown) we find the perturbation is almost entirely concentrated at the
+boundaries. Our radially-local approximation is likely invalid for 
+these small wavenumber perturbations and we do not consider such modes
+further.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvector_iso_kx1} 
+  \includegraphics[width=\linewidth]{figures/eigenvector_iso_kx60}
+  \caption{Eigenfunction of the fundamental VSI for perturbation
+    wavenumber $\khat=1$ (top) and $\khat=20\pi$ (bottom). The real 
+    (imaginary) parts of $W$ are plotted as solid (dotted) lines. 
+    \label{lowfreq_eigenfunc}
+  }
+\end{figure}
+
+In Fig. \ref{lowfreq_eigen} we plot the eigenvalues 
+$\sigma = \omega + \ii\nu$ found for $\khat=20\pi$. This plot is
+similar to Fig. 3 in \cite{mcnally14}. The fundamental
+VSI mode, with one node in $W$, has the smallest $|\sigma|$. The
+eigenvalue calculated from Eq. \ref{sig2_iso} with $L=1$ and the 
+numerically-obtained value are  
 \begin{align*}
-  \nu = 0.2606\epsilon\Omega_k &\quad \text{(from
-    Eq. \ref{simple_growth})},\\
-  \nu = 0.2575 \epsilon \Omega_k &\quad \text{(numerical)}.
+  (\omega,\nu) = (-0.3053,0.2606)\epsilon\Omega_k &\quad \text{(from
+    Eq. \ref{sig2_iso})},\\
+  (\omega,\nu) = (-0.3010,0.2575) \epsilon \Omega_k &\quad \text{(numerical)}.
 \end{align*}
-This agreement is surpsingly good, given that
-Eq. \ref{simple_growth} assumes a thin, nearly-Keplerian disk in the 
-low-frequency approximation, and 
-imposes a different boundary condition to that in the numerical
-calculation. These values are also close to that found by
-\cite{mcnally14}. The eigenfrequency is not sensitive to the
-vertical boundary condition because the expression for $\sigma$, given
-by Eq. \ref{integral_relation1}, involve integrands with $\rho$ as the  
-weight function, which rapidly decays away from the midplane.
+This agreement is surpsingly good given the number of approximations
+used to obtain Eq. \ref{sig2_iso}, which also imposes a different
+boundary condition to that in the numerical calculation. These values
+are also close to that found by \cite{mcnally14}. The eigenfrequency
+is not sensitive to the vertical boundary condition because the
+expression for $\sigma$, given by Eq. \ref{integral_relation1},
+involve integrands with $\rho$ as the   weight function, which rapidly
+decays away from the midplane.
 
@@
   \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
-  \caption{Eigenvalues of the VSI in a neutrally-stratified,
-    nearly-vertically isothermal disk  ($\gamma=\Gamma=1.011$) with 
-    $(p,q,\epsilon)=(-1.5,-1,0.1)$ and $\beta=10^{-3}$. The
-    perturbation wavenumber is $\khat=20\pi$. \label{lowfreq_eigen} 
-  }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/eigenvector_iso}
-  \caption{Eigenfunction of the fundamental VSI,
-    corresponding to the bottom-left eigenvalue displayed in
-    Fig. \ref{lowfreq_eigen} (smallest $|\sigma|$). The real
-    (imaginary) parts of $W$ are plotted as solid (dotted) lines. 
-    \label{lowfreq_eigenfunc}
+  \caption{Eigenvalues of perturbations with wavenumber $\khat=20\pi$
+    in a neutrally-stratified, nearly-vertically isothermal disk
+    ($\gamma=\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+    $\beta=10^{-3}$. \label{lowfreq_eigen} 
   }
@@
 
-Fig. \ref{relax_growth_num} shows the growth rates of the fundamental  
-mode as a function of $\khat$. This plot is largely consistent with  
-Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
-where the low-frequency approximation fails. (In any case, the local model not valid
-for $\khat\ll1$.) For $\khat\gtrsim 10$ introducing thermal
-relaxation rapidly stabilizes the fundamental mode.
-
-% For $\beta=0.063$ at $\khat\gtrsim50$ we found eigenfunctions had 
-% large perturbations near the vertical boundaries, which may be 
-% responsible for the non-smoothness in the curve at large $\khat$. 
-%corrugation mode
-% We first obtain the eigenvalues for the
-% fundamental mode for $k_xH_\mathrm{iso}=0$, then vary parameters
-% slowly to obtain eigenvalues for other values of $k_x$ or $\beta$.  
-%likely attributable to the free surface boundary condition imposed in
-%the numerical calculation (cf. regularity condition at infinity in
-%\S\ref{disp_relax}).
-
-Fig. \ref{relax_growth_num} also confirms our analytical discussion that
-introducing a small but finite cooling time stabilizes the fundamental
-VSI, but there is a preferable wavenumber that maximizes this effect (\S\ref{relax_pert}).
-This occurs at $\khat= O(10)$ for the current disk model.
+Fig. \ref{bcrit_compare} shows the growth rate $\nu$ of the
+fundamental VSI mode as a function of $\beta$ for $\khat=30$. This is
+approximately the perturbation wavenumber observed in the fiducial
+numerical simulation of \cite{nelson13}. We also plot the growth rate 
+obtained by numerically solving the dispersion relation 
+Eq. \ref{relax_disp} and the estimate for the upper limit of
+the thermal relaxation timescale $\beta_\mathrm{crit}$ given by
+Eq. \ref{iso_vsi_cond}. For the present disk parameters
+$\beta_\mathrm{crit}=0.125$. We find good agreement between the
+numerical and analytical growth rates, as well as the critical
+relaxation timescale.
 
  \begin{figure}
-   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
-   0cm]{figures/compare_eigen_imag_bloop} 
-  % \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  % 1cm]{figures/compare_eigen_imag_bloop2} 
-  \caption{Growth rates of the fundamental VSI mode in 
-    nearly-vertically isothermal disks ($\Gamma=1.011$) with
-    $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ under a
-    range of thermal relaxation timescales   
-    $\beta$. The eigenfrequencies are calculated by numerically 
-    solving the full linear eigenvalue problem. \label{relax_growth_num}}   
+   \includegraphics[width=\linewidth]{figures/bcrit_compare} 
+   \caption{Growth rate of the fundamental VSI mode with $\khat=30$ as
+     a function of the thermal relaxation timescale $\beta$. The disk
+     parameters are $\Gamma=1.011$, $\gamma=1.4$, and 
+     $(p,q,\epsilon)=(-1.5,-1,0.05)$. The numerical (analytical) growth
+     rates are show as solid (dotted) lines, while the vertical line
+     is the maximum thermal timescale $\beta_\mathrm{crit}$ for the
+     fundamental mode according to Eq. \ref{iso_vsi_cond}. 
+     \label{bcrit_compare}}   
  \end{figure}
+
+% Fig. \ref{relax_growth_num} shows the growth rates of the fundamental  
+% mode as a function of $\khat$. This plot is largely consistent with  
+% Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
+% where the low-frequency approximation fails. (In any case, the local model not valid
+% for $\khat\ll1$.) For $\khat\gtrsim 10$ introducing thermal
+% relaxation rapidly stabilizes the fundamental mode.
+
+% Fig. \ref{relax_growth_num} also confirms our analytical discussion that
+% introducing a small but finite cooling time stabilizes the fundamental
+% VSI, but there is a preferable wavenumber that maximizes this effect (\S\ref{relax_pert}).
+% This occurs at $\khat= O(10)$ for the current disk model.
+
+% \begin{figure}
+%    \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
+%    0cm]{figures/compare_eigen_imag_bloop} 
+%   % \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+%   % 1cm]{figures/compare_eigen_imag_bloop2} 
+%   \caption{Growth rates of the fundamental VSI mode in 
+%     nearly-vertically isothermal disks ($\Gamma=1.011$) with
+%     $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ under a
+%     range of thermal relaxation timescales   
+%     $\beta$. The eigenfrequencies are calculated by numerically 
+%     solving the full linear eigenvalue problem. \label{relax_growth_num}}   
+% \end{figure}
 
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/eigenvectorW_beta0d125} 
+  0cm]{figures/eigenvectorW_beta0d1} 
   \caption{Numerically-calculated eigenfunction of the fundamental VSI
@@
     thermal relaxation timescale $\beta \equiv t_c\Omega_k = 0.01$
-    (top) and $0.125$ (bottom). The
-    perturbation radial wavenumber is $\khat_\mathrm{iso}=10$. The
+    (top) and $0.1$ (bottom). The
+    perturbation radial wavenumber is $\khat=30$. The
     real (imaginary) part of $W$ is plotted as the solid (dotted)

### 2015-02-07 setup.tex
undergoing major revisions

@@
-\section{Baroclinic disk equilibra}
+\section{Governing equations and disk models}
 We consider  an inviscid, non-self-gravitating three-dimensional (3D)
-disk orbiting a central star of mass $M_*$. 
-We adopt cylindrical
+disk orbiting a central star of mass $M_*$. We adopt cylindrical
 co-ordinates $(r,\phi, z)$ centered on the star to describe the global
-disk. The equilibrium disk is steady and axisymmetric with density,
+disk. The governing fluid equations are
+\begin{align}
+  &\frac{\p\rho}{\p t} + \nabla\cdot{\left(\rho\bm{v}\right)} = 0,\\
+  &\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} =
+  -\frac{1}{\rho}\nabla P - \nabla\Phi_*,\\
+  &\frac{\p P}{\p t} + \bm{v}\cdot\nabla P  = - \gamma P
+  \nabla\cdot\bm{v} - \frac{\rho}{t_c}\left(\frac{P}{\rho} -
+    \left.\frac{P}{\rho}\right|_{t=0}\right), \label{full_energy}
+\end{align}
+where $\rho$ is the mass density, $\bm{v}$ is the velocity field (the
+rotation frequency being $\Omega=v_\phi/r$), $P$
+is the pressure, $\gamma$ is the constant adiabatic index, and $\Phi_*
+= -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational potential of the
+central star and $G$ is the gravitational constant. 
+In the energy equation (Eq. \ref{full_energy}) we have
+included a thermal relaxation term which restores the disk temperature
+$T\propto P/\rho$ to its initial value on a characteristic timescale
+$t_c=\beta\Omega_k^{-1}$ where $\beta$ is a dimensionless coefficient
+and $\Omega_k=\sqrt{GM_*/r^3}$ is the Keplerian frequency.
+
+\subsection{Baroclinic disk equilibria}\label{eqm}
+The equilibrium disk is steady and axisymmetric with density,
 pressure and rotation profiles $\rho(r,z)$, $P(r,z)$ and
@@
   r\Omega^2&= \frac{1}{\rho}\frac{\p P}{\p r} + \frac{\p \Phi_*}{\p
-    r},\label{hori_equilibrium} 
+    r}.\label{hori_equilibrium} 
 \end{align} 
-where $\Phi_* = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational
-potential of the central star.
-
+%where $\Phi_* = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational
+%potential of the central star.
 We examine disks with power-law profiles for the midplane density
@@
 
-\subsection{Vertically isothermal disks}
+\subsubsection{Vertically isothermal disks}
 For the special case $\Gamma=1$ the disk is isothermal in the
@@
 
-\section{The vertically global shearing box} 
-We are interested in the axisymmetric stability of the above system,
-on radial length-scales much smaller than the local radius $r$, but
-may be global in the vertical direction. As such, we work within the
-framework of the \emph{vertically global shearing 
-  box} (VGSB) developed by \cite{mcnally14}. In this approximation
-one sets up a local Cartesian co-ordinate  system $(x,y,z)$ centered
-on a fiducial point $(r,\phi,z)=(r_0,\phi_0,0)$ in the global disk, with $\phi_0(t) =
-\Omega(r_0,0)t$. The local $x,y,z$ axis correspond to the
-global radial, azimuthal and vertical directions.
-
-In this shearing box the equilibrium density, pressure and
-sound-speed are assumed to depend only on $z$, by evaluating these
-quantities at $r=r_0$ in the global disk. The background rotation
-$\Omega$ is approximated by a linear shear in $x$, but left unaltered
-in $z$. However, the vertical shear $\p_z\Omega$ is also assumed to
-only depend on $z$.
-
-The axisymmetric VGSB equations read 
+\section{Linear problem} 
+We consider axisymmetric perturbations to the above equilibria in the
+form 
 \begin{align}
-  & \frac{\p\rho}{\p t} + \frac{\p}{\p x}\left(\rho v_x\right) +
-  \frac{\p}{\p z}\left(\rho v_z\right)=0,\\
-  & \frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
-  v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}} = - 
-  2\Omega_0\hat{\bm{z}}\times\bm{v} - S_0v_x\hat{\bm{y}} \notag\\ 
-  &\phantom{\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
-    v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}}=}-\frac{1}{\rho}\nabla P - \frac{\p\Phi_0}{\p z}\hat{\bm{z}},\\
-  & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P = -\gamma P
-  \nabla\cdot\bm{v} -
-  \frac{\rho}{t_c}\left(\frac{P}{\rho}-\left.\frac{P}{\rho}\right|_{t=0}\right), \label{energy_eq}
-\end{align}
-where $\bm{v}$ is the velocity field relative to the bulk flow
-(i.e. $\bm{v}=\bm{0}$ in equilibrium),  
-subscript $0$ indicates evaluation at $r=r_0$. In the energy equation 
-we have added a thermal relaxation term with timescale
-$t_c=\beta\Omega_k^{-1}$ with $\beta$  being a dimensionless input
-parameter.
-
-\subsection{Linearization}
-We perturb the system such that
-\begin{align}
-  \rho \to \rho(z) + \delta\rho(z)\exp{\left(\ii k_x x - \ii\sigma
+  \rho \to \rho(z) + \delta\rho(r,z)\exp{\left( - \ii\sigma
       t\right)},    
 \end{align}
-and similarly for other fluid variables, where $k_x$ is a horizontal
-wavenumber and $\sigma = \omega + \ii \nu$ is a (generally) complex
-frequency, with $\omega$ amd $\nu$ being the real frequency and growth
-rate, respectively. For clarity we drop the subscript $0$ on the
-frequencies and shear rates below. Then writing $\delta P /\rho \equiv W$
-and $c_s^2\delta\rho/\rho\equiv Q$, the linearized system of equations
-is   
+and similarly for other fluid variables, where $\sigma = \omega + \ii
+\nu$ is a (generally) complex frequency, with $\omega$ amd $\nu$ being
+the real frequency and growth rate, respectively.
+
+The linearized fluid equations in cylindrical co-ordinates are  
 \begin{align}
-   -&\ii\sigma \frac{Q}{c_s^2} + \ii k_x \delta v_x + \frac{d\dd
-     v_z}{dz} + \frac{d\ln{\rho}}{dz}\delta v_z = 0,\label{lin_mass}\\
-   &\ii\sigma \delta v_x + 2\Omega\delta v_y = \ii k_x W,\label{lin_vx}\\
-   -&\frac{\kappa^2}{2\Omega}\delta v_x + \ii \sigma\delta v_y =
-   r\frac{d\Omega}{dz}\delta v_z, \label{lin_vy}\\
-   & \ii\sigma\delta v_z = \frac{dW}{dz} +
-   \frac{d\ln{\rho}}{dz}\left(W-Q\right), \label{lin_vz}\\
-   -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
-   c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
+  \ii\sigma\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
+    z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\\
+  \ii\sigma\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
+  \delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
+  \ii\sigma\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
+  r\frac{\p\Omega}{\p z}\delta v_z,\\
+  \ii \sigma\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
+  \frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},\\
+  \ii\sigma\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
+  \left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
+      \delta v_z}{\p z}\right]\notag \\
+  &+ \frac{1}{t_c}\left(\delta P -  
+    \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,
 \end{align}
-It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa$
-are evaluated at $r=r_0$ using their global expressions above.
+where the coefficients $\hat{g}_c=1$ are inserted to label terms
+associated with the background radial disk structure.
 
-We can eliminate the perturbed velocities between the linearized
-equations to obtain a pair or ordinary differential equations
-for $W,Q$ as
+\subsection{Radially-local approximation}  
+We assume perturbations have spatial dependence in the form
 \begin{align}
-  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
-  &=\left(\frac{\ii
-    k_x r}{D}\frac{d\Omega^2}{dz} -
-  \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
-  \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
-&- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
-  \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
-&\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
-\frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
-&=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
-\left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
+  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r} 
 \end{align}
-where
+such that $|k_rr|\gg1$, so that curvature terms can be ignored. For
+these small radial lengthscale perturbations we also ignore the radial
+variations in the background structure and radial gradients, and
+evaluate all cofficients in the above equations at the fiducial radius
+$r=r_0$. However, the full vertical dependence of the coefficients are
+retained.
+
+In order to emphasize on the radially-local approximation, we
+introduce the local Cartesian co-ordinate $(x,y,z)$ corresponding to
+the the local $(r,\phi,z)$ directions, respectively. Then writing
+$\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
+linearized system of equations are 
 \begin{align}
-  D \equiv \kappa^2 - \sigma^2.
+  \ii\sigma \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
+    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \hat{g}_c
+  \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
+  \ii\sigma \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
+  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
+   \ii \sigma\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
+  \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
+   \ii\sigma\delta v_z &= \frac{dW}{dz} +
+  \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
+  \ii \sigma W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
+  c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
+      v_z}{dz}\right) \notag\\
+  &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
+  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x.\label{lin_energy}
+  % -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
+  % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
 \end{align}
+It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
+\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$.   
+Eq. \ref{lin_mass}---\ref{lin_vz} will form the basis of our numerical
+study of the vertical shear instability, but for analytical discussion
+we will make further simplifications.
 
-\subsection{Limitations}
-Before using the above equations to study the VSI, it is very important to be
-aware of several limitations of the VGSB model. The VGSB momentum  
-equations have the same form as that for the standard shearing box 
-\citep{goldreich65}, except for the appearence of the term 
-$v_zr_0d\Omega_0/dz$. This term appears in the global stability
-problem and is the only term associated with vertical shear that is 
-ported from the global disk into the local model.
+% We are interested in the axisymmetric stability of the above system,
+% on radial length-scales much smaller than the local radius $r$, but
+% may be global in the vertical direction. As such, we work within the
+% framework of the \emph{vertically global shearing 
+%   box} (VGSB) developed by \cite{mcnally14}. In this approximation
+% one sets up a local Cartesian co-ordinate  system $(x,y,z)$ centered
+% on a fiducial point $(r,\phi,z)=(r_0,\phi_0,0)$ in the global disk, with $\phi_0(t) =
+% \Omega(r_0,0)t$. The local $x,y,z$ axis correspond to the
+% global radial, azimuthal and vertical directions.
 
-Note that, within the VGSB framework, the vertical shear 
-$d\Omega_0/dz$ is a free parameter (as are $\Omega_0$ and $S_0$). 
-This is different to the global problem because in that case, the
-vertical shear is a result of the disk radial structure, which is
-ignored in the VGSB equilibria. 
-In effect, we are studying dynamics in the standard shearing box
-subject to an additional azimuthal force $-v_zr_0 d \Omega_0/dz$. This
-force is inspired by the vertical shear in the global disk, and its
-functional form is taken from global equilibrium.
+% In this shearing box the equilibrium density, pressure and
+% sound-speed are assumed to depend only on $z$, by evaluating these
+% quantities at $r=r_0$ in the global disk. The background rotation
+% $\Omega$ is approximated by a linear shear in $x$, but left unaltered
+% in $z$. However, the vertical shear $\p_z\Omega$ is also assumed to
+% only depend on $z$.
 
-%horizontal structure that gives rises 
-%vertical shear is an input force
-%only porting vertical shear  
-%vertical shear doesn't depend on x 
-%not all results can be extrapolated to the global context
+% The axisymmetric VGSB equations read 
+% \begin{align}
+%   & \frac{\p\rho}{\p t} + \frac{\p}{\p x}\left(\rho v_x\right) +
+%   \frac{\p}{\p z}\left(\rho v_z\right)=0,\\
+%   & \frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
+%   v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}} = - 
+%   2\Omega_0\hat{\bm{z}}\times\bm{v} - S_0v_x\hat{\bm{y}} \notag\\ 
+%   &\phantom{\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
+%     v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}}=}-\frac{1}{\rho}\nabla P - \frac{\p\Phi_0}{\p z}\hat{\bm{z}},\\
+%   & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P = -\gamma P
+%   \nabla\cdot\bm{v} -
+%   \frac{\rho}{t_c}\left(\frac{P}{\rho}-\left.\frac{P}{\rho}\right|_{t=0}\right), \label{energy_eq}
+% \end{align}
+% where $\bm{v}$ is the velocity field relative to the bulk flow
+% (i.e. $\bm{v}=\bm{0}$ in equilibrium),  
+% subscript $0$ indicates evaluation at $r=r_0$. In the energy equation 
+% we have added a thermal relaxation term with timescale
+% $t_c=\beta\Omega_k^{-1}$ with $\beta$  being a dimensionless input
+% parameter.
 
-In Appendix \ref{global_corr} we make the connection between the
-linear problem in the VGSB to that in the global disk model. We
-emphasize that not all calculations in the local model can be 
-consistently interpreted in the global context. However, for
-perturbations with small radial lengthscales (large $|k_x|$) and fast
-thermal relaxation ($\beta\ll1$), which characterizes the VSI, the
-VGSB framework is adequate.
+% \subsection{Linearization}
+% We perturb the system such that
+%  For clarity we drop the subscript $0$ on the
+% frequencies and shear rates below.
+
+% We can eliminate the perturbed velocities between the linearized
+% equations to obtain a pair or ordinary differential equations
+% for $W,Q$ as
+% \begin{align}
+%   &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
+%   &=\left(\frac{\ii
+%     k_x r}{D}\frac{d\Omega^2}{dz} -
+%   \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
+%   \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
+% &- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
+%   \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
+% &\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
+% \frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
+% &=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
+% \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
+% \end{align}
+% where
+% \begin{align}
+%   D \equiv \kappa^2 - \sigma^2.
+% \end{align}

### 2015-02-07 numerical.tex
major updating

@@
 parameters as a fiducial case and vary $q\in[-0.2,-1.2]$,
-$\gamma\in[1.2,2.0]$ and $\epsilon\in[]$
+$\gamma\in[1.2,2.0]$ and $\epsilon\in[0.02,0.1]$
 separately. The numerically-obtained $\beta_\mathrm{crit}$ is shown in
@@
   0.8cm]{figures/bcrit_compare_g.ps}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
+  0.8cm]{figures/bcrit_compare_e.ps} 
   \caption{Dependence of the upper limit to the thermal relaxation timescale
@@
     $q\in[-1.2,-0.2]$; middle: varying $\gamma\in[1.2,2.0]$; bottom:
-    varying $\epsilon\in[]$. The perturbation wavenumber is
-    $\khat=20$. {\bf still need variable epsilon case}
+    varying $\epsilon\in[0.02,0.1]$. The perturbation wavenumber is
+    $\khat=10$. 
     \label{bcrit_compare}}  

### 2015-02-07 linear_iso.tex
major updating
updating

@@
 where $\Gamma=1$, then $c_s$ is a constant. We  
-first examine isothermal perturbations ($\gamma\equiv1$ and any value
-of $\beta$, or any value of $\gamma$ with $\beta = 0$) before
+first examine isothermal perturbations ($\beta = 0$) before
 considering the effect of finite thermal relaxation. 
@@
 \subsection{Isothermal perturbations}\label{iso_discuss}
-Isothermal perturbations corresponds to $\gamma=1$ for any thermal
-relaxation timescale $\beta$, or for any adiabatc index $\gamma$ when
-$\beta=0$. In either case $\chi=1$ and Eq. \ref{ode_Q} implies 
+Isothermal perturbations correspond to $\beta\equiv0$, 
+%relaxation timescale $\be
+%or for any adiabatc index $\gamma$ when
+%$\beta=0$. 
+so $\chi=1$ and Eq. \ref{ode_Q} implies 
 \begin{align}
@@
 
-We remark that Eq. \ref{iso_ode1} is also applicable for  $\gamma=\Gamma\neq 1$ in the
-limit $t_c\to\infty$ (i.e. neutrally stratified adiabatic disks),
-since in that case $Q\simeq W$ as well. However, the following
-discussion is only  valid for $\gamma=\Gamma=1$ because $c_s$ is taken
-to be a constant.
+%We remark that Eq. \ref{iso_ode1} is also applicable for  $\gamma=\Gamma\neq 1$ in the
+%limit $t_c\to\infty$ (i.e. neutrally stratified adiabatic disks),
+%since in that case $Q\simeq W$ as well. However, the following
+%discussion is only  valid for $\gamma=\Gamma=1$ because $c_s$ is taken
+%to be a constant.
 

### 2015-02-08 numerical.tex
updating

@@
 
-Fig. \ref{bcrit_compare} shows the growth rate $\nu$ of the
+Fig. \ref{bcrit_compare1} shows the growth rate $\nu$ of the
 fundamental VSI mode as a function of $\beta$ for $\khat=30$. This is
 approximately the perturbation wavenumber observed in the fiducial
-numerical simulation of \cite{nelson13}. We also plot the growth rate 
+numerical simulation of \cite{nelson13}. We also plot the growth rates
 obtained by numerically solving the dispersion relation 
-Eq. \ref{relax_disp} and the estimate for the upper limit of
+Eq. \ref{relax_disp} and the upper limit to 
 the thermal relaxation timescale $\beta_\mathrm{crit}$ given by
@@
 numerical and analytical growth rates, as well as the critical
-relaxation timescale.
+relaxation timescale. Growth rates are insensitive to $\beta$ for
+$\beta\lesssim 0.05$, beyond which growth rates rapidly decrease.
+
+For this disk model, \citeauthor{nelson13} found a thermal relaxation
+timescale of $\beta\gtrsim 0.6$ stabilized the VSI in non-linear
+hydrodynamic simulations. This is consistent with our numerical and
+analytic results.
+
+%implications for nonlinear sims?
 
@@
      fundamental mode according to Eq. \ref{iso_vsi_cond}. 
-     \label{bcrit_compare}}   
+     \label{bcrit_compare1}}   
  \end{figure}
@@
 Fig. \ref{relax_eigenW_num} by comparing the eigenfunction $W$ for
-$\beta=0.01$ and $\beta=0.125$. The latter value is close to the critical
-value beyond which the fundamental isothermal VSI is stabilized
-(\S\ref{iso_vsi_beta_crit}). Hence the latter growth rate is smaller
-than the former by a factor $\sim 20$. Increasing $\beta$ restricts 
-the region in which $W\sim z$ closer to the midplane, with oscillatory
-behavior emerging at the boundaries, where buoyancy first becomes
-important. % We also find this trend for fixed $\beta$ but increasing
+$\beta=0.01$ and $\beta=0.1$. The latter value is close to the
+critical value beyond which the fundamental VSI is stabilized
+(\S\ref{iso_vsi_beta_crit}). Increasing $\beta$
+restricts  the region in which $W\sim z$ closer to the midplane, with
+oscillatory behavior emerging at the boundaries, where buoyancy first
+becomes important.
+
+% We also find this trend for fixed $\beta$ but
+% increasing 
 % $\khat$.
@@
 \end{figure}
-
-For this disk model, \citeauthor{nelson13} found a thermal relaxation
-timescale of $\beta\gtrsim 0.6$ stabilized the VSI in non-linear
-hydrodynamic simulations. This is consistent with our numerical and
-analytic results.
 
@@
 
-\subsubsection{$\gamma=\Gamma \neq 1$, $N_z^2\equiv 0$}
-Consider first the case $\gamma=1.4$ so that there is no vertical
-entropy gradient. We plot in Fig. the growth rates for the
-fundamental VSI for $\beta\in[10^{-2},10^{2}]$. We find growth rates
-do not reach zero, implying instability regardless of the thermal
-relaxation timescale if $\gamma=\Gamma$. This result is consistent
-with \cite{nelson13}.  Note that for $\beta\to\infty$ (adiabatic
-perturbations) and $\gamma=\Gamma$, the disk can be unstable according
-to the Solberg-Hoiland criterion  Eq. \ref{solberg2}.
+% \subsubsection{$\gamma=\Gamma \neq 1$, $N_z^2\equiv 0$}
+% Consider first the case $\gamma=1.4$ so that there is no vertical
+% entropy gradient. We plot in Fig. the growth rates for the
+% fundamental VSI for $\beta\in[10^{-2},10^{2}]$. We find growth rates
+% do not reach zero, implying instability regardless of the thermal
+% relaxation timescale if $\gamma=\Gamma$. This result is consistent
+% with \cite{nelson13}.  Note that for $\beta\to\infty$ (adiabatic
+% perturbations) and $\gamma=\Gamma$, the disk can be unstable according
+% to the Solberg-Hoiland criterion  Eq. \ref{solberg2}.
 
-Growth rates for $\beta = 10^{-2}$ and
-$\beta=10^{2}$ behave similarly. This is not surprising because
-the linearized energy equation (Eq. \ref{lin_energy}) implies that 
-\begin{align}
-  W = \frac{\left(\Gamma^{-1} -
-      \ii\hat{\sigma}\beta\right)}{\left(1-\ii\hat{\sigma}\beta\right)}
-  Q \text{ for } \gamma\equiv \Gamma.   
-\end{align}
-Thus, $W \to Q/\Gamma$ as $\beta\to0$ and $W\to Q$ as
-$\beta\to\infty$. Since $\Gamma=O(1)$, these two limits are expected
-to produce qualitatively similar results. Interestingly,
-Fig. \ref{growth_vnoniso1} show that growth rates 
-maximize for $\beta=O(1)$. However, all growth rates are similar to
-order-of-magnitude for $0.1\lesssim\khat\lesssim 100$.
+% Growth rates for $\beta = 10^{-2}$ and
+% $\beta=10^{2}$ behave similarly. This is not surprising because
+% the linearized energy equation (Eq. \ref{lin_energy}) implies that 
+% \begin{align}
+%   W = \frac{\left(\Gamma^{-1} -
+%       \ii\hat{\sigma}\beta\right)}{\left(1-\ii\hat{\sigma}\beta\right)}
+%   Q \text{ for } \gamma\equiv \Gamma.   
+% \end{align}
+% Thus, $W \to Q/\Gamma$ as $\beta\to0$ and $W\to Q$ as
+% $\beta\to\infty$. Since $\Gamma=O(1)$, these two limits are expected
+% to produce qualitatively similar results. Interestingly,
+% Fig. \ref{growth_vnoniso1} show that growth rates 
+% maximize for $\beta=O(1)$. However, all growth rates are similar to
+% order-of-magnitude for $0.1\lesssim\khat\lesssim 100$.
 
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/growth_vnoniso1} 
-  \caption{Numerically-determined growth rates of the fundamental VSI
-    mode in neutrally-buoyant ($N_z^2\equiv0$), vertically non-isothermal disks
-    ($\Gamma=\gamma=1.4$), with $(p,q,\epsilon)=(0,-1,0.05)$, for
-    a range of thermal relaxation timescales 
-    $\beta$. \label{growth_vnoniso1}}     
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+%   0cm]{figures/growth_vnoniso1} 
+%   \caption{Numerically-determined growth rates of the fundamental VSI
+%     mode in neutrally-buoyant ($N_z^2\equiv0$), vertically non-isothermal disks
+%     ($\Gamma=\gamma=1.4$), with $(p,q,\epsilon)=(0,-1,0.05)$, for
+%     a range of thermal relaxation timescales 
+%     $\beta$. \label{growth_vnoniso1}}     
+% \end{figure}
 
-\subsubsection{$\gamma\neq\Gamma \neq 1$, $N_z^2\geq0$}
-%gamma = 2.0, Gamma = 1.4, vary bcool 
-Finally, we introduce buoyancy in a vertically non-isothermal disk by
-setting $\gamma=2$. This scenario was not considered by  
-\cite{nelson13}, but is similar in spirit with vertically isothermal 
-disks evolved with $\gamma>1$, as considered in the previous section. 
-Fig. \ref{growth_vnoniso2} plots the fundamental VSI growth rates for
-this case. We observe the same trend as for vertically isothermal
-disks with $N_z^2\geq 0$ (Fig. \ref{relax_growth_num}), namely growth
-rates reach zero at finite $\khat$. From Fig.\ref{growth_vnoniso2} we
-expect the VSI to be strongly stabilized even for small, finite
-thermal relaxation times.
+% \subsubsection{$\gamma\neq\Gamma \neq 1$, $N_z^2\geq0$}
+% %gamma = 2.0, Gamma = 1.4, vary bcool 
+% Finally, we introduce buoyancy in a vertically non-isothermal disk by
+% setting $\gamma=2$. This scenario was not considered by  
+% \cite{nelson13}, but is similar in spirit with vertically isothermal 
+% disks evolved with $\gamma>1$, as considered in the previous section. 
+% Fig. \ref{growth_vnoniso2} plots the fundamental VSI growth rates for
+% this case. We observe the same trend as for vertically isothermal
+% disks with $N_z^2\geq 0$ (Fig. \ref{relax_growth_num}), namely growth
+% rates reach zero at finite $\khat$. From Fig.\ref{growth_vnoniso2} we
+% expect the VSI to be strongly stabilized even for small, finite
+% thermal relaxation times.
 
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/growth_vnoniso2}
-  \caption{Same as Fig. \ref{growth_vnoniso1}, but with $\gamma=2$ so that
-    $N_z^2\geq0$. \label{growth_vnoniso2}}
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+%   0cm]{figures/growth_vnoniso2}
+%   \caption{Same as Fig. \ref{growth_vnoniso1}, but with $\gamma=2$ so that
+%     $N_z^2\geq0$. \label{growth_vnoniso2}}
+% \end{figure}
 

### 2015-02-08 appendix.tex
updating appendix

@@
-\section{Limitations of the vertically global shearing box}\label{global_corr}
-Here, we discuss the relation between the linear problem in the
-VGSB framework and the global linear problem.
+\section{Applicability of the vertically global shearing box}\label{global_corr}
+In \S\ref{analytical} we used the vertically global shearing box
+\citep[VGSB,][]{mcnally14} to make analytical progress for vertically
+isothermal disks. Here, we discuss limitations of this approach.
 
- This procedure produces $\phi$ and $z$ momentum equations
-that are equilvalent to those for $y$ and $z$ momenta in the VGSB.
+The main difference between  VGSB framework
 
@@
 \end{align}
-where subscript $0$ denotes evaluation at $r=r_0$. To obtain the
-linear problem in the VGSB, we take the radially-local 
-approximation further and formally set $\hat{g}_c=0$, $r\to x$ and
-$k_r\to k_x$. It is clear that this approximation improves with
-increasing $|k_x|$.
+where subscript $0$ denotes evaluation at $r=r_0$.
+
+%  To obtain the
+% linear problem in the VGSB, we take the radially-local 
+% approximation further and formally set $\hat{g}_c=0$, $r\to x$ and
+% $k_r\to k_x$. It is clear that this approximation improves with
+% increasing $|k_x|$.
+
+\subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
+For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
+simply use the Soilberg-Hoiland critera to assess stability
+(\S\ref{solberg}). Nevertheless, we briefly consider this case in the
+local framework to demonstrate its limitations. For $\beta\to\infty$
+the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
+be written as
+
+\begin{align}
+  0 =\dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
+    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime
+  +\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
+    -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
+          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
+\end{align}
+
+We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
+integrate vertically, assume boundary terms vanish when integrating by
+parts, to obtain
+\begin{align}
+  \hat{\sigma}^2\left(\frac{1}{\gamma} +
+    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
+  v_z|^2 d\zhat 
+  =&  \left(\frac{\gamma-1}{\gamma}\right)
+  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
+  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
+  v_z)^\prime|^2 d\zhat\notag\\
+&+
+  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
+      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
+    2}|\delta v_z|^2 d\zhat
++ \ii\epsilon q \hat{k}
+  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
+  \delta v_z d\zhat.\label{adia_integral}
+\end{align}
+When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
+$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
+expected, a sub-adiabatically stratified disk is stable in the absense
+of vertical shear.
+
+However, in the presence of vertical shear $q\neq0$,
+Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
+$\khat$, which implies the possibility of instability, regardless of
+how much $\gamma$ deviates from unity. This is contradicts the second 
+Solberg-Hoiland criterion (Eq. \ref{solberg2}) and is a result of the
+radially-local approximation (Appendix \ref{global_corr}). To 
+avoid this, we must restrict our attention to short thermal relaxation
+timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
+($\gamma=\Gamma$) when $q\neq0$.
 
@@
 %plot growth rate v.s. bcool
-
-\subsection{Spurious growth of adiabatic perturbations in the local
-  model with vertical shear}\label{analytic_adia} 
-For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
-simply use the Soilberg-Hoiland critera to assess stability
-(\S\ref{solberg}). Nevertheless, we briefly consider this case in the
-local framework to demonstrate its limitations. For $\beta\to\infty$
-the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
-be written as
-
-\begin{align}
-  0 =& \dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
-    \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime\notag\\
-  &+\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right)\right.\notag\\
-  &\phantom{0=}\left.-\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
-          q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
-\end{align}
-
-We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
-integrate vertically, assume boundary terms vanish when integrating by
-parts, to obtain
-\begin{align}
-  &\hat{\sigma}^2\left(\frac{1}{\gamma} +
-    \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
-  v_z|^2 d\zhat \notag\\
-  &=  \left(\frac{\gamma-1}{\gamma}\right)
-  \int_{\zhat_1}^{\zhat_2}\rho|\delta v_z^\prime|^2 d\zhat
-  +\frac{1}{\gamma}\int_{\zhat_1}^{\zhat_2}\frac{1}{\rho}|(\rho\delta
-  v_z)^\prime|^2 d\zhat\notag\\
-  &\phantom{=} +
-  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
-      q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
-    2}|\delta v_z|^2 d\zhat\notag\\
-  &\phantom{=} + \ii\epsilon q \hat{k}
-  \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
-  \delta v_z d\zhat.\label{adia_integral}
-\end{align}
-When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
-$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
-expected, a sub-adiabatically stratified disk is stable in the absense
-of vertical shear.
-
-However, in the presence of vertical shear $q\neq0$,
-Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
-$\khat$, which implies the possibility of instability, regardless of
-how much $\gamma$ deviates from unity. This is contradicts the second 
-Solberg-Hoiland criterion (Eq. \ref{solberg2}) and is a result of the
-radially-local approximation (Appendix \ref{global_corr}). To 
-avoid this, we must restrict our attention to short thermal relaxation
-timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
-($\gamma=\Gamma$) when $q\neq0$.
 

### 2015-02-08 numerical.tex
updating appendix

@@
 Eq. \ref{iso_vsi_cond}. For the present disk parameters
-$\beta_\mathrm{crit}=0.125$. We find good agreement between the
-numerical and analytical growth rates, as well as the critical
-relaxation timescale. Growth rates are insensitive to $\beta$ for
-$\beta\lesssim 0.05$, beyond which growth rates rapidly decrease.
+$\beta_\mathrm{crit}=0.125$.
 
-For this disk model, \citeauthor{nelson13} found a thermal relaxation
-timescale of $\beta\gtrsim 0.6$ stabilized the VSI in non-linear
-hydrodynamic simulations. This is consistent with our numerical and
-analytic results.
+We find good agreement between the numerical and analytical growth
+rates, as well as the critical relaxation timescale. Growth rates are
+insensitive to $\beta$ for $\beta\lesssim 0.05$, beyond which growth
+rates rapidly decrease. For this disk model, \citeauthor{nelson13}
+found a thermal relaxation timescale of $\beta\gtrsim 0.6$ stabilized
+the VSI in non-linear hydrodynamic simulations. This is consistent
+with our results.
 
@@
 oscillatory behavior emerging at the boundaries, where buoyancy first
-becomes important.
+becomes important. Finite thermal relaxation increases $|W|$ near the
+vertical boundaries relative to that about the midplane, but
+perturbations near the disk surface is unlikely important in practice
+because the disk atmosphere contains little mass.
 
@@
 \subsection{Critical thermal relaxation timescale}
+% may need to explain why choose k=10
+% bcrit independent of k for k>>1 
+% for some parameter regimes (e.g. large q, large epsilon) 
+% behavior at larger kx and finite bcool not understood. 
 We test our estimate for the maximum thermal relaxation timescale 
@@
 VSI (Eq. \ref{iso_vsi_cond}). We use the previous disk
-parameters as a fiducial case and vary $q\in[-0.2,-1.2]$,
-$\gamma\in[1.2,2.0]$ and $\epsilon\in[0.02,0.1]$
-separately. The numerically-obtained $\beta_\mathrm{crit}$ is shown in
+parameters as a reference point and vary $q\in[-0.2,-1.2]$,
+$\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$
+separately.
+
+Here we use a wavenumber $\khat=10$.
+
+% , in which case we find $\nu$
+% changes sign as $\beta$ is increased from zero. We find that with
+% $\khat=30$, as used in the previous section, when $\beta$ is not small $\nu$ did not reach zero but instead decreases
+% rapidly after the expected value of $\beta_\mathrm{crit}$.  These
+% modes were found to have large amplitudes near the boundaries.
+
+%we suspect boundary conditions may play a role for
+
+% For large $\khat$ \emph{and} finite
+% thermal relaxation (e.g. $\khat=30$, $\beta\gtrsim 0.2$) we found
+% $\nu$ did not reach zero, but still decreases rapidly after the
+% expected $\beta_\mathrm{crit}$. These modes have large perturbations
+% near the boundaries, which may play a role not accounted for in our
+% analytical discussion.
+
+The numerically-obtained $\beta_\mathrm{crit}$ is shown in
 Fig. \ref{bcrit_compare} in comparison with Eq. \ref{iso_vsi_cond}.
 
-We find Fig. \ref{bcrit_compare} generally agrees with
-Eq. \ref{iso_vsi_cond}.  The match is better for smaller
-$\beta_\mathrm{crit}$. There is a noticeable difference as
-$\gamma\to1$ because the  derivation of $\beta_\mathrm{crit}$ assumed
-$\gamma>1$.
+Our numerical results generally agree with 
+Eq. \ref{iso_vsi_cond}. The agreement improves with decreasing  $|q|$,
+$\epsilon$ and increasing $\gamma$, i.e. weaker instability. 
+There is a noticeable difference as $\gamma\to1$ because the
+derivation of $\beta_\mathrm{crit}$ assumed $\gamma>1$.
 

### 2015-02-09 appendix.tex
major updating

@@
 In \S\ref{analytical} we used the vertically global shearing box
-\citep[VGSB,][]{mcnally14} to make analytical progress for vertically
-isothermal disks. Here, we discuss limitations of this approach.
-
-The main difference between  VGSB framework
-
-Using the notation and disk models in the main text, the
-energy, density and $r$ momentum equations become 
+\citep[VGSB,][]{mcnally14} to make analytical progress. % Here, we discuss limitations of this approach.   
+The main difference between the linear problem in the VGSB 
+and the radially-local approximation
+(Eq. \ref{lin_mass}---\ref{lin_energy}) adopted in our numerical study
+is the absence of  terms associated with the radial disk structure in
+the former framework. These terms appear in the energy, mass and radial momentum
+equations and are reproduced here for convenience,
+% Using the notation and disk models in the main text, the
+% energy, density and $r$ momentum equations become 
 \begin{align}
@@
       z}\right|_0 \delta v_z + \frac{\gamma}{\Gamma} \frac{d\delta
-    v_z}{dz}\right) + \left[\ii k_rc_s^2(r_0,z)
+    v_z}{dz}\right) + \left[\ii k_xc_s^2(r_0,z)
   \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
-      r}\right|_0\right]\delta v_r  +
+      r}\right|_0\right]\delta v_x  +
 \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
 \ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_r + \hat{g}_c
-  \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_r + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
+  \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
       z}\right|_0 \delta v_z + \frac{d\delta
     v_z}{dz}\right),\label{gcorr_terms2} \\
-\ii \sigma \delta v_r & = \ii k_r W  -
+\ii \sigma \delta v_x & = \ii k_x W  -
 \hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
-  r}\right|_0Q - 2\Omega(r_0,z)\delta v_\phi,\label{gcorr_terms3}
+  r}\right|_0Q - 2\Omega(r_0,z)\delta v_y,\label{gcorr_terms3}
 \end{align}
 where subscript $0$ denotes evaluation at $r=r_0$.
+
+In the VGSB, the terms proportional to $\hat{g}_c$ are ignored. For a
+power-law disk, these radial gradients are $O(r^{-1})$, and they
+appear in comparison with terms of $O(k_x)$. The neglected terms 
+therefore have a relative magnitude of $O(\epsilon/\khat)$. This is
+small for thin disks ($\epsilon\ll1$) and/or small radial wavelengths
+($\khat\gg 1$).
 
@@
 \subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
-For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
-simply use the Soilberg-Hoiland critera to assess stability
-(\S\ref{solberg}). Nevertheless, we briefly consider this case in the
-local framework to demonstrate its limitations. For $\beta\to\infty$
-the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
-be written as
-
+An important caveat of the VGSB is that it cannot be used in the
+adiabatic limit. We explain this by setting  $\beta\to\infty$ in
+Eq. \ref{vertiso_gov_nondim}, which can then be written as 
+% For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
+% simply use the Soilberg-Hoiland critera to assess stability
+% (\S\ref{solberg}). Nevertheless, we briefly consider this case in the
+% local framework to demonstrate its limitations. For $\beta\to\infty$
+% the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
+% be written as   
 \begin{align}
@@
 \end{align}
-
 We multiply Eq. \ref{adia_iso3} by $\rho\delta v_z^*$ and
@@
 how much $\gamma$ deviates from unity. This is contradicts the second 
-Solberg-Hoiland criterion (Eq. \ref{solberg2}) and is a result of the
-radially-local approximation (Appendix \ref{global_corr}). To 
-avoid this, we must restrict our attention to short thermal relaxation
-timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
-($\gamma=\Gamma$) when $q\neq0$.
+Solberg-Hoiland criterion (Eq. \ref{solberg2}), which states that
+instability in the adiabatic limit requires the disk to be close to
+neutrally stratified. %also need real freq^2 in adia limit 
+This is because the VGSB retains the destabilizing effect of vertical shear
+($\p_z\Omega$) from the global disk, but ignores the global radial disk
+struture responsible for it. Nevertheless, we demonstrate below that
+this inconsistency is unimportant for the vertical shear instability,
+which occurs for $\beta\ll1$.
+
+% To 
+% avoid this, we must restrict our attention to short thermal relaxation
+% timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
+% ($\gamma=\Gamma$) when $q\neq0$.
 
 \subsection{Effect of global radial gradients}
-We may also reverse the above procedure by reinstating the `global
-correction' terms proportional to $\hat{g}_c$ in the linear VGSB
-equations. Of course, doing so implies the basic state is not uniform
-in $x$, which would be inconsistent with the full VGSB equations in
-equilibrium. Nevertheless, we illustrate in Fig. \ref{gcorr_compare} the effect of
-these global radial gradients on the growth rate of the fundamental
-VSI.
+% We may also reverse the above procedure by reinstating the `global
+% correction' terms proportional to $\hat{g}_c$ in the linear VGSB
+% equations. Of course, doing so implies the basic state is not uniform
+% in $x$, which would be inconsistent with the full VGSB equations in
+% equilibrium. Nevertheless, we illustrate in Fig. \ref{gcorr_compare} the effect of
+% these global radial gradients on the growth rate of the fundamental
+% VSI.
 
-The global correction terms do not change the qualitative dependence
-of the fundamental VSI growth rates on the thermal relaxation
-timescale. The difference is negligble for $\beta\lesssim 1$ (i.e.
-fast thermal relaxation). The critical relaxation timescale for the
-fundamental VSI is unaffected by these global radial radients.
+% The global correction terms do not change the qualitative dependence
+% of the fundamental VSI growth rates on the thermal relaxation
+% timescale. The difference is negligble for $\beta\lesssim 1$ (i.e.
+% fast thermal relaxation). The critical relaxation timescale for the
+% fundamental VSI is unaffected by these global radial radients.
 
-However, in the adiabatic limit $\beta\to\infty$ the linear VGSB
-equations give a positive growth rate as discussed in
-\S\ref{analytic_adia}, which is inconsistent with the second
-Solberg-Hoiland criterion (Eq. \ref{solberg2}). The correct behaviour
-($\nu\to0$ as $\beta\to\infty$) is obtained with global correction.
+% However, in the adiabatic limit $\beta\to\infty$ the linear VGSB
+% equations give a positive growth rate as discussed in
+% \S\ref{analytic_adia}, which is inconsistent with the second
+% Solberg-Hoiland criterion (Eq. \ref{solberg2}). The correct behaviour
+% ($\nu\to0$ as $\beta\to\infty$) is obtained with global correction.
 
-It is interesting to note that there is a thermal relaxation timescale
-that maximizes the mode decay rate. Here, it is $\beta\simeq2$ or
-about $0.3$ orbital periods. This is, in fact, consistent with Fig. 12
-of \cite{nelson13}.
+In  Fig. \ref{gcorr_compare}, we demonstrate the effect of the global radial gradient terms
+proportional to $\hat{g}_c$ in 
+Eq. \ref{gcorr_terms1}---\ref{gcorr_terms3} by calculating the
+fundamental VSI growth rates using three approaches. We compute growth rates from the dispersion
+relation Eq. \ref{relax_disp}, which utilizes the VGSB;  from the
+full linear problem (Eq. \ref{lin_mass}---\ref{lin_energy}) with
+$\hat{g}_c=0$ (i.e. the linear VGSB equations); and from the full
+linear problem with $\hat{g}_c=1$.
+
+All three methods give similar behavior, and growth rates are in close
+agreement for $\beta\lesssim 1$. Differences arise for
+$\beta\gtrsim1$, and as $\beta\to\infty$ the VGSB framework gives a
+(spurious) positive growth rate as expected from the discussion
+above. Inclusion of the global radial gradient terms results in the
+correct behavior  ($\nu\to0$ as $\beta\to\infty$).
 
@@
   0cm]{figures/gcorr_compare} 
-  \caption{Growth rate of
-    the fundamental VSI mode as a function of the thermal
-    relaxatin time $\beta$, computed from the linear VGSB equations used in
-    the main text (dotted) and that including additional global
-    radial gradient terms in  Eq. \ref{gcorr_terms1}---\ref{gcorr_terms3} (solid). 
+  \caption{Growth rate of the fundamental VSI mode as a function of
+    the thermal relaxation time $\beta$. The `Analytical (VGSB)' curve
+    is calculated from the dispersion relation Eq. \ref{relax_disp};
+    the `Numerical (VGSB)' curve is obtained by numerically solving
+    Eq. \ref{lin_mass}---\ref{lin_energy} with $\hat{g}_c=0$
+    (which defines the linear VGSB equations); and the `Numerical
+    (Full)' curve is obtained by numerically solving
+    Eq. \ref{lin_mass}---\ref{lin_energy}  with $\hat{g}_c=1$, which
+    accounts for the radial disk structure.  
+% computed from the linear VGSB equations used in
+%     the main text (dotted) and that including additional global
+%     radial gradient terms in
+%     Eq. \ref{gcorr_terms1}---\ref{gcorr_terms3} (solid). 
     The disk parameters are $\Gamma=1.011$,
     $\gamma=1.4$ and $(p,q,\epsilon)=(-1.5,-1,0.05)$. The perturbation
-    wavenumber is $\khat=20$. 
-    {\bf waiting for
-      high resolution} 
+    wavenumber is $\khat=30$. 
     \label{gcorr_compare}}  
 \end{figure}
+
+Fig. \ref{gcorr_compare} shows that provided we consider $\beta\ll1$,
+then the VGSB framework is an adequate approximation. It is
+interesting to note that there is a thermal relaxation timescale that
+maximizes the mode decay rate. Here, it is $\beta\simeq2$ or about
+$0.3$ orbital periods. This is, in fact, consistent with Fig. 12 of
+\cite{nelson13}.
 

### 2015-02-09 linear_iso.tex
major updating

@@
 considering the effect of finite thermal relaxation. 
-In order to facilitate an analytical discussion, we introduce a number
+In order to facilitate a fully analytic discussion, we introduce a number
 of approximations below that allows a simple governing equation to be
-derived and directly compared with previous works.
+derived and compared with previous works.
 

### 2015-02-09 numerical.tex
major updating

@@
 % choice of kx is O(10), since nelson observe kxH \sim 30
+%not possible to survey all modes/kx/disk parameters
+%focus on effect of thermal relaxation
 
@@
 shown) we find the perturbation is almost entirely concentrated at the
-boundaries. Our radially-local approximation is likely invalid for 
+boundaries. Our radially-local approximation is not valid for 
 these small wavenumber perturbations and we do not consider such modes
@@
 We now consider stably stratified disks ($\gamma > \Gamma$) with
-thermal relaxation $\beta\in[10^{-3},1]$. Here we use a slightly  
+thermal relaxation. Here, we use a slightly  
 different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as 
@@
 Fig. \ref{bcrit_compare1} shows the growth rate $\nu$ of the
-fundamental VSI mode as a function of $\beta$ for $\khat=30$. This is
-approximately the perturbation wavenumber observed in the fiducial
-numerical simulation of \cite{nelson13}. We also plot the growth rates
-obtained by numerically solving the dispersion relation 
-Eq. \ref{relax_disp} and the upper limit to 
-the thermal relaxation timescale $\beta_\mathrm{crit}$ given by
-Eq. \ref{iso_vsi_cond}. For the present disk parameters
-$\beta_\mathrm{crit}=0.125$.
+fundamental VSI as a function of $\beta$ for $\khat\in[1,10^2]$. We
+also plot the critical thermal relaxation timescale
+$\beta_\mathrm{crit}$ given by Eq. \ref{iso_vsi_cond}. For the present
+disk parameters $\beta_\mathrm{crit}=0.125$.
 
-We find good agreement between the numerical and analytical growth
-rates, as well as the critical relaxation timescale. Growth rates are
-insensitive to $\beta$ for $\beta\lesssim 0.05$, beyond which growth
-rates rapidly decrease. For this disk model, \citeauthor{nelson13}
+For $5\lesssim\khat\lesssim30$, we find $\beta_\mathrm{crit}$ provides an
+accurate prediction of the upper limit to the thermal relaxation 
+timescale for  the fundamental mode. For smaller $\khat$,
+$\beta<\beta_\mathrm{crit}$ is a sufficient condition for
+instability. For larger $\khat$, we find $\nu$ does not reach zero but
+rapidly decreases as $\beta$ approaches and increases beyond
+$\beta_\mathrm{crit}$. In the latter case we find the perturbation is
+concentrated at the vertical boundaries, which may play a role not
+accounted by our analytical discussion.
+
+Fig. \ref{bcrit_compare1} shows that 
+$\beta_\mathrm{crit}$ is a good indication for the maximum
+thermal relaxation timescale to permit the fundamental VSI. For
+$\beta\lesssim O(10^{-2})$, growth rates are insensitive to thermal
+relaxation. For this disk model, \citeauthor{nelson13}
 found a thermal relaxation timescale of $\beta\gtrsim 0.6$ stabilized
@@
    \includegraphics[width=\linewidth]{figures/bcrit_compare} 
-   \caption{Growth rate of the fundamental VSI mode with $\khat=30$ as
+   \caption{Growth rate of the fundamental VSI as
      a function of the thermal relaxation timescale $\beta$. The disk
      parameters are $\Gamma=1.011$, $\gamma=1.4$, and 
-     $(p,q,\epsilon)=(-1.5,-1,0.05)$. The numerical (analytical) growth
-     rates are show as solid (dotted) lines, while the vertical line
-     is the maximum thermal timescale $\beta_\mathrm{crit}$ for the
+     $(p,q,\epsilon)=(-1.5,-1,0.05)$. The vertical line
+     is the critical thermal timescale $\beta_\mathrm{crit}$ for the
      fundamental mode according to Eq. \ref{iso_vsi_cond}. 
@@
 
+%approximately the perturbation wavenumber observed in the fiducial
+%numerical simulation of \cite{nelson13}.
+
 We visualize the effect of thermal relaxation in 
@@
     $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ and a dimensionless 
-    thermal relaxation timescale $\beta \equiv t_c\Omega_k = 0.01$
-    (top) and $0.1$ (bottom). The
+    thermal relaxation timescale $\beta = 0.01$
+    (top) and $\beta=0.1$ (bottom). The
     perturbation radial wavenumber is $\khat=30$. The
@@
 % behavior at larger kx and finite bcool not understood. 
-We test our estimate for the maximum thermal relaxation timescale 
-$\beta_\mathrm{crit}$ for the fundamental
-VSI (Eq. \ref{iso_vsi_cond}). We use the previous disk
-parameters as a reference point and vary $q\in[-0.2,-1.2]$,
+We test the dependence of the critical thermal relaxation timescale 
+$\beta_\mathrm{crit}$ for the fundamental VSI on disk parameters
+(Eq. \ref{iso_vsi_cond}).  We use the previous case study as a
+reference point and vary $q\in[-0.2,-1.2]$, 
 $\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$
-separately.
-
-Here we use a wavenumber $\khat=10$.
-
-% , in which case we find $\nu$
-% changes sign as $\beta$ is increased from zero. We find that with
-% $\khat=30$, as used in the previous section, when $\beta$ is not small $\nu$ did not reach zero but instead decreases
-% rapidly after the expected value of $\beta_\mathrm{crit}$.  These
-% modes were found to have large amplitudes near the boundaries.
-
-%we suspect boundary conditions may play a role for
-
-% For large $\khat$ \emph{and} finite
-% thermal relaxation (e.g. $\khat=30$, $\beta\gtrsim 0.2$) we found
-% $\nu$ did not reach zero, but still decreases rapidly after the
-% expected $\beta_\mathrm{crit}$. These modes have large perturbations
-% near the boundaries, which may play a role not accounted for in our
-% analytical discussion.
+separately. Here, we use a wavenumber $\khat=10$, for which we
+find growth rates reach zero for the range of disk parameters
+considered, so that a numerical $\beta_\mathrm{crit}$ can be defined precisely.
 
@@
 Fig. \ref{bcrit_compare} in comparison with Eq. \ref{iso_vsi_cond}.
+Our numerical results generally agree with Eq. \ref{iso_vsi_cond}. The
+agreement improves with decreasing  $|q|$,  $\epsilon$ and increasing
+$\gamma$, i.e. weaker instability, although there is a noticeable difference
+as $\gamma\to1$ because the derivation of $\beta_\mathrm{crit}$
+assumed $\gamma>1$.
 
-Our numerical results generally agree with 
-Eq. \ref{iso_vsi_cond}. The agreement improves with decreasing  $|q|$,
-$\epsilon$ and increasing $\gamma$, i.e. weaker instability. 
-There is a noticeable difference as $\gamma\to1$ because the
-derivation of $\beta_\mathrm{crit}$ assumed $\gamma>1$.
+Fig. \ref{bcrit_compare}, together with the previous section
+(Fig. \ref{bcrit_compare1}), shows that $\beta_\mathrm{crit}$ as given
+by Eq. \ref{iso_vsi_cond} gives a measure of the characteristic
+thermal timescale below which perturbations are effectively isothermal
+so that the fundamental VSI operates.
+
+% From the previous section (Fig. \ref{relax_eigenW_num}) we
+% expect that for other wavenumbers,  still gives a characteristic
+% thermal timescale below which the perturbations are effectively
+% isothermal.
 
@@
 
-% \subsubsection{$\gamma=\Gamma \neq 1$, $N_z^2\equiv 0$}
-% Consider first the case $\gamma=1.4$ so that there is no vertical
-% entropy gradient. We plot in Fig. the growth rates for the
-% fundamental VSI for $\beta\in[10^{-2},10^{2}]$. We find growth rates
-% do not reach zero, implying instability regardless of the thermal
-% relaxation timescale if $\gamma=\Gamma$. This result is consistent
-% with \cite{nelson13}.  Note that for $\beta\to\infty$ (adiabatic
-% perturbations) and $\gamma=\Gamma$, the disk can be unstable according
-% to the Solberg-Hoiland criterion  Eq. \ref{solberg2}.
+Fig. \ref{gcorr_compare_vnoniso} plots the fundamental VSI growth
+rates as a function of $\beta$ for $\gamma\in[1.4,2.5]$. For
+comparison with \cite{nelson13} we consider 
+$\beta\in[10^{-2},10^2]$. In agreement with \citeauthor{nelson13}, in
+the neutrally-stratified case $\gamma=\Gamma$ the disk is unstable
+even for $\beta\gg 1$. (Note that in the adiabatic limit
+$\beta\to\infty$ this disk can be unstable according to the
+Solberg-Hoiland criterion, Eq. \ref{solberg2}.)
 
-% Growth rates for $\beta = 10^{-2}$ and
-% $\beta=10^{2}$ behave similarly. This is not surprising because
-% the linearized energy equation (Eq. \ref{lin_energy}) implies that 
-% \begin{align}
-%   W = \frac{\left(\Gamma^{-1} -
-%       \ii\hat{\sigma}\beta\right)}{\left(1-\ii\hat{\sigma}\beta\right)}
-%   Q \text{ for } \gamma\equiv \Gamma.   
-% \end{align}
-% Thus, $W \to Q/\Gamma$ as $\beta\to0$ and $W\to Q$ as
-% $\beta\to\infty$. Since $\Gamma=O(1)$, these two limits are expected
-% to produce qualitatively similar results. Interestingly,
-% Fig. \ref{growth_vnoniso1} show that growth rates 
-% maximize for $\beta=O(1)$. However, all growth rates are similar to
-% order-of-magnitude for $0.1\lesssim\khat\lesssim 100$.
+For $\gamma>\Gamma$, i.e. stably stratified disks,
+Fig. \ref{gcorr_compare_vnoniso} shows that introducing finite thermal
+relaxation rapidly stabilizes the fundamental VSI mode. This is
+similar to that observed for nearly vertically isothermal disks
+(Fig. \ref{bcrit_compare1}). For $\gamma=1.7,\,2.0,\,2.5$, growth
+rates reach zero for  $\beta\simeq0.17,\,0.083,\,0.045$, respectively.
+Interestingly, these values are equal to
+$\epsilon|q|/(\gamma-\Gamma)$. This suggests that the critial thermal
+relaxation timescale for vertically non-isothermal disks can also be
+estimated by Eq. \ref{iso_vsi_cond} but with $\gamma-1$ replaced by
+$\gamma-\Gamma$.
 
-% \begin{figure}
-%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-%   0cm]{figures/growth_vnoniso1} 
-%   \caption{Numerically-determined growth rates of the fundamental VSI
-%     mode in neutrally-buoyant ($N_z^2\equiv0$), vertically non-isothermal disks
-%     ($\Gamma=\gamma=1.4$), with $(p,q,\epsilon)=(0,-1,0.05)$, for
-%     a range of thermal relaxation timescales 
-%     $\beta$. \label{growth_vnoniso1}}     
-% \end{figure}
-
-% \subsubsection{$\gamma\neq\Gamma \neq 1$, $N_z^2\geq0$}
-% %gamma = 2.0, Gamma = 1.4, vary bcool 
-% Finally, we introduce buoyancy in a vertically non-isothermal disk by
-% setting $\gamma=2$. This scenario was not considered by  
-% \cite{nelson13}, but is similar in spirit with vertically isothermal 
-% disks evolved with $\gamma>1$, as considered in the previous section. 
-% Fig. \ref{growth_vnoniso2} plots the fundamental VSI growth rates for
-% this case. We observe the same trend as for vertically isothermal
-% disks with $N_z^2\geq 0$ (Fig. \ref{relax_growth_num}), namely growth
-% rates reach zero at finite $\khat$. From Fig.\ref{growth_vnoniso2} we
-% expect the VSI to be strongly stabilized even for small, finite
-% thermal relaxation times.
-
-% \begin{figure}
-%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-%   0cm]{figures/growth_vnoniso2}
-%   \caption{Same as Fig. \ref{growth_vnoniso1}, but with $\gamma=2$ so that
-%     $N_z^2\geq0$. \label{growth_vnoniso2}}
-% \end{figure}
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/gcorr_compare_vnoniso}
+  \caption{Growth rate of the fundamental VSI mode as a function of
+    the thermal relaxation timescale $\beta$, in vertically
+    non-isothermal disks with $\Gamma=1.4$ and
+    $\gamma\in[1.4,2.5]$. The disk is neutrally
+    stratified for $\gamma=1.4$ and stably stratified for
+    $\gamma>1.4$. Other disk parameters are
+    $(p,q,\epsilon)=(0,-1,0.05)$ and the perturbation wavenumber is
+    $\khat=30$.  
+    \label{gcorr_compare_vnoniso}}
+\end{figure}
 

### 2015-02-09 paper.tex
major updating

@@
   vertical entropy gradient. Our linear calculations are consistent
-  with previous hydrodynamic simulations. These simulations indicate that the VSI 
+  with previous hydrodynamic simulations, which indicates the VSI 
   requires a short thermal relaxation timscale. For vertically  
@@
   find the MMSN is unstable to the VSI at tens of AU with a
-  characteristic growth times of tens of local orbital periods. 
+  characteristic growth times of tens of local orbital periods. We 
+  suggest the VSI is a viable mechanism for angular momentum
+  transport in the outer parts of protoplanetary disks. 
 \end{abstract}

### 2015-02-09 setup.tex
major updating

@@
 \p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$.   
-Eq. \ref{lin_mass}---\ref{lin_vz} will form the basis of our numerical
+Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
 study of the vertical shear instability, but for analytical discussion

### 2015-02-09 summary.tex
major updating

@@
 explained earlier in the local limit by \cite{urpin98} and
-\cite{urpin03}, simulated by \cite{arlt04} and more  recently by
+\cite{urpin03}; and simulated by \cite{arlt04} and more recently by
 \cite{nelson13}.
@@
 We generalized previous vertically-global linear stability calculations
-of the VSI \citep{nelson13,mcnally14}, which assumed
-isothermal perturbations, to include an energy equation. Our effort
-may also be regarded as an extension of \cite{urpin03}, which 
-included an energy equation, by relaxing the Boussinesq and  
-vertically-local approximations adopted in that analysis.
+of the VSI \citep{nelson13,mcnally14}, which assumed isothermal
+perturbations, to include an energy equation. Our effort may also be
+regarded as an extension of \cite{urpin03}, which included an energy
+equation, by relaxing the Boussinesq and vertically-local
+approximations adopted in that analysis.
 
-In accordance with the above studies, which show that the VSI   
-applies to perturbations with small radial lengthscales, we retained
-the radially-local approximation by using the `vertically-global
-shearing box' (VGSB) formalism developed by \cite{mcnally14}, who 
-showed the linear VSI studied by \cite{nelson13} in a locally
-isothermal disk can be captured by the VGSB.
+We added a source term in the energy equation which relaxes the disk
+temperature to its initial value on a timescale $t_c=\beta\Omega_k^{-1}$. This
+form of thermal relaxation was also employed in the global numerical
+simulations of \cite{nelson13}, and allows the thermodynamic response
+of the disk to be controlled.
 
+In accordance with the above studies, which show that the VSI applies
+to perturbations with small radial lengthscales, we applied the 
+radially-local approximation in the linearized fluid equations. For
+analytic discussion only, we further simplied the linear problem by
+ignoring the background radial disk structure. This is equivalent to adopting
+the `vertically-global shearing box' (VGSB) formalism developed by
+\cite{mcnally14}, who showed the linear VSI studied by
+\cite{nelson13} in a locally isothermal disk can be captured by the
+VGSB. We find the VGSB framework is appropriate for $\beta\lesssim1$. 
+This is convenient because previous works show that rapid thermal
+relaxation ($\beta\ll1$) is generally required for the VSI. However,
+the VGSB framework is not valid in the adiabatic limit.
+
+%OK framework for the regime of interest 
 % The presence of a
@@
 
-We added an energy source term in the VGSB equations which relaxes the disk
-temperature to its initial value on a timescale
-$t_c=\beta\Omega_k^{-1}$. This form of thermal relaxation was also 
-employed in the global numerical simulations of \cite{nelson13}, and allows
-the thermodynamic response of the disk to be controlled.
-
 We mostly focused on vertically isothermal disks ($\Gamma=1$) because
@@
 extension of the linear models considered in \cite{lubow93} by
-adding vertical shear, but in the limit of low-frequency
+adding vertical shear and (rapid) thermal relaxation, but in the limit of low-frequency
 perturbations in nearly-Keplerian disks. These approximations were
@@
 
-For isothermal perturbations ($\gamma=1$ or $\beta=0$) we explicitly
+For isothermal perturbations ($\beta=0$) we explicitly
 demonstrated instability  associated with vertical shear, and that the 
@@
 (i.e. anti-symmetric about $z=0$), which corresponds to a vertical
-velocity perturbation  $\delta v_z\sim\mathrm{constant}$ near $z=0$.
+velocity perturbation  $\delta v_z\sim\mathrm{constant}$ near
+$z=0$.
 
@@
 complicated and requires a numerical treatment. However, for the 
-fundamental VSI some simplification is  possible. In particular, we quantified the critical thermal relaxation
-timescale $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$, below
-which small-scale perturbations ($\khat\gg 1$) may be regarded as isothermal
-and the fundamental VSI operates. We believe this explicit expression for
+fundamental VSI some simplification is  possible. In particular, we
+quantified the critical thermal relaxation timescale
+$\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$, below which small-scale
+perturbations may be regarded as isothermal and the fundamental VSI
+operates.  We expect $\beta<\beta_\mathrm{crit}$ to be a sufficient
+condition for the VSI to operate. We believe this explicit expression for
 $\beta_\mathrm{crit}$ is the most useful result from our
@@
 We also solved the linear stability problem numerically using a
-pseudo-spectral approach. Numerical solutions are mostly consistent
-with our analytical expectations, except for perturbations with small
-radial wavenumber ($\khat\to0$) where the low-frequency approximation
-is inappropriate. The local model is also not valid for such modes.
+pseudo-spectral approach. For numerical calculations we accounted for
+the background radial disk structure for consistency. We focused
+on the fundamental mode to compare with our analytical
+discussion, and because \cite{nelson13} found that this mode 
+eventually dominate their simualtions. Our numerical solutions are
+mostly consistent with our analytic expectations, and we verified the
+dependence of $\beta_\mathrm{crit}$ on disk parameters.
+
+% as given
+% by Eq. \ref{iso_vsi_cond}, provides a good measure of the required
+% th 
+%  except for perturbations with small radial
+% wavenumber ($\khat\to0$) where the low-frequency approximation is
+% inappropriate. The radially local model is also not valid for such modes.
 
@@
 thermal relaxation. When a stable stratification is introduced, the
-rapid reduction in VSI growth rates by a finite thermal relaxation timescale
-is qualitatively similar to that observed for vertically isothermal
-disks.
+rapid reduction in VSI growth rates by a finite thermal relaxation
+timescale is qualitatively similar to that observed for vertically
+isothermal disks. Based on numerical results, we conjecture that the
+critical thermal relaxation timescale is given by
+$\beta_\mathrm{crit}=\epsilon|q|/(\gamma-\Gamma)$ in verticlaly
+non-isothermal disks.
 
@@
 shorten the typical growth times for the VSI. This is also attributed
-to a larger value of $\beta_\mathrm{crit}$, since increasing $T$ also
+to a larger value of $\beta_\mathrm{crit}\propto\epsilon$, since increasing $T$ also
 increases the aspect-ratio $\epsilon$.
@@
 
-\subsection{Future extensions} 
+\subsection{Caveats and outlooks} 
 We highlight below some improvements to the present  
@@
 
-\emph{Geometry.} Extrapolating our results from the shearing box to a global disk  
-has an inconsistency. Namely, the equilibrium temperature in the
-box only depends on height (if at all). However, it is precisely the
-radial temperature (or entropy) gradient in the global disk that
-gives rise to vertical shear. We do not expect this to be a serious issue for 
-perturbations with small radial wavelengths, 
-%explicitly checked by including global terms 
-but will need to be resolved by adopting global disk models  
-if one wishes to generalize the VSI to large-scale perturbations.
+\emph{Geometry.} An important caveat
+
+% Extrapolating our results from the shearing box to a global disk  
+% has an inconsistency. Namely, the equilibrium temperature in the
+% box only depends on height (if at all). However, it is precisely the
+% radial temperature (or entropy) gradient in the global disk that
+% gives rise to vertical shear. We do not expect this to be a serious issue for 
+% perturbations with small radial wavelengths, 
+% but will need to be resolved by adopting global disk models  
+% if one wishes to generalize the VSI to large-scale perturbations.
 

### 2015-02-09 application.tex
major updating
nearly done with major update

@@
 By comparing this expression with the linearized version of the energy
-equation used in our calculations (Eq. \ref{energy_eq}), we identify 
+equation used in our calculations (Eq. \ref{lin_energy}), we identify 
 \begin{align}\label{tc_newton_cool} 
@@
 We recall $\beta < \beta_\mathrm{crit}$ is neccessary for the
-fundamental VSI when $\khat\gg 1$. For $\khat\to0$, this
-becomes a sufficient condition (see Fig. \ref{relax_bcrit} and
-\S\ref{iso_vsi_beta_crit}).
+fundamental VSI when $\khat\gg 1$. %For $\khat\to0$, this
+%becomes a sufficient condition (see Fig. \ref{relax_bcrit} and
+%\S\ref{iso_vsi_beta_crit}).
 
@@
 \end{align*} 
-To estimate a thermal relaxation timescale, we set $\khat=10$ 
+To estimate a thermal relaxation timescale, we set $\khat=30$ 
 in Eq. \ref{beta_mmsn} to obtain 
@@
 
-Using the MMSN disk parameters at $r=50\mathrm{AU}$ and the
-scale-dependent thermal relaxation $\beta(\khat)$ given by
-Eq. \ref{beta_mmsn} in our linear code, we calculate the fundamental
-VSI growth rates in Fig. \ref{mmsn_growth}. Growth rates
-are weakly depedent on $\khat$ until $\khat\gtrsim50$ beyond which the
-mode is rapidly stabilized. A representative growth rate is 
+In Fig. \ref{mmsn_growth}, we calculate the fundamental
+VSI growth rates using the MMSN disk parameters at $r=50\mathrm{AU}$
+and the scale-dependent thermal relaxation $\beta(\khat)$ given by
+Eq. \ref{beta_mmsn} in our linear code.  In this case we found a 
+larger vertical domain with $\zmax=7\epsilon r$ was needed to capture
+modes with $\khat\gtrsim60$, which have perturbations
+concentrated near the vertical boundaries. However, modes with
+$\khat\lesssim 60$ were found to be insensitive to boundary
+conditions.
+
+The most unstable mode occurs at $\khat\simeq30$ with growth rate 
 \begin{align*}
-  \nu =0.059\epsilon\Omega_k = 4.0\times10^{-3}\Omega_k 
+  \nu =0.058\epsilon\Omega_k = 3.9\times10^{-3}\Omega_k.  
 \end{align*}
-for $\khat=10$. This corresponds to a characteristic growth time $\sim
-40$ orbits.
+This corresponds to a characteristic growth time $\sim
+40$ orbits. However, growth rates are weakly depedent on $\khat$ for
+$10\lesssim \khat 60$ so this may be taken as a representative growth
+rate of the fundamental VSI in the MMSN at $r=50\mathrm{AU}$. The
+eigenfunction for $\khat=30$ is shown in Fig. \ref{mmsn_eigenW}.
 
@@
 
-Fig. \ref{mmsn_eigenW} show several eigenfunctions for the modes in
-Fig. \ref{mmsn_growth}. As $\khat$ is increased, the disturbance
-become more concentrated near the vertical boundaries. Thus, modes
-with very large $\khat$ may not be important in practice because they
-predominantly affect the disk atmosphere, which contains little mass.
+% Fig. \ref{mmsn_eigenW} show several eigenfunctions for the modes in
+% Fig. \ref{mmsn_growth}. As $\khat$ is increased, the disturbance
+% become more concentrated near the vertical boundaries. Thus, modes
+% with very large $\khat$ may not be important in practice because they
+% predominantly affect the disk atmosphere, which contains little mass.
 
 \begin{figure}
-   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvectorW_mmsnkx1.ps}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvectorW_mmsnkx10.ps} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/eigenvectorW_mmsnkx50.ps} 
-  \caption{Fundamental VSI modes in the MMSN at
-    $r=50\mathrm{AU}$. The eigenfunction $W$ for $\khat=1$ (top),
-    $\khat=10$ 
-    (middle) and $\khat=50$ (bottom) are shown. The real (imaginary)
+   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
+  0cm]{figures/eigenvectorW_mmsnkx30.ps}
+  % \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  % 0cm]{figures/eigenvectorW_mmsnkx10.ps} 
+  % \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  % 0cm]{figures/eigenvectorW_mmsnkx50.ps} 
+  \caption{Fundamental VSI mode in the MMSN at 
+    $r=50\mathrm{AU}$. The eigenfunction $W$ for $\khat=30$ is shown. The real (imaginary)
     part of $W$ is plotted as the solid (dotted)
@@
 $\hat{\rho}=\rho/\rho(z=0)$.  We have have repeated the example
-calculation above with $\beta\to\tilde{\beta}$. We find growth rates
-were reduced, but at most by $\sim 15\%$ at $\khat=1$. Eigenfunctions
-were similar.
+calculation above with $\beta\to\tilde{\beta}$. 
+%We find growth rates
+%were reduced, but at most by $\sim 15\%$ at $\khat=1$. Eigenfunctions
+%were similar.
 

### 2015-02-09 application.tex
paper updates, code updates

@@
   \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} -
-  \frac{4P\kappa_d\sigma_s}{Tc_v}\left(T^4 - T_{t=0}^4\right). 
+  \frac{4P\kappa_d\sigma_s}{Tc_v}\left(T^4 - T_{t=0}^4\right), 
 \end{align}
-When linearized, the last term on the RHS becomes
+where $c_v$ is the heat capacity at constant volume. When linearized, the last term on the RHS becomes
 \begin{align}
@@
 By comparing this expression with the linearized version of the energy
-equation used in our calculations (Eq. \ref{lin_energy}), we identify 
+equation used in our calculations (Eq. \ref{full_energy1}), we identify 
 \begin{align}\label{tc_newton_cool} 
@@
 as the thermal relaxation timescale for perturbation lengthscales
-$L\ll l_\mathrm{rad}$.
+$l\ll l_\mathrm{rad}$.
 
@@
 This corresponds to a characteristic growth time $\sim
-40$ orbits. However, growth rates are weakly depedent on $\khat$ for
-$10\lesssim \khat 60$ so this may be taken as a representative growth
-rate of the fundamental VSI in the MMSN at $r=50\mathrm{AU}$. The
+40$ orbits. However, since growth rates are weakly depedent on $\khat$ for
+$10\lesssim \khat \lesssim 60$, this may be taken as a representative growth
+time of the fundamental VSI in the MMSN at $r=50\mathrm{AU}$. The
 eigenfunction for $\khat=30$ is shown in Fig. \ref{mmsn_eigenW}.

### 2015-02-09 linear_iso.tex
paper updates, code updates

@@
 The linear problem in the VGSB framework can be derived from 
-Eq. \ref{lin_mass}---\ref{lin_vz} by setting $\hat{g}_c=0$, since 
+Eq. \ref{lin_mass}---\ref{lin_energy} by setting $\hat{g}_c=0$, since 
 the VGSB, like the standard shearing box, is homogeneous in the radial 

### 2015-02-09 setup.tex
paper updates, code updates

@@
 to parametrize the magnitude of the sound-speed, where
-$v_k=r\Omega_k=\sqrt{GM_*/r}$ is the Keplerian azimuthal velocity and
-$G$ is the gravitational constant, and $H_\mathrm{iso}$ is a
-characteristic scale-height.  Thus, $\epsilon$ is also the
-characteristic disk aspect-ratio.
+$v_k=r\Omega_k$ is the Keplerian azimuthal velocity and
+$H_\mathrm{iso}$ is a characteristic scale-height.  Thus, $\epsilon$
+is also the characteristic disk aspect-ratio.
 
@@
   &+ \frac{1}{t_c}\left(\delta P -  
-    \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,
+    \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
 \end{align}

### 2015-02-09 codes/cool_full/bloop_gterms/vsi.f90

@@ -2,8 +2,8 @@ module global
   character*6 :: vbc
   logical :: gterms, eigen_trial, eigenv_out, eigen_refine
   integer :: nz, nzeff, bignz, nkx, nb, ntrials, output_freq
-  real*8, parameter :: pi = 2d0*acos(0d0), mu=2.33d0
-  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls, dist 
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls
   real*8 :: kxmin, kxmax, dlogkx, bmin, bmax, db, ghat
   real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:), gcorr(:)
   real*8, allocatable :: kaxis(:), baxis(:)
@@ -30,7 +30,7 @@ program vsi
   complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
        vr(:,:), rhs(:,:), wtrial(:), wtrial_tot(:)
   real*8, external :: dlogrho_dz, domega2_dz, omega2_z, logrho_z, kappa2_z, d2logrho_dz2, csq_z, gcorr_z 
-  namelist /params/ smallq, smallp, dist, gmma, bgmma, zmax, vbc, gterms
+  namelist /params/ smallq, smallp, eps, gmma, bgmma, zmax, vbc, gterms
   namelist /loop/ kxmin, kxmax, nkx, bmin, bmax, nb, eigen_trial, eigenv_out, eigen_refine
   namelist /grid/ nz
   
@@ -40,8 +40,6 @@ program vsi
   read(7, nml=loop)
   read(7, nml=grid)
   close(7)
-
-  eps = 3.36d-2*dist**(2d0/7d0)/sqrt(mu)
 
   if(mod(nz,2).eq.0) then
      print*, 'Nz needs to be odd but Nz=', nz
@@ -62,7 +60,7 @@ program vsi
 
   smalls = smallq + smallp*(1d0 - bgmma) 
 
-  print*, 'smallh, smalls, eps =', smallh, smalls 
+  print*, 'smallh, smalls=', smallh, smalls 
 
[…]

### 2015-02-09 codes/cool_full/bloop_mmsn3D/vsi.f90

@@ -6,7 +6,7 @@ module global
   real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls, dist 
   real*8 :: kxmin, kxmax, dlogkx, bmin, bmax, db, ghat
   real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:), gcorr(:)
-  real*8, allocatable :: kaxis(:), baxis(:)
+  real*8, allocatable :: kaxis(:), baxis(:), bcool_arr(:)
   complex*16, parameter :: ii = (0d0,1d0)
   complex*16, allocatable :: bigW(:), dbigW(:), bigQ(:), vx(:), vy(:), vz(:), dvz(:) 
 end module global
@@ -101,6 +101,7 @@ program vsi
   allocate(kappa2(nz))
   allocate(csq(nz))
   allocate(gcorr(nz))
+  allocate(bcool_arr(nz))
 
   allocate(bigQ(nz))
   allocate(bigW(nz))
@@ -250,13 +251,15 @@ program vsi
      do k=1, nkx !loop over kx values
         kx = kaxis(k)*smallh !convert input into code units 
         
-        bcool= (4.4d5/(mu*(gmma-1d0)))*dist**(-57./14.)
-        bcool=bcool*(8.3d-9*dist**(33./7.) + 1d0/kaxis(k)**2d0)
+     
+        bcool_arr(:)=(4.4d5/(mu*(gmma-1d0)))*dist**(-57./14.)
+        bcool_arr(:)=bcool_arr(:)*(8.3d-9*dist**(33./7.)+2d0*pi*exp(logrho(:))**2d0/kaxis(k)**2d0)
+
 
         !set up sub-matrices for linear operators
         do i=1, nzeff
-           L1(i,:)  = T(i,:)/bcool
-           L2(i,:)  =-T(i,:)/bcool/bgmma 
+           L1(i,:)  = T(i,:)/bcool_arr(i)
+           L2(i,:)  =-T(i,:)/bcool_arr(i)/bgmma 
            L2d5(i,:)= ii*(gmma/bgmma)*csq(i)*kx/smallh*T(i,:)  + ghat*gcorr(i)*T(i,:)
[…]

### 2015-02-09 codes/cool_full/vsi_gterms.f90

@@ -1,12 +1,10 @@
 module global
   character*6 :: vbc
-  logical :: gterms, eigen_trial, eigenv_out, eigen_refine
-  integer :: nz, nzeff, bignz, nkx, nb, ntrials, output_freq
-  real*8, parameter :: pi = 2d0*acos(0d0), mu=2.33d0
-  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls, dist 
-  real*8 :: kxmin, kxmax, dlogkx, bmin, bmax, db, ghat
+  logical :: gterms
+  integer :: nz, nzeff, bignz
+  real*8, parameter :: pi = 2d0*acos(0d0)
+  real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls, ghat  
   real*8, allocatable :: zaxis(:), logrho(:), dlogrho(:), d2logrho(:), omega2(:), domega2(:), kappa2(:), csq(:), freq(:), growth(:), gcorr(:)
-  real*8, allocatable :: kaxis(:), baxis(:)
   complex*16, parameter :: ii = (0d0,1d0)
   complex*16, allocatable :: bigW(:), dbigW(:), bigQ(:), vx(:), vy(:), vz(:), dvz(:) 
 end module global
@@ -16,32 +14,26 @@ program vsi
   implicit none
   character*1, parameter :: JOBVL = 'N', JOBVR = 'V'
   integer, allocatable :: ipiv(:), eigen_test(:)
-  integer :: i,ip1, j,k, n, loc(1), lmax, nzmid, lwork, info
+  integer :: i,ip1, j,k, loc(1), lmax, nzmid, lwork, info
   real*8 :: T_l, dT_l, d2T_l, T_lp2, dT_lp2, d2T_lp2, m, zbar,lmode
-  real*8 :: w_re, w_im, dummy
   real*8, allocatable :: rwork(:)
-  complex*16 :: wold
   complex*16, allocatable :: T(:,:), Tp(:,:), Tpp(:,:)
-  complex*16, allocatable :: L1(:,:), L2(:,:), L2d5(:,:), L3(:,:), L4(:,:), &
-       L5(:,:), L6(:,:), L7(:,:), L8(:,:), &
+  complex*16, allocatable :: L1(:,:), L2(:,:),L2d5(:,:) ,L3(:,:), L4(:,:), &
[…]

### 2015-02-10 paper.tex
writing introduction

--- 2015-01-31 paper.tex
@@
 accretion disks. For example, there may be regions in protoplanetary  
-disks --- called dead zones ---  which do not support the MRI 
-\citep{gammie96,landry13}. In fact, recent state-of-the-art
+disks (PPDs) which do not support the MRI 
+\citep{gammie96,landry13}. In fact, recent  numerical simulations
+suggest portions of PPDs remain laminar \citep[e.g.][and references
+therein]{lesur14,bai15,gressel15}. In this respect, other mechanisms 
+to angular momentum transport in PPDs remain of interest
+\citep{armitage10}.
 
-%in this respect purely hydrodynamical turbulence should be
+%recent work on vsi 
+One possibiliy is purely hydrodynamic turbulence through the vertical
+shear instability (VSI). The VSI is an axisymmetric linear instability
+in three-dimensional (3D) disks in which the rotation frequency
+$\Omega$ depends on the height from the disk midplane. This
+height-dependence, or vertical shear, enables perturbations with
+large vertical lengthscales (relative to radial lengthscales) to
+become unstable. The VSI is the disk analog of the
+Goldreich-Schubert-Fricke instability
+\citep[GSF,]{goldreich67,fricke68} in differentially rotating stars.   
+\cite{umurhan13} gives a simple explanation for this 
+instability: an exchange of fluid parcels while conserving their
+angular momenta individually, may result in a new configuration of
+lower energy, if the exchange occurs over a sufficiently large
+vertical displacement.
 
-%stuff about vsi
+The VSI has
 
-%recent calculations (nelson, mcnally)
+%history
+
+recently been demonstrated through 
+hydrodynamical simulations \citep{nelson13}
+
+%and kley simulations
+
+%vsi background
+
+%questions for this work
 

### 2015-02-10 summary.tex
writing introduction

@@
 critical thermal relaxation timescale is given by
-$\beta_\mathrm{crit}=\epsilon|q|/(\gamma-\Gamma)$ in verticlaly
+$\beta_\mathrm{crit}=\epsilon|q|/(\gamma-\Gamma)$ in vertically 
 non-isothermal disks.
@@
 We highlight below some improvements to the present  
-linear models that should be pursued:
+linear models in future works:
 
@@
 disk, additional effects may become important. For example, in the
-outer parts of a protoplanetary disk, where we estimate the VSI may
-operate, disk self-gravity (SG) may play a role. Specifically, the vertical
+outer parts of a protoplanetary disk, where we estimate the VSI 
+operates, disk self-gravity (SG) may play a role. Specifically, the vertical
 shear profile depends on the vertical density stratification
-(Eq. \ref{vertical_shear}), which wil be enhanced by vertical SG,
-which will enhance the VSI. Inclusion of the Poisson equation should
+(Eq. \ref{vertical_shear}) which can be enhanced by vertical SG. This
+may promote the VSI. Inclusion of the Poisson equation should
 therefore be considered.
@@
 
-\emph{Geometry.} An important caveat
+\emph{Radial dependence.} We have adopted the radially-local approximation 
+throughout this paper. This is based on previous works, including
+nonlinear simulations, which show that the VSI is relevant for
+perturbations with small radial lengthscales. Although this simplifies
+the linear problem to ordinary differential equations, this approach
+introduces the wavenumber $k_x$ which is a free parameter. The effect
+of radial boundary conditions cannot be explored. Future
+generalizations of the linear problem should retain the radial
+dependence properly and solve the two-dimensional $(r,z)$ partial 
+differential equation.
 

### 2015-02-10 paper.tex
updating intro

@@
 suggest portions of PPDs remain laminar \citep[e.g.][and references
-therein]{lesur14,bai15,gressel15}. In this respect, other mechanisms 
+therein]{lesur14,bai15,gressel15}. In this respect, alternative routes 
 to angular momentum transport in PPDs remain of interest
@@
 in three-dimensional (3D) disks in which the rotation frequency
-$\Omega$ depends on the height from the disk midplane. This
-height-dependence, or vertical shear, enables perturbations with
-large vertical lengthscales (relative to radial lengthscales) to
-become unstable. The VSI is the disk analog of the
-Goldreich-Schubert-Fricke instability
+$\Omega$ depends on the height from the disk midplane. The VSI is the
+disk analog of the Goldreich-Schubert-Fricke instability
 \citep[GSF,]{goldreich67,fricke68} in differentially rotating stars.   
+The height-dependence of $\Omega$, or vertical shear, enables
+perturbations with large vertical lengthscales (relative to radial
+lengthscales) to become unstable.    
 \cite{umurhan13} gives a simple explanation for this 
@@
 
-The VSI has
-
 %history
 
-recently been demonstrated through 
-hydrodynamical simulations \citep{nelson13}
+The VSI has been studied analytically in local approximation
+\citep{urpin98,urpin03}, where a Fourier analysis was performed in 
+both the radial and vertical directions. Simulations of the  
+VSI have been performed by \cite{arlt04}, and more recently by 
+\cite{nelson13}. In these simulations the authors indeed observe the
+growth of vertically-global but radially-localized perturbations when
+the equilibrium rotation varied with height.
 
-%and kley simulations
-
-%vsi background
-
-%questions for this work
+An important finding from these works is that the VSI requires a short
+thermal timescale: perturbations need to be close to
+isothermal, unless the disk is neutrally stratified in the vertical
+direction. This is because vertical motions are stabilized by
+buoyancy if the disk is sub-adiabatically stratified, as required for
+convective stability. When the thermal timescale $t_c$ is parametrized
+in units of the orbital period $P$, \cite{nelson13} find that
+$t_c\lesssim 0.01P$ was needed to observe instability. That is, the
+temperature profile of the disk essentially remains unchanged during
+the disk evolution. This has been confirmed by 
+\cite{stoll14}, who find the 
+VSI is only sustained in radiation hydrodynamic simulations 
+with the inclusion of heating from stellar irradiation.
 
@@
 %can be used as a first estimate
+
+From these works it is clear that the VSI is sensitive 
+to the thermodynamic response of the disk except in the special case of neutrally
+stratified disks. It is therefore important to understand this dependence in order to evaluate
+the viability of the VSI in realistic disks. However,
 

### 2015-02-10 appendix.tex
nearly done with major update
paper updates, code updates
added new appendix - repeat gordon's notes with iso eos

@@
 ($\p_z\Omega$) from the global disk, but ignores the global radial disk
-struture responsible for it. Nevertheless, we demonstrate below that
+structure responsible for it. Nevertheless, we demonstrate below that
 this inconsistency is unimportant for the vertical shear instability,
@@
 
-\section{Full governing equation for vertically isothermal disks}\label{adia_improve}
+\section{Full governing equation for vertically isothermal disks in
+  the VGSB}\label{adia_improve}
 % \section{Improving the nearly-Keplerian approximation for adiabatic
@@
 neglecting these terms has no qualitative effect on the discussion in 
-\S\ref{analytic_adia} and \S\ref{analytic_relax}.  This issue does not
+\S\ref{analytic_relax}.  This issue does not
 arise for isothermal perturbations discussed in \S\ref{iso_discuss}
@@
 
+\section{Eigenfrequency in locally
+  isothermal disks in global cylindrical geometry}
+Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
+all times, so in the perturbed state $\delta P =
+c_s^2\delta\rho$, with $c_s(r)$ begin a prescribed function. 
+The linearized fluid equations in cylindrical
+geometry is then 
+\begin{align}
+ \ii\sigma \delta v_r + 2\Omega\delta v_\phi &= c_s^2\frac{\p}{\p r}\left(\frac{\delta\rho}{\rho}\right),\\
+ \ii\sigma \delta v_\phi &= \frac{1}{r}\delta\bm{v}\cdot\nabla j,\\
+ \ii\sigma \delta v_z &= c_s^2\frac{\p}{\p
+   z}\left(\frac{\delta\rho}{\rho}\right),\\
+ \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right).
+\end{align}
+We can eliminate $\delta v_\phi$ from these equations and obtain
+\begin{align}
+  -\sigma^2\rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) +
+  \kappa^2\rho|\delta v_r|^2 + \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
+  \delta v_r^*\delta v_z =
+  c_s^2\rho\delta\bm{v}^*\cdot\nabla\left(\frac{\ii\sigma\delta\rho}{\rho}\right)
+  =c_s^2\rho\delta\bm{v}^*\cdot\nabla\left[\frac{1}{\rho}\nabla\cdot\left(\rho\delta\bm{v}\right)\right].\label{global_integral}
+\end{align}
+At this point if we make the anelastic approximation by assuming
+$\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this equation \citep{nelson13},
+then the only contribution to instability comes from the third term on
+the left hand side, i.e. the explicit vertical shear term $\p_zj$. The 
+term proportional to $\kappa^2$ is stabilizing. Because radial shear
+is much larger than vertical shear, this implies that significant
+instability can only be achieved with large vertical
+motions (i.e. $|\delta v_z|\gg|\delta v_r|$).
+
+If we integrate Eq. \ref{global_integral} over a volume
+and assume boundary terms vanish when integrating by parts, we obtain 
+\begin{align}
+  -\sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV +
+  \int \kappa^2\rho|\delta v_r|^2 dV + \int \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
+  \delta v_r^*\delta v_z dV = -\int
+  \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
+  dV - \int \nabla\cdot\left(\rho\delta\bm{v}\right) \delta
+  \bm{v}^*\cdot\nabla c_s^2.
+\end{align}
+Recalling $j=r^2\Omega$ and $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r c_s^2$
+for a locally isothermal disk, we obtain
+\begin{align}
+  \sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV =
+  \int \kappa^2\rho|\delta v_r|^2 dV + \int
+  \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
+  dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
+    v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2}{dr}  dV.
+\end{align}
+We see that instability is only possible if
+$dc_s^2/dr\neq0$. Furthermore, since density and temperature typically
+decrease outwards in astrophysical disks, instability is associated
+with the term proportional to $\nabla\cdot\delta\bm{v}$. It is also
+clear that $\sigma^2$ will be complex in general.
+
 \section{Coefficients for the dispersion relation for perturbations
@@
 \begin{align}
-  &c_0 = M(M+1)\widetilde{A}^2,\\
+  &c_0 = M(M+1)A^2,\\
   &c_1 = \ii\beta\left\{\left(1-\gamma\right)\left[1 +
       \khat^2\left(1+2M\right)^2 - 4 \ii\epsilon q\khat M (M+1)\right] 
-    - 2\widetilde{A}^2\gamma M (M+1)\right\},\\
-  &c_2 = \left(\khat^2 + 1\right)\widetilde{A} + \beta^2\left\{(1-\gamma)\left[1
+    - 2A^2\gamma M (M+1)\right\},\\
+  &c_2 = \left(\khat^2 + 1\right)A + \beta^2\left\{(1-\gamma)\left[1
       + \gamma \khat^2(1+2M)^2 - 4\ii\epsilon q \khat \gamma M(M+1)
     \right]
-    -\gamma^2 \widetilde{A}^2 M(M+1)
+    -\gamma^2 A^2 M(M+1)
   \right\},\\

### 2015-02-10 appendix.tex
first rough draft

@@
 above. Inclusion of the global radial gradient terms results in the
-correct behavior  ($\nu\to0$ as $\beta\to\infty$).
+correct behavior  ($\nu\to0$ as $\beta\to\infty$). Despite this
+caveat, Fig. \ref{gcorr_compare} shows that provided we consider
+$\beta\ll1$, then the VGSB framework is an adequate approximation.
 
@@
 \end{figure}
-
-Fig. \ref{gcorr_compare} shows that provided we consider $\beta\ll1$,
-then the VGSB framework is an adequate approximation. It is
-interesting to note that there is a thermal relaxation timescale that
-maximizes the mode decay rate. Here, it is $\beta\simeq2$ or about
-$0.3$ orbital periods. This is, in fact, consistent with Fig. 12 of
-\cite{nelson13}.
 
@@
 
-\section{Eigenfrequency in locally
-  isothermal disks in global cylindrical geometry}
+\section{Axisymmetric instability in vertically 
+  isothermal disks in global cylindrical geometry}\label{global_vertiso}
 Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
 all times, so in the perturbed state $\delta P =
-c_s^2\delta\rho$, with $c_s(r)$ begin a prescribed function. 
+c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
 The linearized fluid equations in cylindrical
@@
 \end{align}
-We can eliminate $\delta v_\phi$ from these equations and obtain
+We can eliminate $\delta v_\phi$ from these equations to obtain
 \begin{align}
@@
 At this point if we make the anelastic approximation by assuming
-$\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this equation \citep{nelson13},
+$\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this
+equation \citep[as argued by][]{nelson13}, 
 then the only contribution to instability comes from the third term on
-the left hand side, i.e. the explicit vertical shear term $\p_zj$. The 
-term proportional to $\kappa^2$ is stabilizing. Because radial shear
+the left hand side, i.e. the explicit vertical shear term
+$\p_zj$. This is the same conclusion reached in
+\S\ref{integral_relation}.
+
+The term proportional to $\kappa^2$ is stabilizing. Because radial shear
 is much larger than vertical shear, this implies that significant
@@
 
-If we integrate Eq. \ref{global_integral} over a volume
+If we integrate Eq. \ref{global_integral} over the fluid volume 
 and assume boundary terms vanish when integrating by parts, we obtain 
@@
 \end{align}
-Recalling $j=r^2\Omega$ and $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r c_s^2$
-for a locally isothermal disk, we obtain
+Note that the treatment so far is in fact valid for any prescribed
+$c_s(r,z)$. However, to proceed further we specialize to the case
+$c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
+disks. Recalling $j=r^2\Omega$ and $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
+c_s^2$  for a vertically isothermal disk, we obtain 
 \begin{align}
@@
   \int \kappa^2\rho|\delta v_r|^2 dV + \int
-  \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
+  \frac{c_s^2(r)}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
   dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
-    v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2}{dr}  dV.
+    v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2(r)}{dr}  dV.
 \end{align}

### 2015-02-10 numerical.tex
paper updates, code updates
first rough draft

@@
-\section{Numerical solutions}
+\section{Numerical solutions}\label{numerical}
 In this section we numerically solve the linear stability problem
@@
 also plot the critical thermal relaxation timescale
-$\beta_\mathrm{crit}$ given by Eq. \ref{iso_vsi_cond}. For the present
-disk parameters $\beta_\mathrm{crit}=0.125$.
+$\beta_\mathrm{crit}=0.125$ given by Eq. \ref{iso_vsi_cond} for the present
+disk parameters.
 
@@
 
+It is interesting to note that there is a thermal relaxation timescale that
+maximizes the mode decay rate for $1\lesssim \khat \lesssim 30$. Here,
+it is $\beta\simeq1$---$2$ or about $0.1$---$0.3$ orbital
+periods. This is in fact consistent with Fig. 12 of \cite{nelson13}
+where the perturbed kinetic energy in their numerical simulations was
+observed to decay most rapidly for a thermal timescale of $\sim 0.1$
+orbits.
+
 %implications for nonlinear sims?
@@
  \begin{figure}
-   \includegraphics[width=\linewidth]{figures/bcrit_compare} 
+   \includegraphics[width=\linewidth]{figures/gcorr_compare2} 
    \caption{Growth rate of the fundamental VSI as
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/gcorr_compare_vnoniso}
+  0cm]{figures/gcorr_compare_vnoniso2}
   \caption{Growth rate of the fundamental VSI mode as a function of

### 2015-02-10 paper.tex
first rough draft

@@
 disk analog of the Goldreich-Schubert-Fricke instability
-\citep[GSF,]{goldreich67,fricke68} in differentially rotating stars.   
+\citep[GSF,][]{goldreich67,fricke68} in differentially rotating stars.   
 The height-dependence of $\Omega$, or vertical shear, enables
@@
 The VSI has been studied analytically in local approximation
-\citep{urpin98,urpin03}, where a Fourier analysis was performed in 
-both the radial and vertical directions. Simulations of the  
+\citep{urpin98,urpin03}. Simulations of the  
 VSI have been performed by \cite{arlt04}, and more recently by 
@@
 
-From these works it is clear that the VSI is sensitive 
-to the thermodynamic response of the disk except in the special case of neutrally
-stratified disks. It is therefore important to understand this dependence in order to evaluate
-the viability of the VSI in realistic disks. However,
+The VSI is sensitive to disk thermodynamics except
+in the special case of neutrally stratified disks. It is therefore
+important to understand and quantify this dependence, for example on
+disk parameters, in order to evaluate the viability of the VSI in
+realistic disks. Numerical simulations of the VSI are challenging,
+however, because global disk simulations should be carried out while
+resolving radial scales much smaller than the local disk scale height.
+
+On the other hand, should a given disk be
+unstable to the VSI, it must pass through the linear instability
+first. Thus, studying the thermodynamic dependence of the linear VSI
+is a useful first step. Furthermore, linear stability calculations are
+essential to complement future numerical simulations.
 
 %what we're doing
+In this work we study the linear VSI in radially-local,
+vertically-global models of astrophysical disks. We include an energy
+equation with a source term that allows the thermodynamic response of
+perturbations between varied from isothermal to adiabatic. Our
+formulation generalizes previous linear calculations of the 
+VSI carried out by \cite{nelson13} and \cite{mcnally14} which assume
+isothermal perturbations. Our models are also an extension of early
+analysis of \cite{urpin98} and \cite{urpin03} who consider
+vertically-localized disturbances. Our vertically-global approach is
+necessary to study disturbances with vertical wavelengths
+comparable to or larger than the disk scaleheight. We thus focus on
+vertically-global perturbations, which were found by \cite{nelson13}
+to dominate their numerical simulations.
 
 %paper organization
+
+%say "grossly simplified
+
+This paper is organized as follows. In \S\ref{setup} we list the
+governing equations and global equilibrium disk models under
+consideration. We formulate the linear stability problem in
+\S\ref{linear} and simplify it by making the radially-local
+approximation. We study the linear problem using two approaches. In
+\S\ref{analytical} we present an analytical discussion of vertically
+isothermal disks. Our treatment is grossly simplified owing to a
+number of approximations made, but closely parallels previous work and 
+allows us to highlight the essential effects of vertical shear. In
+particular, we quantify a characteristic thermal timescale below which
+the VSI is expected to operate. In \S\ref{numerical} we numerically
+solve the linearized fluid equations in the radially-local
+approximation. We find basic agreement with our analysis, including
+the aforementioned characteristic thermal timescale. In
+\S\ref{application} we apply our linear models to the Minimum Mass
+Solar Nebulae (MMSN) and find that the VSI may operate in the outer
+parts of the MMSN. We summarize and discuss our results in
+\S\ref{summary}, including caveats and possible directions for future
+work.
 

### 2015-02-10 setup.tex
first rough draft

@@
-\section{Governing equations and disk models}
+\section{Governing equations and disk models}\label{setup}
 We consider  an inviscid, non-self-gravitating three-dimensional (3D)
@@
   -\frac{\p P}{\p z}\left(\frac{\p j^2}{\p r}\frac{\p S}{ \p z} -
-    \frac{\p j^2}{\p z}\frac{\p S}{\p R} \right)&>0.  
+    \frac{\p j^2}{\p z}\frac{\p S}{\p r} \right)&>0. \label{second_SH} 
 \end{align}
@@
 compared to rotation, so the first criterion is generally
-satisfied. Thus, to assess stability, we evaluate the second criterion
-for the above disk models in $z>0$ where $\p_zP<0$, without loss of
-generality,   
+satisfied. Thus, to assess stability, we consider the second
+criterion. For our disk models,
+\begin{align}
+  &\frac{\p S}{\p r} = c_p\left[\frac{s}{\gamma r} +
+    \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p
+       r}\right],\\
+  &\frac{\p S}{\p z} = c_p
+  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z}. 
+\end{align} 
+Without loss of generality, we evaluate Eq. \ref{second_SH} in $z>0$
+where $\p_zP<0$. Here we approximate $\p_r j^2\simeq r^3\Omega_k^2$ but
+use Eq. \ref{vertical_shear} for $\p_zj^2$. Then 
 \begin{align}\label{solberg2}
-  \gamma - \Gamma > \frac{s\epsilon^2}{\Gamma}
-  \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s
-    -\left(\gamma-\Gamma\right)\frac{\p\ln{P}}{\p\ln{r}} \right]. 
+  % \gamma - \Gamma > \frac{s\epsilon^2}{\Gamma}
+  % \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s
+  %   -\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]. 
+  \gamma - \Gamma > \frac{\epsilon^2}{\Gamma}
+  \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s^2
+    -s\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]. 
 \end{align} 
 For typical model parameters the right hand side is
-$O(\epsilon^2)$. Hence, to have instability in the adiabatic limit 
+$O(\epsilon^2)$. Hence, instability in the adiabatic limit 
 requires $\gamma < \Gamma + O(\epsilon^2)$. Since we consider
@@
 
-\section{Linear problem} 
+\section{Linear problem}\label{linear}
 We consider axisymmetric perturbations to the above equilibria in the
@@
 \subsection{Radially-local approximation}  
-We assume perturbations have spatial dependence in the form
+We consider perturbations with small radial lengthscales 
+so that we can ignore the radial variations in the background
+structure and radial gradients. We thus evaluate all cofficients in
+the above equations at the fiducial radius $r=r_0$. However, the full
+vertical dependence of the coefficients are retained. Then we can
+assume perturbations have spatial dependence in the form
 \begin{align}
-  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r} 
+  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r},
 \end{align}
-such that $|k_rr|\gg1$, so that curvature terms can be ignored. For
-these small radial lengthscale perturbations we also ignore the radial
-variations in the background structure and radial gradients, and
-evaluate all cofficients in the above equations at the fiducial radius
-$r=r_0$. However, the full vertical dependence of the coefficients are
-retained.
+and we assume $|k_rr|\gg1$ so that curvature terms can also be
+ignored.
 
-In order to emphasize on the radially-local approximation, we
+In order to emphasize the radially-local approximation, we
 introduce the local Cartesian co-ordinate $(x,y,z)$ corresponding to
@@
 Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
-study of the vertical shear instability, but for analytical discussion
+study, but for analytical discussion
 we will make further simplifications.

### 2015-02-10 application.tex
first rough draft
spell-checked first rough draft

@@
 approximations. Specifically, we consider vertically-isothermal disks
-and replace explicit $\rho$ depedences in the above expressions by $\Sigma/
+and replace explicit $\rho$ dependencies in the above expressions by $\Sigma/
 H_\mathrm{iso}$, where $\Sigma$ is the surface density at the fiducial
@@
 \end{align}
-We recall $\beta < \beta_\mathrm{crit}$ is neccessary for the
+We recall $\beta < \beta_\mathrm{crit}$ is necessary for the
 fundamental VSI when $\khat\gg 1$. %For $\khat\to0$, this
@@
 Eq. \ref{bcrit_mmsn} for the Minimum Mass Solar Nebula (MMSN) by setting
-the scalings $\hat{T}=\hat{\Sigma}=\hat{\kappa}_d=1$, with $\mu =
+the scaling $\hat{T}=\hat{\Sigma}=\hat{\kappa}_d=1$, with $\mu =
 2.33$ and $\gamma=1.4$. The figure suggests that the fundamental VSI
@@
 This corresponds to a characteristic growth time $\sim
-40$ orbits. However, since growth rates are weakly depedent on $\khat$ for
+40$ orbits. However, since growth rates are weakly dependent on $\khat$ for
 $10\lesssim \khat \lesssim 60$, this may be taken as a representative growth
@@
 
-\subsubsection{Vertically-depedent thermal relaxation} 
+\subsubsection{Vertically-dependent thermal relaxation} 
 Note that for a vertically isothermal disk and opacity $\kappa_d(T)$, 
@@
 \tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
-$\hat{\rho}=\rho/\rho(z=0)$.  We have have repeated the example
-calculation above with $\beta\to\tilde{\beta}$. 
+$\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example
+calculation above with $\beta\to\tilde{\beta}$, but found similar
+results. 
 %We find growth rates

### 2015-02-10 linear_adia.tex
spell-checked first rough draft

@@
 One effect of introducing finite thermal relaxation is buoyancy. 
-The $\zhat^2$ depedence in $\delta v_{z1}$ then makes sense because it
+The $\zhat^2$ dependence in $\delta v_{z1}$ then makes sense because it
 is most significant for large $\zhat$, i.e. away from the midplane
@@
 This dispersion relation generally requires a numerical
-solution. However, for the fundamental mode ($M=0$) some analytical
+solution. However, for the fundamental mode ($M=0$) some analytic
 progress can be made, which we consider next.

### 2015-02-10 linear_iso.tex
first rough draft
spell-checked first rough draft

@@
 replacement after eliminating variables, which requires the vertical
-derivative of $D$, adds unecessary complexity for thin disks.
+derivative of $D$, adds unnecessary complexity for thin disks.
 
@@
 Note that we can generalize Eq. \ref{vertiso_gov_nondim} to other
-(non-Keplerian) rotation profiles by redfining $\khat$
+(non-Keplerian) rotation profiles by redefining $\khat$
 as $k_xc_s/\kappa$, $\hat{\sigma}$ as $\sigma/\kappa$, and 
@@
 
-\subsubsection{Neccessity of vertical shear for
+\subsubsection{Necessity of vertical shear for
   instability}\label{integral_relation} 
-We can establish a necessary condition for instability by multipling
+We can establish a necessary condition for instability by multiplying
 Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
@@
 \end{align}
-It follows that for instability ($\imag\hat{\sigma}>0$), it is neccessary to
-have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical shear.
+It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
+have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
+shear. In Appendix \ref{global_vertiso} we reach a similar conclusion
+in the global linear problem.
 
@@
 For the vertically isothermal disk we have $|f(\hat{z})| =
-|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, which maximises
+|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, which maximizes
 at $|\zhat|=1/\sqrt{2}\epsilon$. For vertical domains smaller 
@@
 \end{align}
-where $n$ is not neccessarily an integer. Indeed, if we demand
+where $n$ is not necessarily an integer. Indeed, if we demand
 a polynomial of degree $L$ as the solution, then the eigenfrequency

### 2015-02-10 numerical.tex
spell-checked first rough draft

@@
 % $\epsilon=0.1$. 
-We first calculate the VSI in a nearly vertcally isothermal, neutrally
-straitified disk ($N_z^2\equiv0$) by setting
+We first calculate the VSI in a nearly vertically isothermal, neutrally
+stratified disk ($N_z^2\equiv0$) by setting
 $\gamma=\Gamma=1.011$ and $\beta=10^{-3}$. Other disk parameters are 
@@
 \end{align*}
-This agreement is surpsingly good given the number of approximations
+This agreement is surprisingly good given the number of approximations
 used to obtain Eq. \ref{sig2_iso}, which also imposes a different
@@
 expression for $\sigma$, given by Eq. \ref{integral_relation1},
-involve integrands with $\rho$ as the   weight function, which rapidly
+involve integrals with $\rho$ as the   weight function, which rapidly
 decays away from the midplane.
@@
 Interestingly, these values are equal to
-$\epsilon|q|/(\gamma-\Gamma)$. This suggests that the critial thermal
+$\epsilon|q|/(\gamma-\Gamma)$. This suggests that the critical thermal
 relaxation timescale for vertically non-isothermal disks can also be

### 2015-02-10 paper.tex
spell-checked first rough draft

@@
   with previous hydrodynamic simulations, which indicates the VSI 
-  requires a short thermal relaxation timscale. For vertically  
+  requires a short thermal relaxation timescale. For vertically  
   isothermal disks, we quantify `fast'
@@
 % stuff about hydrodynamic angular momentum transport 
-Understanding the mechanisms of angular mometum transport is the
+Understanding the mechanisms of angular momentum transport is the
 central problem in the theory of accretion disks 
@@
 %recent work on vsi 
-One possibiliy is purely hydrodynamic turbulence through the vertical
+One possibility is purely hydrodynamic turbulence through the vertical
 shear instability (VSI). The VSI is an axisymmetric linear instability
@@
 necessary to study disturbances with vertical wavelengths
-comparable to or larger than the disk scaleheight. We thus focus on
+comparable to or larger than the disk scale height. We thus focus on
 vertically-global perturbations, which were found by \cite{nelson13}

### 2015-02-10 setup.tex
spell-checked first rough draft

@@
 taking the limit $\Gamma\to 1$. We will, in fact, primarily focus on
-vertically isothermal disks for simplicitly.
+vertically isothermal disks for simplicity.
 
@@
 and similarly for other fluid variables, where $\sigma = \omega + \ii
-\nu$ is a (generally) complex frequency, with $\omega$ amd $\nu$ being
+\nu$ is a (generally) complex frequency, with $\omega$ and $\nu$ being
 the real frequency and growth rate, respectively.
@@
 so that we can ignore the radial variations in the background
-structure and radial gradients. We thus evaluate all cofficients in
+structure and radial gradients. We thus evaluate all coefficients in
 the above equations at the fiducial radius $r=r_0$. However, the full

### 2015-02-10 summary.tex
updating
first rough draft
spell-checked first rough draft

@@
-\section{Summary and discussion}
+\section{Summary and discussion}\label{summary}
 %what have we done, relation to previous work 
@@
 height-dependence of the equilibrium angular velocity profile
-$\Omega(r,z)$ of a baroclinic disk. The VSI was analysed and
+$\Omega(r,z)$ of a baroclinic disk. The VSI was analyzed and
 explained earlier in the local limit by \cite{urpin98} and
@@
 radially-local approximation in the linearized fluid equations. For
-analytic discussion only, we further simplied the linear problem by
+analytic discussion only, we further simplified the linear problem by
 ignoring the background radial disk structure. This is equivalent to adopting
@@
 discussion, and because \cite{nelson13} found that this mode 
-eventually dominate their simualtions. Our numerical solutions are
+eventually dominate their simulations. Our numerical solutions are
 mostly consistent with our analytic expectations, and we verified the
@@
 in the disk model ($\beta$ decreases as $T^{-5}$ for perturbations with small
-radial scale). Thus, increasing the disk temperature even slightly can signifcantly 
+radial scale). Thus, increasing the disk temperature even slightly can significantly 
 shorten the typical growth times for the VSI. This is also attributed
@@
 the linear problem to ordinary differential equations, this approach
-introduces the wavenumber $k_x$ which is a free parameter. The effect
+introduces the wavenumber $k_x$ as a free parameter. The effect
 of radial boundary conditions cannot be explored. Future

### 2015-02-11 numerical.tex
minor updates

@@
 Fig. \ref{gcorr_compare_vnoniso} plots the fundamental VSI growth
-rates as a function of $\beta$ for $\gamma\in[1.4,2.5]$. For
-comparison with \cite{nelson13} we consider 
-$\beta\in[10^{-2},10^2]$. In agreement with \citeauthor{nelson13}, in
+rates as a function of $\beta$ for $\gamma\in[1.4,2.5]$. In agreement with \citeauthor{nelson13}, in 
 the neutrally-stratified case $\gamma=\Gamma$ the disk is unstable 
 even for $\beta\gg 1$. (Note that in the adiabatic limit 
-$\beta\to\infty$ this disk can be unstable according to the
+this disk can be unstable according to the
 Solberg-Hoiland criterion, Eq. \ref{solberg2}.)

### 2015-02-12 appendix.tex
spell-checked first rough draft
minor edits

@@
 In \S\ref{analytical} we used the vertically global shearing box
-\citep[VGSB,][]{mcnally14} to make analytical progress. % Here, we discuss limitations of this approach.   
+\citep[VGSB,][]{mcnally14} to make analytic progress. % Here, we discuss limitations of this approach.   
 The main difference between the linear problem in the VGSB 
@@
 $\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
-expected, a sub-adiabatically stratified disk is stable in the absense
+expected, a sub-adiabatically stratified disk is stable in the absence
 of vertical shear.
@@
 c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
-The linearized fluid equations in cylindrical
+The axisymmetric linearized fluid equations in cylindrical
 geometry is then 
@@
    z}\left(\frac{\delta\rho}{\rho}\right),\\
- \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right).
+ \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right),
 \end{align}
-We can eliminate $\delta v_\phi$ from these equations to obtain
+and we recall $j=r^2\Omega$. We can eliminate $\delta v_\phi$ from these equations to obtain
 \begin{align}
@@
 $c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
-disks. Recalling $j=r^2\Omega$ and $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
-c_s^2$  for a vertically isothermal disk, we obtain 
+disks. Recalling $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
+c_s^2$  for a vertically isothermal disk (Eq. \ref{vertical_shear}), we obtain 
 \begin{align}
@@
 $\hat{T}$ are dimensionless coefficients used to scale the model
-relative to the MMSN. Their nomial values are unity. By assuming a
+relative to the MMSN. Their nominal values are unity. By assuming a
 vertically isothermal disk, we deduce the disk aspect-ratio 

### 2015-02-12 paper.tex
minor edits

@@
 equation with a source term that allows the thermodynamic response of
-perturbations between varied from isothermal to adiabatic. Our
+perturbations to be varied smoothly between isothermal and adiabatic. Our
 formulation generalizes previous linear calculations of the 
@@
 necessary to study disturbances with vertical wavelengths
-comparable to or larger than the disk scale height. We thus focus on
-vertically-global perturbations, which were found by \cite{nelson13}
-to dominate their numerical simulations.
+comparable to or larger than the disk scale height. Such disturbances were 
+found by \cite{nelson13} to dominate their numerical simulations. 
+We thus focus on vertically-global perturbations.
 

### 2015-02-12 summary.tex
minor edits

@@
 
-In accordance with the above studies, which show that the VSI applies
-to perturbations with small radial lengthscales, we applied the 
+In accordance with the above studies, which show that the VSI occurs for 
+perturbations with small radial lengthscales, we applied the 
 radially-local approximation in the linearized fluid equations. For
@@
 eventually dominate their simulations. Our numerical solutions are
-mostly consistent with our analytic expectations, and we verified the
+mostly consistent with our analytic expectations, including the 
 dependence of $\beta_\mathrm{crit}$ on disk parameters.

### 2015-02-23 appendix.tex
new subsection growth time vs rad for mmsn

@@
 neutrally stratified. %also need real freq^2 in adia limit 
-This is because the VGSB retains the destabilizing effect of vertical shear
+This is because the VGSB retains the vertical shear
 ($\p_z\Omega$) from the global disk, but ignores the global radial disk
@@
 
-\section{Axisymmetric instability in vertically 
-  isothermal disks in global cylindrical geometry}\label{global_vertiso}
-Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
-all times, so in the perturbed state $\delta P =
-c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
-The axisymmetric linearized fluid equations in cylindrical
-geometry is then 
-\begin{align}
- \ii\sigma \delta v_r + 2\Omega\delta v_\phi &= c_s^2\frac{\p}{\p r}\left(\frac{\delta\rho}{\rho}\right),\\
- \ii\sigma \delta v_\phi &= \frac{1}{r}\delta\bm{v}\cdot\nabla j,\\
- \ii\sigma \delta v_z &= c_s^2\frac{\p}{\p
-   z}\left(\frac{\delta\rho}{\rho}\right),\\
- \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right),
-\end{align}
-and we recall $j=r^2\Omega$. We can eliminate $\delta v_\phi$ from these equations to obtain
-\begin{align}
-  -\sigma^2\rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) +
-  \kappa^2\rho|\delta v_r|^2 + \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
-  \delta v_r^*\delta v_z =
-  c_s^2\rho\delta\bm{v}^*\cdot\nabla\left(\frac{\ii\sigma\delta\rho}{\rho}\right)
-  =c_s^2\rho\delta\bm{v}^*\cdot\nabla\left[\frac{1}{\rho}\nabla\cdot\left(\rho\delta\bm{v}\right)\right].\label{global_integral} 
-\end{align}
-At this point if we make the anelastic approximation by assuming
-$\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this
-equation \citep[as argued by][]{nelson13}, 
-then the only contribution to instability comes from the third term on
-the left hand side, i.e. the explicit vertical shear term
-$\p_zj$. This is the same conclusion reached in
-\S\ref{integral_relation}.
+% \section{Axisymmetric instability in vertically 
+%   isothermal disks in global cylindrical geometry}\label{global_vertiso}
+% Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
+% all times, so in the perturbed state $\delta P =
+% c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
+% The axisymmetric linearized fluid equations in cylindrical
+% geometry is then 
+% \begin{align}
+%  \ii\sigma \delta v_r + 2\Omega\delta v_\phi &= c_s^2\frac{\p}{\p r}\left(\frac{\delta\rho}{\rho}\right),\\
+%  \ii\sigma \delta v_\phi &= \frac{1}{r}\delta\bm{v}\cdot\nabla j,\\
+%  \ii\sigma \delta v_z &= c_s^2\frac{\p}{\p
+%    z}\left(\frac{\delta\rho}{\rho}\right),\\
+%  \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right),
+% \end{align}
+% and we recall $j=r^2\Omega$. We can eliminate $\delta v_\phi$ from these equations to obtain
+% \begin{align}
+%   -\sigma^2\rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) +
+%   \kappa^2\rho|\delta v_r|^2 + \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
+%   \delta v_r^*\delta v_z =
+%   c_s^2\rho\delta\bm{v}^*\cdot\nabla\left(\frac{\ii\sigma\delta\rho}{\rho}\right)
+%   =c_s^2\rho\delta\bm{v}^*\cdot\nabla\left[\frac{1}{\rho}\nabla\cdot\left(\rho\delta\bm{v}\right)\right].\label{global_integral} 
+% \end{align}
+% At this point if we make the anelastic approximation by assuming
+% $\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this
+% equation \citep[as argued by][]{nelson13}, 
+% then the only contribution to instability comes from the third term on
+% the left hand side, i.e. the explicit vertical shear term
+% $\p_zj$. This is the same conclusion reached in
+% \S\ref{integral_relation}.
 
-The term proportional to $\kappa^2$ is stabilizing. Because radial shear
-is much larger than vertical shear, this implies that significant
-instability can only be achieved with large vertical
-motions (i.e. $|\delta v_z|\gg|\delta v_r|$).
+% The term proportional to $\kappa^2$ is stabilizing. Because radial shear
+% is much larger than vertical shear, this implies that significant
+% instability can only be achieved with large vertical
+% motions (i.e. $|\delta v_z|\gg|\delta v_r|$).
 
-If we integrate Eq. \ref{global_integral} over the fluid volume 
-and assume boundary terms vanish when integrating by parts, we obtain 
-\begin{align}
-  -\sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV +
-  \int \kappa^2\rho|\delta v_r|^2 dV + \int \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
-  \delta v_r^*\delta v_z dV = -\int
-  \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
-  dV - \int \nabla\cdot\left(\rho\delta\bm{v}\right) \delta
-  \bm{v}^*\cdot\nabla c_s^2.
-\end{align}
-Note that the treatment so far is in fact valid for any prescribed
-$c_s(r,z)$. However, to proceed further we specialize to the case
-$c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
-disks. Recalling $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
-c_s^2$  for a vertically isothermal disk (Eq. \ref{vertical_shear}), we obtain 
-\begin{align}
-  \sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV =
-  \int \kappa^2\rho|\delta v_r|^2 dV + \int
-  \frac{c_s^2(r)}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
-  dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
-    v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2(r)}{dr}  dV.
-\end{align}
-We see that instability is only possible if
-$dc_s^2/dr\neq0$. Furthermore, since density and temperature typically
-decrease outwards in astrophysical disks, instability is associated
-with the term proportional to $\nabla\cdot\delta\bm{v}$. It is also
-clear that $\sigma^2$ will be complex in general.
+% If we integrate Eq. \ref{global_integral} over the fluid volume 
+% and assume boundary terms vanish when integrating by parts, we obtain 
+% \begin{align}
+%   -\sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV +
+%   \int \kappa^2\rho|\delta v_r|^2 dV + \int \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
+%   \delta v_r^*\delta v_z dV = -\int
+%   \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
+%   dV - \int \nabla\cdot\left(\rho\delta\bm{v}\right) \delta
+%   \bm{v}^*\cdot\nabla c_s^2.
+% \end{align}
+% Note that the treatment so far is in fact valid for any prescribed
+% $c_s(r,z)$. However, to proceed further we specialize to the case
+% $c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
+% disks. Recalling $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
+% c_s^2$  for a vertically isothermal disk (Eq. \ref{vertical_shear}), we obtain 
+% \begin{align}
+%   \sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV =
+%   \int \kappa^2\rho|\delta v_r|^2 dV + \int
+%   \frac{c_s^2(r)}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
+%   dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
+%     v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2(r)}{dr}  dV.
+% \end{align}
+% We see that instability is only possible if
+% $dc_s^2/dr\neq0$. Furthermore, since density and temperature typically
+% decrease outwards in astrophysical disks, instability is associated
+% with the term proportional to $\nabla\cdot\delta\bm{v}$. It is also
+% clear that $\sigma^2$ will be complex in general.
 

### 2015-02-23 application.tex
new subsection growth time vs rad for mmsn

@@
 
-\subsubsection{Vertically-dependent thermal relaxation} 
-Note that for a vertically isothermal disk and opacity $\kappa_d(T)$, 
-we can restore the $z$ dependence in $\beta$ (due to the $\rho$
-dependence in  $\eta$, see Eq. \ref{eta_def}) by making the transformation
-$\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv 
-\tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
-$\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example
-calculation above with $\beta\to\tilde{\beta}$, but found similar
-results. 
+\subsection{Applicability of the VSI to the MMSN}
+To obtain an overall picture of the relevance of the VSI to the MMSN,
+we calculate characteristic growth times $t_\mathrm{grow} = 1/\nu$ of 
+the fundamental mode in units of the local Keplerian orbital period
+$P_\mathrm{orb}= 2\pi/\Omega_k$. The results are shown in 
+Fig. \ref{mmsn_overall} for perturbation wavenumbers $5\lesssim
+\khat\lesssim 50$. We consider this range because for  
+$\khat=O(1)$ or $\khat=O(10^2)$ we found eigenfunctions have large
+amplitudes near the vertical boundaries, so such modes may be affected
+by boundary conditions.
+
+In the MMSN the fundamental VSI operates at tens of AU, with
+characteristic growth times of tens of orbits. %In our example, the VSI
+% is most favored at $r\simeq 20\mathrm{AU}$ with a growth time
+% $t_\mathrm{grow}\simeq 30$ orbits.
+Fig. \ref{mmsn_overall} indicates that with decreasing radius, the VSI
+shifts to smaller scales (larger $\khat$), since the disk becomes optically
+thick so only small-scale disturbances have sufficiently short thermal
+relaxation timescales. At larger radii ($r\gtrsim 
+40\mathrm{AU}$,) the VSI can operate on larger scales (smaller
+$\khat$) and growth times do not depend strongly on
+$\khat$, since the disk is optically thin. However, approaching
+$r\simeq 100\mathrm{AU}$ growth times become long, at $O(100)$ orbits so
+the VSI may not be dynamically important in the outermost parts of the
+MMSN.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/eigen_compare_grow.ps}
+  \caption{Characteristic growth times of the fundamental VSI mode in
+    the MMSN for a range of perturbation wavenumbers.  
+    \label{mmsn_overall}}    
+\end{figure}
+
+ % \subsection{Vertically-dependent thermal relaxation} 
+% Note that for a vertically isothermal disk and opacity $\kappa_d(T)$, 
+% we can restore the $z$ dependence in $\beta$ (due to the $\rho$
+% dependence in  $\eta$, see Eq. \ref{eta_def}) by making the transformation
+% $\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv 
+% \tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
+% $\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example
+% calculation in \S\ref{mmsn_example} with $\beta\to\tilde{\beta}$, but
+% found similar results.
+
 %We find growth rates

### 2015-02-23 linear_adia.tex
new subsection growth time vs rad for mmsn

@@
 \end{align} 
-Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} in
-the limit $\khat\to\infty$ we find
+Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
+$\khat\gg1$, we find
 \begin{align}
@@
 \end{align}
-We seek solutions with
+This is a pair of simultaneous equations for
+$\hat{\omega}$ and $\beta$.%  A simple solution can be obtained by the
+% following considerations.
+Now, our discussion of stable solutions 
+without vertical shear yield a dispersion
+relation $\hat{\omega}\propto \khat^{-1}$ for $\khat\gg 1$
+(\S\ref{iso_stable}, Eq. \ref{iso_stable_disp}). We  therefore 
+seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with 
+%we wish to seek a criterion independent of k 
+%\begin{align}
+%  \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
+%\end{align}
 \begin{align}
-  \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
+  \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
+  |\hat{\omega}\khat|=O(1). 
 \end{align}

### 2015-02-23 linear_iso.tex
new subsection growth time vs rad for mmsn

@@
 have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
-shear. In Appendix \ref{global_vertiso} we reach a similar conclusion
-in the global linear problem.
+shear. A similar conclusion can be reached in the equivalent linear  
+problem in global cylindrical disk geometry.
 
@@
 
-\subsubsection{Stability in the absence of vertical shear}
+\subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
 When $q=0$, Eq. \ref{iso_ode3} is
@@
   \left|\hat{\sigma}\right| = \sqrt{n}
-  \left(1+\hat{k}^2\right)^{-1/2}. 
+  \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
 \end{align}
@@
 smaller than the local disk scale height.
+
+We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
+Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
+has also been qualitatively observed in numerical simulations performed by 
+\cite{stoll14}.
 

### 2015-02-23 summary.tex
new subsection growth time vs rad for mmsn

@@
 the fact that proper inclusion of viscosity introduces viscous heating
-in the energy equation.
+in the energy equation, as well as a meridional flow in the equilibrium state.
 

### 2015-02-25 application.tex
minor changes, new title

@@
 
-\subsection{Applicability of the VSI to the MMSN}
+\subsection{VSI in the MMSN}
 To obtain an overall picture of the relevance of the VSI to the MMSN,
@@
 relaxation timescales. At larger radii ($r\gtrsim 
-40\mathrm{AU}$,) the VSI can operate on larger scales (smaller
-$\khat$) and growth times do not depend strongly on
-$\khat$, since the disk is optically thin. However, approaching
-$r\simeq 100\mathrm{AU}$ growth times become long, at $O(100)$ orbits so
-the VSI may not be dynamically important in the outermost parts of the
-MMSN.
+40\mathrm{AU}$), the disk is optically thin so growth times do not depend strongly on   
+$\khat$, and the VSI can operate on larger scales (smaller $\khat$).  
+However, approaching 
+$r\simeq 100\mathrm{AU}$ growth times become long at $O(10^2)$ local orbits,
+so the VSI may not be dynamically important in the outermost parts of
+the MMSN.
 

### 2015-02-25 linear_adia.tex
minor changes, new title

@@
 This is a pair of simultaneous equations for
-$\hat{\omega}$ and $\beta$.%  A simple solution can be obtained by the
+$\hat{\omega}$ and $\beta$. % A simple solution can be obtained by the
 % following considerations.
-Now, our discussion of stable solutions 
+ Our discussion of stable solutions 
 without vertical shear yield a dispersion

### 2015-02-26 paper.tex
minor changes, new title
minor changes

@@
 
-\title{Linear analysis of the vertical shear instability in astrophysical disks}
-
+%\title{Linear analysis of the vertical shear instability in astrophysical disks}
+\title{Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula} 
 \author{Min-Kai Lin, Andrew Youdin}
@@
   find the MMSN is unstable to the VSI at tens of AU with a
-  characteristic growth times of tens of local orbital periods. We 
-  suggest the VSI is a viable mechanism for angular momentum
-  transport in the outer parts of protoplanetary disks. 
+  characteristic growth times of tens of local orbital periods. This 
+  suggests the VSI is  dynamically important in the outer parts of protoplanetary disks. 
 \end{abstract}
@@
 convective stability. When the thermal timescale $t_c$ is parametrized
-in units of the orbital period $P$, \cite{nelson13} find that
-$t_c\lesssim 0.01P$ was needed to observe instability. That is, the
+in units of the orbital period $P_\mathrm{orb}$, \cite{nelson13} find that
+$t_c\lesssim 0.01P_\mathrm{orb}$ was needed to observe instability. That is, the
 temperature profile of the disk essentially remains unchanged during

### 2015-02-26 setup.tex
minor changes

@@
 specific entropy, $c_p$ is the heat capacity at constant
-pressure (assumed to be constant), and $j\equiv r^2\Omega$ is the
+pressure, and $j\equiv r^2\Omega$ is the
 specific angular momentum.

### 2015-02-27 linear_iso.tex
minor changes
updated all figs

@@
 However, we find that for perturbations with small radial lengthscales
-(large $|k_x|$) and fast thermal relaxation ($\beta\ll1$), which
+($k_xH_\mathrm{iso}\gg 1$) and fast thermal relaxation ($\beta\ll1$), which
 characterizes the VSI, the VGSB framework is adequate for the purpose
@@
  \end{align}
-from Eq. \ref{lin_vz}. For isothermal perturbations 
-the parameter $\chi=1$ so it is easiest to work with
-$W$ by integrating Eq. \ref{vertiso_gov_nondim}. For later discussion
+from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact 
+simpler to work with $W$ by integrating Eq. \ref{vertiso_gov_nondim}. For later discussion
 it is convenient to write the result as

### 2015-02-27 setup.tex
updated all figs

@@
 \begin{align}
-  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r},
+  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r}.
 \end{align}
-and we assume $|k_rr|\gg1$ so that curvature terms can also be
+We take $k_r>0$ without loss of generality, and further assume $k_rr\gg1$ so that curvature terms can also be 
 ignored.
@@
   &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
-  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x.\label{lin_energy}
+  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x,\label{lin_energy}
   % -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
@@
 \end{align}
+where we have made the notation change $k_r\to k_x$. 
 It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
-\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$.   
+\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$ using their expressions in the global disk. 
 Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical

### 2015-02-28 application.tex
minor changes
redoing application section

@@
 
-In our disk models we have adopted a simple thermal relaxation 
-function in the energy equation. To connect our model with a realistic
-disk, we consider a more physically-motivated form of the energy
-equation with radiative cooling 
+%In our disk models we have adopted a simple thermal relaxation 
+%function in the energy equation. 
+For a realistic disk one can consider the energy equation with
+radiative cooling, 
+%To connect our model with a realistic
+%disk, we consider a more physically-motivated form of the energy
+%equation with radiative cooling 
 \begin{align}\label{real_energy}
-\rho T \frac{DS}{Dt} = - \left(\nabla\cdot\bm{F} - Q_+\right), 
-\end{align}
-where the temperature $T$ is defined through $P=\mathcal{R}\rho
+%\rho T \frac{DS}{Dt} = - \left(\nabla\cdot\bm{F} - Q_+\right), 
+\frac{\p P}{\p t} + \bm{v}\cdot\nabla P +\gamma P \nabla\cdot\bm{v} = - \frac{P}{\rho T
+  c_v}\left(\nabla\cdot\bm{F} - Q_+\right), 
+\end{align}where the temperature $T$ is defined through $P=\mathcal{R}\rho
 T/\mu$, $\mathcal{R}$ is the ideal gas constant and $\mu$ is
-the mean molecular weight, % In Eq. \ref{real_energy} $S$ is the
+the mean molecular weight, $c_v$ is the heat capacity at constant
+volume, 
+% In Eq. \ref{real_energy} $S$ is the
 % specific entropy,
@@
 % specific heat capcity at constant volume
-$D/Dt$ is the convective
-derivative, $\bm{F}$ is the radiative flux, and $Q_+$ is a heating
-term such that an equilibrium solution exists when $D/Dt=0$.
+$\bm{F}$ is the radiative flux, and $Q_+$ is a heating
+term such that an equilibrium solution exists.
 
-Our aim is to relate the right hand side (RHS) of Eq. \ref{real_energy} to
-the thermal relaxation time $t_c$ as used in our linear
-calculations. To do so, we first consider separately perturbations
-with radial lengthscales $l\equiv 2\pi/k_x\ll l_\mathrm{rad}$ and 
-$l\gg l_\mathrm{rad}$, where      
+One should now model in detail each of the terms on the right
+hand side (RHS) of Eq. \ref{real_energy}, linearize them, then solve
+the resulting linear eigenvalue problem. Such an approach is beyond
+the scope of this paper. Here, we simply model the RHS of
+Eq. \ref{real_energy} such that it can be represented by a constant
+thermal relaxation time $t_c$ as used in our linear calculations.  
+%Our aim is to relate the right hand side (RHS) of Eq. \ref{real_energy} to
+%the 
+To do so, we first consider separately perturbations
+with radial lengthscales $l\equiv 2\pi/k_x\gg l_\mathrm{rad}$ and 
+$l\ll l_\mathrm{rad}$, where      
 \begin{align}\label{lrad}
@@
 is the photon mean-free-path and $\kappa_d$ is the (dust) opacity.
+
 %need to comment on kappa_d(rho,T)
 
-\subsection{Newtonian cooling}\label{newton_cool}
-For $l\ll l_\mathrm{rad}$ the radiative cooling is given by {\bf (reference?)}
-\begin{align}
-\nabla\cdot\bm{F} = 4 \rho \kappa_d \sigma_s T^4,
-\end{align}   
-where $\sigma_s$ is the Stefan-Boltzmann constant. We choose the
-heating term 
-\begin{align}
-  Q_+ = 4\rho\kappa_d\sigma_s T^{4}_{t=0}. 
-\end{align}
-Then Eq. \ref{real_energy} becomes
-\begin{align}
-  \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} -
-  \frac{4P\kappa_d\sigma_s}{Tc_v}\left(T^4 - T_{t=0}^4\right), 
-\end{align}
-where $c_v$ is the heat capacity at constant volume. When linearized, the last term on the RHS becomes
-\begin{align}
-  -\frac{16\sigma_s}{c_v}\kappa_d T^3\left(\delta P -
-    \frac{P}{\rho}\delta\rho\right). 
-\end{align}
-By comparing this expression with the linearized version of the energy
-equation used in our calculations (Eq. \ref{full_energy1}), we identify 
-\begin{align}\label{tc_newton_cool} 
-  t_\mathrm{cool} = \frac{c_v}{16\sigma_s\kappa_dT^3}
-\end{align}
-as the thermal relaxation timescale for perturbation lengthscales
-$l\ll l_\mathrm{rad}$.
-
 \subsection{Radiative diffusion}
-For perturbation lengthscales $l\gg l_\mathrm{rad}$, we assume 
-temperature fluctuations are smoothed out by radiative diffusion. To
-model this in linear theory, we choose the heating term such that
-\begin{align}
-  Q_+ - \nabla\cdot\bm{F} \equiv \nabla\cdot\left[k_\mathrm{rad}\nabla
-    \left(T-T_{t=0}\right)\right], 
+Let us first consider perturbations with lengthscales $l\gg
+l_\mathrm{rad}$. In this case, we assume  
+temperature fluctuations are smoothed out by radiative diffusion. We
+mode this in linear theory by writing the linearized RHS of
+Eq. \ref{real_energy} as 
+% we choose the heating term such that
+% \begin{align}
+%   Q_+ - \nabla\cdot\bm{F} \equiv \nabla\cdot\left[k_\mathrm{rad}\nabla
+%     \left(T-T_{t=0}\right)\right], 
+% \end{align} 
+% where $k_\mathrm{rad}$ is the conduction coefficient to be defined
+% implicitly. Eq. \ref{real_energy} becomes
+% \begin{align}
+%   \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} + \frac{P}{\rho T
+%     c_v}\nabla\cdot\left[k_\mathrm{rad}\nabla\left(T-T_{t=0}\right)\right].     
+% \end{align}
+% When linearized, the last term on the RHS becomes
+\begin{align}\label{diff_cool_proper}
+  \frac{P}{\rho T c_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
+    T\right),  
 \end{align}
 where $k_\mathrm{rad}$ is the conduction coefficient to be defined
-implicitly. Eq. \ref{real_energy} becomes
-\begin{align}
-  \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} + \frac{P}{\rho T
-    c_v}\nabla\cdot\left[k_\mathrm{rad}\nabla\left(T-T_{t=0}\right)\right].     
-\end{align}
-When linearized, the last term on the RHS becomes
-\begin{align}\label{diff_cool_proper}
-  \frac{P}{\rho T c_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
-    T\right). 
-\end{align}
-This, of course, fundamentally changes the linear problem by
+implicitly.
+
+However, Eq. \ref{diff_cool_proper} fundamentally changes the linear problem by
 introducing higher order derivatives. We defer a proper analysis of
@@
 \begin{align}\label{eta_def}
-  \eta = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 c_v}. 
+  \eta = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 c_v}, 
 \end{align}
-From Eq. \ref{diff_cool_approx} we identify the thermal relaxation
+where $\sigma_s$ is the Stefan-Boltzmann constant. 
+From Eq. \ref{diff_cool_approx} we identify the scale-dependent thermal relaxation
 time for diffusion as 
@@
 \end{align}
-where $\hat{\eta} = \eta/H_\mathrm{iso}^2\Omega_k$ is the 
+where $\hat{\eta} = \eta/H^2\Omega_k$ is the 
 dimensionless diffusion coefficient.
 
-\subsection{Toy model for realistic thermal relaxation}\label{toy_relax}
+\subsection{Newtonian cooling}\label{newton_cool}
+For perturbations with small lengthscales, $l\ll l_\mathrm{rad}$, 
+the disk is optically thin. In this case, we assume 
+Newtonian cooling can be applied, which is the thermal
+relaxation model as adopted in our linear models. That is, temperature
+flucations $\delta T$ decay on a timescale $t_\mathrm{cool}$,
+independent of $k_x$. For simplicity, we take  
+% the radiative cooling is given by {\bf (reference?)}
+% \begin{align}
+% \nabla\cdot\bm{F} = 4 \rho \kappa_d \sigma_s T^4,
+% \end{align}   
+% where $\sigma_s$ is the Stefan-Boltzmann constant. We choose the
+% heating term 
+% \begin{align}
+%   Q_+ = 4\rho\kappa_d\sigma_s T^{4}_{t=0}. 
+% \end{align}
+% Then Eq. \ref{real_energy} becomes
+% \begin{align}
+%   \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} -
+%   \frac{4P\kappa_d\sigma_s}{Tc_v}\left(T^4 - T_{t=0}^4\right), 
+% \end{align}
+% where $c_v$ is the heat capacity at constant volume. When linearized, the last term on the RHS becomes
+% \begin{align}
+%   -\frac{16\sigma_s}{c_v}\kappa_d T^3\left(\delta P -
+%     \frac{P}{\rho}\delta\rho\right). 
+% \end{align}
+% By comparing this expression with the linearized version of the energy
+% equation used in our calculations (Eq. \ref{full_energy1}), we identify 
+% \begin{align}\label{tc_newton_cool} 
+%   t_\mathrm{cool} = \frac{c_v}{16\sigma_s\kappa_dT^3}
+% \end{align}
+% as the thermal relaxation timescale for optically thin cooling. 
+\begin{align}
+t_\mathrm{cool} = \frac{l_\mathrm{rad}^2}{3\eta}. 
+\end{align}
+
+\subsection{A model for combined thermal relaxation}\label{toy_relax}
 To unify $t_\mathrm{cool}$ and $t_\mathrm{diff}$, we define the
@@
 \begin{align}\label{tc_def}
-  t_c &\equiv t_\mathrm{cool} + t _\mathrm{diff} =
-  \frac{l_\mathrm{rad}^2}{3\eta} + \frac{1}{\eta k_x^2},  
+  t_c &\equiv t_\mathrm{cool} + t _\mathrm{diff}. %=
+ % \frac{l_\mathrm{rad}^2}{3\eta} + \frac{1}{\eta k_x^2},  
 \end{align}
-where the relation between $t_\mathrm{cool}$, $l_\mathrm{rad}$ and
-$\eta$ can be inferred from Eq. \ref{lrad}, Eq. \ref{tc_newton_cool}
-and Eq. \ref{eta_def}. Eq. \ref{tc_def} is a simple prescription so
+%where the relation between $t_\mathrm{cool}$, $l_\mathrm{rad}$ and
+%$\eta$ can be inferred from Eq. \ref{lrad}, Eq. \ref{tc_newton_cool}
+%and Eq. \ref{eta_def}. 
+Eq. \ref{tc_def} is a simple prescription so
 that for small scales (large $k_x$), $t_c\to t_\mathrm{cool}$, while
@@
 and replace explicit $\rho$ dependencies in the above expressions by $\Sigma/
-H_\mathrm{iso}$, where $\Sigma$ is the surface density at the fiducial
+H$, where $\Sigma$ is the surface density at the fiducial
 radius of interest. For example, the dimensionless diffusion
@@
   \beta \equiv t_c\Omega_k =
-  \frac{1}{\hat{\eta}}\left(\frac{l_\mathrm{rad}^2}{3H_\mathrm{iso}^2}
+  \frac{1}{\hat{\eta}}\left(\frac{l_\mathrm{rad}^2}{3H^2}
     + \frac{1}{\khat^2}\right)  \simeq
@@
 described in \cite{chiang10} and summarized in Appendix \ref{mmsn}
-(where an opacity model is also chosen). Defining $\zeta\equiv
+(where an opacity model is also chosen). Defining $r_\mathrm{AU}\equiv
 r/\mathrm{AU}$, we find Eq. \ref{real_beta} becomes
@@
 \beta =
-\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)\zeta^{-57/14}\left(\frac{8.3\times10^{-9}}{\hat{\kappa}_d^2\hat{T}^4\hat{\Sigma}^2}\zeta^{33/7}+\frac{1}{\khat^2}\right).
+\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)r_\mathrm{AU}^{-57/14}\left(\frac{8.3\times10^{-9}}{\hat{\kappa}_d^2\hat{T}^4\hat{\Sigma}^2}r_\mathrm{AU}^{33/7}+\frac{1}{\khat^2}\right).
 \end{align}
-We also evaluate the criterion Eq. \ref{iso_vsi_cond} for the fiducial
+We refer to the case with the scalings
+$\hat{\Sigma}=\hat{\kappa}_d=\hat{T}=1$ as the Minimum Mass Solar
+Nebulae (MMSN).  We also evaluate the criterion Eq. \ref{iso_vsi_cond} for the fiducial
 disk model,  
@@
   \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
-    -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\zeta^{2/7}. 
+    -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
 \end{align}
@@
 In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn} and
-Eq. \ref{bcrit_mmsn} for the Minimum Mass Solar Nebula (MMSN) by setting
-the scaling $\hat{T}=\hat{\Sigma}=\hat{\kappa}_d=1$, with $\mu =
-2.33$ and $\gamma=1.4$. The figure suggests that the fundamental VSI
+Eq. \ref{bcrit_mmsn} with with $\mu =2.33$ and $\gamma=1.4$.  We set
+the scaling $\hat{T}=\hat{\Sigma}=1$ and consider three opacity scales:
+$\hat{\kappa}_d=0.1, \,1 $ and $10$.
+
+The figure suggests that the fundamental VSI 
 operates in the outer parts of the MMSN at radii on 
-the order of tens of AU. However, the thermal relaxation time $\beta$
+the order of tens of AU ($\hat{\kappa}_d=1$, middle panel). 
+However, the thermal relaxation time $\beta$
 is not much shorter than $\beta_\mathrm{crit}$, so the VSI may be 
-weak.
+weak. Lowering the opacity shifts the curves
+inward, but raises also them. In fact, lowering the opacity by an
+order of magnitude renders the fiducial  disk stable to the fundametal
+mode for the range of wavenumbers considered ($\hat{\kappa}_d=0.1$,
+upper panel). Raising the opacity, on
+the other hand, decreases the thermal timescales, but also shifts the
+curves outward, so the fundamental mode is restricted to the outer
+disk ($\hat{\kappa}_d=10$, lower panel).
 
-We remark that it is possible to satisfy $\beta<\beta_\mathrm{crit}$
-at small radii by choosing a sufficiently large $\khat$. However, our
-linear calculations suggest growth rates eventually decay as 
-$\khat\to\infty$. Thus, very small-scale VSI in the inner disk is
-probably unimportant.
+%We remark that it is possible to satisfy $\beta<\beta_\mathrm{crit}$
+%at small radii by choosing a sufficiently large $\khat$. However, our
+%linear calculations suggest growth rates eventually decay as 
+%$\khat\to\infty$. Thus, very small-scale VSI in the inner disk is
+%probably unimportant.
 
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/bcrit_mmsn}  
-  \caption{Dimensionless thermal relaxation timescales $\beta$ in the MMSN
-    using Eq. \ref{beta_mmsn} for three values of the 
+  \includegraphics[scale=.43,clip=true,trim=0cm 1.8cm 0cm 0cm]{figures/bcrit_mmsn0d1} 
+  \includegraphics[scale=.43,clip=true,trim=0cm 1.8cm 0cm 1cm]{figures/bcrit_mmsn}  
+  \includegraphics[scale=.43,clip=true,trim=0cm 0cm 0cm 1cm]{figures/bcrit_mmsn10} 
+  \caption{Dimensionless thermal relaxation timescales $\beta$ in a
+    fiducial protoplanetary disk model. Eq. \ref{beta_mmsn} is plotted
+    for three values of the 
     perturbation radial wavenumber: $\khat=1$ (dotted), $\khat=10$
-    (dashed) and $\khat=100$ (dashed-dot). The solid line is the
-    critical thermal relaxation timescale $\beta_\mathrm{crit}$, and 
-    $\beta<\beta_\mathrm{crit}$
-    is necessary for the fundamental isothermal VSI to operate  
-    for $\khat\gg1$. 
+    (dashed) and $\khat=100$ (dashed-dot), for three values of the
+    opacity relative to the MMSN: $\hat{\kappa}_d=0.1$ (top),
+    $\hat{\kappa}_d=1$ (middle) and $\hat{\kappa}_d=10$ (bottom).  
+    The solid line is the
+    critical thermal relaxation timescale $\beta_\mathrm{crit}$. 
+    % and 
+    % $\beta<\beta_\mathrm{crit}$ is necessary for the fundamental isothermal VSI to operate  
+    % for $\khat\gg1$. 
     \label{mmsn_bcrit_bcool}}   
+\end{figure}
+
+We remark that it is possible, in principle, to satisfy $\beta <
+\beta_\mathrm{crit}$ by choosing a sufficiently large $\khat$,
+equivalently a sufficiently small perturbation lengthscales. We
+illustrate this by plotting at each radius the value of $\khat$ such
+that $\beta = \beta_\mathrm{crit}$. We see that in order for the
+fundamental VSI to operate in the inner disk (few AU), one needs to
+consider large values of $\khat$.
+
+However, our linear
+calculations indicate growth rates eventually decay as one increases
+$\khat$. This, together with our finding of high $\khat$ modes may be
+affected by vertical boundary conditions, suggest the VSI does not
+operate in the inner few AU in protoplanetary disks where thermal
+relaxation is entirely due to radiation \citep[cf.][]{stoll14}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/bcrit_mink} 
+  \caption{The minimum value of the perturbation wavenumber $\khat$ in
+    the MMSN such that the associated dimensionless thermal
+    relaxation time  $\beta$ is  less than the critical timescale
+    $\beta_\mathrm{crit}$. 
+    \label{mmsn_bcrit_bcool_mink}}   
 \end{figure}
@@
 
-\subsection{VSI in the MMSN}
+\subsection{VSI in the outer parts of the MMSN}
 To obtain an overall picture of the relevance of the VSI to the MMSN,

### 2015-02-28 application.tex
redoing app section

@@
 \section{Application to protoplanetary disks}\label{application} 
-As an application of our results, we estimate where in a 
-protoplanetary disk do thermodynamic conditions allow the VSI to 
-operate. Specifically, we consider the fundamental VSI, 
-for which we can use the criterion Eq. \ref{iso_vsi_cond}.
-
-%In our disk models we have adopted a simple thermal relaxation 
-%function in the energy equation. 
-For a realistic disk one can consider the energy equation with
+As a first application of our linear models, we estimate where in a
+protoplanetary disk do the thermodynamic conditions allow the VSI to
+operate. For a realistic disk one can consider the energy equation with
 radiative cooling, 
@@
 the scope of this paper. Here, we simply model the RHS of
-Eq. \ref{real_energy} such that it can be represented by a constant
-thermal relaxation time $t_c$ as used in our linear calculations.  
+Eq. \ref{real_energy} such that it can be represented by a 
+thermal relaxation time $t_c$ for the disk temperature. 
 %Our aim is to relate the right hand side (RHS) of Eq. \ref{real_energy} to
@@
 that for small scales (large $k_x$), $t_c\to t_\mathrm{cool}$, while
-for large scales (small $k_x$), $t_c\to t_\mathrm{diff}$.
-
-In order to utilize our linear results, which assumes $t_c$ is a
-constant, we need to make further 
-approximations. Specifically, we consider vertically-isothermal disks
-and replace explicit $\rho$ dependencies in the above expressions by $\Sigma/
-H$, where $\Sigma$ is the surface density at the fiducial
-radius of interest. For example, the dimensionless diffusion
-coefficient becomes 
-\begin{align}\label{diff_coeff_dimensionless}
-  \hat{\eta} \to  \frac{16\sigma_s T^3}{3\kappa_d\Sigma^2
-    c_v\Omega_k}. 
+for large scales (small $k_x$), $t_c\to t_\mathrm{diff}$. Writing the
+volume density $\rho = \Sigma\hat{\rho}/\sqrt{2\pi}H$, where
+$\hat{\rho} = \rho/\rho_0$ and $\Sigma$ is the surface density at the
+radius of interest, the 
+% In order to utilize our linear results, which assumes $t_c$ is a
+% constant, we need to make further 
+% approximations. Specifically, we consider vertically-isothermal disks
+% and replace explicit $\rho$ dependencies in the above expressions by $\Sigma/
+% H$, where $\Sigma$ is the surface density at the fiducial
+% radius of interest. For example, the dimensionless diffusion
+% coefficient becomes 
+% \begin{align}\label{diff_coeff_dimensionless}
+%   \hat{\eta} \to  \frac{16\sigma_s T^3}{3\kappa_d\Sigma^2
+%     c_v\Omega_k}. 
+% \end{align}
+% Furthermore, we will choose a temperature-dependent opacity model for
+% simplicity, so that $\kappa_d=\kappa_d(T)$.
+dimensionless thermal
+relaxation time $\beta$ becomes 
+\begin{align}\label{real_beta}
+  \beta(z;\khat) \equiv t_c\Omega_k =
+  % \frac{1}{\hat{\eta}}\left[\frac{l_\mathrm{rad}^2}{3H^2}
+  %   + \frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right]  \simeq
+  \frac{1}{\hat{\eta}}\left[\frac{1}{3\kappa_d^2\Sigma^2} 
+    + \frac{\hat{\rho}^2(z)}{2\pi \khat^2}\right].
 \end{align}
-Furthermore, we will choose a temperature-dependent opacity model for
-simplicity, so that $\kappa_d=\kappa_d(T)$. The dimensionless thermal
-relaxation time $\beta$  
-\begin{align}\label{real_beta}
-  \beta \equiv t_c\Omega_k =
-  \frac{1}{\hat{\eta}}\left(\frac{l_\mathrm{rad}^2}{3H^2}
-    + \frac{1}{\khat^2}\right)  \simeq
-  \frac{1}{\hat{\eta}}\left(\frac{1}{3\kappa_d^2\Sigma^2} 
-    + \frac{1}{\khat^2}\right)
-\end{align}
-is then a constant for vertically-isothermal disks, 
-and we can directly apply previous results.
+Note that the thermal relaxation time depends on both the perturbation lengthscale $\khat$ 
+and distance from the midplane. %The latter is due to the density-dependence in the diffusion time. 
+%is then a constant for vertically-isothermal disks, 
+%and we can directly apply previous results.
 
@@
 \begin{align}\label{beta_mmsn}
-\beta =
-\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)r_\mathrm{AU}^{-57/14}\left(\frac{8.3\times10^{-9}}{\hat{\kappa}_d^2\hat{T}^4\hat{\Sigma}^2}r_\mathrm{AU}^{33/7}+\frac{1}{\khat^2}\right).
+  \beta(z; \khat) =
+  &\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)r_\mathrm{AU}^{-57/14}\notag\\
+&\times\left[\frac{8.3\times10^{-9}}{\hat{\kappa}_d^2\hat{T}^4\hat{\Sigma}^2}r_\mathrm{AU}^{33/7}+\frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right].     
 \end{align}
@@
 $\hat{\Sigma}=\hat{\kappa}_d=\hat{T}=1$ as the Minimum Mass Solar
-Nebulae (MMSN).  We also evaluate the criterion Eq. \ref{iso_vsi_cond} for the fiducial
+Nebulae (MMSN). The physical thermal relaxation timescale
+$\beta$ is height-dependent, whilst for analytical discussion we
+assumed it is constant. Nevertheless, it is instructive to evaluate
+and compare with the criterion Eq. \ref{iso_vsi_cond} for the fiducial  
 disk model,  
@@
 
-In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn} and
-Eq. \ref{bcrit_mmsn} with with $\mu =2.33$ and $\gamma=1.4$.  We set
-the scaling $\hat{T}=\hat{\Sigma}=1$ and consider three opacity scales:
-$\hat{\kappa}_d=0.1, \,1 $ and $10$.
+In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn}  with $\mu
+=2.33$ and $\gamma=1.4$.  We set the scaling $\hat{T}=\hat{\Sigma}=1$
+and consider three opacity scales: $\hat{\kappa}_d=0.1, \,1 $ and $10$
+at two heights: $z=0$ and $z=3H$. We also plot Eq. \ref{bcrit_mmsn} in
+each panel for comparison.
 
-The figure suggests that the fundamental VSI 
-operates in the outer parts of the MMSN at radii on 
-the order of tens of AU ($\hat{\kappa}_d=1$, middle panel). 
-However, the thermal relaxation time $\beta$
-is not much shorter than $\beta_\mathrm{crit}$, so the VSI may be 
-weak. Lowering the opacity shifts the curves
-inward, but raises also them. In fact, lowering the opacity by an
-order of magnitude renders the fiducial  disk stable to the fundametal
-mode for the range of wavenumbers considered ($\hat{\kappa}_d=0.1$,
-upper panel). Raising the opacity, on
-the other hand, decreases the thermal timescales, but also shifts the
-curves outward, so the fundamental mode is restricted to the outer
-disk ($\hat{\kappa}_d=10$, lower panel).
+Consider first the MMSN shown in the middle panels of
+Fig. \ref{mmsn_bcrit_bcool}. The critical thermal timescale,
+Eq. \ref{bcrit_mmsn}, as derived from our analytical discussion, is
+most useful to assess the VSI in the optically thin regime (where the
+curves overlap), since in this limit $\beta$ does not depend on
+height. We can therefore conclude that the VSI can operate at tens of
+AU in the MMSN.
+
+In the inner few AU of the MMSN, however, $\beta$  has a
+strong dependence on height because the disk is optically thick. Given
+a fixed $\khat$, one can have $\beta > \beta_\mathrm{crit}$ at the
+midplane, but eventually $\beta<\beta_\mathrm{crit}$ with increasing
+height. Our analytical criteria is not valid for variable $\beta$, but
+given that the VSI depends on vertical shear, an effect that becomes
+important away from the disk midplane, the inner disk may still be
+able to develop the VSI, but it is likely restricted to the disk
+surface.
+
+Now consider lowering the opacity, as shown in the top panels of
+Fig. \ref{mmsn_bcrit_bcool}. The curves shift inward, but raises also
+them. In fact, lowering the opacity by an order of magnitude renders
+$\beta > \beta_\mathrm{crit}$ even away from the midplane. This
+suggests that fiducial  disk stable to the fundametal mode for the
+range of wavenumbers considered.
+
+Raising the opacity, as shown in the lower panels of
+Fig. \ref{mmsn_bcrit_bcool}, decreases the thermal timescales, but
+also shifts the curves outward, so the fundamental mode is restricted
+to larger radii.
 
@@
 
-\begin{figure}
-  \includegraphics[scale=.43,clip=true,trim=0cm 1.8cm 0cm 0cm]{figures/bcrit_mmsn0d1} 
-  \includegraphics[scale=.43,clip=true,trim=0cm 1.8cm 0cm 1cm]{figures/bcrit_mmsn}  
-  \includegraphics[scale=.43,clip=true,trim=0cm 0cm 0cm 1cm]{figures/bcrit_mmsn10} 
-  \caption{Dimensionless thermal relaxation timescales $\beta$ in a
-    fiducial protoplanetary disk model. Eq. \ref{beta_mmsn} is plotted
+\begin{figure*}
+  \includegraphics[scale=.47,clip=true,trim=0cm 1.8cm 0cm
+  0cm]{figures/bcrit_mmsn_kap0d1_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm 1.8cm 0cm
+  0cm]{figures/bcrit_mmsn_kap0d1_z3}\\ 
+  \includegraphics[scale=.47,clip=true,trim=0cm 1.8cm 0cm
+  1cm]{figures/bcrit_mmsn_kap1_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm
+  1.8cm 0cm 1cm]{figures/bcrit_mmsn_kap1_z3}\\
+  \includegraphics[scale=.47,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/bcrit_mmsn_kap10_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm 0cm 0cm
+  1cm]{figures/bcrit_mmsn_kap10_z3} 
+  \caption{Dimensionless thermal relaxation timescales $\beta$,
+    evaluated at the midplane (left) and at $z=3H$ (right) in the
+    fiducial protoplanetary disk. Eq. \ref{beta_mmsn} is plotted  
     for three values of the 
@@
     \label{mmsn_bcrit_bcool}}   
-\end{figure}
+\end{figure*}
 
@@
 
-\subsection{Numerical example}\label{mmsn_example}
-Consider  the MMSN at $r=50\mathrm{AU}$. Eq. \ref{mmsn_epsilon} gives
-$\epsilon \simeq 
-0.067$ which implies, using Eq. \ref{bcrit_mmsn}, a critical 
-relaxation timescale of  
-\begin{align*}
-  \beta_\mathrm{crit} \simeq 0.072. 
-\end{align*} 
-To estimate a thermal relaxation timescale, we set $\khat=30$ 
-in Eq. \ref{beta_mmsn} to obtain 
-\begin{align*}
-  \beta \simeq 0.049.
-\end{align*}
-(In fact, Fig. \ref{mmsn_bcrit_bcool} shows that $\beta$  is not
-sensitive to $\khat$ at $r=50\mathrm{AU}$ for $\khat\gg1$.)
-The MMSN is unstable to the fundamental VSI, but not
-strongly.
+\subsection{Numerical examples}\label{mmsn_example}
+Here we solve the linear stability problem with the
+thermal relaxation function defined by Eq. \ref{beta_mmsn}. We
+consider the MMSN at $r=50$AU and $r=2$AU.
 
-In Fig. \ref{mmsn_growth}, we calculate the fundamental
-VSI growth rates using the MMSN disk parameters at $r=50\mathrm{AU}$
-and the scale-dependent thermal relaxation $\beta(\khat)$ given by
-Eq. \ref{beta_mmsn} in our linear code.  In this case we found a 
-larger vertical domain with $\zmax=7\epsilon r$ was needed to capture
-modes with $\khat\gtrsim60$, which have perturbations
-concentrated near the vertical boundaries. However, modes with
-$\khat\lesssim 60$ were found to be insensitive to boundary
-conditions.
-
-The most unstable mode occurs at $\khat\simeq30$ with growth rate 
-\begin{align*}
-  \nu =0.058\epsilon\Omega_k = 3.9\times10^{-3}\Omega_k.  
-\end{align*}
-This corresponds to a characteristic growth time $\sim
-40$ orbits. However, since growth rates are weakly dependent on $\khat$ for
-$10\lesssim \khat \lesssim 60$, this may be taken as a representative growth
-time of the fundamental VSI in the MMSN at $r=50\mathrm{AU}$. The
-eigenfunction for $\khat=30$ is shown in Fig. \ref{mmsn_eigenW}.
+We first compare $\beta$ at these radii in Fig. \ref{beta_compare}. In
+the optically thin outer disk, $\beta$ is independent of $z$ and
+$\beta<\beta_\mathrm{crit}$ throughout the disk column. We thus expect
+the VSI to operate at $r=50AU$. In the inner disk, $\beta$ depends on 
+height such that $\beta\lesssim\beta_\mathrm{crit}$ only for 
+$|z|\gtrsim 2H$, which suggest that the VSI will be restricted
+restricted to the upper and lower layers in the inner disk.
 
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/growth_mmsn_50AU} 
-  \caption{VSI growth rates in the MMSN at
-    $r=50\mathrm{AU}$. The thermal relaxation timescale $\beta(\khat)$
-    used for this calculation is given by
-    Eq. \ref{beta_mmsn} and is scale-dependent. 
-    \label{mmsn_growth}}  
+  0cm]{figures/beta_compare} 
+  \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
+    (thick solid) and $r=2$AU (thick dotted) for perturbation
+    wavenumbers $\khat=10,\,100$, respectively. The corresponding
+    horizontal lines are the critical thermal relaxation timescales
+    derived in linear theory. 
+    \label{beta_compare}}  
 \end{figure}
 
-% Fig. \ref{mmsn_eigenW} show several eigenfunctions for the modes in
-% Fig. \ref{mmsn_growth}. As $\khat$ is increased, the disturbance
-% become more concentrated near the vertical boundaries. Thus, modes
-% with very large $\khat$ may not be important in practice because they
-% predominantly affect the disk atmosphere, which contains little mass.
+In Fig. \ref{eigen_compare_mmsn} we compute growth rates of the
+fundamental mode at $r=50$AU and $r=2$AU. At $r=50$AU growth rates are
+insensitive to $\khat$ for $1\lesssim \khat\lesssim 60$, and
+corresponds to growth time $\simeq 40$ local orbits. In the inner disk
+at $r=2$AU, we only found unstable modes for $\khat\gtrsim50$. To
+obtain a similar growth timescale at $r=2$AU requires $\khat\sim
+100$. This is consistent with the expectation that, as we proceed into
+the inner disk, it bcomes optically thick, so one needs to consider
+smaller perturbation wavelengths to obtain a sufficiently short
+thermal timescale.
+
+% In both cases we find that increasing $\khat$ increases the
+% perturbation amplitudes near the disk surface. Such modes may be
+% affected by boundary conditions so their robustness is unclear.
 
 \begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/eigen_compare_grow_mmsn} 
+  \caption{Growth rates of the fundamental VSI mode in the MMSN at
+    $r=50$AU (solid) and $r=2$AU (dotted). 
+    \label{eigen_compare_mmsn}}  
+\end{figure}
+
+Finally, we compare eigenfunctions for the vertical velocity at these
+radii in Fig. \ref{mmsn_eigenvz}. We compare cases with similar growth
+rates. At $r=50$AU we have vertical motion throughout the disk column
+because the thermal timescale is less than critical. By contrast, at
+$r=2$AU there is little vertical motion in $|z|\lesssim 2H$. 
+Considering this result with Fig. \ref{beta_compare} suggest 
+that the critical thermal timescale gives an indication of where the 
+VSI operates even when $\beta$ depends on $z$.
+
+%can always go to sufficiently large heights to get beta < beta crit,
+%but getting close to boundaries/lower densities
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvectorvz_mmsn_50AU}
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
-  0cm]{figures/eigenvectorW_mmsnkx30.ps}
-  % \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  % 0cm]{figures/eigenvectorW_mmsnkx10.ps} 
-  % \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  % 0cm]{figures/eigenvectorW_mmsnkx50.ps} 
-  \caption{Fundamental VSI mode in the MMSN at 
-    $r=50\mathrm{AU}$. The eigenfunction $W$ for $\khat=30$ is shown. The real (imaginary)
-    part of $W$ is plotted as the solid (dotted)
-    lines. The thermal relaxation timescale $\beta(\khat)$ is given by
-    Eq. \ref{beta_mmsn} and is scale-dependent. \label{mmsn_eigenW}}    
+  0cm]{figures/eigenvectorvz_mmsn_2AU}
+  \caption{Comparison of vertical velocity perturbations of 
+    fundamental VSI modes of similar growth rates in the MMSN at
+    $r=50$AU (top) and $r=2$AU (bottom). The real (imaginary) part of
+    $\delta v_z$ is shown as 
+    the solid (dotted) curves. The eigenfunction has been normalized
+    such that $\imag(\delta v_z)=0$ at the vertical boundaries, and
+    $\mathrm{max}|\delta v_z|=1$. \label{mmsn_eigenvz}}      
 \end{figure}

### 2015-03-01 application.tex
regular update

@@
 
-We remark that it is possible, in principle, to satisfy $\beta <
-\beta_\mathrm{crit}$ by choosing a sufficiently large $\khat$,
-equivalently a sufficiently small perturbation lengthscales. We
+We remark that it is possible to satisfy $\beta <
+\beta_\mathrm{crit}$ by choosing a sufficiently large $\khat$, or 
+equivalently a sufficiently small perturbation lengthscale. We
 illustrate this by plotting at each radius the value of $\khat$ such
-that $\beta = \beta_\mathrm{crit}$. We see that in order for the
-fundamental VSI to operate in the inner disk (few AU), one needs to
-consider large values of $\khat$.
+that $\beta = \beta_\mathrm{crit}$. In the inner few AU of the MMSN,
+one needs to consider large values of $\khat$ and/or regions away from
+the disk midplane.
 
-However, our linear
-calculations indicate growth rates eventually decay as one increases
-$\khat$. This, together with our finding of high $\khat$ modes may be
-affected by vertical boundary conditions, suggest the VSI does not
-operate in the inner few AU in protoplanetary disks where thermal
-relaxation is entirely due to radiation \citep[cf.][]{stoll14}.
+% However, our linear
+% calculations indicate growth rates eventually decay as one increases
+% $\khat$. This, together with our finding of high $\khat$ modes may be
+% affected by vertical boundary conditions, suggest the VSI does not
+% operate in the inner few AU in protoplanetary disks where thermal
+% relaxation is entirely due to radiation \citep[cf.][]{stoll14}.
 
@@
   \includegraphics[width=\linewidth]{figures/bcrit_mink} 
-  \caption{The minimum value of the perturbation wavenumber $\khat$ in
+  \caption{The minimum perturbation wavenumber $\khat$ in
     the MMSN such that the associated dimensionless thermal
-    relaxation time  $\beta$ is  less than the critical timescale
-    $\beta_\mathrm{crit}$. 
+    relaxation time  $\beta$ at $z=0$ (thick lines) and $z=3H$ (thin
+    lines) is less than the critical timescale $\beta_\mathrm{crit}$.  
     \label{mmsn_bcrit_bcool_mink}}   
@@
 $\beta<\beta_\mathrm{crit}$ throughout the disk column. We thus expect
-the VSI to operate at $r=50AU$. In the inner disk, $\beta$ depends on 
+the VSI to operate at $r=50$AU. In the inner disk, $\beta$ depends on 
 height such that $\beta\lesssim\beta_\mathrm{crit}$ only for 
@@
 $r=2$AU there is little vertical motion in $|z|\lesssim 2H$. 
-Considering this result with Fig. \ref{beta_compare} suggest 
-that the critical thermal timescale gives an indication of where the 
-VSI operates even when $\beta$ depends on $z$.
+Considering this result with Fig. \ref{beta_compare} suggests, but does
+not prove, that the critical thermal timescale gives an indication of
+where the VSI operates even when $\beta$ depends on $z$.
 

### 2015-03-01 numerical.tex
new subsection growth time vs rad for mmsn
regular update

@@
 expected for the fundamental mode according to explicit solutions
-developed in \S\ref{iso_poly}.
+developed in \S\ref{iso_poly}. For completeness we show the
+corresponding eigenfunctions in vertical velocity in
+Fig. \ref{lowfreq_eigenfuncvz}, which show that the fundamental mode
+consists of zero nodes in $\delta v_z$.
 
-However, for $\khat\lesssim 10$, e.g. $\khat=1$ in 
+For $\khat\lesssim 10$, e.g. $\khat=1$ in 
 Fig. \ref{lowfreq_eigenfunc}, the perturbation magnitude near the
@@
   \includegraphics[width=\linewidth]{figures/eigenvector_iso_kx60}
-  \caption{Eigenfunction of the fundamental VSI for perturbation
+  \caption{Eigenfunction $W$ of the fundamental VSI for perturbation
     wavenumber $\khat=1$ (top) and $\khat=20\pi$ (bottom). The real 
-    (imaginary) parts of $W$ are plotted as solid (dotted) lines. 
+    (imaginary) parts of $W$ are plotted as solid (dotted) lines. The
+    eigenfunction has been normalized such that $\imag(W)=0$ at the
+    vertical boundaries and $\mathrm{max}|W|=1$.
     \label{lowfreq_eigenfunc}
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvectorvz_iso_kx1} 
+  \includegraphics[width=\linewidth]{figures/eigenvectorvz_iso_kx60}
+  \caption{Same as Fig. \ref{lowfreq_eigenfunc} but for the vertical
+    velocity eigenfunctions.
+    \label{lowfreq_eigenfuncvz}
   }
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvectorW_beta0d01} 
+  0cm]{figures/eigenvectorvz_beta0d01} 
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/eigenvectorW_beta0d1} 
-  \caption{Numerically-calculated eigenfunction of the fundamental VSI
+  0cm]{figures/eigenvectorvz_beta0d1} 
+  \caption{Numerically-calculated fundamental VSI
     mode in a nearly vertically isothermal disk ($\Gamma=1.011$) with
@@
     (top) and $\beta=0.1$ (bottom). The
-    perturbation radial wavenumber is $\khat=30$. The
-    real (imaginary) part of $W$ is plotted as the solid (dotted)
-    line.
+    real (imaginary) part of the vertical velocity perturbation
+    $\delta v_z$ is plotted as the solid (dotted)
+    line. The eigenfunction has been normalized such that
+    $\imag(\delta v_z)=0$ at the vertical boundaries, and
+    $\mathrm{max}|\delta v_z|=1$.
     \label{relax_eigenW_num}}  
@@
 (Fig. \ref{bcrit_compare1}). For $\gamma=1.7,\,2.0,\,2.5$, growth
-rates reach zero for  $\beta\simeq0.17,\,0.083,\,0.045$, respectively.
+rates reach zero at  $\beta\simeq0.17,\,0.083,\,0.045$, respectively.
 Interestingly, these values are equal to

### 2015-03-02 application.tex
updating

@@
 \end{align} 
-We refer to the case with the scalings
+We refer to the case with $\gamma=1.4$, $\mu=2.33$ and the scalings
 $\hat{\Sigma}=\hat{\kappa}_d=\hat{T}=1$ as the Minimum Mass Solar
-Nebulae (MMSN). The physical thermal relaxation timescale
-$\beta$ is height-dependent, whilst for analytical discussion we
-assumed it is constant. Nevertheless, it is instructive to evaluate
-and compare with the criterion Eq. \ref{iso_vsi_cond} for the fiducial  
-disk model,  
+Nebulae (MMSN) and $\beta=\beta_\mathrm{MMSN}$, 
+\begin{align}\label{beta_mmsn_simp}
+  \beta_\mathrm{MMSN} = 3.9\times10^{-3} r_\mathrm{AU}^{9/14}\left( 1 +
+    1.9\times10^7\hat{\rho}^2r_\mathrm{AU}^{-33/7}\khat^{-2}\right). 
+\end{align}
+% where
+% \begin{align}
+%   \
+% %  \frac{3.7\times10^{-3}}{\mu\left(\gamma-1\right)\hat{\kappa}_d\hat{T}^5}, 
+% \end{align}
+% and
+% \begin{align}
+%   \khat_0 = 4.4\times10^{3}\hat{\kappa}_d\hat{T}^2\hat{\Sigma}.
+% \end{align}
+
+% The physical thermal relaxation timescale $\beta$ is height-dependent,
+% whilst for analytical discussion we assumed it is
+% constant. Nevertheless,
+It is instructive to compare the above thermal timescales
+with the criterion Eq. \ref{iso_vsi_cond} evaluated for the fiducial disk model,    
 \begin{align}\label{bcrit_mmsn}
@@
 fundamental VSI when $\khat\gg 1$. %For $\khat\to0$, this
-%becomes a sufficient condition (see Fig. \ref{relax_bcrit} and
+% becomes a sufficient condition (see Fig. \ref{relax_bcrit} and
 %\S\ref{iso_vsi_beta_crit}).
@@
 at two heights: $z=0$ and $z=3H$. We also plot Eq. \ref{bcrit_mmsn} in
-each panel for comparison.
+each panel.
 
@@
 curves overlap), since in this limit $\beta$ does not depend on
-height. We can therefore conclude that the VSI can operate at tens of
-AU in the MMSN.
+height. We thus infer that the VSI can operate at tens of AU in the
+MMSN.
 
-In the inner few AU of the MMSN, however, $\beta$  has a
-strong dependence on height because the disk is optically thick. Given
-a fixed $\khat$, one can have $\beta > \beta_\mathrm{crit}$ at the
-midplane, but eventually $\beta<\beta_\mathrm{crit}$ with increasing
-height. Our analytical criteria is not valid for variable $\beta$, but
-given that the VSI depends on vertical shear, an effect that becomes
-important away from the disk midplane, the inner disk may still be
-able to develop the VSI, but it is likely restricted to the disk
-surface.
+In the inner few AU of the MMSN, however, $\beta$ depends on $z$. 
+Given a fixed $\khat$, one can have $\beta > \beta_\mathrm{crit}$ at
+the optically thick midplane, but with increasing height the disk
+becomes optically thin and $\beta<\beta_\mathrm{crit}$. Our analytical
+criteria is not valid for variable $\beta$, and we must turn to
+numerical solutions of the linear to determine whether or not the
+fundamental mode operates. We do this in the next
+section.
 
@@
 them. In fact, lowering the opacity by an order of magnitude renders
-$\beta > \beta_\mathrm{crit}$ even away from the midplane. This
-suggests that fiducial  disk stable to the fundametal mode for the
-range of wavenumbers considered.
+$\beta > \beta_\mathrm{crit}$ even away from the midplane. This makes
+disk stable to the fundametal mode. Raising the opacity, as shown in
+the lower panels of Fig. \ref{mmsn_bcrit_bcool}, decreases the thermal
+timescales, but shifts the curves outward, so the fundamental mode is
+restricted to larger radii.
 
-Raising the opacity, as shown in the lower panels of
-Fig. \ref{mmsn_bcrit_bcool}, decreases the thermal timescales, but
-also shifts the curves outward, so the fundamental mode is restricted
-to larger radii.
-
+%but
+%given that the VSI depends on vertical shear, an effect that becomes
+%important away from the disk midplane, the inner disk may still be
+%able to develop the VSI, but it is likely restricted to the disk
+%surface.   
 %We remark that it is possible to satisfy $\beta<\beta_\mathrm{crit}$
@@
 
-We remark that it is possible to satisfy $\beta <
-\beta_\mathrm{crit}$ by choosing a sufficiently large $\khat$, or 
-equivalently a sufficiently small perturbation lengthscale. We
+We remark that it is possible to satisfy $\beta < \beta_\mathrm{crit}$
+by choosing a sufficiently large $\khat$. We
 illustrate this by plotting at each radius the value of $\khat$ such
-that $\beta = \beta_\mathrm{crit}$. In the inner few AU of the MMSN,
-one needs to consider large values of $\khat$ and/or regions away from
-the disk midplane.
+that $\beta = \beta_\mathrm{crit}$. We thus expect that the VSI,
+should it occur in the inner disk, would only occur on very small
+scales. It is not clear if such modes are robust (see below).
 
@@
 
-\subsection{Numerical examples}\label{mmsn_example}
-Here we solve the linear stability problem with the
-thermal relaxation function defined by Eq. \ref{beta_mmsn}. We
-consider the MMSN at $r=50$AU and $r=2$AU.
+\subsection{VSI in the MMSN}
+We have solved the linear problem in the MMSN with the scale and
+height dependent thermal relaxation function defined by
+Eq. \ref{beta_mmsn_simp}.  We use a vertical
+domain $z_\mathrm{max}=5H$. In Fig. \ref{mmsn_overall} we plot growth
+timescales $t_\mathrm{grow} \equiv 1/\nu$ associated with the
+fundamental mode in units of the local orbital time
+$P_\mathrm{orb}\equiv 2\pi/\Omega_k$.
 
-We first compare $\beta$ at these radii in Fig. \ref{beta_compare}. In
-the optically thin outer disk, $\beta$ is independent of $z$ and
-$\beta<\beta_\mathrm{crit}$ throughout the disk column. We thus expect
-the VSI to operate at $r=50$AU. In the inner disk, $\beta$ depends on 
-height such that $\beta\lesssim\beta_\mathrm{crit}$ only for 
-$|z|\gtrsim 2H$, which suggest that the VSI will be restricted
-restricted to the upper and lower layers in the inner disk.
+In fact, for $1\lesssim \khat \lesssim 30$ we either only find the
+fundamental mode, or it is the most unstable. The associated 
+growth rates are also insensitive to $z_\mathrm{max}$. However, for
+large $\khat$ and/or $z_\mathrm{max}$, e.g. $\khat=50$ or 
+$z_\mathrm{max}=10H$, we also found modes with shorter growth times 
+than those shown in Fig. \ref{mmsn_overall}, but these disturbances are 
+concentrated near the vertical boundaries
+\citep{nelson13,mcnally14}. Such modes may not be robust because    
+they are affected by boundary conditions. Furthermore, modes with   
+large $\khat$ are likely subject to viscous decay in a real disk.
+
+We therefore take Fig. \ref{mmsn_overall} as a representation of the
+VSI in the standard MMSN. We see that growth timescales are $\sim
+30$---40 orbits, and the VSI operates from a few AU to tens of
+AU. Importantly, it is necessary to consider smaller scales as $r$
+decreases as the disk generally becomes optically thick.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/eigen_compare_grow.ps}
+  \caption{Characteristic growth times of the fundamental VSI mode in
+    the MMSN.  
+    \label{mmsn_overall}}    
+\end{figure}
+
+\subsubsection{Comparing the VSI in the inner and outer disk} 
+Here we compare the VSI at $r=50$AU and $r=5$AU. We first show in
+Fig. \ref{beta_compare}  the thermal timescale $\beta$ for a
+perturbation wavenumber $\khat=30$. For comparison we also plot
+$\beta_\mathrm{crit}$ at these radii.
 
@@
   \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
-    (thick solid) and $r=2$AU (thick dotted) for perturbation
-    wavenumbers $\khat=10,\,100$, respectively. The corresponding
-    horizontal lines are the critical thermal relaxation timescales
-    derived in linear theory. 
+    (thick solid) and $r=5$AU (thick dotted) for $\khat=30$. The
+    corresponding thin horizontal lines are the critical thermal
+    relaxation timescales derived in linear theory. 
     \label{beta_compare}}
@@
 
-In Fig. \ref{eigen_compare_mmsn} we compute growth rates of the
-fundamental mode at $r=50$AU and $r=2$AU. At $r=50$AU growth rates are
-insensitive to $\khat$ for $1\lesssim \khat\lesssim 60$, and
-corresponds to growth time $\simeq 40$ local orbits. In the inner disk
-at $r=2$AU, we only found unstable modes for $\khat\gtrsim50$. To
-obtain a similar growth timescale at $r=2$AU requires $\khat\sim
-100$. This is consistent with the expectation that, as we proceed into
-the inner disk, it bcomes optically thick, so one needs to consider
-smaller perturbation wavelengths to obtain a sufficiently short
-thermal timescale.
+In the outer disk ($r=50$AU) the disk is optically thin, $\beta$ 
+does not depend on height and $\beta <
+\beta_\mathrm{crit}$. Accordingly we expect, and find, the VSI to
+operate at $r=50$AU. In the inner disk ($r=5$AU) the disk is optically
+thick near the midplane and becomes optically thin away from it, such
+that $\beta < \beta_\mathrm{crit}$ only for $|z|\gtrsim1.25H$. In
+fact, in this case  we also find the VSI to operate, but with
+significantly reduced growth rate (as reflected in
+Fig. \ref{mmsn_overall}).
 
-% In both cases we find that increasing $\khat$ increases the
-% perturbation amplitudes near the disk surface. Such modes may be
-% affected by boundary conditions so their robustness is unclear.
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/eigen_compare_grow_mmsn} 
-  \caption{Growth rates of the fundamental VSI mode in the MMSN at
-    $r=50$AU (solid) and $r=2$AU (dotted). 
-    \label{eigen_compare_mmsn}}  
-\end{figure}
-
-Finally, we compare eigenfunctions for the vertical velocity at these
-radii in Fig. \ref{mmsn_eigenvz}. We compare cases with similar growth
-rates. At $r=50$AU we have vertical motion throughout the disk column
-because the thermal timescale is less than critical. By contrast, at
-$r=2$AU there is little vertical motion in $|z|\lesssim 2H$. 
-Considering this result with Fig. \ref{beta_compare} suggests, but does
-not prove, that the critical thermal timescale gives an indication of
-where the VSI operates even when $\beta$ depends on $z$.
-
-%can always go to sufficiently large heights to get beta < beta crit,
-%but getting close to boundaries/lower densities
+The corresponding vertical velocity perturbations are shown in 
+Fig. \ref{mmsn_eigenvz}. Although at $r=5$AU $\beta$ depends on $z$,
+the disturbance still occupies the fluid column. This is probably due
+to the vertically global nature of the fundamental mode. The fact that
+we obtain the VSI despite $\beta > \beta_\mathrm{crit}$ near the
+midplane suggests, but does not prove, that the VSI can operate
+provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
+fraction of the vertical domain. However, the growth rates are much
+smaller than the case with $\beta < \beta_\mathrm{crit}$ for all $z$.
 
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
-  0cm]{figures/eigenvectorvz_mmsn_2AU}
-  \caption{Comparison of vertical velocity perturbations of 
-    fundamental VSI modes of similar growth rates in the MMSN at
-    $r=50$AU (top) and $r=2$AU (bottom). The real (imaginary) part of
+  0cm]{figures/eigenvectorvz_mmsn_5AU}
+  \caption{Vertical velocity perturbations of the VSI in the MMSN at 
+    $r=50$AU (top) and $r=5$AU (bottom). The real (imaginary) part of 
     $\delta v_z$ is shown as
@@
 \end{figure}
-% only find one unstable mode -> close to marginal 
-% larger kx
-
-\subsection{VSI in the outer parts of the MMSN}
-To obtain an overall picture of the relevance of the VSI to the MMSN,
-we calculate characteristic growth times $t_\mathrm{grow} = 1/\nu$ of 
-the fundamental mode in units of the local Keplerian orbital period
-$P_\mathrm{orb}= 2\pi/\Omega_k$. The results are shown in 
-Fig. \ref{mmsn_overall} for perturbation wavenumbers $5\lesssim
-\khat\lesssim 50$. We consider this range because for  
-$\khat=O(1)$ or $\khat=O(10^2)$ we found eigenfunctions have large
-amplitudes near the vertical boundaries, so such modes may be affected
-by boundary conditions.
-
-In the MMSN the fundamental VSI operates at tens of AU, with
-characteristic growth times of tens of orbits. %In our example, the VSI
-% is most favored at $r\simeq 20\mathrm{AU}$ with a growth time
-% $t_\mathrm{grow}\simeq 30$ orbits.
-Fig. \ref{mmsn_overall} indicates that with decreasing radius, the VSI
-shifts to smaller scales (larger $\khat$), since the disk becomes optically
-thick so only small-scale disturbances have sufficiently short thermal
-relaxation timescales. At larger radii ($r\gtrsim 
-40\mathrm{AU}$), the disk is optically thin so growth times do not depend strongly on   
-$\khat$, and the VSI can operate on larger scales (smaller $\khat$).  
-However, approaching 
-$r\simeq 100\mathrm{AU}$ growth times become long at $O(10^2)$ local orbits,
-so the VSI may not be dynamically important in the outermost parts of
-the MMSN.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/eigen_compare_grow.ps}
-  \caption{Characteristic growth times of the fundamental VSI mode in
-    the MMSN for a range of perturbation wavenumbers.  
-    \label{mmsn_overall}}    
-\end{figure}
-
- % \subsection{Vertically-dependent thermal relaxation} 
-% Note that for a vertically isothermal disk and opacity $\kappa_d(T)$, 
-% we can restore the $z$ dependence in $\beta$ (due to the $\rho$
-% dependence in  $\eta$, see Eq. \ref{eta_def}) by making the transformation
-% $\beta(\khat)\to\beta(\khat/\sqrt{2\pi}\hat{\rho})\equiv 
-% \tilde{\beta}(\khat,z)$ in  Eq. \ref{real_beta}, where
-% $\hat{\rho}=\rho/\rho(z=0)$.  We have repeated the example
-% calculation in \S\ref{mmsn_example} with $\beta\to\tilde{\beta}$, but
-% found similar results.
-
-%We find growth rates
-%were reduced, but at most by $\sim 15\%$ at $\khat=1$. Eigenfunctions
-%were similar.
-
-%at most by 15%

### 2015-03-02 application.tex
revising

@@
 criteria is not valid for variable $\beta$, and we must turn to
-numerical solutions of the linear to determine whether or not the
-fundamental mode operates. We do this in the next
-section.
+numerical solutions of the linear problem to determine whether or not
+the VSI operates. We do this in the next section.
 
@@
 that $\beta = \beta_\mathrm{crit}$. We thus expect that the VSI,
-should it occur in the inner disk, would only occur on very small
+should it develop in the inner disk, would only occur on very small
 scales. It is not clear if such modes are robust (see below).

### 2015-03-02 paper.tex
updated all figs
revising

@@
 \begin{abstract}
-  Baroclinic astrophysical disks with equilibrium rotation profiles
-  $\Omega$ that depends on the distance away from the disk midplane are
-  potentially unstable to the Vertical Shear Instability (VSI). It is
-  is important to understand the linear instability before
-  considering the VSI as a hydrodynamic route to turbulent angular
-  momentum transport in accretion disks. We study the linear,
-  axisymmetric VSI in radially-local but vertically-global models
-  of accretion disks. We include an energy equation with thermal
-  relaxation on a characteristic timescale $t_c=\beta\Omega_k^{-1}$,
-  where $\beta$ is a dimensionless parameter and $\Omega_k$ is the
-  Keplerian rotation frequency, and explore how the VSI depends on the
-  thermodynamic response  of the disk by varying $\beta$. We find
-  increasing $\beta$ away from zero dramatically reduces VSI growth
-  rates because this generally introduces a strongly stabilizing
-  vertical entropy gradient. Our linear calculations are consistent
-  with previous hydrodynamic simulations, which indicates the VSI 
-  requires a short thermal relaxation timescale. For vertically  
-  isothermal disks, we quantify `fast'
+  A robust hydrodynamic mechanism for turbulent angular momentum
+  transport is important for cold accretion disk such as 
+  protoplanetary disks. The vertical shear instability (VSI) has 
+  recently been suggested to provide such a pathway, since it operates in
+  disks with rotation profiles that depends on the height from the
+  disk  midplane, which is generally expected in 
+  astrophysical disks. However, the VSI requires a short thermal
+  timescale. A quantitative assessment of the thermodynamic
+  requirement for the VSI is thus desirable for application to
+  realistic disks. To this end,  
+  we generalize the linear, axisymmetric VSI in radially local but
+  vertically global models of accretion disks to include an energy
+  equation with a thermal relaxation term that restores the disk
+  temperature $T$ over a characteristic timescale
+  $t_c=\beta\Omega_k^{-1}$, where $\beta$ is a dimensionless parameter
+  and $\Omega_k$ is the Keplerian rotation frequency. We explore how
+  the VSI depends on the thermodynamic response  of the disk by
+  varying $\beta$. We find increasing $\beta$ away from zero
+  dramatically reduces VSI growth rates because this generally
+  introduces a strongly stabilizing vertical entropy gradient. Our
+  linear calculations are consistent with previous hydrodynamic
+  simulations. For vertically isothermal disks, we quantify `fast' 
   thermal relaxation as $\beta < \epsilon|q|/(\gamma-1)$, where
-  $\epsilon$ is the disk aspect-ratio, $q$ is the radial power-law
-  index for the equilibrium temperature profile (which gives rise to vertical 
-  shear) and $\gamma$ is the adiabatic index. We apply our linear
+  $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  ($r$
+  being the cylindrical radius from the central object), and 
+  $\gamma$ is the adiabatic index. We apply our linear
   results to a disk model of the Minimum Mass Solar Nebulae (MMSN), and
@@
   characteristic growth times of tens of local orbital periods. This 
-  suggests the VSI is  dynamically important in the outer parts of protoplanetary disks. 
+  suggests the VSI is dynamically important in the outer parts of
+  protoplanetary disks.  
 \end{abstract}
@@
 solve the linearized fluid equations in the radially-local
-approximation. We find basic agreement with our analysis, including
+approximation, but relax all other assumptions made in our analytical approach.   
+We find basic agreement with our analysis, including
 the aforementioned characteristic thermal timescale. In

### 2015-03-02 linear_iso.tex
updating main part of paper

@@
 We begin with a discussion of vertically isothermal disks
-where $\Gamma=1$, then $c_s$ is a constant. We  
-first examine isothermal perturbations ($\beta = 0$) before
-considering the effect of finite thermal relaxation. 
-In order to facilitate a fully analytic discussion, we introduce a number
-of approximations below that allows a simple governing equation to be
-derived and compared with previous works.
+where $\Gamma=1$, then $c_s$ is a constant in the radially-local
+approximation. We  first examine isothermal perturbations ($\beta =
+0$) before considering the effect of finite thermal relaxation. 
+In order to facilitate a fully analytic discussion, we introduce a
+number of approximations below that allows a simple governing equation
+to be derived and compared with previous works.
 
-\subsection{The linear problem in the vertically global
+\subsection{Relation to the vertically global
   shearing box} 
@@
 shearing box (VGSB) formalism developed by \cite{mcnally14}. The VGSB
-is a generalization of the standard shearing box \citep{goldreich65}
-by allowing the shear flow in the shearing box to be height-dependent.
+is an extension of the standard shearing box \citep{goldreich65} to
+allow the shear flow in the shearing box to be height-dependent.  
+The difference between the radially-local approximation and the VGSB is
+that the radial disk structure is ignored in the latter.  That is, the
+equilibrium VGSB is homogeneous in the radial direction, as in the
+standard shearing box.
 
-The linear problem in the VGSB framework can be derived from 
-Eq. \ref{lin_mass}---\ref{lin_energy} by setting $\hat{g}_c=0$, since 
-the VGSB, like the standard shearing box, is homogeneous in the radial 
-direction in equilibrium. The resulting linear problem is very similar
+Thus, the linear problem in the VGSB can be derived from
+Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
+$\hat{g}_c=0$.  
+The resulting linear problem is very similar
 to that for local axisymmetric waves in accretion disks already
@@
 However, we find that for perturbations with small radial lengthscales
-($k_xH_\mathrm{iso}\gg 1$) and fast thermal relaxation ($\beta\ll1$), which
+($k_xH\gg 1$) and fast thermal relaxation ($\beta\ll1$), which
 characterizes the VSI, the VGSB framework is adequate for the purpose
@@
 
-% In effect, we are studying dynamics in the standard shearing box
-% subject to an additional azimuthal force $-v_zr_0 d \Omega_0/dz$.
-
-% $d\Omega_0/dz$ is a free parameter (as are $\Omega_0$ and $S_0$). 
-% This is different to the global problem because in that case, the
-% vertical shear is a result of the disk radial structure, which is
-% ignored in the VGSB equilibria. 
-% In effect, we are studying dynamics in the standard shearing box
-% subject to an additional azimuthal force $-v_zr_0 d \Omega_0/dz$. This
-% force is inspired by the vertical shear in the global disk, and its
-% functional form is taken from global equilibrium.
-
-% \subsection{Limitations}
-% Before using the above equations to study the VSI, it is very important to be
-% aware of several limitations of the VGSB model. The VGSB momentum  
-% equations have the same form as that for the standard shearing box 
-% \citep{goldreich65}, except for the appearence of the term 
-% $v_zr_0d\Omega_0/dz$. This term appears in the global stability
-% problem and is the only term associated with vertical shear that is 
-% ported from the global disk into the local model.
-
-% In Appendix \ref{global_corr} we make the connection between the
-% linear problem in the VGSB to that in the global disk model. We
-% emphasize that not all calculations in the local model can be 
-% consistently interpreted in the global context. However, for
-% perturbations with small radial lengthscales (large $|k_x|$) and fast
-% thermal relaxation ($\beta\ll1$), which characterizes the VSI, the
-% VGSB framework is adequate.
-
-\subsection{Approximate governing equation}\label{approx_gov}
+\subsection{Governing equation in the VGSB}\label{approx_gov}
 We now derive a governing equation with further approximations:

### 2015-03-02 paper.tex
updating main part of paper

@@
   astrophysical disks. However, the VSI requires a short thermal
-  timescale. A quantitative assessment of the thermodynamic
-  requirement for the VSI is thus desirable for application to
-  realistic disks. To this end,  
+  timescale in a stably stratified disk. %which is usually assumed 
+  A quantitative examination of this thermodynamic 
+  requirement is thus necessary to assess the viability of
+  the VSI in realistic disks. To this end,  
   we generalize the linear, axisymmetric VSI in radially local but
@@
   the VSI depends on the thermodynamic response  of the disk by
-  varying $\beta$. We find increasing $\beta$ away from zero
-  dramatically reduces VSI growth rates because this generally
-  introduces a strongly stabilizing vertical entropy gradient. Our
-  linear calculations are consistent with previous hydrodynamic
-  simulations. For vertically isothermal disks, we quantify `fast' 
-  thermal relaxation as $\beta < \epsilon|q|/(\gamma-1)$, where
+  varying $\beta$. We find increasing $\beta$ even slightly away from
+  zero can dramatically reduce VSI growth rates. % , since this generally
+  % introduces a vertical entropy gradient that stabilizes vertical
+  % motions as needed for the VSI.
+  Our  linear calculations are consistent with previous hydrodynamic
+  simulations. For vertically isothermal disks, we quantify the
+  requirement for `fast' thermal relaxation as $\beta < \epsilon|q|/(\gamma-1)$, where
   $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  ($r$
@@
   $\gamma$ is the adiabatic index. We apply our linear
-  results to a disk model of the Minimum Mass Solar Nebulae (MMSN), and
-  find the MMSN is unstable to the VSI at tens of AU with a
-  characteristic growth times of tens of local orbital periods. This 
-  suggests the VSI is dynamically important in the outer parts of
-  protoplanetary disks.  
+  results to a disk model of the Minimum Mass Solar Nebulae (MMSN)
+  with an estimate of the physical thermal timescales in the MMSN
+  based on dust opacity. We find the MMSN is unstable to the VSI at
+  tens of AU with a characteristic growth times of tens of local
+  orbital periods. This suggests the VSI is dynamically important in
+  the outer parts of protoplanetary disks.  
 \end{abstract}
@@
 lengthscales) to become unstable.    
-\cite{umurhan13} gives a simple explanation for this 
-instability: an exchange of fluid parcels while conserving their
-angular momenta individually, may result in a new configuration of
-lower energy, if the exchange occurs over a sufficiently large
-vertical displacement.
+\cite{umurhan13} gives a simple interpretation for the VSI: an
+exchange of fluid parcels while conserving their angular momenta
+individually, may result in a new configuration of lower energy, if
+the exchange occurs over a sufficiently large vertical displacement.
 
@@
 The VSI has been studied analytically in local approximation
-\citep{urpin98,urpin03}. Simulations of the  
+\citep{urpin98,urpin03}. Numerical simulations of the  
 VSI have been performed by \cite{arlt04}, and more recently by 
-\cite{nelson13}. In these simulations the authors indeed observe the
+\cite{nelson13} and \cite{stoll14}. In these simulations the authors indeed observe the
 growth of vertically-global but radially-localized perturbations when
@@
 the disk evolution. This has been confirmed by 
-\cite{stoll14}, who find the 
-VSI is only sustained in radiation hydrodynamic simulations 
-with the inclusion of heating from stellar irradiation.
+\cite{stoll14}, who find the VSI is only sustained in radiation
+hydrodynamic simulations with the inclusion of heating from stellar
+irradiation, which resulted in an effective thermal timescale $t_c\ll
+P_\mathrm{orb}$.
 
@@
 
-The VSI is sensitive to disk thermodynamics except
-in the special case of neutrally stratified disks. It is therefore
-important to understand and quantify this dependence, for example on
-disk parameters, in order to evaluate the viability of the VSI in
-realistic disks. Numerical simulations of the VSI are challenging,
-however, because global disk simulations should be carried out while
-resolving radial scales much smaller than the local disk scale height.
-
-On the other hand, should a given disk be
-unstable to the VSI, it must pass through the linear instability
-first. Thus, studying the thermodynamic dependence of the linear VSI
-is a useful first step. Furthermore, linear stability calculations are
-essential to complement future numerical simulations.
+From these studies it is clear that the VSI is generally sensitive to
+the thermodynamic response of the disk if it is stably
+stratified. Since the latter is usually assumed for typical disk
+models,  
+% disk thermodynamics except
+% in the special case of neutrally stratified disks. 
+it is 
+important to understand and quantify said thermodynamic dependence,
+for example on disk parameters, in order to evaluate the viability of
+the VSI in realistic disks. % Numerical simulations of the VSI are
+% challenging, however, because global disk simulations should be
+% carried out while resolving radial scales much smaller than the local
+% disk scale height.    
+%On the other hand, 
+Should a given disk be unstable to the VSI, it must pass through the
+linear instability first. Thus, studying the thermodynamic dependence
+of the linear VSI is essential. % Furthermore, linear
+% stability calculations are essential to complement future numerical simulations.
 
@@
 isothermal perturbations. Our models are also an extension of early
-analysis of \cite{urpin98} and \cite{urpin03} who consider
-vertically-localized disturbances. Our vertically-global approach is
-necessary to study disturbances with vertical wavelengths
-comparable to or larger than the disk scale height. Such disturbances were 
-found by \cite{nelson13} to dominate their numerical simulations. 
-We thus focus on vertically-global perturbations.
+analysis of \cite{urpin98} and \cite{urpin03} which accounts for thermal
+relaxation, but considered vertically localized disturbances. Our
+vertically global approach is necessary to study disturbances with
+vertical wavelengths comparable to or larger than the disk scale
+height. Such disturbances were  found by \cite{nelson13} and
+\cite{stoll14} to dominate their numerical simulations.  We thus focus
+on vertically global perturbations.
 

### 2015-03-02 setup.tex
updating main part of paper

@@
   &\frac{\p P}{\p t} + \bm{v}\cdot\nabla P  = - \gamma P
-  \nabla\cdot\bm{v} - \frac{\rho}{t_c}\left(\frac{P}{\rho} -
-    \left.\frac{P}{\rho}\right|_{t=0}\right), \label{full_energy}
+  \nabla\cdot\bm{v} - \Lambda % \frac{\rho}{t_c}\left(\frac{P}{\rho} -
+    % \left.\frac{P}{\rho}\right|_{t=0}\right)
+  , \label{full_energy}
 \end{align}
@@
 central star and $G$ is the gravitational constant. 
-In the energy equation (Eq. \ref{full_energy}) we have
-included a thermal relaxation term which restores the disk temperature
-$T\propto P/\rho$ to its initial value on a characteristic timescale
-$t_c=\beta\Omega_k^{-1}$ where $\beta$ is a dimensionless coefficient
-and $\Omega_k=\sqrt{GM_*/r^3}$ is the Keplerian frequency.
+In the energy equation (Eq. \ref{full_energy}) the source term
+$\Lambda$ represent non-adiabatic effects which we
+model below. The gas temperature $T$ is defined through the ideal gas 
+equation of state,  
+\begin{align}
+P = \frac{\mathcal{R}}{\mu}\rho T,
+\end{align}
+where $\mathcal{R}$ is the gas constant and $\mu$ is the mean molecular
+weight.
+
+\subsection{Thermal relaxation}
+We model non-adiabatic effects by defining $\Lambda$ as a thermal
+relaxation term,
+\begin{align}\label{thermal_relax}
+  \Lambda  \equiv \frac{\rho\mathcal{R}}{\mu t_c}\left(T - T_0\right),
+\end{align}
+where $T_0=T_0(r,z)$ is the initial temperature, and $t_c$ is a
+thermal relaxation timescale parametrized via $\beta$ such that 
+\begin{align}\label{beta_def}
+  t_c = \beta \Omega_k^{-1},
+\end{align}
+where $\Omega_k=\sqrt{GM_*/r^3}$ is the Keplerian frequency.
+
+For most of this paper we will take $\beta$ as a constant input
+parameter, so that we can vary the thermodynamic response of the
+disk from isothermal ($\beta\ll 1$) to adiabatic ($\beta \gg 1$) in a
+controlled manner. However, for application to protoplanetary disks in 
+\S\ref{application} we will model $\beta$ based on physical thermal
+timescales, in which case $\beta$ is determined by the equilibrium
+disk properties and the perturbation lengthscales of interest.
 
@@
 pressure and rotation profiles $\rho(r,z)$, $P(r,z)$ and
-$\Omega(r,z)$, respectively. The equilibrium disk satisfies
+$\Omega(r,z)$, respectively. These functions satisfy  
 \begin{align}
@@
 \end{align} 
-%where $\Phi_* = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational
-%potential of the central star.
-We examine disks with power-law profiles for the midplane density
-$\rho_0(r)\equiv\rho(r,0)$: 
+These equations allows us to choose any mid-plane density 
+distribution $\rho_0(r)\equiv \rho(r,0)$. In this work we assume a
+power-law profile, 
 \begin{align}
@@
 where $\rho_{00}$ is the midplane density at the fiducial radius
-$r=r_0$ and $p$ is the power-law index. The equilibrium pressure $P$
-and density $\rho$ are related by 
+$r=r_0$ and $p$ is the power-law index.
+
+We also consider disk equilibria in which the pressure $P$ and density
+$\rho$ are related by   
 \begin{align}
@@
 \end{align}
-where the polytropic index $\Gamma$ is a parameter and 
-the midplane sound-speed $c_{s0}(r)$ is taken to be another power-law: 
+where the polytropic index $\Gamma$ is a  constant parameter  and the
+function $c_{s0}(r)$ is prescribed below. Choosing $\Gamma=1$ gives 
+locally vertically isothermal disk and $c_{s0}(r)$ is the mid-plane
+sound-speed. %For $\Gamma=\gamma$ the disk is adiabatically stratified
+% in the vertical 
+It shall be convenient to define the function $c_s(r,z)$ 
+\begin{align}
+  c_s^2\equiv % \frac{dP}{d\rho} =
+  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} =
+  \frac{\Gamma P}{\rho}, 
+\end{align}
+which is related to the isothermal sound-speed $c_\mathrm{iso} =
+c_s/\sqrt{\Gamma}$. We take $c_{s0}(r)$ to be another power-law:  
 \begin{align}
@@
 \end{align}
-where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the power-law index for
-the midplane temperature. We define
+where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the power-law index, and
+define   
 \begin{align}
   \epsilon(r) \equiv \frac{c_{s0}(r)}{v_k(r)} \equiv
-  \frac{H_\mathrm{iso}}{r} 
+  \frac{H}{r} 
 \end{align}
@@
 $v_k=r\Omega_k$ is the Keplerian azimuthal velocity and
-$H_\mathrm{iso}$ is a characteristic scale-height.  Thus, $\epsilon$
-is also the characteristic disk aspect-ratio.
+$H$ is a characteristic scale-height.  Thus, $\epsilon$
+is also a characteristic disk aspect-ratio.  In practice, we consider
+thin disks with $\epsilon \ll 1$.
 
-The global sound-speed $c_s(r,z)$ is given by
+For discussion purposes it is useful to define the vertical buoyancy
+frequency $N_z$ 
 \begin{align}
-  c_s^2\equiv \frac{dP}{d\rho} =
-  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} = \frac{\Gamma P}{\rho}. 
-\end{align}
-For discussion purposes it is useful to define the vertical buoyancy frequency $N_z$ via
-\begin{align}
-  N_z^2 = c_s^2 \left(\frac{\gamma -
+  N_z^2 \equiv - c_p^{-1}g_z \frac{\p S}{\p z} = c_s^2 \left(\frac{\gamma -
       \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   
 \end{align}
-where $\gamma$ is the constant adiabatic index. We are interested in
+where $c_p$ is the heat capacity at constant pressure, $g_z\equiv
+\rho^{-1}\p_zP $ and $S\equiv c_p^{-1}\ln{P^{1/\gamma}/\rho}$ is the
+specific entropy. The second equality applies to the disk equilibria
+described above. 
+We are interested in
 neutrally or sub-adiabatically stratified 
@@
 
-The explicit expression for the density field for
+%\subsection{Vertical structure}
+
+Solving the equilibrium disk equations, we find the explicit
+expression for the density field for 
 $\Gamma\neq1$ is given  by
@@
   &\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} = 1 +
-  \frac{\left(\Gamma-1\right)}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right),
+  \frac{\left(\Gamma-1\right)}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right).
 \end{align}
-Note that for $\Gamma\neq1$ the disk has a surface $z=H(r)$ such that
-$\rho(r,H)=0$. If we parametrize
-\begin{align}
-  H(r)\equiv h r,
-\end{align}
-then the imposed sound-speed and disk thickness are related by 
-\begin{align}
-  h^2(r) = \left[1-\frac{\epsilon^2(r)}{\Gamma-1}\right]^{-2}-1. 
-\end{align}
+Note that for $\Gamma > 1 + \epsilon^2$ there is a disk surface $H_s$
+such that $\rho(r,H_s)=0$. 
+% Note that for $\Gamma\neq1$ the disk has a surface $z=H(r)$ such that
+% $\rho(r,H)=0$. If we parametrize
+% \begin{align}
+%   H(r)\equiv h r,
+% \end{align}
+% then the imposed sound-speed and disk thickness are related by 
+% \begin{align}
+%   h^2(r) = \left[1-\frac{\epsilon^2(r)}{\Gamma-1}\right]^{-2}-1. 
+% \end{align}
+
 Similarly, the equilibrium rotation is 
@@
 \end{align}
-where $s\equiv q+p(1-\Gamma)$. We also define the radial shear rate  
+where $s\equiv q+p(1-\Gamma)$. We also define % the radial shear rate  
+% \begin{align}
+%   S(r,z) \equiv r\frac{\p\Omega}{\p r},  
+% \end{align}
+% and
+the square of the epicyclic frequency $\kappa$ 
 \begin{align}
-  S(r,z) \equiv r\frac{\p\Omega}{\p r},  
+  \kappa^2(r,z) \equiv \frac{1}{r^3}\frac{\p }{\p
+    r}\left(r^4\Omega^2\right). %2\Omega(2\Omega + S). 
 \end{align}
-and the square of the epicyclic frequency $\kappa$ 
-\begin{align}
-  \kappa^2(r,z) \equiv 2\Omega(2\Omega + S). 
-\end{align}
-
 Note that the disk is in general baroclinic with $\Omega =
@@
   r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
-    z}\frac{dK}{dr}.\label{vertical_shear}
+    z}\frac{dK}{dr} = - \frac{g_z}{\Gamma}\frac{d\ln K}{dr}.\label{vertical_shear}
 \end{align}
@@
 \subsubsection{Vertically isothermal disks}
-For the special case $\Gamma=1$ the disk is isothermal in the
-vertical direction. The equilibrium density $\rho(r,z)$ and rotation
-$\Omega(r,z)$ is then 
+For $\Gamma=1$ the equilibrium density $\rho(r,z)$ and rotation
+$\Omega(r,z)$ are given by 
 \begin{align}
@@
 taking the limit $\Gamma\to 1$. We will, in fact, primarily focus on
-vertically isothermal disks for simplicity.
+vertically isothermal disks which is the relevant case for irradiated
+disks \citep{chiang97}.
 
@@
 \end{align}
-\citep{tassoul78}, where $S = c_p\ln{P^{1/\gamma}/\rho}$ is the
-specific entropy, $c_p$ is the heat capacity at constant
-pressure, and $j\equiv r^2\Omega$ is the
-specific angular momentum.
-
-For rotationally-supported, thin disks, pressure gradients are small 
+\citep{tassoul78}, where $j=r^2\Omega$ is the specific angular
+momentum. For rotationally-supported, thin disks, pressure gradients are small 
 compared to rotation, so the first criterion is generally
@@
 \end{align} 
-For typical model parameters the right hand side is
+For typical model parameters the right hand side (RHS) is
 $O(\epsilon^2)$. Hence, instability in the adiabatic limit 
@@
 entropy gradient in order to be unstable to axisymmetric adiabatic
-perturbations.
+perturbations. This is generally not true for irradiated, stably
+stratified disks \citep{chiang97}, in which the VSI requires rapid
+thermal relaxation, as shown below.
 
@@
 \begin{align}
-  \rho \to \rho(z) + \delta\rho(r,z)\exp{\left( - \ii\sigma
+  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left( - \ii\sigma
       t\right)},    
@@
 It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
-\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$ using their expressions in the global disk. 
+\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$
+using their expressions in the global disk in \S\ref{setup}. 
 Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
@@
 
-% We are interested in the axisymmetric stability of the above system,
-% on radial length-scales much smaller than the local radius $r$, but
-% may be global in the vertical direction. As such, we work within the
-% framework of the \emph{vertically global shearing 
-%   box} (VGSB) developed by \cite{mcnally14}. In this approximation
-% one sets up a local Cartesian co-ordinate  system $(x,y,z)$ centered
-% on a fiducial point $(r,\phi,z)=(r_0,\phi_0,0)$ in the global disk, with $\phi_0(t) =
-% \Omega(r_0,0)t$. The local $x,y,z$ axis correspond to the
-% global radial, azimuthal and vertical directions.
-
-% In this shearing box the equilibrium density, pressure and
-% sound-speed are assumed to depend only on $z$, by evaluating these
-% quantities at $r=r_0$ in the global disk. The background rotation
-% $\Omega$ is approximated by a linear shear in $x$, but left unaltered
-% in $z$. However, the vertical shear $\p_z\Omega$ is also assumed to
-% only depend on $z$.
-
-% The axisymmetric VGSB equations read 
-% \begin{align}
-%   & \frac{\p\rho}{\p t} + \frac{\p}{\p x}\left(\rho v_x\right) +
-%   \frac{\p}{\p z}\left(\rho v_z\right)=0,\\
-%   & \frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
-%   v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}} = - 
-%   2\Omega_0\hat{\bm{z}}\times\bm{v} - S_0v_x\hat{\bm{y}} \notag\\ 
-%   &\phantom{\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} +
-%     v_zr_0\frac{d\Omega_0}{dz} \hat{\bm{y}}=}-\frac{1}{\rho}\nabla P - \frac{\p\Phi_0}{\p z}\hat{\bm{z}},\\
-%   & \frac{\p P}{\p t} + \bm{v}\cdot\nabla P = -\gamma P
-%   \nabla\cdot\bm{v} -
-%   \frac{\rho}{t_c}\left(\frac{P}{\rho}-\left.\frac{P}{\rho}\right|_{t=0}\right), \label{energy_eq}
-% \end{align}
-% where $\bm{v}$ is the velocity field relative to the bulk flow
-% (i.e. $\bm{v}=\bm{0}$ in equilibrium),  
-% subscript $0$ indicates evaluation at $r=r_0$. In the energy equation 
-% we have added a thermal relaxation term with timescale
-% $t_c=\beta\Omega_k^{-1}$ with $\beta$  being a dimensionless input
-% parameter.
-
-% \subsection{Linearization}
-% We perturb the system such that
-%  For clarity we drop the subscript $0$ on the
-% frequencies and shear rates below.
-
-% We can eliminate the perturbed velocities between the linearized
-% equations to obtain a pair or ordinary differential equations
-% for $W,Q$ as
-% \begin{align}
-%   &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
-%   &=\left(\frac{\ii
-%     k_x r}{D}\frac{d\Omega^2}{dz} -
-%   \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
-%   \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
-% &- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
-%   \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
-% &\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
-% \frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
-% &=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
-% \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
-% \end{align}
-% where
-% \begin{align}
-%   D \equiv \kappa^2 - \sigma^2.
-% \end{align}
+% The radially-local approximation above is similar in spirit to the
+% shearing box framework \citep{goldreich65} or the generalization
+% thereof \citep{mcnally14}. The main difference being the terms
+% proportional to $\hat{g}_c$ are neglected in the shearing box.

### 2015-03-03 application.tex
almost done with version 2

@@
 \section{Application to protoplanetary disks}\label{application} 
-As a first application of our linear models, we estimate where in a
+As an application of our linear models, we estimate where in a
 protoplanetary disk do the thermodynamic conditions allow the VSI to
@@
 radiative cooling, 
-%To connect our model with a realistic
-%disk, we consider a more physically-motivated form of the energy
-%equation with radiative cooling 
 \begin{align}\label{real_energy}
@@
   c_v}\left(\nabla\cdot\bm{F} - Q_+\right), 
-\end{align}where the temperature $T$ is defined through $P=\mathcal{R}\rho
-T/\mu$, $\mathcal{R}$ is the ideal gas constant and $\mu$ is
-the mean molecular weight, $c_v$ is the heat capacity at constant
+ \end{align}
+%where the temperature $T$ is defined through $P=\mathcal{R}\rho
+% T/\mu$, $\mathcal{R}$ is the ideal gas constant and $\mu$ is
+% the mean molecular weight, 
+where $c_v$ is the heat capacity at constant
 volume, 
@@
 % specific heat capcity at constant volume
-$\bm{F}$ is the radiative flux, and $Q_+$ is a heating
+$\bm{F}$ is the radiative flux and $Q_+$ is a heating
 term such that an equilibrium solution exists.
 
-One should now model in detail each of the terms on the right
-hand side (RHS) of Eq. \ref{real_energy}, linearize them, then solve
+One should now model in detail each of the terms on the RHS of
+Eq. \ref{real_energy}, linearize them, then solve 
 the resulting linear eigenvalue problem. Such an approach is beyond 
@@
 To do so, we first consider separately perturbations
-with radial lengthscales $l\equiv 2\pi/k_x\gg l_\mathrm{rad}$ and 
+with radial lengthscales $l\gg l_\mathrm{rad}$ and 
 $l\ll l_\mathrm{rad}$, where      
@@
 \subsection{Radiative diffusion}
-Let us first consider perturbations with lengthscales $l\gg
-l_\mathrm{rad}$. In this case, we assume  
+For perturbations with lengthscales $l\gg
+l_\mathrm{rad}$, we assume  
 temperature fluctuations are smoothed out by radiative diffusion. We
@@
 \end{align}
+We note that $t_\mathrm{cool}$ does not explicitly depend on $\rho$.
 
@@
 \end{align}
-%where the relation between $t_\mathrm{cool}$, $l_\mathrm{rad}$ and
-%$\eta$ can be inferred from Eq. \ref{lrad}, Eq. \ref{tc_newton_cool}
-%and Eq. \ref{eta_def}. 
+% where the relation between $t_\mathrm{cool}$, $l_\mathrm{rad}$ and
+% $\eta$ can be inferred from Eq. \ref{lrad}, Eq. \ref{tc_newton_cool}
+% and Eq. \ref{eta_def}. 
 Eq. \ref{tc_def} is a simple prescription so
@@
 \end{align}
-We recall $\beta < \beta_\mathrm{crit}$ is necessary for the
+We recall $\beta < \beta_\mathrm{crit}$ is needed for the
 fundamental VSI when $\khat\gg 1$. %For $\khat\to0$, this
@@
 In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn}  with $\mu
-=2.33$ and $\gamma=1.4$. We set the scaling $\hat{T}=\hat{\Sigma}=1$
-and consider three opacity scales: $\hat{\kappa}_d=0.1, \,1 $ and $10$
-at two heights: $z=0$ and $z=3H$. We also plot Eq. \ref{bcrit_mmsn} in
-each panel.
+=2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for three
+opacity scales: $\hat{\kappa}_d=0.1, \,1 $ and $10$ at two heights:
+$z=0$ and $z=3H$. We also plot Eq. \ref{bcrit_mmsn} in each panel.
 
@@
 Now consider lowering the opacity, as shown in the top panels of
-Fig. \ref{mmsn_bcrit_bcool}. The curves shift inward, but raises also
-them. In fact, lowering the opacity by an order of magnitude renders
-$\beta > \beta_\mathrm{crit}$ even away from the midplane. This makes
-disk stable to the fundametal mode. Raising the opacity, as shown in
+Fig. \ref{mmsn_bcrit_bcool}. The curves shift inward, but are also
+raised. In fact, lowering the opacity by an order of magnitude renders 
+$\beta > \beta_\mathrm{crit}$ even away from the midplane. This is
+expected to stabilize the disk. Raising the opacity, as shown in 
 the lower panels of Fig. \ref{mmsn_bcrit_bcool}, decreases the thermal
@@
 should it develop in the inner disk, would only occur on very small
-scales. It is not clear if such modes are robust (see below).
+scales. However, our linear calculations suggest growth rates
+eventually decay for large $\khat$.
+
+Furthermore, modes with large $\khat$ may be subject to viscous decay
+in a real disk. The viscous time for a perturbation lengthscale
+$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha_s\khat^2\Omega $, where
+$\alpha_s$ is the standard alpha viscosity parameter
+\citep{shakura73}. Setting $t_\mathrm{visc}$ to be longer than the
+typical VSI growth timescale $1/\epsilon\Omega$ places an upper limit
+$\khat \lesssim 30$ in a disk with $\alpha_s\sim 10^{-4}$ and
+$\epsilon \sim 0.1$.
 
@@
 \subsection{VSI in the MMSN}
-We have solved the linear problem in the MMSN with the scale and
+Here we solve the linear problem in the MMSN with the scale and
 height dependent thermal relaxation function defined by
@@
 In fact, for $1\lesssim \khat \lesssim 30$ we either only find the
-fundamental mode, or it is the most unstable. The associated 
+fundamental mode, or that it is the most unstable. The associated 
 growth rates are also insensitive to $z_\mathrm{max}$. However, for
@@
 \citep{nelson13,mcnally14}. Such modes may not be robust because    
-they are affected by boundary conditions. Furthermore, modes with   
-large $\khat$ are likely subject to viscous decay in a real disk.
+they are affected by boundary conditions. % , and high $\khat$ modes are
+% subject to viscous decay in a real disk.
 
 We therefore take Fig. \ref{mmsn_overall} as a representation of the
-VSI in the standard MMSN. We see that growth timescales are $\sim
+VSI in the MMSN. We see that growth timescales are $\sim
 30$---40 orbits, and the VSI operates from a few AU to tens of
 AU. Importantly, it is necessary to consider smaller scales as $r$
-decreases as the disk generally becomes optically thick.
+decreases as the disk generally becomes optically
+thick.
+
+Fig. \ref{mmsn_overall} is roughly consistent with the 
+mid-plane thermal relaxation timescales shown in
+Fig. \ref{mmsn_bcrit_bcool}. For example, for $\khat=10$ we find 
+$\beta \gtrsim \beta_\mathrm{crit}$ for $r\lesssim 10$AU at $z=0$; and
+Fig. \ref{mmsn_overall} show stabilization inside $\sim 8$AU. % The mode
+% operates interior to the expected minimum radius,
+
+% probably because the disk bcomes optically thin away from the
+% midplane.
 
@@
 fact, in this case  we also find the VSI to operate, but with 
-significantly reduced growth rate (as reflected in
-Fig. \ref{mmsn_overall}).
+significantly reduced growth rates (when scaled by the local
+rotation).  
+%: Fig. \ref{mmsn_overall}) show that
+%the growth time at
 
@@
 provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
-fraction of the vertical domain. However, the growth rates are much
-smaller than the case with $\beta < \beta_\mathrm{crit}$ for all $z$.
+fraction of the vertical domain. However, the normalized growth times
+are much longer than the case with $\beta < \beta_\mathrm{crit}$ for
+all $z$: we find $t_\mathrm{grow}\sim 40P_\mathrm{orb}$ at $r=50$AU,
+but $t_\mathrm{grow}\sim 400P_\mathrm{orb}$ at $r=5$AU.
+
+% the case with $\beta < \beta_\mathrm{crit}$ for all $z$.
 

### 2015-03-03 linear_adia.tex
almost done with version 2

@@
-\subsection{Perturbations with thermal relaxation}\label{analytic_relax}
+\subsection{Perturbations with thermal relaxation in the thin-disk limit}\label{analytic_relax}
 We now introduce thermal relaxation ($\beta>0$). We pursue explicit
@@
 We remark that Eq. \ref{nearly_iso_explicit} is in the same 
-form as Eq. 41 in \cite{lubow93}, which describes adiabatic axisymmetric waves in
+form as Eq. 41 in \cite{lubow93}, which describe adiabatic axisymmetric waves in
 a vertically isothermal disk without vertical shear. The two equations
@@
 do the eigenfrequencies and eigenfunctions change when we change
-$t_c$ from zero to a small but finite value? For sufficiently small
-$\beta$, for which the perturbations may be considered 
-nearly-isothermal, we expect the new solutions to only differ slightly
-from the explicit isothermal solutions developed in \S\ref{iso_poly}.
+$t_c=\beta\Omega_k^{-1}$ from zero to a small but finite value? For
+sufficiently small $\beta$, for which the perturbations may be
+considered nearly-isothermal, we expect the new solutions to only
+differ slightly from the explicit isothermal solutions developed in
+\S\ref{iso_poly}.
 
@@
 
-One effect of introducing finite thermal relaxation is buoyancy. 
-The $\zhat^2$ dependence in $\delta v_{z1}$ then makes sense because it
-is most significant for large $\zhat$, i.e. away from the midplane
-where the effect of buoyancy first appears as $t_c$ is increased.
+Introducing a finite thermal relaxation time introduces buoyancy
+forces, which is stabilizing for sub-adiabatically stratified disks
+($\gamma > 1$). The $\zhat^2$ dependence in $\delta v_{z1}$ then makes
+sense because it is most significant for large $\zhat$, i.e. away from
+the midplane where the effect of buoyancy first appears as $\beta$ is
+increased from zero.
 
@@
 vanish, and impose the vertical kinetic energy density
-$\rho|\delta v_z|^2$ to remain finite. Then assuming $g(\zhat)$ is a
-polynomial, we require  
+$\rho|\delta v_z|^2\propto |g|^2 \exp{\left(\real\alpha -
+    1/2\right)\zhat^2}$ to remain finite. Then assuming $g(\zhat)$
+is a polynomial, we require  
 \begin{align}
-  \real\alpha < \frac{1}{2}. 
+  \real\alpha < \frac{1}{2}, 
 \end{align}
-We choose 
+which amounts to choosing 
 \begin{align}
-  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right),  
+  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right).  
 \end{align} 
-by considering the case $q=0$ and $\beta\to\infty$ as examined in
-\cite{lubow93}.
+%by considering the case $q=0$ and $\beta\to\infty$ as examined in
+%\cite{lubow93}.
 
@@
 \begin{align}
-  B + \alpha = M\left(A-2\alpha\right).\label{adia_disp0}
+  B(\hat{\sigma}) + \alpha(\hat{\sigma}) =
+  M\left[A-2\alpha(\hat{\sigma})\right].\label{adia_disp0} 
 \end{align}
@@
 where the coefficients $c_l$ are given in Appendix \ref{relax_coeff}.
-This dispersion relation generally requires a numerical
-solution. However, for the fundamental mode ($M=0$) some analytic
-progress can be made, which we consider next.
+The dispersion relation $\hat{\sigma}=\hat{\sigma}(\khat)$ generally
+requires a numerical solution. However, for the fundamental mode
+($M=0$) some analytic progress can be made, which we consider next.
 
@@
 % following considerations.
- Our discussion of stable solutions 
+Since our discussion of stable solutions  
 without vertical shear yield a dispersion
 relation $\hat{\omega}\propto \khat^{-1}$ for $\khat\gg 1$
-(\S\ref{iso_stable}, Eq. \ref{iso_stable_disp}). We  therefore 
-seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with 
-%we wish to seek a criterion independent of k 
-%\begin{align}
+(\S\ref{iso_stable}, Eq. \ref{iso_stable_disp}), we are motivated to
+seek solutions to 
+Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
+% we wish to seek a criterion independent of k 
+% \begin{align}
 %   \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
@@
 \end{align}
-is required for the fundamental VSI to operate. From our discussion
-for isothermal perturbations (\S\ref{iso_discuss}), we expect
+is required for the fundamental VSI to operate. We will test this
+dependence numerically in \S\ref{bcrit_num_test}. 
+From our discussion for isothermal perturbations
+(\S\ref{iso_discuss}), we expect   
 $\beta<\beta_\mathrm{crit}$ to be a sufficient condition for the
 VSI. That is, when $\beta<\beta_\mathrm{crit}$, perturbations are
-effectively isothermal. For a thin disk 
-$\beta_\mathrm{crit}\ll 1$, which suggests that finite thermal
-relaxation rapidly stabilizes the VSI.
+effectively isothermal.
+
+Rapid thermal relaxation is required to overcome a stabilizing
+vertical entropy gradient \citep{goldreich67,urpin98,urpin03}. These
+studies reached this conclusion by considering vertically-localized
+disturbances. % , although a definitive timescale was not presented. 
+By contrast, our Eq. \ref{iso_vsi_cond} quantifies this requirement
+for vertically global disturbances (i.e. the fundamental mode). We see
+that in a thin disk, a very short thermal timescale
+ is required ($\beta_\mathrm{crit}\ll 1$) because
+$\beta_\mathrm{crit}\propto \epsilon$. 
+This suggests that
+finite thermal relaxation rapidly stabilizes the VSI.
 

### 2015-03-03 linear_iso.tex
updating main part of paper
almost done with version 2

@@
 0$) before considering the effect of finite thermal relaxation. 
-In order to facilitate a fully analytic discussion, we introduce a
+In order to facilitate a fully analytic discussion, we first introduce a
 number of approximations below that allows a simple governing equation
@@
   shearing box} 
-We connect the above linearized equations to the vertically global
+We connect the linearized equations 
+Eq. \ref{lin_mass}---\ref{lin_energy} to the vertically global  
 shearing box (VGSB) formalism developed by \cite{mcnally14}. The VGSB
@@
 that the radial disk structure is ignored in the latter. That is, the 
-equilibrium VGSB is homogeneous in the radial direction, as in the
-standard shearing box.
+equilibrium density and pressure field in the VGSB is homogeneous in
+the radial ($x$) direction, as in the standard shearing box.
 
@@
 Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
-$\hat{g}_c=0$.  
-The resulting linear problem is very similar
+$\hat{g}_c=0$. The resulting linear problem is very similar 
 to that for local axisymmetric waves in accretion disks already
@@
 shear and thermal relaxation. This is convenient as it allows us to
-directly apply their solution methods.
+directly apply some of their solution methods.
 
@@
 azimuthal force $\delta v_zr d \Omega/dz$ with $d\Omega/dz$ as an
-input parameter. This is different to the global problem because in
-that case, the vertical shear is a result of the background radial
-disk structure, which is ignored in the VGSB.
+input parameter. For consistency with the global disk, we choose
+$d\Omega/dz$ in the VGSB to be that due to the thermal structure in
+the global disk (Eq. \ref{vertical_shear}). However, 
+% This is different to the global problem because in
+% that case, the vertical shear is a result of the 
+the background radial disk structure is ignored elsewhere in the VGSB.
 
@@
   \kappa^2$. % The problem simplifies because the eigenvalue now only
-  % appears in the last term in Eq. \ref{iso_ode}.
-  We call this the
-  low-frequency approximation. 
+  % appears in the last term in Eq. \ref{iso_ode}. %stoll kley see low
+  % freq 
+  We call this the low-frequency approximation. 
 \item We are mainly interested in the effect of vertical shear, i.e. $q\neq
@@
 \begin{align}
-  \hat{z} = z/\Hiso,\quad \hat{k}=k_x\Hiso, \quad \hat{\sigma} = \sigma/\Omega_k,
+  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\sigma} = \sigma/\Omega_k,
 \end{align}
@@
 $\delta v_z$. In Appendix \ref{adia_improve} we show that making this
-replacement after eliminating variables, which requires the vertical
-derivative of $D$, adds unnecessary complexity for thin disks.
+replacement after eliminating variables, which involves $dD/dz$,
+introduces terms $O(\epsilon^2)$. Since we are interested in 
+thin disks ($\epsilon\ll 1 $), e.g. protoplaneary disks where 
+$\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
+unnecessary complexity for the present discussion. However, this
+approximation will be relaxed in our numerical study.
 
@@
 
-Note that we can generalize Eq. \ref{vertiso_gov_nondim} to other
-(non-Keplerian) rotation profiles by redefining $\khat$
-as $k_xc_s/\kappa$, $\hat{\sigma}$ as $\sigma/\kappa$, and 
-$\epsilon$ as $c_s/r\kappa$.
+Note that we can generalize Eq. \ref{vertiso_gov_nondim} non-Keplerian
+rotation profiles by redefining $\khat$ as $k_xc_s/\kappa$,
+$\hat{\sigma}$ as $\sigma/\kappa$, and  $\epsilon$ as $c_s/r\kappa$.
 
@@
 \subsection{Isothermal perturbations}\label{iso_discuss}
-Isothermal perturbations correspond to $\beta\equiv0$, 
+Isothermal perturbations correspond to $\beta\equiv0$ 
 %relaxation timescale $\be
@@
 %$\beta=0$. 
-so $\chi=1$ and Eq. \ref{ode_Q} implies 
+($\chi=1$) and Eq. \ref{ode_Q} becomes  
 \begin{align}
-  Q = W, 
+  Q = W 
 \end{align}
- and so 
+(i.e. $\delta P = c_s^2\delta \rho$), and so  
 \begin{align}
@@
 \end{align}
+For a vertically isothermal disk we have 
+\begin{align}
+  f(\zhat) = \frac{\zhat}{\left(1+\epsilon^2\hat{z}^2\right)^{3/2}}, 
+\end{align}
+%For the vertically isothermal disk we have $|f(\hat{z})| =
+%|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, 
+which maximizes at $|\zhat|=1/\sqrt{2}\epsilon$. The existence of a
+maximum vertical shear limits the growth rate of unstable modes, as  
+shown below.
 
-% The linear problem simplifies considerably for the case
-% $\Gamma=\gamma=1$.  Then 
-% \begin{align}
-%   Q=W. 
-% \end{align}
-% Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
-% problem for isothermal perturbations in a vertically isothermal disk, 
-% \begin{align}\label{iso_ode}
-%   \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
-%       r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
-%   \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
-% \end{align} 
-% subject to appropriate boundary conditions. Eq. \ref{iso_ode} also
-% applies to $\gamma>1$ for $t_c\to0$.
+Note that Eq. \ref{iso_ode1} can in fact be derived by substituting
+$Q=W$ in Eq. \ref{ode_w} without approximating $D$ or $\kappa^2$ as
+vertically constant.
 
-Note that for isothermal perturbations the issue of ignoring the
-vertical dependence of $D$ described in \S\ref{approx_gov} and Appendix
-\ref{adia_improve} is irrelevant, since
-Eq. \ref{iso_ode1} may be derived directly by substituting $Q=W$ in
-Eq. \ref{ode_w}, i.e. no derivatives of $D$ needs to be taken.
+% the issue of ignoring the
+% vertical dependence of $D$ described in \S\ref{approx_gov} and Appendix
+% \ref{adia_improve} is irrelevant, since
+% Eq. \ref{iso_ode1} may be derived directly , i.e. no derivatives of $D$ needs to be taken.
 
@@
 satisfying the above boundary conditions, is limited by the maximum
-value of the vertical shear in the domain considered.
+vertical shear in the domain considered.
 
-For the vertically isothermal disk we have $|f(\hat{z})| =
-|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, which maximizes
-at $|\zhat|=1/\sqrt{2}\epsilon$. For vertical domains smaller 
-than this height, the growth rate is limited by $|f|$ at the vertical
-boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
-\hat{z}$, in which case we can expect growth rates to increase at most linearly
-with respect to the vertical domain size.   %no faster than linear
+In practice, one might consider a vertical domain of a few scale 
+heights in a thin disk with $|q|=O(1)$. In this case, $|\epsilon\zhat|\ll1$ and  
+$f\simeq \zhat$, so that $\mathrm{max}|f| = O(1)$, implying a
+maximum growth rate $O(\epsilon \Omega_k)$, consistent with numerical
+results below.
+
+% For vertical domains smaller 
+% than this height, the growth rate is limited by $|f|$ at the vertical
+% boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
+% \hat{z}$, in which case we can expect growth rates to increase at most linearly
+% with respect to the vertical domain size.   %no faster than linear
 
@@
 Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
-that the kinetic energy density remain bounded at infinity, then  
+that the kinetic energy density remain bounded at infinity, then the
+solutions are Hermite polynomials  
 \begin{align}
@@
 n-1$. Physically, this corresponds to radial length-scales much
-smaller than the local disk scale height.
+smaller than the local disk scale height. Note that
+Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
+\cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
+Keplerian disk approximation in that equation.
 
@@
 Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
-has also been qualitatively observed in numerical simulations performed by 
-\cite{stoll14}.
+has also been qualitatively observed in numerical simulations
+\citep{stoll14}.
 
@@
 $L\geq1$. We are interested in 
-unstable modes for which $\imag{\hat{\sigma}}>0$ is the growth rate and is
-given by 
+unstable modes for which $\hat{\nu}=\imag{\hat{\sigma}}>0$ is the
+growth rate and is given by 
 \begin{align}\label{simple_growth}
@@
 For fixed $q$ and $\epsilon$, the growth rate vanishes for both
-$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. The maximum growth rate
+$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
+The maximum growth rate
 occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$

### 2015-03-03 numerical.tex
almost done with version 2

@@
 use the full expressions of $\rho$, $\Omega$ and
-$\kappa$ given in \S\ref{eqm}. We also restore $\hat{g}_c=1$ to
+$\kappa$ given in \S\ref{eqm}. We also set $\hat{g}_c=1$ to
 account for the background radial disk structure.
@@
 \end{align}
-where $\bm{\xi}$ is the Lagrangian displacement. This boundary
-condition is convenient for numerical implementation since it conforms
-to a standard eigenvalue problem. The above discretization procedure
+where $\bm{\xi}$ is the Lagrangian displacement, the meridional
+components of which are given by $\xi_{x,z} = \ii\delta
+v_{x,z}/\sigma$. Calculations with a rigid boundary ($\delta v_z=0$)
+yield similar results. 
+% This boundary
+% condition is convenient for numerical implementation since it conforms
+% to a standard eigenvalue problem.
+
+The above discretization procedure
 converts the linear system of differential equations to a set of 
@@
 
-\subsection{VSI in a locally isothermal, neutrally stratified disk}\label{vertiso_pertiso} 
+\subsection{VSI with rapid thermal relaxation}\label{vertiso_pertiso} 
 % disk model is a nearly-vertically
@@
 % $\epsilon=0.1$. 
-We first calculate the VSI in a nearly vertically isothermal, neutrally
-stratified disk ($N_z^2\equiv0$) by setting
-$\gamma=\Gamma=1.011$ and $\beta=10^{-3}$. Other disk parameters are 
-$(p,q,\epsilon)=(-1.5,-1,0.1)$ and the vertical domain is $\zmax =
-5\epsilon r$. This setup is similar to that adopted in
-\cite{mcnally14}, who calculated the VSI in vertically isothermal
+We first calculate the VSI in a neutrally
+stratified disk ($N_z^2\equiv0$) with rapid thermal relaxation by
+setting $\gamma=\Gamma=1.011$ and $\beta=10^{-3}$. Other disk
+parameters are $(p,q,\epsilon)=(-1.5,-1,0.1)$ and the vertical domain
+is $\zmax = 5H$. The disk is nearly vertically isothermal with
+$c_s(\zmax)\simeq 0.94 c_s(0)$. Fig. \ref{omega_z} shows the
+equilibrium rotation profile away from the disk mid-plane.
+
+This setup is similar to that adopted
+in \cite{mcnally14}, who calculated the VSI in vertically isothermal
 disks subject to isothermal perturbations ($\Gamma=\gamma=1$).  This
 allows us to test our numerical code.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/omega2} 
+  \caption{Equilibrium rotation profile $\Omega(z)$,
+    normalized by its mid-plane value, for a disk with $\Gamma=1.011$
+    and $(p,q,\epsilon)=(-1.5,-1,0.1)$. 
+    \label{omega_z} 
+  }
+\end{figure}
 
@@
 
-% This means
-% that our analytic results should only be applied to perturbations with
-% $\khat\gg 1$, i.e. small radial-lengthscales.
+{\bf ASK. Growth rates are expected to diminish for both $\khat\to 0$
+  and $\khat\to \infty$. This is because for $\khat\ll 1$ the radial
+  wavelenth is large, but such disturbances are 
+  stabilized by by epicyclic motions ($|\omega|\sim
+  \kappa$). Disturbances with $\khat\gg 1$ are radially localized, but  
+  there is no energy change when fluid elements are perturbed at fixed
+  cylindrical radius whilst conserving its angular momentum (since we
+  consider axisymmetric perturbations).  
+}
 
@@
 corresponding eigenfunctions in vertical velocity in
-Fig. \ref{lowfreq_eigenfuncvz}, which show that the fundamental mode
+Fig. \ref{lowfreq_eigenfuncvz}. The fundamental mode
 consists of zero nodes in $\delta v_z$.
@@
 $\sigma = \omega + \ii\nu$ found for $\khat=20\pi$. This plot is
-similar to Fig. 3 in \cite{mcnally14}. The fundamental
-VSI mode, with one node in $W$, has the smallest $|\sigma|$. The
-eigenvalue calculated from Eq. \ref{sig2_iso} with $L=1$ and the 
-numerically-obtained value are  
+similar to Fig. 3 in \cite{mcnally14}. The fundamental has the
+smallest $|\sigma|$, for which the eigenvalue calculated from
+Eq. \ref{sig2_iso} with $L=1$ and the numerically-obtained value are   
 \begin{align*}
@@
      $(p,q,\epsilon)=(-1.5,-1,0.05)$. The vertical line
-     is the critical thermal timescale $\beta_\mathrm{crit}$ for the
-     fundamental mode according to Eq. \ref{iso_vsi_cond}. 
+     is the critical thermal timescale $\beta_\mathrm{crit}$  obtained
+     from Eq. \ref{iso_vsi_cond}. 
      \label{bcrit_compare1}}   
@@
 We visualize the effect of thermal relaxation in 
-Fig. \ref{relax_eigenW_num} by comparing the eigenfunction $W$ for
+Fig. \ref{relax_eigenW_num} by comparing the vertical velocity
+perturbation $\delta v_z$ for
 $\beta=0.01$ and $\beta=0.1$. The latter value is close to the
@@
 (\S\ref{iso_vsi_beta_crit}). Increasing $\beta$ 
-restricts  the region in which $W\sim z$ closer to the midplane, with
+restricts  the region in which $\delta v_z\sim $ constant closer to the midplane, with
 oscillatory behavior emerging at the boundaries, where buoyancy first
-becomes important. Finite thermal relaxation increases $|W|$ near the
-vertical boundaries relative to that about the midplane, but
-perturbations near the disk surface is unlikely important in practice
-because the disk atmosphere contains little mass.
+becomes important. Finite thermal relaxation increases $|\delta v_z|$ near the
+vertical boundaries relative to that about the midplane. However, 
+perturbations near the disk surface may not be energetically important
+in practice because the disk atmosphere contains little mass.
 
@@
 
-\subsection{Critical thermal relaxation timescale}
+\subsection{Critical thermal relaxation timescale}\label{bcrit_num_test}
 % may need to explain why choose k=10
@@
 $\Gamma=1.4$ and $(p,q,\epsilon)=(0,-1,0.05)$ as simulated in
-\cite{nelson13}. In this case we set $\zmax\simeq2.2\epsilon r$, which
+\cite{nelson13}. In this case we set $\zmax\simeq2.2H$, which
 is close to the zero-density surface.
@@
 Fig. \ref{gcorr_compare_vnoniso} shows that introducing finite thermal
-relaxation rapidly stabilizes the fundamental VSI mode. This is
+relaxation rapidly stabilizes the fundamental mode. This is
 similar to that observed for nearly vertically isothermal disks
@@
 % There is good agreement between Fig. \ref{adia_growth_num} and
-% Fig. \ref{adia_growth} for $\khat=k_xH_\mathrm{iso}\gtrsim1$. There is a
+% Fig. \ref{adia_growth} for $\khat=k_xH\gtrsim1$. There is a
 % mismatch for smaller wavenumbers, especially in the growth

### 2015-03-03 paper.tex
almost done with version 2

@@
   being the cylindrical radius from the central object), and 
-  $\gamma$ is the adiabatic index. We apply our linear
-  results to a disk model of the Minimum Mass Solar Nebulae (MMSN)
-  with an estimate of the physical thermal timescales in the MMSN
-  based on dust opacity. We find the MMSN is unstable to the VSI at
-  tens of AU with a characteristic growth times of tens of local
-  orbital periods. This suggests the VSI is dynamically important in
-  the outer parts of protoplanetary disks.  
+  $\gamma$ is the adiabatic index. We apply our linear calculations 
+  to a disk model of the Minimum Mass Solar Nebulae (MMSN) with an 
+  estimate for the physical thermal timescales 
+  in the MMSN based on dust opacity. We find the MMSN is unstable to
+  the VSI at tens of AU with a characteristic growth times of $\sim
+  30$ local orbital periods.  
+  % This suggests the VSI is dynamically important in
+  % the outer parts of protoplanetary disks.  
 \end{abstract}

### 2015-03-03 setup.tex
updating main part of paper
almost done with version 2

@@
 
-We also consider disk equilibria in which the pressure $P$ and density
+We consider disk equilibria in which the pressure $P$ and density
 $\rho$ are related by   
@@
 For discussion purposes it is useful to define the vertical buoyancy
-frequency $N_z$ 
+frequency $N_z$ via
 \begin{align}
@@
 where $c_p$ is the heat capacity at constant pressure, $g_z\equiv
-\rho^{-1}\p_zP $ and $S\equiv c_p^{-1}\ln{P^{1/\gamma}/\rho}$ is the
+\rho^{-1}\p_zP $ and $S\equiv c_p\ln{P^{1/\gamma}/\rho}$ is the
 specific entropy. The second equality applies to the disk equilibria

### 2015-03-03 application.tex
rough version 2 done

@@
 \begin{align}\label{tc_diff_cool} 
-  t_\mathrm{diff} = \frac{1}{\eta k_x^2}\equiv \frac{\Omega_k^{-1}}{\hat{\eta}\khat^2}, 
+  t_\mathrm{diff} = \frac{1}{\eta k_x^2}.%\equiv \frac{\Omega_k^{-1}}{\hat{\eta}\khat^2}, 
 \end{align}
-where $\hat{\eta} = \eta/H^2\Omega_k$ is the 
-dimensionless diffusion coefficient.
+% where $\hat{\eta} = \eta/H^2\Omega_k$ is the 
+% dimensionless diffusion coefficient.
 
@@
   %   + \frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right]  \simeq
-  \frac{1}{\hat{\eta}}\left[\frac{1}{3\kappa_d^2\Sigma^2} 
+  \frac{\Sigma^2\Omega_k}{\eta\rho^2}\left[\frac{1}{3\kappa_d^2\Sigma^2} 
     + \frac{\hat{\rho}^2(z)}{2\pi \khat^2}\right].
@@
 
-We therefore take Fig. \ref{mmsn_overall} as a representation of the
+We therefore consider Fig. \ref{mmsn_overall} as a representation of the
 VSI in the MMSN. We see that growth timescales are $\sim
@@
 AU. Importantly, it is necessary to consider smaller scales as $r$
-decreases as the disk generally becomes optically
-thick.
+decreases as the disk generally becomes optically thick.
 
@@
 $\beta \gtrsim \beta_\mathrm{crit}$ for $r\lesssim 10$AU at $z=0$; and
-Fig. \ref{mmsn_overall} show stabilization inside $\sim 8$AU. % The mode
-% operates interior to the expected minimum radius,
+Fig. \ref{mmsn_overall} show stabilization inside $\sim 8$AU.
 
+%The mode % operates interior to the expected minimum radius, 
 % probably because the disk bcomes optically thin away from the
@@
   \caption{Characteristic growth times of the fundamental VSI mode in
-    the MMSN.  
+    the MMSN. These are calculated using the scale and height
+    dependent thermal relaxation given by Eq. \ref{beta_mmsn_simp}. 
     \label{mmsn_overall}}    
@@
 thick near the midplane and becomes optically thin away from it, such
-that $\beta < \beta_\mathrm{crit}$ only for $|z|\gtrsim1.25H$. In
+that $\beta < \beta_\mathrm{crit}$ for $|z|\gtrsim1.25H$. In
 fact, in this case  we also find the VSI to operate, but with 
@@
 are much longer than the case with $\beta < \beta_\mathrm{crit}$ for
-all $z$: we find $t_\mathrm{grow}\sim 40P_\mathrm{orb}$ at $r=50$AU,
-but $t_\mathrm{grow}\sim 400P_\mathrm{orb}$ at $r=5$AU.
+all $z$: we find $t_\mathrm{grow}\sim 40$ orbits at $r=50$AU,
+but $t_\mathrm{grow}\sim 400$ orbits at $r=5$AU. Thus, introducing an
+optically thick midplane reduces the efficiency of the VSI despite
+$\beta < \beta_\mathrm{crit}$ away from the midplane.
+
+% The grow
+% mode because it involves vertical motion across $z=0$, which is
 
 % the case with $\beta < \beta_\mathrm{crit}$ for all $z$.
+
+% Introducing an optically thick mid-plane may not completely surpress
+% the VSI because the vertical entropy gradient, which usually provides
+% stabilization, vanishes at $z=0$. (Note that $N_z^2\propto z^2$ while
+% vertical shear $d\Omega^2/dz \propto z$)
 
@@
   \caption{Vertical velocity perturbations of the VSI in the MMSN at 
-    $r=50$AU (top) and $r=5$AU (bottom). The real (imaginary) part of 
+    $r=50$AU (top) and $r=5$AU (bottom). These are calculated using the height
+    dependent thermal relaxation given by Eq. \ref{beta_mmsn_simp}. The real (imaginary) part of 
     $\delta v_z$ is shown as

### 2015-03-03 linear_iso.tex
rough version 2 done

@@
 % that case, the vertical shear is a result of the 
-the background radial disk structure is ignored elsewhere in the VGSB.
+the background radial disk structure is ignored elsewhere in the VGSB
+(which is done by setting $\hat{g}_c=0$).
 
 We further discuss caveats of the VGSB in Appendix \ref{global_corr}.
-However, we find that for perturbations with small radial lengthscales
+We find that for perturbations with small radial lengthscales
 ($k_xH\gg 1$) and fast thermal relaxation ($\beta\ll1$), which

### 2015-03-03 numerical.tex
rough version 2 done

@@
 is $\zmax = 5H$. The disk is nearly vertically isothermal with
-$c_s(\zmax)\simeq 0.94 c_s(0)$. Fig. \ref{omega_z} shows the
-equilibrium rotation profile away from the disk mid-plane.
+$c_s(\zmax)\simeq 0.94 c_s(0)$. Fig. \ref{omega_z} shows that the 
+equilibrium rotation decreases slightly away  from the disk mid-plane.
 
@@
 
-{\bf ASK. Growth rates are expected to diminish for both $\khat\to 0$
+{\bf Growth rates are expected to diminish for both $\khat\to 0$
   and $\khat\to \infty$. This is because for $\khat\ll 1$ the radial

### 2015-03-03 paper.tex
rough version 2 done

@@
 Should a given disk be unstable to the VSI, it must pass through the
-linear instability first. Thus, studying the thermodynamic dependence
+linear instability first. Thus, understanding the thermodynamic dependence
 of the linear VSI is essential. % Furthermore, linear
@@
 isothermal perturbations. Our models are also an extension of early
-analysis of \cite{urpin98} and \cite{urpin03} which accounts for thermal
+analysis of \cite{urpin98} and \cite{urpin03}, which accounts for thermal
 relaxation, but considered vertically localized disturbances. Our
@@
 \cite{stoll14} to dominate their numerical simulations.  We thus focus
-on vertically global perturbations.
+on vertically global perturbations which, in fact, we find to be the
+dominant mode in protoplanetary disks with a realistic estimate for
+the thermodynamic timescale.
 
@@
 \S\ref{application} we apply our linear models to the Minimum Mass
-Solar Nebulae (MMSN) and find that the VSI may operate in the outer
-parts of the MMSN. We summarize and discuss our results in
-\S\ref{summary}, including caveats and possible directions for future
-work.
+Solar Nebulae (MMSN) by modeling thermal relaxation
+timescales based on realistic disk properties, and find that the VSI may
+operate in the outer parts of the MMSN. We summarize and discuss our
+results in \S\ref{summary}, including caveats and possible directions
+for future work.
 

### 2015-03-03 summary.tex
rough version 2 done

@@
 height-dependence of the equilibrium angular velocity profile
-$\Omega(r,z)$ of a baroclinic disk. The VSI was analyzed and
-explained earlier in the local limit by \cite{urpin98} and
-\cite{urpin03}; and simulated by \cite{arlt04} and more recently by
-\cite{nelson13}.
+$\Omega=\Omega(r,z)$ in a baroclinic disk. We examined how
+the linear instability depends on disk thermodynamics and applied our
+results to assess the relevance of the VSI in realistic protoplanetary
+disks.
 
-We generalized previous vertically-global linear stability calculations
+% The VSI was analyzed earlier in 
+% the local limit by \cite{urpin98} and 
+% \cite{urpin03} and simulated by \cite{arlt04}. Recent non-linear
+% numerical simulations by \cite{nelson13} and \cite{stoll14} indicate
+% the VSI generally requires a short thermal timescale.
+
+We generalized previous vertically global linear stability calculations
 of the VSI \citep{nelson13,mcnally14}, which assumed isothermal
 perturbations, to include an energy equation. Our effort may also be
-regarded as an extension of \cite{urpin03}, which included an energy
-equation, by relaxing the Boussinesq and vertically-local
+regarded as an extension of \cite{urpin03}, which did include an
+energy equation, by relaxing the Boussinesq and vertically local
 approximations adopted in that analysis.
 
-We added a source term in the energy equation which relaxes the disk
-temperature to its initial value on a timescale $t_c=\beta\Omega_k^{-1}$. This
-form of thermal relaxation was also employed in the global numerical
-simulations of \cite{nelson13}, and allows the thermodynamic response
-of the disk to be controlled.
+We included a source term in the energy equation which relaxes the disk
+temperature to its initial value on a timescale
+$t_c=\beta\Omega_k^{-1}$. This approach to model thermal relaxation was also
+employed in the global numerical simulations of \cite{nelson13}, and
+allows the thermodynamic response of the disk to be controlled.
 
@@
 perturbations with small radial lengthscales, we applied the 
-radially-local approximation in the linearized fluid equations. For
+radially local approximation in the linearized fluid equations. For
 analytic discussion only, we further simplified the linear problem by
@@
 VGSB. We find the VGSB framework is appropriate for $\beta\lesssim1$. 
-This is convenient because previous works show that rapid thermal
-relaxation ($\beta\ll1$) is generally required for the VSI. However,
-the VGSB framework is not valid in the adiabatic limit.
+This is convenient because rapid thermal relaxation ($\beta\ll1$) is
+generally required for the VSI.  However,
+the VGSB framework is not appropriate in the adiabatic limit ($\beta \gg
+1$).
 
@@
 
-We mostly focused on vertically isothermal disks ($\Gamma=1$) because
-it permits a largely analytical discussion. Our analysis is an
+We mostly focused on vertically isothermal disks ($\Gamma=1$), since
+this is the relevant case for protoplanetary disks that are irradiated
+\citep{chiang97}. It also permits a largely analytical discussion.  Our analysis is an
 extension of the linear models considered in \cite{lubow93} by 
@@
 perturbations in nearly-Keplerian disks. These approximations were
-made for simplicity, but also motivated by previous work
-\citep{nelson13}.
+made for simplicity, but also motivated by recent numerical
+simulations \citep{nelson13}.
 
@@
 vertical domain of interest. The fundamental VSI mode is characterized
-by one node in the density perturbation at the mid-plane
+by one node in the density perturbation at the midplane 
 (i.e. anti-symmetric about $z=0$), which corresponds to a vertical
-velocity perturbation  $\delta v_z\sim\mathrm{constant}$ near
-$z=0$.
+velocity perturbation  $\delta v_z\sim\mathrm{constant}$ near the
+mid-plane with zero nodes. These have been referred to as
+`corrugation modes' by \cite{nelson13}, and were found to dominate
+their numerical simulations, as well as that of
+\cite{stoll14}. Furthermore, when computing unstable modes for a
+disk with a physical model for thermal relaxation, we also find
+the fundamental mode to be dominant and robust. Such modes are thus
+likely most relevant to  realistic disks.
 
-% In the opposite limit of adiabatic perturbations ($\gamma>1$
-% and $\beta\to\infty$), we find VSI growth rates are dramatically
-% reduced on account of a stabilizing 
-% vertical entropy gradient. The VSI is not completely
-% surpressed because vertical shear dominates over buoyancy 
-% sufficiently close the disk midplane. However, perturbations grow
-% slowly and is constrained to regions immediately about the disk
-% midplane. Thus, the adiabatic VSI is likely unimportant in practice.
-
-For non-isothermal  perturbations with thermal relaxation ($\gamma>1$,
+For non-isothermal  perturbations with finite thermal relaxation ($\gamma>1$,
 $\beta>0$) , the general dispersion relation (Eq. \ref{relax_disp}) is
-complicated and requires a numerical treatment. However, for the 
+quite complicated. However, for the 
 fundamental VSI some simplification is  possible. In particular, we
@@
 operates.  We expect $\beta<\beta_\mathrm{crit}$ to be a sufficient
-condition for the VSI to operate. We believe this explicit expression for
-$\beta_\mathrm{crit}$ is the most useful result from our
-investigation.
+condition for the VSI to operate. We have therefore provided a
+quantitative measure for the`fast' thermal relaxation requirement for the
+VSI.
 
 We also solved the linear stability problem numerically using a
-pseudo-spectral approach. For numerical calculations we accounted for
-the background radial disk structure for consistency. We focused
-on the fundamental mode to compare with our analytical
-discussion, and because \cite{nelson13} found that this mode 
-eventually dominate their simulations. Our numerical solutions are
-mostly consistent with our analytic expectations, including the 
-dependence of $\beta_\mathrm{crit}$ on disk parameters.
+pseudo-spectral approach. For numerical calculations we relax all the
+approximations made in our analytical discussion, and accounted for
+the background radial disk structure for consistency.  Our
+numerical solutions are mostly consistent with our analytic 
+expectations, including the dependence of $\beta_\mathrm{crit}$ on
+disk parameters.
 
@@
 
-We applied our linear results to assess the applicability of the VSI in
-protoplanetary disks, specifically the Minimum Mass Solar Nebulae (MMSN),
-using disk models described in \cite{chiang10}. We find the MMSN is
-unstable to the fundamental VSI at $10$s of AU, with characteristic
-growth times of $10$s of orbits. However, we note the strong
-temperature dependence of the physical thermal relaxation timescale
-in the disk model ($\beta$ decreases as $T^{-5}$ for perturbations with small
-radial scale). Thus, increasing the disk temperature even slightly can significantly 
-shorten the typical growth times for the VSI. This is also attributed
-to a larger value of $\beta_\mathrm{crit}\propto\epsilon$, since increasing $T$ also
-increases the aspect-ratio $\epsilon$.
+We applied our linear models to assess the applicability of the VSI in
+protoplanetary disks, specifically the Minimum Mass Solar Nebulae
+(MMSN). We consider the disk model described in \cite{chiang10}, and
+modeled the thermal relaxation timescale $\beta$ based on
+dust opacity, which results in a $\beta$ that depends on both the
+lengthscale of interest as well as the height from the midplane.  
+We find the MMSN is unstable to the fundamental VSI at $10$s of AU,
+with characteristic growth times of $\sim 30$ orbits. Although $\beta$
+is not spatially constant as assumed in our analysis, we find that
+evaluating $\beta < \beta_\mathrm{crit}$ at the disk
+midplane gives a rough measure of the radii at which the VSI operates in the
+MMSN.  This simple criteron may be a useful first step to 
+estimate the relevance of the VSI in realistic astrophysical
+disks without resorting to non-linear numerical simulations.
+
+%  However, we note the strong
+% temperature dependence of the physical thermal relaxation timescale
+% in the disk model ($\beta$ decreases as $T^{-5}$ for perturbations with small
+% radial scale). Thus, increasing the disk temperature even slightly can significantly 
+% shorten the typical growth times for the VSI. This is also attributed
+% to a larger value of $\beta_\mathrm{crit}\propto\epsilon$, since increasing $T$ also
+% increases the aspect-ratio $\epsilon$.
 
@@
 \subsection{Caveats and outlooks} 
-We highlight below some improvements to the present  
+We highlight below some extensions to the present  
 linear models in future works:
 
-\emph{Realistic thermal evolution.} We have treated the thermal
-relaxation time $t_c$ as a constant for simplicity. However, a
-physical expression for $t_c$ will depend on 
-$z$. For example, the Newtonian cooling timescale
-(\S\ref{newton_cool}) will depend on $z$ for vertically non-isothermal
-disks, and through the opacity since 
-$\kappa_d=\kappa_d(\rho,T)$ in general. 
-A proper treatment for radiative diffusion, on the other hand, will
-ultimately involve the Laplacian operator. In this case additional
-boundary conditions must be supplemented. We will consider this in a
-future work.
-
-%need to compare kley sims
+\emph{Improved thermal evolution.} We assumed non-adiabatic effects
+can be represented by a thermal relaxation timescale $t_c$. For 
+application to protoplanetary disks we connected $t_c$ to the disk
+properties, but one can improve the realism by generalizing the
+opacity model adopted here to be density-dependent,  i.e. $\kappa_d =
+\kappa_d(\rho, T)$.  However, going beyond the framework of thermal
+relaxation requires a proper treatment for radiative diffusion, which
+involves the Lapacian operator and additional boundary conditions
+must be supplemented. We will consider these improvements in a
+follow-up study.
 
@@
 
-\emph{Radial dependence.} We have adopted the radially-local approximation 
-throughout this paper. This is based on previous works, including
-nonlinear simulations, which show that the VSI is relevant for
-perturbations with small radial lengthscales. Although this simplifies
-the linear problem to ordinary differential equations, this approach
-introduces the wavenumber $k_x$ as a free parameter. The effect
-of radial boundary conditions cannot be explored. Future
+\emph{Radial dependence.} Based on previous works, including
+non-linear simulations, which show that the VSI is relevant for
+perturbations with small radial lengthscales, we have adopted the
+radially-local approximation throughout this paper.  Although this
+simplifies the linear problem to ordinary differential equations, this
+approach introduces the wavenumber $k_x$ as a free parameter. The
+effect of radial boundary conditions cannot be explored. Future 
 generalizations of the linear problem should retain the radial
 dependence properly and solve the two-dimensional $(r,z)$ partial 
-differential equation.
+differential equation eigenvalue problem.
 

### 2015-03-03 linear_adia.tex
updating

@@
 \subsection{Perturbations with thermal relaxation in the thin-disk limit}\label{analytic_relax}
-We now introduce thermal relaxation ($\beta>0$). We pursue explicit
-solutions in the thin-disk approximation. The governing equation,
-Eq. \ref{vertiso_gov_nondim}, is then 
+We now introduce thermal relaxation ($\beta>0$), but pursue explicit
+solutions in the thin-disk approximation for simplicity. The governing
+equation, Eq. \ref{vertiso_gov_nondim}, is then 
 \begin{align}
@@
 \end{align}
-Here $\delta v_z$ corresponds to the explicit solution developed in
+Here $\delta v_z$ corresponds to an explicit solution developed in
 \S\ref{iso_poly} with  $\hat{\sigma}^2$ given by Eq. \ref{sig2_iso}. Inserting
@@
 Rapid thermal relaxation is required to overcome a stabilizing
-vertical entropy gradient \citep{goldreich67,urpin98,urpin03}. These
-studies reached this conclusion by considering vertically-localized
-disturbances. % , although a definitive timescale was not presented. 
+vertical entropy gradient \citep{goldreich67,urpin98,urpin03}, but
+this conclusion was reached in these studies by considering
+vertically-localized disturbances. % , although a definitive timescale
+                                % was not presented.  
 By contrast, our Eq. \ref{iso_vsi_cond} quantifies this requirement

### 2015-03-03 linear_iso.tex
updating

@@
 where $\Gamma=1$, then $c_s$ is a constant in the radially-local
-approximation. We  first examine isothermal perturbations ($\beta =
+approximation. We examine isothermal perturbations ($\beta =
 0$) before considering the effect of finite thermal relaxation. 
@@
 is an extension of the standard shearing box \citep{goldreich65} to
-allow the shear flow in the shearing box to be height-dependent.  
-The difference between the radially-local approximation and the VGSB is
-that the radial disk structure is ignored in the latter. That is, the 
-equilibrium density and pressure field in the VGSB is homogeneous in
-the radial ($x$) direction, as in the standard shearing box.
+allow the shear flow in the shearing box to be height-dependent. This
+is an attempt to capture the effect of vertical shear in a local model
+of accretoin disks. The difference between the radially-local
+approximation and the VGSB is that the radial disk structure is
+ignored in the latter. That is, the  equilibrium density and pressure
+field in the VGSB is homogeneous in the radial ($x$) direction, as in
+the standard shearing box.
 
@@
 Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
-$\hat{g}_c=0$. The resulting linear problem is very similar 
+$\hat{g}_c=0$. The resulting problem is very similar 
 to that for local axisymmetric waves in accretion disks already
@@
 making use of Eq. \ref{vertical_shear}, to
-obtain an equation for $\delta v_z$ when $\Gamma=1$,
+obtain an equation for $\delta v_z$ for $\Gamma=1$,
 \begin{align}

### 2015-03-03 paper.tex
updating

@@
   astrophysical disks. However, the VSI requires a short thermal
-  timescale in a stably stratified disk. %which is usually assumed 
+  timescale in stably stratified disks. %which is usually assumed 
   A quantitative examination of this thermodynamic 
@@
   we generalize the linear, axisymmetric VSI in radially local but
-  vertically global models of accretion disks to include an energy
-  equation with a thermal relaxation term that restores the disk
+  vertically global models of accretion disks to include a thermal
+  relaxation term that restores the disk  
   temperature $T$ over a characteristic timescale
@@
   $\gamma$ is the adiabatic index. We apply our linear calculations 
-  to a disk model of the Minimum Mass Solar Nebulae (MMSN) with an 
-  estimate for the physical thermal timescales 
-  in the MMSN based on dust opacity. We find the MMSN is unstable to
-  the VSI at tens of AU with a characteristic growth times of $\sim
-  30$ local orbital periods.  
-  % This suggests the VSI is dynamically important in
-  % the outer parts of protoplanetary disks.  
+  to a disk model of the Minimum Mass Solar Nebula (MMSN) with an  
+  estimate for the thermal timescale $\beta$ in the MMSN based on dust
+  opacity. We find the MMSN is unstable to the VSI at tens of AU with
+  a characteristic growth times of $\sim 30$ local orbital periods. 
 \end{abstract}
@@
 
-However, MHD turbulence may not be applicable for all  
+However, MHD turbulence may not be applicable to all  
 accretion disks. For example, there may be regions in protoplanetary  
@@
 in three-dimensional (3D) disks in which the rotation frequency
-$\Omega$ depends on the height from the disk midplane. The VSI is the
-disk analog of the Goldreich-Schubert-Fricke instability
+$\Omega$ depends on the height $z$ from the disk midplane. This dependence
+occurs when the equilibrium pressure depends on both the temperature
+and density, making the disk baroclinic. It is thus generally
+expected that $\p_z\Omega\neq 0$ in astrophysical disks.
+
+The VSI can be regarded as the disk analog of the
+Goldreich-Schubert-Fricke instability 
 \citep[GSF,][]{goldreich67,fricke68} in differentially rotating stars.   
@@
 VSI have been performed by \cite{arlt04}, and more recently by 
-\cite{nelson13} and \cite{stoll14}. In these simulations the authors indeed observe the
-growth of vertically-global but radially-localized perturbations when
-the equilibrium rotation varied with height.
+\cite{nelson13} and \cite{stoll14}. In these simulations the authors
+indeed observe the growth of vertically-global but radially-localized
+perturbations when the equilibrium rotation varied with height.
 
@@
 $t_c\lesssim 0.01P_\mathrm{orb}$ was needed to observe instability. That is, the
-temperature profile of the disk essentially remains unchanged during
-the disk evolution. This has been confirmed by 
-\cite{stoll14}, who find the VSI is only sustained in radiation
-hydrodynamic simulations with the inclusion of heating from stellar
-irradiation, which resulted in an effective thermal timescale $t_c\ll
-P_\mathrm{orb}$.
+disk temperature essentially needs to remain unchanged during its
+evolution. This was further supported with radiation hydrodynamic
+simulations by \cite{stoll14}, who find the VSI is only sustained with
+the inclusion of heating from stellar irradiation, which resulted in
+an effective thermal timescale $t_c\ll P_\mathrm{orb}$.
 
@@
 the thermodynamic response of the disk if it is stably
-stratified. Since the latter is usually assumed for typical disk
-models,  
+stratified. Since the latter is usually
+assumed for typical disk models,  
 % disk thermodynamics except
@@
 %On the other hand, 
-Should a given disk be unstable to the VSI, it must pass through the
+Should a given disk become turbulent through the VSI, it must pass through the
 linear instability first. Thus, understanding the thermodynamic dependence
@@
 VSI carried out by \cite{nelson13} and \cite{mcnally14} which assume
-isothermal perturbations. Our models are also an extension of early
-analysis of \cite{urpin98} and \cite{urpin03}, which accounts for thermal
-relaxation, but considered vertically localized disturbances. Our
+isothermal perturbations. Our models are also an extension of earlier 
+analysis by \cite{urpin98} and \cite{urpin03}, which accounts for thermal
+evolution, but considered vertically localized disturbances. Our 
 vertically global approach is necessary to study disturbances with
@@
 dominant mode in protoplanetary disks with a realistic estimate for
-the thermodynamic timescale.
+the thermodynamic timescales.
 

### 2015-03-03 setup.tex
updating

@@
 vertically isothermal disks which is the relevant case for irradiated
-disks \citep{chiang97}.
+protoplanetary disks \citep{chiang97}.
 
@@
   \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s^2
-    -s\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]. 
+    -s\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]
 \end{align} 
-For typical model parameters the right hand side (RHS) is
-$O(\epsilon^2)$. Hence, instability in the adiabatic limit 
+implies stability. For typical model parameters the right hand
+side (RHS) is $O(\epsilon^2)$. Hence, instability in the adiabatic limit  
 requires $\gamma < \Gamma + O(\epsilon^2)$. Since we consider
@@
 In order to emphasize the radially-local approximation, we
-introduce the local Cartesian co-ordinate $(x,y,z)$ corresponding to
+introduce the local Cartesian co-ordinates $(x,y,z)$ corresponding to
 the the local $(r,\phi,z)$ directions, respectively. Then writing
@@
 \p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$
-using their expressions in the global disk in \S\ref{setup}. 
+using their expressions in the global disk given in \S\ref{setup}. 
 Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical

### 2015-03-04 appendix.tex
restructing, moving iso to appendix

@@
 In \S\ref{analytical} we used the vertically global shearing box
-\citep[VGSB,][]{mcnally14} to make analytic progress. % Here, we discuss limitations of this approach.   
+\citep[VGSB,][]{mcnally14} to make analytic progress.   
 The main difference between the linear problem in the VGSB 
@@
 equations and are reproduced here for convenience,
-% Using the notation and disk models in the main text, the
-% energy, density and $r$ momentum equations become 
+
 \begin{align}
@@
 
-%  To obtain the
-% linear problem in the VGSB, we take the radially-local 
-% approximation further and formally set $\hat{g}_c=0$, $r\to x$ and
-% $k_r\to k_x$. It is clear that this approximation improves with
-% increasing $|k_x|$.
-
 \subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
@@
 Eq. \ref{vertiso_gov_nondim}, which can then be written as
-% For adiabatic perturbations ($\gamma>1$, $\beta\to\infty$) we may
-% simply use the Soilberg-Hoiland critera to assess stability
-% (\S\ref{solberg}). Nevertheless, we briefly consider this case in the
-% local framework to demonstrate its limitations. For $\beta\to\infty$
-% the parameter $\chi = 1/\gamma$ and Eq. \ref{vertiso_gov_nondim} can
-% be written as   
+
 \begin{align}
@@
 
-% To 
-% avoid this, we must restrict our attention to short thermal relaxation
-% timescales ($\beta\lesssim1$) and/or neutrally-stratified disks
-% ($\gamma=\Gamma$) when $q\neq0$.
-
 \subsection{Effect of global radial gradients}
-% We may also reverse the above procedure by reinstating the `global
-% correction' terms proportional to $\hat{g}_c$ in the linear VGSB
-% equations. Of course, doing so implies the basic state is not uniform
-% in $x$, which would be inconsistent with the full VGSB equations in
-% equilibrium. Nevertheless, we illustrate in Fig. \ref{gcorr_compare} the effect of
-% these global radial gradients on the growth rate of the fundamental
-% VSI.
-
-% The global correction terms do not change the qualitative dependence
-% of the fundamental VSI growth rates on the thermal relaxation
-% timescale. The difference is negligble for $\beta\lesssim 1$ (i.e.
-% fast thermal relaxation). The critical relaxation timescale for the
-% fundamental VSI is unaffected by these global radial radients.
-
-% However, in the adiabatic limit $\beta\to\infty$ the linear VGSB
-% equations give a positive growth rate as discussed in
-% \S\ref{analytic_adia}, which is inconsistent with the second
-% Solberg-Hoiland criterion (Eq. \ref{solberg2}). The correct behaviour
-% ($\nu\to0$ as $\beta\to\infty$) is obtained with global correction.
 
@@
     accounts for the radial disk structure.  
-% computed from the linear VGSB equations used in
-%     the main text (dotted) and that including additional global
-%     radial gradient terms in
-%     Eq. \ref{gcorr_terms1}---\ref{gcorr_terms3} (solid). 
     The disk parameters are $\Gamma=1.011$,
@@
 
-%nevertheless we have calculated
-%plot growth rate v.s. bcool
-
 \section{Full governing equation for vertically isothermal disks in
   the VGSB}\label{adia_improve}
-% \section{Improving the nearly-Keplerian approximation for adiabatic
-%   disks}\label{adia_improve}
+
 In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
@@
 
-% \section{Alternative inference of instability in the presence of weak
-%   vertical shear}\label{pert_theory}
-% We can also investigate the destabilizing effect of vertical shear
-% without explicitly solving the full governing equation,
-% Eq. \ref{iso_ode3}. We first
-% consider a system with $q\equiv0$, for which the eigenfunctions are
-% Hermite polynomials and the real eigenfrequencies are known. We then
-% perturb this system by introducing a weak vertical shear ($|q|\ll1$)
-% and linearize the governing equation. This procedure is
-% \begin{align}   
-%   q \to 0 + \delta q,\quad
-%   W \to \He_n + \delta W,\quad
-%   \hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}. 
-% \end{align}
-% %where for this exercise we restore $\sigma$ as the eigenfrequency.
-
-% Linearizing the integral relation Eq. \ref{integral_relation1} in the
-% thin-disk limit, and taking the
-% imaginary part, we have
-% \begin{align}
-%   2\hat{\sigma}\imag(\delta\hat{\sigma})
-%   \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-%   \He_n^2(\hat{z}) d\hat{z}
-%   = \delta q \epsilon \hat{k} 
-%   \int_{-\infty}^{\infty}
-%   w(\zhat)\hat{z}\He_n(\hat{z})\He_n^\prime(\zhat) d\hat{z}
-% \end{align}
-% Recognizing $\hat{z} = \He_1(\hat{z})$ and using $\He_n^\prime = n
-% \He_{n-1}$ we have
-%  \begin{align}
-%    2\hat{\sigma}\imag(\delta\hat{\sigma})
-%    \left(1+\hat{k}^2\right) \int_{-\infty}^{\infty} w(\zhat)
-%    \He_n^2(\hat{z}) d\hat{z}
-%    =\delta q \epsilon \hat{k} 
-%    \int_{-\infty}^{\infty}
-%    w(\zhat)\He_1(\hat{z})\He_n(\hat{z})n\He_{n-1}(\hat{z})d\hat{z}. 
-%  \end{align}
-
-% Finally, we evaluate the integrals using the results
-% \begin{align}
-%   \int_{-\infty}^{\infty}
-%   w(\xi) \He_k(\xi)\He_l(\xi) d\xi = \sqrt{2\pi}k!\delta_{kl} \quad
-% % \end{align}
-% \text{and} \quad
-% % \begin{align}
-%   \int_{-\infty}^{\infty}
-%   w(\xi) \He_m(\xi)\He_k(\xi)\He_l(\xi) d\xi =
-%   \frac{\sqrt{2\pi}m!k!l!}{(j-m)!(j-k)!(j-l)!}, 
-% \end{align}
-% where $j = (m+k+l)/2$. We then obtain
-%  \begin{align}
-%    2\hat{\sigma}\imag(\delta\hat{\sigma})
-%   \left(1+\hat{k}^2\right) = \delta q \epsilon \hat{k} n.
-%  \end{align}
-% Inserting the original eigenfrequency $\hat{\sigma} = \pm
-% \sqrt{n}(1+\hat{k}^2)^{-1/2}$ gives
-% \begin{align}
-%   \imag(\delta\hat{\sigma})= \pm \frac{1}{2}\delta q \epsilon
-%   \sqrt{n} \frac{\hat{k}}{\sqrt{1+\hat{k}^2}}. 
-% \end{align}
-% This result agrees with Eq. \ref{simple_growth} in the limit
-% $|q|\to0$. 
-% %for $n=1$, since the
-% %function $\He_1$ solves the governing equation exactly. 
-% For $\hat{k}\gg 1$ we have
-% \begin{align}
-%   \imag(\delta\hat{\sigma})\simeq \pm \frac{1}{2}\delta q \epsilon
-%   \sqrt{n} \sgn{\hat{k}}. 
-% \end{align}
-
-% Differentiating Eq. \ref{adia_iso1} properly gives
-% \begin{align}
-%   \left(D + \gamma k_x^2 c_s^2\right)\frac{d\Delta}{dz}=& D
-%   \frac{d^2\delta v_z}{dz^2} + k_x^2c_s^2\left[\frac{\gamma}{\left(D + \gamma k_x^2 c_s^2\right)}\frac{dD}{dz} + \ii
-%     \frac{d\ln\rho}{dz}\left(\ii +
-%       \frac{q}{k_xr}\right)\right]\frac{d\delta v_z}{dz}\notag\\
-%   &+ \ii k_x^2 c_s^2 \left(\ii +
-%       \frac{q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} -
-%       \frac{1}{\left(D + \gamma
-%           k_x^2c_s^2\right)}\frac{dD}{dz}\frac{d\ln\rho}{dz}\right]\delta
-%     v_z. \label{adia_diso1}
-% \end{align}
-% The terms that were ignored in deriving Eq. \ref{adia_iso3} are those
-% proportional to $dD/dz$ in Eq. \ref{adia_diso1}. Now, in the global
-% disk with $\Gamma=1$ we have
-% \begin{align}\label{dkappa2}
-%   \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
-%     r}\frac{\p\Omega^2}{\p z} = -
-%   \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(2 + q +
-%     \frac{z^2/r^2-2}{z^2/r^2+1}\right). 
-% \end{align}
-% For the local problem we may then write
-% \begin{align}
-%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z;q),
-% \end{align}
-% where the function $F$ corresponds to the bracket in
-% Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
-% 3+q$ as $|z|\to\infty$, i.e. its magnitude is of order unity. 
-% % We estimate the
-% % importance of these terms by noting that 
-% % \begin{align}
-% %   \frac{dD}{dz}\equiv \frac{d\kappa^2}{dz} \simeq \frac{d\Omega^2}{dz}
-% %   = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2},
-% % \end{align}
-% % for thin disks. 
-% Eq. \ref{adia_diso1} becomes
-% \begin{align}
-% \left(D + \gamma k_x^2 c_s^2\right)\frac{d\Delta}{dz}=& D
-%   \frac{d^2\delta v_z}{dz^2} - k_x^2c_s^2\frac{d\ln\rho}{dz}\left[ 1 - 
-%       \frac{\ii q}{k_xr}  +  \frac{\gamma q c_s^2F}{r^2\left(D +
-%         \gamma k_x^2 c_s^2\right)}\right]\frac{d\delta
-%     v_z}{dz}\notag\\ 
-%   &- k_x^2 c_s^2 \left(1 - 
-%       \frac{\ii q}{k_xr}\right)\left[\frac{d^2\ln\rho}{dz^2} + 
-%       \frac{qc_s^2F}{r^2\left(D + \gamma
-%           k_x^2c_s^2\right)}\left(\frac{d\ln\rho}{dz}\right)^2\right]\delta
-%     v_z. \label{adia_diso2}
-% \end{align}
-% Since $D\sim \Omega_k^2$ in the low-frequency limit, we 
-% see from Eq. \ref{adia_diso2} that the neglected terms are
-% $O(\epsilon^2)$. We can make this explicit by combining
-% Eq. \ref{adia_diso1}---\ref{adia_diso2} and Eq. \ref{adia_iso2}, then
-% set $D\to\Omega_k^2$. In terms of non-dimensional variables, the result
-% is  
-% \begin{align}
-%    &\delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon q\hat{k} -
-%     \frac{\gamma q \epsilon^2\hat{k}^2F}{1+\gamma
-%       \hat{k}^2}\right)\ln\rho^{\prime}\delta v_z^\prime +
-%   \left[\left(\frac{1}{\gamma} + \ii \epsilon q
-%       \hat{k}\right)\ln\rho^{\prime\prime} - \hat{k}^2\left(1 -
-%       \frac{\ii\epsilon
-%         q}{\hat{k}}\right)\left(\frac{\gamma-1}{\gamma} +
-%       \frac{q\epsilon^2F}{1+\gamma\hat{k}^2}\right)\ln\rho^{\prime
-%       2}\right]\delta v_z \notag\\&=
-%   -\hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right)\delta v_z\label{adia_diso3}.
-% \end{align}   
-% Eq. \ref{adia_diso3} differs from Eq. \ref{adia_iso3} by terms
-% proportional to $\epsilon^2$. For a thin disk, $\epsilon\ll1$, so
-% neglecting these terms has no qualitative effect on the discussion in
-% \S\ref{analytic_adia}.
-
-% \section{Axisymmetric instability in vertically 
-%   isothermal disks in global cylindrical geometry}\label{global_vertiso}
-% Consider a locally isothermal disk in 3D such that $P=c_s^2\rho$ at
-% all times, so in the perturbed state $\delta P =
-% c_s^2\delta\rho$, with $c_s$ begin a prescribed function of position. 
-% The axisymmetric linearized fluid equations in cylindrical
-% geometry is then 
-% \begin{align}
-%  \ii\sigma \delta v_r + 2\Omega\delta v_\phi &= c_s^2\frac{\p}{\p r}\left(\frac{\delta\rho}{\rho}\right),\\
-%  \ii\sigma \delta v_\phi &= \frac{1}{r}\delta\bm{v}\cdot\nabla j,\\
-%  \ii\sigma \delta v_z &= c_s^2\frac{\p}{\p
-%    z}\left(\frac{\delta\rho}{\rho}\right),\\
-%  \ii\sigma\delta\rho &= \nabla\cdot\left(\rho\delta\bm{v}\right),
-% \end{align}
-% and we recall $j=r^2\Omega$. We can eliminate $\delta v_\phi$ from these equations to obtain
-% \begin{align}
-%   -\sigma^2\rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) +
-%   \kappa^2\rho|\delta v_r|^2 + \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
-%   \delta v_r^*\delta v_z =
-%   c_s^2\rho\delta\bm{v}^*\cdot\nabla\left(\frac{\ii\sigma\delta\rho}{\rho}\right)
-%   =c_s^2\rho\delta\bm{v}^*\cdot\nabla\left[\frac{1}{\rho}\nabla\cdot\left(\rho\delta\bm{v}\right)\right].\label{global_integral} 
-% \end{align}
-% At this point if we make the anelastic approximation by assuming
-% $\nabla\cdot\left(\rho\delta\bm{v}\right)$ is negligible in this
-% equation \citep[as argued by][]{nelson13}, 
-% then the only contribution to instability comes from the third term on
-% the left hand side, i.e. the explicit vertical shear term
-% $\p_zj$. This is the same conclusion reached in
-% \S\ref{integral_relation}.
-
-% The term proportional to $\kappa^2$ is stabilizing. Because radial shear
-% is much larger than vertical shear, this implies that significant
-% instability can only be achieved with large vertical
-% motions (i.e. $|\delta v_z|\gg|\delta v_r|$).
-
-% If we integrate Eq. \ref{global_integral} over the fluid volume 
-% and assume boundary terms vanish when integrating by parts, we obtain 
-% \begin{align}
-%   -\sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV +
-%   \int \kappa^2\rho|\delta v_r|^2 dV + \int \frac{2\Omega}{r}\frac{\p j}{\p z} \rho
-%   \delta v_r^*\delta v_z dV = -\int
-%   \frac{c_s^2}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
-%   dV - \int \nabla\cdot\left(\rho\delta\bm{v}\right) \delta
-%   \bm{v}^*\cdot\nabla c_s^2.
-% \end{align}
-% Note that the treatment so far is in fact valid for any prescribed
-% $c_s(r,z)$. However, to proceed further we specialize to the case
-% $c_s=c_s(r)$ for simplicity, i.e. vertically isothermal
-% disks. Recalling $r\p_z\Omega^2 = -\p_z\ln{\rho}\p_r
-% c_s^2$  for a vertically isothermal disk (Eq. \ref{vertical_shear}), we obtain 
-% \begin{align}
-%   \sigma^2\int \rho\left(|\delta v_r|^2 + |\delta v_z|^2\right) dV =
-%   \int \kappa^2\rho|\delta v_r|^2 dV + \int
-%   \frac{c_s^2(r)}{\rho}\left|\nabla\cdot\left(\rho\delta\bm{v}\right)\right|^2
-%   dV + \int \left[ \rho \left(\nabla\cdot\delta\bm{v}\right)\delta
-%     v_r^* + |\delta v_r|^2\frac{\p \rho}{\p r}\right]\frac{dc_s^2(r)}{dr}  dV.
-% \end{align}
-% We see that instability is only possible if
-% $dc_s^2/dr\neq0$. Furthermore, since density and temperature typically
-% decrease outwards in astrophysical disks, instability is associated
-% with the term proportional to $\nabla\cdot\delta\bm{v}$. It is also
-% clear that $\sigma^2$ will be complex in general.
-
 \section{Coefficients for the dispersion relation for perturbations
@@
 \end{align}
+
+\section{Linear problem in isothermal limit}\label{iso_discuss}  
+Here we examine in more detail the special case of isothermal
+perturbations ($\beta\equiv 0$) in the VGSB framework.   
+%Isothermal perturbations correspond to $\beta\equiv0$ ($\chi=1$) and 
+In this case Eq. \ref{ode_Q} becomes   
+\begin{align} 
+  Q = W 
+\end{align}
+(i.e. $\delta P = c_s^2\delta \rho$), and so $  \ii\hat{\sigma}c_s\delta v_z = W^\prime $
+from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact
+simpler to work with an equation for $W$ by substituting $Q=W$ into
+Eq. \ref{ode_w}. We obtain 
+\begin{align}
+  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
+    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
+\end{align}
+where for discussion purposes we have defined $f(\zhat)$ such that
+\begin{align}\label{fz_shear}
+  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
+\end{align}
+By comparing Eq. \ref{fz_shear} with Eq. \ref{vertical_shear_ex}, we
+see that $f(\zhat)= 
+\zhat\left(1+\epsilon^2\zhat^2\right)^{-3/2}$. More generally, though,
+$f$ may be considered as a representation of the vertical
+shear profile. Physically, we expect there is a maximum value of 
+$|d\Omega/dz|$. The existence of a maximum, as remarked in
+\S\ref{vshear_def}, limits the growth rate of the VSI. We explicitly
+demonstrate this below.
+
+\subsection{Necessity of vertical shear for
+  instability}\label{integral_relation} 
+We can establish a necessary condition for instability by multiplying
+Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
+$\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
+terms when integrating by parts, by assuming $W$ or
+$W^\prime$ vanishes at the boundaries, or that the boundaries are 
+sufficiently far away so that the boundary terms are negligible because of the
+decaying background density with increasing height. Then,
+\begin{align}
+  \hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat
+  =\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
+  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
+\end{align}
+It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
+have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
+shear. A similar conclusion can be reached in the equivalent linear  
+problem in global cylindrical disk geometry.
+
+\subsection{Maximum growth rate}   
+Here we show that the growth rate is limited by the maximum vertical
+shear in the domain. % Writing $\hat{\sigma} = \hat{\omega} +
+% \ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, 
+The real and imaginary parts of 
+Eq. \ref{integral_relation1} are
+\begin{align}
+  \left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
+  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
+  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
+  \notag
+  =\real\left[\ii
+    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
+    f(\hat{z}) W^*W^\prime d\zhat\right], \\
+   2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
+  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}
+  =\imag\left[\ii
+    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
+    f(\hat{z}) W^*W^\prime d\hat{z}\right],
+\end{align}
+where we recall $\hat{\omega} = \real \hat{\sigma}$ and
+$\hat{\nu}=\imag\hat{\sigma}$. 
+Adding the square of these equations give
+\begin{align}
+  \left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
+    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
+    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2
+  +4\hat{\nu}^2\left(1+\hat{k}^2\right) 
+  \int_{\zhat_1}^{\zhat_2}\rho
+  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}
+  =\left|\ii
+    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
+    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
+\end{align}
+It is clear that
+\begin{align}\label{sigma_finite_domain} 
+  4\hat{\nu}^2\left(1+\hat{k}^2\right) 
+  \int_{\zhat_1}^{\zhat_2}\rho
+  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime
+  \right|^2d\hat{z} 
+  \leq\left|
+    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
+    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
+\end{align}
+On the left hand side of this inequality, we apply the Cauchy-Schwarz
+inequality to obtain
+\begin{align}
+  \left( \int_{\zhat_1}^{\zhat_2}\rho
+    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
+  \int_{\zhat_1}^{\zhat_2}\rho 
+  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
+\end{align}
+On the right hand side of Eq. \ref{sigma_finite_domain} we have
+\begin{align}
+  \left|\int_{\zhat_1}^{\zhat_2}\rho
+    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
+  \left|f(\hat{z})W^*W^\prime \right|d\hat{z}
+  \leq
+  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
+  |W|\left|W^\prime \right|d\hat{z},
+\end{align}
+where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
+$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
+Eq. \ref{sigma_finite_domain} gives
+\begin{align}\label{max_growth}
+  |\hat{\nu}|\leq
+  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
+\end{align}
+Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
+follows that the maximum possible growth rate of unstable modes,
+satisfying the above boundary conditions, is limited by the maximum
+vertical shear in the domain considered.
+
+In practice, one might consider a vertical domain of a few scale 
+heights in a thin disk with $|q|=O(1)$. In this case, $|\epsilon\zhat|\ll1$ and  
+$f\simeq \zhat$, so that $\mathrm{max}|f| = O(1)$, implying a
+maximum growth rate $O(\epsilon \Omega_k)$, consistent with numerical
+results.
+
+% For vertical domains smaller 
+% than this height, the growth rate is limited by $|f|$ at the vertical
+% boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
+% \hat{z}$, in which case we can expect growth rates to increase at most linearly
+% with respect to the vertical domain size.   %no faster than linear
+
+\subsection{Thin-disk limit}
+For thin disks ($\epsilon\ll1$), in which 
+$\ln\rho\simeq -\zhat^2/2$ and $f(\zhat)\simeq \zhat$, 
+% by expanding the background density and vertical shear profile in powers
+% of $z/r$. To lowest order we obtain 
+% \begin{align}
+%   &\frac{\rho}{\rho_0} \simeq 
+%   \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv w(\hat{z}),\label{thin_dens}\\
+%   &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}, \label{thin_vshear}
+% \end{align} 
+%With this approximation and those in \S\ref{integral_relation}, 
+Eq. \ref{iso_ode1} becomes  
+\begin{align}\label{iso_ode3}
+  W^{\prime\prime} - \left(1 + \ii q\epsilon
+    \hat{k}\right)\hat{z}W^\prime  +
+  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
+  0.
+\end{align}
+We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
+that derived by \cite{nelson13}, although we have taken a different
+route.
+
+To complete the problem we must specify boundary conditions. The 
+maximum vertical domain size should be limited by the thin
+disk and low-frequency approximations (i.e. $z\lesssim r$).     
+It is, however, common to take $\mathrm{max}|z|\to\infty$ and impose
+that the kinetic  energy remain bounded at infinity. This 
+is appropriate for the density field because both the true density profile 
+and its thin-disk approximation decay rapidly away from the midplane.
+
+However, the thin-disk representation of vertical shear diverges with 
+$\hat{z}$, unlike the true vertical shear profile which eventually
+decays. Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit
+unphysically large growth rates (as demonstrated below). Nevertheless,
+the infinite disk is useful to consider since it permits an analytical
+discussion.
+
+\subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
+When $q=0$, Eq. \ref{iso_ode3} is
+\begin{align}\label{hermite_ode}
+  \left[w(\hat{z})W^\prime \right]^\prime + nW
+  w(\hat{z}) =0, 
+\end{align}
+where $w(\zhat)=\exp{(-\zhat^2/2)}$ and we have defined a new eigenvalue 
+\begin{align}
+  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
+\end{align} 
+Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
+that the kinetic energy density remain bounded at infinity, then the
+solutions are Hermite polynomials  
+\begin{align}
+  W \propto \He_n(\hat{z}),
+\end{align}
+and $n$ is a non-negative integer. This translates to a real
+eigenfrequency $\sigma$ such that
+\begin{align}
+  \left|\hat{\sigma}\right| = \sqrt{n}
+  \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
+\end{align}
+Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
+analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
+n-1$. Physically, this corresponds to radial length-scales much
+smaller than the local disk scale height. Note that
+Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
+\cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
+Keplerian disk approximation in that equation.
+
+We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
+Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
+has also been qualitatively observed in numerical simulations
+\citep{stoll14}.
+
+\subsubsection{Polynomial solutions with vertical
+  shear}\label{iso_poly}
+More generally, we can seek power-series solution to
+Eq. \ref{iso_ode3}, 
+\begin{align}
+  W(\zhat) = \sum_{l=0}^\infty a_l\zhat^l. 
+\end{align}
+Then the coefficients must satisfy the recurrence relation
+\begin{align}
+  (l+2)(l+1)a_{l+2} +
+  \left[n - l\left(1+\ii \epsilon q  \hat{k}\right)\right] a_l = 0, 
+\end{align}
+where $n$ is not necessarily an integer. Indeed, if we demand
+a polynomial of degree $L$ as the solution, then the eigenfrequency
+must satisfy
+\begin{align}\label{sig2_iso}
+  \hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
+      \hat{k}}{1+\hat{k}^2}\right).
+\end{align}
+The eigenfrequency $\hat{\sigma}$ is therefore complex for
+$L\geq1$. The dispersion relation Eq. \ref{sig2_iso} can also be
+derived by setting $\beta=0$ in the more general dispersion relation 
+Eq. \ref{relax_disp} with coefficients given in Appendix
+\ref{relax_coeff}. We are interested in 
+unstable modes for which $\hat{\nu}>0$ and is given by
+\begin{align}\label{simple_growth}
+  \hat{\nu} =\sqrt{
+    \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
+      1\right)}. 
+\end{align}
+For fixed $q$ and $\epsilon$, the growth rate vanishes for both
+$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
+The maximum growth rate
+occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
+\begin{align}
+  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
+\end{align}
+and
+\begin{align}
+  \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
+    q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
+\end{align}
+Since $L\geq1$, there is no limit to the maximum growth rate. This is
+an artifact of the thin-disk approximation because the vertical 
+shear $d\Omega^2/dz\propto z$ increases indefinitely with 
+height. However, Eq. \ref{iso_max_growth} compares well with
+numerical solutions for small $L$ (see \S\ref{vertiso_pertiso}).
 

### 2015-03-04 application.tex
restructing, moving iso to appendix

@@
 fraction of the vertical domain. However, the normalized growth times
-are much longer than the case with $\beta < \beta_\mathrm{crit}$ for
-all $z$: we find $t_\mathrm{grow}\sim 40$ orbits at $r=50$AU,
+are much longer than the case with $\beta < \beta_\mathrm{crit}  \forall z$: we find $t_\mathrm{grow}\sim 40$ orbits at $r=50$AU,
 but $t_\mathrm{grow}\sim 400$ orbits at $r=5$AU. Thus, introducing an

### 2015-03-04 linear_adia.tex
restructing, moving iso to appendix

@@
-\subsection{Perturbations with thermal relaxation in the thin-disk limit}\label{analytic_relax}
-We now introduce thermal relaxation ($\beta>0$), but pursue explicit
-solutions in the thin-disk approximation for simplicity. The governing
-equation, Eq. \ref{vertiso_gov_nondim}, is then 
+\subsection{Thin-disk approximation}\label{analytic_relax}
+Our final approximation apply to thin disks for 
+which $\epsilon\ll 1$. Then we may write $\rho/\rho_0\simeq
+\exp{(-\zhat^2/2)}$. The governing equation,
+Eq. \ref{vertiso_gov_nondim}, is then   
 \begin{align}
@@
 $\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
-and making the low-frequency, Keplerian disk approximation in
+and making the low-frequency, nearly-Keplerian approximation in
 \citeauthor{lubow93} after a change of variables.
 
-\subsubsection{The effect of introducing a small but finite
+The thin disk approximation allows us to develop explicit solutions,
+but some general results may be established without it (see,
+e.g. Appendix \ref{global_corr} and \S\ref{iso_pert}).
+
+\subsection{The effect of introducing a small but finite
   thermal relaxation time}\label{relax_pert}
@@
 sufficiently small $\beta$, for which the perturbations may be 
-considered nearly-isothermal, we expect the new solutions to only
-differ slightly from the explicit isothermal solutions developed in
-\S\ref{iso_poly}.
+considered nearly-isothermal, we expect the solution to only
+differ slightly from a case with $\beta\equiv 0$.
 
@@
 relaxation, we linearize Eq. \ref{nearly_iso_explicit} about a known 
-solution for $\beta\equiv0$  ($\chi\equiv 1$). 
-Let 
+solution for $\beta\equiv0$, which we choose below. Let  
 \begin{align}\label{nearly_iso_pert}
@@
 \end{align}
-Here $\delta v_z$ corresponds to an explicit solution developed in
-\S\ref{iso_poly} with  $\hat{\sigma}^2$ given by Eq. \ref{sig2_iso}. Inserting
-Eq. \ref{nearly_iso_pert} into Eq. \ref{nearly_iso_explicit}  and keeping
-only first order terms, we obtain 
+Inserting Eq. \ref{nearly_iso_pert} into Eq. \ref{nearly_iso_explicit}
+and keeping only first order terms, we obtain  
 \begin{align}
@@
 
-We could proceed by considering the entire set of polynomial solutions
-for $W$, and hence for $\delta v_z$, as derived in \S\ref{iso_poly}, but it
-is simplest to look at the fundamental mode. Thus we consider the `basic state' 
- \begin{align*}
-   \delta v_z = 1,\quad \hat{\sigma}^2\left(1+\hat{k}^2\right) = 1 +
-   \ii\epsilon q \hat{k},
- \end{align*}
- then we may seek 
+In Eq. \ref{nearly_iso_pert} and \ref{nearly_iso4}, $\delta v_z$ is to
+a solution to Eq. \ref{nearly_iso_explicit} with 
+$\beta=0$. For definiteness, let us consider the simple solution 
+\begin{align}
+  \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\sigma}^2 = \frac{1 +
+  \ii\epsilon q \hat{k}}{1+\hat{k}^2}, \label{pert_basic} 
+\end{align}
+which solves Eq. \ref{nearly_iso_explicit} because $\delta
+v_z=$constant, $\chi=1$ so that $C=0$, and the eigenfrequency makes
+$B=0$. Then we may seek 
 \begin{align}
@@
 where $a$, $b$ are constants.
+
+% We could proceed by considering the entire set of polynomial solutions
+% for $W$, and hence for $\delta v_z$, as derived in \S\ref{iso_poly}, but it
+% is simplest to look at the fundamental mode.
 
@@
 
-\subsubsection{Dispersion relation with thermal
-  relaxation}\label{disp_relax}
-To obtain a dispersion relation similar to that for 
-isothermal perturbations (Eq. \ref{sig2_iso}), we first write 
+\subsection{Explicit solutions and dispersion relation}\label{disp_relax}
+%To obtain a dispersion relation similar to that for 
+%isothermal perturbations (Eq. \ref{sig2_iso}), 
+We now solve Eq. \ref{nearly_iso_explicit} explicitly. We first write  
 \begin{align}
@@
 $\rho|\delta v_z|^2\propto |g|^2 \exp{\left(\real\alpha -
-    1/2\right)\zhat^2}$ to remain finite. Then assuming $g(\zhat)$
-is a polynomial, we require  
+    1/2\right)\zhat^2}$ to remain finite as $|\zhat|\to\infty$. 
+Then assuming $g(\zhat)$ is a polynomial, we require  
 \begin{align}
@@
 
-We now have an equation in the same form as that for isothermal
-perturbations (Eq. \ref{iso_ode3}):
+Eq. \ref{nearly_iso_explicit} becomes 
 \begin{align}
@@
 
-\subsubsection{Upper limit on the thermal relaxation timescale for the
+\subsection{Upper limit on the thermal relaxation timescale for the
   fundamental VSI mode}\label{iso_vsi_beta_crit}
@@
 This is a pair of simultaneous equations for
-$\hat{\omega}$ and $\beta$. % A simple solution can be obtained by the
-% following considerations.
-Since our discussion of stable solutions  
-without vertical shear yield a dispersion
-relation $\hat{\omega}\propto \khat^{-1}$ for $\khat\gg 1$
-(\S\ref{iso_stable}, Eq. \ref{iso_stable_disp}), we are motivated to
-seek solutions to 
+$\hat{\omega}$ and $\beta$.  
+% Since our discussion of stable solutions without vertical shear yield
+% a dispersion relation $\hat{\omega}\propto \khat^{-1}$ for $\khat\gg
+% 1$ (\S\ref{iso_stable}, Eq. \ref{iso_stable_disp}).
+We seek solutions to 
 Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
@@
 dependence numerically in \S\ref{bcrit_num_test}. 
-From our discussion for isothermal perturbations
-(\S\ref{iso_discuss}), we expect   
-$\beta<\beta_\mathrm{crit}$ to be a sufficient condition for the
-VSI. That is, when $\beta<\beta_\mathrm{crit}$, perturbations are
+From our discussion in \S\ref{relax_pert} we expect   
+$\beta<\beta_\mathrm{crit}$ is required for the fundamental mode to be
+unstable. That is, when $\beta<\beta_\mathrm{crit}$, perturbations are 
 effectively isothermal.
@@
 finite thermal relaxation rapidly stabilizes the VSI.
+
+\subsection{Isothermal perturbations}\label{iso_pert}
 

### 2015-03-04 linear_iso.tex
restructing, moving iso to appendix

@@
 where $\Gamma=1$, then $c_s$ is a constant in the radially-local
-approximation. We examine isothermal perturbations ($\beta =
-0$) before considering the effect of finite thermal relaxation. 
+approximation. % We examine isothermal perturbations ($\beta =
+% 0$) before considering the effect of finite thermal relaxation. 
 In order to facilitate a fully analytic discussion, we first introduce a
@@
 
-\subsection{Governing equation in the VGSB}\label{approx_gov}
+\subsection{Low-frequency and nearly-Keplerian
+  approximation}\label{approx_gov} 
 We now derive a governing equation with further approximations:
@@
   further by ignoring the vertical dependence of $\Omega$ and $\kappa$
-  elsewhere. For clarity, we replace $\kappa^2$ with $\Omega_k^2$.
+  elsewhere. For clarity, we replace $\kappa^2$ with
+  $\Omega_k^2$. We call this the nearly-Keplerian approximation.
 
@@
 where $^\prime$ denotes $d/d\zhat$. Eq. \ref{vertiso_gov_nondim} will
-be the basis for our discussion of vertically isothermal disks. % ,
-% although further simplifications or maniputations will be performed
-% depending the type of perturbations being considered.
+be the basis for our discussion of vertically isothermal disks.
 
@@
 
-% We will consider sound-speed profiles with $q<0$ in the global 
-% disk. Then $\Omega^2(z)$ decreases away from the midplane,
-% and may reach small values for sufficiently large $|z|$, which
-% invalidates the low-frequency approximation. In principle, then, the
-% above approximations implicitly limits us to consider thin disks ($\epsilon
-% \ll 1$) with vertical domain sizes not too large, so that the rotation
-% is nearly-Keplerian.
-
 Note that we can generalize Eq. \ref{vertiso_gov_nondim} non-Keplerian
@@
 $\hat{\sigma}$ as $\sigma/\kappa$, and  $\epsilon$ as $c_s/r\kappa$.
-
-% Consider rotation profiles with radial dependence
-% $\propto r^{-m}$. Then the same dimensionless governing equation 
-% (Eq. \ref{vertiso_gov_nondim}) holds, provided we redfine $\khat$
-% as $k_xc_s/\kappa$, $\hat{\sigma}$ as $\sigma/\kappa$, and reset
-% $\epsilon$ to $\epsilon/sqrt{4-2m}$.
-
-\subsection{Isothermal perturbations}\label{iso_discuss}
-Isothermal perturbations correspond to $\beta\equiv0$ 
-%relaxation timescale $\be
-%or for any adiabatc index $\gamma$ when
-%$\beta=0$. 
-($\chi=1$) and Eq. \ref{ode_Q} becomes  
-\begin{align}
-  Q = W 
-\end{align}
-(i.e. $\delta P = c_s^2\delta \rho$), and so  
-\begin{align}
-  \ii\hat{\sigma}c_s\delta v_z = W^\prime
-  % \quad   \text{(isothermal perturbations)},
-\end{align}
-from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact 
-simpler to work with $W$ by integrating Eq. \ref{vertiso_gov_nondim}. For later discussion
-it is convenient to write the result as
-
-\begin{align}
-  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
-    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
-\end{align}
-where $f(\zhat)$ is defined such that
-\begin{align}\label{fz_shear}
-  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
-\end{align}
-For a vertically isothermal disk we have 
-\begin{align}
-  f(\zhat) = \frac{\zhat}{\left(1+\epsilon^2\hat{z}^2\right)^{3/2}}, 
-\end{align}
-%For the vertically isothermal disk we have $|f(\hat{z})| =
-%|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, 
-which maximizes at $|\zhat|=1/\sqrt{2}\epsilon$. The existence of a
-maximum vertical shear limits the growth rate of unstable modes, as  
-shown below.
-
-Note that Eq. \ref{iso_ode1} can in fact be derived by substituting
-$Q=W$ in Eq. \ref{ode_w} without approximating $D$ or $\kappa^2$ as
-vertically constant.
-
-% the issue of ignoring the
-% vertical dependence of $D$ described in \S\ref{approx_gov} and Appendix
-% \ref{adia_improve} is irrelevant, since
-% Eq. \ref{iso_ode1} may be derived directly , i.e. no derivatives of $D$ needs to be taken.
-
-%We remark that Eq. \ref{iso_ode1} is also applicable for  $\gamma=\Gamma\neq 1$ in the
-%limit $t_c\to\infty$ (i.e. neutrally stratified adiabatic disks),
-%since in that case $Q\simeq W$ as well. However, the following
-%discussion is only  valid for $\gamma=\Gamma=1$ because $c_s$ is taken
-%to be a constant.
-
-\subsubsection{Necessity of vertical shear for
-  instability}\label{integral_relation} 
-We can establish a necessary condition for instability by multiplying
-Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
-$\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
-terms when integrating by parts, by assuming $W$ or
-$W^\prime$ vanishes at the boundaries, or that the boundaries are 
-sufficiently far away so that the boundary terms are negligible because of the
-decaying background density with increasing height. Then,
-\begin{align}
-  &\hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat \notag\\
-  &=\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
-  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
-\end{align}
-It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
-have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
-shear. A similar conclusion can be reached in the equivalent linear  
-problem in global cylindrical disk geometry.
-
-\subsubsection{Maximum growth rate}   
-Here we show that the growth rate is limited by the maximum vertical
-shear in the domain. Writing $\hat{\sigma} = \hat{\omega} +
-\ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, the real and
-imaginary parts of 
-Eq. \ref{integral_relation1} are
-\begin{align}
-  &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
-  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
-  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
-  \notag\\
-  &=\real\left[\ii
-    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\zhat\right], \notag\\
-  & 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
-  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}\notag\\
-  &=\imag\left[\ii
-    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right].
-\end{align}
-Adding the square of these equations give
-\begin{align}
-  &\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
-    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
-    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2\notag\\
-  &+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
-  \int_{\zhat_1}^{\zhat_2}\rho
-  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
-  &=\left|\ii
-    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
-\end{align}
-It is clear that
-\begin{align}\label{sigma_finite_domain} 
-  &4\hat{\nu}^2\left(1+\hat{k}^2\right) 
-  \int_{\zhat_1}^{\zhat_2}\rho
-  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
-  &\leq\left|
-    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
-\end{align}
-On the left hand side of this inequality, we apply the Cauchy-Schwarz
-inequality to obtain
-\begin{align}
-  &\left( \int_{\zhat_1}^{\zhat_2}\rho
-    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
-  \int_{\zhat_1}^{\zhat_2}\rho 
-  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
-\end{align}
-On the right hand side of Eq. \ref{sigma_finite_domain} we have
-\begin{align}
-  \left|\int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
-  \left|f(\hat{z})W^*W^\prime \right|d\hat{z} \notag\\
-  \leq
-  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
-  |W|\left|W^\prime \right|d\hat{z},
-\end{align}
-where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
-$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
-Eq. \ref{sigma_finite_domain} gives
-\begin{align}\label{max_growth}
-  |\hat{\nu}|\leq
-  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
-\end{align}
-Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
-follows that the maximum possible growth rate of unstable modes,
-satisfying the above boundary conditions, is limited by the maximum
-vertical shear in the domain considered.
-
-In practice, one might consider a vertical domain of a few scale 
-heights in a thin disk with $|q|=O(1)$. In this case, $|\epsilon\zhat|\ll1$ and  
-$f\simeq \zhat$, so that $\mathrm{max}|f| = O(1)$, implying a
-maximum growth rate $O(\epsilon \Omega_k)$, consistent with numerical
-results below.
-
-% For vertical domains smaller 
-% than this height, the growth rate is limited by $|f|$ at the vertical
-% boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
-% \hat{z}$, in which case we can expect growth rates to increase at most linearly
-% with respect to the vertical domain size.   %no faster than linear
-
-\subsubsection{Thin-disk limit}
-We can make further progress for thin disks ($\epsilon\ll1$), 
-by expanding the background density and vertical shear profile in powers
-of $z/r$. To lowest order we obtain 
-\begin{align}
-  &\frac{\rho}{\rho_0} \simeq 
-  \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv w(\hat{z}),\label{thin_dens}\\
-  &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}, \label{thin_vshear}
-\end{align} 
-%With this approximation and those in \S\ref{integral_relation}, 
-and the governing equation becomes 
-\begin{align}\label{iso_ode3}
-  W^{\prime\prime} - \left(1 + \ii q\epsilon
-    \hat{k}\right)\hat{z}W^\prime  +
-  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
-  0.
-\end{align}
-We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
-that derived by \cite{nelson13}, although we have taken a different
-route.
-
-To complete the problem we must specify boundary conditions. The 
-maximum vertical domain size should be limited by the thin
-disk and low-frequency approximations (i.e. $z\lesssim r$).     
-It is, however, common to take $\mathrm{max}|z|\to\infty$ and impose
-that the kinetic  energy remain bounded at infinity. This 
-is appropriate for the density field because both the true density profile 
-and its thin-disk approximation decay rapidly away from the midplane.
-
-However, the thin-disk representation of vertical shear diverges with 
-$\hat{z}$, unlike the true vertical shear profile which eventually
-decays.  Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit unphysically large growth
-rates (as demonstrated below). Nevertheless, the infinite disk is
-useful to consider since it permits an analytical discussion.
-
-\subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
-When $q=0$, Eq. \ref{iso_ode3} is
-\begin{align}\label{hermite_ode}
-  \left[w(\hat{z})W^\prime \right]^\prime + nW
-  w(\hat{z}) =0, 
-\end{align}
-where we have defined a new eigenvalue
-\begin{align}
-  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
-\end{align} 
-Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
-that the kinetic energy density remain bounded at infinity, then the
-solutions are Hermite polynomials  
-\begin{align}
-  W \propto \He_n(\hat{z}),
-\end{align}
-and $n$ is a non-negative integer. This translates to a real
-eigenfrequency $\sigma$ such that
-\begin{align}
-  \left|\hat{\sigma}\right| = \sqrt{n}
-  \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
-\end{align}
-Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
-analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
-n-1$. Physically, this corresponds to radial length-scales much
-smaller than the local disk scale height. Note that
-Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
-\cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
-Keplerian disk approximation in that equation.
-
-We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
-Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
-has also been qualitatively observed in numerical simulations
-\citep{stoll14}.
-
-\subsubsection{Polynomial solutions with vertical shear}\label{iso_poly}
-We can seek power-series solution to Eq. \ref{iso_ode3},
-\begin{align}
-  W(\zhat) = \sum_{l=0}^\infty a_l\zhat^l. 
-\end{align}
-Then the coefficients must satisfy the recurrence relation
-\begin{align}
-  (l+2)(l+1)a_{l+2} +
-  \left[n - l\left(1+\ii \epsilon q  \hat{k}\right)\right] a_l = 0, 
-\end{align}
-where $n$ is not necessarily an integer. Indeed, if we demand
-a polynomial of degree $L$ as the solution, then the eigenfrequency
-must satisfy
-\begin{align}\label{sig2_iso}
-\hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
-    \hat{k}}{1+\hat{k}^2}\right).
-\end{align}
-The eigenfrequency $\hat{\sigma}$ is therefore complex for
-$L\geq1$. We are interested in 
-unstable modes for which $\hat{\nu}=\imag{\hat{\sigma}}>0$ is the
-growth rate and is given by 
-\begin{align}\label{simple_growth}
-  \hat{\nu} =\sqrt{
-   \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
-    1\right)}. 
-\end{align}
-For fixed $q$ and $\epsilon$, the growth rate vanishes for both
-$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
-The maximum growth rate
-occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
-\begin{align}
-  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
-\end{align}
-and
-\begin{align}
-  \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
-    q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
-\end{align}
-Since $L\geq1$, there is no limit to the maximum growth rate. This is
-an artifact of the thin-disk approximation because the vertical 
-shear $d\Omega^2/dz\propto z$ increases indefinitely with
-height. However, we find Eq. \ref{iso_max_growth} compares well with
-numerical solutions of the governing equation for small $L$.
 

### 2015-03-04 setup.tex
restructing, moving iso to appendix

@@
 \end{align}
-Note that the disk is in general baroclinic with $\Omega =
-\Omega(r,z)$. This is explicit from Eq. \ref{eqm_rot}, but is also
-implied by Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} and 
-Eq. \ref{eqm_press},  
-\begin{align}
-  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
-    z}\frac{dK}{dr} = - \frac{g_z}{\Gamma}\frac{d\ln K}{dr}.\label{vertical_shear}
-\end{align}
 
@@
 protoplanetary disks \citep{chiang97}.
+
+\subsection{Vertical shear}\label{vshear_def}
+Note that the disk is in general baroclinic with $\Omega =
+\Omega(r,z)$. This is explicit from Eq. \ref{eqm_rot} and Eq. \ref{vertiso_eqm}, but is also 
+implied by Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} and 
+Eq. \ref{eqm_press},  
+\begin{align}
+  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
+    z}\frac{dK}{dr} = - \frac{g_z}{\Gamma}\frac{d\ln K}{dr}.\label{vertical_shear}
+\end{align}
+Evaluating this expression explicitly,
+\begin{align}
+  r\frac{\p\Omega^2}{\p z} = \frac{sz\Omega_k^2}{\Gamma
+    r\left(1+z^2/r^2\right)^{3/2}}. \label{vertical_shear_ex} 
+\end{align}
+Note that there is a maximum vertical shear rate at $z=r/\sqrt{2}$,
+which is expected to limit the growth rate of unstable modes.
 

### 2015-03-04 codes/cool_full/bloop_mmsn3D/vsi.f90

@@ -1,7 +1,7 @@
 module global
   character*6 :: vbc
   logical :: gterms, eigen_trial, eigenv_out, eigen_refine
-  integer :: nz, nzeff, bignz, nkx, nb, ntrials, output_freq
+  integer :: nz, nzeff, bignz, nkx, nb, ntrials, output_freq, kaphat
   real*8, parameter :: pi = 2d0*acos(0d0), mu=2.33d0
   real*8 :: smallq, smallp, eps, smallh, kx, zmax, gmma, bgmma, bcool, smalls, dist 
   real*8 :: kxmin, kxmax, dlogkx, bmin, bmax, db, ghat
@@ -30,7 +30,7 @@ program vsi
   complex*16,allocatable :: work(:), matrix(:,:), w(:), vl(:,:), &
        vr(:,:), rhs(:,:), wtrial(:), wtrial_tot(:)
   real*8, external :: dlogrho_dz, domega2_dz, omega2_z, logrho_z, kappa2_z, d2logrho_dz2, csq_z, gcorr_z 
-  namelist /params/ smallq, smallp, dist, gmma, bgmma, zmax, vbc, gterms
+  namelist /params/ smallq, smallp, dist, gmma, bgmma, zmax, vbc, gterms, kaphat
   namelist /loop/ kxmin, kxmax, nkx, bmin, bmax, nb, eigen_trial, eigenv_out, eigen_refine
   namelist /grid/ nz
   
@@ -252,10 +252,12 @@ program vsi
         kx = kaxis(k)*smallh !convert input into code units 
         
      
-        bcool_arr(:)=(4.4d5/(mu*(gmma-1d0)))*dist**(-57./14.)
-        bcool_arr(:)=bcool_arr(:)*(8.3d-9*dist**(33./7.)+2d0*pi*exp(logrho(:))**2d0/kaxis(k)**2d0)
+        bcool_arr(:)=(4.4d5*kaphat/(mu*(gmma-1d0)))*dist**(-57./14.)
+        bcool_arr(:)=bcool_arr(:)*(8.3d-9*dist**(33./7.)/kaphat**2d0 + exp(logrho(:))**2d0/kaxis(k)**2d0/2d0/pi)
 
-
+!        bcool_arr(:)=bcool_arr(:)*(8.3d-9*dist**(33./7.)+ 1d0/kaxis(k)**2d0) 
+ 
+ 
         !set up sub-matrices for linear operators
         do i=1, nzeff
[…]

### 2015-03-04 appendix.tex
restructured

@@
 route.  
-
-To complete the problem we must specify boundary conditions. The 
-maximum vertical domain size should be limited by the thin
-disk and low-frequency approximations (i.e. $z\lesssim r$).     
-It is, however, common to take $\mathrm{max}|z|\to\infty$ and impose
-that the kinetic  energy remain bounded at infinity. This 
-is appropriate for the density field because both the true density profile 
+% To complete the problem we must specify boundary conditions. The 
+% maximum vertical domain size should be limited by the thin
+% disk and low-frequency approximations (i.e. $z\lesssim r$).     
+%It is, however, common to take $\mathrm{max}|z|\to\infty$ and 
+We impose the kinetic energy remain bounded at infinity. Even though
+we have adopted the thin-disk approximation, an infinite
+disk is appropriate in the sense that both the true density profile 
 and its thin-disk approximation decay rapidly away from the midplane.
@@
 \cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
-Keplerian disk approximation in that equation.
+nearly-Keplerian approximation in that equation.
 
@@
   shear}\label{iso_poly}
-More generally, we can seek power-series solution to
+In the presence of vertical shear ($q\neq0$) we can still seek power-series solution to
 Eq. \ref{iso_ode3}, 
@@
 Since $L\geq1$, there is no limit to the maximum growth rate. This is
-an artifact of the thin-disk approximation because the vertical 
-shear $d\Omega^2/dz\propto z$ increases indefinitely with 
+an artifact of the thin-disk approximation because the magnitude of vertical 
+shear $|d\Omega^2/dz|\propto |z|$ increases indefinitely with 
 height. However, Eq. \ref{iso_max_growth} compares well with

### 2015-03-04 application.tex
restructured

@@
 
-\subsubsection{Comparing the VSI in the inner and outer disk} 
-Here we compare the VSI at $r=50$AU and $r=5$AU. We first show in
+\subsection{Comparing the VSI in the inner and outer parts of the MMSN} 
+Here we compare the VSI in the MMSN at $r=50$AU and $r=5$AU. We first show in
 Fig. \ref{beta_compare}  the thermal timescale $\beta$ for a
@@
     $\delta v_z$ is shown as
-    the solid (dotted) curves. The eigenfunction has been normalized
-    such that $\imag(\delta v_z)=0$ at the vertical boundaries, and
-    $\mathrm{max}|\delta v_z|=1$. \label{mmsn_eigenvz}}
+    the solid (dotted) curves. % The eigenfunction has been normalized
+    % such that $\imag(\delta v_z)=0$ at the vertical boundaries, and
+    % $\mathrm{max}|\delta v_z|=1$. 
+    \label{mmsn_eigenvz}}
 \end{figure}

### 2015-03-04 linear_adia.tex
restructured

@@
 
-The thin disk approximation allows us to develop explicit solutions,
-but some general results may be established without it (see,
-e.g. Appendix \ref{global_corr} and \S\ref{iso_pert}).
+% The thin disk approximation allows us to develop explicit solutions,
+% but some general results may be established without it (see,
+% e.g. Appendix \ref{global_corr} and \S\ref{iso_pert}).
+
+\subsection{Isothermal perturbations}\label{iso_pert}
+Isothermal perturbations correspond to $\beta = 0$ or $\chi = 1$. This
+case has been investigated numerically by \cite{nelson13} and 
+\cite{mcnally14} in the thin disk limit, the latter of which employed
+the VGSB framework.
+
+In Appendix \S\ref{iso_discuss}, we give an analytical description of the
+linear problem with $\beta=0$. There, we establish more general
+results that do not require the thin disk approximation, namely 
+a necessary condition for instability and an upper bound to the
+instability growth rate. We also develop analytic solutions
+with and without vertical shear in the thin disk limit, including
+explicit expressions for wave frequencies and growth rates, the
+methods and results for which motivates the solution procedure
+employed in the following sections.
 
@@
   thermal relaxation time}\label{relax_pert}
-It is instructive to first ask  the more analytically tractable
-question: how 
+We are interested in the case $\beta\neq 0$, but it is instructive to
+first ask  the more analytically tractable question: how 
 do the eigenfrequencies and eigenfunctions change when we change
@@
 
+For definiteness, let us consider the simple solution 
+\begin{align}
+  \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\sigma}^2 = \frac{1 +
+  \ii\epsilon q \hat{k}}{1+\hat{k}^2}, \label{pert_basic} 
+\end{align}
+which solves Eq. \ref{nearly_iso_explicit} since $\delta
+v_z=$constant and $\chi=1$ so that $B=C=0$. 
+This is the fundamental or `corrugation' mode and represents
+the entire disk column moving vertically. These modes have been
+observed to dominate numerical simulations \citep{nelson13,stoll14},
+and are the ones we find to develop in protoplanetary disks
+with a realistic estimate for thermal timescales
+(\S\ref{application}).
+
 To determine the (assumed) small effect of introducing a short thermal
-relaxation, we linearize Eq. \ref{nearly_iso_explicit} about a known 
-solution for $\beta\equiv0$, which we choose below. Let  
+relaxation, we linearize Eq. \ref{nearly_iso_explicit} about the above
+solution for $\beta\equiv0$. We write 
 \begin{align}\label{nearly_iso_pert}
@@
 \begin{align}
-  \chi \to 1 + \delta\chi = 1 + \ii \hat{\sigma}\left(\gamma-1\right)\delta\beta.
+  \chi \to 1 + \delta\chi = 1 + \ii \hat{\sigma}\left(\gamma-1\right)\delta\beta,
 \end{align}
-Inserting Eq. \ref{nearly_iso_pert} into Eq. \ref{nearly_iso_explicit}
-and keeping only first order terms, we obtain  
-\begin{align}
- 0 =& \delta v_{z1}^{\prime\prime} - z\left(1 + \ii\epsilon q
-   k\right)\delta v_{z1}^\prime\notag\\
- &+ \left[2\hat{\sigma}\delta\hat{\sigma}\left(1 + \khat^2\right) +
-   \left(\hat{\sigma}^2 - 1\right)\delta\chi \right.\notag\\
-&\phantom{++} \left. + \zhat^2\left(\khat^2 -
-     \ii\epsilon q \khat\right)\delta\chi\right]\delta v_z\notag\\
- &+\left[\hat{\sigma}^2\left(1 + \khat^2\right) - \left(1+\ii\epsilon
-     q \khat\right)\right]\delta v_{z1}.\label{nearly_iso4}
-\end{align}
-
-In Eq. \ref{nearly_iso_pert} and \ref{nearly_iso4}, $\delta v_z$ is to
-a solution to Eq. \ref{nearly_iso_explicit} with 
-$\beta=0$. For definiteness, let us consider the simple solution 
-\begin{align}
-  \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\sigma}^2 = \frac{1 +
-  \ii\epsilon q \hat{k}}{1+\hat{k}^2}, \label{pert_basic} 
-\end{align}
-which solves Eq. \ref{nearly_iso_explicit} because $\delta
-v_z=$constant, $\chi=1$ so that $C=0$, and the eigenfrequency makes
-$B=0$. Then we may seek 
+with $\delta v_z$ and $\hat{\sigma}$ given by Eq. \ref{pert_basic}. We
+may then seek  
 \begin{align}
@@
 
+Inserting Eq. \ref{nearly_iso_pert} into Eq. \ref{nearly_iso_explicit}
+and keeping only first order terms yield the requirements, after
+balancing constants and coefficients of $\zhat^2$, 
+% we obtain  
+% \begin{align}
+%  0 =& \delta v_{z1}^{\prime\prime} - z\left(1 + \ii\epsilon q
+%    k\right)\delta v_{z1}^\prime\notag\\
+%  &+ \left[2\hat{\sigma}\delta\hat{\sigma}\left(1 + \khat^2\right) +
+%    \left(\hat{\sigma}^2 - 1\right)\delta\chi \right.\notag\\
+% &\phantom{++} \left. + \zhat^2\left(\khat^2 -
+%      \ii\epsilon q \khat\right)\delta\chi\right]\delta v_z\notag\\
+%  &+\left[\hat{\sigma}^2\left(1 + \khat^2\right) - \left(1+\ii\epsilon
+%      q \khat\right)\right]\delta v_{z1}.\label{nearly_iso4}
+% \end{align}
+% In Eq. \ref{nearly_iso_pert} and \ref{nearly_iso4}, $\delta v_z$ is to
+% a solution to Eq. \ref{nearly_iso_explicit} with 
+% $\beta=0$. 
 % We could proceed by considering the entire set of polynomial solutions
@@
 % is simplest to look at the fundamental mode. 
-
-Inserting the above expressions for $\delta v_z$, $\delta v_{z1}$ and
-$\hat{\sigma}^2$ into Eq. \ref{nearly_iso4}, then balancing
-coefficients of $\zhat^2$ and constants, we require
+%Inserting the above expressions for $\delta v_z$, $\delta v_{z1}$ and
+%$\hat{\sigma}^2$ into Eq. \ref{nearly_iso4}, then balancing
+%coefficients of $\zhat^2$ and constants, we require
 \begin{align}
@@
 This is a pair of simultaneous equations for
-$\hat{\omega}$ and $\beta$.  
-% Since our discussion of stable solutions without vertical shear yield
-% a dispersion relation $\hat{\omega}\propto \khat^{-1}$ for $\khat\gg
-% 1$ (\S\ref{iso_stable}, Eq. \ref{iso_stable_disp}).
-We seek solutions to 
-Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
+$\hat{\omega}$ and $\beta$. In Appendix \ref {iso_stable} we discuss
+stable solutions without vertical shear which yield
+$\hat{\omega}\propto \khat^{-1}$ for $\khat\gg 1$
+(Eq. \ref{iso_stable_disp}),  we are thus motivated to seek solutions
+to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
 % we wish to seek a criterion independent of k 
@@
 that
-\begin{align}
+\begin{align}\label{beta_crit0}
   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
@@
 at marginal stability for perturbations with large $\khat$.
-For a thin disk $\epsilon\ll 1$, we infer 
+
+From our discussion in \S\ref{relax_pert}, which indicate introducing
+a finite thermal relaxation is stabilizing, we expect 
+Eq. \ref{beta_crit0} is an upper limit for instability. For a thin
+disk $\epsilon\ll 1$, we thus infer 
 \begin{align}\label{iso_vsi_cond}
@@
 \end{align}
-is required for the fundamental VSI to operate. We will test this
-dependence numerically in \S\ref{bcrit_num_test}. 
-From our discussion in \S\ref{relax_pert} we expect   
-$\beta<\beta_\mathrm{crit}$ is required for the fundamental mode to be
-unstable. That is, when $\beta<\beta_\mathrm{crit}$, perturbations are 
-effectively isothermal.
+is required for the fundamental VSI to operate. That is, when
+$\beta<\beta_\mathrm{crit}$, perturbations are  effectively isothermal. 
+We will test Eq. \ref{iso_vsi_cond} numerically in
+\S\ref{bcrit_num_test}.
 
@@
 finite thermal relaxation rapidly stabilizes the VSI.
-
-\subsection{Isothermal perturbations}\label{iso_pert}
 

### 2015-03-04 summary.tex
restructured

@@
 allows the thermodynamic response of the disk to be controlled.
+However, it is possible to model $\beta$ self-consistently by
+relating it to thermal timescales expected in realistic disks.
 
@@
 This is convenient because rapid thermal relaxation ($\beta\ll1$) is
-generally required for the VSI.  However,
-the VGSB framework is not appropriate in the adiabatic limit ($\beta \gg
-1$).
+generally required for the VSI.  However, the VGSB framework is not
+appropriate in the adiabatic limit ($\beta \gg 1$).
 
@@
 this is the relevant case for protoplanetary disks that are irradiated
-\citep{chiang97}. It also permits a largely analytical discussion.  Our analysis is an
-extension of the linear models considered in \cite{lubow93} by 
-adding vertical shear and (rapid) thermal relaxation, but in the limit of low-frequency
-perturbations in nearly-Keplerian disks. These approximations were
-made for simplicity, but also motivated by recent numerical
-simulations \citep{nelson13}.
+\citep{chiang97}. It also permits a largely analytical discussion.
+Our analysis is an extension of the linear models considered in
+\cite{lubow93} by  adding vertical shear and (rapid) thermal
+relaxation, but in the limit of low-frequency perturbations in
+nearly-Keplerian disks. These approximations were made for simplicity,
+but also motivated by recent numerical simulations
+\citep{nelson13}. We obtained explicit solutions for the VSI with
+thermal relaxation in the thin disk limit.
 
-For isothermal perturbations ($\beta=0$) we explicitly
-demonstrated instability  associated with vertical shear, and that the 
-VSI growth rate is limited by the maximum vertical shear within the 
-vertical domain of interest. The fundamental VSI mode is characterized
-by one node in the density perturbation at the midplane 
-(i.e. anti-symmetric about $z=0$), which corresponds to a vertical
-velocity perturbation  $\delta v_z\sim\mathrm{constant}$ near the
-mid-plane with zero nodes. These have been referred to as
-`corrugation modes' by \cite{nelson13}, and were found to dominate
-their numerical simulations, as well as that of
+% For isothermal perturbations ($\beta=0$) we explicitly
+% demonstrated instability  associated with vertical shear, and that the 
+% VSI growth rate is limited by the maximum vertical shear within the 
+% vertical domain of interest. 
+% For perturbations with finite thermal relaxation, the general
+% dispersion relation (Eq. \ref{relax_disp}) is quite
+% complicated. However, for the fundamental VSI some simplification is
+% possible. In particular,
+
+The fundamental VSI mode is characterized by a vertical velocity
+perturbation with zero nodes and $\delta v_z\sim\mathrm{constant}$
+near the mid-plane, and by one node in the density perturbation at the
+midplane  (i.e. anti-symmetric about $z=0$).  These have been referred
+to as `corrugation modes' by \cite{nelson13}, and were found to
+dominate their numerical simulations, as well as that of
 \cite{stoll14}. Furthermore, when computing unstable modes for a
-disk with a physical model for thermal relaxation, we also find
-the fundamental mode to be dominant and robust. Such modes are thus
-likely most relevant to  realistic disks.
+protoplanetary disk model with thermal relaxation based on realistic
+disk properties, we find the fundamental mode to be dominant and
+robust against boundary conditions. The fundamental mode is thus
+likely the most relevant VSI mode in protoplanetary disks.
 
-For non-isothermal  perturbations with finite thermal relaxation ($\gamma>1$,
-$\beta>0$) , the general dispersion relation (Eq. \ref{relax_disp}) is
-quite complicated. However, for the 
-fundamental VSI some simplification is  possible. In particular, we
-quantified the critical thermal relaxation timescale
+We quantified the critical thermal relaxation timescale 
 $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$, below which small-scale
 perturbations may be regarded as isothermal and the fundamental VSI
-operates.  We expect $\beta<\beta_\mathrm{crit}$ to be a sufficient
-condition for the VSI to operate. We have therefore provided a
-quantitative measure for the`fast' thermal relaxation requirement for the
-VSI.
+operates. %  We expect $\beta<\beta_\mathrm{crit}$ to be a sufficient
+% condition for the VSI to operate.
+We have therefore provided a quantitative measure for the `fast'
+thermal relaxation requirement for the VSI in astrophysical disks.
 
@@
 dust opacity, which results in a $\beta$ that depends on both the
-lengthscale of interest as well as the height from the midplane.  
+perturbation lengthscale as well as the height from the midplane.  
 We find the MMSN is unstable to the fundamental VSI at $10$s of AU,

### 2015-03-04 application.tex
restructured

@@
     the MMSN such that the associated dimensionless thermal
-    relaxation time  $\beta$ at $z=0$ (thick lines) and $z=3H$ (thin
-    lines) is less than the critical timescale $\beta_\mathrm{crit}$.  
+    relaxation time  $\beta$ at $z=0$ (solid) and $z=3H$ (dashed) is
+    less than the critical timescale $\beta_\mathrm{crit}$.   
     \label{mmsn_bcrit_bcool_mink}}   
@@
   \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
-    (thick solid) and $r=5$AU (thick dotted) for $\khat=30$. The
-    corresponding thin horizontal lines are the critical thermal
+     and $r=5$AU for $\khat=30$ (solid lines). The
+    corresponding thin horizontal dashed lines are the critical thermal
     relaxation timescales derived in linear theory. 

### 2015-03-04 paper.tex
restructured

@@
   opacity. We find the MMSN is unstable to the VSI at tens of AU with 
-  a characteristic growth times of $\sim 30$ local orbital periods. 
+  a characteristic growth time of $\sim 30$ local orbital
+  periods, and the dominant disturbance in the MMSN is the `corrugation' mode
+  which casues the disk column to oscillate vertically. 
 \end{abstract}

### 2015-03-04 application.tex
restructured
appendix.tex

@@
 temperature fluctuations are smoothed out by radiative diffusion. We
-mode this in linear theory by writing the linearized RHS of
+model this in linear theory by writing the linearized RHS of
 Eq. \ref{real_energy} as 
@@
 the lower panels of Fig. \ref{mmsn_bcrit_bcool}, decreases the thermal
-timescales, but shifts the curves outward, so the fundamental mode is
+timescales, but shifts the curves outward, so the instability is
 restricted to larger radii.
@@
 by choosing a sufficiently large $\khat$. We
-illustrate this by plotting at each radius the value of $\khat$ such
+illustrate this in Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at
+each radius the value of $\khat$ such 
 that $\beta = \beta_\mathrm{crit}$. We thus expect that the VSI,
@@
 \citep{shakura73}. Setting $t_\mathrm{visc}$ to be longer than the
-typical VSI growth timescale $1/\epsilon\Omega$ places an upper limit
-$\khat \lesssim 30$ in a disk with $\alpha_s\sim 10^{-4}$ and
-$\epsilon \sim 0.1$.
+minimum VSI growth timescale $1/\epsilon\Omega$ places an upper limit
+$\khat^2\lesssim \epsilon/\alpha_s$. Inserting $\epsilon \sim 0.05$,
+we find only modes with $\khat\lesssim 20$ are relevant for
+$\alpha_s\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha_s\sim
+10^{-5}$. \citep[These low viscosity levels are required for the
+VSI, see][]{nelson13}.
 
@@
 \subsection{VSI in the MMSN}
-Here we solve the linear problem in the MMSN with the scale and
+Here we solve the linear stability problem in the MMSN with the scale and
 height dependent thermal relaxation function defined by
@@
 \citep{nelson13,mcnally14}. Such modes may not be robust because    
-they are affected by boundary conditions. % , and high $\khat$ modes are
-% subject to viscous decay in a real disk.
+they are affected by boundary conditions and viscosity in a real
+disk. % , and high $\khat$ modes are % subject to viscous decay in a real disk.
 
 We therefore consider Fig. \ref{mmsn_overall} as a representation of the
-VSI in the MMSN. We see that growth timescales are $\sim
-30$---40 orbits, and the VSI operates from a few AU to tens of
-AU. Importantly, it is necessary to consider smaller scales as $r$
+VSI in the MMSN. We see the VSI operates from $\sim 5$AU to $\sim
+50$AU with growth timescales $\sim 30$---40 orbits. 
+Importantly, it is necessary to consider smaller scales as $r$
 decreases as the disk generally becomes optically thick.
@@
 $\beta \gtrsim \beta_\mathrm{crit}$ for $r\lesssim 10$AU at $z=0$; and
-Fig. \ref{mmsn_overall} show stabilization inside $\sim 8$AU.
+Fig. \ref{mmsn_overall} show stabilization inside $\sim 8$AU where
+$t_\mathrm{grow}\to\infty$.
 
@@
 \subsection{Comparing the VSI in the inner and outer parts of the MMSN} 
-Here we compare the VSI in the MMSN at $r=50$AU and $r=5$AU. We first show in
+We compare the VSI in the MMSN at $r=50$AU and $r=5$AU. We first show in
 Fig. \ref{beta_compare}  the thermal timescale $\beta$ for a
@@
 that $\beta < \beta_\mathrm{crit}$ for $|z|\gtrsim1.25H$. In
-fact, in this case  we also find the VSI to operate, but with 
-significantly reduced growth rates (when scaled by the local
+fact, in this case  we also find the VSI to operate, but with much
+reduced growth rates (when scaled by the local 
 rotation).  
@@
 fraction of the vertical domain. However, the normalized growth times
-are much longer than the case with $\beta < \beta_\mathrm{crit}  \forall z$: we find $t_\mathrm{grow}\sim 40$ orbits at $r=50$AU,
-but $t_\mathrm{grow}\sim 400$ orbits at $r=5$AU. Thus, introducing an
-optically thick midplane reduces the efficiency of the VSI despite
-$\beta < \beta_\mathrm{crit}$ away from the midplane.
+are much longer at $r=5$AU than at $r=50$AU for which  $\beta <
+\beta_\mathrm{crit}  \forall z$: we find $t_\mathrm{grow}\sim 40$
+orbits at $r=50$AU, but $t_\mathrm{grow}\sim 400$ orbits at
+$r=5$AU. Thus, introducing an optically thick midplane significantly
+reduces the efficiency of the VSI despite $\beta <
+\beta_\mathrm{crit}$ away from the midplane.
 

### 2015-03-04 linear_adia.tex
restructured
appendix.tex

@@
 case has been investigated numerically by \cite{nelson13} and 
-\cite{mcnally14} in the thin disk limit, the latter of which employed
-the VGSB framework.
+\cite{mcnally14}, the latter of which employed the VGSB framework.
 
-In Appendix \S\ref{iso_discuss}, we give an analytical description of the
+In Appendix \ref{iso_discuss}, we give an analytical description of the
 linear problem with $\beta=0$. There, we establish more general
 results that do not require the thin disk approximation, namely 
-a necessary condition for instability and an upper bound to the
-instability growth rate. We also develop analytic solutions
-with and without vertical shear in the thin disk limit, including
-explicit expressions for wave frequencies and growth rates, the
-methods and results for which motivates the solution procedure
-employed in the following sections.
+a necessary condition for instability (there must be vertical shear)
+and an upper bound to the instability growth rate (it is limited by
+the maximum vertical shear rate in the disk). We also develop analytic
+solutions with and without vertical shear in the thin disk limit,
+including explicit expressions for wave frequencies and growth rates. 
+The methods used in Appendix \ref{iso_discuss} and results obtained
+therein motivates the solution procedure employed in the following
+sections.
 
@@
 $t_c=\beta\Omega_k^{-1}$ from zero to a small but finite value? For 
-sufficiently small $\beta$, for which the perturbations may be 
-considered nearly-isothermal, we expect the solution to only
-differ slightly from a case with $\beta\equiv 0$.
+sufficiently small $\beta$ % , for which the perturbations may be 
+% considered nearly-isothermal,
+we expect the solution to only differ slightly from a case with
+$\beta\equiv 0$. We thus perturb a solution for $\beta=0$ to see the
+effect of thermal relaxation.
 
@@
 
-To determine the (assumed) small effect of introducing a short thermal
-relaxation, we linearize Eq. \ref{nearly_iso_explicit} about the above
-solution for $\beta\equiv0$. We write 
+%To determine the (assumed) small effect of introducing a short thermal
+%relaxation, 
+We linearize Eq. \ref{nearly_iso_explicit} about the above
+solution for $\beta\equiv0$ and write 
 \begin{align}\label{nearly_iso_pert}
@@
 
-Introducing a finite thermal relaxation time introduces buoyancy
+A finite thermal relaxation time introduces buoyancy
 forces, which is stabilizing for sub-adiabatically stratified disks
 ($\gamma > 1$). The $\zhat^2$ dependence in $\delta v_{z1}$ then makes
-sense because it is most significant for large $\zhat$, i.e. away from
+sense because it becomes significant for large $|\zhat|$, i.e. away from
 the midplane where the effect of buoyancy first appears as $\beta$ is
@@
 Here we estimate the maximum thermal relaxation timescale
-$\beta$ that allows the fundamental VSI for $\khat\gg1$.   
+$\beta$ that allows the fundamental VSI mode for $\khat\gg1$.   
 % Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
@@
 $\hat{\omega}\propto \khat^{-1}$ for $\khat\gg 1$
-(Eq. \ref{iso_stable_disp}),  we are thus motivated to seek solutions
+(Eq. \ref{iso_stable_disp}).  We are thus motivated to seek solutions
 to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
@@
 a finite thermal relaxation is stabilizing, we expect
-Eq. \ref{beta_crit0} is an upper limit for instability. For a thin
+Eq. \ref{beta_crit0} is an upper limit. For a thin  
 disk $\epsilon\ll 1$, we thus infer 
@@
 $\beta<\beta_\mathrm{crit}$, perturbations are  effectively isothermal. 
-We will test Eq. \ref{iso_vsi_cond} numerically in
+We will confirm Eq. \ref{iso_vsi_cond} numerically in
 \S\ref{bcrit_num_test}.

### 2015-03-04 numerical.tex
restructured
appendix.tex

@@
 disks subject to isothermal perturbations ($\Gamma=\gamma=1$).  This
-allows us to test our numerical code.
+allows us to test our numerical code. We also compare results with our 
+analytical discussion of isothermal perturbations in Appendix
+\ref{iso_discuss}.
 
@@
 
-{\bf Growth rates are expected to diminish for both $\khat\to 0$
+ Growth rates are expected to diminish for both $\khat\to 0$
   and $\khat\to \infty$. This is because for $\khat\ll 1$ the radial
@@
   consider axisymmetric perturbations).
-}
 
@@
 expected for the fundamental mode according to explicit solutions
-developed in \S\ref{iso_poly}. For completeness we show the
+developed in Appendix \ref{iso_poly}. For completeness we show the
 corresponding eigenfunctions in vertical velocity in
@@
 shown) we find the perturbation is almost entirely concentrated at the
-boundaries. Our radially-local approximation is not valid for 
-these small wavenumber perturbations and we do not consider such modes
-further.
+boundaries. However, our radially-local approximation is not valid for 
+such small wavenumber perturbations.
 
@@
     wavenumber $\khat=1$ (top) and $\khat=20\pi$ (bottom). The real 
-    (imaginary) parts of $W$ are plotted as solid (dotted) lines. The
-    eigenfunction has been normalized such that $\imag(W)=0$ at the
-    vertical boundaries and $\mathrm{max}|W|=1$. 
+    (imaginary) parts of $W$ are plotted as solid (dotted) lines. % The
+    % eigenfunction has been normalized such that $\imag(W)=0$ at the
+    % vertical boundaries and $\mathrm{max}|W|=1$. 
     \label{lowfreq_eigenfunc}
@@
 $\sigma = \omega + \ii\nu$ found for $\khat=20\pi$. This plot is
-similar to Fig. 3 in \cite{mcnally14}. The fundamental has the
+similar to Fig. 3 in \cite{mcnally14}. The fundamental mode has the
 smallest $|\sigma|$, for which the eigenvalue calculated from
@@
 rapidly decreases as $\beta$ approaches and increases beyond
-$\beta_\mathrm{crit}$. In the latter case we find the perturbation is
-concentrated at the vertical boundaries, which may play a role not
+$\beta_\mathrm{crit}$. In the latter case we find the perturbation
+acquires large amplitudes near the vertical boundaries. Boundary
+conditions may thus may play a role for these modes that is not
 accounted by our analytical discussion.
@@
 perturbation $\delta v_z$ for
-$\beta=0.01$ and $\beta=0.1$. The latter value is close to the
-critical value beyond which the fundamental VSI is stabilized
-(\S\ref{iso_vsi_beta_crit}). Increasing $\beta$ 
+$\beta=0.01$ and $\beta=0.1$. The latter value is close to
+$\beta_\mathrm{crit}=0.125$. 
+% the
+% critical value beyond which the fundamental VSI is stabilized
+% (\S\ref{iso_vsi_beta_crit}). 
+Increasing $\beta$ 
 restricts  the region in which $\delta v_z\sim $ constant closer to the midplane, with
@@
     $\delta v_z$ is plotted as the solid (dotted)
-    line. The eigenfunction has been normalized such that
-    $\imag(\delta v_z)=0$ at the vertical boundaries, and
-    $\mathrm{max}|\delta v_z|=1$.
+    line. % The eigenfunction has been normalized such that
+    % $\imag(\delta v_z)=0$ at the vertical boundaries, and
+    % $\mathrm{max}|\delta v_z|=1$.
     \label{relax_eigenW_num}}  

### 2015-03-04 paper.tex
appendix.tex

--- 2015-03-03 paper.tex
@@
   the VSI in realistic disks. To this end,  
-  we generalize the linear, axisymmetric VSI in radially local but
-  vertically global models of accretion disks to include a thermal
-  relaxation term that restores the disk  
-  temperature $T$ over a characteristic timescale
+  we study the linear, axisymmetric VSI in radially local but
+  vertically global models of accretion disks in which the disk  
+  temperature $T$ is restored to its initial value over a characteristic timescale 
   $t_c=\beta\Omega_k^{-1}$, where $\beta$ is a dimensionless parameter
@@
   the VSI depends on the thermodynamic response  of the disk by
-  varying $\beta$. We find increasing $\beta$ even slightly away from
+  varying thermal timescale $\beta$. We find increasing $\beta$ even slightly away from 
   zero can dramatically reduce VSI growth rates. % , since this generally 
@@
   % motions as needed for the VSI.
-  Our  linear calculations are consistent with previous hydrodynamic
-  simulations. For vertically isothermal disks, we quantify the
+  For vertically isothermal disks, we quantify the
   requirement for `fast' thermal relaxation as $\beta < \epsilon|q|/(\gamma-1)$, where
@@
   being the cylindrical radius from the central object), and 
-  $\gamma$ is the adiabatic index. We apply our linear calculations 
+  $\gamma$ is the adiabatic index. Our  linear calculations are consistent with previous hydrodynamic 
+  simulations. We apply our linear calculations  
   to a disk model of the Minimum Mass Solar Nebula (MMSN) with an  
   estimate for the thermal timescale $\beta$ in the MMSN based on dust
-  opacity. We find the MMSN is unstable to the VSI at tens of AU with
-  a characteristic growth times of $\sim 30$ local orbital periods. 
+  opacity. We find the MMSN is unstable to the VSI from $\sim 5$AU to $\sim 50$AU with  
+  a characteristic growth time of $\sim 30$ local orbital
+  periods, and the dominant disturbance in the MMSN is the `corrugation' mode
+  which casues the disk column to oscillate vertically. 
 \end{abstract}
@@
 important to understand and quantify said thermodynamic dependence,
-for example on disk parameters, in order to evaluate the viability of
+for example on disk parameters, in order to evaluate the relevance of
 the VSI in realistic disks. % Numerical simulations of the VSI are

### 2015-03-04 setup.tex
appendix.tex

@@
 $H$ is a characteristic scale-height.  Thus, $\epsilon$
-is also a characteristic disk aspect-ratio.  In practice, we consider
-thin disks with $\epsilon \ll 1$.
+is also a characteristic disk aspect-ratio. We are 
+interested in thin disks with $\epsilon \ll 1$.%, as relevant for protoplanetary disks \citep{chiang97}.
 
@@
 taking the limit $\Gamma\to 1$. We will, in fact, primarily focus on
-vertically isothermal disks which is the relevant case for irradiated
+vertically isothermal disks which is the relevant case for passively irradiated
 protoplanetary disks \citep{chiang97}.
@@
 Note that there is a maximum vertical shear rate at $z=r/\sqrt{2}$,
-which is expected to limit the growth rate of unstable modes.
+which is expected to limit the growth rate of instabilities associated with vertical shear.
 
@@
 entropy gradient in order to be unstable to axisymmetric adiabatic
-perturbations. This is generally not true for irradiated, stably
+perturbations. This is generally not true for passively irradiated, stably
 stratified disks \citep{chiang97}, in which the VSI requires rapid
@@
 \p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$
-using their expressions in the global disk given in \S\ref{setup}. 
+using their expressions in the global disk given in \S\ref{setup}. Thus, they are functions of $z$ only. 
 Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical

### 2015-03-04 summary.tex
appendix.tex

@@
 employed in the global numerical simulations of \cite{nelson13}, and
-allows the thermodynamic response of the disk to be controlled.
-However, it is possible to model $\beta$ self-consistently by
-relating it to thermal timescales expected in realistic disks.
+allows the thermodynamic response of the disk to be controlled by
+varying the parameter $\beta$. However, it is possible to model
+$\beta$ self-consistently by relating it to thermal timescales
+expected in realistic disks.
 
@@
 \cite{nelson13} in a locally isothermal disk can be captured by the
-VGSB. We find the VGSB framework is appropriate for $\beta\lesssim1$. 
+VGSB. We find the VGSB framework is appropriate for non-isothermal
+perturbations provided $\beta\lesssim1$.  
 This is convenient because rapid thermal relaxation ($\beta\ll1$) is
@@
 We mostly focused on vertically isothermal disks ($\Gamma=1$), since
-this is the relevant case for protoplanetary disks that are irradiated
+this is the relevant case for protoplanetary disks that are passively irradiated by its central star 
 \citep{chiang97}. It also permits a largely analytical discussion.
@@
 protoplanetary disk model with thermal relaxation based on realistic
-disk properties, we find the fundamental mode to be dominant and
-robust against boundary conditions. The fundamental mode is thus
-likely the most relevant VSI mode in protoplanetary disks.
+disk properties, we find the fundamental mode to be the dominant mode,
+ and it is robust against boundary conditions. The fundamental mode is
+ thus likely the most relevant VSI mode in protoplanetary disks.
 
@@
 pseudo-spectral approach. For numerical calculations we relax all the
-approximations made in our analytical discussion, and accounted for
+approximations made in our analysis, and accounted for
 the background radial disk structure for consistency.  Our
@@
 expectations, including the dependence of $\beta_\mathrm{crit}$ on
-disk parameters.
+disk parameters.  %for regimes of interest
 
@@
 code. We confirm the finding by \cite{nelson13} that in the absence
-of vertical entropy gradients, the VSI operates and is insensitive to
+of vertical entropy gradients, the VSI is insensitive to
 thermal relaxation. When a stable stratification is introduced, the
@@
 We applied our linear models to assess the applicability of the VSI in
-protoplanetary disks, specifically the Minimum Mass Solar Nebulae
-(MMSN). We consider the disk model described in \cite{chiang10}, and
-modeled the thermal relaxation timescale $\beta$ based on
-dust opacity, which results in a $\beta$ that depends on both the
-perturbation lengthscale as well as the height from the midplane.  
-We find the MMSN is unstable to the fundamental VSI at $10$s of AU,
-with characteristic growth times of $\sim 30$ orbits. Although $\beta$
-is not spatially constant as assumed in our analysis, we find that
-evaluating $\beta < \beta_\mathrm{crit}$ at the disk
-midplane gives a rough measure of the radii at which the VSI operates in the
-MMSN.  This simple criteron may be a useful first step to 
-estimate the relevance of the VSI in realistic astrophysical
-disks without resorting to non-linear numerical simulations.
+protoplanetary disks, using a model for the the Minimum Mass Solar Nebulae
+(MMSN) described in \cite{chiang10}. We  
+modeled the thermal relaxation timescale $\beta$ based on realistic
+disk properties and dust opacity, which results in a $\beta$ that
+depends on both the perturbation lengthscale as well as the height
+from the midplane. We find the MMSN is unstable to the VSI
+from $\sim 5$AU to $\sim50$AU with characteristic growth times of
+$\sim 30$ orbits. Although $\beta$ is not vertically constant in the
+MMSN as assumed in our analysis, we find that evaluating $\beta <
+\beta_\mathrm{crit}$ at the disk midplane gives a rough measure of the
+radii at which the VSI operates in the MMSN.  This simple criteron may
+be a useful first step to estimate the relevance of the VSI in
+realistic astrophysical disks without resorting to non-linear
+numerical simulations.
 
@@
 application to protoplanetary disks we connected $t_c$ to the disk
-properties, but one can improve the realism by generalizing the
+properties, but one can improve the realism further by generalizing the
 opacity model adopted here to be density-dependent,  i.e. $\kappa_d =
-\kappa_d(\rho, T)$.  However, going beyond the framework of thermal
+\kappa_d(\rho, T)$ as described in \cite{bell94}. %two temperature
+% structure  
+However, going beyond the framework of thermal
 relaxation requires a proper treatment for radiative diffusion, which
@@
 laminar to develop the VSI.  This problem is, however, complicated by
-the fact that proper inclusion of viscosity introduces viscous heating
-in the energy equation, as well as a meridional flow in the equilibrium state.
+the fact that proper inclusion of viscosity introduces a meridional
+flow in the equilibrium state, as well as viscous heating
+in the energy equation.
 

### 2015-03-04 summary.tex
appendix.tex

@@
 
-The fundamental VSI mode is characterized by a vertical velocity
-perturbation with zero nodes and $\delta v_z\sim\mathrm{constant}$
-near the mid-plane, and by one node in the density perturbation at the
-midplane  (i.e. anti-symmetric about $z=0$).  These have been referred
+The fundamental VSI mode is characterized by a symmetric vertical
+velocity perturbation with zero nodes and $\delta
+v_z\sim\mathrm{constant}$ near the mid-plane, and by an anti-symmetric
+density perturbation with one node at $z=0$.  These have been referred
 to as `corrugation modes' by \cite{nelson13}, and were found to

### 2015-03-05 appendix.tex
revised

@@
 \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
-\ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_r + \hat{g}_c
+\ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_x + \hat{g}_c
   \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
@@
 \end{align}
-where subscript $0$ denotes evaluation at $r=r_0$.
+where subscript $0$ denotes evaluation at $r=r_0$. We recall terms
+proportional to $\hat{g}_c$ are associated with the background density
+and pressure radial structures, and nominally $\hat{g}_c=1$.
 
-In the VGSB, the terms proportional to $\hat{g}_c$ are ignored. For a
-power-law disk, these radial gradients are $O(r^{-1})$, and they
-appear in comparison with terms of $O(k_x)$. The neglected terms 
-therefore have a relative magnitude of $O(\epsilon/\khat)$. This is
-small for thin disks ($\epsilon\ll1$) and/or small radial wavelengths
-($\khat\gg 1$).
+However, in the VGSB these terms are neglected by 
+setting $\hat{g}_c=0$. For a power-law disk, these radial gradients
+are $O(r^{-1})$, and they appear in comparison with terms of
+$O(k_x)$. The neglected terms  therefore have a relative magnitude of
+$O(\epsilon/\khat)$. This is small for thin disks ($\epsilon\ll1$)
+and/or small radial wavelengths ($\khat\gg 1$).
 
@@
 An important caveat of the VGSB is that it cannot be used in the
-adiabatic limit. We explain this by setting  $\beta\to\infty$ in
+adiabatic limit. We explain this by considering vertically isothermal
+disks and setting  $\beta\to\infty$ in 
 Eq. \ref{vertiso_gov_nondim}, which can then be written as
@@
 Solberg-Hoiland criterion (Eq. \ref{solberg2}), which states that 
-instability in the adiabatic limit requires the disk to be close to
-neutrally stratified. %also need real freq^2 in adia limit 
+instability in a thin, adiabatic disk requires it to be close to 
+neutrally stratified, i.e. $|\gamma-1|\ll1$ for vertically isothermal disks
+evolved adiabatically. 
+%also need real freq^2 in adia limit 
 This is because the VGSB retains the vertical shear
-($\p_z\Omega$) from the global disk, but ignores the global radial disk
-structure responsible for it. Nevertheless, we demonstrate below that
-this inconsistency is unimportant for the vertical shear instability,
-which occurs for $\beta\ll1$.
+($\p_z\Omega$) from the global disk, but ignores the global radial
+disk structure responsible for it. Nevertheless, we demonstrate below
+that this inconsistency is unimportant for the VSI because it occurs
+for $\beta\ll1$.
 
 \subsection{Effect of global radial gradients}
-
-In  Fig. \ref{gcorr_compare}, we demonstrate the effect of the global radial gradient terms
+In  Fig. \ref{gcorr_compare}, we plot the effect of the global radial gradient terms
 proportional to $\hat{g}_c$ in 
@@
 $\beta\gtrsim1$, and as $\beta\to\infty$ the VGSB framework gives a
-(spurious) positive growth rate as expected from the discussion
+(spurious) growth rate as expected from the discussion
 above. Inclusion of the global radial gradient terms results in the
@@
     the `Numerical (VGSB)' curve is obtained by numerically solving
-    Eq. \ref{lin_mass}---\ref{lin_energy} with $\hat{g}_c=0$
+    the linear problem, Eq. \ref{lin_mass}---\ref{lin_energy}, with $\hat{g}_c=0$
     (which defines the linear VGSB equations); and the `Numerical
@@
 In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
-eliminating variables to obtain a single equation for $\delta v_z$,
-Eq. \ref{vertiso_gov}.  This procedure ignores the vertical dependence of
-$D=\kappa^2(z) - \sigma^2$. We show here that this has no significant
+eliminating variables to obtain a single equation for $\delta v_z$ in
+vertically isothermal disks, Eq. \ref{vertiso_gov}.  
+This procedure ignores the vertical dependence of
+$D=\kappa^2(z) - \sigma^2$. We show here that this has no serious 
 consequence for thin disks.
@@
 Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
-3+q$ as $|z|\to\infty$, i.e. its magnitude is of order unity.
+3+q$ as $|z|\to\infty$. Since we consider $q=O(1)$, $|F|$ is of
+order unity.
 
- Eliminating $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
+Now, eliminating $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
 Eq. \ref{lin_vz} for vertically isothermal disks, retaining the
@@
 neglecting these terms has no qualitative effect on the discussion in 
-\S\ref{analytic_relax}.  This issue does not
-arise for isothermal perturbations discussed in \S\ref{iso_discuss}
-since in that case we work with a governing equation for $W$ instead.
+\S\ref{analytic_relax}.  % This issue does not
+% arise for isothermal perturbations discussed in \S\ref{iso_discuss}
+% since in that case we work with a governing equation for $W$ instead.
 
@@
 
-\subsection{Thin-disk limit}
-For thin disks ($\epsilon\ll1$), in which 
+\subsection{Explicit solutions in the thin-disk limit}
+For thin disks ($\epsilon\ll1$), we write 
 $\ln\rho\simeq -\zhat^2/2$ and $f(\zhat)\simeq \zhat$, 
@@
 %With this approximation and those in \S\ref{integral_relation}, 
-Eq. \ref{iso_ode1} becomes  
+so Eq. \ref{iso_ode1} becomes  
 \begin{align}\label{iso_ode3}
@@
 %0.033576258
-In addition, we
-use the opacity model {\bf(reference?)}
+In addition, we use the opacity model
  \begin{align}
    \kappa_d &= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2
-   \mathrm{cm}^2\,\mathrm{g}
+   \mathrm{cm}^2\,\mathrm{g}^{-1}
     =
@@
  \end{align}
-where the second equality follows from the model temperature profile
-above, and $\hat{\kappa}_d$ has similar meaning as $\hat{\Sigma}$ and
-$\hat{T}$.
+\citep{bell94} where the second equality follows from the model
+temperature profile above, and $\hat{\kappa}_d$ has similar meaning as
+$\hat{\Sigma}$ and $\hat{T}$.

### 2015-03-05 appendix.tex
revised

@@
 
-\section{Linear problem in isothermal limit}\label{iso_discuss}  
+\section{Linear problem in the isothermal limit}\label{iso_discuss}  
 Here we examine in more detail the special case of isothermal

### 2015-03-07 numerical.tex
commented out kx<1 comment, there is convec overstability

@@
 
- Growth rates are expected to diminish for both $\khat\to 0$
+ With rapid thermal relaxation, growth rates are expected to diminish for both $\khat\to 0$
  and $\khat\to \infty$. This is because for $\khat\ll 1$ the radial
@@
 
-For $\khat\lesssim 10$, e.g. $\khat=1$ in 
-Fig. \ref{lowfreq_eigenfunc}, the perturbation magnitude near the
-boundaries begin to dominate. For even smaller $\khat\lesssim 1$ (not
-shown) we find the perturbation is almost entirely concentrated at the
-boundaries. However, our radially-local approximation is not valid for 
-such small wavenumber perturbations.
+%For $\khat\lesssim 10$, e.g. $\khat=1$ in 
+%Fig. \ref{lowfreq_eigenfunc}, the perturbation magnitude near the
+%boundaries begin to dominate. For even smaller $\khat\lesssim 1$ (not
+%shown) we find the perturbation is almost entirely concentrated at the
+%boundaries. However, our radially-local approximation is not valid for 
+%such small wavenumber perturbations.
 

### 2015-03-07 paper.tex
minor corrections

@@
   a characteristic growth time of $\sim 30$ local orbital
-  periods, and the dominant disturbance in the MMSN is the `corrugation' mode
+  periods. We find the dominant VSI mode in the MMSN is the `corrugation' mode
   which casues the disk column to oscillate vertically. 

### 2015-03-09 summary.tex
added convective overstability discussion

@@
 We highlight below some extensions to the present  
-linear models in future works:
+linear models in future works.
 
-\emph{Improved thermal evolution.} We assumed non-adiabatic effects
+\subsubsection{Parameter study and convective overstability}
+We have focused on parameter regimes relevant to the VSI,
+namely $\khat\gtrsim 1$ and $\beta\ll 1$. This excludes the
+`convective overstability' recently discussed by \cite{klahr14} and
+\cite{lyra14} in vertically unstratified disk models.
+
+The convective overstability corresponds to growing epicyclic motions.
+It is most effective for disturbances with $|k_x|\ll |k_z|$, where
+$k_z$ is a vertical wavenumber, and $\beta\sim 1$. These conditions
+are mutually exclusive with the VSI. 
+The maximum growth 
+rate for convective overstability is given by 
+\begin{align}
+  \nu_\mathrm{conv,max} = -\frac{N^2_r}{4\Omega}\label{max_conv_gen}
+\end{align}
+\citep{lyra14}, where $N_r^2 = -(c_p\rho)^{-1}\p_rP\p_rS$ is the
+square of the radial buoyancy frequency
+(cf. Eq. \ref{nzsq_def}). Thus, growth requires $N_r^2<0$.
+
+Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
+we find  
+\begin{align}
+  \nu_\mathrm{conv,max} =
+  -\frac{\epsilon^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
+  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
+\end{align}
+% We see that convective overstability is generally slower 
+% than the VSI because $\nu_\mathrm{conv,max} = O(\epsilon^2\Omega_k)$,
+% whereas the maximum VSI growth rate is $O(\epsilon\Omega_k)$. 
+For the MMSN considered in this work (\S\ref{application} and Appendix
+\ref{mmsn}), convective overstability does not apply because
+$N_r^2>0$.
+
+However, for more general disk structures convective overstability is
+possible. In fact, the models adopted for our numerical study in
+\S\ref{numerical} do satisfy $N_r^2<0$. %show early result
+We will explore convective overstability more fully in a follow-up
+study.  Although our linear code can be used for convective
+overstability in principle, the fact convective overstability is
+relevant for perturbations with $k_x\to0$ suggest radially global disk
+models may be more appropriate.
+
+% Nevertheless, it will be important to expand the parameter space covered in
+% this study to include convective overstability,
+
+%self-consistence may require radially global disk models
+
+\subsubsection{Improved thermal evolution.} 
+We assumed non-adiabatic effects
 can be represented by a thermal relaxation timescale $t_c$. For 
@@
 opacity model adopted here to be density-dependent,  i.e. $\kappa_d =
-\kappa_d(\rho, T)$ as described in \cite{bell94}. %two temperature
+\kappa_d(\rho, T)$. 
+%two temperature
 % structure
+
 However, going beyond the framework of thermal
@@
 
-\emph{Additional physics.} Depending on where the VSI operates in a
+\subsubsection{Additional physics.}
+
+Depending on where the VSI operates in a
 disk, additional effects may become important. For example, in the
@@
 
-\emph{Radial dependence.} Based on previous works, including
-non-linear simulations, which show that the VSI is relevant for
-perturbations with small radial lengthscales, we have adopted the
-radially-local approximation throughout this paper.  Although this
-simplifies the linear problem to ordinary differential equations, this
-approach introduces the wavenumber $k_x$ as a free parameter. The
-effect of radial boundary conditions cannot be explored. Future 
-generalizations of the linear problem should retain the radial
-dependence properly and solve the two-dimensional $(r,z)$ partial 
-differential equation eigenvalue problem.
+% \emph{Radial dependence.} Based on previous works, including
+% non-linear simulations, which show that the VSI is relevant for
+% perturbations with small radial lengthscales, we have adopted the
+% radially-local approximation throughout this paper.  Although this
+% simplifies the linear problem to ordinary differential equations, this
+% approach introduces the wavenumber $k_x$ as a free parameter. The
+% effect of radial boundary conditions cannot be explored. Future 
+% generalizations of the linear problem should retain the radial
+% dependence properly and solve the two-dimensional $(r,z)$ partial 
+% differential equation eigenvalue problem.
 

### 2015-03-10 appendix.tex
adding citations to isothermal discussion without vshear

@@
 \subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
-When $q=0$, Eq. \ref{iso_ode3} is
+The case of $q=0$ correspond to axisymmetric waves in a vertically isothermal disks, and has been discussed in detail 
+by several authors \citep[e.g.][]{}
+
+Eq. \ref{iso_ode3} is
 \begin{align}\label{hermite_ode}

### 2015-03-10 appendix.tex
updated iso discussion in appendix

--- 2015-03-05 appendix.tex
@@
 
-\section{Linear problem in isothermal limit}\label{iso_discuss}  
+\section{Linear problem in the isothermal limit}\label{iso_discuss}  
 Here we examine in more detail the special case of isothermal
-perturbations ($\beta\equiv 0$) in the VGSB framework.   
+perturbations ($\beta\equiv 0$) for vertically isothermal disks 
+($\Gamma=1$) in the VGSB framework.   
 %Isothermal perturbations correspond to $\beta\equiv0$ ($\chi=1$) and 
@@
 simpler to work with an equation for $W$ by substituting $Q=W$ into
-Eq. \ref{ode_w}. We obtain 
+Eq. \ref{ode_w}. In this case, we shall not yet make the low-frequency
+approximation, but first make the nearly-Keplerian approximation. We
+obtain, in terms of dimensionless variables, 
 \begin{align}
-  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
-    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
+  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \frac{\ii \epsilon q \hat{k} f(\zhat)
+      }{1-\hat{\sigma}^2} \right]W^\prime +
+  \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W,\label{iso_ode0} 
 \end{align}
@@
 
-\subsection{Necessity of vertical shear for
+\subsection{Stability conditions in the low-frequency limit}
+In the low-frequency limit we assume $|\hat{\sigma}|\ll 1$ and
+approximate Eq. \ref{iso_ode0} as 
+\begin{align}
+  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
+    f(\zhat)\right]W^\prime +
+  \hat{\sigma}^2\left(1+\hat{k}^2\right)W. \label{iso_ode1}
+\end{align}
+\subsubsection{Necessity of vertical shear for
   instability}\label{integral_relation} 
@@
 
-\subsection{Explicit solutions in the thin-disk limit}
-For thin disks ($\epsilon\ll1$), we write 
-$\ln\rho\simeq -\zhat^2/2$ and $f(\zhat)\simeq \zhat$, 
-% by expanding the background density and vertical shear profile in powers
-% of $z/r$. To lowest order we obtain 
-% \begin{align}
-%   &\frac{\rho}{\rho_0} \simeq 
-%   \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv w(\hat{z}),\label{thin_dens}\\
-%   &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}, \label{thin_vshear}
-% \end{align} 
-%With this approximation and those in \S\ref{integral_relation}, 
-so Eq. \ref{iso_ode1} becomes  
+\subsection{Explicit solutions in the thin-disk limit}\label{iso_explicit}
+Here we assume a thin disk ($\epsilon\ll1$) so that $\ln\rho\simeq
+-\zhat^2/2$ and $f(\zhat)\simeq \zhat$. However, we do not assume
+low-frequency from the onset. Then Eq. \ref{iso_ode0} becomes 
 \begin{align}\label{iso_ode3}
-  W^{\prime\prime} - \left(1 + \ii q\epsilon
-    \hat{k}\right)\hat{z}W^\prime  +
-  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
-  0.
+  W^{\prime\prime} - \left(1 + \frac{\ii q\epsilon \hat{k}}{1-\hat{\sigma}^2}\right)\hat{z}W^\prime  +
+  \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W = 0.
 \end{align}
-We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
-that derived by \cite{nelson13}, although we have taken a different
-route.  
+We remark that for $\hat{k}^2\gg 1$ and taking low-frequency limit,
+Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in \cite{nelson13},
+although we have taken a different route.
+
 % To complete the problem we must specify boundary conditions. The 
@@
 %It is, however, common to take $\mathrm{max}|z|\to\infty$ and
-We impose the kinetic energy remain bounded at infinity. Even though
-we have adopted the thin-disk approximation, an infinite
-disk is appropriate in the sense that both the true density profile 
-and its thin-disk approximation decay rapidly away from the midplane.
 
-However, the thin-disk representation of vertical shear diverges with 
-$\hat{z}$, unlike the true vertical shear profile which eventually
-decays. Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit 
-unphysically large growth rates (as demonstrated below). Nevertheless,
-the infinite disk is useful to consider since it permits an analytical
-discussion.
-
-\subsubsection{Stability in the absence of vertical shear}\label{iso_stable}
-When $q=0$, Eq. \ref{iso_ode3} is
-\begin{align}\label{hermite_ode}
-  \left[w(\hat{z})W^\prime \right]^\prime + nW
-  w(\hat{z}) =0, 
-\end{align}
-where $w(\zhat)=\exp{(-\zhat^2/2)}$ and we have defined a new eigenvalue 
-\begin{align}
-  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
-\end{align} 
-Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
-that the kinetic energy density remain bounded at infinity, then the
-solutions are Hermite polynomials  
-\begin{align}
-  W \propto \He_n(\hat{z}),
-\end{align}
-and $n$ is a non-negative integer. This translates to a real
-eigenfrequency $\sigma$ such that
-\begin{align}
-  \left|\hat{\sigma}\right| = \sqrt{n}
-  \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
-\end{align}
-Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
-analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
-n-1$. Physically, this corresponds to radial length-scales much
-smaller than the local disk scale height. Note that
-Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
-\cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
-nearly-Keplerian approximation in that equation.
-
-We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
-Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
-has also been qualitatively observed in numerical simulations
-\citep{stoll14}.
-
-\subsubsection{Polynomial solutions with vertical
-  shear}\label{iso_poly}
-In the presence of vertical shear ($q\neq0$) we can still seek power-series solution to
-Eq. \ref{iso_ode3}, 
+We seek power series solutions to Eq. \ref{iso_ode3}, 
 \begin{align}
@@
   (l+2)(l+1)a_{l+2} +
-  \left[n - l\left(1+\ii \epsilon q  \hat{k}\right)\right] a_l = 0, 
+  \left[\hat{\sigma}^2\left(1+\frac{\khat^2}{1-\hat{\sigma}^2}\right)
+    - l\left(1+\frac{\ii \epsilon q
+        \hat{k}}{1-\hat{\sigma}^2}\right)\right] a_l = 0. 
 \end{align}
-where $n$ is not necessarily an integer. Indeed, if we demand
-a polynomial of degree $L$ as the solution, then the eigenfrequency
-must satisfy
-\begin{align}\label{sig2_iso}
-  \hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
-      \hat{k}}{1+\hat{k}^2}\right).
+We demand the series to terminate  at $l=L$, i.e. a polynomial
+solution, so that the vertical kinetic energy remain bounded as
+$|\zhat|\to\infty$.  Then the eigenfrequency $\hat{\sigma}$ is given via 
+\begin{align}
+\hat{\sigma}^4 - \left(L+1+\khat^2\right)\hat{\sigma}^2 + L\left(1 +
+  \ii\epsilon q \khat\right) = 0. \label{simple_growth}
+\end{align}
+
+\subsubsection{Stability without vertical shear}
+In the absence of vertical shear $q=0$, Eq. \ref{simple_growth} can be
+written as \begin{align}
+  \hat{\sigma}^2\khat^2 =
+  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right). 
+\end{align}
+This is just the dispersion relation for axisymmetric waves in a
+vertically isothermal disk already discussed in previous works 
+\citep[e.g.][]{okazaki87,takeuchi98,tanaka02,zhang06,ogilvie13,barker14}. In
+this case the solutions are Hermite polynomials, $W\propto
+\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma}$ is real so the disk
+is stable, consistent with \S\ref{integral_relation}.
+
+\subsubsection{Instability with vertical shear}
+A simple solution to Eq. \ref{simple_growth} may be obtained by
+considering small radial lengthscales $\khat\gg 1$. Then for the
+low-frequency branch of Eq. \ref{simple_growth} 
+\begin{align}
+  \hat{\sigma}^2 \simeq L\left(\frac{1+\ii q \epsilon
+       \hat{k}}{L+1+\hat{k}^2}\right).\label{simple_growth2}
 \end{align}
 The eigenfrequency $\hat{\sigma}$ is therefore complex for 
-$L\geq1$. The dispersion relation Eq. \ref{sig2_iso} can also be
-derived by setting $\beta=0$ in the more general dispersion relation 
-Eq. \ref{relax_disp} with coefficients given in Appendix
-\ref{relax_coeff}. We are interested in 
-unstable modes for which $\hat{\nu}>0$ and is given by
-\begin{align}\label{simple_growth}
-  \hat{\nu} =\sqrt{
-    \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
-      1\right)}. 
-\end{align}
-For fixed $q$ and $\epsilon$, the growth rate vanishes for both
-$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
-The maximum growth rate
-occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
-\begin{align}
-  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
-\end{align}
-and
-\begin{align}
-  \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
-    q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
-\end{align}
-Since $L\geq1$, there is no limit to the maximum growth rate. This is
-an artifact of the thin-disk approximation because the magnitude of vertical 
-shear $|d\Omega^2/dz|\propto |z|$ increases indefinitely with 
-height. However, Eq. \ref{iso_max_growth} compares well with
-numerical solutions for small $L$ (see \S\ref{vertiso_pertiso}).
+$L\geq1$, which signifies instability since we can choose
+the sign of the square root to make $\hat{\nu} = \imag\hat{\sigma}>0$.
+
+% \begin{align}\label{sig2_iso}
+%   \hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
+%       \hat{k}}{1+\hat{k}^2}\right).
+% \end{align}
+% The eigenfrequency $\hat{\sigma}$ is therefore complex for
+% $L\geq1$. The dispersion relation Eq. \ref{sig2_iso} can also be
+% derived by setting $\beta=0$ in the more general dispersion relation 
+% Eq. \ref{relax_disp} with coefficients given in Appendix
+% \ref{relax_coeff}. We are interested in 
+% unstable modes for which $\hat{\nu}>0$ and is given by
+% \begin{align}
+%   \hat{\nu} =\sqrt{
+%     \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
+%       1\right)}. 
+% \end{align}
+
+% We impose the kinetic energy remain bounded at infinity. Even though
+% we have adopted the thin-disk approximation, an infinite
+% disk is appropriate in the sense that both the true density profile 
+% and its thin-disk approximation decay rapidly away from the midplane.
+
+% However, the thin-disk representation of vertical shear diverges with 
+% $\hat{z}$, unlike the true vertical shear profile which eventually
+% decays. Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit 
+% unphysically large growth rates (as demonstrated below). Nevertheless,
+% the infinite disk is useful to consider since it permits an analytical
+% discussion.
+
+% \subsubsection{Stability in the absence of vertical
+%   shear}\label{iso_stable}
+% \
+% Eq. \ref{iso_ode3} is
+% \begin{align}\label{hermite_ode}
+%   \left[w(\hat{z})W^\prime \right]^\prime + nW
+%   w(\hat{z}) =0, 
+% \end{align}
+% where $w(\zhat)=\exp{(-\zhat^2/2)}$ and we have defined a new eigenvalue 
+% \begin{align}
+%   n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
+% \end{align} 
+% Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
+% that the kinetic energy density remain bounded at infinity, then the
+% solutions are Hermite polynomials  
+% \begin{align}
+%   W \propto \He_n(\hat{z}),
+% \end{align}
+% and $n$ is a non-negative integer. This translates to a real
+% eigenfrequency $\sigma$ such that
+% \begin{align}
+%   \left|\hat{\sigma}\right| = \sqrt{n}
+%   \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
+% \end{align}
+% Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
+% analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
+% n-1$. Physically, this corresponds to radial length-scales much
+% smaller than the local disk scale height. Note that
+% Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
+% \cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
+% nearly-Keplerian approximation in that equation.
+
+% We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
+% Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
+% has also been qualitatively observed in numerical simulations
+% \citep{stoll14}.
+
+% \subsubsection{Polynomial solutions with vertical
+%   shear}\label{iso_poly}
+
+% For fixed $q$ and $\epsilon$, the growth rate vanishes for both
+% $\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
+% The maximum growth rate
+% occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
+% \begin{align}
+%   |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
+% \end{align}
+% and
+% \begin{align}
+%   \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
+%     q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
+% \end{align}
+% Since $L\geq1$, there is no limit to the maximum growth rate. This is
+% an artifact of the thin-disk approximation because the magnitude of vertical 
+% shear $|d\Omega^2/dz|\propto |z|$ increases indefinitely with 
+% height. However, Eq. \ref{iso_max_growth} compares well with
+% numerical solutions for small $L$ (see \S\ref{vertiso_pertiso}).
 

### 2015-03-10 linear_adia.tex
appendix.tex
corrected two typos (eq 67 and eq 46)
updated iso discussion in appendix

@@
 \begin{align}
-  0 = \delta v_z^{\prime\prime} - z A\delta v_z^\prime +
+  \delta v_z^{\prime\prime} - z A\delta v_z^\prime +
   (B - C\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
@@
 
-In Appendix \ref{iso_discuss}, we give an analytical description of the
+In Appendix \ref{iso_discuss}, we give an analytical discussion of the
 linear problem with $\beta=0$. There, we establish more general
@@
 the maximum vertical shear rate in the disk). We also develop analytic
-solutions with and without vertical shear in the thin disk limit,
-including explicit expressions for wave frequencies and growth rates. 
-The methods used in Appendix \ref{iso_discuss} and results obtained
-therein motivates the solution procedure employed in the following
-sections.
+solutions in the thin disk limit. The methods used in Appendix
+\ref{iso_discuss} and results obtained therein motivates the solution
+procedure employed in the following sections.
 
@@
   0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
-  - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3.\label{relax_cond2}
+  - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
+    &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
+\label{relax_cond2}
 \end{align}

### 2015-03-10 numerical.tex
updated iso discussion in appendix

@@
 for the fundamental mode and that calculated from
-Eq. \ref{sig2_iso} (with $L=1$). There is good agreement for 
-$\khat\gtrsim10$. The match worsens for smaller $\khat$ because the 
-the low-frequency approximation used to obtain Eq. \ref{sig2_iso} 
-is no longer valid, as $|\omega|$ is not small compared to $\Omega_k$
-for $\khat\lesssim 10$.
+Eq. \ref{simple_growth} (with $L=1$). There is good agreement for  
+$\khat\gtrsim10$. The match worsens for decreasing $\khat$ due to the
+neglect of global radial structure in the analysis.
+
+% the low-frequency approximation used to obtain Eq. \ref{sig2_iso} 
+% is no longer valid, as $|\omega|$ is not small compared to $\Omega_k$
+% for $\khat\lesssim 10$.
 
@@
 expected for the fundamental mode according to explicit solutions
-developed in Appendix \ref{iso_poly}. For completeness we show the
+developed in Appendix \ref{iso_explicit}. For completeness we show the
 corresponding eigenfunctions in vertical velocity in
@@
 smallest $|\sigma|$, for which the eigenvalue calculated from
-Eq. \ref{sig2_iso} with $L=1$ and the numerically-obtained value are   
+Eq. \ref{simple_growth2} with $L=1$ and the numerically-obtained value are   
 \begin{align*}
-  (\omega,\nu) = (-0.3053,0.2606)\epsilon\Omega_k &\quad \text{(from
-    Eq. \ref{sig2_iso})},\\
+  (\omega,\nu) = (-0.3053,0.2605)\epsilon\Omega_k &\quad \text{(from
+    Eq. \ref{simple_growth2})},\\
   (\omega,\nu) = (-0.3010,0.2575) \epsilon \Omega_k &\quad \text{(numerical)}.
@@
 This agreement is surprisingly good given the number of approximations
-used to obtain Eq. \ref{sig2_iso}, which also imposes a different
+used to obtain Eq. \ref{simple_growth2}, which also imposes a different
 boundary condition to that in the numerical calculation. These values

### 2015-03-12 appendix.tex
updating results

@@
 
-\subsubsection{Maximum growth rate}   
+\subsubsection{Maximum growth rate}\label{max_growth1}
 Here we show that the growth rate is limited by the maximum vertical
@@
 \end{align}
-We remark that for $\hat{k}^2\gg 1$ and taking low-frequency limit,
-Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in \cite{nelson13},
-although we have taken a different route.
+We remark that taking the low-frequency limit and considering
+$\hat{k}^2\gg 1$, Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in
+\cite{nelson13}, although we have taken a different route.
 
@@
   \hat{\sigma}^2\khat^2 =
-  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right). 
+  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right). \label{iso_disp_full}
 \end{align}
@@
 \He_L(\zhat)$.  The eigenfrequency $\hat{\sigma}$ is real so the disk
-is stable, consistent with \S\ref{integral_relation}.
+is stable, consistent with \S\ref{integral_relation}. The
+low-frequency solution to Eq. \ref{iso_disp_full} correspond to
+inertial waves \citep{balbus03}.
 

### 2015-03-12 numerical.tex
updating results

--- 2015-03-07 numerical.tex
@@
 
-%, so that we may consistently consider vertical domains up to
-%$z\sim r$.
-
 We solve the linear problem by expanding the
@@
 and discretizing the equations on a grid with
-$z\in[-\zmax,\zmax]$. At the vertical boundaries we impose a free
-surface, 
+$z\in[-\zmax,\zmax]$. Our standard boundary conditions at the vertical
+boundaries is a free surface, 
 \begin{align}
@@
 components of which are given by $\xi_{x,z} = \ii\delta
-v_{x,z}/\sigma$. Calculations with a rigid boundary ($\delta v_z=0$)
-yield similar results. 
-% This boundary
-% condition is convenient for numerical implementation since it conforms
-% to a standard eigenvalue problem.
+v_{x,z}/\sigma$.
 
@@
 
-% mention mode selection -> focus on fundamental corrugation ->
-% simulations eventually dominated by them 
-% choice of kx is O(10), since nelson observe kxH \sim 30
-%not possible to survey all modes/kx/disk parameters
-%focus on effect of thermal relaxation
-
-\subsection{VSI with rapid thermal relaxation}\label{vertiso_pertiso} 
-% disk model is a nearly-vertically
-% isothermal disk with $\Gamma=1.011$, $p=-1.5$, $q=-1$,
-% $\epsilon=0.1$. 
-We first calculate the VSI in a neutrally
-stratified disk ($N_z^2\equiv0$) with rapid thermal relaxation by
-setting $\gamma=\Gamma=1.011$ and $\beta=10^{-3}$. Other disk
-parameters are $(p,q,\epsilon)=(-1.5,-1,0.1)$ and the vertical domain
-is $\zmax = 5H$. The disk is nearly vertically isothermal with
-$c_s(\zmax)\simeq 0.94 c_s(0)$. Fig. \ref{omega_z} shows that the 
-equilibrium rotation decreases slightly away  from the disk mid-plane.
-
-This setup is similar to that adopted
-in \cite{mcnally14}, who calculated the VSI in vertically isothermal
-disks subject to isothermal perturbations ($\Gamma=\gamma=1$).  This
-allows us to test our numerical code. We also compare results with our 
-analytical discussion of isothermal perturbations in Appendix
-\ref{iso_discuss}.
+Unless otherwise stated, our fiducial disk model is: a nearly 
+vertically isothermal disk with $\Gamma=1.011$, stable
+stratification $\gamma=1.4$, $(p,q,\epsilon)=(-1.5,-1,0.05)$, and 
+vertical domain size $\zmax=5H$. In Fig. \ref{omega_z} we plot the
+corresponding equilibrium rotation profile, which shows that $\Omega$
+decreases slightly away from the mid-plane.
 
@@
     normalized by its mid-plane value, for a disk with $\Gamma=1.011$
-    and $(p,q,\epsilon)=(-1.5,-1,0.1)$. 
+    and $(p,q,\epsilon)=(-1.5,-1,0.05)$. 
     \label{omega_z} 
@@
 
-%behaviour as func of kx
+\subsection{VSI with rapid thermal relaxation}\label{vertiso_pertiso} 
+We first calculate the VSI in a disk with rapid thermal relaxation by
+setting $\beta=10^{-3}$. (Smaller values yield the same results.) This
+is similar to strictly isothermal perturbations in a vertically
+isothermal disk considered in \cite{nelson13} and \cite{mcnally14}. % We
+% compare results with our analytical discussion of isothermal 
+% perturbations in Appendix \ref{iso_discuss}.
 
-Fig. \ref{iso_eigen_kx} compares the numerically-obtained eigenvalues
-for the fundamental mode and that calculated from 
-Eq. \ref{sig2_iso} (with $L=1$). There is good agreement for 
-$\khat\gtrsim10$. The match worsens for smaller $\khat$ because the 
-the low-frequency approximation used to obtain Eq. \ref{sig2_iso} 
-is no longer valid, as $|\omega|$ is not small compared to $\Omega_k$
-for $\khat\lesssim 10$.
+Fig. \ref{iso_eigen_kx} compares the numerically obtained eigenvalues 
+for the fundamental mode and that calculated analytically from our
+analysis of isothermal perturbations in Appendix \ref{iso_discuss}
+(Eq. \ref{simple_growth} with $L=1$). We find good agreement in the
+wave frequency $\omega$ at all $\khat$ and in growth rates $\nu$ for
+$\khat\gtrsim 10$. Growth rates are $O(\epsilon\Omega_k)$, 
+consistent with our discussion in Appendix \ref{max_growth1}, but
+are over-estimated by our analysis. The match in $\nu$ worsens for
+decreasing $\khat$ due to the neglect of global radial structure in
+the  analysis. Nevertheless, given the number of approximations made
+to calculate growth rates analytically, this agreement is
+satisfactory, especially at large $\khat$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/compare_eigen_imag_iso} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/compare_eigen_real_iso}
+  \caption{Growth rate (top) and real frequency (bottom) of the
+    fundamental VSI mode in a disk with $(\gamma,
+    \Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+    $\beta=10^{-3}$. The solid (dotted) line is obtained numerically
+    (analytically).  
+    \label{iso_eigen_kx} 
+  }
+\end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_eigen_imag_iso} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  1cm]{figures/compare_eigen_real_iso}
-  \caption{Growth rate (top) and real frequency (bottom) of the
-    fundamental VSI mode in a disk with $\gamma=\Gamma=1.011$,
-    $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
-    $\beta=10^{-3}$. The solid (dotted) line is obtained numerically
-    (analytically).  
-    \label{iso_eigen_kx} 
-  }
-\end{figure}
-
-Fig. \ref{lowfreq_eigenfunc} compares the fundamental VSI eigenfunction
-$W$ for $\khat=1$ and $\khat=20\pi$. The latter value, corresponding
-to $k_x= 200\pi/r$, was used by \cite{mcnally14}. We find $W$ is
-similar for all $\khat\gtrsim10$ and is proportional to $z$, as
-expected for the fundamental mode according to explicit solutions
-developed in Appendix \ref{iso_poly}. For completeness we show the
-corresponding eigenfunctions in vertical velocity in
-Fig. \ref{lowfreq_eigenfuncvz}. The fundamental mode
-consists of zero nodes in $\delta v_z$.
-
-%For $\khat\lesssim 10$, e.g. $\khat=1$ in 
-%Fig. \ref{lowfreq_eigenfunc}, the perturbation magnitude near the
-%boundaries begin to dominate. For even smaller $\khat\lesssim 1$ (not
-%shown) we find the perturbation is almost entirely concentrated at the
-%boundaries. However, our radially-local approximation is not valid for 
-%such small wavenumber perturbations.
+Fig. \ref{lowfreq_eigenfunc} plots the eigenfunctions $W$ and $\delta 
+v_z$ for the fundamental mode with $\khat=10$ (i.e. the mode with
+maximum growth rate in Fig. \ref{iso_eigen_kx}). We find $W\propto z$
+throughout most of the disk, corresponding to a constant vertical
+velocity perturbation, as expected for the fundamental mode according
+to explicit solutions developed in Appendix \ref{iso_explicit}. We
+show a meridional visualization of this mode in
+Fig. \ref{lowfreq_eigenfunc_2d}. Although there is a constant vertical
+velocity throughout the fluid column, most of the momentum is
+contained within $\sim 2H$ of the midplane due to the density
+stratification.
 
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvector_iso_kx1} 
-  \includegraphics[width=\linewidth]{figures/eigenvector_iso_kx60}
-  \caption{Eigenfunction $W$ of the fundamental VSI for perturbation
-    wavenumber $\khat=1$ (top) and $\khat=20\pi$ (bottom). The real 
-    (imaginary) parts of $W$ are plotted as solid (dotted) lines. % The
-    % eigenfunction has been normalized such that $\imag(W)=0$ at the
-    % vertical boundaries and $\mathrm{max}|W|=1$. 
+  0cm]{figures/eigenvectorW_iso} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/eigenvectorvz_iso}
+  \caption{Pressure perturbation $W$ (top) and vertical velocity
+    perturbation $\delta v_z$ of the fundamental VSI for a
+    perturbation  wavenumber $\khat=10$  The real 
+    (imaginary) parts of $W$ and $\delta v_z$ are plotted as solid
+    (dotted) lines. 
     \label{lowfreq_eigenfunc}
@@
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvectorvz_iso_kx1} 
-  \includegraphics[width=\linewidth]{figures/eigenvectorvz_iso_kx60}
-  \caption{Same as Fig. \ref{lowfreq_eigenfunc} but for the vertical
-    velocity eigenfunctions.
-    \label{lowfreq_eigenfuncvz}
+  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 3.1cm
+  0cm]{figures/result2d_vel}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
+  0cm]{figures/result2d_mom} 
+  \caption{Visualization of the fundamental VSI mode in
+    Fig. \ref{lowfreq_eigenfunc}. The color scale corresponds to the
+    real pressure perturbation $W$ scaled by its maximum value.
+    Arrows in the left panel correspond to the meridional velocity
+    field $(\delta v_x,\delta v_z)$, and arrows in the right panel
+    correspond to to the meridional momentum $\rho(\delta v_x,\delta
+    v_z)$. 
+    \label{lowfreq_eigenfunc_2d}
   }
@@
 
-In Fig. \ref{lowfreq_eigen} we plot the eigenvalues 
-$\sigma = \omega + \ii\nu$ found for $\khat=20\pi$. This plot is
-similar to Fig. 3 in \cite{mcnally14}. The fundamental mode has the
-smallest $|\sigma|$, for which the eigenvalue calculated from
-Eq. \ref{sig2_iso} with $L=1$ and the numerically-obtained value are   
-\begin{align*}
-  (\omega,\nu) = (-0.3053,0.2606)\epsilon\Omega_k &\quad \text{(from
-    Eq. \ref{sig2_iso})},\\
-  (\omega,\nu) = (-0.3010,0.2575) \epsilon \Omega_k &\quad \text{(numerical)}.
-\end{align*}
-This agreement is surprisingly good given the number of approximations
-used to obtain Eq. \ref{sig2_iso}, which also imposes a different
-boundary condition to that in the numerical calculation. These values
-are also close to that found by \cite{mcnally14}. The eigenfrequency
-is not sensitive to the vertical boundary condition because the
-expression for $\sigma$, given by Eq. \ref{integral_relation1},
-involve integrals with $\rho$ as the   weight function, which rapidly
-decays away from the midplane.
+\subsubsection{Case study: $\khat=10$}
+In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of unstable
+modes found for our standard setup with $\khat=10$. This case is shown
+by red crosses. We also plot several other calculations with different
+vertical boundary conditions: free surface at $\zmax = 3H$ (black
+dots), free surface at $\zmax=7H$ (green triangles) and a rigid
+boundary at $\zmax=5H$ (blue cross).
+
+The fundamental mode, marked by `f', corresponds to the eigenvalue
+with smallest magnitude $|\sigma|$. We see that the fundamental mode,
+while not the most unstable in this case, is more robust against
+vertical boundary conditions. Higher order modes with larger growth
+rates (i.e. the branch where $\nu$ increases with $|\omega|$) require
+larger vertical domains.
 
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
-  \caption{Eigenvalues of perturbations with wavenumber $\khat=20\pi$
-    in a neutrally-stratified, nearly-vertically isothermal disk
-    ($\gamma=\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
-    $\beta=10^{-3}$. \label{lowfreq_eigen} 
+  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10.ps}
+  \caption{Unstable modes for a disk with 
+    $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+    $\beta=10^{-3}$. The perturbation wavenumber is $\khat=10$. 
+    The standard setup with a free surface ($\Delta P=0$) and 
+    vertical domain size $\zmax=5H$ is shown by the red cross. Other
+    cases correspond to different boundary condtions applied at
+    different heights. The fundamental mode is marked by `f'.   
+    \label{compare_modes_iso_kx10} 
+  }
+\end{figure}
+
+\subsubsection{Surface modes: $\khat=30$} 
+Our standard setup with $\khat=10$ in 
+Fig. \ref{compare_modes_iso_kx10} (red crosses) differs to that from
+\cite{nelson13} and \cite{mcnally14} in the lack of `surface modes' in
+our case. These modes lie parallel to the vertical axis in mode
+diagrams such as Fig. \ref{compare_modes_iso_kx10}. We find them for
+sufficiently large $\zmax$ and/or $\khat$. % For example, the two
+% largest growth rates found for the case with $\zmax=7H$ in
+% Fig. \ref{compare_modes_iso_kx10} are surface modes.
+
+To illustrate this, we plot the mode diagram for $\khat=30$ in  
+Fig. \ref{compare_modes_iso_kx30} with different boundary 
+conditions. Surface modes can be seen with
+$\zmax\geq 5H$, examples of which are marked by `s'. These modes have
+been discussed in the above studies. They typically have the largest
+growth rates but require sufficiently large vertical domain size and their eigenvalues
+vary with $\zmax$.
+
+Due to their dependence on boundary conditions, we
+do not consider surface modes further in detail. In fact, we will see
+in the next section that surface modes are more effectively stabilized
+by finite thermal relaxation than the fundamental mode. %  Thus, surface
+% modes may not be the most relevant in realistic disks.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_tags.ps}
+  \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for $\khat=30$
+    \label{compare_modes_iso_kx30}. Examples of surface modes are
+    marked with `s'. These are found for sufficiently large vertical
+    domain sizes. 
   }
@@
 \subsection{Effect of thermal relaxation}
-We now consider stably stratified disks ($\gamma > \Gamma$) with
-thermal relaxation. Here, we use a slightly  
-different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as 
-adopted in simulations performed by \cite{nelson13}. Other parameters 
-are unchanged from the previous section.
+We now consider the effect of thermal relaxation by increasing
+$\beta$. Note that for the standard disk model the critical thermal 
+timescale $\beta_\mathrm{crit} = 0.125$.
+
+Fig. \ref{compare_modes_cool_kx10} show unstable modes with $\khat=10$
+and $\beta\in[0.01,0.1]$. In addition to our standard setup with 
+$\zmax=5H$, we also plot results for $\zmax=3H$ and $\zmax=7H$. In all
+three cases the number of unstable modes decrease with increasing
+$\beta$. The fundamental mode is not the most unstable for 
+$\beta=0.01$ as there is a large number of higher order modes, but
+these depend on boundary conditions.
+
+However, the fundamental mode is dominant for $\beta \geq 
+0.03$. Higher order modes, with larger growth rates than the
+fundamental mode, are more effectively stabilized with increasing 
+$\beta$. For $\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we
+only find the fundamental mode.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/compare_modes_cool_kx10_z3.ps} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/compare_modes_cool_kx10_z5.ps}
+  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx10_z7.ps}
+  \caption{Unstable modes with $\khat=10$ and thermal
+    relaxation timescales $\beta=0.01$ (black dots), $\beta=0.03$ (red
+    crosses), $\beta=0.05$ (green triangles) and $\beta=0.1$ (blue
+    crosses); for different vertical domain sizes $\zmax=3H$ (top),
+    $\zmax=5H$ (middle, standard setup) and $\zmax=7H$ (bottom). The
+    fundamental mode is marked with `f'. 
+    \label{compare_modes_cool_kx10} 
+  }
+\end{figure}
+
+Fig. \ref{compare_modes_cool_kx30} show unstable modes with
+$\khat=30$. In this case surface modes appear for sufficiently large   
+$\zmax$. However, they are effectively stabilized with increasing 
+$\beta$. For $\zmax=3H$, the fundamental mode becomes dominant for
+$\beta\geq 0.03$, as before. For $\zmax=5H$ and $\zmax=7H$, we find
+the appearence of modes slightly more unstable than the fundamental
+mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
+panels). However, these modes are clearly associated with the vertical
+domain size as they are absent for $\zmax=3H$. % Thus, for $\zmax=5H$ and 
+% $\zmax=7H$  the fundamental mode is likely dominant for 
+% $\beta\geq0.05$
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/compare_modes_cool_kx30_z3.ps} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/compare_modes_cool_kx30_z5.ps}
+  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx30_z7.ps}
+  \caption{Same as Fig. \ref{compare_modes_cool_kx10} but for
+    $\khat=30$. Examples of surface modes are marked with `s', and 
+    appear for sufficiently large $\zmax$.
+    \label{compare_modes_cool_kx30} 
+  }
+\end{figure}
+
+Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
+show that the fundamental mode is not sensitive to boundary
+conditions. This suggests that tracking fundamental mode would be a
+more robust way of assessing stability as a function of $\beta$,
+rather than searching for the most 
+unstable mode, because the latter may be associated with boundary
+conditions. In fact, the fundamental mode bcomes dominant 
+(or close to it) with increasing $\beta$.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvectorW_beta0d1} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/eigenvectorvz_beta0d1}
+  \caption{Same as Fig. \ref{lowfreq_eigenfunc_cool} but with a
+    thermal relaxation timescale $\beta=0.1$. 
+    \label{lowfreq_eigenfunc_cool}
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 3.1cm
+  0cm]{figures/result2d_vel_cool}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
+  0cm]{figures/result2d_mom_cool} 
+  \caption{Same as  Fig. \ref{lowfreq_eigenfunc_2d} but with a thermal
+    relaxation timescale $\beta=0.1$. 
+    \label{lowfreq_eigenfunc_2d_cool}
+  }
+\end{figure}
+
+\subsection{Critical thermal relaxation
+  timescale}\label{bcrit_num_test}
 
@@
 
-% Fig. \ref{relax_growth_num} shows the growth rates of the fundamental  
-% mode as a function of $\khat$. This plot is largely consistent with  
-% Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
-% where the low-frequency approximation fails. (In any case, the local model not valid
-% for $\khat\ll1$.) For $\khat\gtrsim 10$ introducing thermal
-% relaxation rapidly stabilizes the fundamental mode.
-
-% Fig. \ref{relax_growth_num} also confirms our analytical discussion that
-% introducing a small but finite cooling time stabilizes the fundamental
-% VSI, but there is a preferable wavenumber that maximizes this effect (\S\ref{relax_pert}).
-% This occurs at $\khat= O(10)$ for the current disk model.
-
-% \begin{figure}
-%    \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
-%    0cm]{figures/compare_eigen_imag_bloop} 
-%   % \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-%   % 1cm]{figures/compare_eigen_imag_bloop2} 
-%   \caption{Growth rates of the fundamental VSI mode in 
-%     nearly-vertically isothermal disks ($\Gamma=1.011$) with
-%     $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ under a
-%     range of thermal relaxation timescales   
-%     $\beta$. The eigenfrequencies are calculated by numerically 
-%     solving the full linear eigenvalue problem. \label{relax_growth_num}}   
-% \end{figure}
-
-%approximately the perturbation wavenumber observed in the fiducial
-%numerical simulation of \cite{nelson13}.
-
-We visualize the effect of thermal relaxation in 
-Fig. \ref{relax_eigenW_num} by comparing the vertical velocity
-perturbation $\delta v_z$ for
-$\beta=0.01$ and $\beta=0.1$. The latter value is close to
-$\beta_\mathrm{crit}=0.125$. 
-% the
-% critical value beyond which the fundamental VSI is stabilized
-% (\S\ref{iso_vsi_beta_crit}). 
-Increasing $\beta$ 
-restricts  the region in which $\delta v_z\sim $ constant closer to the midplane, with
-oscillatory behavior emerging at the boundaries, where buoyancy first
-becomes important. Finite thermal relaxation increases $|\delta v_z|$ near the
-vertical boundaries relative to that about the midplane. However, 
-perturbations near the disk surface may not be energetically important
-in practice because the disk atmosphere contains little mass.
-
-% We also find this trend for fixed $\beta$ but
-% increasing 
-% $\khat$.
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvectorvz_beta0d01} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/eigenvectorvz_beta0d1} 
-  \caption{Numerically-calculated fundamental VSI
-    mode in a nearly vertically isothermal disk ($\Gamma=1.011$) with
-    $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ and a dimensionless 
-    thermal relaxation timescale $\beta = 0.01$
-    (top) and $\beta=0.1$ (bottom). The
-    real (imaginary) part of the vertical velocity perturbation
-    $\delta v_z$ is plotted as the solid (dotted)
-    line. % The eigenfunction has been normalized such that
-    % $\imag(\delta v_z)=0$ at the vertical boundaries, and
-    % $\mathrm{max}|\delta v_z|=1$.
-    \label{relax_eigenW_num}}  
-\end{figure}
-
-\subsection{Critical thermal relaxation timescale}\label{bcrit_num_test}
 % may need to explain why choose k=10

### 2015-03-12 appendix.tex
rough revision done

@@
 
-\subsubsection{Stability without vertical shear}
+\subsubsection{Stability without vertical shear}\label{stable_novshear}
 In the absence of vertical shear $q=0$, Eq. \ref{simple_growth} can be
@@
 this case the solutions are Hermite polynomials, $W\propto
-\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma}$ is real so the disk
+\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma} = \hat{\omega}$ is real so the disk
 is stable, consistent with \S\ref{integral_relation}. The
 low-frequency solution to Eq. \ref{iso_disp_full} correspond to
-inertial waves \citep{balbus03}.
+inertial waves \citep{balbus03}. In the low frequency limit
+$|\hat{\omega}|\ll 1$ and assuming $L\geq 1 $, the 
+dispersion relation is $\hat{\omega}^2\khat^2 = L$, or
+$\hat{\omega}\propto \khat^{-1}$ for fixed $L$. (Low frequency thus 
+require $\khat\gg \sqrt{L}$.) This inverse relation has been qualitatively
+observed in numerical simulations of \cite{stoll14}.
 

### 2015-03-12 linear_adia.tex
updating results
rough revision done

@@
 % find the adiabatic dispersion relation, 
-This is a $6^\mathrm{th}$ order 
-polynomial in $\hat{\sigma}$:  
+For ease of analysis, we square Eq. \ref{adia_disp0} to obtain a 
+a polynomial in $\hat{\sigma}$,  
 \begin{align}
@@
 This is a pair of simultaneous equations for
-$\hat{\omega}$ and $\beta$. In Appendix \ref {iso_stable} we discuss
-stable solutions without vertical shear which yield
-$\hat{\omega}\propto \khat^{-1}$ for $\khat\gg 1$
-(Eq. \ref{iso_stable_disp}).  We are thus motivated to seek solutions
-to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
+$\hat{\omega}$ and $\beta$. Now, low-frequency axisymmetric waves in
+the absence of vertical shear are stable and  satisfy
+$\hat{\omega}\propto \khat^{-1}$ (Appendix \ref
+{stable_novshear}, Eq. \ref{iso_disp_full}).  We are thus motivated to
+seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
 % we wish to seek a criterion independent of k 
@@
 finite thermal relaxation rapidly stabilizes the VSI.
-
-% In Fig. \ref{relax_bcrit} we plot the critical thermal relaxation
-% timescale obtained from solving Eq. \ref{relax_disp_fund} numerically 
-% for a disk with $\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$, for which
-% we have $\beta_\mathrm{crit}=0.125$ from Eq. \ref{iso_vsi_cond}. Our
-% upper estimate agrees well with the numerical solution for $\khat\gtrsim
-% 10$.
-
-% \begin{figure}
-%   \includegraphics[width=\linewidth]{figures/bcrit_theory}
-%   \caption{Condition on the thermal relaxation timescale $\beta$ for
-%     the existence of unstable fundamental VSI mode in a vertically
-%     isothermal disk with $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$.  
-%     The dotted line denotes the asymptotic value
-%     $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$ for large $\khat$. 
-%     \label{relax_bcrit}}  
-% \end{figure}
-
-% \subsubsection{Lower limit on the thermal relaxation timescale for the
-%   fundamental adiabatic VSI mode}
-% {\bf Note: for reference only. Not to be included in paper.} 
-% Fig. \ref{relax_bcrit} indicates there exists a lower limit on the
-% thermal relaxation time for the system to behave adiabatically and
-% support the fundamental adiabatic VSI. This critical
-% timescale satisfies Eq. \ref{relax_cond1}---\ref{relax_cond2} with 
-% \begin{align}
-%   \beta \sim \beta_0\khat^{1/2},\quad \hat{\omega}  \sim
-%   \hat{\omega}_0\khat^{-1/2} \text{ as } \khat\to\infty.
-% \end{align}
-% The constants $\beta_0$ and $\hat{\omega}_0$ are determined by the
-% simultaneous equations
-% \begin{align}
-%   & 0 = \gamma\beta_0^2(1-\gamma) + 2\epsilon q \gamma
-%   \beta_0\hat{\omega}_0 - \hat{\omega}_0^2 +
-%   \beta_0^2\gamma^2\hat{\omega}_0^4.\\
-%   & 0= \beta_0(1-\gamma) + \epsilon q \hat{\omega}_0 - \epsilon q
-%   \gamma^2 \beta_0^2\hat{\omega}_0^3. 
-% \end{align}
-% For $\epsilon = 0.05$, $q=-1$ and $\gamma=1.4$, the appropriate
-% solution is $\beta_0 = 10.33$ and $\hat{\omega}_0 = 0.74$.
-
-% If we follow through the solution method as before, we obtain a
-% dispersion relation 
-% \begin{align}
-%   &\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) \notag\\ &=
-%   \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
-%   \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}\label{adia_disp}.
-% \end{align}
-% This equation is the same as Eq. 54 in \cite{lubow93} when $q=0$ and
-% the low-frequency, Keplerian approximation is made for the latter.
-
-% Note that when $\gamma>1$ and $q=0$, the term with $\ln\rho^{\prime
-%   2}$ in the integrand (third term on the RHS) represents
-% stabilization by the vertical entropy gradient (real buoyancy frequency). 
-% It gives an imaginary contribution to $\hat{\sigma}$ when $q\neq0$,
-% but this is a factor $|\epsilon/\hat{k}|$ smaller than the real
-% (stabilizing) contribution, 
-% since $\epsilon\ll 1$ for a thin disk and we (usually) consider
-% $\hat{k}\gg1$. Thus, instability is expected to be associated with the 
-% last term on the RHS of Eq. \ref{adia_integral}.
-
-% \subsubsection{Explicit solutions in the thin-disk limit}\label{adia_explicit}
-% We can construct solutions in the thin-disk limit, in which 
-% $\ln(\rho/\rho_0) \simeq -\zhat^2/2$. Eq. \ref{adia_iso3} becomes   
-% \begin{align}
-%   0 = \delta v_z^{\prime\prime} - \zhat A \delta v_z^\prime + \left(B
-%     - C \zhat^2\right)\delta v_z, \label{adia_iso4} 
-% \end{align}
-% with
-% \begin{align}
-%   &A \equiv 1 + \ii \epsilon q \hat{k},\\
-%   &B \equiv \hat{\sigma}^2\left(\frac{1}{\gamma} + \hat{k}^2\right) -
-%   \left(\frac{1}{\gamma} + \ii \hat{k} \epsilon q\right)\\
-%   &C \equiv \frac{\left(\gamma-1\right)}{\gamma}\hat{k}^2\left(1 - \frac{\ii
-%       \epsilon q}{\hat{k}}\right).\label{adia_thin}
-% \end{align}
-
-% Inserting the definition of $B$ we get
-% \begin{align}
-%   &\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) \notag\\ &=
-%   \left(\frac{1}{\gamma}-\frac{1}{2}\right) +\frac{\ii}{2}\epsilon q
-%   \hat{k} + \frac{1}{2}\left(2M+1\right)\left(A^2 + 4C\right)^{1/2}\label{adia_disp}. 
-% \end{align}
-% This gives the eigenfrequency $\hat{\sigma}$ and completes the solution
-% description.
-
-% %Recalling that $B$ depends $\hat{\sigma}$, Eq. \ref{adia_disp} gives
-% %the eigenfrequency as a function of disk paramters.
-
-% \subsubsection{The case of $\gamma=2$}
-% In principle, we can solve Eq. \ref{adia_disp} for the mode
-% frequencies and growth rates as a function of the disk parameters and
-% the radial wavenumber. This is complicated in 
-% general, but simplifies for $\gamma=2$, for which the quantity
-% $A^2+4C$ is real. In this case we find
-% \begin{align}
-%   \hat{\nu}^2 =
-%   \frac{1}{2}\frac{(2M+1)}{\sqrt{1+2\hat{k}^2}}&\left[
-%     \sqrt{1-\frac{4M(M+1)\hat{k}^2\epsilon^2q^2}{(2M+1)^2(1+2\hat{k}^2)}}\right.\notag\\
-%   &\phantom{=}
-%   \left.- \sqrt{1 - \frac{\hat{k}^2\epsilon^2q^2}{(1+2\hat{k}^2)}}\right].
-% \end{align}
-% In the limit of a thin disk $\epsilon\to0$ or weak shear $|q|\to0$, we
-% have 
-% \begin{align}
-%   \hat{\nu}^2 = \frac{\epsilon^2
-%     q^2}{4(2M+1)}\frac{\hat{k}^2}{(1+2\hat{k}^2)^{3/2}},\label{gam2_growth}
-% \end{align}
-% and the maximum growth rate occurs at $\hat{k}=\hat{k}_\mathrm{opt}=1$,  
-% \begin{align}
-%   \mathrm{max}(\hat{\nu}) &= \frac{|\epsilon
-%     q|}{2(3^{3/4}\sqrt{2M+1})}\label{gam2_max_growth}.
-%   % & \leq \frac{|\epsilon
-%   % q|}{2\times3^{3/4}}\notag.
-% \end{align}
-% Unlike isothermal perturbations, Eq. \ref{gam2_max_growth} shows that
-% maximum growth rate occurs at $M=0$, and growth rates decrease with increasing $M$. This
-% is despite the fact that in the thin-disk approximation 
-% $d\Omega^2/dz\propto z$, so there is infinite vertical shear
-% available.
-
-% The acceptable solution for $\alpha$ when $\gamma=2$ is
-% \begin{align}
-%   \alpha = \frac{1}{2}\left[\left(1+\ii\epsilon q \hat{k}\right) -
-%     \sqrt{1 + \hat{k}^2\left(2-\epsilon^2 q^2\right)}\right].\label{gam2_alpha}
-% \end{align}
-% Since $\epsilon\ll 1$, $\real\alpha<0$, so the vertical velocity
-% eventually decays for large heights. For $\hat{k}\gg 1$ the
-% perturbation decays rapidly away from the midplane. This is again
-% unlike isothermal perturbations, whose magnitude increase with $|z|$.
-
-% %gauss decay means boundary conditions have no influence, all perts go
-% %to zero
-
-% \subsubsection{Comparison with $\gamma=1$}\label{adia_compare_gam1}
-% We can compare growth rates obtained for $\gamma=2$ to those for
-% $\gamma=1$ in the previous section. If we compare growth rates at their 
-% respective optimum radial wavelengths, we find
-% that for $\gamma=2$ the \emph{maximum} growth rate, occuring
-% at $M=0$, is still $\sim2$ times smaller than the \emph{minimum} growth
-% rate for $\gamma=1$. That is,
-% \begin{align}
-%   &\mathrm{max}(\hat{\nu}) \geq \frac{|\epsilon q|}{2\sqrt{1+|\epsilon
-%       q|}} & \text{at }\hat{k} = \hat{k}_\mathrm{opt} \text{ for }   \gamma=1,  \\
-%   &\mathrm{max}(\hat{\nu}) \leq \frac{|\epsilon q|}{2\times 3^{3/4}}
-%            & \text{at }\hat{k} = \hat{k}_\mathrm{opt} \text{ for } \gamma=2.
-% \end{align}
-
-% We can also compare growth rates in the limit $\hat{k}\gg1$ using
-% Eq. \ref{simple_growth} and Eq. \ref{gam2_growth},
-% \begin{align}
-%   &\hat{\nu} \geq \sqrt{\frac{|\epsilon q|}{2\hat{k}}}
-%   &  \text{as }  \hat{k}\to\infty  \text{ for }  \gamma=1, \\
-%   &\hat{\nu} \leq \frac{|\epsilon q|}{2^{7/4}\sqrt{\hat{k}}}
-%   &\text{as }  \hat{k}\to\infty  \text{ for }  \gamma=2.\label{gam2_growth_rate}
-% \end{align}
-% Then the growth rate of perturbations with large $|\hat{k}|$ in a
-% $\gamma=2$ disk is at most $2^{-5/4}\sqrt{|\epsilon q|}$ times that in a
-% $\gamma=1$ disk. For $\epsilon = 0.1$ and $|q|=1$, this factor is
-% $\sim 0.1$.
-
-% In Fig. \ref{adia_growth} we plot the eigenfrequencies as a
-% function of the radial wavenumber $\hat{k}$, for a range of adiabatic
-% indices, using Eq. \ref{adia_disp}. The low-frequency approximation
-% is questionable for $\hat{k}\lesssim4$ since
-% $|\hat{\sigma}|^2\gtrsim0.2$. For large $\hat{k}$ the presence of a
-% positive vertical entropy gradient is strongly stabilizing. This is
-% because the vertical shear grows as $z$ away from the midplane (and
-% has a maximum when the thin-disk approximation is relaxed), but square
-% of the bouyancy frequency, which is stabilizing, grows as $z^2$ away
-% from the midplane. %The entropy $P/\rho^\gamma$ formally diverges   
-% %for $z\to\infty$ even without the thin-disk approximation.
-
-% \begin{figure}
-%   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/rate_theory_om}
-%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_nu}
-%   \caption{Real frequency (top) and growth rate (bottom) of the fundamental VSI mode in vertically 
-%     isothermal disks ($\Gamma=1$) subject to adiabatic perturbations,
-%     calculated in thin disk approximation from Eq. \ref{adia_disp}
-%     with $M=0$, $\epsilon=0.1$ and  
-%     $|q|=1$. Frequencies are shown as a function of the radial
-%     wavenumber for different values of the adiabatic index
-%     $\gamma$. For $\gamma\equiv 1$ the perturbations are isothermal
-%     and there is no vertical entropy gradient.\label{adia_growth}}  
-% \end{figure}

### 2015-03-12 numerical.tex
rough revision done

@@
 show a meridional visualization of this mode in
-Fig. \ref{lowfreq_eigenfunc_2d}. Although there is a constant vertical
+Fig. \ref{lowfreq_eigenfunc_2d}. Radial velocities are typically much
+smaller than vertical velocities. Although there is a constant vertical
 velocity throughout the fluid column, most of the momentum is
@@
 
-\subsubsection{Surface modes: $\khat=30$} 
+\subsubsection{Comment on `surface modes'} 
 Our standard setup with $\khat=10$ in 
@@
 $\zmax\geq 5H$, examples of which are marked by `s'. These modes have
-been discussed in the above studies. They typically have the largest
-growth rates but require sufficiently large vertical domain size and their eigenvalues
-vary with $\zmax$.
+been discussed in the aforementioned studies. They typically have the
+largest growth rates but require sufficiently large vertical domain
+size and their eigenvalues vary with $\zmax$.
 
@@
 
-\subsection{Effect of thermal relaxation}
+\subsection{Effect of thermal relaxation}\label{therm_relax_eff}
 We now consider the effect of thermal relaxation by increasing
@@
 show that the fundamental mode is not sensitive to boundary
-conditions. This suggests that tracking fundamental mode would be a
+conditions. Thus, tracking fundamental mode is a
 more robust way of assessing stability as a function of $\beta$,
@@
 
+Fig. \ref{lowfreq_eigenfunc_cool}---\ref{lowfreq_eigenfunc_2d_cool}
+shows the fundamental mode with $\khat=10$ in a disk with 
+$\beta=0.1$. These plots can be compared the case with rapid thermal
+relaxation (Fig. \ref{lowfreq_eigenfunc}---\ref{lowfreq_eigenfunc_2d}). We see 
+that increasing $\beta$ causes the region characterized by $W\sim z$ and  
+$\delta v_z\sim$ constant closer to the mid-plane, with oscillatory
+behavior emerging near vertical boundaries where buoyancy becomes
+important. This leads to the `tilted' pressure perturbations seen in
+Fig. \ref{lowfreq_eigenfunc_2d_cool}
+(cf. Fig. \ref{lowfreq_eigenfunc_2d}). However, most of the meridional
+momentum is still contained within $|z|\lesssim 2H$ as for 
+the case with $\beta=10^{-3}$.
+
 \begin{figure}
@@
   1cm]{figures/eigenvectorvz_beta0d1}
-  \caption{Same as Fig. \ref{lowfreq_eigenfunc_cool} but with a
+  \caption{Same as Fig. \ref{lowfreq_eigenfunc} but with a
     thermal relaxation timescale $\beta=0.1$. 
@@
   timescale}\label{bcrit_num_test}
+We now test the critical thermal relaxation timescale given by
+Eq. \ref{iso_vsi_cond}. We follow the fundamental mode by starting
+from eigenvalues for $\beta=10^{-3}$ and slowly increase the thermal
+relaxation timescale up to $\beta=10^3$. Fig. \ref{bcrit_compare1}
+show the obtained growth rates, together with the critical 
+thermal relaxation timescale $\beta_\mathrm{crit}=0.125$ for our
+fiducial disk parameters.
 
-Fig. \ref{bcrit_compare1} shows the growth rate $\nu$ of the
-fundamental VSI as a function of $\beta$ for $\khat\in[1,10^2]$. We
-also plot the critical thermal relaxation timescale
-$\beta_\mathrm{crit}=0.125$ given by Eq. \ref{iso_vsi_cond} for the present
-disk parameters.
+For $5\lesssim\khat\lesssim30$, we find $\beta_\mathrm{crit}$ provides
+an accurate prediction of the upper limit to the thermal relaxation 
+timescale for the fundamental mode. For smaller $\khat$, 
+$\beta<\beta_\mathrm{crit}$ the VSI can operate slightly beyond
+$\beta_\mathrm{crit}$ (recall that $\beta_\mathrm{crit}$ was derived
+by assuming $\khat\gg 1$). For $\khat\geq 50$, we find perturbations
+acquire large amplitudes near the vertical boundaries. Complications
+% to our simple solution procedure 
+may arise from surface modes. Nevertheless, growth rates rapidly
+decrease as $\beta$ approach and increase beyond
+$\beta_\mathrm{crit}$.
 
-For $5\lesssim\khat\lesssim30$, we find $\beta_\mathrm{crit}$ provides an
-accurate prediction of the upper limit to the thermal relaxation 
-timescale for  the fundamental mode. For smaller $\khat$,
-$\beta<\beta_\mathrm{crit}$ is a sufficient condition for
-instability. For larger $\khat$, we find $\nu$ does not reach zero but
-rapidly decreases as $\beta$ approaches and increases beyond
-$\beta_\mathrm{crit}$. In the latter case we find the perturbation
-acquires large amplitudes near the vertical boundaries. Boundary
-conditions may thus may play a role for these modes that is not
-accounted by our analytical discussion.
-
-Fig. \ref{bcrit_compare1} shows that 
-$\beta_\mathrm{crit}$ is a good indication for the maximum
-thermal relaxation timescale to permit the fundamental VSI. For
-$\beta\lesssim O(10^{-2})$, growth rates are insensitive to thermal
-relaxation. For this disk model, \citeauthor{nelson13}
-found a thermal relaxation timescale of $\beta\gtrsim 0.6$ stabilized
-the VSI in non-linear hydrodynamic simulations. This is consistent
-with our results.
+Fig. \ref{bcrit_compare1} shows that $\beta_\mathrm{crit}$ provides a  
+practical indication for the maximum thermal relaxation timescale to
+permit the fundamental VSI. % From our discussion in 
+% \S\ref{therm_relax_eff}, which indicate the fundamental mode typically
+% dominant 
+\cite{nelson13} found a thermal relaxation timescale of $\beta\gtrsim
+0.6$ stabilized the VSI in non-linear hydrodynamic simulations. This 
+is consistent with our results.
 
@@
    \includegraphics[width=\linewidth]{figures/gcorr_compare2} 
-   \caption{Growth rate of the fundamental VSI as
-     a function of the thermal relaxation timescale $\beta$. The disk
-     parameters are $\Gamma=1.011$, $\gamma=1.4$, and 
-     $(p,q,\epsilon)=(-1.5,-1,0.05)$. The vertical line
+   \caption{Growth rate of the fundamental VSI in the fiducial disk
+     model as a function of the thermal relaxation timescale $\beta$
+     The vertical line
      is the critical thermal timescale $\beta_\mathrm{crit}$  obtained
@@
 % behavior at larger kx and finite bcool not understood. 
-We test the dependence of the critical thermal relaxation timescale 
-$\beta_\mathrm{crit}$ for the fundamental VSI on disk parameters
+We further test the dependence of the critical thermal relaxation timescale 
+$\beta_\mathrm{crit}$ on disk parameters
 (Eq. \ref{iso_vsi_cond}).  We use the previous case study as a
@@
 
-The numerically-obtained $\beta_\mathrm{crit}$ is shown in
+The numerically obtained $\beta_\mathrm{crit}$ is shown in
 Fig. \ref{bcrit_compare} in comparison with Eq. \ref{iso_vsi_cond}.
@@
 agreement improves with decreasing  $|q|$,  $\epsilon$ and increasing
-$\gamma$, i.e. weaker instability, although there is a noticeable difference
+$\gamma$, i.e. weaker instability. There is a noticeable difference
 as $\gamma\to1$ because the derivation of $\beta_\mathrm{crit}$
@@
 (Fig. \ref{bcrit_compare1}), shows that $\beta_\mathrm{crit}$ as given
-by Eq. \ref{iso_vsi_cond} gives a measure of the characteristic
-thermal timescale below which perturbations are effectively isothermal
-so that the fundamental VSI operates.
+by Eq. \ref{iso_vsi_cond} characterizes the thermal timescale below
+which perturbations are effectively isothermal 
+so that the VSI operates.
 
@@
 $\Gamma=1.4$ and $(p,q,\epsilon)=(0,-1,0.05)$ as simulated in
-\cite{nelson13}. In this case we set $\zmax\simeq2.2H$, which
-is close to the zero-density surface.
+\cite{nelson13}. In this case we set $\zmax=0.99H_s$.
 
 Fig. \ref{gcorr_compare_vnoniso} plots the fundamental VSI growth
-rates as a function of $\beta$ for $\gamma\in[1.4,2.5]$. In agreement with \citeauthor{nelson13}, in 
+rates as a function of $\beta$ for $\gamma\in[1.4,2.5]$. In agreement
+with \citeauthor{nelson13}, in  
 the neutrally-stratified case $\gamma=\Gamma$ the disk is unstable 
@@
 \end{figure}
-
-% \subsection{VSI in a stably stratified disk}
-% We demonstrate the (strong) stabilizing effect of a  
-% positive vertical entropy gradient ($N_z^2\geq0$) by setting $\gamma=2$.  
-% Since we expect the perturbations to decay rapidly
-% away from the midplane (see \S\ref{analytic_adia}), we consider a
-% smaller domain with $\zmax = 2\epsilon r$. 
-% Other disk and perturbation parameters are the same as in
-% \S\ref{vertiso_pertiso}.
-
-% Fig. \ref{lowfreq_eigen_adia} show the eigenvalues for this
-% problem. The growth rates are much smaller than
-% those in the previous section. The fundamental VSI mode is that with
-% $\mathrm{max}\nu$. Its expected and numerically-calculated growth
-% rates compares well, with 
-% \begin{align*}
-%   \nu = 0.03751 \epsilon \Omega_k &\quad \text{(from
-%     Eq. \ref{gam2_growth_rate})},\\
-%   \nu = 0.03672 \epsilon \Omega_k &\quad \text{(numerical)}.
-% \end{align*}
-% The fundamental mode is plotted in Fig. \ref{lowfreq_eigenfunc_adia},
-% and clearly show that perturbations are rapidly stabilized away from
-% the midplane. For the adopted disk and perturbation parameters,
-% Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$, corresponding
-% to a characteristic decay length scale of $\sqrt{2/|\alpha|}\epsilon r\simeq 0.2\epsilon r$,
-% which is consistent with that observed in  Fig. \ref{lowfreq_eigenfunc_adia}.
-
-% This decay lengthscale is rougly the height within which
-% vertical shear is larger than the buoyancy frequency squared. That is,
-% \begin{align*}
-% \left|r\frac{d\Omega^2}{dz}\right|\gtrsim N_z^2
-% \end{align*}
-% for
-% \begin{align*}
-%   \left|\frac{z}{\epsilon r}\right| \lesssim \frac{\epsilon|q|\gamma}{\gamma-1},
-% \end{align*}
-% in a thin, vertically isothermal disk. This evaluates to $|z|\lesssim
-% 0.2\epsilon r$ for the current disk model, consistent with
-% Fig. \ref{lowfreq_eigenfunc_adia}.
-
-% Furthermore, the growth rate to should be no larger than the
-% maximum vertical shear available within this decay lengthscale. Using 
-% Eq. \ref{max_growth} as a rough guide, 
-% \begin{align}
-%   \frac{\nu}{\epsilon\Omega_k}<\frac{\epsilon |q|^2\gamma}{2(\gamma-1)},
-% \end{align}
-% or $\nu< 0.1\epsilon\Omega_k$ in our disk model,
-% consistent with numerical results.
-
-% \begin{figure}
-%   \includegraphics[width=\linewidth]{figures/eigenvalues_adia}
-%   \caption{Eigenvalues in for the
-%     VSI in a nearly vertically isothermal disk
-%     ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
-%     evolved adiabatically with $\gamma=2$. The perturbation radial
-%     wavenumber is $\khat=20\pi$. \label{lowfreq_eigen_adia}
-%   }
-% \end{figure}
-
-% \begin{figure}
-%   \includegraphics[width=\linewidth]{figures/eigenvectorvz_adia}
-%   \caption{Fundamental VSI mode in a nearly vertically
-%     isothermal disk ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
-%     evolved adiabatically
-%     with $\gamma=2$. This eigenfunction 
-%     corresponds to the top-left eigenvalue displayed in 
-%     Fig. \ref{lowfreq_eigen_adia} (largest $|\nu|$).  
-%     The  real (imaginary) part of $\delta v_z$ are shown as solid
-%     (dotted) lines. 
-%     \label{lowfreq_eigenfunc_adia}
-%   }
-% \end{figure}
-
-% Although a positive vertical entropy gradient is stabilizing, we note
-% that $N_z^2\propto z^2$  while vertical shear $rd\Omega^2/dz\propto
-% z$, so that there is always a region sufficiently close to the
-% midplane in which vertical shear dominates over bouyancy and the VSI
-% can operate. However, the growth rates are small and the instability
-% only affects the regions close the midplane. These modes may not be
-% important in practice.
-
-% \subsection{Other adiabatic indices}
-% To see the transition from neutrally-bouyant to stably stratified cases above, 
-% we repeated the previous calculation with $\gamma\in[1,2]$. 
-% We restore the vertical domain to $\zmax=10\epsilon r$. 
-% Fig. \ref{adia_growth_num} shows the eigenfrequencies of the
-% fundamental VSI mode.
-
-% %To select the appropriate
-% %eigenvalues we vary the parameters slowly away from a reference solution.   
-% % The numerically-determined eigenfrequencies are qualitatively
-% % consistent with the discussion in \S\ref{adia_compare_gam1}.
-% There is good agreement between Fig. \ref{adia_growth_num} and
-% Fig. \ref{adia_growth} for $\khat=k_xH\gtrsim1$. There is a
-% mismatch for smaller wavenumbers, especially in the growth
-% rates. Actual growth rates decrease slower as $\khat\to0$
-% than that in Fig. \ref{adia_growth}. Disagreement at small
-% $\khat$ is not unexpected because the low-frequency
-% approximation used to obtain Fig. \ref{adia_growth} becomes invalid since
-% $|\omega|\sim\Omega_k$ for small $\khat$.
-
-% Given the number of approximations made in \S\ref{adia_compare_gam1},
-% Fig. \ref{adia_growth_num} compares well with 
-% Fig. \ref{adia_growth}. The main qualitative features of the full
-% numerical solution in Fig. \ref{adia_growth_num} are captured by
-% Fig. \ref{adia_growth}: stabilization for 
-% $\khat\to0$ and $\khat\to\infty$,
-% stabilization for increasing $\gamma$ (but ineffective at small $\khat$), 
-% and the shift of the most unstable wavenumber to smaller values as
-% $\gamma$ is increased.
-
-% \begin{figure}
-%   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0cm]{figures/compare_eigen_real1}
-%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/compare_eigen_imag1}
-%   \caption{Real frequency (top) and growth rate 
-%     (bottom) of the fundamental VSI in a nearly vertically   
-%     isothermal disks ($\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$, 
-%     subject to adiabatic perturbations with different values of $\gamma$. 
-%     The eigenfrequencies are calculated by numerically
-%     solving the full linear eigenvalue problem. This plot is to
-%     be compared with Fig. \ref{adia_growth}. \label{adia_growth_num}}   
-% \end{figure}

### 2015-03-12 appendix.tex
generalized bcrit (attempt 1)

@@
 \end{align}
+
+\subsection{Limiting form with real frequencies and $\khat\gg 1$}\label{disp_neut_limit}
+For neutral modes with real $\hat{\sigma}=\hat{\omega}$ and large
+wavenumber $\khat\gg 1$, the real and imaginary parts of
+Eq. \ref{relax_disp} are approximately 
+\begin{align}
+  &0 =M(M+1) + 4\beta\epsilon q M(M+1)\left(\hat{\omega}\khat\right) 
+ +\left[1 +
+    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
+ + 2\epsilon q \gamma\beta \left(\hat{\omega}\khat\right)^3 -  \left(\hat{\omega}\khat\right)^4 
+  + \beta^2\gamma^2\hat{\omega}^2
+  \left(\hat{\omega}\khat\right)^2,\label{relax_cond1} \\
+   &0= \epsilon q M (M+1)\khat +
+   \beta(1-\gamma)(1+2M)^2\khat\left(\hat{\omega}\khat\right) 
+   + \epsilon q \khat \left(\hat{\omega}\khat\right)^2 -
+   2\beta\hat{\omega}\left(\hat{\omega}\khat\right)^2
+   - \epsilon q
+   \gamma^2\beta^2\hat{\omega}\left(\hat{\omega}\khat\right)^3 
+   +
+   2\beta\gamma\hat{\omega}\left(\hat{\omega}\khat\right)^4 \label{relax_cond2}. 
+\end{align}
+These equations are used to derive a critical thermal timescale in
+\S\ref{iso_vsi_beta_crit}. For a given $M$, we assume
+$|\hat{\omega}\khat|$ remain constant but $\khat\gg 1$ and
+$|\hat{\omega}|\ll 1$.  Then for a thin disk $\epsilon \ll 1$,
+we neglect the second, fourth and last term in Eq. \ref{relax_cond1};
+and neglect the last three terms in Eq. \ref{relax_cond2}. This gives
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}.
 

### 2015-03-12 linear_adia.tex
generalized bcrit (attempt 1)

@@
 balancing constants and coefficients of $\zhat^2$,
-% we obtain  
-% \begin{align}
-%  0 =& \delta v_{z1}^{\prime\prime} - z\left(1 + \ii\epsilon q
-%    k\right)\delta v_{z1}^\prime\notag\\
-%  &+ \left[2\hat{\sigma}\delta\hat{\sigma}\left(1 + \khat^2\right) +
-%    \left(\hat{\sigma}^2 - 1\right)\delta\chi \right.\notag\\
-% &\phantom{++} \left. + \zhat^2\left(\khat^2 -
-%      \ii\epsilon q \khat\right)\delta\chi\right]\delta v_z\notag\\
-%  &+\left[\hat{\sigma}^2\left(1 + \khat^2\right) - \left(1+\ii\epsilon
-%      q \khat\right)\right]\delta v_{z1}.\label{nearly_iso4}
-% \end{align}
-% In Eq. \ref{nearly_iso_pert} and \ref{nearly_iso4}, $\delta v_z$ is to
-% a solution to Eq. \ref{nearly_iso_explicit} with 
-% $\beta=0$. 
-% We could proceed by considering the entire set of polynomial solutions
-% for $W$, and hence for $\delta v_z$, as derived in \S\ref{iso_poly}, but it
-% is simplest to look at the fundamental mode. 
-%Inserting the above expressions for $\delta v_z$, $\delta v_{z1}$ and
-%$\hat{\sigma}^2$ into Eq. \ref{nearly_iso4}, then balancing
-%coefficients of $\zhat^2$ and constants, we require
+
 \begin{align}
@@
 very small scales.
-% We plot Eq. \ref{nearly_iso_dsig} in
-% Fig. \ref{domegadbeta} with 
-% $\epsilon=0.05$, $q=-1.0$ and $\gamma=1.4$. For this disk model the
-% optimum wavenumber for stabilization is $\hat{k}\sim 5$---$6$.
-
-% This contrasts to the case without vertical shear
-% ($q=0$), in which $\p\hat{\sigma}/\p\beta\to\mathrm{constant}$ as 
-% $\hat{k}\to\infty$.
-
-% \begin{figure}
-%   \includegraphics[width=\linewidth]{figures/domegadbeta}
-%   \caption{Rate of change of the fundamental VSI eigenfrequency
-%   $\sigma$ with respect to the introduction of a small but finite
-%   cooling time. The plotted quantity is $\p\hat{\sigma}/\p\beta$ at $\beta=0$, evaluated using
-%   Eq. \ref{nearly_iso_dsig} with 
-%   disk parameters $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The real
-%   (imaginary) part of $\p\hat{\sigma}/\p\beta$ is shown as the solid
-%   (dotted) lines. 
-%   \label{domegadbeta}}  
-% \end{figure}
 
@@
 \end{align}
-% We remark that for $q=0$ and $\beta\to\infty$, Eq. \ref{adia_disp0} is
-% the same as Eq. 54 in \cite{lubow93} when the low-frequency, Keplerian
-% approximation is made in the latter. 
-% Eq. \ref{nearly_iso_explicit} is same form
-% as the governing equation for the adiabatic case
-% (Eq. \ref{adia_iso4}). However, the coefficients $\widetilde{B}$ and $\widetilde{C}$
-% have a complicated dependence on $\hat{\sigma}$. 
-% Applying the same solution method as in \S\ref{adia_explicit}, we 
-% find the adiabatic dispersion relation, 
 For ease of analysis, we square Eq. \ref{adia_disp0} to obtain a 
@@
 The dispersion relation $\hat{\sigma}=\hat{\sigma}(\khat)$ generally
-requires a numerical solution. However, for the fundamental mode
-($M=0$) some analytic progress can be made, which we consider next.
+requires a numerical solution. However, simple solutions may be
+obtained in limiting cases which we consider below.
 
-% It is simplest to solve Eq. \ref{relax_disp} numerically. In
-% Fig. \ref{relax_disp_fig} we plot the growth rate of the fundamental
-% ($M=0$) VSI mode as a function of $\khat$ for a range of
-% dimensionless thermal relaxation timescales $\beta$, in a disk with
-% $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. Increasing $\beta$ even
-% slightly above zero significantly reduces the growth rates for
-% $\khat\gg 1$. (There is little effect at small $\khat$ but this case is
-% not of interest because the local model breaks down.) Thus, 
-% a growing fundamental VSI mode exists only for nearly
-% isothermal perturbations ($\beta\ll 1$).  The plot is also consistent with our previous
-% discussion that there exists an intermediate wavenumber for which
-% finite thermal relaxation has maximal effect (for small $\beta$).
-
-% \begin{figure}
-%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm 0cm]{figures/rate_theory_grow_relax}
-% %  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm 1cm]{figures/rate_theory_grow_relax2}
-%   \caption{Growth rate of the fundamental VSI mode in a disk with
-%     dimensionless thermal relaxation timescales
-%     $\beta\in[10^{-3},1]$.  Other disk parameters are
-%     $\epsilon=0.05$, $q=-1$ and $\gamma=1.4$. The growth rates are
-%     calculated from the dispersion relation Eq. \ref{relax_disp}. 
-%     \label{relax_disp_fig}}  
-% \end{figure}
-
-% ate_theory, eps=0.05, smallq=-1d0, gmma=1.4, bmax=1.0, ncases=11,
-%mmode=0.0, root=1, legend=[0.102,0.14,5.4,0.355], yrange=[0,6]
+% However, for the fundamental mode
+% ($M=0$) some analytic progress can be made, which we consider next.
 
 \subsection{Upper limit on the thermal relaxation timescale for the
-  fundamental VSI mode}\label{iso_vsi_beta_crit}
+  VSI}\label{iso_vsi_beta_crit}
+Here we estimate the maximum thermal relaxation timescale
+$\beta$ that allows the VSI. We consider wavenumbers such that
+$\khat^2\gg M+1$ which, as we will find, correspond to low frequency
+modes.
 
-Here we estimate the maximum thermal relaxation timescale
-$\beta$ that allows the fundamental VSI mode for $\khat\gg1$.   
-% Notice in Fig. \ref{relax_bcrit} the upper limit on $\beta$ for the
-% fundamental isothermal VSI tends to a constant for large
-% $\khat$. Using this clue, we can calculate this critical $\beta$ by
-% considering $\khat\to\infty$. We proceed as follows. 
-At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
-is real. The dispersion relation for the fundamental mode is then
-\begin{align}\label{relax_disp_fund}
-  0 = \sum_{l=1}^{6}c_l\hat{\omega}^l \quad \text{for $M=\hat{\nu}=0$}.  
+%At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$ is real.  
+%This is a pair of simultaneous
+%equations for $\hat{\omega}$ and $\beta$. To proceed, we
+%consider further limits.
+
+Our starting point is Eq. \ref{relax_cond1}---\ref{relax_cond2} which
+approximates the dispersion relation Eq. \ref{relax_disp} at marginal
+stability with $\khat\gg 1$. We consider modes with $M=0$ (the
+fundamental mode) or  $M=O(1)$ and low frequency $|\hat{\omega}|\ll
+1$ such that $|\hat{\omega}\khat|$ is $O(1)$ because $\khat\gg 1$. We
+also assume $|q|$ and $\beta$ are $O(1)$ or less. For a thin disk
+($\epsilon \ll 1$), we have approximately
+\begin{align}
+ 0 = &M(M+1) + \left[1 +
+    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2\notag\\
+   &-  \left(\hat{\omega}\khat\right)^4,\label{relax_cond_simp1}\\
+   0= &M(M+1) + \frac{\beta}{\epsilon q} (1-\gamma)(1+2M)^2
+   \left(\hat{\omega}\khat\right) + \left(\hat{\omega}\khat\right)^2. \label{relax_cond_simp2}
 \end{align}
-Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
-$\khat\gg1$, we find
+(We explain this in Appendix \ref{disp_neut_limit}.)
+
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} is a pair of
+simultaneous equations for $\hat{\omega}$ and the critical thermal
+timescale $\beta$ for marginal stability. If $\beta\ll 1$,
+then a simple solution to Eq. \ref{relax_cond_simp1} is 
 \begin{align}
-  0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
-  \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
-  &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
-  0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
-  - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
-    &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
-\label{relax_cond2}
+  \left(\hat{\omega}\khat\right)^2 \simeq M+1, 
 \end{align}
-This is a pair of simultaneous equations for
-$\hat{\omega}$ and $\beta$. Now, low-frequency axisymmetric waves in
-the absence of vertical shear are stable and  satisfy
-$\hat{\omega}\propto \khat^{-1}$ (Appendix \ref
-{stable_novshear}, Eq. \ref{iso_disp_full}).  We are thus motivated to
-seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
-% we wish to seek a criterion independent of k 
-% \begin{align}
-%   \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
-%\end{align}
-\begin{align}
-  \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
-  |\hat{\omega}\khat|=O(1). 
+which is the dispersion relation for low-frequency waves in vertically
+isothermal disks without vertical shear (see Appendix
+\ref{stable_novshear}). Then  
+\begin{align}\label{bcrit_gen}
+  \beta = \frac{\epsilon|q|}{(\gamma - 1)}
+  \frac{(1+M)^{3/2}}{(1+2M)^2} \leq \beta_\mathrm{crit} 
 \end{align}
-Then
-\begin{align}
-  \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
+where 
+\begin{align}\label{iso_vsi_cond}
+  \beta_\mathrm{crit}  \equiv \frac{\epsilon|q|}{\gamma-1}
 \end{align}
-from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
-that
-\begin{align}\label{beta_crit0}
-  \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
-        q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
-\end{align}
-at marginal stability for perturbations with large $\khat$.
+is the critical thermal timescale for the fundamental mode
+($M=0$). However, according to Eq. \ref{bcrit_gen}  the critical
+thermal timescale decreases with increasing  $M$. We thus expect
+higher order modes to be stabilized more rapidly by finite thermal
+relaxation $\beta$. When $\beta>\beta_\mathrm{crit}$, all modes
+(satisfying the above assumptions) are stabilized. We can interpret
+$\beta_\mathrm{crit}$ as the thermal timescale below which
+perturbations are effectively isothermal to allow the VSI.  We will confirm
+Eq. \ref{iso_vsi_cond} numerically in \S\ref{bcrit_num_test}.
 
-From our discussion in \S\ref{relax_pert}, which indicate introducing
-a finite thermal relaxation is stabilizing, we expect
-Eq. \ref{beta_crit0} is an upper limit. For a thin  
-disk $\epsilon\ll 1$, we thus infer 
-\begin{align}\label{iso_vsi_cond}
-  \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
-  \beta_\mathrm{crit} 
-\end{align}
-is required for the fundamental VSI to operate. That is, when
-$\beta<\beta_\mathrm{crit}$, perturbations are  effectively isothermal. 
-We will confirm Eq. \ref{iso_vsi_cond} numerically in
-\S\ref{bcrit_num_test}.
+(We remark that the critical thermal timescale $\beta_\mathrm{crit}$
+for the fundamental mode may be obtained more readily with fewer
+assumptions by setting $M=0$ in
+Eq. \ref{relax_cond1}---\ref{relax_cond2} from the onset. )
 
 Rapid thermal relaxation is required to overcome a stabilizing
-vertical entropy gradient \citep{goldreich67,urpin98,urpin03}, but
-this conclusion was reached in these studies by considering
+vertical entropy gradient \citep{goldreich67,urpin98,urpin03}. This 
+conclusion was reached in these studies by considering
 vertically-localized disturbances. % , although a definitive timescale
@@
 By contrast, our Eq. \ref{iso_vsi_cond} quantifies this requirement
-for vertically global disturbances (i.e. the fundamental mode). We see
-that in a thin disk, a very short thermal timescale
- is required ($\beta_\mathrm{crit}\ll 1$) because
-$\beta_\mathrm{crit}\propto \epsilon$. 
-This suggests that
-finite thermal relaxation rapidly stabilizes the VSI.
+for vertically global disturbances. We see that in a thin disk, a very
+short thermal timescale is required ($\beta_\mathrm{crit}\ll 1$)
+because $\beta_\mathrm{crit}\propto \epsilon$. 
+This suggests that finite thermal relaxation rapidly stabilizes the
+VSI.
+
+% Here we estimate the maximum thermal relaxation timescale
+% $\beta$ that allows the fundamental VSI mode for $\khat\gg1$.
+
+% At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
+% is real. The dispersion relation for the fundamental mode is then
+% \begin{align}\label{relax_disp_fund}
+%   0 = \sum_{l=1}^{6}c_l\hat{\omega}^l \quad \text{for $M=\hat{\nu}=0$}.  
+% \end{align} 
+% Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
+% $\khat\gg1$, we find
+% \begin{align}
+%   0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
+%   \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
+%   &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
+%   0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
+%   - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
+%     &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
+% \label{relax_cond2}
+% \end{align}
+% This is a pair of simultaneous equations for
+% $\hat{\omega}$ and $\beta$. Now, low-frequency axisymmetric waves in
+% the absence of vertical shear are stable and  satisfy
+% $\hat{\omega}\propto \khat^{-1}$ (Appendix \ref
+% {stable_novshear}, Eq. \ref{iso_disp_full}).  We are thus motivated to
+% seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
+% % we wish to seek a criterion independent of k 
+% % \begin{align}
+% %   \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
+% %\end{align}
+% \begin{align}
+%   \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
+%   |\hat{\omega}\khat|=O(1). 
+% \end{align}
+% Then
+% \begin{align}
+%   \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
+% \end{align}
+% from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
+% that
+% \begin{align}\label{beta_crit0}
+%   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
+%         q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
+% \end{align}
+% at marginal stability for perturbations with large $\khat$.
+
+% From our discussion in \S\ref{relax_pert}, which indicate introducing
+% a finite thermal relaxation is stabilizing, we expect
+% Eq. \ref{beta_crit0} is an upper limit. For a thin  
+% disk $\epsilon\ll 1$, we thus infer 
+% \begin{align}\label{iso_vsi_cond}
+%   \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
+%   \beta_\mathrm{crit} 
+% \end{align}
+% is required for the fundamental VSI to operate. That is, when
+% $\beta<\beta_\mathrm{crit}$, perturbations are  effectively isothermal. 
+% We will confirm Eq. \ref{iso_vsi_cond} numerically in
+% \S\ref{bcrit_num_test}.

### 2015-03-12 linear_adia.tex
generalized bcrit (attempt 1)

@@
 higher order modes to be stabilized more rapidly by finite thermal
-relaxation $\beta$. When $\beta>\beta_\mathrm{crit}$, all modes
+relaxation. When $\beta>\beta_\mathrm{crit}$, all modes
 (satisfying the above assumptions) are stabilized. We can interpret
@@
 assumptions by setting $M=0$ in
-Eq. \ref{relax_cond1}---\ref{relax_cond2} from the onset. )
+Eq. \ref{relax_cond1}---\ref{relax_cond2} from the onset.)
 

### 2015-03-13 linear_adia.tex
corrected generalized bcrit

@@
 \begin{align}
- 0 = &M(M+1) + \left[1 +
+ 0 = &M(M+1)\left(1-\epsilon^2q^2\khat^2\right) + \left[1 +
     \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2\notag\\
    &-  \left(\hat{\omega}\khat\right)^4,\label{relax_cond_simp1}\\
-   0= &M(M+1) + \frac{\beta}{\epsilon q} (1-\gamma)(1+2M)^2
+   0= &2M(M+1) + \frac{\beta}{\epsilon q} (1-\gamma)(1+2M)^2
    \left(\hat{\omega}\khat\right) + \left(\hat{\omega}\khat\right)^2. \label{relax_cond_simp2}
@@
 simultaneous equations for $\hat{\omega}$ and the critical thermal
-timescale $\beta$ for marginal stability. If $\beta\ll 1$,
+timescale $\beta$ for marginal stability. If $\beta\ll 1$ and the disk
+is sufficiently thin so that $|\epsilon q \khat |\ll 1$
 then a simple solution to Eq. \ref{relax_cond_simp1} is 
@@
   \beta = \frac{\epsilon|q|}{(\gamma - 1)}
-  \frac{(1+M)^{3/2}}{(1+2M)^2} \leq \beta_\mathrm{crit} 
+  \frac{\sqrt{1+M}}{(1+2M)} \leq \beta_\mathrm{crit} 
 \end{align}
@@
 
-(We remark that the critical thermal timescale $\beta_\mathrm{crit}$
-for the fundamental mode may be obtained more readily with fewer
-assumptions by setting $M=0$ in
+(We remark that the critical thermal timescale for the fundamental
+mode, $\beta_\mathrm{crit}$, may be obtained more readily without 
+assuming $|\epsilon q \khat|\ll 1$ by setting $M=0$ in
 Eq. \ref{relax_cond1}---\ref{relax_cond2} from the onset.)

### 2015-03-13 numerical.tex
corrected generalized bcrit

@@
 \begin{figure}
-  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 3.1cm
-  0cm]{figures/result2d_vel}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
-  0cm]{figures/result2d_mom} 
+%  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 2.5cm
+%  0cm]{figures/result2d_vel}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
+%  0cm]{figures/result2d_mom} 
+  \includegraphics[width=\linewidth]{figures/result2d_iso}
   \caption{Visualization of the fundamental VSI mode in
     Fig. \ref{lowfreq_eigenfunc}. The color scale corresponds to the
-    real pressure perturbation $W$ scaled by its maximum value.
-    Arrows in the left panel correspond to the meridional velocity
-    field $(\delta v_x,\delta v_z)$, and arrows in the right panel
-    correspond to to the meridional momentum $\rho(\delta v_x,\delta
-    v_z)$. 
+    real pressure perturbation $\delta P$ scaled by its maximum value.
+    Arrows correspond to the vector field 
+    field $\sqrt{\rho}(\delta v_x,\delta v_z)$. Note that the horizontal scale has been scretched for clarity. 
     \label{lowfreq_eigenfunc_2d}
@@
 \begin{figure}
-  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 3.1cm
-  0cm]{figures/result2d_vel_cool}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
-  0cm]{figures/result2d_mom_cool} 
+  % \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 2.5cm
+  % 0cm]{figures/result2d_vel_cool}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
+  % 0cm]{figures/result2d_mom_cool} 
+  \includegraphics[width=\linewidth]{figures/result2d_cool}
   \caption{Same as  Fig. \ref{lowfreq_eigenfunc_2d} but with a thermal

### 2015-03-13 appendix.tex
generalized bcrit (attempt 1)
corrected generalized bcrit
updating

--- 2015-03-12 appendix.tex
@@
 \section{Applicability of the vertically global shearing box}\label{global_corr}
-In \S\ref{analytical} we used the vertically global shearing box
+In \S\ref{analytical} we effectively employed the vertically global shearing box
 \citep[VGSB,][]{mcnally14} to make analytic progress.   
@@
 (Eq. \ref{lin_mass}---\ref{lin_energy}) adopted in our numerical study
-is the absence of  terms associated with the radial disk structure in
+is the absence of terms explicitly associated with the radial disk structure in
 the former framework. These terms appear in the energy, mass and radial momentum
-equations and are reproduced here for convenience,
+equations and are labeled by $\hat{g}_c$ in
+Eq. \ref{lin_mass}---\ref{lin_energy}. Nominally $\hat{g}_c=1$.
 
-\begin{align}
-  \ii \sigma W  &=c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
-      z}\right|_0 \delta v_z + \frac{\gamma}{\Gamma} \frac{d\delta
-    v_z}{dz}\right) + \left[\ii k_xc_s^2(r_0,z)
-  \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
-      r}\right|_0\right]\delta v_x  +
-\frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
-\ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_x + \hat{g}_c
-  \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
-      z}\right|_0 \delta v_z + \frac{d\delta
-    v_z}{dz}\right),\label{gcorr_terms2} \\
-\ii \sigma \delta v_x & = \ii k_x W  -
-\hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
-  r}\right|_0Q - 2\Omega(r_0,z)\delta v_y,\label{gcorr_terms3}
-\end{align}
-where subscript $0$ denotes evaluation at $r=r_0$. We recall terms
-proportional to $\hat{g}_c$ are associated with the background density
-and pressure radial structures, and nominally $\hat{g}_c=1$.
+% \begin{align}
+%   \ii \sigma W  &=c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
+%       z}\right|_0 \delta v_z + \frac{\gamma}{\Gamma} \frac{d\delta
+%     v_z}{dz}\right) + \left[\ii k_xc_s^2(r_0,z)
+%   \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
+%       r}\right|_0\right]\delta v_x  +
+% \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
+% \ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_x + \hat{g}_c
+%   \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
+%       z}\right|_0 \delta v_z + \frac{d\delta
+%     v_z}{dz}\right),\label{gcorr_terms2} \\
+% \ii \sigma \delta v_x & = \ii k_x W  -
+% \hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
+%   r}\right|_0Q - 2\Omega(r_0,z)\delta v_y,\label{gcorr_terms3}
+% \end{align}
+% where subscript $0$ denotes evaluation at $r=r_0$. We recall terms
+% proportional to $\hat{g}_c$ are associated with the background density
+% and pressure radial structures, and nominally $\hat{g}_c=1$.
 
@@
 disks and setting  $\beta\to\infty$ in 
-Eq. \ref{vertiso_gov_nondim}, which can then be written as
-
+Eq. \ref{vertiso_gov_nondim} (which itself is obtained by setting
+$\hat{g}_c=0$ in Eq. \ref{lin_mass}---\ref{lin_energy}) to give 
 \begin{align}
@@
 \end{align}
-When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
-$\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
-expected, a sub-adiabatically stratified disk is stable in the absence
-of vertical shear.
+% When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
+% $\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
+% expected, a sub-adiabatically stratified disk is stable in the absence
+% of vertical shear.
 
-However, in the presence of vertical shear $q\neq0$,
-Eq. \ref{adia_integral} suggests $\hat{\sigma}$ is complex for real
-$\khat$, which implies the possibility of instability, regardless of
-how much $\gamma$ deviates from unity. This is contradicts the second 
-Solberg-Hoiland criterion (Eq. \ref{solberg2}), which states that 
-instability in a thin, adiabatic disk requires it to be close to 
-neutrally stratified, i.e. $|\gamma-1|\ll1$ for vertically isothermal disks
-evolved adiabatically. 
+In the presence of vertical shear $q\neq0$, Eq. \ref{adia_integral}
+suggests $\hat{\sigma}$ is complex for real $\khat$, which implies the
+possibility of instability, regardless of the value of
+$|\gamma-1|$. This is contradicts the second Solberg-Hoiland criterion 
+(Eq. \ref{solberg2}), which imply instability requires 
+$|\gamma-1|\ll1$ for thin, vertically isothermal disks evolved adiabatically. 
 %also need real freq^2 in adia limit 
@@
 In  Fig. \ref{gcorr_compare}, we plot the effect of the global radial gradient terms
-proportional to $\hat{g}_c$ in 
-Eq. \ref{gcorr_terms1}---\ref{gcorr_terms3} by calculating the
+proportional to $\hat{g}_c$ in Eq. \ref{lin_mass}---\ref{lin_energy}
+by calculating the
 fundamental VSI growth rates using three approaches. We compute growth rates from the dispersion
-relation Eq. \ref{relax_disp}, which utilizes the VGSB;  from the
-full linear problem (Eq. \ref{lin_mass}---\ref{lin_energy}) with
-$\hat{g}_c=0$ (i.e. the linear VGSB equations); and from the full
-linear problem with $\hat{g}_c=1$.
+relation Eq. \ref{relax_disp}, which assumes $\hat{g}_c=0$;  from
+Eq. \ref{lin_mass}---\ref{lin_energy} with $\hat{g}_c=0$; and from
+Eq. \ref{lin_mass}---\ref{lin_energy}  with $\hat{g}_c=1$.
 
@@
 caveat, Fig. \ref{gcorr_compare} shows that provided we consider
-$\beta\ll1$, then the VGSB framework is an adequate approximation.
+$\beta\ll1$, then the VGSB framework is an adequate approximation to
+study the VSI.
 
@@
 \end{align}
+
+\subsection{Limiting form with real frequencies and $\khat^2\gg 1$}\label{disp_neut_limit}
+Assuming neutral modes with real $\hat{\sigma}=\hat{\omega}$ and $\khat^2\gg 1$, the real and imaginary parts of
+Eq. \ref{relax_disp} are approximately 
+\begin{align}
+  0 =&M(M+1)(1-\epsilon^2 q^2 \khat^2) + 4\beta\epsilon q M(M+1)\left(\hat{\omega}\khat\right) 
+ +\left[1 +
+    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
+ + 2\epsilon q \gamma\beta \left(\hat{\omega}\khat\right)^3 -  \left(\hat{\omega}\khat\right)^4\notag \\
+  &+ \beta^2\gamma^2\hat{\omega}^2
+  \left(\hat{\omega}\khat\right)^4,\label{relax_cond1} \\
+   0=& 2\epsilon q M (M+1)\khat +
+   \beta(1-\gamma)(1+2M)^2\khat\left(\hat{\omega}\khat\right) 
+   + \epsilon q \khat \left(\hat{\omega}\khat\right)^2 -
+   2\beta\hat{\omega}\left(\hat{\omega}\khat\right)^2
+   - \epsilon q
+   \gamma^2\beta^2\hat{\omega}\left(\hat{\omega}\khat\right)^3 
+   +
+   2\beta\gamma\hat{\omega}\left(\hat{\omega}\khat\right)^4 \label{relax_cond2}. 
+\end{align}
+These equations are used to derive a critical thermal timescale in
+\S\ref{iso_vsi_beta_crit}. For fixed $M$ of $O(1)$,  we assume
+$|\hat{\omega}\khat|$ remain constant but consider $\khat\gg 1$ so
+that $|\hat{\omega}|\ll 1$.  Then for a thin disk $\epsilon \ll 1$
+we neglect the second, fourth and last term in Eq. \ref{relax_cond1};
+and neglect the last three terms in Eq. \ref{relax_cond2}. This
+procedure gives Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2},
+but is not valid if $\beta$ or $|q|$ is large.
+
+For the fundamental mode $M=0$ and Eq. \ref{relax_cond1} ---
+\ref{relax_cond2} simply significantly. It is then straight
+forward to eliminate $\hat{\omega}\khat$ between 
+Eq. \ref{relax_cond_simp1} and \ref{relax_cond_simp2} to obtain
+$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}) without the restrictive
+assumption of $|\epsilon q \khat|\ll1$, as used in the main text to obtain the
+equivalent expression for $M>0$ (Eq. \ref{bcrit_gen}).
+
+%  In this case the fourth term
+% in Eq. \ref{relax_cond1} may be retained but other aforementioned terms
+% are still neglected. One then obtains an expression for $\beta$ by
+% eliminating $\hat{\omega}\khat$ between Eq. \ref{relax_cond1} ---
+% \ref{relax_cond2}. The expression for the critical thermal timescale
+% for the fundamental mode, $\beta_\mathrm{crit}$
+% (Eq. \ref{iso_vsi_con}, is then obtained by taking the thin disk
+% limit.
+
+%Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
+% $\khat\gg1$, we find
+% \begin{align}
+%   0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
+%   \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
+%   &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
+%   0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
+%   - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
+%     &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
+% \label{relax_cond2}
+% \end{align}
+% This is a pair of simultaneous equations for
+% $\hat{\omega}$ and $\beta$. Now, low-frequency axisymmetric waves in
+% the absence of vertical shear are stable and  satisfy
+% $\hat{\omega}\propto \khat^{-1}$ (Appendix \ref
+% {stable_novshear}, Eq. \ref{iso_disp_full}).  We are thus motivated to
+% seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
+% % we wish to seek a criterion independent of k 
+% % \begin{align}
+% %   \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
+% %\end{align}
+% \begin{align}
+%   \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
+%   |\hat{\omega}\khat|=O(1). 
+% \end{align}
+% Then
+% \begin{align}
+%   \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
+% \end{align}
+% from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
+% that
+% \begin{align}\label{beta_crit0}
+%   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
+%         q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
+% \end{align}
 
@@
 this case the solutions are Hermite polynomials, $W\propto
-\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma}$ is real so the disk
-is stable, consistent with \S\ref{integral_relation}. The
-low-frequency solution to Eq. \ref{iso_disp_full} correspond to
-inertial waves \citep{balbus03}.
+\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma} = \hat{\omega}$ is real and the disk
+is stable. The low-frequency branch of Eq. \ref{iso_disp_full} are 
+inertial waves \citep{balbus03}. For 
+$|\hat{\omega}|\ll 1$ and $L\geq 1$ the dispersion relation is
+$\hat{\omega}^2\khat^2 = L$, or $\hat{\omega}\propto \khat^{-1}$ for
+fixed $L$.  This inverse relation has been qualitatively
+observed in numerical simulations of \cite{stoll14}.
 
 \subsubsection{Instability with vertical shear}
-A simple solution to Eq. \ref{simple_growth} may be obtained by
-considering small radial lengthscales $\khat\gg 1$. Then for the
-low-frequency branch of Eq. \ref{simple_growth} 
+A simple solution for $\hat{\sigma}$ may be obtained by balancing
+the  last two terms of Eq. \ref{simple_growth} to obtain the
+low-frequency branch,  
 \begin{align}
@@
 \end{align}
-The eigenfrequency $\hat{\sigma}$ is therefore complex for 
-$L\geq1$, which signifies instability since we can choose
-the sign of the square root to make $\hat{\nu} = \imag\hat{\sigma}>0$.
+This signifies instability for $L\geq1$ 
+since we can choose the sign of the square root to make $\hat{\nu} =
+\imag\hat{\sigma}>0$.
 

### 2015-03-13 linear_adia.tex
updating

--- 2015-03-12 linear_adia.tex
@@
 v_z=$constant and $\chi=1$ so that $B=C=0$. 
-This is the fundamental or `corrugation' mode and represents
-the entire disk column moving vertically. These modes have been
+This is the fundamental or `corrugation' mode where the vertical
+velocity is constant throughout the disk column. These modes have been
 observed to dominate numerical simulations \citep{nelson13,stoll14},  
-and are the ones we find to develop in protoplanetary disks
+and are the ones we find to dominate in numerical calculations with 
+increasing $\beta$, as well as in protoplanetary disks 
 with a realistic estimate for thermal timescales
-(\S\ref{application}).
+(\S\ref{application}).  %Their importance is highlighted
 
-%To determine the (assumed) small effect of introducing a short thermal
-%relaxation, 
 We linearize Eq. \ref{nearly_iso_explicit} about the above
@@
   0 &= 2a + 2\hat{\sigma}\left(1+\khat^2\right)\delta\hat{\sigma} +
-  \left(\hat{\sigma}^2 -1\right)\delta\chi,\notag\\
+  \left(\hat{\sigma}^2 -1\right)\delta\chi, \\
   0&= 2a\left(1+\ii\epsilon q \khat\right) - \left(\khat^2 - \ii\epsilon
@@
 Using the expressions for $\delta\chi$ and $\hat{\sigma}$ we can solve
-for $\delta\hat{\sigma}$. This gives 
-\begin{align}\label{nearly_iso_dsig}
-  \delta \hat{\sigma} =
-  -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii\epsilon
-      q - \hat{k}\right)^2}{2\left(1+\ii\epsilon q \hat{k}\right)\left(1+\hat{k}^2\right)^2}\delta\beta.
-\end{align}
-The imaginary part of Eq. \ref{nearly_iso_dsig} is
+for $\delta\hat{\sigma}$. We find imaginary part of $\delta\hat{\sigma}$
+is 
+% This gives 
+% \begin{align}\label{nearly_iso_dsig}
+%   \delta \hat{\sigma} =
+%   -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii\epsilon
+%       q - \hat{k}\right)^2}{2\left(1+\ii\epsilon q \hat{k}\right)\left(1+\hat{k}^2\right)^2}\delta\beta.
+% \end{align}
+% The imaginary part of Eq. \ref{nearly_iso_dsig} is
 \begin{align}
@@
 \end{align}
-For ease of analysis, we square Eq. \ref{adia_disp0} to obtain a 
-a polynomial in $\hat{\sigma}$,  
+For ease of analysis, we re-arrange Eq. \ref{adia_disp0} and square it
+to obtain a polynomial in $\hat{\sigma}$,  
 \begin{align}
@@
 where the coefficients $c_l$ are given in Appendix \ref{relax_coeff}.
-The dispersion relation $\hat{\sigma}=\hat{\sigma}(\khat)$ generally
-requires a numerical solution. However, simple solutions may be
-obtained in limiting cases which we consider below.
+The dispersion relation $\hat{\sigma}=\hat{\sigma}(\khat)$ is
+complicated and generally requires a numerical solution. However,
+simple solutions may be obtained in limiting cases which we consider
+below.
 
@@
 Here we estimate the maximum thermal relaxation timescale 
-$\beta$ that allows the VSI. We consider wavenumbers such that
-$\khat^2\gg M+1$ which, as we will find, correspond to low frequency
-modes.
-
-%At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$ is real.  
-%This is a pair of simultaneous
-%equations for $\hat{\omega}$ and $\beta$. To proceed, we
-%consider further limits.
-
+$\beta$ that allows the VSI for modes with $\khat^2\gg M+1$.  
+% which, as we will find, correspond to low frequency
+% modes. 
 Our starting point is Eq. \ref{relax_cond1}---\ref{relax_cond2} which
-approximates the dispersion relation Eq. \ref{relax_disp} at marginal
-stability with $\khat\gg 1$. We consider modes with $M=0$ (the
+approximates the dispersion relation Eq. \ref{relax_disp} assuming marginal
+stability and $\khat^2\gg 1$. We consider modes with $M=0$ (the
 fundamental mode) or  $M=O(1)$ and low frequency $|\hat{\omega}|\ll
-1$ such that $|\hat{\omega}\khat|$ is $O(1)$ because $\khat\gg 1$. We
-also assume $|q|$ and $\beta$ are $O(1)$ or less. For a thin disk
-($\epsilon \ll 1$), we have approximately
+1$ such that $|\hat{\omega}\khat|$ is $O(1)$. We
+also assume $|q|$ and $\beta$ are $O(1)$ or smaller. For a thin disk
+($\epsilon \ll 1$) we have 
 \begin{align}
- 0 = &M(M+1) + \left[1 +
-    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2\notag\\
-   &-  \left(\hat{\omega}\khat\right)^4,\label{relax_cond_simp1}\\
-   0= &M(M+1) + \frac{\beta}{\epsilon q} (1-\gamma)(1+2M)^2
+ 0 = &M(M+1)\left(1-\epsilon^2q^2\khat^2\right) \notag\\ 
+ &+ \left[1 +
+    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
+   -  \left(\hat{\omega}\khat\right)^4,\label{relax_cond_simp1}\\
+   0= &2M(M+1) + \frac{\beta}{\epsilon q} (1-\gamma)(1+2M)^2
    \left(\hat{\omega}\khat\right) + \left(\hat{\omega}\khat\right)^2. \label{relax_cond_simp2}
@@
 Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} is a pair of
-simultaneous equations for $\hat{\omega}$ and the critical thermal
-timescale $\beta$ for marginal stability. If $\beta\ll 1$,
-then a simple solution to Eq. \ref{relax_cond_simp1} is 
+simultaneous equations for $\hat{\omega}$ and $\beta$ for marginal
+stability. If $\beta\ll 1$ and the disk is sufficiently thin so that
+$|\epsilon q \khat |\ll 1$ 
+then an approximate solution to Eq. \ref{relax_cond_simp1} is 
 \begin{align}
@@
 is the critical thermal timescale for the fundamental mode
-($M=0$). However, according to Eq. \ref{bcrit_gen}  the critical
-thermal timescale decreases with increasing  $M$. We thus expect
-higher order modes to be stabilized more rapidly by finite thermal
-relaxation. When $\beta>\beta_\mathrm{crit}$, all modes
-(satisfying the above assumptions) are stabilized. We can interpret
-$\beta_\mathrm{crit}$ as the thermal timescale below which
-perturbations are effectively isothermal to allow the VSI.  We will confirm
-Eq. \ref{iso_vsi_cond} numerically in \S\ref{bcrit_num_test}.
+($M=0$). We comment that the expression for $\beta_\mathrm{crit}$ may
+in fact be obtained without assuming $|\epsilon q \khat|\ll 1$.  We
+will confirm Eq. \ref{iso_vsi_cond} numerically in \S\ref{bcrit_num_test}.
 
-(We remark that the critical thermal timescale $\beta_\mathrm{crit}$
-for the fundamental mode may be obtained more readily with fewer
-assumptions by setting $M=0$ in
-Eq. \ref{relax_cond1}---\ref{relax_cond2} from the onset.)
+Eq. \ref{bcrit_gen} suggests that as the thermal timescale is
+increased from zero, higher order modes with $M>0$
+are stabilized before the fundamental mode. We indeed observe this
+trend in numerical calculations, despite adopting somewhat restrictive
+assumptions in order to arrive at the simple expression 
+Eq. \ref{bcrit_gen} for $M>0$.
 
 Rapid thermal relaxation is required to overcome a stabilizing
-vertical entropy gradient \citep{goldreich67,urpin98,urpin03}. This 
-conclusion was reached in these studies by considering
-vertically-localized disturbances. % , although a definitive timescale
-                                % was not presented.  
-By contrast, our Eq. \ref{iso_vsi_cond} quantifies this requirement
-for vertically global disturbances. We see that in a thin disk, a very
-short thermal timescale is required ($\beta_\mathrm{crit}\ll 1$)
-because $\beta_\mathrm{crit}\propto \epsilon$. 
-This suggests that finite thermal relaxation rapidly stabilizes the
-VSI.
+vertical entropy gradient \citep{goldreich67,urpin98,urpin03}, but these
+studies consider vertically localized disturbances. Our 
+Eq. \ref{iso_vsi_cond} quantifies this requirement for vertically
+global disturbances (which have small $M$).  We can interpret
+$\beta_\mathrm{crit}$ as the thermal timescale below which
+perturbations are effectively isothermal and the VSI operates.  We see
+that in a thin disk, a very short thermal timescale is required
+($\beta_\mathrm{crit}\ll 1$) because $\beta_\mathrm{crit}\propto \epsilon$.
 

### 2015-03-13 linear_iso.tex
restructured
appendix.tex
updating

@@
 number of approximations below that allows a simple governing equation
-to be derived and compared with previous works.
+to be derived and compared with previous works.  However, all of the 
+approximations described in
+\S\ref{vgsb_approx}---\S\ref{analytic_relax} are relaxed in our
+numerical study.
 
 \subsection{Relation to the vertically global
-  shearing box} 
+  shearing box}\label{vgsb_approx} 
 We connect the linearized equations 
@@
 is an extension of the standard shearing box \citep{goldreich65} to
-allow the shear flow in the shearing box to be height-dependent. This
-is an attempt to capture the effect of vertical shear in a local model
-of accretoin disks. The difference between the radially-local
-approximation and the VGSB is that the radial disk structure is
-ignored in the latter. That is, the  equilibrium density and pressure
-field in the VGSB is homogeneous in the radial ($x$) direction, as in
-the standard shearing box.
+allow the shear flow in the shearing box to be height-dependent. 
+The difference between the radially local
+approximation and the VGSB is that the equilibrium disk
+structure is assumed to be homogeneous in the radial ($x$), as in the
+standard shearing box.
 
@@
 
-% but modified by the vertical shear term
-% $r\p_z\Omega\delta \delta v_z$.
-
 After setting $\hat{g}_c=0$ in Eq. \ref{lin_mass}, \ref{lin_vx} and
-\ref{lin_energy}, we can eliminate the perturbed velocities between
-the linearized equations to obtain a pair or ordinary differential
+\ref{lin_energy}, we can eliminate variables % eliminate the perturbed velocities between
+% the linearized equations 
+to obtain a pair or ordinary differential
 equations for $W,Q$ as
@@
 
-The VGSB is a self-consistent framework but extrapolating results 
-obtained from the VGSB to the global disk has limitations. This is
-because within the VGSB framework, $d\Omega/dz$ can be freely
-specified. In effect, by adopting the VGSB framework, we are studying
-dynamics in the standard shearing box subject to an additional
-azimuthal force $\delta v_zr d \Omega/dz$ with $d\Omega/dz$ as an
-input parameter. For consistency with the global disk, we choose
-$d\Omega/dz$ in the VGSB to be that due to the thermal structure in
-the global disk (Eq. \ref{vertical_shear}). However, 
-% This is different to the global problem because in
-% that case, the vertical shear is a result of the 
-the background radial disk structure is ignored elsewhere in the VGSB
-(which is done by setting $\hat{g}_c=0$).
+Extrapolating results obtained from the VGSB to the global disk has
+limitations. This is because within the VGSB framework, $d\Omega/dz$
+can be freely specified. %  In effect, by adopting the VGSB framework, we are studying
+% dynamics in the standard shearing box subject to an additional
+% azimuthal force $\delta v_zr d \Omega/dz$ with $d\Omega/dz$ as an
+% input parameter. 
+%For consistency with the global disk, 
+We choose $d\Omega/dz$ to be that due to the thermal structure in 
+the global disk (Eq. \ref{vertical_shear}). However, background radial
+gradient terms are ignored elsewhere in the VGSB (which is done by
+setting $\hat{g}_c=0$).
 
-We further discuss caveats of the VGSB in Appendix \ref{global_corr}.
-We find that for perturbations with small radial lengthscales
-($k_xH\gg 1$) and fast thermal relaxation ($\beta\ll1$), which
-characterizes the VSI, the VGSB framework is adequate for the purpose
-of a simple analytical discussion.
+%the above inconsistency
+
+We further discuss caveats of the VGSB in Appendix
+\ref{global_corr}. We find that for perturbations with small radial
+lengthscales ($k_xH\gg 1$) and fast thermal relaxation ($\beta\ll1$),
+which characterizes the VSI, the VGSB framework is adequate for the
+purpose of a simple analytical treatment of the VSI.
 
@@
 making use of Eq. \ref{vertical_shear}, to
-obtain an equation for $\delta v_z$ for $\Gamma=1$,
+obtain an equation for $\delta v_z$. The result for $\Gamma=1$ is
 \begin{align}
@@
 \begin{align}
-  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\sigma} = \sigma/\Omega_k,
+  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\sigma} =\hat{\omega} +
+  \ii\hat{\nu}= \sigma/\Omega_k,
 \end{align}
-then Eq. \ref{vertiso_gov} becomes
+where $\hat{\omega}$ and $\hat{\nu}$ are real. 
+Eq. \ref{vertiso_gov} becomes 
 \begin{align}
@@
 \end{align}
-where $^\prime$ denotes $d/d\zhat$. Eq. \ref{vertiso_gov_nondim} will
-be the basis for our discussion of vertically isothermal disks.
+where $^\prime$ denotes $d/d\zhat$.%  Eq. \ref{vertiso_gov_nondim} will
+% be the basis for our discussion of vertically isothermal disks.
 
@@
 $\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
-unnecessary complexity for the present discussion. However, this
-approximation will be relaxed in our numerical study.
+unnecessary complexity for the present discussion. % However, this
+% approximation will be relaxed in our numerical study.
 
-Note that we can generalize Eq. \ref{vertiso_gov_nondim} non-Keplerian
-rotation profiles by redefining $\khat$ as $k_xc_s/\kappa$,
-$\hat{\sigma}$ as $\sigma/\kappa$, and  $\epsilon$ as $c_s/r\kappa$.
+% Note that we can generalize Eq. \ref{vertiso_gov_nondim} non-Keplerian
+% rotation profiles by redefining $\khat$ as $k_xc_s/\kappa$,
+% $\hat{\sigma}$ as $\sigma/\kappa$, and  $\epsilon$ as $c_s/r\kappa$.
 

### 2015-03-13 numerical.tex
updating

@@
 \end{align}
-where $\bm{\xi}$ is the Lagrangian displacement, the meridional
-components of which are given by $\xi_{x,z} = \ii\delta
-v_{x,z}/\sigma$.
+where $\bm{\xi}$ is the Lagrangian displacement with meridional 
+components $\xi_{x,z} = \ii\delta v_{x,z}/\sigma$.
 
@@
 Fig. \ref{iso_eigen_kx} compares the numerically obtained eigenvalues 
-for the fundamental mode and that calculated analytically from our
-analysis of isothermal perturbations in Appendix \ref{iso_discuss}
-(Eq. \ref{simple_growth} with $L=1$). We find good agreement in the
+for the fundamental mode and that calculated from 
+Eq. \ref{simple_growth} (with $L=1$). We find good agreement in the
 wave frequency $\omega$ at all $\khat$ and in growth rates $\nu$ for
-$\khat\gtrsim 10$. Growth rates are $O(\epsilon\Omega_k)$, 
-consistent with our discussion in Appendix \ref{max_growth1}, but
-are over-estimated by our analysis. The match in $\nu$ worsens for
+$\khat\gtrsim 10$. Growth rates are over-estimated by our analysis but 
+are $O(\epsilon\Omega_k)$, which is consistent with the discussion in
+Appendix \ref{max_growth1}. The match in $\nu$ worsens for
 decreasing $\khat$ due to the neglect of global radial structure in
@@
   \caption{Growth rate (top) and real frequency (bottom) of the
-    fundamental VSI mode in a disk with $(\gamma,
-    \Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
-    $\beta=10^{-3}$. The solid (dotted) line is obtained numerically
+    fundamental VSI mode in the fiducial disk model with $\beta =
+    10^{-3}$. 
+    % a disk with $(\gamma,
+    % \Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+    % $\beta=10^{-3}$. 
+    The solid (dotted) line is obtained numerically
     (analytically).  
@@
 show a meridional visualization of this mode in
-Fig. \ref{lowfreq_eigenfunc_2d}. Radial velocities are typically much
-smaller than vertical velocities. Although there is a constant vertical
-velocity throughout the fluid column, most of the momentum is
+Fig. \ref{lowfreq_eigenfunc_2d}. The $x$ axis is stretched for
+clarity; radial velocities are in fact typically much
+smaller than vertical velocities. Most of the meridional kinetic energy is
 contained within $\sim 2H$ of the midplane due to the density 
@@
     Fig. \ref{lowfreq_eigenfunc}. The color scale corresponds to the
-    real pressure perturbation $\delta P$ scaled by its maximum value.
-    Arrows correspond to the vector field 
-    field $\sqrt{\rho}(\delta v_x,\delta v_z)$. Note that the horizontal scale has been scretched for clarity. 
+    pressure perturbation $\delta P=\rho W$ scaled by its maximum value.
+    Arrows correspond to the vector field field $\sqrt{\rho}(\delta
+    v_x,\delta v_z)$. Note that the horizontal axis has been scretched
+    for clarity.  
     \label{lowfreq_eigenfunc_2d}
@@
 In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of unstable
-modes found for our standard setup with $\khat=10$. This case is shown
-by red crosses. We also plot several other calculations with different
-vertical boundary conditions: free surface at $\zmax = 3H$ (black
-dots), free surface at $\zmax=7H$ (green triangles) and a rigid
-boundary at $\zmax=5H$ (blue cross).
+modes with $\khat=10$. The result for our fiducial setup is shown
+in black dots. We also plot other calculations with different
+vertical boundary conditions: rigid boundary and $\zmax=5H$ (red
+cross),  free surface and $\zmax = 3H$ (blue triangles), free surface
+and $\zmax=7H$ (green cross).
 
 The fundamental mode, marked by `f', corresponds to the eigenvalue
-with smallest magnitude $|\sigma|$. We see that the fundamental mode,
-while not the most unstable in this case, is more robust against
-vertical boundary conditions. Higher order modes with larger growth
-rates (i.e. the branch where $\nu$ increases with $|\omega|$) require
-larger vertical domains.
+with the smallest $|\sigma|$. We see that the fundamental mode is
+robust against vertical boundary conditions as it appears at the same
+position in all cases. Higher order modes with larger growth rates
+(i.e. the branch where $\nu$ increases with $|\omega|$) require larger
+vertical domains.  The `tail' end with decreasing $\nu$ to the right
+of the plot differ significantly between the cases, and are thus
+likely associated with boundary conditions.
 
@@
     $\beta=10^{-3}$. The perturbation wavenumber is $\khat=10$. 
-    The standard setup with a free surface ($\Delta P=0$) and 
-    vertical domain size $\zmax=5H$ is shown by the red cross. Other
+    The fiducial setup with a free surface ($\Delta P=0$) and 
+    vertical domain size $\zmax=5H$ is shown by black dots Other
     cases correspond to different boundary condtions applied at
@@
 \subsubsection{Comment on `surface modes'} 
-Our standard setup with $\khat=10$ in 
-Fig. \ref{compare_modes_iso_kx10} (red crosses) differs to that from
-\cite{nelson13} and \cite{mcnally14} in the lack of `surface modes' in
-our case. These modes lie parallel to the vertical axis in mode
+Our fiducial setup with $\khat=10$ in 
+Fig. \ref{compare_modes_iso_kx10} (black dots) differs to that from
+\cite{nelson13} and \cite{mcnally14} in the lack of `surface modes'. 
+These modes lie almost parallel to the vertical axis in mode 
 diagrams such as Fig. \ref{compare_modes_iso_kx10}. We find them for 
@@
 Fig. \ref{compare_modes_iso_kx30} with different boundary 
-conditions. Surface modes can be seen with
-$\zmax\geq 5H$, examples of which are marked by `s'. These modes have
-been discussed in the aforementioned studies. They typically have the
-largest growth rates but require sufficiently large vertical domain
-size and their eigenvalues vary with $\zmax$.
+conditions. Surface modes are readily identified for 
+$\zmax\geq 5H$ and are marked by `s'. These modes have
+already been discussed in the aforementioned studies. They typically
+have the largest growth rates but require sufficiently large vertical
+domain size and their eigenvalues vary with $\zmax$.
 
@@
 Fig. \ref{compare_modes_cool_kx10} show unstable modes with $\khat=10$
-and $\beta\in[0.01,0.1]$. In addition to our standard setup with 
+and $\beta\in[0.01,0.1]$. In addition to our fiducial setup with 
 $\zmax=5H$, we also plot results for $\zmax=3H$ and $\zmax=7H$. In all

### 2015-03-13 numerical.tex
updating

@@
 fundamental mode, are more effectively stabilized with increasing 
-$\beta$. For $\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we
+$\beta$.
+
+For $\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we
 only find the fundamental mode.
@@
 Fig. \ref{lowfreq_eigenfunc_2d_cool}
-(cf. Fig. \ref{lowfreq_eigenfunc_2d}). However, most of the meridional
-momentum is still contained within $|z|\lesssim 2H$ as for 
-the case with $\beta=10^{-3}$.
+(cf. Fig. \ref{lowfreq_eigenfunc_2d}). % However, most of the meridional
+% momentum is still contained within $|z|\lesssim 2H$ as for 
+% the case with $\beta=10^{-3}$.
 
@@
 
-Fig. \ref{bcrit_compare1} shows that $\beta_\mathrm{crit}$ provides a  
-practical indication for the maximum thermal relaxation timescale to
-permit the fundamental VSI. % From our discussion in 
-% \S\ref{therm_relax_eff}, which indicate the fundamental mode typically
-% dominant 
-\cite{nelson13} found a thermal relaxation timescale of $\beta\gtrsim
-0.6$ stabilized the VSI in non-linear hydrodynamic simulations. This 
-is consistent with our results.
-
 It is interesting to note that there is a thermal relaxation timescale that
@@
 $\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$
-separately. Here, we use a wavenumber $\khat=10$, for which we
-find growth rates reach zero for the range of disk parameters
-considered, so that a numerical $\beta_\mathrm{crit}$ can be defined precisely.
+separately.%  Here, we use a wavenumber $\khat=10$, for which we
+% find growth rates reach zero for the range of disk parameters
+% considered, so that a numerical $\beta_\mathrm{crit}$ can be defined precisely.
 
@@
 so that the VSI operates.
+
+Fig. \ref{bcrit_compare1} shows that $\beta_\mathrm{crit}$ provides a  
+practical indication for the maximum thermal relaxation timescale to
+permit the fundamental VSI. % From our discussion in 
+% \S\ref{therm_relax_eff}, which indicate the fundamental mode typically
+% dominant 
+\cite{nelson13} found a thermal relaxation timescale of $\beta\gtrsim
+0.6$ stabilized the VSI in non-linear hydrodynamic simulations. This 
+is consistent with our results.
 

### 2015-03-13 appendix.tex
updating

@@
 Eq. \ref{relax_cond_simp1} and \ref{relax_cond_simp2} to obtain
-$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}) without the restrictive
-assumption of $|\epsilon q \khat|\ll1$, as used in the main text to obtain the
+$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}) without
+assuming $|\epsilon q \khat|\lesssim 1$, as done in the main text to obtain the
 equivalent expression for $M>0$ (Eq. \ref{bcrit_gen}).
@@
 \end{align}
-This is just the dispersion relation for axisymmetric waves in a
-vertically isothermal disk already discussed in previous works 
+This is just the dispersion relation for axisymmetric isothermal waves in a
+vertically isothermal disk
 \citep[e.g.][]{okazaki87,takeuchi98,tanaka02,zhang06,ogilvie13,barker14}. In 

### 2015-03-13 linear_adia.tex
updating

@@
 Here we estimate the maximum thermal relaxation timescale 
-$\beta$ that allows the VSI for modes with $\khat^2\gg M+1$.  
+$\beta$ that allows the VSI.  
+%for modes with $\khat^2\gg M+1$.  
 % which, as we will find, correspond to low frequency
@@
 approximates the dispersion relation Eq. \ref{relax_disp} assuming marginal
-stability and $\khat^2\gg 1$. We consider modes with $M=0$ (the
+stability and $\khat^2\gg 1$. We consider $M=0$ (the 
 fundamental mode) or  $M=O(1)$ and low frequency $|\hat{\omega}|\ll
-1$ such that $|\hat{\omega}\khat|$ is $O(1)$. We
+1$ modes such that $|\hat{\omega}\khat|$ is $O(1)$. We 
 also assume $|q|$ and $\beta$ are $O(1)$ or smaller. For a thin disk
@@
 stability. If $\beta\ll 1$ and the disk is sufficiently thin so that  
-$|\epsilon q \khat |\ll 1$ 
+$|\epsilon q \khat |^2\lesssim 1$ (despite $\khat^2\gg 1$), 
 then an approximate solution to Eq. \ref{relax_cond_simp1} is 
 \begin{align}
-  \left(\hat{\omega}\khat\right)^2 \simeq M+1, 
+  \left(\hat{\omega}\khat\right)^2 \simeq (M+1)\left(1 -
+    \frac{M|\epsilon q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
 \end{align}
-which is the dispersion relation for low-frequency waves in vertically
+Note that for $q=0$ Eq. \ref{marg_freq_approx} reduces to 
+the dispersion relation for low-frequency waves in vertically
 isothermal disks without vertical shear (see Appendix
-\ref{stable_novshear}). Then  
+\ref{stable_novshear}). Inserting Eq. \ref{marg_freq_approx} into
+Eq. \ref{relax_cond_simp2}, we obtain 
 \begin{align}\label{bcrit_gen}
-  \beta = \frac{\epsilon|q|}{(\gamma - 1)}
-  \frac{\sqrt{1+M}}{(1+2M)} \leq \beta_\mathrm{crit} 
+  \beta &= \frac{\epsilon|q|}{(\gamma - 1)}
+  \frac{\sqrt{1+M}}{(1+2M)}\left[1 + \frac{M(2M-1)|\epsilon q
+      \khat|^2}{2(2M+1)^2}\right].
+%  &\lesssim \beta_\mathrm{crit}\notag
 \end{align}
-where 
+For modes satisfying the above assumptions (or $M=0$) so that the
+second term in square brackets is small, we have 
 \begin{align}\label{iso_vsi_cond}
-  \beta_\mathrm{crit}  \equiv \frac{\epsilon|q|}{\gamma-1} 
+  \beta\leq \beta_\mathrm{crit}  \equiv
+  \frac{\epsilon|q|}{\gamma-1}, 
 \end{align}
+where $\beta_\mathrm{crit}$
 is the critical thermal timescale for the fundamental mode
 ($M=0$). We comment that the expression for $\beta_\mathrm{crit}$ may
-in fact be obtained without assuming $|\epsilon q \khat|\ll 1$.  We
-will confirm Eq. \ref{iso_vsi_cond} numerically in \S\ref{bcrit_num_test}.
+in fact be obtained without limiting the magnitude of $|\epsilon q   
+\khat|$.  We will confirm Eq. \ref{iso_vsi_cond} numerically in
+\S\ref{bcrit_num_test}.
 
-Eq. \ref{bcrit_gen} suggests that as the thermal timescale is
+Eq. \ref{iso_vsi_cond} suggests that as the thermal timescale is
 increased from zero, higher order modes with $M>0$

### 2015-03-13 linear_iso.tex
updating

@@
 approximation and the VGSB is that the equilibrium disk
-structure is assumed to be homogeneous in the radial ($x$), as in the
+structure is assumed to be homogeneous in the radial ($x$) direction,
+as in the 
 standard shearing box.
@@
 which characterizes the VSI, the VGSB framework is adequate for the 
-purpose of a simple analytical treatment of the VSI.
+purpose of a simple analytical treatment.
 

### 2015-03-13 numerical.tex
updating

@@
 corresponding to Eq. \ref{lin_mass}---\ref{lin_energy}. 
-We relax the thin-disk, low-frequency and nearly-Keplerian
-approximations used in the analytical discussion above, and  
+We relax the  low-frequency, nearly-Keplerian and thin-disk 
+approximations used in the analytical discussion above. We 
 use the full expressions of $\rho$, $\Omega$ and 
-$\kappa$ given in \S\ref{eqm}. We also set $\hat{g}_c=1$ to
+$\kappa$ and $c_s$ given in \S\ref{eqm}. We also set $\hat{g}_c=1$ to  
 account for the background radial disk structure.
@@
 
-Unless otherwise stated, our fiducial disk model is: a nearly 
+Unless otherwise stated, our fiducial disk model is a nearly 
 vertically isothermal disk with $\Gamma=1.011$, stable
@@
 is similar to strictly isothermal perturbations in a vertically
-isothermal disk considered in \cite{nelson13} and \cite{mcnally14}. % We
-% compare results with our analytical discussion of isothermal 
-% perturbations in Appendix \ref{iso_discuss}.
+isothermal disk considered in \cite{nelson13} and \cite{mcnally14}.
 
@@
 the analysis. Nevertheless, given the number of approximations made
-to calculate growth rates analytically, this agreement is
-satisfactory, especially at large $\khat$.
+in our analysis, the agreement is satisfactory, especially at large
+$\khat$.
 
@@
 
-With rapid thermal relaxation, growth rates are expected to diminish for both $\khat\to 0$
-and $\khat\to \infty$. This is because for $\khat\ll 1$ the radial
-wavelenth is large, but such disturbances are 
+With rapid thermal relaxation, growth rates are expected to diminish
+for both $\khat\to 0$ and $\khat\to \infty$. This is because for
+$\khat\ll 1$ the radial wavelenth is large, but such disturbances are 
 stabilized by by epicyclic motions ($|\omega|\sim
@@
 throughout most of the disk, corresponding to a constant vertical
-velocity perturbation, as expected for the fundamental mode according
-to explicit solutions developed in Appendix \ref{iso_explicit}. We
-show a meridional visualization of this mode in
-Fig. \ref{lowfreq_eigenfunc_2d}. The $x$ axis is stretched for
-clarity; radial velocities are in fact typically much
+velocity perturbation, as expected for the fundamental mode (see
+Appendix \ref{iso_explicit}). We show a meridional visualization of
+this mode in Fig. \ref{lowfreq_eigenfunc_2d}. The $x$ axis is
+stretched for clarity; radial velocities are in fact typically much 
 smaller than vertical velocities. Most of the meridional kinetic energy is
@@
   1cm]{figures/eigenvectorvz_iso}
-  \caption{Pressure perturbation $W$ (top) and vertical velocity
-    perturbation $\delta v_z$ of the fundamental VSI for a
-    perturbation  wavenumber $\khat=10$  The real 
+  \caption{Pseudo-enthalpy perturbation $W$ (top) and vertical velocity
+    perturbation $\delta v_z$ of the fundamental VSI with 
+    $\khat=10$ in the fiducial disk with $\beta=10^{-3}$. The real 
     (imaginary) parts of $W$ and $\delta v_z$ are plotted as solid
@@
     Arrows correspond to the vector field field $\sqrt{\rho}(\delta
-    v_x,\delta v_z)$. Note that the horizontal axis has been scretched
+    v_x,\delta v_z)$. Note that the horizontal axis has been stretched 
     for clarity.  
@@
 vertical boundary conditions: rigid boundary and $\zmax=5H$ (red
-cross),  free surface and $\zmax = 3H$ (blue triangles), free surface
-and $\zmax=7H$ (green cross).
+crosses),  free surface and $\zmax = 3H$ (blue triangles), free surface
+and $\zmax=7H$ (green crosses).
 
@@
 (i.e. the branch where $\nu$ increases with $|\omega|$) require larger
-vertical domains.  The `tail' end with decreasing $\nu$ to the right
+vertical domains, since they have increasing number of nodes along 
+$z$. The `tail' end with decreasing $\nu$ to the right
 of the plot differ significantly between the cases, and are thus
@@
   \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10.ps}
-  \caption{Unstable modes for a disk with 
-    $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
-    $\beta=10^{-3}$. The perturbation wavenumber is $\khat=10$. 
-    The fiducial setup with a free surface ($\Delta P=0$) and 
-    vertical domain size $\zmax=5H$ is shown by black dots Other
+  \caption{Unstable modes with $\khat=10$ in the fiducial disk model
+    with $\beta=10^{-3}$. %  a disk with 
+    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+    % $\beta=10^{-3}$. 
+    The fiducial boundary condition is a free surface ($\Delta P=0$)
+    with vertical domain size $\zmax=5H$  (black dots).  Other
     cases correspond to different boundary condtions applied at
@@
 conditions. Surface modes are readily identified for 
-$\zmax\geq 5H$ and are marked by `s'. These modes have
-already been discussed in the aforementioned studies. They typically
+$\zmax\geq 5H$ and are marked by `s'. They typically
 have the largest growth rates but require sufficiently large vertical
@@
 Due to their dependence on boundary conditions, we 
-do not consider surface modes further in detail. In fact, we will see
-in the next section that surface modes are more effectively stabilized
-by finite thermal relaxation than the fundamental mode. %  Thus, surface
+do not consider surface modes further in detail. These modes have
+already been discussed in the aforementioned studies. In fact, we will
+see in the next section that surface modes are more effectively
+stabilized by finite thermal relaxation than the fundamental mode. %  Thus, surface 
 % modes may not be the most relevant in realistic disks.
@@
     \label{compare_modes_iso_kx30}. Examples of surface modes are
-    marked with `s'. These are found for sufficiently large vertical
-    domain sizes. 
+    marked by `s'. 
   }
@@
 \subsection{Effect of thermal relaxation}\label{therm_relax_eff}
-We now consider the effect of thermal relaxation by increasing
+We now consider the effect of finite thermal relaxation by increasing 
 $\beta$. Note that for the standard disk model the critical thermal 
@@
 $\beta$. The fundamental mode is not the most unstable for 
-$\beta=0.01$ as there is a large number of higher order modes, but
-these depend on boundary conditions.
+$\beta=0.01$ as there is a number of higher order modes, but 
+their existence depends on $\zmax$.
 
 However, the fundamental mode is dominant for $\beta \geq 
-0.03$. Higher order modes, with larger growth rates than the
-fundamental mode, are more effectively stabilized with increasing 
-$\beta$.
-
-For $\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we
-only find the fundamental mode.
+0.03$. Higher order modes, with larger growth rates, are more
+effectively stabilized with increasing $\beta$. For $\beta=0.1$, which
+is close to $\beta_\mathrm{crit}$, we only find the fundamental
+mode. This is consistent with the discussion in 
+\S\ref{iso_vsi_beta_crit}.
 
@@
     crosses); for different vertical domain sizes $\zmax=3H$ (top),
-    $\zmax=5H$ (middle, standard setup) and $\zmax=7H$ (bottom). The
-    fundamental mode is marked with `f'. 
+    $\zmax=5H$ (middle, our fiducial setup) and $\zmax=7H$
+    (bottom). The fundamental mode is marked with `f'.  
     \label{compare_modes_cool_kx10} 
@@
 
-Fig. \ref{compare_modes_cool_kx30} show unstable modes with
-$\khat=30$. In this case surface modes appear for sufficiently large   
-$\zmax$. However, they are effectively stabilized with increasing 
-$\beta$. For $\zmax=3H$, the fundamental mode becomes dominant for
-$\beta\geq 0.03$, as before. For $\zmax=5H$ and $\zmax=7H$, we find
-the appearence of modes slightly more unstable than the fundamental
-mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
-panels). However, these modes are clearly associated with the vertical
-domain size as they are absent for $\zmax=3H$. % Thus, for $\zmax=5H$ and 
-% $\zmax=7H$  the fundamental mode is likely dominant for 
-% $\beta\geq0.05$
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_modes_cool_kx30_z3.ps} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_modes_cool_kx30_z5.ps}
-  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx30_z7.ps}
-  \caption{Same as Fig. \ref{compare_modes_cool_kx10} but for
-    $\khat=30$. Examples of surface modes are marked with `s', and 
-    appear for sufficiently large $\zmax$.
-    \label{compare_modes_cool_kx30} 
-  }
-\end{figure}
-
-Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
-show that the fundamental mode is not sensitive to boundary
-conditions. Thus, tracking fundamental mode is a
-more robust way of assessing stability as a function of $\beta$,
-rather than searching for the most 
-unstable mode, because the latter may be associated with boundary
-conditions. In fact, the fundamental mode bcomes dominant 
-(or close to it) with increasing $\beta$.
-
 Fig. \ref{lowfreq_eigenfunc_cool}---\ref{lowfreq_eigenfunc_2d_cool}
 shows the fundamental mode with $\khat=10$ in a disk with 
-$\beta=0.1$. These plots can be compared the case with rapid thermal
-relaxation (Fig. \ref{lowfreq_eigenfunc}---\ref{lowfreq_eigenfunc_2d}). We see 
-that increasing $\beta$ causes the region characterized by $W\sim z$ and  
-$\delta v_z\sim$ constant closer to the mid-plane, with oscillatory
-behavior emerging near vertical boundaries where buoyancy becomes
-important. This leads to the `tilted' pressure perturbations seen in
-Fig. \ref{lowfreq_eigenfunc_2d_cool}
+$\beta=0.1$. Comparing these plots with
+Fig. \ref{lowfreq_eigenfunc}---\ref{lowfreq_eigenfunc_2d}, which have
+$\beta=10^{-3}$, we see that increasing $\beta$ restricts the region 
+characterized by $W\sim z$ and $\delta v_z\sim$ constant closer to the
+mid-plane, with oscillatory behavior emerging near vertical boundaries
+where buoyancy becomes important. This leads to the `tilted' pressure
+perturbations seen in Fig. \ref{lowfreq_eigenfunc_2d_cool} 
 (cf. Fig. \ref{lowfreq_eigenfunc_2d}). % However, most of the meridional
@@
 
+In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable
+modes with $\khat=30$. In this case surface modes appear for
+$\zmax\geq5H$. However, they are effectively stabilized with
+increasing $\beta$. For $\zmax=3H$, the fundamental mode becomes
+dominant for $\beta\geq 0.03$, as observed for $\khat=10$. For
+$\zmax=5H$ and $\zmax=7H$, we find the emergence modes slightly more
+unstable than the fundamental mode even when $\beta=0.1$ (the leftmost
+eigenvalues in the lower two panels). However, these modes are clearly associated
+with the vertical domain size as they are absent for $\zmax=3H$. 
+% Thus, for $\zmax=5H$ and  
+% $\zmax=7H$  the fundamental mode is likely dominant for 
+% $\beta\geq0.05$
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/compare_modes_cool_kx30_z3.ps} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/compare_modes_cool_kx30_z5.ps}
+  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx30_z7.ps}
+  \caption{Same as Fig. \ref{compare_modes_cool_kx10} but for
+    $\khat=30$. Examples of surface modes are marked with `s'. 
+    \label{compare_modes_cool_kx30} 
+  }
+\end{figure}
+
+Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
+show that the fundamental mode is not sensitive to boundary
+conditions. The fundamental mode also becomes dominant as $\beta$
+approaches $\beta_\mathrm{crit}$.  
+Thus, tracking the fundamental mode is a more robust way of 
+assessing stability as $\beta$ is increased from zero, rather than
+searching for the most unstable mode, which are affected by boundary 
+conditions.
+
 \subsection{Critical thermal relaxation
@@
 We now test the critical thermal relaxation timescale given by
-Eq. \ref{iso_vsi_cond}. We follow the fundamental mode by starting
-from eigenvalues for $\beta=10^{-3}$ and slowly increase the thermal
+Eq. \ref{iso_vsi_cond}. We begin from eigenvalues for the fundamental
+mode with $\beta=10^{-3}$ and slowly increase the thermal
 relaxation timescale up to $\beta=10^3$. Fig. \ref{bcrit_compare1}
@@
 an accurate prediction of the upper limit to the thermal relaxation 
-timescale for the fundamental mode. For smaller $\khat$, 
-$\beta<\beta_\mathrm{crit}$ the VSI can operate slightly beyond
-$\beta_\mathrm{crit}$ (recall that $\beta_\mathrm{crit}$ was derived
-by assuming $\khat\gg 1$). For $\khat\geq 50$, we find perturbations
-acquire large amplitudes near the vertical boundaries. Complications
-% to our simple solution procedure 
-may arise from surface modes. Nevertheless, growth rates rapidly
-decrease as $\beta$ approach and increase beyond
-$\beta_\mathrm{crit}$.
+timescale for the fundamental mode. For smaller $\khat$, the VSI can
+operate slightly beyond $\beta_\mathrm{crit}$ (recall that
+$\beta_\mathrm{crit}$ was derived by assuming $\khat^2\gg 1$). For
+$\khat\geq 50$, we find perturbations have large amplitudes near the
+vertical boundaries, and complications to our numerical search of
+eigenvalues arise from emergence of modes associated with boundary
+conditions. Nevertheless, growth rates rapidly decrease as $\beta$
+approaches and increase beyond $\beta_\mathrm{crit}$.
 
-It is interesting to note that there is a thermal relaxation timescale that
-maximizes the mode decay rate for $1\lesssim \khat \lesssim 30$. Here,
-it is $\beta\simeq1$---$2$ or about $0.1$---$0.3$ orbital
-periods. This is in fact consistent with Fig. 12 of \cite{nelson13}
-where the perturbed kinetic energy in their numerical simulations was
-observed to decay most rapidly for a thermal timescale of $\sim 0.1$
-orbits.
-
-%implications for nonlinear sims?
+Fig. \ref{bcrit_compare1} is consistent with numerical simulations
+performed by \cite{nelson13} for which a thermal relaxation timescale
+$\beta\gtrsim 0.6$ stablizes the VSI. This can be anticipated from our
+estimate of the critical thermal timescale $\beta_\mathrm{crit}\simeq
+0.1$ for our fiducial disk parameters (which is the same as that in
+\citeauthor{nelson13}). It is interesting to note that there is a
+thermal relaxation timescale that maximizes the mode decay rate for
+$1\lesssim \khat \lesssim 30$. Here, it is $\beta\simeq1$---$2$ or
+about $0.1$---$0.3$ orbital periods. This is in fact consistent with 
+Fig. 12 of \citeauthor{nelson13} where the perturbed kinetic energy in their
+numerical simulations were observed to decay most rapidly for a
+thermal timescale of $\sim 0.1$ orbits.
 
@@
    \includegraphics[width=\linewidth]{figures/gcorr_compare2} 
-   \caption{Growth rate of the fundamental VSI in the fiducial disk
+   \caption{Growth rate of the fundamental VSI mode in the fiducial disk
      model as a function of the thermal relaxation timescale $\beta$
-     The vertical line
-     is the critical thermal timescale $\beta_\mathrm{crit}$  obtained
+     The vertical line is the critical thermal timescale
+     $\beta_\mathrm{crit}$ obtained 
      from Eq. \ref{iso_vsi_cond}. 
@@
 
-% may need to explain why choose k=10
-% bcrit independent of k for k>>1 
-% for some parameter regimes (e.g. large q, large epsilon) 
-% behavior at larger kx and finite bcool not understood. 
  We further test the dependence of the critical thermal relaxation timescale 
@@
  $\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$
-separately.%  Here, we use a wavenumber $\khat=10$, for which we
+ separately.
+ % Here, we use a wavenumber $\khat=10$, for which we
  % find growth rates reach zero for the range of disk parameters
  % considered, so that a numerical $\beta_\mathrm{crit}$ can be defined precisely. 
-
  The numerically obtained $\beta_\mathrm{crit}$ is shown in
@@
 
-Fig. \ref{bcrit_compare}, together with the previous section
-(Fig. \ref{bcrit_compare1}), shows that $\beta_\mathrm{crit}$ as given
-by Eq. \ref{iso_vsi_cond} characterizes the thermal timescale below
-which perturbations are effectively isothermal 
-so that the VSI operates.
-
-Fig. \ref{bcrit_compare1} shows that $\beta_\mathrm{crit}$ provides a  
-practical indication for the maximum thermal relaxation timescale to
-permit the fundamental VSI. % From our discussion in 
-% \S\ref{therm_relax_eff}, which indicate the fundamental mode typically
-% dominant 
-\cite{nelson13} found a thermal relaxation timescale of $\beta\gtrsim
-0.6$ stabilized the VSI in non-linear hydrodynamic simulations. This 
-is consistent with our results.
-
-% From the previous section (Fig. \ref{relax_eigenW_num}) we
-% expect that for other wavenumbers,  still gives a characteristic
-% thermal timescale below which the perturbations are effectively
-% isothermal.
+ Fig. \ref{bcrit_compare}, together with Fig. \ref{bcrit_compare1} and 
+ our finding that the fundamental mode becomes dominant as one
+ increases $\beta$, shows that $\beta_\mathrm{crit}$ as given 
+ by Eq. \ref{iso_vsi_cond} provides a  
+ practical measure for the thermal relaxation timescale below which 
+ VSI will operate.
 
@@
     $\beta_\mathrm{crit}$ for the fundamental VSI mode on disk
-    parameters. The fiducial parameters are $\Gamma=1.011$,
-    $\gamma=1.4$ and $(p,q,\epsilon)=(-1.5,-1,0.05)$. Top: varying
+    parameters. Our fiducial setup is $(\gamma, \Gamma= (1.4, 1.011)$
+    and $(p,q,\epsilon)=(-1.5,-1,0.05)$. Top: varying 
     $q\in[-1.2,-0.2]$; middle: varying $\gamma\in[1.2,2.0]$; bottom:
@@
 Fig. \ref{gcorr_compare_vnoniso} shows that introducing finite thermal
-relaxation rapidly stabilizes the fundamental mode. This is
-similar to that observed for nearly vertically isothermal disks
-(Fig. \ref{bcrit_compare1}). For $\gamma=1.7,\,2.0,\,2.5$, growth
-rates reach zero at  $\beta\simeq0.17,\,0.083,\,0.045$, respectively.
-Interestingly, these values are equal to
-$\epsilon|q|/(\gamma-\Gamma)$. This suggests that the critical thermal
-relaxation timescale for vertically non-isothermal disks can also be
-estimated by Eq. \ref{iso_vsi_cond} but with $\gamma-1$ replaced by
-$\gamma-\Gamma$.
+relaxation rapidly stabilizes the disk, similar to that observed for
+nearly vertically isothermal disks (Fig. \ref{bcrit_compare1}). For
+$\gamma=1.7,\,2.0,\,2.5$, growth rates reach zero at
+$\beta\simeq0.17,\,0.083,\,0.045$, respectively. Interestingly, these
+values are equal to $\epsilon|q|/(\gamma-\Gamma)$. This suggests that
+the critical thermal relaxation timescale for vertically
+non-isothermal disks can also be estimated by Eq. \ref{iso_vsi_cond}
+but with $\gamma-1$ replaced by $\gamma-\Gamma$.
 

### 2015-03-14 application.tex
updating

@@
 protoplanetary disk do the thermodynamic conditions allow the VSI to
-operate. For a realistic disk one can consider the energy equation with
-radiative cooling, 
+operate. For a realistic disk one can consider the energy equation
+with radiative cooling, 
 \begin{align}\label{real_energy}
-%\rho T \frac{DS}{Dt} = - \left(\nabla\cdot\bm{F} - Q_+\right), 
+  % \rho T \frac{DS}{Dt} = - \left(\nabla\cdot\bm{F} - Q_+\right), 
   \frac{\p P}{\p t} + \bm{v}\cdot\nabla P +\gamma P \nabla\cdot\bm{v} = - \frac{P}{\rho T
@@
 \end{align}
-%where the temperature $T$ is defined through $P=\mathcal{R}\rho
-% T/\mu$, $\mathcal{R}$ is the ideal gas constant and $\mu$ is
-% the mean molecular weight, 
 where $c_v$ is the heat capacity at constant
 volume, 
-% In Eq. \ref{real_energy} $S$ is the
-% specific entropy,
-% \begin{align}
-%   S \equiv c_v\ln{\left(\frac{p}{\rho^{\gamma}}\right)} 
-% \end{align}
-% where $c_v \equiv \left(\gamma-1\right)^{-1}\mathcal{R}/\mu$ is the
-% specific heat capcity at constant volume
 $\bm{F}$ is the radiative flux and $Q_+$ is a heating
-term such that an equilibrium solution exists.
+term to permit an equilibrium solution.
 
@@
 Eq. \ref{real_energy}, linearize them, then solve 
-the resulting linear eigenvalue problem. Such an approach is beyond 
-the scope of this paper. Here, we simply model the RHS of 
-Eq. \ref{real_energy} such that it can be represented by a 
-thermal relaxation time $t_c$ for the disk temperature. 
-%Our aim is to relate the right hand side (RHS) of Eq. \ref{real_energy} to
-%the 
+the resulting linear eigenvalue problem. This is beyond 
+the scope of the present work. Here, we take the simpler approach of
+modeling the linearized RHS of Eq. \ref{real_energy} such that it can 
+be represented by a thermal relaxation time $t_c$. 
 To do so, we first consider separately perturbations
@@
 
-%need to comment on kappa_d(rho,T)
-
 \subsection{Radiative diffusion}
 For perturbations with lengthscales $l\gg
-l_\mathrm{rad}$, we assume  
-temperature fluctuations are smoothed out by radiative diffusion. We
-model this in linear theory by writing the linearized RHS of
-Eq. \ref{real_energy} as 
+l_\mathrm{rad}$, we assume temperature fluctuations are smoothed out
+by radiative diffusion and model this in linear theory by writing the
+linearized RHS of Eq. \ref{real_energy} as 
 % we choose the heating term such that
@@
 
-However, Eq. \ref{diff_cool_proper} fundamentally changes the linear problem by
-introducing higher order derivatives. We defer a proper analysis of
-this problem to a follow-up study. Here, we are interested in
-order-of-magnitude estimates for perturbations with radial
+However, Eq. \ref{diff_cool_proper} fundamentally changes the linear
+problem by introducing higher order derivatives. We defer a proper
+analysis of this problem to a follow-up study. Here, we are interested
+in order-of-magnitude estimates for perturbations with radial
 lengthscales much smaller than vertical lengthscales. We thus proceed 
@@
 independent of $k_x$. For simplicity, we take  
-% the radiative cooling is given by {\bf (reference?)}
-% \begin{align}
-% \nabla\cdot\bm{F} = 4 \rho \kappa_d \sigma_s T^4,
-% \end{align}   
-% where $\sigma_s$ is the Stefan-Boltzmann constant. We choose the
-% heating term 
-% \begin{align}
-%   Q_+ = 4\rho\kappa_d\sigma_s T^{4}_{t=0}. 
-% \end{align}
-% Then Eq. \ref{real_energy} becomes
-% \begin{align}
-%   \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} -
-%   \frac{4P\kappa_d\sigma_s}{Tc_v}\left(T^4 - T_{t=0}^4\right), 
-% \end{align}
-% where $c_v$ is the heat capacity at constant volume. When linearized, the last term on the RHS becomes
-% \begin{align}
-%   -\frac{16\sigma_s}{c_v}\kappa_d T^3\left(\delta P -
-%     \frac{P}{\rho}\delta\rho\right). 
-% \end{align}
-% By comparing this expression with the linearized version of the energy
-% equation used in our calculations (Eq. \ref{full_energy1}), we identify 
-% \begin{align}\label{tc_newton_cool} 
-%   t_\mathrm{cool} = \frac{c_v}{16\sigma_s\kappa_dT^3}
-% \end{align}
-% as the thermal relaxation timescale for optically thin cooling. 
 \begin{align}
@@
 \end{align}
-% where the relation between $t_\mathrm{cool}$, $l_\mathrm{rad}$ and
-% $\eta$ can be inferred from Eq. \ref{lrad}, Eq. \ref{tc_newton_cool}
-% and Eq. \ref{eta_def}. 
 Eq. \ref{tc_def} is a simple prescription so
@@
 $\hat{\rho} = \rho/\rho_0$ and $\Sigma$ is the surface density at the
-radius of interest, the 
-% In order to utilize our linear results, which assumes $t_c$ is a
-% constant, we need to make further 
-% approximations. Specifically, we consider vertically-isothermal disks
-% and replace explicit $\rho$ dependencies in the above expressions by $\Sigma/
-% H$, where $\Sigma$ is the surface density at the fiducial
-% radius of interest. For example, the dimensionless diffusion
-% coefficient becomes 
-% \begin{align}\label{diff_coeff_dimensionless}
-%   \hat{\eta} \to  \frac{16\sigma_s T^3}{3\kappa_d\Sigma^2
-%     c_v\Omega_k}. 
-% \end{align}
-% Furthermore, we will choose a temperature-dependent opacity model for
-% simplicity, so that $\kappa_d=\kappa_d(T)$.
-dimensionless thermal
+radius of interest, the dimensionless thermal
 relaxation time $\beta$ becomes 
@@
 \end{align}
-Note that the thermal relaxation time depends on both the perturbation lengthscale $\khat$ 
-and distance from the midplane. %The latter is due to the density-dependence in the diffusion time. 
-%is then a constant for vertically-isothermal disks, 
-%and we can directly apply previous results.
+Note that $\beta$ depends on both the perturbation lengthscale $\khat$ 
+and distance from the midplane.
 
@@
 % \end{align}
-
-% The physical thermal relaxation timescale $\beta$ is height-dependent,
-% whilst for analytical discussion we assumed it is
-% constant. Nevertheless,
-It is instructive to compare the above thermal timescales
-with the criterion Eq. \ref{iso_vsi_cond} evaluated for the fiducial disk model,    
+It is instructive to compare the above thermal timescale 
+with the criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
 \begin{align}\label{bcrit_mmsn}
-  \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
-    -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
+  % \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
+  % -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
+  \beta_\mathrm{crit,MMSN} = 0.024r_\mathrm{AU}^{2/7}. 
 \end{align}
-We recall $\beta < \beta_\mathrm{crit}$ is needed for the
-fundamental VSI when $\khat\gg 1$. %For $\khat\to0$, this
-% becomes a sufficient condition (see Fig. \ref{relax_bcrit} and
-%\S\ref{iso_vsi_beta_crit}).
 
@@
 =2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for three
-opacity scales: $\hat{\kappa}_d=0.1, \,1 $ and $10$ at two heights:
-$z=0$ and $z=3H$. We also plot Eq. \ref{bcrit_mmsn} in each panel.
+opacity scales: $\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$ at two
+heights: $z=0$ and $z=3H$. We also plot Eq. \ref{bcrit_mmsn} in each
+panel.
 
@@
 Fig. \ref{mmsn_bcrit_bcool}. The critical thermal timescale,
-Eq. \ref{bcrit_mmsn}, as derived from our analytical discussion, is
-most useful to assess the VSI in the optically thin regime (where the
+Eq. \ref{bcrit_mmsn}, is useful in the optically thin regime (where the
 curves overlap), since in this limit $\beta$ does not depend on
-height. We thus infer that the VSI can operate at tens of AU in the
-MMSN.
+height, as assumed in our analysis. We thus infer that the VSI can
+operate at tens of AU in the MMSN.
 
-In the inner few AU of the MMSN, however, $\beta$ depends on $z$. 
-Given a fixed $\khat$, one can have $\beta > \beta_\mathrm{crit}$ at
-the optically thick midplane, but with increasing height the disk 
-becomes optically thin and $\beta<\beta_\mathrm{crit}$. Our analytical
-criteria is not valid for variable $\beta$, and we must turn to
-numerical solutions of the linear problem to determine whether or not
-the VSI operates. We do this in the next section.
+In the inner few AU of the MMSN, however, $\beta$ 
+may be smaller or greater than $\beta_\mathrm{crit}$ depending on
+$z$ as the disk transitions from an optically thick midplane to an
+optically thin atmosphere. 
+%Given a fixed $\khat$, one can have $\beta > \beta_\mathrm{crit}$ at
+%the optically thick midplane, but with increasing height the disk 
+%becomes optically thin and $\beta<\beta_\mathrm{crit}$. 
+Our analytical criteria is not valid for variable $\beta$, and we must
+turn to numerical solutions of the linear problem to determine whether
+or not the VSI operates. We do this in the next section.
 
-Now consider lowering the opacity, as shown in the top panels of
-Fig. \ref{mmsn_bcrit_bcool}. The curves shift inward, but are also
-raised. In fact, lowering the opacity by an order of magnitude renders 
+Lowering the opacity, as shown in the top panels of
+Fig. \ref{mmsn_bcrit_bcool}, shifts the curves inwards but also raises
+them. In fact, lowering the opacity by an order of magnitude renders  
 $\beta > \beta_\mathrm{crit}$ even away from the midplane. This is 
@@
 We remark that it is possible to satisfy $\beta < \beta_\mathrm{crit}$
-by choosing a sufficiently large $\khat$. We
+for sufficiently large $\khat$. We
 illustrate this in Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at 
 each radius the value of $\khat$ such  
-that $\beta = \beta_\mathrm{crit}$. We thus expect that the VSI,
-should it develop in the inner disk, would only occur on very small
-scales. However, our linear calculations suggest growth rates
-eventually decay for large $\khat$.
+that $\beta = \beta_\mathrm{crit}$. We thus expect that the VSI would only
+occur on very small radial scales in the inner disk. % However, these
+% may not be important as 
+% our
+% linear calculations suggest growth rates eventually decay for large
+% $\khat$.
 
-Furthermore, modes with large $\khat$ may be subject to viscous decay
-in a real disk. The viscous time for a perturbation lengthscale
+However, disturbances with large $\khat$ may be subject to viscous
+decay in a real disk. The viscous timescale for a perturbation lengthscale 
 $l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha_s\khat^2\Omega $, where
@@
 \citep{shakura73}. Setting $t_\mathrm{visc}$ to be longer than the
-minimum VSI growth timescale $1/\epsilon\Omega$ places an upper limit
+minimum VSI growth timescale $1/\epsilon\Omega$ implies 
 $\khat^2\lesssim \epsilon/\alpha_s$. Inserting $\epsilon \sim 0.05$,
@@
 VSI, see][]{nelson13}.
-
-% However, our linear
-% calculations indicate growth rates eventually decay as one increases
-% $\khat$. This, together with our finding of high $\khat$ modes may be
-% affected by vertical boundary conditions, suggest the VSI does not
-% operate in the inner few AU in protoplanetary disks where thermal
-% relaxation is entirely due to radiation \citep[cf.][]{stoll14}.
 
@@
 Here we solve the linear stability problem in the MMSN with the scale and
-height dependent thermal relaxation function defined by
+height dependent thermal relaxation timescale defined by
 Eq. \ref{beta_mmsn_simp}.  We use a vertical 
@@
 
-In fact, for $1\lesssim \khat \lesssim 30$ we either only find the
-fundamental mode, or that it is the most unstable. The associated 
-growth rates are also insensitive to $z_\mathrm{max}$. However, for
-large $\khat$ and/or $z_\mathrm{max}$, e.g. $\khat=50$ or 
-$z_\mathrm{max}=10H$, we also found modes with shorter growth times 
-than those shown in Fig. \ref{mmsn_overall}, but these disturbances are 
-concentrated near the vertical boundaries
-\citep{nelson13,mcnally14}. Such modes may not be robust because    
-they are affected by boundary conditions and viscosity in a real
-disk. % , and high $\khat$ modes are % subject to viscous decay in a real disk.
+In fact, for $\khat \lesssim 30$ we either only find the
+fundamental mode, or that it is the most unstable, and the associated
+growth rates are insensitive to $z_\mathrm{max}$. This is consistent with our
+analytical discussion (\S\ref{iso_vsi_beta_crit}) and previous
+numerical results (\S\ref{therm_relax_eff}), which indicate  
+higher order modes are more easily stabilized by finite thermal
+relaxation.
+
+However, for large $\khat$ and/or $z_\mathrm{max}$, we also found
+surface modes with shorter growth times than those shown in
+Fig. \ref{mmsn_overall}. Such modes may not be robust because    
+they are affected by boundary conditions (\S\ref{surf_comment}) and/or
+viscosity in a real disk.
 
@@
 VSI in the MMSN. We see the VSI operates from $\sim 5$AU to $\sim
-50$AU with growth timescales $\sim 30$---40 orbits. 
-Importantly, it is necessary to consider smaller scales as $r$
-decreases as the disk generally becomes optically thick.
+50$AU with growth timescales $\sim 30$---40 orbits. It is necessary to
+consider smaller scales as $r$ decreases as the disk generally becomes
+optically thick.
 
 Fig. \ref{mmsn_overall} is roughly consistent with the 
-mid-plane thermal relaxation timescales shown in
-Fig. \ref{mmsn_bcrit_bcool}. For example, for $\khat=10$ we find 
-$\beta \gtrsim \beta_\mathrm{crit}$ for $r\lesssim 10$AU at $z=0$; and
-Fig. \ref{mmsn_overall} show stabilization inside $\sim 8$AU where
-$t_\mathrm{grow}\to\infty$.
-
-%The mode % operates interior to the expected minimum radius, 
-% probably because the disk bcomes optically thin away from the
-% midplane.
+midplane thermal relaxation timescales for the MMSN ( 
+Fig. \ref{mmsn_bcrit_bcool}, left middle panel). For example, for 
+$\khat=10$ we find $\beta \gtrsim \beta_\mathrm{crit}$ for $r\lesssim
+10$AU at $z=0$; and Fig. \ref{mmsn_overall} show stabilization inside
+$\sim 8$AU where $t_\mathrm{grow}\to\infty$.
 
@@
   \includegraphics[width=\linewidth]{figures/eigen_compare_grow.ps}
-  \caption{Characteristic growth times of the fundamental VSI mode in
+  \caption{Characteristic growth times of the VSI in
     the MMSN. These are calculated using the scale and height
@@
 rotation).
-%: Fig. \ref{mmsn_overall}) show that
-%the growth time at
 
@@
 Fig. \ref{mmsn_eigenvz}. Although at $r=5$AU $\beta$ depends on $z$,
-the disturbance still occupies the fluid column. This is probably due
-to the vertically global nature of the fundamental mode. The fact that
+the disturbance still occupies the fluid column due to the vertically
+global nature of the fundamental mode. The fact that
 we obtain the VSI despite $\beta > \beta_\mathrm{crit}$ near the
-midplane suggests, but does not prove, that the VSI can operate
+midplane suggests that the VSI can operate 
 provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
 fraction of the vertical domain. However, the normalized growth times
-are much longer at $r=5$AU than at $r=50$AU for which  $\beta <
-\beta_\mathrm{crit}  \forall z$: we find $t_\mathrm{grow}\sim 40$
-orbits at $r=50$AU, but $t_\mathrm{grow}\sim 400$ orbits at
-$r=5$AU. Thus, introducing an optically thick midplane significantly
-reduces the efficiency of the VSI despite $\beta <
+are much longer at $r=5$AU than at $r=50$AU: %  for which  $\beta <
+% \beta_\mathrm{crit}  \forall z$: 
+$t_\mathrm{grow}\sim 40$ orbits at $r=50$AU, but $t_\mathrm{grow}\sim
+400$ orbits at $r=5$AU. Thus, introducing an optically thick midplane
+significantly reduces the efficiency of the VSI despite $\beta <
 \beta_\mathrm{crit}$ away from the midplane.
-
-% The grow
-% mode because it involves vertical motion across $z=0$, which is
-
-% the case with $\beta < \beta_\mathrm{crit}$ for all $z$.
-
-% Introducing an optically thick mid-plane may not completely surpress
-% the VSI because the vertical entropy gradient, which usually provides
-% stabilization, vanishes at $z=0$. (Note that $N_z^2\propto z^2$ while
-% vertical shear $d\Omega^2/dz \propto z$)
 
@@
   \caption{Vertical velocity perturbations of the VSI in the MMSN at 
-    $r=50$AU (top) and $r=5$AU (bottom). These are calculated using the height
-    dependent thermal relaxation given by Eq. \ref{beta_mmsn_simp}. The real (imaginary) part of 
+    $r=50$AU (top) and $r=5$AU (bottom). The real (imaginary) part of 
     $\delta v_z$ is shown as
-    the solid (dotted) curves. % The eigenfunction has been normalized
-    % such that $\imag(\delta v_z)=0$ at the vertical boundaries, and
-    % $\mathrm{max}|\delta v_z|=1$. 
+    the solid (dotted) curves. 
     \label{mmsn_eigenvz}}

### 2015-03-14 numerical.tex
updating

@@
 
-\subsubsection{Comment on `surface modes'} 
+\subsubsection{Comment on `surface modes'}\label{surf_comment}
 Our fiducial setup with $\khat=10$ in 
@@
 $\khat\geq 50$, we find perturbations have large amplitudes near the
-vertical boundaries, and complications to our numerical search of
-eigenvalues arise from emergence of modes associated with boundary
+ vertical boundaries, and 
+complications to our numerical search of
+eigenvalues arise from the emergence of modes associated with boundary
 conditions. Nevertheless, growth rates rapidly decrease as $\beta$
@@
    \includegraphics[width=\linewidth]{figures/gcorr_compare2} 
-   \caption{Growth rate of the fundamental VSI mode in the fiducial disk
-     model as a function of the thermal relaxation timescale $\beta$
-     The vertical line is the critical thermal timescale
-     $\beta_\mathrm{crit}$ obtained 
+   \caption{VSI growth rates in the fiducial disk
+     model as a function of the thermal relaxation timescale
+     $\beta$. These are obtained by following the eigenvalues of the
+     fundamental mode at $\beta=10^{-3}$. The vertical line is the
+     critical thermal timescale $\beta_\mathrm{crit}$ obtained  
      from Eq. \ref{iso_vsi_cond}. 

### 2015-03-14 appendix.tex
updating

@@
 \citep[VGSB,][]{mcnally14} to make analytic progress.   
-The main difference between the linear problem in the VGSB 
-and the radially-local approximation
-(Eq. \ref{lin_mass}---\ref{lin_energy}) adopted in our numerical study
+The difference between the linear problem in the VGSB 
+and the radially-local approximation adopted in our numerical study
 is the absence of terms explicitly associated with the radial disk structure in
-the former framework. These terms appear in the energy, mass and radial momentum
+the VGSB. These terms appear in the energy, mass and radial momentum
 equations and are labeled by $\hat{g}_c$ in
@@
 % of vertical shear. 
-
 In the presence of vertical shear $q\neq0$, Eq. \ref{adia_integral}
-suggests $\hat{\sigma}$ is complex for real $\khat$, which implies the
+suggests $\hat{\sigma}^2$ is complex for real $\khat$, which implies the
 possibility of instability, regardless of the value of
 $|\gamma-1|$. This is contradicts the second Solberg-Hoiland criterion 
-(Eq. \ref{solberg2}), which imply instability requires 
-$|\gamma-1|\ll1$ for thin, vertically isothermal disks evolved adiabatically. 
+(Eq. \ref{solberg2}), which imply instability for adiabatic
+perturbations in a thin, vertically isothermal disk would require 
+$|\gamma-1|\ll1$. 
 %also need real freq^2 in adia limit 
@@
 caveat, Fig. \ref{gcorr_compare} shows that provided we consider
-$\beta\ll1$, then the VGSB framework is an adequate approximation to
-study the VSI.
+$\beta\ll1$, then the VGSB framework is adequate for studying the
+VSI.
 
@@
     accounts for the radial disk structure.  
-    The disk parameters are $\Gamma=1.011$,
-    $\gamma=1.4$ and $(p,q,\epsilon)=(-1.5,-1,0.05)$. The perturbation
+    The disk parameters are $(\gamma, \Gamma)=(1.4, 1.011)$ and
+    $(p,q,\epsilon)=(-1.5,-1,0.05)$. The perturbation
     wavenumber is $\khat=30$. 
@@
 Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
-3+q$ as $|z|\to\infty$. Since we consider $q=O(1)$, $|F|$ is of
-order unity.
+3+q$ as $|z|\to\infty$. Since we consider $|q|=O(1)$, this impies
+$|F|=O(1)$ as well.
 
@@
 neglecting these terms has no qualitative effect on the discussion in 
-\S\ref{analytic_relax}.  % This issue does not
-% arise for isothermal perturbations discussed in \S\ref{iso_discuss}
-% since in that case we work with a governing equation for $W$ instead.
+\S\ref{analytic_relax}.
 
@@
 These equations are used to derive a critical thermal timescale in
-\S\ref{iso_vsi_beta_crit}. For fixed $M$ of $O(1)$,  we assume
+\S\ref{iso_vsi_beta_crit}. For fixed $M=O(1)$,  we assume
 $|\hat{\omega}\khat|$ remain constant but consider $\khat\gg 1$ so
@@
 equivalent expression for $M>0$ (Eq. \ref{bcrit_gen}).
-
-%  In this case the fourth term
-% in Eq. \ref{relax_cond1} may be retained but other aforementioned terms
-% are still neglected. One then obtains an expression for $\beta$ by
-% eliminating $\hat{\omega}\khat$ between Eq. \ref{relax_cond1} ---
-% \ref{relax_cond2}. The expression for the critical thermal timescale
-% for the fundamental mode, $\beta_\mathrm{crit}$
-% (Eq. \ref{iso_vsi_con}, is then obtained by taking the thin disk
-% limit.
 
@@
 Here we show that the growth rate is limited by the maximum vertical
-shear in the domain. % Writing $\hat{\sigma} = \hat{\omega} +
+shear rate in the domain. % Writing $\hat{\sigma} = \hat{\omega} +
 % \ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, 
@@
 satisfying the above boundary conditions, is limited by the maximum
-vertical shear in the domain considered.
+vertical shear rate in the domain considered.
 
@@
 \end{align}
-Then the coefficients must satisfy the recurrence relation
+Then the coefficients $a_l$ are given by the recurrence relation
 \begin{align}
@@
 \end{align}
-We demand the series to terminate  at $l=L$, i.e. a polynomial
-solution, so that the vertical kinetic energy remain bounded as
+We demand the series to terminate  at $l=L$, i.e. a polynomial of
+order $L$, so that the vertical kinetic energy remain bounded as 
 $|\zhat|\to\infty$.  Then the eigenfrequency $\hat{\sigma}$ is given via 
@@
   \hat{\sigma}^2\khat^2 =
-  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right). \label{iso_disp_full}
+  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right), \label{iso_disp_full}
 \end{align}
-This is just the dispersion relation for axisymmetric isothermal waves in a
+which is just the dispersion relation for axisymmetric isothermal waves in a
 vertically isothermal disk
@@
 \subsubsection{Instability with vertical shear}
-A simple solution for $\hat{\sigma}$ may be obtained by balancing
-the  last two terms of Eq. \ref{simple_growth} to obtain the
-low-frequency branch,  
+For large $\khat^2$, the low-frequency branch may be obtained by balancing
+the last two terms of Eq. \ref{simple_growth} . Then 
 \begin{align}
@@
 \end{align}
-This signifies instability for $L\geq1$ 
+This signifies instability for $L\geq1$ --- the VSI ---   
 since we can choose the sign of the square root to make $\hat{\nu} =
-\imag\hat{\sigma}>0$.
+\imag\hat{\sigma}>0$.  These are the low frequency unstable modes observed
+in our numerical calculations in \S\ref{vertiso_pertiso}.
+
+%growth rate of high frequency branch is much smaller
 

### 2015-03-14 application.tex
updating

@@
 In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn}  with $\mu
-=2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for three
-opacity scales: $\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$ at two
-heights: $z=0$ and $z=3H$. We also plot Eq. \ref{bcrit_mmsn} in each
-panel.
+=2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for opacity scales
+$\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$, at $z=0$ and $z=3H$. We
+also plot Eq. \ref{bcrit_mmsn} in each panel.
 
@@
 
-In fact, for $\khat \lesssim 30$ we either only find the
-fundamental mode, or that it is the most unstable, and the associated
-growth rates are insensitive to $z_\mathrm{max}$. This is consistent with our
-analytical discussion (\S\ref{iso_vsi_beta_crit}) and previous
-numerical results (\S\ref{therm_relax_eff}), which indicate  
-higher order modes are more easily stabilized by finite thermal
-relaxation.
+In fact, we either only find the fundamental mode, or that it is the
+most unstable, except for the appearence of some surface modes when considering large
+$\khat\gtrsim 50$. However, surface modes are dependent on
+boundary conditions and disturbances with large $\khat$ are more likely
+subject to viscous decay, as discussed above. The dominance of the
+fundamental mode is also observed in hydrodynamic
+simulations of \cite{stoll14} which treat radiative cooling properly.
 
-However, for large $\khat$ and/or $z_\mathrm{max}$, we also found
-surface modes with shorter growth times than those shown in
-Fig. \ref{mmsn_overall}. Such modes may not be robust because    
-they are affected by boundary conditions (\S\ref{surf_comment}) and/or
-viscosity in a real disk.
+% The dominance of the
+% fundamental mode is consistent with our analytical discussion
+% (\S\ref{iso_vsi_beta_crit}) and previous numerical results (\S\ref{therm_relax_eff}), which indicate  
+% higher order modes are more easily stabilized by finite thermal
+% relaxation.
+
+% In fact, for $\khat \lesssim 30$ we either only find the
+% fundamental mode, or that it is the most unstable, and the associated
+% growth rates are insensitive to $z_\mathrm{max}$. This is consistent with our
+% analytical discussion (\S\ref{iso_vsi_beta_crit}) and previous
+% numerical results (\S\ref{therm_relax_eff}), which indicate  
+% higher order modes are more easily stabilized by finite thermal
+% relaxation.   
+% However, for large $\khat$ and/or $z_\mathrm{max}$, we also found
+% surface modes with shorter growth times than those shown in
+% Fig. \ref{mmsn_overall}. Such modes may not be robust because    
+% they are affected by boundary conditions (\S\ref{surf_comment}) and/or
+% viscosity in a real disk.
 
@@
 
-\subsection{Comparing the VSI in the inner and outer parts of the MMSN} 
+\subsection{Numerical examples} 
 We compare the VSI in the MMSN at $r=50$AU and $r=5$AU. We first show in
@@
 
-In the outer disk ($r=50$AU) the disk is optically thin, $\beta$ 
-does not depend on height and $\beta <
-\beta_\mathrm{crit}$. Accordingly we expect, and find, the VSI to
-operate at $r=50$AU. In the inner disk ($r=5$AU) the disk is optically
+The outer disk ($r=50$AU) is optically thin,
+$\beta<\beta_\mathrm{crit}$  and  does not depend on height. 
+Accordingly we expect, and find, the VSI 
+operates at $r=50$AU. The inner disk ($r=5$AU) is optically
 thick near the midplane and becomes optically thin away from it, such
@@
 we obtain the VSI despite $\beta > \beta_\mathrm{crit}$ near the
-midplane suggests that the VSI can operate 
+midplane suggests that the instability can operate 
 provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large

### 2015-03-14 numerical.tex
updating

@@
 We now consider the effect of finite thermal relaxation by increasing 
-$\beta$. Note that for the standard disk model the critical thermal 
+$\beta$. Note that for the fiducial disk model the critical thermal 
 timescale $\beta_\mathrm{crit} = 0.125$.
@@
 assessing stability as $\beta$ is increased from zero, rather than
-searching for the most unstable mode, which are affected by boundary 
-conditions.
+searching for the most unstable mode, which may correspond to modes
+that depend on boundary conditions.
 
@@
 the neutrally-stratified case $\gamma=\Gamma$ the disk is unstable 
-even for $\beta\gg 1$. (Note that in the adiabatic limit 
-this disk can be unstable according to the
-Solberg-Hoiland criterion, Eq. \ref{solberg2}.)
+even for $\beta\gg 1$. This is due to the absence of a stabilizing
+vertical entropy gradient ($N_z^2\equiv 0$). % (Note that in the adiabatic limit 
+% this disk can be unstable according to the
+% Solberg-Hoiland criterion, Eq. \ref{solberg2}.)
 

### 2015-03-14 summary.tex
updating

@@
 \section{Summary and discussion}\label{summary}
-%what have we done, relation to previous work 
-In this paper we have studied the linear axisymmetric vertical shear
+In this paper we have studied the  axisymmetric vertical shear
 instability (VSI) in astrophysical disks that arises from the
-height-dependence of the equilibrium angular velocity profile
+height dependence of the equilibrium angular velocity profile
 $\Omega=\Omega(r,z)$ in a baroclinic disk. We examined how
@@
 
-% The VSI was analyzed earlier in 
-% the local limit by \cite{urpin98} and 
-% \cite{urpin03} and simulated by \cite{arlt04}. Recent non-linear
-% numerical simulations by \cite{nelson13} and \cite{stoll14} indicate
-% the VSI generally requires a short thermal timescale.
-
-We generalized previous vertically global linear stability calculations
-of the VSI \citep{nelson13,mcnally14}, which assumed isothermal
+We generalized previous vertically global linear 
+calculations of the VSI \citep{nelson13,mcnally14}, which assumed isothermal
 perturbations, to include an energy equation. Our effort may also be
@@
 temperature to its initial value on a timescale
-$t_c=\beta\Omega_k^{-1}$. This approach to model thermal relaxation was also
-employed in the global numerical simulations of \cite{nelson13}, and
-allows the thermodynamic response of the disk to be controlled by
-varying the parameter $\beta$. However, it is possible to model
-$\beta$ self-consistently by relating it to thermal timescales
-expected in realistic disks.
+$t_c=\beta\Omega_k^{-1}$. This approach allows the thermodynamic
+response of the disk to be controlled by varying the parameter
+$\beta$, and has been employed in previous numerical simulations of
+the VSI \citep{nelson13}. However, it is possible to 
+relate  $\beta$ to physical thermal timescales expected in realistic disks.
 
@@
 the `vertically-global shearing box' (VGSB) formalism developed by
-\cite{mcnally14}, who showed the linear VSI studied by
-\cite{nelson13} in a locally isothermal disk can be captured by the
-VGSB. We find the VGSB framework is appropriate for non-isothermal
-perturbations provided $\beta\lesssim1$.  
-This is convenient because rapid thermal relaxation ($\beta\ll1$) is
+\cite{mcnally14}. % who showed the linear VSI studied by
+% \cite{nelson13} in a locally isothermal disk can be captured by the
+% VGSB. 
+We find the VGSB framework is appropriate for non-isothermal
+perturbations provided $\beta\lesssim1$.  This is convenient because
+rapid thermal relaxation ($\beta\ll1$) is 
 generally required for the VSI.  However, the VGSB framework is not
@@
 
-We mostly focused on vertically isothermal disks ($\Gamma=1$), since
+We mostly focused on vertically isothermal disks ($\Gamma=1$) as 
 this is the relevant case for protoplanetary disks that are passively irradiated by its central star 
@@
 \cite{lubow93} by  adding vertical shear and (rapid) thermal
-relaxation, but in the limit of low-frequency perturbations in
-nearly-Keplerian disks. These approximations were made for simplicity,
-but also motivated by recent numerical simulations
-\citep{nelson13}. We obtained explicit solutions for the VSI with
+relaxation. We obtained explicit solutions describing the VSI with 
 thermal relaxation in the thin disk limit.
-
-% For isothermal perturbations ($\beta=0$) we explicitly
-% demonstrated instability  associated with vertical shear, and that the 
-% VSI growth rate is limited by the maximum vertical shear within the 
-% vertical domain of interest. 
-% For perturbations with finite thermal relaxation, the general
-% dispersion relation (Eq. \ref{relax_disp}) is quite
-% complicated. However, for the fundamental VSI some simplification is
-% possible. In particular,
 
@@
 protoplanetary disk model with thermal relaxation based on realistic
-disk properties, we find the fundamental mode to be the dominant mode,
- and it is robust against boundary conditions. The fundamental mode is
+disk properties, we find the fundamental mode to be the dominant
+mode. The fundamental mode is
 thus likely the most relevant VSI mode in protoplanetary disks.
 
-We quantified the critical thermal relaxation timescale 
+We determined the critical thermal relaxation timescale 
 $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$, below which small-scale
-perturbations may be regarded as isothermal and the fundamental VSI
-operates. %  We expect $\beta<\beta_\mathrm{crit}$ to be a sufficient
-% condition for the VSI to operate.
-We have therefore provided a quantitative measure for the `fast'
+perturbations may be regarded as isothermal and the VSI 
+operates. We have therefore provided a quantitative measure for the `fast'
 thermal relaxation requirement for the VSI in astrophysical disks.
@@
 approximations made in our analysis, and accounted for
-the background radial disk structure for consistency.  Our
-numerical solutions are mostly consistent with our analytic 
+the background radial disk structure.  Our
+numerical results are mostly consistent with our analytic 
 expectations, including the dependence of $\beta_\mathrm{crit}$ on
-disk parameters.  %for regimes of interest
-
-% as given
-% by Eq. \ref{iso_vsi_cond}, provides a good measure of the required
-% th 
-%  except for perturbations with small radial
-% wavenumber ($\khat\to0$) where the low-frequency approximation is
-% inappropriate. The radially local model is also not valid for such modes.
+disk parameters.  We find the fundamental mode is robust against
+boundary conditions, and higher order modes are more rapidly
+stabilized with increasing $\beta$.
 
@@
 code. We confirm the finding by \cite{nelson13} that in the absence
-of vertical entropy gradients, the VSI is insensitive to
-thermal relaxation. When a stable stratification is introduced, the
+of vertical entropy gradients, the VSI does not require rapid thermal
+relaxation.  When a stable stratification is introduced, the
 rapid reduction in VSI growth rates by a finite thermal relaxation
@@
 We applied our linear models to assess the applicability of the VSI in
-protoplanetary disks, using a model for the the Minimum Mass Solar Nebulae
-(MMSN) described in \cite{chiang10}. We  
+protoplanetary disks. We considered the Minimum Mass Solar Nebulae
+(MMSN) described in \cite{chiang10} and   
 modeled the thermal relaxation timescale $\beta$ based on realistic
-disk properties and dust opacity, which results in a $\beta$ that
+disk properties and dust opacity. This results in a $\beta$ function that
 depends on both the perturbation lengthscale as well as the height 
@@
 $\sim 30$ orbits. Although $\beta$ is not vertically constant in the
-MMSN as assumed in our analysis, we find that evaluating $\beta <
-\beta_\mathrm{crit}$ at the disk midplane gives a rough measure of the
-radii at which the VSI operates in the MMSN.  This simple criteron may
-be a useful first step to estimate the relevance of the VSI in
+MMSN as assumed in our analysis, we find that evaluating the condition
+$\beta <\beta_\mathrm{crit}$ at the disk midplane gives a rough measure of the
+radii at which the VSI operates in the MMSN. This simple criteron is
+thus a useful first step to estimate the relevance of the VSI in
 realistic astrophysical disks without resorting to non-linear
 numerical simulations.
-
-%  However, we note the strong
-% temperature dependence of the physical thermal relaxation timescale
-% in the disk model ($\beta$ decreases as $T^{-5}$ for perturbations with small
-% radial scale). Thus, increasing the disk temperature even slightly can significantly 
-% shorten the typical growth times for the VSI. This is also attributed
-% to a larger value of $\beta_\mathrm{crit}\propto\epsilon$, since increasing $T$ also
-% increases the aspect-ratio $\epsilon$.
-
-%relation with kley simulations
 
@@
 
-The convective overstability corresponds to growing epicyclic motions.
-It is most effective for disturbances with $|k_x|\ll |k_z|$, where
-$k_z$ is a vertical wavenumber, and $\beta\sim 1$. These conditions
-are mutually exclusive with the VSI. 
-The maximum growth 
-rate for convective overstability is given by 
+The convective overstability corresponds to slowly growing epicyclic motions.
+It is most effective for $\beta\sim 1$ and disturbances with $|k_x|\ll |k_z|$, where
+$k_z$ is a vertical wavenumber. These conditions are mutually exclusive with the VSI. 
+The maximum growth rate for convective overstability is given by 
 \begin{align}
@@
 % whereas the maximum VSI growth rate is $O(\epsilon\Omega_k)$. 
-For the MMSN considered in this work (\S\ref{application} and Appendix
+In fact, for the MMSN considered in this work (\S\ref{application} and Appendix
 \ref{mmsn}), convective overstability does not apply because
@@
 However, for more general disk structures convective overstability is
-possible. In fact, the models adopted for our numerical study in
-\S\ref{numerical} do satisfy $N_r^2<0$. %show early result
-We will explore convective overstability more fully in a follow-up
-study.  Although our linear code can be used for convective
-overstability in principle, the fact convective overstability is
-relevant for perturbations with $k_x\to0$ suggest radially global disk
-models may be more appropriate.
+possible. The models adopted for our numerical study in
+\S\ref{numerical} do satisfy $N_r^2<0$, but convective overstability
+requires considering perturbations with long radial lengthscales,
+$\khat\ll 1$, for which radially global disk models would be more 
+appropriate.  
+% We will explore convective overstability more fully in a
+% follow-up study.
 
-% Nevertheless, it will be important to expand the parameter space covered in
-% this study to include convective overstability,
+\subsubsection{Improved thermal evolution} 
+We assumed non-adiabatic effects
+can be represented by a thermal relaxation or cooling timescale $t_c$. For 
+application to protoplanetary disks we connected $t_c$ to the disk
+properties. One one can improve the realism further by generalizing the 
+opacity model adopted here to be density (and hence height) dependent. 
+%here to be density-dependent,  i.e. $\kappa_d =
+%% \kappa_d(\rho, T)$. 
+However, going beyond thermal relaxation requires a
+proper treatment for radiative diffusion. The linear problem then
+involves a Laplacian operator, which require additional boundary
+conditions. We will consider these improvements in a
+follow-up study. %additional heat source
 
-%self-consistence may require radially global disk models
-
-\subsubsection{Improved thermal evolution.} 
-We assumed non-adiabatic effects
-can be represented by a thermal relaxation timescale $t_c$. For 
-application to protoplanetary disks we connected $t_c$ to the disk
-properties, but one can improve the realism further by generalizing the
-opacity model adopted here to be density-dependent,  i.e. $\kappa_d =
-\kappa_d(\rho, T)$. 
-%two temperature
-% structure
-
-However, going beyond the framework of thermal
-relaxation requires a proper treatment for radiative diffusion, which
-involves the Lapacian operator and additional boundary conditions
-must be supplemented. We will consider these improvements in a
-follow-up study.
-
-\subsubsection{Additional physics.}
-
+\subsubsection{Additional physics} 
 Depending on where the VSI operates in a
@@
 outer parts of a protoplanetary disk, where we estimate the VSI 
-operates, disk self-gravity (SG) may play a role. Specifically, the vertical
-shear profile depends on the vertical density stratification
-(Eq. \ref{vertical_shear}) which can be enhanced by vertical SG. This
-may promote the VSI. Inclusion of the Poisson equation should
-therefore be considered.
+operates, disk self-gravity (SG) may play a role. Vertical
+self-gravity may favor the VSI by increasing $|\p_z\Omega|$ through
+enhancing the density stratification 
+(Eq. \ref{vertical_shear}). %sg in pert? no because kx>>1. but maybe for kx~1
 
-We have neglected viscosity. \cite{nelson13} have shown that an
-alpha viscosity of $\lesssim O(10^{-4})$ is required for the VSI to
-operate. It is of interest to quantify the effect of viscosity
+We have neglected viscosity. \cite{nelson13} find that an 
+alpha viscosity $\lesssim O(10^{-4})$ is required for the VSI. 
+It is of interest to quantify the effect of viscosity 
 on the linear VSI. This would help identify, more accurately, the
@@
 in the energy equation.
-
-% \emph{Radial dependence.} Based on previous works, including
-% non-linear simulations, which show that the VSI is relevant for
-% perturbations with small radial lengthscales, we have adopted the
-% radially-local approximation throughout this paper.  Although this
-% simplifies the linear problem to ordinary differential equations, this
-% approach introduces the wavenumber $k_x$ as a free parameter. The
-% effect of radial boundary conditions cannot be explored. Future 
-% generalizations of the linear problem should retain the radial
-% dependence properly and solve the two-dimensional $(r,z)$ partial 
-% differential equation eigenvalue problem.
-
-% Extrapolating our results from the shearing box to a global disk  
-% has an inconsistency. Namely, the equilibrium temperature in the
-% box only depends on height (if at all). However, it is precisely the
-% radial temperature (or entropy) gradient in the global disk that
-% gives rise to vertical shear. We do not expect this to be a serious issue for 
-% perturbations with small radial wavelengths, 
-% but will need to be resolved by adopting global disk models  
-% if one wishes to generalize the VSI to large-scale perturbations.
-
-% \emph{Theoretical generalizations.} Our analytical discussion
-% (\S\ref{analytical}) can be extended in several ways. One could
-% relax the low-frequency approximation to improve the agreement between
-% numerical and analytical eigenfrequencies at small $\khat$. However,
-% such modes are radially more global, and would need a global
-% cylindrical disk model to study them. Another minor improvement is to
-
-%analytical extensions -> nonisothermal, all frequency, non-keplerian,
-%diffusion 
-%viscosity self-gravity 
-%disk models: polytropic connected with isothermal

### 2015-03-15 numerical.tex
updating

@@
 We first calculate the VSI in a disk with rapid thermal relaxation by
-setting $\beta=10^{-3}$. (Smaller values yield the same results.) This
+setting $\beta=10^{-3}$. (Smaller values yield the similar results.) This
 is similar to strictly isothermal perturbations in a vertically
@@
 for the fundamental mode and that calculated from 
-Eq. \ref{simple_growth} (with $L=1$). We find good agreement in the
+Eq. \ref{simple_growth}. We find good agreement in the
 wave frequency $\omega$ at all $\khat$ and in growth rates $\nu$ for
@@
 the analysis. Nevertheless, given the number of approximations made
-in our analysis, the agreement is satisfactory, especially at large
-$\khat$.
+in our analysis, the agreement is satisfactory.
 
@@
 position in all cases. Higher order modes with larger growth rates
-(i.e. the branch where $\nu$ increases with $|\omega|$) require larger
-vertical domains, since they have increasing number of nodes along 
-$z$. The `tail' end with decreasing $\nu$ to the right
-of the plot differ significantly between the cases, and are thus
-likely associated with boundary conditions.
+(along the analytic lines) require larger vertical domains, since they
+have increasing number of nodes along $z$.
+
+The `tail' end with decreasing $\nu$ to the right of
+Fig. \ref{compare_modes_iso_kx10} is not predicted from
+analysis. These modes differ significantly between cases and are thus
+likely associated with boundary conditions. % We do not consider them 
+% further.
 
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10.ps}
+  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10_analytic.ps}
   \caption{Unstable modes with $\khat=10$ in the fiducial disk model
@@
     cases correspond to different boundary condtions applied at
-    different heights. The fundamental mode is marked by `f'.   
+    different heights. The fundamental mode is marked by `f'. Lines
+    are computed from the analytic dispersion relation 
+    with thermal relaxation Eq. \ref{relax_disp} (dotted), and that for
+    isothermal perturbations, Eq. \ref{simple_growth} (dashed). 
     \label{compare_modes_iso_kx10} 
@@
 conditions. Surface modes are readily identified for 
-$\zmax\geq 5H$ and are marked by `s'. They typically
-have the largest growth rates but require sufficiently large vertical
-domain size and their eigenvalues vary with $\zmax$.
+$\zmax\geq 5H$ and are marked by `s'. They have the largest growth
+rates but require sufficiently large vertical domain size and their
+eigenvalues vary with $\zmax$.
 
 Due to their dependence on boundary conditions, we 
-do not consider surface modes further in detail. These modes have
+do not consider surface modes further in detail. Surface modes have
 already been discussed in the aforementioned studies. In fact, we will
@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_tags.ps}
+  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_analytic.ps}
   \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for $\khat=30$
@@
 and $\beta\in[0.01,0.1]$. In addition to our fiducial setup with 
-$\zmax=5H$, we also plot results for $\zmax=3H$ and $\zmax=7H$. In all 
-three cases the number of unstable modes decrease with increasing
-$\beta$. The fundamental mode is not the most unstable for 
-$\beta=0.01$ as there is a number of higher order modes, but 
-their existence depends on $\zmax$.
+$\zmax=5H$, we also plot results for $\zmax=3H$, $\zmax=7H$ and 
+analytic results computed from Eq. \ref{relax_disp}.
+
+We find the number of unstable modes decrease with increasing
+$\beta$. Consider the fiducial setup with $\zmax=5H$ 
+(middle panel). The fundamental mode ($M=0$) is not the most unstable for   
+$\beta=0.01$ as there are modes with $M>0$ that have larger growth 
+rates. However, by comparing with results for $\zmax=3H$, we see these
+higher order modes are strongly influenced by boundary conditions, and
+there is a significant mismatch between numerical and analytical
+results for $\zmax=3H$ for $M>1$.
 
 However, the fundamental mode is dominant for $\beta \geq 
-0.03$. Higher order modes, with larger growth rates, are more
-effectively stabilized with increasing $\beta$. For $\beta=0.1$, which
-is close to $\beta_\mathrm{crit}$, we only find the fundamental
-mode. This is consistent with the discussion in 
+0.03$. Higher order modes, with larger growth rates at small $\beta$,
+are more effectively stabilized with increasing $\beta$. For
+$\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we only find the
+fundamental mode. This is consistent with the discussion in 
 \S\ref{iso_vsi_beta_crit}.
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_modes_cool_kx10_z3.ps} 
+  0cm]{figures/compare_modes_cool_kx10_z3_analytic.ps} 
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_modes_cool_kx10_z5.ps}
-  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx10_z7.ps}
+  0cm]{figures/compare_modes_cool_kx10_z5_analytic.ps}
+  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx10_z7_analytic.ps}
   \caption{Unstable modes with $\khat=10$ and thermal
@@
     $\zmax=5H$ (middle, our fiducial setup) and $\zmax=7H$
-    (bottom). The fundamental mode is marked with `f'.  
+    (bottom). Lines are computed analytically from
+    Eq. \ref{relax_disp} for  $M=0$ (solid, fundamental mode) and
+    $M=1,\,2,\,3$ (dotted, dashed, dash-dot, respectively). Along each
+    line, $\beta$ increases continously from $0.01$ to $0.1$ from top
+    to bottom, and squares mark corresponding $\beta$ values with
+    numerical results. 
     \label{compare_modes_cool_kx10} 
@@
 
+%fundamental mode bcomes the dominant mode as we approach beta
+%critical, therefore can use beta crit as criteria
+
 In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable
-modes with $\khat=30$. In this case surface modes appear for
-$\zmax\geq5H$. However, they are effectively stabilized with
+modes with $\khat=30$. For clarity we only plot analytical results for
+the fundamental mode, which is again robust against boundary
+conditions compared to higher order modes. We find surface modes
+appear for $\zmax\geq5H$. However, they are effectively stabilized with
 increasing $\beta$. For $\zmax=3H$, the fundamental mode becomes
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_modes_cool_kx30_z3.ps} 
+  0cm]{figures/compare_modes_cool_kx30_z3_analytic.ps} 
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_modes_cool_kx30_z5.ps}
-  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx30_z7.ps}
+  0cm]{figures/compare_modes_cool_kx30_z5_analytic.ps}
+  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx30_z7_analytic.ps}
   \caption{Same as Fig. \ref{compare_modes_cool_kx10} but for
@@
 conditions. The fundamental mode also becomes dominant as $\beta$
-approaches $\beta_\mathrm{crit}$.  
+approaches $\beta_\mathrm{crit}$. %beta<beta_crit is necessary and sufficient  
 Thus, tracking the fundamental mode is a more robust way of 
@@
 operate slightly beyond $\beta_\mathrm{crit}$ (recall that
-$\beta_\mathrm{crit}$ was derived by assuming $\khat^2\gg 1$). For
+$\beta_\mathrm{crit}$ was derived for $\khat^2\gg 1$). For
 $\khat\geq 50$, we find perturbations have large amplitudes near the
@@
 eigenvalues arise from the emergence of modes associated with boundary
-conditions. Nevertheless, growth rates rapidly decrease as $\beta$
+conditions. %We generally find boundary effects become more significant
+%with increasing $\khat$.  
+Nevertheless, growth rates rapidly decrease as $\beta$
 approaches and increase beyond $\beta_\mathrm{crit}$.
@@
 performed by \cite{nelson13} for which a thermal relaxation timescale
-$\beta\gtrsim 0.6$ stablizes the VSI. This can be anticipated from our
+$\beta\gtrsim 0.6$ stablized the VSI. This can be anticipated from our
 estimate of the critical thermal timescale $\beta_\mathrm{crit}\simeq
@@
  $\beta_\mathrm{crit}$ on disk parameters
- (Eq. \ref{iso_vsi_cond}).  We use the previous case study as a
+ (Eq. \ref{iso_vsi_cond}).  We use the fiducial setup as a
  reference point and vary $q\in[-0.2,-1.2]$, 

### 2015-03-15 appendix.tex
about to edit numerical

@@
 \He_L(\zhat)$.  The eigenfrequency $\hat{\sigma} = \hat{\omega}$ is real and the disk
-is stable. The low-frequency branch of Eq. \ref{iso_disp_full} are 
+is stable. The low frequency branch of Eq. \ref{iso_disp_full} are 
 inertial waves \citep{balbus03}. For 
@@
 \subsubsection{Instability with vertical shear}
-For large $\khat^2$, the low-frequency branch may be obtained by balancing
-the last two terms of Eq. \ref{simple_growth} . Then 
+The VSI corresponds to unstable inertial waves. This is readily
+obtained for large $\khat^2$ by balancing
+the last two terms of Eq. \ref{simple_growth}  to obtain the low
+frequency branch. Then 
 \begin{align}
@@
 \end{align}
-This signifies instability for $L\geq1$ --- the VSI ---   
-since we can choose the sign of the square root to make $\hat{\nu} =
-\imag\hat{\sigma}>0$.  These are the low frequency unstable modes observed
-in our numerical calculations in \S\ref{vertiso_pertiso}.
+This signifies instability for $L\geq1$ since we can choose the sign
+of the square root to make $\hat{\nu} = \imag\hat{\sigma}>0$.  These
+are the low frequency unstable modes seen in
+Fig. \ref{compare_modes_iso_kx10} and
+Fig. \ref{compare_modes_iso_kx30} , for which $\nu\propto\omega$.
 

### 2015-03-15 application.tex
about to edit numerical

@@
 Here we solve the linear stability problem in the MMSN with the scale and
-height dependent thermal relaxation timescale defined by
-Eq. \ref{beta_mmsn_simp}.  We use a vertical 
-domain $z_\mathrm{max}=5H$. In Fig. \ref{mmsn_overall} we plot growth
-timescales $t_\mathrm{grow} \equiv 1/\nu$ associated with the
-fundamental mode in units of the local orbital time
-$P_\mathrm{orb}\equiv 2\pi/\Omega_k$.
+height dependent thermal relaxation timescale given by
+Eq. \ref{beta_mmsn_simp}.  Fig. \ref{beta_compare} show examples of
+the thermal timescales at $r=5$AU and $r=50$AU for a
+perturbation wavenumber $\khat=30$. We also plot the 
+corresponding critical thermal timescales evaluated for the MMSN
+(Eq. \ref{bcrit_mmsn}).  
+ \begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/beta_compare}
+  \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
+     and $r=5$AU for $\khat=30$ (solid lines). The
+    corresponding horizontal dashed lines are the critical thermal
+    relaxation timescales derived in linear theory. 
+    \label{beta_compare}}
+\end{figure}
 
-In fact, we either only find the fundamental mode, or that it is the
-most unstable, except for the appearence of some surface modes when considering large
-$\khat\gtrsim 50$. However, surface modes are dependent on
-boundary conditions and disturbances with large $\khat$ are more likely
-subject to viscous decay, as discussed above. The dominance of the
-fundamental mode is also observed in hydrodynamic
-simulations of \cite{stoll14} which treat radiative cooling properly.
+The outer disk ($r=50$AU) is optically thin,
+$\beta<\beta_\mathrm{crit}$  and  does not depend on height.
+Accordingly we expect, and find, the VSI operates at $r=50$AU. The
+inner disk ($r=5$AU) is optically thick near the midplane but becomes
+optically thin away from it, such that $\beta < \beta_\mathrm{crit}$
+for $|z|\gtrsim1.25H$. In fact, we also find the VSI at $r=5$AU, but
+with much reduced growth rates (when scaled by the local rotation).
 
-% The dominance of the
-% fundamental mode is consistent with our analytical discussion
-% (\S\ref{iso_vsi_beta_crit}) and previous numerical results (\S\ref{therm_relax_eff}), which indicate  
-% higher order modes are more easily stabilized by finite thermal
-% relaxation.
+In Fig. \ref{mmsn_overall} we plot characteristic growth timescales
+$t_\mathrm{grow} \equiv 1/\nu$ of the fundamental mode with a range of
+wavenumbers, in units of the local orbital time $P_\mathrm{orb}\equiv
+2\pi/\Omega_k$.  In fact, we either only find the fundamental mode, or
+that it is the most unstable, except for the appearence of some
+surface modes towards the inner disk when considering large
+wavenumbers $\khat\gtrsim 50$. However,
+surface modes are affected by boundary conditions and disturbances
+with large $\khat$ are more likely subject to viscous decay, as
+discussed above. The dominance of the fundamental mode is also
+observed in hydrodynamic simulations of \cite{stoll14} which treat
+radiative cooling properly.
 
-% In fact, for $\khat \lesssim 30$ we either only find the
-% fundamental mode, or that it is the most unstable, and the associated
-% growth rates are insensitive to $z_\mathrm{max}$. This is consistent with our
-% analytical discussion (\S\ref{iso_vsi_beta_crit}) and previous
-% numerical results (\S\ref{therm_relax_eff}), which indicate  
-% higher order modes are more easily stabilized by finite thermal
-% relaxation.   
-% However, for large $\khat$ and/or $z_\mathrm{max}$, we also found
-% surface modes with shorter growth times than those shown in
-% Fig. \ref{mmsn_overall}. Such modes may not be robust because    
-% they are affected by boundary conditions (\S\ref{surf_comment}) and/or
-% viscosity in a real disk.
-
-We therefore consider Fig. \ref{mmsn_overall} as a representation of the
+We can therefore consider Fig. \ref{mmsn_overall} as a representation of the
 VSI in the MMSN. We see the VSI operates from $\sim 5$AU to $\sim
 50$AU with growth timescales $\sim 30$---40 orbits. It is necessary to
-consider smaller scales as $r$ decreases as the disk generally becomes
-optically thick.
+consider smaller scales with decreasing $r$ as the disk generally becomes
+optically thick, and growth times rise rapidly towards the inner disk.
 
-Fig. \ref{mmsn_overall} is roughly consistent with the 
-midplane thermal relaxation timescales for the MMSN ( 
-Fig. \ref{mmsn_bcrit_bcool}, left middle panel). For example, for 
-$\khat=10$ we find $\beta \gtrsim \beta_\mathrm{crit}$ for $r\lesssim
-10$AU at $z=0$; and Fig. \ref{mmsn_overall} show stabilization inside
-$\sim 8$AU where $t_\mathrm{grow}\to\infty$.
+The fact that we obtain the VSI in the inner disk despite $\beta >
+\beta_\mathrm{crit}$ near the midplane suggests that the instability
+can operate provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
+fraction of the vertical domain. However, the normalized growth times
+become much longer. For example, for $\khat=30$ we find
+$t_\mathrm{grow}\sim 40$ orbits at $r=50$AU, but $t_\mathrm{grow}\sim
+400$ orbits at $r=5$AU. Thus, introducing an optically thick midplane
+significantly reduces the efficiency of the VSI despite $\beta <
+\beta_\mathrm{crit}$ away from the midplane. %because of fundamental mode
+
+We thus find Fig. \ref{mmsn_overall} to be roughly consistent with
+comparing the midplane thermal relaxation timescales to the critical
+thermal timescale. For example, 
+Fig. \ref{mmsn_bcrit_bcool} show that $\beta(\khat=10,z=0) \gtrsim
+\beta_\mathrm{crit}$ for $r\lesssim 10$AU in the MMSN (left middle panel); and Fig. \ref{mmsn_overall}
+show that modes with $\khat=10$ are stabilized inside $\sim 8$AU (where
+$t_\mathrm{grow}\to\infty$). Our numerical results suggest the VSI
+is efficient at radii of few tens of AU in the MMSN, consistent with
+estimates made by \cite{nelson13}.
 
@@
 \end{figure}
-
-\subsection{Numerical examples} 
-We compare the VSI in the MMSN at $r=50$AU and $r=5$AU. We first show in
-Fig. \ref{beta_compare}  the thermal timescale $\beta$ for a
-perturbation wavenumber $\khat=30$. For comparison we also plot
-$\beta_\mathrm{crit}$ at these radii.
-
- \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/beta_compare}
-  \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
-     and $r=5$AU for $\khat=30$ (solid lines). The
-    corresponding thin horizontal dashed lines are the critical thermal
-    relaxation timescales derived in linear theory. 
-    \label{beta_compare}}
-\end{figure}
-
-The outer disk ($r=50$AU) is optically thin,
-$\beta<\beta_\mathrm{crit}$  and  does not depend on height. 
-Accordingly we expect, and find, the VSI 
-operates at $r=50$AU. The inner disk ($r=5$AU) is optically
-thick near the midplane and becomes optically thin away from it, such
-that $\beta < \beta_\mathrm{crit}$ for $|z|\gtrsim1.25H$. In
-fact, in this case  we also find the VSI to operate, but with much
-reduced growth rates (when scaled by the local 
-rotation).
-
-The corresponding vertical velocity perturbations are shown in 
-Fig. \ref{mmsn_eigenvz}. Although at $r=5$AU $\beta$ depends on $z$,
-the disturbance still occupies the fluid column due to the vertically
-global nature of the fundamental mode. The fact that
-we obtain the VSI despite $\beta > \beta_\mathrm{crit}$ near the
-midplane suggests that the instability can operate 
-provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
-fraction of the vertical domain. However, the normalized growth times
-are much longer at $r=5$AU than at $r=50$AU: %  for which  $\beta <
-% \beta_\mathrm{crit}  \forall z$: 
-$t_\mathrm{grow}\sim 40$ orbits at $r=50$AU, but $t_\mathrm{grow}\sim
-400$ orbits at $r=5$AU. Thus, introducing an optically thick midplane
-significantly reduces the efficiency of the VSI despite $\beta <
-\beta_\mathrm{crit}$ away from the midplane.
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvectorvz_mmsn_50AU}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
-  0cm]{figures/eigenvectorvz_mmsn_5AU}
-  \caption{Vertical velocity perturbations of the VSI in the MMSN at 
-    $r=50$AU (top) and $r=5$AU (bottom). The real (imaginary) part of 
-    $\delta v_z$ is shown as
-    the solid (dotted) curves. 
-    \label{mmsn_eigenvz}}
-\end{figure}

### 2015-03-15 linear_adia.tex
about to edit numerical

@@
 
-% The thin disk approximation allows us to develop explicit solutions,
-% but some general results may be established without it (see,
-% e.g. Appendix \ref{global_corr} and \S\ref{iso_pert}).
-
 \subsection{Isothermal perturbations}\label{iso_pert}
-Isothermal perturbations correspond to $\beta = 0$ or $\chi = 1$. This
-case has been investigated numerically by \cite{nelson13} and 
-\cite{mcnally14}, the latter of which employed the VGSB framework.
+Isothermal perturbations correspond to $\beta = 0$ or $\gamma=1$ so 
+that $\chi = 1$. This case has been investigated numerically
+by \cite{nelson13} and \cite{mcnally14}.
 
 In Appendix \ref{iso_discuss}, we give an analytical discussion of the
-linear problem with $\beta=0$. There, we establish more general
+linear problem for isothermal perturbations. There, we establish more general
 results that do not require the thin disk approximation, namely 
@@
 and an upper bound to the instability growth rate (it is limited by
-the maximum vertical shear rate in the disk). We also develop analytic
-solutions in the thin disk limit. The methods used in Appendix
-\ref{iso_discuss} and results obtained therein motivates the solution
-procedure employed in the following sections.
+the maximum vertical shear rate in the disk). We also develop 
+solutions in the thin disk limit. The 
+methods used in Appendix \ref{iso_discuss} and results obtained
+therein motivates the solution procedure employed in the following
+sections.
 
@@
 $t_c=\beta\Omega_k^{-1}$ from zero to a small but finite value? For 
-sufficiently small $\beta$ % , for which the perturbations may be 
-% considered nearly-isothermal,
-we expect the solution to only differ slightly from a case with
-$\beta\equiv 0$. We thus perturb a solution for $\beta=0$ to see the
-effect of thermal relaxation.
+sufficiently small $\beta$ we expect the solution to only differ
+slightly from a case with $\beta\equiv 0$. We thus perturb a solution
+for $\beta=0$ to see the effect of thermal relaxation.
 
@@
 and are the ones we find to dominate in numerical calculations with 
-increasing $\beta$, as well as in protoplanetary disks 
+increasing $\beta$ (\S\ref{therm_relax_eff}), as well as in protoplanetary disks 
 with a realistic estimate for thermal timescales
-(\S\ref{application}).  %Their importance is highlighted
+(\S\ref{application}).
 
@@
 
-Inserting Eq. \ref{nearly_iso_pert} into Eq. \ref{nearly_iso_explicit}
-and keeping only first order terms yield the requirements, after
-balancing constants and coefficients of $\zhat^2$,
-
-\begin{align}
-  0 &= 2a + 2\hat{\sigma}\left(1+\khat^2\right)\delta\hat{\sigma} +
-  \left(\hat{\sigma}^2 -1\right)\delta\chi, \\
-  0&= 2a\left(1+\ii\epsilon q \khat\right) - \left(\khat^2 - \ii\epsilon
-    q \khat\right)\delta\chi.
-\end{align}
-
-Using the expressions for $\delta\chi$ and $\hat{\sigma}$ we can solve
-for $\delta\hat{\sigma}$. We find imaginary part of $\delta\hat{\sigma}$
+We insert Eq. \ref{nearly_iso_pert} into
+Eq. \ref{nearly_iso_explicit}, keeping only first order terms, and
+solve for $\delta\hat{\sigma}$ using the above xpressions for $\delta\chi$
+and $\hat{\sigma}$. We find imaginary part of $\delta\hat{\sigma}$
 is 
+% yield the requirements, after
+% balancing constants and coefficients of $\zhat^2$, 
+% \begin{align}
+%   0 &= 2a + 2\hat{\sigma}\left(1+\khat^2\right)\delta\hat{\sigma} +
+%   \left(\hat{\sigma}^2 -1\right)\delta\chi, \\
+%   0&= 2a\left(1+\ii\epsilon q \khat\right) - \left(\khat^2 - \ii\epsilon
+%     q \khat\right)\delta\chi.
+% \end{align}
 % This gives 
@@
 Since $\epsilon \ll 1$, introducing finite cooling $\delta\beta>0$
-implies $\delta\hat{\nu} < 0$, i.e. stabilization. Interestingly, for both
-$\hat{k}\to0$ and $\hat{k}\to\infty$ we find $\delta\hat{\nu}\to0$, so
-stabilization by small finite cooling is ineffective at very large or
-very small scales.
+implies $\delta\hat{\nu} < 0$, i.e. stabilization, since $\gamma>1$. 
+% Interestingly, for both
+% $\hat{k}\to0$ and $\hat{k}\to\infty$ we find $\delta\hat{\nu}\to0$, so
+% stabilization by small finite cooling is ineffective at very large or
+% very small scales.
 
-A finite thermal relaxation time introduces buoyancy
-forces, which is stabilizing for sub-adiabatically stratified disks
-($\gamma > 1$). The $\zhat^2$ dependence in $\delta v_{z1}$ then makes
-sense because it becomes significant for large $|\zhat|$, i.e. away from
+A finite thermal relaxation time allow buoyancy
+forces to stabilize vertical motions in sub-adiabatically stratified
+disks. The $\zhat^2$ dependence in $\delta v_{z1}$ makes  
+sense because it becomes significant at large $|\zhat|$, i.e. away from
 the midplane where the effect of buoyancy first appears as $\beta$ is
-increased from zero.
+increased from zero (since $N_z^2\propto z^2$ for a thin disk, see
+Eq. \ref{nzsq_def}).
 
 \subsection{Explicit solutions and dispersion relation}\label{disp_relax}
-%To obtain a dispersion relation similar to that for 
-%isothermal perturbations (Eq. \ref{sig2_iso}), 
 We now solve Eq. \ref{nearly_iso_explicit} explicitly. We first write  
@@
 \end{align}
-%by considering the case $q=0$ and $\beta\to\infty$ as examined in
-%\cite{lubow93}.
 
@@
 
-% However, for the fundamental mode
-% ($M=0$) some analytic progress can be made, which we consider next.
-
 \subsection{Upper limit on the thermal relaxation timescale for the
@@
 $\beta$ that allows the VSI.  
-%for modes with $\khat^2\gg M+1$.  
-% which, as we will find, correspond to low frequency
-% modes. 
 Our starting point is Eq. \ref{relax_cond1}---\ref{relax_cond2} which
@@
 fundamental mode) or  $M=O(1)$ and low frequency $|\hat{\omega}|\ll
-1$ modes such that $|\hat{\omega}\khat|$ is $O(1)$. We 
+1$ such that $|\hat{\omega}\khat|$ is $O(1)$. We 
 also assume $|q|$ and $\beta$ are $O(1)$ or smaller. For a thin disk
@@
 stability. If $\beta\ll 1$ and the disk is sufficiently thin so that  
-$|\epsilon q \khat |^2\lesssim 1$ (despite $\khat^2\gg 1$), 
+$|\epsilon q \khat |^2\lesssim 1$, 
 then an approximate solution to Eq. \ref{relax_cond_simp1} is 
@@
 \end{align}
-For modes satisfying the above assumptions (or $M=0$) so that the
-second term in square brackets is small, we have 
+Since the second term in square brackets is small by assumption, we have 
 \begin{align}\label{iso_vsi_cond}
@@
 ($M=0$). We comment that the expression for $\beta_\mathrm{crit}$ may
-in fact be obtained without limiting the magnitude of $|\epsilon q   
-\khat|$.  We will confirm Eq. \ref{iso_vsi_cond} numerically in
+in fact be obtained without assuming $|\epsilon q\khat|\lesssim 1$.  We will confirm Eq. \ref{iso_vsi_cond} numerically in 
 \S\ref{bcrit_num_test}.
 
-Eq. \ref{iso_vsi_cond} suggests that as the thermal timescale is
-increased from zero, higher order modes with $M>0$
-are stabilized before the fundamental mode. We indeed observe this
-trend in numerical calculations, despite adopting somewhat restrictive
-assumptions in order to arrive at the simple expression 
-Eq. \ref{bcrit_gen} for $M>0$.
+Eq. \ref{iso_vsi_cond} suggests that as the thermal timescale $\beta$
+is increased from zero, higher order modes with $M>0$
+will be stabilized before the fundamental mode. We indeed observe this
+trend in numerical calculations (\S\ref{therm_relax_eff}), despite
+adopting somewhat restrictive assumptions for $M>0$ in order to arrive
+at the simple expression Eq. \ref{bcrit_gen}.
 

### 2015-03-15 linear_iso.tex
about to edit numerical

@@
 where $\Gamma=1$, then $c_s$ is a constant in the radially-local
-approximation. % We examine isothermal perturbations ($\beta =
-% 0$) before considering the effect of finite thermal relaxation. 
+approximation. 
 In order to facilitate a fully analytic discussion, we first introduce a
@@
 to be derived and compared with previous works.  However, all of the 
-approximations described in
-\S\ref{vgsb_approx}---\S\ref{analytic_relax} are relaxed in our
+approximations described below are relaxed in our
 numerical study.
@@
 is an extension of the standard shearing box \citep{goldreich65} to 
-allow the shear flow in the shearing box to be height-dependent. 
-The difference between the radially local
-approximation and the VGSB is that the equilibrium disk
-structure is assumed to be homogeneous in the radial ($x$) direction,
-as in the 
-standard shearing box.
+background shear flows that are height dependent. 
+As in the  standard shearing box, in the VGSB the equilibrium disk
+structure is assumed to be homogeneous in the radial ($x$) direction.
 
@@
 After setting $\hat{g}_c=0$ in Eq. \ref{lin_mass}, \ref{lin_vx} and
-\ref{lin_energy}, we can eliminate variables % eliminate the perturbed velocities between
-% the linearized equations 
+\ref{lin_energy}, we can eliminate variables 
 to obtain a pair or ordinary differential
@@
 limitations. This is because within the VGSB framework, $d\Omega/dz$
-can be freely specified. %  In effect, by adopting the VGSB framework, we are studying
-% dynamics in the standard shearing box subject to an additional
-% azimuthal force $\delta v_zr d \Omega/dz$ with $d\Omega/dz$ as an
-% input parameter. 
-%For consistency with the global disk, 
+can be freely specified. 
 We choose $d\Omega/dz$ to be that due to the thermal structure in 
@@
 
-%the above inconsistency
-
 We further discuss caveats of the VGSB in Appendix
-\ref{global_corr}. We find that for perturbations with small radial
-lengthscales ($k_xH\gg 1$) and fast thermal relaxation ($\beta\ll1$),
-which characterizes the VSI, the VGSB framework is adequate for the 
-purpose of a simple analytical treatment.
+\ref{global_corr}. We find that for fast thermal relaxation
+($\beta\ll1$) and perturbations with small radial
+lengthscales ($k_xH\gg 1$), which characterizes the VSI, the VGSB
+framework is adequate for the purpose of a simple analytical
+treatment.
 
@@
   $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
-  \kappa^2$. % The problem simplifies because the eigenvalue now only
-  % appears in the last term in Eq. \ref{iso_ode}. %stoll kley see low
-  % freq 
-  We call this the low-frequency approximation. 
+  \kappa^2$. %\citep[e.g.][]{lin10}. 
+  We call this the low frequency approximation. %physically inertial
+                                %waves 
 \item We are mainly interested in the effect of vertical shear, i.e. $q\neq
@@
 \end{align}
-where $^\prime$ denotes $d/d\zhat$.%  Eq. \ref{vertiso_gov_nondim} will
-% be the basis for our discussion of vertically isothermal disks.
+where $^\prime$ denotes $d/d\zhat$.
 
@@
 $\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
-unnecessary complexity for the present discussion. % However, this
-% approximation will be relaxed in our numerical study.
-
-% Note that we can generalize Eq. \ref{vertiso_gov_nondim} non-Keplerian
-% rotation profiles by redefining $\khat$ as $k_xc_s/\kappa$,
-% $\hat{\sigma}$ as $\sigma/\kappa$, and  $\epsilon$ as $c_s/r\kappa$.
-
-% This expression can be simplified in the limit $|q|\ll 1$, 
-%  \begin{align}\label{simple_growth}
-%    \frac{\imag(\sigma)}{\Omega_k} = \pm \sqrt{M} 
-%    \frac{q\epsilon\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
-%  \end{align}
-%  with $M$ being an integer.
-
-% In Appendix \ref{pert_theory} we give an alternative method to
-% infer instability in the presence of vertical shear without explicitly
-% solving the governing equation.
+unnecessary complexity for the present discussion.

### 2015-03-15 numerical.tex
about to edit numerical

@@
 We first calculate the VSI in a disk with rapid thermal relaxation by
-setting $\beta=10^{-3}$. (Smaller values yield the similar results.) This
+setting $\beta=10^{-3}$. (Smaller values yield similar results.) This
 is similar to strictly isothermal perturbations in a vertically
@@
   \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_analytic.ps}
-  \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for $\khat=30$
-    \label{compare_modes_iso_kx30}. Examples of surface modes are
-    marked by `s'. 
+  \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for $\khat=30$. Examples of surface modes are
+    marked by `s'. \label{compare_modes_iso_kx30}
   }
@@
     relaxation timescales $\beta=0.01$ (black dots), $\beta=0.03$ (red
-    crosses), $\beta=0.05$ (green triangles) and $\beta=0.1$ (blue
+    crosses), $\beta=0.05$ (blue triangles) and $\beta=0.1$ (green 
     crosses); for different vertical domain sizes $\zmax=3H$ (top),
-    $\zmax=5H$ (middle, our fiducial setup) and $\zmax=7H$
-    (bottom). Lines are computed analytically from
+    $\zmax=5H$ (middle, the fiducial setup) and $\zmax=7H$
+    (bottom). Lines are computed from
     Eq. \ref{relax_disp} for  $M=0$ (solid, fundamental mode) and
@@
 modes with $\khat=30$. For clarity we only plot analytical results for
-the fundamental mode, which is again robust against boundary
-conditions compared to higher order modes. We find surface modes
+the fundamental mode, which is again insensitive boundary
+conditions compared to higher order modes and thus likely more
+robust in realistic disks. We find surface modes
 appear for $\zmax\geq5H$. However, they are effectively stabilized with
@@
 eigenvalues in the lower two panels). However, these modes are clearly associated
-with the vertical domain size as they are absent for $\zmax=3H$. 
+with the boundary condition as they are absent for $\zmax=3H$. 
 % Thus, for $\zmax=5H$ and  
@@
 approaches $\beta_\mathrm{crit}$. %beta<beta_crit is necessary and sufficient  
-Thus, tracking the fundamental mode is a more robust way of 
-assessing stability as $\beta$ is increased from zero, rather than
+Thus, tracking the fundamental mode is a more robust assessment of
+stability as $\beta$ is increased from zero, rather than 
 searching for the most unstable mode, which may correspond to modes

### 2015-03-15 paper.tex
about to edit numerical

--- 2015-03-04 paper.tex
@@
 \newcommand{\w}{ \widetilde{W}}
-%\newcommand{\sbar}{\bar{\sigma}}
 \newcommand{\dphi}{\dd \Phi}
@@
 
-%\title{Linear analysis of the vertical shear instability in astrophysical disks}
 \title{Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula} 
@@
   recently been suggested to provide such a pathway, since it operates in
-  disks with rotation profiles that depends on the height from the 
+  disks with rotation profiles that depend on the height from the 
   disk midplane, which is generally expected in  
   astrophysical disks. However, the VSI requires a short thermal
-  timescale in stably stratified disks. %which is usually assumed 
+  timescale in stably stratified disks.  
   A quantitative examination of this thermodynamic 
@@
   the VSI in realistic disks. To this end,  
-  we generalize the linear, axisymmetric VSI in radially local but
-  vertically global models of accretion disks to include a thermal
-  relaxation term that restores the disk  
-  temperature $T$ over a characteristic timescale
+  we study the linear, axisymmetric VSI in radially local, 
+  vertically global models of accretion disks in which the disk  
+  temperature $T$ is relaxed its initial value over a characteristic timescale 
   $t_c=\beta\Omega_k^{-1}$, where $\beta$ is a dimensionless parameter 
@@
   varying $\beta$. We find increasing $\beta$ even slightly away from 
-  zero can dramatically reduce VSI growth rates. % , since this generally
-  % introduces a vertical entropy gradient that stabilizes vertical
-  % motions as needed for the VSI.
-  Our  linear calculations are consistent with previous hydrodynamic
-  simulations. For vertically isothermal disks, we quantify the
-  requirement for `fast' thermal relaxation as $\beta < \epsilon|q|/(\gamma-1)$, where
+  zero dramatically reduce VSI growth rates. 
+  For vertically isothermal disks, we quantify the
+  `short' thermal timescale requirement for the VSI as $\beta < \epsilon|q|/(\gamma-1)$, where
   $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  ($r$
   being the cylindrical radius from the central object), and 
-  $\gamma$ is the adiabatic index. We apply our linear calculations 
+  $\gamma$ is the adiabatic index. 
+  We apply our linear calculations  
   to a disk model of the Minimum Mass Solar Nebula (MMSN) with an  
   estimate for the thermal timescale $\beta$ in the MMSN based on dust
-  opacity. We find the MMSN is unstable to the VSI at tens of AU with 
+  opacity. We find the MMSN is unstable to the VSI from $\sim 5$AU to $\sim 50$AU with  
   a characteristic growth time of $\sim 30$ local orbital 
-  periods, and the dominant disturbance in the MMSN is the `corrugation' mode
-  which casues the disk column to oscillate vertically. 
+  periods, and the dominant VSI mode in the MMSN is the `fundamental
+  corrugation mode' which casues the disk column to oscillate
+  vertically. We suggest the VSI is dynamically imporant in the outer
+  parts of protoplanetary disks. 
 \end{abstract}
@@
 \section{Introduction}
-% stuff about hydrodynamic angular momentum transport 
 Understanding the mechanisms of angular momentum transport is the
@@
 \citep{gammie96,landry13}. In fact, recent  numerical simulations
-suggest portions of PPDs remain laminar \citep[e.g.][and references
-therein]{lesur14,bai15,gressel15}. In this respect, alternative routes 
+suggest significant portions of PPDs remain laminar \citep[e.g.][and references
+therein]{lesur14,bai15,gressel15}. Thus, alternative routes 
 to angular momentum transport in PPDs remain of interest
@@
 
-%recent work on vsi 
 One possibility is purely hydrodynamic turbulence through the vertical
@@
 in three-dimensional (3D) disks in which the rotation frequency
-$\Omega$ depends on the height $z$ from the disk midplane. This dependence
-occurs when the equilibrium pressure depends on both the temperature
-and density, making the disk baroclinic. It is thus generally
-expected that $\p_z\Omega\neq 0$ in astrophysical disks.
+$\Omega$ depends on the height $z$ from the disk midplane. This
+dependence occurs when the equilibrium pressure depends on both the
+temperature and density, making the disk baroclinic. It is therefore
+generally expected that $\p_z\Omega\neq 0$ in astrophysical disks.
 
@@
 perturbations with large vertical lengthscales (relative to radial 
-lengthscales) to become unstable.    
-\cite{umurhan13} gives a simple interpretation for the VSI: an
-exchange of fluid parcels while conserving their angular momenta
-individually, may result in a new configuration of lower energy, if
-the exchange occurs over a sufficiently large vertical displacement.
-
+lengthscales) to become unstable. \cite{umurhan13} gives a simple
+interpretation for the VSI: an exchange of fluid parcels while
+conserving their angular momenta individually, may result in a new
+configuration of lower energy, if the exchange occurs over a
+sufficiently large vertical displacement. In practice, the VSI
+corresponds to unstable inertial waves.  
+%inertial waves 
 %history
@@
 VSI have been performed by \cite{arlt04}, and more recently by 
-\cite{nelson13} and \cite{stoll14}. In these simulations the authors
-indeed observe the growth of vertically-global but radially-localized
+\cite{nelson13} and \cite{stoll14}. In these new simulations the authors
+indeed observe the growth of vertically global but radially localized
 perturbations when the equilibrium rotation varied with height.
@@
 isothermal, unless the disk is neutrally stratified in the vertical
-direction. This is because vertical motions are stabilized by
-buoyancy if the disk is sub-adiabatically stratified, as required for
+direction. This is because vertical motions, required for the VSI, are stabilized by
+buoyancy if the disk is sub-adiabatically stratified for
 convective stability. When the thermal timescale $t_c$ is parametrized
@@
 
-%why linear approach? computationally fast (need very high resolution) 
-%can be used as a first estimate
+From these studies it is clear that the VSI is sensitive to
+the thermodynamic response of the disk if it is stably
+stratified. Since the latter is usually assumed for typical disk models,  
+it is important to understand and quantify said thermodynamic dependence,
+for example on disk parameters, in order to evaluate the relevance of
+the VSI in realistic disks. Should a given disk become turbulent
+through the VSI, it must first pass through the
+linear instability. Thus, understanding the thermodynamic dependence
+of the linear VSI is essential.
 
-From these studies it is clear that the VSI is generally sensitive to
-the thermodynamic response of the disk if it is stably
-stratified. Since the latter is usually
-assumed for typical disk models,  
-% disk thermodynamics except
-% in the special case of neutrally stratified disks. 
-it is 
-important to understand and quantify said thermodynamic dependence,
-for example on disk parameters, in order to evaluate the viability of
-the VSI in realistic disks. % Numerical simulations of the VSI are
-% challenging, however, because global disk simulations should be
-% carried out while resolving radial scales much smaller than the local
-% disk scale height.    
-%On the other hand, 
-Should a given disk become turbulent through the VSI, it must pass through the
-linear instability first. Thus, understanding the thermodynamic dependence
-of the linear VSI is essential. % Furthermore, linear
-% stability calculations are essential to complement future numerical simulations.
-
-%what we're doing
-In this work we study the linear VSI in radially-local,
-vertically-global models of astrophysical disks. We include an energy
+In this work we study the linear VSI in radially local,
+vertically global models of astrophysical disks. We include an energy
 equation with a source term that allows the thermodynamic response of
@@
 VSI carried out by \cite{nelson13} and \cite{mcnally14} which assume
-isothermal perturbations. Our models are also an extension of earlier 
+isothermal perturbations and therefore did not include an energy
+equation. Our models are also an extension of earlier 
 analysis by \cite{urpin98} and \cite{urpin03}, which accounts for thermal
@@
 
-%paper organization
-
-%say "grossly simplified
-
 This paper is organized as follows. In \S\ref{setup} we list the
@@
 consideration. We formulate the linear stability problem in
-\S\ref{linear} and simplify it by making the radially-local
+\S\ref{linear} and simplify it by making the radially local
 approximation. We study the linear problem using two approaches. In
@@
 allows us to highlight the essential effects of vertical shear. In
-particular, we quantify a characteristic thermal timescale below which
-the VSI is expected to operate. In \S\ref{numerical} we numerically
-solve the linearized fluid equations in the radially-local
-approximation, but relax all other assumptions made in our analytical approach.   
+particular, we quantify a characteristic upper limit to the thermal
+timescale for the VSI to operate. In \S\ref{numerical} we numerically
+solve the linear problem in the radially local approximation, but
+relax all other approximations made in our analytical approach.    
 We find basic agreement with our analysis, including
@@
 \S\ref{application} we apply our linear models to the Minimum Mass
-Solar Nebulae (MMSN) by modeling thermal relaxation
+Solar Nebula (MMSN) by modeling thermal relaxation
 timescales based on realistic disk properties, and find that the VSI may
-operate in the outer parts of the MMSN. We summarize and discuss our
-results in \S\ref{summary}, including caveats and possible directions
+operate in the outer parts of the MMSN. We summarize in
+\S\ref{summary} with a discussion of caveats and possible directions
 for future work.
@@
 
-%\acknowledgements
 \appendix

### 2015-03-15 setup.tex
added convective overstability discussion
about to edit numerical

@@
 \S\ref{application} we will model $\beta$ based on physical thermal
-timescales, in which case $\beta$ is determined by the equilibrium
-disk properties and the perturbation lengthscales of interest.
+timescales, in which case $\beta$ is non-constant and is determined by
+the equilibrium disk properties and the perturbation lengthscales of
+interest.
 
@@
 where the polytropic index $\Gamma$ is a  constant parameter  and the
-function $c_{s0}(r)$ is prescribed below. Choosing $\Gamma=1$ gives 
+function $c_{s0}(r)$ is prescribed below. Choosing $\Gamma=1$ gives  a 
 locally vertically isothermal disk and $c_{s0}(r)$ is the mid-plane
-sound-speed. %For $\Gamma=\gamma$ the disk is adiabatically stratified
-% in the vertical 
+sound-speed. 
 It shall be convenient to define the function $c_s(r,z)$ 
@@
 is also a characteristic disk aspect-ratio. We are 
-interested in thin disks with $\epsilon \ll 1$.%, as relevant for protoplanetary disks \citep{chiang97}.
+interested in thin disks with $\epsilon \ll 1$.%,%  as relevant for
+% protoplanetary disks \citep{chiang97}.
 
@@
   N_z^2 \equiv - c_p^{-1}g_z \frac{\p S}{\p z} = c_s^2 \left(\frac{\gamma -
-      \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   
+      \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   \label{nzsq_def}
 \end{align}
@@
 \end{align}
-Evaluating this expression explicitly,
+For the above disk models,
 \begin{align}
@@
 \citep{tassoul78}, where $j=r^2\Omega$ is the specific angular
-momentum. For rotationally-supported, thin disks, pressure gradients are small 
-compared to rotation, so the first criterion is generally
-satisfied. Thus, to assess stability, we consider the second
+momentum. For rotationally-supported, thin disks %  pressure gradients are small 
+% compared to rotation, so 
+the first criterion is generally satisfied. 
+Thus, to assess stability, we consider the second
 criterion. For our disk models,
@@
   \ii\sigma\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
-    z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\\
+    z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
   \ii\sigma\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
@@
 structure and radial gradients. We thus evaluate all coefficients in
-the above equations at the fiducial radius $r=r_0$. However, the full
-vertical dependence of the coefficients are retained. Then we can
-assume perturbations have spatial dependence in the form
+Eq. \ref{full_mass1}---\ref{full_energy1} at the fiducial radius
+$r=r_0$. However, the full vertical dependence of the coefficients are
+retained. Then we can assume perturbations have spatial dependence in
+the form 
 \begin{align}
@@
 
-In order to emphasize the radially-local approximation, we
+In order to emphasize the radially local approximation, we
 introduce the local Cartesian co-ordinates $(x,y,z)$ corresponding to
@@
 we will make further simplifications.
-
-% The radially-local approximation above is similar in spirit to the
-% shearing box framework \citep{goldreich65} or the generalization
-% thereof \citep{mcnally14}. The main difference being the terms
-% proportional to $\hat{g}_c$ are neglected in the shearing box.

### 2015-03-15 summary.tex
about to edit numerical

@@
 
-\subsubsection{Parameter study and convective overstability}
+\subsubsection{Convective overstability}
 We have focused on parameter regimes relevant to the VSI,
@@
 \end{align}
-% We see that convective overstability is generally slower 
-% than the VSI because $\nu_\mathrm{conv,max} = O(\epsilon^2\Omega_k)$,
-% whereas the maximum VSI growth rate is $O(\epsilon\Omega_k)$. 
-In fact, for the MMSN considered in this work (\S\ref{application} and Appendix
-\ref{mmsn}), convective overstability does not apply because
+In fact, for the MMSN considered in this work, convective
+overstability does not apply because 
 $N_r^2>0$.
 
-However, for more general disk structures convective overstability is
+For more general disk structures convective overstability is
 possible. The models adopted for our numerical study in
@@
 appropriate.
-% We will explore convective overstability more fully in a
-% follow-up study.
 
@@
 application to protoplanetary disks we connected $t_c$ to the disk
-properties. One one can improve the realism further by generalizing the 
+properties. One can improve the realism further by generalizing the 
 opacity model adopted here to be density (and hence height) dependent. 
@@
 However, going beyond thermal relaxation requires a
-proper treatment for radiative diffusion. The linear problem then
+proper treatment of radiative diffusion. The linear problem then
 involves a Laplacian operator, which require additional boundary

### 2015-03-16 _region_.tex
minor updates
revised

--- 2014-12-17 linear_iso.tex
@@
-\section{Analytical discussion of vertically isothermal disks} 
-We begin with a discussion of vertically isothermal disks
-where $\Gamma=1$, then $c_s$ is a constant in the shearing box. We
-examine separately the limiting case of isothermal perturbations
-($\gamma\equiv1$) and adiabatic perturbations ($\gamma\neq 1$ and
-$t_c\to\infty$).   
-\subsection{Isothermal perturbations}
-The linear problem simplifies considerably for the case
-$\Gamma=\gamma=1$. Then 
+\message{ !name(linear_adia.tex)}
+\message{ !name(linear_adia.tex) !offset(-2) }
+\subsection{Thin-disk approximation}\label{analytic_relax}
+Our final approximation apply to thin disks for 
+which $\epsilon\ll 1$. Then we may write $\rho/\rho_0\simeq
+\exp{(-\zhat^2/2)}$. The governing equation,
+Eq. \ref{vertiso_gov_nondim}, is then   
 \begin{align}
-  Q=W. 
+  \delta v_z^{\prime\prime} - z A\delta v_z^\prime +
+  (B - C\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
 \end{align}
-Inserting this into Eq. \ref{ode_w} gives a single equation for the linear stability
-problem for isothermal perturbations in a vertically isothermal disk, 
-\begin{align}\label{iso_ode}
-  \frac{d^2W}{dz^2} + \left(\frac{d\ln{\rho}}{dz} - \frac{\ii k_x
-      r}{D}\frac{d\Omega^2}{dz}\right) \frac{dW}{dz} +
-  \sigma^2\left(\frac{1}{c_s^2} + \frac{k_x^2}{D}\right)W=0, 
+with
+\begin{align}
+  &A \equiv 1 + \ii \epsilon q \hat{k},\\
+  &B \equiv \hat{\sigma}^2\left(\chi + \hat{k}^2\right) -
+  \left(\chi + \ii \epsilon q \hat{k}\right),\\
+  &C \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
+    \epsilon q\hat{k}\right), 
 \end{align}
-subject to appropriate boundary conditions. We remark that Eq
-.\ref{iso_ode} is also applicable for    
-$\gamma=\Gamma\neq 1$ in the limit $t_c\to\infty$ (i.e. neutrally
-stratified adiabatic disks), since in that case  
-$Q\simeq W$ as well. However, the following discussion is only
-valid for $\gamma=\Gamma=1$ because $c_s$ is taken to be a constant.
+and $\chi =
+\left(1-\ii\hat{\sigma}\beta\right)/\left(1-\ii\hat{\sigma}\beta\gamma\right)
+$ in terms of dimensionless variables.
 
-\subsubsection{Approximations}
-To facilitate an analytical discussion, we simplify the governing
-equation Eq. \ref{iso_ode} with the following approximations:
+We remark that Eq. \ref{nearly_iso_explicit} is in the same 
+form as Eq. 41 in \cite{lubow93}, which describe adiabatic axisymmetric waves in
+a vertically isothermal disk without vertical shear. The two equations
+become identical after  setting $q=0$ (no vertical shear) and
+$\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
+and making the low-frequency, nearly-Keplerian approximation in
+\citeauthor{lubow93} after a change of variables.
 
-\begin{enumerate}
-\item We consider eigenvalues such that
-  $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
-  \kappa^2$. The problem simplifies because the eigenvalue now only
-  appears in the last term in Eq. \ref{iso_ode}. 
-\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
-  0$. This is already captured explicitly in Eq. \ref{iso_ode} through
-  the factor $d\Omega^2/dz$. We therefore simplify further by ignoring
-  the vertical dependence of $\kappa^2$ elsewhere. For clarity, we
-  will replace $\kappa^2$ with $\Omega_k^2$.%, although the discussion
-  % is unchanged if we replaced 
-  % $\kappa^2$ with $\kappa^2(0)$, for example.    
-\end{enumerate}
-These assumptions amount to making the replacement $D\to\Omega_k^2$.  
-We will consider sound-speed profiles with $q<0$ in the global 
-disk. Then $\Omega^2(z)$ decreases away from the midplane,
-and may reach small values for sufficiently large $|z|$, which
-invalidates the low-frequency approximation. In principle, then, the
-above approximations implicitly limits us to consider thin disks ($\epsilon
-\ll 1$) with vertical domain sizes not too large, so that the rotation
-is nearly-Keplerian.
+\subsection{Isothermal perturbations}\label{iso_pert}
+Isothermal perturbations correspond to $\beta = 0$ or $\gamma=1$ so 
+that $\chi = 1$. This case has been investigated numerically
+by \cite{nelson13} and \cite{mcnally14}.
 
-\subsubsection{Neccessity of vertical shear for
-  instability}\label{integral_relation} 
-With the above approximations and introducing the dimensionless variables
+In Appendix \ref{iso_discuss}, we give an analytical discussion of the
+linear problem for isothermal perturbations. There, we establish more general
+results that do not require the thin disk approximation, namely 
+a necessary condition for instability (there must be vertical shear)
+and an upper bound to the instability growth rate (it is limited by
+the maximum vertical shear rate in the disk). We also develop 
+solutions in the thin disk limit. The 
+methods used in Appendix \ref{iso_discuss} and results obtained
+therein motivates the solution procedure employed in the following
+sections.
+
+\subsection{The effect of introducing a small but finite
+  thermal relaxation time}\label{relax_pert}
+We are interested in the case $\beta\neq 0$, but it is instructive to
+first ask  the more analytically tractable question: how 
+do the eigenfrequencies and eigenfunctions change when we change
+$t_c=\beta\Omega_k^{-1}$ from zero to a small but finite value? For 
+sufficiently small $\beta$ we expect the solution to only differ
+slightly from a case with $\beta\equiv 0$. We thus perturb a solution
+for $\beta=0$ to see the effect of thermal relaxation.
+
+For definiteness, let us consider the simple solution 
 \begin{align}
-  \hat{z} = z/\Hiso,\quad \hat{k}=k_x\Hiso, \quad \hat{\sigma} = \sigma/\Omega_k,
+  \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\sigma}^2 = \frac{1 +
+  \ii\epsilon q \hat{k}}{1+\hat{k}^2}, \label{pert_basic} 
 \end{align}
-the governing equation can be re-written as
+which solves Eq. \ref{nearly_iso_explicit} since $\delta
+v_z=$constant and $\chi=1$ so that $B=C=0$. 
+This is the fundamental or `corrugation' mode where the vertical
+velocity is constant throughout the disk column. These modes have been
+observed to dominate numerical simulations \citep{nelson13,stoll14},  
+and are the ones we find to dominate in numerical calculations with 
+increasing $\beta$ (\S\ref{therm_relax_eff}), as well as in protoplanetary disks 
+with a realistic estimate for thermal timescales
+(\S\ref{application}).
+
+We linearize Eq. \ref{nearly_iso_explicit} about the above
+solution for $\beta\equiv0$ and write 
+\begin{align}\label{nearly_iso_pert}
+  \beta \to 0 + \delta\beta,\, \delta v_z\to \delta v_z+\delta
+  v_{z1},\,\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}, 
+\end{align}
+which implies 
 \begin{align}
-  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
-    f(\zhat)\right]W^\prime + \hat{\sigma}^2\left(1+\hat{k}^2\right)W,\label{iso_ode1}
+  \chi \to 1 + \delta\chi = 1 + \ii \hat{\sigma}\left(\gamma-1\right)\delta\beta,
 \end{align}
-where the primes denote $d/d\zhat$ and $f(\zhat)$ is defined such that
-\begin{align}\label{fz_shear}
-  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
+with $\delta v_z$ and $\hat{\sigma}$ given by Eq. \ref{pert_basic}. We
+may then seek  
+\begin{align}
+  \delta v_{z1} = a \zhat^2 + b,
+\end{align}
+where $a$, $b$ are constants.
+
+We insert Eq. \ref{nearly_iso_pert} into
+Eq. \ref{nearly_iso_explicit}, keeping only first order terms, and
+solve for $\delta\hat{\sigma}$ using the above xpressions for $\delta\chi$
+and $\hat{\sigma}$. We find imaginary part of $\delta\hat{\sigma}$
+is 
+% yield the requirements, after
+% balancing constants and coefficients of $\zhat^2$, 
+% \begin{align}
+%   0 &= 2a + 2\hat{\sigma}\left(1+\khat^2\right)\delta\hat{\sigma} +
+%   \left(\hat{\sigma}^2 -1\right)\delta\chi, \\
+%   0&= 2a\left(1+\ii\epsilon q \khat\right) - \left(\khat^2 - \ii\epsilon
+%     q \khat\right)\delta\chi.
+% \end{align}
+% This gives 
+% \begin{align}\label{nearly_iso_dsig}
+%   \delta \hat{\sigma} =
+%   -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii\epsilon
+%       q - \hat{k}\right)^2}{2\left(1+\ii\epsilon q \hat{k}\right)\left(1+\hat{k}^2\right)^2}\delta\beta.
+% \end{align}
+% The imaginary part of Eq. \ref{nearly_iso_dsig} is
+\begin{align}
+  \delta\hat{\nu} =
+  -\frac{\left(\gamma-1\right)\hat{k}^2 \left(\hat{k}^2 -
+      2\epsilon^2q^2\hat{k}^2 - \epsilon^2q^2\right)}{2\left(1+\epsilon^2 q^2
+      \hat{k}^2\right)\left(1+\hat{k}^2\right)^2}\delta\beta.  
+\end{align}
+Since $\epsilon \ll 1$, introducing finite cooling $\delta\beta>0$
+implies $\delta\hat{\nu} < 0$, i.e. stabilization, since $\gamma>1$. 
+% Interestingly, for both
+% $\hat{k}\to0$ and $\hat{k}\to\infty$ we find $\delta\hat{\nu}\to0$, so
+% stabilization by small finite cooling is ineffective at very large or
+% very small scales.
+
+A finite thermal relaxation time allow buoyancy
+forces to stabilize vertical motions in sub-adiabatically stratified
+disks. The $\zhat^2$ dependence in $\delta v_{z1}$ makes  
+sense because it becomes significant at large $|\zhat|$, i.e. away from
+the midplane where the effect of buoyancy first appears as $\beta$ is
+increased from zero (since $N_z^2\propto z^2$ for a thin disk, see
+Eq. \ref{nzsq_def}).
+
+\subsection{Explicit solutions and dispersion relation}\label{disp_relax}
+We now solve Eq. \ref{nearly_iso_explicit} explicitly. We first write  
+\begin{align}
+  \delta v_z(\zhat) =
+  g(\zhat)\exp\left(\frac{\alpha\zhat^2}{2}\right), \label{adia_ansatz}
+\end{align}
+where $\alpha$ is a constant to be chosen for convenience. Inserting
+Eq. \ref{adia_ansatz} into Eq. \ref{nearly_iso_explicit} gives
+\begin{align}
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
+    \alpha\right)g
+  +\left(\alpha^2 - \alpha A - C\right)\zhat^2 g.
+\end{align}
+We choose $\alpha$ to make the coefficient of $\zhat^2g$
+vanish, and impose the vertical kinetic energy density
+$\rho|\delta v_z|^2\propto |g|^2 \exp{\left(\real\alpha -
+    1/2\right)\zhat^2}$ to remain finite as $|\zhat|\to\infty$. 
+Then assuming $g(\zhat)$ is a polynomial, we require  
+\begin{align}
+  \real\alpha < \frac{1}{2}, 
+\end{align}
+which amounts to choosing 
+\begin{align}
+  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right).  
 \end{align}
 
-We can establish a necessary condition for instability by multipling
-Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
-$\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
-terms when integrating by parts, by assuming $W$ or
-$W^\prime$ vanishes at the boundaries, or that the boundaries are 
-sufficiently far away so that the boundary terms are negligible because of the
-decaying background density with increasing height. Then,
+Eq. \ref{nearly_iso_explicit} becomes 
 \begin{align}
-  &\hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat \notag\\
-  &=\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
-  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime +
+  \left(B + \alpha\right)g.
 \end{align}
-It follows that for instability ($\imag\hat{\sigma}>0$), it is neccessary to
-have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical shear.
+We seek polynomial solutions 
+\begin{align}
+  g(\zhat) = \sum_{m=0}^M b_m \zhat^m,
+\end{align}
+which requires
+\begin{align}
+  B(\hat{\sigma}) + \alpha(\hat{\sigma}) =
+  M\left[A-2\alpha(\hat{\sigma})\right].\label{adia_disp0} 
+\end{align}
+For ease of analysis, we re-arrange Eq. \ref{adia_disp0} and square it
+to obtain a polynomial in $\hat{\sigma}$,  
+\begin{align}
+  0 = \sum_{l=0}^{6}c_l\hat{\sigma}^l,\label{relax_disp}
+\end{align}
+where the coefficients $c_l$ are given in Appendix \ref{relax_coeff}.
+The dispersion relation $\hat{\sigma}=\hat{\sigma}(\khat)$ is
+complicated and generally requires a numerical solution. However,
+simple solutions may be obtained in limiting cases which we consider
+below.
 
-\subsubsection{Maximum growth rate}   
-Here we show that the growth rate is limited by the maximum vertical
-shear in the domain. Writing $\hat{\sigma} = \hat{\omega} +
-\ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, the real and
-imaginary parts of 
-Eq. \ref{integral_relation1} are
+\subsection{Upper limit on the thermal relaxation timescale for the
+  VSI}\label{iso_vsi_beta_crit}
+Here we estimate the maximum thermal relaxation timescale 
+$\beta$ that allows the VSI.  
+Our starting point is Eq. \ref{relax_cond1}---\ref{relax_cond2} which
+approximates the dispersion relation Eq. \ref{relax_disp} assuming marginal
+stability and $\khat^2\gg 1$. We consider $M=0$ (the 
+fundamental mode) or  $M=O(1)$ and low frequency $|\hat{\omega}|\ll
+1$ such that $|\hat{\omega}\khat|$ is $O(1)$. We 
+also assume $|q|$ and $\beta$ are $O(1)$ or smaller. For a thin disk
+($\epsilon \ll 1$) we have 
 \begin{align}
-  &\left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
-  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
-  \int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\hat{z}
-  \notag\\
-  &=\real\left[\ii
-    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\zhat\right], \notag\\
-  & 2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
-  \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}\notag\\
-  &=\imag\left[\ii
-    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right].
+ 0 = &M(M+1)\left(1-\epsilon^2q^2\khat^2\right) \notag\\ 
+ &+ \left[1 +
+    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
+   -  \left(\hat{\omega}\khat\right)^4,\label{relax_cond_simp1}\\
+   0= &2M(M+1) + \frac{\beta}{\epsilon q} (1-\gamma)(1+2M)^2
+   \left(\hat{\omega}\khat\right) + \left(\hat{\omega}\khat\right)^2. \label{relax_cond_simp2}
 \end{align}
-Adding the square of these equations give
+(We explain this in Appendix \ref{disp_neut_limit}.)
+
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} is a pair of
+simultaneous equations for $\hat{\omega}$ and $\beta$ for marginal
+stability. If $\beta\ll 1$ and the disk is sufficiently thin so that  
+$|\epsilon q \khat |^2\lesssim 1$, 
+then an approximate solution to Eq. \ref{relax_cond_simp1} is 
 \begin{align}
-  &\left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
-    \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
-    \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2\notag\\
-  &+4\hat{\nu}^2\left(1+\hat{k}^2\right) 
-  \int_{\zhat_1}^{\zhat_2}\rho
-  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
-  &=\left|\ii
-    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
+  \left(\hat{\omega}\khat\right)^2 \simeq (M+1)\left(1 -
+    \frac{M|\epsilon q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
 \end{align}
-It is clear that
-\begin{align}\label{sigma_finite_domain} 
-  &4\hat{\nu}^2\left(1+\hat{k}^2\right) 
-  \int_{\zhat_1}^{\zhat_2}\rho
-  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\notag\\
-  &\leq\left|
-    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
+Note that for $q=0$, Eq. \ref{marg_freq_approx} reduces to 
+the dispersion relation for low-frequency waves in vertically
+isothermal disks without vertical shear (see Appendix
+\ref{stable_novshear}). Inserting Eq. \ref{marg_freq_approx} into
+Eq. \ref{relax_cond_simp2}, we obtain 
+\begin{align}\label{bcrit_gen}
+  \beta &= \frac{\epsilon|q|}{(\gamma - 1)}
+  \frac{\sqrt{1+M}}{(1+2M)}\left[1 + \frac{M(2M-1)|\epsilon q
+      \khat|^2}{2(2M+1)^2}\right].
+%  &\lesssim \beta_\mathrm{crit}\notag
 \end{align}
-On the left hand side of this inequality, we apply the Cauchy-Schwarz
-inequality to obtain
-\begin{align}
-  &\left( \int_{\zhat_1}^{\zhat_2}\rho
-    |W|\left|W^\prime \right|d\hat{z}\right)^2\leq
-  \int_{\zhat_1}^{\zhat_2}\rho 
-  |W|^2d\hat{z}\int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}.
+Since $|\epsilon q \khat|^2\lesssim 1$ by assumption, we have  
+\begin{align}\label{iso_vsi_cond}
+  \beta\leq \beta_\mathrm{crit}  \equiv
+  \frac{\epsilon|q|}{\gamma-1}, 
 \end{align}
-On the right hand side of Eq. \ref{sigma_finite_domain} we have
-\begin{align}
-  \left|\int_{\zhat_1}^{\zhat_2}\rho
-    f(\hat{z}) W^*W^\prime d\hat{z}\right|\leq \int_{\zhat_1}^{\zhat_2}\rho
-  \left|f(\hat{z})W^*W^\prime \right|d\hat{z} \notag\\
-  \leq
-  \mathrm{max}\left(|f|\right)\int_{\zhat_1}^{\zhat_2}\rho
-  |W|\left|W^\prime \right|d\hat{z},
-\end{align}
-where $\mathrm{max}(|f|)$ is the maximum value of $|f|$ in
-$\zhat\in[\zhat_1,\zhat_2]$. Inserting these inequalities into
-Eq. \ref{sigma_finite_domain} gives
-\begin{align}\label{max_growth}
-  |\hat{\nu}|\leq
-  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
-\end{align}
-Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
-follows that the maximum possible growth rate of unstable modes,
-satisfying the above boundary conditions, is limited by the maximum
-value of the vertical shear in the domain considered.
+where $\beta_\mathrm{crit}$
+is the critical thermal timescale for the fundamental mode
+($M=0$). We comment that the expression for $\beta_\mathrm{crit}$ may
+in fact be obtained without assuming $|\epsilon q\khat|\lesssim 1$.
+We will confirm $\beta_\mathrm{crit}$ numerically in 
+\S\ref{bcrit_num_test}.
 
-For the vertically isothermal disk we have $|f(\hat{z})| =
-|\hat{z}|\left(1+\epsilon^2\hat{z}^2\right)^{-3/2}$, which maximises
-at $|\zhat|=1/\sqrt{2}\epsilon$. For vertical domains smaller 
-than this height, the growth rate is limited by $|f|$ at the vertical
-boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
-\hat{z}$, in which case we can expect growth rates to increase linearly
-with the vertical domain size.   %no faster than linear
+Eq. \ref{iso_vsi_cond} suggests that as the thermal timescale $\beta$
+is increased from zero, higher order modes with $M>0$
+will be stabilized before the fundamental mode. We indeed observe this
+trend in numerical calculations (\S\ref{therm_relax_eff}), despite
+adopting somewhat restrictive assumptions for $M>0$ in order to arrive
+at the simple expression Eq. \ref{bcrit_gen}.
 
-\subsubsection{Thin-disk limit}
-We can make further progress for thin disks ($\epsilon\ll1$), 
-by expanding the background density and vertical shear profile in powers
-of $z/r$. To lowest order we obtain 
-\begin{align}
-  &\frac{\rho}{\rho_0} \simeq 
-  \exp{\left(-\frac{\hat{z}^2}{2}\right)} \equiv w(\hat{z}),\label{thin_dens}\\
-  &\frac{d\Omega^2}{d\hat{z}} \simeq \epsilon^2q\Omega_k^2\hat{z}, \label{thin_vshear}
-\end{align} 
-%With this approximation and those in \S\ref{integral_relation}, 
-and the governing equation becomes 
-\begin{align}\label{iso_ode3}
-  W^{\prime\prime} - \left(1 + \ii q\epsilon
-    \hat{k}\right)\hat{z}W^\prime  +
-  \hat{\sigma}^2\left(1+\hat{k}^2\right)W = 
-  0.
-\end{align}
-We remark that for large $|\hat{k}|$, Eq. \ref{iso_ode3} is the same as
-that derived by \cite{nelson13}, although we have taken a different
-route.
+Rapid thermal relaxation is required to overcome a stabilizing
+vertical entropy gradient \citep{goldreich67,urpin98,urpin03}, but these
+studies consider vertically localized disturbances. Our 
+Eq. \ref{iso_vsi_cond} quantifies this requirement for vertically
+global disturbances (which have small $M$).  We can interpret
+$\beta_\mathrm{crit}$ as the thermal timescale below which
+perturbations are effectively isothermal and the VSI operates.  We see
+that in a thin disk, a very short thermal timescale is required
+($\beta_\mathrm{crit}\ll 1$) because $\beta_\mathrm{crit}\propto \epsilon$.
 
-To complete the problem we must specify boundary conditions. The 
-maximum vertical domain size should be limited by the thin
-disk and low-frequency approximations (i.e. $z\lesssim r$).     
-It is, however, common to take $\mathrm{max}|z|\to\infty$ and impose
-that the kinetic  energy remain bounded at infinity. This 
-is appropriate for the density field because both the true density profile 
-and its thin-disk approximation decay rapidly away from the midplane.
+% Here we estimate the maximum thermal relaxation timescale
+% $\beta$ that allows the fundamental VSI mode for $\khat\gg1$.
 
-However, the thin-disk representation of vertical shear diverges with 
-$\hat{z}$, unlike the true vertical shear profile which eventually
-decays.  Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit unphysically large growth
-rates (as demonstrated below). Nevertheless, the infinite disk is
-useful to consider since it permits an analytical discussion.
+% At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
+% is real. The dispersion relation for the fundamental mode is then
+% \begin{align}\label{relax_disp_fund}
+%   0 = \sum_{l=1}^{6}c_l\hat{\omega}^l \quad \text{for $M=\hat{\nu}=0$}.  
+% \end{align} 
+% Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
+% $\khat\gg1$, we find
+% \begin{align}
+%   0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
+%   \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
+%   &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
+%   0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
+%   - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
+%     &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
+% \label{relax_cond2}
+% \end{align}
+% This is a pair of simultaneous equations for
+% $\hat{\omega}$ and $\beta$. Now, low-frequency axisymmetric waves in
+% the absence of vertical shear are stable and  satisfy
+% $\hat{\omega}\propto \khat^{-1}$ (Appendix \ref
+% {stable_novshear}, Eq. \ref{iso_disp_full}).  We are thus motivated to
+% seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
+% % we wish to seek a criterion independent of k 
+% % \begin{align}
+% %   \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
+% %\end{align}
+% \begin{align}
+%   \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
+%   |\hat{\omega}\khat|=O(1). 
+% \end{align}
+% Then
+% \begin{align}
+%   \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
+% \end{align}
+% from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
+% that
+% \begin{align}\label{beta_crit0}
+%   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
+%         q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
+% \end{align}
+% at marginal stability for perturbations with large $\khat$.
 
-\subsubsection{Stability in the absence of vertical shear}
-When $q=0$, Eq. \ref{iso_ode3} is
-\begin{align}\label{hermite_ode}
-  \left[w(\hat{z})W^\prime \right]^\prime + nW
-  w(\hat{z}) =0, 
-\end{align}
-where we have defined a new eigenvalue
-\begin{align}
-  n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
-\end{align} 
-Eq. \ref{iso_ode3} is Hermite's differential equation. If we impose
-that the kinetic energy density remain bounded at infinity, then  
-\begin{align}
-  W \propto \He_n(\hat{z}),
-\end{align}
-and $n$ is a non-negative integer. This translates to a real
-eigenfrequency $\sigma$ such that
-\begin{align}
-  \left|\hat{\sigma}\right| = \sqrt{n}
-  \left(1+\hat{k}^2\right)^{-1/2}. 
-\end{align}
-Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
-analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
-n-1$. Physically, this corresponds to radial length-scales much
-smaller than the local disk scale height.
-
-\subsubsection{Polynomial solutions with vertical shear}
-We can seek power-series solution to Eq. \ref{iso_ode3},
-\begin{align}
-  W(\zhat) = \sum_{l=0}^\infty a_l\zhat^l. 
-\end{align}
-Then the coefficients must satisfy the recurrence relation
-\begin{align}
-  (l+2)(l+1)a_{l+2} +
-  \left[n - l\left(1+\ii \epsilon q  \hat{k}\right)\right] a_l = 0, 
-\end{align}
-where $n$ is not neccessarily an integer. Indeed, if we demand
-a polynomial of degree $L$ as the solution, then the eigenfrequency
-must satisfy
-\begin{align}
-\hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
-    \hat{k}}{1+\hat{k}^2}\right).
-\end{align}
-The eigenfrequency $\hat{\sigma}$ is therefore complex for
-$L\geq1$. We are interested in 
-unstable modes for which $\imag{\hat{\sigma}}>0$ is the growth rate and is
-given by 
-\begin{align}\label{simple_growth}
-  \hat{\nu} =\sqrt{
-   \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
-    1\right)}. 
-\end{align}
-For fixed $q$ and $\epsilon$, the growth rate vanishes for both
-$\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. The maximum growth rate
-occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
-\begin{align}
-  |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
-\end{align}
-and
-\begin{align}
-  \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
-    q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
-\end{align}
-Since $L\geq1$, there is no limit to the maximum growth rate. This is
-an artifact of the thin-disk approximation because the vertical 
-shear $d\Omega^2/dz\propto z$ increases indefinitely with
-height. However, we find Eq. \ref{iso_max_growth} compares well with
-numerical solutions of the governing equation for small $L$.
-
-% This expression can be simplified in the limit $|q|\ll 1$, 
-%  \begin{align}\label{simple_growth}
-%    \frac{\imag(\sigma)}{\Omega_k} = \pm \sqrt{M} 
-%    \frac{q\epsilon\hat{k}}{2\sqrt{1+\hat{k}^2}} \quad\quad (|q|\ll 1), 
+% From our discussion in \S\ref{relax_pert}, which indicate introducing
+% a finite thermal relaxation is stabilizing, we expect
+% Eq. \ref{beta_crit0} is an upper limit. For a thin  
+% disk $\epsilon\ll 1$, we thus infer 
+% \begin{align}\label{iso_vsi_cond}
+%   \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
+%   \beta_\mathrm{crit} 
 % \end{align}
-%  with $M$ being an integer.
-
-In Appendix \ref{pert_theory} we give an alternative method to
-infer instability in the presence of vertical shear without explicitly
-solving the governing equation.
+% is required for the fundamental VSI to operate. That is, when
+% $\beta<\beta_\mathrm{crit}$, perturbations are  effectively isothermal. 
+% We will confirm Eq. \ref{iso_vsi_cond} numerically in
+% \S\ref{bcrit_num_test}. 
+\message{ !name(linear_adia.tex) !offset(-315) }

### 2015-03-16 appendix.tex
revised

@@
 An important caveat of the VGSB is that it cannot be used in the
-adiabatic limit. We explain this by considering vertically isothermal
+adiabatic limit, even if the above conditions are satisfied. We explain this by considering vertically isothermal
 disks and setting  $\beta\to\infty$ in 
@@
 that this inconsistency is unimportant for the VSI because it occurs
-for $\beta\ll1$.
+for $\beta\ll1$.   %and small scale
 
@@
     r}\frac{\p\Omega^2}{\p z} = -
-  \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(2 + q +
-    \frac{z^2/r^2-2}{z^2/r^2+1}\right). 
+   \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(
+  \frac{3z^2}{r^2+z^2}-1\right) \equiv - \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2}F(r,z).
 \end{align}
-For the local problem we may then write
-\begin{align}
-  \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z;q),
-\end{align}
-where the function $F$ corresponds to the bracket in
-Eq. \ref{dkappa2}. This function varies from $F=q$ at $z=0$ to $F\to
-3+q$ as $|z|\to\infty$. Since we consider $|q|=O(1)$, this impies
-$|F|=O(1)$ as well.
+% For the local problem we may then write
+% \begin{align}
+%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z),
+% \end{align}
+The function $F$ varies monotonically from $F=-1$ at $z=0$ to $F\to2$
+as $|z|\to\infty $, so $|F|=O(1)$. 
+% For the local problem we set
+% $dD/dz = \p\kappa^2/\p z$. 
+% This function varies from $F=q$ at $z=0$ to $F\to
+% 3+q$ as $|z|\to\infty$. Since we consider $|q|=O(1)$, this impies
+% $|F|=O(1)$ as well.
 
-Now, eliminating $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
-Eq. \ref{lin_vz} for vertically isothermal disks, retaining the
-vertical dependence of $D$ and including thermal relaxation, we
-obtain 
+For the radially local problem, we eliminate $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
+Eq. \ref{lin_vz}  (with $\Gamma=1$), but now retain the vertical dependence
+of $D$ using Eq. \ref{dkappa2}, to obtain  
 \begin{align}
@@
 where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
-t_c \gamma\right)$. Making the low-frequency approximation with the
-replacement $D\to \Omega_k^2$ gives, in terms of dimensionless
-variables,
+t_c \gamma\right)$. Making the low frequency, nearly Keplerian
+approximations with the replacement $D\to \Omega_k^2$ gives, in terms
+of dimensionless variables,
 \begin{align}
@@
 Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
-proportional to $\epsilon^2$. For a thin disk, $\epsilon\ll1$, so
-neglecting these terms has no qualitative effect on the discussion in 
-\S\ref{analytic_relax}.
+proportional to $\epsilon^2$. For a thin disk ($\epsilon\ll1$) with
+$|q|=O(1)$, as considered in this work, neglecting these terms has no
+qualitative effect on the discussion in \S\ref{analytic_relax}.
 
@@
 and neglect the last three terms in Eq. \ref{relax_cond2}. This
-procedure gives Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2},
+approximation gives Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2},
 but is not valid if $\beta$ or $|q|$ is large.
@@
 For the fundamental mode $M=0$ and Eq. \ref{relax_cond1} ---
-\ref{relax_cond2} simply significantly. It is then straight
+\ref{relax_cond2} simplify significantly. It is then straight
 forward to eliminate $\hat{\omega}\khat$ between 
@@
 simpler to work with an equation for $W$ by substituting $Q=W$ into
-Eq. \ref{ode_w}. In this case, we shall not yet make the low-frequency
-approximation, but first make the nearly-Keplerian approximation. We
+Eq. \ref{ode_w}. In this case, we shall not yet make the low frequency
+approximation, but first make the nearly Keplerian approximation. We
 obtain, in terms of dimensionless variables, 
@@
 \zhat\left(1+\epsilon^2\zhat^2\right)^{-3/2}$. More generally, though,
-$f$ may be considered as a representation of the vertical
+$f$ may be regarded as a representation of the vertical
 shear profile. Physically, we expect there is a maximum value of 
-$|d\Omega/dz|$. The existence of a maximum, as remarked in
-\S\ref{vshear_def}, limits the growth rate of the VSI. We explicitly
-demonstrate this below.
+$|d\Omega/dz|$, the existence of which should limit the growth rate of
+the VSI,  as remarked in \S\ref{vshear_def}. We explicitly demonstrate
+this below.
 
@@
 It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
-have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. vertical
-shear. A similar conclusion can be reached in the equivalent linear  
-problem in global cylindrical disk geometry.
+have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. there must
+be vertical shear. A similar conclusion can be reached in the
+equivalent linear problem in global cylindrical disk geometry.
 
@@
 -\zhat^2/2$ and $f(\zhat)\simeq \zhat$. However, we do not assume 
-low-frequency from the onset. Then Eq. \ref{iso_ode0} becomes 
+low frequency from the onset. Then Eq. \ref{iso_ode0} becomes 
 \begin{align}\label{iso_ode3}

### 2015-03-16 application.tex
revised

@@
 linearized RHS of Eq. \ref{real_energy} as
-% we choose the heating term such that
-% \begin{align}
-%   Q_+ - \nabla\cdot\bm{F} \equiv \nabla\cdot\left[k_\mathrm{rad}\nabla
-%     \left(T-T_{t=0}\right)\right], 
-% \end{align} 
-% where $k_\mathrm{rad}$ is the conduction coefficient to be defined
-% implicitly. Eq. \ref{real_energy} becomes
-% \begin{align}
-%   \frac{DP}{Dt} = -\gamma P \nabla\cdot\bm{v} + \frac{P}{\rho T
-%   c_v}\nabla\cdot\left[k_\mathrm{rad}\nabla\left(T-T_{t=0}\right)\right].     
-% \end{align}
-% When linearized, the last term on the RHS becomes
+
 \begin{align}\label{diff_cool_proper}
@@
 \end{align}
-% we are interested in radially-thin perturbations  
 where $\eta=k_\mathrm{rad}/\rho c_v$ is the diffusion coefficient and
@@
 \end{align}
-% where $\hat{\eta} = \eta/H^2\Omega_k$ is the 
-% dimensionless diffusion coefficient.
 
@@
 relaxation model as adopted in our linear models. That is, temperature
-flucations $\delta T$ decay on a timescale $t_\mathrm{cool}$,
+fluctuations $\delta T$ decay on a timescale $t_\mathrm{cool}$,
 independent of $k_x$. For simplicity, we take  
@@
 
-%but
-%given that the VSI depends on vertical shear, an effect that becomes
-%important away from the disk midplane, the inner disk may still be
-%able to develop the VSI, but it is likely restricted to the disk
-%surface.   
-%We remark that it is possible to satisfy $\beta<\beta_\mathrm{crit}$
-%at small radii by choosing a sufficiently large $\khat$. However, our
-%linear calculations suggest growth rates eventually decay as 
-%$\khat\to\infty$. Thus, very small-scale VSI in the inner disk is
-%probably unimportant.
-
 \begin{figure*}
@@
     critical thermal relaxation timescale $\beta_\mathrm{crit}$.  
-    % and 
-    % $\beta<\beta_\mathrm{crit}$ is necessary for the fundamental isothermal VSI to operate  
-    % for $\khat\gg1$. 
     \label{mmsn_bcrit_bcool}}   
@@
 that $\beta = \beta_\mathrm{crit}$. We thus expect that the VSI would only
-occur on very small radial scales in the inner disk. % However, these
-% may not be important as 
-% our
-% linear calculations suggest growth rates eventually decay for large
-% $\khat$.
+occur on very small radial scales in the inner disk.
 
@@
 optically thin away from it, such that $\beta < \beta_\mathrm{crit}$
-for $|z|\gtrsim1.25H$. In fact, we also find the VSI at $r=5$AU, but
-with much reduced growth rates (when scaled by the local rotation).
+for $|z|\gtrsim1.25H$. In fact, we also find the VSI in this case, but
+with much reduced growth rates (when scaled by the local rotation), as
+described below.
 
@@
 2\pi/\Omega_k$.  In fact, we either only find the fundamental mode, or
-that it is the most unstable, except for the appearence of some
+that it is the most unstable, except for the appearance of some
 surface modes towards the inner disk when considering large
@@
 significantly reduces the efficiency of the VSI despite $\beta <
-\beta_\mathrm{crit}$ away from the midplane. %because of fundamental mode
+\beta_\mathrm{crit}$ away from the midplane.
 

### 2015-03-16 linear_adia.tex
about to edit numerical
revised

@@
 Eq. \ref{nearly_iso_explicit}, keeping only first order terms, and
-solve for $\delta\hat{\sigma}$ using the above xpressions for $\delta\chi$
+solve for $\delta\hat{\sigma}$ using the above expressions for $\delta\chi$
 and $\hat{\sigma}$. We find imaginary part of $\delta\hat{\sigma}$
@@
 \end{align}
-Note that for $q=0$ Eq. \ref{marg_freq_approx} reduces to 
+Note that for $q=0$, Eq. \ref{marg_freq_approx} reduces to 
 the dispersion relation for low-frequency waves in vertically
@@
 \end{align}
-Since the second term in square brackets is small by assumption, we have 
+Since $|\epsilon q \khat|^2\lesssim 1$ by assumption, we have  
 \begin{align}\label{iso_vsi_cond}
@@
 ($M=0$). We comment that the expression for $\beta_\mathrm{crit}$ may
-in fact be obtained without assuming $|\epsilon q\khat|\lesssim 1$.  We will confirm Eq. \ref{iso_vsi_cond} numerically in 
+in fact be obtained without assuming $|\epsilon q\khat|\lesssim 1$.
+We will confirm $\beta_\mathrm{crit}$ numerically in 
 \S\ref{bcrit_num_test}.

### 2015-03-16 numerical.tex
revised

@@
 \end{align}
+%note that there are g_c terms in the boundary condition
 where $\bm{\xi}$ is the Lagrangian displacement with meridional 
-components $\xi_{x,z} = \ii\delta v_{x,z}/\sigma$.
+components $\xi_{x,z} = \ii\delta v_{x,z}/\sigma$. In some cases we
+impose a rigid boundary at which $\delta v_z=0$.
 
@@
 for both $\khat\to 0$ and $\khat\to \infty$. This is because for
-$\khat\ll 1$ the radial wavelenth is large, but such disturbances are 
+$\khat\ll 1$ the radial wavelength is large, but such disturbances are 
 stabilized by by epicyclic motions ($|\omega|\sim
@@
 there is no energy change when fluid elements are perturbed at fixed
-cylindrical radius whilst conserving its angular momentum (since we
-consider axisymmetric perturbations).
+cylindrical radius whilst conserving its angular momentum (applicable
+to axisymmetric perturbations).
 
@@
 In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of unstable 
-modes with $\khat=10$. The result for our fiducial setup is shown
-in black dots. We also plot other calculations with different
+modes with $\khat=10$. Black dots correspond to our fiducial setup. 
+We also plot other calculations with different
 vertical boundary conditions: rigid boundary and $\zmax=5H$ (red
@@
 The fundamental mode, marked by `f', corresponds to the eigenvalue
-with the smallest $|\sigma|$. We see that the fundamental mode is
-robust against vertical boundary conditions as it appears at the same
-position in all cases. Higher order modes with larger growth rates
-(along the analytic lines) require larger vertical domains, since they
-have increasing number of nodes along $z$.
+with the smallest $|\sigma|$. The fundamental mode is
+robust against the vertical boundary conditions considered as it
+appears at the same position in all cases. Higher order modes along
+the analytic lines have larger growth rates, but require larger
+vertical domains, since they have increasing number of nodes  $z$.
 
-The `tail' end with decreasing $\nu$ to the right of
-Fig. \ref{compare_modes_iso_kx10} is not predicted from
-analysis. These modes differ significantly between cases and are thus
-likely associated with boundary conditions. % We do not consider them 
+The `tail' end of modes with decreasing $\nu$ to the right of
+Fig. \ref{compare_modes_iso_kx10} is not predicted by our 
+analysis. They have lower growth rates than modes consistent with our
+analysis (i.e. those along the lines). These modes differ significantly between
+cases and are thus likely associated with boundary conditions. % We do
+                                % not consider them  
                                 % further.
@@
     with vertical domain size $\zmax=5H$  (black dots).  Other
-    cases correspond to different boundary condtions applied at
+    cases correspond to different boundary conditions applied at
     different heights. The fundamental mode is marked by `f'. Lines
@@
 diagrams such as Fig. \ref{compare_modes_iso_kx10}. We find them for 
-sufficiently large $\zmax$ and/or $\khat$. % For example, the two
-% largest growth rates found for the case with $\zmax=7H$ in
-% Fig. \ref{compare_modes_iso_kx10} are surface modes.
+sufficiently large $\zmax$ and/or $\khat$.
 
@@
 
-Due to their dependence on boundary conditions, we 
-do not consider surface modes further in detail. Surface modes have
+Due to their dependence on boundary conditions, we do not consider
+surface modes further in detail, which have
 already been discussed in the aforementioned studies. In fact, we will
 see in the next section that surface modes are more effectively
-stabilized by finite thermal relaxation than the fundamental mode. %  Thus, surface 
-% modes may not be the most relevant in realistic disks.
+stabilized by finite thermal relaxation than the fundamental mode.
 
@@
 $\zmax=5H$, we also plot results for $\zmax=3H$, $\zmax=7H$ and 
-analytic results computed from Eq. \ref{relax_disp}.
+analytic results computed from Eq. \ref{relax_disp}.  %analytic
+                                %results for low order modes
 
@@
     $M=1,\,2,\,3$ (dotted, dashed, dash-dot, respectively). Along each
-    line, $\beta$ increases continously from $0.01$ to $0.1$ from top
+    line, $\beta$ increases continuously from $0.01$ to $0.1$ from top
     to bottom, and squares mark corresponding $\beta$ values with
@@
 perturbations seen in Fig. \ref{lowfreq_eigenfunc_2d_cool} 
-(cf. Fig. \ref{lowfreq_eigenfunc_2d}). % However, most of the meridional
-% momentum is still contained within $|z|\lesssim 2H$ as for 
-% the case with $\beta=10^{-3}$.
+(cf. Fig. \ref{lowfreq_eigenfunc_2d}).
 
@@
 
-%fundamental mode bcomes the dominant mode as we approach beta
-%critical, therefore can use beta crit as criteria
-
 In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable
-modes with $\khat=30$. For clarity we only plot analytical results for
-the fundamental mode, which is again insensitive boundary
-conditions compared to higher order modes and thus likely more
-robust in realistic disks. We find surface modes
-appear for $\zmax\geq5H$. However, they are effectively stabilized with
-increasing $\beta$. For $\zmax=3H$, the fundamental mode becomes
-dominant for $\beta\geq 0.03$, as observed for $\khat=10$. For
-$\zmax=5H$ and $\zmax=7H$, we find the emergence modes slightly more
-unstable than the fundamental mode even when $\beta=0.1$ (the leftmost
-eigenvalues in the lower two panels). However, these modes are clearly associated
-with the boundary condition as they are absent for $\zmax=3H$. 
-% Thus, for $\zmax=5H$ and  
-% $\zmax=7H$  the fundamental mode is likely dominant for 
-% $\beta\geq0.05$
+modes with $\khat=30$. For clarity, we only plot analytic results 
+for the fundamental mode ($M=0$) and $M=1,\,2$, and omit the
+squares. The fundamental mode is again insensitive boundary
+conditions compared to higher order modes, and are in close agreement
+with analytic results. As before we observe more effective
+stabilization for higher order modes, including the absence of surface
+modes (which only occur for $\zmax\geq 5H$) as $\beta$ is increased.  
+For $\zmax=3H$, the fundamental mode becomes dominant for $\beta\geq
+0.03$. For $\zmax=5H$ and $\zmax=7H$, we
+find modes slightly more unstable than the fundamental
+mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
+panels), but these are likely associated  with the boundary
+condition as they are absent for $\zmax=3H$. These observations
+suggest the fundamental mode to be more robust in realistic disks.
 
@@
 conditions. The fundamental mode also becomes dominant as $\beta$
-approaches $\beta_\mathrm{crit}$. %beta<beta_crit is necessary and sufficient  
-Thus, tracking the fundamental mode is a more robust assessment of
-stability as $\beta$ is increased from zero, rather than 
-searching for the most unstable mode, which may correspond to modes
-that depend on boundary conditions.
+approaches $\beta_\mathrm{crit}$, except with the possible emergence
+of modes associated with boundary conditions. Thus, tracking the
+fundamental mode is a more preferable assessment of stability as $\beta$
+is increased from zero, rather than searching for the most unstable
+mode, because the latter may correspond to modes that depend on boundary
+conditions.
+
+%boundary conditions play an increasing role for higher M
 
@@
 performed by \cite{nelson13} for which a thermal relaxation timescale
-$\beta\gtrsim 0.6$ stablized the VSI. This can be anticipated from our
+$\beta\gtrsim 0.6$ stabilized the VSI. This can be anticipated from our
 estimate of the critical thermal timescale $\beta_\mathrm{crit}\simeq
-0.1$ for our fiducial disk parameters (which is the same as that in
-\citeauthor{nelson13}). It is interesting to note that there is a
-thermal relaxation timescale that maximizes the mode decay rate for
-$1\lesssim \khat \lesssim 30$. Here, it is $\beta\simeq1$---$2$ or
+0.1$ for our fiducial disk parameters (which is the same as that
+adopted by \citeauthor{nelson13}). It is interesting to note that there is a
+thermal timescale that maximizes the mode decay rate. Here,
+it is $\beta\simeq1$---$2$ or 
 about $0.1$---$0.3$ orbital periods. This is in fact consistent with 
-Fig. 12 of \citeauthor{nelson13} where the perturbed kinetic energy in their
+Fig. 12 of \cite{nelson13} where the perturbed kinetic energy in their
 numerical simulations were observed to decay most rapidly for a
@@
      $\beta$. These are obtained by following the eigenvalues of the
-     fundamental mode at $\beta=10^{-3}$. The vertical line is the
+     fundamental mode starting from $\beta=10^{-3}$ and increasing
+     $\beta$. The vertical line is the
      critical thermal timescale $\beta_\mathrm{crit}$ obtained  
@@
  We further test the dependence of the critical thermal relaxation timescale 
- $\beta_\mathrm{crit}$ on disk parameters
- (Eq. \ref{iso_vsi_cond}).  We use the fiducial setup as a
- reference point and vary $q\in[-0.2,-1.2]$, 
- $\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$
+ $\beta_\mathrm{crit}$ on disk parameters (Eq. \ref{iso_vsi_cond}).
+ We use the fiducial setup as a  reference point and vary
+ $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$ 
  separately.
- % Here, we use a wavenumber $\khat=10$, for which we
- % find growth rates reach zero for the range of disk parameters
- % considered, so that a numerical $\beta_\mathrm{crit}$ can be defined precisely. 
+
  The numerically obtained $\beta_\mathrm{crit}$ is shown in
@@
  our finding that the fundamental mode becomes dominant as one
- increases $\beta$, shows that $\beta_\mathrm{crit}$ as given 
- by Eq. \ref{iso_vsi_cond} provides a  
- practical measure for the thermal relaxation timescale below which 
- VSI will operate.
+ increases $\beta$, and that it is not sensitive to boundary
+ conditions, shows that $\beta_\mathrm{crit}$ as given  by
+ Eq. \ref{iso_vsi_cond} provides a practical measure for the thermal
+ relaxation timescale below which  VSI will operate.
 

### 2015-03-16 paper.tex
revised

@@
   periods, and the dominant VSI mode in the MMSN is the `fundamental
-  corrugation mode' which casues the disk column to oscillate
-  vertically. We suggest the VSI is dynamically imporant in the outer
+  corrugation mode' which causes the disk column to oscillate
+  vertically. We suggest the VSI is dynamically important in the outer
   parts of protoplanetary disks. 
@@
 The height-dependence of $\Omega$, or vertical shear, enables
-perturbations with large vertical lengthscales (relative to radial 
-lengthscales) to become unstable. \cite{umurhan13} gives a simple
+perturbations with large vertical length-scales (relative to radial 
+length-scales) to become unstable. \cite{umurhan13} gives a simple
 interpretation for the VSI: an exchange of fluid parcels while

### 2015-03-16 setup.tex
revised

@@
 timescales, in which case $\beta$ is non-constant and is determined by
-the equilibrium disk properties and the perturbation lengthscales of
+the equilibrium disk properties and the perturbation length-scales of
 interest.
@@
 \subsection{Radially-local approximation}  
-We consider perturbations with small radial lengthscales 
+We consider perturbations with small radial length-scales 
 so that we can ignore the radial variations in the background

### 2015-03-16 summary.tex
revised

@@
 In this paper we have studied the  axisymmetric vertical shear
-instability (VSI) in astrophysical disks that arises from the
+instability (VSI) in astrophysical disks that arise from the
 height dependence of the equilibrium angular velocity profile
@@
 regarded as an extension of \cite{urpin03}, which did include an
-energy equation, by relaxing the Boussinesq and vertically local
-approximations adopted in that analysis.
+energy equation, but adopted the Boussinesq and vertically local
+approximations. We considered vertically global and fully
+compressible perturbations.
 
@@
 ignoring the background radial disk structure. This is equivalent to adopting
-the `vertically-global shearing box' (VGSB) formalism developed by
-\cite{mcnally14}. % who showed the linear VSI studied by
-% \cite{nelson13} in a locally isothermal disk can be captured by the
-% VGSB. 
-We find the VGSB framework is appropriate for non-isothermal
-perturbations provided $\beta\lesssim1$.  This is convenient because
+the `vertically global shearing box' (VGSB) formalism developed by
+\cite{mcnally14}. 
+We find the VGSB framework is appropriate for analyzing the VSI because 
 rapid thermal relaxation ($\beta\ll1$) is 
-generally required for the VSI.  However, the VGSB framework is not
+generally required for the instability.   However, the VGSB framework is not
 appropriate in the adiabatic limit ($\beta \gg 1$).
-
-%OK framework for the regime of interest 
-% The presence of a
-% vertical shear, but 
-% is not consistent with the
-% global
 
@@
 
-We briefly studied vertically non-isothermal disks using our linear 
-code. We confirm the finding by \cite{nelson13} that in the absence
+We briefly studied vertically non-isothermal disks. We confirm the finding by \cite{nelson13} that in the absence
 of vertical entropy gradients, the VSI does not require rapid thermal
@@
 We applied our linear models to assess the applicability of the VSI in
-protoplanetary disks. We considered the Minimum Mass Solar Nebulae
+protoplanetary disks. We considered the Minimum Mass Solar Nebula 
 (MMSN) described in \cite{chiang10} and   
@@
 $\beta <\beta_\mathrm{crit}$ at the disk midplane gives a rough measure of the
-radii at which the VSI operates in the MMSN. This simple criteron is
+radii at which the VSI operates in the MMSN. This simple criterion is
 thus a useful first step to estimate the relevance of the VSI in
@@
 For more general disk structures convective overstability is
-possible. The models adopted for our numerical study in
+possible in principle. The models adopted for our numerical study in
 \S\ref{numerical} do satisfy $N_r^2<0$, but convective overstability
-requires considering perturbations with long radial lengthscales,
+require perturbations with long radial lengthscales,
 $\khat\ll 1$, for which radially global disk models would be more 
-appropriate.
+self-consistent.
 
@@
 We assumed non-adiabatic effects
-can be represented by a thermal relaxation or cooling timescale $t_c$. For 
+can be represented by a thermal relaxation timescale $t_c$. For 
 application to protoplanetary disks we connected $t_c$ to the disk
@@
 proper treatment of radiative diffusion. The linear problem then
-involves a Laplacian operator, which require additional boundary
-conditions. We will consider these improvements in a
+involves a Laplacian operator and additional boundary
+conditions must be supplemented. We will consider these improvements in a
 follow-up study. %additional heat source

### 2015-03-16 numerical.tex
revised

@@
 mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
-panels), but these are likely associated  with the boundary
-condition as they are absent for $\zmax=3H$. These observations
+panels), but these are clearly associated with boundary
+conditions as they are absent for $\zmax=3H$. These observations
 suggest the fundamental mode to be more robust in realistic disks.

### 2015-03-28 appendix.tex
revised
major revision in process because of barker & latter

@@
 % \end{align}
-The function $F$ varies monotonically from $F=-1$ at $z=0$ to $F\to2$
+The function $F$ increases monotonically from $F=-1$ at $z=0$ to $F\to2$
 as $|z|\to\infty $, so $|F|=O(1)$. 
@@
 \section{Linear problem in the isothermal limit}\label{iso_discuss}  
-Here we examine in more detail the special case of isothermal
-perturbations ($\beta\equiv 0$) for vertically isothermal disks 
-($\Gamma=1$) in the VGSB framework.   
-%Isothermal perturbations correspond to $\beta\equiv0$ ($\chi=1$) and 
-In this case Eq. \ref{ode_Q} becomes   
+Here we summarize some results for isothermal
+perturbations ($\beta\equiv 0$) in vertically isothermal disks 
+($\Gamma=1$) in the VGSB framework. In this case Eq. \ref{ode_Q}
+becomes    
 \begin{align} 
@@
 $\hat{k}^2\gg 1$, Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in
-\cite{nelson13}, although we have taken a different route.
+\cite{nelson13} or Eq. 28 in \cite{barker15}, although we have taken a
+different route.
 
@@
 
+Note that we have applied a regularity condition at infinity, since
+the vertically isothermal disk has no surface. If vertical boundaries
+are imposed at finite height, as done in numerical calculations, then
+the above solution needs to be modified to match the desired boundary
+conditions. This enables the `surface modes' seen in numerical
+calculations \citep{barker15}.
+
 \subsubsection{Stability without vertical shear}\label{stable_novshear}
@@
 vertically isothermal disk
-\citep[e.g.][]{okazaki87,takeuchi98,tanaka02,zhang06,ogilvie13,barker14}. In 
+\citep[e.g.][]{okazaki87,takeuchi98,tanaka02,zhang06,ogilvie13,barker14,barker15}. In 
 this case the solutions are Hermite polynomials, $W\propto
@@
 \subsubsection{Instability with vertical shear}
-The VSI corresponds to unstable inertial waves. This is readily
-obtained for large $\khat^2$ by balancing
-the last two terms of Eq. \ref{simple_growth}  to obtain the low
+The VSI corresponds to unstable inertial waves \citep{barker15}. This
+is readily obtained for large $\khat^2$ by balancing
+the last two terms of Eq. \ref{simple_growth} to give the low
 frequency branch. Then 
@@
   \hat{\sigma}^2 \simeq L\left(\frac{1+\ii q \epsilon
-       \hat{k}}{L+1+\hat{k}^2}\right).\label{simple_growth2}
+       \hat{k}}{L+1+\hat{k}^2}\right), \label{simple_growth2}
 \end{align}
-This signifies instability for $L\geq1$ since we can choose the sign
+which is equivalent to Eq. 34 of \cite{barker15} in the limit
+$\khat^2\gg L$. This signifies instability for
+$L\geq1$ since we can choose the sign 
 of the square root to make $\hat{\nu} = \imag\hat{\sigma}>0$.  These
@@
 Fig. \ref{compare_modes_iso_kx10} and
-Fig. \ref{compare_modes_iso_kx30} , for which $\nu\propto\omega$.
+Fig. \ref{compare_modes_iso_kx30}, for which $\nu\propto\omega$.
 

### 2015-03-28 linear_adia.tex
major revision in process because of barker & latter

@@
 Isothermal perturbations correspond to $\beta = 0$ or $\gamma=1$ so 
-that $\chi = 1$. This case has been investigated numerically
-by \cite{nelson13} and \cite{mcnally14}.
+that $\chi = 1$. This case has been investigated by \cite{nelson13},
+\cite{mcnally14} and \cite{barker15}.
 
-In Appendix \ref{iso_discuss}, we give an analytical discussion of the
-linear problem for isothermal perturbations. There, we establish more general
-results that do not require the thin disk approximation, namely 
-a necessary condition for instability (there must be vertical shear)
-and an upper bound to the instability growth rate (it is limited by
-the maximum vertical shear rate in the disk). We also develop 
-solutions in the thin disk limit. The 
-methods used in Appendix \ref{iso_discuss} and results obtained
+In Appendix \ref{iso_discuss}, we collect some useful results for
+isothermal  perturbations. This includes a necessary condition for
+instability (there must be vertical shear) and an upper bound to the
+instability growth rate (it is limited by the maximum vertical shear
+rate in the disk). We also develop solutions in the thin disk
+limit. The methods used in Appendix \ref{iso_discuss} and results obtained
 therein motivates the solution procedure employed in the following
-sections.
+sections, and are provided for ease of reference. We refer the reader
+to \cite{barker15} for a more detailed treatment of isothermal 
+perturbations.
 

### 2015-03-28 linear_iso.tex
revised
major revision in process because of barker & latter

@@
 number of approximations below that allows a simple governing equation
-to be derived and compared with previous works.  However, all of the 
-approximations described below are relaxed in our
-numerical study.
+to be derived and compared with previous works. Similar approximations
+are employed by \cite{nelson13} and \cite{barker15}. However, all of
+the approximations described below are relaxed in our numerical study.
+
+%vertically isothermal disk lacks a surface
 
@@
 introduces terms $O(\epsilon^2)$. Since we are interested in 
-thin disks ($\epsilon\ll 1 $), e.g. protoplaneary disks where 
+thin disks ($\epsilon\ll 1 $), e.g. protoplanetary disks where 
 $\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce

### 2015-03-28 paper.tex
major revision in process because of barker & latter

@@
   the VSI depends on the thermodynamic response of the disk by
-  varying $\beta$. We find increasing $\beta$ even slightly away from 
+  varying $\beta$. We find that in a stably stratified disk,
+  increasing $\beta$ even slightly away from    
   zero dramatically reduce VSI growth rates. 
@@
 \citep[GSF,][]{goldreich67,fricke68} in differentially rotating stars.   
-The height-dependence of $\Omega$, or vertical shear, enables
-perturbations with large vertical length-scales (relative to radial 
-length-scales) to become unstable. \cite{umurhan13} gives a simple
-interpretation for the VSI: an exchange of fluid parcels while
-conserving their angular momenta individually, may result in a new
-configuration of lower energy, if the exchange occurs over a
-sufficiently large vertical displacement. In practice, the VSI
-corresponds to unstable inertial waves.  
-%inertial waves 
-%history
+The height-dependence of $\Omega$, or vertical shear, can render
+inertial waves unstable. \cite{barker15} gives a simple 
+interpretation of the VSI: in an axisymmetric inviscid disk, the 
+exchange of fluid parcels vertically, while conserving their specific angular
+momentum, can result in a new configuration of lower energy if the
+distribution of specific angular momentum varies with $z$. The VSI is
+characterized by perturbations with large vertical length-scales
+(relative to radial length-scales).
 
-The VSI has been studied analytically in local approximation
-\citep{urpin98,urpin03}.  Numerical simulations of the  
-VSI have been performed by \cite{arlt04}, and more recently by 
-\cite{nelson13} and \cite{stoll14}. In these new simulations the authors
-indeed observe the growth of vertically global but radially localized
-perturbations when the equilibrium rotation varied with height.
+Early analysis of the VSI in radially and vertically local 
+models of accretion disks \citep{urpin98,urpin03} showed that the VSI
+requires a short thermal timescale so that perturbations evolve 
+isothermally, unless the disk is neutrally stratified in the vertical
+direction. This is because vertical motions associated with the VSI
+are stabilized by buoyancy forces if the disk is sub-adiabatically
+stratified (convectively stable). A sufficiently short thermal
+timescale renders buoyancy ineffective.
 
-An important finding from these works is that the VSI requires a short
-thermal timescale: perturbations need to be close to
-isothermal, unless the disk is neutrally stratified in the vertical
-direction. This is because vertical motions, required for the VSI, are stabilized by
-buoyancy if the disk is sub-adiabatically stratified for
-convective stability. When the thermal timescale $t_c$ is parametrized
-in units of the orbital period $P_\mathrm{orb}$, \cite{nelson13} find that
-$t_c\lesssim 0.01P_\mathrm{orb}$ was needed to observe instability. That is, the
-disk temperature essentially needs to remain unchanged during its
-evolution. This was further supported with radiation hydrodynamic
-simulations by \cite{stoll14}, who find the VSI is only sustained with
-the inclusion of heating from stellar irradiation, which resulted in
-an effective thermal timescale $t_c\ll P_\mathrm{orb}$.
+The need for a short thermal timescale in stably stratified disks has
+been demonstrated in numerical simulations of the VSI
+\citep{nelson13,stoll14}. When the  thermal timescale $t_c$ is
+parametrized in units of the orbital period  $P_\mathrm{orb}$,
+\cite{nelson13} find that $t_c\lesssim  0.01P_\mathrm{orb}$ was needed
+to observe the VSI. That is, the  disk temperature needs to remain
+effectively unchanged. Radiation hydrodynamic simulations performed by
+\cite{stoll14} also showed that the VSI is only sustained with the
+inclusion of heating from stellar irradiation, which results in an
+effective thermal timescale $t_c\ll P_\mathrm{orb}$ that maintains a
+fixed temperature profile.
 
-From these studies it is clear that the VSI is sensitive to
-the thermodynamic response of the disk if it is stably
-stratified. Since the latter is usually assumed for typical disk models,  
-it is important to understand and quantify said thermodynamic dependence,
-for example on disk parameters, in order to evaluate the relevance of
-the VSI in realistic disks. Should a given disk become turbulent
-through the VSI, it must first pass through the
-linear instability. Thus, understanding the thermodynamic dependence
-of the linear VSI is essential.
+Given that the VSI is generally sensitive to disk thermodynamics,  
+it is important to understand and quantify this dependence, for
+example on disk parameters, in order to evaluate the 
+relevance of the VSI to realistic disks, in which one may expect
+non-zero vertical entropy gradients and thermal timescales. Should a
+given disk become turbulent through the VSI, it must first pass
+through the linear instability. Thus, understanding the thermodynamic
+dependence of the  linear VSI is essential.
 
-In this work we study the linear VSI in radially local,
-vertically global models of astrophysical disks. We include an energy
-equation with a source term that allows the thermodynamic response of
-perturbations to be varied smoothly between isothermal and adiabatic. Our
-formulation generalizes previous linear calculations of the 
-VSI carried out by \cite{nelson13} and \cite{mcnally14} which assume
-isothermal perturbations and therefore did not include an energy
-equation. Our models are also an extension of earlier 
-analysis by \cite{urpin98} and \cite{urpin03}, which accounts for thermal
-evolution, but considered vertically localized disturbances. Our 
-vertically global approach is necessary to study disturbances with
-vertical wavelengths comparable to or larger than the disk scale
-height. Such disturbances were  found by \cite{nelson13} and
-\cite{stoll14} to dominate their numerical simulations.  We thus focus
-on vertically global perturbations which, in fact, we find to be the
-dominant mode in protoplanetary disks with a realistic estimate for
-the thermodynamic timescales.
+In this work we study the linear VSI in radially local, vertically
+global models of astrophysical disks. This approach has
+been adopted in recent studies of the linear VSI
+\citep{nelson13,mcnally14,barker15}. However, these studies consider 
+isothermal perturbations, for which the thermal timescale is formally
+zero, and/or neutrally stratified disks in which the stabilizing
+effect of bouyancy is absent. We account for both effects to study the
+VSI in more general disk models. To do so, we examine 
+sub-adiabatically stratified disks and include an energy equation 
+with a source term that allows the thermodynamic response of 
+perturbations to be varied smoothly between isothermal and 
+adiabatic.
+
+% Our models are also an extension of earlier 
+% analysis by \cite{urpin98} and \cite{urpin03}, which accounts for thermal
+% evolution, but considered vertically localized disturbances.
+% Our 
+% vertically global approach is necessary to study disturbances with
+% vertical wavelengths comparable to or larger than the disk scale
+% height. Such disturbances were found by \cite{nelson13} and
+% \cite{stoll14} to dominate their numerical simulations. We thus focus
+% on vertically global perturbations which, in fact, we find to be the
+% dominant mode in protoplanetary disks with a realistic estimate for
+% the thermodynamic timescales.
 

### 2015-03-28 numerical.tex
updating due to barker latter 2015

@@
 We first calculate the VSI in a disk with rapid thermal relaxation by
-setting $\beta=10^{-3}$. (Smaller values yield similar results.) This
-is similar to strictly isothermal perturbations in a vertically
-isothermal disk considered in \cite{nelson13} and \cite{mcnally14}.
+setting $\beta=10^{-3}$. (Smaller values yield similar results.) 
+This setup mimics strictly isothermal perturbations in a vertically
+isothermal disk, as considered by \cite{nelson13}, \cite{mcnally14} and
+\cite{barker15}. %Note, however, that \citeauthor{nelson13} and
+% \citeauthor{barker15} employed the  approximations described in \S\ref{analytical} throughout their analysis, which are relaxed in our
+% numerical study.  
+For completeness we review this case before considering the effect of
+thermal relaxation.
+
+%  In particular, we focus
+% on results most relevant to our eventual application to protoplanetary disks.
+
+% The reader is referred
+% to the above works for a detailed look at different VSI modes. Here,
+% we present selected results most relevant to our eventual application
+% to protoplanetary disks.
 
@@
 
-With rapid thermal relaxation, growth rates are expected to diminish
-for both $\khat\to 0$ and $\khat\to \infty$. This is because for
-$\khat\ll 1$ the radial wavelength is large, but such disturbances are 
-stabilized by by epicyclic motions ($|\omega|\sim
-\kappa$). Disturbances with $\khat\gg 1$ are radially localized, but  
-there is no energy change when fluid elements are perturbed at fixed
-cylindrical radius whilst conserving its angular momentum (applicable
-to axisymmetric perturbations).
+% With rapid thermal relaxation, growth rates are expected to diminish
+% for both $\khat\to 0$ and $\khat\to \infty$. This is because for
+% $\khat\ll 1$ the radial wavelength is large, but such disturbances are 
+% stabilized by by epicyclic motions ($|\omega|\sim
+% \kappa$). Disturbances with $\khat\gg 1$ are radially localized, but  
+% there is no energy change when fluid elements are perturbed at fixed
+% cylindrical radius whilst conserving its angular momentum (applicable
+% to axisymmetric perturbations).
 
 Fig. \ref{lowfreq_eigenfunc} plots the eigenfunctions $W$ and $\delta 
-v_z$ for the fundamental mode with $\khat=10$ (i.e. the mode with
+v_z$ for the fundamental mode with $\khat=10$ (which has the 
 maximum growth rate in Fig. \ref{iso_eigen_kx}). We find $W\propto z$
@@
 \subsubsection{Case study: $\khat=10$}
-In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of unstable 
+In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of all unstable 
 modes with $\khat=10$. Black dots correspond to our fiducial setup. 
@@
 The `tail' end of modes with decreasing $\nu$ to the right of
-Fig. \ref{compare_modes_iso_kx10} is not predicted by our 
-analysis. They have lower growth rates than modes consistent with our
-analysis (i.e. those along the lines). These modes differ significantly between
-cases and are thus likely associated with boundary conditions. % We do
+Fig. \ref{compare_modes_iso_kx10} are in fact also unstable inertial modes, like those
+along the analytic lines, but are strongly affected by the presence of 
+the imposed vertical boundaries \citep{barker15}.
+
+% They have lower growth rates than modes consistent with our
+% analysis (i.e. those along the lines). These modes differ significantly between
+% cases and are thus likely associated with boundary conditions. 
+% We do
 % not consider them  
@@
 Our fiducial setup with $\khat=10$ in 
-Fig. \ref{compare_modes_iso_kx10} (black dots) differs to that from
-\cite{nelson13} and \cite{mcnally14} in the lack of `surface modes'. 
-These modes lie almost parallel to the vertical axis in mode 
-diagrams such as Fig. \ref{compare_modes_iso_kx10}. We find them for 
-sufficiently large $\zmax$ and/or $\khat$.
+Fig. \ref{compare_modes_iso_kx10} (black dots)  lack `surface modes'
+found by \cite{nelson13}, \cite{mcnally14}  and 
+by \cite{barker15}.  We illustrate this with the mode
+diagram for $\khat=30$ in Fig. \ref{compare_modes_iso_kx30}. Surface
+modes lie almost parallel to the vertical axis
+in such diagrams. They are readily identified for  $\zmax\geq 5H$ and
+are marked by `s'.  We typically find an increasing
+number of them with increasing $\zmax$ and/or $\khat$,
+consistent with \cite{barker15}.
 
-To illustrate this, we plot the mode diagram for $\khat=30$ in  
-Fig. \ref{compare_modes_iso_kx30} with different boundary 
-conditions. Surface modes are readily identified for 
-$\zmax\geq 5H$ and are marked by `s'. They have the largest growth
-rates but require sufficiently large vertical domain size and their
-eigenvalues vary with $\zmax$.
+As explained by \cite{barker15}, the occurence of surface modes in the
+vertically isothermal disk is due to imposing a finite vertical domain
+size. The fastest growing mode is typically a surface mode, if they
+exist, but their association with vertical boundaries means that
+$\mathrm{max}(\nu)$ depends on the choice of $\zmax$. Thus, the
+vertically isothermal disk is not useful for studying surface modes. 
+%Thus, seeking the most unstable mode in a vertically isothermal disk
+%may not yield robust results because this
 
-Due to their dependence on boundary conditions, we do not consider
-surface modes further in detail, which have
-already been discussed in the aforementioned studies. In fact, we will
-see in the next section that surface modes are more effectively
-stabilized by finite thermal relaxation than the fundamental mode.
+We will see in the next section that surface modes, as well
+as other higher order modes, are more effectively stabilized by finite
+thermal relaxation than the fundamental mode. Since we are primarily
+interested in characterizing a maximum thermal timescale that allows
+the VSI, we can focus on the fundamental mode. For a detailed analysis of 
+surface modes, see \cite{barker15}.
+
+%refer the reader to BL15 for surface modes
 

### 2015-03-28 paper.tex
updating due to barker latter 2015

@@
 zero, and/or neutrally stratified disks in which the stabilizing
-effect of bouyancy is absent. We account for both effects to study the
+effect of buoyancy is absent. We account for both effects to study the
 VSI in more general disk models. To do so, we examine 

### 2015-03-28 numerical.tex
updating due to barker latter 2015

@@
 found by \cite{nelson13}, \cite{mcnally14}  and 
-by \cite{barker15}.  We illustrate this with the mode
+by \cite{barker15}.  We illustrate surface modes with the mode
 diagram for $\khat=30$ in Fig. \ref{compare_modes_iso_kx30}. Surface
@@
 vertically isothermal disk is due to imposing a finite vertical domain
-size. The fastest growing mode is typically a surface mode, if they
+size. The fastest growing mode is a surface mode, if they
 exist, but their association with vertical boundaries means that
@@
 $\beta=0.01$ as there are modes with $M>0$ that have larger growth 
-rates. However, by comparing with results for $\zmax=3H$, we see these
+rates. However, by comparing with results for $\zmax=3H$, we see that these
 higher order modes are strongly influenced by boundary conditions, and
@@
 
-However, the fundamental mode is dominant for $\beta \geq 
+However, the fundamental mode becomes dominant for $\beta \geq 
 0.03$. Higher order modes, with larger growth rates at small $\beta$,
@@
 characterized by $W\sim z$ and $\delta v_z\sim$ constant closer to the
-mid-plane, with oscillatory behavior emerging near vertical boundaries
+midplane, with oscillatory behavior emerging near vertical boundaries
 where buoyancy becomes important. This leads to the `tilted' pressure
@@
 panels), but these are clearly associated with boundary
-conditions as they are absent for $\zmax=3H$. These observations
-suggest the fundamental mode to be more robust in realistic disks.
+conditions as they are absent for $\zmax=3H$.
 
@@
 is increased from zero, rather than searching for the most unstable
-mode, because the latter may correspond to modes that depend on boundary
-conditions.
+mode. This is because the latter may correspond to modes that depend on boundary
+conditions, e.g. surface modes, whose growth rate varies with
+$\zmax$.
 

### 2015-03-28 numerical.tex
urgent revision

@@
 mode. This is because the latter may correspond to modes that depend on boundary
-conditions, e.g. surface modes, whose growth rate varies with
-$\zmax$.
+conditions, e.g. surface modes whose growth rate varies with
+$\zmax$ in a vertically isothermal disk.
 
@@
 We now test the critical thermal relaxation timescale given by
-Eq. \ref{iso_vsi_cond}. We begin from eigenvalues for the fundamental
+Eq. \ref{iso_vsi_cond}. We iterate starting from eigenvalues for the fundamental
 mode with $\beta=10^{-3}$ and slowly increase the thermal
 relaxation timescale up to $\beta=10^3$. Fig. \ref{bcrit_compare1}
-show the obtained growth rates, together with the critical  
+show the growth rates obtained this way, together with the critical  
 thermal relaxation timescale $\beta_\mathrm{crit}=0.125$ for our
@@
  Eq. \ref{iso_vsi_cond} provides a practical measure for the thermal
- relaxation timescale below which  VSI will operate.
+ relaxation timescale below which  VSI will operate in a vertically
+ isothermal disk.
 
@@
 \subsection{Vertically non-isothermal disks}
-We briefly consider vertically non-isothermal disks with 
-$\Gamma=1.4$ and $(p,q,\epsilon)=(0,-1,0.05)$ as simulated in
-\cite{nelson13}. In this case we set $\zmax=0.99H_s$.
 
-Fig. \ref{gcorr_compare_vnoniso} plots the fundamental VSI growth
-rates as a function of $\beta$ for $\gamma\in[1.4,2.5]$. In agreement
-with \citeauthor{nelson13}, in   
-the neutrally-stratified case $\gamma=\Gamma$ the disk is unstable 
-even for $\beta\gg 1$. This is due to the absence of a stabilizing
-vertical entropy gradient ($N_z^2\equiv 0$). % (Note that in the adiabatic limit 
-% this disk can be unstable according to the
-% Solberg-Hoiland criterion, Eq. \ref{solberg2}.)
+% We briefly consider vertically non-isothermal disks with 
+% $\Gamma=1.4$ and $(p,q,\epsilon)=(0,-1,0.05)$ as simulated in
+% \cite{nelson13}. In this case we set $\zmax=0.99H_s$.
 
-For $\gamma>\Gamma$, i.e. stably stratified disks,
-Fig. \ref{gcorr_compare_vnoniso} shows that introducing finite thermal
-relaxation rapidly stabilizes the disk, similar to that observed for
-nearly vertically isothermal disks (Fig. \ref{bcrit_compare1}). For
-$\gamma=1.7,\,2.0,\,2.5$, growth rates reach zero at
-$\beta\simeq0.17,\,0.083,\,0.045$, respectively. Interestingly, these
-values are equal to $\epsilon|q|/(\gamma-\Gamma)$. This suggests that
-the critical thermal relaxation timescale for vertically
-non-isothermal disks can also be estimated by Eq. \ref{iso_vsi_cond}
-but with $\gamma-1$ replaced by $\gamma-\Gamma$.
+% Fig. \ref{gcorr_compare_vnoniso} plots the fundamental VSI growth
+% rates as a function of $\beta$ for $\gamma\in[1.4,2.5]$. In agreement
+% with \citeauthor{nelson13}, in   
+% the neutrally-stratified case $\gamma=\Gamma$ the disk is unstable 
+% even for $\beta\gg 1$. This is due to the absence of a stabilizing
+% vertical entropy gradient ($N_z^2\equiv 0$). % (Note that in the adiabatic limit 
+% % this disk can be unstable according to the
+% % Solberg-Hoiland criterion, Eq. \ref{solberg2}.)
 
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/gcorr_compare_vnoniso2}
-  \caption{Growth rate of the fundamental VSI mode as a function of
-    the thermal relaxation timescale $\beta$, in vertically
-    non-isothermal disks with $\Gamma=1.4$ and
-    $\gamma\in[1.4,2.5]$. The disk is neutrally
-    stratified for $\gamma=1.4$ and stably stratified for
-    $\gamma>1.4$. Other disk parameters are
-    $(p,q,\epsilon)=(0,-1,0.05)$ and the perturbation wavenumber is
-    $\khat=30$.   
-    \label{gcorr_compare_vnoniso}}
-\end{figure}
+% For $\gamma>\Gamma$, i.e. stably stratified disks,
+% Fig. \ref{gcorr_compare_vnoniso} shows that introducing finite thermal
+% relaxation rapidly stabilizes the disk, similar to that observed for
+% nearly vertically isothermal disks (Fig. \ref{bcrit_compare1}). For
+% $\gamma=1.7,\,2.0,\,2.5$, growth rates reach zero at
+% $\beta\simeq0.17,\,0.083,\,0.045$, respectively. Interestingly, these
+% values are equal to $\epsilon|q|/(\gamma-\Gamma)$. This suggests that
+% the critical thermal relaxation timescale for vertically
+% non-isothermal disks can also be estimated by Eq. \ref{iso_vsi_cond}
+% but with $\gamma-1$ replaced by $\gamma-\Gamma$.
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+%   0cm]{figures/gcorr_compare_vnoniso2}
+%   \caption{Growth rate of the fundamental VSI mode as a function of
+%     the thermal relaxation timescale $\beta$, in vertically
+%     non-isothermal disks with $\Gamma=1.4$ and
+%     $\gamma\in[1.4,2.5]$. The disk is neutrally
+%     stratified for $\gamma=1.4$ and stably stratified for
+%     $\gamma>1.4$. Other disk parameters are
+%     $(p,q,\epsilon)=(0,-1,0.05)$ and the perturbation wavenumber is
+%     $\khat=30$.   
+%     \label{gcorr_compare_vnoniso}}
+% \end{figure}

### 2015-03-29 numerical.tex
major revisions

@@
 \subsection{Vertically non-isothermal disks} 
+%We briefly consider vertically non-isothermal disks in which there is
+%a disk surface at finite height. 
+\cite{barker15} have argued that vertically isothermal disks are not
+ideal for studing the VSI because $\zmax$ is a free
+parameter in numerical calcluations. On the other hand, vertically
+non-isothermal 
+disks have a disk surface at finite height so there is a natural
+choice for $\zmax$. This means that higher order modes
+and surface modes are better defined in a vertically non-isothermal
+disk. However, note that this issue is unimportant for low order modes 
+such as the fundamental mode, which are not sensitive to boundary
+conditions.
+
+Here we briefly consider a vertically non-isothermal disk model with  
+$(p,q,\epsilon)=(0,-1,0.05)$ and $\Gamma=1.0837$. This value of 
+$\Gamma$ is chosen so that the standard vertical domain size $\zmax =
+5H$ effectively corresponds to the disk surface ($\zmax\simeq
+H_s$). The disk temperature decreases rapidly with height such that 
+$c_s(\zmax)\simeq 10^{-4}c_s(0)$.
+
+In Fig. \ref{compare_modes_vnoniso_kx10} we plot the mode diagram for
+$\khat=10$ for several values of $\beta$. This plot is
+qualitatively similar to Fig. \ref{compare_modes_cool_kx10}. As before
+we find as $\beta$ is increased, higher order modes are rapidly
+stabilized, eventually only the fundamental mode can operate before
+the VSI is entirely stabilized.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/compare_modes_Gam1.084_kx10.ps}
+  \caption{Unstable modes in a vertically non-isothermal disk with
+    $\khat=10$ and a range of thermal relaxation timescales.  
+    \label{compare_modes_vnoniso_kx10}}
+\end{figure}
+
+We do not find surface modes for $\khat=10$. In the vertically
+non-isothermal disk, surface modes develop for sufficiently large 
+$\khat$ \citep{barker15}. We demonstrate them in
+Fig. \ref{compare_modes_vnoniso_kx100} which show eigenvalues for
+$\khat=100$. While growth rates generally decrease with increasing 
+$\beta$, we find eigenvalues can overlap, making it difficult to 
+distinguish between different modes. However, since there 
+is a well-defined surface, we may seek the 
+most unstable mode without the boundary-dependent issue that would
+arise in a vertically isothermal disk.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/compare_modes_Gam1.084_kx100.ps}
+  \caption{Same as Fig. \ref{compare_modes_vnoniso_kx10} but with
+    $\khat=100$. In this case surface modes appear and are readily identified for
+    $\beta=10^{-3}$ and nearly
+    parallel to the vertical axis
+    \label{compare_modes_vnoniso_kx100}}
+\end{figure}
+
+Fig. \ref{compare_eigenvz_vnoniso_kx100}
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.73cm 0cm
+  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d001.ps}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.73cm 0cm
+  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d01.ps}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d1.ps}
+  \caption{Vertical velocity eigenfunction of the most unstable mode
+    with $\khat=100$ found in the vertically non-isothermal disk with
+    $\beta = 10^{-3}$ (top), $\beta=10^{-2}$ (middle) and $\beta=0.1$
+    (bottom). 
+    \label{compare_eigenvz_vnoniso_kx100}}
+\end{figure}
 

### 2015-03-29 numerical.tex
updating

@@
 
-%  In particular, we focus
-% on results most relevant to our eventual application to protoplanetary disks.
+% Fig. \ref{iso_eigen_kx} compares the numerically obtained eigenvalues 
+% for the fundamental mode and that calculated from 
+% Eq. \ref{simple_growth}. We find good agreement in the
+% wave frequency $\omega$ at all $\khat$ and in growth rates $\nu$ for
+% $\khat\gtrsim 10$. Growth rates are over-estimated by our analysis but 
+% are $O(\epsilon\Omega_k)$, which is consistent with the discussion in
+% Appendix \ref{max_growth1}. The match in $\nu$ worsens for 
+% decreasing $\khat$ due to the neglect of global radial structure in
+% the analysis. Nevertheless, given the number of approximations made
+% in our analysis, the agreement is satisfactory.
 
-% The reader is referred
-% to the above works for a detailed look at different VSI modes. Here,
-% we present selected results most relevant to our eventual application
-% to protoplanetary disks.
-
-Fig. \ref{iso_eigen_kx} compares the numerically obtained eigenvalues 
-for the fundamental mode and that calculated from 
-Eq. \ref{simple_growth}. We find good agreement in the
-wave frequency $\omega$ at all $\khat$ and in growth rates $\nu$ for
-$\khat\gtrsim 10$. Growth rates are over-estimated by our analysis but 
-are $O(\epsilon\Omega_k)$, which is consistent with the discussion in
-Appendix \ref{max_growth1}. The match in $\nu$ worsens for 
-decreasing $\khat$ due to the neglect of global radial structure in
-the analysis. Nevertheless, given the number of approximations made
-in our analysis, the agreement is satisfactory.
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_eigen_imag_iso} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  1cm]{figures/compare_eigen_real_iso}
-  \caption{Growth rate (top) and real frequency (bottom) of the
-    fundamental VSI mode in the fiducial disk model with $\beta =
-    10^{-3}$. 
-    % a disk with $(\gamma,
-    % \Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
-    % $\beta=10^{-3}$. 
-    The solid (dotted) line is obtained numerically
-    (analytically).  
-    \label{iso_eigen_kx} 
-  }
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+%   0cm]{figures/compare_eigen_imag_iso} 
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+%   1cm]{figures/compare_eigen_real_iso}
+%   \caption{Growth rate (top) and real frequency (bottom) of the
+%     fundamental VSI mode in the fiducial disk model with $\beta =
+%     10^{-3}$. 
+%     % a disk with $(\gamma,
+%     % \Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+%     % $\beta=10^{-3}$. 
+%     The solid (dotted) line is obtained numerically
+%     (analytically).  
+%     \label{iso_eigen_kx} 
+%   }
+% \end{figure}
 
@@
 % to axisymmetric perturbations).
+
+\subsubsection{Case study: $\khat=10$}
+In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of all unstable 
+modes with $\khat=10$. Black dots correspond to our fiducial setup. 
+We also plot other calculations with different
+vertical boundary conditions: rigid boundary and $\zmax=5H$ (red
+crosses),  free surface and $\zmax = 3H$ (blue triangles), free surface
+and $\zmax=7H$ (green crosses).
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10_analytic.ps}
+  \caption{Unstable modes with $\khat=10$ in the fiducial disk model
+    with $\beta=10^{-3}$. %  a disk with 
+    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+    % $\beta=10^{-3}$. 
+    The fiducial boundary condition is a free surface ($\Delta P=0$)
+    with vertical domain size $\zmax=5H$  (black dots).  Other
+    cases correspond to different boundary conditions applied at
+    different heights. The fundamental mode is marked by `f'. Lines
+    are computed from the analytic dispersion relation 
+    with thermal relaxation Eq. \ref{relax_disp} (dotted), and that for
+    isothermal perturbations, Eq. \ref{simple_growth} (dashed). 
+    \label{compare_modes_iso_kx10} 
+  }
+\end{figure}
+
+The fundamental mode, marked by `f', corresponds to the eigenvalue
+with the smallest $|\sigma|$. The fundamental mode is
+robust against the vertical boundary conditions considered as it
+appears at the same position in all cases.
 
@@
 
-\subsubsection{Case study: $\khat=10$}
-In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of all unstable 
-modes with $\khat=10$. Black dots correspond to our fiducial setup. 
-We also plot other calculations with different
-vertical boundary conditions: rigid boundary and $\zmax=5H$ (red
-crosses),  free surface and $\zmax = 3H$ (blue triangles), free surface
-and $\zmax=7H$ (green crosses).
-
-The fundamental mode, marked by `f', corresponds to the eigenvalue
-with the smallest $|\sigma|$. The fundamental mode is
-robust against the vertical boundary conditions considered as it
-appears at the same position in all cases. Higher order modes along
+Higher order modes along
 the analytic lines have larger growth rates, but require larger
@@
 % further.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10_analytic.ps}
-  \caption{Unstable modes with $\khat=10$ in the fiducial disk model
-    with $\beta=10^{-3}$. %  a disk with 
-    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
-    % $\beta=10^{-3}$. 
-    The fiducial boundary condition is a free surface ($\Delta P=0$)
-    with vertical domain size $\zmax=5H$  (black dots).  Other
-    cases correspond to different boundary conditions applied at
-    different heights. The fundamental mode is marked by `f'. Lines
-    are computed from the analytic dispersion relation 
-    with thermal relaxation Eq. \ref{relax_disp} (dotted), and that for
-    isothermal perturbations, Eq. \ref{simple_growth} (dashed). 
-    \label{compare_modes_iso_kx10} 
-  }
-\end{figure}
 
@@
 
-Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
-show that the fundamental mode is not sensitive to boundary
-conditions. The fundamental mode also becomes dominant as $\beta$
-approaches $\beta_\mathrm{crit}$, except with the possible emergence
-of modes associated with boundary conditions. Thus, tracking the
-fundamental mode is a more preferable assessment of stability as $\beta$
-is increased from zero, rather than searching for the most unstable
-mode. This is because the latter may correspond to modes that depend on boundary
-conditions, e.g. surface modes whose growth rate varies with
-$\zmax$ in a vertically isothermal disk.
+% Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
+% show that the fundamental mode is not sensitive to boundary
+% conditions. The fundamental mode also becomes dominant as $\beta$
+% approaches $\beta_\mathrm{crit}$, except with the possible emergence
+% of modes associated with boundary conditions. Thus, tracking the
+% fundamental mode is a more preferable assessment of stability as $\beta$
+% is increased from zero, rather than searching for the most unstable
+% mode. This is because the latter may correspond to modes that depend on boundary
+% conditions, e.g. surface modes whose growth rate varies with
+% $\zmax$ in a vertically isothermal disk.
 
@@
   timescale}\label{bcrit_num_test}
-We now test the critical thermal relaxation timescale given by
-Eq. \ref{iso_vsi_cond}. We iterate starting from eigenvalues for the fundamental
-mode with $\beta=10^{-3}$ and slowly increase the thermal
-relaxation timescale up to $\beta=10^3$. Fig. \ref{bcrit_compare1}
-show the growth rates obtained this way, together with the critical  
-thermal relaxation timescale $\beta_\mathrm{crit}=0.125$ for our
-fiducial disk parameters.
+In Fig. \ref{bcrit_compare1} we plot the maximum VSI growth rates
+found in the fiducial disk model as a function of the thermal
+relaxation timescale $\beta$ for several values of $\khat$. 
+We also plot the critical timescale $\beta_\mathrm{crit}=0.125$ for
+the fiducial disk parameters.
 
-For $5\lesssim\khat\lesssim30$, we find $\beta_\mathrm{crit}$ provides
-an accurate prediction of the upper limit to the thermal relaxation 
-timescale for the fundamental mode. For smaller $\khat$, the VSI can
-operate slightly beyond $\beta_\mathrm{crit}$ (recall that
-$\beta_\mathrm{crit}$ was derived for $\khat^2\gg 1$). For
-$\khat\geq 50$, we find perturbations have large amplitudes near the
-vertical boundaries, and 
-complications to our numerical search of
-eigenvalues arise from the emergence of modes associated with boundary
-conditions. %We generally find boundary effects become more significant
-%with increasing $\khat$.  
-Nevertheless, growth rates rapidly decrease as $\beta$
-approaches and increase beyond $\beta_\mathrm{crit}$.
+Note that the curves are not neccessarily smooth because the most unstable
+mode may not always correspond to the same type of disturbance. For example,
+as seen above, for $\khat \leq 10$ higher order body modes are
+dominant as $\beta\to 0$, but the fundamental corrugation
+mode becomes dominant as $\beta$ is increased. For $\khat=1$ the VSI
+can operate slightly beyond $\beta_\mathrm{crit}$  (recall that
+$\beta_\mathrm{crit}$ was derived assuming $\khat^2\gg 1$). For
+$\khat=5,\,10$ we find $\beta_\mathrm{crit}$ accurately predicts the
+thermal timescal beyond which the VSI is suppressed.
 
-Fig. \ref{bcrit_compare1} is consistent with numerical simulations
-performed by \cite{nelson13} for which a thermal relaxation timescale
-$\beta\gtrsim 0.6$ stabilized the VSI. This can be anticipated from our
-estimate of the critical thermal timescale $\beta_\mathrm{crit}\simeq
-0.1$ for our fiducial disk parameters (which is the same as that
-adopted by \citeauthor{nelson13}). It is interesting to note that there is a
-thermal timescale that maximizes the mode decay rate. Here,
-it is $\beta\simeq1$---$2$ or 
-about $0.1$---$0.3$ orbital periods. This is in fact consistent with 
-Fig. 12 of \cite{nelson13} where the perturbed kinetic energy in their
-numerical simulations were observed to decay most rapidly for a
-thermal timescale of $\sim 0.1$ orbits.
+Complications from vertical boundaries arise when considering large
+$\khat$. For $\khat=30$ we find for $\beta\gtrsim 0.1$ the dominant
+disturbance correspond to the leftmost eigenvalues
+seen in the lower two panels of Fig. \ref{compare_modes_cool_kx30},
+which depend on $\zmax$. For $\khat=50,\,100$ we find surface modes
+dominate.
+
+Our result is consistent with numerical simulations performed by
+\cite{nelson13} for which a thermal relaxation timescale $\beta\gtrsim
+0.6$ stabilized the VSI; as Fig. \ref{bcrit_compare1} show that growth
+rates for $\beta\gtrsim 0.6$ are small. One might consider 
+$\khat > 100$ to increase the maximum growth rate (note that the most unstable
+mode for large $\khat$ are typically surface modes), but in a realistic disk
+very small scale disturbances will likely be subject to viscous decay
+(\S\ref{application}). Numerical simulations are also unlikely to
+resolve such small scales.
+
+As pointed out by \cite{barker15} surface modes
+only appear in vertically isothermal disks when vertical boundaries
+are imposed, so they are strongly dependent on $\zmax$.  Thus, seeking
+the most unstable mode in a vertically isothermal disk may not yield
+robust results because such modes, which are often surface modes, may
+depend on boundary conditions.
 
 \begin{figure}
-   \includegraphics[width=\linewidth]{figures/gcorr_compare2} 
-   \caption{VSI growth rates in the fiducial disk
+   \includegraphics[width=\linewidth]{figures/gcorr_compare_iso_maxrate} 
+   \caption{Maximum VSI growth rates in the fiducial disk
      model as a function of the thermal relaxation timescale
-     $\beta$. These are obtained by following the eigenvalues of the
-     fundamental mode starting from $\beta=10^{-3}$ and increasing
      $\beta$. The vertical line is the
@@
 
- We further test the dependence of the critical thermal relaxation timescale 
- $\beta_\mathrm{crit}$ on disk parameters (Eq. \ref{iso_vsi_cond}).
- We use the fiducial setup as a  reference point and vary
- $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$ 
- separately.
+Despite this caveat, and that surface modes are not accounted in our
+analysis,  we see from Fig. \ref{bcrit_compare1} that 
+$\beta_\mathrm{crit}$ nevertheless provides a reasonable estimate for
+an upper limit to the thermal timescale that allows the VSI. We thus
+test the dependence of $\beta_\mathrm{crit}$ on disk parameters
+(Eq. \ref{iso_vsi_cond}).  We use the fiducial setup as a  reference
+point and vary $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and
+$\epsilon\in[0.02,0.1]$ separately. We consider the fundamental
+corrugation mode for which $\beta_\crit$ is the actual critical
+thermal timescale.%  We use $\khat=10$ to avoid complications to our
+% eigenvalue search from surface modes.
 
@@
 
- Fig. \ref{bcrit_compare}, together with Fig. \ref{bcrit_compare1} and 
- our finding that the fundamental mode becomes dominant as one
- increases $\beta$, and that it is not sensitive to boundary
- conditions, shows that $\beta_\mathrm{crit}$ as given  by
- Eq. \ref{iso_vsi_cond} provides a practical measure for the thermal
- relaxation timescale below which  VSI will operate in a vertically
- isothermal disk.
+Having validated
+
+ % Fig. \ref{bcrit_compare}, together with Fig. \ref{bcrit_compare1} and 
+ % our finding that the fundamental mode becomes dominant as one
+ % increases $\beta$, and that it is not sensitive to boundary
+ % conditions, shows that $\beta_\mathrm{crit}$ as given  by
+ % Eq. \ref{iso_vsi_cond} provides a practical measure for the thermal
+ % relaxation timescale below which  VSI will operate in a vertically
+ % isothermal disk.
 
@@
 
-Fig. \ref{compare_eigenvz_vnoniso_kx100}
+Fig. \ref{compare_eigenvz_vnoniso_kx100} show the most unstable modes
+with $\khat=100$ and $\beta=10^{-3},\, 10^{-2}$ and
+$0.1$. At such large $\khat$, we find with $\beta=10^{-3}$ the most
+unstable mode is an asymmetric surface mode \citep[also seen in Fig. 7
+of][]{barker15}. However, increasing the thermal timescale to
+$\beta=10^{-2}$ an anti-symmetric surface mode becomes
+dominant. Further increasing to $\beta=0.1$ surface modes are
+suppressed and the dominant mode is the fundamental corrugation
+mode. Notice as $\beta$ is increased the perturbation amplitude near
+the vertical boundaries become negligible. This is not surprsing since
+the stabilizing effect of buoyancy is largest at the disk surface.
 

### 2015-03-29 paper.tex
updating

@@
   we study the linear, axisymmetric VSI in radially local, 
-  vertically global models of accretion disks in which the disk  
+  vertically global models of stably stratified accretion disks in which the disk  
   temperature $T$ is relaxed its initial value over a characteristic timescale 
@@
   the VSI depends on the thermodynamic response of the disk by
-  varying $\beta$. We find that in a stably stratified disk,
+  varying $\beta$. We find that 
   increasing $\beta$ even slightly away from    

### 2015-03-29 linear_adia.tex
major revisions

@@
 v_z=$constant and $\chi=1$ so that $B=C=0$. 
-This is the fundamental or `corrugation' mode where the vertical
+This is the `fundamental corrugation mode' where the vertical
 velocity is constant throughout the disk column. These modes have been
 observed to dominate numerical simulations \citep{nelson13,stoll14},  
-and are the ones we find to dominate in numerical calculations with 
+and are the ones we find to typically dominate in numerical calculations with 
 increasing $\beta$ (\S\ref{therm_relax_eff}), as well as in protoplanetary disks 

### 2015-03-29 numerical.tex
major revisions

@@
   \caption{Equilibrium rotation profile $\Omega(z)$,
-    normalized by its mid-plane value, for a disk with $\Gamma=1.011$
+    normalized by its mid-plane value, for  the fiducial disk model with $\Gamma=1.011$
     and $(p,q,\epsilon)=(-1.5,-1,0.05)$. 
@@
 % numerical study.  
-For completeness we review this case before considering the effect of
+For completeness, and as a way of introducing terminology, 
+we review this case before considering the effect of 
 thermal relaxation.
@@
 \subsubsection{Case study: $\khat=10$}
-In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of all unstable 
+In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of unstable 
 modes with $\khat=10$. Black dots correspond to our fiducial setup. 
-We also plot other calculations with different
+We also plot results obtained with different
 vertical boundary conditions: rigid boundary and $\zmax=5H$ (red
@@
 and $\zmax=7H$ (green crosses).
+
+The modes shown in Fig. \ref{compare_modes_iso_kx10} are inertial
+waves made unstable by vertical shear and have been termed `body
+modes' \citep{barker15}.  The `fundamental corrugation mode', marked by `f',
+corresponds to the eigenvalue with the smallest $|\sigma|$. Hereafter
+we will simply refer to this as the fundamental mode. The
+fundamental mode is robust against vertical boundary conditions 
+as it appears at the same position in
+Fig. \ref{compare_modes_iso_kx10} for all cases. Higher  
+order modes, corresponding to increasing $|\omega|$, increasingly
+deviates from the analytic line due to the imposed boundaries, 
+unless the vertical domain  size is increased \citep{barker15}.
 
@@
 
-The fundamental mode, marked by `f', corresponds to the eigenvalue
-with the smallest $|\sigma|$. The fundamental mode is
-robust against the vertical boundary conditions considered as it
-appears at the same position in all cases.
-
-Fig. \ref{lowfreq_eigenfunc} plots the eigenfunctions $W$ and $\delta 
-v_z$ for the fundamental mode with $\khat=10$ (which has the 
-maximum growth rate in Fig. \ref{iso_eigen_kx}). We find $W\propto z$
-throughout most of the disk, corresponding to a constant vertical
-velocity perturbation, as expected for the fundamental mode (see
-Appendix \ref{iso_explicit}). We show a meridional visualization of
-this mode in Fig. \ref{lowfreq_eigenfunc_2d}. The $x$ axis is
+For later comparison, we plot in Fig. \ref{lowfreq_eigenfunc} the 
+eigenfunctions $W$ and $\delta   v_z$ for the fundamental mode. We also show a 
+meridional visualization in Fig. \ref{lowfreq_eigenfunc_2d}. The $x$ axis is
 stretched for clarity; radial velocities are in fact typically much 
@@
 
-Higher order modes along
-the analytic lines have larger growth rates, but require larger
-vertical domains, since they have increasing number of nodes  $z$.
+\subsubsection{Comment on `surface modes'}\label{surf_comment} 
+Fig. \ref{compare_modes_iso_kx10} lack `surface modes' found by 
+\cite{nelson13}, \cite{mcnally14}  and by \cite{barker15}.  These are
+a class of unstable modes associated with vertical shear near the disk
+boundaries.
 
-The `tail' end of modes with decreasing $\nu$ to the right of
-Fig. \ref{compare_modes_iso_kx10} are in fact also unstable inertial modes, like those
-along the analytic lines, but are strongly affected by the presence of 
-the imposed vertical boundaries \citep{barker15}.
-
-% They have lower growth rates than modes consistent with our
-% analysis (i.e. those along the lines). These modes differ significantly between
-% cases and are thus likely associated with boundary conditions. 
-% We do
-% not consider them  
-% further.
-
-\subsubsection{Comment on `surface modes'}\label{surf_comment} 
-Our fiducial setup with $\khat=10$ in 
-Fig. \ref{compare_modes_iso_kx10} (black dots)  lack `surface modes'
-found by \cite{nelson13}, \cite{mcnally14}  and 
-by \cite{barker15}.  We illustrate surface modes with the mode
+We illustrate surface modes with the mode
 diagram for $\khat=30$ in Fig. \ref{compare_modes_iso_kx30}. Surface
@@
 in such diagrams. They are readily identified for  $\zmax\geq 5H$ and
-are marked by `s'.  We typically find an increasing
-number of them with increasing $\zmax$ and/or $\khat$,
+are marked by `s'.  Fig. \ref{lowfreq_eigenfunc_surf} shows an example
+of a surface mode in the fiducial disk. We typically find an increasing
+number of surface modes with increasing $\zmax$ and/or $\khat$,
 consistent with \cite{barker15}.
 
-As explained by \cite{barker15}, the occurence of surface modes in the
-vertically isothermal disk is due to imposing a finite vertical domain
-size. The fastest growing mode is a surface mode, if they
-exist, but their association with vertical boundaries means that
-$\mathrm{max}(\nu)$ depends on the choice of $\zmax$. Thus, the
-vertically isothermal disk is not useful for studying surface modes. 
+As explained by \cite{barker15}, the occurence of surface modes 
+occurence in the vertically isothermal disk is due to imposing a
+finite vertical domain size. The fastest growing mode is a surface
+mode, if they exist, but their association with vertical boundaries
+means that $\mathrm{max}(\nu)$ depends on the choice of $\zmax$.   
+We refer the reader to \citeauthor{barker15} for a
+detailed analysis of surface modes.
+
+% Thus, the
+% vertically isothermal disk is not useful for studying surface modes. 
 %Thus, seeking the most unstable mode in a vertically isothermal disk
 %may not yield robust results because this
-
-We will see in the next section that surface modes, as well
-as other higher order modes, are more effectively stabilized by finite
-thermal relaxation than the fundamental mode. Since we are primarily
-interested in characterizing a maximum thermal timescale that allows
-the VSI, we can focus on the fundamental mode. For a detailed analysis of 
-surface modes, see \cite{barker15}.
-
-%refer the reader to BL15 for surface modes
 
@@
     marked by `s'. \label{compare_modes_iso_kx30}
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/eigenvectorvz_iso_surf}
+  \caption{Example of a `surface mode' in the fiducial disk model.
+    \label{lowfreq_eigenfunc_surf}
   }
@@
 $\beta_\mathrm{crit}$ nevertheless provides a reasonable estimate for
-an upper limit to the thermal timescale that allows the VSI. We thus
+an upper limit to the thermal timescale for the VSI. We further 
 test the dependence of $\beta_\mathrm{crit}$ on disk parameters
@@
 point and vary $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and
-$\epsilon\in[0.02,0.1]$ separately. We consider the fundamental
-corrugation mode for which $\beta_\crit$ is the actual critical
-thermal timescale.%  We use $\khat=10$ to avoid complications to our
+$\epsilon\in[0.02,0.1]$ separately.
+
+% We consider the fundamental
+% corrugation mode for which $\beta_\mathrm{crit}$ is the actual critical
+% thermal timescale. 
+% We use $\khat=10$ to avoid complications to our
 % eigenvalue search from surface modes.
 
- The numerically obtained $\beta_\mathrm{crit}$ is shown in
- Fig. \ref{bcrit_compare} in comparison with Eq. \ref{iso_vsi_cond}.
- Our numerical results generally agree with Eq. \ref{iso_vsi_cond}. The
- agreement improves with decreasing  $|q|$,  $\epsilon$ and increasing
- $\gamma$, i.e. weaker instability. There is a noticeable difference
- as $\gamma\to1$ because the derivation of $\beta_\mathrm{crit}$
- assumed $\gamma>1$.
+The numerically obtained $\beta_\mathrm{crit}$ are shown in
+Fig. \ref{bcrit_compare} in comparison with Eq. \ref{iso_vsi_cond},
+which generally agree. This result, together with 
+Fig. \ref{bcrit_compare1}, shows that 
+$\beta \lesssim \beta_\mathrm{crit}$ is a practical condition under  
+which the VSI will operate effectively in stably stratified,
+vertically isothermal disks.
 
-Having validated
-
- % Fig. \ref{bcrit_compare}, together with Fig. \ref{bcrit_compare1} and 
- % our finding that the fundamental mode becomes dominant as one
- % increases $\beta$, and that it is not sensitive to boundary
- % conditions, shows that $\beta_\mathrm{crit}$ as given  by
- % Eq. \ref{iso_vsi_cond} provides a practical measure for the thermal
- % relaxation timescale below which  VSI will operate in a vertically
- % isothermal disk.
+% The agreement improves with decreasing  $|q|$,
+% $\epsilon$ and increasing $\gamma$, i.e. weaker instability. There is
+% a noticeable difference as $\gamma\to1$ because the derivation of
+% $\beta_\mathrm{crit}$ assumed $\gamma>1$.
 
@@
 
-\subsection{Vertically non-isothermal disks} 
+\subsection{Vertically non-isothermal disk model} 
 %We briefly consider vertically non-isothermal disks in which there is
@@
 
+Fig. \ref{bcrit_compare2} plots the maximum VSI growth rates in the
+vertically non-isothermal disk model.  Based on
+Eq. \ref{iso_vsi_cond}, we conjecture a critical
+thermal timescale given by $\epsilon|s|/(\gamma - \Gamma)$. This
+evaluates to 0.158, which is close to the cut-off thermal timescale
+for $\khat\in[5,30]$. Fig. \ref{bcrit_compare2} is qualitatively quite
+similar to the fiducial disk in  Fig. \ref{bcrit_compare1}.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/gcorr_compare_vnoniso_maxrate} 
+  \caption{Maximum VSI growth rates in the vertically non-isothermal disk
+    model as a function of the thermal relaxation timescale
+    $\beta$. The vertical line is a conjectured 
+    critical thermal timescale given by
+    $\epsilon|s|/(\gamma-\Gamma)$.  
+    \label{bcrit_compare2}}   
+\end{figure}
+
   % We briefly consider vertically non-isothermal disks with 

### 2015-03-30 application.tex
waiting for zmax3 max rate fig

@@
 the scope of the present work. Here, we take the simpler approach of
-modeling the linearized RHS of Eq. \ref{real_energy} such that it can 
-be represented by a thermal relaxation time $t_c$. 
-To do so, we first consider separately perturbations
-with radial lengthscales $l\gg l_\mathrm{rad}$ and  
+representing the linearized RHS of Eq. \ref{real_energy} by a thermal
+relaxation time $t_c$. To do so, we first consider separately
+perturbations with radial lengthscales $l\gg l_\mathrm{rad}$ and  
 $l\ll l_\mathrm{rad}$, where      
@@
 \end{align}
-where $\eta=k_\mathrm{rad}/\rho c_v$ is the diffusion coefficient and
+\citep[see also][]{barker15} where $\eta=k_\mathrm{rad}/\rho c_v$ is
+the diffusion coefficient and 
 is given in terms of physical disk parameters as 
@@
 \subsection{A model for combined thermal relaxation}\label{toy_relax}
-To unify $t_\mathrm{cool}$ and $t_\mathrm{diff}$, we define the
-effective thermal relaxation timescale in linear theory as
+%To unify $t_\mathrm{cool}$ and $t_\mathrm{diff}$, 
+We define the effective thermal relaxation timescale in linear theory as
 \begin{align}\label{tc_def}
@@
 %becomes optically thin and $\beta<\beta_\mathrm{crit}$. 
-Our analytical criteria is not valid for variable $\beta$, and we must
+Our analytical criteria is not valid for non-uniform $\beta$, and we must
 turn to numerical solutions of the linear problem to determine whether
@@
 We remark that it is possible to satisfy $\beta < \beta_\mathrm{crit}$ 
-for sufficiently large $\khat$. We
-illustrate this in Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at 
-each radius the value of $\khat$ such  
-that $\beta = \beta_\mathrm{crit}$. We thus expect that the VSI would only
-occur on very small radial scales in the inner disk.
+for sufficiently large $\khat$. We illustrate this in 
+Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at each radius the value
+of $\khat$ such that $\beta = \beta_\mathrm{crit}$, which show that in
+the inner disk, say $r\sim 1$AU, the VSI would only occur on very
+small radial lengthscales. 
+%surface modes can develop for large $\khat$, 
+%which have small vertical lengthscales scales. 
+%also small vertical scales if they are surface modes
 
@@
 surface modes towards the inner disk when considering large 
-wavenumbers $\khat\gtrsim 50$. However,
-surface modes are affected by boundary conditions and disturbances
-with large $\khat$ are more likely subject to viscous decay, as
-discussed above. The dominance of the fundamental mode is also
-observed in hydrodynamic simulations of \cite{stoll14} which treat
-radiative cooling properly.
+wavenumbers $\khat\gtrsim 50$. However, surface modes are dependent on
+boundary conditions in a vertically isothermal disk \citep{barker15},  
+and disturbances with large $\khat$ are more likely subject to viscous
+decay, as discussed above. % It is also unclear if surface modes are
+% dynamically important given they are concentrated in the disk
+% atmosphere where the density is small. 
+% The dominance of the fundamental mode is also
+% observed in hydrodynamic simulations of \cite{stoll14} which treat
+% radiative cooling properly.   unclear if surface modes are
+% dynamically important because the there is little mass in the disk at
 
-We can therefore consider Fig. \ref{mmsn_overall} as a representation of the
+We therefore consider Fig. \ref{mmsn_overall} as a representation of the
 VSI in the MMSN. We see the VSI operates from $\sim 5$AU to $\sim
-50$AU with growth timescales $\sim 30$---40 orbits. It is necessary to
-consider smaller scales with decreasing $r$ as the disk generally becomes
-optically thick, and growth times rise rapidly towards the inner disk.
+50$AU with growth timescales $\sim 30$---40 orbits and
+$t_\mathrm{grow}$ rises rapidly towards the inner disk. It is
+necessary to consider smaller scales with decreasing $r$ as the disk
+generally becomes optically thick.
 
@@
 \beta_\mathrm{crit}$ away from the midplane.
+
+%surface modes may develop for large kx, they are not affected
 

### 2015-03-30 numerical.tex
waiting for zmax3 max rate fig

@@
 \end{align}
-%note that there are g_c terms in the boundary condition
+% note that there are g_c terms in the boundary condition
 where $\bm{\xi}$ is the Lagrangian displacement with meridional 
@@
 
-Unless otherwise stated, our fiducial disk model is a nearly 
-vertically isothermal disk with $\Gamma=1.011$, stable
-stratification $\gamma=1.4$, $(p,q,\epsilon)=(-1.5,-1,0.05)$, and 
-vertical domain size $\zmax=5H$. In Fig. \ref{omega_z} we plot the
-corresponding equilibrium rotation profile, which shows that $\Omega$
-decreases slightly away from the mid-plane.
+Unless otherwise stated, our fiducial disk parameters are 
+$(\gamma, \Gamma) = (1.4, 1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.05)$, and 
+vertical domain size $\zmax=5H$. This setup is effectively vertically
+isothermal since $c_s(\zmax) \simeq 0.9c_s(0)$. In Fig. \ref{omega_z}
+we plot the corresponding equilibrium rotation profile, which shows
+that $\Omega$ decreases slightly away from the midplane.
 
@@
 we review this case before considering the effect of 
-thermal relaxation.
+thermal relaxation. A more detailed analysis is given by \cite{barker15}.
 
@@
 waves made unstable by vertical shear and have been termed `body
-modes' \citep{barker15}.  The `fundamental corrugation mode', marked by `f',
+modes' by \cite{barker15}.  The `fundamental corrugation mode', marked by `f',
 corresponds to the eigenvalue with the smallest $|\sigma|$. Hereafter
-we will simply refer to this as the fundamental mode. The
-fundamental mode is robust against vertical boundary conditions 
-as it appears at the same position in
-Fig. \ref{compare_modes_iso_kx10} for all cases. Higher  
+we will simply refer to this as the fundamental mode, and it is robust
+against vertical boundary conditions as it appears at the same
+position in Fig. \ref{compare_modes_iso_kx10} for all cases. Higher  
 order modes, corresponding to increasing $|\omega|$, increasingly
 deviates from the analytic line due to the imposed boundaries, 
-unless the vertical domain  size is increased \citep{barker15}.
+unless $\zmax$ is increased \citep{barker15}.
 
@@
     perturbation $\delta v_z$ of the fundamental VSI with 
-    $\khat=10$ in the fiducial disk with $\beta=10^{-3}$. The real 
+    $\khat=10$ in the fiducial disk with rapid thermal relaxation $\beta=10^{-3}$. The real  
     (imaginary) parts of $W$ and $\delta v_z$ are plotted as solid 
@@
 Fig. \ref{compare_modes_iso_kx10} lack `surface modes' found by 
-\cite{nelson13}, \cite{mcnally14}  and by \cite{barker15}.  These are
-a class of unstable modes associated with vertical shear near the disk
-boundaries.
+\cite{nelson13}, \cite{mcnally14}  and \cite{barker15}. 
+This a class of unstable modes associated with vertical shear near the
+disk boundaries \citep{barker15}. They are also absent in our  
+analysis because there we assumed a disk with infinite 
+vertical extent, as the vertically isothermal disk has no
+surface. However, surface modes can occur in the present calculations 
+because there is an imposed vertical boundary.
 
-We illustrate surface modes with the mode
-diagram for $\khat=30$ in Fig. \ref{compare_modes_iso_kx30}. Surface
+We illustrate surface modes with the mode diagram for $\khat=30$ in
+Fig. \ref{compare_modes_iso_kx30}. Surface 
 modes lie almost parallel to the vertical axis
@@
 are marked by `s'.  Fig. \ref{lowfreq_eigenfunc_surf} shows an example
-of a surface mode in the fiducial disk. We typically find an increasing
+of a surface mode. We typically find an increasing
 number of surface modes with increasing $\zmax$ and/or $\khat$,
@@
 
-As explained by \cite{barker15}, the occurence of surface modes 
-occurence in the vertically isothermal disk is due to imposing a
-finite vertical domain size. The fastest growing mode is a surface
-mode, if they exist, but their association with vertical boundaries
-means that $\mathrm{max}(\nu)$ depends on the choice of $\zmax$.   
-We refer the reader to \citeauthor{barker15} for a
-detailed analysis of surface modes.
+As explained by \cite{barker15}, while surface modes have a physical
+origin, their occurence in the vertically isothermal disk is due to
+imposing a finite vertical domain size. The fastest growing mode is a
+surface mode, if they exist, but their association with vertical
+boundaries means that $\mathrm{max}(\nu)$ depends on the choice of
+$\zmax$. This has important implications when interpreting results
+obtained from the vertically isothermal disk model. We return to this
+issue later.
 
-% Thus, the
-% vertically isothermal disk is not useful for studying surface modes. 
+%Thus, the
+%vertically isothermal disk is not useful for studying surface modes. 
 %Thus, seeking the most unstable mode in a vertically isothermal disk
@@
 We now consider the effect of finite thermal relaxation by increasing 
-$\beta$. Note that for the fiducial disk model the critical thermal 
-timescale $\beta_\mathrm{crit} = 0.125$.
-
+$\beta$. % Note that for the fiducial disk model the critical thermal 
+% timescale $\beta_\mathrm{crit} = 0.125$.
 Fig. \ref{compare_modes_cool_kx10} show unstable modes with $\khat=10$ 
@@
 analytic results computed from Eq. \ref{relax_disp}.  %analytic
-                                %results for low order modes
+% results for low order modes
 
@@
 $\beta=0.01$ as there are modes with $M>0$ that have larger growth 
-rates. However, by comparing with results for $\zmax=3H$, we see that these
-higher order modes are strongly influenced by boundary conditions, and
-there is a significant mismatch between numerical and analytical
-results for $\zmax=3H$ for $M>1$.
+rates. As remarked above, these higher order modes are affected by
+$\zmax$, e.g. there is a significant mismatch between numerical and 
+analytical results for $M\geq1$ when $\zmax=3H$ (upper panel).
+
+%However, by comparing with results for $\zmax=3H$, we see that
 
 However, the fundamental mode becomes dominant for $\beta \geq 
-0.03$. Higher order modes, with larger growth rates at small $\beta$,
+0.03$. Higher order modes, which would be dominant at small $\beta$, 
 are more effectively stabilized with increasing $\beta$. For 
-$\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we only find the
-fundamental mode. This is consistent with the discussion in 
-\S\ref{iso_vsi_beta_crit}.
+$\beta=0.1$, which is close to the critical thermal timescale
+$\beta_\mathrm{crit} (=0.125$ for the fiducial disk
+parameters), we only find the fundamental mode. This is consistent with
+the discussion in \S\ref{iso_vsi_beta_crit}.
 
@@
 found in the fiducial disk model as a function of the thermal
-relaxation timescale $\beta$ for several values of $\khat$. 
-We also plot the critical timescale $\beta_\mathrm{crit}=0.125$ for
-the fiducial disk parameters.
+relaxation timescale $\beta$ for several values of $\khat$ and
+$\zmax$. We also plot the critical timescale
+$\beta_\mathrm{crit}=0.125$ for the fiducial disk parameters.
 
@@
 $\beta_\mathrm{crit}$ was derived assuming $\khat^2\gg 1$). For
-$\khat=5,\,10$ we find $\beta_\mathrm{crit}$ accurately predicts the
-thermal timescal beyond which the VSI is suppressed.
+$\khat=5,\,10$ (and $30$ for $\zmax=3H$), we find
+$\beta_\mathrm{crit}$ accurately predicts the thermal timescal beyond
+which the VSI is suppressed.
 
 Complications from vertical boundaries arise when considering large
-$\khat$. For $\khat=30$ we find for $\beta\gtrsim 0.1$ the dominant
-disturbance correspond to the leftmost eigenvalues
+$\khat$. For $\khat=30$ and $\zmax=5H,\,7H$ we find for $\beta\gtrsim
+0.1$ the dominant disturbance correspond to the leftmost eigenvalues
 seen in the lower two panels of Fig. \ref{compare_modes_cool_kx30},
-which depend on $\zmax$. For $\khat=50,\,100$ we find surface modes
-dominate.
+which do not exist for $\zmax=3H$. For $\khat=50,\,100$ we find
+surface modes dominate.
 
@@
 0.6$ stabilized the VSI; as Fig. \ref{bcrit_compare1} show that growth
-rates for $\beta\gtrsim 0.6$ are small. One might consider 
-$\khat > 100$ to increase the maximum growth rate (note that the most unstable
-mode for large $\khat$ are typically surface modes), but in a realistic disk
+rates for $\beta\gtrsim 0.6$ are small. To destablize the disk with
+$\beta\gtrsim O(1)$, one might consider 
+$\khat > 100$ to increase the maximum growth rate by enabling surface
+modes, but in a realistic disk
 very small scale disturbances will likely be subject to viscous decay
-(\S\ref{application}). Numerical simulations are also unlikely to
+(see \S\ref{application}). Numerical simulations are also unlikely to
 resolve such small scales.
 
-As pointed out by \cite{barker15} surface modes
-only appear in vertically isothermal disks when vertical boundaries
-are imposed, so they are strongly dependent on $\zmax$.  Thus, seeking
-the most unstable mode in a vertically isothermal disk may not yield
-robust results because such modes, which are often surface modes, may
-depend on boundary conditions.
-
 \begin{figure}
-   \includegraphics[width=\linewidth]{figures/gcorr_compare_iso_maxrate} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z5}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z7} 
   \caption{Maximum VSI growth rates in the fiducial disk
@@
 
-Despite this caveat, and that surface modes are not accounted in our
-analysis,  we see from Fig. \ref{bcrit_compare1} that 
-$\beta_\mathrm{crit}$ nevertheless provides a reasonable estimate for
-an upper limit to the thermal timescale for the VSI. We further 
-test the dependence of $\beta_\mathrm{crit}$ on disk parameters
+As pointed out by \cite{barker15}, in vertically isothermal disks with
+an imposed boundary, surface modes are dependent on $\zmax$, but they
+are often the most unstable. Despite this caveat, and that surface
+modes are not accounted in our analysis, we see from
+Fig. \ref{bcrit_compare1} that $\beta_\mathrm{crit}$ nevertheless
+provides a reasonable estimate for an upper limit to the thermal
+timescale for the VSI. 
+%It becomes an accurate criter when considering fixed
+%$\khat$
+
+We further  test the dependence of $\beta_\mathrm{crit}$ on disk parameters
 (Eq. \ref{iso_vsi_cond}).  We use the fiducial setup as a  reference
 point and vary $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and
-$\epsilon\in[0.02,0.1]$ separately.
-
-% We consider the fundamental
-% corrugation mode for which $\beta_\mathrm{crit}$ is the actual critical
-% thermal timescale. 
-% We use $\khat=10$ to avoid complications to our
-% eigenvalue search from surface modes.
-
-The numerically obtained $\beta_\mathrm{crit}$ are shown in
-Fig. \ref{bcrit_compare} in comparison with Eq. \ref{iso_vsi_cond},
-which generally agree. This result, together with 
-Fig. \ref{bcrit_compare1}, shows that 
-$\beta \lesssim \beta_\mathrm{crit}$ is a practical condition under  
-which the VSI will operate effectively in stably stratified,
+$\epsilon\in[0.02,0.1]$ separately. The numerically obtained
+$\beta_\mathrm{crit}$ are shown in Fig. \ref{bcrit_compare} in
+comparison with Eq. \ref{iso_vsi_cond}, which generally agree. This
+result, together with Fig. \ref{bcrit_compare1}, shows that  
+$\beta \lesssim \beta_\mathrm{crit}$ can be used as a practical
+criteria for the VSI to operate effectively in stably stratified,
 vertically isothermal disks.
@@
 parameter in numerical calcluations. On the other hand, vertically
-non-isothermal 
-disks have a disk surface at finite height so there is a natural
-choice for $\zmax$. This means that higher order modes
-and surface modes are better defined in a vertically non-isothermal
-disk. However, note that this issue is unimportant for low order modes 
-such as the fundamental mode, which are not sensitive to boundary
-conditions.
+non-isothermal disks have a disk surface at finite height so there is
+a natural choice for $\zmax$. This means that higher order body modes and
+surface modes are better defined in a vertically non-isothermal disk.
+
+% However, note that this issue is unimportant for low order modes 
+% such as the fundamental mode, which are not sensitive to boundary
+% conditions. This is useful because when the fundamental mode is
+% stabilized as $\beta$ is increased, higher order modes will have
+% already been stabilized, the the growth rate of surface modes also
+% significantly reduced. Then as far as a characteristic upper limit to
+% the thermal timescale is conconcered, we may just consider the
+% fundamental mode.
 
@@
 qualitatively similar to Fig. \ref{compare_modes_cool_kx10}. As before
-we find as $\beta$ is increased, higher order modes are rapidly
+we find as $\beta$ is increased, higher order body modes are rapidly
 stabilized, eventually only the fundamental mode can operate before
@@
 is a well-defined surface, we may seek the 
-most unstable mode without the boundary-dependent issue that would
-arise in a vertically isothermal disk.
+most unstable mode without the boundary-dependent issue that arose 
+with a vertically isothermal disk.
 
@@
   \caption{Same as Fig. \ref{compare_modes_vnoniso_kx10} but with
-    $\khat=100$. In this case surface modes appear and are readily identified for
-    $\beta=10^{-3}$ and nearly
-    parallel to the vertical axis
+    $\khat=100$. In this case surface modes appear nearly parallel to
+    the vertical axis
+    (cf. Fig. \ref{compare_modes_vnoniso_kx10}). However, this
+    distinction becomes difficult for increasing $\beta$ as modes
+    overlap. 
+    % but
+    % overlap with body modes (those that are roughly horizontal when 
+    % $\beta=10^{-3}$) as $\beta$ is further increased. 
     \label{compare_modes_vnoniso_kx100}}
@@
 with $\khat=100$ and $\beta=10^{-3},\, 10^{-2}$ and
-$0.1$. At such large $\khat$, we find with $\beta=10^{-3}$ the most
+$0.1$. We find with $\beta=10^{-3}$ the most 
 unstable mode is an asymmetric surface mode \citep[also seen in Fig. 7
@@
 
-Fig. \ref{bcrit_compare2} plots the maximum VSI growth rates in the
-vertically non-isothermal disk model.  Based on
+Finally, in Fig. \ref{bcrit_compare2} we plot the maximum VSI growth
+rates in the vertically non-isothermal disk model.  Based on 
 Eq. \ref{iso_vsi_cond}, we conjecture a critical
@@
 for $\khat\in[5,30]$. Fig. \ref{bcrit_compare2} is qualitatively quite
-similar to the fiducial disk in  Fig. \ref{bcrit_compare1}.
+similar to the fiducial disk model in 
+Fig. \ref{bcrit_compare1}. Importantly, the maximum thermal
+timescale for $\khat=O(10)$ in fact provides a 
+characteristic upper limit for other values of $\khat$.
 
@@
     critical thermal timescale given by
-    $\epsilon|s|/(\gamma-\Gamma)$.  
+    $\epsilon|s|/(\gamma-\Gamma)=0.158$.  
     \label{bcrit_compare2}}   

### 2015-03-30 paper.tex
waiting for zmax3 max rate fig

@@
   astrophysical disks. However, the VSI requires a short thermal
-  timescale in stably stratified disks.  
+  timescale in sub-adabatically or stably stratified disks.  
   A quantitative examination of this thermodynamic 
-  requirement is thus necessary to assess the viability of
+  requirement is needed to assess the viability of
   the VSI in realistic disks. To this end,  
@@
   vertically global models of stably stratified accretion disks in which the disk  
-  temperature $T$ is relaxed its initial value over a characteristic timescale 
+  temperature $T$ is relaxed to its initial value over a characteristic timescale 
   $t_c=\beta\Omega_k^{-1}$, where $\beta$ is a dimensionless parameter 
@@
   $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  ($r$
-  being the cylindrical radius from the central object), and 
+  being the cylindrical radius from the central star), and 
   $\gamma$ is the adiabatic index. 
@@
   estimate for the thermal timescale $\beta$ in the MMSN based on dust
-  opacity. We find the MMSN is unstable to the VSI from $\sim 5$AU to $\sim 50$AU with  
-  a characteristic growth time of $\sim 30$ local orbital 
-  periods, and the dominant VSI mode in the MMSN is the `fundamental
-  corrugation mode' which causes the disk column to oscillate
-  vertically. We suggest the VSI is dynamically important in the outer
-  parts of protoplanetary disks. 
+  opacity. We find the MMSN is unstable to the VSI from $\sim 5$AU to $\sim 50$AU
+  with a characteristic growth time of $\sim 30$ local orbital periods. 
+  % ,and the dominant VSI mode in the MMSN is the `fundamental
+  % corrugation mode' which causes the disk column to oscillate
+  % vertically.
+  We thus suggest the VSI is dynamically important, at least in the
+  outer of protoplanetary disks at tens of AU. 
 \end{abstract}
@@
 isothermal perturbations, for which the thermal timescale is formally
-zero, and/or neutrally stratified disks in which the stabilizing
-effect of buoyancy is absent. We account for both effects to study the
+zero, and/or neutrally stratified disks in which buoyancy forces are
+absent by construction. We account for both effects to study the
 VSI in more general disk models. To do so, we examine 

### 2015-03-30 summary.tex
waiting for zmax3 max rate fig

@@
 height dependence of the equilibrium angular velocity profile
-$\Omega=\Omega(r,z)$ in a baroclinic disk. We examined how
-the linear instability depends on disk thermodynamics and applied our
-results to assess the relevance of the VSI in realistic protoplanetary
-disks.
+$\Omega=\Omega(r,z)$ in a baroclinic disk. Vertical shear is generally
+expected in astrophysical disks due to the radial dependence of the
+disk temperature, or more generally the entropy. However, the VSI
+requires a short thermal timescale in sub-adiabatically stratified
+disks, which is also typical of astrophysical disks. Physically, a
+short thermal timescale is needed to overcome the stabilizing
+influence of buoynancy forces.
+
+We thus examined how the linear VSI depend on disk
+thermodynamics. Our primary goal was to quantify a characteristic
+thermal timescale below which the VSI can operate in stably stratified
+disks, thereby provide a simple way to assess whether or not the VSI
+is relevant in realistic disks.
 
 We generalized previous vertically global linear 
-calculations of the VSI \citep{nelson13,mcnally14}, which assumed isothermal
-perturbations, to include an energy equation. Our effort may also be
-regarded as an extension of \cite{urpin03}, which did include an
-energy equation, but adopted the Boussinesq and vertically local
-approximations. We considered vertically global and fully
-compressible perturbations.
-
-We included a source term in the energy equation which relaxes the disk
+calculations of the VSI \citep{nelson13,mcnally14,barker15}, which
+assumed isothermal perturbations, to include an energy equation with % Our 
+% effort may also be regarded as an extension of the original VSI study
+% by \cite{urpin03} who did account for thermal evolution, but adopted 
+% the Boussinesq and vertically local approximations. We considered
+% vertically global and fully compressible perturbations. 
+a source term in that relaxes the disk
 temperature to its initial value on a timescale
@@
 perturbations with small radial lengthscales, we applied the 
-radially local approximation in the linearized fluid equations. For
-analytic discussion only, we further simplified the linear problem by
-ignoring the background radial disk structure. This is equivalent to adopting
-the `vertically global shearing box' (VGSB) formalism developed by
-\cite{mcnally14}. 
-We find the VGSB framework is appropriate for analyzing the VSI because 
-rapid thermal relaxation ($\beta\ll1$) is 
-generally required for the instability.   However, the VGSB framework is not
-appropriate in the adiabatic limit ($\beta \gg 1$).
+radially local approximation in the linearized fluid equations. % For
+% analytic discussion only, we further simplified the linear problem by
+% ignoring the background radial disk structure. This is equivalent to adopting
+% the `vertically global shearing box' (VGSB) formalism developed by
+% \cite{mcnally14}. 
+% We find the VGSB framework is appropriate for analyzing the VSI because 
+% rapid thermal relaxation ($\beta\ll1$) is generally required for the
+% instability to operate in stably stratified disks 
+% However, the VGSB framework is not
+% appropriate in the adiabatic limit ($\beta \gg 1$). 
+We mostly focused on vertically isothermal disks ($\Gamma=1$) as 
+this is the relevant case for protoplanetary disks that are passively
+irradiated by its central star \citep{chiang97}. It also permits a
+largely analytical discussion. Our analysis is an extension of the
+linear models considered in \cite{lubow93} by adding vertical shear
+and (rapid) thermal relaxation. We obtained explicit solutions
+describing the VSI with thermal relaxation in the thin disk limit.
 
-We mostly focused on vertically isothermal disks ($\Gamma=1$) as 
-this is the relevant case for protoplanetary disks that are passively irradiated by its central star 
-\citep{chiang97}. It also permits a largely analytical discussion.
-Our analysis is an extension of the linear models considered in
-\cite{lubow93} by  adding vertical shear and (rapid) thermal
-relaxation. We obtained explicit solutions describing the VSI with 
-thermal relaxation in the thin disk limit.
+% The fundamental VSI mode is characterized by a symmetric vertical
+% velocity perturbation with zero nodes and $\delta
+% v_z\sim\mathrm{constant}$ near the mid-plane, and by an anti-symmetric
+% density perturbation with one node at $z=0$.  These have been referred
+% to as `corrugation modes' by \cite{nelson13}, and were found to
+% dominate their numerical simulations, as well as that of
+% \cite{stoll14}. Furthermore, when computing unstable modes for a
+% protoplanetary disk model with thermal relaxation based on realistic
+% disk properties, we find the fundamental mode to be the dominant
+% mode. The fundamental mode is
+% thus likely the most relevant VSI mode in protoplanetary disks.
 
-The fundamental VSI mode is characterized by a symmetric vertical
-velocity perturbation with zero nodes and $\delta
-v_z\sim\mathrm{constant}$ near the mid-plane, and by an anti-symmetric
-density perturbation with one node at $z=0$.  These have been referred
-to as `corrugation modes' by \cite{nelson13}, and were found to
-dominate their numerical simulations, as well as that of
-\cite{stoll14}. Furthermore, when computing unstable modes for a
-protoplanetary disk model with thermal relaxation based on realistic
-disk properties, we find the fundamental mode to be the dominant
-mode. The fundamental mode is
-thus likely the most relevant VSI mode in protoplanetary disks.
+For a vertically isothermal disk we have determined a thermal
+relaxation timescale $\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$,
+below which perturbations may be regarded as isothermal and the VSI 
+operates effectively. While $\beta_\mathrm{crit}$ was derived for the
+lowest order VSI mode --- the `fundamental 
+corrugation mode' --- we find from numerical solutions to the  linear
+problem that $\beta_\mathrm{crit}$ in fact provides a useful way to
+characterize the stabilization of the VSI by a finite thermal 
+timescale. That is, all VSI growth rates are reduced significantly as
+$\beta$ is increased from zero to $\beta_\mathrm{crit}$.  
+% This is despite the fact the occurence of `surface
+% modes' associated with imposed vertical boundaries in numerical
+% models of vertically isothermal disks. These modes are absent in our
+% analysis since the vertically isothermal disk has no surface. 
+We have therefore provided a quantitative measure for the 
+`short' thermal timescale requirement for the VSI in astrophysical
+disks.
 
-We determined the critical thermal relaxation timescale 
-$\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$, below which small-scale
-perturbations may be regarded as isothermal and the VSI 
-operates. We have therefore provided a quantitative measure for the `fast'
-thermal relaxation requirement for the VSI in astrophysical disks.
+%Despite complications from imposing a vertical boundary in numerical
+%calculations in a vertically isothermal disk, which gives rise to
+%surface modes \citep{barker15} (and other boundary effects).
 
-We also solved the linear stability problem numerically using a
-pseudo-spectral approach. For numerical calculations we relax all the
-approximations made in our analysis, and accounted for
-the background radial disk structure.  Our
-numerical results are mostly consistent with our analytic 
-expectations, including the dependence of $\beta_\mathrm{crit}$ on
-disk parameters.  We find the fundamental mode is robust against
-boundary conditions, and higher order modes are more rapidly
-stabilized with increasing $\beta$.
-
-We briefly studied vertically non-isothermal disks. We confirm the finding by \cite{nelson13} that in the absence
-of vertical entropy gradients, the VSI does not require rapid thermal
-relaxation.  When a stable stratification is introduced, the
-rapid reduction in VSI growth rates by a finite thermal relaxation
-timescale is qualitatively similar to that observed for vertically
-isothermal disks. Based on numerical results, we conjecture that the
-critical thermal relaxation timescale is given by 
-$\beta_\mathrm{crit}=\epsilon|q|/(\gamma-\Gamma)$ in vertically 
-non-isothermal disks.
-
+% We also solved the linear stability problem numerically using a
+% pseudo-spectral approach. For numerical calculations we relax all the
+% approximations made in our analysis, and accounted for
+% the background radial disk structure.  Our
+% numerical results are mostly consistent with our analytic 
+% expectations, including the dependence of $\beta_\mathrm{crit}$ on
+% disk parameters.  
+% We find the fundamental mode is robust against
+% boundary conditions, and higher order modes are more rapidly
+% stabilized with increasing $\beta$. 
+% We briefly studied vertically non-isothermal disks. We confirm the finding by \cite{nelson13} that in the absence
+% of vertical entropy gradients, the VSI does not require rapid thermal
+% relaxation.  When a stable stratification is introduced, the
+% rapid reduction in VSI growth rates by a finite thermal relaxation
+% timescale is qualitatively similar to that observed for vertically
+% isothermal disks. Based on numerical results, we conjecture that the
+% critical thermal relaxation timescale is given by 
+% $\beta_\mathrm{crit}=\epsilon|s|/(\gamma-\Gamma)$ in vertically 
+% non-isothermal disks.  
 %application to MMSN
@@
 (MMSN) described in \cite{chiang10} and   
-modeled the thermal relaxation timescale $\beta$ based on realistic
+modeled the thermal timescale $\beta$ based on realistic
 disk properties and dust opacity. This results in a $\beta$ function that
@@
 from $\sim 5$AU to $\sim50$AU with characteristic growth times of
-$\sim 30$ orbits. Although $\beta$ is not vertically constant in the
+$\sim 30$ orbits. While it is possible to have the VSI at even
+smaller radii, this would require very small radial lengthscales, 
+which may be subject to viscous decay in a real disk. 
+%or  
+% higher growth rates by having
+% surface modes --- these are dependent
+% on boundary conditions in the
+% vertically isothermal disk
+
+Although $\beta$ is not vertically constant in the
 MMSN as assumed in our analysis, we find that evaluating the condition
@@
 radii at which the VSI operates in the MMSN. This simple criterion is
-thus a useful first step to estimate the relevance of the VSI in
-realistic astrophysical disks without resorting to non-linear
-numerical simulations.
+thus a practical first step in estimating the relevance of the VSI in
+realistic astrophysical disks without resorting to detailed linear
+analysis or non-linear numerical simulations.
 
@@
 \subsubsection{Improved thermal evolution} 
-We assumed non-adiabatic effects
-can be represented by a thermal relaxation timescale $t_c$. For 
+We assumed non-adiabatic effects can be represented by a thermal
+relaxation timescale $t_c$. For  
 application to protoplanetary disks we connected $t_c$ to the disk
@@
 
-We have neglected viscosity. \cite{nelson13} find that an 
-alpha viscosity $\lesssim O(10^{-4})$ is required for the VSI. 
-It is of interest to quantify the effect of viscosity 
-on the linear VSI. This would help identify, more accurately, the
-appropriate regions in a protoplanetary disk that are sufficiently
-laminar to develop the VSI.  This problem is, however, complicated by
-the fact that proper inclusion of viscosity introduces a meridional
-flow in the equilibrium state, as well as viscous heating
-in the energy equation.
+We have neglected viscosity. \cite{nelson13} find that an alpha
+viscosity $\lesssim O(10^{-4})$ is required for the
+VSI. \cite{barker15} found that inclusion of viscosity is needed for 
+convergence when studying surface modes. Viscosity may also affect the
+thermal timescale requirement of the VSI, because viscous heating
+needs to be self-consistently included in the energy equation. This
+should be investigated in future works.
+
+% It is of interest to quantify the effect of viscosity 
+% on the linear VSI. This would help identify, more accurately, the
+% appropriate regions in a protoplanetary disk that are sufficiently
+% laminar to develop the VSI.  This problem is, however, complicated by
+% the fact that proper inclusion of viscosity introduces a meridional
+% flow in the equilibrium state, as well as viscous heating
+% in the energy equation.

### 2015-03-30 numerical.tex
revising

@@
 vertical domain size $\zmax=5H$. This setup is effectively vertically
-isothermal since $c_s(\zmax) \simeq 0.9c_s(0)$. In Fig. \ref{omega_z}
+isothermal since $c_s^2(\zmax) \simeq 0.9c_s^2(0)$. In Fig. \ref{omega_z}
 we plot the corresponding equilibrium rotation profile, which shows
@@
 we review this case before considering the effect of 
-thermal relaxation. A more detailed analysis is given by \cite{barker15}.
+thermal relaxation. %A more detailed analysis is given by \cite{barker15}.
 
@@
 \begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z3}
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z5}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z7} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z7} 
   \caption{Maximum VSI growth rates in the fiducial disk 
      model as a function of the thermal relaxation timescale
-     $\beta$. The vertical line is the
+     $\beta$, for $\zmax=3H$ (top), $\zmax=5H$ (middle, standard
+     setup) and $\zmax=7H$ (bottom). The vertical line is the
      critical thermal timescale $\beta_\mathrm{crit}$ obtained  
@@
 H_s$). The disk temperature decreases rapidly with height such that 
-$c_s(\zmax)\simeq 10^{-4}c_s(0)$.
+$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$.
 

### 2015-03-31 paper.tex
revising

@@
   A robust hydrodynamic mechanism for turbulent angular momentum
-  transport is important for cold accretion disk such as 
+  transport is important for cold accretion disks such as 
   protoplanetary disks. The vertical shear instability (VSI) has 
@@
   disk midplane, which is generally expected in  
-  astrophysical disks. However, the VSI requires a short thermal
-  timescale in sub-adabatically or stably stratified disks.  
+  astrophysical disks. % with, for example, radial variations in temperature. 
+  However, the VSI requires a short thermal or cooling 
+  timescale in sub-adabatically (stably) stratified disks.  
   A quantitative examination of this thermodynamic 
-  requirement is needed to assess the viability of
+  requirement is thus needed to assess the viability of
   the VSI in realistic disks. To this end,  
@@
   `short' thermal timescale requirement for the VSI as $\beta < \epsilon|q|/(\gamma-1)$, where
-  $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  ($r$
+  $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  is the radial temperature gradient ($r$
   being the cylindrical radius from the central star), and 
   $\gamma$ is the adiabatic index. 
-  We apply our linear calculations  
+  We apply our results  
   to a disk model of the Minimum Mass Solar Nebula (MMSN) with an  
   estimate for the thermal timescale $\beta$ in the MMSN based on dust
-  opacity. We find the MMSN is unstable to the VSI from $\sim 5$AU to $\sim 50$AU
-  with a characteristic growth time of $\sim 30$ local orbital periods. 
+  opacity. In the inner parts of the MMSN, e.g. at $\lesssim 5$AU, 
+  the VSI will neccessarily have extremely short lengthscales $\ll 
+  O(10^{-2}H)$, where $H$ is the disk scale height. In the outer disk, 
+  we find the MMSN is unstable to the VSI from $\sim 5$AU to $\sim
+  50$AU with a characteristic lengthscale $O(10^{-2}H)$ to
+  $O(10^{-1}H)$ and a growth time of $\sim 30$ local orbital periods.  
   % ,and the dominant VSI mode in the MMSN is the `fundamental
@@
   We thus suggest the VSI is dynamically important, at least in the 
-  outer of protoplanetary disks at tens of AU. 
+  outer parts of protoplanetary disks at tens of AU. 
 \end{abstract}

### 2015-03-31 paper.tex
revised intro and abstract

@@
   `short' thermal timescale requirement for the VSI as $\beta < \epsilon|q|/(\gamma-1)$, where
-  $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  is the radial temperature gradient ($r$
-  being the cylindrical radius from the central star), and 
+  $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  is the radial temperature gradient, and 
   $\gamma$ is the adiabatic index. 
@@
   opacity. In the inner parts of the MMSN, e.g. at $\lesssim 5$AU, 
-  the VSI will neccessarily have extremely short lengthscales $\ll 
-  O(10^{-2}H)$, where $H$ is the disk scale height. In the outer disk, 
-  we find the MMSN is unstable to the VSI from $\sim 5$AU to $\sim
-  50$AU with a characteristic lengthscale $O(10^{-2}H)$ to
-  $O(10^{-1}H)$ and a growth time of $\sim 30$ local orbital periods.  
+  we find the VSI would only operate on extremely short lengthscales $\ll 
+  O(10^{-2}H)$, where $H$ is the disk scale height. On longer
+  lengthscales $O(10^{-2}H)$ to $O(10^{-1}H)$, we find the MMSN is
+  unstable to the VSI from $\sim 5$AU to $\sim 50$AU with a
+  characteristic growth time of $\sim 30$ local orbital periods.  
   % ,and the dominant VSI mode in the MMSN is the `fundamental
@@
   % vertically.
-  We thus suggest the VSI is dynamically important, at least in the 
+  We thus suggest the VSI is at least dynamically important in the 
   outer parts of protoplanetary disks at tens of AU. 
@@
 disks (PPDs) which do not support the MRI 
-\citep{gammie96,landry13}. In fact, recent  numerical simulations
+\citep{gammie96,landry13}. In fact, recent MHD simulations
 suggest significant portions of PPDs remain laminar \citep[e.g.][and references
-therein]{lesur14,bai15,gressel15}. Thus, alternative routes 
-to angular momentum transport in PPDs remain of interest
-\citep{armitage10}.
+therein]{lesur14,bai15,gressel15}. Thus, alternative mechanisms 
+to angular momentum transport in PPDs, such as purely hydrodynamic
+turbulence,  remain of interest \citep{armitage10}. In addition to
+mass accretion, hydrodynamic turbulence in PPDs will also
+impact dust evolution and planetesimal formation \citep{chiang10}.
 
-One possibility is purely hydrodynamic turbulence through the vertical
-shear instability (VSI). The VSI is an axisymmetric linear instability
+One possible route to hydrodynamic turbulence is the vertical shear
+instability (VSI). The VSI is an 
+axisymmetric linear instability 
 in three-dimensional (3D) disks in which the rotation frequency
@@
 \citep[GSF,][]{goldreich67,fricke68} in differentially rotating stars.   
-The height-dependence of $\Omega$, or vertical shear, can render
+A height-dependence in $\Omega$, or vertical shear, can render
 inertial waves unstable. \cite{barker15} gives a simple 
@@
 characterized by perturbations with large vertical length-scales 
-(relative to radial length-scales).
+relative to radial length-scales.
 
@@
 isothermally, unless the disk is neutrally stratified in the vertical
-direction. This is because vertical motions associated with the VSI
+direction. This is because
+vertical motions associated with the VSI 
 are stabilized by buoyancy forces if the disk is sub-adiabatically
@@
 timescale renders buoyancy ineffective.    
+%efficient exchang of heat with surroundings
 
-The need for a short thermal timescale in stably stratified disks has
-been demonstrated in numerical simulations of the VSI
+The need for a short thermal timescale for the VSI to operate in
+stably stratified disks has been demonstrated in recent numerical simulations 
 \citep{nelson13,stoll14}. When the  thermal timescale $t_c$ is
@@
 
-Given that the VSI is generally sensitive to disk thermodynamics,  
-it is important to understand and quantify this dependence, for
-example on disk parameters, in order to evaluate the 
-relevance of the VSI to realistic disks, in which one may expect
-non-zero vertical entropy gradients and thermal timescales. Should a
-given disk become turbulent through the VSI, it must first pass
-through the linear instability. Thus, understanding the thermodynamic
-dependence of the  linear VSI is essential.
+These studies show that the VSI is generally sensitive to disk
+thermodynamics. It is thus important to quantify this dependence, for
+example on disk parameters, in order to evaluate the relevance of the
+VSI to realistic disks, which one may expect to be stably
+stratified and have non-zero thermal timescales. For example, we show in  
+Fig. \ref{bcrit_mmsn2d} an estimate of the thermal timescales in a
+typical protoplanetary disk model (see below). In order to determine
+where in the disk the VSI may operate, and hence lead to hydrodynamic
+turbulence, understanding the thermodynamic dependence of the linear VSI is essential. 
+% since the transition
+% Should a disk
+% become turbulent through the VSI, it must first pass through the linear
+% instability.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/bcrit_mmsn2d} 
+  \caption{Estimate of the thermal timescale  in a fiducial
+    protoplanetary disk model, the Minimum Mass Solar Nebula, for
+    perturbations with radial lengthscales $\sim 10^{-2}H$, where $H$
+    is the disk scale-height. The red line corresponds to a 
+    characteristic thermal timescale, derived in this work, 
+    below which the VSI is expected to operate effectively. 
+    \label{bcrit_mmsn2d} 
+  }
+\end{figure}
 
@@
 particular, we quantify a characteristic upper limit to the thermal
-timescale for the VSI to operate. In \S\ref{numerical} we numerically
+timescale for the VSI. In \S\ref{numerical} we numerically
 solve the linear problem in the radially local approximation, but
@@
 the aforementioned characteristic thermal timescale. In
-\S\ref{application} we apply our linear models to the Minimum Mass
-Solar Nebula (MMSN) by modeling thermal relaxation
-timescales based on realistic disk properties, and find that the VSI may
-operate in the outer parts of the MMSN. We summarize in
+\S\ref{application} we apply our results to the Minimum Mass
+Solar Nebula (MMSN) by modeling thermal 
+timescales based on realistic disk properties, and find that the VSI
+can operate effectively in the outer parts of the MMSN. We summarize in
 \S\ref{summary} with a discussion of caveats and possible directions

### 2015-03-31 setup.tex
major revision in process because of barker & latter
revised intro and abstract

@@
 the equilibrium disk properties and the perturbation length-scales of
-interest.
+interest (e.g. Fig. \ref{bcrit_mmsn2d}).
 
@@
 \end{align} 
-These equations allows us to choose any mid-plane density 
+These equations allows us to choose any midplane density 
 distribution $\rho_0(r)\equiv \rho(r,0)$. In this work we assume a
@@
 specific entropy. The second equality applies to the disk equilibria
-described above. 
-We are interested in
-neutrally or sub-adiabatically stratified 
-disks with $\Gamma\leq \gamma$ so that $N_z^2\geq0$.
+described above. We are interested in sub-adiabatically stratified 
+disks with $\Gamma < \gamma$ so that $N_z^2\geq0$. % , and the disk is
+% convectively stable in the vertical direction.
 
@@
 
+% This approach has been adopted
+% in recent linear calculations of the VSI
+% \cite{nelson13,mcnally14,barker15} because unstable modes have short
+% radial lengthscales.
+
 \subsection{Radially-local approximation}  
@@
 so that we can ignore the radial variations in the background
-structure and radial gradients. We thus evaluate all coefficients in
+structure and radial gradients. \cite{barker15} have validated this
+approximation by comparison with the linear VSI in global cylindrical
+disk geometry. We thus evaluate all coefficients in
 Eq. \ref{full_mass1}---\ref{full_energy1} at the fiducial radius

### 2015-03-31 appendix.tex
middle of revising numerical section

@@
 \subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
-An important caveat of the VGSB is that it cannot be used in the
+A limitation of the VGSB is that it cannot be used in the
 adiabatic limit, even if the above conditions are satisfied. We explain this by considering vertically isothermal

### 2015-03-31 linear_adia.tex
middle of revising numerical section

--- 2015-03-28 linear_adia.tex
@@
 $\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
-and making the low-frequency, nearly-Keplerian approximation in
+and making the low-frequency, Keplerian approximation in
 \citeauthor{lubow93} after a change of variables.
@@
 to \cite{barker15} for a more detailed treatment of isothermal 
-perturbations.
+perturbations. %which also account for more general boundary
+               %conditions than that considered below
 
@@
   thermal relaxation time}\label{relax_pert}
-We are interested in the case $\beta\neq 0$, but it is instructive to
+We are interested in finite thermal relaxation
+timescales $\beta > 0$, but it is instructive to
 first ask  the more analytically tractable question: how 
 do the eigenfrequencies and eigenfunctions change when we change
-$t_c=\beta\Omega_k^{-1}$ from zero to a small but finite value? For 
+$\beta$ from zero to a small but finite value? For 
 sufficiently small $\beta$ we expect the solution to only differ
@@
 v_z=$constant and $\chi=1$ so that $B=C=0$. 
-This is the fundamental or `corrugation' mode where the vertical
-velocity is constant throughout the disk column. These modes have been
-observed to dominate numerical simulations \citep{nelson13,stoll14},  
-and are the ones we find to dominate in numerical calculations with 
-increasing $\beta$ (\S\ref{therm_relax_eff}), as well as in protoplanetary disks 
+This is the lowest order VSI mode, the `fundamental corrugation mode',
+where the vertical velocity is constant throughout the disk
+column. Hereafter we shall simply refer to them as the fundamental
+mode.
+
+The fundamental mode has been observed to dominate numerical simulations
+\citep{nelson13,stoll14}, and are the ones we find to typically 
+dominate in numerical calculations with increasing $\beta$
+(\S\ref{therm_relax_eff}) for moderately large    
+values of $\khat$, as well as in protoplanetary disks 
 with a realistic estimate for thermal timescales
-(\S\ref{application}).
+(\S\ref{application}). We will see later that consideration of low
+order modes in fact provides a useful way to charactertize the effect
+of increasing the thermal timescale on the VSI.
 
@@
 complicated and generally requires a numerical solution. However,
-simple solutions may be obtained in limiting cases which we consider
+simple results may be obtained in limiting cases which we consider
 below.
@@
 simultaneous equations for $\hat{\omega}$ and $\beta$ for marginal
-stability. If $\beta\ll 1$ and the disk is sufficiently thin so that  
+stability. (It is important to note that marginal stability is an
+assumption.) If $\beta\ll 1$ and the disk is sufficiently thin so that  
 $|\epsilon q \khat |^2\lesssim 1$, 
@@
 Note that for $q=0$, Eq. \ref{marg_freq_approx} reduces to 
-the dispersion relation for low-frequency waves in vertically
+the dispersion relation for low-frequency inertial waves in vertically
 isothermal disks without vertical shear (see Appendix
@@
 \end{align}
-where $\beta_\mathrm{crit}$
-is the critical thermal timescale for the fundamental mode
-($M=0$). We comment that the expression for $\beta_\mathrm{crit}$ may
-in fact be obtained without assuming $|\epsilon q\khat|\lesssim 1$.
-We will confirm $\beta_\mathrm{crit}$ numerically in 
-\S\ref{bcrit_num_test}.
+where $\beta_\mathrm{crit}$ is the critical thermal timescale for
+the fundamental mode ($M=0$). 
+We comment the expression for $\beta_\mathrm{crit}$ may in fact be
+obtained without assuming it is much less than unity from the onset or
+$|\epsilon q\khat|\lesssim 1$. % This is because for $M=0$, 
+% Eq. \ref{relax_cond_simp1} --- \ref{relax_cond_simp2} simplify 
+We will  
+confirm $\beta_\mathrm{crit}$ numerically in \S\ref{bcrit_num_test}.
 
@@
 Rapid thermal relaxation is required to overcome a stabilizing
-vertical entropy gradient \citep{goldreich67,urpin98,urpin03}, but these
-studies consider vertically localized disturbances. Our 
+vertical entropy gradient \citep{goldreich67,urpin98,urpin03}. 
 Eq. \ref{iso_vsi_cond} quantifies this requirement for vertically
@@
 that in a thin disk, a very short thermal timescale is required
-($\beta_\mathrm{crit}\ll 1$) because $\beta_\mathrm{crit}\propto \epsilon$.
+($\beta_\mathrm{crit}\ll 1$) because $\beta_\mathrm{crit}\propto
+\epsilon$.
 

### 2015-03-31 linear_iso.tex
middle of revising numerical section

@@
 
-\subsection{Low-frequency and nearly-Keplerian
+\subsection{Low-frequency and Keplerian
   approximation}\label{approx_gov} 
@@
   \kappa^2$. %\citep[e.g.][]{lin10}. 
-  We call this the low frequency approximation. %physically inertial
-                                %waves 
+  We call this the low frequency approximation. We are therefore
+  considering inertial waves \citep{barker15}.
+
 \item We are mainly interested in the effect of vertical shear, i.e. $q\neq
@@
   elsewhere. For clarity, we replace $\kappa^2$ with
-  $\Omega_k^2$. We call this the nearly-Keplerian approximation.
+  $\Omega_k^2$. We call this the Keplerian approximation.
 
@@
 replacement after eliminating variables, which involves $dD/dz$,
-introduces terms $O(\epsilon^2)$. Since we are interested in 
+introduces terms of $O(\epsilon^2)$. Since we are interested in 
 thin disks ($\epsilon\ll 1 $), e.g. protoplanetary disks where 

### 2015-03-31 numerical.tex
middle of revising numerical section

@@
 corresponding to Eq. \ref{lin_mass}---\ref{lin_energy}. 
-We relax the  low-frequency, nearly-Keplerian and thin-disk 
-approximations used in the analytical discussion above. We 
+We relax the  low-frequency, Keplerian and thin-disk 
+approximations used in the analytical discussion above and 
 use the full expressions of $\rho$, $\Omega$ and 
@@
 This setup mimics strictly isothermal perturbations in a vertically
-isothermal disk, as considered by \cite{nelson13}, \cite{mcnally14} and
-\cite{barker15}. %Note, however, that \citeauthor{nelson13} and
-% \citeauthor{barker15} employed the  approximations described in \S\ref{analytical} throughout their analysis, which are relaxed in our
-% numerical study.  
+isothermal disk, already considered by \cite{nelson13}, \cite{mcnally14} and
+\cite{barker15}. 
 For completeness, and as a way of introducing terminology, 
@@
 
-The modes shown in Fig. \ref{compare_modes_iso_kx10} are inertial
-waves made unstable by vertical shear and have been termed `body
-modes' by \cite{barker15}.  The `fundamental corrugation mode', marked by `f',
-corresponds to the eigenvalue with the smallest $|\sigma|$. Hereafter
-we will simply refer to this as the fundamental mode, and it is robust
+The modes shown in Fig. \ref{compare_modes_iso_kx10} are `body modes'
+and are inertial waves destabilized by vertical shear
+\citep{barker15}. (The possible exception being the fastest growing
+mode for $\zmax=7H$ which is a surface mode --- see the discussion
+below.)  The `fundamental corrugation mode', marked by `f',
+corresponds to the eigenvalue with the smallest $|\sigma|$. It is robust
 against vertical boundary conditions as it appears at the same
@@
 \subsubsection{Comment on `surface modes'}\label{surf_comment} 
-Fig. \ref{compare_modes_iso_kx10} lack `surface modes' found by 
+Fig. \ref{compare_modes_iso_kx10} lack the distinct `surface modes' found by 
 \cite{nelson13}, \cite{mcnally14}  and \cite{barker15}. 
 This a class of unstable modes associated with vertical shear near the
-disk boundaries \citep{barker15}. They are also absent in our  
+disk boundaries \citep{barker15}. They are absent in our  
 analysis because there we assumed a disk with infinite 
-vertical extent, as the vertically isothermal disk has no
+vertical extent, since the vertically isothermal disk has no
 surface. However, surface modes can occur in the present calculations 
@@
 origin, their occurence in the vertically isothermal disk is due to
-imposing a finite vertical domain size. The fastest growing mode is a
-surface mode, if they exist, but their association with vertical
-boundaries means that $\mathrm{max}(\nu)$ depends on the choice of
-$\zmax$. This has important implications when interpreting results
-obtained from the vertically isothermal disk model. We return to this
-issue later.
+imposing a finite vertical domain size. This has important
+implications when interpreting results obtained from the vertically
+isothermal disk model. For example, seeking the most unstable mode may
+result in a surface mode, but its growth rate would depend on
+$\zmax$. We return to this issue later.
 

### 2015-03-31 paper.tex
revised intro and abstract
middle of revising numerical section

@@
   A quantitative examination of this thermodynamic 
-  requirement is thus needed to assess the viability of
+  requirement is needed to assess the viability of
   the VSI in realistic disks. To this end,  
@@
   opacity. In the inner parts of the MMSN, e.g. at $\lesssim 5$AU, 
-  we find the VSI would only operate on extremely short lengthscales $\ll 
-  O(10^{-2}H)$, where $H$ is the disk scale height. On longer
-  lengthscales $O(10^{-2}H)$ to $O(10^{-1}H)$, we find the MMSN is
+  we find the VSI would only operate on extremely short radial lengthscales, $\ll 
+  O(10^{-2}H)$, where $H$ is the disk scale height. On longer radial 
+  lengthscales of $O(10^{-2}H)$ to $O(10^{-1}H)$, we find the MMSN is
   unstable to the VSI from $\sim 5$AU to $\sim 50$AU with a
@@
     characteristic thermal timescale, derived in this work, 
-    below which the VSI is expected to operate effectively. 
+    below which the VSI is expected to operate effectively (the region
+    outside the line).   
     \label{bcrit_mmsn2d} 

### 2015-03-31 setup.tex
middle of revising numerical section

@@
 timescales, in which case $\beta$ is non-constant and is determined by
-the equilibrium disk properties and the perturbation length-scales of
+the equilibrium disk properties and the perturbation lengthscales of
 interest (e.g. Fig. \ref{bcrit_mmsn2d}).
@@
 
-For discussion purposes it is useful to define the vertical buoyancy
-frequency $N_z$ via
-\begin{align}
-  N_z^2 \equiv - c_p^{-1}g_z \frac{\p S}{\p z} = c_s^2 \left(\frac{\gamma -
-      \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   \label{nzsq_def}
-\end{align}
-where $c_p$ is the heat capacity at constant pressure, $g_z\equiv
-\rho^{-1}\p_zP $ and $S\equiv c_p\ln{P^{1/\gamma}/\rho}$ is the
-specific entropy. The second equality applies to the disk equilibria
-described above. We are interested in sub-adiabatically stratified 
-disks with $\Gamma < \gamma$ so that $N_z^2\geq0$. % , and the disk is
-% convectively stable in the vertical direction.
-
 %\subsection{Vertical structure}
@@
   \frac{\rho(r,z)}{\rho_0(r)} &=
-  \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]}\\    
+  \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},\\    
   \frac{\Omega^2(r,z)}{\Omega_k^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
@@
 \end{align}
-Note that there is a maximum vertical shear rate at $z=r/\sqrt{2}$,
+Note that there is a maximum vertical shear rate at $|z|=r/\sqrt{2}$,
 which is expected to limit the growth rate of instabilities associated with vertical shear.
 
-\subsection{Stability condition against axisymmetric adiabatic
+\subsection{Entropy gradients and vertical buoyancy}
+The specific entropy is defined as $S\equiv
+c_p\ln{P^{1/\gamma}/\rho}$, where $c_p$ is the heat capacity at
+constant pressure. For our disk models,
+\begin{align}
+  &\frac{\p S}{\p r} = c_p\left[\frac{s}{\gamma r} +
+    \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p
+       r}\right],\\
+  &\frac{\p S}{\p z} = c_p
+  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z},\label{dSdz}
+\end{align} 
+which are used below.
+
+For discussion purposes it is useful to define the vertical buoyancy
+frequency $N_z$ via 
+\begin{align}
+  N_z^2 \equiv - c_p^{-1}g_z \frac{\p S}{\p z} = c_s^2 \left(\frac{\gamma -
+      \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   \label{nzsq_def}
+\end{align}
+where $g_z\equiv \rho^{-1}\p_zP $. The second equality applies to the
+disk equilibria described above (using Eq. \ref{dSdz}). We are
+interested in sub-adiabatically stratified disks with $\Gamma <
+\gamma$ so that $N_z^2\geq0$. % , and the disk is 
+% convectively stable in the vertical direction.
+
+\subsection{Stability against axisymmetric adiabatic
   perturbations}\label{solberg}
@@
 Thus, to assess stability, we consider the second
-criterion. For our disk models,
-\begin{align}
-  &\frac{\p S}{\p r} = c_p\left[\frac{s}{\gamma r} +
-    \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p
-       r}\right],\\
-  &\frac{\p S}{\p z} = c_p
-  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z}. 
-\end{align} 
+criterion. 
 Without loss of generality, we evaluate Eq. \ref{second_SH} in $z>0$

### 2015-03-31 numerical.tex
revising in progress by mkl

@@
 and are inertial waves destabilized by vertical shear
-\citep{barker15}. (The possible exception being the fastest growing
+\citep{barker15}. (The exception being the fastest growing
 mode for $\zmax=7H$ which is a surface mode --- see the discussion
@@
 
+%presence of buoyancy forces at the boundary may make BC more
+%influential
+
 In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable 
@@
 for the fundamental mode ($M=0$) and $M=1,\,2$, and omit the 
-squares. The fundamental mode is again insensitive boundary
-conditions compared to higher order modes, and are in close agreement
-with analytic results. As before we observe more effective
-stabilization for higher order modes, including the absence of surface
-modes (which only occur for $\zmax\geq 5H$) as $\beta$ is increased.  
-For $\zmax=3H$, the fundamental mode becomes dominant for $\beta\geq
-0.03$. For $\zmax=5H$ and $\zmax=7H$, we
+squares. The fundamental mode is again insensitive to $\zmax$ and are
+in close agreement with analytic results. However, 
+there is increasing mismatch between analytical and numerical results
+for higher order modes as $\beta$ is increased. This suggests that, in
+the presence of buoyancy, boundary effects become important for these
+modes, which have larger $\khat$ than that considered previously. For
+example, for $\zmax=5H$ and $\zmax=7H$, we 
 find modes slightly more unstable than the fundamental
 mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
-panels), but these are clearly associated with boundary
+panels), but these are clearly affected by boundary
 conditions as they are absent for $\zmax=3H$.
+
+Nevertheless, as before we observe more effective stabilization for
+higher order modes, including the absence of surface modes, for
+example as $\beta$ is increased from $0.01$ to $0.03$ (for $\zmax\geq
+5H$). If we discard modes associated with boundary conditions, then
+the fundamental mode would again become the dominant mode as $\beta$ is
+increased.
 
@@
 seen in the lower two panels of Fig. \ref{compare_modes_cool_kx30},
-which do not exist for $\zmax=3H$. For $\khat=50,\,100$ we find
-surface modes dominate.
+which do not exist for $\zmax=3H$. For $\khat=50,\,100$ and
+$\beta\to0$, we find surface modes dominate so the maximum growth rate
+depends on $\zmax$. However, the maximum growth rate is sensitive to
+$\zmax$ for moderately large $\beta$ ($\gtrsim 0.1$), i.e. close to
+the stabilization of the fundamental mode. 
+%not surprising since we enable strongly stabilizing buoyancy forces
+%at large zmax
 

### 2015-04-01 paper.tex
revising, added soft intro

@@
   disk midplane, which is generally expected in  
-  astrophysical disks. % with, for example, radial variations in temperature. 
-  However, the VSI requires a short thermal or cooling 
-  timescale in sub-adabatically (stably) stratified disks.  
+  astrophysical disks. However, vertical shear is weak in thin disks, 
+  so the VSI can be easily stabilized by buoyancy forces. Consequently,  
+  the VSI requires a rather short thermal or cooling timescale in
+  sub-adabatically (stably) stratified disks.     
   A quantitative examination of this thermodynamic 
@@
 
-\section{Introduction}
+\section{Introduction}\label{intro}
 Understanding the mechanisms of angular momentum transport is the
@@
 inertial waves unstable. \cite{barker15} gives a simple 
-interpretation of the VSI: in an axisymmetric inviscid disk, the 
-exchange of fluid parcels vertically, while conserving their specific angular
-momentum, can result in a new configuration of lower energy if the
-distribution of specific angular momentum varies with $z$. The VSI is
-characterized by perturbations with large vertical length-scales 
-relative to radial length-scales.
+interpretation of the VSI: in an axisymmetric, inviscid, neutrally
+stratified disk, the exchange of fluid parcels vertically, while
+conserving their specific angular momentum, can result in a new
+configuration of lower energy if the distribution of specific angular
+momentum varies with $z$. The VSI is characterized by perturbations
+with large vertical length-scales  relative to radial length-scales.
 
@@
 stratified (convectively stable). A sufficiently short thermal
-timescale renders buoyancy ineffective.    
+timescale renders buoyancy ineffective, but a quantitative criteria is
+desirable. This is the central theme of this work.   
 %efficient exchang of heat with surroundings
@@
 VSI to realistic disks, which one may expect to be stably
-stratified and have non-zero thermal timescales. For example, we show in  
-Fig. \ref{bcrit_mmsn2d} an estimate of the thermal timescales in a
-typical protoplanetary disk model (see below). In order to determine
+stratified and have non-zero thermal timescales.  In order to determine
 where in the disk the VSI may operate, and hence lead to hydrodynamic
@@
 % instability.
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/bcrit_mmsn2d} 
-  \caption{Estimate of the thermal timescale  in a fiducial 
-    protoplanetary disk model, the Minimum Mass Solar Nebula, for
-    perturbations with radial lengthscales $\sim 10^{-2}H$, where $H$
-    is the disk scale-height. The red line corresponds to a 
-    characteristic thermal timescale, derived in this work, 
-    below which the VSI is expected to operate effectively (the region
-    outside the line).   
-    \label{bcrit_mmsn2d} 
-  }
-\end{figure}
 
@@
 
-This paper is organized as follows. In \S\ref{setup} we list the
-governing equations and global equilibrium disk models under
+This paper is organized as follows. We first highlight in
+\S\ref{vsi_require} the need of a very short thermal timescale for the
+VSI to operate in stably stratified disks. There we also preview our
+main result --- a quantitative thermodynamic criteria for the VSI ---
+and a first application to protoplanetary disks. The remaining sections
+are devoted to deriving these results from linear theory.
+
+In \S\ref{setup} we list the governing equations and global
+equilibrium disk models under   
 consideration. We formulate the linear stability problem in 
@@
 number of approximations made, but closely parallels previous work and 
-allows us to highlight the essential effects of vertical shear. In
-particular, we quantify a characteristic upper limit to the thermal
-timescale for the VSI. In \S\ref{numerical} we numerically
+allows us to highlight the essential effects of vertical shear, and to
+derive a characteristic upper limit to the thermal timescale for the
+VSI. In \S\ref{numerical} we numerically
 solve the linear problem in the radially local approximation, but
@@
 We find basic agreement with our analysis, including
-the aforementioned characteristic thermal timescale. In
+the aforementioned characteristic thermal timescale.
+
+In
 \S\ref{application} we apply our results to the Minimum Mass
@@
 for future work.
+
+\section{Thermodynamic requirement for the vertical shear instability}\label{vsi_require}    
+In this paper we will find that in a stably stratified disk, the VSI
+requires a \emph{very} short thermal timescale $t_c$, much less than the
+disk dynamical timescale,
+\begin{align}
+  t _c \ll \Omega^{-1},  
+\end{align}
+consistent with numerical experiments \citep{nelson13}. In subsequent
+sections, we quantify this requirement more 
+precisely. For this preliminary discussion we shall 
+consider a vertically isothermal disk with aspect-ratio $\epsilon$,
+dimensionless radial temperature gradient $q$ and adiabatic index
+$\gamma$.
+
+In a thin disk, the specific angular momentum varies with height as 
+\begin{align}\label{vshear_thin}
+  r \frac{\p \Omega^2 }{\p z} \simeq \epsilon q \hat{z} \Omega^2,
+\end{align}
+where $\hat{z}$ is a normalized vertical co-ordinate. As noted in
+\S\ref{intro} this vertical variation is a source of free energy and
+thus vertical shear ($\p_z\Omega \neq0$) is destabilizing.
+
+If $\gamma>1$ then the disk is stably stratified with a real vertical
+buoyancy frequency $N_z$.  For a thin disk,     
+\begin{align}\label{nz_thin}
+  N_z^2 \simeq \left(\frac{\gamma-1}{\gamma}\right) \hat{z}^2
+  \Omega^2.  
+\end{align}
+This provides a stabilizing effect as buoyancy opposes vertical
+motion, which is required for the VSI. In fact, if we consider
+axisymmetric adiabatic disturbances, then because the 
+Solberg-Hoiland criteria \citep{tassoul78} are usually satisified in a
+thin disk, the disk would be stable.
+
+However, the stabilizing effect of buoyancy can be mitigated for
+non-adiabatic disturbances. Consider a fluid element displaced upwards, away from
+the disk midplane. In a stably stratified adiabatic disk, it would
+sink because of buoyancy. However, if thermal timescales are short, in
+that temperature flucations decay rapidly (i.e. non-adiabatic), then
+the fluid element would attain its surrounding density and 
+%(assuming an ideal gas and it reaches pressure 
+%equilibrium with its surroundings). 
+thus buoyancy is rendered ineffective. Qualitatively, a short thermal
+timescale $t_c$ reduces the effective (stabilizing) buoyancy force.
+
+How short must $t_c$ be for vertical shear to prevail? We see
+from above that the destabilizing effect of vertical shear is rather
+weak because $|\p_z\Omega|\propto \epsilon \ll 1$. 
+Then even a small amount of buoyancy can significantly stabilize
+the VSI. It follows that $t_c$ must be \emph{extremely} short. The
+central result of this work is the requirement 
+\begin{align}\label{prelim_bcrit}
+  (\gamma - 1)t_c < \epsilon |q| \Omega^{-1}
+\end{align}   
+for the VSI in a vertically isothermal, thin disk. Physically, the  
+left-hand side represents an effective buoyancy, while the right-hand
+side represents the (weakly)  destabilizing effect of vertical
+shear. \emph{In a thin disk the thermal timescale must be shorter than
+  the dynamical timescale by $O(\epsilon)\ll 1$ for the VSI to operate.}
+
+% While assumptions and simplifications will be made 
+% to derive this,
+
+The simplicity of Eq. \ref{prelim_bcrit} means it can be 
+readily applied to evaluate where in a disk the VSI is likely to
+operate. For example, we show in Fig. \ref{bcrit_mmsn2d} an estimate
+of the thermal timescales in a typical protoplanetary disk model (see 
+below). Here, $H$ is the disk scale height and $k_x$ is a radial
+wavenumber of the disturbance being considered, thus larger $k_x$
+correspond to smaller perturbations. The region outside the red line
+satisfies Eq. \ref{prelim_bcrit}.
+
+The derivation of Eq. \ref{prelim_bcrit} in fact
+assumes $t_c$ is vertically uniform. While this is not generally true
+in the inner disk, we find Eq. \ref{prelim_bcrit} still provides a
+useful guide. We see that the VSI is likely to  operate in the outer 
+disk at tens of AU, but that smaller scale disturbances may operate at
+smaller radii.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.7cm 0cm
+  0.73cm]{figures/bcrit_mmsn2d_kx10}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.46cm 0cm
+  0.73cm]{figures/bcrit_mmsn2d_kx50}
+  \caption{Estimate of the thermal timescale  in a fiducial 
+    protoplanetary disk model, the Minimum Mass Solar Nebula. The red
+    line corresponds to a characteristic thermal timescale, derived in
+    this work, below which the VSI is expected to operate effectively
+    (the region outside the line).    
+    \label{bcrit_mmsn2d} 
+  }
+\end{figure}
+
+We now proceed through the formal linear stability problem to arrive
+at Eq. \ref{prelim_bcrit} and verify it numerically. The interested 
+reader may skip forward \S\ref{application} for the application of our
+results to protoplanetary disks.
 

### 2015-04-01 linear_adia.tex
revising in progress by mkl
revising

@@
 
-Rapid thermal relaxation is required to overcome a stabilizing
-vertical entropy gradient \citep{goldreich67,urpin98,urpin03}. 
-Eq. \ref{iso_vsi_cond} quantifies this requirement for vertically
-global disturbances (which have small $M$).  We can interpret 
-$\beta_\mathrm{crit}$ as the thermal timescale below which 
-perturbations are effectively isothermal and the VSI operates. We see
-that in a thin disk, a very short thermal timescale is required
-($\beta_\mathrm{crit}\ll 1$) because $\beta_\mathrm{crit}\propto
-\epsilon$.
+%Rapid thermal relaxation is required to overcome a stabilizing
+%vertical entropy gradient \citep{goldreich67,urpin98,urpin03}. 
+%Eq. \ref{iso_vsi_cond} quantifies this requirement for vertically
+%global disturbances (which have small $M$).
+
+We can interpret  $\beta_\mathrm{crit}$ as the thermal timescale below
+which buoyancy becomes ineffective and perturbations are effectively
+isothermal so the VSI can operate.  We see that in a thin disk, a very short thermal timescale
+is required ($\beta_\mathrm{crit}\ll 1$) because
+$\beta_\mathrm{crit}\propto \epsilon$. Physically, this is because
+vertical shear is weak in a thin disk, so even a small entropy
+gradient can significantly reduce VSI growth rates.
 

### 2015-04-01 numerical.tex
revising, added soft intro
revising

@@
 $\beta\to0$, we find surface modes dominate so the maximum growth rate
-depends on $\zmax$. However, the maximum growth rate is sensitive to
+depends on $\zmax$. However, the maximum growth rate is insensitive to
 $\zmax$ for moderately large $\beta$ ($\gtrsim 0.1$), i.e. close to
@@
 0.6$ stabilized the VSI; as Fig. \ref{bcrit_compare1} show that growth
-rates for $\beta\gtrsim 0.6$ are small. To destablize the disk with
-$\beta\gtrsim O(1)$, one might consider 
-$\khat > 100$ to increase the maximum growth rate by enabling surface
-modes, but in a realistic disk
-very small scale disturbances will likely be subject to viscous decay
-(see \S\ref{application}). Numerical simulations are also unlikely to
-resolve such small scales.
+rates for $\beta\gtrsim 0.6$ are small, and are restricted to very
+small radial lengthscales $\sim 10^{-2}H$, for which current numerical
+simulations may not be able to resolve.
+%  To destablize the disk with
+% $\beta\gtrsim O(1)$, one might consider 
+% $\khat > 100$ to increase the maximum growth rate by enabling surface
+% modes, but in a realistic disk
+% very small scale disturbances will likely be subject to viscous decay
+% (see \S\ref{application}). 
+%Numerical simulations are also unlikely to
+%resolve such small scales.
 

### 2015-04-02 appendix.tex
first complete revision after barker & latter

@@
 where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
-t_c \gamma\right)$. Making the low frequency, nearly Keplerian
+t_c \gamma\right)$. Making the low frequency, Keplerian
 approximations with the replacement $D\to \Omega_k^2$ gives, in terms
@@
 \section{Linear problem in the isothermal limit}\label{iso_discuss}  
-Here we summarize some results for isothermal
+Here we summarize selected results for isothermal
 perturbations ($\beta\equiv 0$) in vertically isothermal disks 
@@
 (i.e. $\delta P = c_s^2\delta \rho$), and so $  \ii\hat{\sigma}c_s\delta v_z = W^\prime $
-from Eq. \ref{lin_vz}. For isothermal perturbations it is in fact
+from Eq. \ref{lin_vz}. For isothermal perturbations it is 
 simpler to work with an equation for $W$ by substituting $Q=W$ into
 Eq. \ref{ode_w}. In this case, we shall not yet make the low frequency
-approximation, but first make the nearly Keplerian approximation. We
+approximation, but first make the  Keplerian approximation. We
 obtain, in terms of dimensionless variables, 
@@
   |\hat{\nu}|\leq
-  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|). 
+  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|) < \frac{\epsilon |q|}{2}\mathrm{max}(|f|) . 
 \end{align}
@@
 \end{align}
-We remark that taking the low-frequency limit and considering
+We remark that taking the low frequency limit and considering
 $\hat{k}^2\gg 1$, Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in
@@
 \subsubsection{Instability with vertical shear}
-The VSI corresponds to unstable inertial waves \citep{barker15}. This
+The VSI corresponds to unstable inertial waves. This
 is readily obtained for large $\khat^2$ by balancing
@@
 Fig. \ref{compare_modes_iso_kx10} and
-Fig. \ref{compare_modes_iso_kx30}, for which $\nu\propto\omega$.
+Fig. \ref{compare_modes_iso_kx30}, for which $\nu\propto|\omega|$.
 

### 2015-04-02 application.tex
first complete revision after barker & latter

@@
 protoplanetary disk do the thermodynamic conditions allow the VSI to
-operate. For a realistic disk one can consider the energy equation
+operate. For a realistic disk one might consider an energy equation
 with radiative cooling, 
@@
 volume density $\rho = \Sigma\hat{\rho}/\sqrt{2\pi}H$, where
-$\hat{\rho} = \rho/\rho_0$ and $\Sigma$ is the surface density at the
+$\hat{\rho} = \rho/\rho_0$ and $\Sigma(r)$ is the surface density at the
 radius of interest, the dimensionless thermal
@@
 \begin{align}\label{real_beta}
-  \beta(z;\khat) \equiv t_c\Omega_k =
+  \beta(r,z;\khat) \equiv t_c\Omega_k =
   % \frac{1}{\hat{\eta}}\left[\frac{l_\mathrm{rad}^2}{3H^2}
@@
 \end{align}
-Note that $\beta$ depends on both the perturbation lengthscale $\khat$ 
-and distance from the midplane.
+Note that $\beta$ depends the perturbation lengthscale $\khat$,
+distance from the midplane, as well as the distance from the star.
 
@@
 \begin{align}\label{beta_mmsn}
-  \beta(z; \khat) =
+  \beta(r, z; \khat) =
   &\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)r_\mathrm{AU}^{-57/14}\notag\\ 
@@
 $z$ as the disk transitions from an optically thick midplane to an
-optically thin atmosphere. 
+optically thin atmosphere (see also Fig. \ref{bcrit_mmsn2d}). 
 %Given a fixed $\khat$, one can have $\beta > \beta_\mathrm{crit}$ at
@@
 %becomes optically thin and $\beta<\beta_\mathrm{crit}$. 
-Our analytical criteria is not valid for non-uniform $\beta$, and we must
-turn to numerical solutions of the linear problem to determine whether
-or not the VSI operates. We do this in the next section.
+Our analytical criteria $\beta < \beta_\mathrm{crit}$ is not valid for
+vertically non-uniform $\beta$, and we must turn to numerical
+solutions of the linear problem to determine whether or not the VSI
+operates. We do this in the next section.
 
@@
 Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at each radius the value
-of $\khat$ such that $\beta = \beta_\mathrm{crit}$, which show that in
-the inner disk, say $r\sim 1$AU, the VSI would only occur on very
-small radial lengthscales. 
+of $\khat$ such that $\beta = \beta_\mathrm{crit}$. If we consider the
+midplane of the inner disk, say $r\sim 1$AU, the VSI would only occur
+on very small radial lengthscales $\ll O(10^{-2}H)$. In the disk
+atmosphere where $\beta$ is small, surface modes may develop, 
+%However, at such
+%small scales, surface modes are likely to dominate (and are
+%restricted to the disk atmosphere), 
+but they are not well-defined in vertically isothermal disk models
+with an imposed boundary, and they also require large $\khat$.
+
 %surface modes can develop for large $\khat$, 
@@
 that it is the most unstable, except for the appearance of some 
-surface modes towards the inner disk when considering large 
-wavenumbers $\khat\gtrsim 50$. However, surface modes are dependent on
-boundary conditions in a vertically isothermal disk \citep{barker15},  
-and disturbances with large $\khat$ are more likely subject to viscous
-decay, as discussed above. % It is also unclear if surface modes are
+surface modes for large wavenumbers $\khat\gtrsim 50$, which are 
+boundary-dependent.  
+%However, surface modes are dependent on
+%boundary conditions in a vertically isothermal disk \citep{barker15},  
+%and 
+Disturbances with large $\khat$ are also more likely subject to viscous 
+decay, as discussed above.
+
+% It is also unclear if surface modes are
 % dynamically important given they are concentrated in the disk
@@
 We therefore consider Fig. \ref{mmsn_overall} as a representation of the
-VSI in the MMSN. We see the VSI operates from $\sim 5$AU to $\sim
+VSI in outer parts of the MMSN. We see the VSI operates from $\sim 5$AU to $\sim
 50$AU with growth timescales $\sim 30$---40 orbits and
@@
 thermal timescale. For example, 
-Fig. \ref{mmsn_bcrit_bcool} show that $\beta(\khat=10,z=0) \gtrsim
-\beta_\mathrm{crit}$ for $r\lesssim 10$AU in the MMSN (left middle panel); and Fig. \ref{mmsn_overall}
+Fig. \ref{mmsn_bcrit_bcool} show that $\beta(z=0,\khat=10) \gtrsim
+\beta_\mathrm{crit}$ for $r\lesssim 10$AU in the MMSN (left middle
+panel); and Fig. \ref{mmsn_overall} 
 show that modes with $\khat=10$ are stabilized inside $\sim 8$AU (where
@@
   \caption{Characteristic growth times of the VSI in 
-    the MMSN. These are calculated using the scale and height
-    dependent thermal relaxation given by Eq. \ref{beta_mmsn_simp}. 
+    the MMSN. These are calculated using the thermal relaxation
+    function given by Eq. \ref{beta_mmsn_simp}. 
     \label{mmsn_overall}}    

### 2015-04-02 numerical.tex
first complete revision after barker & latter

@@
 below.)  The `fundamental corrugation mode', marked by `f',
-corresponds to the eigenvalue with the smallest $|\sigma|$. It is robust
+corresponds to the eigenvalue with the smallest $|\sigma|$. The
+fundamental mode is robust  
 against vertical boundary conditions as it appears at the same
@@
 deviates from the analytic line due to the imposed boundaries, 
-unless $\zmax$ is increased \citep{barker15}.
+unless $\zmax$ is increased.
 
@@
 This a class of unstable modes associated with vertical shear near the
-disk boundaries \citep{barker15}. They are absent in our  
+disk boundaries. They are absent in our  
 analysis because there we assumed a disk with infinite 
@@
 
-We illustrate surface modes with the mode diagram for $\khat=30$ in
-Fig. \ref{compare_modes_iso_kx30}. Surface 
+We illustrate surface modes in Fig. \ref{compare_modes_iso_kx30} 
+which show unstable modes with $\khat=30$. Surface 
 modes lie almost parallel to the vertical axis 
@@
 of a surface mode. We typically find an increasing
-number of surface modes with increasing $\zmax$ and/or $\khat$,
-consistent with \cite{barker15}.
+number of surface modes with increasing $\zmax$ and/or $\khat$. 
+%consistent with \cite{barker15}.
 
@@
   \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_analytic.ps}
-  \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for $\khat=30$. Examples of surface modes are
+  \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for
+    $\khat=30$. Examples of surface modes are 
     marked by `s'. \label{compare_modes_iso_kx30}
@@
 (middle panel). The fundamental mode ($M=0$) is not the most unstable for   
-$\beta=0.01$ as there are modes with $M>0$ that have larger growth 
-rates. As remarked above, these higher order modes are affected by
+$\beta=0.01$ as there area number of modes with $M>0$ that have larger
+growth rates. As noted above, these higher order modes are affected by
 $\zmax$, e.g. there is a significant mismatch between numerical and 
@@
 midplane, with oscillatory behavior emerging near vertical boundaries
-where buoyancy becomes important. This leads to the `tilted' pressure
-perturbations seen in Fig. \ref{lowfreq_eigenfunc_2d_cool} 
+where buoyancy first becomes important. This leads to the `tilted'
+pressure perturbations seen in Fig. \ref{lowfreq_eigenfunc_2d_cool} 
 (cf. Fig. \ref{lowfreq_eigenfunc_2d}).
@@
 as seen above, for $\khat \leq 10$ higher order body modes are
-dominant as $\beta\to 0$, but the fundamental corrugation
+dominant as $\beta\to 0$, but the fundamental 
 mode becomes dominant as $\beta$ is increased. For $\khat=1$ the VSI
@@
 $\beta\to0$, we find surface modes dominate so the maximum growth rate
-depends on $\zmax$. However, the maximum growth rate is insensitive to
+depends on $\zmax$ in the limit of rapid thermal relaxation. 
+However, the maximum growth rate is insensitive to 
 $\zmax$ for moderately large $\beta$ ($\gtrsim 0.1$), i.e. close to
@@
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z3}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z5}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm 0.9cm]{figures/gcorr_compare_iso_maxrate_z7} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0.9cm]{figures/gcorr_compare_iso_maxrate_z3} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0.9cm]{figures/gcorr_compare_iso_maxrate_z5} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
+  0.9cm]{figures/gcorr_compare_iso_maxrate_z7}  
   \caption{Maximum VSI growth rates in the fiducial disk 
@@
 result, together with Fig. \ref{bcrit_compare1}, shows that  
-$\beta \lesssim \beta_\mathrm{crit}$ can be used as a practical
+$\beta \lesssim \beta_\mathrm{crit}$ is a practical
 criteria for the VSI to operate effectively in stably stratified,
@@
     $\beta_\mathrm{crit}$ for the fundamental VSI mode on disk
-    parameters. Our fiducial setup is $(\gamma, \Gamma= (1.4, 1.011)$
+    parameters. The fiducial setup is $(\gamma, \Gamma= (1.4, 1.011)$
     and $(p,q,\epsilon)=(-1.5,-1,0.05)$. Top: varying 
@@
 
-% However, note that this issue is unimportant for low order modes 
-% such as the fundamental mode, which are not sensitive to boundary
-% conditions. This is useful because when the fundamental mode is
-% stabilized as $\beta$ is increased, higher order modes will have
-% already been stabilized, the the growth rate of surface modes also
-% significantly reduced. Then as far as a characteristic upper limit to
-% the thermal timescale is conconcered, we may just consider the
-% fundamental mode.
+However, as we have seen in Fig. \ref{bcrit_compare1}, the maximum
+growth rate becomes insensitive to $\zmax$ for $\beta\gtrsim
+\beta_\mathrm{crit}$. This suggests that the characteristic upper
+limit on the thermal timescale for the VSI is not sensitive to
+$\zmax$, so that vertically isothermal disks are in fact suitable for
+the purpose of this paper.
 
-Here we briefly consider a vertically non-isothermal disk model with  
-$(p,q,\epsilon)=(0,-1,0.05)$ and $\Gamma=1.0837$. This value of 
-$\Gamma$ is chosen so that the standard vertical domain size $\zmax =
+%Physically this is because modes that depend on vertical boundaries
+%are also subject to the strongest stabilization by buoyancy, which
+%maximize at $\zmax$. These modes are theremore always strongly damped
+%as we increase the thermal timescale.
+
+Nevertheless, here we briefly consider a vertically non-isothermal
+disk model with $(p,q,\epsilon)=(0,-1,0.05)$ and $\Gamma=1.0837$. This
+value of $\Gamma$ is chosen so that the standard vertical domain size $\zmax =
 5H$ effectively corresponds to the disk surface ($\zmax\simeq
@@
 stabilized, eventually only the fundamental mode can operate before
-the VSI is entirely stabilized.
+the VSI is entirely suppressed.
 
@@
 
-We do not find surface modes for $\khat=10$. In the vertically
-non-isothermal disk, surface modes develop for sufficiently large 
-$\khat$ \citep{barker15}. We demonstrate them in
+We do not find surface modes for $\khat=10$, but do find them for
+sufficiently large $\khat$. We demonstrate this in  % In the vertically
+% non-isothermal disk, surface modes develop for sufficiently large 
+% $\khat$ \citep{barker15}. We demonstrate them in
 Fig. \ref{compare_modes_vnoniso_kx100} which show eigenvalues for
@@
 most unstable mode without the boundary-dependent issue that arose 
-with a vertically isothermal disk.
+with a vertically isothermal disk in the limit $\beta\to0$.
 
@@
 with $\khat=100$ and $\beta=10^{-3},\, 10^{-2}$ and
-$0.1$. We find with $\beta=10^{-3}$ the most 
+$0.1$. For $\beta=10^{-3}$ the most 
 unstable mode is an asymmetric surface mode \citep[also seen in Fig. 7
@@
 $\beta=10^{-2}$ an anti-symmetric surface mode becomes
-dominant. Further increasing to $\beta=0.1$ surface modes are
-suppressed and the dominant mode is the fundamental corrugation
+dominant. Further increasing to $\beta=0.1$, the dominant mode has a
+mixed character between a surface mode and the fundamental
 mode. Notice as $\beta$ is increased the perturbation amplitude near
@@
 the stabilizing effect of buoyancy is largest at the disk surface. 
+%formally diverges
 
@@
 rates in the vertically non-isothermal disk model. Based on 
-Eq. \ref{iso_vsi_cond}, we conjecture a critical
-thermal timescale given by $\epsilon|s|/(\gamma - \Gamma)$. This
+Eq. \ref{iso_vsi_cond}, we conjecture the critical 
+thermal timescale in the vertically non-isothermal disk is 
+given by $\epsilon|s|/(\gamma - \Gamma)$. This 
 evaluates to 0.158, which is close to the cut-off thermal timescale
-for $\khat\in[5,30]$. Fig. \ref{bcrit_compare2} is qualitatively quite
+for $\khat\in[5,30]$. Fig. \ref{bcrit_compare2} is qualitatively 
 similar to the fiducial disk model in
-Fig. \ref{bcrit_compare1}. Importantly, the maximum thermal
-timescale for $\khat=O(10)$ in fact provides a 
-characteristic upper limit for other values of $\khat$.
+Fig. \ref{bcrit_compare1}. Importantly, for $\khat=O(10)$ there is a
+precise cut-off thermal timescale, the value of which in fact provides a  
+characteristic upper limit for other $\khat$.
 

### 2015-04-02 summary.tex
first complete revision after barker & latter

--- 2015-03-16 summary.tex
@@
 height dependence of the equilibrium angular velocity profile 
-$\Omega=\Omega(r,z)$ in a baroclinic disk. We examined how
-the linear instability depends on disk thermodynamics and applied our
-results to assess the relevance of the VSI in realistic protoplanetary
-disks.
+$\Omega=\Omega(r,z)$ in a baroclinic disk. Vertical shear is generally
+expected in astrophysical disks due to the radial variation in the
+disk temperature, or more generally the entropy. However, 
+astrophysical disks are also typically sub-adiabatically stratified. 
+In this case, the VSI requires a short thermal timescale $t_c$ in
+order to operate.
 
-We generalized previous vertically global linear 
-calculations of the VSI \citep{nelson13,mcnally14}, which assumed isothermal
-perturbations, to include an energy equation. Our effort may also be
-regarded as an extension of \cite{urpin03}, which did include an
-energy equation, but adopted the Boussinesq and vertically local
-approximations. We considered vertically global and fully
-compressible perturbations.
+Physically, a short thermal timescale is needed to overcome the 
+stabilizing influence of buoynancy forces in a stably stratified
+disk. Because vertical shear is weak in thin disks, $t_c\Omega\ll 1$
+is generally needed for the VSI to operate, otherwise buoyancy is
+strongly stabilizing. Our primary goal was to quantify this
+requirement, thereby provide a simple way to assess 
+whether or not the VSI is relevant in realistic disks.
 
-We included a source term in the energy equation which relaxes the disk
+To do so, we generalized previous vertically global linear 
+calculations of the VSI \citep{nelson13,mcnally14,barker15}, which
+assumed isothermal perturbations, to include an energy equation with 
+a source term in that relaxes the disk
 temperature to its initial value on a timescale
-$t_c=\beta\Omega_k^{-1}$. This approach allows the thermodynamic
-response of the disk to be controlled by varying the parameter
-$\beta$, and has been employed in previous numerical simulations of
-the VSI \citep{nelson13}. However, it is possible to 
-relate  $\beta$ to physical thermal timescales expected in realistic disks.
+$t_c$. This approach of `thermal relaxation' allows
+the thermodynamic response of the disk to be controlled by varying
+$t_c$, and has been employed in previous numerical simulations of the
+VSI \citep{nelson13}.
 
@@
 perturbations with small radial lengthscales, we applied the 
-radially local approximation in the linearized fluid equations. For
-analytic discussion only, we further simplified the linear problem by
-ignoring the background radial disk structure. This is equivalent to adopting
-the `vertically global shearing box' (VGSB) formalism developed by
-\cite{mcnally14}. 
-We find the VGSB framework is appropriate for analyzing the VSI because 
-rapid thermal relaxation ($\beta\ll1$) is 
-generally required for the instability.   However, the VGSB framework is not
-appropriate in the adiabatic limit ($\beta \gg 1$).
+radially local approximation in the linearized fluid equations. 
+We mostly focused on vertically isothermal disks as 
+this is the relevant case for protoplanetary disks passively
+irradiated by its central star \citep{chiang97}. It also permits a
+largely analytical discussion. Our analysis is an extension of the
+linear models considered in \cite{lubow93} by adding vertical shear
+and (rapid) thermal relaxation. We obtained explicit solutions
+describing the VSI with thermal relaxation in the thin disk limit.
 
-We mostly focused on vertically isothermal disks ($\Gamma=1$) as 
-this is the relevant case for protoplanetary disks that are passively irradiated by its central star 
-\citep{chiang97}. It also permits a largely analytical discussion.
-Our analysis is an extension of the linear models considered in
-\cite{lubow93} by  adding vertical shear and (rapid) thermal
-relaxation. We obtained explicit solutions describing the VSI with 
-thermal relaxation in the thin disk limit.
-
-The fundamental VSI mode is characterized by a symmetric vertical
-velocity perturbation with zero nodes and $\delta
-v_z\sim\mathrm{constant}$ near the mid-plane, and by an anti-symmetric
-density perturbation with one node at $z=0$.  These have been referred
-to as `corrugation modes' by \cite{nelson13}, and were found to
-dominate their numerical simulations, as well as that of
-\cite{stoll14}. Furthermore, when computing unstable modes for a
-protoplanetary disk model with thermal relaxation based on realistic
-disk properties, we find the fundamental mode to be the dominant
-mode. The fundamental mode is
-thus likely the most relevant VSI mode in protoplanetary disks.
-
-We determined the critical thermal relaxation timescale 
-$\beta_\mathrm{crit}=\epsilon|q|/(\gamma-1)$, below which small-scale
-perturbations may be regarded as isothermal and the VSI 
-operates. We have therefore provided a quantitative measure for the `fast'
-thermal relaxation requirement for the VSI in astrophysical disks.
-
-We also solved the linear stability problem numerically using a
-pseudo-spectral approach. For numerical calculations we relax all the
-approximations made in our analysis, and accounted for
-the background radial disk structure.  Our
-numerical results are mostly consistent with our analytic 
-expectations, including the dependence of $\beta_\mathrm{crit}$ on
-disk parameters.  We find the fundamental mode is robust against
-boundary conditions, and higher order modes are more rapidly
-stabilized with increasing $\beta$.
-
-We briefly studied vertically non-isothermal disks. We confirm the finding by \cite{nelson13} that in the absence
-of vertical entropy gradients, the VSI does not require rapid thermal
-relaxation.  When a stable stratification is introduced, the
-rapid reduction in VSI growth rates by a finite thermal relaxation
-timescale is qualitatively similar to that observed for vertically
-isothermal disks. Based on numerical results, we conjecture that the
-critical thermal relaxation timescale is given by 
-$\beta_\mathrm{crit}=\epsilon|q|/(\gamma-\Gamma)$ in vertically 
-non-isothermal disks.
-
-%application to MMSN
+For a vertically isothermal, sub-adiabatically stratified disk we
+determined the thermal timescale   
+\begin{align*}
+  t_\mathrm{crit} = \frac{\epsilon |q|}{\gamma-1}\Omega^{-1}, 
+\end{align*}
+below which the VSI operates effectively. While $t_\mathrm{crit}$ was
+derived for the lowest order VSI mode --- the `fundamental  
+corrugation mode' --- we find from numerical calculations that
+$t_\mathrm{crit}$ can be used to characterize the general
+stabilization of the VSI by a finite thermal timescale. %That is, all
+%VSI growth rates are reduced significantly as $\beta$ is increased
+%from 
+%zero to $\beta_\mathrm{crit}$.    
+We have therefore provided a quantitative measure for the thermal
+timescale requirement for the VSI in astrophysical 
+disks. The thermal timescale must be shorter than the disk
+dynamical timescale by $O(\epsilon)$, which is
+quite short for thin disks ($\epsilon\ll 1$), because vertical shear is weak.
 
@@
 (MMSN) described in \cite{chiang10} and   
-modeled the thermal relaxation timescale $\beta$ based on realistic
-disk properties and dust opacity. This results in a $\beta$ function that
-depends on both the perturbation lengthscale as well as the height 
-from the midplane. We find the MMSN is unstable to the VSI 
+modeled the thermal timescale $t_c$ based on realistic
+disk properties and dust opacity. 
+%This results in $t_c$ being a
+% function of position and the perturbation lengthscale of interest. 
+We find the MMSN is unstable to the VSI 
 from $\sim 5$AU to $\sim50$AU with characteristic growth times of
-$\sim 30$ orbits. Although $\beta$ is not vertically constant in the
+$\sim 30$ orbits. It is possible to have the VSI at smaller
+radii, but only on very small radial lengthscales (signficantly shorter
+than the disk scaleheight), and such perturbations may
+be subject to viscous decay in a real disk.
+
+Although $t_c$ is not vertically constant in the 
 MMSN as assumed in our analysis, we find that evaluating the condition
-$\beta <\beta_\mathrm{crit}$ at the disk midplane gives a rough measure of the
+$t_c< t_\mathrm{crit}$ at the disk midplane gives an indication of the
 radii at which the VSI operates in the MMSN. This simple criterion is
-thus a useful first step to estimate the relevance of the VSI in
-realistic astrophysical disks without resorting to non-linear
-numerical simulations.
+thus a practical first step in estimating the relevance of the VSI in
+realistic astrophysical disks without resorting to linear
+calculations or non-linear numerical simulations.
 
@@
 We have focused on parameter regimes relevant to the VSI,
-namely $\khat\gtrsim 1$ and $\beta\ll 1$. This excludes the
-`convective overstability' recently discussed by \cite{klahr14} and
-\cite{lyra14} in vertically unstratified disk models.
+namely perturbations with small radial lengthscales 
+($k_xH\gtrsim 1$) in disks with short thermal timescales
+($t_c\Omega\ll 1$). This excludes the `convective overstability'
+recently discussed by \cite{klahr14} and \cite{lyra14} in vertically
+unstratified disk models.
 
 The convective overstability corresponds to slowly growing epicyclic motions.
-It is most effective for $\beta\sim 1$ and disturbances with $|k_x|\ll |k_z|$, where
-$k_z$ is a vertical wavenumber. These conditions are mutually exclusive with the VSI. 
+It is most effective for $t_c\Omega \sim 1$ and disturbances with
+$|k_x|\ll |k_z|$, where $k_z$ is a vertical wavenumber. These
+conditions are mutually exclusive with the VSI.  
 The maximum growth rate for convective overstability is given by 
@@
 require perturbations with long radial lengthscales,
-$\khat\ll 1$, for which radially global disk models would be more 
+$k_xH\ll 1$, for which radially global disk models would be more 
 self-consistent.
@@
 \subsubsection{Improved thermal evolution} 
-We assumed non-adiabatic effects
-can be represented by a thermal relaxation timescale $t_c$. For 
+We assumed non-adiabatic effects can be represented by a thermal
+relaxation timescale $t_c$. For  
 application to protoplanetary disks we connected $t_c$ to the disk
-properties. One can improve the realism further by generalizing the 
-opacity model adopted here to be density (and hence height) dependent. 
+properties such as surface density, temperature and opacity, making
+$t_c$ depend on position and the perturbation 
+lengthscales. One might improve the realism further by generalizing
+the opacity model adopted here to be density (and hence height)
+dependent.  
 %here to be density-dependent,  i.e. $\kappa_d =
@@
 However, going beyond thermal relaxation requires a
-proper treatment of radiative diffusion. The linear problem then
-involves a Laplacian operator and additional boundary
-conditions must be supplemented. We will consider these improvements in a
+proper treatment of radiative diffusion. % The linear problem then
+% involves a Laplacian operator and additional boundary
+% conditions must be supplemented. 
+We will consider these improvements in a
 follow-up study. %additional heat source
@@
 
-We have neglected viscosity. \cite{nelson13} find that an 
-alpha viscosity $\lesssim O(10^{-4})$ is required for the VSI. 
-It is of interest to quantify the effect of viscosity 
-on the linear VSI. This would help identify, more accurately, the
-appropriate regions in a protoplanetary disk that are sufficiently
-laminar to develop the VSI.  This problem is, however, complicated by
-the fact that proper inclusion of viscosity introduces a meridional
-flow in the equilibrium state, as well as viscous heating
-in the energy equation.
+We have neglected viscosity. \cite{nelson13} find that an alpha
+viscosity $\lesssim O(10^{-4})$ is required for the
+VSI. \cite{barker15} found that inclusion of viscosity is needed for 
+convergence when studying surface modes. %but not body modes? 
+Viscosity may also affect the
+thermal timescale requirement of the VSI, because viscous heating
+needs to be self-consistently included in the energy equation. This
+should be investigated in future works.
+
+% It is of interest to quantify the effect of viscosity 
+% on the linear VSI. This would help identify, more accurately, the
+% appropriate regions in a protoplanetary disk that are sufficiently
+% laminar to develop the VSI.  This problem is, however, complicated by
+% the fact that proper inclusion of viscosity introduces a meridional
+% flow in the equilibrium state, as well as viscous heating
+% in the energy equation.

### 2015-04-02 summary.tex
first complete revision after barker & latter
minor changes to summary

@@
 
-Physically, a short thermal timescale is needed to overcome the 
-stabilizing influence of buoynancy forces in a stably stratified
+Physically, a short thermal timescale is needed to reduce the  
+stabilizing effect of buoynancy forces in a stably stratified
 disk. Because vertical shear is weak in thin disks, $t_c\Omega\ll 1$
-is generally needed for the VSI to operate, otherwise buoyancy is
-strongly stabilizing. Our primary goal was to quantify this
+is generally needed for the VSI to operate, otherwise buoyancy would
+be strongly stabilizing. Our primary goal was to quantify this
 requirement, thereby provide a simple way to assess 
@@
 temperature to its initial value on a timescale
-$t_c$. This approach of `thermal relaxation' allows
+$t_c$. This `thermal relaxation' model allows
 the thermodynamic response of the disk to be controlled by varying
@@
 irradiated by its central star \citep{chiang97}. It also permits a
-largely analytical discussion. Our analysis is an extension of the
+largely analytical discussion. Our analysis is then an extension of the
 linear models considered in \cite{lubow93} by adding vertical shear
@@
 $\sim 30$ orbits. It is possible to have the VSI at smaller
-radii, but only on very small radial lengthscales (signficantly shorter
-than the disk scaleheight), and such perturbations may
+radii, but only on radial lengthscales signficantly shorter
+than the disk scaleheight, and such perturbations may
 be subject to viscous decay in a real disk.
@@
 (cf. Eq. \ref{nzsq_def}). Thus, growth requires $N_r^2<0$.  
-
-Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
-we find  
-\begin{align}
-  \nu_\mathrm{conv,max} =
-  -\frac{\epsilon^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
-  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
-\end{align}
+%Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
+%we find  
+%\begin{align}
+%  \nu_\mathrm{conv,max} =
+%  -\frac{\epsilon^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
+%  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
+%\end{align}
 In fact, for the MMSN considered in this work, convective

### 2015-04-02 appendix.tex
minor changes to appendix

@@
 \end{align}
-Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), it
+%Recalling that $f$ represents vertical shear (Eq. \ref{fz_shear}), 
+It
 follows that the maximum possible growth rate of unstable modes,
 satisfying the above boundary conditions, is limited by the maximum
-vertical shear rate in the domain considered.
+vertical shear rate in the domain considered,
+\begin{align}
+  \nu < \mathrm{max} \left|r\frac{d\Omega}{dz}\right|, 
+\end{align}
+as expected on physical grounds.
 

### 2015-04-02 paper.tex
revising
minor revision abstract

@@
   astrophysical disks. However, vertical shear is weak in thin disks, 
-  so the VSI can be easily stabilized by buoyancy forces. Consequently,  
-  the VSI requires a rather short thermal or cooling timescale in
-  sub-adabatically (stably) stratified disks.     
+  so the VSI can easily be stabilized by buoyancy forces in the
+  presence of a stable vertical entropy  gradient. % , which is also
+  % expected in astrophysical disks. 
+  Consequently,  the VSI
+  requires a rather short thermal or cooling timescale in
+  sub-adabatically stratified disks.      
   A quantitative examination of this thermodynamic 
@@
   $t_c=\beta\Omega_k^{-1}$, where $\beta$ is a dimensionless parameter 
-  and $\Omega_k$ is the Keplerian rotation frequency. We explore how
-  the VSI depends on the thermodynamic response of the disk by
-  varying $\beta$. We find that 
+  and $\Omega_k$ is the Keplerian rotation frequency. % We explore how
+  % the VSI depends on the thermodynamic response of the disk by
+  % varying $\beta$.
+  We find that 
   increasing $\beta$ even slightly away from    
@@
 timescale renders buoyancy ineffective, but a quantitative criteria is
-desirable. This is the central theme of this work.   
+desirable. This is our primary goal.  
 %efficient exchang of heat with surroundings

### 2015-04-02 paper.tex
more physical arguments in intro

@@
 \end{align}
-consistent with numerical experiments \citep{nelson13}. In subsequent
-sections, we quantify this requirement more 
-precisely. For this preliminary discussion we shall 
-consider a vertically isothermal disk with aspect-ratio $\epsilon$,
+consistent with numerical experiments \citep{nelson13}. For this
+preliminary discussion we shall consider a vertically isothermal disk
+with aspect-ratio $\epsilon$, 
 dimensionless radial temperature gradient $q$ and adiabatic index
@@
 sink because of buoyancy. However, if thermal timescales are short, in
-that temperature flucations decay rapidly (i.e. non-adiabatic), then
+that temperature flucations decay rapidly (i.e. isothermal perturbations), then 
 the fluid element would attain its surrounding density and 
@@
 
-How short must $t_c$ be for vertical shear to prevail? We see
-from above that the destabilizing effect of vertical shear is rather
-weak because $|\p_z\Omega|\propto \epsilon \ll 1$. 
-Then even a small amount of buoyancy can significantly stabilize
-the VSI. It follows that $t_c$ must be \emph{extremely} short. The
-central result of this work is the requirement 
+How short must $t_c$ be for vertical shear to prevail? For a rough
+comparison of vertical shear to buoyancy for non-adiabatic
+perturbations, we apply Eq. 47 of \cite{urpin03}, which is based on
+\cite{townsend58},   
+\begin{align}\label{bcrit_est}
+  N_z^2 t_c \lesssim \left|r\frac{\p\Omega}{\p z}\right|. 
+\end{align}
+The left-hand side represents an effective buoyancy, which is
+stabilizing; while the right-hand side represents vertical shear,
+which is destabilizing. In a thin disk, the destabilizing effect of
+vertical shear is rather weak because $|\p_z\Omega|\propto \epsilon
+\ll 1$. Then even a small amount of buoyancy can significantly
+stabilize the VSI. It follows that $t_c$ must be quite short to permit
+instability.
+
+Note that in astrophysical disks both $N_z^2$ and $\p_z\Omega$ varies
+with height. Eq. \ref{bcrit_est} then suggest that with 
+increasing height, buoyancy would eventually stabilize the VSI (since
+$N_z^2\propto z^2$ and $\p_z\Omega\propto z$). 
+%This means that 
+%disturbances localized away from the disk midplane are easily stabilized by
+%buoyancy. 
+This argument assumes vertically-localized disturbances. It is,
+however, not clear how to apply Eq. \ref{bcrit_est} to vertically
+global disturbances, as considered in this work.
+
+%it would imply you can always have VSI sufficiently close to the
+%midplane
+
+Our central result is the requirement 
 \begin{align}\label{prelim_bcrit}
@@
 \end{align}   
-for the VSI in a vertically isothermal, thin disk. Physically, the  
-left-hand side represents an effective buoyancy, while the right-hand
-side represents the (weakly)  destabilizing effect of vertical
-shear. \emph{In a thin disk the thermal timescale must be shorter than
-  the dynamical timescale by $O(\epsilon)\ll 1$ for the VSI to operate.}
+for the VSI in a vertically isothermal, thin disk. This equation
+reflects Eq. \ref{bcrit_est} but is simpler since there is no
+$z$ dependence. We find that \emph{in a thin 
+  disk the thermal timescale must be shorter than the dynamical
+  timescale by $O(\epsilon)\ll 1$ for the VSI to operate.}
 
@@
 
-The simplicity of Eq. \ref{prelim_bcrit} means it can be 
+Eq. \ref{prelim_bcrit} can be 
 readily applied to evaluate where in a disk the VSI is likely to
@@
 correspond to smaller perturbations. The region outside the red line
-satisfies Eq. \ref{prelim_bcrit}.
-
-The derivation of Eq. \ref{prelim_bcrit} in fact
+satisfies Eq. \ref{prelim_bcrit}. The criterion Eq. \ref{prelim_bcrit} in fact
 assumes $t_c$ is vertically uniform. While this is not generally true

### 2015-04-02 paper.tex
more physical arguments in intro

@@
   $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  is the radial temperature gradient, and 
-  $\gamma$ is the adiabatic index. 
+  $\gamma$ is the adiabatic index. Thus, in a thin disk the thermal timescale must be shorter than 
+  the dyamical timescale by $O(\epsilon)\ll 1$ for the VSI to operate.  
   We apply our results  
@@
 reflects Eq. \ref{bcrit_est} but is simpler since there is no
-$z$ dependence. We find that \emph{in a thin 
+$z$ dependence. \emph{In a thin 
   disk the thermal timescale must be shorter than the dynamical

### 2015-04-03 paper.tex
new fig from youdin, incoportated in text

@@
 \begin{align}\label{vshear_thin}
-  r \frac{\p \Omega^2 }{\p z} \simeq \epsilon q \hat{z} \Omega^2,
+  r \frac{\p \Omega^2 }{\p z} \simeq \epsilon q \Omega^2\left(\frac{z}{H}\right),
 \end{align}
-where $\hat{z}$ is a normalized vertical co-ordinate. As noted in
+where $H$ is the disk scale height. As noted in
 \S\ref{intro} this vertical variation is a source of free energy and
@@
 \begin{align}\label{nz_thin}
-  N_z^2 \simeq \left(\frac{\gamma-1}{\gamma}\right) \hat{z}^2
+  N_z^2 \simeq \left(\frac{\gamma-1}{\gamma}\right) \left(\frac{z}{H}\right)^2
   \Omega^2.  
@@
 sink because of buoyancy. However, if thermal timescales are short, in
-that temperature flucations decay rapidly (i.e. isothermal perturbations), then 
-the fluid element would attain its surrounding density and 
+that temperature flucations decay rapidly (i.e. nearly isothermal
+perturbations), then the fluid element would attain its surrounding
+density and  
 %(assuming an ideal gas and it reaches pressure 
@@
 which is destabilizing. In a thin disk, the destabilizing effect of
-vertical shear is rather weak because $|\p_z\Omega|\propto \epsilon
-\ll 1$. Then even a small amount of buoyancy can significantly
-stabilize the VSI. It follows that $t_c$ must be quite short to permit
-instability.
+vertical shear is rather weak because $|r\p_z\Omega|\propto \epsilon
+\ll 1$. By comparison the buoyancy frequency $N_z/\Omega\sim 1$ at
+$z\sim H$.
+
+To illustrate this point, we show in Fig. \ref{eqm_structure} the
+equilibrium structure of a typical protoplanetary disk. Here $\rho_0$
+is the midplane density and vertical shear rate and buoyancy
+frequencies are normalized by the local Keplerian frequency. It is
+clear that vertical shear is 
+generally weak compared to buoyancy. Then even a small (positive)
+vertical entropy gradient can significantly stabilize the VSI. It
+follows that $t_c$ must be quite short to permit instability.
+
+\begin{figure*}
+  \includegraphics[width=\linewidth]{figures/rhoshearNz}
+  \caption{Typical protoplanetary disk structure (the Minimum Mass
+    Solar Nebla): density (left), vertical shear rate (middle) and
+    buoyancy frequency (right).  
+    \label{eqm_structure} 
+  }
+\end{figure*}
 
 Note that in astrophysical disks both $N_z^2$ and $\p_z\Omega$ varies
-with height. Eq. \ref{bcrit_est} then suggest that with 
+with height, as evident in
+Fig. \ref{eqm_structure}. Eq. \ref{bcrit_est} then suggest that with
 increasing height, buoyancy would eventually stabilize the VSI (since 
 $N_z^2\propto z^2$ and $\p_z\Omega\propto z$). 
-%This means that 
-%disturbances localized away from the disk midplane are easily stabilized by
-%buoyancy. 
-This argument assumes vertically-localized disturbances. It is,
+% This means that 
+% disturbances localized away from the disk midplane are easily stabilized by
+% buoyancy. 
+This argument assumes vertically localized disturbances. It is,
 however, not clear how to apply Eq. \ref{bcrit_est} to vertically
@@
 operate. For example, we show in Fig. \ref{bcrit_mmsn2d} an estimate
-of the thermal timescales in a typical protoplanetary disk model (see 
-below). Here, $H$ is the disk scale height and $k_x$ is a radial
+of the thermal timescales the protoplanetary disk model shown in
+Fig. \ref{eqm_structure}. Here $k_x$ is a radial
 wavenumber of the disturbance being considered, thus larger $k_x$
@@
 in the inner disk, we find Eq. \ref{prelim_bcrit} still provides a
-useful guide. We see that the VSI is likely to  operate in the outer 
+useful guide. We see that the VSI will operate in the outer 
 disk at tens of AU, but that smaller scale disturbances may operate at

### 2015-04-03 paper.tex
minor writing edits in intro

@@
 isothermal disks. Our treatment is grossly simplified owing to a
-number of approximations made, but closely parallels previous work and 
-allows us to highlight the essential effects of vertical shear, and to
+number of approximations made, but closely parallels previous work, 
+allowing us to highlight the essential effects of vertical shear, and to
 derive a characteristic upper limit to the thermal timescale for the
-VSI. In \S\ref{numerical} we numerically
+VSI. This timescale is \emph{much shorter} than the dynamical
+timescale. In \S\ref{numerical} we numerically
 solve the linear problem in the radially local approximation, but
@@
 
-\section{Thermodynamic requirement for the vertical shear instability}\label{vsi_require}    
-In this paper we will find that in a stably stratified disk, the VSI
-requires a \emph{very} short thermal timescale $t_c$, much less than the
+\section{Thermodynamic requirement for the vertical shear instability}\label{vsi_require}     In this paper we will find that in a stably stratified thin disk, the VSI  
+requires a thermal timescale $t_c$ much less than the
 disk dynamical timescale, 
 \begin{align}
-  t _c \ll \Omega^{-1},  
+  t _c \ll \Omega^{-1}. 
 \end{align}
-consistent with numerical experiments \citep{nelson13}. For this
-preliminary discussion we shall consider a vertically isothermal disk
-with aspect-ratio $\epsilon$, 
+To see why this should be expected, we need to compare vertical shear
+with vertical buoyancy. For this preliminary discussion we consider a
+vertically isothermal disk with aspect-ratio $\epsilon$,  
 dimensionless radial temperature gradient $q$ and adiabatic index
-$\gamma$.
+$\gamma$. For a protoplanetary disk, $\epsilon \sim 0.05$, $|q|\sim 1$ and $\gamma\simeq 1.4$.
 
-In a thin disk, the specific angular momentum varies with height as 
+For a thin disk, the specific angular momentum varies with height as 
 \begin{align}\label{vshear_thin}
@@
 
-If $\gamma>1$ then the disk is stably stratified with a real vertical
-buoyancy frequency $N_z$.  For a thin disk,     
+For $\gamma>1$ the disk is stably stratified with a real vertical
+buoyancy frequency $N_z$.  In a thin disk,     
 \begin{align}\label{nz_thin}
@@
 However, the stabilizing effect of buoyancy can be mitigated for 
-non-adiabatic disturbances. Consider a fluid element displaced upwards, away from
+effectively isothermal disturbances. Consider a fluid element displaced upwards, away from
 the disk midplane. In a stably stratified adiabatic disk, it would
 sink because of buoyancy. However, if thermal timescales are short, in
-that temperature flucations decay rapidly (i.e. nearly isothermal
-perturbations), then the fluid element would attain its surrounding
+that temperature flucations decay rapidly, then the fluid element would attain its surrounding 
 density and  
@@
 
-How short must $t_c$ be for vertical shear to prevail? For a rough
+How short must $t_c$ be for vertical shear to prevail? A rough 
 comparison of vertical shear to buoyancy for non-adiabatic
-perturbations, we apply Eq. 47 of \cite{urpin03}, which is based on
+perturbations can be obtained by applying Eq. 47 of \cite{urpin03}, which is based on 
 \cite{townsend58},   
@@
 stabilizing; while the right-hand side represents vertical shear,
-which is destabilizing. In a thin disk, the destabilizing effect of
-vertical shear is rather weak because $|r\p_z\Omega|\propto \epsilon
-\ll 1$. By comparison the buoyancy frequency $N_z/\Omega\sim 1$ at
-$z\sim H$.
+which is destabilizing. In a thin disk $\epsilon \ll 1$, so the destabilizing effect of 
+vertical shear is rather weak because $|r\p_z\Omega|\sim\epsilon\Omega
+\ll \Omega$. By comparison, the buoyancy frequency $N_z\sim
+\Omega$. It follows that $t_c\lesssim O(\epsilon\Omega^{-1})$ is
+required for instability.  \emph{In a thin  
+  disk the thermal timescale must be shorter than the dynamical
+  timescale by $O(\epsilon)\ll 1$ for the VSI to operate.}
 
@@
 generally weak compared to buoyancy. Then even a small (positive)
-vertical entropy gradient can significantly stabilize the VSI. It
-follows that $t_c$ must be quite short to permit instability.
+vertical entropy gradient can significantly stabilize the VSI.
 
@@
 for the VSI in a vertically isothermal, thin disk. This equation
-reflects Eq. \ref{bcrit_est} but is simpler since there is no
-$z$ dependence. \emph{In a thin 
-  disk the thermal timescale must be shorter than the dynamical
-  timescale by $O(\epsilon)\ll 1$ for the VSI to operate.}
-
+reflects Eq. \ref{bcrit_est} but there is no $z$ dependence. 
 % While assumptions and simplifications will be made 
 % to derive this, 
-
 Eq. \ref{prelim_bcrit} can be 
@@
 correspond to smaller perturbations. The region outside the red line
-satisfies Eq. \ref{prelim_bcrit}. The criterion Eq. \ref{prelim_bcrit} in fact
+satisfies Eq. \ref{prelim_bcrit}.
+
+The criterion Eq. \ref{prelim_bcrit} in fact
 assumes $t_c$ is vertically uniform. While this is not generally true

### 2015-04-04 paper.tex
minor notation update

@@
 \begin{align}
-  t _c \ll \Omega^{-1}. 
+  t _c \ll \Omega_k^{-1}. 
 \end{align}
+where $\Omega_k$ is the Keplerian frequency. 
 To see why this should be expected, we need to compare vertical shear
@@
 \begin{align}\label{vshear_thin}
-  r \frac{\p \Omega^2 }{\p z} \simeq \epsilon q \Omega^2\left(\frac{z}{H}\right),
+  r \frac{\p \Omega^2 }{\p z} \simeq \epsilon q \Omega_k^2\left(\frac{z}{H}\right),
 \end{align}
@@
   N_z^2 \simeq \left(\frac{\gamma-1}{\gamma}\right) \left(\frac{z}{H}\right)^2
-  \Omega^2.  
+  \Omega_k^2.  
 \end{align}
@@
 which is destabilizing. In a thin disk $\epsilon \ll 1$, so the destabilizing effect of 
-vertical shear is rather weak because $|r\p_z\Omega|\sim\epsilon\Omega
-\ll \Omega$. By comparison, the buoyancy frequency $N_z\sim
-\Omega$. It follows that $t_c\lesssim O(\epsilon\Omega^{-1})$ is
+vertical shear is rather weak because $|r\p_z\Omega|\sim\epsilon\Omega_k
+\ll \Omega_k$. By comparison, the buoyancy frequency $N_z\sim
+\Omega_k$. It follows that $t_c\lesssim O(\epsilon\Omega_k^{-1})$ is
 required for instability.  \emph{In a thin  
@@
 To illustrate this point, we show in Fig. \ref{eqm_structure} the
-equilibrium structure of a typical protoplanetary disk. Here $\rho_0$
-is the midplane density and vertical shear rate and buoyancy
-frequencies are normalized by the local Keplerian frequency. It is
+equilibrium structure of a typical protoplanetary disk. %Here $\rho_0$
+%is the midplane density and vertical shear rate and buoyancy
+%frequencies are normalized by the local Keplerian frequency. 
+It is
 clear that vertical shear is 
@@
     Solar Nebla): density (left), vertical shear rate (middle) and
-    buoyancy frequency (right).  
+    buoyancy frequency (right). Here $\rho_0$ is the midplane density.  
     \label{eqm_structure} 
@@
 \begin{align}\label{prelim_bcrit}
-  (\gamma - 1)t_c < \epsilon |q| \Omega^{-1}
+  (\gamma - 1)t_c < \epsilon |q| \Omega_k^{-1}
 \end{align}   

### 2015-04-08 linear_iso.tex
moving more equations to appendix

@@
 
-\subsection{Relation to the vertically global
-  shearing box}\label{vgsb_approx} 
-We connect the linearized equations 
-Eq. \ref{lin_mass}---\ref{lin_energy} to the vertically global  
-shearing box (VGSB) formalism developed by \cite{mcnally14}. The VGSB
-is an extension of the standard shearing box \citep{goldreich65} to 
-background shear flows that are height dependent. 
-As in the  standard shearing box, in the VGSB the equilibrium disk
-structure is assumed to be homogeneous in the radial ($x$) direction.
+% \subsection{Relation to the vertically global
+%   shearing box}\label{vgsb_approx} 
+% We connect the linearized equations 
+% Eq. \ref{lin_mass}---\ref{lin_energy} to the vertically global  
+% shearing box (VGSB) formalism developed by \cite{mcnally14}. The VGSB
+% is an extension of the standard shearing box \citep{goldreich65} to 
+% background shear flows that are height dependent. 
+% As in the  standard shearing box, in the VGSB the equilibrium disk
+% structure is assumed to be homogeneous in the radial ($x$) direction.
 
-Thus, the linear problem in the VGSB can be derived from
-Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
-$\hat{g}_c=0$. The resulting problem is very similar 
-to that for local axisymmetric waves in accretion disks already
-discussed by \cite{lubow93}, but modified by the presence of vertical
-shear and thermal relaxation. This is convenient as it allows us to
-directly apply some of their solution methods.
+% Thus, the linear problem in the VGSB can be derived from
+% Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
+% $\hat{g}_c=0$. The resulting problem is very similar 
+% to that for local axisymmetric waves in accretion disks already
+% discussed by \cite{lubow93}, but modified by the presence of vertical
+% shear and thermal relaxation. This is convenient as it allows us to
+% directly apply some of their solution methods.
 
-After setting $\hat{g}_c=0$ in Eq. \ref{lin_mass}, \ref{lin_vx} and
-\ref{lin_energy}, we can eliminate variables 
-to obtain a pair or ordinary differential
-equations for $W,Q$ as
-\begin{align}
-  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
-  &=\left(\frac{\ii
-      k_x r}{D}\frac{d\Omega^2}{dz} -
-    \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
-    \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
-  &- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
-    \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
-  &\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
-  \frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
-  &=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
-  \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
-\end{align}
-where
-\begin{align}
-  D \equiv \kappa^2 - \sigma^2,
-\end{align} 
-and we have replaced the notation $\p_z$ by $d/dz$ since we are
-considering solutions at a fixed radius.
+% After setting $\hat{g}_c=0$ in Eq. \ref{lin_mass}, \ref{lin_vx} and
+% \ref{lin_energy}, we can eliminate variables 
+% to obtain a pair or ordinary differential
+% equations for $W,Q$ as
+% \begin{align}
+%   &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
+%   &=\left(\frac{\ii
+%       k_x r}{D}\frac{d\Omega^2}{dz} -
+%     \frac{d\ln{\rho}}{dz}\right)\left[\frac{dW}{dz} +
+%     \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] \notag\\
+%   &- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
+%     \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
+%   &\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
+%   \frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
+%   &=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
+%   \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],\label{ode_Q} 
+% \end{align}
+% where
+% \begin{align}
+%   D \equiv \kappa^2 - \sigma^2,
+% \end{align} 
+% and we have replaced the notation $\p_z$ by $d/dz$ since we are
+% considering solutions at a fixed radius.
 
-Extrapolating results obtained from the VGSB to the global disk has
-limitations. This is because within the VGSB framework, $d\Omega/dz$
-can be freely specified. 
-We choose $d\Omega/dz$ to be that due to the thermal structure in 
-the global disk (Eq. \ref{vertical_shear}). However, background radial
-gradient terms are ignored elsewhere in the VGSB (which is done by
-setting $\hat{g}_c=0$).
+% Extrapolating results obtained from the VGSB to the global disk has
+% limitations. This is because within the VGSB framework, $d\Omega/dz$
+% can be freely specified. 
+% We choose $d\Omega/dz$ to be that due to the thermal structure in 
+% the global disk (Eq. \ref{vertical_shear}). However, background radial
+% gradient terms are ignored elsewhere in the VGSB (which is done by
+% setting $\hat{g}_c=0$).
 
-We further discuss caveats of the VGSB in Appendix
-\ref{global_corr}. We find that for fast thermal relaxation
-($\beta\ll1$) and perturbations with small radial
-lengthscales ($k_xH\gg 1$), which characterizes the VSI, the VGSB
-framework is adequate for the purpose of a simple analytical
-treatment.
+% We further discuss caveats of the VGSB in Appendix
+% \ref{global_corr}. We find that for fast thermal relaxation
+% ($\beta\ll1$) and perturbations with small radial
+% lengthscales ($k_xH\gg 1$), which characterizes the VSI, the VGSB
+% framework is adequate for the purpose of a simple analytical
+% treatment.
 
-\subsection{Low-frequency and Keplerian
-  approximation}\label{approx_gov} 
-We now derive a governing equation with further approximations:
+\subsection{Shearing box, low-frequency and Keplerian
+  approximations}\label{approx_gov}
+
+In Appendix \ref{adia_improve}, we derive a governing equation with the following approximations:
+
 \begin{enumerate}
+
+\item%  We adopt the  `vertically global shearing box' (VGSB) formalism
+%   developed by \cite{mcnally14}, which is an extension to the standard
+%   shearing box \citep{goldreich65} background shear flows that are
+%   height dependent.  As in the standard shearing box, in the VGSB the equilibrium disk
+%   structure is assumed to be homogeneous in the radial ($x$) direction.
+
+% % Thus, the linear problem in the VGSB can be derived from
+% % Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
+% % $\hat{g}_c=0$.
+
+% in which explicit dependences on the background radial disk structure
+%   in Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
+%   $\hat{g}_c=0$.
+
+% However, the global radial temperature structure,
+%   which gives rise to vertical shear, is retained.
+
+%   We further discuss caveats of the VGSB in Appendix
+%   \ref{global_corr}. We find that for fast thermal relaxation
+%   ($\beta\ll1$) and perturbations with small radial
+%   lengthscales ($k_xH\gg 1$), which characterizes the VSI, the VGSB
+%   framework is adequate for the purpose of a simple analytical
+%   treatment.
+
 \item We consider eigenvalues such that
@@
 \end{enumerate}
-Making both of these approximations is equivalent to setting 
-$D\to\Omega_k^2$.  After making this replacement, we eliminate $W$ and
-$Q$ between  Eq. \ref{lin_vz} and Eq. \ref{ode_w}---\ref{ode_Q},
-making use of Eq. \ref{vertical_shear}, to
-obtain an equation for $\delta v_z$. The result for $\Gamma=1$ is
+% Making both of these approximations is equivalent to setting 
+% $D\to\Omega_k^2$.  After making this replacement, we eliminate $W$ and
+% $Q$ between  Eq. \ref{lin_vz} and Eq. \ref{ode_w}---\ref{ode_Q},
+% making use of Eq. \ref{vertical_shear}, to
+% obtain an equation for $\delta v_z$.
+
+%The result for $\Gamma=1$ is
 \begin{align}

### 2015-04-08 paper.tex
minor notation update
a punchier title and reworded abstract (with ultimately very similar content).  I think these versions are less technical and more inviting to the average arXiv reader, but still loyal to the work.   We can discuss iterations.

@@
 \defcitealias{lin12}{L12} 
+\def \OmK {\Omega_{\rm K}}
 
@@
 
-\title{Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula} 
-\author{Min-Kai Lin, Andrew Youdin}
-\affil{ $^1$Department of Astronomy and Steward Observatory, University of
+\title{Cooling Requirements for the Vertical Shear Instability, \\
+ with Implications for Protoplanetary Disks}
+%\title{Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula} 
+\author{Min-Kai Lin \& Andrew Youdin}
+\affil{Department of Astronomy and Steward Observatory, \\
+University of
   Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
@@
 \begin{abstract}
-  A robust hydrodynamic mechanism for turbulent angular momentum
-  transport is important for cold accretion disks such as 
-  protoplanetary disks. The vertical shear instability (VSI) has 
-  recently been suggested to provide such a pathway, since it operates in
-  disks with rotation profiles that depend on the height from the 
-  disk midplane, which is generally expected in  
-  astrophysical disks. However, vertical shear is weak in thin disks, 
-  so the VSI can easily be stabilized by buoyancy forces in the
-  presence of a stable vertical entropy  gradient. % , which is also
-  % expected in astrophysical disks. 
-  Consequently,  the VSI
-  requires a rather short thermal or cooling timescale in
-  sub-adabatically stratified disks.      
-  A quantitative examination of this thermodynamic 
-  requirement is needed to assess the viability of
-  the VSI in realistic disks. To this end,  
-  we study the linear, axisymmetric VSI in radially local, 
-  vertically global models of stably stratified accretion disks in which the disk  
-  temperature $T$ is relaxed to its initial value over a characteristic timescale 
-  $t_c=\beta\Omega_k^{-1}$, where $\beta$ is a dimensionless parameter 
-  and $\Omega_k$ is the Keplerian rotation frequency. % We explore how
-  % the VSI depends on the thermodynamic response of the disk by
-  % varying $\beta$.
-  We find that 
-  increasing $\beta$ even slightly away from    
-  zero dramatically reduce VSI growth rates. 
-  For vertically isothermal disks, we quantify the
-  `short' thermal timescale requirement for the VSI as $\beta < \epsilon|q|/(\gamma-1)$, where
-  $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  is the radial temperature gradient, and 
-  $\gamma$ is the adiabatic index. Thus, in a thin disk the thermal timescale must be shorter than 
-  the dyamical timescale by $O(\epsilon)\ll 1$ for the VSI to operate.  
-  We apply our results  
-  to a disk model of the Minimum Mass Solar Nebula (MMSN) with an  
-  estimate for the thermal timescale $\beta$ in the MMSN based on dust
-  opacity. In the inner parts of the MMSN, e.g. at $\lesssim 5$AU, 
-  we find the VSI would only operate on extremely short radial lengthscales, $\ll 
-  O(10^{-2}H)$, where $H$ is the disk scale height. On longer radial 
-  lengthscales of $O(10^{-2}H)$ to $O(10^{-1}H)$, we find the MMSN is
-  unstable to the VSI from $\sim 5$AU to $\sim 50$AU with a
-  characteristic growth time of $\sim 30$ local orbital periods.  
-  % ,and the dominant VSI mode in the MMSN is the `fundamental
-  % corrugation mode' which causes the disk column to oscillate
-  % vertically.
-  We thus suggest the VSI is at least dynamically important in the 
-  outer parts of protoplanetary disks at tens of AU. 
+  It is difficult to understand how cold circumstellar disks accrete onto their central stars.  Due to 
+  lower levels of ionization, the magnetic transport mechanisms that operate in warmer disks are less effective. 
+  A hydrodynamic mechanism, the vertical shear instability (VSI), offers a means to drive angular momentum 
+  transport and accretion flows in cold protoplanetary disks.  
+  The VSI is driven by a weak vertical gradient in the disk's orbital motion.
+  In order to grow, the VSI must overcome vertical buoyancy, a strong stabilizing influence in cold disks, where 
+  heating is dominated by external irradiation.  Rapid cooling, via radiative losses, reduces the effective buoyancy
+  and allows the VSI to operate.  In this paper, we quantify the cooling timescale, $t_c$, needed for growth of the VSI
+  and consider the implications for protoplanetary disks where $t_c$ is determined by the opacity of dust 
+  grains.  We perform a linear analysis of the VSI with cooling in vertically global and radially local disk models.
+  For irradiated disks, we find that the VSI is most vigorous for rapid cooling with $t_c < \Omega_K^{-1} \epsilon |q| / (\gamma -1)$
+  in terms of the Keplerian orbital frequency, $\OmK$; the disk's aspect ratio, $\epsilon \ll 1$; the radial 
+  powerlaw temperature gradient, $q$; and the adiabatic index, $\gamma$.  For longer cooling times, the VSI 
+  is much less effective because growth slows and shifts to smaller length scales, which are more prone to 
+  viscous or turbulent decay.  We apply our results to a  Minimum Mass Solar Nebula (MMSN) disk model.  We
+  find that the VSI is most effective at intermediate radii, from $\sim 5$AU to $\sim 50$AU with a
+  characteristic growth time of $\sim 30$ local orbital periods.  Growth is suppressed by long cooling times both in the opaque 
+  inner disk and the optically thin outer disk.  A reduction in the dust opacity by a factor of 10 increases cooling times 
+  enough to quench the VSI at all disk radii.  Thus the formation of solid protoplanets, a sink for dust grains, can impede the VSI.
+
+%
+
+%  A robust hydrodynamic mechanism for turbulent angular momentum
+%  transport is important for cold accretion disks such as 
+%  protoplanetary disks. The vertical shear instability (VSI) has 
+%  recently been suggested to provide such a pathway, since it operates in
+%  disks with rotation profiles that depend on the height from the 
+%  disk midplane, which is generally expected in  
+%  astrophysical disks. However, vertical shear is weak in thin disks, 
+%  so the VSI can easily be stabilized by buoyancy forces in the
+%  presence of a stable vertical entropy  gradient. % , which is also
+%  % expected in astrophysical disks. 
+%  Consequently,  the VSI
+%  requires a rather short thermal or cooling timescale in
+%  sub-adabatically stratified disks.      
+%  A quantitative examination of this thermodynamic 
+%  requirement is needed to assess the viability of
+%  the VSI in realistic disks. To this end,  
+%  we study the linear, axisymmetric VSI in radially local, 
+%  vertically global models of stably stratified accretion disks in which the disk  
+%  temperature $T$ is relaxed to its initial value over a characteristic timescale 
+%  $t_c=\beta\Omega_k^{-1}$, where $\beta$ is a dimensionless parameter 
+%  and $\Omega_k$ is the Keplerian rotation frequency. % We explore how
+%  % the VSI depends on the thermodynamic response of the disk by
+%  % varying $\beta$.
+%  We find that 
+%  increasing $\beta$ even slightly away from    
+%  zero dramatically reduce VSI growth rates. 
+%  For vertically isothermal disks, we quantify the
+%  `short' thermal timescale requirement for the VSI as $\beta < \epsilon|q|/(\gamma-1)$, where
+%  $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  is the radial temperature gradient, and 
+%  $\gamma$ is the adiabatic index. Thus, in a thin disk the thermal timescale must be shorter than 
+%  the dyamical timescale by $O(\epsilon)\ll 1$ for the VSI to operate.  
+%  We apply our results  
+%  to a disk model of the Minimum Mass Solar Nebula (MMSN) with an  
+%  estimate for the thermal timescale $\beta$ in the MMSN based on dust
+%  opacity. In the inner parts of the MMSN, e.g. at $\lesssim 5$AU, 
+%  we find the VSI would only operate on extremely short radial lengthscales, $\ll 
+%  O(10^{-2}H)$, where $H$ is the disk scale height. On longer radial 
+%  lengthscales of $O(10^{-2}H)$ to $O(10^{-1}H)$, we find the MMSN is
+%  unstable to the VSI from $\sim 5$AU to $\sim 50$AU with a
+%  characteristic growth time of $\sim 30$ local orbital periods.  
+%  % ,and the dominant VSI mode in the MMSN is the `fundamental
+%  % corrugation mode' which causes the disk column to oscillate
+%  % vertically.
+%  We thus suggest the VSI is at least dynamically important in the 
+%  outer parts of protoplanetary disks at tens of AU. 
 \end{abstract}
@@
 vertical entropy gradient can significantly stabilize the VSI. 
+Note that in astrophysical disks both $N_z^2$ and $\p_z\Omega$ varies
+with height. Eq. \ref{bcrit_est} then suggest that with
+increasing height, buoyancy would eventually stabilize the VSI (since
+$N_z^2\propto z^2$ and $\p_z\Omega\propto z$).
+% This means that 
+% disturbances localized away from the disk midplane are easily stabilized by
+% buoyancy. 
+This argument assumes vertically localized disturbances. It is,
+however, not clear how to apply Eq. \ref{bcrit_est} to vertically
+global disturbances, as considered in this work.
 
@@
 \end{figure*}
-
-Note that in astrophysical disks both $N_z^2$ and $\p_z\Omega$ varies
-with height, as evident in
-Fig. \ref{eqm_structure}. Eq. \ref{bcrit_est} then suggest that with
-increasing height, buoyancy would eventually stabilize the VSI (since 
-$N_z^2\propto z^2$ and $\p_z\Omega\propto z$). 
-% This means that 
-% disturbances localized away from the disk midplane are easily stabilized by
-% buoyancy. 
-This argument assumes vertically localized disturbances. It is,
-however, not clear how to apply Eq. \ref{bcrit_est} to vertically
-global disturbances, as considered in this work.
-
-%it would imply you can always have VSI sufficiently close to the
-%midplane
 

### 2015-04-08 appendix.tex
moving more equations to appendix
restructured: moved more equations to appendix, re-ordered appendix, slightly shortened analytical section

@@
-\section{Applicability of the vertically global shearing box}\label{global_corr}
-In \S\ref{analytical} we effectively employed the vertically global shearing box
-\citep[VGSB,][]{mcnally14} to make analytic progress.   
-The difference between the linear problem in the VGSB 
-and the radially-local approximation adopted in our numerical study
-is the absence of terms explicitly associated with the radial disk structure in
-the VGSB. These terms appear in the energy, mass and radial momentum
-equations and are labeled by $\hat{g}_c$ in
-Eq. \ref{lin_mass}---\ref{lin_energy}. Nominally $\hat{g}_c=1$.
+\section{Full governing equation for vertically isothermal disks in
+  the vertically global shearing box}\label{adia_improve}
+Here we explain the derivation of Eq. \ref{vertiso_gov} used in our 
+analytical discussion in \S\ref{analytical}. We shall need the
+vertical derivative of the epicyclic frequency $\kappa$. In the global
+disk with $\Gamma=1$ we have 
+\begin{align}\label{dkappa2}
+  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
+    r}\frac{\p\Omega^2}{\p z} = -
+  \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(
+    \frac{3z^2}{r^2+z^2}-1\right) \equiv - \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2}F(r,z).
+\end{align}
+% For the local problem we may then write
+% \begin{align}
+%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z),
+% \end{align}
+The function $F$ increases monotonically from $F=-1$ at $z=0$ to $F\to2$
+as $|z|\to\infty $, so $|F|=O(1)$.
+
+In the vertically global shearing box (VGSB) approximation, we set
+$\hat{g}_c=0$ in Eq. \ref{lin_mass} --- \ref{lin_energy}. We then eliminate
+the horizonatal velocity perturbations ($\delta v_x,\, \delta v_y$) to
+obtain  
+%we can eliminate variables 
+%to obtain a pair or ordinary differential
+%equations for $W,Q$ as 
+\begin{align}
+  & \ii\sigma\delta v_z = \frac{dW}{dz} + \frac{d\ln{\rho}}{d z}\left(W-Q\right),\label{ode_vz} \\
+  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W =\ii\sigma \left(\frac{\ii
+      k_x r}{D}\frac{d\Omega^2}{dz} -
+    \frac{d\ln{\rho}}{dz}\right)\delta v_z - \ii\sigma \frac{d\delta v_z}{dz},
+  % \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] 
+  % \notag\\
+  % &- \frac{d^2W}{dz^2} - \frac{d\ln{\rho}}{dz}\left(\frac{dW}{dz} -
+  %   \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),
+  \label{ode_w}\\
+%  &\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
+%  \frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right) 
+%  =\ii \sigma c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} -
+%    1\right)\delta v_z,  
+  &\sigma^2W - \gamma\sigma^2 Q +
+  \frac{\ii\sigma}{t_c}\left(W-Q\right) 
+  =\ii \sigma c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma -
+    1\right)\delta v_z,
+  % 
+  % \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right],
+  \label{ode_Q} 
+\end{align}
+(note that $\Gamma=1$ has been used) where
+\begin{align}
+  D \equiv \kappa^2 - \sigma^2,
+\end{align} 
+and we have replaced the notation $\p_z$ by $d/dz$ since we are
+considering solutions at a fixed radius.
+
+% In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
+% eliminating variables to obtain a single equation for $\delta v_z$ in
+% vertically isothermal disks, Eq. \ref{vertiso_gov}.  
+% This procedure ignores the vertical dependence of
+% $D=\kappa^2(z) - \sigma^2$. We show here that this has no serious 
+% consequence for thin disks.
+
+%For the radially local problem, we eliminate $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
+%Eq. \ref{lin_vz}  (with $\Gamma=1$), but now retain the vertical dependence
+%of $D$ using Eq. \ref{dkappa2}, to obtain
+
+We now eliminate $W,\, Q$ from Eq. \ref{ode_vz} --- \ref{ode_Q} in
+favor of $\delta v_z$, making use of Eq. \ref{dkappa2} to evaluate
+$dD/dz$, to obtain
+
+\begin{align}\label{full_ode}
+  0 =& \frac{d^2\delta v_z}{dz^2} + \left[1 + \frac{\ii k_x c_s^2
+      q}{Dr} - \frac{k_x^2c_s^2}{\left(k_x^2c_s^2 + \chi
+        D\right)}\frac{qc_s^2F}{Dr^2}\right]\frac{d\ln\rho}{dz}\frac{d\delta
+    v_z}{dz} \notag\\
+  &+ \left\{\sigma^2\left(\frac{k_x^2}{D} +
+      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
+        q}{Dr}\right)\frac{d^2\ln\rho}{dz^2} -
+    \frac{c_s^2}{D}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
+      \frac{\ii k_x q}{r}\right)
+   \left[\left(1-\chi\right) +
+     \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
+   \right\}\delta v_z,
+\end{align}
+where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
+t_c \gamma\right)$. Making the low frequency and Keplerian
+approximations is equivalent to the replacement $D\to
+\Omega_k^2$. This gives, in terms of dimensionless variables,  
+\begin{align}
+  &\delta v_z ^{\prime\prime} + \left[1 + \ii\epsilon q\hat{k} -
+    \frac{ \hat{k}^2}{
+      \left(\hat{k}^2+\chi\right)}q \epsilon^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
+  \left\{\left(\chi + \ii \epsilon q
+      \hat{k}\right)\ln\rho^{\prime\prime} - \ln\rho^{\prime
+      2}\left(\khat^2 -
+      \ii\epsilon
+      q\hat{k}\right)\left[1 - \chi +
+      \frac{\chi}{\left(\hat{k}^2+\chi\right)}q\epsilon^2F\right]\right\}\delta v_z \notag\\&=
+  -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
+\end{align} 
+Since we are interested in thin disks with $\epsilon \ll 1$ and $|q| =
+O(1)$, we neglect the terms proportional to $\epsilon^2$, which gives 
+Eq. \ref{vertiso_gov_nondim} (the dimensional version being
+Eq. \ref{vertiso_gov}). This is equivalent to neglecting the   
+vertical dependence of $D$ by setting $D\to\Omega_k^2$ in
+Eq. \ref{ode_w} before eliminating variables in favor of $\delta
+v_z$, which is the simpler way to obtain Eq. \ref{vertiso_gov}.
+
+%Thus, a simpler route to the governing
+%equation (Eq. )
+
+%Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
+%proportional to $\epsilon^2$. For a thin disk ($\epsilon\ll1$) with
+%$|q|=O(1)$, as considered in this work, neglecting these terms has no
+%qualitative effect on the discussion in \S\ref{analytic_relax}. 
+% In deriving Eq. \ref{vertiso_gov}, the vertical dependence of
+% $D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
+% $D\to\Omega_k^2$ was made before eliminating variables in favor of
+% $\delta v_z$. In Appendix \ref{adia_improve} we show that making this
+% replacement after eliminating variables, which involves $dD/dz$,
+% introduces terms of $O(\epsilon^2)$. Since we are interested in 
+% thin disks ($\epsilon\ll 1 $), e.g. protoplanetary disks where 
+% $\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
+% unnecessary complexity for the present discussion.
+
+\section{Applicability of the  VGSB}\label{global_corr}
+%In \S\ref{analytical} we effectively employed the vertically global
+%shearing box \citep[VGSB,][]{mcnally14} to make analytic progress. 
+The vertically global shearing box \citep[VGSB,][]{mcnally14} is an
+extension of the standard shearing box \citep[SSB,][]{goldreich65} to
+background shear flows that are height dependent. 
+%As in the SSB, in
+%the VGSB the equilibrium disk structure is assumed to be homogeneous
+%in the radial ($x$) direction.  
+Within the VGSB framework, the vertical shear $d\Omega/dz$
+can be freely specified. We choose $d\Omega/dz$ to be that due to the
+radial temperature gradient in the global disk
+(Eq. \ref{vertical_shear}). However, explicit dependencies on the
+radial disk structure are ignored elsewhere in the linearized equations
+as described below. We show in the following sections that this
+inconsistency only becomes important in the adiabatic limit, which is
+not the relevant regime for the VSI.
+
+%is only important when
+%considering adiabatic limit, which is not the relevant regime for the
+%VSI.
+
+The background radial disk structure explicitly appears 
+in the linearized mass, radial momentum and energy equations
+(Eq. \ref{lin_mass}, \ref{lin_vx} and \ref{lin_energy}) as the terms
+proportional to $\hat{g}_c$. Nominally $\hat{g}_c=1$.      
+In the VGSB (and SSB) these terms are neglected by setting
+$\hat{g}_c=0$. For a power-law disk, these radial gradients  
+are $O(r^{-1})$, and they appear in comparison with terms of
+$O(k_x)$. The neglected terms therefore have a relative magnitude of
+$O(\epsilon/\khat)$, which is small for thin disks ($\epsilon\ll1$)
+and/or small radial wavelengths ($\khat\gg 1$).
+
+% The difference between the linear problem in the VGSB 
+% and the radially-local approximation adopted in our numerical study
+% is the absence of terms explicitly associated with the radial disk structure in
+% the VGSB.
+
+%However, background radial gradient terms 
+%are ignored elsewhere in the VGSB (which is done by setting
+%$\hat{g}_c=0$).
 
@@
 
-However, in the VGSB these terms are neglected by 
-setting $\hat{g}_c=0$. For a power-law disk, these radial gradients
-are $O(r^{-1})$, and they appear in comparison with terms of
-$O(k_x)$. The neglected terms therefore have a relative magnitude of
-$O(\epsilon/\khat)$. This is small for thin disks ($\epsilon\ll1$)
-and/or small radial wavelengths ($\khat\gg 1$).
-
 \subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
@@
 disks and setting  $\beta\to\infty$ in 
-Eq. \ref{vertiso_gov_nondim} (which itself is obtained by setting
-$\hat{g}_c=0$ in Eq. \ref{lin_mass}---\ref{lin_energy}) to give 
+Eq. \ref{vertiso_gov_nondim} % (which itself is obtained by setting
+% $\hat{g}_c=0$ in Eq. \ref{lin_mass}---\ref{lin_energy})
+to give 
 \begin{align}
@@
 \end{figure}
-
-\section{Full governing equation for vertically isothermal disks in
-  the VGSB}\label{adia_improve}
-
-In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
-eliminating variables to obtain a single equation for $\delta v_z$ in
-vertically isothermal disks, Eq. \ref{vertiso_gov}.  
-This procedure ignores the vertical dependence of
-$D=\kappa^2(z) - \sigma^2$. We show here that this has no serious 
-consequence for thin disks.
-
-In the global disk with $\Gamma=1$ we have
-\begin{align}\label{dkappa2}
-  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
-    r}\frac{\p\Omega^2}{\p z} = -
-   \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(
-  \frac{3z^2}{r^2+z^2}-1\right) \equiv - \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2}F(r,z).
-\end{align}
-% For the local problem we may then write
-% \begin{align}
-%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z),
-% \end{align}
-The function $F$ increases monotonically from $F=-1$ at $z=0$ to $F\to2$
-as $|z|\to\infty $, so $|F|=O(1)$. 
-% For the local problem we set
-% $dD/dz = \p\kappa^2/\p z$. 
-% This function varies from $F=q$ at $z=0$ to $F\to
-% 3+q$ as $|z|\to\infty$. Since we consider $|q|=O(1)$, this impies
-% $|F|=O(1)$ as well.
-
-For the radially local problem, we eliminate $W$ and $Q$ from Eq. \ref{ode_w}---\ref{ode_Q} and
-Eq. \ref{lin_vz}  (with $\Gamma=1$), but now retain the vertical dependence
-of $D$ using Eq. \ref{dkappa2}, to obtain  
-\begin{align}
-  0 =& \frac{d^2\delta v_z}{dz^2} + \left[1 + \frac{\ii k_x c_s^2
-      q}{Dr} - \frac{k_x^2c_s^2}{\left(k_x^2c_s^2 + \chi
-        D\right)}\frac{qc_s^2F}{Dr^2}\right]\frac{d\ln\rho}{dz}\frac{d\delta
-    v_z}{dz} \notag\\
-  &+ \left\{\sigma^2\left(\frac{k_x^2}{D} +
-      \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
-        q}{Dr}\right)\frac{d^2\ln\rho}{dz^2} -
-    \frac{c_s^2}{D}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
-      \frac{\ii k_x q}{r}\right)
-   \left[\left(1-\chi\right) +
-     \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
-   \right\}\delta v_z,
-\end{align}
-where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
-t_c \gamma\right)$. Making the low frequency, Keplerian
-approximations with the replacement $D\to \Omega_k^2$ gives, in terms
-of dimensionless variables,
-\begin{align}
-   &\delta v_z ^{\prime\prime} + \left[1 + \ii\epsilon q\hat{k} -
-    \frac{ \hat{k}^2}{
-      \left(\hat{k}^2+\chi\right)}q \epsilon^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
-  \left\{\left(\chi + \ii \epsilon q
-      \hat{k}\right)\ln\rho^{\prime\prime} - \ln\rho^{\prime
-      2}\left(\khat^2 -
-      \ii\epsilon
-      q\hat{k}\right)\left[1 - \chi +
-      \frac{\chi}{\left(\hat{k}^2+\chi\right)}q\epsilon^2F\right]\right\}\delta v_z \notag\\&=
-  -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
-\end{align}   
-Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
-proportional to $\epsilon^2$. For a thin disk ($\epsilon\ll1$) with
-$|q|=O(1)$, as considered in this work, neglecting these terms has no
-qualitative effect on the discussion in \S\ref{analytic_relax}.
-
-\section{Coefficients for the dispersion relation for perturbations
-with thermal relaxation}\label{relax_coeff}
-The coefficients of the dispersion relation Eq. \ref{relax_disp} is
-given by:
-\begin{align}
-  &c_0 = M(M+1)A^2,\\
-  &c_1 = \ii\beta\left\{\left(1-\gamma\right)\left[1 +
-      \khat^2\left(1+2M\right)^2 - 4 \ii\epsilon q\khat M (M+1)\right] 
-    - 2A^2\gamma M (M+1)\right\},\\
-  &c_2 = \left(\khat^2 + 1\right)A + \beta^2\left\{(1-\gamma)\left[1
-      + \gamma \khat^2(1+2M)^2 - 4\ii\epsilon q \khat \gamma M(M+1)
-    \right]
-    -\gamma^2 A^2 M(M+1)
-  \right\},\\
-  &c_3 = \beta\left\{\epsilon q \khat + \gamma \left[\ii + \epsilon q
-      k \left(1+2\khat^2\right)\right] - 3\ii - 2\ii
-    \khat^2\right\},\\
-  &c_4 =
-  \beta^2\left(1+\gamma\khat^2\right)\left[\gamma\left(1-\ii\epsilon q
-    \khat\right)-2\right] - \left(1+\khat^2\right)^2,\\
-&c_5 = 2\ii\beta\left(1+\khat^2\right)\left(1+\gamma\khat^2\right),\\
-&c_6 = \beta^2\left(1+\gamma\khat^2\right)^2.
-\end{align}
-
-\subsection{Limiting form with real frequencies and $\khat^2\gg 1$}\label{disp_neut_limit}
-Assuming neutral modes with real $\hat{\sigma}=\hat{\omega}$ and $\khat^2\gg 1$, the real and imaginary parts of
-Eq. \ref{relax_disp} are approximately 
-\begin{align}
-  0 =&M(M+1)(1-\epsilon^2 q^2 \khat^2) + 4\beta\epsilon q M(M+1)\left(\hat{\omega}\khat\right) 
- +\left[1 +
-    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
- + 2\epsilon q \gamma\beta \left(\hat{\omega}\khat\right)^3 -  \left(\hat{\omega}\khat\right)^4\notag \\
-  &+ \beta^2\gamma^2\hat{\omega}^2
-  \left(\hat{\omega}\khat\right)^4,\label{relax_cond1} \\
-   0=& 2\epsilon q M (M+1)\khat +
-   \beta(1-\gamma)(1+2M)^2\khat\left(\hat{\omega}\khat\right) 
-   + \epsilon q \khat \left(\hat{\omega}\khat\right)^2 -
-   2\beta\hat{\omega}\left(\hat{\omega}\khat\right)^2
-   - \epsilon q
-   \gamma^2\beta^2\hat{\omega}\left(\hat{\omega}\khat\right)^3 
-   +
-   2\beta\gamma\hat{\omega}\left(\hat{\omega}\khat\right)^4 \label{relax_cond2}. 
-\end{align}
-These equations are used to derive a critical thermal timescale in
-\S\ref{iso_vsi_beta_crit}. For fixed $M=O(1)$,  we assume
-$|\hat{\omega}\khat|$ remain constant but consider $\khat\gg 1$ so
-that $|\hat{\omega}|\ll 1$.  Then for a thin disk $\epsilon \ll 1$
-we neglect the second, fourth and last term in Eq. \ref{relax_cond1};
-and neglect the last three terms in Eq. \ref{relax_cond2}. This
-approximation gives Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2},
-but is not valid if $\beta$ or $|q|$ is large.
-
-For the fundamental mode $M=0$ and Eq. \ref{relax_cond1} ---
-\ref{relax_cond2} simplify significantly. It is then straight
-forward to eliminate $\hat{\omega}\khat$ between 
-Eq. \ref{relax_cond_simp1} and \ref{relax_cond_simp2} to obtain
-$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}) without
-assuming $|\epsilon q \khat|\lesssim 1$, as done in the main text to obtain the
-equivalent expression for $M>0$ (Eq. \ref{bcrit_gen}).
-
-%Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
-% $\khat\gg1$, we find
-% \begin{align}
-%   0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
-%   \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
-%   &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
-%   0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
-%   - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
-%     &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
-% \label{relax_cond2}
-% \end{align}
-% This is a pair of simultaneous equations for
-% $\hat{\omega}$ and $\beta$. Now, low-frequency axisymmetric waves in
-% the absence of vertical shear are stable and  satisfy
-% $\hat{\omega}\propto \khat^{-1}$ (Appendix \ref
-% {stable_novshear}, Eq. \ref{iso_disp_full}).  We are thus motivated to
-% seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
-% % we wish to seek a criterion independent of k 
-% % \begin{align}
-% %   \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
-% %\end{align}
-% \begin{align}
-%   \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
-%   |\hat{\omega}\khat|=O(1). 
-% \end{align}
-% Then
-% \begin{align}
-%   \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
-% \end{align}
-% from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
-% that
-% \begin{align}\label{beta_crit0}
-%   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
-%         q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
-% \end{align}
 
@@
 
+\section{Coefficients for the dispersion relation for perturbations
+with thermal relaxation}\label{relax_coeff}
+The coefficients of the dispersion relation Eq. \ref{relax_disp} is
+given by:
+\begin{align}
+  &c_0 = M(M+1)A^2,\\
+  &c_1 = \ii\beta\left\{\left(1-\gamma\right)\left[1 +
+      \khat^2\left(1+2M\right)^2 - 4 \ii\epsilon q\khat M (M+1)\right] 
+    - 2A^2\gamma M (M+1)\right\},\\
+  &c_2 = \left(\khat^2 + 1\right)A + \beta^2\left\{(1-\gamma)\left[1
+      + \gamma \khat^2(1+2M)^2 - 4\ii\epsilon q \khat \gamma M(M+1)
+    \right]
+    -\gamma^2 A^2 M(M+1)
+  \right\},\\
+  &c_3 = \beta\left\{\epsilon q \khat + \gamma \left[\ii + \epsilon q
+      k \left(1+2\khat^2\right)\right] - 3\ii - 2\ii
+    \khat^2\right\},\\
+  &c_4 =
+  \beta^2\left(1+\gamma\khat^2\right)\left[\gamma\left(1-\ii\epsilon q
+    \khat\right)-2\right] - \left(1+\khat^2\right)^2,\\
+&c_5 = 2\ii\beta\left(1+\khat^2\right)\left(1+\gamma\khat^2\right),\\
+&c_6 = \beta^2\left(1+\gamma\khat^2\right)^2.
+\end{align}
+
+\subsection{Limiting form with real frequencies and $\khat^2\gg 1$}\label{disp_neut_limit}
+Assuming neutral modes with real $\hat{\sigma}=\hat{\omega}$ and $\khat^2\gg 1$, the real and imaginary parts of
+Eq. \ref{relax_disp} are approximately 
+\begin{align}
+  0 =&M(M+1)(1-\epsilon^2 q^2 \khat^2) + 4\beta\epsilon q M(M+1)\left(\hat{\omega}\khat\right) 
+ +\left[1 +
+    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
+ + 2\epsilon q \gamma\beta \left(\hat{\omega}\khat\right)^3 -  \left(\hat{\omega}\khat\right)^4\notag \\
+  &+ \beta^2\gamma^2\hat{\omega}^2
+  \left(\hat{\omega}\khat\right)^4,\label{relax_cond1} \\
+   0=& 2\epsilon q M (M+1)\khat +
+   \beta(1-\gamma)(1+2M)^2\khat\left(\hat{\omega}\khat\right) 
+   + \epsilon q \khat \left(\hat{\omega}\khat\right)^2 -
+   2\beta\hat{\omega}\left(\hat{\omega}\khat\right)^2
+   - \epsilon q
+   \gamma^2\beta^2\hat{\omega}\left(\hat{\omega}\khat\right)^3 
+   +
+   2\beta\gamma\hat{\omega}\left(\hat{\omega}\khat\right)^4 \label{relax_cond2}. 
+\end{align}
+These equations are used to derive a critical thermal timescale in
+\S\ref{iso_vsi_beta_crit}. For fixed $M=O(1)$,  we assume
+$|\hat{\omega}\khat|$ remain constant but consider $\khat\gg 1$ so
+that $|\hat{\omega}|\ll 1$.  Then for a thin disk $\epsilon \ll 1$
+we neglect the second, fourth and last term in Eq. \ref{relax_cond1};
+and neglect the last three terms in Eq. \ref{relax_cond2}. This
+approximation gives Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2},
+but is not valid if $\beta$ or $|q|$ is large.
+
+For the fundamental mode $M=0$ and Eq. \ref{relax_cond1} ---
+\ref{relax_cond2} simplify significantly. It is then straight
+forward to eliminate $\hat{\omega}\khat$ between 
+Eq. \ref{relax_cond_simp1} and \ref{relax_cond_simp2} to obtain
+$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}) without
+assuming $|\epsilon q \khat|\lesssim 1$, as done in the main text to obtain the
+equivalent expression for $M>0$ (Eq. \ref{bcrit_gen}).
+
+%Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
+% $\khat\gg1$, we find
+% \begin{align}
+%   0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
+%   \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
+%   &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
+%   0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
+%   - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
+%     &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
+% \label{relax_cond2}
+% \end{align}
+% This is a pair of simultaneous equations for
+% $\hat{\omega}$ and $\beta$. Now, low-frequency axisymmetric waves in
+% the absence of vertical shear are stable and  satisfy
+% $\hat{\omega}\propto \khat^{-1}$ (Appendix \ref
+% {stable_novshear}, Eq. \ref{iso_disp_full}).  We are thus motivated to
+% seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
+% % we wish to seek a criterion independent of k 
+% % \begin{align}
+% %   \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
+% %\end{align}
+% \begin{align}
+%   \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
+%   |\hat{\omega}\khat|=O(1). 
+% \end{align}
+% Then
+% \begin{align}
+%   \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
+% \end{align}
+% from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
+% that
+% \begin{align}\label{beta_crit0}
+%   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
+%         q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
+% \end{align}
+
 \section{Fiducial model for a protoplanetary disk}\label{mmsn}

### 2015-04-08 linear_iso.tex
restructured: moved more equations to appendix, re-ordered appendix, slightly shortened analytical section

@@
 
-In Appendix \ref{adia_improve}, we derive a governing equation with the following approximations:
+In Appendix \ref{adia_improve} we derive a governing equation with the following simplifications:
 
@@
 
-\item%  We adopt the  `vertically global shearing box' (VGSB) formalism
-%   developed by \cite{mcnally14}, which is an extension to the standard
-%   shearing box \citep{goldreich65} background shear flows that are
-%   height dependent.  As in the standard shearing box, in the VGSB the equilibrium disk
-%   structure is assumed to be homogeneous in the radial ($x$) direction.
+\item  We adopt the  `vertically global shearing box' (VGSB) formalism 
+  developed by \cite{mcnally14}, in which explicit dependencies on the
+  background radial disk structure in 
+  Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
+  $\hat{g}_c=0$. However, the global radial temperature 
+  gradient in the expression for the 
+  vertical shear (Eq. \ref{vertical_shear}) is retained.
 
-% % Thus, the linear problem in the VGSB can be derived from
-% % Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
-% % $\hat{g}_c=0$.
-
-% in which explicit dependences on the background radial disk structure
-%   in Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
-%   $\hat{g}_c=0$.
-
-% However, the global radial temperature structure,
-%   which gives rise to vertical shear, is retained.
-
-%   We further discuss caveats of the VGSB in Appendix
-%   \ref{global_corr}. We find that for fast thermal relaxation
-%   ($\beta\ll1$) and perturbations with small radial
-%   lengthscales ($k_xH\gg 1$), which characterizes the VSI, the VGSB
-%   framework is adequate for the purpose of a simple analytical
-%   treatment.
+  In Appendix \ref{global_corr}  we discuss limitations of using
+  the VGSB, where we show that this approximation is valid provided we 
+  consider consider fast thermal relaxation times
+  ($\beta\lesssim 1$) and perturbations with small radial lengthscales
+  ($k_xH\gtrsim 1$), as appropriate for the VSI.
 
 \item We consider eigenvalues such that
-  $|\sigma^2|\ll \kappa^2$. Then we can replace $D=\kappa^2 -\sigma^2\to
-  \kappa^2$. %\citep[e.g.][]{lin10}. 
-  We call this the low frequency approximation. We are therefore
-  considering inertial waves \citep{barker15}.
+  $|\sigma^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency
+  approximation. We are therefore considering inertial waves
+  \citep{barker15}.
 
@@
 
+These approximations lead to Eq. \ref{full_ode} which, for a thin disk
+$\epsilon\ll 1$, simplifies further to  
 %The result for $\Gamma=1$ is
@@
 where $^\prime$ denotes $d/d\zhat$.
-
-In deriving Eq. \ref{vertiso_gov}, the vertical dependence of
-$D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
-$D\to\Omega_k^2$ was made before eliminating variables in favor of
-$\delta v_z$. In Appendix \ref{adia_improve} we show that making this
-replacement after eliminating variables, which involves $dD/dz$,
-introduces terms of $O(\epsilon^2)$. Since we are interested in 
-thin disks ($\epsilon\ll 1 $), e.g. protoplanetary disks where 
-$\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
-unnecessary complexity for the present discussion.

### 2015-04-08 paper.tex
minor tweaks to the early part of intro.  the intro is quite good already.  my current concern is that we should not ONLY give credit to BL15 for the physical interpretation of VSI/GSFI, as this potentially gives the false impression that it was not understood before.  I'll tweak that a bit next.

@@
 turbulence,  remain of interest \citep{armitage10}. In addition to
-mass accretion, hydrodynamic turbulence in PPDs will also
+mass accretion, turbulence in PPDs will also
 impact dust evolution and planetesimal formation \citep{chiang10}.
@@
 in three-dimensional (3D) disks in which the rotation frequency
-$\Omega$ depends on the height $z$ from the disk midplane. This
-dependence occurs when the equilibrium pressure depends on both the
-temperature and density, making the disk baroclinic. It is therefore
-generally expected that $\p_z\Omega\neq 0$ in astrophysical disks.
+$\Omega$ depends on the height $z$ from the disk midplane. 
+Vertical shear occurs whenever the disk is baroclinic, i.e.\ when constant 
+density and constant pressure surfaces are misaligned.  Baroclinicity, and thus vertical shear, is 
+practically unavoidable in disks, except at special locations like the midplane.
 
-The VSI can be regarded as the disk analog of the
+The VSI  in disks is fundamentally equivalent to the
 Goldreich-Schubert-Fricke instability 
-\citep[GSF,][]{goldreich67,fricke68} in differentially rotating stars.   
+\citep[GSFI,][]{goldreich67,fricke68} in differentially rotating stars.   
 A height-dependence in $\Omega$, or vertical shear, can render

### 2015-04-09 appendix.tex
restructed, merged with youdin's abstract

@@
 \subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
-A limitation of the VGSB is that it cannot be used in the
-adiabatic limit, even if the above conditions are satisfied. We explain this by considering vertically isothermal
+A limitation of the VGSB is that it cannot be used for 
+adiabatic flow, even if the above conditions are satisfied. We explain this by considering vertically isothermal
 disks and setting  $\beta\to\infty$ in 
@@
 %also need real freq^2 in adia limit 
-This is because the VGSB retains the vertical shear
-($\p_z\Omega$) from the global disk, but ignores the global radial
-disk structure responsible for it. Nevertheless, we demonstrate below
+This is because the VGSB retains the global radial
+disk structure responsible for vertical shear
+($\p_z\Omega$), but ignores it elsewhere in the linear problem. 
+Nevertheless, we demonstrate below
 that this inconsistency is unimportant for the VSI because it occurs
@@
 \end{align}
-(i.e. $\delta P = c_s^2\delta \rho$), and so $  \ii\hat{\sigma}c_s\delta v_z = W^\prime $
-from Eq. \ref{lin_vz}. For isothermal perturbations it is 
+(i.e. $\delta P = c_s^2\delta \rho$)% , and so $  \ii\hat{\sigma}c_s\delta v_z = W^\prime $
+% from Eq. \ref{lin_vz}
+. For isothermal perturbations it is 
 simpler to work with an equation for $W$ by substituting $Q=W$ into
-Eq. \ref{ode_w}. In this case, we shall not yet make the low frequency
+Eq. \ref{ode_w} and using Eq. \ref{ode_vz} to eliminate $\delta v_z$. 
+In this case, we shall not yet make the low frequency
 approximation, but first make the  Keplerian approximation. We

### 2015-04-09 paper.tex
restructed, merged with youdin's abstract

--- 2015-04-08 paper.tex
@@
 
-\title{Cooling Requirements for the Vertical Shear Instability, \\
- with Implications for Protoplanetary Disks}
+\title{Cooling Requirements for the Vertical Shear Instability in
+  astrophysical disks \\ and Implications for Protoplanetary Disks}
 %\title{Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula} 
@@
   A hydrodynamic mechanism, the vertical shear instability (VSI), offers a means to drive angular momentum 
-  transport and accretion flows in cold protoplanetary disks.  
+  transport in cold accretion disks such as protoplanetary disks.  
   The VSI is driven by a weak vertical gradient in the disk's orbital motion.

### 2015-04-16 paper.tex
new title
editing intro.  still in progress so don't criticize yet.  sorry taking so long...

@@
 
-\title{Cooling Requirements for the Vertical Shear Instability in
-  astrophysical disks \\ and Implications for Protoplanetary Disks}
+\title{Cooling Requirements for the Vertical Shear Instability 
+  in Protoplanetary Disks}
 %\title{Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula} 
@@
 mass accretion, turbulence in PPDs will also
-impact dust evolution and planetesimal formation \citep{chiang10}.
+impact dust evolution and planetesimal formation \citep{yl07, chiang10}.
 
 One possible route to hydrodynamic turbulence is the vertical shear
-instability (VSI). The VSI is an 
-axisymmetric linear instability 
+instability (VSI; citealp{urpin98, nelson13}).  The VSI  in disks is the same
+basic mechanism as the Goldreich-Schubert-Fricke instability 
+\citep[GSFI,][]{goldreich67,fricke68} in differentially rotating stars.   
+The VSI arises when vertical shear, i.e. a variation in orbital motion with 
+height, destabilizes inertial-gravity waves, oscillations 
+with rotation and buoyancy as restoring forces.  Vertical shear
+occurs wherever the disk is baroclinic, i.e.\ when constant 
+density and constant pressure surfaces are misaligned.  Baroclinicity, and thus vertical shear, is 
+practically unavoidable in disks, except at special locations like the midplane.
+
+To overcome the stabilizing effects of rotation, the VSI triggers motions which are vertically elongated 
+and radially narrow.  This geometry is only minimally affected by the stabilizing radial gradient of angular momentum.
+The vertical elongation taps vertical shear, but is also subject to the stabilizing effects of vertical buoyancy.
+Vertical buoyancy could be weak if there is strong internal heating, for instance close to the star (refs).  However, 
+realistic disk models with external irradiation tend to have significant vertical buoyancy, at last outside $\sim 1 AU$ (refs).
+
+To overcome vertical buoyancy, the VSI requires a short cooling time \citep{goldreich67, nelson13}.  Rapid cooling via radiation
+brings a moving fluid into thermal equilibrium with its surroundings diminishing buoyant stabilization.  An isothermal equation
+of state implies instantaneous cooling, and is the ideal
+
+The VSI is an axisymmetric linear instability 
 in three-dimensional (3D) disks in which the rotation frequency
@@
 
-The VSI  in disks is fundamentally equivalent to the
-Goldreich-Schubert-Fricke instability 
-\citep[GSFI,][]{goldreich67,fricke68} in differentially rotating stars.   
 A height-dependence in $\Omega$, or vertical shear, can render

### 2015-04-17 paper.tex
Reorganized the introduction to be a bit less technical and a bit better (I hope) organized from the broad to the specific.  Added some more classic references to go with the broader context.  Let me know what you think...

--- 2015-04-09 paper.tex
@@
   enough to quench the VSI at all disk radii.  Thus the formation of solid protoplanets, a sink for dust grains, can impede the VSI.
-
-%
-
-%  A robust hydrodynamic mechanism for turbulent angular momentum
-%  transport is important for cold accretion disks such as 
-%  protoplanetary disks. The vertical shear instability (VSI) has 
-%  recently been suggested to provide such a pathway, since it operates in
-%  disks with rotation profiles that depend on the height from the 
-%  disk midplane, which is generally expected in  
-%  astrophysical disks. However, vertical shear is weak in thin disks, 
-%  so the VSI can easily be stabilized by buoyancy forces in the
-%  presence of a stable vertical entropy  gradient. % , which is also
-%  % expected in astrophysical disks. 
-%  Consequently,  the VSI
-%  requires a rather short thermal or cooling timescale in
-%  sub-adabatically stratified disks.      
-%  A quantitative examination of this thermodynamic 
-%  requirement is needed to assess the viability of
-%  the VSI in realistic disks. To this end,  
-%  we study the linear, axisymmetric VSI in radially local, 
-%  vertically global models of stably stratified accretion disks in which the disk  
-%  temperature $T$ is relaxed to its initial value over a characteristic timescale 
-%  $t_c=\beta\Omega_k^{-1}$, where $\beta$ is a dimensionless parameter 
-%  and $\Omega_k$ is the Keplerian rotation frequency. % We explore how
-%  % the VSI depends on the thermodynamic response of the disk by
-%  % varying $\beta$.
-%  We find that 
-%  increasing $\beta$ even slightly away from    
-%  zero dramatically reduce VSI growth rates. 
-%  For vertically isothermal disks, we quantify the
-%  `short' thermal timescale requirement for the VSI as $\beta < \epsilon|q|/(\gamma-1)$, where
-%  $\epsilon$ is the disk aspect-ratio, $q = \p\ln T/\p\ln r$  is the radial temperature gradient, and 
-%  $\gamma$ is the adiabatic index. Thus, in a thin disk the thermal timescale must be shorter than 
-%  the dyamical timescale by $O(\epsilon)\ll 1$ for the VSI to operate.  
-%  We apply our results  
-%  to a disk model of the Minimum Mass Solar Nebula (MMSN) with an  
-%  estimate for the thermal timescale $\beta$ in the MMSN based on dust
-%  opacity. In the inner parts of the MMSN, e.g. at $\lesssim 5$AU, 
-%  we find the VSI would only operate on extremely short radial lengthscales, $\ll 
-%  O(10^{-2}H)$, where $H$ is the disk scale height. On longer radial 
-%  lengthscales of $O(10^{-2}H)$ to $O(10^{-1}H)$, we find the MMSN is
-%  unstable to the VSI from $\sim 5$AU to $\sim 50$AU with a
-%  characteristic growth time of $\sim 30$ local orbital periods.  
-%  % ,and the dominant VSI mode in the MMSN is the `fundamental
-%  % corrugation mode' which causes the disk column to oscillate
-%  % vertically.
-%  We thus suggest the VSI is at least dynamically important in the 
-%  outer parts of protoplanetary disks at tens of AU. 
 \end{abstract}
@@
 \section{Introduction}\label{intro}
-Understanding the mechanisms of angular momentum transport is the
-central problem in the theory of accretion disks 
-\citep{lyndenbell74}. The   
-magneto-rotational instability in the context of differentially
-rotating disks \citep[MRI,][]{balbus91} is now accepted as a robust
-pathway to magneto-hydrodynamic (MHD) turbulence, which transports
-angular momentum outwards and thus enables mass accretion.
+Understanding how disks transport mass and angular momentum underlies 
+many problems in astrophysics, including star and planet formation  
+\citep[][]{lyndenbell74, armitage10}.  The turbulence associated with many transport mechanisms
+is particularly important for dust evolution and planetesimal formation \citep{yl07, chiang10}.
 
-However, MHD turbulence may not be applicable to all  
-accretion disks. For example, there may be regions in protoplanetary  
-disks (PPDs) which do not support the MRI 
-\citep{gammie96,landry13}. In fact, recent MHD simulations
-suggest significant portions of PPDs remain laminar \citep[e.g.][and references
-therein]{lesur14,bai15,gressel15}. Thus, alternative mechanisms 
-to angular momentum transport in PPDs, such as purely hydrodynamic
-turbulence,  remain of interest \citep{armitage10}. In addition to
-mass accretion, turbulence in PPDs will also
-impact dust evolution and planetesimal formation \citep{chiang10}.
+Magneto-hydrodynamic (MHD) turbulence driven by
+the magneto-rotational instability \citep[MRI,][]{balbus91} has long been the most
+promising transport mechanism in low mass disks with weak self-gravity. 
+However, many parts of protoplanetary  
+disks (PPDs) are cold, have low levels of ionization, and do not support the MRI 
+\citep{blaes94,salmeron03}. Recent simulations
+suggest that significant portions of PPDs fail to develop MHD turbulence \citep[e.g.][]{simon13, lesur14,bai15,gressel15}.
 
+A purely hydrodynamic mechanism could circumvent difficulties with non-ideal MHD, but must overcome
+the strong centrifugal stability imposed the positive radial angular momentum gradient in nearly Keplerian disks \citep{balbus96}.
 One possible route to hydrodynamic turbulence is the vertical shear
-instability (VSI). The VSI is an 
-axisymmetric linear instability 
-in three-dimensional (3D) disks in which the rotation frequency
-$\Omega$ depends on the height $z$ from the disk midplane. 
-Vertical shear occurs whenever the disk is baroclinic, i.e.\ when constant 
+instability (VSI, \citealp{urpin98, nelson13}, hereafter N13).  The basic mechanism of the VSI  in disks was first
+identified in the context of differentially rotating stars \citep[][]{goldreich67,fricke68}.   
+The VSI arises when vertical shear, i.e.\ a variation in orbital motion along the rotation axis, 
+destabilizes inertial-gravity waves, which are oscillations 
+with rotation and buoyancy as restoring forces.  Vertical shear
+occurs wherever the disk is baroclinic, i.e.\ when constant 
 density and constant pressure surfaces are misaligned.  Baroclinicity, and thus vertical shear, is 
@@
 
-The VSI  in disks is fundamentally equivalent to the
-Goldreich-Schubert-Fricke instability 
-\citep[GSFI,][]{goldreich67,fricke68} in differentially rotating stars.   
-A height-dependence in $\Omega$, or vertical shear, can render
-inertial waves unstable. \cite{barker15} gives a simple 
-interpretation of the VSI: in an axisymmetric, inviscid, neutrally
-stratified disk, the exchange of fluid parcels vertically, while
-conserving their specific angular momentum, can result in a new
-configuration of lower energy if the distribution of specific angular
-momentum varies with $z$. The VSI is characterized by perturbations
-with large vertical length-scales  relative to radial length-scales.
+To overcome centrifugal stabilization, the VSI triggers motions which are vertically elongated 
+and radially narrow.  %This geometry minimizes the effects of the radial gradient of angular momentum.
+Vertical elongation taps the free energy of the vertical shear \citep{umurhan13}, but is also subject to 
+the stabilizing effects of vertical buoyancy. In principle, vertical buoyancy can be eliminated by strong internal heating, i.e.\ by the onset of convection.  
+Real PPDs should be vertically stably stratified in the outer regions, beyond $\sim 1$--5 AU, where heating is 
+irradiation dominated \citep{bitsch15}.  In the inner disk, strong vertical buoyancy is possible if accretion heating is weak
+or is concentrated in surface layers \citep{gammie96,lesniak11}.
 
-Early analysis of the VSI in radially and vertically local 
-models of accretion disks \citep{urpin98,urpin03} showed that the VSI
-requires a short thermal timescale so that perturbations evolve 
-isothermally, unless the disk is neutrally stratified in the vertical
-direction. This is because
-vertical motions associated with the VSI 
-are stabilized by buoyancy forces if the disk is sub-adiabatically
-stratified (convectively stable). A sufficiently short thermal
-timescale renders buoyancy ineffective, but a quantitative criteria is
-desirable. This is our primary goal.  
-%efficient exchang of heat with surroundings
+To overcome vertical buoyancy, the VSI requires a short cooling time (\citealp{goldreich67}; N13).  Rapid radiative cooling 
+brings a moving fluid into thermal equilibrium with its surroundings, thereby diminishing buoyancy.  An isothermal equation
+of state implies instantaneous cooling, and is the ideal context for studying the VSI (\citealp{urpin03}, N13, \citealp{barker15}).
 
-The need for a short thermal timescale for the VSI to operate in
-stably stratified disks has been demonstrated in recent numerical simulations 
-\citep{nelson13,stoll14}. When the  thermal timescale $t_c$ is
-parametrized in units of the orbital period  $P_\mathrm{orb}$,
-\cite{nelson13} find that $t_c\lesssim  0.01P_\mathrm{orb}$ was needed
-to observe the VSI. That is, the  disk temperature needs to remain
-effectively unchanged. Radiation hydrodynamic simulations performed by
-\cite{stoll14} also showed that the VSI is only sustained with the
-inclusion of heating from stellar irradiation, which results in an
-effective thermal timescale $t_c\ll P_\mathrm{orb}$ that maintains a
-fixed temperature profile.
+A full understanding of the VSI in real disks requires considering finite cooling times, $t_{\rm c}$.
+Simulations with a fixed, finite  $t_{\rm c}$ find that  VSI turbulence in stably stratified disks requires rapid cooling 
+with $t_{\rm c} \lesssim 0.01/\Omega$, in terms of the orbital frequency $\Omega$ (N13).
+N13 further find that the VSI can drive moderately strong transport, with Reynolds stresses of $\alpha \sim 10^{-3}$ times the mean pressure.
+Simulations with realistic radiative transfer, in lieu of a constant $t_{\rm c}$ approximation,  find that the VSI in irradiated disks drive transport with $\alpha \sim 10^{-4}$
+ in a  $\sim 2$ -- 10 AU PPD model \citep{stoll14}.
 
-These studies show that the VSI is generally sensitive to disk
-thermodynamics. It is thus important to quantify this dependence, for
-example on disk parameters, in order to evaluate the relevance of the
-VSI to realistic disks, which one may expect to be stably
-stratified and have non-zero thermal timescales.  In order to determine
-where in the disk the VSI may operate, and hence lead to hydrodynamic
-turbulence, understanding the thermodynamic dependence of the linear VSI is essential. 
-% since the transition
-% Should a disk
-% become turbulent through the VSI, it must first pass through the linear
-% instability.
+This paper presents the first comprehensive study of how finite cooling times affect the linear development of the VSI.
+We consider models with constant $t_{\rm c}$  and  also with more realistic cooling from a dust opacity in both the optically thick (diffusive) and thin regimes, where appropriate.
+Coupled with the results from numerical simulations, our results can be used to understand how the VSI operates over a wide range of model PPDs.
+This paper is organized as follows.
+In \S\ref{vsi_require} we introduce and qualitatively explain our main result, the critical cooling time for the VSI.
 
-In this work we study the linear VSI in radially local, vertically
-global models of astrophysical disks. This approach has
-been adopted in recent studies of the linear VSI
-\citep{nelson13,mcnally14,barker15}. However, these studies consider 
-isothermal perturbations, for which the thermal timescale is formally
-zero, and/or neutrally stratified disks in which buoyancy forces are
-absent by construction. We account for both effects to study the
-VSI in more general disk models. To do so, we examine 
-sub-adiabatically stratified disks and include an energy equation 
-with a source term that allows the thermodynamic response of 
-perturbations to be varied smoothly between isothermal and 
-adiabatic.
-
-% Our models are also an extension of earlier 
-% analysis by \cite{urpin98} and \cite{urpin03}, which accounts for thermal
-% evolution, but considered vertically localized disturbances.
-% Our 
-% vertically global approach is necessary to study disturbances with
-% vertical wavelengths comparable to or larger than the disk scale
-% height. Such disturbances were found by \cite{nelson13} and
-% \cite{stoll14} to dominate their numerical simulations. We thus focus
-% on vertically global perturbations which, in fact, we find to be the
-% dominant mode in protoplanetary disks with a realistic estimate for
-% the thermodynamic timescales.
-
-This paper is organized as follows. We first highlight in
-\S\ref{vsi_require} the need of a very short thermal timescale for the
-VSI to operate in stably stratified disks. There we also preview our
-main result --- a quantitative thermodynamic criteria for the VSI ---
-and a first application to protoplanetary disks. The remaining sections
-are devoted to deriving these results from linear theory.
+\emph{TODO: the following is still a bit too wordy IMO but we can redo as a final step after these sections are finalized.}
 
@@
 for future work.
+
+%The VSI is an axisymmetric linear instability 
+%in three-dimensional (3D) disks in which the rotation frequency
+%$\Omega$ depends on the height $z$ from the disk midplane. 
+%Vertical shear occurs whenever the disk is baroclinic, i.e.\ when constant 
+%density and constant pressure surfaces are misaligned.  Baroclinicity, and thus vertical shear, is 
+%practically unavoidable in disks, except at special locations like the midplane.
+%
+%A height-dependence in $\Omega$, or vertical shear, can render
+%inertial waves unstable. \cite{barker15} gives a simple 
+%interpretation of the VSI: in an axisymmetric, inviscid, neutrally
+%stratified disk, the exchange of fluid parcels vertically, while
+%conserving their specific angular momentum, can result in a new
+%configuration of lower energy if the distribution of specific angular
+%momentum varies with $z$. The VSI is characterized by perturbations
+%with large vertical length-scales  relative to radial length-scales.   
+%
+%Early analysis of the VSI in radially and vertically local 
+%models of accretion disks \citep{urpin98,urpin03} showed that the VSI
+%requires a short thermal timescale so that perturbations evolve 
+%isothermally, unless the disk is neutrally stratified in the vertical
+%direction. This is because
+%vertical motions associated with the VSI 
+%are stabilized by buoyancy forces if the disk is sub-adiabatically
+%stratified (convectively stable). A sufficiently short thermal
+%timescale renders buoyancy ineffective, but a quantitative criteria is
+%desirable. This is our primary goal.  
+%
+%The need for a short thermal timescale for the VSI to operate in
+%stably stratified disks has been demonstrated in recent numerical simulations 
+%\citep{nelson13,stoll14}. When the  thermal timescale $t_c$ is
+%parametrized in units of the orbital period  $P_\mathrm{orb}$,
+%\cite{nelson13} find that $t_c\lesssim  0.01P_\mathrm{orb}$ was needed
+%to observe the VSI. That is, the  disk temperature needs to remain
+%effectively unchanged. Radiation hydrodynamic simulations performed by
+%\cite{stoll14} also showed that the VSI is only sustained with the
+%inclusion of heating from stellar irradiation, which results in an
+%effective thermal timescale $t_c\ll P_\mathrm{orb}$ that maintains a
+%fixed temperature profile.     
+%
+%These studies show that the VSI is generally sensitive to disk
+%thermodynamics. It is thus important to quantify this dependence, for
+%example on disk parameters, in order to evaluate the relevance of the
+%VSI to realistic disks, which one may expect to be stably
+%stratified and have non-zero thermal timescales.  In order to determine
+%where in the disk the VSI may operate, and hence lead to hydrodynamic
+%turbulence, understanding the thermodynamic dependence of the linear VSI is essential. 
+%
+%In this work we study the linear VSI in radially local, vertically
+%global models of astrophysical disks. This approach has
+%been adopted in recent studies of the linear VSI
+%\citep{nelson13,mcnally14,barker15}. However, these studies consider 
+%isothermal perturbations, for which the thermal timescale is formally
+%zero, and/or neutrally stratified disks in which buoyancy forces are
+%absent by construction. We account for both effects to study the
+%VSI in more general disk models. To do so, we examine 
+%sub-adiabatically stratified disks and include an energy equation 
+%with a source term that allows the thermodynamic response of 
+%perturbations to be varied smoothly between isothermal and 
+%adiabatic.   
+%
+%
+%This paper is organized as follows. We first highlight in
+%\S\ref{vsi_require} the need of a very short thermal timescale for the
+%VSI to operate in stably stratified disks. There we also preview our
+%main result --- a quantitative thermodynamic criteria for the VSI ---
+%and a first application to protoplanetary disks. The remaining sections
+%are devoted to deriving these results from linear theory. 
+%
+%In \S\ref{setup} we list the governing equations and global
+%equilibrium disk models under   
+%consideration. We formulate the linear stability problem in 
+%\S\ref{linear} and simplify it by making the radially local
+%approximation. We study the linear problem using two approaches. In
+%\S\ref{analytical} we present an analytical discussion of vertically
+%isothermal disks. Our treatment is grossly simplified owing to a
+%number of approximations made, but closely parallels previous work, 
+%allowing us to highlight the essential effects of vertical shear, and to
+%derive a characteristic upper limit to the thermal timescale for the
+%VSI. This timescale is \emph{much shorter} than the dynamical
+%timescale. In \S\ref{numerical} we numerically
+%solve the linear problem in the radially local approximation, but
+%relax all other approximations made in our analytical approach.    
+%We find basic agreement with our analysis, including
+%the aforementioned characteristic thermal timescale. 
+%
+%In
+%\S\ref{application} we apply our results to the Minimum Mass
+%Solar Nebula (MMSN) by modeling thermal 
+%timescales based on realistic disk properties, and find that the VSI
+%can operate effectively in the outer parts of the MMSN. We summarize in
+%\S\ref{summary} with a discussion of caveats and possible directions
+%for future work.
 

### 2015-04-19 appendix.tex
minor edits to youdin's new intro, updated notations and reference styles

@@
 %shearing box \citep[VGSB,][]{mcnally14} to make analytic progress. 
-The vertically global shearing box \citep[VGSB,][]{mcnally14} is an
+The vertically global shearing box \citepalias[VGSB,][]{mcnally14} is an
 extension of the standard shearing box \citep[SSB,][]{goldreich65} to
@@
 $\hat{k}^2\gg 1$, Eq. \ref{iso_ode3} becomes equivalent to Eq. 39 in
-\cite{nelson13} or Eq. 28 in \cite{barker15}, although we have taken a
+\citetalias{nelson13} or Eq. 28 in \citetalias{barker15}, although we have taken a
 different route.
@@
 conditions. This enables the `surface modes' seen in numerical
-calculations \citep{barker15}.
+calculations \citepalias{barker15}.
 
@@
 which is just the dispersion relation for axisymmetric isothermal waves in a
-vertically isothermal disk
-\citep[e.g.][]{okazaki87,takeuchi98,tanaka02,zhang06,ogilvie13,barker14,barker15}. In 
+vertically isothermal disk (e.g. \citealt{okazaki87}; \citealt{takeuchi98}; \citealt{tanaka02}; 
+\citealt{zhang06}; \citealt{ogilvie13}; \citealt{barker14}; \citetalias{barker15}). In 
 this case the solutions are Hermite polynomials, $W\propto
@@
 \end{align}
-which is equivalent to Eq. 34 of \cite{barker15} in the limit
+which is equivalent to Eq. 34 of \citetalias{barker15} in the limit
 $\khat^2\gg L$. This signifies instability for

### 2015-04-19 application.tex
minor edits to youdin's new intro, updated notations and reference styles

@@
 \end{align}
-\citep[see also][]{barker15} where $\eta=k_\mathrm{rad}/\rho c_v$ is
+\citepalias[see also][]{barker15} where $\eta=k_\mathrm{rad}/\rho c_v$ is
 the diffusion coefficient and 
@@
 decay in a real disk. The viscous timescale for a perturbation lengthscale 
-$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha_s\khat^2\Omega $, where
-$\alpha_s$ is the standard alpha viscosity parameter
+$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $, where
+$\alpha$ is the standard alpha viscosity parameter
 \citep{shakura73}. Setting $t_\mathrm{visc}$ to be longer than the
 minimum VSI growth timescale $1/\epsilon\Omega$ implies 
-$\khat^2\lesssim \epsilon/\alpha_s$. Inserting $\epsilon \sim 0.05$,
+$\khat^2\lesssim \epsilon/\alpha$. Inserting $\epsilon \sim 0.05$,
 we find only modes with $\khat\lesssim 20$ are relevant for
-$\alpha_s\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha_s\sim
-10^{-5}$. \citep[These low viscosity levels are required for the
+$\alpha\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha\sim
+10^{-5}$. \citepalias[These low viscosity levels are required for the
 VSI, see][]{nelson13}.
@@
 is efficient at radii of few tens of AU in the MMSN, consistent with
-estimates made by \cite{nelson13}.
+estimates made by \citetalias{nelson13}.
 

### 2015-04-19 linear_adia.tex
minor edits to youdin's new intro, updated notations and reference styles

@@
 Isothermal perturbations correspond to $\beta = 0$ or $\gamma=1$ so 
-that $\chi = 1$. This case has been investigated by \cite{nelson13},
-\cite{mcnally14} and \cite{barker15}.
+that $\chi = 1$. This case has been investigated by \citetalias{nelson13},
+\citetalias{mcnally14} and \citetalias{barker15}.
 
@@
 sections, and are provided for ease of reference. We refer the reader
-to \cite{barker15} for a more detailed treatment of isothermal 
+to \citetalias{barker15} for a more detailed treatment of isothermal 
 perturbations. %which also account for more general boundary
@@
 The fundamental mode has been observed to dominate numerical simulations
-\citep{nelson13,stoll14}, and are the ones we find to typically 
+\citep[\citetalias{nelson13}]{stoll14}, and are the ones we find to typically 
 dominate in numerical calculations with increasing $\beta$
@@
   \delta v_z(\zhat) =
-  g(\zhat)\exp\left(\frac{\alpha\zhat^2}{2}\right), \label{adia_ansatz}
+  g(\zhat)\exp\left(\frac{\lambda\zhat^2}{2}\right), \label{adia_ansatz}
 \end{align}
-where $\alpha$ is a constant to be chosen for convenience. Inserting
+where $\lambda$ is a constant to be chosen for convenience. Inserting
 Eq. \ref{adia_ansatz} into Eq. \ref{nearly_iso_explicit} gives
 \begin{align}
-  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime + \left(B +
-    \alpha\right)g
-  +\left(\alpha^2 - \alpha A - C\right)\zhat^2 g.
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\lambda\right)g^\prime + \left(B +
+    \lambda\right)g
+  +\left(\lambda^2 - \lambda A - C\right)\zhat^2 g.
 \end{align}
-We choose $\alpha$ to make the coefficient of $\zhat^2g$
+We choose $\lambda$ to make the coefficient of $\zhat^2g$
 vanish, and impose the vertical kinetic energy density
-$\rho|\delta v_z|^2\propto |g|^2 \exp{\left(\real\alpha -
+$\rho|\delta v_z|^2\propto |g|^2 \exp{\left(\real\lambda -
     1/2\right)\zhat^2}$ to remain finite as $|\zhat|\to\infty$. 
@@
 \begin{align}
-  \real\alpha < \frac{1}{2}, 
+  \real\lambda < \frac{1}{2}, 
 \end{align}
@@
 \begin{align}
-  \alpha = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right).  
+  \lambda = \frac{1}{2}\left(A - \sqrt{A^2 + 4C}\right).  
 \end{align}
@@
 \begin{align}
-  0 = g^{\prime\prime} - \hat{z}\left(A - 2\alpha\right)g^\prime +
-  \left(B + \alpha\right)g.
+  0 = g^{\prime\prime} - \hat{z}\left(A - 2\lambda\right)g^\prime +
+  \left(B + \lambda\right)g.
 \end{align}
@@
 \begin{align}
-  B(\hat{\sigma}) + \alpha(\hat{\sigma}) =
-  M\left[A-2\alpha(\hat{\sigma})\right].\label{adia_disp0} 
+  B(\hat{\sigma}) + \lambda(\hat{\sigma}) =
+  M\left[A-2\lambda(\hat{\sigma})\right].\label{adia_disp0} 
 \end{align}

### 2015-04-19 linear_iso.tex
minor edits to youdin's new intro, updated notations and reference styles

@@
 to be derived and compared with previous works. Similar approximations
-are employed by \cite{nelson13} and \cite{barker15}. However, all of
+are employed by \citetalias{nelson13} and \citetalias{barker15}. However, all of
 the approximations described below are relaxed in our numerical study.
@@
 \item  We adopt the  `vertically global shearing box' (VGSB) formalism 
-  developed by \cite{mcnally14}, in which explicit dependencies on the
+  developed by \citetalias{mcnally14}, in which explicit dependencies on the
   background radial disk structure in 
@@
   approximation. We are therefore considering inertial waves
-  \citep{barker15}.
+  \citepalias{barker15}.
 

### 2015-04-19 numerical.tex
minor edits to youdin's new intro, updated notations and reference styles

@@
 This setup mimics strictly isothermal perturbations in a vertically
-isothermal disk, already considered by \cite{nelson13}, \cite{mcnally14} and
-\cite{barker15}. 
+isothermal disk, already considered previously \citetalias{nelson13,mcnally14,barker15}. 
 For completeness, and as a way of introducing terminology, 
@@
 and are inertial waves destabilized by vertical shear
-\citep{barker15}. (The exception being the fastest growing
+\citepalias{barker15}. (The exception being the fastest growing
 mode for $\zmax=7H$ which is a surface mode --- see the discussion
@@
 Fig. \ref{compare_modes_iso_kx10} lack the distinct `surface modes' found by 
-\cite{nelson13}, \cite{mcnally14}  and \cite{barker15}. 
+\citetalias{nelson13}, \citetalias{mcnally14}  and \citetalias{barker15}. 
 This a class of unstable modes associated with vertical shear near the
@@
 
-As explained by \cite{barker15}, while surface modes have a physical
+As explained by \citetalias{barker15}, while surface modes have a physical
 origin, their occurence in the vertically isothermal disk is due to
@@
 Our result is consistent with numerical simulations performed by
-\cite{nelson13} for which a thermal relaxation timescale $\beta\gtrsim
+\citetalias{nelson13} for which a thermal relaxation timescale $\beta\gtrsim
 0.6$ stabilized the VSI; as Fig. \ref{bcrit_compare1} show that growth
@@
 
-As pointed out by \cite{barker15}, in vertically isothermal disks with
+As pointed out by \citetalias{barker15}, in vertically isothermal disks with
 an imposed boundary, surface modes are dependent on $\zmax$, but they
@@
 %a disk surface at finite height. 
-\cite{barker15} have argued that vertically isothermal disks are not
+\citetalias{barker15} have argued that vertically isothermal disks are not
 ideal for studing the VSI because $\zmax$ is a free 
@@
 $0.1$. For $\beta=10^{-3}$ the most 
-unstable mode is an asymmetric surface mode \citep[also seen in Fig. 7
+unstable mode is an asymmetric surface mode \citepalias[also seen in Fig. 7
 of][]{barker15}. However, increasing the thermal timescale to

### 2015-04-19 paper.tex
minor edits to youdin's new intro, updated notations and reference styles

@@
 \defcitealias{lin12}{L12} 
+\defcitealias{nelson13}{N13}
+\defcitealias{barker15}{BL15}
+\defcitealias{mcnally14}{MP14}
 \def \OmK {\Omega_{\rm K}}
@@
   in Protoplanetary Disks}
-%\title{Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula} 
+%\title{Vertical shear instability in astrophysical disks: vertically
+%global linear calculations with thermal relaxation and applications
+%to the Solar Nebula}  
 \author{Min-Kai Lin \& Andrew Youdin}
@@
   It is difficult to understand how cold circumstellar disks accrete onto their central stars.  Due to 
-  lower levels of ionization, the magnetic transport mechanisms that operate in warmer disks are less effective. 
+  lower levels of ionization, the magnetic transport mechanisms that
+  operate in warmer disks are less effective.  
   A hydrodynamic mechanism, the vertical shear instability (VSI), offers a means to drive angular momentum 
@@
   The VSI is driven by a weak vertical gradient in the disk's orbital motion.
-  In order to grow, the VSI must overcome vertical buoyancy, a strong stabilizing influence in cold disks, where 
-  heating is dominated by external irradiation.  Rapid cooling, via radiative losses, reduces the effective buoyancy
-  and allows the VSI to operate.  In this paper, we quantify the cooling timescale, $t_c$, needed for growth of the VSI
+  In order to grow, the VSI must overcome vertical buoyancy, a strong
+  stabilizing influence in cold disks, where  
+  heating is dominated by external irradiation.  Rapid cooling, via
+  radiative losses, reduces the effective buoyancy 
+  and allows the VSI to operate.  In this paper, we quantify the
+  cooling timescale, $t_c$, needed for growth of the VSI 
   and consider the implications for protoplanetary disks where $t_c$ is determined by the opacity of dust 
-  grains.  We perform a linear analysis of the VSI with cooling in vertically global and radially local disk models.
-  For irradiated disks, we find that the VSI is most vigorous for rapid cooling with $t_c < \Omega_K^{-1} \epsilon |q| / (\gamma -1)$
-  in terms of the Keplerian orbital frequency, $\OmK$; the disk's aspect ratio, $\epsilon \ll 1$; the radial 
-  powerlaw temperature gradient, $q$; and the adiabatic index, $\gamma$.  For longer cooling times, the VSI 
+  grains.  We perform a linear analysis of the VSI with cooling in
+  vertically global and radially local disk models. 
+  For irradiated disks, we find that the VSI is most vigorous for
+  rapid cooling with $t_c < \OmK^{-1} \epsilon |q| / (\gamma -1)$ 
+  in terms of the Keplerian orbital frequency, $\OmK$; the disk's
+  aspect ratio, $\epsilon \ll 1$; the radial  
+  powerlaw temperature gradient, $q$; and the adiabatic index,
+  $\gamma$.  For longer cooling times, the VSI  
   is much less effective because growth slows and shifts to smaller length scales, which are more prone to 
@@
   find that the VSI is most effective at intermediate radii, from $\sim 5$AU to $\sim 50$AU with a
-  characteristic growth time of $\sim 30$ local orbital periods.  Growth is suppressed by long cooling times both in the opaque 
-  inner disk and the optically thin outer disk.  A reduction in the dust opacity by a factor of 10 increases cooling times 
-  enough to quench the VSI at all disk radii.  Thus the formation of solid protoplanets, a sink for dust grains, can impede the VSI.
+  characteristic growth time of $\sim 30$ local orbital periods.
+  Growth is suppressed by long cooling times both in the opaque  
+  inner disk and the optically thin outer disk.  A reduction in the
+  dust opacity by a factor of 10 increases cooling times  
+  enough to quench the VSI at all disk radii.  Thus the formation of
+  solid protoplanets, a sink for dust grains, can impede the VSI. 
 \end{abstract}
@@
 \citep{blaes94,salmeron03}. Recent simulations
-suggest that significant portions of PPDs fail to develop MHD turbulence \citep[e.g.][]{simon13, lesur14,bai15,gressel15}.
+suggest that significant portions of PPDs fail to develop MHD
+turbulence \citep[e.g.][]{simon13, lesur14,bai15,gressel15}.
 
 A purely hydrodynamic mechanism could circumvent difficulties with non-ideal MHD, but must overcome
-the strong centrifugal stability imposed the positive radial angular momentum gradient in nearly Keplerian disks \citep{balbus96}.
+the strong centrifugal stability imposed by the positive radial specific angular
+momentum gradient in nearly Keplerian disks \citep{balbus96}.
 One possible route to hydrodynamic turbulence is the vertical shear
-instability (VSI, \citealp{urpin98, nelson13}, hereafter N13).  The basic mechanism of the VSI  in disks was first
+instability (VSI, \citealp{urpin98, urpin03, nelson13}, hereafter \citetalias{nelson13}).  The
+basic mechanism of the VSI  in disks was first 
 identified in the context of differentially rotating stars \citep[][]{goldreich67,fricke68}.   
-The VSI arises when vertical shear, i.e.\ a variation in orbital motion along the rotation axis, 
+The VSI arises when vertical shear, i.e. \ a variation in orbital motion along the rotation axis, 
 destabilizes inertial-gravity waves, which are oscillations 
@@
 density and constant pressure surfaces are misaligned.  Baroclinicity, and thus vertical shear, is 
-practically unavoidable in disks, except at special locations like the midplane.
+practically unavoidable in astrophysical disks, except at special locations like the
+midplane.
 
@@
 Vertical elongation taps the free energy of the vertical shear \citep{umurhan13}, but is also subject to 
-the stabilizing effects of vertical buoyancy. In principle, vertical buoyancy can be eliminated by strong internal heating, i.e.\ by the onset of convection.  
-Real PPDs should be vertically stably stratified in the outer regions, beyond $\sim 1$--5 AU, where heating is 
-irradiation dominated \citep{bitsch15}.  In the inner disk, strong vertical buoyancy is possible if accretion heating is weak
-or is concentrated in surface layers \citep{gammie96,lesniak11}.
+the stabilizing effects of vertical buoyancy if the disk is stably
+stratified. To overcome vertical buoyancy, the VSI requires a short cooling time
+(\citealp{goldreich67}; \citetalias{nelson13}). Rapid radiative 
+cooling, equivalently a short thermal relaxation timescale, brings a
+moving fluid element into thermal equilibrium with 
+its surroundings, thereby diminishing buoyancy. %not obvious actually
+                                %- requires assumption of ideal gas
+                                %e.o.s. and pressure eqm established
+                                %quickly 
+An isothermal equation of state implies instantaneous cooling (and
+heating), and is the ideal context for studying the VSI
+(\citetalias{nelson13}; \citet{mcnally14}, hereafter \citetalias{mcnally14}; \citealp{barker15},
+hereafter \citetalias{barker15}).
 
-To overcome vertical buoyancy, the VSI requires a short cooling time (\citealp{goldreich67}; N13).  Rapid radiative cooling 
-brings a moving fluid into thermal equilibrium with its surroundings, thereby diminishing buoyancy.  An isothermal equation
-of state implies instantaneous cooling, and is the ideal context for studying the VSI (\citealp{urpin03}, N13, \citealp{barker15}).
+Alternatively, vertical buoyancy can be eliminated by strong internal
+heating, i.e.\ by the onset of convection, so that the disk becomes 
+neutrally stratified in the vertical direction
+\citepalias{nelson13,barker15}. However, realistic PPDs should be vertically
+stably stratified in the outer regions, beyond $\sim 1$--5 AU, where
+heating is  irradiation dominated \citep{bitsch15}.  In the inner
+disk, strong vertical buoyancy is possible if accretion heating is
+weak or is concentrated in surface layers \citep{gammie96,lesniak11}.
 
-A full understanding of the VSI in real disks requires considering finite cooling times, $t_{\rm c}$.
-Simulations with a fixed, finite  $t_{\rm c}$ find that  VSI turbulence in stably stratified disks requires rapid cooling 
-with $t_{\rm c} \lesssim 0.01/\Omega$, in terms of the orbital frequency $\Omega$ (N13).
-N13 further find that the VSI can drive moderately strong transport, with Reynolds stresses of $\alpha \sim 10^{-3}$ times the mean pressure.
-Simulations with realistic radiative transfer, in lieu of a constant $t_{\rm c}$ approximation,  find that the VSI in irradiated disks drive transport with $\alpha \sim 10^{-4}$
- in a  $\sim 2$ -- 10 AU PPD model \citep{stoll14}.
+A full understanding of the VSI in real disks therefore requires
+considering finite, non-zero cooling times, $t_{\rm c}$. Non-linear
+hydrodynamical simulations with a prescribed  $t_{\rm c}$ find that VSI
+turbulence in stably stratified disks requires rapid cooling  
+with $t_{\rm c} \lesssim 0.01P_{\mathrm{orb}}$, where $P_\mathrm{orb}$ 
+%=2\pi/\Omega$ 
+is the local orbital period %and $\Omega$ is the orbital
+%frequency 
+ \citepalias{nelson13}. \citetalias{nelson13} further find
+that the VSI can drive moderately strong transport, with Reynolds
+stresses of $\alpha \sim 10^{-3}$ times the mean pressure. Simulations
+with realistic radiative transfer, in lieu of a parameterized $t_{\rm
+  c}$, find that the VSI in irradiated disks drive transport with
+$\alpha \sim 10^{-4}$  in a  $\sim 2$ -- 10AU PPD model \citep{stoll14}.
 
-This paper presents the first comprehensive study of how finite cooling times affect the linear development of the VSI.
-We consider models with constant $t_{\rm c}$  and  also with more realistic cooling from a dust opacity in both the optically thick (diffusive) and thin regimes, where appropriate.
-Coupled with the results from numerical simulations, our results can be used to understand how the VSI operates over a wide range of model PPDs.
-This paper is organized as follows.
-In \S\ref{vsi_require} we introduce and qualitatively explain our main result, the critical cooling time for the VSI.
+To reach hydrodynamic turbulence via the VSI, a disk should  
+first develop the linear instability. A quantitative 
+examination of the thermodynamic requirement of the linear VSI is thus essential in assessing its 
+relevance to PPDs. In this work, we study the linear VSI in vertically global disk models 
+with non-zero cooling timescales, which permit buoyancy forces. 
+Our effort generalizes both early analyses of the VSI that account for buoyancy but in vertically localized disk models \citep{urpin98,urpin03}, 
+as well as recent calculations in vertically global disks but without buoyancy forces \citepalias{nelson13,mcnally14,barker15}. 
+%to study of how finite cooling times, and thus buoyancy, 
+%affect the VSI. Our effort may also be regarded
+%as a extension of early analyses of the VSI in 
+%vertically localized disk models with cooling \citep{urpin98,urpin03} to
+%vertically global disks with cooling. 
+We consider disk   
+models with vertically constant $t_{\rm c}$, as well as more realistic
+cooling timescales based on dust opacity in both the optically thick
+(diffusive) and thin regimes, where appropriate. Coupled with the
+results from numerical simulations, our results can be used to 
+understand how the VSI operates over a wide range of model PPDs.
 
-\emph{TODO: the following is still a bit too wordy IMO but we can redo as a final step after these sections are finalized.}
+%\emph{TODO: the following is still a bit too wordy IMO but we can redo as a final step after these sections are finalized.}
 
+This paper is organized as follows. In \S\ref{vsi_require} we 
+introduce and qualitatively explain our main result of a critical or
+characteristic cooling timescale, beyond which the VSI is either
+suppressed or inefficient. We then present the  
+formal linear problem in \S\ref{setup}---\ref{numerical}.    
 In \S\ref{setup} we list the governing equations and global
-equilibrium disk models under   
-consideration. We formulate the linear stability problem in 
-\S\ref{linear} and simplify it by making the radially local
-approximation. We study the linear problem using two approaches. In
-\S\ref{analytical} we present an analytical discussion of vertically
-isothermal disks. Our treatment is grossly simplified owing to a
-number of approximations made, but closely parallels previous work, 
-allowing us to highlight the essential effects of vertical shear, and to
-derive a characteristic upper limit to the thermal timescale for the
-VSI. This timescale is \emph{much shorter} than the dynamical
-timescale. In \S\ref{numerical} we numerically
-solve the linear problem in the radially local approximation, but
-relax all other approximations made in our analytical approach.    
-We find basic agreement with our analysis, including
-the aforementioned characteristic thermal timescale.
-
+equilibrium disk models under consideration, and formulate the linear
+stability problem in \S\ref{linear}. In 
+\S\ref{analytical} we give a simplified analytical discussion for
+vertically isothermal disks, where we derive the aforementioned
+characteristic upper limit on the cooling or thermal timescale for the VSI. 
+% Our treatment is grossly simplified owing to a
+% number of approximations made, but closely parallels previous work, 
+% allowing us to highlight the essential effects of vertical shear, and to
+% derive a characteristic upper limit to the thermal timescale for the
+% VSI. This timescale is \emph{much shorter} than the dynamical
+% timescale. 
+In \S\ref{numerical} we numerically
+solve the linear problem, and  
+find basic agreement with our analysis.
+% , including
+% the aforementioned characteristic thermal timescale. 
 In 
 \S\ref{application} we apply our results to the Minimum Mass
-Solar Nebula (MMSN) by modeling thermal 
-timescales based on realistic disk properties, and find that the VSI
-can operate effectively in the outer parts of the MMSN. We summarize in
-\S\ref{summary} with a discussion of caveats and possible directions
-for future work.
+Solar Nebula (MMSN) by modeling cooling timescales based on realistic
+disk properties, and find that the VSI can operate effectively in the
+outer parts of the MMSN. We summarize in \S\ref{summary} with a
+discussion of caveats and possible directions for future work.
 

### 2015-04-19 summary.tex
restructed, merged with youdin's abstract
minor edits to youdin's new intro, updated notations and reference styles

@@
 To do so, we generalized previous vertically global linear 
-calculations of the VSI \citep{nelson13,mcnally14,barker15}, which
+calculations of the VSI \citepalias{nelson13,mcnally14,barker15}, which
 assumed isothermal perturbations, to include an energy equation with 
@@
 $t_c$, and has been employed in previous numerical simulations of the
-VSI \citep{nelson13}.
+VSI \citepalias{nelson13}.
 
@@
 \begin{align*}
-  t_\mathrm{crit} = \frac{\epsilon |q|}{\gamma-1}\Omega^{-1}, 
+  t_\mathrm{crit} = \frac{\epsilon |q|}{\gamma-1}\Omega_k^{-1}, 
 \end{align*}
@@
 
-We have neglected viscosity. \cite{nelson13} find that an alpha
+We have neglected viscosity. \citetalias{nelson13} find that an alpha
 viscosity $\lesssim O(10^{-4})$ is required for the
-VSI. \cite{barker15} found that inclusion of viscosity is needed for 
+VSI. \citetalias{barker15} found that inclusion of viscosity is needed for 
 convergence when studying surface modes. %but not body modes? 

### 2015-04-19 paper.tex
minor edits to youdin's new intro, updated notations and reference styles

@@
 with non-zero cooling timescales, which permit buoyancy forces. 
-Our effort generalizes both early analyses of the VSI that account for buoyancy but in vertically localized disk models \citep{urpin98,urpin03}, 
-as well as recent calculations in vertically global disks but without buoyancy forces \citepalias{nelson13,mcnally14,barker15}. 
+Our effort generalizes both early analyses of the VSI that account for buoyancy but in vertically localized disk models \citep{urpin98,urpin03}, and
+recent calculations in vertically global disks but without buoyancy forces \citepalias{nelson13,mcnally14,barker15}. 
 %to study of how finite cooling times, and thus buoyancy, 

### 2015-04-19 application.tex
minor edits to intro

@@
 As an application of our linear models, we estimate where in a
-protoplanetary disk do the thermodynamic conditions allow the VSI to
+PPD do the thermodynamic conditions allow the VSI to
 operate. For a realistic disk one might consider an energy equation
@@
 \subsection{Evaluation for a fiducial disk model}
-We evaluate Eq. \ref{real_beta} for the protoplanetary disk model
+We evaluate Eq. \ref{real_beta} for the PPD model
 described in \cite{chiang10} and summarized in Appendix \ref{mmsn}
@@
     evaluated at the midplane (left) and at $z=3H$ (right) in the
-    fiducial protoplanetary disk. Eq. \ref{beta_mmsn} is plotted  
+    fiducial PPD. Eq. \ref{beta_mmsn} is plotted  
     for three values of the 

### 2015-04-19 linear_adia.tex
minor edits to intro

@@
 (\S\ref{therm_relax_eff}) for moderately large    
-values of $\khat$, as well as in protoplanetary disks 
+values of $\khat$, as well as in PPDs 
 with a realistic estimate for thermal timescales

### 2015-04-19 numerical.tex
minor edits to intro

@@
 This setup mimics strictly isothermal perturbations in a vertically
-isothermal disk, already considered previously \citetalias{nelson13,mcnally14,barker15}. 
+isothermal disk, already considered previously \citepalias{nelson13,mcnally14,barker15}. 
 For completeness, and as a way of introducing terminology, 

### 2015-04-19 paper.tex
minor edits to intro

@@
   A hydrodynamic mechanism, the vertical shear instability (VSI), offers a means to drive angular momentum 
-  transport in cold accretion disks such as protoplanetary disks.  
+  transport in cold accretion disks such as protoplanetary disks (PPDs).  
   The VSI is driven by a weak vertical gradient in the disk's orbital motion.
@@
   cooling timescale, $t_c$, needed for growth of the VSI 
-  and consider the implications for protoplanetary disks where $t_c$ is determined by the opacity of dust 
+  and consider the implications for PPDs where $t_c$ is determined by the opacity of dust 
   grains.  We perform a linear analysis of the VSI with cooling in
@@
   is much less effective because growth slows and shifts to smaller length scales, which are more prone to 
-  viscous or turbulent decay.  We apply our results to a  Minimum Mass Solar Nebula (MMSN) disk model.  We
+  viscous or turbulent decay.  We apply our results to a  Minimum Mass Solar Nebula disk model.  We
   find that the VSI is most effective at intermediate radii, from $\sim 5$AU to $\sim 50$AU with a
@@
 heating), and is the ideal context for studying the VSI
-(\citetalias{nelson13}; \citet{mcnally14}, hereafter \citetalias{mcnally14}; \citealp{barker15},
+(\citealp{urpin03}; \citetalias{nelson13}; \citealp{mcnally14}, hereafter \citetalias{mcnally14}; \citealp{barker15},
 hereafter \citetalias{barker15}).
@@
 first develop the linear instability. A quantitative 
-examination of the thermodynamic requirement of the linear VSI is thus essential in assessing its 
-relevance to PPDs. In this work, we study the linear VSI in vertically global disk models 
+examination of the thermodynamic requirement for the linear VSI is
+thus essential in assessing its 
+relevance to PPDs. In this work, we present a
+comprehensive study of the linear VSI in vertically global disk models 
 with non-zero cooling timescales, which permit buoyancy forces. 
-Our effort generalizes both early analyses of the VSI that account for buoyancy but in vertically localized disk models \citep{urpin98,urpin03}, and
-recent calculations in vertically global disks but without buoyancy forces \citepalias{nelson13,mcnally14,barker15}. 
+% In this work, we present a comprehensive study of the linear VSI in vertically global disk models 
+% with non-zero cooling timescales $t_c > 0$, which permit buoyancy forces. 
+% Our effort generalizes both early analyses of the VSI in vertically
+% localized disk models \citep{urpin98,urpin03} and recent calculations in vertically global 
+% disks \citepalias{nelson13,mcnally14,barker15}. These works either
+% consider the limit of instantaneous cooling ($t_c\to0$) or neutrally
+% stratified disks, for which buoyancy forces play no role. 
+Our effort generalizes previous studies of the linear VSI in 
+vertically local \citep{urpin98,urpin03} and vertically global 
+\citepalias{nelson13,mcnally14,barker15} disk models that consider the limit
+of instantaneous cooling, for which buoyancy plays no role.  
 %to study of how finite cooling times, and thus buoyancy, 
@@
 dimensionless radial temperature gradient $q$ and adiabatic index
-$\gamma$. For a protoplanetary disk, $\epsilon \sim 0.05$, $|q|\sim 1$ and $\gamma\simeq 1.4$.
+$\gamma$. For a PPD, $\epsilon \sim 0.05$, $|q|\sim 1$ and $\gamma\simeq 1.4$.
 
@@
 To illustrate this point, we show in Fig. \ref{eqm_structure} the
-equilibrium structure of a typical protoplanetary disk. %Here $\rho_0$
+equilibrium structure of a typical PPD. %Here $\rho_0$
 %is the midplane density and vertical shear rate and buoyancy
@@
   \includegraphics[width=\linewidth]{figures/rhoshearNz}
-  \caption{Typical protoplanetary disk structure (the Minimum Mass
+  \caption{Typical PPD structure (the Minimum Mass
     Solar Nebla): density (left), vertical shear rate (middle) and
@@
 operate. For example, we show in Fig. \ref{bcrit_mmsn2d} an estimate
-of the thermal timescales the protoplanetary disk model shown in
+of the thermal timescales the PPD model shown in
 Fig. \ref{eqm_structure}. Here $k_x$ is a radial
@@
   \caption{Estimate of the thermal timescale  in a fiducial 
-    protoplanetary disk model, the Minimum Mass Solar Nebula. The red
+    PPD model, the Minimum Mass Solar Nebula. The red
     line corresponds to a characteristic thermal timescale, derived in
@@
 reader may skip forward \S\ref{application} for the application of our
-results to protoplanetary disks.
+results to PPD.
 

### 2015-04-19 setup.tex
minor edits to youdin's new intro, updated notations and reference styles
minor edits to intro

@@
 disk from isothermal ($\beta\ll 1$) to adiabatic ($\beta \gg 1$) in a
-controlled manner. However, for application to protoplanetary disks in 
+controlled manner. However, for application to PPDs in 
 \S\ref{application} we will model $\beta$ based on physical thermal
@@
 vertically isothermal disks which is the relevant case for passively irradiated
-protoplanetary disks \citep{chiang97}.
+PPDs \citep{chiang97}.
 
@@
 so that we can ignore the radial variations in the background
-structure and radial gradients. \cite{barker15} have validated this
+structure and radial gradients. \citetalias{barker15} have validated this
 approximation by comparison with the linear VSI in global cylindrical

### 2015-04-19 summary.tex
minor edits to intro

@@
 Physically, a short thermal timescale is needed to reduce the  
-stabilizing effect of buoynancy forces in a stably stratified
+stabilizing effect of buoyancy forces in a stably stratified
 disk. Because vertical shear is weak in thin disks, $t_c\Omega\ll 1$
@@
 We mostly focused on vertically isothermal disks as 
-this is the relevant case for protoplanetary disks passively
+this is the relevant case for protoplanetary disks (PPDs) passively
 irradiated by its central star \citep{chiang97}. It also permits a
@@
 We applied our linear models to assess the applicability of the VSI in
-protoplanetary disks. We considered the Minimum Mass Solar Nebula 
+PPDs. We considered the Minimum Mass Solar Nebula 
 (MMSN) described in \cite{chiang10} and   
@@
 relaxation timescale $t_c$. For  
-application to protoplanetary disks we connected $t_c$ to the disk
+application to PPDs we connected $t_c$ to the disk
 properties such as surface density, temperature and opacity, making
@@
 disk, additional effects may become important. For example, in the
-outer parts of a protoplanetary disk, where we estimate the VSI 
+outer parts of a PPD, where we estimate the VSI 
 operates, disk self-gravity (SG) may play a role. Vertical

### 2015-04-23 paper.tex

@@
 axisymmetric adiabatic disturbances, then because the 
-Solberg-Hoiland criteria \citep{tassoul78} are usually satisified in a
+Solberg-Hoiland criteria \citep{tassoul78} are usually satisfied in a
 thin disk, the disk would be stable.

### 2015-04-23 setup.tex

@@
 We consider  an inviscid, non-self-gravitating three-dimensional (3D)
-disk orbiting a central star of mass $M_*$. We adopt cylindrical
-co-ordinates $(r,\phi, z)$ centered on the star to describe the global
-disk. The governing fluid equations are
+disk orbiting a central star of mass $M_*$.  Cylindrical
+co-ordinates $(r,\phi, z)$ are centered on the star. The governing fluid equations are
 \begin{align}
@@
 central star and $G$ is the gravitational constant. 
-In the energy equation (Eq. \ref{full_energy}) the source term
-$\Lambda$ represent non-adiabatic effects which we
-model below. The gas temperature $T$ is defined through the ideal gas 
+In the energy equation (Eq.\ \ref{full_energy}) the sink term
+$\Lambda$ includes non-adiabatic cooling and (negative) heating. 
+The gas temperature $T$ obeys the ideal gas 
 equation of state,  
@@
 \subsection{Thermal relaxation}
-We model non-adiabatic effects by defining $\Lambda$ as a thermal
-relaxation term,
+We model radiative effects as thermal
+relaxation with a timescale $t_c$, i.e.,
 \begin{align}\label{thermal_relax}
-  \Lambda  \equiv \frac{\rho\mathcal{R}}{\mu t_c}\left(T - T_0\right),
+  \Lambda  \equiv \frac{\rho\mathcal{R}}{\mu}\frac{T - T_0}{t_c},
 \end{align}
-where $T_0=T_0(r,z)$ is the initial temperature, and $t_c$ is a
-thermal relaxation timescale parametrized via $\beta$ such that 
+where $T_0=T_0(r,z)$ is the equilibrium temperature.  We 
+define the dimensionless cooling time 
 \begin{align}\label{beta_def}
-  t_c = \beta \Omega_k^{-1},
+  \beta \equiv \Omega_k t_c
 \end{align}
-where $\Omega_k=\sqrt{GM_*/r^3}$ is the Keplerian frequency.
+relative to the Keplerian frequency, $\Omega_k=\sqrt{GM_*/r^3}$.
 
@@
 disk from isothermal ($\beta\ll 1$) to adiabatic ($\beta \gg 1$) in a
-controlled manner. However, for application to PPDs in 
-\S\ref{application} we will model $\beta$ based on physical thermal
-timescales, in which case $\beta$ is non-constant and is determined by
-the equilibrium disk properties and the perturbation lengthscales of
-interest (e.g. Fig. \ref{bcrit_mmsn2d}).
+controlled manner.
+
+In \S\ref{application} we analyze the VSI using a realistic model for 
+$\beta$ in PPDs with a dust opacity, using Eq.\ (\ref{beta_mmsn_simp}).  
+ Fig. \ref{bcrit_mmsn2d} shows that in this model $\beta >> 1$ in the 
+ dense inner disk and near the midplane, but that rapid cooling ($\beta \ll 1$)
+ is possible elsewhere.  Furthermore short wavelength perturbations can cool 
+ quicker ...
+
+ \emph{TODO: $k_x$ is in Fig. \ref{bcrit_mmsn2d} but not explained yet...}
+% based on physical thermal
+%timescales, in which case $\beta$ is non-constant and is determined by
+%the equilibrium disk properties and the perturbation lengthscales of
+%interest (e.g. Fig. \ref{bcrit_mmsn2d}).
 

### 2015-04-23 appendix.tex
notation change from epsilon to h

@@
 
-% In \S\ref{approx_gov} we made the replacement $D\to\Omega_k^2$ before
+% In \S\ref{approx_gov} we made the replacement $D\to\Omega_K^2$ before
 % eliminating variables to obtain a single equation for $\delta v_z$ in
@@
 approximations is equivalent to the replacement $D\to
-\Omega_k^2$. This gives, in terms of dimensionless variables,  
+\Omega_K^2$. This gives, in terms of dimensionless variables,  
 \begin{align}
-  &\delta v_z ^{\prime\prime} + \left[1 + \ii\epsilon q\hat{k} -
+  &\delta v_z ^{\prime\prime} + \left[1 + \ii h q\hat{k} -
     \frac{ \hat{k}^2}{
-      \left(\hat{k}^2+\chi\right)}q \epsilon^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
-  \left\{\left(\chi + \ii \epsilon q
+      \left(\hat{k}^2+\chi\right)}q h^2F\right]\ln\rho^{\prime}\delta v_z^\prime +
+  \left\{\left(\chi + \ii h q
       \hat{k}\right)\ln\rho^{\prime\prime} - \ln\rho^{\prime
       2}\left(\khat^2 -
-      \ii\epsilon
+      \ii h
       q\hat{k}\right)\left[1 - \chi +
-      \frac{\chi}{\left(\hat{k}^2+\chi\right)}q\epsilon^2F\right]\right\}\delta v_z \notag\\&=
+      \frac{\chi}{\left(\hat{k}^2+\chi\right)}qh^2F\right]\right\}\delta v_z \notag\\&=
   -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
 \end{align} 
-Since we are interested in thin disks with $\epsilon \ll 1$ and $|q| =
-O(1)$, we neglect the terms proportional to $\epsilon^2$, which gives 
+Since we are interested in thin disks with $h \ll 1$ and $|q| =
+O(1)$, we neglect the terms proportional to $h^2$, which gives 
 Eq. \ref{vertiso_gov_nondim} (the dimensional version being
 Eq. \ref{vertiso_gov}). This is equivalent to neglecting the   
-vertical dependence of $D$ by setting $D\to\Omega_k^2$ in
+vertical dependence of $D$ by setting $D\to\Omega_K^2$ in
 Eq. \ref{ode_w} before eliminating variables in favor of $\delta
@@
 %Eq. \ref{adia_diso3} differs from Eq. \ref{vertiso_gov_nondim} by terms
-%proportional to $\epsilon^2$. For a thin disk ($\epsilon\ll1$) with
+%proportional to $h^2$. For a thin disk ($h\ll1$) with
 %$|q|=O(1)$, as considered in this work, neglecting these terms has no
@@
 % $D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
-% $D\to\Omega_k^2$ was made before eliminating variables in favor of
+% $D\to\Omega_K^2$ was made before eliminating variables in favor of
 % $\delta v_z$. In Appendix \ref{adia_improve} we show that making this
 % replacement after eliminating variables, which involves $dD/dz$,
-% introduces terms of $O(\epsilon^2)$. Since we are interested in 
-% thin disks ($\epsilon\ll 1 $), e.g. protoplanetary disks where 
-% $\epsilon\lesssim 0.1$ \citep{chiang10}, these terms introduce
+% introduces terms of $O(h^2)$. Since we are interested in 
+% thin disks ($h\ll 1 $), e.g. protoplanetary disks where 
+% $h\lesssim 0.1$ \citep{chiang10}, these terms introduce
 % unnecessary complexity for the present discussion.
@@
 $O(k_x)$. The neglected terms therefore have a relative magnitude of
-$O(\epsilon/\khat)$, which is small for thin disks ($\epsilon\ll1$)
+$O(h/\khat)$, which is small for thin disks ($h\ll1$)
 and/or small radial wavelengths ($\khat\gg 1$).
@@
 \begin{align}
-  0 =\dd v_z^{\prime\prime} + \left(1 + \ii \epsilon q
+  0 =\dd v_z^{\prime\prime} + \left(1 + \ii h q
     \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime
   +\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
-    -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii\epsilon  
+    -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii h  
           q}{\hat{k}}\right)\ln\rho^{\prime 2}\right]\right\}\delta v_z.\label{adia_iso3}
@@
 &+
-  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii\epsilon
+  \left(\frac{\gamma-1}{\gamma}\right)\hat{k}^2\left(1-\frac{\ii h
       q}{\hat{k}}\right) \int_{\zhat_1}^{\zhat_2}\rho\ln\rho^{\prime
     2}|\delta v_z|^2 d\zhat
-+ \ii\epsilon q \hat{k}
++ \ii h q \hat{k}
   \int_{\zhat_1}^{\zhat_2}\ln\rho^\prime(\rho\delta v_z^*)^\prime
@@
     The disk parameters are $(\gamma, \Gamma)=(1.4, 1.011)$ and
-    $(p,q,\epsilon)=(-1.5,-1,0.05)$. The perturbation
+    $(p,q,h)=(-1.5,-1,0.05)$. The perturbation
     wavenumber is $\khat=30$. 
@@
 \begin{align}
-  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \frac{\ii \epsilon q \hat{k} f(\zhat)
+  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \frac{\ii h q \hat{k} f(\zhat)
       }{1-\hat{\sigma}^2} \right]W^\prime +
@@
 \begin{align}\label{fz_shear}
-  \frac{d\Omega^2}{d\hat{z}} = \epsilon^2q f(\hat{z})\Omega_k^2.
+  \frac{d\Omega^2}{d\hat{z}} = h^2q f(\hat{z})\Omega_K^2.
 \end{align}
@@
 see that $f(\zhat)= 
-\zhat\left(1+\epsilon^2\zhat^2\right)^{-3/2}$. More generally, though,
+\zhat\left(1+h^2\zhat^2\right)^{-3/2}$. More generally, though,
 $f$ may be regarded as a representation of the vertical
@@
 \begin{align}
-  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii \epsilon q \hat{k}
+  0 = W^{\prime\prime} + \left[\ln\rho^\prime - \ii h q \hat{k}
     f(\zhat)\right]W^\prime +
@@
   =\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
-  +\ii \epsilon q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
+  +\ii h q \hat{k}\int_{\zhat_1}^{\zhat_2}\rho f(\zhat) W^*W^\prime d\zhat.\label{integral_relation1}
 \end{align}
@@
   =\real\left[\ii
-    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
+    h\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
     f(\hat{z}) W^*W^\prime d\zhat\right], \\
@@
   =\imag\left[\ii
-    \epsilon\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
+    h\hat{k}q \int_{\zhat_1}^{\zhat_2}\rho
     f(\hat{z}) W^*W^\prime d\hat{z}\right],
@@
   =\left|\ii
-    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
+    h\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
     f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
@@
   \leq\left|
-    \epsilon\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
+    h\hat{k}q\int_{\zhat_1}^{\zhat_2}\rho
     f(\hat{z}) W^*W^\prime d\hat{z}\right|^2.
@@
   |\hat{\nu}|\leq
-  \frac{\epsilon |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|) < \frac{\epsilon |q|}{2}\mathrm{max}(|f|) . 
+  \frac{h |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|) < \frac{h |q|}{2}\mathrm{max}(|f|) . 
 \end{align}
@@
 In practice, one might consider a vertical domain of a few scale 
-heights in a thin disk with $|q|=O(1)$. In this case, $|\epsilon\zhat|\ll1$ and  
+heights in a thin disk with $|q|=O(1)$. In this case, $|h\zhat|\ll1$ and  
 $f\simeq \zhat$, so that $\mathrm{max}|f| = O(1)$, implying a
-maximum growth rate $O(\epsilon \Omega_k)$, consistent with numerical
+maximum growth rate $O(h \Omega_K)$, consistent with numerical
 results.
@@
 % than this height, the growth rate is limited by $|f|$ at the vertical
-% boundaries considered. For $|\epsilon\hat{z}|\ll1$ we have $f\simeq
+% boundaries considered. For $|h\hat{z}|\ll1$ we have $f\simeq
 % \hat{z}$, in which case we can expect growth rates to increase at most linearly
@@
 \subsection{Explicit solutions in the thin-disk limit}\label{iso_explicit}
-Here we assume a thin disk ($\epsilon\ll1$) so that $\ln\rho\simeq
+Here we assume a thin disk ($h\ll1$) so that $\ln\rho\simeq
 -\zhat^2/2$ and $f(\zhat)\simeq \zhat$. However, we do not assume 
@@
 \begin{align}\label{iso_ode3}
-  W^{\prime\prime} - \left(1 + \frac{\ii q\epsilon \hat{k}}{1-\hat{\sigma}^2}\right)\hat{z}W^\prime  +
+  W^{\prime\prime} - \left(1 + \frac{\ii qh \hat{k}}{1-\hat{\sigma}^2}\right)\hat{z}W^\prime  +
   \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W = 0.
@@
   \left[\hat{\sigma}^2\left(1+\frac{\khat^2}{1-\hat{\sigma}^2}\right)
-    - l\left(1+\frac{\ii \epsilon q
+    - l\left(1+\frac{\ii h q
         \hat{k}}{1-\hat{\sigma}^2}\right)\right] a_l = 0. 
@@
 \hat{\sigma}^4 - \left(L+1+\khat^2\right)\hat{\sigma}^2 + L\left(1 +
-  \ii\epsilon q \khat\right) = 0. \label{simple_growth}
+  \ii h q \khat\right) = 0. \label{simple_growth}
 \end{align}
@@
 \begin{align}
-  \hat{\sigma}^2 \simeq L\left(\frac{1+\ii q \epsilon
+  \hat{\sigma}^2 \simeq L\left(\frac{1+\ii q h
        \hat{k}}{L+1+\hat{k}^2}\right), \label{simple_growth2}
@@
 % \begin{align}\label{sig2_iso}
-%   \hat{\sigma}^2 = L\left(\frac{1+\ii q \epsilon
+%   \hat{\sigma}^2 = L\left(\frac{1+\ii q h
 %       \hat{k}}{1+\hat{k}^2}\right).
@@
 %   \hat{\nu} =\sqrt{
-%     \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2\epsilon^2\hat{k}^2} - 
+%     \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} - 
 %       1\right)}. 
@@
 % \end{align}
-% Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_k^2$, our
+% Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_K^2$, our
 % analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
@@
 
-% For fixed $q$ and $\epsilon$, the growth rate vanishes for both
+% For fixed $q$ and $h$, the growth rate vanishes for both
 % $\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
@@
 % \begin{align}
-%   |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|\epsilon q|}{|\epsilon q|}},
+%   |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|h q|}{|h q|}},
 % \end{align}
@@
 % \begin{align}
-%   \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|\epsilon
-%     q|}{2\sqrt{1+|\epsilon q|}}. \label{iso_max_growth}
+%   \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|h
+%     q|}{2\sqrt{1+|h q|}}. \label{iso_max_growth}
 % \end{align}
@@
   &c_1 = \ii\beta\left\{\left(1-\gamma\right)\left[1 +
-      \khat^2\left(1+2M\right)^2 - 4 \ii\epsilon q\khat M (M+1)\right] 
+      \khat^2\left(1+2M\right)^2 - 4 \ii h q\khat M (M+1)\right] 
     - 2A^2\gamma M (M+1)\right\},\\
   &c_2 = \left(\khat^2 + 1\right)A + \beta^2\left\{(1-\gamma)\left[1
-      + \gamma \khat^2(1+2M)^2 - 4\ii\epsilon q \khat \gamma M(M+1)
+      + \gamma \khat^2(1+2M)^2 - 4\ii h q \khat \gamma M(M+1)
     \right]
@@
   \right\},\\
-  &c_3 = \beta\left\{\epsilon q \khat + \gamma \left[\ii + \epsilon q
+  &c_3 = \beta\left\{h q \khat + \gamma \left[\ii + h q
       k \left(1+2\khat^2\right)\right] - 3\ii - 2\ii
@@
   &c_4 =
-  \beta^2\left(1+\gamma\khat^2\right)\left[\gamma\left(1-\ii\epsilon q
+  \beta^2\left(1+\gamma\khat^2\right)\left[\gamma\left(1-\ii h q
     \khat\right)-2\right] - \left(1+\khat^2\right)^2,\\
@@
 \begin{align}
-  0 =&M(M+1)(1-\epsilon^2 q^2 \khat^2) + 4\beta\epsilon q M(M+1)\left(\hat{\omega}\khat\right) 
+  0 =&M(M+1)(1-h^2 q^2 \khat^2) + 4\beta h q M(M+1)\left(\hat{\omega}\khat\right) 
  +\left[1 +
     \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
- + 2\epsilon q \gamma\beta \left(\hat{\omega}\khat\right)^3 -  \left(\hat{\omega}\khat\right)^4\notag \\
+ + 2h q \gamma\beta \left(\hat{\omega}\khat\right)^3 -  \left(\hat{\omega}\khat\right)^4\notag \\
   &+ \beta^2\gamma^2\hat{\omega}^2
   \left(\hat{\omega}\khat\right)^4,\label{relax_cond1} \\
-   0=& 2\epsilon q M (M+1)\khat +
+   0=& 2h q M (M+1)\khat +
    \beta(1-\gamma)(1+2M)^2\khat\left(\hat{\omega}\khat\right) 
-   + \epsilon q \khat \left(\hat{\omega}\khat\right)^2 -
+   + h q \khat \left(\hat{\omega}\khat\right)^2 -
    2\beta\hat{\omega}\left(\hat{\omega}\khat\right)^2
-   - \epsilon q
+   - h q
    \gamma^2\beta^2\hat{\omega}\left(\hat{\omega}\khat\right)^3 
@@
 $|\hat{\omega}\khat|$ remain constant but consider $\khat\gg 1$ so
-that $|\hat{\omega}|\ll 1$.  Then for a thin disk $\epsilon \ll 1$
+that $|\hat{\omega}|\ll 1$.  Then for a thin disk $h \ll 1$
 we neglect the second, fourth and last term in Eq. \ref{relax_cond1};
@@
 $\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}) without
-assuming $|\epsilon q \khat|\lesssim 1$, as done in the main text to obtain the
+assuming $|h q \khat|\lesssim 1$, as done in the main text to obtain the
 equivalent expression for $M>0$ (Eq. \ref{bcrit_gen}).
@@
 % \begin{align}
-%   0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2\epsilon q
+%   0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2h q
 %   \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
 %   &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
-%   0=& \beta(1-\gamma)\khat^2 + \epsilon q \khat^2 (\hat{\omega}\khat)
-%   - 2\beta (\hat{\omega}\khat)^2 - \epsilon q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
+%   0=& \beta(1-\gamma)\khat^2 + h q \khat^2 (\hat{\omega}\khat)
+%   - 2\beta (\hat{\omega}\khat)^2 - h q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
 %     &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
@@
 % \begin{align}
-%   \hat{\omega}\khat = \frac{(\gamma-1)\beta}{\epsilon q}
+%   \hat{\omega}\khat = \frac{(\gamma-1)\beta}{h q}
 % \end{align}
@@
 % \begin{align}\label{beta_crit0}
-%   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(\epsilon
+%   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(h
 %         q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
@@
 \begin{align}\label{mmsn_epsilon}
-  \epsilon =
+  h =
   3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}, 

### 2015-04-23 application.tex
notation change from epsilon to h

@@
 \begin{align}\label{tc_diff_cool} 
-  t_\mathrm{diff} = \frac{1}{\eta k_x^2}.%\equiv \frac{\Omega_k^{-1}}{\hat{\eta}\khat^2}, 
+  t_\mathrm{diff} = \frac{1}{\eta k_x^2}.%\equiv \frac{\Omega_K^{-1}}{\hat{\eta}\khat^2}, 
 \end{align}
@@
 \begin{align}\label{real_beta}
-  \beta(r,z;\khat) \equiv t_c\Omega_k =
+  \beta(r,z;\khat) \equiv t_c\Omega_K =
   % \frac{1}{\hat{\eta}}\left[\frac{l_\mathrm{rad}^2}{3H^2}
   %   + \frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right]  \simeq
-  \frac{\Sigma^2\Omega_k}{\eta\rho^2}\left[\frac{1}{3\kappa_d^2\Sigma^2} 
+  \frac{\Sigma^2\Omega_K}{\eta\rho^2}\left[\frac{1}{3\kappa_d^2\Sigma^2} 
     + \frac{\hat{\rho}^2(z)}{2\pi \khat^2}\right].
@@
 \citep{shakura73}. Setting $t_\mathrm{visc}$ to be longer than the
-minimum VSI growth timescale $1/\epsilon\Omega$ implies 
-$\khat^2\lesssim \epsilon/\alpha$. Inserting $\epsilon \sim 0.05$,
+minimum VSI growth timescale $1/h\Omega$ implies 
+$\khat^2\lesssim h/\alpha$. Inserting $h \sim 0.05$,
 we find only modes with $\khat\lesssim 20$ are relevant for
@@
 wavenumbers, in units of the local orbital time $P_\mathrm{orb}\equiv
-2\pi/\Omega_k$. In fact, we either only find the fundamental mode, or 
+2\pi/\Omega_K$. In fact, we either only find the fundamental mode, or 
 that it is the most unstable, except for the appearance of some 

### 2015-04-23 linear_adia.tex
notation change from epsilon to h

@@
 Our final approximation apply to thin disks for 
-which $\epsilon\ll 1$. Then we may write $\rho/\rho_0\simeq
+which $h\ll 1$. Then we may write $\rho/\rho_0\simeq
 \exp{(-\zhat^2/2)}$. The governing equation,
@@
 \begin{align}
-  &A \equiv 1 + \ii \epsilon q \hat{k},\\
+  &A \equiv 1 + \ii h q \hat{k},\\
   &B \equiv \hat{\sigma}^2\left(\chi + \hat{k}^2\right) -
-  \left(\chi + \ii \epsilon q \hat{k}\right),\\
+  \left(\chi + \ii h q \hat{k}\right),\\
   &C \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
-    \epsilon q\hat{k}\right), 
+    h q\hat{k}\right), 
 \end{align}
@@
   \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\sigma}^2 = \frac{1 +
-  \ii\epsilon q \hat{k}}{1+\hat{k}^2}, \label{pert_basic} 
+  \ii h q \hat{k}}{1+\hat{k}^2}, \label{pert_basic} 
 \end{align}
@@
 %   \left(\hat{\sigma}^2 -1\right)\delta\chi, \\
-%   0&= 2a\left(1+\ii\epsilon q \khat\right) - \left(\khat^2 - \ii\epsilon
+%   0&= 2a\left(1+\ii h q \khat\right) - \left(\khat^2 - \ii h
 %     q \khat\right)\delta\chi.
@@
 %   \delta \hat{\sigma} =
-%   -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii\epsilon
-%       q - \hat{k}\right)^2}{2\left(1+\ii\epsilon q \hat{k}\right)\left(1+\hat{k}^2\right)^2}\delta\beta.
+%   -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii h
+%       q - \hat{k}\right)^2}{2\left(1+\ii h q \hat{k}\right)\left(1+\hat{k}^2\right)^2}\delta\beta.
 % \end{align}
@@
   -\frac{\left(\gamma-1\right)\hat{k}^2 \left(\hat{k}^2 -
-      2\epsilon^2q^2\hat{k}^2 - \epsilon^2q^2\right)}{2\left(1+\epsilon^2 q^2
+      2h^2q^2\hat{k}^2 - h^2q^2\right)}{2\left(1+h^2 q^2
       \hat{k}^2\right)\left(1+\hat{k}^2\right)^2}\delta\beta.  
 \end{align}
-Since $\epsilon \ll 1$, introducing finite cooling $\delta\beta>0$
+Since $h \ll 1$, introducing finite cooling $\delta\beta>0$
 implies $\delta\hat{\nu} < 0$, i.e. stabilization, since $\gamma>1$. 
@@
 also assume $|q|$ and $\beta$ are $O(1)$ or smaller. For a thin disk
-($\epsilon \ll 1$) we have 
+($h \ll 1$) we have 
 \begin{align}
- 0 = &M(M+1)\left(1-\epsilon^2q^2\khat^2\right) \notag\\ 
+ 0 = &M(M+1)\left(1-h^2q^2\khat^2\right) \notag\\ 
  &+ \left[1 +
@@
    -  \left(\hat{\omega}\khat\right)^4,\label{relax_cond_simp1}\\
-   0= &2M(M+1) + \frac{\beta}{\epsilon q} (1-\gamma)(1+2M)^2
+   0= &2M(M+1) + \frac{\beta}{h q} (1-\gamma)(1+2M)^2
    \left(\hat{\omega}\khat\right) + \left(\hat{\omega}\khat\right)^2. \label{relax_cond_simp2}
@@
 assumption.) If $\beta\ll 1$ and the disk is sufficiently thin so that  
-$|\epsilon q \khat |^2\lesssim 1$, 
+$|h q \khat |^2\lesssim 1$, 
 then an approximate solution to Eq. \ref{relax_cond_simp1} is 
@@
   \left(\hat{\omega}\khat\right)^2 \simeq (M+1)\left(1 -
-    \frac{M|\epsilon q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
+    \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
 \end{align}
@@
 \begin{align}\label{bcrit_gen}
-  \beta &= \frac{\epsilon|q|}{(\gamma - 1)}
-  \frac{\sqrt{1+M}}{(1+2M)}\left[1 + \frac{M(2M-1)|\epsilon q
+  \beta &= \frac{h|q|}{(\gamma - 1)}
+  \frac{\sqrt{1+M}}{(1+2M)}\left[1 + \frac{M(2M-1)|h q
       \khat|^2}{2(2M+1)^2}\right].
@@
 \end{align}
-Since $|\epsilon q \khat|^2\lesssim 1$ by assumption, we have  
+Since $|h q \khat|^2\lesssim 1$ by assumption, we have  
 \begin{align}\label{iso_vsi_cond}
   \beta\leq \beta_\mathrm{crit}  \equiv
-  \frac{\epsilon|q|}{\gamma-1}, 
+  \frac{h|q|}{\gamma-1}, 
 \end{align}
@@
 obtained without assuming it is much less than unity from the onset or
-$|\epsilon q\khat|\lesssim 1$. % This is because for $M=0$, 
+$|h q\khat|\lesssim 1$. % This is because for $M=0$, 
 % Eq. \ref{relax_cond_simp1} --- \ref{relax_cond_simp2} simplify 
@@
 is required ($\beta_\mathrm{crit}\ll 1$) because
-$\beta_\mathrm{crit}\propto \epsilon$. Physically, this is because
+$\beta_\mathrm{crit}\propto h$. Physically, this is because
 vertical shear is weak in a thin disk, so even a small entropy
@@
 % Eq. \ref{beta_crit0} is an upper limit. For a thin  
-% disk $\epsilon\ll 1$, we thus infer 
+% disk $h\ll 1$, we thus infer 
 % \begin{align}\label{iso_vsi_cond}
-%   \beta \lesssim \frac{|\epsilon q|}{\gamma-1} \equiv
+%   \beta \lesssim \frac{|h q|}{\gamma-1} \equiv
 %   \beta_\mathrm{crit} 

### 2015-04-23 numerical.tex
notation change from epsilon to h

@@
 Unless otherwise stated, our fiducial disk parameters are 
-$(\gamma, \Gamma) = (1.4, 1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.05)$, and 
+$(\gamma, \Gamma) = (1.4, 1.011)$, $(p,q, h)=(-1.5,-1,0.05)$, and 
 vertical domain size $\zmax=5H$. This setup is effectively vertically
@@
     normalized by its mid-plane value, for  the fiducial disk model with $\Gamma=1.011$
-    and $(p,q,\epsilon)=(-1.5,-1,0.05)$. 
+    and $(p,q, h)=(-1.5,-1,0.05)$. 
     \label{omega_z} 
@@
 % $\khat\gtrsim 10$. Growth rates are over-estimated by our analysis but 
-% are $O(\epsilon\Omega_k)$, which is consistent with the discussion in
+% are $O( h\Omega_k)$, which is consistent with the discussion in
 % Appendix \ref{max_growth1}. The match in $\nu$ worsens for 
@@
 %     % a disk with $(\gamma,
-%     % \Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+%     % \Gamma)=(1.4,1.011)$, $(p,q, h)=(-1.5,-1,0.1)$ and
 %     % $\beta=10^{-3}$. 
@@
     with $\beta=10^{-3}$. %  a disk with 
-    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
+    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q, h)=(-1.5,-1,0.1)$ and
     % $\beta=10^{-3}$. 
@@
 point and vary $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and
-$\epsilon\in[0.02,0.1]$ separately. The numerically obtained
+$ h\in[0.02,0.1]$ separately. The numerically obtained
 $\beta_\mathrm{crit}$ are shown in Fig. \ref{bcrit_compare} in
@@
 % The agreement improves with decreasing  $|q|$,
-% $\epsilon$ and increasing $\gamma$, i.e. weaker instability. There is
+% $ h$ and increasing $\gamma$, i.e. weaker instability. There is
 % a noticeable difference as $\gamma\to1$ because the derivation of
@@
     parameters. The fiducial setup is $(\gamma, \Gamma= (1.4, 1.011)$
-    and $(p,q,\epsilon)=(-1.5,-1,0.05)$. Top: varying 
+    and $(p,q, h)=(-1.5,-1,0.05)$. Top: varying 
     $q\in[-1.2,-0.2]$; middle: varying $\gamma\in[1.2,2.0]$; bottom:
-    varying $\epsilon\in[0.02,0.1]$. The perturbation wavenumber is
+    varying $ h\in[0.02,0.1]$. The perturbation wavenumber is
     $\khat=10$.  
@@
 Nevertheless, here we briefly consider a vertically non-isothermal
-disk model with $(p,q,\epsilon)=(0,-1,0.05)$ and $\Gamma=1.0837$. This
+disk model with $(p,q, h)=(0,-1,0.05)$ and $\Gamma=1.0837$. This
 value of $\Gamma$ is chosen so that the standard vertical domain size $\zmax =
@@
 thermal timescale in the vertically non-isothermal disk is 
-given by $\epsilon|s|/(\gamma - \Gamma)$. This 
+given by $ h|s|/(\gamma - \Gamma)$. This 
 evaluates to 0.158, which is close to the cut-off thermal timescale
@@
     critical thermal timescale given by
-    $\epsilon|s|/(\gamma-\Gamma)=0.158$.  
+    $ h|s|/(\gamma-\Gamma)=0.158$.  
     \label{bcrit_compare2}}   
@@
   % We briefly consider vertically non-isothermal disks with 
-% $\Gamma=1.4$ and $(p,q,\epsilon)=(0,-1,0.05)$ as simulated in
+% $\Gamma=1.4$ and $(p,q, h)=(0,-1,0.05)$ as simulated in
 % \cite{nelson13}. In this case we set $\zmax=0.99H_s$.
@@
 % $\beta\simeq0.17,\,0.083,\,0.045$, respectively. Interestingly, these
-% values are equal to $\epsilon|q|/(\gamma-\Gamma)$. This suggests that
+% values are equal to $ h|q|/(\gamma-\Gamma)$. This suggests that
 % the critical thermal relaxation timescale for vertically
@@
 %     $\gamma>1.4$. Other disk parameters are
-%     $(p,q,\epsilon)=(0,-1,0.05)$ and the perturbation wavenumber is
+%     $(p,q, h)=(0,-1,0.05)$ and the perturbation wavenumber is
 %     $\khat=30$.   

### 2015-04-23 summary.tex
notation change from epsilon to h

@@
 \begin{align*}
-  t_\mathrm{crit} = \frac{\epsilon |q|}{\gamma-1}\Omega_k^{-1}, 
+  t_\mathrm{crit} = \frac{ h |q|}{\gamma-1}\Omega_k^{-1}, 
 \end{align*}
@@
 disks. The thermal timescale must be shorter than the disk
-dynamical timescale by $O(\epsilon)$, which is
-quite short for thin disks ($\epsilon\ll 1$), because vertical shear is weak.
+dynamical timescale by $O( h)$, which is
+quite short for thin disks ($ h\ll 1$), because vertical shear is weak.
 
@@
 %  \nu_\mathrm{conv,max} =
-%  -\frac{\epsilon^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
+%  -\frac{ h^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
 %  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}

### 2015-04-23 appendix.tex
messed up notation change. need to fix by inspection.

@@
 \begin{align}
-  & \ii\sigma\delta v_z = \frac{dW}{dz} + \frac{d\ln{\rho}}{d z}\left(W-Q\right),\label{ode_vz} \\
-  &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W =\ii\sigma \left(\frac{\ii
+  & \ii\zeta\delta v_z = \frac{dW}{dz} + \frac{d\ln{\rho}}{d z}\left(W-Q\right),\label{ode_vz} \\
+  &\frac{\zeta^2}{c_s^2}Q + \frac{\zeta^2k_x^2}{D}W =\ii\zeta \left(\frac{\ii
       k_x r}{D}\frac{d\Omega^2}{dz} -
-    \frac{d\ln{\rho}}{dz}\right)\delta v_z - \ii\sigma \frac{d\delta v_z}{dz},
+    \frac{d\ln{\rho}}{dz}\right)\delta v_z - \ii\zeta \frac{d\delta v_z}{dz},
   % \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] 
@@
 %    1\right)\delta v_z,  
-  &\sigma^2W - \gamma\sigma^2 Q +
-  \frac{\ii\sigma}{t_c}\left(W-Q\right) 
-  =\ii \sigma c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma -
+  &\zeta^2W - \gamma\zeta^2 Q +
+  \frac{\ii\zeta}{t_c}\left(W-Q\right) 
+  =\ii \zeta c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma -
     1\right)\delta v_z,
@@
 \begin{align}
-  D \equiv \kappa^2 - \sigma^2,
+  D \equiv \kappa^2 - \zeta^2,
 \end{align} 
@@
     v_z}{dz} \notag\\
-  &+ \left\{\sigma^2\left(\frac{k_x^2}{D} +
+  &+ \left\{\zeta^2\left(\frac{k_x^2}{D} +
       \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
@@
 \end{align}
-where we recall $\chi = \left(1-\ii\sigma t_c\right)/\left(1-\ii\sigma
+where we recall $\chi = \left(1-\ii\zeta t_c\right)/\left(1-\ii\zeta
 t_c \gamma\right)$. Making the low frequency and Keplerian
@@
       \frac{\chi}{\left(\hat{k}^2+\chi\right)}qh^2F\right]\right\}\delta v_z \notag\\&=
-  -\hat{\sigma}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
+  -\hat{\zeta}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
 \end{align} 
@@
     \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime
-  +\left\{\hat{\sigma}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
+  +\left\{\hat{\zeta}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
     -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii h  
@@
 \begin{align}
-  \hat{\sigma}^2\left(\frac{1}{\gamma} +
+  \hat{\zeta}^2\left(\frac{1}{\gamma} +
     \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
@@
 % When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
-% $\hat{\sigma}^2$ is real, and  $\hat{\sigma}^2>0$ if $\gamma>1$. As
+% $\hat{\zeta}^2$ is real, and  $\hat{\zeta}^2>0$ if $\gamma>1$. As
 % expected, a sub-adiabatically stratified disk is stable in the absence
@@
 In the presence of vertical shear $q\neq0$, Eq. \ref{adia_integral}
-suggests $\hat{\sigma}^2$ is complex for real $\khat$, which implies the
+suggests $\hat{\zeta}^2$ is complex for real $\khat$, which implies the
 possibility of instability, regardless of the value of
@@
 above. Inclusion of the global radial gradient terms results in the
-correct behavior  ($\nu\to0$ as $\beta\to\infty$). Despite this
+correct behavior  ($\sigma\to0$ as $\beta\to\infty$). Despite this
 caveat, Fig. \ref{gcorr_compare} shows that provided we consider
@@
 \end{align}
-(i.e. $\delta P = c_s^2\delta \rho$)% , and so $  \ii\hat{\sigma}c_s\delta v_z = W^\prime $
+(i.e. $\delta P = c_s^2\delta \rho$)% , and so $  \ii\hat{\zeta}c_s\delta v_z = W^\prime $
 % from Eq. \ref{lin_vz}
@@
   0 = W^{\prime\prime} + \left[\ln\rho^\prime - \frac{\ii h q \hat{k} f(\zhat)
-      }{1-\hat{\sigma}^2} \right]W^\prime +
-  \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W,\label{iso_ode0} 
+      }{1-\hat{\zeta}^2} \right]W^\prime +
+  \hat{\zeta}^2\left(1+\frac{\hat{k}^2}{1-\hat{\zeta}^2}\right)W,\label{iso_ode0} 
 \end{align}
@@
 \subsection{Stability conditions in the low-frequency limit}
-In the low-frequency limit we assume $|\hat{\sigma}|\ll 1$ and
+In the low-frequency limit we assume $|\hat{\zeta}|\ll 1$ and
 approximate Eq. \ref{iso_ode0} as 
@@
     f(\zhat)\right]W^\prime +
-  \hat{\sigma}^2\left(1+\hat{k}^2\right)W. \label{iso_ode1}
+  \hat{\zeta}^2\left(1+\hat{k}^2\right)W. \label{iso_ode1}
 \end{align}
@@
 \begin{align}
-  \hat{\sigma}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat
+  \hat{\zeta}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat
   =\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
@@
 \end{align}
-It follows that for instability ($\imag\hat{\sigma}>0$), it is necessary to
+It follows that for instability ($\imag\hat{\zeta}>0$), it is necessary to
 have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. there must
@@
 Here we show that the growth rate is limited by the maximum vertical
-shear rate in the domain. % Writing $\hat{\sigma} = \hat{\omega} +
-% \ii\hat{\nu}$ with real $\hat{\omega}$ and $\hat{\nu}$, 
+shear rate in the domain. % Writing $\hat{\zeta} = \hat{\omega} +
+% \ii\hat{\zeta}$ with real $\hat{\omega}$ and $\hat{\zeta}$, 
 The real and imaginary parts of 
@@
 \begin{align}
-  \left(\hat{\omega}^2-\hat{\nu}^2\right)\left(1+\hat{k}^2\right)
+  \left(\hat{\omega}^2-\hat{sigma}^2\right)\left(1+\hat{k}^2\right) 
   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
@@
     f(\hat{z}) W^*W^\prime d\zhat\right], \\
-   2\hat{\omega}\hat{\nu}\left(1+\hat{k}^2\right)
+   2\hat{\omega}\hat{\sigma}\left(1+\hat{k}^2\right)
   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z}
@@
 \end{align}
-where we recall $\hat{\omega} = \real \hat{\sigma}$ and
-$\hat{\nu}=\imag\hat{\sigma}$. 
+where we recall $\hat{\omega} = \real \hat{\zeta}$ and
+$\hat{\sigma}=\imag\hat{\zeta}$. 
 Adding the square of these equations give
 \begin{align}
-  \left[|\hat{\sigma}|^2\left(1+\hat{k}^2\right)
+  \left[|\hat{\zeta}|^2\left(1+\hat{k}^2\right)
     \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
     \int_{\zhat_1}^{\zhat_2}\rho\left|W^\prime \right|^2d\hat{z}\right]^2
-  +4\hat{\nu}^2\left(1+\hat{k}^2\right) 
+  +4\hat{\sigma}^2\left(1+\hat{k}^2\right) 
   \int_{\zhat_1}^{\zhat_2}\rho
@@
 \begin{align}\label{sigma_finite_domain} 
-  4\hat{\nu}^2\left(1+\hat{k}^2\right) 
+  4\hat{\sigma}^2\left(1+\hat{k}^2\right) 
   \int_{\zhat_1}^{\zhat_2}\rho
@@
 \begin{align}\label{max_growth}
-  |\hat{\nu}|\leq
+  |\hat{\sigma}|\leq
   \frac{h |\hat{k} q|}{2\sqrt{1+\hat{k}^2}}\mathrm{max}(|f|) < \frac{h |q|}{2}\mathrm{max}(|f|) . 
@@
 \begin{align}
-  \nu < \mathrm{max} \left|r\frac{d\Omega}{dz}\right|, 
+  \sigma < \mathrm{max} \left|r\frac{d\Omega}{dz}\right|, 
 \end{align}
@@
 \begin{align}\label{iso_ode3}
-  W^{\prime\prime} - \left(1 + \frac{\ii qh \hat{k}}{1-\hat{\sigma}^2}\right)\hat{z}W^\prime  +
-  \hat{\sigma}^2\left(1+\frac{\hat{k}^2}{1-\hat{\sigma}^2}\right)W = 0.
+  W^{\prime\prime} - \left(1 + \frac{\ii qh \hat{k}}{1-\hat{\zeta}^2}\right)\hat{z}W^\prime  +
+  \hat{\zeta}^2\left(1+\frac{\hat{k}^2}{1-\hat{\zeta}^2}\right)W = 0.
 \end{align}
@@
   (l+2)(l+1)a_{l+2} +
-  \left[\hat{\sigma}^2\left(1+\frac{\khat^2}{1-\hat{\sigma}^2}\right)
+  \left[\hat{\zeta}^2\left(1+\frac{\khat^2}{1-\hat{\zeta}^2}\right)
     - l\left(1+\frac{\ii h q
-        \hat{k}}{1-\hat{\sigma}^2}\right)\right] a_l = 0. 
+        \hat{k}}{1-\hat{\zeta}^2}\right)\right] a_l = 0. 
 \end{align}
@@
 order $L$, so that the vertical kinetic energy remain bounded as 
-$|\zhat|\to\infty$.  Then the eigenfrequency $\hat{\sigma}$ is given via 
+$|\zhat|\to\infty$.  Then the eigenfrequency $\hat{\zeta}$ is given via 
 \begin{align}
-\hat{\sigma}^4 - \left(L+1+\khat^2\right)\hat{\sigma}^2 + L\left(1 +
+\hat{\zeta}^4 - \left(L+1+\khat^2\right)\hat{\zeta}^2 + L\left(1 +
   \ii h q \khat\right) = 0. \label{simple_growth}
@@
 written as \begin{align}
-  \hat{\sigma}^2\khat^2 =
-  \left(L-\hat{\sigma}^2\right)\left(1-\hat{\sigma}^2\right), \label{iso_disp_full}
+  \hat{\zeta}^2\khat^2 =
+  \left(L-\hat{\zeta}^2\right)\left(1-\hat{\zeta}^2\right), \label{iso_disp_full}
 \end{align}
@@
 this case the solutions are Hermite polynomials, $W\propto
-\He_L(\zhat)$.  The eigenfrequency $\hat{\sigma} = \hat{\omega}$ is real and the disk
+\He_L(\zhat)$.  The eigenfrequency $\hat{\zeta} = \hat{\omega}$ is real and the disk
 is stable. The low frequency branch of Eq. \ref{iso_disp_full} are 
@@
 \begin{align}
-  \hat{\sigma}^2 \simeq L\left(\frac{1+\ii q h
+  \hat{\zeta}^2 \simeq L\left(\frac{1+\ii q h
        \hat{k}}{L+1+\hat{k}^2}\right), \label{simple_growth2}
@@
 $L\geq1$ since we can choose the sign 
-of the square root to make $\hat{\nu} = \imag\hat{\sigma}>0$.  These
+of the square root to make $\hat{\sigma} = \imag\hat{\zeta}>0$.  These
 are the low frequency unstable modes seen in
 Fig. \ref{compare_modes_iso_kx10} and
-Fig. \ref{compare_modes_iso_kx30}, for which $\nu\propto|\omega|$.
+Fig. \ref{compare_modes_iso_kx30}, for which $\sigma\propto|\omega|$.
 
@@
 % \begin{align}\label{sig2_iso}
-%   \hat{\sigma}^2 = L\left(\frac{1+\ii q h
+%   \hat{\zeta}^2 = L\left(\frac{1+\ii q h
 %       \hat{k}}{1+\hat{k}^2}\right).
 % \end{align}
-% The eigenfrequency $\hat{\sigma}$ is therefore complex for
+% The eigenfrequency $\hat{\zeta}$ is therefore complex for
 % $L\geq1$. The dispersion relation Eq. \ref{sig2_iso} can also be
@@
 % \ref{relax_coeff}. We are interested in 
-% unstable modes for which $\hat{\nu}>0$ and is given by
+% unstable modes for which $\hat{\zeta}>0$ and is given by
 % \begin{align}
-%   \hat{\nu} =\sqrt{
+%   \hat{\zeta} =\sqrt{
 %     \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} - 
@@
 % \begin{align}
-%   n \equiv \hat{\sigma}^2(1+\hat{k}^2). 
+%   n \equiv \hat{\zeta}^2(1+\hat{k}^2). 
 % \end{align} 
@@
 % \begin{align}
-%   \left|\hat{\sigma}\right| = \sqrt{n}
+%   \left|\hat{\zeta}\right| = \sqrt{n}
 %   \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
@@
 % \begin{align}
-%   \mathrm{max}\left(\hat{\nu}\right) =\frac{\sqrt{L}|h
+%   \mathrm{max}\left(\hat{\zeta}\right) =\frac{\sqrt{L}|h
 %     q|}{2\sqrt{1+|h q|}}. \label{iso_max_growth}
@@
 \subsection{Limiting form with real frequencies and $\khat^2\gg 1$}\label{disp_neut_limit}
-Assuming neutral modes with real $\hat{\sigma}=\hat{\omega}$ and $\khat^2\gg 1$, the real and imaginary parts of
+Assuming neutral modes with real $\hat{\zeta}=\hat{\omega}$ and $\khat^2\gg 1$, the real and imaginary parts of
 Eq. \ref{relax_disp} are approximately 
@@
 %  \rho_0 = 2.7\times10^{-9}
-%  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
+%  \hat{\zeta}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
 \rho_0 = 1.7\times10^{-9}

### 2015-04-23 application.tex
messed up notation change. need to fix by inspection.

@@
 In Fig. \ref{mmsn_overall} we plot characteristic growth timescales
-$t_\mathrm{grow} \equiv 1/\nu$ of the fundamental mode with a range of
+$t_\mathrm{grow} \equiv 1/\sigma$ of the fundamental mode with a range of
 wavenumbers, in units of the local orbital time $P_\mathrm{orb}\equiv

### 2015-04-23 linear_adia.tex
messed up notation change. need to fix by inspection.

@@
-\subsection{Thin-disk approximation}\label{analytic_relax}
-Our final approximation apply to thin disks for 
-which $h\ll 1$. Then we may write $\rho/\rho_0\simeq
-\exp{(-\zhat^2/2)}$. The governing equation,
-Eq. \ref{vertiso_gov_nondim}, is then   
-\begin{align}
-  \delta v_z^{\prime\prime} - z A\delta v_z^\prime +
-  (B - C\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
-\end{align}
-with
-\begin{align}
-  &A \equiv 1 + \ii h q \hat{k},\\
-  &B \equiv \hat{\sigma}^2\left(\chi + \hat{k}^2\right) -
-  \left(\chi + \ii h q \hat{k}\right),\\
-  &C \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
-    h q\hat{k}\right), 
-\end{align}
-and $\chi =
-\left(1-\ii\hat{\sigma}\beta\right)/\left(1-\ii\hat{\sigma}\beta\gamma\right)
-$ in terms of dimensionless variables.
-
-We remark that Eq. \ref{nearly_iso_explicit} is in the same 
-form as Eq. 41 in \cite{lubow93}, which describe adiabatic axisymmetric waves in
-a vertically isothermal disk without vertical shear. The two equations
-become identical after  setting $q=0$ (no vertical shear) and
-$\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
-and making the low-frequency, Keplerian approximation in
-\citeauthor{lubow93} after a change of variables.
-
 \subsection{Isothermal perturbations}\label{iso_pert}
@@
 \begin{align}
-  \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\sigma}^2 = \frac{1 +
+  \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\zeta}^2 = \frac{1 +
   \ii h q \hat{k}}{1+\hat{k}^2}, \label{pert_basic} 
@@
   \beta \to 0 + \delta\beta,\, \delta v_z\to \delta v_z+\delta
-  v_{z1},\,\hat{\sigma} \to \hat{\sigma} + \delta\hat{\sigma}, 
+  v_{z1},\,\hat{\zeta} \to \hat{\zeta} + \delta\hat{\zeta}, 
 \end{align}
@@
 \begin{align}
-  \chi \to 1 + \delta\chi = 1 + \ii \hat{\sigma}\left(\gamma-1\right)\delta\beta,
+  \chi \to 1 + \delta\chi = 1 + \ii \hat{\zeta}\left(\gamma-1\right)\delta\beta,
 \end{align}
-with $\delta v_z$ and $\hat{\sigma}$ given by Eq. \ref{pert_basic}. We
+with $\delta v_z$ and $\hat{\zeta}$ given by Eq. \ref{pert_basic}. We
 may then seek  
@@
 Eq. \ref{nearly_iso_explicit}, keeping only first order terms, and
-solve for $\delta\hat{\sigma}$ using the above expressions for $\delta\chi$
-and $\hat{\sigma}$. We find imaginary part of $\delta\hat{\sigma}$
+solve for $\delta\hat{\zeta}$ using the above expressions for $\delta\chi$
+and $\hat{\zeta}$. We find imaginary part of $\delta\hat{\zeta}$
 is 
@@
 % \begin{align}
-%   0 &= 2a + 2\hat{\sigma}\left(1+\khat^2\right)\delta\hat{\sigma} +
-%   \left(\hat{\sigma}^2 -1\right)\delta\chi, \\
+%   0 &= 2a + 2\hat{\zeta}\left(1+\khat^2\right)\delta\hat{\zeta} +
+%   \left(\hat{\zeta}^2 -1\right)\delta\chi, \\
 %   0&= 2a\left(1+\ii h q \khat\right) - \left(\khat^2 - \ii h
@@
 % \begin{align}\label{nearly_iso_dsig}
-%   \delta \hat{\sigma} =
+%   \delta \hat{\zeta} =
 %   -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii h
@@
 \begin{align}
-  \delta\hat{\nu} =
+  \delta\hat{\sigma} =
   -\frac{\left(\gamma-1\right)\hat{k}^2 \left(\hat{k}^2 -
@@
 \begin{align}
-  B(\hat{\sigma}) + \lambda(\hat{\sigma}) =
-  M\left[A-2\lambda(\hat{\sigma})\right].\label{adia_disp0} 
+  B(\hat{\zeta}) + \lambda(\hat{\zeta}) =
+  M\left[A-2\lambda(\hat{\zeta})\right].\label{adia_disp0} 
 \end{align}
 For ease of analysis, we re-arrange Eq. \ref{adia_disp0} and square it
-to obtain a polynomial in $\hat{\sigma}$,  
+to obtain a polynomial in $\hat{\zeta}$,  
 \begin{align}
-  0 = \sum_{l=0}^{6}c_l\hat{\sigma}^l,\label{relax_disp}
+  0 = \sum_{l=0}^{6}c_l\hat{\zeta}^l,\label{relax_disp}
 \end{align}
 where the coefficients $c_l$ are given in Appendix \ref{relax_coeff}.
-The dispersion relation $\hat{\sigma}=\hat{\sigma}(\khat)$ is
+The dispersion relation $\hat{\zeta}=\hat{\zeta}(\khat)$ is
 complicated and generally requires a numerical solution. However,
@@
 
-% At marginal stability, the eigenfrequency $\hat{\sigma}=\hat{\omega}$
+% At marginal stability, the eigenfrequency $\hat{\zeta}=\hat{\omega}$
 % is real. The dispersion relation for the fundamental mode is then

### 2015-04-23 linear_iso.tex
notation change from epsilon to h
messed up notation change. need to fix by inspection.

@@
+\subsection{Shearing box, low-frequency and Keplerian
+  approximations}\label{approx_gov}
+
+In Appendix \ref{adia_improve} we derive a governing equation with the following simplifications:
+
+\begin{enumerate}
+
+\item  We adopt the  `vertically global shearing box' (VGSB) formalism 
+  developed by \citetalias{mcnally14}, in which explicit dependencies on the
+  background radial disk structure in 
+  Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
+  $\hat{g}_c=0$. However, the global radial temperature 
+  gradient in the expression for the 
+  vertical shear (Eq. \ref{vertical_shear}) is retained.
+
+  In Appendix \ref{global_corr}  we discuss limitations of using
+  the VGSB, where we show that this approximation is valid provided we 
+  consider consider fast thermal relaxation times
+  ($\beta\lesssim 1$) and perturbations with small radial lengthscales
+  ($k_xH\gtrsim 1$), as appropriate for the VSI.
+
+\item We consider eigenvalues such that
+  $|\zeta^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency
+  approximation. We are therefore considering inertial waves
+  \citepalias{barker15}.
+
+\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
+  0$. This is already captured explicitly in the linearized equations through
+  the term $d\Omega/dz$. We therefore simplify 
+  further by ignoring the vertical dependence of $\Omega$ and $\kappa$
+  elsewhere. For clarity, we replace $\kappa^2$ with
+  $\Omega_K^2$. We call this the Keplerian approximation.
+
+\end{enumerate}
+% Making both of these approximations is equivalent to setting 
+% $D\to\Omega_K^2$.  After making this replacement, we eliminate $W$ and
+% $Q$ between  Eq. \ref{lin_vz} and Eq. \ref{ode_w}---\ref{ode_Q},
+% making use of Eq. \ref{vertical_shear}, to
+% obtain an equation for $\delta v_z$.
+
+These approximations lead to Eq. \ref{full_ode} which, for a thin disk
+$h\ll 1$, simplifies further to  
+%The result for $\Gamma=1$ is
+\begin{align}
+  0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
+      q}{\Omega_K^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
+    v_z}{dz} \notag\\
+  % &+ \left[\zeta^2\left(\frac{k_x^2}{\Omega_K^2} +
+  %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
+  &+ \left[\frac{\zeta^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_K^2} +
+      \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
+        q}{\Omega_K^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
+  &\phantom{+=}\left.-
+    \frac{c_s^2}{\Omega_K^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
+      \frac{\ii k_x q}{r}\right)
+    \left(1-\chi\right) 
+  \right]\delta v_z,\label{vertiso_gov}
+\end{align}
+where
+\begin{align}
+  \chi \equiv \frac{1-\ii\zeta t_c}{1-\ii\zeta t_c\gamma}.
+\end{align}
+
+It is convenient to introduce the dimensionless variables
+\begin{align}
+  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\zeta} =\hat{\omega} +
+  \ii\hat{\sigma}= \zeta/\Omega_K,
+\end{align}
+where $\hat{\omega}$ and $\hat{\sigma}$ are real. 
+Eq. \ref{vertiso_gov} becomes 
+\begin{align}
+  0=& \delta v_z ^{\prime\prime} + \left(1 + \ii h
+    q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime \notag\\
+  &+
+  \left[\hat{\zeta}^2\left(\hat{k}^2+\chi\right) +
+    \left(  \chi + \ii h q\hat{k}\right)\ln\rho^{\prime\prime}\right.
+  \notag\\
+  &\left.- \ln\rho^{\prime
+      2}\left(\khat^2 -
+      \ii h
+      q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
+\end{align}
+where $^\prime$ denotes $d/d\zhat$.
+
+\subsection{Thin-disk approximation}\label{analytic_relax}
+Our final approximation apply to thin disks for 
+which $h\ll 1$. Then we may write $\rho/\rho_0\simeq
+\exp{(-\zhat^2/2)}$. The governing equation,
+Eq. \ref{vertiso_gov_nondim}, is then   
+\begin{align}
+  \delta v_z^{\prime\prime} - z A\delta v_z^\prime +
+  (B - C\zhat^2)\delta v_z = 0,\label{nearly_iso_explicit}
+\end{align}
+with
+\begin{align}
+  &A \equiv 1 + \ii h q \hat{k},\\
+  &B \equiv \hat{\zeta}^2\left(\chi + \hat{k}^2\right) -
+  \left(\chi + \ii h q \hat{k}\right),\\
+  &C \equiv \left(1-\chi\right)\left(\hat{k}^2 - \ii
+    h q\hat{k}\right), 
+\end{align}
+and $\chi =
+\left(1-\ii\hat{\zeta}\beta\right)/\left(1-\ii\hat{\zeta}\beta\gamma\right)
+$ in terms of dimensionless variables.
+
+We remark that Eq. \ref{nearly_iso_explicit} is in the same 
+form as Eq. 41 in \cite{lubow93}, which describe adiabatic axisymmetric waves in
+a vertically isothermal disk without vertical shear. The two equations
+become identical after  setting $q=0$ (no vertical shear) and
+$\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
+and making the low-frequency, Keplerian approximation in
+\citeauthor{lubow93} after a change of variables.
+
 \section{Analytical discussion of vertically isothermal
@@
 % \begin{align}
-%   &\frac{\sigma^2}{c_s^2}Q + \frac{\sigma^2k_x^2}{D}W \notag\\ 
+%   &\frac{\zeta^2}{c_s^2}Q + \frac{\zeta^2k_x^2}{D}W \notag\\ 
 %   &=\left(\frac{\ii
@@
 %     \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
-%   &\sigma^2W - \frac{\gamma}{\Gamma}\sigma^2 Q +
-%   \frac{\ii\sigma}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
+%   &\zeta^2W - \frac{\gamma}{\Gamma}\zeta^2 Q +
+%   \frac{\ii\zeta}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
 %   &=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
@@
 % \begin{align}
-%   D \equiv \kappa^2 - \sigma^2,
+%   D \equiv \kappa^2 - \zeta^2,
 % \end{align} 
@@
 % treatment.
-
-\subsection{Shearing box, low-frequency and Keplerian
-  approximations}\label{approx_gov}
-
-In Appendix \ref{adia_improve} we derive a governing equation with the following simplifications:
-
-\begin{enumerate}
-
-\item  We adopt the  `vertically global shearing box' (VGSB) formalism 
-  developed by \citetalias{mcnally14}, in which explicit dependencies on the
-  background radial disk structure in 
-  Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
-  $\hat{g}_c=0$. However, the global radial temperature 
-  gradient in the expression for the 
-  vertical shear (Eq. \ref{vertical_shear}) is retained.
-
-  In Appendix \ref{global_corr}  we discuss limitations of using
-  the VGSB, where we show that this approximation is valid provided we 
-  consider consider fast thermal relaxation times
-  ($\beta\lesssim 1$) and perturbations with small radial lengthscales
-  ($k_xH\gtrsim 1$), as appropriate for the VSI.
-
-\item We consider eigenvalues such that
-  $|\sigma^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency
-  approximation. We are therefore considering inertial waves
-  \citepalias{barker15}.
-
-\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
-  0$. This is already captured explicitly in the linearized equations through
-  the term $d\Omega/dz$. We therefore simplify 
-  further by ignoring the vertical dependence of $\Omega$ and $\kappa$
-  elsewhere. For clarity, we replace $\kappa^2$ with
-  $\Omega_k^2$. We call this the Keplerian approximation.
-
-\end{enumerate}
-% Making both of these approximations is equivalent to setting 
-% $D\to\Omega_k^2$.  After making this replacement, we eliminate $W$ and
-% $Q$ between  Eq. \ref{lin_vz} and Eq. \ref{ode_w}---\ref{ode_Q},
-% making use of Eq. \ref{vertical_shear}, to
-% obtain an equation for $\delta v_z$.
-
-These approximations lead to Eq. \ref{full_ode} which, for a thin disk
-$\epsilon\ll 1$, simplifies further to  
-%The result for $\Gamma=1$ is
-\begin{align}
-  0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
-      q}{\Omega_k^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
-    v_z}{dz} \notag\\
-  % &+ \left[\sigma^2\left(\frac{k_x^2}{\Omega_k^2} +
-  %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
-  &+ \left[\frac{\sigma^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_k^2} +
-      \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
-        q}{\Omega_k^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
-  &\phantom{+=}\left.-
-    \frac{c_s^2}{\Omega_k^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
-      \frac{\ii k_x q}{r}\right)
-    \left(1-\chi\right) 
-  \right]\delta v_z,\label{vertiso_gov}
-\end{align}
-where
-\begin{align}
-  \chi \equiv \frac{1-\ii\sigma t_c}{1-\ii\sigma t_c\gamma}.
-\end{align}
-
-It is convenient to introduce the dimensionless variables
-\begin{align}
-  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\sigma} =\hat{\omega} +
-  \ii\hat{\nu}= \sigma/\Omega_k,
-\end{align}
-where $\hat{\omega}$ and $\hat{\nu}$ are real. 
-Eq. \ref{vertiso_gov} becomes 
-\begin{align}
-  0=& \delta v_z ^{\prime\prime} + \left(1 + \ii\epsilon
-    q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime \notag\\
-  &+
-  \left[\hat{\sigma}^2\left(\hat{k}^2+\chi\right) +
-    \left(  \chi + \ii \epsilon q\hat{k}\right)\ln\rho^{\prime\prime}\right.
-  \notag\\
-  &\left.- \ln\rho^{\prime
-      2}\left(\khat^2 -
-      \ii\epsilon
-      q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
-\end{align}
-where $^\prime$ denotes $d/d\zhat$.

### 2015-04-23 paper.tex
messed up notation change. need to fix by inspection.

@@
   For irradiated disks, we find that the VSI is most vigorous for
-  rapid cooling with $t_c < \OmK^{-1} \epsilon |q| / (\gamma -1)$ 
+  rapid cooling with $t_c < \OmK^{-1}  h |q| / (\gamma -1)$ 
   in terms of the Keplerian orbital frequency, $\OmK$; the disk's
-  aspect ratio, $\epsilon \ll 1$; the radial  
+  aspect ratio, $ h \ll 1$; the radial  
   powerlaw temperature gradient, $q$; and the adiabatic index,
@@
 \begin{align}
-  t _c \ll \Omega_k^{-1}. 
+  t _c \ll \Omega_K^{-1}. 
 \end{align}
-where $\Omega_k$ is the Keplerian frequency. 
+where $\Omega_K$ is the Keplerian frequency. 
 To see why this should be expected, we need to compare vertical shear
 with vertical buoyancy. For this preliminary discussion we consider a
-vertically isothermal disk with aspect-ratio $\epsilon$,  
+vertically isothermal disk with aspect-ratio $ h$,  
 dimensionless radial temperature gradient $q$ and adiabatic index
-$\gamma$. For a PPD, $\epsilon \sim 0.05$, $|q|\sim 1$ and $\gamma\simeq 1.4$.
+$\gamma$. For a PPD, $ h \sim 0.05$, $|q|\sim 1$ and $\gamma\simeq 1.4$.
 
@@
 \begin{align}\label{vshear_thin}
-  r \frac{\p \Omega^2 }{\p z} \simeq \epsilon q \Omega_k^2\left(\frac{z}{H}\right),
+  r \frac{\p \Omega^2 }{\p z} \simeq  h q \Omega_K^2\left(\frac{z}{H}\right),
 \end{align}
@@
   N_z^2 \simeq \left(\frac{\gamma-1}{\gamma}\right) \left(\frac{z}{H}\right)^2
-  \Omega_k^2.  
+  \Omega_K^2.  
 \end{align}
@@
 stabilizing; while the right-hand side represents vertical shear,
-which is destabilizing. In a thin disk $\epsilon \ll 1$, so the destabilizing effect of 
-vertical shear is rather weak because $|r\p_z\Omega|\sim\epsilon\Omega_k
-\ll \Omega_k$. By comparison, the buoyancy frequency $N_z\sim
-\Omega_k$. It follows that $t_c\lesssim O(\epsilon\Omega_k^{-1})$ is
+which is destabilizing. In a thin disk $ h \ll 1$, so the destabilizing effect of 
+vertical shear is rather weak because $|r\p_z\Omega|\sim h\Omega_K
+\ll \Omega_K$. By comparison, the buoyancy frequency $N_z\sim
+\Omega_K$. It follows that $t_c\lesssim O( h\Omega_K^{-1})$ is
 required for instability.  \emph{In a thin  
   disk the thermal timescale must be shorter than the dynamical
-  timescale by $O(\epsilon)\ll 1$ for the VSI to operate.}
+  timescale by $O( h)\ll 1$ for the VSI to operate.}
 
@@
 \begin{align}\label{prelim_bcrit}
-  (\gamma - 1)t_c < \epsilon |q| \Omega_k^{-1}
+  (\gamma - 1)t_c <  h |q| \Omega_K^{-1}
 \end{align}   

### 2015-04-23 setup.tex
messed up notation change. need to fix by inspection.

@@
 \section{Linear problem}\label{linear}
+
 We consider axisymmetric perturbations to the above equilibria in the
@@
 \begin{align}
-  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left( - \ii\sigma
+  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left( - \ii\zeta
       t\right)},    
 \end{align}
-and similarly for other fluid variables, where $\sigma = \omega + \ii
-\nu$ is a (generally) complex frequency, with $\omega$ and $\nu$ being
+and similarly for other fluid variables, where $\zeta = \omega + \ii
+\sigma$ is a (generally) complex frequency, with $\omega$ and $\sigma$ being
 the real frequency and growth rate, respectively.
 
-The linearized fluid equations in cylindrical co-ordinates are  
-\begin{align}
-  \ii\sigma\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
-    z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
-  \ii\sigma\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
-  \delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
-  \ii\sigma\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
-  r\frac{\p\Omega}{\p z}\delta v_z,\\
-  \ii \sigma\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
-  \frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},\\
-  \ii\sigma\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
-  \left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
-      \delta v_z}{\p z}\right]\notag \\
-  &+ \frac{1}{t_c}\left(\delta P -  
-    \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
-\end{align}
-where the coefficients $\hat{g}_c=1$ are inserted to label terms
-associated with the background radial disk structure.
+% The linearized fluid equations in cylindrical co-ordinates are  
+% \begin{align}
+%   \ii\zeta\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
+%     z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
+%   \ii\zeta\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
+%   \delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
+%   \ii\zeta\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
+%   r\frac{\p\Omega}{\p z}\delta v_z,\\
+%   \ii \zeta\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
+%   \frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},\\
+%   \ii\zeta\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
+%   \left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
+%       \delta v_z}{\p z}\right]\notag \\
+%   &+ \frac{1}{t_c}\left(\delta P -  
+%     \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
+% \end{align}
+% where the coefficients $\hat{g}_c=1$ are inserted to label terms
+% associated with the background radial disk structure.
 
@@
 
-\subsection{Radially-local approximation}  
 We consider perturbations with small radial length-scales 
@@
 \begin{align}
-  \ii\sigma \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
+  \ii\zeta \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
     v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \hat{g}_c
   \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
-  \ii\sigma \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
+  \ii\zeta \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
   \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
-   \ii \sigma\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
+   \ii \zeta\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
   \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
-   \ii\sigma\delta v_z &= \frac{dW}{dz} +
+   \ii\zeta\delta v_z &= \frac{dW}{dz} +
   \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
-  \ii \sigma W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
+  \ii \zeta W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
   c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
@@
   \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x,\label{lin_energy}
-  % -&\ii\sigma W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\sigma\frac{\gamma}{\Gamma} Q +
+  % -&\ii\zeta W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\zeta\frac{\gamma}{\Gamma} Q +
   % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}

### 2015-04-23 appendix.tex
restructured, moved early part of sec 5 to sec 4 as discussed. notation changes need further checking

@@
   the vertically global shearing box}\label{adia_improve}
-Here we explain the derivation of Eq. \ref{vertiso_gov} used in our 
+Here we explain the derivation of Eq. \ref{nearly_iso_explicit} used in our 
 analytical discussion in \S\ref{analytical}. We shall need the
@@
 Since we are interested in thin disks with $h \ll 1$ and $|q| =
-O(1)$, we neglect the terms proportional to $h^2$, which gives 
-Eq. \ref{vertiso_gov_nondim} (the dimensional version being
-Eq. \ref{vertiso_gov}). This is equivalent to neglecting the   
-vertical dependence of $D$ by setting $D\to\Omega_K^2$ in
-Eq. \ref{ode_w} before eliminating variables in favor of $\delta
-v_z$, which is the simpler way to obtain Eq. \ref{vertiso_gov}.
+O(1)$, we neglect the terms proportional to $h^2$, giving 
+\begin{align}
+  0= \delta v_z ^{\prime\prime} + \left(1 + \ii h
+     q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime 
+   +
+   \left[\hat{\zeta}^2\left(\hat{k}^2+\chi\right) +
+     \left(  \chi + \ii h q\hat{k}\right)\ln\rho^{\prime\prime}
+     - \ln\rho^{\prime
+       2}\left(\khat^2 -
+       \ii h
+       q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
+ \end{align}
+ which is used in the discussion below.  Eq. \ref{nearly_iso_explicit}, as
+ used in our analytical discussion, is obtained by inserting the 
+ thin-disk expression for the density field. We comment that the simpler way
+ of obtaining Eq. \ref{vertiso_gov_nondim} is to neglect the vertical 
+ dependence of $D$ by setting $D\to\Omega_K^2$ in Eq. \ref{ode_w} before eliminating
+ variables in favor of $\delta v_z$.
 
@@
 \begin{align}
-  \left(\hat{\omega}^2-\hat{sigma}^2\right)\left(1+\hat{k}^2\right) 
+  \left(\hat{\omega}^2-\hat{\sigma}^2\right)\left(1+\hat{k}^2\right) 
   \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -

### 2015-04-23 linear_adia.tex
restructured, moved early part of sec 5 to sec 4 as discussed. notation changes need further checking

@@
+\section{Analytic results for vertically isothermal disks}\label{analytical}
+
 \subsection{Isothermal perturbations}\label{iso_pert}
@@
 instability growth rate (it is limited by the maximum vertical shear
-rate in the disk). We also develop solutions in the thin disk
-limit. The methods used in Appendix \ref{iso_discuss} and results obtained
+rate in the disk). Neither of these actually require the thin-disk
+form of the density field described above.  
+We also develop solutions in the thin-disk
+limit, but without assuming low frequency from the onset. 
+The methods used in Appendix \ref{iso_discuss} and results obtained
 therein motivates the solution procedure employed in the following
@@
 Since $h \ll 1$, introducing finite cooling $\delta\beta>0$
-implies $\delta\hat{\nu} < 0$, i.e. stabilization, since $\gamma>1$. 
+implies $\delta\hat{\sigma} < 0$, i.e. stabilization, since $\gamma>1$. 
 % Interestingly, for both

### 2015-04-23 linear_iso.tex
restructured, moved early part of sec 5 to sec 4 as discussed. notation changes need further checking

@@
-\subsection{Shearing box, low-frequency and Keplerian
-  approximations}\label{approx_gov}
+\subsection{Reduced equation for vertically isothermal disks}
+%We begin with a discussion of vertically isothermal disks
+%where $\Gamma=1$, then $c_s$ is a constant in the radially-local
+%approximation. 
+%we first introduce a
+%number of approximations below that allows a simple governing equation
+%to be derived and compared with previous works. 
+%However, all of
+%the approximations described below are relaxed in our numerical
+%study.
 
-In Appendix \ref{adia_improve} we derive a governing equation with the following simplifications:
+In preparation for a numerical study of the above equations, we will 
+first present a fully analytic discussion for vertically isothermal disks with
+$\Gamma=1$, so that $c_s$ is vertically constant. To this end, 
+we introduce the following simplifications:
+%In Appendix \ref{adia_improve} we derive  a with the following simplifications:
 
@@
 
-  In Appendix \ref{global_corr}  we discuss limitations of using
-  the VGSB, where we show that this approximation is valid provided we 
-  consider consider fast thermal relaxation times
-  ($\beta\lesssim 1$) and perturbations with small radial lengthscales
-  ($k_xH\gtrsim 1$), as appropriate for the VSI.
-
 \item We consider eigenvalues such that
   $|\zeta^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency 
-  approximation. We are therefore considering inertial waves
-  \citepalias{barker15}.
+  approximation
 
 \item We are mainly interested in the effect of vertical shear, i.e. $q\neq
-  0$. This is already captured explicitly in the linearized equations through
-  the term $d\Omega/dz$. We therefore simplify 
-  further by ignoring the vertical dependence of $\Omega$ and $\kappa$
-  elsewhere. For clarity, we replace $\kappa^2$ with
-  $\Omega_K^2$. We call this the Keplerian approximation.
+  0$ for vertically isothermal disks. This is already captured
+  explicitly in the linearized equations through the term
+  $d\Omega/dz$. We thus ignore the vertical dependence of $\Omega$ and
+  $\kappa$ elsewhere, and replace them with
+  $\Omega_K$.  We call this the Keplerian approximation.
+
+\item We assume thin disks for which $h\ll 1$, and write 
+  $\rho/\rho_0\simeq  \exp{(-z^2/2H^2)}$.
 
 \end{enumerate}
-% Making both of these approximations is equivalent to setting 
-% $D\to\Omega_K^2$.  After making this replacement, we eliminate $W$ and
-% $Q$ between  Eq. \ref{lin_vz} and Eq. \ref{ode_w}---\ref{ode_Q},
-% making use of Eq. \ref{vertical_shear}, to
-% obtain an equation for $\delta v_z$.
+% These approximations lead to Eq. \ref{full_ode} which, for a thin disk
+% $h\ll 1$, simplifies further to  
+% %The result for $\Gamma=1$ is
+% \begin{align}
+%   0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
+%       q}{\Omega_K^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
+%     v_z}{dz} \notag\\
+%   % &+ \left[\zeta^2\left(\frac{k_x^2}{\Omega_K^2} +
+%   %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
+%   &+ \left[\frac{\zeta^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_K^2} +
+%       \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
+%         q}{\Omega_K^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
+%   &\phantom{+=}\left.-
+%     \frac{c_s^2}{\Omega_K^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
+%       \frac{\ii k_x q}{r}\right)
+%     \left(1-\chi\right) 
+%   \right]\delta v_z,\label{vertiso_gov}
+% \end{align}
+% where
+% \begin{align}
+%   \chi \equiv \frac{1-\ii\zeta t_c}{1-\ii\zeta t_c\gamma}.
+% \end{align}
 
-These approximations lead to Eq. \ref{full_ode} which, for a thin disk
-$h\ll 1$, simplifies further to  
-%The result for $\Gamma=1$ is
-\begin{align}
-  0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
-      q}{\Omega_K^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
-    v_z}{dz} \notag\\
-  % &+ \left[\zeta^2\left(\frac{k_x^2}{\Omega_K^2} +
-  %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
-  &+ \left[\frac{\zeta^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_K^2} +
-      \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
-        q}{\Omega_K^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
-  &\phantom{+=}\left.-
-    \frac{c_s^2}{\Omega_K^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
-      \frac{\ii k_x q}{r}\right)
-    \left(1-\chi\right) 
-  \right]\delta v_z,\label{vertiso_gov}
-\end{align}
-where
-\begin{align}
-  \chi \equiv \frac{1-\ii\zeta t_c}{1-\ii\zeta t_c\gamma}.
-\end{align}
-
-It is convenient to introduce the dimensionless variables
+ In terms of the dimensionless variables
 \begin{align}
@@
 \end{align}
-where $\hat{\omega}$ and $\hat{\sigma}$ are real. 
-Eq. \ref{vertiso_gov} becomes 
-\begin{align}
-  0=& \delta v_z ^{\prime\prime} + \left(1 + \ii h
-    q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime \notag\\
-  &+
-  \left[\hat{\zeta}^2\left(\hat{k}^2+\chi\right) +
-    \left(  \chi + \ii h q\hat{k}\right)\ln\rho^{\prime\prime}\right.
-  \notag\\
-  &\left.- \ln\rho^{\prime
-      2}\left(\khat^2 -
-      \ii h
-      q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
-\end{align}
-where $^\prime$ denotes $d/d\zhat$.
-
-\subsection{Thin-disk approximation}\label{analytic_relax}
-Our final approximation apply to thin disks for 
-which $h\ll 1$. Then we may write $\rho/\rho_0\simeq
-\exp{(-\zhat^2/2)}$. The governing equation,
-Eq. \ref{vertiso_gov_nondim}, is then   
+where $\hat{\omega}$ and $\hat{\sigma}$ are real, the above
+approximations lead to a single second order ordinary differential
+equation, 
+% Eq. \ref{vertiso_gov} becomes 
+% \begin{align}
+%   0=& \delta v_z ^{\prime\prime} + \left(1 + \ii h
+%     q\hat{k}\right)\ln\rho^{\prime}\delta v_z^\prime \notag\\
+%   &+
+%   \left[\hat{\zeta}^2\left(\hat{k}^2+\chi\right) +
+%     \left(  \chi + \ii h q\hat{k}\right)\ln\rho^{\prime\prime}\right.
+%   \notag\\
+%   &\left.- \ln\rho^{\prime
+%       2}\left(\khat^2 -
+%       \ii h
+%       q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
+% \end{align}
+% where $^\prime$ denotes $d/d\zhat$.
+%\subsection{Thin-disk approximation}\label{analytic_relax}
+%The governing equation, Eq. \ref{vertiso_gov_nondim}, is then   
 \begin{align}
@@
 \end{align}
-with
+where $^\prime$ denotes $d/d\zhat$ and 
 \begin{align}
@@
 \end{align}
-and $\chi =
-\left(1-\ii\hat{\zeta}\beta\right)/\left(1-\ii\hat{\zeta}\beta\gamma\right)
-$ in terms of dimensionless variables.
+where
+\begin{align}
+\chi = \frac{1-\ii\hat{\zeta}\beta}{1-\ii\hat{\zeta}\beta\gamma}.
+\end{align}
+The derivation of Eq. \ref{nearly_iso_explicit}  is outlined in
+Appendix \ref{adia_improve}.
 
@@
 
-\section{Analytical discussion of vertically isothermal
-  disks in the framework of the vertically global shearing
-  box}\label{analytical}  
-We begin with a discussion of vertically isothermal disks
-where $\Gamma=1$, then $c_s$ is a constant in the radially-local
-approximation. 
-In order to facilitate a fully analytic discussion, we first introduce a
-number of approximations below that allows a simple governing equation
-to be derived and compared with previous works. Similar approximations
-are employed by \citetalias{nelson13} and \citetalias{barker15}. However, all of
-the approximations described below are relaxed in our numerical study.
+ In Appendix \ref{global_corr}  we discuss limitations of using
+  the VGSB, where we show that this approximation is valid provided we 
+  consider consider fast thermal relaxation times
+  ($\beta\lesssim 1$) and perturbations with small radial lengthscales
+  ($\khat\gtrsim 1$), as appropriate for the VSI.
 

### 2015-04-23 numerical.tex
messed up notation change. need to fix by inspection.
restructured, moved early part of sec 5 to sec 4 as discussed. notation changes need further checking

@@
 where $\bm{\xi}$ is the Lagrangian displacement with meridional 
-components $\xi_{x,z} = \ii\delta v_{x,z}/\sigma$. In some cases we
+components $\xi_{x,z} = \ii\delta v_{x,z}/\zeta$. In some cases we
 impose a rigid boundary at which $\delta v_z=0$.
@@
 % Eq. \ref{simple_growth}. We find good agreement in the
-% wave frequency $\omega$ at all $\khat$ and in growth rates $\nu$ for
+% wave frequency $\omega$ at all $\khat$ and in growth rates $\sigma$ for
 % $\khat\gtrsim 10$. Growth rates are over-estimated by our analysis but 
-% are $O( h\Omega_k)$, which is consistent with the discussion in
-% Appendix \ref{max_growth1}. The match in $\nu$ worsens for 
+% are $O( h\Omega_K)$, which is consistent with the discussion in
+% Appendix \ref{max_growth1}. The match in $\sigma$ worsens for 
 % decreasing $\khat$ due to the neglect of global radial structure in
@@
 below.)  The `fundamental corrugation mode', marked by `f',
-corresponds to the eigenvalue with the smallest $|\sigma|$. The
+corresponds to the eigenvalue with the smallest $|\zeta|$. The
 fundamental mode is robust  

### 2015-04-23 setup.tex
restructured, moved early part of sec 5 to sec 4 as discussed. notation changes need further checking

@@
 \begin{align}
-  &\frac{\p\rho}{\p t} + \nabla\cdot{\left(\rho\bm{v}\right)} = 0,\\
+  &\frac{\p\rho}{\p t} + \nabla\cdot{\left(\rho\bm{v}\right)} = 0,\label{full_mass}\\
   &\frac{\p\bm{v}}{\p t} + \bm{v}\cdot\nabla\bm{v} =
@@
 \section{Linear problem}\label{linear} 
+In this section we present the two sets of equations used to study the
+axisymmetric linear stability of the above equilibria. We first give
+the linearized equations in the radially-local approximation, which
+are solved in our numerical study. From this set, we further derive a
+simplified equation for vertically isothermal disks used for analytic
+discussion in the next section.
 
+\subsection{Radially local approximation}
 We consider axisymmetric perturbations to the above equilibria in the
@@
 \begin{align}
-  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left( - \ii\zeta
+  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left(\ii k_rr - \ii\zeta
       t\right)},    
@@
 \sigma$ is a (generally) complex frequency, with $\omega$ and $\sigma$ being
-the real frequency and growth rate, respectively.
+the real frequency and growth rate, respectively, and $k_r$ is a real 
+radial wavenumber. We take $k_r>0$ without loss of generality.
 
+The VSI is expected to operate for perturbations with small radial
+wavelengths \citepalias{nelson13,barker15}. We thus linearize 
+Eq. \ref{full_mass}---\ref{full_energy} assuming perturbations 
+in the above form with radially slowly-varying
+amplitudes and $k_r r \gg 1$, so that the radial dependence of $\delta\rho(r,z)$, etc.,
+may be neglected, as well as curvature terms. Henceforth we are
+considering solutions at some fiducial radius $r=r_0$.  
+%\citetalias{barker15} have validated this 
+% approximation for the VSI by comparison with the linear problem in
+% global cylindrical disk geometry.
+
+In order to emphasize the radially local approximation, we introduce
+the local Cartesian co-ordinates $(x,y,z)$ corresponding to the the
+local $(r,\phi,z)$ directions, respectively.  
+%We thus linearize 
+%and replace  
 % The linearized fluid equations in cylindrical co-ordinates are  
@@
 % associated with the background radial disk structure. 
-
 % This approach has been adopted
@@
 % radial lengthscales. 
-
-We consider perturbations with small radial length-scales 
-so that we can ignore the radial variations in the background
-structure and radial gradients. \citetalias{barker15} have validated this
-approximation by comparison with the linear VSI in global cylindrical
-disk geometry. We thus evaluate all coefficients in
-Eq. \ref{full_mass1}---\ref{full_energy1} at the fiducial radius
-$r=r_0$. However, the full vertical dependence of the coefficients are
-retained. Then we can assume perturbations have spatial dependence in
-the form 
-\begin{align}
-  \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r}.
-\end{align}
-We take $k_r>0$ without loss of generality, and further assume $k_rr\gg1$ so that curvature terms can also be 
-ignored.
-
-In order to emphasize the radially local approximation, we
-introduce the local Cartesian co-ordinates $(x,y,z)$ corresponding to
-the the local $(r,\phi,z)$ directions, respectively. Then writing
-$\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
+%We consider perturbations with small radial length-scales 
+%so that we can ignore the radial variations in the background
+%structure and radial gradients. 
+%We thus evaluate all coefficients in
+%Eq. \ref{full_mass1}---\ref{full_energy1} at the fiducial radius
+%$r=r_0$. However, the full vertical dependence of the coefficients are
+%retained. % Then we can assume perturbations have spatial dependence in
+% the form 
+% \begin{align}
+%   \delta \rho(r,z) \to \delta\rho(z)\exp{\ii k_r r}.
+% \end{align}
+% We take $k_r>0$ without loss of generality, and further assume $k_rr\gg1$ so that curvature terms can also be 
+% ignored.       
+Then writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
 linearized system of equations are 
@@
 study, but for analytical discussion
-we will make further simplifications.
+we will make further simplifications, as discussed next.

### 2015-04-23 appendix.tex
going through re-structured paper

@@
 \begin{align}
-  & \ii\zeta\delta v_z = \frac{dW}{dz} + \frac{d\ln{\rho}}{d z}\left(W-Q\right),\label{ode_vz} \\
-  &\frac{\zeta^2}{c_s^2}Q + \frac{\zeta^2k_x^2}{D}W =\ii\zeta \left(\frac{\ii
+  & \ii\upsilon\delta v_z = \frac{dW}{dz} + \frac{d\ln{\rho}}{d z}\left(W-Q\right),\label{ode_vz} \\
+  &\frac{\upsilon^2}{c_s^2}Q + \frac{\upsilon^2k_x^2}{D}W =\ii\upsilon \left(\frac{\ii
       k_x r}{D}\frac{d\Omega^2}{dz} -
-    \frac{d\ln{\rho}}{dz}\right)\delta v_z - \ii\zeta \frac{d\delta v_z}{dz},
+    \frac{d\ln{\rho}}{dz}\right)\delta v_z - \ii\upsilon \frac{d\delta v_z}{dz},
   % \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] 
@@
 %    1\right)\delta v_z,  
-  &\zeta^2W - \gamma\zeta^2 Q +
-  \frac{\ii\zeta}{t_c}\left(W-Q\right) 
-  =\ii \zeta c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma -
+  &\upsilon^2W - \gamma\upsilon^2 Q +
+  \frac{\ii\upsilon}{t_c}\left(W-Q\right) 
+  =\ii \upsilon c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma -
     1\right)\delta v_z,
@@
 \begin{align}
-  D \equiv \kappa^2 - \zeta^2,
+  D \equiv \kappa^2 - \upsilon^2,
 \end{align} 
@@
     v_z}{dz} \notag\\
-  &+ \left\{\zeta^2\left(\frac{k_x^2}{D} +
+  &+ \left\{\upsilon^2\left(\frac{k_x^2}{D} +
       \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
@@
 \end{align}
-where we recall $\chi = \left(1-\ii\zeta t_c\right)/\left(1-\ii\zeta
+where we recall $\chi = \left(1-\ii\upsilon t_c\right)/\left(1-\ii\upsilon
 t_c \gamma\right)$. Making the low frequency and Keplerian
@@
       \frac{\chi}{\left(\hat{k}^2+\chi\right)}qh^2F\right]\right\}\delta v_z \notag\\&=
-  -\hat{\zeta}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
+  -\hat{\upsilon}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
 \end{align} 
@@
    +
-   \left[\hat{\zeta}^2\left(\hat{k}^2+\chi\right) +
+   \left[\hat{\upsilon}^2\left(\hat{k}^2+\chi\right) +
      \left(  \chi + \ii h q\hat{k}\right)\ln\rho^{\prime\prime}
@@
     \hat{k}\right)\left(\ln\rho^{\prime}\delta v_z\right)^\prime
-  +\left\{\hat{\zeta}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
+  +\left\{\hat{\upsilon}^2\left(\frac{1}{\gamma}+\hat{k}^2\right) 
     -\left(\frac{\gamma-1}{\gamma}\right)\left[\ln\rho^{\prime\prime}+\hat{k}^2\left(1-\frac{\ii h  
@@
 \begin{align}
-  \hat{\zeta}^2\left(\frac{1}{\gamma} +
+  \hat{\upsilon}^2\left(\frac{1}{\gamma} +
     \hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|\delta
@@
 % When $q\equiv0$, all the terms on the right-hand-side (RHS) are real. Then
-% $\hat{\zeta}^2$ is real, and  $\hat{\zeta}^2>0$ if $\gamma>1$. As
+% $\hat{\upsilon}^2$ is real, and  $\hat{\upsilon}^2>0$ if $\gamma>1$. As
 % expected, a sub-adiabatically stratified disk is stable in the absence
@@
 In the presence of vertical shear $q\neq0$, Eq. \ref{adia_integral}
-suggests $\hat{\zeta}^2$ is complex for real $\khat$, which implies the
+suggests $\hat{\upsilon}^2$ is complex for real $\khat$, which implies the
 possibility of instability, regardless of the value of
@@
 \end{align}
-(i.e. $\delta P = c_s^2\delta \rho$)% , and so $  \ii\hat{\zeta}c_s\delta v_z = W^\prime $
+(i.e. $\delta P = c_s^2\delta \rho$)% , and so $  \ii\hat{\upsilon}c_s\delta v_z = W^\prime $
 % from Eq. \ref{lin_vz}
@@
   0 = W^{\prime\prime} + \left[\ln\rho^\prime - \frac{\ii h q \hat{k} f(\zhat)
-      }{1-\hat{\zeta}^2} \right]W^\prime +
-  \hat{\zeta}^2\left(1+\frac{\hat{k}^2}{1-\hat{\zeta}^2}\right)W,\label{iso_ode0} 
+      }{1-\hat{\upsilon}^2} \right]W^\prime +
+  \hat{\upsilon}^2\left(1+\frac{\hat{k}^2}{1-\hat{\upsilon}^2}\right)W,\label{iso_ode0} 
 \end{align}
@@
 \subsection{Stability conditions in the low-frequency limit}
-In the low-frequency limit we assume $|\hat{\zeta}|\ll 1$ and
+In the low-frequency limit we assume $|\hat{\upsilon}|\ll 1$ and
 approximate Eq. \ref{iso_ode0} as 
@@
     f(\zhat)\right]W^\prime +
-  \hat{\zeta}^2\left(1+\hat{k}^2\right)W. \label{iso_ode1}
+  \hat{\upsilon}^2\left(1+\hat{k}^2\right)W. \label{iso_ode1}
 \end{align}
@@
 \begin{align}
-  \hat{\zeta}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat
+  \hat{\upsilon}^2\left(1+\hat{k}^2\right)\int_{\zhat_1}^{\zhat_2}\rho|W|^2d\zhat
   =\int_{\zhat_1}^{\zhat_2}\rho|W^\prime|^2d\zhat 
@@
 \end{align}
-It follows that for instability ($\imag\hat{\zeta}>0$), it is necessary to
+It follows that for instability ($\imag\hat{\upsilon}>0$), it is necessary to
 have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. there must
@@
 Here we show that the growth rate is limited by the maximum vertical
-shear rate in the domain. % Writing $\hat{\zeta} = \hat{\omega} +
-% \ii\hat{\zeta}$ with real $\hat{\omega}$ and $\hat{\zeta}$, 
+shear rate in the domain. % Writing $\hat{\upsilon} = \hat{\omega} +
+% \ii\hat{\upsilon}$ with real $\hat{\omega}$ and $\hat{\upsilon}$, 
 The real and imaginary parts of 
@@
 \end{align}
-where we recall $\hat{\omega} = \real \hat{\zeta}$ and
-$\hat{\sigma}=\imag\hat{\zeta}$. 
+where we recall $\hat{\omega} = \real \hat{\upsilon}$ and
+$\hat{\sigma}=\imag\hat{\upsilon}$. 
 Adding the square of these equations give
 \begin{align}
-  \left[|\hat{\zeta}|^2\left(1+\hat{k}^2\right)
+  \left[|\hat{\upsilon}|^2\left(1+\hat{k}^2\right)
     \int_{\zhat_1}^{\zhat_2}\rho|W|^2d\hat{z} -
@@
 \begin{align}\label{iso_ode3}
-  W^{\prime\prime} - \left(1 + \frac{\ii qh \hat{k}}{1-\hat{\zeta}^2}\right)\hat{z}W^\prime  +
-  \hat{\zeta}^2\left(1+\frac{\hat{k}^2}{1-\hat{\zeta}^2}\right)W = 0.
+  W^{\prime\prime} - \left(1 + \frac{\ii qh \hat{k}}{1-\hat{\upsilon}^2}\right)\hat{z}W^\prime  +
+  \hat{\upsilon}^2\left(1+\frac{\hat{k}^2}{1-\hat{\upsilon}^2}\right)W = 0.
 \end{align}
@@
   (l+2)(l+1)a_{l+2} +
-  \left[\hat{\zeta}^2\left(1+\frac{\khat^2}{1-\hat{\zeta}^2}\right)
+  \left[\hat{\upsilon}^2\left(1+\frac{\khat^2}{1-\hat{\upsilon}^2}\right)
     - l\left(1+\frac{\ii h q
-        \hat{k}}{1-\hat{\zeta}^2}\right)\right] a_l = 0. 
+        \hat{k}}{1-\hat{\upsilon}^2}\right)\right] a_l = 0. 
 \end{align}
@@
 order $L$, so that the vertical kinetic energy remain bounded as 
-$|\zhat|\to\infty$.  Then the eigenfrequency $\hat{\zeta}$ is given via 
+$|\zhat|\to\infty$.  Then the eigenfrequency $\hat{\upsilon}$ is given via 
 \begin{align}
-\hat{\zeta}^4 - \left(L+1+\khat^2\right)\hat{\zeta}^2 + L\left(1 +
+\hat{\upsilon}^4 - \left(L+1+\khat^2\right)\hat{\upsilon}^2 + L\left(1 +
   \ii h q \khat\right) = 0. \label{simple_growth}
@@
 written as \begin{align}
-  \hat{\zeta}^2\khat^2 =
-  \left(L-\hat{\zeta}^2\right)\left(1-\hat{\zeta}^2\right), \label{iso_disp_full}
+  \hat{\upsilon}^2\khat^2 =
+  \left(L-\hat{\upsilon}^2\right)\left(1-\hat{\upsilon}^2\right), \label{iso_disp_full}
 \end{align}
@@
 this case the solutions are Hermite polynomials, $W\propto
-\He_L(\zhat)$.  The eigenfrequency $\hat{\zeta} = \hat{\omega}$ is real and the disk
+\He_L(\zhat)$.  The eigenfrequency $\hat{\upsilon} = \hat{\omega}$ is real and the disk
 is stable. The low frequency branch of Eq. \ref{iso_disp_full} are 
@@
 \begin{align}
-  \hat{\zeta}^2 \simeq L\left(\frac{1+\ii q h
+  \hat{\upsilon}^2 \simeq L\left(\frac{1+\ii q h
        \hat{k}}{L+1+\hat{k}^2}\right), \label{simple_growth2}
@@
 $L\geq1$ since we can choose the sign 
-of the square root to make $\hat{\sigma} = \imag\hat{\zeta}>0$.  These
+of the square root to make $\hat{\sigma} = \imag\hat{\upsilon}>0$.  These
 are the low frequency unstable modes seen in
@@
 % \begin{align}\label{sig2_iso}
-%   \hat{\zeta}^2 = L\left(\frac{1+\ii q h
+%   \hat{\upsilon}^2 = L\left(\frac{1+\ii q h
 %       \hat{k}}{1+\hat{k}^2}\right).
 % \end{align}
-% The eigenfrequency $\hat{\zeta}$ is therefore complex for
+% The eigenfrequency $\hat{\upsilon}$ is therefore complex for
 % $L\geq1$. The dispersion relation Eq. \ref{sig2_iso} can also be
@@
 % \ref{relax_coeff}. We are interested in 
-% unstable modes for which $\hat{\zeta}>0$ and is given by
+% unstable modes for which $\hat{\upsilon}>0$ and is given by
 % \begin{align}
-%   \hat{\zeta} =\sqrt{
+%   \hat{\upsilon} =\sqrt{
 %     \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} - 
@@
 % \begin{align}
-%   n \equiv \hat{\zeta}^2(1+\hat{k}^2). 
+%   n \equiv \hat{\upsilon}^2(1+\hat{k}^2). 
 % \end{align} 
@@
 % \begin{align}
-%   \left|\hat{\zeta}\right| = \sqrt{n}
+%   \left|\hat{\upsilon}\right| = \sqrt{n}
 %   \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
@@
 % \begin{align}
-%   \mathrm{max}\left(\hat{\zeta}\right) =\frac{\sqrt{L}|h
+%   \mathrm{max}\left(\hat{\upsilon}\right) =\frac{\sqrt{L}|h
 %     q|}{2\sqrt{1+|h q|}}. \label{iso_max_growth}
@@
 \subsection{Limiting form with real frequencies and $\khat^2\gg 1$}\label{disp_neut_limit}
-Assuming neutral modes with real $\hat{\zeta}=\hat{\omega}$ and $\khat^2\gg 1$, the real and imaginary parts of
+Assuming neutral modes with real $\hat{\upsilon}=\hat{\omega}$ and $\khat^2\gg 1$, the real and imaginary parts of
 Eq. \ref{relax_disp} are approximately 
@@
 %  \rho_0 = 2.7\times10^{-9}
-%  \hat{\zeta}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
+%  \hat{\upsilon}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
 \rho_0 = 1.7\times10^{-9}

### 2015-04-23 linear_adia.tex
going through re-structured paper

@@
 \begin{align}
-  \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\zeta}^2 = \frac{1 +
+  \beta\equiv 0, \quad \delta v_z = 1,\quad \hat{\upsilon}^2 = \frac{1 +
   \ii h q \hat{k}}{1+\hat{k}^2}, \label{pert_basic} 
@@
   \beta \to 0 + \delta\beta,\, \delta v_z\to \delta v_z+\delta
-  v_{z1},\,\hat{\zeta} \to \hat{\zeta} + \delta\hat{\zeta}, 
+  v_{z1},\,\hat{\upsilon} \to \hat{\upsilon} + \delta\hat{\upsilon}, 
 \end{align}
@@
 \begin{align}
-  \chi \to 1 + \delta\chi = 1 + \ii \hat{\zeta}\left(\gamma-1\right)\delta\beta,
+  \chi \to 1 + \delta\chi = 1 + \ii \hat{\upsilon}\left(\gamma-1\right)\delta\beta,
 \end{align}
-with $\delta v_z$ and $\hat{\zeta}$ given by Eq. \ref{pert_basic}. We
+with $\delta v_z$ and $\hat{\upsilon}$ given by Eq. \ref{pert_basic}. We
 may then seek  
@@
 Eq. \ref{nearly_iso_explicit}, keeping only first order terms, and
-solve for $\delta\hat{\zeta}$ using the above expressions for $\delta\chi$
-and $\hat{\zeta}$. We find imaginary part of $\delta\hat{\zeta}$
+solve for $\delta\hat{\upsilon}$ using the above expressions for $\delta\chi$
+and $\hat{\upsilon}$. We find imaginary part of $\delta\hat{\upsilon}$
 is 
@@
 % \begin{align}
-%   0 &= 2a + 2\hat{\zeta}\left(1+\khat^2\right)\delta\hat{\zeta} +
-%   \left(\hat{\zeta}^2 -1\right)\delta\chi, \\
+%   0 &= 2a + 2\hat{\upsilon}\left(1+\khat^2\right)\delta\hat{\upsilon} +
+%   \left(\hat{\upsilon}^2 -1\right)\delta\chi, \\
 %   0&= 2a\left(1+\ii h q \khat\right) - \left(\khat^2 - \ii h
@@
 % \begin{align}\label{nearly_iso_dsig}
-%   \delta \hat{\zeta} =
+%   \delta \hat{\upsilon} =
 %   -\frac{\ii\left(\gamma-1\right)\hat{k}^2\left(\ii h
@@
 \begin{align}
-  B(\hat{\zeta}) + \lambda(\hat{\zeta}) =
-  M\left[A-2\lambda(\hat{\zeta})\right].\label{adia_disp0} 
+  B(\hat{\upsilon}) + \lambda(\hat{\upsilon}) =
+  M\left[A-2\lambda(\hat{\upsilon})\right].\label{adia_disp0} 
 \end{align}
 For ease of analysis, we re-arrange Eq. \ref{adia_disp0} and square it
-to obtain a polynomial in $\hat{\zeta}$,  
+to obtain a polynomial in $\hat{\upsilon}$,  
 \begin{align}
-  0 = \sum_{l=0}^{6}c_l\hat{\zeta}^l,\label{relax_disp}
+  0 = \sum_{l=0}^{6}c_l\hat{\upsilon}^l,\label{relax_disp}
 \end{align}
 where the coefficients $c_l$ are given in Appendix \ref{relax_coeff}.
-The dispersion relation $\hat{\zeta}=\hat{\zeta}(\khat)$ is
+The dispersion relation $\hat{\upsilon}=\hat{\upsilon}(\khat)$ is
 complicated and generally requires a numerical solution. However,
@@
 
-% At marginal stability, the eigenfrequency $\hat{\zeta}=\hat{\omega}$
+% At marginal stability, the eigenfrequency $\hat{\upsilon}=\hat{\omega}$
 % is real. The dispersion relation for the fundamental mode is then

### 2015-04-23 linear_iso.tex
going through re-structured paper

@@
 In preparation for a numerical study of the above equations, we will 
-first present a fully analytic discussion for vertically isothermal disks with
-$\Gamma=1$, so that $c_s$ is vertically constant. To this end, 
+first present a fully analytic discussion for vertically isothermal
+disks, for which $\Gamma=1$ and $c_s$ is vertically constant. To this end, 
 we introduce the following simplifications:
@@
 \item We consider eigenvalues such that
-  $|\zeta^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency 
+  $|\upsilon^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency 
   approximation
@@
 %     v_z}{dz} \notag\\
-%   % &+ \left[\zeta^2\left(\frac{k_x^2}{\Omega_K^2} +
+%   % &+ \left[\upsilon^2\left(\frac{k_x^2}{\Omega_K^2} +
 %   %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
-%   &+ \left[\frac{\zeta^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_K^2} +
+%   &+ \left[\frac{\upsilon^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_K^2} +
 %       \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
@@
 % \begin{align}
-%   \chi \equiv \frac{1-\ii\zeta t_c}{1-\ii\zeta t_c\gamma}.
+%   \chi \equiv \frac{1-\ii\upsilon t_c}{1-\ii\upsilon t_c\gamma}.
 % \end{align}
@@
 \begin{align}
-  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\zeta} =\hat{\omega} +
-  \ii\hat{\sigma}= \zeta/\Omega_K,
+  \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\upsilon} =\hat{\omega} +
+  \ii\hat{\sigma}= \upsilon/\Omega_K,
 \end{align}
@@
 %   &+
-%   \left[\hat{\zeta}^2\left(\hat{k}^2+\chi\right) +
+%   \left[\hat{\upsilon}^2\left(\hat{k}^2+\chi\right) +
 %     \left(  \chi + \ii h q\hat{k}\right)\ln\rho^{\prime\prime}\right.
@@
   &A \equiv 1 + \ii h q \hat{k},\\
-  &B \equiv \hat{\zeta}^2\left(\chi + \hat{k}^2\right) -
+  &B \equiv \hat{\upsilon}^2\left(\chi + \hat{k}^2\right) -
   \left(\chi + \ii h q \hat{k}\right),\\
@@
 \begin{align}
-\chi = \frac{1-\ii\hat{\zeta}\beta}{1-\ii\hat{\zeta}\beta\gamma}.
+\chi = \frac{1-\ii\hat{\upsilon}\beta}{1-\ii\hat{\upsilon}\beta\gamma}.
 \end{align}
@@
 % \begin{align}
-%   &\frac{\zeta^2}{c_s^2}Q + \frac{\zeta^2k_x^2}{D}W \notag\\ 
+%   &\frac{\upsilon^2}{c_s^2}Q + \frac{\upsilon^2k_x^2}{D}W \notag\\ 
 %   &=\left(\frac{\ii
@@
 %     \frac{dQ}{dz}\right) - \frac{d^2\ln{\rho}}{dz^2}\left(W-Q\right),\label{ode_w}\\
-%   &\zeta^2W - \frac{\gamma}{\Gamma}\zeta^2 Q +
-%   \frac{\ii\zeta}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
+%   &\upsilon^2W - \frac{\gamma}{\Gamma}\upsilon^2 Q +
+%   \frac{\ii\upsilon}{t_c}\left(W-\frac{Q}{\Gamma}\right)\notag\\
 %   &=c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma} - 1\right) 
@@
 % \begin{align}
-%   D \equiv \kappa^2 - \zeta^2,
+%   D \equiv \kappa^2 - \upsilon^2,
 % \end{align} 

### 2015-04-23 numerical.tex
going through re-structured paper

@@
 where $\bm{\xi}$ is the Lagrangian displacement with meridional 
-components $\xi_{x,z} = \ii\delta v_{x,z}/\zeta$. In some cases we
+components $\xi_{x,z} = \ii\delta v_{x,z}/\upsilon$. In some cases we
 impose a rigid boundary at which $\delta v_z=0$.
@@
 below.)  The `fundamental corrugation mode', marked by `f',
-corresponds to the eigenvalue with the smallest $|\zeta|$. The
+corresponds to the eigenvalue with the smallest $|\upsilon|$. The
 fundamental mode is robust  

### 2015-04-23 setup.tex
going through re-structured paper

@@
 \begin{align}
-  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left(\ii k_rr - \ii\zeta
+  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left(\ii k_rr - \ii\upsilon
       t\right)},    
 \end{align}
-and similarly for other fluid variables, where $\zeta = \omega + \ii
+and similarly for other fluid variables, where $\upsilon = \omega + \ii
 \sigma$ is a (generally) complex frequency, with $\omega$ and $\sigma$ being
@@
 % \begin{align}
-%   \ii\zeta\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
+%   \ii\upsilon\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
 %     z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
-%   \ii\zeta\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
+%   \ii\upsilon\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
 %   \delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
-%   \ii\zeta\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
+%   \ii\upsilon\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
 %   r\frac{\p\Omega}{\p z}\delta v_z,\\
-%   \ii \zeta\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
+%   \ii \upsilon\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
 %   \frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},\\
-%   \ii\zeta\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
+%   \ii\upsilon\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
 %   \left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
@@
 % ignored.       
-Then writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
+Then writing $\delta P /\rho \equiv W$, $c_s^2\delta\rho/\rho\equiv
+Q$ and making the notation change $k_r\to k_x$,  the
 linearized system of equations are 
 \begin{align}
-  \ii\zeta \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
+  \ii\upsilon \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
     v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \hat{g}_c
   \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
-  \ii\zeta \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
+  \ii\upsilon \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
   \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
-   \ii \zeta\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
+   \ii \upsilon\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
   \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
-   \ii\zeta\delta v_z &= \frac{dW}{dz} +
+   \ii\upsilon\delta v_z &= \frac{dW}{dz} +
   \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
-  \ii \zeta W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
+  \ii \upsilon W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
   c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
@@
   \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x,\label{lin_energy}
-  % -&\ii\zeta W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\zeta\frac{\gamma}{\Gamma} Q +
+  % -&\ii\upsilon W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\upsilon\frac{\gamma}{\Gamma} Q +
   % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
 \end{align}
-where we have made the notation change $k_r\to k_x$. 
+where the coefficients $\hat{g}_c=1$ are inserted to label terms
+associated with the background radial disk structure. 
 It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 

### 2015-04-23 appendix.tex
fig updates, notation updates except for sec 3

@@
 analytical discussion in \S\ref{analytical}. We shall need the
-vertical derivative of the epicyclic frequency $\kappa$. In the global
+vertical derivative of the epicyclic frequency $\kappa$. In the
+vertically isothermal global
 disk with $\Gamma=1$ we have 
@@
 adiabatic flow, even if the above conditions are satisfied. We explain this by considering vertically isothermal
-disks and setting  $\beta\to\infty$ in 
+disks and setting  $\beta\to\infty$ and hence $\chi = 1/\gamma$ in 
 Eq. \ref{vertiso_gov_nondim} % (which itself is obtained by setting
@@
 Nevertheless, we demonstrate below
-that this inconsistency is unimportant for the VSI because it occurs
+that this inconsistency is unimportant for the VSI, which occurs 
 for $\beta\ll1$.   %and small scale
@@
 We demand the series to terminate  at $l=L$, i.e. a polynomial of
-order $L$, so that the vertical kinetic energy remain bounded as 
+order $L$, so that the vertical kinetic energy density remain bounded as 
 $|\zhat|\to\infty$.  Then the eigenfrequency $\hat{\upsilon}$ is given via 

### 2015-04-23 linear_iso.tex
fig updates, notation updates except for sec 3

@@
 In preparation for a numerical study of the above equations, we will 
-first present a fully analytic discussion for vertically isothermal
+first give a fully analytic discussion for vertically isothermal
 disks, for which $\Gamma=1$ and $c_s$ is vertically constant. To this end, 
@@
 \item We consider eigenvalues such that
-  $|\upsilon^2|\ll \kappa^2,\,\Omega^2$. We call this the low frequency 
-  approximation
+  $|\upsilon^2|\ll \kappa^2,\,\Omega^2$; which correspond to inertial
+  waves \citepalias{barker15}. We call this the low frequency  
+  approximation.
 
-\item We are mainly interested in the effect of vertical shear, i.e. $q\neq
-  0$ for vertically isothermal disks. This is already captured
+\item We are mainly interested in the effect of vertical shear, 
+%  i.e. $q\neq
+%  0$ for vertically isothermal disks. 
+  which is already captured
   explicitly in the linearized equations through the term
@@
   the VGSB, where we show that this approximation is valid provided we 
-  consider consider fast thermal relaxation times
+  consider fast cooling times
   ($\beta\lesssim 1$) and perturbations with small radial lengthscales

### 2015-04-23 numerical.tex
fig updates, notation updates except for sec 3

@@
 $\zmax$, so that vertically isothermal disks are in fact suitable for
-the purpose of this paper.
+the purpose of characterizing the thermodynamic dependence of the
+VSI.
 
@@
     model as a function of the thermal relaxation timescale
-    $\beta$. The vertical line is a conjectured 
-    critical thermal timescale given by
-    $ h|s|/(\gamma-\Gamma)=0.158$.  
+    $\beta$. The vertical dashed line is the critical thermal
+    timescale given by Eq. \ref{iso_vsi_cond}. A conjectured critical thermal
+    timescale accounting for a non-isothermal background, given by 
+    $\beta_\mathrm{crit,noniso}= h|s|/(\gamma-\Gamma)=0.158$,  
+    is also shown as the vertical dotted line. 
     \label{bcrit_compare2}}   

### 2015-04-23 setup.tex
notation change: g_c to zeta
temp commit to do a notation merge

--- 2015-04-23 setup.tex
@@
 \section{Linear problem}\label{linear} 
-In this section we present the two sets of equations used to study the
-axisymmetric linear stability of the above equilibria. We first give
-the linearized equations in the radially-local approximation, which
-are solved in our numerical study. From this set, we further derive a
-simplified equation for vertically isothermal disks used for analytic
-discussion in the next section.
+This section presents the two sets of equations we use to study the
+linear development of the VSI.  Both sets are radially local and vertically global 
+with a finite cooling time.
+The first set, presented in \S\ref{sec:radlocal}, is more general and used 
+for precise numerical calculations.  The second set, in \S\ref{sec:simplified}, makes
+additional approximations  about disk structure and wave frequency.  This simplified
+set is used to obtain analytic results.
 
-\subsection{Radially local approximation}
-We consider axisymmetric perturbations to the above equilibria in the
-form 
+\subsection{Radially local approximation}\label{sec:radlocal}
+We consider axisymmetric perturbations to the above disk equilibria.    
+The growth of the VSI is strongest for short radial wavelengths, as compared
+ to the disk radius \citepalias{nelson13,barker15}.  We thus perform a
+ standard two step process to obtain linearized equations in the radially local 
+ approximation (e.g., \citealp{goldreich67}, who also consider vertically local perturbations).
+
+ We first expand all fluid variables 
+ into the equilibrium value plus a perturbation denoted by $\delta$, e.g.\ 
+ $\delta \rho$ for the perturbed density, and drop all non-linear perturbations.  Second, we perform
+ a Taylor expansion about a fiducial radius $r_0$ with the local radial coordinate
+ $x = r - r_0$, keeping only the leading order terms in $x/r_0$.  We also relabel 
+ (trivially for axisymmetric perturbations) the azimuthal direction as $y$.
+
+ Perturbations take the form of a radial plane wave with arbitrary vertical dependence, 
+ i.e.\
  \begin{align}
-  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left(\ii k_rr - \ii\zeta
+  \delta\rho \rightarrow \delta\rho(z)\exp{\left(\ii k_x x - \ii\zeta
       t\right)},    
 \end{align}
-and similarly for other fluid variables, where $\zeta = \omega + \ii
-\sigma$ is a (generally) complex frequency, with $\omega$ and $\sigma$ being
-the real frequency and growth rate, respectively, and $k_r$ is a real 
-radial wavenumber. We take $k_r>0$ without loss of generality.
+where $\zeta = \omega + \ii \sigma$ is a (generally) complex frequency, 
+with $\omega$ and $\sigma$ being
+the real frequency and growth rate, respectively, and $k_x$ is a real 
+radial wavenumber.  Henceforth all unperturbed fluid variables, including 
+gradients suc as $\p \Omega/ \p z$, refer to  equilibrium values at $(r_0, z)$.
 
-The VSI is expected to operate for perturbations with small radial
-wavelengths \citepalias{nelson13,barker15}. We thus linearize 
-Eq. \ref{full_mass}---\ref{full_energy} assuming perturbations 
-in the above form with radially slowly-varying
-amplitudes and $k_r r \gg 1$, so that the radial dependence of $\delta\rho(r,z)$, etc.,
-may be neglected, as well as curvature terms. Henceforth we are
-considering solutions at some fiducial radius $r=r_0$.  
+We further define the perturbation variables  $W \equiv \delta P /\rho$ 
+and $Q \equiv c_s^2\delta\rho/\rho$.  With this procedure the linearized
+system of equations are
+\begin{align}
+  \ii\zeta \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
+    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \ksi
+  \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
+  \ii\zeta \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
+  \ksi\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
+   \ii \zeta\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
+  \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
+   \ii\zeta\delta v_z &= \frac{dW}{dz} +
+  \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
+  \ii \zeta W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
+  c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
+      v_z}{dz}\right) \notag\\
+  &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
+  \ksi\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x.\label{lin_energy}
+  % -&\ii\zeta W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\zeta\frac{\gamma}{\Gamma} Q +
+  % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
+\end{align}
+The above equations are a set of ODEs (ordinary differential equations) in $z$.  Solutions to these 
+are presented in \S\ref{numerical} and \S\ref{application}.   The coefficient $ \ksi = 1$ is 
+introduced simply to label terms with radial gradients of the equilibrium state, 
+which again are evaluated at $(r_o, z)$.   We will consider the effects of ignoring these radial gradients below.
+
+%We consider axisymmetric perturbations to the above equilibria in the
+%form 
+%\begin{align}
+%  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left(\ii k_rr - \ii\zeta
+%      t\right)},    
+%\end{align}
+%and similarly for other fluid variables, where $\zeta = \omega + \ii
+%\sigma$ is a (generally) complex frequency, with $\omega$ and $\sigma$ being
+%the real frequency and growth rate, respectively, and $k_r$ is a real 
+%radial wavenumber. We take $k_r>0$ without loss of generality.
+
+%The VSI is expected to operate for perturbations with small radial
+%wavelengths \citepalias{nelson13,barker15}. We thus linearize 
+%Eq. \ref{full_mass}---\ref{full_energy} assuming perturbations 
+%in the above form with radially slowly-varying
+%amplitudes and $k_r r \gg 1$, so that the radial dependence of $\delta\rho(r,z)$, etc.,
+%may be neglected, as well as curvature terms. Henceforth we are
+%considering solutions at some fiducial radius $r=r_0$.  
 %\citetalias{barker15} have validated this 
@@
 
-In order to emphasize the radially local approximation, we introduce
-the local Cartesian co-ordinates $(x,y,z)$ corresponding to the the
-local $(r,\phi,z)$ directions, respectively.  
+%In order to emphasize the radially local approximation, we introduce
+%the local Cartesian co-ordinates $(x,y,z)$ corresponding to the the
+%local $(r,\phi,z)$ directions, respectively.  
 %We thus linearize 
@@
 %   \ii\zeta\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
-%     z}\left(\rho\delta v_z\right)  + \hat{g}_c\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
+%     z}\left(\rho\delta v_z\right)  + \ksi\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
 %   \ii\zeta\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
-%   \delta v_\phi - \hat{g}_c\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
+%   \delta v_\phi - \ksi\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
 %   \ii\zeta\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
@@
 %   &+ \frac{1}{t_c}\left(\delta P -  
-%     \frac{P}{\rho}\delta\rho\right) + \hat{g}_c \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
+%     \frac{P}{\rho}\delta\rho\right) + \ksi \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
 % \end{align}
-% where the coefficients $\hat{g}_c=1$ are inserted to label terms
+% where the coefficients $\zeta=1$ are inserted to label terms
 % associated with the background radial disk structure. 
@@
 % ignored.       
-Then writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
-linearized system of equations are 
-\begin{align}
-  \ii\zeta \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
-    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \hat{g}_c
-  \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
-  \ii\zeta \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
-  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
-   \ii \zeta\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
-  \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
-   \ii\zeta\delta v_z &= \frac{dW}{dz} +
-  \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
-  \ii \zeta W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
-  c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
-      v_z}{dz}\right) \notag\\
-  &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
-  \hat{g}_c\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x,\label{lin_energy}
-  % -&\ii\zeta W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\zeta\frac{\gamma}{\Gamma} Q +
-  % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
-\end{align}
-where we have made the notation change $k_r\to k_x$. 
-It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
-\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$
-using their expressions in the global disk given in \S\ref{setup}. Thus, they are functions of $z$ only. 
-Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
-study, but for analytical discussion
-we will make further simplifications, as discussed next.
+%Then writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
+%linearized system of equations are 
+%where we have made the notation change $k_r\to k_x$. 
+%It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 
+%\p_z\Omega,\, \p_r\ln{\rho} $ and $ \p_r P$ are evaluated at $r=r_0$
+%using their expressions in the global disk given in \S\ref{setup}. Thus, they are functions of $z$ only. 
+%Eq. \ref{lin_mass}---\ref{lin_energy} will form the basis of our numerical
+%study, but for analytical discussion
+%we will make further simplifications, as discussed next.

### 2015-04-23 appendix.tex

@@
 In the vertically global shearing box (VGSB) approximation, we set
-$\hat{g}_c=0$ in Eq. \ref{lin_mass} --- \ref{lin_energy}. We then eliminate
+$\zeta=0$ in Eq. \ref{lin_mass} --- \ref{lin_energy}. We then eliminate
 the horizonatal velocity perturbations ($\delta v_x,\, \delta v_y$) to
@@
 (Eq. \ref{lin_mass}, \ref{lin_vx} and \ref{lin_energy}) as the terms
-proportional to $\hat{g}_c$. Nominally $\hat{g}_c=1$.      
+proportional to $\zeta$. Nominally $\zeta=1$.      
 In the VGSB (and SSB) these terms are neglected by setting
-$\hat{g}_c=0$. For a power-law disk, these radial gradients  
+$\zeta=0$. For a power-law disk, these radial gradients  
 are $O(r^{-1})$, and they appear in comparison with terms of
@@
 %are ignored elsewhere in the VGSB (which is done by setting
-%$\hat{g}_c=0$).
+%$\zeta=0$).
 
@@
 %     v_z}{dz}\right) + \left[\ii k_xc_s^2(r_0,z)
-%   \frac{\gamma}{\Gamma} + \hat{g}_c\left.\frac{1}{\rho}\frac{\p P}{\p
+%   \frac{\gamma}{\Gamma} + \zeta\left.\frac{1}{\rho}\frac{\p P}{\p
 %       r}\right|_0\right]\delta v_x  +
 % \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right),\label{gcorr_terms1}\\
-% \ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_x + \hat{g}_c
+% \ii \sigma Q &= c_s^2(r_0,z)\left(\ii k_x + \zeta
 %   \left.\frac{\p\ln{\rho}}{\p r}\right|_0\right)\delta v_x + c _s^2(r_0, z)\left( \left.\frac{\p\ln\rho}{\p
@@
 % \ii \sigma \delta v_x & = \ii k_x W  -
-% \hat{g}_c\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
+% \zeta\frac{1}{c_s^2(r_0,z)}\left.\frac{1}{\rho}\frac{\p P}{\p
 %   r}\right|_0Q - 2\Omega(r_0,z)\delta v_y,\label{gcorr_terms3}
@@
 % where subscript $0$ denotes evaluation at $r=r_0$. We recall terms
-% proportional to $\hat{g}_c$ are associated with the background density
-% and pressure radial structures, and nominally $\hat{g}_c=1$.
+% proportional to $\zeta$ are associated with the background density
+% and pressure radial structures, and nominally $\zeta=1$.
 
@@
 Eq. \ref{vertiso_gov_nondim} % (which itself is obtained by setting
-% $\hat{g}_c=0$ in Eq. \ref{lin_mass}---\ref{lin_energy})
+% $\zeta=0$ in Eq. \ref{lin_mass}---\ref{lin_energy})
 to give 
@@
 In  Fig. \ref{gcorr_compare}, we plot the effect of the global radial gradient terms
-proportional to $\hat{g}_c$ in Eq. \ref{lin_mass}---\ref{lin_energy}
+proportional to $\zeta$ in Eq. \ref{lin_mass}---\ref{lin_energy}
 by calculating the
 fundamental VSI growth rates using three approaches. We compute growth rates from the dispersion
-relation Eq. \ref{relax_disp}, which assumes $\hat{g}_c=0$;  from
-Eq. \ref{lin_mass}---\ref{lin_energy} with $\hat{g}_c=0$; and from
-Eq. \ref{lin_mass}---\ref{lin_energy}  with $\hat{g}_c=1$.
+relation Eq. \ref{relax_disp}, which assumes $\zeta=0$;  from
+Eq. \ref{lin_mass}---\ref{lin_energy} with $\zeta=0$; and from
+Eq. \ref{lin_mass}---\ref{lin_energy}  with $\zeta=1$.
 
@@
     the `Numerical (VGSB)' curve is obtained by numerically solving
-    the linear problem, Eq. \ref{lin_mass}---\ref{lin_energy}, with $\hat{g}_c=0$
+    the linear problem, Eq. \ref{lin_mass}---\ref{lin_energy}, with $\zeta=0$
     (which defines the linear VGSB equations); and the `Numerical
     (Full)' curve is obtained by numerically solving 
-    Eq. \ref{lin_mass}---\ref{lin_energy}  with $\hat{g}_c=1$, which
+    Eq. \ref{lin_mass}---\ref{lin_energy}  with $\zeta=1$, which
     accounts for the radial disk structure.  

### 2015-04-23 application.tex

@@
 However, disturbances with large $\khat$ may be subject to viscous
-decay in a real disk. The viscous timescale for a perturbation lengthscale 
-$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $, where
-$\alpha$ is the standard alpha viscosity parameter
-\citep{shakura73}. Setting $t_\mathrm{visc}$ to be longer than the
+decay in a real disk. Adopting the standard precription for the
+kinematic viscosity $\nu = \alpha c_s H$ \citep{shakura73}, where
+$\alpha$ is the dimensionless viscosity parameter, 
+the viscous timescale for a perturbation lengthscale 
+$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $. 
+Setting $t_\mathrm{visc}$ to be longer than the
 minimum VSI growth timescale $1/h\Omega$ implies 

### 2015-04-23 linear_adia.tex

@@
 Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} is a pair of
-simultaneous equations for $\hat{\omega}$ and $\beta$ for marginal
-stability. (It is important to note that marginal stability is an
-assumption.) If $\beta\ll 1$ and the disk is sufficiently thin so that  
+simultaneous equations for $\hat{\omega}$ and $\beta$ \emph{assuming} marginal 
+stability. %  (It is important to note that marginal stability is an
+% assumption.)
+If $\beta\ll 1$ and the disk is sufficiently thin so that  
 $|h q \khat |^2\lesssim 1$, 

### 2015-04-23 setup.tex

@@
  Perturbations take the form of a radial plane wave with arbitrary vertical dependence, 
- i.e.\
+ e.g.\
  \begin{align}
-  \delta\rho \rightarrow \delta\rho(z)\exp{\left(\ii k_x x - \ii\zeta
+  \delta\rho \rightarrow \delta\rho(z)\exp{\left(\ii k_x x - \ii\upsilon
       t\right)},    
@@
 radial wavenumber.  Henceforth all unperturbed fluid variables, including 
-gradients suc as $\p \Omega/ \p z$, refer to  equilibrium values at $(r_0, z)$.
+gradients such as $\p \Omega/ \p z$, refer to  equilibrium values at $(r_0, z)$.
 
@@
 \begin{align}
-  \ii\zeta \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
-    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \ksi
+  \ii\upsilon \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
+    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \zeta
   \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
-  \ii\zeta \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
-  \ksi\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
-   \ii \zeta\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
+  \ii\upsilon \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
+  \zeta\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
+   \ii \upsilon\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
   \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
-   \ii\zeta\delta v_z &= \frac{dW}{dz} +
+   \ii\upsilon\delta v_z &= \frac{dW}{dz} +
   \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
-  \ii \zeta W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
+  \ii \upsilon W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
   c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
@@
   &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
-  \ksi\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x.\label{lin_energy}
-  % -&\ii\zeta W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\zeta\frac{\gamma}{\Gamma} Q +
+  \zeta\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x.\label{lin_energy}
+  % -&\ii\upsilon W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\upsilon\frac{\gamma}{\Gamma} Q +
   % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
@@
 The above equations are a set of ODEs (ordinary differential equations) in $z$.  Solutions to these 
-are presented in \S\ref{numerical} and \S\ref{application}.   The coefficient $ \ksi = 1$ is 
-introduced simply to label terms with radial gradients of the equilibrium state, 
+are presented in \S\ref{numerical} and \S\ref{application}.   The coefficient $ \zeta = 1$ is 
+introduced simply to label terms containing radial gradients of the equilibrium state, 
 which again are evaluated at $(r_o, z)$.   We will consider the effects of ignoring these radial gradients below.
@@
 %\begin{align}
-%  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left(\ii k_rr - \ii\zeta
+%  \rho \to \rho(r, z) + \delta\rho(r,z)\exp{\left(\ii k_rr - \ii\upsilon
 %      t\right)},    
 %\end{align}
-%and similarly for other fluid variables, where $\zeta = \omega + \ii
+%and similarly for other fluid variables, where $\upsilon = \omega + \ii
 %\sigma$ is a (generally) complex frequency, with $\omega$ and $\sigma$ being
@@
 % \begin{align}
-%   \ii\zeta\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
-%     z}\left(\rho\delta v_z\right)  + \ksi\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
-%   \ii\zeta\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
-%   \delta v_\phi - \ksi\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
-%   \ii\zeta\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
+%   \ii\upsilon\delta\rho &= \frac{\rho}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p}{\p
+%     z}\left(\rho\delta v_z\right)  + \zeta\frac{\p\rho}{\p r}\delta v_r,\label{full_mass1}\\
+%   \ii\upsilon\delta v_r &= \frac{1}{\rho}\frac{\p\delta P}{\p r} - 2\Omega
+%   \delta v_\phi - \zeta\frac{\delta\rho}{\rho^2}\frac{\p P}{\p r},\\ 
+%   \ii\upsilon\delta v_\phi  &= \frac{\kappa^2}{2\Omega}\delta v_r +
 %   r\frac{\p\Omega}{\p z}\delta v_z,\\
-%   \ii \zeta\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
+%   \ii \upsilon\delta v_z &= \frac{1}{\rho}\frac{\p\delta P}{\p z} -
 %   \frac{\delta \rho}{\rho^2}\frac{\p P}{\p z},\\
-%   \ii\zeta\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
+%   \ii\upsilon\delta P &=\frac{\p P}{\p z}\delta v_z + \gamma P
 %   \left[\frac{1}{r}\frac{\p}{\p r}\left(r\delta v_r\right) + \frac{\p
@@
 %   &+ \frac{1}{t_c}\left(\delta P -  
-%     \frac{P}{\rho}\delta\rho\right) + \ksi \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
+%     \frac{P}{\rho}\delta\rho\right) + \zeta \frac{\p P}{\p r}\delta v_r,\label{full_energy1}
 % \end{align}
@@
 % ignored.       
-%Then writing $\delta P /\rho \equiv W$ and $c_s^2\delta\rho/\rho\equiv Q$, the
+%Then writing $\delta P /\rho \equiv W$, $c_s^2\delta\rho/\rho\equiv
+%Q$ and making the notation change $k_r\to k_x$,  the
 %linearized system of equations are 
-%where we have made the notation change $k_r\to k_x$. 
+%\begin{align}
+%  \ii\upsilon \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
+%    v_z}{dz} + \frac{\p\ln{\rho}}{\p z}\delta v_z + \zeta
+%  \frac{\p\ln{\rho}}{\p r} \delta v_x,\label{lin_mass}\\
+%  \ii\upsilon \delta v_x  &= \ii k_x W - 2\Omega\delta v_y -
+%  \zeta\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
+%   \ii \upsilon\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
+%  \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
+%   \ii\upsilon\delta v_z &= \frac{dW}{dz} +
+%  \frac{\p\ln{\rho}}{\p z}\left(W-Q\right), \label{lin_vz}\\
+%  \ii \upsilon W &= c_s^2\frac{\p\ln{\rho}}{\p z}\delta v_z +
+%  c_s^2\frac{\gamma}{\Gamma} \left(\ii k_x\delta v_x + \frac{d\delta
+%      v_z}{dz}\right) \notag\\
+%  &+ \frac{1}{t_c}\left(W - \frac{Q}{\Gamma}\right) +
+%  \zeta\frac{1}{\rho}\frac{\p P}{\p r}\delta v_x,\label{lin_energy}
+%  % -&\ii\upsilon W + \frac{1}{t_c}\left(W-\frac{Q}{\Gamma}\right) = -\ii\upsilon\frac{\gamma}{\Gamma} Q +
+%  % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
+%\end{align}
+%where the coefficients $\zeta=1$ are inserted to label terms
+%associated with the background radial disk structure. 
 %It is understood that the quantities $\rho,\,c_s^2,\,\Omega,\kappa,\, 

### 2015-04-23 summary.tex
messed up notation change. need to fix by inspection.

@@
 $\Omega=\Omega(r,z)$ in a baroclinic disk. Vertical shear is generally
-expected in astrophysical disks due to the radial variation in the
-disk temperature, or more generally the entropy. However, 
-astrophysical disks are also typically sub-adiabatically stratified. 
-In this case, the VSI requires a short thermal timescale $t_c$ in
-order to operate.
+expected in astrophysical disks due to baroclinicity. This may arise 
+from a radial variation in the disk temperature, or more generally in
+the disk entropy. However, astrophysical disks are also typically
+sub-adiabatically stratified for stability against vertical
+convection. In this case, the VSI requires a short thermal or cooling
+timescale $t_c$ in order to operate.
 
@@
 irradiated by its central star \citep{chiang97}. It also permits a
-largely analytical discussion. Our analysis is then an extension of the
+simplified analytical discussion by extending the
 linear models considered in \cite{lubow93} by adding vertical shear
@@
 \begin{align*}
-  t_\mathrm{crit} = \frac{ h |q|}{\gamma-1}\Omega_k^{-1}, 
+  t_\mathrm{crit} = \frac{ h |q|}{\gamma-1}\Omega_K^{-1}, 
 \end{align*}
@@
 radii, but only on radial lengthscales signficantly shorter
-than the disk scaleheight, and such perturbations may
+than the disk scaleheight. However, such small-scale perturbations may 
 be subject to viscous decay in a real disk.
@@
 \begin{align}
-  \nu_\mathrm{conv,max} = -\frac{N^2_r}{4\Omega}\label{max_conv_gen}
+  \sigma_\mathrm{conv,max} = -\frac{N^2_r}{4\Omega}\label{max_conv_gen}
 \end{align}
@@
 %\begin{align}
-%  \nu_\mathrm{conv,max} =
-%  -\frac{ h^2\Omega_k}{4\gamma\Gamma}\left(p+q\right)
+%  \sigma_\mathrm{conv,max} =
+%  -\frac{ h^2\Omega_K}{4\gamma\Gamma}\left(p+q\right)
 %  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
@@
 For more general disk structures convective overstability is
-possible in principle. The models adopted for our numerical study in
+possible, in principle. The models adopted for our numerical study in
 \S\ref{numerical} do satisfy $N_r^2<0$, but convective overstability
@@
 operates, disk self-gravity (SG) may play a role. Vertical
-self-gravity may favor the VSI by increasing $|\p_z\Omega|$ through
+SG may favor the VSI by increasing $|\p_z\Omega|$ through
 enhancing the density stratification 
@@
 
-We have neglected viscosity. \citetalias{nelson13} find that an alpha
-viscosity $\lesssim O(10^{-4})$ is required for the
-VSI. \citetalias{barker15} found that inclusion of viscosity is needed for 
-convergence when studying surface modes. %but not body modes? 
+We have neglected viscosity. \citetalias{nelson13} find that a
+dimensionless viscosity of $\alpha\lesssim O(10^{-4})$ is required for the
+VSI; while \citetalias{barker15} find that inclusion of viscosity is needed for 
+convergence when studying surface modes in radially global disk
+models. %but not body modes? 
 Viscosity may also affect the
-thermal timescale requirement of the VSI, because viscous heating
+thermodynamic requirement for the VSI, because viscous heating
 needs to be self-consistently included in the energy equation. This

### 2015-04-23 application.tex
introduce alias for Omega_K,  \OmK, think I got all...

@@
 \begin{align}\label{tc_diff_cool} 
-  t_\mathrm{diff} = \frac{1}{\eta k_x^2}.%\equiv \frac{\Omega_K^{-1}}{\hat{\eta}\khat^2}, 
+  t_\mathrm{diff} = \frac{1}{\eta k_x^2}.%\equiv \frac{\OmK^{-1}}{\hat{\eta}\khat^2}, 
 \end{align}
@@
 \begin{align}\label{real_beta}
-  \beta(r,z;\khat) \equiv t_c\Omega_K =
+  \beta(r,z;\khat) \equiv t_c\OmK =
   % \frac{1}{\hat{\eta}}\left[\frac{l_\mathrm{rad}^2}{3H^2}
   %   + \frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right]  \simeq
-  \frac{\Sigma^2\Omega_K}{\eta\rho^2}\left[\frac{1}{3\kappa_d^2\Sigma^2} 
+  \frac{\Sigma^2\OmK}{\eta\rho^2}\left[\frac{1}{3\kappa_d^2\Sigma^2} 
     + \frac{\hat{\rho}^2(z)}{2\pi \khat^2}\right].
@@
 wavenumbers, in units of the local orbital time $P_\mathrm{orb}\equiv
-2\pi/\Omega_K$. In fact, we either only find the fundamental mode, or 
+2\pi/\OmK$. In fact, we either only find the fundamental mode, or 
 that it is the most unstable, except for the appearance of some 

### 2015-04-23 linear_iso.tex
notation change: g_c to zeta
temp commit to do a notation merge
introduce alias for Omega_K,  \OmK, think I got all...

@@
-\subsection{Reduced equation for vertically isothermal disks}
+\subsection{Reduced equation for vertically isothermal disks}\label{sec:simplified}
 %We begin with a discussion of vertically isothermal disks
@@
   Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
-  $\hat{g}_c=0$. However, the global radial temperature 
+  $\zeta=0$. However, the global radial temperature 
   gradient in the expression for the 
@@
   $\kappa$ elsewhere, and replace them with
-  $\Omega_K$.  We call this the Keplerian approximation.
+  $\OmK$.  We call this the Keplerian approximation.
 
@@
 %   0 =& \frac{d^2\delta v_z}{dz^2} + \left(1 + \frac{\ii k_x c_s^2
-%       q}{\Omega_K^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
+%       q}{\OmK^2r}\right)\frac{d\ln\rho}{dz}\frac{d\delta
 %     v_z}{dz} \notag\\
-%   % &+ \left[\upsilon^2\left(\frac{k_x^2}{\Omega_K^2} +
+%   % &+ \left[\upsilon^2\left(\frac{k_x^2}{\OmK^2} +
 %   %     \frac{\chi}{c_s^2}\right) + \left(\chi + \frac{\ii k_x c_s^2
-%   &+ \left[\frac{\upsilon^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\Omega_K^2} +
+%   &+ \left[\frac{\upsilon^2}{c_s^2}\left(\frac{k_x^2c_s^2}{\OmK^2} +
 %       \chi\right) + \left(\chi + \frac{\ii k_x c_s^2
-%         q}{\Omega_K^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
+%         q}{\OmK^2r}\right)\frac{d^2\ln\rho}{dz^2}\right. \notag\\
 %   &\phantom{+=}\left.-
-%     \frac{c_s^2}{\Omega_K^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
+%     \frac{c_s^2}{\OmK^2}\left(\frac{d\ln\rho}{dz}\right)^2\left(k_x^2 -
 %       \frac{\ii k_x q}{r}\right)
@@
   \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\upsilon} =\hat{\omega} +
-  \ii\hat{\sigma}= \upsilon/\Omega_K,
+  \ii\hat{\sigma}= \upsilon/\OmK,
 \end{align}
@@
 % Eq. \ref{lin_mass}---\ref{lin_energy} by setting 
-% $\hat{g}_c=0$. The resulting problem is very similar 
+% $\zeta=0$. The resulting problem is very similar 
 % to that for local axisymmetric waves in accretion disks already
@@
 
-% After setting $\hat{g}_c=0$ in Eq. \ref{lin_mass}, \ref{lin_vx} and
+% After setting $\zeta=0$ in Eq. \ref{lin_mass}, \ref{lin_vx} and
 % \ref{lin_energy}, we can eliminate variables 
@@
 % gradient terms are ignored elsewhere in the VGSB (which is done by
-% setting $\hat{g}_c=0$).
+% setting $\zeta=0$).
 

### 2015-04-23 numerical_lowfreq.tex
introduce alias for Omega_K,  \OmK, think I got all...

\section{Numerical solutions in the low frequency limit}
In this section we numerically solve the linear stability problem.  
We relax the thin-disk assumption, using the full, non-Keplerian
expressions of $\rho$, $\Omega$ and $\kappa$, so that we may consider
vertical domains up to $z\sim r$. However, we retain the low-frequency
approximation $|\sigma^2|\ll\kappa^2$ to obtain a linear eigenvalue
problem, which limits us to consider large radial wavenumbers
$k_x$. We will relax this in the next section when considering
moderate values of $k_x$. % This level of approximation lies in between
% the full problem Eq. \ref{iso_ode}, solved by \cite{mcnally14}, and
% the thin-disk limit Eq. \ref{iso_ode3}, solved by \cite{nelson13}.

%We will solve different versions of the linearized equations depending
%on the disk properties, but in all cases we use

\subsection{Vertically isothermal disks with isothermal
  perturbations}\label{vertiso_pertiso} 
We first validate the low-frequency approximation by re-producing
results from \cite{mcnally14} who solved the full problem for
vertically isothermal disks ($\Gamma=1$) subject to isothermal
perturbations ($\gamma=1$). The parameters are $q=-1$,
$p=-1.5$, $\epsilon=0.1$ and $k_x = 200\pi/r$ (or $\hat{k} = 20\pi$),
with solid vertical boundaries.

For this problem we solve Eq. \ref{iso_ode}, with $D$ replaced by
$\kappa^2$, using a pseudo-spectral
method by expanding $W$ in Chebyshev polynomials $T_l$ up to $l=512$
and discretizing the equation on a grid with
$\zmax=10\epsilon r$. This procedure converts the linear problem
to a standard matrix eigenvalue problem.

Fig. \ref{lowfreq_eigen} shows the eigenvalues $\sigma = \omega +
\ii\nu$. This plot is effectively identical to that in
\cite{mcnally14}. The fundamental VSI mode, with one node in $W$, has
the smallest $|\sigma|$ and is shown in Fig. \ref{lowfreq_eigenfunc}.  
The expected growth rate from Eq. \ref{simple_growth} with 
$L=1$ is  
\begin{align*}
  \nu = 0.2606\epsilon\OmK,
\end{align*}
as obtained numerically. This agreement is surpsingly good, given that
Eq. \ref{simple_growth} assumes the thin-disk approximation and
imposes a different boundary condition to that in the numerical
calculation. The eigenfunction $W$ is qualitatively proportional to
$z$, as expected for the fundamental mode according to explicit
solutions discussed in \S\ref{iso_poly}.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
  \caption{Eigenvalues in the low-frequency approximation for the
    vertical shear instability in a vertically isothermal disk evolved
    isothermally ($\gamma=\Gamma=1$). The disk parameters are $q=-1$,
    $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
    wavenumber is $k_x=200\pi/r$. This is the set up considered in
    \cite{mcnally14}. \label{lowfreq_eigen}
  }
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvector_iso}
  \caption{Eigenfunction of the fundamental VSI,
    corresponding to the bottom-left eigenvalue displayed in
    Fig. \ref{lowfreq_eigen} (smallest $|\sigma|$). The
    solid and dashed lines are $\real W$ and $\imag W$, respectively. 
    \label{lowfreq_eigenfunc}
  }
\end{figure}

\subsection{Stabilization of the VSI by a positive vertical entropy
  gradient}
We demonstrate the (strong) stabilizing effect of a  
positive vertical entropy gradient. We solve the more general form of
the linear problem, Eq. \ref{ode_w}---\ref{ode_Q}, again with the
replacement $D\to\kappa^2$. We consider an effectively vertically
isothermal disk by setting $\Gamma=1.011$ with adiabatic index
$\gamma=2$. We also set $\beta=10^9$ so the cooling term is
effectively switched off. Other disk and perturbation parameters are
the same as in \S\ref{vertiso_pertiso}.

We use the same pseudo-spectral method as before. Here we set
$\zmax = 2\epsilon r$ and impose a free surface. This boundary
condition was found to be most convenient for numerical
implementation. Since we expect the perturbations to decay rapidly
away from the midplane (see \S\ref{analytic_adia}), we do not expect 
boundary conditions to play an important role, provided that $\zmax$ is
sufficiently large.

Fig. \ref{lowfreq_eigen_adia} show the eigenvalues for this
problem. The growth rates are about an order of magnitude smaller than
the previous case. The fundamental VSI mode is that with
$\mathrm{max}|\nu|$. Its expected and numerically-calculated growth
rates compares well, with 
\begin{align*}
  \nu = 0.03751 \epsilon \OmK &\quad \text{(from
    Eq. \ref{gam2_growth_rate})},\\
  \nu = 0.03682 \epsilon \OmK &\quad \text{(numerical)}.
\end{align*}
The fundamental mode is plotted in Fig. \ref{lowfreq_eigenfunc_adia},
and clearly show that perturbations are rapidly stabilized away from
the midplane. For the adopted disk and perturbation parameters 
($|q|=1$ and $\hat{k} = 20\pi$), Eq. \ref{gam2_alpha} gives $\real{\alpha}\simeq - 43.8$,
corresponding to a characteristic decay lengthscale of $\simeq
0.15\epsilon r$, which is consistent with that observed in 
Fig. \ref{lowfreq_eigenfunc_adia}.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvalues_adia}
  \caption{Eigenvalues in the low-frequency approximation for the
    vertical shear instability in a nearly-vertically isothermal disk
    ($\Gamma=1.011$) evolved adiabatically with $\gamma=2$ and
    $\beta=10^9$. The disk
    parameters are $q=-1$, 
    $p=-1.5$ and $\epsilon=0.1$, while the perturbation radial
    wavenumber is $k_x=200\pi/r$. \label{lowfreq_eigen_adia}
  }
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvectorvz_adia}
  \caption{Fundamental VSI mode in a nearly-vertically
    isothermal disk ($\Gamma=1.011$) evolved adiabatically
    ($\gamma=2$, $\beta=10^9$). This eigenfunction 
    corresponds to the top-left eigenvalue displayed in 
    Fig. \ref{lowfreq_eigen_adia} (largest $|\nu|$).  
    The solid and dashed lines are $\real \delta v_z$ and $\imag
    \delta v_z$, respectively.  
    \label{lowfreq_eigenfunc_adia}
  }
\end{figure}

\subsection{Effect of cooling}

### 2015-04-23 paper.tex
introduce alias for Omega_K,  \OmK, think I got all...

@@
 \begin{align}
-  t _c \ll \Omega_K^{-1}. 
+  t _c \ll \OmK^{-1}. 
 \end{align}
-where $\Omega_K$ is the Keplerian frequency. 
+where $\OmK$ is the Keplerian frequency. 
 To see why this should be expected, we need to compare vertical shear
@@
 \begin{align}\label{vshear_thin}
-  r \frac{\p \Omega^2 }{\p z} \simeq  h q \Omega_K^2\left(\frac{z}{H}\right),
+  r \frac{\p \Omega^2 }{\p z} \simeq  h q \OmK^2\left(\frac{z}{H}\right),
 \end{align}
@@
   N_z^2 \simeq \left(\frac{\gamma-1}{\gamma}\right) \left(\frac{z}{H}\right)^2
-  \Omega_K^2.  
+  \OmK^2.  
 \end{align}
@@
 which is destabilizing. In a thin disk $ h \ll 1$, so the destabilizing effect of 
-vertical shear is rather weak because $|r\p_z\Omega|\sim h\Omega_K
-\ll \Omega_K$. By comparison, the buoyancy frequency $N_z\sim
-\Omega_K$. It follows that $t_c\lesssim O( h\Omega_K^{-1})$ is
+vertical shear is rather weak because $|r\p_z\Omega|\sim h\OmK
+\ll \OmK$. By comparison, the buoyancy frequency $N_z\sim
+\OmK$. It follows that $t_c\lesssim O( h\OmK^{-1})$ is
 required for instability.  \emph{In a thin  
@@
 \begin{align}\label{prelim_bcrit}
-  (\gamma - 1)t_c <  h |q| \Omega_K^{-1}
+  (\gamma - 1)t_c <  h |q| \OmK^{-1}
 \end{align}   

### 2015-04-23 setup.tex
introduce alias for Omega_K,  \OmK, think I got all...

@@
 \begin{align}\label{beta_def}
-  \beta \equiv \Omega_k t_c
+  \beta \equiv \OmK t_c
 \end{align}
-relative to the Keplerian frequency, $\Omega_k=\sqrt{GM_*/r^3}$.
+relative to the Keplerian frequency, $\OmK=\sqrt{GM_*/r^3}$.
 
@@
 to parametrize the magnitude of the sound-speed, where
-$v_k=r\Omega_k$ is the Keplerian azimuthal velocity and
+$v_k=r\OmK$ is the Keplerian azimuthal velocity and
 $H$ is a characteristic scale-height.  Thus, $\epsilon$
@@
 \begin{align}\label{eqm_rot}
-  \frac{\Omega^2(r,z)}{\Omega_k^2(r)}=1 +
+  \frac{\Omega^2(r,z)}{\OmK^2(r)}=1 +
   \left(p+\frac{s}{\Gamma}\right)\epsilon^2(r) 
@@
   \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},\\    
-  \frac{\Omega^2(r,z)}{\Omega_k^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
+  \frac{\Omega^2(r,z)}{\OmK^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
     \frac{1}{\sqrt{1+z^2/r^2}}\right),\label{vertiso_eqm}
@@
 \begin{align}
-  r\frac{\p\Omega^2}{\p z} = \frac{sz\Omega_k^2}{\Gamma
+  r\frac{\p\Omega^2}{\p z} = \frac{sz\OmK^2}{\Gamma
     r\left(1+z^2/r^2\right)^{3/2}}. \label{vertical_shear_ex} 
@@
 Without loss of generality, we evaluate Eq. \ref{second_SH} in $z>0$
-where $\p_zP<0$. Here we approximate $\p_r j^2\simeq r^3\Omega_k^2$ but
+where $\p_zP<0$. Here we approximate $\p_r j^2\simeq r^3\OmK^2$ but
 use Eq. \ref{vertical_shear} for $\p_zj^2$. Then 

### 2015-04-24 paper.tex
exceedingly minor end of workday commit.

@@
 \section{Thermodynamic requirement for the vertical shear instability}\label{vsi_require}     In this paper we will find that in a stably stratified thin disk, the VSI  
-requires a thermal timescale $t_c$ much less than the
+requires a thermal timescale $t_c$ much shorter than the
 disk dynamical timescale, 
@@
 \end{align}
-where $\OmK$ is the Keplerian frequency. 
+where $\OmK$ is the Keplerian frequency, consistent with N13. 
 To see why this should be expected, we need to compare vertical shear

### 2015-04-24 appendix.tex
minor notation corrections

@@
 
-% In \S\ref{approx_gov} we made the replacement $D\to\Omega_K^2$ before
+% In \S\ref{approx_gov} we made the replacement $D\to\OmK^2$ before
 % eliminating variables to obtain a single equation for $\delta v_z$ in
@@
 approximations is equivalent to the replacement $D\to
-\Omega_K^2$. This gives, in terms of dimensionless variables,  
+\OmK^2$. This gives, in terms of dimensionless variables,  
 \begin{align}
@@
  of obtaining Eq. \ref{vertiso_gov_nondim} is to neglect the vertical 
- dependence of $D$ by setting $D\to\Omega_K^2$ in Eq. \ref{ode_w} before eliminating
+ dependence of $D$ by setting $D\to\OmK^2$ in Eq. \ref{ode_w} before eliminating
  variables in favor of $\delta v_z$.
@@
 % $D=\kappa^2(z)-\sigma^2$ was ignored because the replacement
-% $D\to\Omega_K^2$ was made before eliminating variables in favor of
+% $D\to\OmK^2$ was made before eliminating variables in favor of
 % $\delta v_z$. In Appendix \ref{adia_improve} we show that making this
@@
 \begin{align}\label{fz_shear}
-  \frac{d\Omega^2}{d\hat{z}} = h^2q f(\hat{z})\Omega_K^2.
+  \frac{d\Omega^2}{d\hat{z}} = h^2q f(\hat{z})\OmK^2.
 \end{align}
@@
 $f\simeq \zhat$, so that $\mathrm{max}|f| = O(1)$, implying a
-maximum growth rate $O(h \Omega_K)$, consistent with numerical
+maximum growth rate $O(h \OmK)$, consistent with numerical
 results.
@@
 % \end{align}
-% Since we have assumed $|\sigma^2|\ll \kappa^2\sim \Omega_K^2$, our
+% Since we have assumed $|\sigma^2|\ll \kappa^2\sim \OmK^2$, our
 % analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   

### 2015-04-24 setup.tex
minor notation corrections

@@
 \end{align}
-where $\zeta = \omega + \ii \sigma$ is a (generally) complex frequency, 
+where $\upsilon = \omega + \ii \sigma$ is a (generally) complex frequency, 
 with $\omega$ and $\sigma$ being
@@
 introduced simply to label terms containing radial gradients of the equilibrium state, 
-which again are evaluated at $(r_o, z)$.   We will consider the effects of ignoring these radial gradients below.
+which again are evaluated at $(r_0, z)$.   We will consider the effects of ignoring these radial gradients below.
 

### 2015-04-24 setup.tex
notation updates across entire manuscript

@@
 \begin{align}
-  \epsilon(r) \equiv \frac{c_{s0}(r)}{v_k(r)} \equiv
+   h(r) \equiv \frac{c_{s0}(r)}{v_k(r)} \equiv
   \frac{H}{r} 
@@
 $v_k=r\OmK$ is the Keplerian azimuthal velocity and
-$H$ is a characteristic scale-height.  Thus, $\epsilon$
+$H$ is a characteristic scale-height.  Thus, $ h$
 is also a characteristic disk aspect-ratio. We are 
-interested in thin disks with $\epsilon \ll 1$.%,%  as relevant for
+interested in thin disks with $ h \ll 1$.%,%  as relevant for
 % protoplanetary disks \citep{chiang97}.
@@
   &\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} = 1 +
-  \frac{\left(\Gamma-1\right)}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right).
+  \frac{\left(\Gamma-1\right)}{ h^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right).
 \end{align}
-Note that for $\Gamma > 1 + \epsilon^2$ there is a disk surface $H_s$
+Note that for $\Gamma > 1 +  h^2$ there is a disk surface $H_s$
 such that $\rho(r,H_s)=0$. 
@@
 % \begin{align}
-%   h^2(r) = \left[1-\frac{\epsilon^2(r)}{\Gamma-1}\right]^{-2}-1. 
+%   h^2(r) = \left[1-\frac{ h^2(r)}{\Gamma-1}\right]^{-2}-1. 
 % \end{align}
@@
   \frac{\Omega^2(r,z)}{\OmK^2(r)}=1 +
-  \left(p+\frac{s}{\Gamma}\right)\epsilon^2(r) 
+  \left(p+\frac{s}{\Gamma}\right) h^2(r) 
   +\frac{s}{\Gamma} \left(1-\frac{1}{\sqrt{1+z^2/r^2}}\right), 
@@
   \frac{\rho(r,z)}{\rho_0(r)} &=
-  \exp{\left[\frac{1}{\epsilon^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},\\    
-  \frac{\Omega^2(r,z)}{\OmK^2(r)}& =1+ (p+q)\epsilon^2(r) + q\left(1 -
+  \exp{\left[\frac{1}{ h^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},\\    
+  \frac{\Omega^2(r,z)}{\OmK^2(r)}& =1+ (p+q) h^2(r) + q\left(1 -
     \frac{1}{\sqrt{1+z^2/r^2}}\right),\label{vertiso_eqm}
@@
 \begin{align}\label{solberg2}
-  % \gamma - \Gamma > \frac{s\epsilon^2}{\Gamma}
+  % \gamma - \Gamma > \frac{s h^2}{\Gamma}
   % \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s
   %   -\left(\gamma-\Gamma\right)\frac{\p\ln{\rho}}{\p\ln{r}} \right]. 
-  \gamma - \Gamma > \frac{\epsilon^2}{\Gamma}
+  \gamma - \Gamma > \frac{ h^2}{\Gamma}
   \left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} \left[s^2
@@
 implies stability. For typical model parameters the right hand
-side (RHS) is $O(\epsilon^2)$. Hence, instability in the adiabatic limit  
-requires $\gamma < \Gamma + O(\epsilon^2)$. Since we consider
-$\epsilon\ll1$, this implies the equilibrium disk must have a small
+side (RHS) is $O( h^2)$. Hence, instability in the adiabatic limit  
+requires $\gamma < \Gamma + O( h^2)$. Since we consider
+$ h\ll1$, this implies the equilibrium disk must have a small
 entropy gradient in order to be unstable to axisymmetric adiabatic

### 2015-04-24 paper.tex
Minor tweak to abs: mention dust opacity where the PPD results are mentioned (not before).

@@
   and allows the VSI to operate.  In this paper, we quantify the
-  cooling timescale, $t_c$, needed for growth of the VSI 
-  and consider the implications for PPDs where $t_c$ is determined by the opacity of dust 
-  grains.  We perform a linear analysis of the VSI with cooling in
+  cooling timescale, $t_c$, needed for growth of the VSI.
+   We perform a linear analysis of the VSI with cooling in
   vertically global and radially local disk models. 
@@
   is much less effective because growth slows and shifts to smaller length scales, which are more prone to 
-  viscous or turbulent decay.  We apply our results to a  Minimum Mass Solar Nebula disk model.  We
+  viscous or turbulent decay.  We apply our results to PPD models 
+  where $t_c$ is determined by the opacity of dust grains.  We
   find that the VSI is most effective at intermediate radii, from $\sim 5$AU to $\sim 50$AU with a

### 2015-04-25 paper.tex
edits mainly to section 2.  Please proofread and check.  It should be obvious but I mainly wanted to (aO cut ideas already in the intro (2) expand on the Richardson like argument (i.e not force the reader to go to Urpin to follow the flow) (c) put the figs a bit earlier (e.g. fig 1 where we first give shear and buoyancy) and (d) expand a bit the physical discussion of t_c figure and implications, but I could be convinced that some of this belongs more in the concluding discussion (on this is was just worried, maybe too much, about giving some wrong impressions).

@@
 
-\section{Thermodynamic requirement for the vertical shear instability}\label{vsi_require}     In this paper we will find that in a stably stratified thin disk, the VSI  
+\section{Thermodynamic requirement for the vertical shear instability}\label{vsi_require}     
+In a stably stratified thin disk, the VSI  
 requires a thermal timescale $t_c$ much shorter than the
@@
 \end{align}
-where $\OmK$ is the Keplerian frequency, consistent with N13. 
-To see why this should be expected, we need to compare vertical shear
-with vertical buoyancy. For this preliminary discussion we consider a
+where $\OmK$ is the Keplerian frequency (N13).    
+This requirement -- that the cooling time be much shorter than an orbital period, 
+which in turn is much smaller than the relevant oscillation period of vertically elongated gravity waves  -- is quite stringent.
+It highlights the fact that vertical buoyancy is a very strong stabilizing influence.
+Very strong radiative damping is thus required to weaken buoyancy and allow the weak vertical shear to drive instability.
+
+To roughly quantify the required smallness of $t_c$, we consider a
 vertically isothermal disk with aspect-ratio $ h$, 
-dimensionless radial temperature gradient $q$ and adiabatic index
-$\gamma$. For a PPD, $ h \sim 0.05$, $|q|\sim 1$ and $\gamma\simeq 1.4$.
+ radial temperature profile $T \propto r^q$ and adiabatic index
+$\gamma$. For a PPD, $ h \sim 0.05$, $q\sim -0.5$ and $\gamma\simeq 1.4$.
 
-For a thin disk, the specific angular momentum varies with height as 
+\begin{figure*}
+  \includegraphics[width=\linewidth]{figures/rhoshearNz}
+  \caption{The radial and vertical structure of a ``minimum mass" PPD model showing gas density (left, relative to the midplane density at 1 AU), vertical shear rate (middle) and
+    buoyancy frequency (right).
+    \label{eqm_structure} 
+  }
+\end{figure*}
+
+For a thin disk, the vertical shear increases with height, $z$, from the midplane as 
 \begin{align}\label{vshear_thin}
-  r \frac{\p \Omega^2 }{\p z} \simeq  h q \OmK^2\left(\frac{z}{H}\right),
+  r \frac{\p \Omega }{\p z} \simeq  h q \OmK \left(\frac{z}{2 H}\right),
 \end{align}
-where $H$ is the disk scale height. As noted in
-\S\ref{intro} this vertical variation is a source of free energy and
-thus vertical shear ($\p_z\Omega \neq0$) is destabilizing.
-
-For $\gamma>1$ the disk is stably stratified with a real vertical
+for several disk scale heights, $H = hr$. 
+%As noted in
+%\S\ref{intro} this vertical variation is a source of free energy and
+%thus vertical shear ($\p_z\Omega \neq0$) is destabilizing. 
+%
+%For $\gamma>1$ the disk is stably stratified with a real 
+This destabilizing shear completes with the stabilizing vertical
 buoyancy frequency $N_z$.  In a thin disk,     
@@
 \end{align}
-This provides a stabilizing effect as buoyancy opposes vertical
-motion, which is required for the VSI. In fact, if we consider
-axisymmetric adiabatic disturbances, then because the 
-Solberg-Hoiland criteria \citep{tassoul78} are usually satisfied in a
-thin disk, the disk would be stable.
-
-However, the stabilizing effect of buoyancy can be mitigated for 
-effectively isothermal disturbances. Consider a fluid element displaced upwards, away from
-the disk midplane. In a stably stratified adiabatic disk, it would
-sink because of buoyancy. However, if thermal timescales are short, in
-that temperature flucations decay rapidly, then the fluid element would attain its surrounding 
-density and  
-%(assuming an ideal gas and it reaches pressure 
-%equilibrium with its surroundings). 
-thus buoyancy is rendered ineffective. Qualitatively, a short thermal
-timescale $t_c$ reduces the effective (stabilizing) buoyancy force.
-
-How short must $t_c$ be for vertical shear to prevail? A rough 
-comparison of vertical shear to buoyancy for non-adiabatic
-perturbations can be obtained by applying Eq. 47 of \cite{urpin03}, which is based on 
-\cite{townsend58},   
-\begin{align}\label{bcrit_est}
-  N_z^2 t_c \lesssim \left|r\frac{\p\Omega}{\p z}\right|. 
-\end{align}
-The left-hand side represents an effective buoyancy, which is
-stabilizing; while the right-hand side represents vertical shear,
-which is destabilizing. In a thin disk $ h \ll 1$, so the destabilizing effect of 
-vertical shear is rather weak because $|r\p_z\Omega|\sim h\OmK
-\ll \OmK$. By comparison, the buoyancy frequency $N_z\sim
-\OmK$. It follows that $t_c\lesssim O( h\OmK^{-1})$ is
-required for instability.  \emph{In a thin  
-  disk the thermal timescale must be shorter than the dynamical
-  timescale by $O( h)\ll 1$ for the VSI to operate.}
-
-To illustrate this point, we show in Fig. \ref{eqm_structure} the
-equilibrium structure of a typical PPD. %Here $\rho_0$
+Fig. \ref{eqm_structure} maps shear and buoyancy (as well as the gas density) 
+in a fiducial PPD model, whose parameters are in Appendix \ref{mmsn} . %Here $\rho_0$
 %is the midplane density and vertical shear rate and buoyancy
 %frequencies are normalized by the local Keplerian frequency. 
-It is
-clear that vertical shear is 
-generally weak compared to buoyancy. Then even a small (positive)
-vertical entropy gradient can significantly stabilize the VSI. 
-Note that in astrophysical disks both $N_z^2$ and $\p_z\Omega$ varies
-with height. Eq. \ref{bcrit_est} then suggest that with
-increasing height, buoyancy would eventually stabilize the VSI (since
-$N_z^2\propto z^2$ and $\p_z\Omega\propto z$).
-% This means that 
-% disturbances localized away from the disk midplane are easily stabilized by
-% buoyancy. 
-This argument assumes vertically localized disturbances. It is,
-however, not clear how to apply Eq. \ref{bcrit_est} to vertically
-global disturbances, as considered in this work.
+Vertical shear is 
+generally weak compared to buoyancy, suggesting stability.
 
-\begin{figure*}
-  \includegraphics[width=\linewidth]{figures/rhoshearNz}
-  \caption{Typical PPD structure (the Minimum Mass
-    Solar Nebla): density (left), vertical shear rate (middle) and
-    buoyancy frequency (right). Here $\rho_0$ is the midplane density.  
-    \label{eqm_structure} 
-  }
-\end{figure*}
+Without any cooling, the Solberg-Hoiland criteria confirms that vertical buoyancy is strong
+enough to ensure (axisymmetric) stability, see \S\ref{solberg}.  With radiative cooling, 
+thermal fluctuations decay which, combined with pressure equilibration, reduces the effective buoyancy.
 
-Our central result is the requirement 
+%However, the stabilizing effect of buoyancy can be mitigated for 
+%effectively isothermal disturbances. Consider a fluid element displaced upwards, away from
+%the disk midplane. In a stably stratified adiabatic disk, it would
+%sink because of buoyancy. However, if thermal timescales are short, in
+%that temperature flucations decay rapidly, then the fluid element would attain its surrounding 
+%density and  
+%%(assuming an ideal gas and it reaches pressure 
+%%equilibrium with its surroundings). 
+%thus buoyancy is rendered ineffective. Qualitatively, a short thermal
+%timescale $t_c$ reduces the effective (stabilizing) buoyancy force.
+
+How short must $t_c$ be for vertical shear to prevail?  We start by 
+considering the Richardson number $\mathrm{Ri} = N_z^2/(r\p_z\Omega)^2$, a 
+ratio of buoyant to shear energies.  Though not precisely applicable to our problem,
+non-rotating shear flows are stable if $\mathrm{Ri} > 1/4$ (\citealp{chandrasekhar61}, 
+also see \citealp{ys02, lee10} for applications to thin dust layers in PPDs).
+Following \cite{urpin03} and \cite{townsend58},  we reduce the buoyant energy by the 
+ratio of cooling to forcing timescales, $t_c |r \p_z\Omega| \leq 1$ (with no reduction in buoyant energy 
+expected when this inequality fails).  With this reduction, the Richardson like criterion for 
+shear \emph{instability} becomes
+\begin{align}\label{bcrit_est}
+  t_c \lesssim \frac{\left|r\p\Omega/\p z\right|} {N_z^2}. 
+\end{align}
+
+This result is far from rigorous, but it contains some useful insights.
+ If we interpret Eq.\ \ref{bcrit_est} as a local criterion, then we see that 
+ the minimum cooling time decreases with height as $t_c \propto 1/z$.
+ This finding is relevant to the radiative damping of so called ``surface 
+ modes," which we consider in \S\ref{numerical}.
+
+ To roughly obtain a global cooling time criterion, we evaluate Eq.\ \ref{bcrit_est}
+ at $z \sim H$.  To get the precise form we want, we choose $z = \gamma H / 2$ to 
+ obtain the cooling requirement
  \begin{align}\label{prelim_bcrit}
-  (\gamma - 1)t_c <  h |q| \OmK^{-1}
+ t_c <  h |q| \OmK^{-1}/ (\gamma - 1)
 \end{align}   
-for the VSI in a vertically isothermal, thin disk. This equation
-reflects Eq. \ref{bcrit_est} but there is no $z$ dependence. 
-% While assumptions and simplifications will be made 
-% to derive this, 
-Eq. \ref{prelim_bcrit} can be 
-readily applied to evaluate where in a disk the VSI is likely to
-operate. For example, we show in Fig. \ref{bcrit_mmsn2d} an estimate
-of the thermal timescales the PPD model shown in
-Fig. \ref{eqm_structure}. Here $k_x$ is a radial
-wavenumber of the disturbance being considered, thus larger $k_x$
-correspond to smaller perturbations. The region outside the red line
-satisfies Eq. \ref{prelim_bcrit}.
+for the VSI in a vertically isothermal, thin disk.  This estimate agrees with 
+our central result, the cooling criterion that we derive more rigorously in \S\ref{iso_vsi_beta_crit}.
 
-The criterion Eq. \ref{prelim_bcrit} in fact
-assumes $t_c$ is vertically uniform. While this is not generally true
-in the inner disk, we find Eq. \ref{prelim_bcrit} still provides a
-useful guide. We see that the VSI will operate in the outer 
-disk at tens of AU, but that smaller scale disturbances may operate at
-smaller radii.
+%
+%The left-hand side represents an effective buoyancy, which is
+%stabilizing; while the right-hand side represents vertical shear,
+%which is destabilizing. In a thin disk $ h \ll 1$, so the destabilizing effect of 
+%vertical shear is rather weak because $|r\p_z\Omega|\sim h\OmK
+%\ll \OmK$. By comparison, the buoyancy frequency $N_z\sim
+%\OmK$. It follows that $t_c\lesssim O( h\OmK^{-1})$ is
+%required for instability.  \emph{In a thin  
+%  disk the thermal timescale must be shorter than the dynamical
+%  timescale by $O( h)\ll 1$ for the VSI to operate.}   
+%
+%
+%Then even a small (positive)
+%vertical entropy gradient can significantly stabilize the VSI. 
+%Note that in astrophysical disks both $N_z^2$ and $\p_z\Omega$ varies
+%with height. Eq. \ref{bcrit_est} then suggest that with
+%increasing height, buoyancy would eventually stabilize the VSI (since
+%$N_z^2\propto z^2$ and $\p_z\Omega\propto z$).
+%% This means that 
+%% disturbances localized away from the disk midplane are easily stabilized by
+%% buoyancy. 
+%This argument assumes vertically localized disturbances. It is,
+%however, not clear how to apply Eq. \ref{bcrit_est} to vertically
+%global disturbances, as considered in this work.
+%
+%
+%
+%
+%Our central result is the requirement 
+%\begin{align}\label{prelim_bcrit}
+%  (\gamma - 1)t_c <  h |q| \OmK^{-1}
+%\end{align}   
+%for the VSI in a vertically isothermal, thin disk. This equation
+%reflects Eq. \ref{bcrit_est} but there is no $z$ dependence. 
+%% While assumptions and simplifications will be made 
+%% to derive this, 
+%Eq. \ref{prelim_bcrit} can be 
+%readily applied to evaluate where in a disk the VSI is likely to
+%operate.
 
@@
 
-We now proceed through the formal linear stability problem to arrive
+We now consider whether, and where, thermal timescales in realistic PPDs
+are short enough to support VSI growth.  Fig. \ref{bcrit_mmsn2d} plots the
+ thermal timescales in our fiducial PPD model with a small grain dust opacity. 
+ Since cooling by radiative diffusion depends on a perturbation's lengthscale, we
+ plot $t_c$ for two values of the radial wavenumber,  $k_x$,  the relevant 
+ component for VSI modes that are radially narrow.   The red line
+marks the boundary implied by Eq. \ref{prelim_bcrit}.  We caution that 
+ Eq. \ref{prelim_bcrit} applies when $t_c$ is vertically uniform.  This section is 
+ aimed at basic insights, while \S\ref{application}
+ analyzes the VSI with realistic spatial variations in $t_c$.
+
+Due to the smaller values of $t_c$ in the outer disk, we generally expect the VSI to 
+operate outside $\sim 10$ AU.  This conclusion changes however, if the dust opacity is reduced, 
+slowing down cooling, as we show later (e.g.\ Fig.\ \ref{mmsn_bcrit_bcool}).
+
+In the inner disk we see that thermal times are long in the midplane, but short at higher altitutdes. 
+Again, this structure necessitates the more detailed analyses that follow.  Nevertheless, we can already 
+see that due to radiative diffusion higher wavenumbers can cool more rapidly.  Thus smaller scale disturbances 
+may trigger VSI at smaller disk radii.
+
+However we cannot and do not find that VSI will 
+always be relevant for sufficiently small perturbations.  For starters, very small wavelengths are in the optically 
+thin cooling regime,  where cooling is independent of lengthscale.  Moreover, small scale perturbations 
+are prone to viscous damping, and should drive weaker transport.
+
+We now proceed with our detailed analysis of the VSI to arrive
 at Eq. \ref{prelim_bcrit} and verify it numerically. The interested 
-reader may skip forward \S\ref{application} for the application of our
-results to PPD.
+reader may jump to \S\ref{application} for the application of our
+results to PPDs.
 

### 2015-04-25 linear_adia.tex
minor edits to sec 2, 3, 4

@@
 slightly from a case with $\beta\equiv 0$. We thus perturb a solution
-for $\beta=0$ to see the effect of thermal relaxation.
+for $\beta=0$ to see the effect of finite cooling.
 
@@
 where the vertical velocity is constant throughout the disk
-column. Hereafter we shall simply refer to them as the fundamental
+column. Hereafter we shall simply call it the fundamental
 mode.
@@
 
-A finite thermal relaxation time allow buoyancy
+A finite cooling time allow buoyancy
 forces to stabilize vertical motions in sub-adiabatically stratified

### 2015-04-25 linear_iso.tex
minor notation corrections
minor edits to sec 2, 3, 4

@@
 
-\item We consider eigenvalues such that
-  $|\upsilon^2|\ll \kappa^2,\,\Omega^2$; which correspond to inertial
-  waves \citepalias{barker15}. We call this the low frequency  
-  approximation.
+\item We consider inertial waves with eigenfrequencies such that
+  $|\upsilon^2|\ll \kappa^2,\,\Omega^2$ \citepalias{barker15}. We call
+  this the low-frequency approximation.
 
@@
   explicitly in the linearized equations through the term
-  $d\Omega/dz$. We thus ignore the vertical dependence of $\Omega$ and
+  $\p_z\Omega$. We thus ignore the vertical dependence of $\Omega$ and
   $\kappa$ elsewhere, and replace them with
@@
 where $\hat{\omega}$ and $\hat{\sigma}$ are real, the above
-approximations lead to a single second order ordinary differential
-equation, 
+approximations lead to a single second order ODE, 
 % Eq. \ref{vertiso_gov} becomes 

### 2015-04-25 paper.tex
minor edits to sec 2, 3, 4

@@
 where $\OmK$ is the Keplerian frequency (N13).    
-This requirement -- that the cooling time be much shorter than an orbital period, 
-which in turn is much smaller than the relevant oscillation period of vertically elongated gravity waves  -- is quite stringent.
-It highlights the fact that vertical buoyancy is a very strong stabilizing influence.
-Very strong radiative damping is thus required to weaken buoyancy and allow the weak vertical shear to drive instability.
+This requirement --- that the cooling time be much shorter than an orbital period, 
+which in turn is much smaller than the relevant oscillation period of vertically elongated gravity waves  --- is quite stringent.
+It highlights the fact that vertical buoyancy is strongly 
+stabilizing. Rapid radiative damping is thus required to weaken
+buoyancy and allow the weak vertical shear to drive instability.
 
@@
  radial temperature profile $T \propto r^q$ and adiabatic index
-$\gamma$. For a PPD, $ h \sim 0.05$, $q\sim -0.5$ and $\gamma\simeq 1.4$.
+$\gamma>1$ so the disk is stably stratified.  For a PPD, $ h \sim 0.05$,
+$q\sim -0.5$ and $\gamma\simeq 1.4$.
 
@@
   \includegraphics[width=\linewidth]{figures/rhoshearNz}
-  \caption{The radial and vertical structure of a ``minimum mass" PPD model showing gas density (left, relative to the midplane density at 1 AU), vertical shear rate (middle) and
+  \caption{The radial and vertical structure of a `minimum mass' PPD model showing gas density (left, relative to the midplane density at 1 AU), vertical shear rate (middle) and
     buoyancy frequency (right).
@@
 
-For a thin disk, the vertical shear increases with height, $z$, from the midplane as 
+For a thin disk ($h\ll 1$), the vertical shear rate varies with height, $z$, from the midplane as 
 \begin{align}\label{vshear_thin}
@@
 \end{align}
-for several disk scale heights, $H = hr$. 
+where $H=hr$ is the characteristic disk scale height. 
+%for several disk scale heights, $H = hr$. 
 %As noted in
@@
 %For $\gamma>1$ the disk is stably stratified with a real 
-This destabilizing shear completes with the stabilizing vertical
+This destabilizing shear competes with the stabilizing vertical
 buoyancy frequency $N_z$.  In a thin disk,     
@@
 \end{align}
-Fig. \ref{eqm_structure} maps shear and buoyancy (as well as the gas density) 
-in a fiducial PPD model, whose parameters are in Appendix \ref{mmsn} . %Here $\rho_0$
+Fig. \ref{eqm_structure} maps the vertical shear rate and buoyancy
+frequency (as well as the gas density) 
+in a fiducial PPD model considered in \S\ref{application}. 
+%, whose parameters are in Appendix \ref{mmsn} . %Here $\rho_0$
 %is the midplane density and vertical shear rate and buoyancy
@@
 Vertical shear is 
-generally weak compared to buoyancy, suggesting stability.
-
-Without any cooling, the Solberg-Hoiland criteria confirms that vertical buoyancy is strong
+generally weak compared to buoyancy, suggesting stability. In fact, 
+without any cooling, the Solberg-Hoiland criteria \citep{tassoul78}  
+confirms that vertical buoyancy is strong
 enough to ensure (axisymmetric) stability, see \S\ref{solberg}.  With radiative cooling, 
@@
 ratio of cooling to forcing timescales, $t_c |r \p_z\Omega| \leq 1$ (with no reduction in buoyant energy 
-expected when this inequality fails).  With this reduction, the Richardson like criterion for 
+expected when this inequality fails).  With this reduction, the Richardson-like criterion for 
 shear \emph{instability} becomes
 \begin{align}\label{bcrit_est}
-  t_c \lesssim \frac{\left|r\p\Omega/\p z\right|} {N_z^2}. 
+  t_c \lesssim \frac{\left|r\p\Omega/\p z\right|} {N_z^2}, 
 \end{align}
+which is Eq.  47 of \cite{urpin03}. %This result is far from rigorous, but it contains some useful insights.
+If we interpret Eq.\ \ref{bcrit_est} as a local criterion, then we see that 
+the maximum cooling time which permits instability decreases with height as $1/|z|$;  
+the stabilizing effect of vertical buoyancy increases more rapidly away from
+the midplane than the destabilizing effect of vertical shear.  
+This finding is relevant to the radiative damping of so called `surface 
+modes', which we consider in \S\ref{numerical}.
 
-This result is far from rigorous, but it contains some useful insights.
- If we interpret Eq.\ \ref{bcrit_est} as a local criterion, then we see that 
- the minimum cooling time decreases with height as $t_c \propto 1/z$.
- This finding is relevant to the radiative damping of so called ``surface 
- modes," which we consider in \S\ref{numerical}.
-
- To roughly obtain a global cooling time criterion, we evaluate Eq.\ \ref{bcrit_est}
- at $z \sim H$.  To get the precise form we want, we choose $z = \gamma H / 2$ to 
- obtain the cooling requirement
+For vertically global disturbances considered in this work, 
+% To roughly obtain a global cooling time criterion, we evaluate Eq.\ \ref{bcrit_est}
+% at $z \sim H$.  To get the precise form we want, we choose $z = \gamma H / 2$ to 
+we obtain the cooling requirement 
 \begin{align}\label{prelim_bcrit}
@@
 \end{align}   
-for the VSI in a vertically isothermal, thin disk.  This estimate agrees with 
-our central result, the cooling criterion that we derive more rigorously in \S\ref{iso_vsi_beta_crit}.
+for the VSI to operate effectively in a vertically isothermal, thin
+disk. We derive this result more formally in
+\S\ref{iso_vsi_beta_crit}, but notice that Eq. \ref{prelim_bcrit}
+may be obtained from Eq. \ref{bcrit_est} by evaulating it at the
+height $|z|=\gamma H/2$. %i.e., at $z=O(H)$
+
+%This estimate agrees with 
+%our central result, the cooling criterion that we derive more rigorously in \S\ref{iso_vsi_beta_crit}.
 
@@
  Eq. \ref{prelim_bcrit} applies when $t_c$ is vertically uniform.  This section is 
- aimed at basic insights, while \S\ref{application}
- analyzes the VSI with realistic spatial variations in $t_c$.
+ aimed at basic insights; while more realistic spatial variations in
+ $t_c$ is considered in \S\ref{application}. 
+ %analyzes the VSI with realistic spatial variations in.
 
  Due to the smaller values of $t_c$ in the outer disk, we generally expect the VSI to 
-operate outside $\sim 10$ AU.  This conclusion changes however, if the dust opacity is reduced, 
-slowing down cooling, as we show later (e.g.\ Fig.\ \ref{mmsn_bcrit_bcool}).
-
-In the inner disk we see that thermal times are long in the midplane, but short at higher altitutdes. 
+ operate outside $\sim 10$ AU.  This conclusion changes however, if
+ the dust opacity is reduced,  
+slowing down cooling, as we show in \S\ref{application}. % (e.g.\ Fig.\
+% \ref{mmsn_bcrit_bcool}). 
+In the inner disk, thermal times are long in the midplane, but short at higher altitutdes. 
 Again, this structure necessitates the more detailed analyses that follow.  Nevertheless, we can already 
@@
 
-However we cannot and do not find that VSI will 
-always be relevant for sufficiently small perturbations.  For starters, very small wavelengths are in the optically 
-thin cooling regime,  where cooling is independent of lengthscale.  Moreover, small scale perturbations 
-are prone to viscous damping, and should drive weaker transport.
+On the other hand, one cannot obtain arbitrarily short cooling times
+by invoking arbitrarily small scales (cf. \citealt{urpin03}, 
+\citetalias{barker15}), because very small wavelengths are in the
+optically thin cooling regime,  where cooling is independent of
+lengthscale. Furthermore, while we do not include viscosity in our
+idealized disk models, in reality small-scale perturbations will be 
+prone to viscous damping, and should drive weaker transport.
+
+% However, we cannot and do not find that VSI will always be relevant for
+% sufficiently small perturbations because \begin{inparaenum}[(i)] \item very small wavelengths are in the optically 
+% thin cooling regime,  where cooling is independent of lengthscale;
+% and \item small scale perturbations 
+% are more prone to viscous damping, and should drive weaker transport.
+% \end{inparaenum}
 

### 2015-04-25 setup.tex
edits mainly to section 2.  Please proofread and check.  It should be obvious but I mainly wanted to (aO cut ideas already in the intro (2) expand on the Richardson like argument (i.e not force the reader to go to Urpin to follow the flow) (c) put the figs a bit earlier (e.g. fig 1 where we first give shear and buoyancy) and (d) expand a bit the physical discussion of t_c figure and implications, but I could be convinced that some of this belongs more in the concluding discussion (on this is was just worried, maybe too much, about giving some wrong impressions).
minor edits to sec 2, 3, 4

@@
 \section{Governing equations and disk models}\label{setup}
-We consider  an inviscid, non-self-gravitating three-dimensional (3D)
+We consider  an inviscid, non-self-gravitating three-dimensional 
 disk orbiting a central star of mass $M_*$.  Cylindrical
@@
 \end{align}
-relative to the Keplerian frequency, $\OmK=\sqrt{GM_*/r^3}$.
+relative to the Keplerian frequency, $\OmK=\sqrt{GM_*/r^3}$.  We shall
+use the terms `thermal relaxation' and `cooling' interchangeably
+throughout this paper.
 
@@
 disk from isothermal ($\beta\ll 1$) to adiabatic ($\beta \gg 1$) in a
-controlled manner.
+controlled manner. However, in \S\ref{application} 
+we will consider a more realistic model for $\beta$ in
+PPDs with a dust opacity (which gives Fig.  \ref{bcrit_mmsn2d}).
 
-In \S\ref{application} we analyze the VSI using a realistic model for 
-$\beta$ in PPDs with a dust opacity, using Eq.\ (\ref{beta_mmsn_simp}).  
- Fig. \ref{bcrit_mmsn2d} shows that in this model $\beta >> 1$ in the 
- dense inner disk and near the midplane, but that rapid cooling ($\beta \ll 1$)
- is possible elsewhere.  Furthermore short wavelength perturbations can cool 
- quicker ...
+% using Eq.\(\ref{beta_mmsn_simp}), which   
+% Fig. \ref{bcrit_mmsn2d} shows that in this model $\beta >> 1$ in the 
+% dense inner disk and near the midplane, but that rapid cooling ($\beta \ll 1$)
+%is possible elsewhere.  Furthermore short wavelength perturbations can cool 
+%quicker ...
 
- \emph{TODO: $k_x$ is in Fig. \ref{bcrit_mmsn2d} but not explained yet...}
+% \emph{TODO: $k_x$ is in Fig. \ref{bcrit_mmsn2d} but not explained yet...}
+
 % based on physical thermal
@@
 \end{align}
-where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the power-law index, and
+where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the temperature power-law index, and
 define   
@@
 \subsubsection{Vertically isothermal disks}
-For $\Gamma=1$ the equilibrium density $\rho(r,z)$ and rotation
-$\Omega(r,z)$ are given by 
+For $\Gamma=1$ the equilibrium density and rotation profiles are
+given, respectivley, by 
 \begin{align}
@@
 \end{align}
-Note that there is a maximum vertical shear rate at $|z|=r/\sqrt{2}$,
+Note that at fixed $r$ there is a maximum vertical shear rate at $|z|=r/\sqrt{2}$,
 which is expected to limit the growth rate of instabilities associated with vertical shear.
@@
 Without loss of generality, we evaluate Eq. \ref{second_SH} in $z>0$
-where $\p_zP<0$. Here we approximate $\p_r j^2\simeq r^3\OmK^2$ but
+where $\p_zP<0$. Here we approximate $\p_r j^2\simeq r^3\OmK^2$ and
 use Eq. \ref{vertical_shear} for $\p_zj^2$. Then 
@@
 stratified disks \citep{chiang97}, in which the VSI requires rapid
-thermal relaxation, as shown below.
+cooling, as shown below.
 
@@
  We first expand all fluid variables 
- into the equilibrium value plus a perturbation denoted by $\delta$, e.g.\ 
+ into the equilibrium value plus an Eulerian perturbation denoted by $\delta$, e.g.\ 
  $\delta \rho$ for the perturbed density, and drop all non-linear perturbations.  Second, we perform
@@
 the real frequency and growth rate, respectively, and $k_x$ is a real 
-radial wavenumber.  Henceforth all unperturbed fluid variables, including 
+radial wavenumber.  We take $k_x>0$ without loss of generality, 
+assume $k_xr_0 \gg 1$ and neglect curvature terms. 
+Henceforth all unperturbed fluid variables, including 
 gradients such as $\p \Omega/ \p z$, refer to  equilibrium values at $(r_0, z)$.
@@
   \zeta\frac{1}{\rho}\frac{\p P}{\p r}\frac{Q}{c_s^2}\label{lin_vx}\\
-   \ii \upsilon\delta v_y &= r\frac{\p \Omega}{\p z}\delta v_z +
+   \ii \upsilon\delta v_y &= r_0\frac{\p \Omega}{\p z}\delta v_z +
   \frac{\kappa^2}{2\Omega}\delta v_x, \label{lin_vy}\\
@@
 \end{align}
-The above equations are a set of ODEs (ordinary differential equations) in $z$.  Solutions to these 
+This is a set of ordinary differential equations (ODEs) in
+$z$.  %boundary conditions for well posed problem 
+Solutions to these 
 are presented in \S\ref{numerical} and \S\ref{application}.   The coefficient $ \zeta = 1$ is 
 introduced simply to label terms containing radial gradients of the equilibrium state, 
-which again are evaluated at $(r_0, z)$.   We will consider the effects of ignoring these radial gradients below.
+which again are evaluated at $(r_0, z)$. For clarity, hereafter we drop the subcript $0$.    
+We will consider the effects of ignoring these radial gradients below.
 

### 2015-04-28 paper.tex
more edits to section 3, still in progress.  compactifying a bit with some clarification on how to derive.  will let you know when done with this round.

@@
 optically thin cooling regime,  where cooling is independent of
-lengthscale. Furthermore, while we do not include viscosity in our
-idealized disk models, in reality small-scale perturbations will be 
-prone to viscous damping, and should drive weaker transport.
+lengthscale. Furthermore, 
+%while we do not include viscosity in our
+%idealized disk models, in reality 
+small-scale perturbations will be 
+prone to viscous damping, including the effective viscosity from turbulent motions.
+While our linear, inviscid analysis does not address this issue directly, we expect that 
+shorter wavelength modes will drive weaker non-linear (and non-axisymmetric) transport.
 

### 2015-04-28 setup.tex
more edits to section 3, still in progress.  compactifying a bit with some clarification on how to derive.  will let you know when done with this round.

@@
 \section{Governing equations and disk models}\label{setup}
-We consider  an inviscid, non-self-gravitating three-dimensional 
-disk orbiting a central star of mass $M_*$.  Cylindrical
-co-ordinates $(r,\phi, z)$ are centered on the star. The governing fluid equations are
+An inviscid, non-self-gravitating disk orbiting a central star of mass $M_*$
+obeys the three-dimensional fluid equations:
 \begin{align}
@@
 = -GM_*(r^2 + z^2)^{-1/2}$ is the gravitational potential of the
-central star and $G$ is the gravitational constant. 
+central star with $G$ as the gravitational constant. The cylindrical
+co-ordinates $(r,\phi, z)$ are centered on the star. 
 In the energy equation (Eq.\ \ref{full_energy}) the sink term
@@
 \end{align}
-relative to the Keplerian frequency, $\OmK=\sqrt{GM_*/r^3}$.  We shall
+relative to the Keplerian frequency, $\OmK=\sqrt{GM_*/r^3}$.  We 
 use the terms `thermal relaxation' and `cooling' interchangeably
@@
 we will consider a more realistic model for $\beta$ in
-PPDs with a dust opacity (which gives Fig.  \ref{bcrit_mmsn2d}).
+PPDs with a dust opacity (as in Fig.\  \ref{bcrit_mmsn2d}).
 
@@
 $\Omega(r,z)$, respectively. These functions satisfy  
-\begin{align}
+\begin{subequations}\begin{align}
   0 &= \frac{1}{\rho}\frac{\p P}{\p z} + \frac{\p \Phi_*}{\p z},\label{vert_equilibrium}\\
@@
     r}.\label{hori_equilibrium} 
+\end{align} \end{subequations}
+A unique solution requires additional assumptions about disk structure.  We choose
+\begin{subequations}\begin{align}
+ \rho(r,0) &\equiv \rho_0(r) = \rho_{00}\left(\frac{r}{r_0}\right)^p, \\
+ T(r,0) &\equiv T_0(r) = T_{00}\left(\frac{r}{r_0}\right)^q, \\
+ P(r,z) & = K(r)\rho^{\Gamma}(r,z). \label{eqm_press}
+\end{align} \end{subequations}
+where $r_0$ is a fiducial radius and $p$ and $q$ are the standard powerlaw indices
+for midplane density and temperature, respectively.  The polytropic index $\Gamma$ 
+parametrizes the vertical stratification with $\Gamma = 1$ ($\Gamma = \gamma$)
+describing vertically isothermal (adiabatic) disks, respectively.  The ideal gas
+law requires $K = \rho_0^{1-\Gamma}\mathcal{R} T_0/\mu$.
+
+%These equations allows us to choose any midplane density 
+%distribution $\rho_0(r)\equiv \rho(r,0)$. In this work we assume a
+%power-law profile, 
+%\begin{align}
+%  \rho_0(r) = \rho_{00}\left(\frac{r}{r_0}\right)^p,
+%\end{align}
+%where $\rho_{00}$ is the midplane density at the fiducial radius
+%$r=r_0$ and $p$ is the power-law index.
+
+%We consider disk equilibria in which the pressure $P$ and density
+%$\rho$ are related by   
+%\begin{align}
+%  P(r,z) &= 
+%  \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z),\notag\\
+%  & \equiv K(r)\rho^{\Gamma}(r,z). \label{eqm_press}
+%\end{align}
+%where the polytropic index $\Gamma$ is a  constant parameter  and the
+%function $c_{s0}(r)$ is prescribed below. Choosing $\Gamma=1$ gives  a 
+%locally vertically isothermal disk and $c_{s0}(r)$ is the mid-plane
+%sound-speed. 
+%It shall be convenient to define the function $c_s(r,z)$
+
+We further define a modified sound speed
+\begin{align}
+	c_s \equiv \sqrt{\Gamma P /\rho } , 
+% c_s^2\equiv % \frac{dP}{d\rho} =
+%  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} =
+%  \frac{\Gamma P}{\rho},  
 \end{align}
-These equations allows us to choose any midplane density 
-distribution $\rho_0(r)\equiv \rho(r,0)$. In this work we assume a
-power-law profile, 
+which in general differs from the  isothermal, $c_\mathrm{iso} =
+c_s/\sqrt{\Gamma}$, and adiabatic $c_\mathrm{ad} =
+c_s\sqrt{\gamma/\Gamma}$, sound speeds.
+
+We also introduce the characteristic scale-height, $H = c_s(r,0)/\OmK$
+and disk aspect-ratio
+%We take $c_{s0}(r)$ to be another power-law:  
+%\begin{align}
+%  c_{s0}^2(r)=c_{s00}^2\left(\frac{r}{r_0}\right)^q, 
+%\end{align}
+%where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the temperature power-law index, and
+%define   
 \begin{align}
-  \rho_0(r) = \rho_{00}\left(\frac{r}{r_0}\right)^p,
+   h(r) \equiv %\frac{c_{s0}(r)}{v_k(r)} \equiv
+  \frac{H}{r} \propto r^{(1+q)/2}
 \end{align}
-where $\rho_{00}$ is the midplane density at the fiducial radius
-$r=r_0$ and $p$ is the power-law index.
+We are interested in thin disks with $ h \ll 1$.
 
-We consider disk equilibria in which the pressure $P$ and density
-$\rho$ are related by   
-\begin{align}
-  P(r,z) &= 
-  \frac{c_{s0}^2(r)}{\Gamma}\rho_0(r)^{1-\Gamma}\rho^\Gamma(r,z),\notag\\
-  & \equiv K(r)\rho^{\Gamma}(r,z). \label{eqm_press}
-\end{align}
-where the polytropic index $\Gamma$ is a  constant parameter  and the
-function $c_{s0}(r)$ is prescribed below. Choosing $\Gamma=1$ gives  a 
-locally vertically isothermal disk and $c_{s0}(r)$ is the mid-plane
-sound-speed. 
-It shall be convenient to define the function $c_s(r,z)$ 
-\begin{align}
-  c_s^2\equiv % \frac{dP}{d\rho} =
-  c_{s0}^2(r)\left(\frac{\rho}{\rho_0}\right)^{\Gamma-1} =
-  \frac{\Gamma P}{\rho}, 
-\end{align}
-which is related to the isothermal sound-speed $c_\mathrm{iso} =
-c_s/\sqrt{\Gamma}$. We take $c_{s0}(r)$ to be another power-law:  
-\begin{align}
-  c_{s0}^2(r)=c_{s00}^2\left(\frac{r}{r_0}\right)^q, 
-\end{align}
-where $c_{s00}\equiv c_{s0}(r_0)$  and $q$ is the temperature power-law index, and
-define   
-\begin{align}
-   h(r) \equiv \frac{c_{s0}(r)}{v_k(r)} \equiv
-  \frac{H}{r} 
-\end{align}
-to parametrize the magnitude of the sound-speed, where
-$v_k=r\OmK$ is the Keplerian azimuthal velocity and
-$H$ is a characteristic scale-height.  Thus, $ h$
-is also a characteristic disk aspect-ratio. We are 
-interested in thin disks with $ h \ll 1$.%,%  as relevant for
-% protoplanetary disks \citep{chiang97}.
-
-%\subsection{Vertical structure}
-
-Solving the equilibrium disk equations, we find the explicit
-expression for the density field for 
-$\Gamma\neq1$ is given  by
+\subsubsection{Density structure}
+The equilibrium density field follows from vertical hydrostatic equilibrium (Eq.\ \ref{vert_equilibrium}).
+The solution for $\Gamma\neq1$ is
 \begin{align}\label{eqm_dens}
@@
 Note that for $\Gamma > 1 +  h^2$ there is a disk surface $H_s$
-such that $\rho(r,H_s)=0$. 
+where $\rho(r,H_s)=0$. 
 % Note that for $\Gamma\neq1$ the disk has a surface $z=H(r)$ such that
@@
 
-Similarly, the equilibrium rotation is 
+The vertically isothermal case, $\Gamma = 1$, can be found either 
+independently or as a limit of Eq. \ref{eqm_dens} to be 
+\begin{align}
+  \frac{\rho(r,z)}{\rho_0(r)} &=
+  \exp{\left[\frac{1}{ h^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},   \\
+  & \approx \exp {\left[ - {z^2 \over 2 H^2}\right]} \label{rhoisothin},
+  \end{align}
+  where Eq.\ \ref{rhoisothin} takes the $z \ll r$ thin disk limit.
+
+ \subsubsection{Rotation and vertical shear profiles}
+The equilbrium rotation field follows from the density field and centrifugal balance (Eq.\ \ref{hori_equilibrium}),
+giving 
 \begin{align}\label{eqm_rot}
   \frac{\Omega^2(r,z)}{\OmK^2(r)}=1 +
-  \left(p+\frac{s}{\Gamma}\right) h^2(r) 
+  \left(\frac{p+q}{\Gamma}\right) h^2(r) 
   +\frac{s}{\Gamma} \left(1-\frac{1}{\sqrt{1+z^2/r^2}}\right), 
 \end{align}
-where $s\equiv q+p(1-\Gamma)$. We also define % the radial shear rate  
-% \begin{align}
-%   S(r,z) \equiv r\frac{\p\Omega}{\p r},  
-% \end{align}
-% and
-the square of the epicyclic frequency $\kappa$ 
-\begin{align}
-  \kappa^2(r,z) \equiv \frac{1}{r^3}\frac{\p }{\p
-    r}\left(r^4\Omega^2\right). %2\Omega(2\Omega + S). 
-\end{align}
+for all $\gamma$, where $s\equiv q+p(1-\Gamma)$.
 
-\subsubsection{Vertically isothermal disks}
-For $\Gamma=1$ the equilibrium density and rotation profiles are
-given, respectivley, by 
-\begin{align}
-  \frac{\rho(r,z)}{\rho_0(r)} &=
-  \exp{\left[\frac{1}{ h^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},\\    
-  \frac{\Omega^2(r,z)}{\OmK^2(r)}& =1+ (p+q) h^2(r) + q\left(1 -
-    \frac{1}{\sqrt{1+z^2/r^2}}\right),\label{vertiso_eqm}
-\end{align}
-which may be obtained from Eq. \ref{eqm_dens} and Eq. \ref{eqm_rot} by
-taking the limit $\Gamma\to 1$. We will, in fact, primarily focus on
-vertically isothermal disks which is the relevant case for passively irradiated
-PPDs \citep{chiang97}.
+Notice that for $\Gamma = 1$, $s/\Gamma = q$...
+
+...
+%\subsubsection{Vertically isothermal disks}
+%For $\Gamma=1$ the equilibrium density and rotation profiles are
+%given, respectivley, by 
+%\begin{align}
+%  \frac{\rho(r,z)}{\rho_0(r)} &=
+%  \exp{\left[\frac{1}{ h^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},\\    
+%  \frac{\Omega^2(r,z)}{\OmK^2(r)}& =1+ (p+q) h^2(r) + q\left(1 -
+%    \frac{1}{\sqrt{1+z^2/r^2}}\right),\label{vertiso_eqm}
+%\end{align}
+%which may be obtained from Eq. \ref{eqm_dens} and Eq. \ref{eqm_rot} by
+%taking the limit $\Gamma\to 1$. We will, in fact, primarily focus on
+%vertically isothermal disks which is the relevant case for passively irradiated
+%PPDs \citep{chiang97}.
 

### 2015-04-28 paper.tex
edits to section 4 with corresponding tweaks to appendix A.  General editing for clarity and conciseness.  Also attempted to blunt perceived criticism of VGSB (which contains many aspects) by focusing on the 'fully radially local approximation'.

@@
 %\documentclass[12pt]{aastex}
-\documentclass[iop]{emulateapj}
+\documentclass[iop, numberedappendix]{emulateapj}
 %\documentclass{article}
 %\usepackage{draftwatermark}
-\usepackage{paralist}
+%\usepackage{paralist}
 \shorttitle{Vertical shear instability}
@@
 \defcitealias{mcnally14}{MP14}
+\defcitealias{goldreich67}{GS67}
+
 \def \OmK {\Omega_{\rm K}}
@@
 basic mechanism of the VSI  in disks was first 
-identified in the context of differentially rotating stars \citep[][]{goldreich67,fricke68}.   
+identified in the context of differentially rotating stars (\citealp{goldreich67}, hereafter \citetalias{goldreich67}, \citealp{fricke68}).   
 The VSI arises when vertical shear, i.e. \ a variation in orbital motion along the rotation axis, 
@@
 stratified. To overcome vertical buoyancy, the VSI requires a short cooling time
-(\citealp{goldreich67}; \citetalias{nelson13}). Rapid radiative 
+(\citetalias{goldreich67}; \citetalias{nelson13}). Rapid radiative 
 cooling, equivalently a short thermal relaxation timescale, brings a

### 2015-04-29 paper.tex
minor edits, mainly to section 5 where I thought the intro should (a) deal more briefly with the isothermal limit and (b) summarize the results of the section (as this is probably the hardest section for most readers to follow).

@@
 
-\section{Thermodynamic requirement for the vertical shear instability}\label{vsi_require}     
+\section{Why must cooling be so fast?}\label{vsi_require}     
 In a stably stratified thin disk, the VSI  

### 2015-04-29 setup.tex
edits to section 4 with corresponding tweaks to appendix A.  General editing for clarity and conciseness.  Also attempted to blunt perceived criticism of VGSB (which contains many aspects) by focusing on the 'fully radially local approximation'.
more edits to 3.3 will require merge

@@
 Note that for $\Gamma > 1 +  h^2$ there is a disk surface $H_s$
-where $\rho(r,H_s)=0$. 
+where $\rho(r,H_s)=0$.  For $h \ll 1$, this surface is at $H_s/H \approx \sqrt{2/(\Gamma-1)}$.
 % Note that for $\Gamma\neq1$ the disk has a surface $z=H(r)$ such that
@@
 
-The vertically isothermal case, $\Gamma = 1$, can be found either 
-independently or as a limit of Eq. \ref{eqm_dens} to be 
-\begin{align}
+The vertically isothermal case, $\Gamma = 1$, can be calculated either 
+as a special case or as the $\Gamma \rightarrow 1$ limit of Eq. \ref{eqm_dens} to be 
+\begin{subequations}\begin{align}
   \frac{\rho(r,z)}{\rho_0(r)} &=
@@
   & \approx \exp {\left[ - {z^2 \over 2 H^2}\right]} \label{rhoisothin},
-  \end{align}
-  where Eq.\ \ref{rhoisothin} takes the $z \ll r$ thin disk limit.
+  \end{align}\end{subequations}
+  where Eq.\ \ref{rhoisothin} takes the $z \ll r$ thin disk limit.  We will primarily focus on
+vertically isothermal disks, the relevant case for passively irradiated
+PPDs \citep{chiang97}.
 
- \subsubsection{Rotation and vertical shear profiles}
-The equilbrium rotation field follows from the density field and centrifugal balance (Eq.\ \ref{hori_equilibrium}),
+ \subsubsection{Rotation and vertical shear profiles}\label{vshear_def}
+The equilibrium rotation field, $\Omega(r,z)$ follows from the density field and centrifugal balance (Eq.\ \ref{hori_equilibrium}),
 giving 
 \begin{align}\label{eqm_rot}
-  \frac{\Omega^2(r,z)}{\OmK^2(r)}=1 +
-  \left(\frac{p+q}{\Gamma}\right) h^2(r) 
+  \frac{\Omega^2}{\OmK^2}=1 +
+  \frac{p+q}{\Gamma} h^2(r) 
   +\frac{s}{\Gamma} \left(1-\frac{1}{\sqrt{1+z^2/r^2}}\right), 
 \end{align}
-for all $\gamma$, where $s\equiv q+p(1-\Gamma)$.
+for all $\Gamma$, where $s\equiv q+p(1-\Gamma)$.
 
-Notice that for $\Gamma = 1$, $s/\Gamma = q$...
+The vertical shear shear follows as
+\begin{align}
+  r\frac{\p\Omega^2}{\p z} = \frac{sz}{\Gamma r}  \frac{\OmK^2}{
+    \left(1+z^2/r^2\right)^{3/2}} . \label{vertical_shear_ex} 
+\end{align}
+For $\Gamma = 1$, notice that $s/\Gamma = q$ is the only disk parameter that affects vertical shear.
+Vertical shear increases linearly with height near the mid plane, and $\p_z \Omega^2$ is maximized at
+the smaller of $r/\sqrt{2}$ or $H_s$.
 
-...
+A more general expression for vertical shear holds for vertical polytopes 
+(which satisfy Eq.\ \ref{eqm_press}), with no assumed radial structure:
+\begin{align}
+  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
+    z}\frac{dK}{dr} = - \frac{g_z}{\Gamma}\frac{d\ln K}{dr}, \label{vertical_shear}
+\end{align}
+where $g_z = \p_z P/\rho$.
+
 %\subsubsection{Vertically isothermal disks}
@@
 
-\subsection{Vertical shear}\label{vshear_def}
-Note that the disk is in general baroclinic with $\Omega =
-\Omega(r,z)$. This is explicit from Eq. \ref{eqm_rot} and Eq. \ref{vertiso_eqm}, but is also 
-implied by Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} and 
-Eq. \ref{eqm_press},  
-\begin{align}
-  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
-    z}\frac{dK}{dr} = - \frac{g_z}{\Gamma}\frac{d\ln K}{dr}.\label{vertical_shear}
-\end{align}
-For the above disk models,
-\begin{align}
-  r\frac{\p\Omega^2}{\p z} = \frac{sz\OmK^2}{\Gamma
-    r\left(1+z^2/r^2\right)^{3/2}}. \label{vertical_shear_ex} 
-\end{align}
-Note that at fixed $r$ there is a maximum vertical shear rate at $|z|=r/\sqrt{2}$,
-which is expected to limit the growth rate of instabilities associated with vertical shear.
+%\subsection{Vertical shear}\label{vshear_def}
+%Note that the disk is in general baroclinic with $\Omega =
+%\Omega(r,z)$. This is explicit from Eq. \ref{eqm_rot} and Eq. \ref{vertiso_eqm}, but is also 
+%implied by Eq. \ref{vert_equilibrium}---\ref{hori_equilibrium} and 
+%Eq. \ref{eqm_press},  
+%\begin{align}
+%  r\frac{\p\Omega^2}{\p z} = - \rho^{\Gamma-1}\frac{\p\ln\rho}{\p
+%    z}\frac{dK}{dr} = - \frac{g_z}{\Gamma}\frac{d\ln K}{dr}.\label{vertical_shear}
+%\end{align}
+%For the above disk models,
+%\begin{align}
+%  r\frac{\p\Omega^2}{\p z} = \frac{sz\OmK^2}{\Gamma
+%    r\left(1+z^2/r^2\right)^{3/2}}. \label{vertical_shear_ex} 
+%\end{align}
+%Note that at fixed $r$ there is a maximum vertical shear rate at $|z|=r/\sqrt{2}$,
+%which is expected to limit the growth rate of instabilities associated with vertical shear.
 
 \subsection{Entropy gradients and vertical buoyancy}
-The specific entropy is defined as $S\equiv
-c_p\ln{P^{1/\gamma}/\rho}$, where $c_p$ is the heat capacity at
-constant pressure. For our disk models,
+The gradients of entropy, $S\equiv
+C_p\ln{P^{1/\gamma}/\rho}$, in our disk models are
+\begin{subequations}
 \begin{align}
-  &\frac{\p S}{\p r} = c_p\left[\frac{s}{\gamma r} +
+  &\frac{\p S}{\p r} = C_p\left[\frac{s}{\gamma r} +
     \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p
        r}\right],\\
-  &\frac{\p S}{\p z} = c_p
-  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z},\label{dSdz}
+  &\frac{\p S}{\p z} = C_p
+  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z} = {C_p g_z \over c_s^2} {\Gamma - \gamma \over \gamma},\label{dSdz}
 \end{align} 
-which are used below.
+\end{subequations}
+ where $C_p$ is the heat capacity at
+constant pressure.
 
-For discussion purposes it is useful to define the vertical buoyancy
-frequency $N_z$ via 
+The vertical buoyancy frequency $N_z$ is 
 \begin{align}
-  N_z^2 \equiv - c_p^{-1}g_z \frac{\p S}{\p z} = c_s^2 \left(\frac{\gamma -
-      \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,   \label{nzsq_def}
+  N_z^2 \equiv - C_p^{-1}g_z \frac{\p S}{\p z} = {g_z^2 \over c_s^2}{\gamma - \Gamma \over \gamma}. %c_s^2 \left(\frac{\gamma -
+   %   \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,  
+    \label{nzsq_def}
 \end{align}
-where $g_z\equiv \rho^{-1}\p_zP $. The second equality applies to the
-disk equilibria described above (using Eq. \ref{dSdz}). We are
-interested in sub-adiabatically stratified disks with $\Gamma <
-\gamma$ so that $N_z^2\geq0$. % , and the disk is 
+We only consider convectively stable disks with with $\Gamma <
+\gamma$ and  $N_z^2\geq0$. % , and the disk is 
 % convectively stable in the vertical direction.
 
-\subsection{Stability against axisymmetric adiabatic
-  perturbations}\label{solberg}
-A rotating flow is stable against axisymmetric adiabatic perturbations
+\subsection{Stability without cooling}\label{solberg}
+In the absence of cooling, axisymmetric stability is ensured
 if both Solberg-Hoiland criteria  are satisfied:
 \begin{align}
-  \kappa^2 - \frac{1}{\rho c_p}\nabla P \cdot \nabla S &> 0,\\
+  \kappa^2 - \frac{1}{\rho C_p}\nabla P \cdot \nabla S &> 0,\\
   -\frac{\p P}{\p z}\left(\frac{\p j^2}{\p r}\frac{\p S}{ \p z} -
@@
 \citep{tassoul78}, where $j=r^2\Omega$ is the specific angular
-momentum. For rotationally-supported, thin disks %  pressure gradients are small 
+momentum and $\kappa = r^{-3} \p_r j^2 $ is the epicyclic frequency.  For rotationally-supported, thin disks %  pressure gradients are small 
 % compared to rotation, so 

### 2015-04-29 appendix.tex

@@
-\section{Full governing equation for vertically isothermal disks in
-  the vertically global shearing box}\label{adia_improve}
-Here we explain the derivation of Eq. \ref{nearly_iso_explicit} used in our 
-analytical discussion in \S\ref{analytical}. We shall need the
-vertical derivative of the epicyclic frequency $\kappa$. In the
-vertically isothermal global
-disk with $\Gamma=1$ we have 
-\begin{align}\label{dkappa2}
-  \frac{\p\kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
-    r}\frac{\p\Omega^2}{\p z} = -
-  \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2} \left(
-    \frac{3z^2}{r^2+z^2}-1\right) \equiv - \frac{\p\ln\rho}{\p z}\frac{qc_s^2}{r^2}F(r,z).
-\end{align}
-% For the local problem we may then write
-% \begin{align}
-%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z),
-% \end{align}
-The function $F$ increases monotonically from $F=-1$ at $z=0$ to $F\to2$
-as $|z|\to\infty $, so $|F|=O(1)$.
-
-In the vertically global shearing box (VGSB) approximation, we set
-$\zeta=0$ in Eq. \ref{lin_mass} --- \ref{lin_energy}. We then eliminate
-the horizonatal velocity perturbations ($\delta v_x,\, \delta v_y$) to
+\section{Derivation of the approximate equations}\label{adia_improve}
+Here we detail the derivation of Eq. \ref{nearly_iso_explicit} used in the 
+analytical derivations of \S\ref{analytical}.  Starting with Eqs.\ \ref{lin_mass} --- \ref{lin_energy},
+we set $\Gamma = 1$ and $\zeta = 0$ for the vertically isothermal limit and the fully-radially-local 
+approximation, respectively.  We  eliminate
+the horizontal velocity perturbations ($\delta v_x,\, \delta v_y$) to
 obtain  
@@
 %equations for $W,Q$ as 
-\begin{align}
+\begin{subequations}\label{A1}\begin{align}
   & \ii\upsilon\delta v_z = \frac{dW}{dz} + \frac{d\ln{\rho}}{d z}\left(W-Q\right),\label{ode_vz} \\
@@
   \label{ode_Q} 
-\end{align}
-(note that $\Gamma=1$ has been used) where
+\end{align}\end{subequations}
+where
 \begin{align}
@@
 \end{align} 
-and we have replaced the notation $\p_z$ by $d/dz$ since we are
-considering solutions at a fixed radius.
+and we have replaced $\p/\p z$ by $d/dz$ for the fully radially local treatment.
+
+Reduction to a single ODE requires vertical derivatives.  At this point we could use the 
+low frequency and Keplerian approximations to set $D \rightarrow \kappa^2 \rightarrow \OmK^2$,
+ and we would wind up with Eq. \ref{nearly_iso_explicit} more directly.
+However to demonstrate that the order of approximation is irrelevant, we will retain
+the vertical derivatives of the epicyclic frequency $\kappa$,
+\begin{align}\label{dkappa2}
+  \frac{d\kappa^2}{d z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
+    r}\frac{\p\Omega^2}{\p z} = -
+  \frac{d\ln\rho}{d z}\frac{qc_s^2}{r^2} \left(
+    \frac{3z^2}{r^2+z^2}-1\right) \equiv - \frac{d\ln\rho}{d z}\frac{qc_s^2}{r^2}F(r,z).
+\end{align}
+% For the local problem we may then write
+% \begin{align}
+%   \frac{dD}{dz}  = - \frac{d\ln\rho}{dz}\frac{qc_s^2}{r^2}F(z),
+% \end{align}
+The function $F$ increases monotonically from $F=-1$ at $z=0$ to $F\to2$
+as $|z|\to\infty $, so $|F|=O(1)$.
 
@@
 
-We now eliminate $W,\, Q$ from Eq. \ref{ode_vz} --- \ref{ode_Q} in
-favor of $\delta v_z$, making use of Eq. \ref{dkappa2} to evaluate
-$dD/dz$, to obtain
-
+We now eliminate $W,\, Q$ from Eqs.\ \ref{A1}, using  Eqs. \ref{chi} and \ref{dkappa2} to obtain  
 \begin{align}\label{full_ode}
@@
      \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
-   \right\}\delta v_z,
+   \right\}\delta v_z.
 \end{align}
-where we recall $\chi = \left(1-\ii\upsilon t_c\right)/\left(1-\ii\upsilon
-t_c \gamma\right)$. Making the low frequency and Keplerian
-approximations is equivalent to the replacement $D\to
-\OmK^2$. This gives, in terms of dimensionless variables,  
+
+Now we make the low frequency and Keplerian
+approximations, setting $D\to
+\OmK^2$, to give  
 \begin{align}
@@
       \frac{\chi}{\left(\hat{k}^2+\chi\right)}qh^2F\right]\right\}\delta v_z \notag\\&=
-  -\hat{\upsilon}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3}.
+  -\hat{\upsilon}^2\left(\hat{k}^2+\chi\right)\delta v_z\label{adia_diso3},
 \end{align} 
-Since we are interested in thin disks with $h \ll 1$ and $|q| =
-O(1)$, we neglect the terms proportional to $h^2$, giving 
+in terms of dimensionless variables of Eq.\ \ref{nondim}.
+
+Retaining terms to first order in the disk aspect ratio, $h \ll 1$, gives
 \begin{align}
@@
        \ii h
-       q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z,\label{vertiso_gov_nondim}
+       q\hat{k}\right)\left(1 - \chi\right)\right]\delta v_z.\label{vertiso_gov_nondim}
  \end{align}
- which is used in the discussion below.  Eq. \ref{nearly_iso_explicit}, as
- used in our analytical discussion, is obtained by inserting the 
- thin-disk expression for the density field. We comment that the simpler way
- of obtaining Eq. \ref{vertiso_gov_nondim} is to neglect the vertical 
- dependence of $D$ by setting $D\to\OmK^2$ in Eq. \ref{ode_w} before eliminating
- variables in favor of $\delta v_z$.
+Applying Eq.\ \ref{rhoisothin} for the density field gives the desired result, Eq. \ref{nearly_iso_explicit}.
+
+We can establish a correspondence between our  Eq. \ref{nearly_iso_explicit} and
+ Eq. 41 in \cite{lubow93}, which describes adiabatic axisymmetric waves in
+a vertically isothermal disk without vertical shear.   Accounting for the required change of variables, 
+the correspondence is exact after setting $q=0$ (no vertical shear) and
+$\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
+and applying the approximations in our \S \ref{sec:simplified} to \citeauthor{lubow93}.
+
+%We comment that the simpler way
+ %of obtaining Eq. \ref{vertiso_gov_nondim} is to neglect the vertical 
+ %dependence of $D$ by setting $D\to\OmK^2$ in Eq. \ref{ode_w} before eliminating
+ %variables in favor of $\delta v_z$.
 

### 2015-04-29 linear_adia.tex

@@
-\section{Analytic results for vertically isothermal disks}\label{analytical}
+\section{Analytic results with finite cooling times}\label{analytical}
 
-\subsection{Isothermal perturbations}\label{iso_pert}
-Isothermal perturbations correspond to $\beta = 0$ or $\gamma=1$ so 
-that $\chi = 1$. This case has been investigated by \citetalias{nelson13},
-\citetalias{mcnally14} and \citetalias{barker15}.
+%\subsection{Isothermal perturbations}\label{iso_pert}
+Previous analytic studies of the VSI have largely focused on isothermal
+perturbations, with infinitely rapid cooling, as discussed in the introduction.  We further explore
+this $\beta = 0$ limit in Appendix \ref{iso_discuss} both to further develop intuition for this 
+idealized case and to establish a connection with previous 
+works.  However our main interest is the effect of finite cooling times, which
+we explore analytically in this section.
 
-In Appendix \ref{iso_discuss}, we collect some useful results for
-isothermal  perturbations. This includes a necessary condition for
-instability (there must be vertical shear) and an upper bound to the
-instability growth rate (it is limited by the maximum vertical shear
-rate in the disk). Neither of these actually require the thin-disk
-form of the density field described above.  
-We also develop solutions in the thin-disk
-limit, but without assuming low frequency from the onset. 
-The methods used in Appendix \ref{iso_discuss} and results obtained
-therein motivates the solution procedure employed in the following
-sections, and are provided for ease of reference. We refer the reader
-to \citetalias{barker15} for a more detailed treatment of isothermal 
-perturbations. %which also account for more general boundary
-               %conditions than that considered below
+In \S\ref{relax_pert} we prove that even an infinitesimal increase in the cooling time,
+starting from $\beta = 0$, slows the growth rate of the VSI.  In \S\ref{disp_relax} we 
+derive exact solutions to the simplified VSI model developed above (Eq.\ \ref{nearly_iso_explicit}).
+In \S\ref{iso_vsi_beta_crit} we reach our main result, the maximum cooling time above which 
+VSI growth is suppressed.  We heuristically argue for this result in \S\ref{vsi_require} and numerically
+verify it in \S\ref{numerical}.
 
-\subsection{The effect of introducing a small but finite
-  thermal relaxation time}\label{relax_pert}
+%Isothermal perturbations correspond to $\beta = 0$ or $\gamma=1$ so 
+%that $\chi = 1$. This case has been investigated by \citetalias{nelson13},
+%\citetalias{mcnally14} and \citetalias{barker15}.  
+%
+%In Appendix \ref{iso_discuss}, we collect some useful results for
+%isothermal  perturbations. This includes a necessary condition for
+%instability (there must be vertical shear) and an upper bound to the
+%instability growth rate (it is limited by the maximum vertical shear
+%rate in the disk). Neither of these actually require the thin-disk
+%form of the density field described above.  
+%We also develop solutions in the thin-disk
+%limit, but without assuming low frequency from the onset. 
+%The methods used in Appendix \ref{iso_discuss} and results obtained
+%therein motivates the solution procedure employed in the following
+%sections, and are provided for ease of reference. We refer the reader
+%to \citetalias{barker15} for a more detailed treatment of isothermal 
+%perturbations. %which also account for more general boundary
+%               %conditions than that considered below
+
+\subsection{Introducing a small but finite
+  cooling time}\label{relax_pert}
 We are interested in finite thermal relaxation
@@
 (\S\ref{application}). We will see later that consideration of low
-order modes in fact provides a useful way to charactertize the effect
+order modes in fact provides a useful way to characterize the effect
 of increasing the thermal timescale on the VSI.
@@
 \end{align}
-where $\lambda$ is a constant to be chosen for convenience. Inserting
+where the constant $\lambda$ is a constant to be chosen for convenience. Inserting
 Eq. \ref{adia_ansatz} into Eq. \ref{nearly_iso_explicit} gives
@@
 
-\subsection{Upper limit on the thermal relaxation timescale for the
-  VSI}\label{iso_vsi_beta_crit}
+\subsection{An upper limit to the cooling time}\label{iso_vsi_beta_crit}
 Here we estimate the maximum thermal relaxation timescale 
@@
    0= &2M(M+1) + \frac{\beta}{h q} (1-\gamma)(1+2M)^2
-   \left(\hat{\omega}\khat\right) + \left(\hat{\omega}\khat\right)^2. \label{relax_cond_simp2}
+   \left(\hat{\omega}\khat\right) + \left(\hat{\omega}\khat\right)^2, \label{relax_cond_simp2}
 \end{align}
-(We explain this in Appendix \ref{disp_neut_limit}.)
+as explained in  Appendix \ref{disp_neut_limit}.
 

### 2015-04-29 linear_iso.tex

@@
 
-In preparation for a numerical study of the above equations, we will 
-first give a fully analytic discussion for vertically isothermal
-disks, for which $\Gamma=1$ and $c_s$ is vertically constant. To this end, 
-we introduce the following simplifications:
-%In Appendix \ref{adia_improve} we derive  a with the following simplifications:
+Our simplified model starts with  Eqs.\ \ref{lin_all} 
+and makes the following additional approximations:
 
@@
 
-\item  We adopt the  `vertically global shearing box' (VGSB) formalism 
-  developed by \citetalias{mcnally14}, in which explicit dependencies on the
-  background radial disk structure in 
-  Eq. \ref{lin_mass}---\ref{lin_energy} are neglected by setting
-  $\zeta=0$. However, the global radial temperature 
-  gradient in the expression for the 
-  vertical shear (Eq. \ref{vertical_shear}) is retained.
+ \item We set $\Gamma = 1$, focusing on vertically isothermal disks.
 
-\item We consider inertial waves with eigenfrequencies such that
-  $|\upsilon^2|\ll \kappa^2,\,\Omega^2$ \citepalias{barker15}. We call
-  this the low-frequency approximation.
+\item  We set $\zeta = 0$,  dropping terms with an explicit dependence on the 
+radial structure of the equilibrium disk.  Vertical shear, which implicitly depends on the radial
+temperature gradient, is retained.  This fully-radially-local approximation is also made in the
+ `vertically global shearing box' (VGSB) of \citetalias{mcnally14}.
 
-\item We are mainly interested in the effect of vertical shear, 
-%  i.e. $q\neq
-%  0$ for vertically isothermal disks. 
-  which is already captured
-  explicitly in the linearized equations through the term
-  $\p_z\Omega$. We thus ignore the vertical dependence of $\Omega$ and
-  $\kappa$ elsewhere, and replace them with
-  $\OmK$.  We call this the Keplerian approximation.
+\item We make the low frequency approximation, $|\upsilon^2|\ll \kappa^2,\,\Omega^2$,
+keeping only the lowest order terms in $|\upsilon|/\kappa$.  Similar to the 
+incompressible \citepalias{goldreich67} or anelastic \citepalias{barker15} approximations,
+the low frequency approximation filters p-modes in favor of g-modes, i.e.\ inertial
+gravity waves \citep{lubow93}.
 
-\item We assume thin disks for which $h\ll 1$, and write 
-  $\rho/\rho_0\simeq  \exp{(-z^2/2H^2)}$.
+\item We make the Keplerian approximation, setting $\Omega \rightarrow \OmK$ 
+and $\kappa \rightarrow \OmK$, but retaining the vertical dependence 
+in the crucial vertical shear term,
+  $\p_z\Omega$.
+
+\item We assume thin disk with $h\ll 1$ and take the $z \ll r$ limit to 
+  obtain Eq.\ \ref{rhoisothin} for the equilibrium $\rho(z)$.
 
@@
  In terms of the dimensionless variables
-\begin{align}
+\begin{align}\label{nondim}
   \hat{z} = z/H,\quad \hat{k}=k_xH, \quad \hat{\upsilon} =\hat{\omega} +
@@
 where $^\prime$ denotes $d/d\zhat$ and 
-\begin{align}
+\begin{subequations}\begin{align}
   &A \equiv 1 + \ii h q \hat{k},\\
@@
     h q\hat{k}\right), 
-\end{align}
+\end{align}\end{subequations}
 where
-\begin{align}
+\begin{align}\label{chi}
 \chi = \frac{1-\ii\hat{\upsilon}\beta}{1-\ii\hat{\upsilon}\beta\gamma}.
 \end{align}
-The derivation of Eq. \ref{nearly_iso_explicit}  is outlined in
+The derivation of Eq. \ref{nearly_iso_explicit}  is detailed in
 Appendix \ref{adia_improve}.
 
-We remark that Eq. \ref{nearly_iso_explicit} is in the same 
-form as Eq. 41 in \cite{lubow93}, which describe adiabatic axisymmetric waves in
-a vertically isothermal disk without vertical shear. The two equations
-become identical after  setting $q=0$ (no vertical shear) and
-$\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
-and making the low-frequency, Keplerian approximation in
-\citeauthor{lubow93} after a change of variables.
+%We remark that Eq. \ref{nearly_iso_explicit} is in the same 
+%form as Eq. 41 in \cite{lubow93}, which describe adiabatic axisymmetric waves in
+%a vertically isothermal disk without vertical shear. The two equations
+%become identical after  setting $q=0$ (no vertical shear) and
+%$\chi=1/\gamma$ (adiabatic flow) in our Eq. \ref{nearly_iso_explicit},
+%and making the low-frequency, Keplerian approximation in
+%\citeauthor{lubow93} after a change of variables.
 
- In Appendix \ref{global_corr}  we discuss limitations of using
-  the VGSB, where we show that this approximation is valid provided we 
-  consider fast cooling times
-  ($\beta\lesssim 1$) and perturbations with small radial lengthscales
-  ($\khat\gtrsim 1$), as appropriate for the VSI.
+ In Appendix \ref{global_corr}  we discuss the limits of these approximations
+ and show that they are valid in the parameter space of interest for the VSI.
+In particular we point out that the fully-radially-local approximation is only valid for
+short cooling times ($\beta \lesssim 1$).
 

### 2015-04-29 appendix.tex
minute changes to sec 3, 4, 5, appendix A; partial re-write of appendix B; new file for fig 1 not found, please update

@@
 \begin{subequations}\label{A1}\begin{align}
-  & \ii\upsilon\delta v_z = \frac{dW}{dz} + \frac{d\ln{\rho}}{d z}\left(W-Q\right),\label{ode_vz} \\
+  & \ii\upsilon\delta v_z = \frac{dW}{dz} + \frac{\p\ln{\rho}}{\p z}\left(W-Q\right),\label{ode_vz} \\
   &\frac{\upsilon^2}{c_s^2}Q + \frac{\upsilon^2k_x^2}{D}W =\ii\upsilon \left(\frac{\ii
-      k_x r}{D}\frac{d\Omega^2}{dz} -
-    \frac{d\ln{\rho}}{dz}\right)\delta v_z - \ii\upsilon \frac{d\delta v_z}{dz},
+      k_x r}{D}\frac{\p\Omega^2}{\p z} -
+    \frac{\p \ln{\rho}}{\p z}\right)\delta v_z - \ii\upsilon \frac{d\delta v_z}{dz},
   % \left[\frac{dW}{dz} + \frac{d\ln{\rho}}{dz}\left(W-Q\right)\right] 
@@
   \frac{\ii\upsilon}{t_c}\left(W-Q\right) 
-  =\ii \upsilon c_s^2\frac{d\ln{\rho}}{dz}\left(\gamma -
+  =\ii \upsilon c_s^2\frac{\p\ln{\rho}}{\p z}\left(\gamma -
     1\right)\delta v_z,
@@
 \begin{align}
-  D \equiv \kappa^2 - \upsilon^2,
+  D \equiv \kappa^2 - \upsilon^2.
 \end{align} 
-and we have replaced $\p/\p z$ by $d/dz$ for the fully radially local treatment.
+%
 
-Reduction to a single ODE requires vertical derivatives.  At this point we could use the 
-low frequency and Keplerian approximations to set $D \rightarrow \kappa^2 \rightarrow \OmK^2$,
- and we would wind up with Eq. \ref{nearly_iso_explicit} more directly.
-However to demonstrate that the order of approximation is irrelevant, we will retain
-the vertical derivatives of the epicyclic frequency $\kappa$,
+Reduction to a single ODE requires 
+$\p_zD = \p_z\kappa^2$.  At this point we could apply the 
+low frequency and Keplerian approximations to set $D \rightarrow
+\kappa^2 \rightarrow \OmK^2$, then $D$ is vertically constant, 
+and we can obtain Eq. \ref{nearly_iso_explicit} more directly.
+However, to demonstrate that the order of approximation is irrelevant,
+we will retain $\p_z\kappa^2$ initially. Using 
+Eq. \ref{kap2_def}---\ref{vertical_shear},  
 \begin{align}\label{dkappa2}
-  \frac{d\kappa^2}{d z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
+  \frac{\p \kappa^2}{\p z} = 4 \frac{\p\Omega^2}{\p z} + r\frac{\p}{\p
     r}\frac{\p\Omega^2}{\p z} = -
@@
 
-We now eliminate $W,\, Q$ from Eqs.\ \ref{A1}, using  Eqs. \ref{chi} and \ref{dkappa2} to obtain  
+We  eliminate $W,\, Q$ from Eqs.\ \ref{A1}, using  Eqs. \ref{chi} and \ref{dkappa2} to obtain  
 \begin{align}\label{full_ode}
@@
      \frac{\chi}{\left(k_x^2c_s^2 + \chi D\right)}\frac{qc_s^2 F}{r^2}\right] 
-   \right\}\delta v_z.
+   \right\}\delta v_z, 
 \end{align}
-
-Now we make the low frequency and Keplerian
+where we have replaced $\p/\p z$ by $d/dz$ for the
+fully-radially-local treatment. Now we make the low frequency and Keplerian
 approximations, setting $D\to 
@@
 \end{align}
-Applying Eq.\ \ref{rhoisothin} for the density field gives the desired result, Eq. \ref{nearly_iso_explicit}.
+Approximating the density field by Eq.\ \ref{rhoisothin} then gives 
+Eq. \ref{nearly_iso_explicit}.
 
@@
 
-\section{Applicability of the  VGSB}\label{global_corr}
-%In \S\ref{analytical} we effectively employed the vertically global
-%shearing box \citep[VGSB,][]{mcnally14} to make analytic progress. 
-The vertically global shearing box \citepalias[VGSB,][]{mcnally14} is an
-extension of the standard shearing box \citep[SSB,][]{goldreich65} to
-background shear flows that are height dependent. 
-%As in the SSB, in
-%the VGSB the equilibrium disk structure is assumed to be homogeneous
-%in the radial ($x$) direction.  
-Within the VGSB framework, the vertical shear $d\Omega/dz$
-can be freely specified. We choose $d\Omega/dz$ to be that due to the
-radial temperature gradient in the global disk
-(Eq. \ref{vertical_shear}). However, explicit dependencies on the
-radial disk structure are ignored elsewhere in the linearized equations
-as described below. We show in the following sections that this
-inconsistency only becomes important in the adiabatic limit, which is
-not the relevant regime for the VSI.
+\section{Applicability of the fully-radially-local approximation}\label{global_corr}
+In the fully-radially-local approximation, background radial
+gradients are ignored except where it appears
+implicitly in the expression for the vertical shear rate $\p_z\Omega$
+(via Eq. \ref{vertical_shear}).  This is done by neglecting the terms proportional to $\zeta$
+in Eq. \ref{lin_mass}, \ref{lin_vx} and \ref{lin_energy}, i.e. setting 
+$\zeta=0$. (Nomially $\zeta=1$, which is used in our numerical study.)
 
-%is only important when
-%considering adiabatic limit, which is not the relevant regime for the
-%VSI.
-
-The background radial disk structure explicitly appears 
-in the linearized mass, radial momentum and energy equations
-(Eq. \ref{lin_mass}, \ref{lin_vx} and \ref{lin_energy}) as the terms
-proportional to $\zeta$. Nominally $\zeta=1$.      
-In the VGSB (and SSB) these terms are neglected by setting
-$\zeta=0$. For a power-law disk, these radial gradients  
+For a power-law disk, the neglected radial gradients   
 are $O(r^{-1})$, and they appear in comparison with terms of
@@
 $O(h/\khat)$, which is small for thin disks ($h\ll1$) 
-and/or small radial wavelengths ($\khat\gg 1$).
+and/or small radial wavelengths ($\khat\gg 1$).  We show in the
+following sections that the fully-radially-local approximation only 
+becomes invalid in the adiabatic limit, which is not the relevant
+regime for the VSI.
 
-% The difference between the linear problem in the VGSB 
-% and the radially-local approximation adopted in our numerical study
-% is the absence of terms explicitly associated with the radial disk structure in
-% the VGSB.
+We comment that this approximtaion is equivalent to 
+that adopted in the vertically global shearing box formalism
+\citepalias[VGSB,][]{mcnally14}, which is an 
+extension of the standard shearing box \citep{goldreich65} to
+background shear flows that are height dependent.
 
@@
 
-\subsection{Spurious growth of adiabatic perturbations in the VGSB}\label{analytic_adia} 
-A limitation of the VGSB is that it cannot be used for 
-adiabatic flow, even if the above conditions are satisfied. We explain this by considering vertically isothermal
-disks and setting  $\beta\to\infty$ and hence $\chi = 1/\gamma$ in 
+\subsection{Spurious growth of adiabatic perturbations when
+  $\zeta=0$}\label{analytic_adia}  
+A limitation of the reduced model described in Appendix
+\ref{adia_improve}, \S\ref{sec:simplified} and used in \S\ref{analytical}, is that it
+cannot be employed for adiabatic flow when there is vertical shear, even
+if $h/\khat\ll1$.  We explain this by setting  $\beta\to\infty$ and hence $\chi = 1/\gamma$ in 
 Eq. \ref{vertiso_gov_nondim} % (which itself is obtained by setting
@@
 %also need real freq^2 in adia limit 
-This is because the VGSB retains the global radial
-disk structure responsible for vertical shear
-($\p_z\Omega$), but ignores it elsewhere in the linear problem. 
-Nevertheless, we demonstrate below
+This is because we have retained the global radial temperature 
+gradient to obtain vertical shear, 
+but have ignored it elsewhere in the linear problem (as well as the
+background radial density gradient). Nevertheless, we demonstrate below
 that this inconsistency is unimportant for the VSI, which occurs 
@@
 agreement for $\beta\lesssim 1$. Differences arise for 
-$\beta\gtrsim1$, and as $\beta\to\infty$ the VGSB framework gives a
+$\beta\gtrsim1$, and as $\beta\to\infty$ the fully-radially-local
+approximation, where $\zeta=0$, gives a
 (spurious) growth rate as expected from the discussion
 above. Inclusion of the global radial gradient terms results in the
-correct behavior  ($\sigma\to0$ as $\beta\to\infty$). Despite this
+expected behavior  ($\sigma\to0$ as $\beta\to\infty$). Despite this
 caveat, Fig. \ref{gcorr_compare} shows that provided we consider 
-$\beta\ll1$, then the VGSB framework is adequate for studying the
-VSI.
+$\beta\ll1$, then setting $\zeta=0$ does not affect the VSI
+significantly.
 
@@
   \caption{Growth rate of the fundamental VSI mode as a function of
-    the thermal relaxation time $\beta$. The `Analytical (VGSB)' curve
-    is calculated from the dispersion relation Eq. \ref{relax_disp};
-    the `Numerical (VGSB)' curve is obtained by numerically solving
-    the linear problem, Eq. \ref{lin_mass}---\ref{lin_energy}, with $\zeta=0$
-    (which defines the linear VGSB equations); and the `Numerical
-    (Full)' curve is obtained by numerically solving 
-    Eq. \ref{lin_mass}---\ref{lin_energy}  with $\zeta=1$, which
-    accounts for the radial disk structure.  
+    the thermal relaxation time $\beta$. The `Analytical ($\zeta=0$)' curve
+    is calculated from the dispersion relation Eq. \ref{relax_disp}. 
+    The `Numerical ($\zeta=0$)' curve is obtained from the numerical
+    solution to Eq. \ref{lin_mass}---\ref{lin_energy} with
+    $\zeta=0$,  which 
+    neglects explicit dependencies on the radial disk structure.  
+    The `Numerical
+    ($\zeta=1$)' curve is obtained from Eq. \ref{lin_mass}---\ref{lin_energy} with
+    $\zeta=1$, which accounts for the radial disk structure self-consistently.   
     The disk parameters are $(\gamma, \Gamma)=(1.4, 1.011)$ and
@@
 perturbations ($\beta\equiv 0$) in vertically isothermal disks 
-($\Gamma=1$) in the VGSB framework. In this case Eq. \ref{ode_Q}
+($\Gamma=1$) in the fully-radially local approximation
+($\zeta=0$). In this case Eq. \ref{ode_Q}
 becomes    

### 2015-04-29 application.tex
minute changes to sec 3, 4, 5, appendix A; partial re-write of appendix B; new file for fig 1 not found, please update

@@
   \frac{\p P}{\p t} + \bm{v}\cdot\nabla P +\gamma P \nabla\cdot\bm{v} = - \frac{P}{\rho T
-    c_v}\left(\nabla\cdot\bm{F} - Q_+\right), 
+    C_v}\left(\nabla\cdot\bm{F} - Q_+\right), 
 \end{align}
-where $c_v$ is the heat capacity at constant
+where $C_v$ is the heat capacity at constant
 volume, 
@@
 \begin{align}\label{diff_cool_proper}
-  \frac{P}{\rho T c_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
+  \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
     T\right),  
@@
 \begin{align}\label{diff_cool_approx}
-  \frac{P}{\rho T c_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
+  \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
     T\right) &\to -\frac{k_\mathrm{rad}}{\rho
-    c_v}k_x^2P\left(\frac{\delta T}{T}\right)\notag\\
+    C_v}k_x^2P\left(\frac{\delta T}{T}\right)\notag\\
   &\equiv -\eta k_x^2 \left(\delta P - \frac{P}{\rho}\delta\rho\right), 
 \end{align}
-\citepalias[see also][]{barker15} where $\eta=k_\mathrm{rad}/\rho c_v$ is
+\citepalias[see also][]{barker15} where $\eta=k_\mathrm{rad}/\rho C_v$ is
 the diffusion coefficient and 
@@
 \begin{align}\label{eta_def}
-  \eta = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 c_v}, 
+  \eta = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 C_v}, 
 \end{align}

### 2015-04-29 linear_adia.tex
minute changes to sec 3, 4, 5, appendix A; partial re-write of appendix B; new file for fig 1 not found, please update

@@
 
-In \S\ref{relax_pert} we prove that even an infinitesimal increase in the cooling time,
+In \S\ref{relax_pert} we show that even an infinitesimal increase in the cooling time,
 starting from $\beta = 0$, slows the growth rate of the VSI.  In \S\ref{disp_relax} we 
@@
 In \S\ref{iso_vsi_beta_crit} we reach our main result, the maximum cooling time above which 
-VSI growth is suppressed.  We heuristically argue for this result in \S\ref{vsi_require} and numerically
+VSI growth is suppressed.  We heuristically argued for this result in
+\S\ref{vsi_require}, and will numerically
 verify it in \S\ref{numerical}.

### 2015-04-29 linear_iso.tex
minute changes to sec 3, 4, 5, appendix A; partial re-write of appendix B; new file for fig 1 not found, please update

@@
 Our simplified model starts with  Eqs.\ \ref{lin_all} 
-and makes the following additional approximations:
+and makes the following additional simplifications:
 
@@
 
-\item  We set $\zeta = 0$,  dropping terms with an explicit dependence on the 
+\item  We set $\zeta = 0$,  neglecting terms with an explicit dependence on the 
 radial structure of the equilibrium disk.  Vertical shear, which implicitly depends on the radial
 temperature gradient, is retained.  This fully-radially-local approximation is also made in the
- `vertically global shearing box' (VGSB) of \citetalias{mcnally14}.
+ `vertically global shearing box' of \citetalias{mcnally14}.
 
-\item We make the low frequency approximation, $|\upsilon^2|\ll \kappa^2,\,\Omega^2$,
-keeping only the lowest order terms in $|\upsilon|/\kappa$.  Similar to the 
-incompressible \citepalias{goldreich67} or anelastic \citepalias{barker15} approximations,
+\item We make the low frequency approximation, assuming $|\upsilon^2|\ll \kappa^2,\,\Omega^2$. 
+%keeping only the lowest order terms in $|\upsilon|/\kappa$.  
+Similar to the incompressible \citepalias{goldreich67} or anelastic \citepalias{nelson13,barker15} approximations,
 the low frequency approximation filters p-modes in favor of g-modes, i.e.\ inertial
@@
 
-\item We assume thin disk with $h\ll 1$ and take the $z \ll r$ limit to 
-  obtain Eq.\ \ref{rhoisothin} for the equilibrium $\rho(z)$.
+\item We consider thin disks with $h\ll 1$ 
+  %and take the $z \ll r$ limit to 
+  %obtain 
+  and use the Gaussian approximation, Eq.\ \ref{rhoisothin}, for the equilibrium density field.
 

### 2015-04-29 paper.tex
minute changes to sec 3, 4, 5, appendix A; partial re-write of appendix B; new file for fig 1 not found, please update

@@
 prone to viscous damping, including the effective viscosity from turbulent motions.
-While our linear, inviscid analysis does not address this issue directly, we expect that 
+While our linear, inviscid calculations do not address this issue directly, we expect that 
 shorter wavelength modes will drive weaker non-linear (and non-axisymmetric) transport.

### 2015-04-29 setup.tex
minute changes to sec 3, 4, 5, appendix A; partial re-write of appendix B; new file for fig 1 not found, please update

@@
 c_s\sqrt{\gamma/\Gamma}$, sound speeds. 
-
 We also introduce the characteristic scale-height, $H = c_s(r,0)/\OmK$
@@
 Note that for $\Gamma > 1 +  h^2$ there is a disk surface $H_s$
-where $\rho(r,H_s)=0$.  For $h \ll 1$, this surface is at $H_s/H \approx \sqrt{2/(\Gamma-1)}$.
+where $\rho(r,H_s)=0$; and $H_s \simeq \sqrt{2/(\Gamma-1)}H$ for $h\ll 1$.
+%For $h \ll 1$, this surface is located at $H_s \simeq \sqrt{2/(\Gamma-1)}H$. 
 % Note that for $\Gamma\neq1$ the disk has a surface $z=H(r)$ such that
@@
 The vertically isothermal case, $\Gamma = 1$, can be calculated either 
-as a special case or as the $\Gamma \rightarrow 1$ limit of Eq. \ref{eqm_dens} to be 
+as a special case or as the $\Gamma \rightarrow 1$ limit of Eq. \ref{eqm_dens},    
 \begin{subequations}\begin{align}
-  \frac{\rho(r,z)}{\rho_0(r)} &=
+\lim_{\Gamma\to1}  \frac{\rho(r,z)}{\rho_0(r)} &=
   \exp{\left[\frac{1}{ h^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},   \\
-  & \approx \exp {\left[ - {z^2 \over 2 H^2}\right]} \label{rhoisothin},
+  & \simeq \exp {\left( - {z^2 \over 2 H^2}\right)} \quad\text{for } |z|\ll r \label{rhoisothin}. 
   \end{align}\end{subequations}
-  where Eq.\ \ref{rhoisothin} takes the $z \ll r$ thin disk limit.  We will primarily focus on
+Eq.\ \ref{rhoisothin} is the approximate form of the density field in the thin-disk limit.  
+We will primarily focus on 
 vertically isothermal disks, the relevant case for passively irradiated
@@
 \begin{align}\label{eqm_rot}
-  \frac{\Omega^2}{\OmK^2}=1 +
+  \frac{\Omega^2(r,z)}{\OmK^2(r)}=1 +
   \frac{p+q}{\Gamma} h^2(r) 
@@
 \end{align}
-for all $\Gamma$, where $s\equiv q+p(1-\Gamma)$.
+for all $\Gamma$, where $s\equiv q+p(1-\Gamma)$.  We also refer to the
+epicyclic frequency $\kappa$ defined via
+\begin{align}\label{kap2_def}
+\kappa^2 = \frac{1}{r^3}\frac{\p}{\p r}\left(r^4\Omega^2\right). 
+\end{align}
 
-The vertical shear shear follows as
+The vertical shear rate follows from Eq. \ref{eqm_rot}, 
 \begin{align}
@@
 For $\Gamma = 1$, notice that $s/\Gamma = q$ is the only disk parameter that affects vertical shear.
-Vertical shear increases linearly with height near the mid plane, and $\p_z \Omega^2$ is maximized at
-the smaller of $r/\sqrt{2}$ or $H_s$.
+Vertical shear increases linearly with height near the midplane, and $|\p_z \Omega^2|$ is maximized at
+$\mathrm{min}(r/\sqrt{2},H_s)$, which is expected to limit the maximum VSI growth rate. 
+%the smaller of $r/\sqrt{2}$ or $H_s$.
 
@@
 Henceforth all unperturbed fluid variables, including 
-gradients such as $\p \Omega/ \p z$, refer to  equilibrium values at $(r_0, z)$.
+gradients such as $\p_z \Omega$, refer to  equilibrium values at $(r_0, z)$.
 
@@
 system of equations are
-\begin{align}
+\begin{subequations}\label{lin_all}\begin{align}
   \ii\upsilon \frac{Q}{c_s^2}  &=  \ii k_x \delta v_x + \frac{d\dd
@@
   % c_s^2\frac{d\ln{\rho}}{dz}\left(\frac{\gamma}{\Gamma}-1\right)\delta v_z.\label{lin_energy}
-\end{align}
+\end{align}\end{subequations}
 This is a set of ordinary differential equations (ODEs) in

### 2015-05-01 appendix.tex
edits in progress to 5.3 and appendix D...

@@
 
-\section{Coefficients for the dispersion relation for perturbations
-with thermal relaxation}\label{relax_coeff}
-The coefficients of the dispersion relation Eq. \ref{relax_disp} is
-given by:
-\begin{align}
+\section{Analytic dispersion relation
+with thermal relaxation}
+\subsection{Coefficients}\label{relax_coeff}
+The coefficients of the dispersion relation Eq.\ \ref{relax_disp} are:
+\begin{subequations}\label{coeffs}\begin{align}
   &c_0 = M(M+1)A^2,\\
@@
 &c_6 = \beta^2\left(1+\gamma\khat^2\right)^2.
+\end{align}\end{subequations}
+
+\subsection{Finding marginal stability}\label{disp_neut_limit}
+To investigate marginal stability we set $\hat{\sigma} = 0$ so the frequency, $\hat{\upsilon}=\hat{\omega}_c$, is real.  
+We take the short radial wavelength limit, $\khat^2\gg 1$, of the coefficients in Eq.\ \ref{coeffs}.  The real and 
+imaginary parts of Eq. \ref{relax_disp} then give relations for $\hat{\omega}_c$ and $\beta_c$, the cooling time for 
+marginal stability, as 
+\begin{align}\
+  0 =&M(M+1)(1-h^2 q^2 \khat^2) + 4\beta_c h q M(M+1)\left(\hat{\omega}_c\khat\right) 
+ +\left[1 +
+    \gamma\beta_c^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}_c\khat\right)^2 \label{relax_cond1} \\
+ &+ 2h q \gamma\beta_c \left(\hat{\omega}_c\khat\right)^3 -  \left(\hat{\omega}_c\khat\right)^4 
+  + \beta_c^2\gamma^2\hat{\omega}_c^2
+  \left(\hat{\omega}_c\khat\right)^4, \notag \\
+   0=& 2h q M (M+1)\khat +
+   \beta_c(1-\gamma)(1+2M)^2\khat\left(\hat{\omega}_c\khat\right) 
+   + h q \khat \left(\hat{\omega}_c\khat\right)^2 -
+   2\beta_c\hat{\omega}_c\left(\hat{\omega}_c\khat\right)^2
+   - h q
+   \gamma^2\beta_c^2\hat{\omega}_c\left(\hat{\omega}_c\khat\right)^3 \label{relax_cond2} \\
+   &+
+   2\beta_c\gamma\hat{\omega}_c\left(\hat{\omega}_c\khat\right)^4 \notag. 
 \end{align}
+We next (re)-apply the low frequency and thin approximations, $h, \hat{\omega}_c \ll 1$ and 
+further apply the rapid cooling limit $\beta_c \ll 1$.  We retain terms with the combinations 
+$\hat{\omega}_c \hat{k}$, $\beta_c/h$, and $h \hat{k}$ which could be order unity.  For inertial waves we know that
+$\hat{\omega}_c \sim 1/\hat{k}$ holds, while $\beta_c/h \sim O(1)$ will be justified \emph{a posteriori}
+and the magnitude of $h \hat{k}$ depends on the choice of wavenumber.  Applying this ordering we 
+arrive at Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}.
 
-\subsection{Limiting form with real frequencies and $\khat^2\gg 1$}\label{disp_neut_limit}
-Assuming neutral modes with real $\hat{\upsilon}=\hat{\omega}$ and $\khat^2\gg 1$, the real and imaginary parts of
-Eq. \ref{relax_disp} are approximately 
-\begin{align}
-  0 =&M(M+1)(1-h^2 q^2 \khat^2) + 4\beta h q M(M+1)\left(\hat{\omega}\khat\right) 
- +\left[1 +
-    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
- + 2h q \gamma\beta \left(\hat{\omega}\khat\right)^3 -  \left(\hat{\omega}\khat\right)^4\notag \\
-  &+ \beta^2\gamma^2\hat{\omega}^2
-  \left(\hat{\omega}\khat\right)^4,\label{relax_cond1} \\
-   0=& 2h q M (M+1)\khat +
-   \beta(1-\gamma)(1+2M)^2\khat\left(\hat{\omega}\khat\right) 
-   + h q \khat \left(\hat{\omega}\khat\right)^2 -
-   2\beta\hat{\omega}\left(\hat{\omega}\khat\right)^2
-   - h q
-   \gamma^2\beta^2\hat{\omega}\left(\hat{\omega}\khat\right)^3 
-   +
-   2\beta\gamma\hat{\omega}\left(\hat{\omega}\khat\right)^4 \label{relax_cond2}. 
-\end{align}
-These equations are used to derive a critical thermal timescale in
-\S\ref{iso_vsi_beta_crit}. For fixed $M=O(1)$,  we assume
-$|\hat{\omega}\khat|$ remain constant but consider $\khat\gg 1$ so
-that $|\hat{\omega}|\ll 1$.  Then for a thin disk $h \ll 1$
-we neglect the second, fourth and last term in Eq. \ref{relax_cond1};
-and neglect the last three terms in Eq. \ref{relax_cond2}. This
-approximation gives Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2},
-but is not valid if $\beta$ or $|q|$ is large.
+%For fixed $M=O(1)$,  we assume
+%$|\hat{\omega}\khat|$ remain constant but consider $\khat\gg 1$ so
+%that $|\hat{\omega}|\ll 1$.  Then for a thin disk $h \ll 1$
+%we neglect the second, fourth and last term in Eq. \ref{relax_cond1};
+%and neglect the last three terms in Eq. \ref{relax_cond2}. This
+%approximation gives Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2},
+%but is not valid if $\beta$ or $|q|$ is large.
 
-For the fundamental mode $M=0$ and Eq. \ref{relax_cond1} ---
-\ref{relax_cond2} simplify significantly. It is then straight
-forward to eliminate $\hat{\omega}\khat$ between 
-Eq. \ref{relax_cond_simp1} and \ref{relax_cond_simp2} to obtain
-$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}) without
-assuming $|h q \khat|\lesssim 1$, as done in the main text to obtain the
-equivalent expression for $M>0$ (Eq. \ref{bcrit_gen}).
+%For the fundamental mode $M=0$ and Eq. \ref{relax_cond1} ---
+%\ref{relax_cond2} simplify significantly. It is then straight
+%forward to eliminate $\hat{\omega}\khat$ between 
+%Eq. \ref{relax_cond_simp1} and \ref{relax_cond_simp2} to obtain
+%$\beta_\mathrm{crit}$ (Eq. \ref{iso_vsi_cond}) without
+%assuming $|h q \khat|\lesssim 1$, as done in the main text to obtain the
+%equivalent expression for $M>0$ (Eq. \ref{bcrit_gen}).
 

### 2015-05-01 linear_adia.tex
edits in progress to 5.3 and appendix D...

@@
 Here we estimate the maximum thermal relaxation timescale 
-$\beta$ that allows the VSI.  
-Our starting point is Eq. \ref{relax_cond1}---\ref{relax_cond2} which
-approximates the dispersion relation Eq. \ref{relax_disp} assuming marginal
-stability and $\khat^2\gg 1$. We consider $M=0$ (the 
-fundamental mode) or  $M=O(1)$ and low frequency $|\hat{\omega}|\ll
-1$ such that $|\hat{\omega}\khat|$ is $O(1)$. We 
-also assume $|q|$ and $\beta$ are $O(1)$ or smaller. For a thin disk
-($h \ll 1$) we have 
-\begin{align}
- 0 = &M(M+1)\left(1-h^2q^2\khat^2\right) \notag\\ 
- &+ \left[1 +
-    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2
-   -  \left(\hat{\omega}\khat\right)^4,\label{relax_cond_simp1}\\
-   0= &2M(M+1) + \frac{\beta}{h q} (1-\gamma)(1+2M)^2
-   \left(\hat{\omega}\khat\right) + \left(\hat{\omega}\khat\right)^2, \label{relax_cond_simp2}
-\end{align}
-as explained in  Appendix \ref{disp_neut_limit}.
+$\beta_c$ that allows growth the VSI.  In Appendix \ref{disp_neut_limit}
+we derive the following relation between $\beta_c$ and the wave frequency,
+\hat{\omega}_c$  at marginal stability to the VSI
+
+%Our starting point is Eq. \ref{relax_cond1}---\ref{relax_cond2} which
+%approximates the dispersion relation Eq. \ref{relax_disp} assuming marginal
+%stability and $\khat^2\gg 1$. We consider $M=0$ (the 
+%fundamental mode) or  $M=O(1)$ and low frequency $|\hat{\omega}|\ll
+%1$ such that $|\hat{\omega}\khat|$ is $O(1)$. We 
+%also assume $|q|$ and $\beta$ are $O(1)$ or smaller. 
+%
+%For a thin disk
+%($h \ll 1$) we have 
+\begin{subequations}\begin{align}
+ \left(\hat{\omega}\khat\right)^4  = &  \left[1 +
+    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2 \label{relax_cond_simp1}\\ 
+ &+ M(M+1)\left(1-h^2q^2\khat^2\right)
+    , \notag \\
+   \left(\hat{\omega}\khat\right)^2 = & \frac{\beta}{h q} (\gamma-1)(1+2M)^2
+   \left(\hat{\omega}\khat\right) \label{relax_cond_simp2}\\
+   $- 2M(M+1) , \notag
+\end{align}\end{subequations}
+These equations hold for $\beta_c \ll 1$ and $\hat{k} \sim 1/|\
 
@@
 $|h q \khat |^2\lesssim 1$,
-then an approximate solution to Eq. \ref{relax_cond_simp1} is 
+
+An approximate solution to Eq. \ref{relax_cond_simp1} for the frequency
 \begin{align}
@@
 \end{align}
+holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
 Note that for $q=0$, Eq. \ref{marg_freq_approx} reduces to 

### 2015-05-01 linear_adia.tex
edits to 5, mainly 5.3.  Please check, especially the footnote about the correspondence with appendix C which maybe isn't exact?

@@
 In \S\ref{iso_vsi_beta_crit} we reach our main result, the maximum cooling time above which 
-VSI growth is suppressed.  We heuristically argued for this result in
-\S\ref{vsi_require}, and will numerically
-verify it in \S\ref{numerical}.
+VSI growth is suppressed.
 
@@
 we derive the following relation between $\beta_c$ and the wave frequency,
-\hat{\omega}_c$  at marginal stability to the VSI
+$\hat{\omega}_c$  at marginal stability to the VSI:
+\begin{subequations}\begin{align}
+ \left(\hat{\omega}_c\khat\right)^4  = &  
+% \left[1 + \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}_c\khat\right)^2 \label{relax_cond_simp1}\\ 
+ \left(\hat{\omega}_c\khat\right)^2  + M(M+1)\left(1-h^2q^2\khat^2\right) , \label{relax_cond_simp1} \\
+   \left(\hat{\omega}_c\khat\right)^2 = & \frac{\beta}{h q} (\gamma-1)(1+2M)^2
+   \left(\hat{\omega}_c\khat\right) \label{relax_cond_simp2}\\
+   &- 2M(M+1) . \notag
+\end{align}\end{subequations}
+These equations hold for $\beta_c \ll 1$ and $\khat \sim 1/\hat{\omega}_c \gg1$,
+as explained in Appendix \ref{disp_neut_limit}.
 
-%Our starting point is Eq. \ref{relax_cond1}---\ref{relax_cond2} which
-%approximates the dispersion relation Eq. \ref{relax_disp} assuming marginal
-%stability and $\khat^2\gg 1$. We consider $M=0$ (the 
-%fundamental mode) or  $M=O(1)$ and low frequency $|\hat{\omega}|\ll
-%1$ such that $|\hat{\omega}\khat|$ is $O(1)$. We 
-%also assume $|q|$ and $\beta$ are $O(1)$ or smaller. 
-%
-%For a thin disk
-%($h \ll 1$) we have 
-\begin{subequations}\begin{align}
- \left(\hat{\omega}\khat\right)^4  = &  \left[1 +
-    \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}\khat\right)^2 \label{relax_cond_simp1}\\ 
- &+ M(M+1)\left(1-h^2q^2\khat^2\right)
-    , \notag \\
-   \left(\hat{\omega}\khat\right)^2 = & \frac{\beta}{h q} (\gamma-1)(1+2M)^2
-   \left(\hat{\omega}\khat\right) \label{relax_cond_simp2}\\
-   $- 2M(M+1) , \notag
-\end{align}\end{subequations}
-These equations hold for $\beta_c \ll 1$ and $\hat{k} \sim 1/|\
+Eq. \ref{relax_cond_simp1} gives an approximate result for the frequency
+\begin{align}
+ \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left[1  + O( M|h q \khat|^2)\right]  \label{marg_freq_approx}
+%  \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left(1 -
+%    \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
+\end{align}
+which holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
+These distinct conditions mean that our results are most robust for the fundamental vertical mode ($M = 0$) but also
+hold for higher order vertical modes when radial wavelengths are not too short.
+Note that Eq. \ref{marg_freq_approx} 
+reduces (\emph{nearly}?)\footnote{\emph The correspondence is only exact 
+if $L = M+1$.  Is this the case (in which case at least note in the appendix) or is there a discrepancy (in which case 
+we can't say it reduces)?}
+ to the dispersion relation for low-frequency inertial waves 
+%in vertically isothermal disks without vertical shear 
+(see Appendix\ref{stable_novshear} for details).
 
-Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} is a pair of
-simultaneous equations for $\hat{\omega}$ and $\beta$ \emph{assuming} marginal 
-stability. %  (It is important to note that marginal stability is an
-% assumption.)
-If $\beta\ll 1$ and the disk is sufficiently thin so that  
-$|h q \khat |^2\lesssim 1$,
-
-An approximate solution to Eq. \ref{relax_cond_simp1} for the frequency
-\begin{align}
-  \left(\hat{\omega}\khat\right)^2 \simeq (M+1)\left(1 -
-    \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
-\end{align}
-holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
-Note that for $q=0$, Eq. \ref{marg_freq_approx} reduces to 
-the dispersion relation for low-frequency inertial waves in vertically
-isothermal disks without vertical shear (see Appendix
-\ref{stable_novshear}). Inserting Eq. \ref{marg_freq_approx} into
+Inserting Eq. \ref{marg_freq_approx} into
 Eq. \ref{relax_cond_simp2}, we obtain 
 \begin{align}\label{bcrit_gen}
-  \beta &= \frac{h|q|}{(\gamma - 1)}
-  \frac{\sqrt{1+M}}{(1+2M)}\left[1 + \frac{M(2M-1)|h q
-      \khat|^2}{2(2M+1)^2}\right].
+  \beta_c &\simeq  \frac{h|q|}{(\gamma - 1)}
+  \frac{\sqrt{1+M}}{(1+2M)}
+%  \left[1 + \frac{M(2M-1)|h q   \khat|^2}{2(2M+1)^2}\right].
 %  &\lesssim \beta_\mathrm{crit}\notag
 \end{align}
-Since $|h q \khat|^2\lesssim 1$ by assumption, we have  
+where the above-mentioned approximations still hold, but the correction term ($O(M |h q \khat|^2)$) is now suppressed.
+Eq. \ref{bcrit_gen} says that higher $M$ modes have lower $\beta_c$, i.e.\ require faster cooling for VSI growth.
+This trend is confirmed numerically in \S\ref{therm_relax_eff} (and the trend holds even when the restrictive assumption
+on $|h q \hat{k}|$ is dropped).
+
+We are most interested in the longest cooling time which allows growth for any $M$.  
+This criterion is clearly set by the fundamental, $M = 0$, mode.  Our cooling criterion for 
+VSI growth is thus 
  \begin{align}\label{iso_vsi_cond}
   \beta\leq \beta_\mathrm{crit}  \equiv
-  \frac{h|q|}{\gamma-1}, 
+  \frac{h|q|}{\gamma-1} . 
 \end{align}
-where $\beta_\mathrm{crit}$ is the critical thermal timescale for
-the fundamental mode ($M=0$). 
-We comment the expression for $\beta_\mathrm{crit}$ may in fact be
-obtained without assuming it is much less than unity from the onset or
-$|h q\khat|\lesssim 1$. % This is because for $M=0$, 
-% Eq. \ref{relax_cond_simp1} --- \ref{relax_cond_simp2} simplify 
-We will  
-confirm $\beta_\mathrm{crit}$ numerically in \S\ref{bcrit_num_test}.
+The thin disk approximation, $h \ll 1$, justifies the $\beta_c \ll 1$ approximation.
+ We heuristically explain $\beta_\mathrm{crit}$ in
+\S\ref{vsi_require}, and numerically test its validity in \S\ref{bcrit_num_test}.
 
-Eq. \ref{iso_vsi_cond} suggests that as the thermal timescale $\beta$
-is increased from zero, higher order modes with $M>0$
-will be stabilized before the fundamental mode. We indeed observe this
-trend in numerical calculations (\S\ref{therm_relax_eff}), despite
-adopting somewhat restrictive assumptions for $M>0$ in order to arrive
-at the simple expression Eq. \ref{bcrit_gen}.
+%Eq. \ref{iso_vsi_cond} suggests that as the thermal timescale $\beta$
+%is increased from zero, higher order modes with $M>0$
+%will be stabilized before the fundamental mode. We indeed observe this
+%trend in numerical calculations (\S\ref{therm_relax_eff}), despite
+%adopting somewhat restrictive assumptions for $M>0$ in order to arrive
+%at the simple expression Eq. \ref{bcrit_gen}.
 
-%Rapid thermal relaxation is required to overcome a stabilizing
-%vertical entropy gradient \citep{goldreich67,urpin98,urpin03}. 
-%Eq. \ref{iso_vsi_cond} quantifies this requirement for vertically
-%global disturbances (which have small $M$).
-
-We can interpret  $\beta_\mathrm{crit}$ as the thermal timescale below
-which buoyancy becomes ineffective and perturbations are effectively
-isothermal so the VSI can operate.  We see that in a thin disk, a very short thermal timescale
-is required ($\beta_\mathrm{crit}\ll 1$) because
-$\beta_\mathrm{crit}\propto h$. Physically, this is because
-vertical shear is weak in a thin disk, so even a small entropy
-gradient can significantly reduce VSI growth rates.
-
-% Here we estimate the maximum thermal relaxation timescale
-% $\beta$ that allows the fundamental VSI mode for $\khat\gg1$.
-
-% At marginal stability, the eigenfrequency $\hat{\upsilon}=\hat{\omega}$
-% is real. The dispersion relation for the fundamental mode is then
-% \begin{align}\label{relax_disp_fund}
-%   0 = \sum_{l=1}^{6}c_l\hat{\omega}^l \quad \text{for $M=\hat{\nu}=0$}.  
-% \end{align} 
-% Taking the real and imaginary parts of Eq. \ref{relax_disp_fund} and considering 
-% $\khat\gg1$, we find
-% \begin{align}
-%   0 =& \left[1 + \gamma\beta^2\left(1-\gamma\right)\right] + 2h q
-%   \gamma\beta (\hat{\omega}\khat) -  (\hat{\omega}\khat)^2 \notag\\
-%   &+ \beta^2\gamma^2\hat{\omega}^2 (\hat{\omega}\khat)^2,\label{relax_cond1}\\
-%   0=& \beta(1-\gamma)\khat^2 + h q \khat^2 (\hat{\omega}\khat)
-%   - 2\beta (\hat{\omega}\khat)^2 - h q \gamma^2\beta^2 (\hat{\omega}\khat)^3 \notag\\
-%     &+ 2\beta\gamma(\hat{\omega}\khat)^4.  
-% \label{relax_cond2}
-% \end{align}
-% This is a pair of simultaneous equations for
-% $\hat{\omega}$ and $\beta$. Now, low-frequency axisymmetric waves in
-% the absence of vertical shear are stable and  satisfy
-% $\hat{\omega}\propto \khat^{-1}$ (Appendix \ref
-% {stable_novshear}, Eq. \ref{iso_disp_full}).  We are thus motivated to
-% seek solutions to Eq. \ref{relax_cond1}---\ref{relax_cond2} with  
-% % we wish to seek a criterion independent of k 
-% % \begin{align}
-% %   \khat \to \infty, \quad \hat{\omega}\to 0 ,\quad \hat{\omega} \khat\text{ finite}.
-% %\end{align}
-% \begin{align}
-%   \khat \gg1, \quad |\hat{\omega}|\ll 1 ,\quad
-%   |\hat{\omega}\khat|=O(1). 
-% \end{align}
-% Then
-% \begin{align}
-%   \hat{\omega}\khat = \frac{(\gamma-1)\beta}{h q}
-% \end{align}
-% from Eq. \ref{relax_cond2}, which implies, from Eq. \ref{relax_cond1}
-% that
-% \begin{align}\label{beta_crit0}
-%   \beta = \frac{1}{(\gamma-1)}\left[\frac{1}{\left(h
-%         q\right)^2} - \frac{\gamma}{(\gamma-1)}\right]^{-1/2} 
-% \end{align}
-% at marginal stability for perturbations with large $\khat$.
-
-% From our discussion in \S\ref{relax_pert}, which indicate introducing
-% a finite thermal relaxation is stabilizing, we expect
-% Eq. \ref{beta_crit0} is an upper limit. For a thin  
-% disk $h\ll 1$, we thus infer 
-% \begin{align}\label{iso_vsi_cond}
-%   \beta \lesssim \frac{|h q|}{\gamma-1} \equiv
-%   \beta_\mathrm{crit} 
-% \end{align}
-% is required for the fundamental VSI to operate. That is, when
-% $\beta<\beta_\mathrm{crit}$, perturbations are  effectively isothermal. 
-% We will confirm Eq. \ref{iso_vsi_cond} numerically in
-% \S\ref{bcrit_num_test}.
+%We can interpret  $\beta_\mathrm{crit}$ as the thermal timescale below
+%which buoyancy becomes ineffective and perturbations are effectively
+%isothermal so the VSI can operate.  We see that in a thin disk, a very short thermal timescale
+%is required ($\beta_\mathrm{crit}\ll 1$) because
+%$\beta_\mathrm{crit}\propto h$. Physically, this is because
+%vertical shear is weak in a thin disk, so even a small entropy
+%gradient can significantly reduce VSI growth rates.

### 2015-05-01 linear_adia.tex
merge with youdin's update

--- 2015-05-01 linear_adia.tex
@@
 we derive the following relation between $\beta_c$ and the wave frequency,
-\hat{\omega}_c$  at marginal stability to the VSI
+$\hat{\omega}_c$  at marginal stability to the VSI
 
@@
    \left(\hat{\omega}\khat\right)^2 = & \frac{\beta}{h q} (\gamma-1)(1+2M)^2
-   \left(\hat{\omega}\khat\right) \label{relax_cond_simp2}\\
-   $- 2M(M+1) , \notag
-\end{align}\end{subequations}
-These equations hold for $\beta_c \ll 1$ and $\hat{k} \sim 1/|\
+   \left(\hat{\omega}\khat\right)  
+    - 2M(M+1).\label{relax_cond_simp2} %\notag 
+\end{align}
+\end{subequations}
+These equations hold for $\beta_c \ll 1$ and $\hat{k} \gg1 $
 

### 2015-05-02 appendix.tex
merge with youdin's updates
additional changes to sec 5.3, added new appendix D2, minor addition to appendix C2.2

@@
 Here we detail the derivation of Eq. \ref{nearly_iso_explicit} used in the 
-analytical derivations of \S\ref{analytical}.  Starting with Eqs.\ \ref{lin_mass} --- \ref{lin_energy},
+analytical discussion of \S\ref{analytical}.  Starting with Eqs.\ \ref{lin_mass} --- \ref{lin_energy},
 we set $\Gamma = 1$ and $\zeta = 0$ for the vertically isothermal limit and the fully-radially-local 
@@
 in terms of dimensionless variables of Eq.\ \ref{nondim}. 
-
 Retaining terms to first order in the disk aspect ratio, $h \ll 1$, gives
@@
 observed in numerical simulations of \cite{stoll14}. 
+Note that $L=M+1$, where $M\geq 0$ is the mode number used for analytical
+discussion (\S\ref{analytical}) based on the reduced equation for
+$\delta v_z$, rather than for $W$ as considered here. 
+%Note that for
+%analytical discussions (\S\ref{analytical}) we work with an equation
+%for $\delta v_z$ so that $L=M+1$.
 
@@
 \end{align}
-We next (re)-apply the low frequency and thin approximations, $h, \hat{\omega}_c \ll 1$ and 
-further apply the rapid cooling limit $\beta_c \ll 1$.  We retain terms with the combinations 
-$\hat{\omega}_c \hat{k}$, $\beta_c/h$, and $h \hat{k}$ which could be order unity.  For inertial waves we know that
-$\hat{\omega}_c \sim 1/\hat{k}$ holds, while $\beta_c/h \sim O(1)$ will be justified \emph{a posteriori}
-and the magnitude of $h \hat{k}$ depends on the choice of wavenumber.  Applying this ordering we 
-arrive at Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}.
+We recall that in the low frequency and thin-disk approximations, $h,
+|\hat{\omega}_c| \ll 1$. We note that $|q|=O(1)$ and 
+$\gamma>1$ but is $O(1)$. We consider $M$ to be $O(1)$ or smaller, so 
+that $|\hat{\omega_c}\khat|=O(1)$, since for inertial waves
+$|\hat{\omega}\khat|^2\simeq 1 + M $. Finally, we assume 
+$\beta_c \ll 1$, to be justified \emph{a posteriori}. Applying these
+approximations give
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}.
+
+\subsection{Marginal stability for the fundamental mode}\label{bcrit_alt}
+If we apply the above approximations for the fundamental mode $M=0$ to
+Eq. \ref{relax_cond1}---\ref{relax_cond2}, but with the less
+restrictive assumption $\beta_c \lesssim O(1)$, we expect the balance  
+\begin{subequations}\begin{align}
+    &0  = \left[1 + \gamma\beta_c^2\left(1-\gamma\right)\right] + 2h q
+    \gamma \beta_c \left(\hat{\omega}_c\khat\right) - \left(\hat{\omega}_c\khat\right)^2,\\
+    &0 =  \beta_c(1-\gamma) + hq \left(\hat{\omega}_c\khat\right). 
+  \end{align}\end{subequations}
+Then
+\begin{align}
+  \beta_c = \frac{h|q|}{\gamma-1}\left[1 - \frac{\gamma h^2 
+      q^2}{\gamma-1}\right]^{-1/2},
+\end{align}
+which reduces to Eq. \ref{iso_vsi_cond} for thin-disks ($h\ll1$). We
+thus deduce that in fact $\beta_c\ll 1$, which motivates us to make this
+strong assumption from the onset.
 

### 2015-05-02 linear_adia.tex
very minor tweaks to 5 (done for now) and starting with 6.
update with youdin's update
additional changes to sec 5.3, added new appendix D2, minor addition to appendix C2.2

--- 2015-05-01 linear_adia.tex
@@
 Here we estimate the maximum thermal relaxation timescale 
-$\beta_c$ that allows growth the VSI.  In Appendix \ref{disp_neut_limit}
-we derive the following relation between $\beta_c$ and the wave frequency,
-$\hat{\omega}_c$  at marginal stability to the VSI:
+$\beta_c$ that allows growth of the VSI. In Appendix \ref{disp_neut_limit}
+we derive the following relation between $\beta_c$ and the wave
+frequency $\hat{\omega}_c$, \emph{assuming} marginal stability to the
+VSI:  
 \begin{subequations}\begin{align}
- \left(\hat{\omega}_c\khat\right)^4  = &  
-% \left[1 + \gamma\beta^2\left(1-\gamma\right)(1+2M)^2\right]\left(\hat{\omega}_c\khat\right)^2 \label{relax_cond_simp1}\\ 
- \left(\hat{\omega}_c\khat\right)^2  + M(M+1)\left(1-h^2q^2\khat^2\right) , \label{relax_cond_simp1} \\
-   \left(\hat{\omega}_c\khat\right)^2 = & \frac{\beta}{h q} (\gamma-1)(1+2M)^2
+    \left(\hat{\omega}_c\khat\right)^4  =& \left(\hat{\omega}_c\khat\right)^2 
+  + M(M+1)\left(1-h^2q^2\khat^2\right) ,\label{relax_cond_simp1}  \\
+  \left(\hat{\omega}_c\khat\right)^2 = & \frac{\beta_c}{h q} (\gamma-1)(1+2M)^2
   \left(\hat{\omega}_c\khat\right) \label{relax_cond_simp2}\\
@@
 \end{align}\end{subequations}
-These equations hold for $\beta_c \ll 1$ and $\khat \sim 1/\hat{\omega}_c \gg1$,
-as explained in Appendix \ref{disp_neut_limit}.
+These equations consider $\khat^2\gg 1$, and assumes 
+$|\hat{\omega}\khat| = O(1)$, $M\lesssim O(1)$ and $\beta_c\ll 1$.
 
-Eq. \ref{relax_cond_simp1} gives an approximate result for the frequency
+Eq. \ref{relax_cond_simp1} admits the following   
+approximate solution for the frequency if $M|hq\khat|^2/(2M+1)\lesssim
+1$, 
 \begin{align}
- \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left[1  + O( M|h q \khat|^2)\right]  \label{marg_freq_approx}
-%  \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left(1 -
-%    \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
+  % \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left[1  + O( M|h q \khat|^2)\right]  \label{marg_freq_approx}
+  \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left(1 -
+    \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
 \end{align}  
-which holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
-These distinct conditions mean that our results are most robust for the fundamental vertical mode ($M = 0$) but also
-hold for higher order vertical modes when radial wavelengths are not too short.
-Note that Eq. \ref{marg_freq_approx} 
-reduces (\emph{nearly}?)\footnote{\emph The correspondence is only exact 
-if $L = M+1$.  Is this the case (in which case at least note in the appendix) or is there a discrepancy (in which case 
-we can't say it reduces)?}
- to the dispersion relation for low-frequency inertial waves 
-%in vertically isothermal disks without vertical shear 
-(see Appendix\ref{stable_novshear} for details).
+% which holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
+% These distinct conditions mean that our results are most robust for the fundamental vertical mode ($M = 0$) but also
+% hold for higher order vertical modes when radial wavelengths are not
+% too short. 
+Note that Eq. \ref{marg_freq_approx} reduces to the dispersion
+relation for low-frequency inertial waves in the absence of vertical
+shear (see Appendix \ref{stable_novshear} for details).
 
-Inserting Eq. \ref{marg_freq_approx} into
-Eq. \ref{relax_cond_simp2}, we obtain 
+If we further assume $|hq\khat|^2\ll 1$, we obtain from
+Eq. \ref{relax_cond_simp2}---\ref{marg_freq_approx} that  
 \begin{align}\label{bcrit_gen} 
-  \beta_c &\simeq  \frac{h|q|}{(\gamma - 1)}
-  \frac{\sqrt{1+M}}{(1+2M)}
+  \beta_c &\simeq  \frac{h|q|}{(\gamma - 1)}\frac{\sqrt{1+M}}{(1+2M)}.   
+%  \frac{\sqrt{1+M}}{(1+2M)}
 %  \left[1 + \frac{M(2M-1)|h q   \khat|^2}{2(2M+1)^2}\right].
-%  &\lesssim \beta_\mathrm{crit}\notag
+%  &\lesssim \beta_\mathrm{crit}%\notag
+%>>>>>>> 8816838e4a06974fc61d74b84a2474adaeee3ebc
 \end{align}
-where the above-mentioned approximations still hold, but the correction term ($O(M |h q \khat|^2)$) is now suppressed.
-Eq. \ref{bcrit_gen} says that higher $M$ modes have lower $\beta_c$, i.e.\ require faster cooling for VSI growth.
-This trend is confirmed numerically in \S\ref{therm_relax_eff} (and the trend holds even when the restrictive assumption
-on $|h q \hat{k}|$ is dropped).
+% but the correction term ($O(M |h q \khat|^2)$) is now suppressed.    
+Eq. \ref{bcrit_gen} indicates that higher $M$ modes have lower $\beta_c$,
+i.e.\ require faster cooling for VSI growth\footnote{In fact,   
+  this conclusion does not require $|hq\khat|^2$ to be very 
+  small. The may be seen from the more complete expression for
+  $\beta_c$ obtained by substituting Eq. \ref{marg_freq_approx} into  
+  Eq. \ref{relax_cond_simp2} and retaining the term proportional to 
+  $|hq\khat|^2$.}.    
+This trend is confirmed numerically in \S\ref{therm_relax_eff}.
+
+%(and the trend holds even when the restrictive assumption
+%on $|h q \hat{k}|$ is dropped).
 
 We are most interested in the longest cooling time which allows growth for any $M$.  
-This criterion is clearly set by the fundamental, $M = 0$, mode.  Our cooling criterion for 
+This criterion is clearly set by the fundamental, $M = 0$, mode.  Our
+cooling criterion for  
 VSI growth is thus  
  \begin{align}\label{iso_vsi_cond}
-  \beta\leq \beta_\mathrm{crit}  \equiv
+  \beta_c\leq \beta_\mathrm{crit}  \equiv
   \frac{h|q|}{\gamma-1} . 
 \end{align}
-The thin disk approximation, $h \ll 1$, justifies the $\beta_c \ll 1$ approximation.
- We heuristically explain $\beta_\mathrm{crit}$ in
-\S\ref{vsi_require}, and numerically test its validity in \S\ref{bcrit_num_test}.
+The thin disk approximation, $h \ll 1$, indicates that $\beta_c \ll
+1$, as assumed to obtain Eq. \ref{marg_freq_approx}. We heuristically
+explain $\beta_\mathrm{crit}$ in \S\ref{vsi_require}, and numerically
+test its validity in \S\ref{bcrit_num_test}.
 
-%Eq. \ref{iso_vsi_cond} suggests that as the thermal timescale $\beta$
-%is increased from zero, higher order modes with $M>0$
-%will be stabilized before the fundamental mode. We indeed observe this
-%trend in numerical calculations (\S\ref{therm_relax_eff}), despite
-%adopting somewhat restrictive assumptions for $M>0$ in order to arrive
-%at the simple expression Eq. \ref{bcrit_gen}.
+In Appendix \ref{bcrit_alt}, we give an alternative derivation for 
+$\beta_\mathrm{crit}$ which does not assume it is much less
+than unity from the onset. This derivation also independent of $|h q
+\khat|$.
 
-%We can interpret  $\beta_\mathrm{crit}$ as the thermal timescale below
-%which buoyancy becomes ineffective and perturbations are effectively
-%isothermal so the VSI can operate.  We see that in a thin disk, a very short thermal timescale
-%is required ($\beta_\mathrm{crit}\ll 1$) because
-%$\beta_\mathrm{crit}\propto h$. Physically, this is because
-%vertical shear is weak in a thin disk, so even a small entropy
-%gradient can significantly reduce VSI growth rates.
+% For $M=0$, it is straight foward to solve Eq. \ref{relax_cond_simp1} 
+% --- \ref{relax_cond_simp2} exactly and obtain the maximum cooling time
+% for the fundamental mode, $\beta_\mathrm{crit}$
+% (Eq. \ref{iso_vsi_cond}), which shows that $\beta_\mathrm{crit}\ll 1$
+% because we consider thin disks.

### 2015-05-02 numerical.tex

@@
-\section{Numerical solutions}\label{numerical}
-In this section we numerically solve the linear stability problem
-corresponding to Eq. \ref{lin_mass}---\ref{lin_energy}. 
-We relax the  low-frequency, Keplerian and thin-disk 
-approximations used in the analytical discussion above and 
-use the full expressions of $\rho$, $\Omega$ and 
-$\kappa$ and $c_s$ given in \S\ref{eqm}. We also set $\hat{g}_c=1$ to   
-account for the background radial disk structure.
+\section{Fixed cooling time results}\label{numerical}
+%comment: "numerical solutions' was too vague, especially since the next section contains numerical solutions.
+This section investigates the linear growth of the VSI with fixed, finite cooling times, $\beta$. 
+We solve the perturbation equations \ref{lin_all} with $\zeta=1$, i.e. including the background radial disk structure.  
+The disk's equilbrium  structure given in \S\ref{eqm}.  This general analysis does not make the approximations 
+described in \S\ref{sec:simplified}.
 
@@
 components $\xi_{x,z} = \ii\delta v_{x,z}/\upsilon$. In some cases we
-impose a rigid boundary at which $\delta v_z=0$.
+impose a rigid boundary with $\delta v_z(\pm\zmax)=0$.
 
@@
 converts the linear system of differential equations to a set of 
-algebraic equations, for which we use matrix routines in the LAPACK
-package to solve.
+algebraic equations,  which we solve using LAPACK matrix routines.\footnote{provide a reference of some sort to LAPACK}
 
-Unless otherwise stated, our fiducial disk parameters are 
+Unless otherwise stated, we adopt fiducial disk parameters are 
 $(\gamma, \Gamma) = (1.4, 1.011)$, $(p,q, h)=(-1.5,-1,0.05)$, and 
 vertical domain size $\zmax=5H$. This setup is effectively vertically
-isothermal since $c_s^2(\zmax) \simeq 0.9c_s^2(0)$. In Fig. \ref{omega_z}
+isothermal since $c_s^2(\zmax) \simeq 0.9c_s^2(0)$.\footnote{While this model
+has a zero density surface at $z = \pm H_s \simeq \pm 16.4 H$ (see Eq.\ \ref{eqm_rot}), the surface is 
+outside the numerical domain.} In Fig. \ref{omega_z}
 we plot the corresponding equilibrium rotation profile, which shows
-that $\Omega$ decreases slightly away from the midplane.
+that $\Omega$ decreases slightly away from the midplane.   From Eq.\ \ref{iso_vsi_cond},
+we expect that $\beta < \beta_\mathrm{crit} \approx 0.13$ is needed for VSI growth in our fiducial model.
 
@@
 
-\subsection{VSI with rapid thermal relaxation}\label{vertiso_pertiso} 
+\subsection{Rapid thermal relaxation}\label{vertiso_pertiso} 
 We first calculate the VSI in a disk with rapid thermal relaxation by
-setting $\beta=10^{-3}$. (Smaller values yield similar results.) 
-This setup mimics strictly isothermal perturbations in a vertically
-isothermal disk, already considered previously \citepalias{nelson13,mcnally14,barker15}. 
-For completeness, and as a way of introducing terminology, 
-we review this case before considering the effect of 
-thermal relaxation. %A more detailed analysis is given by \cite{barker15}.
+setting $\beta=10^{-3}$. Since this cooling is much faster than the predicted
+$\beta_\mathrm{crit}$, we expect and that this case should give 
+similar results to the well studied case of isothermal perturbations
+ \citepalias[e.g.][]{nelson13,mcnally14,barker15}.  Indeed we found that smaller
+ $\beta$ values gave similar results. This case allows us to explore and test our
+finite cooling time model in a familiar context.
+
+%For completeness, and as a way of introducing terminology, 
+%we review this case before considering the effect of 
+%thermal relaxation.\footnote{\emph{not accurate, even $\beta = 0$ is instant thermal relaxation...}}
 
@@
 
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10_analytic.ps}
+  \caption{Growth rate, $\sigma$, vs.\ osciallation frequency, $\omega$, for
+  VSI modes with wavenumber $k_x = \khat/H=10/H$ in the fiducial disk model
+    with rapid cooling, $\beta=10^{-3}$. %  a disk with 
+    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q, h)=(-1.5,-1,0.1)$ and
+    % $\beta=10^{-3}$. 
+We investigate the effect of vertical box size, from $\zmax=3H$ (\emph{blue triangles}) to $7H$ (\emph{green x's})
+    with a free surface boundary ($\Delta P=0$).  For $\zmax=5H$  (black dots) we also compare free surface (\emph{black dots}) and rigid (\emph{red plusses}), finding little difference.
+      The fundamental mode is marked by `f'. Lines show the analytic predictions for
+      isothermal perturbations ($\beta = 0$, \emph{dashed}) and for $\beta = 10^{-3}$ (\emph{dotted}).
+      See text for discussion.
+   % are computed from the analytic dispersion relation 
+    %with thermal relaxation Eq. \ref{relax_disp} (dotted), and that for
+    %isothermal perturbations, Eq. \ref{simple_growth} (dashed). 
+    \label{compare_modes_iso_kx10} 
+  }
+\end{figure}
+
 \subsubsection{Case study: $\khat=10$}
-In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of unstable 
-modes with $\khat=10$. Black dots correspond to our fiducial setup.  
-We also plot results obtained with different 
-vertical boundary conditions: rigid boundary and $\zmax=5H$ (red
-crosses),  free surface and $\zmax = 3H$ (blue triangles), free surface
-and $\zmax=7H$ (green crosses).
+In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues -- growth rates and
+wave frequencies -- of unstable 
+modes for $\khat=10$.  We compare our numerical results to the analytic dispersion relations for 
+$\beta = 0$ and $10^{-3}$, from Eqs.\ \ref{simple_growth} and \ref{relax_disp}, respectively. Each symbol 
+corresponds to a different order mode, i.e.\ a different number of vertical node crossings.  Lower order modes
+have lower (in magnitude) frequencies, with the lowest order `fundamental' modes further to the left in Fig. \ref{compare_modes_iso_kx10}.
+
+For our fiducial case (black dots), the lower order and lower frequency modes closely follow the analytic prediction with growth rates increasing with $|\omega|$, the magnitude of the frequency.
+However, for $|\omega| \gtrsim 4 h \OmK$ the growth rate declines with $|\omega|$.
+This break from the analytic prediction is not due the choice of boundary condition, as we demonstrate by considering rigid boundaries (red plusses).
+
+Rather, the decline in growth rate for high order models is due to the vertical box size, as seen by comparing the $\zmax/H = 3, 5$ and 7 cases in Fig.
+\ref{compare_modes_iso_kx10}.
+Larger boxes give better agreement with the analytic theory, which are infinite in vertical extent.
+Larger boxes include regions of larger vertical shear (see Eqs.
+\ref{vertical_shear_thin} and \ref{vertical_shear_ex}), which higher order modes can tap to give larger growth rates.
+The need for vertically extended domains to capture the largest VSI growth rates is problematic 
+For the $\zmax = 7 H$ case, we begin to see another branch of modes, the so-called ``surface modes" described below.
+
+%comment: the following reads like a figure caption, not a description of results.
+%Black dots correspond to our fiducial setup.  
+%We also plot results obtained with different 
+%vertical boundary conditions: rigid boundary and $\zmax=5H$ (red
+%crosses),  free surface and $\zmax = 3H$ (blue triangles), free surface
+%and $\zmax=7H$ (green crosses).
 
@@
 unless $\zmax$ is increased.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10_analytic.ps}
-  \caption{Unstable modes with $\khat=10$ in the fiducial disk model
-    with $\beta=10^{-3}$. %  a disk with 
-    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q, h)=(-1.5,-1,0.1)$ and
-    % $\beta=10^{-3}$. 
-    The fiducial boundary condition is a free surface ($\Delta P=0$)
-    with vertical domain size $\zmax=5H$  (black dots).  Other
-    cases correspond to different boundary conditions applied at
-    different heights. The fundamental mode is marked by `f'. Lines
-    are computed from the analytic dispersion relation 
-    with thermal relaxation Eq. \ref{relax_disp} (dotted), and that for
-    isothermal perturbations, Eq. \ref{simple_growth} (dashed). 
-    \label{compare_modes_iso_kx10} 
-  }
-\end{figure}
 
@@
     pressure perturbation $\delta P=\rho W$ scaled by its maximum value.
-    Arrows correspond to the vector field field $\sqrt{\rho}(\delta
-    v_x,\delta v_z)$. Note that the horizontal axis has been stretched 
+    Arrows show the flow, scaled as $\sqrt{\rho}(\delta
+    v_x,\delta v_z)$. The horizontal axis has been stretched 
     for clarity.  

### 2015-05-03 linear_adia.tex
very slight correction to the sentence before eq 47
finally resolved disagreement over sec 5.3, eq 48

@@
 Eq. \ref{relax_cond_simp1} admits the following   
-approximate solution for the frequency if $M|hq\khat|^2/(2M+1)\lesssim
-1$, 
+approximate solution for the frequency 
+if $\mathrm{min}(1,M)\times|hq\khat|^2\lesssim 1$, 
+%if $4M(M+1)|hq\khat|^2/(2M+1)^2 < 1$,  
 \begin{align}
@@
 \end{align}  
+(Equality holds for $M=0$.)  
 % which holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
@@
 \end{align}
-% but the correction term ($O(M |h q \khat|^2)$) is now suppressed.    
 Eq. \ref{bcrit_gen} indicates that higher $M$ modes have lower $\beta_c$,
-i.e.\ require faster cooling for VSI growth\footnote{In fact,   
-  this conclusion does not require $|hq\khat|^2$ to be very 
-  small. The may be seen from the more complete expression for
-  $\beta_c$ obtained by substituting Eq. \ref{marg_freq_approx} into  
-  Eq. \ref{relax_cond_simp2} and retaining the term proportional to 
-  $|hq\khat|^2$.}.    
+i.e.\ require faster cooling for VSI growth\footnote{
+In fact, this conclusion does not require $\theta\equiv|hq\khat|^2$ to be extremely 
+small. By retaining terms proportional to
+  $\theta$, we find $\beta_c\to \beta_c \times 
+  \left[1 + M(2M-1)\theta/2(2M+1)^2 \right]$. Since $\theta\lesssim 1$ by assumption, 
+ $\mathrm{max}(\beta_c)$ still occurs at $M=0$.}.     
 This trend is confirmed numerically in \S\ref{therm_relax_eff}.
@@
 The thin disk approximation, $h \ll 1$, indicates that $\beta_c \ll
-1$, as assumed to obtain Eq. \ref{marg_freq_approx}. We heuristically
-explain $\beta_\mathrm{crit}$ in \S\ref{vsi_require}, and numerically
+1$, as assumed to obtain Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. 
+We heuristically explain $\beta_\mathrm{crit}$ in \S\ref{vsi_require}, and numerically
 test its validity in \S\ref{bcrit_num_test}.

### 2015-05-03 appendix.tex
further refinements to appendix D2, D3. plus minor notation update

@@
 caveat, Fig. \ref{gcorr_compare} shows that provided we consider 
-$\beta\ll1$, then setting $\zeta=0$ does not affect the VSI
+$\beta\lesssim O(1)$, then setting $\zeta=0$ does not affect the VSI
 significantly.
@@
 \subsection{Finding marginal stability}\label{disp_neut_limit}
-To investigate marginal stability we set $\hat{\sigma} = 0$ so the frequency, $\hat{\upsilon}=\hat{\omega}_c$, is real.  
-We take the short radial wavelength limit, $\khat^2\gg 1$, of the coefficients in Eq.\ \ref{coeffs}.  The real and 
-imaginary parts of Eq. \ref{relax_disp} then give relations for $\hat{\omega}_c$ and $\beta_c$, the cooling time for 
-marginal stability, as 
-\begin{align}\
+To investigate marginal stability we set $\hat{\sigma} = 0$ so the
+frequency, $\hat{\upsilon}=\hat{\omega}_c$, is real; and $\beta =
+\beta_c$,  the cooling time for marginal stability. 
+We take the short radial wavelength limit, $\khat^2\gg 1$, of the
+coefficients in Eq.\ \ref{coeffs}. We consider $M \lesssim O(1)$ and
+assume $\beta_c \ll \khat$.  The real and 
+imaginary parts of Eq. \ref{relax_disp} then give relations for $\hat{\omega}_c$ and $\beta_c$ as 
+\begin{align}
   0 =&M(M+1)(1-h^2 q^2 \khat^2) + 4\beta_c h q M(M+1)\left(\hat{\omega}_c\khat\right) 
@@
 |\hat{\omega}_c| \ll 1$. We note that $|q|=O(1)$ and 
-$\gamma>1$ but is $O(1)$. We consider $M$ to be $O(1)$ or smaller, so 
-that $|\hat{\omega_c}\khat|=O(1)$, since for inertial waves
-$|\hat{\omega}\khat|^2\simeq 1 + M $. Finally, we assume 
-$\beta_c \ll 1$, to be justified \emph{a posteriori}. Applying these
-approximations give
+$\gamma>1$ but is $O(1)$. We assume $|\hat{\omega_c}\khat|=O(1)$,
+since for inertial waves $|\hat{\omega}\khat|\simeq \sqrt{1 + M}$. Finally, we further assume 
+$\beta_c \ll 1$, to be justified \emph{a posteriori}, to give 
 Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}.
 
-\subsection{Marginal stability for the fundamental mode}\label{bcrit_alt}
-If we apply the above approximations for the fundamental mode $M=0$ to
-Eq. \ref{relax_cond1}---\ref{relax_cond2}, but with the less
-restrictive assumption $\beta_c \lesssim O(1)$, we expect the balance  
+\subsection{Marginal stability for the fundamental
+  mode}\label{bcrit_alt}
+For the fundamental mode ($M=0$) we can deduce $\beta_c\ll 1$, which
+motivates us to make this strong assumption from the onset. In this
+case, Eq. \ref{relax_cond1} --- \ref{relax_cond2} 
+imply the following balance for large $\khat$ and small
+$\hat{\omega_c}$ while $\hat{\omega}_c\khat$ remain fixed:  
 \begin{subequations}\begin{align}
@@
 \begin{align}
-  \beta_c = \frac{h|q|}{\gamma-1}\left[1 - \frac{\gamma h^2 
-      q^2}{\gamma-1}\right]^{-1/2},
+  \beta_c = \frac{h|q|}{\gamma-1}\left(1 - \frac{\gamma h^2 
+      q^2}{\gamma-1}\right)^{-1/2},
 \end{align}
-which reduces to Eq. \ref{iso_vsi_cond} for thin-disks ($h\ll1$). We
-thus deduce that in fact $\beta_c\ll 1$, which motivates us to make this
-strong assumption from the onset.
+which reduces to Eq. \ref{iso_vsi_cond} at $O(h)$.
 

### 2015-05-03 appendix.tex
further refinements to appendix D2, D3. plus minor notation update

@@
 imply the following balance for large $\khat$ and small
-$\hat{\omega_c}$ while $\hat{\omega}_c\khat$ remain fixed:  
+$\hat{\omega}_c$ while $\hat{\omega}_c\khat$ remain fixed:  
 \begin{subequations}\begin{align}

### 2015-05-04 appendix.tex
hopefully final resolution to sec 5.3, major revision, please double check (esp. the new appendix D3)

@@
 
-%growth rate of high frequency branch is much smaller
-
-% \begin{align}\label{sig2_iso}
-%   \hat{\upsilon}^2 = L\left(\frac{1+\ii q h
-%       \hat{k}}{1+\hat{k}^2}\right).
-% \end{align}
-% The eigenfrequency $\hat{\upsilon}$ is therefore complex for
-% $L\geq1$. The dispersion relation Eq. \ref{sig2_iso} can also be
-% derived by setting $\beta=0$ in the more general dispersion relation 
-% Eq. \ref{relax_disp} with coefficients given in Appendix
-% \ref{relax_coeff}. We are interested in 
-% unstable modes for which $\hat{\upsilon}>0$ and is given by
-% \begin{align}
-%   \hat{\upsilon} =\sqrt{
-%     \frac{L}{2\left(1+\hat{k}^2\right)}\left(\sqrt{1+q^2h^2\hat{k}^2} - 
-%       1\right)}. 
-% \end{align}
-
-% We impose the kinetic energy remain bounded at infinity. Even though
-% we have adopted the thin-disk approximation, an infinite
-% disk is appropriate in the sense that both the true density profile 
-% and its thin-disk approximation decay rapidly away from the midplane.
-
-% However, the thin-disk representation of vertical shear diverges with 
-% $\hat{z}$, unlike the true vertical shear profile which eventually
-% decays. Thus, taking $\zhat_{1,2}\to \pm\infty$ will permit 
-% unphysically large growth rates (as demonstrated below). Nevertheless,
-% the infinite disk is useful to consider since it permits an analytical
-% discussion.
-
-% \subsubsection{Stability in the absence of vertical
-%   shear}\label{iso_stable}
-% \
-% Eq. \ref{iso_ode3} is
-% \begin{align}\label{hermite_ode}
-%   \left[w(\hat{z})W^\prime \right]^\prime + nW
-%   w(\hat{z}) =0, 
-% \end{align}
-% where $w(\zhat)=\exp{(-\zhat^2/2)}$ and we have defined a new eigenvalue 
-% \begin{align}
-%   n \equiv \hat{\upsilon}^2(1+\hat{k}^2). 
-% \end{align} 
-% Eq. \ref{hermite_ode} is Hermite's differential equation. If we impose
-% that the kinetic energy density remain bounded at infinity, then the
-% solutions are Hermite polynomials  
-% \begin{align}
-%   W \propto \He_n(\hat{z}),
-% \end{align}
-% and $n$ is a non-negative integer. This translates to a real
-% eigenfrequency $\sigma$ such that
-% \begin{align}
-%   \left|\hat{\upsilon}\right| = \sqrt{n}
-%   \left(1+\hat{k}^2\right)^{-1/2}. \label{iso_stable_disp}
-% \end{align}
-% Since we have assumed $|\sigma^2|\ll \kappa^2\sim \OmK^2$, our
-% analysis is only valid for large wavenumbers such that $\hat{k}^2\gg   
-% n-1$. Physically, this corresponds to radial length-scales much
-% smaller than the local disk scale height. Note that
-% Eq. \ref{iso_stable_disp} is implied by Eq. 54 in
-% \cite{lubow93} after setting $\gamma=1$ and making the low-frequency,
-% nearly-Keplerian approximation in that equation.
-
-% We remark that for fixed $n$ and $\khat\gg 1$ the dispersion relation 
-% Eq. \ref{iso_stable_disp} imply $\hat{\omega}\propto 1/\khat$. This inverse relation 
-% has also been qualitatively observed in numerical simulations
-% \citep{stoll14}.
-
-% \subsubsection{Polynomial solutions with vertical
-%   shear}\label{iso_poly}
-
-% For fixed $q$ and $h$, the growth rate vanishes for both
-% $\hat{k}^2\to0$ and $\hat{k}^2\to\infty$. 
-% The maximum growth rate
-% occurs at the optimum wavenumber $\hat{k}_\mathrm{opt}$
-% \begin{align}
-%   |\hat{k}_\mathrm{opt}| = \sqrt{\frac{2+|h q|}{|h q|}},
-% \end{align}
-% and
-% \begin{align}
-%   \mathrm{max}\left(\hat{\upsilon}\right) =\frac{\sqrt{L}|h
-%     q|}{2\sqrt{1+|h q|}}. \label{iso_max_growth}
-% \end{align}
-% Since $L\geq1$, there is no limit to the maximum growth rate. This is
-% an artifact of the thin-disk approximation because the magnitude of vertical 
-% shear $|d\Omega^2/dz|\propto |z|$ increases indefinitely with 
-% height. However, Eq. \ref{iso_max_growth} compares well with
-% numerical solutions for small $L$ (see \S\ref{vertiso_pertiso}).
-
 \section{Analytic dispersion relation
@@
 Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}.
+
+\subsection{Maximum critical cooling time}\label{max_cool}
+ % {\bf [at the moment for reference
+ %  only, not to be included in final version]}\\
+Consider the simplified dispersion relations for marginal stability,
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. Writing 
+$X= \hat{\omega}\khat$, $\theta = (hq\khat)^2\leq1$ and  treating $M$ as a
+continuous variable, we find from Eq. \ref{relax_cond_simp1} that 
+%\begin{align}
+%  &x^4  = x^2 + M(M+1)(1-\theta), \label{simp1}\\
+%  &x^2  = \frac{\gamma-1}{h q}\beta_c (1+2M)^2x - 2M(M+1)\label{simp2},
+%\end{align}
+%where and
+\begin{align}
+  2 X \frac{d X}{dM} = \frac{X^2(1-\theta)(2M+1)}{X^2 +
+  2M(M+1)(1-\theta)}, 
+\end{align}
+and from Eq. \ref{relax_cond_simp2} that
+\begin{align}
+  2X\frac{dX}{dM} = \left[X^2 +
+  2M(M+1)\right]\left(\frac{d\ln{\beta_c}}{dM} + \frac{4}{2M+1} +
+  \frac{d\ln{X}}{dM}\right) - 2(2M+1). 
+\end{align}
+We eliminate $dX/dM$, making use of Eq. \ref{relax_cond_simp1} in the process, to obtain 
+\begin{align}
+  (2M+1)\frac{d\ln{\beta_c}}{dM} = \frac{\mathcal{N}}{\mathcal{D}}, 
+\end{align}
+where the denominator $\mathcal{D}$ is
+\begin{align}
+\mathcal{D} = 2 \left[X^2 +
+  2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]>0, 
+\end{align}
+and the numerator $\mathcal{N}$ is 
+\begin{align}
+  \mathcal{N} = -(1-\theta)\left[ 2M(M+1)(2M+1)^2 + 12X^2M(M+1)\right]
+  - (3+\theta)X^2 < 0.
+\end{align}
+Hence,
+\begin{align}
+  \frac{d\beta_c}{dM}<0, 
+\end{align}
+for all $M$ provided that $\theta\leq 1$. This implies 
+$\mathrm{max}(\beta_c)$ occurs at $M=0$.
 

### 2015-05-04 appendix.tex
hopefully final resolution to sec 5.3, major revision, please double check (esp. the new appendix D3)

@@
 
-\subsection{Stability conditions in the low-frequency limit}
-In the low-frequency limit we assume $|\hat{\upsilon}|\ll 1$ and
-approximate Eq. \ref{iso_ode0} as 
+\subsection{Maximum growth rate  in the low-frequency limit}
+Here we consider the low-frequency limit $|\hat{\upsilon}|\ll 1$ and 
+show that the growth rate is limited by the maximum vertical
+shear rate in the domain. 
+%In the low-frequency limit we assume $|\hat{\upsilon}|\ll 1$ and
+We approximate Eq. \ref{iso_ode0} as 
 \begin{align}
@@
 \end{align}
-\subsubsection{Necessity of vertical shear for
-  instability}\label{integral_relation} 
-We can establish a necessary condition for instability by multiplying
-Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
+%\subsubsection{Necessity of vertical shear for
+%  instability}\label{integral_relation} 
+%We can establish a necessary condition for instability 
+We multiply Eq. \ref{iso_ode1} by $\rho W^*$ and integrate vertically from
 $\zhat=\zhat_1$ to $z=\zhat_2$. We neglect boundary 
@@
 It follows that for instability ($\imag\hat{\upsilon}>0$), it is necessary to
-have $q\neq0$ or more generally $d\Omega^2/dz\neq 0$, i.e. there must
-be vertical shear. A similar conclusion can be reached in the
-equivalent linear problem in global cylindrical disk geometry.
+have $q\neq0$ or more generally $d\Omega/dz\neq 0$, i.e. there must
+be vertical shear. 
+%A similar conclusion can be reached in the
+%equivalent linear problem in global cylindrical disk geometry.  
+%\subsubsection{Maximum growth rate}\label{max_growth1}
 
-\subsubsection{Maximum growth rate}\label{max_growth1}
-Here we show that the growth rate is limited by the maximum vertical
-shear rate in the domain. % Writing $\hat{\upsilon} = \hat{\omega} +
-% \ii\hat{\upsilon}$ with real $\hat{\omega}$ and $\hat{\upsilon}$, 
 The real and imaginary parts of 
@@
 \begin{align}
-  \sigma < \mathrm{max} \left|r\frac{d\Omega}{dz}\right|, 
+  |\sigma| < \mathrm{max} \left|r\frac{d\Omega}{dz}\right|, 
 \end{align}
-as expected on physical grounds.
+as expected on physical grounds. %A similar conclusion can be reached in the
+%equivalent linear problem in global cylindrical disk geometry.
 
@@
 results.
-
-% For vertical domains smaller 
-% than this height, the growth rate is limited by $|f|$ at the vertical
-% boundaries considered. For $|h\hat{z}|\ll1$ we have $f\simeq
-% \hat{z}$, in which case we can expect growth rates to increase at most linearly
-% with respect to the vertical domain size.   %no faster than linear
 
@@
 \begin{align}
-  (2M+1)\frac{d\ln{\beta_c}}{dM} = \frac{\mathcal{N}}{\mathcal{D}}, 
+  (2M+1)\frac{d\ln{\beta_c}}{dM} %= \frac{\mathcal{N}}{\mathcal{D}}, 
+  = -\frac{(1-\theta)M(M+1)\left[ 2(2M+1)^2 + 12X^2\right]
+  + (3+\theta)X^2}{2 \left[X^2 +
+  2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]}.  
 \end{align}
-where the denominator $\mathcal{D}$ is
-\begin{align}
-\mathcal{D} = 2 \left[X^2 +
-  2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]>0, 
-\end{align}
-and the numerator $\mathcal{N}$ is 
-\begin{align}
-  \mathcal{N} = -(1-\theta)\left[ 2M(M+1)(2M+1)^2 + 12X^2M(M+1)\right]
-  - (3+\theta)X^2 < 0.
-\end{align}
+%where the denominator $\mathcal{D}$ is
+%\begin{align}
+%\mathcal{D} = 2 \left[X^2 +
+ % 2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]>0, 
+%\end{align}
+%and the numerator $\mathcal{N}$ is 
+%\begin{align}
+ % \mathcal{N} = -(1-\theta)\left[ 2M(M+1)(2M+1)^2 + 12X^2M(M+1)\right]
+ % - (3+\theta)X^2 < 0.
+%\end{align}
 Hence,
@@
 \end{align}
-for all $M$ provided that $\theta\leq 1$. This implies 
+for all $M\geq 0$ if $\theta\leq 1$. This implies 
 $\mathrm{max}(\beta_c)$ occurs at $M=0$.

### 2015-05-04 application.tex
edits in progress to 6 plus some suggestions to 7, not done yet...

@@
 \section{Application to protoplanetary disks}\label{application} 
+\emph{I suggest the following edits
+\begin{itemize}
+\item Instead of starting over again with the full energy equation focus on giving a phyiscal values for $\Lambda$ and $t_c$ already introduced.
+\item Don't describe what we don't do, especially at the outset.  Say what's done, and then if essential list caveats.
+\end{itemize}
+
 As an application of our linear models, we estimate where in a
@@
 \subsection{Radiative diffusion}
+\emph{the following sentence is too long and has a dangling "this".}
+
 For perturbations with lengthscales $l\gg
@@
 
+\emph{I would have this be an equation for something, like $\Gamma$ or $t_c$.}
 \begin{align}\label{diff_cool_proper}
@@
 implicitly.
+
+\emph{To get to the point more directly and intuitively, I would start below with "We include only the radial derivatives in Eq.\ blah, neglecting vertical derivatives.
+This approximation is valid for the vertically elongated modes we consider."  I think we've already discussed vertical elongation enough that this can be brief.
+Then maybe say that keeping the vertical derivates would be bad, if you really want to.
+"With this simplification (give revised equation)}
 
@@
 For perturbations with small lengthscales, $l\ll l_\mathrm{rad}$, 
-the disk is optically thin. In this case, we assume 
+the disk is optically thin.
+\emph{the disk isn't (necessarily) opt. thin, the pert. is}
+ In this case, we assume 
 Newtonian cooling can be applied, which is the thermal

### 2015-05-04 numerical.tex
edits in progress to 6 plus some suggestions to 7, not done yet...

@@
 We first calculate the VSI in a disk with rapid thermal relaxation by
-setting $\beta=10^{-3}$. Since this cooling is much faster than the predicted
-$\beta_\mathrm{crit}$, we expect and that this case should give 
-similar results to the well studied case of isothermal perturbations
- \citepalias[e.g.][]{nelson13,mcnally14,barker15}.  Indeed we found that smaller
- $\beta$ values gave similar results. This case allows us to explore and test our
+setting $\beta=10^{-3}$.  Since $\beta \ll \beta_\mathrm{crit}$, this case
+gives similar results to the well studied case of isothermal perturbations
+ \citepalias[e.g.][]{nelson13,mcnally14,barker15}.\footnote{To be sure, we confirmed that smaller
+ $\beta$ values gave similar results.} This case allows us to explore and test our
 finite cooling time model in a familiar context.
@@
 corresponds to a different order mode, i.e.\ a different number of vertical node crossings.  Lower order modes
-have lower (in magnitude) frequencies, with the lowest order `fundamental' modes further to the left in Fig. \ref{compare_modes_iso_kx10}.
+have lower (in magnitude) frequencies for a given radial wavenumber, as expected for inertial-gravity waves (see \S\ref{stable_novshear}).
 
@@
 \ref{compare_modes_iso_kx10}.
-Larger boxes give better agreement with the analytic theory, which are infinite in vertical extent.
+Larger boxes give better agreement with the analytic theory, which has infinite vertical extent.
 Larger boxes include regions of larger vertical shear (see Eqs.
-\ref{vertical_shear_thin} and \ref{vertical_shear_ex}), which higher order modes can tap to give larger growth rates.
-The need for vertically extended domains to capture the largest VSI growth rates is problematic 
-For the $\zmax = 7 H$ case, we begin to see another branch of modes, the so-called ``surface modes" described below.
+\ref{vshear_thin} and \ref{vertical_shear_ex}), which higher order modes can tap to give larger growth rates.
+For the $\zmax = 7 H$ case, we begin to see another branch of modes with the highest growth rates at $\omega \sim -6 h\OmK$.
+This branch contains the  ``surface modes" described below.
+
+The need for vertically extended domains to capture the largest VSI growth rates is problematic, especially for hydrodynamic simulations.
+This  complication is mitigated by at least two factors.
+First, we show below that with slower cooling, the growth of higher order modes is preferentially damped, consistent with the analytic analysis in \S\ref{iso_vsi_beta_crit}.
+Second, the fastest growing modes may not dominate transport when they operate in very low density  surface layers, i.e.
+at many $H$.
+Quantifying which modes will contribute most to non-linear transport is an important issue, but beyond the scope of this work.
 
@@
 
-The modes shown in Fig. \ref{compare_modes_iso_kx10} are `body modes'
-and are inertial waves destabilized by vertical shear
-\citepalias{barker15}. (The exception being the fastest growing
-mode for $\zmax=7H$ which is a surface mode --- see the discussion
-below.)  The `fundamental corrugation mode', marked by `f',
-corresponds to the eigenvalue with the smallest $|\upsilon|$. The
-fundamental mode is robust  
-against vertical boundary conditions as it appears at the same
-position in Fig. \ref{compare_modes_iso_kx10} for all cases. Higher  
-order modes, corresponding to increasing $|\omega|$, increasingly
-deviates from the analytic line due to the imposed boundaries, 
-unless $\zmax$ is increased.
+%The modes shown in Fig. \ref{compare_modes_iso_kx10} are `body modes'
+%and are inertial waves destabilized by vertical shear
+%\citepalias{barker15}. (The exception being the fastest growing
+%mode for $\zmax=7H$ which is a surface mode --- see the discussion
+%below.)  The `fundamental corrugation mode', marked by `f',
+%corresponds to the eigenvalue with the smallest $|\upsilon|$. The
+%fundamental mode is robust  
+%against vertical boundary conditions as it appears at the same
+%position in Fig. \ref{compare_modes_iso_kx10} for all cases. Higher  
+%order modes, corresponding to increasing $|\omega|$, increasingly
+%deviates from the analytic line due to the imposed boundaries, 
+%unless $\zmax$ is increased.
 
-For later comparison, we plot in Fig. \ref{lowfreq_eigenfunc} the 
-eigenfunctions $W$ and $\delta   v_z$ for the fundamental mode. We also show a 
-meridional visualization in Fig. \ref{lowfreq_eigenfunc_2d}. The $x$ axis is 
-stretched for clarity; radial velocities are in fact typically much 
-smaller than vertical velocities. Most of the meridional kinetic energy is
-contained within $\sim 2H$ of the midplane due to the density 
+Fig. \ref{lowfreq_eigenfunc} shows 
+eigenfunctions $W = \delta P/\rho$ and $\delta   v_z$ for the lowest order fundamental mode.  
+Fig. \ref{lowfreq_eigenfunc_2d} maps the pressure perturbation and meridional flow,
+scale by $\sqrt{\rho}$ to reflect the contribution to kinetic energy. Due to the stretched $x$ axis, 
+radial velocities are in fact typically much 
+smaller than vertical velocities, as expected for a vertically elongated, incompressible mode. 
+Most of the kinetic energy is contained within $\sim 2H$ of the midplane due to the density 
 stratification.
@@
 Fig. \ref{compare_modes_cool_kx10} show unstable modes with $\khat=10$ 
-and $\beta\in[0.01,0.1]$. In addition to our fiducial setup with 
+and $\beta$ from $0.01$ to  $0.1$. In addition to our fiducial setup with 
 $\zmax=5H$, we also plot results for $\zmax=3H$, $\zmax=7H$ and 
@@
 (middle panel). The fundamental mode ($M=0$) is not the most unstable for   
-$\beta=0.01$ as there area number of modes with $M>0$ that have larger
+$\beta=0.01$ as higher order modes have larger
 growth rates. As noted above, these higher order modes are affected by
@@
 example as $\beta$ is increased from $0.01$ to $0.03$ (for $\zmax\geq
-5H$). If we discard modes associated with boundary conditions, then
-the fundamental mode would again become the dominant mode as $\beta$ is
-increased.
+5H$). 
+%\emph{not a clear and possible not valid argument.  All modes are associated with
+%a BC and the disappearance of a mode for smaller zmax could be the artificial effect.}
+%If we discard modes associated with boundary conditions, then
+%the fundamental mode would again become the dominant mode as $\beta$ is
+%increased.
 
@@
 %a disk surface at finite height. 
+\emph{fix this rambling discussion..}
 \citetalias{barker15} have argued that vertically isothermal disks are not
@@
 H_s$). The disk temperature decreases rapidly with height such that 
-$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$.
+$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$.    \emph{note that this would imply midplane
+temperatures in excess of $10^6$ K so it's completely un-astrophyscial.}
 

### 2015-05-04 linear_adia.tex

@@
 \end{align}\end{subequations}
-These equations consider $\khat^2\gg 1$, and assumes 
-$|\hat{\omega}\khat| = O(1)$, $M\lesssim O(1)$ and $\beta_c\ll 1$.
+These equations consider $\khat^2\gg 1$, $M\lesssim O(1)$; and assume 
+$|\hat{\omega}\khat| = O(1)$ and $\beta_c\ll 1$.   Note that
+Eq. \ref{relax_cond_simp1} reduces to the
+dispersion relation for low-frequency inertial waves in the absence of
+vertical shear (see Appendix \ref{stable_novshear} for details).
 
-Eq. \ref{relax_cond_simp1} admits the following   
-approximate solution for the frequency 
-if $\mathrm{min}(1,M)\times|hq\khat|^2\lesssim 1$, 
-%if $4M(M+1)|hq\khat|^2/(2M+1)^2 < 1$,  
+In Appendix \ref{max_cool}, we find that %if $|hq\khat|^2\leq 1$ then 
 \begin{align}
-  % \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left[1  + O( M|h q \khat|^2)\right]  \label{marg_freq_approx}
-  \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left(1 -
-    \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
-\end{align}  
-(Equality holds for $M=0$.)  
-% which holds for either (a) $M= 0$ and arbitrary $|h q \hat{k}|$ or (b) $|h q \hat{k}| \ll 1$ and arbirtary $M \sim O(1)$. 
-% These distinct conditions mean that our results are most robust for the fundamental vertical mode ($M = 0$) but also
-% hold for higher order vertical modes when radial wavelengths are not
-% too short. 
-Note that Eq. \ref{marg_freq_approx} reduces to the dispersion
-relation for low-frequency inertial waves in the absence of vertical
-shear (see Appendix \ref{stable_novshear} for details).
-
-If we further assume $|hq\khat|^2\ll 1$, we obtain from
-Eq. \ref{relax_cond_simp2}---\ref{marg_freq_approx} that  
-\begin{align}\label{bcrit_gen} 
-  \beta_c &\simeq  \frac{h|q|}{(\gamma - 1)}\frac{\sqrt{1+M}}{(1+2M)}.   
-%  \frac{\sqrt{1+M}}{(1+2M)}
-%  \left[1 + \frac{M(2M-1)|h q   \khat|^2}{2(2M+1)^2}\right].
-%  &\lesssim \beta_\mathrm{crit}%\notag
-%>>>>>>> 8816838e4a06974fc61d74b84a2474adaeee3ebc
+  |hq\khat|^2 \leq 1 \Rightarrow \frac{d\beta_c}{dM} < 0 \quad \forall
+  M\geq 0. \label{bcrit_gen}
 \end{align}
 Eq. \ref{bcrit_gen} indicates that higher $M$ modes have lower $\beta_c$,
-i.e.\ require faster cooling for VSI growth\footnote{
-In fact, this conclusion does not require $\theta\equiv|hq\khat|^2$ to be extremely 
-small. By retaining terms proportional to
-  $\theta$, we find $\beta_c\to \beta_c \times 
-  \left[1 + M(2M-1)\theta/2(2M+1)^2 \right]$. Since $\theta\lesssim 1$ by assumption, 
- $\mathrm{max}(\beta_c)$ still occurs at $M=0$.}.     
-This trend is confirmed numerically in \S\ref{therm_relax_eff}.
+i.e.\ require faster cooling for VSI growth. This trend is confirmed
+numerically in \S\ref{therm_relax_eff}.
 
+% Eq. \ref{relax_cond_simp1} admits the following   
+% approximate solution for the frequency 
+% if $\mathrm{min}(1,M)\times|hq\khat|^2\lesssim 1$, 
+% \begin{align}
+%   % \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left[1  + O( M|h q \khat|^2)\right]  \label{marg_freq_approx}
+%   \left(\hat{\omega}_c\khat\right)^2 \simeq (M+1)\left(1 -
+%     \frac{M|h q \khat|^2}{2M+1}\right).  \label{marg_freq_approx}
+% \end{align}  
+% (Equality holds for $M=0$.)  
+% Note that Eq. \ref{marg_freq_approx} reduces to the dispersion
+% relation for low-frequency inertial waves in the absence of vertical
+% shear (see Appendix \ref{stable_novshear} for details). 
+% If we further assume $|hq\khat|^2\ll 1$, we obtain from
+% Eq. \ref{relax_cond_simp2}---\ref{marg_freq_approx} that  
+% \begin{align}\label{bcrit_gen} 
+%   \beta_c &\simeq  \frac{h|q|}{(\gamma - 1)}\frac{\sqrt{1+M}}{(1+2M)}.   
+% %  \frac{\sqrt{1+M}}{(1+2M)}
+% %  \left[1 + \frac{M(2M-1)|h q   \khat|^2}{2(2M+1)^2}\right].
+% %  &\lesssim \beta_\mathrm{crit}%\notag
+% %>>>>>>> 8816838e4a06974fc61d74b84a2474adaeee3ebc
+% \end{align}
+% Eq. \ref{bcrit_gen} indicates that higher $M$ modes have lower $\beta_c$,
+% i.e.\ require faster cooling for VSI growth. 
+%\footnote{
+%In fact, this conclusion does not require $\theta\equiv|hq\khat|^2$ to be extremely 
+%small. By retaining terms proportional to
+% $\theta$, we find $\beta_c\to \beta_c \times 
+%  \left[1 + M(2M-1)\theta/2(2M+1)^2 \right]$. Since $\theta\lesssim 1$ by assumption, 
+% $\mathrm{max}(\beta_c)$ still occurs at $M=0$.}.     
 %(and the trend holds even when the restrictive assumption
@@
 We are most interested in the longest cooling time which allows growth for any $M$.  
-This criterion is clearly set by the fundamental, $M = 0$, mode.  Our
-cooling criterion for  
-VSI growth is thus  
+This criterion is clearly set by the fundamental mode, $M = 0$. It is straight
+forward to solve Eq. \ref{relax_cond_simp1} --- \ref{relax_cond_simp2}
+for $\beta_c(M=0)  \equiv \beta_\mathrm{crit}$ in this special case. We thus find the cooling criterion for  
+VSI growth is  
  \begin{align}\label{iso_vsi_cond}
-  \beta_c\leq \beta_\mathrm{crit}  \equiv
+   \beta <   \beta_\mathrm{crit}  \equiv
    \frac{h|q|}{\gamma-1} . 
  \end{align}
-The thin disk approximation, $h \ll 1$, indicates that $\beta_c \ll
+ The thin disk approximation, $h \ll 1$, indicates that $\beta_\mathrm{crit} \ll
  1$, as assumed to obtain Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. 
@@
 
-In Appendix \ref{bcrit_alt}, we give an alternative derivation for 
-$\beta_\mathrm{crit}$ which does not assume it is much less
-than unity from the onset. This derivation also independent of $|h q
-\khat|$.
+We comment that the expression for $\beta_\mathrm{crit}$ may in fact 
+be obtained without assuming it is much less than unity from the onset or 
+restrictions on $| h q \khat|$.  This is clarified in Appendix
+\ref{bcrit_alt}.
 

### 2015-05-05 application.tex
Comments on Section 7 for suggested edits.  Hopefully they are coherent, but ask if not.

@@
 \section{Application to protoplanetary disks}\label{application} 
-\emph{I suggest the following edits
+\emph{For this section. There are 3 main ideas, and I think they should be the subsections, more comments related to this are in the text itself:
+\begin{itemize}
+\item Developing the cooling model in opt thin and this regimes, including MMSN value
+\item Comparison to expected $\beta_{crit}$
+\item Linear growth with MMSN cooling law
+\end{itemize}}
+
+\emph{Suggested edits for this intro
 \begin{itemize}
@@
 \item Don't describe what we don't do, especially at the outset.  Say what's done, and then if essential list caveats.
-\end{itemize}
+\end{itemize}}
 
@@
 \subsection{Radiative diffusion}
+\emph{I think 7.1-7.4 should be combined to a single section on the "Cooling Model".
+As that clearly is a single idea and osme of the subsections are too trivial.
+Conversion to subsubsections as needed.
+However the application to beta-crit could be split off as 7.2.
+Otherwise that material which has a crucial figure discussion is a bit buried.}
+
 \emph{the following sentence is too long and has a dangling "this".}
@@
 \subsection{Newtonian cooling}\label{newton_cool}
+\emph{There are no rules against subsections this short, but I think it's bad style.}
+
 For perturbations with small lengthscales, $l\ll l_\mathrm{rad}$, 
-the disk is optically thin.
-\emph{the disk isn't (necessarily) opt. thin, the pert. is}
+the disk is optically thin.\\
+\emph{Above, the disk isn't (necessarily) opt. thin, the pert. is} \\
  In this case, we assume 
@@
 $\hat{\Sigma}=\hat{\kappa}_d=\hat{T}=1$ as the Minimum Mass Solar
-Nebulae (MMSN) and $\beta=\beta_\mathrm{MMSN}$, 
+Nebulae (MMSN) and $\beta=\beta_\mathrm{MMSN}$, \\
+\emph{I would keep $\kappa$ scaling in this eqn, since it's used in plots.  Then just say  that $\kappa_d = 1$ is fiducial case. }
 \begin{align}\label{beta_mmsn_simp}
@@
 \end{align}
+\emph{I would explain here that when $k < c\rho \kappa^m r^n$ (with correct values for $c, m, n$) cooling is optically thin, as this is useful for explaining figure.}
 % where
@@
 % \end{align}
-It is instructive to compare the above thermal timescale 
+
+\emph{Sub-section break  ( 7.2) for the "$\beta_{crit}$ in the MMSN" as this is break from the derivation (and evaluation) of the cooling time (which is fairly general) to the application to the VSI.}
+
+It is instructive (\emph{advice: don't say that "it is instructive", just be instructive}) to compare the above thermal timescale 
 with the criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
@@
 
+\emph{This paragraph currently merely copies the figure caption, which is a waste and not illuminating to read.  Better to give some description of what the figure is showing. (see below for ideas)}\\
 In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn}  with $\mu
 =2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for opacity scales
-$\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$, at $z=0$ and $z=3H$. We
-also plot Eq. \ref{bcrit_mmsn} in each panel.
+$\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$, at $z=0$ and $z=3H$. \\
+\emph{Here you can describe the general features e.g. diffusive and thin cooling regions, which will help later discussions.}\\
+We also plot Eq.
+\ref{bcrit_mmsn} in each panel.
+\emph{This sentence explains nothing and is likely to be of little help to the reader.
+Instead wait to describe the science, e.g.
+after describing the general shape of the $\beta$ curves,  you can start the discussion below with "Fig () also compares the MMSN cooling times to $\beta_{crit}$.}
 
@@
 curves overlap), since in this limit $\beta$ does not depend on
-height, as assumed in our analysis. We thus infer that the VSI can
+height, as assumed in our analysis. \\
+\emph{above sentence too long and confusing.  "useful" is a vague term.  
+"where the curves overlap" is also confusing, forcing the reader to look at the figure in the middle of a complex sentence.
+Better to have already explained above how the $\beta$ curves show that opt.
+thin cooling is height indep.
+(even though should be obvious from equations).
+Then the comparison to $\beta_{crit}$ can be made as something like "Since beta-crit was derived for a height-independent cooling time,  the direct comparison our MMSN cooling model is valid only for modes with optically thin cooling.
+"}
+We thus infer that the VSI can
 operate at tens of AU in the MMSN.
@@
 solutions of the linear problem to determine whether or not the VSI
-operates. We do this in the next section.
+operates. We do this in the next section.    \\
+\emph{a better organized discussion might end on this point.  Might not be easy to rearrange to do this, but it would ideal to have the seamless flow to what you do next.}
 
@@
 Fig. \ref{mmsn_bcrit_bcool}, shifts the curves inwards but also raises
-them. In fact, lowering the opacity by an order of magnitude renders  
+them.\\
+\emph{this graphical discussion is not very inuitive (the main problem) and even arbitrary (lowering vs.
+shifting to the left).
+Better to say what's happening physically, i.e.
+`With higher opacity, optically thin cooling faster and diffusive cooling is slower, explaining how the curves shift.} \\
+In fact, lowering the opacity by an order of magnitude renders  
 $\beta > \beta_\mathrm{crit}$ even away from the midplane. This is 
@@
 
+\emph{the following statement is not generally true, so it should not be stated without caveat}
+
 We remark that it is possible to satisfy $\beta < \beta_\mathrm{crit}$ 
@@
 %also small vertical scales if they are surface modes
+
+\emph{The following viscosity discussion might be better placed elsewhere.
+Since the issue also comes up for numerical solutions (not just the comparison to beta-crit here), maybe a short concluding subsection to summarize all the arguments on the "vialibity of short wavelength VSI".}
 
@@
 height dependent thermal relaxation timescale given by
-Eq. \ref{beta_mmsn_simp}.  Fig. \ref{beta_compare} show examples of
+Eq. \ref{beta_mmsn_simp}. 
+ Fig. \ref{beta_compare} show examples of
 the thermal timescales at $r=5$AU and $r=50$AU for a
@@
 corresponding critical thermal timescales evaluated for the MMSN
-(Eq. \ref{bcrit_mmsn}).  
+(Eq. \ref{bcrit_mmsn}).  \\
+\emph{For this figure to belong here (and not in the previous subsection, with the other comparisons of $\beta$ and $\beta_{crit}$) the discussion should be be more focused on the linear analysis.
+Again it's better to start by discussing what the figure tells us, than repeating the figure caption.
+E.g.\ you could use Fig 21 to explain when a height varying $\beta$ is required.
+I would also explain (especially if it stays in this subsection) what the growth time is for the r = 5 AU case.
+Looks to be $\sim 1000$ orbits, which is remakrable given that $\beta < \beta_{\rm crit}$ outside $z \gtrsim  1.6 H$.?
+That kind of detail could be illuminating, as it's nominally surprising given previous evidence that behavior near 5 -- 7 $H$ can affect growth.}
+
  \begin{figure}
@@
 surface modes for large wavenumbers $\khat\gtrsim 50$, which are 
-boundary-dependent.  
+boundary-dependent.  \\
+\emph{above sentence too long and complex, could e.g.\ split off that surface modes are neglected and why.}
 %However, surface modes are dependent on

### 2015-05-05 application.tex
merge with youdin's updates
re-writing sec 7 in progress

@@
 \section{Application to protoplanetary disks}\label{application} 
-\emph{For this section. There are 3 main ideas, and I think they should be the subsections, more comments related to this are in the text itself:
-\begin{itemize}
-\item Developing the cooling model in opt thin and this regimes, including MMSN value
-\item Comparison to expected $\beta_{crit}$
-\item Linear growth with MMSN cooling law
-\end{itemize}}
+% \emph{I suggest the following edits}
+% \emph{For this section. There are 3 main ideas, and I think they should be the subsections, more comments related to this are in the text itself}:
+% \begin{itemize}
+% \item Developing the cooling model in opt thin and this regimes, including MMSN value
+% \item Comparison to expected $\beta_{crit}$
+% \item Linear growth with MMSN cooling law
+% \end{itemize}
 
-\emph{Suggested edits for this intro
-\begin{itemize}
-\item Instead of starting over again with the full energy equation focus on giving a phyiscal values for $\Lambda$ and $t_c$ already introduced.
-\item Don't describe what we don't do, especially at the outset.  Say what's done, and then if essential list caveats.
-\end{itemize}}
+% \emph{Suggested edits for this intro}
+% \begin{itemize}
+% \item Instead of starting over again with the full energy equation focus on giving a phyiscal values for $\Lambda$ and $t_c$ already introduced.
+% \item Don't describe what we don't do, especially at the outset.  Say what's done, and then if essential list caveats.
+% \end{itemize}
 
@@
 PPD do the thermodynamic conditions allow the VSI to
-operate. For a realistic disk one might consider an energy equation
-with radiative cooling, 
-\begin{align}\label{real_energy}
-  % \rho T \frac{DS}{Dt} = - \left(\nabla\cdot\bm{F} - Q_+\right), 
-  \frac{\p P}{\p t} + \bm{v}\cdot\nabla P +\gamma P \nabla\cdot\bm{v} = - \frac{P}{\rho T
-    C_v}\left(\nabla\cdot\bm{F} - Q_+\right), 
-\end{align}
-where $C_v$ is the heat capacity at constant
-volume, 
-$\bm{F}$ is the radiative flux and $Q_+$ is a heating
-term to permit an equilibrium solution.
+operate. We first model the cooling time $t_c$ based on realistic
+disk properties, which will yield a non-uniform dimensionless cooling
+time $\beta(r,z;\khat)$. We then compare this to the critical cooling time
+$\beta_\mathrm{crit}$ derived above. Finally, we explicitly compute 
+linear growth rates numerically in order to characterize 
+growth times expected in PPDs.
 
-One should now model in detail each of the terms on the RHS of
-Eq. \ref{real_energy}, linearize them, then solve 
-the resulting linear eigenvalue problem. This is beyond 
-the scope of the present work. Here, we take the simpler approach of
-representing the linearized RHS of Eq. \ref{real_energy} by a thermal
-relaxation time $t_c$. To do so, we first consider separately
-perturbations with radial lengthscales $l\gg l_\mathrm{rad}$ and  
-$l\ll l_\mathrm{rad}$, where      
-\begin{align}\label{lrad}
-  l_\mathrm{rad} \equiv \frac{1}{\kappa_d\rho} 
-\end{align} 
-is the photon mean-free-path and $\kappa_d$ is the (dust) opacity.
+%linearized RHS of Eq. \ref{real_energy} by a thermal
+%relaxation time $t_c$.
+% For a realistic disk one might consider an energy equation
+% with radiative cooling, 
+% \begin{align}\label{real_energy}
+%   % \rho T \frac{DS}{Dt} = - \left(\nabla\cdot\bm{F} - Q_+\right), 
+%   \frac{\p P}{\p t} + \bm{v}\cdot\nabla P +\gamma P \nabla\cdot\bm{v} = - \frac{P}{\rho T
+%     C_v}\left(\nabla\cdot\bm{F} - Q_+\right), 
+% \end{align}
+% where $C_v$ is the heat capacity at constant
+% volume, 
+% $\bm{F}$ is the radiative flux and $Q_+$ is a heating
+% term to permit an equilibrium solution. 
+% One should now model in detail each of the terms on the RHS of
+% Eq. \ref{real_energy}, linearize them, then solve 
+% the resulting linear eigenvalue problem. This is beyond 
+% the scope of the present work.
 
-\subsection{Radiative diffusion}
-\emph{I think 7.1-7.4 should be combined to a single section on the "Cooling Model".
+\subsection{Cooling model}
+
+% \emph{I think 7.1-7.4 should be combined to a single section on the "Cooling Model".
 As that clearly is a single idea and osme of the subsections are too trivial.
@@
 
-\emph{the following sentence is too long and has a dangling "this".}
+\subsubsection{Radiative diffusion and Newtonian cooling}
 
-For perturbations with lengthscales $l\gg
-l_\mathrm{rad}$, we assume temperature fluctuations are smoothed out
-by radiative diffusion and model this in linear theory by writing the
-linearized RHS of Eq. \ref{real_energy} as
+We first consider separately perturbations with radial lengthscales
+$l\gg l_\mathrm{rad}$ and $l\ll l_\mathrm{rad}$, where       
+\begin{align}\label{lrad}
+  l_\mathrm{rad} \equiv \frac{1}{\kappa_d\rho} 
+\end{align} is the photon mean-free-path and $\kappa_d$ is the (dust)
+opacity.
 
-\emph{I would have this be an equation for something, like $\Gamma$ or $t_c$.}
+For perturbations with lengthscales $l\gg l_\mathrm{rad}$, we assume
+temperature fluctuations are smoothed out by radiative diffusion and
+model this by writing the linearized cooling function as
+
 \begin{align}\label{diff_cool_proper}
-  \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
+  \delta \Lambda = \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
     T\right),  
@@
 where $k_\mathrm{rad}$ is the conduction coefficient to be defined
-implicitly.
+implicitly, and $C_v$ is the heat capacity as constant volume.
 
-\emph{To get to the point more directly and intuitively, I would start below with "We include only the radial derivatives in Eq.\ blah, neglecting vertical derivatives.
-This approximation is valid for the vertically elongated modes we consider."  I think we've already discussed vertical elongation enough that this can be brief.
-Then maybe say that keeping the vertical derivates would be bad, if you really want to.
-"With this simplification (give revised equation)}
+% However, Eq. \ref{diff_cool_proper} fundamentally changes the linear
+% problem by introducing higher order derivatives. We defer a proper
+% analysis of this problem to a follow-up study.
 
-However, Eq. \ref{diff_cool_proper} fundamentally changes the linear
-problem by introducing higher order derivatives. We defer a proper
-analysis of this problem to a follow-up study. Here, we are interested
-in order-of-magnitude estimates for perturbations with radial
-lengthscales much smaller than vertical lengthscales. We thus proceed 
-by assuming vertical derivatives can be neglected in 
-Eq. \ref{diff_cool_proper} and write  
+Since the VSI is characterized by vertically-elongated,
+radially-narrow disturbances, we simplify Eq. \ref{diff_cool_proper} 
+by retaining only radial derivatives of the perturbations, so that 
 \begin{align}\label{diff_cool_approx}
-  \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
-    T\right) &\to -\frac{k_\mathrm{rad}}{\rho
-    C_v}k_x^2P\left(\frac{\delta T}{T}\right)\notag\\
-  &\equiv -\eta k_x^2 \left(\delta P - \frac{P}{\rho}\delta\rho\right), 
+  % \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
+  %  T\right) &
+  \delta\Lambda \to %-\frac{k_\mathrm{rad}}{\rho
+  % C_v}k_x^2P\left(\frac{\delta T}{T}\right)%\notag\\
+  % \equiv 
+  %-\eta k_x^2 \left(\delta P - \frac{P}{\rho}\delta\rho\right) =
+  -\eta k_x^2 P \frac{\delta T}{T}, %=
+  %-\frac{P}{t_\mathrm{diff}}\frac{\delta T}{T},
+  %\frac{P}{t_\mathrm{diff}}\left(\frac{delta T}{T}\right), 
 \end{align}
@@
 \end{align}
-where $\sigma_s$ is the Stefan-Boltzmann constant. 
-From Eq. \ref{diff_cool_approx} we identify the scale-dependent thermal relaxation
+where $\sigma_s$ is the Stefan-Boltzmann constant. From our
+definition of thermal relaxation (Eq. \ref{thermal_relax}), 
+we thus identify the scale-dependent thermal relaxation
 time for diffusion as 
@@
 
-\subsection{Newtonian cooling}\label{newton_cool}
-\emph{There are no rules against subsections this short, but I think it's bad style.}
-
-For perturbations with small lengthscales, $l\ll l_\mathrm{rad}$, 
-the disk is optically thin.\\
-\emph{Above, the disk isn't (necessarily) opt. thin, the pert. is} \\
- In this case, we assume 
-Newtonian cooling can be applied, which is the thermal
+%\subsubsection{Newtonian cooling}\label{newton_cool}
+%\emph{There are no rules against subsections this short, but I think it's bad style.}
+Perturbations with small lengthscales, $l\ll
+l_\mathrm{rad}$, are in the optically thin regime. In this case, we assume 
+Newtonian cooling can be applied, which is exactly the thermal 
 relaxation model as adopted in our linear models. That is, temperature
@@
 
-\subsection{A model for combined thermal relaxation}\label{toy_relax}
+\subsubsection{A model for combined thermal relaxation}\label{toy_relax}
 %To unify $t_\mathrm{cool}$ and $t_\mathrm{diff}$, 
@@
 
-\subsection{Evaluation for a fiducial disk model}
-We evaluate Eq. \ref{real_beta} for the PPD model
-described in \cite{chiang10} and summarized in Appendix \ref{mmsn}
-(where an opacity model is also chosen). Defining $r_\mathrm{AU}\equiv 
-r/\mathrm{AU}$, we find Eq. \ref{real_beta} becomes
-\begin{align}\label{beta_mmsn}
-  \beta(r, z; \khat) =
-  &\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)r_\mathrm{AU}^{-57/14}\notag\\ 
-&\times\left[\frac{8.3\times10^{-9}}{\hat{\kappa}_d^2\hat{T}^4\hat{\Sigma}^2}r_\mathrm{AU}^{33/7}+\frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right].         
+%\subsubsection{Evaluation for a fiducial disk model}
+
+We evaluate Eq. \ref{real_beta} for the Minimum Mass Solar Nebula
+(MMSN) disk model described in \cite{chiang10} and summarized in
+Appendix \ref{mmsn} (where an opacity model is also chosen). Defining
+$r_\mathrm{AU}\equiv r/\mathrm{AU}$, we find Eq. \ref{real_beta}
+becomes 
+% \begin{align}\label{beta_mmsn}
+%   \beta(r, z; \khat) =
+%   &\frac{4.4\times10^5}{\mu(\gamma-1)}\left(\frac{\hat{\kappa}_d\hat{\Sigma}^2}{\hat{T}}\right)r_\mathrm{AU}^{-57/14}\notag\\ 
+%   &\times\left[\frac{8.3\times10^{-9}}{\hat{\kappa}_d^2\hat{T}^4\hat{\Sigma}^2}r_\mathrm{AU}^{33/7}+\frac{\hat{\rho}^2(z)}{2\pi\khat^2}\right].         
+% \end{align} 
+% We shall consider the case $\gamma=1.4$, $\mu=2.33$ with scalings
+% $\hat{\Sigma}=\hat{T}=1$, and refer to this as the Minimum Mass Solar 
+% Nebulae (MMSN) where $\beta=\beta_\mathrm{MMSN}$, \\
+% \emph{I would keep $\kappa$ scaling in this eqn, since it's used in plots.  Then just say  that $\kappa_d = 1$ is fiducial case. }
+\begin{align}\label{beta_mmsn_simp}
+  \beta(r,z;\khat)\notag = &3.9\times10^{-3}\hat{\kappa}_d
+  r_\mathrm{AU}^{9/14}
+  \notag\\ &\times \left[\hat{\kappa}_d^{-2} +
+    1.9\times10^7\hat{\rho}^2(z)r_\mathrm{AU}^{-33/7}\khat^{-2}\right]. 
 \end{align}
-We refer to the case with $\gamma=1.4$, $\mu=2.33$ and the scalings
-$\hat{\Sigma}=\hat{\kappa}_d=\hat{T}=1$ as the Minimum Mass Solar
-Nebulae (MMSN) and $\beta=\beta_\mathrm{MMSN}$, \\
-\emph{I would keep $\kappa$ scaling in this eqn, since it's used in plots.  Then just say  that $\kappa_d = 1$ is fiducial case. }
-\begin{align}\label{beta_mmsn_simp}
-  \beta_\mathrm{MMSN} = 3.9\times10^{-3} r_\mathrm{AU}^{9/14}\left( 1 +
-    1.9\times10^7\hat{\rho}^2r_\mathrm{AU}^{-33/7}\khat^{-2}\right). 
+The opacity scale for our fiducial case is $\hat{\kappa}_d=1$.
+
+Perturbations are in the optically thin regime,
+$l\sim 1/k_x \lesssim\l_\mathrm{rad}$, when 
+\begin{align}
+  \khat\gtrsim 2.5\times10^3\hat{\kappa}_d\hat{\rho}(z)r_\mathrm{AU}^{-33/14}.  
 \end{align}
-\emph{I would explain here that when $k < c\rho \kappa^m r^n$ (with correct values for $c, m, n$) cooling is optically thin, as this is useful for explaining figure.}
+In the inner disk, e.g. $r\sim 1$AU, only
+extremely small-scale perturabtions are in the optically-thin
+regime. However, in the outer disk, e.g. $r\sim 10$AU, the transition
+occurs at more moderate wavenumbers $\khat\sim 10$.
+%, though the
+%wavelength is still small compared to $H$.
+
+%\emph{I would explain here that when $k < c\rho \kappa^m r^n$ (with correct values for $c, m, n$) cooling is optically thi%n, as this is useful for explaining figure.}
 % where
@@
 
-\emph{Sub-section break  ( 7.2) for the "$\beta_{crit}$ in the MMSN" as this is break from the derivation (and evaluation) of the cooling time (which is fairly general) to the application to the VSI.}
-
-It is instructive (\emph{advice: don't say that "it is instructive", just be instructive}) to compare the above thermal timescale 
-with the criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
+\subsection{Comparing $\beta$ to $\beta_\mathrm{crit}$ in the MMSN} 
+Here we compare the above thermal timescale, Eq. \ref{beta_mmsn_simp}, to  
+the VSI criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
 \begin{align}\label{bcrit_mmsn}
@@
 
-\emph{This paragraph currently merely copies the figure caption, which is a waste and not illuminating to read.  Better to give some description of what the figure is showing. (see below for ideas)}\\
-In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn}  with $\mu
-=2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for opacity scales
-$\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$, at $z=0$ and $z=3H$. \\
-\emph{Here you can describe the general features e.g. diffusive and thin cooling regions, which will help later discussions.}\\
-We also plot Eq.
+% \emph{This paragraph currently merely copies the figure caption, which is a waste and not illuminating to read.
+Better to give some description of what the figure is showing.
+(see below for ideas)}\\
+% In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn}  with $\mu
+% =2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for opacity scales
+% $\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$, at $z=0$ and $z=3H$. \\
+% \emph{Here you can describe the general features e.g. diffusive and thin cooling regions, which will help later discussions.}\\
+% We also plot Eq.
 \ref{bcrit_mmsn} in each panel.

### 2015-05-05 atlanta/poster/makepdf.sh
atlanta poster files

@@ -1,0 +1,6 @@
+#!/bin/bash
+latex poster.tex
+dvips -o poster.ps poster.dvi
+###ps2pdf -sPAPERSIZE=a0 poster.ps
+ps2pdf poster.ps
+echo done

### 2015-05-05 appendix.tex
re-writing sec 7 in progress

@@
 
-\subsection{Marginal stability for the fundamental
-  mode}\label{bcrit_alt}
-For the fundamental mode ($M=0$) we can deduce $\beta_c\ll 1$, which
-motivates us to make this strong assumption from the onset. In this
-case, Eq. \ref{relax_cond1} --- \ref{relax_cond2} 
-imply the following balance for large $\khat$ and small
-$\hat{\omega}_c$ while $\hat{\omega}_c\khat$ remain fixed:  
-\begin{subequations}\begin{align}
-    &0  = \left[1 + \gamma\beta_c^2\left(1-\gamma\right)\right] + 2h q
-    \gamma \beta_c \left(\hat{\omega}_c\khat\right) - \left(\hat{\omega}_c\khat\right)^2,\\
-    &0 =  \beta_c(1-\gamma) + hq \left(\hat{\omega}_c\khat\right). 
-  \end{align}\end{subequations}
-Then
-\begin{align}
-  \beta_c = \frac{h|q|}{\gamma-1}\left(1 - \frac{\gamma h^2 
-      q^2}{\gamma-1}\right)^{-1/2},
-\end{align}
-which reduces to Eq. \ref{iso_vsi_cond} at $O(h)$.
+% \subsection{Marginal stability for the fundamental
+%   mode}\label{bcrit_alt}
+% For the fundamental mode ($M=0$) we can deduce $\beta_c\ll 1$, which
+% motivates us to make this strong assumption from the onset. In this
+% case, Eq. \ref{relax_cond1} --- \ref{relax_cond2} 
+% imply the following balance for large $\khat$ and small
+% $\hat{\omega}_c$ while $\hat{\omega}_c\khat$ remain fixed:  
+% \begin{subequations}\begin{align}
+%     &0  = \left[1 + \gamma\beta_c^2\left(1-\gamma\right)\right] + 2h q
+%     \gamma \beta_c \left(\hat{\omega}_c\khat\right) - \left(\hat{\omega}_c\khat\right)^2,\\
+%     &0 =  \beta_c(1-\gamma) + hq \left(\hat{\omega}_c\khat\right). 
+%   \end{align}\end{subequations}
+% Then
+% \begin{align}
+%   \beta_c = \frac{h|q|}{\gamma-1}\left(1 - \frac{\gamma h^2 
+%       q^2}{\gamma-1}\right)^{-1/2},
+% \end{align}
+% which reduces to Eq. \ref{iso_vsi_cond} at $O(h)$.
 
@@
 $\hat{T}$ are dimensionless coefficients used to scale the model
-relative to the MMSN. Their nominal values are unity. By assuming a
+relative to the MMSN. By assuming a
 vertically isothermal disk, we deduce the disk aspect-ratio 
@@
 temperature profile above, and $\hat{\kappa}_d$ has similar meaning as
-$\hat{\Sigma}$ and $\hat{T}$.
+$\hat{\Sigma}$ and $\hat{T}$. In \S\ref{application} we fix
+$\hat{\Sigma} = \hat{T}=1$, $\mu=2.33$ and use $\gamma=1.4$.

### 2015-05-05 application.tex
re-writing sec 7 in progress

@@
 \section{Application to protoplanetary disks}\label{application}
-% \emph{I suggest the following edits}
-% \emph{For this section. There are 3 main ideas, and I think they should be the subsections, more comments related to this are in the text itself}:
-% \begin{itemize}
-% \item Developing the cooling model in opt thin and this regimes, including MMSN value
-% \item Comparison to expected $\beta_{crit}$
-% \item Linear growth with MMSN cooling law
-% \end{itemize}
-
-% \emph{Suggested edits for this intro}
-% \begin{itemize}
-% \item Instead of starting over again with the full energy equation focus on giving a phyiscal values for $\Lambda$ and $t_c$ already introduced.
-% \item Don't describe what we don't do, especially at the outset.  Say what's done, and then if essential list caveats.
-% \end{itemize}
 
@@
 \subsection{Cooling model}
-
-% \emph{I think 7.1-7.4 should be combined to a single section on the "Cooling Model".
-As that clearly is a single idea and osme of the subsections are too trivial.
-Conversion to subsubsections as needed.
-However the application to beta-crit could be split off as 7.2.
-Otherwise that material which has a crucial figure discussion is a bit buried.}
 
@@
 
-\subsubsection{A model for combined thermal relaxation}\label{toy_relax}
-%To unify $t_\mathrm{cool}$ and $t_\mathrm{diff}$, 
+\subsubsection{Evaluation for a fiducial disk model}\label{toy_relax}
 We define the effective thermal relaxation timescale in linear theory as
@@
 \end{align}
-Note that $\beta$ depends the perturbation lengthscale $\khat$,
+Note that $\beta$ depends the perturbation lengthscale, 
 distance from the midplane, as well as the distance from the star.
@@
 \subsection{Comparing $\beta$ to $\beta_\mathrm{crit}$ in the MMSN}
-Here we compare the above thermal timescale, Eq. \ref{beta_mmsn_simp}, to  
+
+Fig. \ref{beta_compare} show examples of
+$\beta$ at $r=5$AU and $r=50$AU for a perturbation wavenumber
+$\khat=30$.
+
+% In the outer disk  at $r=50$AU, $\beta$ is vertically-uniform and is less
+% than $\beta_\mathrm{crit}$.  Accordingly we expect, and find, the VSI
+% operates.    
+% In the inner disk ($r=5$AU), the perturbation is in the optically
+% thick regime near the midplane, but becomes optically-thin away from
+% the midplane.  In this case, $\beta < \beta_\mathrm{crit}$
+% for $|z|\gtrsim1.25H$. In fact, we also find the VSI in this case, but
+% with much reduced growth rates (when scaled by the local rotation), as
+% described below.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/beta_compare}
+  \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
+    and $r=5$AU for $\khat=30$ (solid lines). The
+    corresponding horizontal dashed lines are the critical thermal
+    relaxation timescales derived in linear theory. 
+    \label{beta_compare}}
+\end{figure}
+
+In Fig. \ref{mmsn_bcrit_bcool} we compare the above thermal timescale,
+Eq. \ref{beta_mmsn_simp},  for several opacity scales, to  
 the VSI criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
@@
   % -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
-  \beta_\mathrm{crit,MMSN} = 0.024r_\mathrm{AU}^{2/7}. 
+  \beta_\mathrm{crit} = 0.024r_\mathrm{AU}^{2/7}. 
 \end{align}
-
-% \emph{This paragraph currently merely copies the figure caption, which is a waste and not illuminating to read.
-Better to give some description of what the figure is showing.
-(see below for ideas)}\\
-% In Fig. \ref{mmsn_bcrit_bcool} we plot Eq. \ref{beta_mmsn}  with $\mu
-% =2.33$ $\gamma=1.4$, $\hat{T}=\hat{\Sigma}=1$ for opacity scales
-% $\hat{\kappa}_d=0.1, \,1 $ (MMSN) and $10$, at $z=0$ and $z=3H$. \\
-% \emph{Here you can describe the general features e.g. diffusive and thin cooling regions, which will help later discussions.}\\
-% We also plot Eq.
-\ref{bcrit_mmsn} in each panel.
-\emph{This sentence explains nothing and is likely to be of little help to the reader.
-Instead wait to describe the science, e.g.
-after describing the general shape of the $\beta$ curves,  you can start the discussion below with "Fig () also compares the MMSN cooling times to $\beta_{crit}$.}
-
-Consider first the MMSN shown in the middle panels of 
-Fig. \ref{mmsn_bcrit_bcool}. The critical thermal timescale,
-Eq. \ref{bcrit_mmsn}, is useful in the optically thin regime (where the
-curves overlap), since in this limit $\beta$ does not depend on
-height, as assumed in our analysis. \\
-\emph{above sentence too long and confusing.  "useful" is a vague term.  
-"where the curves overlap" is also confusing, forcing the reader to look at the figure in the middle of a complex sentence.
-Better to have already explained above how the $\beta$ curves show that opt.
-thin cooling is height indep.
-(even though should be obvious from equations).
-Then the comparison to $\beta_{crit}$ can be made as something like "Since beta-crit was derived for a height-independent cooling time,  the direct comparison our MMSN cooling model is valid only for modes with optically thin cooling.
-"}
-We thus infer that the VSI can
-operate at tens of AU in the MMSN.
-
-In the inner few AU of the MMSN, however, $\beta$ 
-may be smaller or greater than $\beta_\mathrm{crit}$ depending on
-$z$ as the disk transitions from an optically thick midplane to an
-optically thin atmosphere (see also Fig. \ref{bcrit_mmsn2d}). 
-%Given a fixed $\khat$, one can have $\beta > \beta_\mathrm{crit}$ at
-%the optically thick midplane, but with increasing height the disk 
-%becomes optically thin and $\beta<\beta_\mathrm{crit}$. 
-Our analytical criteria $\beta < \beta_\mathrm{crit}$ is not valid for
-vertically non-uniform $\beta$, and we must turn to numerical
-solutions of the linear problem to determine whether or not the VSI
-operates. We do this in the next section.    \\
-\emph{a better organized discussion might end on this point.  Might not be easy to rearrange to do this, but it would ideal to have the seamless flow to what you do next.}
-
-Lowering the opacity, as shown in the top panels of
-Fig. \ref{mmsn_bcrit_bcool}, shifts the curves inwards but also raises
-them.\\
-\emph{this graphical discussion is not very inuitive (the main problem) and even arbitrary (lowering vs.
-shifting to the left).
-Better to say what's happening physically, i.e.
-`With higher opacity, optically thin cooling faster and diffusive cooling is slower, explaining how the curves shift.} \\
-In fact, lowering the opacity by an order of magnitude renders  
-$\beta > \beta_\mathrm{crit}$ even away from the midplane. This is 
-expected to stabilize the disk. Raising the opacity, as shown in 
-the lower panels of Fig. \ref{mmsn_bcrit_bcool}, decreases the thermal
-timescales, but shifts the curves outward, so the instability is
-restricted to larger radii.
 
@@
     evaluated at the midplane (left) and at $z=3H$ (right) in the
-    fiducial PPD. Eq. \ref{beta_mmsn} is plotted  
+    fiducial PPD. Eq. \ref{beta_mmsn_simp} is plotted  
     for three values of the 
@@
 
-\emph{the following statement is not generally true, so it should not be stated without caveat}
+For disturbances in the optically-thin cooling regime, $\beta$ does
+not depend on the height $z$ or perturbation lengthscale $\khat$. This
+generally occurs at sufficiently large radii. In this case we may
+directly compare $\beta$ to $\beta_\mathrm{crit}$, which was derived
+assuming it is vertically-uniform.  For the case with MMSN
+opacity ($\hat{\kappa}_d=1$), we  infer that the VSI can operate at
+10s of AU.
+
+For disturbances in the optically-thick or diffusive regime, $\beta$
+increases with decreasing $\khat$ (larger scales) and decreases with
+increasing height, on account of the density-dependence of
+$t_\mathrm{diff}$.  This generally occurs at small radii. In this case 
+$\beta$ may be smaller or greater than $\beta_\mathrm{crit}$ depending
+on $z$. Our analytical criteria $\beta < \beta_\mathrm{crit}$ is not valid for
+vertically non-uniform $\beta$, and we must turn to numerical
+solutions of the linear problem to determine whether or not the VSI
+operates. We do this in \S\ref{vsi_mmsn_grow}.
+
+%\emph{a better organized discussion might end on this point.  Might not be easy to rearrange to do this, but it would ideal to have the seamless flow to what you do next.}
+
+Next consider the effect of the opacity scale $\hat{\kappa}_d$. The previous section
+(see also Eq. \ref{beta_mmsn_simp}) indicates that increasing $\hat{\kappa}_d$
+increases (decreases) the cooling time in the optically thick (thin)
+limits. The overall effect of  increasing (decreasing) the opacity
+scale by an order of magnitude is illustrated in the lower (upper)
+panels of Fig. \ref{mmsn_bcrit_bcool}.
+
+For increasing opacity, $\hat{\kappa}_d=1\to 10$,  a comparison
+between the middle and lower panels shows that the optically-thick
+regions shift outward (where the three curves are distinct); while the
+optically-thin regions have lowered cooling times (where the curves
+overlap). Thus, instability is expected to be restricted to larger
+radii.
+
+For decreasing opacity, $\hat{\kappa}_d=1\to 0.1$,  the
+optically-thick regions shift inwards; while cooling times in the
+optically-thin limit are increased. In fact, the latter effect is
+significant enough to raise $\beta > \beta_\mathrm{crit}$ over all
+radii, and even away from the midplane. This is expected to stabilize
+the disk against the VSI.
+
+%\emph{the following statement is not generally true, so it should not be stated without caveat}
 
 We remark that it is possible to satisfy $\beta < \beta_\mathrm{crit}$
-for sufficiently large $\khat$. We illustrate this in 
+at small radii for sufficiently large $\khat$. We illustrate this in 
 Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at each radius the value
 of $\khat$ such that $\beta = \beta_\mathrm{crit}$. If we consider the
-midplane of the inner disk, say $r\sim 1$AU, the VSI would only occur
-on very small radial lengthscales $\ll O(10^{-2}H)$. In the disk
-atmosphere where $\beta$ is small, surface modes may develop, 
-%However, at such
-%small scales, surface modes are likely to dominate (and are
-%restricted to the disk atmosphere), 
-but they are not well-defined in vertically isothermal disk models
-with an imposed boundary, and they also require large $\khat$.
+midplane at, say $r\sim 1$AU, the VSI would only occur on very small
+radial lengthscales $\ll O(10^{-2}H)$. It is, however, unclear if such
+small-scale disturbances are dynamically important.
 
-%surface modes can develop for large $\khat$, 
-%which have small vertical lengthscales scales. 
-%also small vertical scales if they are surface modes
-
-\emph{The following viscosity discussion might be better placed elsewhere.
-Since the issue also comes up for numerical solutions (not just the comparison to beta-crit here), maybe a short concluding subsection to summarize all the arguments on the "vialibity of short wavelength VSI".}
-
-However, disturbances with large $\khat$ may be subject to viscous
-decay in a real disk. Adopting the standard precription for the
-kinematic viscosity $\nu = \alpha c_s H$ \citep{shakura73}, where
-$\alpha$ is the dimensionless viscosity parameter, 
-the viscous timescale for a perturbation lengthscale 
-$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $. 
-Setting $t_\mathrm{visc}$ to be longer than the
-minimum VSI growth timescale $1/h\Omega$ implies 
-$\khat^2\lesssim h/\alpha$. Inserting $h \sim 0.05$,
-we find only modes with $\khat\lesssim 20$ are relevant for
-$\alpha\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha\sim
-10^{-5}$. \citepalias[These low viscosity levels are required for the
-VSI, see][]{nelson13}.
+%In the disk
+% atmosphere where $\beta$ is small, surface modes may develop, 
+% but they are not well-defined in vertically isothermal disk models
+% with an imposed boundary, and they also require large $\khat$.
 
@@
 
-\subsection{VSI in the MMSN}
+\subsection{VSI growth in the MMSN}\label{vsi_mmsn_grow}
 Here we solve the linear stability problem in the MMSN with the scale and
@@
 Eq. \ref{beta_mmsn_simp}.
- Fig. \ref{beta_compare} show examples of
-the thermal timescales at $r=5$AU and $r=50$AU for a
-perturbation wavenumber $\khat=30$. We also plot the 
-corresponding critical thermal timescales evaluated for the MMSN
-(Eq. \ref{bcrit_mmsn}).  \\
-\emph{For this figure to belong here (and not in the previous subsection, with the other comparisons of $\beta$ and $\beta_{crit}$) the discussion should be be more focused on the linear analysis.
+
+% \\
+% \emph{For this figure to belong here (and not in the previous subsection, with the other comparisons of $\beta$ and $\beta_{crit}$) the discussion should be be more focused on the linear analysis.
 Again it's better to start by discussing what the figure tells us, than repeating the figure caption.
@@
 
- \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/beta_compare}
-  \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
-    and $r=5$AU for $\khat=30$ (solid lines). The
-    corresponding horizontal dashed lines are the critical thermal
-    relaxation timescales derived in linear theory. 
-    \label{beta_compare}}
-\end{figure}
-
-The outer disk ($r=50$AU) is optically thin,
-$\beta<\beta_\mathrm{crit}$  and  does not depend on height.
-Accordingly we expect, and find, the VSI operates at $r=50$AU. The
-inner disk ($r=5$AU) is optically thick near the midplane but becomes
-optically thin away from it, such that $\beta < \beta_\mathrm{crit}$
-for $|z|\gtrsim1.25H$. In fact, we also find the VSI in this case, but
-with much reduced growth rates (when scaled by the local rotation), as
-described below.
-
 In Fig. \ref{mmsn_overall} we plot characteristic growth timescales
@@
 that it is the most unstable, except for the appearance of some 
-surface modes for large wavenumbers $\khat\gtrsim 50$, which are 
-boundary-dependent.  \\
-\emph{above sentence too long and complex, could e.g.\ split off that surface modes are neglected and why.}
+surface modes for large wavenumbers $\khat\gtrsim 50$. We do not
+consider surface modes since they are boundary-dependent and likely
+subject to viscous decay (see). 
+%  which are 
+% boundary-dependent. 
+% Disturbances with large $\khat$ are also more likely subject to viscous 
+% decay, as discussed above. 
+ %\\
+%\emph{above sentence too long and complex, could e.g.\ split off that surface modes are neglected and why.}
 %However, surface modes are dependent on
@@
 %and
-Disturbances with large $\khat$ are also more likely subject to viscous 
-decay, as discussed above.
-
-% It is also unclear if surface modes are
-% dynamically important given they are concentrated in the disk
-% atmosphere where the density is small. 
-% The dominance of the fundamental mode is also
-% observed in hydrodynamic simulations of \cite{stoll14} which treat
-% radiative cooling properly.   unclear if surface modes are
-% dynamically important because the there is little mass in the disk at
 
 We therefore consider Fig. \ref{mmsn_overall} as a representation of the
-VSI in outer parts of the MMSN. We see the VSI operates from $\sim 5$AU to $\sim
+VSI in the fiducial MMSN. We see the VSI operates from $\sim 5$AU to $\sim
 50$AU with growth timescales $\sim 30$---40 orbits and
 $t_\mathrm{grow}$ rises rapidly towards the inner disk. It is
-necessary to consider smaller scales with decreasing $r$ as the disk
-generally becomes optically thick.
+necessary to consider smaller scales with decreasing $r$ as
+perturbations generally become optically thick.
 
@@
 $t_\mathrm{grow}\sim 40$ orbits at $r=50$AU, but $t_\mathrm{grow}\sim
-400$ orbits at $r=5$AU. Thus, introducing an optically thick midplane
+400$ orbits at $r=5$AU (and rapidly increases to $> 10^3P_0$ for
+smaller $r$). Thus, introducing an optically thick midplane
 significantly reduces the efficiency of the VSI despite $\beta <

### 2015-05-05 summary.tex
introduce alias for Omega_K,  \OmK, think I got all...
re-writing sec 7 in progress

@@
 temperature to its initial value on a timescale
-$t_c$. This `thermal relaxation' model allows
-the thermodynamic response of the disk to be controlled by varying
-$t_c$, and has been employed in previous numerical simulations of the
-VSI \citepalias{nelson13}.
+$t_c$. % This `thermal relaxation' model allows
+% the thermodynamic response of the disk to be controlled by varying
+% $t_c$, and has been employed in previous numerical simulations of the
+% VSI \citepalias{nelson13}.
 
@@
 \begin{align*}
-  t_\mathrm{crit} = \frac{ h |q|}{\gamma-1}\Omega_K^{-1}, 
+  t_\mathrm{crit} = \frac{ h |q|}{\gamma-1}\OmK^{-1}, 
 \end{align*}
@@
 \subsection{Caveats and outlooks} 
-We highlight below some extensions to the present  
-linear models in future works.
+We highlight below some outstanding issues that should be addressed in
+future works.
+
+\subsubsection{Viscosity} 
+% Depending on where the VSI operates in a
+% disk, additional effects may become important. For example, in the
+% outer parts of a PPD, where we estimate the VSI 
+% operates, disk self-gravity (SG) may play a role. Vertical
+% SG may favor the VSI by increasing $|\p_z\Omega|$ through
+% enhancing the density stratification 
+% (Eq. \ref{vertical_shear}). %sg in pert? no because kx>>1. but maybe for kx~1
+
+We have neglected viscosity. \citetalias{nelson13} find that a
+dimensionless viscosity of $\alpha\lesssim O(10^{-4})$ is required for the
+VSI; while \citetalias{barker15} find that inclusion of viscosity is needed for 
+convergence when studying surface modes in radially global disk
+models. %but not body modes? 
+Viscosity may also affect the
+thermodynamic requirement for the VSI, because viscous heating
+needs to be self-consistently included in the energy equation. This
+should be investigated in future works.
+
+However, disturbances with large $\khat$ may be subject to viscous
+decay in a real disk. Adopting the standard precription for the
+kinematic viscosity $\nu = \alpha c_s H$ \citep{shakura73}, where
+$\alpha$ is the dimensionless viscosity parameter, 
+the viscous timescale for a perturbation lengthscale 
+$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $. 
+Setting $t_\mathrm{visc}$ to be longer than the
+minimum VSI growth timescale $1/h\Omega$ implies 
+$\khat^2\lesssim h/\alpha$. Inserting $h \sim 0.05$,
+we find only modes with $\khat\lesssim 20$ are relevant for
+$\alpha\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha\sim
+10^{-5}$. \citepalias[These low viscosity levels are required for the
+VSI, see][]{nelson13}.
+
+% It is of interest to quantify the effect of viscosity 
+% on the linear VSI. This would help identify, more accurately, the
+% appropriate regions in a protoplanetary disk that are sufficiently
+% laminar to develop the VSI.  This problem is, however, complicated by
+% the fact that proper inclusion of viscosity introduces a meridional
+% flow in the equilibrium state, as well as viscous heating
+% in the energy equation.
 
@@
 %  \sigma_\mathrm{conv,max} =
-%  -\frac{ h^2\Omega_K}{4\gamma\Gamma}\left(p+q\right)
+%  -\frac{ h^2\OmK}{4\gamma\Gamma}\left(p+q\right)
 %  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
@@
 follow-up study. %additional heat source
-
-\subsubsection{Additional physics} 
-Depending on where the VSI operates in a
-disk, additional effects may become important. For example, in the
-outer parts of a PPD, where we estimate the VSI 
-operates, disk self-gravity (SG) may play a role. Vertical
-SG may favor the VSI by increasing $|\p_z\Omega|$ through
-enhancing the density stratification 
-(Eq. \ref{vertical_shear}). %sg in pert? no because kx>>1. but maybe for kx~1
-
-We have neglected viscosity. \citetalias{nelson13} find that a
-dimensionless viscosity of $\alpha\lesssim O(10^{-4})$ is required for the
-VSI; while \citetalias{barker15} find that inclusion of viscosity is needed for 
-convergence when studying surface modes in radially global disk
-models. %but not body modes? 
-Viscosity may also affect the
-thermodynamic requirement for the VSI, because viscous heating
-needs to be self-consistently included in the energy equation. This
-should be investigated in future works.
-
-% It is of interest to quantify the effect of viscosity 
-% on the linear VSI. This would help identify, more accurately, the
-% appropriate regions in a protoplanetary disk that are sufficiently
-% laminar to develop the VSI.  This problem is, however, complicated by
-% the fact that proper inclusion of viscosity introduces a meridional
-% flow in the equilibrium state, as well as viscous heating
-% in the energy equation.

### 2015-05-05 application.tex
rough re-write of sec 7 done by mkl

@@
 \end{align}
-In the inner disk, e.g. $r\sim 1$AU, only
+In the inner disk, e.g. $r_\mathrm{AU}\sim 1$, only
 extremely small-scale perturabtions are in the optically-thin
-regime. However, in the outer disk, e.g. $r\sim 10$AU, the transition
+regime. However, in the outer disk, e.g. $r_\mathrm{AU}\sim 10$, the transition
 occurs at more moderate wavenumbers $\khat\sim 10$.
@@
 \subsection{Comparing $\beta$ to $\beta_\mathrm{crit}$ in the MMSN} 
+In this section we compare the above thermal timescale, Eq. \ref{beta_mmsn_simp}, to 
+the VSI criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
+\begin{align}\label{bcrit_mmsn}
+  % \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
+  % -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
+  \beta_\mathrm{crit} = 0.024r_\mathrm{AU}^{2/7}. 
+\end{align}
+Fig. \ref{beta_compare} show examples of $\beta$ in the fiducial disk
+at $r_\mathrm{AU}=5$ and $r_\mathrm{AU}=50$ for a perturbation wavenumber $\khat=30$.  
+At $r_\mathrm{AU}=50$, the disturbance is in the optically-thin regime,  
+$\beta$ is vertically uniform, as assumed for our analysis, and $\beta
+< \beta_\mathrm{crit}\, \forall z$. Accordingly we expect, and find, the VSI
+operates.
 
-Fig. \ref{beta_compare} show examples of
-$\beta$ at $r=5$AU and $r=50$AU for a perturbation wavenumber
-$\khat=30$.
+On the other hand, at $r_\mathrm{AU}=5$, the perturbation is in the optically
+thick (thin) regime near (away from) the midplane. In this case,
+$\beta < \beta_\mathrm{crit}$ for $|z|\gtrsim1.25H$. Our analytic 
+criteria is not valid for vertically non-uniform $\beta$ so we must
+turn to numerical solutions of the linear problem. We do this in
+\S\ref{vsi_mmsn_grow}. % and find the VSI also operates in this case, but
+% with significantly reduced efficiency.
 
-% In the outer disk  at $r=50$AU, $\beta$ is vertically-uniform and is less
-% than $\beta_\mathrm{crit}$.  Accordingly we expect, and find, the VSI
-% operates.    
-% In the inner disk ($r=5$AU), the perturbation is in the optically
-% thick regime near the midplane, but becomes optically-thin away from
-% the midplane.  In this case, $\beta < \beta_\mathrm{crit}$
-% for $|z|\gtrsim1.25H$. In fact, we also find the VSI in this case, but
-% with much reduced growth rates (when scaled by the local rotation), as
+%, as
 % described below.
@@
   0cm]{figures/beta_compare}
-  \caption{Thermal relaxation timescales in the MMSN at $r=50$AU
+  \caption{Thermal relaxation timescales in the fiducial MMSN at $r=50$AU
     and $r=5$AU for $\khat=30$ (solid lines). The
@@
 
-In Fig. \ref{mmsn_bcrit_bcool} we compare the above thermal timescale,
-Eq. \ref{beta_mmsn_simp},  for several opacity scales, to  
-the VSI criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
-\begin{align}\label{bcrit_mmsn}
-  % \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
-  % -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
-  \beta_\mathrm{crit} = 0.024r_\mathrm{AU}^{2/7}. 
-\end{align}
+In Fig. \ref{mmsn_bcrit_bcool}, we compare $\beta$ to
+$\beta_\mathrm{crit}$ for $r_\mathrm{AU}\in[1,100]$, as well as for different
+opacity scales. Notice that $\beta$ increases for either large or
+small radii. For sufficiently large $r$, disturbances are generally
+in the optically-thin cooling regime, characterized graphically by the
+over-lapping curves. 
+%, where $\beta$ is independent of
+%$z$ and the perturbation lengthscale $\khat$.
+In this case, we may
+directly compare $\beta$ to $\beta_\mathrm{crit}$ and infer that the
+VSI can operate at 10s of AU in the fiducial MMSN disk (with
+$\hat{\kappa}_d=1$).  However, at small radii, $\beta$ depends on both
+$z$ and $\khat$, as the disturbance enters the optically-thick limit
+at the midplane. This regime is characterized by the distinct curves for each
+$\khat$ toward decreasing $r$.
+
+%\emph{a better organized discussion might end on this point.  Might not be easy to rearrange to do this, but it would ideal to have the seamless flow to what you do next.}
+
+Next consider the effect of the opacity scale $\hat{\kappa}_d$. The previous section
+(see also Eq. \ref{beta_mmsn_simp}) indicates that increasing $\hat{\kappa}_d$
+increases (decreases) the cooling time in the optically thick (thin)
+limits. The overall effect of  increasing (decreasing) the opacity
+scale by an order of magnitude is illustrated in the lower (upper)
+panels of Fig. \ref{mmsn_bcrit_bcool}.
+
+For increasing opacity, $\hat{\kappa}_d=1\to 10$,  a comparison
+between the middle and lower panels shows that the optically-thick
+regions shift outward; while the
+optically-thin regions have lowered cooling times. Thus, instability
+is expected to be restricted to larger radii.
+
+For decreasing opacity, $\hat{\kappa}_d=1\to 0.1$,  the
+optically-thick regions shift inwards; while cooling times in the
+optically-thin limit are increased. In fact, the latter effect is
+significant enough to raise $\beta > \beta_\mathrm{crit}$ over all
+radii, and even away from the midplane. This is expected to stabilize
+the disk against the VSI.
 
@@
 
-For disturbances in the optically-thin cooling regime, $\beta$ does
-not depend on the height $z$ or perturbation lengthscale $\khat$. This
-generally occurs at sufficiently large radii. In this case we may
-directly compare $\beta$ to $\beta_\mathrm{crit}$, which was derived
-assuming it is vertically-uniform.  For the case with MMSN
-opacity ($\hat{\kappa}_d=1$), we  infer that the VSI can operate at
-10s of AU.
-
-For disturbances in the optically-thick or diffusive regime, $\beta$
-increases with decreasing $\khat$ (larger scales) and decreases with
-increasing height, on account of the density-dependence of
-$t_\mathrm{diff}$.  This generally occurs at small radii. In this case 
-$\beta$ may be smaller or greater than $\beta_\mathrm{crit}$ depending
-on $z$. Our analytical criteria $\beta < \beta_\mathrm{crit}$ is not valid for
-vertically non-uniform $\beta$, and we must turn to numerical
-solutions of the linear problem to determine whether or not the VSI
-operates. We do this in \S\ref{vsi_mmsn_grow}.
-
-%\emph{a better organized discussion might end on this point.  Might not be easy to rearrange to do this, but it would ideal to have the seamless flow to what you do next.}
-
-Next consider the effect of the opacity scale $\hat{\kappa}_d$. The previous section
-(see also Eq. \ref{beta_mmsn_simp}) indicates that increasing $\hat{\kappa}_d$
-increases (decreases) the cooling time in the optically thick (thin)
-limits. The overall effect of  increasing (decreasing) the opacity
-scale by an order of magnitude is illustrated in the lower (upper)
-panels of Fig. \ref{mmsn_bcrit_bcool}.
-
-For increasing opacity, $\hat{\kappa}_d=1\to 10$,  a comparison
-between the middle and lower panels shows that the optically-thick
-regions shift outward (where the three curves are distinct); while the
-optically-thin regions have lowered cooling times (where the curves
-overlap). Thus, instability is expected to be restricted to larger
-radii.
-
-For decreasing opacity, $\hat{\kappa}_d=1\to 0.1$,  the
-optically-thick regions shift inwards; while cooling times in the
-optically-thin limit are increased. In fact, the latter effect is
-significant enough to raise $\beta > \beta_\mathrm{crit}$ over all
-radii, and even away from the midplane. This is expected to stabilize
-the disk against the VSI.
-
 %\emph{the following statement is not generally true, so it should not be stated without caveat}
@@
 We remark that it is possible to satisfy $\beta < \beta_\mathrm{crit}$
-at small radii for sufficiently large $\khat$. We illustrate this in 
+for sufficiently large $\khat$ and small $r$. We illustrate this in 
 Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at each radius the value
 of $\khat$ such that $\beta = \beta_\mathrm{crit}$. If we consider the
-midplane at, say $r\sim 1$AU, the VSI would only occur on very small
+midplane at, say $r_\mathrm{AU}\sim 1$, the VSI would only occur on very small
 radial lengthscales $\ll O(10^{-2}H)$. It is, however, unclear if such
@@
 
-% \\
-% \emph{For this figure to belong here (and not in the previous subsection, with the other comparisons of $\beta$ and $\beta_{crit}$) the discussion should be be more focused on the linear analysis.
-Again it's better to start by discussing what the figure tells us, than repeating the figure caption.
-E.g.\ you could use Fig 21 to explain when a height varying $\beta$ is required.
-I would also explain (especially if it stays in this subsection) what the growth time is for the r = 5 AU case.
-Looks to be $\sim 1000$ orbits, which is remakrable given that $\beta < \beta_{\rm crit}$ outside $z \gtrsim  1.6 H$.?
-That kind of detail could be illuminating, as it's nominally surprising given previous evidence that behavior near 5 -- 7 $H$ can affect growth.}
-
 In Fig. \ref{mmsn_overall} we plot characteristic growth timescales
@@
 that it is the most unstable, except for the appearance of some 
-surface modes for large wavenumbers $\khat\gtrsim 50$. We do not
-consider surface modes since they are boundary-dependent and likely
-subject to viscous decay (see). 
+surface modes for large wavenumbers $\khat\gtrsim 50$. We discard
+surface modes for this discussion since they are not well-defined 
+in vertically isothermal disk models \citepalias{barker15} and may be 
+subject to viscous decay in real disks (see \S\ref{caveats_visc}).
+
 %  which are 
@@
 We therefore consider Fig. \ref{mmsn_overall} as a representation of the
-VSI in the fiducial MMSN. We see the VSI operates from $\sim 5$AU to $\sim
-50$AU with growth timescales $\sim 30$---40 orbits and
+VSI in the fiducial MMSN. We see the VSI operates from $r_\mathrm{AU}\sim 5$ to $r_\mathrm{AU}\sim
+50$ with growth timescales $\sim 30$---40 orbits and 
 $t_\mathrm{grow}$ rises rapidly towards the inner disk. It is
@@
 
-The fact that we obtain the VSI in the inner disk despite $\beta >
-\beta_\mathrm{crit}$ near the midplane suggests that the instability
+The fact that we can obtain the VSI in the despite $\beta >
+\beta_\mathrm{crit}$ near the midplane (e.g. $\khat = 30$ at
+$r_\mathrm{AU} =5$, see Fig. \ref{beta_compare}) suggests  that the instability
 can operate provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
@@
 become much longer. For example, for $\khat=30$ we find
-$t_\mathrm{grow}\sim 40$ orbits at $r=50$AU, but $t_\mathrm{grow}\sim
-400$ orbits at $r=5$AU (and rapidly increases to $> 10^3P_0$ for
-smaller $r$). Thus, introducing an optically thick midplane
+$t_\mathrm{grow}\sim 40$ orbits at $r_\mathrm{AU}=50$, but $t_\mathrm{grow}\sim
+400$ orbits at $r_\mathrm{AU}=5$ (and rapidly increases to $> 10^3$ orbits for
+smaller $r$). Thus, introducing an optically-thick midplane 
 significantly reduces the efficiency of the VSI despite $\beta <
@@
 Fig. \ref{mmsn_bcrit_bcool} show that $\beta(z=0,\khat=10) \gtrsim
-\beta_\mathrm{crit}$ for $r\lesssim 10$AU in the MMSN (left middle
+\beta_\mathrm{crit}$ for $r_\mathrm{AU}\lesssim 10$ in the MMSN (left middle
 panel); and Fig. \ref{mmsn_overall} 
-show that modes with $\khat=10$ are stabilized inside $\sim 8$AU (where
+show that modes with $\khat=10$ are stabilized for $r_\mathrm{AU}\lesssim 8$ (where
 $t_\mathrm{grow}\to\infty$). Our numerical results suggest the VSI

### 2015-05-05 application.tex
slight revision of sec 7, shortened sec 8 (moved viscosity disscussion here), shortened sec 6.4 (vert non-iso)

@@
 operate. We first model the cooling time $t_c$ based on realistic
-disk properties, which will yield a non-uniform dimensionless cooling
+disk properties, which yields a non-uniform dimensionless cooling
 time $\beta(r,z;\khat)$. We then compare this to the critical cooling time
 $\beta_\mathrm{crit}$ derived above. Finally, we explicitly compute 
-linear growth rates numerically in order to characterize 
-growth times expected in PPDs.
+linear growth rates numerically, using $\beta(r,z;\khat)$, to
+characterize growth times expected in a typical PPD.
 
@@
 For perturbations with lengthscales $l\gg l_\mathrm{rad}$, we assume
-temperature fluctuations are smoothed out by radiative diffusion and
+temperature fluctuations are smoothed out by radiative diffusion. We 
 model this by writing the linearized cooling function as
@@
 definition of thermal relaxation (Eq. \ref{thermal_relax}), 
-we thus identify the scale-dependent thermal relaxation
+we identify the scale-dependent thermal relaxation
 time for diffusion as 
@@
 %$z$ and the perturbation lengthscale $\khat$.
-In this case, we may
-directly compare $\beta$ to $\beta_\mathrm{crit}$ and infer that the
-VSI can operate at 10s of AU in the fiducial MMSN disk (with
+In this case, we may use the criterion $\beta<\beta_\mathrm{crit}$ to
+infer that the VSI can operate at 10s of AU in the fiducial MMSN disk (with
 $\hat{\kappa}_d=1$).  However, at small radii, $\beta$ depends on both
@@
 of $\khat$ such that $\beta = \beta_\mathrm{crit}$. If we consider the
-midplane at, say $r_\mathrm{AU}\sim 1$, the VSI would only occur on very small
+midplane at, say $r_\mathrm{AU}\sim 1$, the VSI would only occur on 
 radial lengthscales $\ll O(10^{-2}H)$. It is, however, unclear if such
@@
 In Fig. \ref{mmsn_overall} we plot characteristic growth timescales
-$t_\mathrm{grow} \equiv 1/\sigma$ of the fundamental mode with a range of
+$t_\mathrm{grow} \equiv 1/\sigma$ of the VSI with a range of
 wavenumbers, in units of the local orbital time $P_\mathrm{orb}\equiv
-2\pi/\OmK$. In fact, we either only find the fundamental mode, or 
+2\pi/\OmK$. We either only find the fundamental mode, or 
 that it is the most unstable, except for the appearance of some 
@@
 surface modes for this discussion since they are not well-defined 
-in vertically isothermal disk models \citepalias{barker15} and may be 
-subject to viscous decay in real disks (see \S\ref{caveats_visc}).
+in vertically isothermal disks with an imposed boundary,
+\citepalias{barker15} and may be subject to viscous decay in reality
+(see \S\ref{caveats_visc}).
 
@@
 We therefore consider Fig. \ref{mmsn_overall} as a representation of the
-VSI in the fiducial MMSN. We see the VSI operates from $r_\mathrm{AU}\sim 5$ to $r_\mathrm{AU}\sim
+VSI in the fiducial MMSN. We see the VSI operates from
+$r_\mathrm{AU}\sim 5$ to $r_\mathrm{AU}\sim 
 50$ with growth timescales $\sim 30$---40 orbits and  
@@
 
-The fact that we can obtain the VSI in the despite $\beta >
+The fact that we can obtain the VSI at small radii, despite $\beta >
 \beta_\mathrm{crit}$ near the midplane (e.g. $\khat = 30$ at
-$r_\mathrm{AU} =5$, see Fig. \ref{beta_compare}) suggests  that the instability
-can operate provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
-fraction of the vertical domain. However, the normalized growth times
-become much longer. For example, for $\khat=30$ we find
-$t_\mathrm{grow}\sim 40$ orbits at $r_\mathrm{AU}=50$, but $t_\mathrm{grow}\sim
-400$ orbits at $r_\mathrm{AU}=5$ (and rapidly increases to $> 10^3$ orbits for
+$r_\mathrm{AU} =5$, see Fig. \ref{beta_compare}), suggests  that the
+instability can operate provided $\beta < \beta_\mathrm{crit}$ for a
+sufficiently large fraction of the vertical domain. However, the
+normalized growth times become much longer. For example, for
+$\khat=30$ we find $t_\mathrm{grow}\sim 40$ orbits at
+$r_\mathrm{AU}=50$, but $t_\mathrm{grow}\sim 400$ orbits at
+$r_\mathrm{AU}=5$ (and rapidly increases to $> 10^3$ orbits for
 smaller $r$). Thus, introducing an optically-thick midplane
@@
   \caption{Characteristic growth times of the VSI in 
-    the MMSN. These are calculated using the thermal relaxation
+    the fiducial MMSN. These are calculated using the thermal relaxation
     function given by Eq. \ref{beta_mmsn_simp}. 

### 2015-05-05 summary.tex
rough re-write of sec 7 done by mkl
slight revision of sec 7, shortened sec 8 (moved viscosity disscussion here), shortened sec 6.4 (vert non-iso)

@@
 instability (VSI) in astrophysical disks that arise from the 
-height dependence of the equilibrium angular velocity profile 
-$\Omega=\Omega(r,z)$ in a baroclinic disk. Vertical shear is generally
-expected in astrophysical disks due to baroclinicity. This may arise 
-from a radial variation in the disk temperature, or more generally in
+height dependence of the equilibrium angular velocity profile,  
+$\p_z\Omega\neq 0$. Vertical shear is generally
+expected in astrophysical disks due to baroclinicity ($\nabla
+P\times\nabla \rho \neq 0$). This may be due
+to a radial variation in the disk temperature, or more generally in
 the disk entropy. However, astrophysical disks are also typically
 sub-adiabatically stratified for stability against vertical
-convection. In this case, the VSI requires a short thermal or cooling
-timescale $t_c$ in order to operate.
+convection ($N_z^2\geq 0$). In this case, the VSI requires a short
+thermal or cooling timescale $t_c$ in order to operate.
 
 Physically, a short thermal timescale is needed to reduce the  
-stabilizing effect of buoyancy forces in a stably stratified
-disk. Because vertical shear is weak in thin disks, $t_c\Omega\ll 1$
-is generally needed for the VSI to operate, otherwise buoyancy would
-be strongly stabilizing. Our primary goal was to quantify this
-requirement, thereby provide a simple way to assess 
-whether or not the VSI is relevant in realistic disks.
+stabilizing effect of buoyancy forces. Because vertical shear is weak
+in thin disks, very rapid cooling, corresponding to $t_c\Omega\ll 1$, 
+is generally needed for the VSI, otherwise buoyancy is strongly
+stabilizing. Our primary goal was to quantify this requirement,
+thereby provide a simple way to assess whether or not the VSI is
+relevant in realistic disks.
 
-To do so, we generalized previous vertically global linear 
+To do so, we generalized previous vertically global, radially local linear 
 calculations of the VSI \citepalias{nelson13,mcnally14,barker15}, which
@@
 temperature to its initial value on a timescale
-$t_c$. % This `thermal relaxation' model allows
+$t_c$. 
+% This `thermal relaxation' model allows
 % the thermodynamic response of the disk to be controlled by varying
@@
 % VSI \citepalias{nelson13}.  
-
-In accordance with the above studies, which show that the VSI occurs for 
-perturbations with small radial lengthscales, we applied the 
-radially local approximation in the linearized fluid equations. 
-We mostly focused on vertically isothermal disks as 
-this is the relevant case for protoplanetary disks (PPDs) passively
+%In accordance with the above studies, which show that the VSI occurs for 
+%perturbations with small radial lengthscales, we applied the 
+%radially local approximation in the linearized fluid equations. 
+We mostly focused on vertically isothermal disks as this is the
+relevant case for protoplanetary disks (PPDs) passively 
 irradiated by its central star \citep{chiang97}. It also permits a
@@
 linear models considered in \cite{lubow93} by adding vertical shear
-and (rapid) thermal relaxation. We obtained explicit solutions 
+and (rapid) cooling. We obtained explicit solutions 
 describing the VSI with thermal relaxation in the thin disk limit.
@@
 \end{align*}
-below which the VSI operates effectively. While $t_\mathrm{crit}$ was
-derived for the lowest order VSI mode --- the `fundamental  
-corrugation mode' --- we find from numerical calculations that
-$t_\mathrm{crit}$ can be used to characterize the general
-stabilization of the VSI by a finite thermal timescale. %That is, all
+below which the VSI operates effectively. 
+%While $t_\mathrm{crit}$ was
+%derived for the lowest order VSI mode --- the `fundamental  
+%corrugation mode' --- we find from numerical calculations that
+%$t_\mathrm{crit}$ can be used to characterize the general
+%stabilization of the VSI by a finite thermal timescale. %That is, all
 %VSI growth rates are reduced significantly as $\beta$ is increased
@@
 %zero to $\beta_\mathrm{crit}$.    
-We have therefore provided a quantitative measure for the thermal
+We have thus provided a quantitative measure for the thermal
 timescale requirement for the VSI in astrophysical 
@@
 radii, but only on radial lengthscales signficantly shorter
-than the disk scaleheight. However, such small-scale perturbations may 
-be subject to viscous decay in a real disk.
+than the disk scaleheight. %However, such small-scale perturbations may 
+%be subject to viscous decay in a real disk.
 
@@
 
-\subsubsection{Viscosity} 
+\subsubsection{Viscosity}\label{caveats_visc}
 % Depending on where the VSI operates in a
@@
 
-We have neglected viscosity. \citetalias{nelson13} find that a
-dimensionless viscosity of $\alpha\lesssim O(10^{-4})$ is required for the
-VSI; while \citetalias{barker15} find that inclusion of viscosity is needed for 
-convergence when studying surface modes in radially global disk
-models. %but not body modes? 
-Viscosity may also affect the
-thermodynamic requirement for the VSI, because viscous heating
-needs to be self-consistently included in the energy equation. This
-should be investigated in future works.
-
-However, disturbances with large $\khat$ may be subject to viscous
-decay in a real disk. Adopting the standard precription for the
-kinematic viscosity $\nu = \alpha c_s H$ \citep{shakura73}, where
-$\alpha$ is the dimensionless viscosity parameter, 
-the viscous timescale for a perturbation lengthscale 
+We have neglected viscosity. 
+% while \citetalias{barker15} find that inclusion of viscosity is needed for 
+% convergence when studying surface modes in radially global disk
+% models. %but not body modes? 
+% Viscosity may also affect the
+% thermodynamic requirement for the VSI, because viscous heating
+% needs to be self-consistently included in the energy equation. This
+% should be investigated in future works.   
+An important effect of viscosity is setting a minimum wavelength (or 
+maximum wavenumber) for which the VSI is relevant. Disturbances with
+large $\khat\equiv k_xH$ may be subject to viscous decay. Adopting the standard
+precription for the kinematic viscosity $\nu = \alpha c_s H$
+\citep{shakura73}, where $\alpha$ is the dimensionless viscosity
+parameter, the viscous timescale for a perturbation lengthscale  
 $l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $. 
@@
 $\alpha\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha\sim
-10^{-5}$. \citepalias[These low viscosity levels are required for the
-VSI, see][]{nelson13}.
+10^{-5}$. \citepalias[These small viscosity values are required for
+the VSI to operate, see][]{nelson13}.
+
+%\citetalias{nelson13} find that a
+%dimensionless viscosity of $\alpha\lesssim O(10^{-4})$ is required for the
+%VSI.
+
+The above argument justifies our focus on moderate
+wavenumbers, $\khat=O(10)$, which precludes extremely small-scale
+disturbances such as surface modes. Nevertheless, future calculations
+should explicitly account for viscosity. In particular, viscous
+heating may influence the thermodynamic dependence of the VSI.
+
+% needs to be self-consistently included in the energy equation.
 
@@
 However, going beyond thermal relaxation requires a
-proper treatment of radiative diffusion. % The linear problem then
+proper treatment of radiative diffusion. 
+% done roughly in this work -
+% a proper study 
 % involves a Laplacian operator and additional boundary

### 2015-05-06 appendix.tex
another round of edits to 7 with a few italics comments remaining.  shortened version of appendix E moved to main text.

@@
 
-\section{Fiducial model for a protoplanetary disk}\label{mmsn}
-For results application in \S\ref{application} we use the disk model
-described in \cite{chiang10}. This disk model orbits a Solar-mass star and 
-has the surface density distribution
-\begin{align}\label{mmsn_sigma}
-  \Sigma = 2200
-  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-3/2}\mathrm{g}\,\mathrm{cm}^{-2},  
-\end{align}
-and the temperature profile
-\begin{align}\label{mmsn_temp}
-  T = 120\hat{T}\left(\frac{r}{\mathrm{AU}}\right)^{-3/7} \mathrm{K},
-\end{align}
-which implies $q=-3/7$. In the above expressions, $\hat{\Sigma}$ and
-$\hat{T}$ are dimensionless coefficients used to scale the model
-relative to the MMSN. By assuming a
-vertically isothermal disk, we deduce the disk aspect-ratio 
-\begin{align}\label{mmsn_epsilon}
-  h =
-  3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}, 
-\end{align}
-and the mid-plane density distribution 
-\begin{align}
-%  \rho_0 = 2.7\times10^{-9}
-%  \hat{\upsilon}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
-\rho_0 = 1.7\times10^{-9}
-  \hat{\Sigma}\left(\frac{\hat{T}}{\mu}\right)^{-1/2}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},
-\end{align}
-which implies $p=-39/14$. 
-%0.033576258
-In addition, we use the opacity model
- \begin{align}
-   \kappa_d &= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2
-   \mathrm{cm}^2\,\mathrm{g}^{-1}
-    =
-   2.88\hat{\kappa}_d\hat{T}^2\left(\frac{r}{\mathrm{AU}}\right)^{-6/7}\mathrm{cm}^2\,\mathrm{g}^{-1},   
- \end{align}
-\citep{bell94} where the second equality follows from the model
-temperature profile above, and $\hat{\kappa}_d$ has similar meaning as
-$\hat{\Sigma}$ and $\hat{T}$. In \S\ref{application} we fix
-$\hat{\Sigma} = \hat{T}=1$, $\mu=2.33$ and use $\gamma=1.4$.
+%\section{Fiducial model for a protoplanetary disk}\label{mmsn}
+%For results application in \S\ref{application} we use the disk model
+%described in \cite{chiang10}. This disk model orbits a Solar-mass star and 
+%has the surface density distribution
+%\begin{align}\label{mmsn_sigma}
+%  \Sigma = 2200
+%  \hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-3/2}\mathrm{g}\,\mathrm{cm}^{-2},  
+%\end{align}
+%and the temperature profile
+%\begin{align}\label{mmsn_temp}
+%  T = 120\hat{T}\left(\frac{r}{\mathrm{AU}}\right)^{-3/7} \mathrm{K},
+%\end{align}
+%which implies $q=-3/7$. In the above expressions, $\hat{\Sigma}$ and
+%$\hat{T}$ are dimensionless coefficients used to scale the model
+%relative to the MMSN. By assuming a
+%vertically isothermal disk, we deduce the disk aspect-ratio 
+%\begin{align}\label{mmsn_epsilon}
+%  h =
+%  3.36\times10^{-2}\left(\frac{\hat{T}}{\mu}\right)^{1/2}\left(\frac{r}{\mathrm{AU}}\right)^{2/7}, 
+%\end{align}
+%and the mid-plane density distribution 
+%\begin{align}
+%%  \rho_0 = 2.7\times10^{-9}
+%%  \hat{\upsilon}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},  
+%\rho_0 = 1.7\times10^{-9}
+%  \hat{\Sigma}\left(\frac{\hat{T}}{\mu}\right)^{-1/2}\left(\frac{r}{\mathrm{AU}}\right)^{-39/14}\mathrm{g}\,\mathrm{cm}^{-3},
+%\end{align}
+%which implies $p=-39/14$. 
+%%0.033576258
+%In addition, we use the opacity model
+% \begin{align}
+%   \kappa_d &= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2
+%   \mathrm{cm}^2\,\mathrm{g}^{-1}
+%    =
+%   2.88\hat{\kappa}_d\hat{T}^2\left(\frac{r}{\mathrm{AU}}\right)^{-6/7}\mathrm{cm}^2\,\mathrm{g}^{-1},   
+% \end{align}
+%\citep{bell94} where the second equality follows from the model
+%temperature profile above, and $\hat{\kappa}_d$ has similar meaning as
+%$\hat{\Sigma}$ and $\hat{T}$. In \S\ref{application} we fix
+%$\hat{\Sigma} = \hat{T}=1$, $\mu=2.33$ and use $\gamma=1.4$.

### 2015-05-06 application.tex
another round of edits to 7 with a few italics comments remaining.  shortened version of appendix E moved to main text.

--- 2015-05-05 application.tex
@@
-\section{Application to protoplanetary disks}\label{application}
-
-As an application of our linear models, we estimate where in a
-PPD do the thermodynamic conditions allow the VSI to
-operate. We first model the cooling time $t_c$ based on realistic
-disk properties, which will yield a non-uniform dimensionless cooling
-time $\beta(r,z;\khat)$. We then compare this to the critical cooling time
-$\beta_\mathrm{crit}$ derived above. Finally, we explicitly compute 
-linear growth rates numerically in order to characterize 
-growth times expected in PPDs.
+\section{The VSI with realistic  protoplanetary disk cooling}\label{application} 
+To this point we have treated the cooling time, $\beta$, as a free parameter that, moreover, is independent of height. 
+To understand the applicability of the VSI to PPDs, we now consider $\beta$ values that are consistent with PPD models.
+Section \ref{cooling_model}  develops our PPD cooling law, which depends on disk radius, disk height and perturbation wavelength.
+We compare these PPD $\beta$ values to $\beta_{\rm crit}$ in \S\ref{bcritPPD}.
+While this comparison is instructive, it is not complete since $\beta_{\rm crit}$ was derived assuming a vertically constant $\beta$.
+Thus in \S\ref{vsi_mmsn_grow} we analyze the linear growth of VSI with our PPD $\beta$ values.
 
@@
 
-\subsection{Cooling model}
+\subsection{Cooling model} \label{cooling_model}
 
 \subsubsection{Radiative diffusion and Newtonian cooling}
-
-We first consider separately perturbations with radial lengthscales
-$l\gg l_\mathrm{rad}$ and $l\ll l_\mathrm{rad}$, where       
+The thermal relaxation of a perturbation depends on the relative sizes of the perturbation's 
+lengthscale, $l$, and the photon mean-free-path,
 \begin{align}\label{lrad}
   l_\mathrm{rad} \equiv \frac{1}{\kappa_d\rho} 
-\end{align} is the photon mean-free-path and $\kappa_d$ is the (dust)
-opacity.
+\end{align}
+where $\kappa_d$ is the opacity, specifically the dust opacity appropriate for cold PPDs.
 
-For perturbations with lengthscales $l\gg l_\mathrm{rad}$, we assume
-temperature fluctuations are smoothed out by radiative diffusion and
-model this by writing the linearized cooling function as
+In the optically thick limit, $l\gg l_\mathrm{rad}$, radiative diffusion smooths out thermal perturbations.
+In this regime, the linearized cooling function is
 
 \begin{align}\label{diff_cool_proper}
-  \delta \Lambda = \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
+  \delta \Lambda_\mathrm{diff} = \frac{P}{\rho T C_v} \nabla\cdot\left(k_\mathrm{rad}\nabla\delta
     T\right),  
 \end{align}
-where $k_\mathrm{rad}$ is the conduction coefficient to be defined
-implicitly, and $C_v$ is the heat capacity as constant volume.
-
-% However, Eq. \ref{diff_cool_proper} fundamentally changes the linear
-% problem by introducing higher order derivatives. We defer a proper
-% analysis of this problem to a follow-up study.
+where $k_\mathrm{rad}$ is the radiative conduction coefficient defined
+below, and $C_v$ is the specific heat capacity at constant volume.
 
 Since the VSI is characterized by vertically-elongated,
-radially-narrow disturbances, we simplify Eq. \ref{diff_cool_proper} 
-by retaining only radial derivatives of the perturbations, so that 
+radially-narrow disturbances, we retain only the radial derivatives in Eq. \ref{diff_cool_proper}.
+Thus, in the radially local approximation, we have
 \begin{align}\label{diff_cool_approx}
@@
   %  T\right) &
-  \delta\Lambda \to %-\frac{k_\mathrm{rad}}{\rho
+  \delta\Lambda_\mathrm{diff} \approx %-\frac{k_\mathrm{rad}}{\rho
   % C_v}k_x^2P\left(\frac{\delta T}{T}\right)%\notag\\
@@
 \end{align}
-\citepalias[see also][]{barker15} where $\eta=k_\mathrm{rad}/\rho C_v$ is
-the diffusion coefficient and 
-is given in terms of physical disk parameters as 
+where the radiative diffusion coefficient is
 \begin{align}\label{eta_def}
-  \eta = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 C_v}, 
+  \eta \equiv {k_\mathrm{rad} \over \rho C_v} = \frac{16\sigma_s T^3}{3\kappa_d\rho^2 C_v}, 
 \end{align}
-where $\sigma_s$ is the Stefan-Boltzmann constant. From our
-definition of thermal relaxation (Eq. \ref{thermal_relax}), 
-we thus identify the scale-dependent thermal relaxation
-time for diffusion as 
+and $\sigma_s$ is the Stefan-Boltzmann constant.
+
+The thermal relaxation  timescale (defined in Eq. \ref{thermal_relax}) 
+for radiative diffusion is thus 
 \begin{align}\label{tc_diff_cool} 
@@
 
-%\subsubsection{Newtonian cooling}\label{newton_cool}
-%\emph{There are no rules against subsections this short, but I think it's bad style.}
-Perturbations with small lengthscales, $l\ll 
-l_\mathrm{rad}$, are in the optically thin regime. In this case, we assume 
-Newtonian cooling can be applied, which is exactly the thermal 
-relaxation model as adopted in our linear models. That is, temperature
-fluctuations $\delta T$ decay on a timescale $t_\mathrm{cool}$,
-independent of $k_x$. For simplicity, we take  
+In the optically thin regime, $l\ll 
+l_\mathrm{rad}$, thermal relaxation operates by ``Newtonian cooling."
+The cooling time is independent of $l$ and 
+inversely proportional to the opacity, $t_\mathrm{thin} \propto 1/\kappa_d$. 
+Specifically 
 \begin{align}
-  t_\mathrm{cool} = \frac{l_\mathrm{rad}^2}{3\eta}. 
+  t_\mathrm{thin} = \frac{l_\mathrm{rad}^2}{3\eta},
 \end{align}
-We note that $t_\mathrm{cool}$ does not explicitly depend on $\rho$.
+and $t_\mathrm{thin}$ does not depend on $\rho$ (because $\kappa_d$ depends on $T$ only).
 
-\subsubsection{Evaluation for a fiducial disk model}\label{toy_relax}
-We define the effective thermal relaxation timescale in linear theory as
+Our general cooling time for all perturbations,
 \begin{align}\label{tc_def}
-  t_c &\equiv t_\mathrm{cool} + t _\mathrm{diff}. %=
+  t_c &\equiv t_\mathrm{thin} + t _\mathrm{diff},%=
   % \frac{l_\mathrm{rad}^2}{3\eta} + \frac{1}{\eta k_x^2},  
 \end{align}
-Eq. \ref{tc_def} is a simple prescription so
-that for small scales (large $k_x$), $t_c\to t_\mathrm{cool}$, while
-for large scales (small $k_x$), $t_c\to t_\mathrm{diff}$. Writing the
-volume density $\rho = \Sigma\hat{\rho}/\sqrt{2\pi}H$, where
-$\hat{\rho} = \rho/\rho_0$ and $\Sigma(r)$ is the surface density at the
-radius of interest, the dimensionless thermal
+is a simple prescription to smoothly match the optically thick and thin limits.
+
+\subsubsection{PPD cooling times}\label{toy_relax}
+In a vertically isothermal disk with surface density $\Sigma = \rho_0/\sqrt{2\pi}H$, 
+the dimensionless thermal
 relaxation time $\beta$ becomes 
@@
 \end{align}
-Note that $\beta$ depends the perturbation lengthscale, 
-distance from the midplane, as well as the distance from the star.
+where $\hat{\rho} = \rho/\rho_0$.  The first and second terms in square brackets represent the optically thin and thick cooling regimes, respectively.
 
-%\subsubsection{Evaluation for a fiducial disk model}
+We adopt the Minimum Mass Solar Nebula
+(MMSN) disk model of \cite{chiang10} which specifies
+\emph{note that the normalizations are suppressed b/c I can't see where we used them.  We should only add back in if we decide to use them.}
+\begin{subequations}
+\begin{align}
+\label{mmsn_sigma}
+  \Sigma &= 2200 %\hat{\Sigma}\left(\frac{r}{\mathrm{AU}}\right)^{-3/2}
+ r_\mathrm{AU}^{-3/2} \mathrm{g}\,\mathrm{cm}^{-2},  \\
+ T &= 120%\hat{T}\left(\frac{r}{\mathrm{AU}}\right)
+  r_\mathrm{AU}^{-3/7} \mathrm{K}, \label{mmsn_temp}  
+\end{align}\end{subequations}
+with $r_\mathrm{AU}\equiv r/\mathrm{AU}$.
+This model has $q = -3/7$, $p = -39/14$ and $h  = 0.034 r_\mathrm{AU}^{2/7}$, using $\mu = 2.33$ and the above relation between $\Sigma$ and $\rho_0 \propto r^p$.
 
-We evaluate Eq. \ref{real_beta} for the Minimum Mass Solar Nebula
-(MMSN) disk model described in \cite{chiang10} and summarized in
-Appendix \ref{mmsn} (where an opacity model is also chosen). Defining
-$r_\mathrm{AU}\equiv r/\mathrm{AU}$, we find Eq. \ref{real_beta}
-becomes 
+For the dust opacity we adopt the \citet{bell94} law with $\kappa_d \propto T^2$, giving
+\begin{align}
+ \kappa_d %&= 2 \hat{\kappa}_d \left(\frac{T}{100\mathrm{K}}\right)^2 \mathrm{cm}^2\,\mathrm{g}^{-1}
+    =
+   2.88\hat{\kappa}_d r_\mathrm{AU}^{-6/7}\mathrm{cm}^2\,\mathrm{g}^{-1},   
+\end{align}
+where the opacity normalication factor, $\hat{\kappa}_d$, scales with the ratio of small dust to gas;  $\hat{\kappa}_d = 1$ is our fiducial case.
+
+%and summarized in
+%Appendix \ref{mmsn} (where an opacity model is also chosen). Defining
+%$r_\mathrm{AU}\equiv r/\mathrm{AU}$, we find Eq. \ref{real_beta}
+%becomes 
 % \begin{align}\label{beta_mmsn}
@@
 % \emph{I would keep $\kappa$ scaling in this eqn, since it's used in plots.  Then just say  that $\kappa_d = 1$ is fiducial case. }
+
+For this disk and opacity model, the cooling time becomes
 \begin{align}\label{beta_mmsn_simp}
-  \beta(r,z;\khat)\notag = &3.9\times10^{-3}\hat{\kappa}_d
+  \beta(r,z;\khat)\notag = &3.9\times10^{-3}\hat{\kappa}_d^{-1}
   r_\mathrm{AU}^{9/14}
-  \notag\\ &\times \left[\hat{\kappa}_d^{-2} +
-    1.9\times10^7\hat{\rho}^2(z)r_\mathrm{AU}^{-33/7}\khat^{-2}\right]. 
+  \notag\\ &\times \left[1 +
+    1.9\times10^7\hat{\kappa}_d^{2}\hat{\rho}^2(z)r_\mathrm{AU}^{-33/7}\khat^{-2}\right]. 
 \end{align}
-The opacity scale for our fiducial case is $\hat{\kappa}_d=1$.
-
-Perturbations are in the optically thin regime,
-$l\sim 1/k_x \lesssim\l_\mathrm{rad}$, when 
+Perturbations are in the optically thin regime for
 \begin{align}
@@
 extremely small-scale perturabtions are in the optically-thin
-regime. However, in the outer disk, e.g. $r_\mathrm{AU}\sim 10$, the transition
-occurs at more moderate wavenumbers $\khat\sim 10$.
-%, though the
-%wavelength is still small compared to $H$.
-
-%\emph{I would explain here that when $k < c\rho \kappa^m r^n$ (with correct values for $c, m, n$) cooling is optically thi%n, as this is useful for explaining figure.}
-% where
-% \begin{align}
-%   \
-% %  \frac{3.7\times10^{-3}}{\mu\left(\gamma-1\right)\hat{\kappa}_d\hat{T}^5}, 
-% \end{align}
-% and
-% \begin{align}
-%   \khat_0 = 4.4\times10^{3}\hat{\kappa}_d\hat{T}^2\hat{\Sigma}.
-% \end{align}
-
-\subsection{Comparing $\beta$ to $\beta_\mathrm{crit}$ in the MMSN} 
-In this section we compare the above thermal timescale, Eq. \ref{beta_mmsn_simp}, to 
-the VSI criterion Eq. \ref{iso_vsi_cond} evaluated for the MMSN,     
-\begin{align}\label{bcrit_mmsn}
-  % \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
-  % -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
-  \beta_\mathrm{crit} = 0.024r_\mathrm{AU}^{2/7}. 
-\end{align}
-Fig. \ref{beta_compare} show examples of $\beta$ in the fiducial disk
-at $r_\mathrm{AU}=5$ and $r_\mathrm{AU}=50$ for a perturbation wavenumber $\khat=30$.  
-At $r_\mathrm{AU}=50$, the disturbance is in the optically-thin regime,  
-$\beta$ is vertically uniform, as assumed for our analysis, and $\beta
-< \beta_\mathrm{crit}\, \forall z$. Accordingly we expect, and find, the VSI
-operates.
-
-On the other hand, at $r_\mathrm{AU}=5$, the perturbation is in the optically
-thick (thin) regime near (away from) the midplane. In this case,
-$\beta < \beta_\mathrm{crit}$ for $|z|\gtrsim1.25H$. Our analytic 
-criteria is not valid for vertically non-uniform $\beta$ so we must
-turn to numerical solutions of the linear problem. We do this in
-\S\ref{vsi_mmsn_grow}. % and find the VSI also operates in this case, but
-% with significantly reduced efficiency.
-
-%, as
-% described below.
+regime near the midplane. However, in the outer disk, e.g. $r_\mathrm{AU}\sim 10$, 
+ more moderate wavenumbers $\khat\gtrsim 10$ experience optically-thin cooling.
 
@@
 
-In Fig. \ref{mmsn_bcrit_bcool}, we compare $\beta$ to
-$\beta_\mathrm{crit}$ for $r_\mathrm{AU}\in[1,100]$, as well as for different
-opacity scales. Notice that $\beta$ increases for either large or
-small radii. For sufficiently large $r$, disturbances are generally
-in the optically-thin cooling regime, characterized graphically by the
-over-lapping curves. 
-%, where $\beta$ is independent of
-%$z$ and the perturbation lengthscale $\khat$.
-In this case, we may
-directly compare $\beta$ to $\beta_\mathrm{crit}$ and infer that the
-VSI can operate at 10s of AU in the fiducial MMSN disk (with
-$\hat{\kappa}_d=1$).  However, at small radii, $\beta$ depends on both
-$z$ and $\khat$, as the disturbance enters the optically-thick limit
-at the midplane. This regime is characterized by the distinct curves for each
-$\khat$ toward decreasing $r$.
+\subsection{$\beta$ vs.\ $\beta_\mathrm{crit}$ in PPDs}\label{bcritPPD} 
+The simplest way to estimate whether the VSI can operate at a given radius in a PPD is to 
+compare the cooling time (Eq. \ref{beta_mmsn_simp}) to 
+its critical value (Eq. \ref{iso_vsi_cond}), evaluated for the MMSN as    
+\begin{align}\label{bcrit_mmsn}
+  % \beta_\mathrm{crit} = \frac{1.44\times10^{-2}}{(\gamma
+  % -1)}\left(\frac{\hat{T}}{\mu}\right)^{1/2}r_\mathrm{AU}^{2/7}. 
+  \beta_\mathrm{crit} = 0.024r_\mathrm{AU}^{2/7}. 
+\end{align}
 
-%\emph{a better organized discussion might end on this point.  Might not be easy to rearrange to do this, but it would ideal to have the seamless flow to what you do next.}
+This comparison strictly valid only for optically thin cooling which is independent of height,
+as assumed in the derivation of $\beta_\mathrm{crit}$.  This condition typically holds in the outer disk.  
+Fig. \ref{beta_compare} shows that a  $\khat=30$ perturbation at 50 AU has vertically constant $\beta$. 
+ Furthermore, since $\beta < \beta_\mathrm{crit}$, the mode is unstable.
 
-Next consider the effect of the opacity scale $\hat{\kappa}_d$. The previous section
-(see also Eq. \ref{beta_mmsn_simp}) indicates that increasing $\hat{\kappa}_d$
-increases (decreases) the cooling time in the optically thick (thin)
-limits. The overall effect of  increasing (decreasing) the opacity
-scale by an order of magnitude is illustrated in the lower (upper)
-panels of Fig. \ref{mmsn_bcrit_bcool}.
+For optically thick cooling, $\beta$ varies with height, complicating the comparison with $\beta_\mathrm{crit}$.   
+At 5 AU, Fig. \ref{beta_compare} shows that cooling times are long in the midplane with $\beta > \beta_\mathrm{crit}$.
+However, since $\beta < \beta_\mathrm{crit}$ away from the midplane, we require the analysis of \S\ref{vsi_mmsn_grow} to determine if the VSI can grow.
+(That analysis will show that growth is stongly suppressed for this case.)  We proceed with the awareness that optically thick regions in the inner disk are a complication, but that we can reasonably expect VSI growth if $\beta < \beta_\mathrm{crit}$ at all heights.
 
-For increasing opacity, $\hat{\kappa}_d=1\to 10$,  a comparison
-between the middle and lower panels shows that the optically-thick
-regions shift outward; while the
-optically-thin regions have lowered cooling times. Thus, instability
-is expected to be restricted to larger radii.
-
-For decreasing opacity, $\hat{\kappa}_d=1\to 0.1$,  the
-optically-thick regions shift inwards; while cooling times in the
-optically-thin limit are increased. In fact, the latter effect is
-significant enough to raise $\beta > \beta_\mathrm{crit}$ over all
-radii, and even away from the midplane. This is expected to stabilize
-the disk against the VSI.
+%Fig. \ref{beta_compare} show examples of $\beta$ in the fiducial disk
+%at $r_\mathrm{AU}=5$ and $r_\mathrm{AU}=50$ for a perturbation wavenumber $\khat=30$.  
+%At $r_\mathrm{AU}=50$, the disturbance is in the optically-thin regime,  
+%$\beta$ is vertically uniform, as assumed for our analysis, and $\beta
+%< \beta_\mathrm{crit}\, \forall z$. Accordingly we expect, and find, the VSI
+%operates.     
+%
+%On the other hand, at $r_\mathrm{AU}=5$, the perturbation is in the optically
+%thick (thin) regime near (away from) the midplane. In this case,
+%$\beta < \beta_\mathrm{crit}$ for $|z|\gtrsim1.25H$. Our analytic 
+%criteria is not valid for vertically non-uniform $\beta$ so we must
+%turn to numerical solutions of the linear problem. We do this in
+%\S\ref{vsi_mmsn_grow}. % and find the VSI also operates in this case, but
+%% with significantly reduced efficiency. 
+%
+%%, as
+%% described below.
 
@@
 
-%\emph{the following statement is not generally true, so it should not be stated without caveat}
+In Fig. \ref{mmsn_bcrit_bcool}, we compare $\beta$ to
+$\beta_\mathrm{crit}$ across a range of disk radii for different heights, opacity values and wavenumbers.
+For optically thin cooling in the outer disk, curves for different wavenumbers overlap, as expected from Eq.\ \ref{beta_mmsn_simp}.
+Since the (optically thin) slope of $\beta$ is slightly steeper than for $\beta_\mathrm{crit}$, VSI growth can be suppressed at large radii (for the chosen opacity law).
+This effect is seen for $\hat{\kappa}_d = 1$ in the central panels of Fig.\ \ref{mmsn_bcrit_bcool}, where VSI is damped outside $\sim 150$ AU.
 
-We remark that it is possible to satisfy $\beta < \beta_\mathrm{crit}$
-for sufficiently large $\khat$ and small $r$. We illustrate this in 
-Fig. \ref{mmsn_bcrit_bcool_mink} by plotting at each radius the value
-of $\khat$ such that $\beta = \beta_\mathrm{crit}$. If we consider the
-midplane at, say $r_\mathrm{AU}\sim 1$, the VSI would only occur on very small
-radial lengthscales $\ll O(10^{-2}H)$. It is, however, unclear if such
-small-scale disturbances are dynamically important.
+We find that the most important factor for VSI growth is the opacity.
+With a smaller opacity, $\hat{\kappa}_d = 0.1$, growth is suppressed at all radii, as shown in the top panels of Fig.\ \ref{mmsn_bcrit_bcool}.
+Since optically thin cooling is too slow in this case, optically thick cooling -- above the floor set by optically thin cooling -- is also too slow.
 
-%In the disk
-% atmosphere where $\beta$ is small, surface modes may develop, 
-% but they are not well-defined in vertically isothermal disk models
-% with an imposed boundary, and they also require large $\khat$.
+Larger opacities make optically thin cooling much faster than $\beta_\mathrm{crit}$, as shown in the top panels of Fig.\ \ref{mmsn_bcrit_bcool}.
+However with larger opacities, optically thick cooling affects larger disk radii, slowing the cooling.
+Remarkably, the adopted MMSN value of opacity hits the a sweet spot where optically thin cooling is fast enough, yet slower optically thick can be restricted to inner disk radii.
+
+At smaller disk radii,  Fig.
+\ref{mmsn_bcrit_bcool} shows the hallmark of diffusive cooling, that larger wavenumbers can cool faster, but not below the floor set by optically thin cooling.
+Optically thick cooling times also rise sharply toward smaller radii (as $\beta \propto r^{-57/14}$) due to high densities and short orbital times.
+This effect suppresses VSI growth at small radii, but with a strong wavenumber dependence.
+
+%for $r_\mathrm{AU}\in[1,100]$, as well as for different
+%opacity scales. Notice that $\beta$ increases for either large or
+%small radii. For sufficiently large $r$, disturbances are generally
+%in the optically-thin cooling regime, characterized graphically by the
+%over-lapping curves. 
+%%, where $\beta$ is independent of
+%%$z$ and the perturbation lengthscale $\khat$.
+%In this case, we may
+%directly compare $\beta$ to $\beta_\mathrm{crit}$ and infer that the
+%VSI can operate at 10s of AU in the fiducial MMSN disk (with
+%$\hat{\kappa}_d=1$).  However, at small radii, $\beta$ depends on both
+%$z$ and $\khat$, as the disturbance enters the optically-thick limit
+%at the midplane. This regime is characterized by the distinct curves for each
+%$\khat$ toward decreasing $r$.  
+%  
+%%\emph{a better organized discussion might end on this point.  Might not be easy to rearrange to do this, but it would ideal to have the seamless flow to what you do next.}
+% 
+%Next consider the effect of the opacity scale $\hat{\kappa}_d$. The previous section
+%(see also Eq. \ref{beta_mmsn_simp}) indicates that increasing $\hat{\kappa}_d$
+%increases (decreases) the cooling time in the optically thick (thin)
+%limits. The overall effect of  increasing (decreasing) the opacity
+%scale by an order of magnitude is illustrated in the lower (upper)
+%panels of Fig. \ref{mmsn_bcrit_bcool}. 
+%
+%For increasing opacity, $\hat{\kappa}_d=1\to 10$,  a comparison
+%between the middle and lower panels shows that the optically-thick
+%regions shift outward; while the
+%optically-thin regions have lowered cooling times. Thus, instability
+%is expected to be restricted to larger radii. 
+%
+%For decreasing opacity, $\hat{\kappa}_d=1\to 0.1$,  the
+%optically-thick regions shift inwards; while cooling times in the
+%optically-thin limit are increased. In fact, the latter effect is
+%significant enough to raise $\beta > \beta_\mathrm{crit}$ over all
+%radii, and even away from the midplane. This is expected to stabilize
+%the disk against the VSI.
 
@@
 
+Fig. \ref{mmsn_bcrit_bcool_mink} highlights this wavenumber dependence by plotting
+the wavenumbers for which $\beta = \beta_\mathrm{crit}$.
+Above the solid curves (i.e.\ for larger wavenumbers), $\beta <  \beta_\mathrm{crit}$ at all disk heights, implying linear VSI growth.
+Below the dashed curves, VSI growth is strongly suppressed since $\beta >  \beta_\mathrm{crit}$ for all $|z| > 3H$.
+Between the solid and dashed curves some growth is possible, but only fairly close to the solid curve (according to \S\ref{vsi_mmsn_grow}).
+Thus linear growth of the VSI near 1 AU is only possible if $k_x \gtrsim 10^3/H$.
+As argued elsewhere, we doubt that such small-scale disturbances will drive significant turbulence or transport.
+
+We thus doubt that the VSI is significant at 1 AU in MMSN-like PPDs, for any opacity.
+At higher opacities, the required wavenumbers become shorter and even more problematic, as shown by the red curves in Fig.
+\ref{mmsn_bcrit_bcool_mink}.
+At lower opacties, even optically thin cooling (the fastest possible) is too slow, as discussed above.
+
+\emph{Big question, do we need to include disk mass as another parameter to vary.  Or not in this study.  Has been included in SK14.}
+
+%In the disk
+% atmosphere where $\beta$ is small, surface modes may develop, 
+% but they are not well-defined in vertically isothermal disk models
+% with an imposed boundary, and they also require large $\khat$.
+
 \subsection{VSI growth in the MMSN}\label{vsi_mmsn_grow}
-Here we solve the linear stability problem in the MMSN with the scale and
-height dependent thermal relaxation timescale given by
+We now consider the growth of the VSI in a MMSN disk model.  The linear analysis is similar to 
+\S\ref{numerical} but with different disk parameters and with the self-consistent cooling times of 
 Eq. \ref{beta_mmsn_simp}.
 
-In Fig. \ref{mmsn_overall} we plot characteristic growth timescales
-$t_\mathrm{grow} \equiv 1/\sigma$ of the fundamental mode with a range of
-wavenumbers, in units of the local orbital time $P_\mathrm{orb}\equiv
-2\pi/\OmK$. In fact, we either only find the fundamental mode, or 
-that it is the most unstable, except for the appearance of some 
-surface modes for large wavenumbers $\khat\gtrsim 50$. We discard
-surface modes for this discussion since they are not well-defined 
-in vertically isothermal disk models \citepalias{barker15} and may be 
-subject to viscous decay in real disks (see \S\ref{caveats_visc}).
+We consider the growth timescales of the fundamental mode for a range of
+wavenumbers and disk radii. We focus on the fundamental, i.e. lowest order vertical, mode 
+because it is the fastest growing mode except for some surface modes at high 
+wavenumbers.   We neglect these surface modes for reasons disucssed in \S\ref{numerical} and \S\ref{caveats_visc}.
 
@@
 
-We therefore consider Fig. \ref{mmsn_overall} as a representation of the
-VSI in the fiducial MMSN. We see the VSI operates from $r_\mathrm{AU}\sim 5$ to $r_\mathrm{AU}\sim
-50$ with growth timescales $\sim 30$---40 orbits and 
-$t_\mathrm{grow}$ rises rapidly towards the inner disk. It is
-necessary to consider smaller scales with decreasing $r$ as
-perturbations generally become optically thick.
+Fig. \ref{mmsn_overall} shows that in the MMSN, the VSI  is active from $r_\mathrm{AU}\sim 5$ to $r_\mathrm{AU}\sim
+50$ with growth timescales $\sim 30$---40 orbits.
+A small radius cufoff exists, inside of which growth is strongly suppressed.
+The cutoff occurs at smaller radii for larger wavenumbers, as expected from Fig.\ \ref{mmsn_bcrit_bcool_mink}.
+Growth at smaller disk radii is possible for yet larger wavenumbers, yet we remain skeptical about the non-linear viability of such short modes.
 
-The fact that we can obtain the VSI in the despite $\beta >
-\beta_\mathrm{crit}$ near the midplane (e.g. $\khat = 30$ at
-$r_\mathrm{AU} =5$, see Fig. \ref{beta_compare}) suggests  that the instability
-can operate provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
-fraction of the vertical domain. However, the normalized growth times
-become much longer. For example, for $\khat=30$ we find
-$t_\mathrm{grow}\sim 40$ orbits at $r_\mathrm{AU}=50$, but $t_\mathrm{grow}\sim
-400$ orbits at $r_\mathrm{AU}=5$ (and rapidly increases to $> 10^3$ orbits for
-smaller $r$). Thus, introducing an optically-thick midplane 
-significantly reduces the efficiency of the VSI despite $\beta <
-\beta_\mathrm{crit}$ away from the midplane.
+The growth times rise gradually as $r$ increases towards 100 AU in Fig.
+\ref{mmsn_overall}.
+This trend is expected as the outer radius cutoff at $\sim 200$AU (from Fig.\ \ref{mmsn_bcrit_bcool}) is approached.
 
-%surface modes may develop for large kx, they are not affected
+%
+%The fact that we can obtain the VSI in the despite $\beta >
+%\beta_\mathrm{crit}$ near the midplane (e.g. $\khat = 30$ at
+%$r_\mathrm{AU} =5$, see Fig. \ref{beta_compare}) suggests  that the instability
+%can operate provided $\beta < \beta_\mathrm{crit}$ for a sufficiently large
+%fraction of the vertical domain. However, the normalized growth times
+%become much longer. For example, for $\khat=30$ we find
+%$t_\mathrm{grow}\sim 40$ orbits at $r_\mathrm{AU}=50$, but $t_\mathrm{grow}\sim
+%400$ orbits at $r_\mathrm{AU}=5$ (and rapidly increases to $> 10^3$ orbits for
+%smaller $r$). Thus, introducing an optically-thick midplane 
+%significantly reduces the efficiency of the VSI despite $\beta <
+%\beta_\mathrm{crit}$ away from the midplane.
 
-We thus find Fig. \ref{mmsn_overall} to be roughly consistent with
-comparing the midplane thermal relaxation timescales to the critical
-thermal timescale. For example, 
-Fig. \ref{mmsn_bcrit_bcool} show that $\beta(z=0,\khat=10) \gtrsim
-\beta_\mathrm{crit}$ for $r_\mathrm{AU}\lesssim 10$ in the MMSN (left middle
-panel); and Fig. \ref{mmsn_overall} 
-show that modes with $\khat=10$ are stabilized for $r_\mathrm{AU}\lesssim 8$ (where
-$t_\mathrm{grow}\to\infty$). Our numerical results suggest the VSI
+%We thus find Fig. \ref{mmsn_overall} to be roughly consistent with
+%comparing the midplane thermal relaxation timescales to the critical
+%thermal timescale. For example, 
+%Fig. \ref{mmsn_bcrit_bcool} show that $\beta(z=0,\khat=10) \gtrsim
+%\beta_\mathrm{crit}$ for $r_\mathrm{AU}\lesssim 10$ in the MMSN (left middle
+%panel); and Fig. \ref{mmsn_overall} 
+%show that modes with $\khat=10$ are stabilized for $r_\mathrm{AU}\lesssim 8$ (where
+%$t_\mathrm{grow}\to\infty$).
+
+\emph{I don't understand the specific comparison being made here, especially since N13 is for vert.\ constant beta.
+Maybe remind me of which specific result/figure.
+I would think the more relevant comparison is to SK14 (either here or in discussion).}
+ Our numerical results suggest the VSI
 is efficient at radii of few tens of AU in the MMSN, consistent with
@@
   \includegraphics[width=\linewidth]{figures/eigen_compare_grow.ps}
-  \caption{Characteristic growth times of the VSI in 
-    the MMSN. These are calculated using the thermal relaxation
-    function given by Eq. \ref{beta_mmsn_simp}. 
+  \caption{VSI growth times ($t_{\rm grow} = 1/\sigma$) in orbital units ($P_\mathrm{orb} = 2\pi/\OmK$) for
+    the reference MMSN disk model with self-consistent dust cooling. 
     \label{mmsn_overall}}    

### 2015-05-06 application.tex

@@
 Thus in \S\ref{vsi_mmsn_grow} we analyze the linear growth of VSI with our PPD $\beta$ values.
+
+As an application of our linear models, we estimate where in a
+PPD do the thermodynamic conditions allow the VSI to
+operate. We first model the cooling time $t_c$ based on realistic
+disk properties, which yields a non-uniform dimensionless cooling
+time $\beta(r,z;\khat)$. We then compare this to the critical cooling time
+$\beta_\mathrm{crit}$ derived above. Finally, we explicitly compute 
+linear growth rates numerically, using $\beta(r,z;\khat)$, to
+characterize growth times expected in a typical PPD.
 

### 2015-05-06 numerical.tex

@@
 \subsection{Vertically non-isothermal disk model} 
-%We briefly consider vertically non-isothermal disks in which there is
-%a disk surface at finite height. 
-\emph{fix this rambling discussion..}
-\citetalias{barker15} have argued that vertically isothermal disks are not
-ideal for studing the VSI because $\zmax$ is a free 
-parameter in numerical calcluations. On the other hand, vertically 
-non-isothermal disks have a disk surface at finite height so there is
-a natural choice for $\zmax$. This means that higher order body modes and
-surface modes are better defined in a vertically non-isothermal disk.
+%\emph{fix this rambling discussion..}
 
-However, as we have seen in Fig. \ref{bcrit_compare1}, the maximum
-growth rate becomes insensitive to $\zmax$ for $\beta\gtrsim
-\beta_\mathrm{crit}$. This suggests that the characteristic upper
-limit on the thermal timescale for the VSI is not sensitive to
-$\zmax$, so that vertically isothermal disks are in fact suitable for
-the purpose of characterizing the thermodynamic dependence of the
-VSI.
+% \citetalias{barker15} have argued that vertically isothermal disks are not
+% ideal for studing the VSI because $\zmax$ is a free 
+% parameter in numerical calcluations. On the other hand, vertically 
+% non-isothermal disks have a disk surface at finite height so there is
+% a natural choice for $\zmax$. This means that higher order body modes and
+% surface modes are better defined in a vertically non-isothermal disk.
+
+% However, as we have seen in Fig. \ref{bcrit_compare1}, the maximum
+% growth rate becomes insensitive to $\zmax$ for $\beta\gtrsim
+% \beta_\mathrm{crit}$. This suggests that the characteristic upper
+% limit on the thermal timescale for the VSI is not sensitive to
+% $\zmax$, so that vertically isothermal disks are in fact suitable for
+% the purpose of characterizing the thermodynamic dependence of the
+% VSI.
 
@@
 
-Nevertheless, here we briefly consider a vertically non-isothermal
+Here we briefly consider a vertically non-isothermal
 disk model with $(p,q, h)=(0,-1,0.05)$ and $\Gamma=1.0837$. This
 value of $\Gamma$ is chosen so that the standard vertical domain size $\zmax =
-5H$ effectively corresponds to the disk surface ($\zmax\simeq
-H_s$). The disk temperature decreases rapidly with height such that 
-$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$.    \emph{note that this would imply midplane
-temperatures in excess of $10^6$ K so it's completely un-astrophyscial.}
+5H\simeq H_s$ effectively corresponds to the disk surface. The disk
+temperature decreases rapidly with height such that
+$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$. A disk surface at finite height
+means that higher order body modes and surface modes are better
+defined (cf. in vertically isothermal disk models where the disk
+boundary $\zmax$ is a free parameter, see \citetalias{barker15}).
+
+% \emph{note that this would
+% imply midplane 
+% temperatures in excess of $10^6$ K so it's completely un-astrophyscial.}
 

### 2015-05-06 numerical.tex

@@
 \section{Fixed cooling time results}\label{numerical}
-%comment: "numerical solutions' was too vague, especially since the next section contains numerical solutions.
-This section investigates the linear growth of the VSI with fixed, finite cooling times, $\beta$. 
-We solve the perturbation equations \ref{lin_all} with $\zeta=1$, i.e. including the background radial disk structure.  
-The disk's equilbrium  structure given in \S\ref{eqm}.  This general analysis does not make the approximations 
-described in \S\ref{sec:simplified}.
+This section investigates the linear growth of the VSI with fixed cooling times, $\beta/\OmK$, that are independent of height. 
+We solve the radially local model of Eqs.\ \ref{lin_all} with $\zeta=1$, i.e.\ retaining  radial gradients of the background disk structure.  
+The disk's equilbrium  structure given in \S\ref{eqm}.  This general analysis does not make the simplifying approximations 
+used in \S\ref{analytical}.
 
@@
 
-\subsubsection{Case study: $\khat=10$}
-In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues -- growth rates and
+\subsubsection{Case study: $\khat=10$}\label{k10}
+Fig. \ref{compare_modes_iso_kx10} maps eigenvalues -- growth rates and
 wave frequencies -- of unstable 
-modes for $\khat=10$.  We compare our numerical results to the analytic dispersion relations for 
-$\beta = 0$ and $10^{-3}$, from Eqs.\ \ref{simple_growth} and \ref{relax_disp}, respectively. Each symbol 
+modes for $\khat=10$.  Each symbol 
 corresponds to a different order mode, i.e.\ a different number of vertical node crossings.  Lower order modes
 have lower (in magnitude) frequencies for a given radial wavenumber, as expected for inertial-gravity waves (see \S\ref{stable_novshear}).
+
+ We compare our numerical results to the analytic dispersion relations for 
+$\beta = 0$ and $10^{-3}$, from Eqs.\ \ref{simple_growth} and \ref{relax_disp}, respectively.
+The analytic models also have discrete modes, which lie on the continuous curves that are plotted for clarity in Fig.
+\ref{compare_modes_iso_kx10}.
+While the analytic models have many simplifications, they crucially lack an artificial vertical surface (as they include an infinite vertical domain).
 
@@
 Fig. \ref{lowfreq_eigenfunc_2d} maps the pressure perturbation and meridional flow,
-scale by $\sqrt{\rho}$ to reflect the contribution to kinetic energy. Due to the stretched $x$ axis, 
+scaled by $\sqrt{\rho}$ to reflect the contribution to kinetic energy. Due to the stretched $x$ axis, 
 radial velocities are in fact typically much 
@@
 
-\subsubsection{Comment on `surface modes'}\label{surf_comment} 
-Fig. \ref{compare_modes_iso_kx10} lack the distinct `surface modes' found by 
-\citetalias{nelson13}, \citetalias{mcnally14}  and \citetalias{barker15}. 
-This a class of unstable modes associated with vertical shear near the
-disk boundaries. They are absent in our  
-analysis because there we assumed a disk with infinite 
-vertical extent, since the vertically isothermal disk has no
-surface. However, surface modes can occur in the present calculations 
-because there is an imposed vertical boundary.
+\subsubsection{Surface Modes of the VSI}\label{surf_comment} 
+The surface modes mentioned above are more prominent for larger wavenumbers.  
+Fig. \ref{compare_modes_iso_kx30} maps the $\khat = 30$
+eigenvalues, with the surface modes labelled.  The surface modes strongly depend 
+on the location of the imposed vertical boundary, and disagree with the analytic models, which lack an imposed surface. 
+We thus discount the physical significance of surface modes for our model, as we explain further below.
 
-We illustrate surface modes in Fig. \ref{compare_modes_iso_kx30} 
-which show unstable modes with $\khat=30$. Surface 
-modes lie almost parallel to the vertical axis 
-in such diagrams. They are readily identified for  $\zmax\geq 5H$ and
-are marked by `s'.  Fig. \ref{lowfreq_eigenfunc_surf} shows an example
-of a surface mode. We typically find an increasing
-number of surface modes with increasing $\zmax$ and/or $\khat$. 
-%consistent with \cite{barker15}.
+Surface modes are a well known feature of the VSI in finite vertical domains \citepalias{nelson13, mcnally14}.
+\citetalias{barker15} show that surface modes arise whether the surface is artificial
+(as for our imposed $\zmax$) or natural (for models with a zero density surface).  Physical surface modes
+are in principle possible; BL15 mention surfaces between a disk's interior and corona.  In models where the surface is imposed,
+like ours, surface modes are a numerical artifact.
 
-As explained by \citetalias{barker15}, while surface modes have a physical
-origin, their occurence in the vertically isothermal disk is due to
-imposing a finite vertical domain size. This has important
-implications when interpreting results obtained from the vertically
-isothermal disk model. For example, seeking the most unstable mode may
-result in a surface mode, but its growth rate would depend on
-$\zmax$. We return to this issue later.
+The possibility of artificial surface modes seeding growth in numerical simulations of the VSI  merits further study.
+Indeed N13 find that the initial growth of perturbations primarily occurs near the vertical boundaries.
+
+The motion in surface modes is indeed concentrated near the surface, as shown in Fig. \ref{lowfreq_eigenfunc_surf}.  
+Thus even if surface modes were physical, their contribution to transport might (by themselves) be weak,  following the arguments in
+\S\ref{k10}.  Moreover surface modes, like all modes with large wavenumbers, are more prone to viscous damping, as discussed in \S\ref{caveats_visc}.
+
+While the relatively large growth rates of surface modes is tantalizing, we dismiss them as ``fool's gold"  in models that lack a physical surface.
+
+%...
+%This a class of unstable modes associated with vertical shear near the
+%disk boundaries. They are absent in our  
+%analysis because there we assumed a disk with infinite 
+%vertical extent, since the vertically isothermal disk has no
+%surface. However, surface modes can occur in the present calculations 
+%because there is an imposed vertical boundary.   
+%
+%We illustrate surface modes in Fig. \ref{compare_modes_iso_kx30} 
+%which show unstable modes with $\khat=30$. Surface 
+%modes lie almost parallel to the vertical axis 
+%in such diagrams. They are readily identified for  $\zmax\geq 5H$ and
+%are marked by `s'.  Fig. \ref{lowfreq_eigenfunc_surf} shows an example
+%of a surface mode. We typically find an increasing
+%number of surface modes with increasing $\zmax$ and/or $\khat$. 
+%%consistent with \cite{barker15}. 
+% 
+%As explained by \citetalias{barker15}, while surface modes have a physical
+%origin, their occurence in the vertically isothermal disk is due to
+%imposing a finite vertical domain size. This has important
+%implications when interpreting results obtained from the vertically
+%isothermal disk model. For example, seeking the most unstable mode may
+%result in a surface mode, but its growth rate would depend on
+%$\zmax$. We return to this issue later.
 

### 2015-05-07 numerical.tex
6.2 edited.

@@
 
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvectorW_iso} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/eigenvectorvz_iso}
+  \caption{Pseudo-enthalpy perturbation $W$ (top) and vertical velocity
+    perturbation $\delta v_z$ of the fundamental VSI with 
+    $\khat=10$ in the fiducial disk with rapid thermal relaxation $\beta=10^{-3}$. The real  
+    (imaginary) parts of $W$ and $\delta v_z$ are plotted as solid 
+    (dotted) lines.  
+    \label{lowfreq_eigenfunc}
+  }
+\end{figure}
+
+\begin{figure}
+%  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 2.5cm
+%  0cm]{figures/result2d_vel}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
+%  0cm]{figures/result2d_mom} 
+  \includegraphics[width=\linewidth]{figures/result2d_iso}
+  \caption{Visualization of the fundamental VSI mode in
+    Fig. \ref{lowfreq_eigenfunc}. The color scale corresponds to the
+    pressure perturbation $\delta P=\rho W$ scaled by its maximum value.
+    Arrows show the flow, scaled as $\sqrt{\rho}(\delta
+    v_x,\delta v_z)$. The horizontal axis has been stretched 
+    for clarity.  
+    \label{lowfreq_eigenfunc_2d}
+  }
+\end{figure}
+
 \subsubsection{Case study: $\khat=10$}\label{k10}
@@
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvectorW_iso} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  1cm]{figures/eigenvectorvz_iso}
-  \caption{Pseudo-enthalpy perturbation $W$ (top) and vertical velocity
-    perturbation $\delta v_z$ of the fundamental VSI with 
-    $\khat=10$ in the fiducial disk with rapid thermal relaxation $\beta=10^{-3}$. The real  
-    (imaginary) parts of $W$ and $\delta v_z$ are plotted as solid 
-    (dotted) lines.  
-    \label{lowfreq_eigenfunc}
+  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_analytic.ps}
+  \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for
+    $\khat=30$. Examples of surface modes are 
+    marked by `s'. \label{compare_modes_iso_kx30}
   }
@@
 \begin{figure}
-%  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 2.5cm
-%  0cm]{figures/result2d_vel}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
-%  0cm]{figures/result2d_mom} 
-  \includegraphics[width=\linewidth]{figures/result2d_iso}
-  \caption{Visualization of the fundamental VSI mode in
-    Fig. \ref{lowfreq_eigenfunc}. The color scale corresponds to the
-    pressure perturbation $\delta P=\rho W$ scaled by its maximum value.
-    Arrows show the flow, scaled as $\sqrt{\rho}(\delta
-    v_x,\delta v_z)$. The horizontal axis has been stretched 
-    for clarity.  
-    \label{lowfreq_eigenfunc_2d}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  0cm]{figures/eigenvectorvz_iso_surf}
+  \caption{Example of a `surface mode' in the fiducial disk model.
+    \label{lowfreq_eigenfunc_surf}
   }
@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_analytic.ps}
-  \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for
-    $\khat=30$. Examples of surface modes are 
-    marked by `s'. \label{compare_modes_iso_kx30}
-  }
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/eigenvectorvz_iso_surf}
-  \caption{Example of a `surface mode' in the fiducial disk model.
-    \label{lowfreq_eigenfunc_surf}
-  }
-\end{figure}
-
-\subsection{Effect of thermal relaxation}\label{therm_relax_eff}
-We now consider the effect of finite thermal relaxation by increasing 
-$\beta$. % Note that for the fiducial disk model the critical thermal 
-% timescale $\beta_\mathrm{crit} = 0.125$.
-Fig. \ref{compare_modes_cool_kx10} show unstable modes with $\khat=10$ 
-and $\beta$ from $0.01$ to  $0.1$. In addition to our fiducial setup with 
-$\zmax=5H$, we also plot results for $\zmax=3H$, $\zmax=7H$ and 
-analytic results computed from Eq. \ref{relax_disp}.  %analytic
-% results for low order modes
-
-We find the number of unstable modes decrease with increasing
-$\beta$. Consider the fiducial setup with $\zmax=5H$ 
-(middle panel). The fundamental mode ($M=0$) is not the most unstable for   
-$\beta=0.01$ as higher order modes have larger
-growth rates. As noted above, these higher order modes are affected by
-$\zmax$, e.g. there is a significant mismatch between numerical and 
-analytical results for $M\geq1$ when $\zmax=3H$ (upper panel).
-
-%However, by comparing with results for $\zmax=3H$, we see that
-
-However, the fundamental mode becomes dominant for $\beta \geq 
-0.03$. Higher order modes, which would be dominant at small $\beta$, 
-are more effectively stabilized with increasing $\beta$. For 
-$\beta=0.1$, which is close to the critical thermal timescale
-$\beta_\mathrm{crit} (=0.125$ for the fiducial disk
-parameters), we only find the fundamental mode. This is consistent with
-the discussion in \S\ref{iso_vsi_beta_crit}.
-
-\begin{figure}
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
@@
 
-Fig. \ref{lowfreq_eigenfunc_cool}---\ref{lowfreq_eigenfunc_2d_cool}
-shows the fundamental mode with $\khat=10$ in a disk with 
-$\beta=0.1$. Comparing these plots with
-Fig. \ref{lowfreq_eigenfunc}---\ref{lowfreq_eigenfunc_2d}, which have
-$\beta=10^{-3}$, we see that increasing $\beta$ restricts the region 
-characterized by $W\sim z$ and $\delta v_z\sim$ constant closer to the
-midplane, with oscillatory behavior emerging near vertical boundaries
-where buoyancy first becomes important. This leads to the `tilted'
-pressure perturbations seen in Fig. \ref{lowfreq_eigenfunc_2d_cool} 
-(cf. Fig. \ref{lowfreq_eigenfunc_2d}).
+\subsection{Slower thermal relaxation}\label{therm_relax_eff}
+We now consider the effect of longer cooling timescales by gradually increasing 
+$\beta$. We expect VSI growth for  $\beta < \beta_\mathrm{crit} = 0.125$.  We also expect that higher order modes will damp at even lower $\beta$ values, as argued in \S\ref{iso_vsi_beta_crit}.
+
+Fig. \ref{compare_modes_cool_kx10} largely confirms these expectations by plotting eigenvalues for $\khat=10$ 
+and $\beta$ from $0.01$ to  $0.1$. 
+The analytic results from Eq. \ref{relax_disp} are now plotted as different curves for each mode order $M$, with $\beta$ varying along each curve.  
+We see the standard increase in frequency, $|\omega|$, with mode order.
+
+As expected, the higher order modes are preferentially damped.
+For $\beta \geq 0.03$ the fundamental ($M = 0$) mode is the fastest growing.
+For $\beta = 0.1$, only the $M = 0$ mode grows.
+This growth is slow since $\beta$ is near $\beta_\mathrm{crit}$.
+
+Fig.
+\ref{compare_modes_cool_kx10} also shows how cooling times affect the dependence on $\zmax$, the size of the vertical domain.
+For all cooling times, the fundamental mode depends only weakly on $\zmax$, and there is good agreement with analytic values.
+This convergence is reassuring given the importance of the fundamental mode at longer cooling times.
+For higher order modes and short cooling times, however, the eigenvalues vary strongly with $\zmax$.
+The disagreement with analytic theory is strongest for the smallest domain, with $\zmax = 3 H$.
+For $\beta \geq 0.03$, the $\zmax/H = 5, 7$ and the analytic results are well converged (aided by the fact that only $M < 5$ modes grow).
+
+%We find the number of unstable modes decrease with increasing
+%$\beta$. Consider the fiducial setup with $\zmax=5H$ 
+%(middle panel). The fundamental mode ($M=0$) is not the most unstable for   
+%$\beta=0.01$ as higher order modes have larger
+%growth rates. As noted above, these higher order modes are affected by
+%$\zmax$, e.g. there is a significant mismatch between numerical and 
+%analytical results for $M\geq1$ when $\zmax=3H$ (upper panel). 
+%
+%However, the fundamental mode becomes dominant for $\beta \geq 
+%0.03$. Higher order modes, which would be dominant at small $\beta$, 
+%are more effectively stabilized with increasing $\beta$. For 
+%$\beta=0.1$, which is close to the critical thermal timescale
+%$\beta_\mathrm{crit} (=0.125$ for the fiducial disk
+%parameters), we only find the fundamental mode. This is consistent with
+%the discussion in \S\ref{iso_vsi_beta_crit}.
 
@@
 
-%presence of buoyancy forces at the boundary may make BC more
-%influential
-
-In Fig. \ref{compare_modes_cool_kx30} we plot eigenvalues for unstable 
-modes with $\khat=30$. For clarity, we only plot analytic results 
-for the fundamental mode ($M=0$) and $M=1,\,2$, and omit the 
-squares. The fundamental mode is again insensitive to $\zmax$ and are
-in close agreement with analytic results. However, 
-there is increasing mismatch between analytical and numerical results
-for higher order modes as $\beta$ is increased. This suggests that, in
-the presence of buoyancy, boundary effects become important for these
-modes, which have larger $\khat$ than that considered previously. For
-example, for $\zmax=5H$ and $\zmax=7H$, we 
-find modes slightly more unstable than the fundamental
-mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
-panels), but these are clearly affected by boundary
-conditions as they are absent for $\zmax=3H$.
-
-Nevertheless, as before we observe more effective stabilization for
-higher order modes, including the absence of surface modes, for
-example as $\beta$ is increased from $0.01$ to $0.03$ (for $\zmax\geq
-5H$). 
-%\emph{not a clear and possible not valid argument.  All modes are associated with
-%a BC and the disappearance of a mode for smaller zmax could be the artificial effect.}
-%If we discard modes associated with boundary conditions, then
-%the fundamental mode would again become the dominant mode as $\beta$ is
-%increased.
+Figures \ref{lowfreq_eigenfunc_cool} and \ref{lowfreq_eigenfunc_2d_cool}
+illustrate the eigenfunction of the fundamental mode with $\khat=10$ and
+$\beta=0.1$. Compared with the small $\beta$ case in 
+Figs. \ref{lowfreq_eigenfunc} and \ref{lowfreq_eigenfunc_2d}, the eigenfuctions 
+show a more complex variation of phase with height.  This dependence yields 
+the `tilted' appearance of pressure field in  Fig. \ref{lowfreq_eigenfunc_2d_cool}.
+Physically, the increased role of buoyancy, which increases with height, explains 
+why larger $\beta$ values produce more complex vertical structure.
 
@@
 
-% Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
-% show that the fundamental mode is not sensitive to boundary
-% conditions. The fundamental mode also becomes dominant as $\beta$
-% approaches $\beta_\mathrm{crit}$, except with the possible emergence
-% of modes associated with boundary conditions. Thus, tracking the
-% fundamental mode is a more preferable assessment of stability as $\beta$
-% is increased from zero, rather than searching for the most unstable
-% mode. This is because the latter may correspond to modes that depend on boundary
-% conditions, e.g. surface modes whose growth rate varies with
-% $\zmax$ in a vertically isothermal disk.
+Fig. \ref{compare_modes_cool_kx30} shows how cooling affects higher wavenumber
+modes, specifically $\khat=30$.   We focus on the $\zmax/H = 5, 7$ and analytic results for $\beta \geq 0.03$. 
+These cases  show good convergence, though not quite as good as for $\khat = 10$.
 
-%boundary conditions play an increasing role for higher M
+With slower cooling, the wave frequency shows a more complex 
+dependence on mode order, both analytically and numerically. For $\beta = 0.1$,
+ the fundamental mode (which lies on the $M = 0$ curve) no longer has the smallest $|\omega|$ value.
+
+Moreover, the fundamental mode is no longer the fastest growing mode for $\beta \geq 0.03$, or even 
+for $\beta = 0.1$.  This complication is not actually surprising since $|hq\khat| = 1.5$ is no longer less than unity,
+as required in the analytic derivation of \S\ref{iso_vsi_beta_crit}.  Though  inconvenient,  given that the 
+fundemental mode is easy to identify and the most numerically converged, this complication is not ultimately significant
+for the operability of the VSI.
+
+Fig. \ref{compare_modes_cool_kx30} also demonstrates that artificial surface modes are damped for $\beta \geq 0.03$.
+
+%For clarity, we only plot analytic results 
+%for the fundamental mode ($M=0$) and $M=1,\,2$, and omit the 
+%squares. The fundamental mode is again insensitive to $\zmax$ and are
+%in close agreement with analytic results. However, 
+%there is increasing mismatch between analytical and numerical results
+%for higher order modes as $\beta$ is increased. This suggests that, in
+%the presence of buoyancy, boundary effects become important for these
+%modes, which have larger $\khat$ than that considered previously. For
+%example, for $\zmax=5H$ and $\zmax=7H$, we 
+%find modes slightly more unstable than the fundamental
+%mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
+%panels), but these are clearly affected by boundary
+%conditions as they are absent for $\zmax=3H$. 
+%
+%Nevertheless, as before we observe more effective stabilization for
+%higher order modes, including the absence of surface modes, for
+%example as $\beta$ is increased from $0.01$ to $0.03$ (for $\zmax\geq
+%5H$).
 
@@
   timescale}\label{bcrit_num_test}
+Having explored the behavior of VSI modes with cooling, we turn to the numerical vailidity of the 
+analytic cooling criterion, $\beta < \beta_\mathrm{crit} = h|q|/(\gamma -1)$.
+
+...
+
 In Fig. \ref{bcrit_compare1} we plot the maximum VSI growth rates

### 2015-05-07 appendix.tex
mkl edit of sec 5.3, 6.0, 6.1, partial 7

@@
 \subsection{Maximum critical cooling time}\label{max_cool}   
- % {\bf [at the moment for reference
- %  only, not to be included in final version]}\\
+Here we show that, under appropriate conditions, the maximum critical 
+cooling time is that for the $M=0$ or fundamental mode. This allows
+us to focus on the fundamental mode to obtain an overall cooling
+requirement for the VSI.
+
 Consider the simplified dispersion relations for marginal stability,
-Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. Writing 
-$X= \hat{\omega}\khat$, $\theta = (hq\khat)^2\leq1$ and  treating $M$ as a
-continuous variable, we find from Eq. \ref{relax_cond_simp1} that 
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. We write 
+$X= \hat{\omega}\khat$, $\theta = (hq\khat)^2$ and  
+treat $M$ as a continuous variable. We find from
+Eq. \ref{relax_cond_simp1} that  
 %\begin{align}
@@
 Hence,
-\begin{align}
+\begin{align}\label{dbeta_dm}
   \frac{d\beta_c}{dM}<0, 
 \end{align}
-for all $M\geq 0$ if $\theta\leq 1$. This implies 
-$\mathrm{max}(\beta_c)$ occurs at $M=0$.
+for all $M\geq 0$ if $\theta\leq 1$, which imply 
+$\mathrm{max}(\beta_c)$ occurs at $M=0$. For fixed $\khat$ the
+condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
+This means that if $\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$, then
+all such modes will be stabilized.
+
+This result highlights the importance of the fundamental mode  
+--- it is the most difficult mode to stabilize with finite
+cooling. Furthermore, for $M=0$ we may obtain the expression  
+for $\beta_\mathrm{crit}$ from the dispersion relations 
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} without  
+assuming it is much less than unity from the onset or place
+restrictions on $\theta$.
+
+% restrictions on $| h q \khat|$.
 

### 2015-05-07 application.tex
mkl edit of sec 5.3, 6.0, 6.1, partial 7

--- 2015-05-06 application.tex
@@
 \section{The VSI with realistic  protoplanetary disk cooling}\label{application} 
-To this point we have treated the cooling time, $\beta$, as a free parameter that, moreover, is independent of height. 
-To understand the applicability of the VSI to PPDs, we now consider $\beta$ values that are consistent with PPD models.
-Section \ref{cooling_model}  develops our PPD cooling law, which depends on disk radius, disk height and perturbation wavelength.
-We compare these PPD $\beta$ values to $\beta_{\rm crit}$ in \S\ref{bcritPPD}.
-While this comparison is instructive, it is not complete since $\beta_{\rm crit}$ was derived assuming a vertically constant $\beta$.
-Thus in \S\ref{vsi_mmsn_grow} we analyze the linear growth of VSI with our PPD $\beta$ values.
+To this point we have treated the cooling time, $\beta$, as a free
+parameter that, moreover, is independent of height.  
+To understand the applicability of the VSI to PPDs, we now consider
+$\beta$ values that are consistent with PPD models. 
+Section \ref{cooling_model}  develops our PPD cooling law, which
+depends on disk radius, disk height and perturbation wavelength. 
+We compare these PPD $\beta$ values to $\beta_{\rm crit}$ in
+\S\ref{bcritPPD}.  While this comparison is instructive, it is not
+complete since $\beta_{\rm crit}$ was derived assuming a vertically
+constant $\beta$.  Thus in \S\ref{vsi_mmsn_grow} we analyze the linear
+growth of VSI with our PPD $\beta$ values.
+
+% As an application of our linear models, we estimate where in a
+% PPD do the thermodynamic conditions allow the VSI to
+% operate. We first model the cooling time $t_c$ based on realistic
+% disk properties, which yields a non-uniform dimensionless cooling
+% time $\beta(r,z;\khat)$. We then compare this to the critical cooling time
+% $\beta_\mathrm{crit}$ derived above. Finally, we explicitly compute 
+% linear growth rates numerically, using $\beta(r,z;\khat)$, to
+% characterize growth times expected in a typical PPD.
 
@@
 Since the VSI is characterized by vertically-elongated,
-radially-narrow disturbances, we retain only the radial derivatives in Eq. \ref{diff_cool_proper}.
+radially-narrow disturbances, we retain only the radial derivatives of
+the perturbations in Eq. \ref{diff_cool_proper}. 
 Thus, in the radially local approximation, we have
@@
   %  T\right) &
-  \delta\Lambda_\mathrm{diff} \approx %-\frac{k_\mathrm{rad}}{\rho
+  \delta\Lambda_\mathrm{diff} \simeq %-\frac{k_\mathrm{rad}}{\rho
   % C_v}k_x^2P\left(\frac{\delta T}{T}\right)%\notag\\
@@
 and $\sigma_s$ is the Stefan-Boltzmann constant. 
-
 The thermal relaxation  timescale (defined in Eq. \ref{thermal_relax}) 
@@
 In the optically thin regime, $l\ll 
-l_\mathrm{rad}$, thermal relaxation operates by ``Newtonian cooling."
+l_\mathrm{rad}$, thermal relaxation operates by `Newtonian cooling'.
 The cooling time is independent of $l$ and 
@@
 \end{align}
-and $t_\mathrm{thin}$ does not depend on $\rho$ (because $\kappa_d$ depends on $T$ only).
+and $t_\mathrm{thin}$ does not depend on $\rho$ (because our adopted
+$\kappa_d$ depends on $T$ only, see below).
 
@@
 \subsubsection{PPD cooling times}\label{toy_relax}
-In a vertically isothermal disk with surface density $\Sigma = \rho_0/\sqrt{2\pi}H$, 
+In a vertically isothermal disk with surface density $\Sigma = \rho_0\sqrt{2\pi}H$, 
 the dimensionless thermal
@@
 \end{align}
-where $\hat{\rho} = \rho/\rho_0$.  The first and second terms in square brackets represent the optically thin and thick cooling regimes, respectively.
+where $\hat{\rho} = \rho/\rho_0$.  The first and second terms in
+square brackets represent the optically thin and thick cooling
+regimes, respectively.
 
@@
 (MMSN) disk model of \cite{chiang10} which specifies
-\emph{note that the normalizations are suppressed b/c I can't see where we used them.  We should only add back in if we decide to use them.}
+% \emph{note that the normalizations are suppressed b/c I can't see where we used them.  We should only add back in if we decide to use them.}
 \begin{subequations}
@@
 \end{align}\end{subequations}
-with $r_\mathrm{AU}\equiv r/\mathrm{AU}$.
-This model has $q = -3/7$, $p = -39/14$ and $h  = 0.034 r_\mathrm{AU}^{2/7}$, using $\mu = 2.33$ and the above relation between $\Sigma$ and $\rho_0 \propto r^p$.
+with $r_\mathrm{AU}\equiv r/\mathrm{AU}$. This model has $q = -3/7$,
+$p = -39/14$ and $h  = 0.022 r_\mathrm{AU}^{2/7}$, using $\mu = 2.33$
+and the above relation between $\Sigma$ and $\rho_0 \propto r^p$.
 
-For the dust opacity we adopt the \citet{bell94} law with $\kappa_d \propto T^2$, giving
+For the dust opacity we adopt the \citet{bell94} law with $\kappa_d
+\propto T^2$, giving 
 \begin{align}
@@
 \end{align}
-where the opacity normalication factor, $\hat{\kappa}_d$, scales with the ratio of small dust to gas;  $\hat{\kappa}_d = 1$ is our fiducial case.
+where the opacity normalization factor, $\hat{\kappa}_d$, scales with
+the ratio of small dust to gas;  $\hat{\kappa}_d = 1$ is our fiducial
+case.
 
@@
 
-This comparison strictly valid only for optically thin cooling which is independent of height,
-as assumed in the derivation of $\beta_\mathrm{crit}$.  This condition typically holds in the outer disk.  
-Fig. \ref{beta_compare} shows that a  $\khat=30$ perturbation at 50 AU has vertically constant $\beta$. 
- Furthermore, since $\beta < \beta_\mathrm{crit}$, the mode is unstable.
+This comparison is strictly valid only for optically thin cooling
+which is independent of height, as assumed in the derivation of
+$\beta_\mathrm{crit}$.  This condition typically holds in the outer
+disk.   Fig. \ref{beta_compare} shows that a  $\khat=30$ perturbation
+at 50AU has vertically constant $\beta$. Furthermore, since $\beta <
+\beta_\mathrm{crit}$, the disturbance would be unstable.
 
-For optically thick cooling, $\beta$ varies with height, complicating the comparison with $\beta_\mathrm{crit}$.   
-At 5 AU, Fig. \ref{beta_compare} shows that cooling times are long in the midplane with $\beta > \beta_\mathrm{crit}$.
-However, since $\beta < \beta_\mathrm{crit}$ away from the midplane, we require the analysis of \S\ref{vsi_mmsn_grow} to determine if the VSI can grow.
-(That analysis will show that growth is stongly suppressed for this case.)  We proceed with the awareness that optically thick regions in the inner disk are a complication, but that we can reasonably expect VSI growth if $\beta < \beta_\mathrm{crit}$ at all heights.
+For optically thick cooling, $\beta$ varies with height, complicating
+the comparison with $\beta_\mathrm{crit}$.    
+At 5AU, Fig. \ref{beta_compare} shows that cooling times are long in
+the midplane with $\beta > \beta_\mathrm{crit}$. 
+However, since $\beta < \beta_\mathrm{crit}$ away from the midplane,
+we require the analysis of \S\ref{vsi_mmsn_grow} to determine if the
+VSI can grow.  (That calculation will show that growth is stongly
+suppressed for this case.)  We proceed with the awareness that
+optically thick regions in the inner disk are a complication, but that
+we can reasonably expect VSI growth if $\beta < \beta_\mathrm{crit}$
+at all heights.
 
@@
 In Fig. \ref{mmsn_bcrit_bcool}, we compare $\beta$ to
-$\beta_\mathrm{crit}$ across a range of disk radii for different heights, opacity values and wavenumbers.
-For optically thin cooling in the outer disk, curves for different wavenumbers overlap, as expected from Eq.\ \ref{beta_mmsn_simp}.
-Since the (optically thin) slope of $\beta$ is slightly steeper than for $\beta_\mathrm{crit}$, VSI growth can be suppressed at large radii (for the chosen opacity law).
-This effect is seen for $\hat{\kappa}_d = 1$ in the central panels of Fig.\ \ref{mmsn_bcrit_bcool}, where VSI is damped outside $\sim 150$ AU.
+$\beta_\mathrm{crit}$ across a range of disk radii for different
+heights, opacity values and wavenumbers. 
+For optically thin cooling in the outer disk, curves for different
+wavenumbers overlap, as expected from Eq.\ \ref{beta_mmsn_simp}. 
+Since the (optically thin) slope of $\beta$ is slightly steeper than
+for $\beta_\mathrm{crit}$, VSI growth can be suppressed at large radii
+(for the chosen opacity law).  This effect is seen for $\hat{\kappa}_d
+= 1$ in the central panels of Fig.\ \ref{mmsn_bcrit_bcool}, where VSI
+is damped outside $\sim 150$ AU.
 
 We find that the most important factor for VSI growth is the opacity.
-With a smaller opacity, $\hat{\kappa}_d = 0.1$, growth is suppressed at all radii, as shown in the top panels of Fig.\ \ref{mmsn_bcrit_bcool}.
-Since optically thin cooling is too slow in this case, optically thick cooling -- above the floor set by optically thin cooling -- is also too slow.
+With a smaller opacity, $\hat{\kappa}_d = 0.1$, growth is suppressed
+at all radii, as shown in the top panels of Fig.\
+\ref{mmsn_bcrit_bcool}.  Since optically thin cooling is too slow in
+this case, optically thick cooling --- above the floor set by optically
+thin cooling --- is also too slow.
 

### 2015-05-07 linear_adia.tex
re-writing sec 7 in progress
mkl edit of sec 5.3, 6.0, 6.1, partial 7

@@
 \end{align}\end{subequations}
-These equations consider $\khat^2\gg 1$, $M\lesssim O(1)$; and assume 
-$|\hat{\omega}\khat| = O(1)$ and $\beta_c\ll 1$.   Note that
+These equations consider $\khat^2\gg 1$ and %, $M\lesssim O(1)$; 
+assumes $|\hat{\omega}\khat| $ is $O(1)$ and $\beta_c\ll 1$.   Note that 
 Eq. \ref{relax_cond_simp1} reduces to the  
@@
 vertical shear (see Appendix \ref{stable_novshear} for details). 
+%marginal stability requires hqk \leq 1
 
-In Appendix \ref{max_cool}, we find that %if $|hq\khat|^2\leq 1$ then 
-\begin{align}
-  |hq\khat|^2 \leq 1 \Rightarrow \frac{d\beta_c}{dM} < 0 \quad \forall
-  M\geq 0. \label{bcrit_gen}
+We are most interested in the longest cooling time which allows growth
+for any $M$. In Appendix \ref{max_cool}, we find that if
+$|hq\khat|\leq 1$, then $\beta_c$ decreases with increasing $M$.   
+%\begin{align}
+%  |hq\khat|^2 \leq 1 \Rightarrow \frac{d\beta_c}{dM} < 0 \quad \forall
+%  M\geq 0. \label{bcrit_gen}
+%\end{align}
+%Eq. \ref{bcrit_gen} indicates that higher $M$ modes have lower $\beta_c$,
+%i.e.\ require faster cooling for VSI growth. This trend is confirmed
+%numerically in \S\ref{therm_relax_eff}.  
+The VSI criterion is then set by $\beta_c$ for the
+fundamental mode ($M = 0$), for which the exact solution to 
+Eq. \ref{relax_cond_simp1} --- \ref{relax_cond_simp2} is
+$\hat{\omega}_c\khat = \beta_c(\gamma-1)/hq = -1$. We thus find the
+cooling criterion for VSI growth is    
+ \begin{align}\label{iso_vsi_cond}
+   \beta <   \beta_\mathrm{crit}  \equiv
+   \frac{h|q|}{\gamma-1} . 
  \end{align}
-Eq. \ref{bcrit_gen} indicates that higher $M$ modes have lower $\beta_c$,
-i.e.\ require faster cooling for VSI growth. This trend is confirmed
-numerically in \S\ref{therm_relax_eff}.
+ The thin disk approximation, $h \ll 1$, indicates that $\beta_\mathrm{crit} \ll
+ 1$, as assumed to obtain Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. 
+ We heuristically explain $\beta_\mathrm{crit}$ in \S\ref{vsi_require}, and numerically
+ test its validity in \S\ref{bcrit_num_test}.
+
+% This is clarified in Appendix
+% \ref{bcrit_alt}.
+
+% For $M=0$, it is straight foward to solve Eq. \ref{relax_cond_simp1} 
+% --- \ref{relax_cond_simp2} exactly and obtain the maximum cooling time
+% for the fundamental mode, $\beta_\mathrm{crit}$
+% (Eq. \ref{iso_vsi_cond}), which shows that $\beta_\mathrm{crit}\ll 1$
+% because we consider thin disks.
 
@@
 %on $|h q \hat{k}|$ is dropped).
-
-We are most interested in the longest cooling time which allows growth for any $M$.  
-This criterion is clearly set by the fundamental mode, $M = 0$. It is straight
-forward to solve Eq. \ref{relax_cond_simp1} --- \ref{relax_cond_simp2}
-for $\beta_c(M=0)  \equiv \beta_\mathrm{crit}$ in this special case. We thus find the cooling criterion for  
-VSI growth is  
- \begin{align}\label{iso_vsi_cond}
-   \beta <   \beta_\mathrm{crit}  \equiv
-   \frac{h|q|}{\gamma-1} . 
- \end{align}
- The thin disk approximation, $h \ll 1$, indicates that $\beta_\mathrm{crit} \ll
- 1$, as assumed to obtain Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. 
- We heuristically explain $\beta_\mathrm{crit}$ in \S\ref{vsi_require}, and numerically
- test its validity in \S\ref{bcrit_num_test}.
-
-We comment that the expression for $\beta_\mathrm{crit}$ may in fact 
-be obtained without assuming it is much less than unity from the onset or 
-restrictions on $| h q \khat|$.  This is clarified in Appendix
-\ref{bcrit_alt}.
-
-% For $M=0$, it is straight foward to solve Eq. \ref{relax_cond_simp1} 
-% --- \ref{relax_cond_simp2} exactly and obtain the maximum cooling time
-% for the fundamental mode, $\beta_\mathrm{crit}$
-% (Eq. \ref{iso_vsi_cond}), which shows that $\beta_\mathrm{crit}\ll 1$
-% because we consider thin disks.

### 2015-05-07 numerical.tex
mkl edit of sec 5.3, 6.0, 6.1, partial 7

--- 2015-05-06 numerical.tex
@@
-\section{Fixed cooling time results}\label{numerical}
-This section investigates the linear growth of the VSI with fixed cooling times, $\beta/\OmK$, that are independent of height. 
-We solve the radially local model of Eqs.\ \ref{lin_all} with $\zeta=1$, i.e.\ retaining  radial gradients of the background disk structure.  
-The disk's equilbrium  structure given in \S\ref{eqm}.  This general analysis does not make the simplifying approximations 
+\section{Numerical results for fixed cooling times}\label{numerical}
+This section investigates the linear growth of the VSI with fixed
+cooling times, $\beta\OmK^{-1}$, that are independent of height.  
+We solve the radially local model of Eqs.\ \ref{lin_all} with
+$\zeta=1$, i.e.\ retaining  radial gradients of the background disk
+structure. The equilbrium disk structure given is in \S\ref{eqm}. This
+numerical study does not make the simplifying approximations  
 used in \S\ref{analytical}.
@@
 converts the linear system of differential equations to a set of 
-algebraic equations,  which we solve using LAPACK matrix routines.\footnote{provide a reference of some sort to LAPACK}
+algebraic equations,  which we solve using LAPACK matrix
+routines\footnote{Available at \url{http://www.netlib.org/lapack/}.}.
 
@@
 that $\Omega$ decreases slightly away from the midplane.   From Eq.\ \ref{iso_vsi_cond},
-we expect that $\beta < \beta_\mathrm{crit} \approx 0.13$ is needed for VSI growth in our fiducial model.
+we expect that $\beta < \beta_\mathrm{crit} = 0.125$ is needed for VSI growth in our fiducial model.
 
@@
 gives similar results to the well studied case of isothermal perturbations
- \citepalias[e.g.][]{nelson13,mcnally14,barker15}.\footnote{To be sure, we confirmed that smaller
+ \citepalias[e.g.][]{nelson13,mcnally14,barker15}.\footnote{We confirmed that smaller
  $\beta$ values gave similar results.} This case allows us to explore and test our
@@
     % $\beta=10^{-3}$. 
-We investigate the effect of vertical box size, from $\zmax=3H$ (\emph{blue triangles}) to $7H$ (\emph{green x's})
+We investigate the effect of vertical box size, from $\zmax=3H$
+(\emph{blue triangles}) to $7H$ (\emph{green} x's)
     with a free surface boundary ($\Delta P=0$).  For $\zmax=5H$  (black dots) we also compare free surface (\emph{black dots}) and rigid (\emph{red plusses}), finding little difference.
@@
 \subsubsection{Case study: $\khat=10$}\label{k10}
-Fig. \ref{compare_modes_iso_kx10} maps eigenvalues -- growth rates and
-wave frequencies -- of unstable 
+Fig. \ref{compare_modes_iso_kx10} maps the growth rates $\sigma$ and
+wave frequencies $\omega$ of unstable 
 modes for $\khat=10$.  Each symbol 
 corresponds to a different order mode, i.e.\ a different number of vertical node crossings.  Lower order modes
-have lower (in magnitude) frequencies for a given radial wavenumber, as expected for inertial-gravity waves (see \S\ref{stable_novshear}).
+have smaller $|\omega|$ for fixed radial wavenumber, as expected for
+inertial-gravity waves (see \S\ref{stable_novshear}).
 
 We compare our numerical results to the analytic dispersion relations for 
-$\beta = 0$ and $10^{-3}$, from Eqs.\ \ref{simple_growth} and \ref{relax_disp}, respectively.
-The analytic models also have discrete modes, which lie on the continuous curves that are plotted for clarity in Fig.
-\ref{compare_modes_iso_kx10}.
-While the analytic models have many simplifications, they crucially lack an artificial vertical surface (as they include an infinite vertical domain).
+$\beta = 0$ and $10^{-3}$, from Eqs.\ \ref{simple_growth} and
+\ref{relax_disp}, respectively.   The analytic models also have
+discrete modes, which lie on the continuous curves that are plotted
+for clarity in Fig. \ref{compare_modes_iso_kx10}.   While the analytic
+models have many simplifications, they crucially lack an artificial
+vertical surface (as they assume an infinite vertical domain).
 
-For our fiducial case (black dots), the lower order and lower frequency modes closely follow the analytic prediction with growth rates increasing with $|\omega|$, the magnitude of the frequency.
-However, for $|\omega| \gtrsim 4 h \OmK$ the growth rate declines with $|\omega|$.
-This break from the analytic prediction is not due the choice of boundary condition, as we demonstrate by considering rigid boundaries (red plusses).
+For our fiducial case (black dots), the lower order and lower
+frequency modes closely follow the analytic prediction with growth
+rates increasing with $|\omega|$. %, the magnitude of the frequency.
+However, for $|\omega| \gtrsim 4 h \OmK$ the growth rate declines with
+$|\omega|$.  This break from the analytic prediction is not due the
+choice of boundary condition, as we demonstrate by considering rigid
+boundaries (red plusses).
 
-Rather, the decline in growth rate for high order models is due to the vertical box size, as seen by comparing the $\zmax/H = 3, 5$ and 7 cases in Fig.
-\ref{compare_modes_iso_kx10}.
-Larger boxes give better agreement with the analytic theory, which has infinite vertical extent.
-Larger boxes include regions of larger vertical shear (see Eqs.
-\ref{vshear_thin} and \ref{vertical_shear_ex}), which higher order modes can tap to give larger growth rates.
-For the $\zmax = 7 H$ case, we begin to see another branch of modes with the highest growth rates at $\omega \sim -6 h\OmK$.
-This branch contains the  ``surface modes" described below.
+Rather, the decline in growth rate for high order models is due to the
+vertical box size, as seen by comparing the $\zmax/H = 3, 5$ and 7
+cases in Fig. \ref{compare_modes_iso_kx10}. Increasing $\zmax$ give
+better agreement with the analytic theory, which have $\zmax \to  
+\infty$. Larger boxes include regions of larger vertical shear (see
+Eqs. \ref{vshear_thin} and \ref{vertical_shear_ex}), which higher
+order modes can tap to give larger growth rates. For the $\zmax = 7
+H$ case, we begin to see another branch of modes with the highest
+growth rates at $\omega \sim -6 h\OmK$. This branch contains the
+`surface modes' described below.
 
-The need for vertically extended domains to capture the largest VSI growth rates is problematic, especially for hydrodynamic simulations.
-This  complication is mitigated by at least two factors.
-First, we show below that with slower cooling, the growth of higher order modes is preferentially damped, consistent with the analytic analysis in \S\ref{iso_vsi_beta_crit}.
-Second, the fastest growing modes may not dominate transport when they operate in very low density  surface layers, i.e.
-at many $H$.
-Quantifying which modes will contribute most to non-linear transport is an important issue, but beyond the scope of this work.
+The need for vertically extended domains to capture the largest VSI
+growth rates is problematic, especially for hydrodynamic simulations.
+This  complication is mitigated by at least two factors.  First, we
+show below that with slower cooling, the growth of higher order modes
+is preferentially damped, consistent with the analytic analysis in
+\S\ref{iso_vsi_beta_crit}.  Second, the fastest growing modes may not
+dominate transport when they operate in very low density  surface
+layers, i.e. at many $H$. Quantifying which modes will contribute
+most to non-linear transport is an important issue, but beyond the
+scope of this work.
 
@@
 Fig. \ref{lowfreq_eigenfunc_2d} maps the pressure perturbation and meridional flow,
-scaled by $\sqrt{\rho}$ to reflect the contribution to kinetic energy. Due to the stretched $x$ axis, 
-radial velocities are in fact typically much 
-smaller than vertical velocities, as expected for a vertically elongated, incompressible mode. 
-Most of the kinetic energy is contained within $\sim 2H$ of the midplane due to the density 
-stratification.
+scaled by $\sqrt{\rho}$ to reflect the contribution to kinetic
+energy. Notice the stretched $x$ axis. Radial velocities are in fact typically much 
+smaller than vertical velocities, as expected for a vertically
+elongated, anelastic mode \citepalias{nelson13}. Most of the kinetic
+energy is contained within $\sim 2H$ of the midplane due to the
+density stratification.
 
@@
   1cm]{figures/eigenvectorvz_iso}
-  \caption{Pseudo-enthalpy perturbation $W$ (top) and vertical velocity
-    perturbation $\delta v_z$ of the fundamental VSI with 
-    $\khat=10$ in the fiducial disk with rapid thermal relaxation $\beta=10^{-3}$. The real  
-    (imaginary) parts of $W$ and $\delta v_z$ are plotted as solid 
-    (dotted) lines.  
+  \caption{Pseudo-enthalpy perturbation $W=\delta P/\rho$ (top) and
+    vertical velocity perturbation $\delta v_z$ of the fundamental VSI with 
+    $\khat=10$ in the fiducial disk with rapid thermal relaxation
+    $\beta=10^{-3}$. The real (imaginary) parts of $W$ and $\delta
+    v_z$ are plotted as solid (dotted) lines.  
     \label{lowfreq_eigenfunc}
@@
 
-\subsubsection{Surface Modes of the VSI}\label{surf_comment} 
-The surface modes mentioned above are more prominent for larger wavenumbers.  
-Fig. \ref{compare_modes_iso_kx30} maps the $\khat = 30$
-eigenvalues, with the surface modes labelled.  The surface modes strongly depend 
-on the location of the imposed vertical boundary, and disagree with the analytic models, which lack an imposed surface. 
-We thus discount the physical significance of surface modes for our model, as we explain further below.
+\subsubsection{Surface modes of the VSI}\label{surf_comment} 
+The surface modes mentioned above are more prominent for larger
+wavenumbers. Fig. \ref{compare_modes_iso_kx30} maps the $\khat = 30$
+eigenvalues, with the surface modes labelled.  The surface modes
+strongly depend on the location of the imposed vertical boundary, and
+disagree with the analytic models, which lack an imposed surface. We
+thus discount the physical significance of surface modes for our
+model, as we explain further below.
 
-Surface modes are a well known feature of the VSI in finite vertical domains \citepalias{nelson13, mcnally14}.
-\citetalias{barker15} show that surface modes arise whether the surface is artificial
-(as for our imposed $\zmax$) or natural (for models with a zero density surface).  Physical surface modes
-are in principle possible; BL15 mention surfaces between a disk's interior and corona.  In models where the surface is imposed,
-like ours, surface modes are a numerical artifact.
+Surface modes are a well known feature of the VSI in finite vertical
+domains \citepalias{nelson13, mcnally14}. \citetalias{barker15} show
+that surface modes arise whether the surface is artificial (as for our
+imposed $\zmax$) or natural (for disk models with a zero density surface).
+Physical surface modes are in principle possible; BL15 mention
+surfaces between a disk's interior and corona.  In models where the
+surface is imposed, like ours, surface modes are a numerical
+artifact.
 
-The possibility of artificial surface modes seeding growth in numerical simulations of the VSI  merits further study.
-Indeed N13 find that the initial growth of perturbations primarily occurs near the vertical boundaries.
+The possibility of artificial surface modes seeding growth in
+numerical simulations of the VSI  merits further study. Indeed
+\citetalias{nelson13} find that the initial growth of perturbations
+primarily occurs near the vertical boundaries. 
+The motion in surface modes is indeed concentrated near the surface,
+as shown in Fig. \ref{lowfreq_eigenfunc_surf}.   
+Thus even if surface modes were physical, their contribution to
+transport might (by themselves) be weak,  following the arguments in 
+\S\ref{k10}.  Moreover surface modes, like all modes with large
+wavenumbers, are more prone to viscous damping, as discussed in
+\S\ref{caveats_visc}.
 
-The motion in surface modes is indeed concentrated near the surface, as shown in Fig. \ref{lowfreq_eigenfunc_surf}.  
-Thus even if surface modes were physical, their contribution to transport might (by themselves) be weak,  following the arguments in
-\S\ref{k10}.  Moreover surface modes, like all modes with large wavenumbers, are more prone to viscous damping, as discussed in \S\ref{caveats_visc}.
-
-While the relatively large growth rates of surface modes is tantalizing, we dismiss them as ``fool's gold"  in models that lack a physical surface.
+While the relatively large growth rates of surface modes is
+tantalizing, we dismiss them in disk models that lack a physical
+surface.
 

### 2015-05-07 numerical.tex

--- 2015-05-07 numerical.tex
@@
 Fig. \ref{lowfreq_eigenfunc_2d} maps the pressure perturbation and meridional flow,
-scaled by $\sqrt{\rho}$ to reflect the contribution to kinetic energy. Due to the stretched $x$ axis, 
-radial velocities are in fact typically much 
-smaller than vertical velocities, as expected for a vertically elongated, incompressible mode. 
-Most of the kinetic energy is contained within $\sim 2H$ of the midplane due to the density 
-stratification.
+scaled by $\sqrt{\rho}$ to reflect the contribution to kinetic
+energy. Notice the stretched $x$ axis. Radial velocities are in fact typically much 
+smaller than vertical velocities, as expected for a vertically
+elongated, anelastic mode \citepalias{nelson13}. Most of the kinetic
+energy is contained within $\sim 2H$ of the midplane due to the
+density stratification.
+
+\begin{figure}
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  0cm]{figures/eigenvectorW_iso} 
+  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/eigenvectorvz_iso}
+  \caption{Pseudo-enthalpy perturbation $W=\delta P/\rho$ (top) and
+    vertical velocity perturbation $\delta v_z$ of the fundamental VSI with 
+    $\khat=10$ in the fiducial disk with rapid thermal relaxation
+    $\beta=10^{-3}$. The real (imaginary) parts of $W$ and $\delta
+    v_z$ are plotted as solid (dotted) lines.  
+    \label{lowfreq_eigenfunc}
+\end{figure}
 

### 2015-05-07 application.tex
mkl edits of 5.3, 6.0, 6.1, 7; atlanta file updates

@@
     relaxation time $\beta$ at $z=0$ (solid) and $z=3H$ (dashed) is
-    less than the critical timescale $\beta_\mathrm{crit}$.   
+    less than the critical timescale $\beta_\mathrm{crit}$. Asterisks
+    correspond to the inner cut-off radii shown in
+    Fig. \ref{mmsn_overall}.   
     \label{mmsn_bcrit_bcool_mink}}   
@@
 Fig. \ref{mmsn_bcrit_bcool_mink} highlights this wavenumber dependence by plotting
-the wavenumbers for which $\beta = \beta_\mathrm{crit}$.
-Above the solid curves (i.e.\ for larger wavenumbers), $\beta <  \beta_\mathrm{crit}$ at all disk heights, implying linear VSI growth.
-Below the dashed curves, VSI growth is strongly suppressed since $\beta >  \beta_\mathrm{crit}$ for all $|z| > 3H$.
-Between the solid and dashed curves some growth is possible, but only fairly close to the solid curve (according to \S\ref{vsi_mmsn_grow}).
-Thus linear growth of the VSI near 1 AU is only possible if $k_x \gtrsim 10^3/H$.
-As argued elsewhere, we doubt that such small-scale disturbances will drive significant turbulence or transport.
+the wavenumbers for which $\beta = \beta_\mathrm{crit}$.  Above the
+solid curves (i.e.\ for larger wavenumbers), $\beta <
+\beta_\mathrm{crit}$ at all disk heights, implying linear VSI growth.
+Below the dashed curves, VSI growth is strongly suppressed since
+$\beta >  \beta_\mathrm{crit}$ for all $|z| < 3H$.  Between the solid
+and dashed curves some growth is possible, but only fairly close to
+the solid curve (according to \S\ref{vsi_mmsn_grow}).  Thus linear
+growth of the VSI near 1 AU is only possible if $k_x \gtrsim 10^3/H$.
+As argued elsewhere, such small-scale disturbances may not 
+drive significant turbulence or transport.
 
-We thus doubt that the VSI is significant at 1 AU in MMSN-like PPDs, for any opacity.
-At higher opacities, the required wavenumbers become shorter and even more problematic, as shown by the red curves in Fig.
-\ref{mmsn_bcrit_bcool_mink}.
-At lower opacties, even optically thin cooling (the fastest possible) is too slow, as discussed above.
+We thus doubt that the VSI is significant at 1 AU in MMSN-like PPDs,
+for any opacity.  At higher opacities, the required wavenumbers become
+shorter and even more problematic, as shown by the red curves in
+Fig. \ref{mmsn_bcrit_bcool_mink}.  At lower opacties, even optically
+thin cooling (the fastest possible) is too slow, as discussed above.
 
-\emph{Big question, do we need to include disk mass as another parameter to vary.  Or not in this study.  Has been included in SK14.}
+% \emph{Big question, do we need to include disk mass as another parameter to vary.  Or not in this study.  Has been included in SK14.}
 
@@
 \subsection{VSI growth in the MMSN}\label{vsi_mmsn_grow}
-We now consider the growth of the VSI in a MMSN disk model.  The linear analysis is similar to 
-\S\ref{numerical} but with different disk parameters and with the self-consistent cooling times of 
-Eq. \ref{beta_mmsn_simp}.
+We now consider the growth of the VSI in a MMSN disk model.  The
+linear analysis is similar to  
+\S\ref{numerical} but with different disk parameters and with the
+self-consistent cooling times of  Eq. \ref{beta_mmsn_simp}.
 
@@
 because it is the fastest growing mode except for some surface modes at high 
-wavenumbers.   We neglect these surface modes for reasons disucssed in \S\ref{numerical} and \S\ref{caveats_visc}.
+wavenumbers. We neglect these surface modes for reasons disucssed in
+\S\ref{numerical} and \S\ref{caveats_visc}.
 
@@
 
-Fig. \ref{mmsn_overall} shows that in the MMSN, the VSI  is active from $r_\mathrm{AU}\sim 5$ to $r_\mathrm{AU}\sim
-50$ with growth timescales $\sim 30$---40 orbits.
-A small radius cufoff exists, inside of which growth is strongly suppressed.
-The cutoff occurs at smaller radii for larger wavenumbers, as expected from Fig.\ \ref{mmsn_bcrit_bcool_mink}.
-Growth at smaller disk radii is possible for yet larger wavenumbers, yet we remain skeptical about the non-linear viability of such short modes.
+Fig. \ref{mmsn_overall} shows that in the MMSN, the VSI  is active
+from $r_\mathrm{AU}\sim 5$ to $r_\mathrm{AU}\sim 
+50$ with growth timescales $\sim 30$---40 orbits.  A small radius
+cut-off exists, inside of which growth is strongly suppressed.  The 
+cut-off occurs at smaller radii for larger wavenumbers, as expected
+from Fig.\ \ref{mmsn_bcrit_bcool_mink}.  Growth at smaller disk radii
+is possible for yet larger wavenumbers, yet we remain skeptical about
+the non-linear viability of such short modes.
 
-The growth times rise gradually as $r$ increases towards 100 AU in Fig.
-\ref{mmsn_overall}.
-This trend is expected as the outer radius cutoff at $\sim 200$AU (from Fig.\ \ref{mmsn_bcrit_bcool}) is approached.
+The growth times rise gradually as $r$ increases towards 100AU in
+Fig. \ref{mmsn_overall}.  This trend is expected as the outer radius
+cutoff at $\sim 150$AU (from Fig.\ \ref{mmsn_bcrit_bcool}) is
+approached. Our numerical results suggest the VSI is efficient at
+radii of few tens of AU in the MMSN, consistent with estimates made by
+\citetalias{nelson13} using the same disk model.
 
@@
 
-\emph{I don't understand the specific comparison being made here, especially since N13 is for vert.\ constant beta.
+% \emph{I don't understand the specific comparison being made here, especially since N13 is for vert.\ constant beta.
 Maybe remind me of which specific result/figure.
 I would think the more relevant comparison is to SK14 (either here or in discussion).}
- Our numerical results suggest the VSI
-is efficient at radii of few tens of AU in the MMSN, consistent with
-estimates made by \citetalias{nelson13}.
 

### 2015-05-07 numerical.tex
mkl edits of 5.3, 6.0, 6.1, 7; atlanta file updates
edits to 6.3.  Also removed max = 3H cases from two figures.  The fist one (beta = 10^-3) is enough to show that it's less convergent.  From there the 5 & 7 cases are clearly the most relevant and we need to narrow the focus onto what's most important.  Minor text changes in 6.2 and fig caption to accommodate this change.

@@
     \label{omega_z} 
+  }
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10_analytic.ps}
+  \caption{Growth rate, $\sigma$, vs.\ osciallation frequency, $\omega$, for
+  VSI modes with wavenumber $k_x = \khat/H=10/H$ in the fiducial disk model
+    with rapid cooling, $\beta=10^{-3}$. %  a disk with 
+    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q, h)=(-1.5,-1,0.1)$ and
+    % $\beta=10^{-3}$. 
+We investigate the effect of vertical box size, from $\zmax=3H$
+(\emph{blue triangles}) to $7H$ (\emph{green} x's)
+    with a free surface boundary ($\Delta P=0$).  For $\zmax=5H$  (black dots) we also compare free surface (\emph{black dots}) and rigid (\emph{red plusses}), finding little difference.
+      The fundamental mode is marked by `f'. Lines show the analytic predictions for
+      isothermal perturbations ($\beta = 0$, \emph{dashed}) and for $\beta = 10^{-3}$ (\emph{dotted}).
+      See text for discussion.
+   % are computed from the analytic dispersion relation 
+    %with thermal relaxation Eq. \ref{relax_disp} (dotted), and that for
+    %isothermal perturbations, Eq. \ref{simple_growth} (dashed). 
+    \label{compare_modes_iso_kx10} 
   }
@@
 % to axisymmetric perturbations).
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10_analytic.ps}
-  \caption{Growth rate, $\sigma$, vs.\ osciallation frequency, $\omega$, for
-  VSI modes with wavenumber $k_x = \khat/H=10/H$ in the fiducial disk model
-    with rapid cooling, $\beta=10^{-3}$. %  a disk with 
-    % $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q, h)=(-1.5,-1,0.1)$ and
-    % $\beta=10^{-3}$. 
-We investigate the effect of vertical box size, from $\zmax=3H$
-(\emph{blue triangles}) to $7H$ (\emph{green} x's)
-    with a free surface boundary ($\Delta P=0$).  For $\zmax=5H$  (black dots) we also compare free surface (\emph{black dots}) and rigid (\emph{red plusses}), finding little difference.
-      The fundamental mode is marked by `f'. Lines show the analytic predictions for
-      isothermal perturbations ($\beta = 0$, \emph{dashed}) and for $\beta = 10^{-3}$ (\emph{dotted}).
-      See text for discussion.
-   % are computed from the analytic dispersion relation 
-    %with thermal relaxation Eq. \ref{relax_disp} (dotted), and that for
-    %isothermal perturbations, Eq. \ref{simple_growth} (dashed). 
-    \label{compare_modes_iso_kx10} 
-  }
-\end{figure}
 
@@
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/eigenvectorW_iso} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  1cm]{figures/eigenvectorvz_iso}
-  \caption{Pseudo-enthalpy perturbation $W=\delta P/\rho$ (top) and
-    vertical velocity perturbation $\delta v_z$ of the fundamental VSI with 
-    $\khat=10$ in the fiducial disk with rapid thermal relaxation
-    $\beta=10^{-3}$. The real (imaginary) parts of $W$ and $\delta
-    v_z$ are plotted as solid (dotted) lines.  
-    \label{lowfreq_eigenfunc}
-\end{figure}
-
-\begin{figure}
   \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_analytic.ps}
@@
 \subsection{Slower thermal relaxation}\label{therm_relax_eff}
-We now consider the effect of longer cooling timescales by gradually increasing 
-$\beta$. We expect VSI growth for  $\beta < \beta_\mathrm{crit} = 0.125$.  We also expect that higher order modes will damp at even lower $\beta$ values, as argued in \S\ref{iso_vsi_beta_crit}.
+We now consider the effect of longer cooling timescales by gradually
+increasing  
+$\beta$. We expect VSI growth for  $\beta < \beta_\mathrm{crit} =
+0.125$.  We also expect that higher order modes will damp at even
+lower $\beta$ values, as argued in \S\ref{iso_vsi_beta_crit}.
 
-Fig. \ref{compare_modes_cool_kx10} largely confirms these expectations by plotting eigenvalues for $\khat=10$ 
+Fig. \ref{compare_modes_cool_kx10} largely confirms these expectations
+by plotting eigenvalues for $\khat=10$  
 and $\beta$ from $0.01$ to  $0.1$.  
-The analytic results from Eq. \ref{relax_disp} are now plotted as different curves for each mode order $M$, with $\beta$ varying along each curve.  
-We see the standard increase in frequency, $|\omega|$, with mode order.
+The analytic results from Eq. \ref{relax_disp} are now plotted as
+different curves for each mode order $M$, with $\beta$ varying along
+each curve. We see the standard increase in frequency, $|\omega|$,
+with mode order.
 
-As expected, the higher order modes are preferentially damped.
-For $\beta \geq 0.03$ the fundamental ($M = 0$) mode is the fastest growing.
-For $\beta = 0.1$, only the $M = 0$ mode grows.
-This growth is slow since $\beta$ is near $\beta_\mathrm{crit}$.
+As expected, the higher order modes are preferentially damped.  For
+$\beta \geq 0.03$ the fundamental ($M = 0$) mode is the fastest
+growing. For $\beta = 0.1$, only the $M = 0$ mode grows. This growth
+is slow since $\beta$ is near $\beta_\mathrm{crit}$.
 
-Fig.
-\ref{compare_modes_cool_kx10} also shows how cooling times affect the dependence on $\zmax$, the size of the vertical domain.
-For all cooling times, the fundamental mode depends only weakly on $\zmax$, and there is good agreement with analytic values.
-This convergence is reassuring given the importance of the fundamental mode at longer cooling times.
-For higher order modes and short cooling times, however, the eigenvalues vary strongly with $\zmax$.
-The disagreement with analytic theory is strongest for the smallest domain, with $\zmax = 3 H$.
-For $\beta \geq 0.03$, the $\zmax/H = 5, 7$ and the analytic results are well converged (aided by the fact that only $M < 5$ modes grow).
+Fig. \ref{compare_modes_cool_kx10} also shows how cooling times affect
+the dependence on $\zmax$, the size of the vertical domain.   For all
+cooling times, the fundamental mode depends only weakly on $\zmax$,
+and there is good agreement with analytic values.  This convergence is
+reassuring given the importance of the fundamental mode at longer
+cooling times.  For higher order modes and short cooling times,
+however, the eigenvalues vary strongly with $\zmax$.  The disagreement
+with analytic theory is strongest for the smallest domain, with $\zmax
+= 3 H$.  For $\beta \geq 0.03$, the $\zmax/H = 5, 7$ and the analytic
+results are well converged (aided by the fact that only $M < 5$ modes
+grow).
 
@@
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0cm]{figures/compare_modes_cool_kx30_z3_analytic.ps} 
+ % \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  %0cm]{figures/compare_modes_cool_kx30_z3_analytic.ps} 
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
@@
   \caption{Same as Fig. \ref{compare_modes_cool_kx10} but for
-    $\khat=30$. Examples of surface modes are marked with `s'. 
+    $\khat=30$ and domain sizes $\zmax = 5H$ (\emph{top}) and $7H$ (\emph{bottom}). Examples of surface modes are marked with `s'. 
     \label{compare_modes_cool_kx30} 
@@
 Fig. \ref{compare_modes_cool_kx30} shows how cooling affects higher wavenumber
-modes, specifically $\khat=30$.   We focus on the $\zmax/H = 5, 7$ and analytic results for $\beta \geq 0.03$. 
-These cases  show good convergence, though not quite as good as for $\khat = 10$.
+modes, specifically $\khat=30$.   We only show the larger domains with $\zmax = 5H, 7H$. 
+These cases again show good convergence for $\beta \geq 0.03$, which the
+the smaller domain with  $\zmax = 3H$ lacks.
 
@@
 Fig. \ref{compare_modes_cool_kx30} also demonstrates that artificial surface modes are damped for $\beta \geq 0.03$.
+We are thus confident that artificial surface modes should not affect our determination of the critical cooling time.
 
@@
 
-\subsection{Critical thermal relaxation
-  timescale}\label{bcrit_num_test}
-Having explored the behavior of VSI modes with cooling, we turn to the numerical vailidity of the 
-analytic cooling criterion, $\beta < \beta_\mathrm{crit} = h|q|/(\gamma -1)$.
-
-...
-
-In Fig. \ref{bcrit_compare1} we plot the maximum VSI growth rates
-found in the fiducial disk model as a function of the thermal
-relaxation timescale $\beta$ for several values of $\khat$ and
-$\zmax$. We also plot the critical timescale
-$\beta_\mathrm{crit}=0.125$ for the fiducial disk parameters.
-
-Note that the curves are not neccessarily smooth because the most unstable
-mode may not always correspond to the same type of disturbance. For example,
-as seen above, for $\khat \leq 10$ higher order body modes are
-dominant as $\beta\to 0$, but the fundamental 
-mode becomes dominant as $\beta$ is increased. For $\khat=1$ the VSI
-can operate slightly beyond $\beta_\mathrm{crit}$  (recall that
-$\beta_\mathrm{crit}$ was derived assuming $\khat^2\gg 1$). For
-$\khat=5,\,10$ (and $30$ for $\zmax=3H$), we find
-$\beta_\mathrm{crit}$ accurately predicts the thermal timescal beyond
-which the VSI is suppressed.
-
-Complications from vertical boundaries arise when considering large
-$\khat$. For $\khat=30$ and $\zmax=5H,\,7H$ we find for $\beta\gtrsim
-0.1$ the dominant disturbance correspond to the leftmost eigenvalues 
-seen in the lower two panels of Fig. \ref{compare_modes_cool_kx30},
-which do not exist for $\zmax=3H$. For $\khat=50,\,100$ and
-$\beta\to0$, we find surface modes dominate so the maximum growth rate
-depends on $\zmax$ in the limit of rapid thermal relaxation. 
-However, the maximum growth rate is insensitive to 
-$\zmax$ for moderately large $\beta$ ($\gtrsim 0.1$), i.e. close to
-the stabilization of the fundamental mode.  
-%not surprising since we enable strongly stabilizing buoyancy forces
-%at large zmax
-
-Our result is consistent with numerical simulations performed by
-\citetalias{nelson13} for which a thermal relaxation timescale $\beta\gtrsim
-0.6$ stabilized the VSI; as Fig. \ref{bcrit_compare1} show that growth
-rates for $\beta\gtrsim 0.6$ are small, and are restricted to very
-small radial lengthscales $\sim 10^{-2}H$, for which current numerical
-simulations may not be able to resolve.
-%  To destablize the disk with
-% $\beta\gtrsim O(1)$, one might consider 
-% $\khat > 100$ to increase the maximum growth rate by enabling surface
-% modes, but in a realistic disk
-% very small scale disturbances will likely be subject to viscous decay
-% (see \S\ref{application}). 
-%Numerical simulations are also unlikely to
-%resolve such small scales.
-
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
-  0.9cm]{figures/gcorr_compare_iso_maxrate_z3} 
+  %\includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  %0.9cm]{figures/gcorr_compare_iso_maxrate_z3} 
   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
@@
      model as a function of the thermal relaxation timescale
-     $\beta$, for $\zmax=3H$ (top), $\zmax=5H$ (middle, standard
-     setup) and $\zmax=7H$ (bottom). The vertical line is the
+     $\beta$, for %$\zmax=3H$ (top), 
+     $\zmax=5H$ (top) and $\zmax=7H$ (bottom). The vertical line is the
      critical thermal timescale $\beta_\mathrm{crit}$ obtained  
@@
 
-As pointed out by \citetalias{barker15}, in vertically isothermal disks with
-an imposed boundary, surface modes are dependent on $\zmax$, but they
-are often the most unstable. Despite this caveat, and that surface
-modes are not accounted in our analysis, we see from
-Fig. \ref{bcrit_compare1} that $\beta_\mathrm{crit}$ nevertheless
-provides a reasonable estimate for an upper limit to the thermal
-timescale for the VSI.  
-%It becomes an accurate criter when considering fixed
-%$\khat$
+\subsection{Critical thermal relaxation timescale}\label{bcrit_num_test}
+Having explored the behavior of VSI modes with cooling, we turn to the numerical validity of the 
+analytic cooling criterion for vertically isothermal disks, $\beta < \beta_\mathrm{crit} = h|q|/(\gamma -1)$.
 
-We further  test the dependence of $\beta_\mathrm{crit}$ on disk parameters
-(Eq. \ref{iso_vsi_cond}).  We use the fiducial setup as a  reference
-point and vary $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and
-$ h\in[0.02,0.1]$ separately. The numerically obtained
-$\beta_\mathrm{crit}$ are shown in Fig. \ref{bcrit_compare} in
-comparison with Eq. \ref{iso_vsi_cond}, which generally agree. This
-result, together with Fig. \ref{bcrit_compare1}, shows that  
-$\beta \lesssim \beta_\mathrm{crit}$ is a practical
-criteria for the VSI to operate effectively in stably stratified,
-vertically isothermal disks.
+Fig. \ref{bcrit_compare1} shows how VSI growth rates vary with cooling time 
+in our fiducial model with $\beta_\mathrm{crit} = 0.125$.  
+Curves are for a fixed horizontal wavenumber, and show the maximum growth rate for all vertical mode orders.  The 
+discontinuity in some curves occurs when the fastest growth switches to a different mode order.
 
-% The agreement improves with decreasing  $|q|$,
-% $ h$ and increasing $\gamma$, i.e. weaker instability. There is
-% a noticeable difference as $\gamma\to1$ because the derivation of
-% $\beta_\mathrm{crit}$ assumed $\gamma>1$.
+For $\khat = 5, 10$, the growth rate drops to zero at the expected $\beta = \beta_\mathrm{crit}$.
+For longer wavelength modes, with $\khat = 1$, growth persists to slightly longer cooling times.  
+This difference is not surprising since our analytic derivation assumed $\khat \gg1$, and 
+the change is quantitatively minor.
+
+For shorter wavelengths, with $\khat \geq 30$, growth persists for $\beta$ significantly larger
+than $\beta_\mathrm{crit}$.  This tail of growth is partly explained by the breaking of the $\khat < 1/|qh| = 20$ 
+approximation, used in  the analytic derivation.  Despite the lack of a clear stability boundary 
+at high $\khat$, the $\beta_\mathrm{crit}$ threshold remains useful, since growth rates drop to 
+$\lesssim 10\%$ of  their maximum value at  $\beta_\mathrm{crit}$ and continue to fall for larger $\beta$. 
+ Moreover, we expect that longer wavelength modes with $\khat \lesssim 20$
+are more significant for disk transport, see \S\ref{caveats_visc}.
+
+Comparing the $\zmax/H = 5, 7$ cases in Fig. \ref{bcrit_compare1}, we see that the location of the vertical
+boundary has little effect on the critical cooling time.  This agreement occurs despite the fact that peak growth rates
+differ as $\beta \to 0$.  We are thus confident that artificial boundary effects, including surface modes, do not
+affect our analysis of the critical cooling time.
+
+Our results agree with the vertically isothermal analysis of \citetalias{nelson13} 
+which used the same disk parameters as our fiducial model.  The expected $\beta_\mathrm{crit} = 0.125$ 
+is consistent with the simulations shown in their Fig.\ 12.  \citetalias{nelson13}  
+found nonlinear VSI growth for $\beta = 0.06< \beta_\mathrm{crit}$ but no growth for 
+  $\beta = 0.6 > \beta_\mathrm{crit}$.\footnote{Since the dimensionless cooling time $T_{\rm relax}$ in \citetalias{nelson13}
+  is normalized to the orbital period, we convert $\beta = \OmK P_{\rm orb} T_{\rm relax} = 2 \pi T_{\rm relax}$.}
 
@@
 \end{figure}
+
+Fig.\ \ref{bcrit_compare} confirms that the numerically determined
+$\beta_\mathrm{crit}$ shows the expected scaling with disk parameters $h$, $q$, and $\gamma$.
+For this test we fix $\khat = 10$ (an appropriate value for all the reasons discussed above) and measure
+the smallest $\beta$ value where growth vanishes.  The agreement with the analytic scaling of 
+Eq.\ \ref{iso_vsi_cond} is quite good, confirming the applicability of our critical cooling time in
+vertically isothermal disks.
+
+%we plot the maximum VSI growth rates
+%found in the fiducial disk model as a function of the thermal
+%relaxation timescale $\beta$ for several values of $\khat$ and
+%$\zmax$. We also plot the critical timescale
+%$\beta_\mathrm{crit}=0.125$ for the fiducial disk parameters.   
+%
+%Note that the curves are not neccessarily smooth because the most unstable
+%mode may not always correspond to the same type of disturbance. For example,
+%as seen above, for $\khat \leq 10$ higher order body modes are
+%dominant as $\beta\to 0$, but the fundamental 
+%mode becomes dominant as $\beta$ is increased. For $\khat=1$ the VSI
+%can operate slightly beyond $\beta_\mathrm{crit}$  (recall that
+%$\beta_\mathrm{crit}$ was derived assuming $\khat^2\gg 1$). For
+%$\khat=5,\,10$ (and $30$ for $\zmax=3H$), we find
+%$\beta_\mathrm{crit}$ accurately predicts the thermal timescal beyond
+%which the VSI is suppressed.   
+%
+%Complications from vertical boundaries arise when considering large
+%$\khat$. For $\khat=30$ and $\zmax=5H,\,7H$ we find for $\beta\gtrsim
+%0.1$ the dominant disturbance correspond to the leftmost eigenvalues 
+%seen in the lower two panels of Fig. \ref{compare_modes_cool_kx30},
+%which do not exist for $\zmax=3H$. For $\khat=50,\,100$ and
+%$\beta\to0$, we find surface modes dominate so the maximum growth rate
+%depends on $\zmax$ in the limit of rapid thermal relaxation. 
+%However, the maximum growth rate is insensitive to 
+%$\zmax$ for moderately large $\beta$ ($\gtrsim 0.1$), i.e. close to
+%the stabilization of the fundamental mode.
+
+%Our result is consistent with numerical simulations performed by
+%\citetalias{nelson13} for which a thermal relaxation timescale $\beta\gtrsim
+%0.6$ stabilized the VSI; as Fig. \ref{bcrit_compare1} show that growth
+%rates for $\beta\gtrsim 0.6$ are small, and are restricted to very
+%small radial lengthscales $\sim 10^{-2}H$, for which current numerical
+%simulations may not be able to resolve.
+%
+%As pointed out by \citetalias{barker15}, in vertically isothermal disks with
+%an imposed boundary, surface modes are dependent on $\zmax$, but they
+%are often the most unstable. Despite this caveat, and that surface
+%modes are not accounted in our analysis, we see from
+%Fig. \ref{bcrit_compare1} that $\beta_\mathrm{crit}$ nevertheless
+%provides a reasonable estimate for an upper limit to the thermal
+%timescale for the VSI.  
+%
+%We further  test the dependence of $\beta_\mathrm{crit}$ on disk parameters
+%(Eq. \ref{iso_vsi_cond}).  We use the fiducial setup as a  reference
+%point and vary $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and
+%$ h\in[0.02,0.1]$ separately. The numerically obtained
+%$\beta_\mathrm{crit}$ are shown in Fig. \ref{bcrit_compare} in
+%comparison with Eq. \ref{iso_vsi_cond}, which generally agree. This
+%result, together with Fig. \ref{bcrit_compare1}, shows that  
+%$\beta \lesssim \beta_\mathrm{crit}$ is a practical
+%criteria for the VSI to operate effectively in stably stratified,
+%vertically isothermal disks.
 

### 2015-05-07 summary.tex
edits to caveats, including moving them above the summary.  Don't want to end on caveats and neglected effects, even if some of them are future looking.  minor tweaks elsewhere to match.

@@
+\section{Caveats and Neglected Effects} 
+%We highlight below some outstanding issues that should be addressed in
+%future works.
+
+\subsection{Viscosity}\label{caveats_visc}
+% Depending on where the VSI operates in a
+% disk, additional effects may become important. For example, in the
+% outer parts of a PPD, where we estimate the VSI 
+% operates, disk self-gravity (SG) may play a role. Vertical
+% SG may favor the VSI by increasing $|\p_z\Omega|$ through
+% enhancing the density stratification 
+% (Eq. \ref{vertical_shear}). %sg in pert? no because kx>>1. but maybe for kx~1
+
+Our neglect of viscosity is valid in a laminar disk, because molecular viscosity is so small.
+However turbulence could act as an effective viscosity which preferentially damps 
+small scale modes.   Since the goal of VSI is to drive turbulent transport,
+the most relevant modes for sustaining the VSI should be able to overcome
+turbulent damping.
+
+To roughly estimate the wavelengths which are prone to damping, we
+consider 
+%An important effect of viscosity is setting a minimum wavelength (or 
+%maximum wavenumber) for which the VSI is relevant. Disturbances with
+%large $\khat\equiv k_xH$ may be subject to viscous decay. Adopting 
+the standard
+precription for the kinematic viscosity $\nu = \alpha c_s H$
+\citep{shakura73}, in terms of the dimsionless $\alpha$ 
+parameter. The viscous timescale for a perturbation lengthscale  
+$l\sim 1/k_x = H/\khat$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $. 
+For significant growth, $t_\mathrm{visc}$ should be longer than the
+characterstic VSI growth timescale $1/(h|q|\Omega)$, i.e.\ 
+$\khat^2\lesssim |q|h/\alpha$. With $h \sim 0.05$, $q = -1$  and
+ $\alpha$ from simulations \citepalias{nelson13},
+we estimate that growth requires $\khat\lesssim 20$ for
+$\alpha\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha\sim
+10^{-5}$.
+
+This argument justifies our focus on moderate
+wavenumbers, $\khat=O(10)$. Future work should consider
+the viscous effects in more detail, to better understand
+how the VSI operates in nature and in simulations.
+
+%suggestion: Keep this idea to yourself, since you don't have a result yet.
+% In particular, viscous
+%heating may influence the thermodynamic dependence of the VSI.
+
+\subsection{Convective overstability}
+The VSI bears some similarity to the `convective overstability'
+ (CONO, \citealp{klahr14, lyra14}).   Both instabilities rely on thermal relaxation to 
+overcome the stability of adiabatic perturbations, i.e.\ the Solberg-Hoiland
+criteria (see \S\ref{solberg}).
+
+The key differences are that the CONO neglects vertical buoyancy
+and requires an imaginary horizontal buoyancy,
+\begin{align}\label{Nr}
+N_r^2 \equiv -{1 \over \rho C_P}{\p P \over \p r} {\p S \over \p r} < 0\, .
+\end{align}
+Since vertical buoyancy is an important stabilitizing influnce for the VSI, it should
+be considered in future studies of CONO.
+
+The sign of $N_r^2$ depends on disk parameters.  To demonstrate that the required parameters 
+are somewhat extreme, we consider the midplane of vertically isothermal disks with 
+$\Sigma \propto r^n$ so that $n = p+q/2 + 3/2$. Using Eqs. \ref{dSdr} and \ref{Nr} with $\gamma = 1.4$,
+$N_r^2 < 0$ requires $n> 3(1/2 + q)$.  For standard disk temperature laws this requirement implies
+a flat or rising surface density profile, e.g.\ $n = 3/14$ for $q = -3/7$ or $n = 0$ for $q = -1/2$.
+Thus CONO is most like to operate at special locations, like the outside edges of 
+disk gaps or holes and/or shadowed regions where $q < -1/2$.
+
+Perhaps due to these physical differences, CONO operates in different regimes of parameter
+space than the VSI.  CONO grows best for longer wavelengths $\khat \gtrsim 1$ and longer 
+cooling times $\beta \sim 1$.
+
+Future work should aim to understand these related instabilities in the same framework, thereby 
+explaining the key differences.
+
+%We have focused on parameter regimes relevant to the VSI,
+%namely perturbations with small radial lengthscales 
+%($k_xH\gtrsim 1$) in disks with short thermal timescales
+%($t_c\Omega\ll 1$). This excludes the `convective overstability'
+%recently discussed by \cite{klahr14} and \cite{lyra14} in vertically
+%unstratified disk models.   
+%
+%The convective overstability corresponds to slowly growing epicyclic motions.
+%It is most effective for $t_c\Omega \sim 1$ and disturbances with
+%$|k_x|\ll |k_z|$, where $k_z$ is a vertical wavenumber. These
+%conditions are mutually exclusive with the VSI.  
+%The maximum growth rate for convective overstability is given by 
+%\begin{align}
+%  \sigma_\mathrm{conv,max} = -\frac{N^2_r}{4\Omega}\label{max_conv_gen}
+%\end{align}
+%\citep{lyra14}, where $N_r^2 = -(c_p\rho)^{-1}\p_rP\p_rS$ is the
+%square of the radial buoyancy frequency
+%(cf. Eq. \ref{nzsq_def}). Thus, growth requires $N_r^2<0$.  
+%%Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
+%%we find  
+%%\begin{align}
+%%  \sigma_\mathrm{conv,max} =
+%%  -\frac{ h^2\OmK}{4\gamma\Gamma}\left(p+q\right)
+%%  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
+%%\end{align}
+%In fact, for the MMSN considered in this work, convective
+%overstability does not apply because 
+%$N_r^2>0$.   
+%
+%For more general disk structures convective overstability is
+%possible, in principle. The models adopted for our numerical study in
+%\S\ref{numerical} do satisfy $N_r^2<0$, but convective overstability
+%require perturbations with long radial lengthscales,
+%$k_xH\ll 1$, for which radially global disk models would be more 
+%self-consistent.
+
+\subsubsection{Radiative transfer} 
+Our relatively simple treatment of cooling with an idealized dust opacity
+could certainly be generalized in future works.  In hotter disks, gas phase opacities
+must be considered.   In cold disks there are many choices for a dust opacity which 
+varies with grain sizes and compositions. Dust properties could change the viability 
+of the VSI for better or worse.  The radiative transfer itself
+could be calculated with higher levels of sophistication, as is already being done in
+numerical simulations \citep{stoll14}.
+
 \section{Summary and discussion}\label{summary}
@@
 calculations or non-linear numerical simulations.
-
-\subsection{Caveats and outlooks} 
-We highlight below some outstanding issues that should be addressed in
-future works.
-
-\subsubsection{Viscosity}\label{caveats_visc}
-% Depending on where the VSI operates in a
-% disk, additional effects may become important. For example, in the
-% outer parts of a PPD, where we estimate the VSI 
-% operates, disk self-gravity (SG) may play a role. Vertical
-% SG may favor the VSI by increasing $|\p_z\Omega|$ through
-% enhancing the density stratification 
-% (Eq. \ref{vertical_shear}). %sg in pert? no because kx>>1. but maybe for kx~1
-
-We have neglected viscosity. 
-% while \citetalias{barker15} find that inclusion of viscosity is needed for 
-% convergence when studying surface modes in radially global disk
-% models. %but not body modes? 
-% Viscosity may also affect the
-% thermodynamic requirement for the VSI, because viscous heating
-% needs to be self-consistently included in the energy equation. This
-% should be investigated in future works.   
-An important effect of viscosity is setting a minimum wavelength (or 
-maximum wavenumber) for which the VSI is relevant. Disturbances with
-large $\khat\equiv k_xH$ may be subject to viscous decay. Adopting the standard
-precription for the kinematic viscosity $\nu = \alpha c_s H$
-\citep{shakura73}, where $\alpha$ is the dimensionless viscosity
-parameter, the viscous timescale for a perturbation lengthscale  
-$l\sim 1/k_x$ is $t_\mathrm{visc} = 1/\alpha\khat^2\Omega $. 
-Setting $t_\mathrm{visc}$ to be longer than the
-minimum VSI growth timescale $1/h\Omega$ implies 
-$\khat^2\lesssim h/\alpha$. Inserting $h \sim 0.05$,
-we find only modes with $\khat\lesssim 20$ are relevant for
-$\alpha\sim 10^{-4}$ or $\khat\lesssim 70$ for $\alpha\sim
-10^{-5}$. \citepalias[These small viscosity values are required for
-the VSI to operate, see][]{nelson13}.
-
-%\citetalias{nelson13} find that a
-%dimensionless viscosity of $\alpha\lesssim O(10^{-4})$ is required for the
-%VSI.
-
-The above argument justifies our focus on moderate
-wavenumbers, $\khat=O(10)$, which precludes extremely small-scale
-disturbances such as surface modes. Nevertheless, future calculations
-should explicitly account for viscosity. In particular, viscous
-heating may influence the thermodynamic dependence of the VSI.
-
-% needs to be self-consistently included in the energy equation.
-
-% It is of interest to quantify the effect of viscosity 
-% on the linear VSI. This would help identify, more accurately, the
-% appropriate regions in a protoplanetary disk that are sufficiently
-% laminar to develop the VSI.  This problem is, however, complicated by
-% the fact that proper inclusion of viscosity introduces a meridional
-% flow in the equilibrium state, as well as viscous heating
-% in the energy equation.
-
-\subsubsection{Convective overstability}
-We have focused on parameter regimes relevant to the VSI,
-namely perturbations with small radial lengthscales 
-($k_xH\gtrsim 1$) in disks with short thermal timescales
-($t_c\Omega\ll 1$). This excludes the `convective overstability'
-recently discussed by \cite{klahr14} and \cite{lyra14} in vertically
-unstratified disk models.
-
-The convective overstability corresponds to slowly growing epicyclic motions.
-It is most effective for $t_c\Omega \sim 1$ and disturbances with
-$|k_x|\ll |k_z|$, where $k_z$ is a vertical wavenumber. These
-conditions are mutually exclusive with the VSI.  
-The maximum growth rate for convective overstability is given by 
-\begin{align}
-  \sigma_\mathrm{conv,max} = -\frac{N^2_r}{4\Omega}\label{max_conv_gen}
-\end{align}
-\citep{lyra14}, where $N_r^2 = -(c_p\rho)^{-1}\p_rP\p_rS$ is the
-square of the radial buoyancy frequency
-(cf. Eq. \ref{nzsq_def}). Thus, growth requires $N_r^2<0$.  
-%Evaluating Eq. \ref{max_conv_gen} at the midplane of our disk models,
-%we find  
-%\begin{align}
-%  \sigma_\mathrm{conv,max} =
-%  -\frac{ h^2\OmK}{4\gamma\Gamma}\left(p+q\right)
-%  \left[\left(\gamma-1\right)p-q\right]. \label{max_conv}
-%\end{align}
-In fact, for the MMSN considered in this work, convective
-overstability does not apply because 
-$N_r^2>0$.
-
-For more general disk structures convective overstability is
-possible, in principle. The models adopted for our numerical study in
-\S\ref{numerical} do satisfy $N_r^2<0$, but convective overstability
-require perturbations with long radial lengthscales,
-$k_xH\ll 1$, for which radially global disk models would be more 
-self-consistent.
-
-\subsubsection{Improved thermal evolution} 
-We assumed non-adiabatic effects can be represented by a thermal
-relaxation timescale $t_c$. For  
-application to PPDs we connected $t_c$ to the disk
-properties such as surface density, temperature and opacity, making
-$t_c$ depend on position and the perturbation 
-lengthscales. One might improve the realism further by generalizing
-the opacity model adopted here to be density (and hence height)
-dependent.  
-%here to be density-dependent,  i.e. $\kappa_d =
-%% \kappa_d(\rho, T)$. 
-However, going beyond thermal relaxation requires a
-proper treatment of radiative diffusion. 
-% done roughly in this work -
-% a proper study 
-% involves a Laplacian operator and additional boundary
-% conditions must be supplemented. 
-We will consider these improvements in a
-follow-up study. %additional heat source

### 2015-05-07 numerical.tex
merge with youdin's updates
new 6.4, merge with youdin's updates

@@
   %0.9cm]{figures/gcorr_compare_iso_maxrate_z3} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
+  \includegraphics[scale=0.4415,clip=true,trim=0cm 1.75cm 0cm
   0.9cm]{figures/gcorr_compare_iso_maxrate_z5} 
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
+  \includegraphics[scale=0.4415,clip=true,trim=0cm 0.0cm 0cm
   0.9cm]{figures/gcorr_compare_iso_maxrate_z7}  
@@
 For longer wavelength modes, with $\khat = 1$, growth persists to slightly longer cooling times.  
-This difference is not surprising since our analytic derivation assumed $\khat \gg1$, and 
+This difference is not surprising since our analytic derivation assumed $\khat^2 \gg1$, and 
 the change is quantitatively minor.
@@
 \subsection{Vertically non-isothermal disk model} 
-%\emph{fix this rambling discussion..}
+We briefly consider a vertically non-isothermal disk model with
+$(p,q, h)=(-0.5,-1,0.05)$ and $(\gamma, \Gamma)=(1.4,1.3)$. We set the
+vertical domain size close to the physical disk surface, $\zmax =
+0.99H_s\simeq 2.6H$, which is smaller than our standard domain
+size. In this case the vertical shear rate $\p_z\Omega\propto s$, the
+radial entropy gradient. Here, $s=-0.85$.
 
-% \citetalias{barker15} have argued that vertically isothermal disks are not
-% ideal for studing the VSI because $\zmax$ is a free 
-% parameter in numerical calcluations. On the other hand, vertically 
-% non-isothermal disks have a disk surface at finite height so there is
-% a natural choice for $\zmax$. This means that higher order body modes and
-% surface modes are better defined in a vertically non-isothermal disk.
+From the expression for the
+critical cooling timescale, Eq. \ref{iso_vsi_cond}, it is reasonable
+to expect a similar criteria for vertically non-isothermal 
+disks in the form
+\begin{align}\label{bcrit_noniso}
+ \beta_\mathrm{crit}\to\beta_\mathrm{crit, noniso} =
+ \frac{h|s|}{\gamma - \Gamma},
+\end{align}
+and $\beta_\mathrm{crit,noniso}=0.425$ for the chosen parameters. This
+is $\sim 3$ times larger than our fiducial, vertically isothermal
+disk with $\beta_\mathrm{crit}=0.125$.
 
-% However, as we have seen in Fig. \ref{bcrit_compare1}, the maximum
-% growth rate becomes insensitive to $\zmax$ for $\beta\gtrsim
-% \beta_\mathrm{crit}$. This suggests that the characteristic upper
-% limit on the thermal timescale for the VSI is not sensitive to
-% $\zmax$, so that vertically isothermal disks are in fact suitable for
-% the purpose of characterizing the thermodynamic dependence of the
-% VSI.
-
-%Physically this is because modes that depend on vertical boundaries
-%are also subject to the strongest stabilization by buoyancy, which
-%maximize at $\zmax$. These modes are theremore always strongly damped
-%as we increase the thermal timescale.
-
-Here we briefly consider a vertically non-isothermal
-disk model with $(p,q, h)=(0,-1,0.05)$ and $\Gamma=1.0837$. This
-value of $\Gamma$ is chosen so that the standard vertical domain size $\zmax =
-5H\simeq H_s$ effectively corresponds to the disk surface. The disk
-temperature decreases rapidly with height such that
-$c_s^2(\zmax)\simeq 10^{-4}c_s^2(0)$. A disk surface at finite height
-means that higher order body modes and surface modes are better
-defined (cf. in vertically isothermal disk models where the disk
-boundary $\zmax$ is a free parameter, see \citetalias{barker15}).
-
-% \emph{note that this would
-% imply midplane 
-% temperatures in excess of $10^6$ K so it's completely un-astrophyscial.}
+% A disk surface at finite height
+% means that higher order body modes and surface modes are better
+% defined (cf. in vertically isothermal disk models where the disk
+% boundary $\zmax$ is a free parameter, see \citetalias{barker15}).
 
 In Fig. \ref{compare_modes_vnoniso_kx10} we plot the mode diagram for
-$\khat=10$ for several values of $\beta$. This plot is 
+$\khat=30$ for several values of $\beta$. This plot is 
 qualitatively similar to Fig. \ref{compare_modes_cool_kx10}. As before
@@
 stabilized, eventually only the fundamental mode can operate before
-the VSI is entirely suppressed.
+the VSI is entirely suppressed. (Note the significant reduction in
+growth rates for $\beta=0.4$, which is close to
+$\beta_\mathrm{crit,noniso}$.) However, this behavior is not seen for 
+large wavenumbers (see below).
 
@@
   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/compare_modes_Gam1.084_kx10.ps}
+  0cm]{figures/compare_modes_Gam1.3_kx30.ps}
   \caption{Unstable modes in a vertically non-isothermal disk with
-    $\khat=10$ and a range of thermal relaxation timescales.  
+    $\khat=30$ and a range of thermal relaxation timescales.  
     \label{compare_modes_vnoniso_kx10}}
@@
 
-We do not find surface modes for $\khat=10$, but do find them for
-sufficiently large $\khat$. We demonstrate this in  % In the vertically
-% non-isothermal disk, surface modes develop for sufficiently large 
-% $\khat$ \citep{barker15}. We demonstrate them in
-Fig. \ref{compare_modes_vnoniso_kx100} which show eigenvalues for
-$\khat=100$. While growth rates generally decrease with increasing 
-$\beta$, we find eigenvalues can overlap, making it difficult to 
-distinguish between different modes. However, since there 
-is a well-defined surface, we may seek the  
-most unstable mode without the boundary-dependent issue that arose 
-with a vertically isothermal disk in the limit $\beta\to0$.
+In Fig. \ref{bcrit_compare2} we plot the maximum VSI growth
+rates in the vertically non-isothermal disk model, which is
+qualitatively similar to Fig. \ref{bcrit_compare1} for the vertically
+isothermal disk. For $\khat=O(10)$, there is a   
+precise cut-off thermal timescale close to that given by
+Eq. \ref{bcrit_noniso}. We found that very large wavenumbers, 
+$\khat > O(10^2)$, were needed to enable surface modes, and modes with
+such values of $\khat$ do not have a maximum cooling time. However,
+these small lengthscales may not be important in practice
+(see \S\ref{caveats_visc}). Thus, $\beta<\beta_\mathrm{crit,noniso}$ 
+provides an excellent criteria for instability.
 
 \begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/compare_modes_Gam1.084_kx100.ps}
-  \caption{Same as Fig. \ref{compare_modes_vnoniso_kx10} but with
-    $\khat=100$. In this case surface modes appear nearly parallel to
-    the vertical axis
-    (cf. Fig. \ref{compare_modes_vnoniso_kx10}). However, this
-    distinction becomes difficult for increasing $\beta$ as modes
-    overlap. 
-    % but
-    % overlap with body modes (those that are roughly horizontal when 
-    % $\beta=10^{-3}$) as $\beta$ is further increased. 
-    \label{compare_modes_vnoniso_kx100}}
-\end{figure}
-
-Fig. \ref{compare_eigenvz_vnoniso_kx100} show the most unstable modes
-with $\khat=100$ and $\beta=10^{-3},\, 10^{-2}$ and
-$0.1$. For $\beta=10^{-3}$ the most 
-unstable mode is an asymmetric surface mode \citepalias[also seen in Fig. 7
-of][]{barker15}. However, increasing the thermal timescale to
-$\beta=10^{-2}$ an anti-symmetric surface mode becomes
-dominant. Further increasing to $\beta=0.1$, the dominant mode has a
-mixed character between a surface mode and the fundamental
-mode. Notice as $\beta$ is increased the perturbation amplitude near
-the vertical boundaries become negligible. This is not surprsing since  
-the stabilizing effect of buoyancy is largest at the disk surface. 
-%formally diverges
-
-\begin{figure}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.73cm 0cm
-  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d001.ps}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.73cm 0cm
-  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d01.ps}
-  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
-  0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d1.ps}
-  \caption{Vertical velocity eigenfunction of the most unstable mode
-    with $\khat=100$ found in the vertically non-isothermal disk with
-    $\beta = 10^{-3}$ (top), $\beta=10^{-2}$ (middle) and $\beta=0.1$
-    (bottom). 
-    \label{compare_eigenvz_vnoniso_kx100}}
-\end{figure}
-
-Finally, in Fig. \ref{bcrit_compare2} we plot the maximum VSI growth
-rates in the vertically non-isothermal disk model. Based on 
-Eq. \ref{iso_vsi_cond}, we conjecture the critical 
-thermal timescale in the vertically non-isothermal disk is 
-given by $ h|s|/(\gamma - \Gamma)$. This 
-evaluates to 0.158, which is close to the cut-off thermal timescale
-for $\khat\in[5,30]$. Fig. \ref{bcrit_compare2} is qualitatively 
-similar to the fiducial disk model in
-Fig. \ref{bcrit_compare1}. Importantly, for $\khat=O(10)$ there is a
-precise cut-off thermal timescale, the value of which in fact provides a  
-characteristic upper limit for other $\khat$.
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{figures/gcorr_compare_vnoniso_maxrate} 
+  \includegraphics[scale=0.4415]{figures/gcorr_compare_vnoniso_maxrate2} 
   \caption{Maximum VSI growth rates in the vertically non-isothermal disk
     model as a function of the thermal relaxation timescale
-    $\beta$. The vertical dashed line is the critical thermal
+    $t_c=\beta\OmK^{-1}$. The vertical dashed line is the critical thermal
     timescale given by Eq. \ref{iso_vsi_cond}. A conjectured critical thermal
     timescale accounting for a non-isothermal background, given by 
-    $\beta_\mathrm{crit,noniso}= h|s|/(\gamma-\Gamma)=0.158$,  
+     $\beta_\mathrm{crit,noniso}= 0.425$ (see Eq. \ref{bcrit_noniso}),   
     is also shown as the vertical dotted line. 
@@
 
+The above calculation demonstrates that the effect of introducing
+finite cooling on the VSI operates very similarly in vertically
+isothermal disks with an imposed surface, and in vertically
+non-isothermal disks with a physical surface. This is despite the fact
+that the verticaly buoyancy frequency, $N_z$, formally diverges in
+latter case as $|z|\to H_s$.
+
 % We briefly consider vertically non-isothermal disks with 
@@
 % \cite{nelson13}. In this case we set $\zmax=0.99H_s$.
+
+% We do not find surface modes for $\khat=10$, but do find them for
+% sufficiently large $\khat$. We demonstrate this in  % In the vertically
+% % non-isothermal disk, surface modes develop for sufficiently large 
+% % $\khat$ \citep{barker15}. We demonstrate them in
+% Fig. \ref{compare_modes_vnoniso_kx100} which show eigenvalues for
+% $\khat=100$. While growth rates generally decrease with increasing 
+% $\beta$, we find eigenvalues can overlap, making it difficult to 
+% distinguish between different modes. However, since there 
+% is a well-defined surface, we may seek the  
+% most unstable mode without the boundary-dependent issue that arose 
+% with a vertically isothermal disk in the limit $\beta\to0$.
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+%   0cm]{figures/compare_modes_Gam1.084_kx100.ps}
+%   \caption{Same as Fig. \ref{compare_modes_vnoniso_kx10} but with
+%     $\khat=100$. In this case surface modes appear nearly parallel to
+%     the vertical axis
+%     (cf. Fig. \ref{compare_modes_vnoniso_kx10}). However, this
+%     distinction becomes difficult for increasing $\beta$ as modes
+%     overlap. 
+%     % but
+%     % overlap with body modes (those that are roughly horizontal when 
+%     % $\beta=10^{-3}$) as $\beta$ is further increased. 
+%     \label{compare_modes_vnoniso_kx100}}
+% \end{figure}
+
+% Fig. \ref{compare_eigenvz_vnoniso_kx100} show the most unstable modes
+% with $\khat=100$ and $\beta=10^{-3},\, 10^{-2}$ and
+% $0.1$. For $\beta=10^{-3}$ the most 
+% unstable mode is an asymmetric surface mode \citepalias[also seen in Fig. 7
+% of][]{barker15}. However, increasing the thermal timescale to
+% $\beta=10^{-2}$ an anti-symmetric surface mode becomes
+% dominant. Further increasing to $\beta=0.1$, the dominant mode has a
+% mixed character between a surface mode and the fundamental
+% mode. Notice as $\beta$ is increased the perturbation amplitude near
+% the vertical boundaries become negligible. This is not surprsing since  
+% the stabilizing effect of buoyancy is largest at the disk surface. 
+% %formally diverges
+
+% \begin{figure}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.73cm 0cm
+%   0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d001.ps}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 1.73cm 0cm
+%   0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d01.ps}
+%   \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
+%   0cm]{figures/eigenvectorvz_vnoniso_kx100_beta0d1.ps}
+%   \caption{Vertical velocity eigenfunction of the most unstable mode
+%     with $\khat=100$ found in the vertically non-isothermal disk with
+%     $\beta = 10^{-3}$ (top), $\beta=10^{-2}$ (middle) and $\beta=0.1$
+%     (bottom). 
+%     \label{compare_eigenvz_vnoniso_kx100}}
+% \end{figure}
 

### 2015-05-07 setup.tex

@@
 \begin{align}\label{kap2_def}
-\kappa^2 = \frac{1}{r^3}\frac{\p}{\p r}\left(r^4\Omega^2\right). 
+\kappa^2 \equiv \frac{1}{r^3}\frac{\p}{\p r}\left(r^4\Omega^2\right) = \frac{1}{r^3}\frac{\p j^2}{\p r}, 
 \end{align}
+where $j$ is the specific angular momentum.
 
@@
 The gradients of entropy, $S\equiv
-C_p\ln{P^{1/\gamma}/\rho}$, in our disk models are
+C_P\ln{P^{1/\gamma}/\rho}$, in our disk models are
 \begin{subequations}
 \begin{align}
-  &\frac{\p S}{\p r} = C_p\left[\frac{s}{\gamma r} +
+  &\frac{\p S}{\p r} = C_P\left[\frac{s}{\gamma r} +
     \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p
-       r}\right],\\
-  &\frac{\p S}{\p z} = C_p
-  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z} = {C_p g_z \over c_s^2} {\Gamma - \gamma \over \gamma},\label{dSdz}
+       r}\right], \label{dSdr}\\
+  &\frac{\p S}{\p z} = C_P
+  \left(\frac{\Gamma}{\gamma}-1\right)\frac{\p \ln{\rho}}{\p z} = {C_P g_z \over c_s^2} {\Gamma - \gamma \over \gamma},\label{dSdz}
 \end{align} 
 \end{subequations}
- where $C_p$ is the heat capacity at
+ where $C_P$ is the heat capacity at
 constant pressure.
@@
 \begin{align}
-  N_z^2 \equiv - C_p^{-1}g_z \frac{\p S}{\p z} = {g_z^2 \over c_s^2}{\gamma - \Gamma \over \gamma}. %c_s^2 \left(\frac{\gamma -
+  N_z^2 \equiv - C_P^{-1}g_z \frac{\p S}{\p z} = {g_z^2 \over c_s^2}{\gamma - \Gamma \over \gamma}. %c_s^2 \left(\frac{\gamma -
    %   \Gamma}{\gamma}\right)\left(\frac{\p\ln\rho}{\p z}\right)^2,  
@@
 \begin{align}
-  \kappa^2 - \frac{1}{\rho C_p}\nabla P \cdot \nabla S &> 0,\\
+  \kappa^2 - \frac{1}{\rho C_P}\nabla P \cdot \nabla S &> 0,\\
   -\frac{\p P}{\p z}\left(\frac{\p j^2}{\p r}\frac{\p S}{ \p z} -
@@
 \end{align}
-\citep{tassoul78}, where $j=r^2\Omega$ is the specific angular
-momentum and $\kappa = r^{-3} \p_r j^2 $ is the epicyclic frequency.  For rotationally-supported, thin disks %  pressure gradients are small 
+\citep{tassoul78}. %, where $j=r^2\Omega$ is the specific angular
+%momentum. %and $\kappa = r^{-3} \p_r j^2 $ is the epicyclic frequency.  
+For rotationally-supported, thin disks %  pressure gradients are small 
 % compared to rotation, so 

### 2015-05-08 summary.tex
starting to edit summary (not done yet)

@@
 \section{Summary and discussion}\label{summary}
+In this paper we study the vertical shear instability (VSI) with a focus on 
+the radiative cooling requirements.  In turn we address the viability of the VSI 
+as an angular momentum transport mechanism in protoplanetary disks (PPDs).
+
+Our linear axisymmetric analysis of the VSI 
+considers (uniquely to our knowledge) finite cooling times in a vertically global model.  
+Our main analytical finding, which we confirm numerically, is the critical cooling 
+timescale above which VSI growth is suppressed, Eqs.\ \ref{prelim_bcrit} and {iso_vsi_cond}.
+Short cooling times are needed to weaken the stabilizing effects of vertical buoyancy, thereby allowing 
+the free energy in vertical shear to drive instability.
+
+Our main focus is irradiated, vertically isothermal disks which have strong vertical buoyancy.
+The critical cooling time is thus short, shorter than the orbital time by a factor of the disk
+aspect ratio.  This finding is consistent with, and helps explain, the results of 3D numerical simulations \citepalias{N13}.
+
+In applying our results to PPDs, we pay particular attention to the transition from 
+optically thick radiative diffusion to optically thin Newtonian cooling.  The largest obstacle to VSI occurs in 
+high density inner disk, $\lesssim 1$ -- 5 AU, where radiative diffusion times are slow.  Shorter wavelength disturbances
+ cool faster, though only if they remain optically thick.  In the inner disk cooling can only be fast enough
+ if wavelengths are too short to drive significant transport.  The best hope for the VSI in inner disk is a low surface density.
+
+...
+
 In this paper we have studied the axisymmetric vertical shear 

### 2015-05-08 appendix.tex
new 6.4, merge with youdin's updates
cleaned up figs folder

@@
 \subsection{Maximum critical cooling time}\label{max_cool}   
-Here we show that, under appropriate conditions, the maximum critical 
+Here we show that, under appropriate conditions, the longest critical 
 cooling time is that for the $M=0$ or fundamental mode. This allows
@@
 for all $M\geq 0$ if $\theta\leq 1$, which imply 
-$\mathrm{max}(\beta_c)$ occurs at $M=0$. For fixed $\khat$ the
+$\mathrm{max}(\beta_c)$ occurs at $M=0$. This result may also be
+obtained by explicitly solving 
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} with $\theta$
+treated as a small parameter. For fixed $\khat$ the 
 condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
-This means that if $\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$, then
-all such modes will be stabilized.
+All all such modes are stabilized if
+$\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$.
 

### 2015-05-08 summary.tex
cleaned up figs folder

--- 2015-05-07 summary.tex
@@
 Perhaps due to these physical differences, CONO operates in different regimes of parameter
-space than the VSI.  CONO grows best for longer wavelengths $\khat \gtrsim 1$ and longer 
+space than the VSI.  CONO grows best for longer wavelengths $\khat \lesssim 1$ and longer 
 cooling times $\beta \sim 1$.
@@
 
-\subsubsection{Radiative transfer} 
+\subsection{Radiative transfer} 
 Our relatively simple treatment of cooling with an idealized dust opacity

### 2015-05-08 summary.tex
my compilation problem

--- 2015-05-08 summary.tex
@@
 Our main analytical finding, which we confirm numerically, is the critical cooling 
-timescale above which VSI growth is suppressed, Eqs.\ \ref{prelim_bcrit} and {iso_vsi_cond}.
+timescale above which VSI growth is suppressed, Eqs.\ \ref{prelim_bcrit} and \ref{iso_vsi_cond}.
 Short cooling times are needed to weaken the stabilizing effects of vertical buoyancy, thereby allowing 

### 2015-05-08 numerical.tex
figure compilation problems (fixed by commenting out for now)
re-shuffled 6.4, updated fig 2, poster files update

@@
 \subsection{Vertically non-isothermal disk model} 
-We briefly consider a vertically non-isothermal disk model with
-$(p,q, h)=(-0.5,-1,0.05)$ and $(\gamma, \Gamma)=(1.4,1.3)$. We set the
-vertical domain size close to the physical disk surface, $\zmax =
-0.99H_s\simeq 2.6H$, which is smaller than our standard domain
-size. In this case the vertical shear rate $\p_z\Omega\propto s$, the
-radial entropy gradient. Here, $s=-0.85$.
-
-From the expression for the
-critical cooling timescale, Eq. \ref{iso_vsi_cond}, it is reasonable
-to expect a similar criteria for vertically non-isothermal 
-disks in the form
+We briefly examine a vertically non-isothermal disk model with 
+a physical surface. Recall from Eq. \ref{vertical_shear_ex} that in
+general the vertical shear rate $\p_z\Omega\propto s$, the radial
+entropy gradient. From the expression for the critical cooling timescale,
+Eq. \ref{iso_vsi_cond}, it is reasonable to expect a similar criteria
+for vertically non-isothermal disks in the form 
 \begin{align}\label{bcrit_noniso}
  \beta_\mathrm{crit}\to\beta_\mathrm{crit, noniso} =
- \frac{h|s|}{\gamma - \Gamma},
+ \frac{h|s|}{\gamma - \Gamma}. 
 \end{align}
-and $\beta_\mathrm{crit,noniso}=0.425$ for the chosen parameters. This
-is $\sim 3$ times larger than our fiducial, vertically isothermal
-disk with $\beta_\mathrm{crit}=0.125$.
+
+Here, we consider a disk model with $(p,q, h)=(-0.5,-1,0.05)$ and $(\gamma,
+\Gamma)=(1.4,1.3)$. We set the vertical domain size close to the
+physical disk surface, $\zmax = 0.99H_s\simeq 2.6H$, which is smaller
+than our standard domain size. For this setup, $s=-0.85$ and
+$\beta_\mathrm{crit,noniso}=0.425$, which is $\sim 3$ times larger than
+our fiducial, vertically isothermal disk with $\beta_\mathrm{crit}=0.125$.
 
@@
 non-isothermal disks with a physical surface. This is despite the fact
-that the verticaly buoyancy frequency, $N_z$, formally diverges in
-latter case as $|z|\to H_s$.
+that the vertical buoyancy frequency, $N_z$, formally diverges as $|z|\to H_s$.
 

### 2015-05-08 paper.tex
summary done, please check.  added middle initial

@@
 \shorttitle{Vertical shear instability}
-\shortauthors{M-K. Lin, A. Youdin}
+\shortauthors{M-K.\ Lin, A.\ N.\ Youdin}
 
@@
 %to the Solar Nebula}  
-\author{Min-Kai Lin \& Andrew Youdin}
+\author{Min-Kai Lin \& Andrew N Youdin}
 \affil{Department of Astronomy and Steward Observatory, \\
@@
   It is difficult to understand how cold circumstellar disks accrete onto their central stars.  Due to 
-  lower levels of ionization, the magnetic transport mechanisms that
+  lower levels of ionization, magnetic transport mechanisms that
   operate in warmer disks are less effective.  
@@
 many problems in astrophysics, including star and planet formation  
-\citep[][]{lyndenbell74, armitage10}.  The turbulence associated with many transport mechanisms
+\citep[][]{armitage10}.  The turbulence associated with many transport mechanisms
 is particularly important for dust evolution and planetesimal formation \citep{yl07, chiang10}.
@@
 identified in the context of differentially rotating stars (\citealp{goldreich67}, hereafter \citetalias{goldreich67}, \citealp{fricke68}).   
-The VSI arises when vertical shear, i.e. \ a variation in orbital motion along the rotation axis, 
+The VSI arises when vertical shear, i.e.\ a variation in orbital motion along the rotation axis, 
 destabilizes inertial-gravity waves, which are oscillations 

### 2015-05-08 summary.tex
re-shuffled 6.4, updated fig 2, poster files update
summary done, please check.  added middle initial

@@
  (CONO, \citealp{klahr14, lyra14}).   Both instabilities rely on thermal relaxation to 
-overcome the stability of adiabatic perturbations, i.e.\ the Solberg-Hoiland
+overcome the stability of disks to adiabatic perturbations, i.e.\ the Solberg-Hoiland
 criteria (see \S\ref{solberg}).
@@
 
-The sign of $N_r^2$ depends on disk parameters.  To demonstrate that the required parameters 
-are somewhat extreme, we consider the midplane of vertically isothermal disks with 
+The sign of $N_r^2$ depends on disk parameters.  To demonstrate that the parameters needed for $N_r^2 <0$
+are somewhat extreme, or at least non-standard, we consider the midplane of vertically isothermal disks with 
 $\Sigma \propto r^n$ so that $n = p+q/2 + 3/2$. Using Eqs. \ref{dSdr} and \ref{Nr} with $\gamma = 1.4$,
-$N_r^2 < 0$ requires $n> 3(1/2 + q)$.  For standard disk temperature laws this requirement implies
+$N_r^2 < 0$ requires $n> 3(1/2 + q)$.  For standard disk temperature laws, this requirement implies
 a flat or rising surface density profile, e.g.\ $n = 3/14$ for $q = -3/7$ or $n = 0$ for $q = -1/2$.
-Thus CONO is most like to operate at special locations, like the outside edges of 
-disk gaps or holes and/or shadowed regions where $q < -1/2$.
+Since standard $\Sigma$ profiles decline with radius, CONO is most like to operate at 
+special locations, like the outside edges of disk gaps or holes and/or shadowed regions where $q < -1/2$.
 
 Perhaps due to these physical differences, CONO operates in different regimes of parameter
-space than the VSI.  CONO grows best for longer wavelengths $\khat \lesssim 1$ and longer 
+space than the VSI.  The CONO grows best for longer wavelengths $\khat \lesssim 1$ and longer 
 cooling times $\beta \sim 1$.
@@
 could certainly be generalized in future works.  In hotter disks, gas phase opacities
-must be considered.   In cold disks there are many choices for a dust opacity which 
-varies with grain sizes and compositions. Dust properties could change the viability 
-of the VSI for better or worse.  The radiative transfer itself
+must be considered.   In cold disks, there are many choices for the dust opacity, which 
+varies with grain sizes and compositions. Changing dust properties would alter 
+the viability of the VSI for better or worse.  The radiative transfer itself
 could be calculated with higher levels of sophistication, as is already being done in
-numerical simulations \citep{stoll14}.
+numerical simulations of the VSI \citep{stoll14}.
 
 \section{Summary and discussion}\label{summary}
-In this paper we study the vertical shear instability (VSI) with a focus on 
-the radiative cooling requirements.  In turn we address the viability of the VSI 
+In this paper we study the vertical shear instability (VSI) with a focus the 
+role of radiative cooling.  In turn we assess the viability of the VSI 
 as an angular momentum transport mechanism in protoplanetary disks (PPDs).
 
-Our linear axisymmetric analysis of the VSI 
+Our linear, axisymmetric analysis of the VSI 
 considers (uniquely to our knowledge) finite cooling times in a vertically global model.  
-Our main analytical finding, which we confirm numerically, is the critical cooling 
-timescale above which VSI growth is suppressed, Eqs.\ \ref{prelim_bcrit} and \ref{iso_vsi_cond}.
-Short cooling times are needed to weaken the stabilizing effects of vertical buoyancy, thereby allowing 
-the free energy in vertical shear to drive instability.
+In order for vertical shear to drive the VSI, short cooling times are needed to weaken 
+the stabilizing effects of vertical buoyancy. Our main analytical finding, which we confirm numerically, 
+is the critical cooling  timescale above which VSI growth is suppressed, Eqs.\ \ref{prelim_bcrit} and \ref{iso_vsi_cond}.
 
@@
 The critical cooling time is thus short, shorter than the orbital time by a factor of the disk
-aspect ratio.  This finding is consistent with, and helps explain, the results of 3D numerical simulations \citepalias{N13}.
+aspect ratio.  This finding is consistent with, and helps explain, the results of 3D numerical simulations \citepalias{nelson13}.
 
@@
 high density inner disk, $\lesssim 1$ -- 5 AU, where radiative diffusion times are slow.  Shorter wavelength disturbances
- cool faster, though only if they remain optically thick.  In the inner disk cooling can only be fast enough
- if wavelengths are too short to drive significant transport.  The best hope for the VSI in inner disk is a low surface density.
+ cool faster, so long as they remain optically thick.  In the inner disk, however, our VSI cooling criterion requires 
+ wavelengths that are too short to drive significant transport.  The best hope for the VSI in inner disk appears to 
+ be a low surface density, which speeds radiative diffusion by lengthening the photon mean free path.  This option naturally begs 
+ the question of what accretion mechanism lowered the surface density in the first place.
 
-...
+ In the outer disk, the VSI tends to cool in the optically thin limit.  The main issue is whether the opacity 
+ is high enough for optically thin cooling to be sufficiently fast.  Our standard opacity assumes a Solar abundances of small dust.
+ In this case, the VSI can operate from $\sim 5$ -- 100 AU.  A factor of 10 reduction in the opacity, for instance by locking
+ small dust into planetesimals and planetary cores \citep{youdin13}, makes cooling times too slow for VSI growth.  
+ In this case cooling is too slow not just in the outer disk, but into  $< 1$ AU.
 
-In this paper we have studied the axisymmetric vertical shear 
-instability (VSI) in astrophysical disks that arise from the 
-height dependence of the equilibrium angular velocity profile,  
-$\p_z\Omega\neq 0$. Vertical shear is generally
-expected in astrophysical disks due to baroclinicity ($\nabla
-P\times\nabla \rho \neq 0$). This may be due
-to a radial variation in the disk temperature, or more generally in
-the disk entropy. However, astrophysical disks are also typically
-sub-adiabatically stratified for stability against vertical
-convection ($N_z^2\geq 0$). In this case, the VSI requires a short
-thermal or cooling timescale $t_c$ in order to operate.
+ An enhanced opacity makes optically thin cooling faster and radiative diffusion slower.  This shift favors VSI growth 
+ in the outer disk at the expense of the inner disk, pushing the inner limit of VSI growth further out.
+ The standard choice -- corresponding to Solar abundances in a standard MMSN disk -- allows the VSI 
+ to grow over the widest range of relevant disk radii.
 
-Physically, a short thermal timescale is needed to reduce the  
-stabilizing effect of buoyancy forces. Because vertical shear is weak
-in thin disks, very rapid cooling, corresponding to $t_c\Omega\ll 1$, 
-is generally needed for the VSI, otherwise buoyancy is strongly
-stabilizing. Our primary goal was to quantify this requirement,
-thereby provide a simple way to assess whether or not the VSI is
-relevant in realistic disks.
+ Our detailed study of the spectrum of VSI modes confirms that some artificial `surface modes' are 
+ triggered by imposed vertical boundaries \citepalias{nelson13, barker15}.  Domain size
+ convergence studies are thus essential.  Fortunately, our results show that longer cooling times 
+  stifle the growth of surface modes.  Thus, at least in some cases, more realistic radiative transfer also produces more reliable dynamics.
 
-To do so, we generalized previous vertically global, radially local linear 
-calculations of the VSI \citepalias{nelson13,mcnally14,barker15}, which
-assumed isothermal perturbations, to include an energy equation with 
-a source term in that relaxes the disk 
-temperature to its initial value on a timescale
-$t_c$. 
-% This `thermal relaxation' model allows
-% the thermodynamic response of the disk to be controlled by varying
-% $t_c$, and has been employed in previous numerical simulations of the
-% VSI \citepalias{nelson13}.  
-%In accordance with the above studies, which show that the VSI occurs for 
-%perturbations with small radial lengthscales, we applied the 
-%radially local approximation in the linearized fluid equations. 
-We mostly focused on vertically isothermal disks as this is the
-relevant case for protoplanetary disks (PPDs) passively 
-irradiated by its central star \citep{chiang97}. It also permits a
-simplified analytical discussion by extending the 
-linear models considered in \cite{lubow93} by adding vertical shear
-and (rapid) cooling. We obtained explicit solutions 
-describing the VSI with thermal relaxation in the thin disk limit.
-
-For a vertically isothermal, sub-adiabatically stratified disk we
-determined the thermal timescale   
-\begin{align*}
-  t_\mathrm{crit} = \frac{ h |q|}{\gamma-1}\OmK^{-1}, 
-\end{align*}
-below which the VSI operates effectively. 
-%While $t_\mathrm{crit}$ was
-%derived for the lowest order VSI mode --- the `fundamental  
-%corrugation mode' --- we find from numerical calculations that
-%$t_\mathrm{crit}$ can be used to characterize the general
-%stabilization of the VSI by a finite thermal timescale. %That is, all
-%VSI growth rates are reduced significantly as $\beta$ is increased
-%from 
-%zero to $\beta_\mathrm{crit}$.    
-We have thus provided a quantitative measure for the thermal
-timescale requirement for the VSI in astrophysical 
-disks. The thermal timescale must be shorter than the disk
-dynamical timescale by $O( h)$, which is 
-quite short for thin disks ($ h\ll 1$), because vertical shear is weak.
-
-We applied our linear models to assess the applicability of the VSI in
-PPDs. We considered the Minimum Mass Solar Nebula 
-(MMSN) described in \cite{chiang10} and   
-modeled the thermal timescale $t_c$ based on realistic
-disk properties and dust opacity. 
-%This results in $t_c$ being a
-% function of position and the perturbation lengthscale of interest. 
-We find the MMSN is unstable to the VSI 
-from $\sim 5$AU to $\sim50$AU with characteristic growth times of
-$\sim 30$ orbits. It is possible to have the VSI at smaller
-radii, but only on radial lengthscales signficantly shorter
-than the disk scaleheight. %However, such small-scale perturbations may 
-%be subject to viscous decay in a real disk.
-
-Although $t_c$ is not vertically constant in the 
-MMSN as assumed in our analysis, we find that evaluating the condition
-$t_c< t_\mathrm{crit}$ at the disk midplane gives an indication of the
-radii at which the VSI operates in the MMSN. This simple criterion is
-thus a practical first step in estimating the relevance of the VSI in
-realistic astrophysical disks without resorting to linear
-calculations or non-linear numerical simulations.
+ The VSI deserves further study as a viable mechanism to drive at least low levels of accretion in cold disks.

### 2015-05-08 paper.tex
edits to 1 and 2.  tightened up the summary in the intro.  and the cooling time discusion also improved with outer stability limit now mentioned.

@@
 destabilizes inertial-gravity waves, which are oscillations 
-with rotation and buoyancy as restoring forces. Vertical shear
+with rotation and buoyancy as restoring forces. 
+% \citep{lubow93}. 
+Vertical shear
 occurs wherever the disk is baroclinic, i.e.\ when constant 
@@
 (\citetalias{goldreich67}; \citetalias{nelson13}). Rapid radiative 
-cooling, equivalently a short thermal relaxation timescale, brings a
+cooling, i.e.\ a short thermal relaxation timescale, brings a
 moving fluid element into thermal equilibrium with 
@@
                                 %quickly 
-An isothermal equation of state implies instantaneous cooling (and
-heating), and is the ideal context for studying the VSI
+An isothermal equation of state implies instantaneous thermal relaxation, and is the ideal context for studying the VSI
 (\citealp{urpin03}; \citetalias{nelson13}; \citealp{mcnally14}, hereafter \citetalias{mcnally14}; \citealp{barker15},
@@
 stably stratified in the outer regions, beyond $\sim 1$--5 AU, where
-heating is  irradiation dominated \citep{bitsch15}.  In the inner
+heating is  irradiation dominated \citep{bitsch15}.  Even in the inner
 disk, strong vertical buoyancy is possible if accretion heating is
@@
 
-A full understanding of the VSI in real disks therefore requires
+Understanding the VSI in real disks therefore requires
 considering finite, non-zero cooling times, $t_{\rm c}$. Non-linear
@@
 turbulence in stably stratified disks requires rapid cooling  
-with $t_{\rm c} \lesssim 0.01P_{\mathrm{orb}}$, where $P_\mathrm{orb}$ 
-%=2\pi/\Omega$ 
-is the local orbital period %and $\Omega$ is the orbital
-%frequency 
- \citepalias{nelson13}. \citetalias{nelson13} further find
-that the VSI can drive moderately strong transport, with Reynolds
-stresses of $\alpha \sim 10^{-3}$ times the mean pressure. Simulations
-with realistic radiative transfer, in lieu of a parameterized $t_{\rm
+with $t_{\rm c}$ shorter than orbital timescales
+ \citepalias{nelson13}. When the cooling time is short enough, 
+ the VSI can drive moderately strong transport, with Reynolds
+stresses of $\alpha \sim 10^{-3}$ times the mean pressure \citepalias{nelson13}. Simulations
+with realistic radiative transfer, in lieu of a fixed $t_{\rm
   c}$, find that the VSI in irradiated disks drive transport with
@@
 
-To reach hydrodynamic turbulence via the VSI, a disk should  
-first develop the linear instability. A quantitative 
-examination of the thermodynamic requirement for the linear VSI is
-thus essential in assessing its 
-relevance to PPDs. In this work, we present a
-comprehensive study of the linear VSI in vertically global disk models 
-with non-zero cooling timescales, which permit buoyancy forces. 
-% In this work, we present a comprehensive study of the linear VSI in vertically global disk models 
-% with non-zero cooling timescales $t_c > 0$, which permit buoyancy forces. 
-% Our effort generalizes both early analyses of the VSI in vertically
-% localized disk models \citep{urpin98,urpin03} and recent calculations in vertically global 
-% disks \citepalias{nelson13,mcnally14,barker15}. These works either
-% consider the limit of instantaneous cooling ($t_c\to0$) or neutrally
-% stratified disks, for which buoyancy forces play no role. 
-Our effort generalizes previous studies of the linear VSI in 
-vertically local \citep{urpin98,urpin03} and vertically global 
-\citepalias{nelson13,mcnally14,barker15} disk models that consider the limit
-of instantaneous cooling, for which buoyancy plays no role.  
-%to study of how finite cooling times, and thus buoyancy, 
-%affect the VSI. Our effort may also be regarded
-%as a extension of early analyses of the VSI in 
-%vertically localized disk models with cooling \citep{urpin98,urpin03} to
-%vertically global disks with cooling. 
-We consider disk  
-models with vertically constant $t_{\rm c}$, as well as more realistic
-cooling timescales based on dust opacity in both the optically thick
-(diffusive) and thin regimes, where appropriate. Coupled with the 
-results from numerical simulations, our results can be used to 
-understand how the VSI operates over a wide range of model PPDs.
+Studying the linear growth of the VSI is necessary for 
+understanding how it can ultimately drive turbulent transport.
+The pioneering linear analyses of the VSI considered vertically 
+local disturbances \citep{urpin98,urpin03}.  A vertical global analysis 
+(e.g.\ \citetalias{nelson13,mcnally14,barker15}) 
+is essential for understanding how vertically elongated disturbances 
+interact with the disk's vertical structure.  Moreover a vertically global analysis
+allows more direct comparisons with modern numerical simulations.  This work 
+generalizes previous vertically global analyses by including a finite cooling time.
 
-%\emph{TODO: the following is still a bit too wordy IMO but we can redo as a final step after these sections are finalized.}
+%NOTE: no need to reexplain importance of cooling or to say more about our work before the "organized as follows" summary
 
-This paper is organized as follows. In \S\ref{vsi_require} we 
-introduce and qualitatively explain our main result of a critical or
-characteristic cooling timescale, beyond which the VSI is either
-suppressed or inefficient. We then present the  
-formal linear problem in \S\ref{setup}---\ref{numerical}.    
-In \S\ref{setup} we list the governing equations and global
-equilibrium disk models under consideration, and formulate the linear
-stability problem in \S\ref{linear}. In 
-\S\ref{analytical} we give a simplified analytical discussion for
-vertically isothermal disks, where we derive the aforementioned
-characteristic upper limit on the cooling or thermal timescale for the VSI. 
-% Our treatment is grossly simplified owing to a
-% number of approximations made, but closely parallels previous work, 
-% allowing us to highlight the essential effects of vertical shear, and to
-% derive a characteristic upper limit to the thermal timescale for the
-% VSI. This timescale is \emph{much shorter} than the dynamical
-% timescale. 
-In \S\ref{numerical} we numerically
-solve the linear problem, and  
-find basic agreement with our analysis.
-% , including
-% the aforementioned characteristic thermal timescale. 
-In 
-\S\ref{application} we apply our results to the Minimum Mass
-Solar Nebula (MMSN) by modeling cooling timescales based on realistic
-disk properties, and find that the VSI can operate effectively in the
-outer parts of the MMSN. We summarize in \S\ref{summary} with a
-discussion of caveats and possible directions for future work.
+This paper is organized as follows. In \S\ref{vsi_require} we explain, without derivation,  
+our main result for the critical cooling timescale, beyond which VSI growth is 
+suppressed. %We then present the  
+%formal linear problem in \S\ref{setup}---\ref{numerical}.    
+ We develop our disk model in \S\ref{setup} and derive linearized
+ perturbation equations in  \S\ref{linear}.
+ Section \ref{analytical} contains our main analytic results, leading to 
+ the derivation of the critical cooling time in \S\ref{iso_vsi_beta_crit}.
+ In \S\ref{numerical} we analyze linear VSI growth, and numerically confirm the critical cooling time.
+We apply our results to PPDs in \S\ref{application}, with cooling times derived from dust opacities.
+We discuss caveats and extensions in \S\ref{caveats}.
+ We conlude in \S\ref{summary} with a summary.  Some technical and background 
+ developments are explained in the appendices.
 
@@
 In a stably stratified thin disk, the VSI  
-requires a thermal timescale $t_c$ much shorter than the
+requires a thermal timescale $t_c$ significantly shorter than the
 disk dynamical timescale, 
 \begin{align}
-  t _c \ll \OmK^{-1}. 
+  t _c \ll \OmK^{-1},
 \end{align}
@@
 This destabilizing shear competes with the stabilizing vertical
-buoyancy frequency $N_z$.  In a thin disk,     
+buoyancy frequency, $N_z$.  In a thin disk,     
 \begin{align}\label{nz_thin}
@@
 frequency (as well as the gas density) 
-in a fiducial PPD model considered in \S\ref{application}. 
+in a fiducial PPD model, see \S\ref{toy_relax} for details. 
 %, whose parameters are in Appendix \ref{mmsn} . %Here $\rho_0$
@@
 
-How short must $t_c$ be for vertical shear to prevail?  We start by 
-considering the Richardson number $\mathrm{Ri} = N_z^2/(r\p_z\Omega)^2$, a 
+How short must $t_c$ be for vertical shear to prevail?  We start with 
+ the Richardson number $\mathrm{Ri} = N_z^2/(r\p_z\Omega)^2$, a 
 ratio of buoyant to shear energies.  Though not precisely applicable to our problem,
@@
 expected when this inequality fails).  With this reduction, the Richardson-like criterion for 
-shear \emph{instability} becomes
+shear instability becomes  \citep{urpin03}
 \begin{align}\label{bcrit_est}
-  t_c \lesssim \frac{\left|r\p\Omega/\p z\right|} {N_z^2}, 
+  t_c \lesssim \frac{\left|r\p\Omega/\p z\right|} {N_z^2}.
 \end{align}
-which is Eq.  47 of \cite{urpin03}. %This result is far from rigorous, but it contains some useful insights.
+ %This result is far from rigorous, but it contains some useful insights.
 If we interpret Eq.\ \ref{bcrit_est} as a local criterion, then we see that 
@@
 
-For vertically global disturbances considered in this work, 
-% To roughly obtain a global cooling time criterion, we evaluate Eq.\ \ref{bcrit_est}
-% at $z \sim H$.  To get the precise form we want, we choose $z = \gamma H / 2$ to 
-we obtain the cooling requirement 
+We are more interested, however, the the global instability criterion for disturbances at all heights.
+Our desired result 
+is the cooling requirement 
 \begin{align}\label{prelim_bcrit}
@@
 \end{align}   
-for the VSI to operate effectively in a vertically isothermal, thin
-disk. We derive this result more formally in
-\S\ref{iso_vsi_beta_crit}, but notice that Eq. \ref{prelim_bcrit}
-may be obtained from Eq. \ref{bcrit_est} by evaulating it at the
-height $|z|=\gamma H/2$. %i.e., at $z=O(H)$
+for VSI growth that we derive in \S\ref{iso_vsi_beta_crit}.  We can simply 
+(but non-rigourously) obtain Eq. \ref{prelim_bcrit}
+  by evaulating  Eq. \ref{bcrit_est} at $|z|=\gamma H/2$. %i.e., at $z=O(H)$
 
@@
 are short enough to support VSI growth.  Fig. \ref{bcrit_mmsn2d} plots the
- thermal timescales in our fiducial PPD model with a small grain dust opacity. 
- Since cooling by radiative diffusion depends on a perturbation's lengthscale, we
- plot $t_c$ for two values of the radial wavenumber,  $k_x$,  the relevant 
- component for VSI modes that are radially narrow.  The red line
-marks the boundary implied by Eq. \ref{prelim_bcrit}.  We caution that 
- Eq. \ref{prelim_bcrit} applies when $t_c$ is vertically uniform.  This section is 
- aimed at basic insights; while more realistic spatial variations in
- $t_c$ is considered in \S\ref{application}. 
- %analyzes the VSI with realistic spatial variations in.
+ thermal timescales in our fiducial PPD model, see \S\ref{toy_relax} for details.
+ The regions between the red lines satisfy the cooling requirement of Eq.\ \ref{prelim_bcrit}.
+ Even without a detailed analysis we can expect VSI growth between $\sim 10$ -- 150 AU.
+ The inner disk is complicated by the fact that the optically thick midplane has longer cooling times. 
+ Thus the cooling criterion fails in the midplane but is satisfied above.  This case requires the 
+ more detailed analysis of \S\ref{application}.
 
- Due to the smaller values of $t_c$ in the outer disk, we generally expect the VSI to 
- operate outside $\sim 10$ AU.  This conclusion changes however, if
- the dust opacity is reduced,  
-slowing down cooling, as we show in \S\ref{application}. % (e.g.\ Fig.\
-% \ref{mmsn_bcrit_bcool}). 
-In the inner disk, thermal times are long in the midplane, but short at higher altitutdes. 
-Again, this structure necessitates the more detailed analyses that follow.  Nevertheless, we can already 
-see that due to radiative diffusion higher wavenumbers can cool more rapidly.  Thus smaller scale disturbances 
-may trigger VSI at smaller disk radii.
+  Fig. \ref{bcrit_mmsn2d} also compares two different radial wavenumbers, 
+  $k_x = 10/H$ and $50/H$.  In the inner disk, the higher wavenumber mode cools faster, due to radiative diffusion.
+  Thus in the inner disk, shorter wavelength VSI modes are more like to grow. In the outer disk, the cooling time is the 
+  same for different $k_x$ and $z$, because optically thin cooling is independent of both lengthscale and density.
+  Thus the outer barrier to growth, beyond $\sim 150$ AU in this model, is independent of  wavelength. 
+This example highlights the fact that optically thin cooling sets a lower limit to the cooling time; a fact that is sometimes obscured when 
+the radiative diffusion approximation is made at the outset.
 
-On the other hand, one cannot obtain arbitrarily short cooling times
-by invoking arbitrarily small scales (cf. \citealt{urpin03}, 
-\citetalias{barker15}), because very small wavelengths are in the
-optically thin cooling regime,  where cooling is independent of
-lengthscale. Furthermore, 
-%while we do not include viscosity in our
-%idealized disk models, in reality 
-small-scale perturbations will be 
-prone to viscous damping, including the effective viscosity from turbulent motions.
-While our linear, inviscid calculations do not address this issue directly, we expect that 
-shorter wavelength modes will drive weaker non-linear (and non-axisymmetric) transport.
-
-% However, we cannot and do not find that VSI will always be relevant for
-% sufficiently small perturbations because \begin{inparaenum}[(i)] \item very small wavelengths are in the optically 
-% thin cooling regime,  where cooling is independent of lengthscale;
-% and \item small scale perturbations 
-% are more prone to viscous damping, and should drive weaker transport.
-% \end{inparaenum}
-
-We now proceed with our detailed analysis of the VSI to arrive
-at Eq. \ref{prelim_bcrit} and verify it numerically. The interested 
-reader may jump to \S\ref{application} for the application of our
-results to PPDs.
+% Since cooling by radiative diffusion depends on a perturbation's lengthscale, we
+% plot $t_c$ for two values of the radial wavenumber,  $k_x$,  the relevant 
+% component for VSI modes that are radially narrow.  The red line
+%marks the boundary implied by Eq. \ref{prelim_bcrit}.  We caution that 
+% Eq. \ref{prelim_bcrit} applies when $t_c$ is vertically uniform.  This section is 
+% aimed at basic insights; while more realistic spatial variations in
+% $t_c$ are considered in \S\ref{application}. 
+% %analyzes the VSI with realistic spatial variations in.
+% 
+% Due to the smaller values of $t_c$ in the outer disk, we generally expect the VSI to 
+% operate outside $\sim 10$ AU.  This conclusion changes however, if
+% the dust opacity is reduced,  
+%slowing down cooling, as we show in \S\ref{application}. % (e.g.\ Fig.\
+%% \ref{mmsn_bcrit_bcool}). 
+%In the inner disk, thermal times are long in the midplane, but short at higher altitutdes. 
+%Again, this structure necessitates the more detailed analyses that follow.  Nevertheless, we can already 
+%see that due to radiative diffusion higher wavenumbers can cool more rapidly.  Thus smaller scale disturbances 
+%may trigger VSI at smaller disk radii. 
+%
+%On the other hand, one cannot obtain arbitrarily short cooling times
+%by invoking arbitrarily small scales (cf. \citealt{urpin03}, 
+%\citetalias{barker15}), because very small wavelengths are in the
+%optically thin cooling regime,  where cooling is independent of
+%lengthscale. Furthermore, 
+%%while we do not include viscosity in our
+%%idealized disk models, in reality 
+%small-scale perturbations will be 
+%prone to viscous damping, including the effective viscosity from turbulent motions.
+%While our linear, inviscid calculations do not address this issue directly, we expect that 
+%shorter wavelength modes will drive weaker non-linear (and non-axisymmetric) transport. 
+%
+%% However, we cannot and do not find that VSI will always be relevant for
+%% sufficiently small perturbations because \begin{inparaenum}[(i)] \item very small wavelengths are in the optically 
+%% thin cooling regime,  where cooling is independent of lengthscale;
+%% and \item small scale perturbations 
+%% are more prone to viscous damping, and should drive weaker transport.
+%% \end{inparaenum} 
+%
+%We now proceed with our detailed analysis of the VSI to arrive
+%at Eq. \ref{prelim_bcrit} and verify it numerically. The interested 
+%reader may jump to \S\ref{application} for the application of our
+%results to PPDs.
 

### 2015-05-08 linear_iso.tex
further refinements to appendix D2, D3. plus minor notation update
mkl going through entire manucript, merge with youdin's updates

@@
 Similar to the incompressible \citepalias{goldreich67} or anelastic \citepalias{nelson13,barker15} approximations,
-the low frequency approximation filters p-modes in favor of g-modes, i.e.\ inertial
-gravity waves \citep{lubow93}.
+the low frequency approximation filters acoustic waves in favor of
+inertial-gravity waves \citep{lubow93}.
 
@@
 In particular we point out that the fully-radially-local approximation is only valid for
-short cooling times ($\beta \lesssim 1$).
+short cooling times, $\beta \lesssim O(1)$.
 

### 2015-05-08 numerical.tex
mkl going through entire manucript, merge with youdin's updates

@@
 qualitatively similar to Fig. \ref{compare_modes_cool_kx10}. As before
-we find as $\beta$ is increased, higher order body modes are rapidly
+we find as $\beta$ is increased, higher order modes are rapidly
 stabilized, eventually only the fundamental mode can operate before

### 2015-05-08 paper.tex
mkl going through entire manucript, merge with youdin's updates

--- 2015-05-08 paper.tex
@@
 %to the Solar Nebula}  
-\author{Min-Kai Lin \& Andrew N Youdin}
+\author{Min-Kai Lin\altaffilmark{$\dagger$} \& Andrew N Youdin}
 \affil{Department of Astronomy and Steward Observatory, \\
@@
   Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
+\altaffiltext{$\dagger$}{Steward Theory Fellow}
 \email{minkailin@email.arizona.edu, youdin@email.arizona.edu}
@@
 \begin{align}\label{prelim_bcrit}
-  t_c <  h |q| \OmK^{-1}/ (\gamma - 1)
+  t_c <  &\underbrace{\frac{h |q| }{\gamma - 1}}_{\beta_\mathrm{crit}}\OmK^{-1}
+ % &\equiv \beta_\mathrm{crit}\OmK^{-1},\notag
 \end{align}   
@@
     PPD model, the Minimum Mass Solar Nebula. The red
-    line corresponds to a characteristic thermal timescale, derived in
+    lines corresponds to a maximum thermal timescale, derived in
     this work, below which the VSI is expected to operate effectively
-    (the region outside the line).    
+    (the region bounded by the red lines).     
     \label{bcrit_mmsn2d} 
@@
 plot $t_c$ for two values of the radial wavenumber,  $k_x$,  the relevant 
- component for VSI modes that are radially narrow.  The red line
-marks the boundary implied by Eq. \ref{prelim_bcrit}.  We caution that 
+component for VSI modes that are radially narrow. In the
+optically-thin cooling regime, $t_c$ is independent of lengthscales. 
+The red lines mark the boundaries implied by
+Eq. \ref{prelim_bcrit}.  We caution that 
 Eq. \ref{prelim_bcrit} applies when $t_c$ is vertically uniform.  This section is 
@@
 $t_c$ is considered in \S\ref{application}.
- %analyzes the VSI with realistic spatial variations in.
 
 Due to the smaller values of $t_c$ in the outer disk, we generally expect the VSI to 
- operate outside $\sim 10$ AU.  This conclusion changes however, if
+operate outside $\sim 10$AU, but inside $\sim 150$AU since $t_c$
+exceeds the critical value beyond this radius. This conclusion changes however, if
 the dust opacity is reduced,  
@@
 \citetalias{barker15}), because very small wavelengths are in the
-optically thin cooling regime,  where cooling is independent of
-lengthscale. Furthermore, 
+optically thin cooling regime and there exists a minimum thermal
+timescale. Furthermore, 
 %while we do not include viscosity in our

### 2015-05-08 setup.tex
mkl going through entire manucript, merge with youdin's updates

@@
 \begin{align}\label{thermal_relax}
-  \Lambda  \equiv \frac{\rho\mathcal{R}}{\mu}\frac{T - T_0}{t_c},
+  \Lambda  \equiv \frac{\rho\mathcal{R}}{\mu}\frac{T - T_\mathrm{eq}}{t_c},
 \end{align}
-where $T_0=T_0(r,z)$ is the equilibrium temperature.  We 
+where $T_\mathrm{eq}=T_\mathrm{eq}(r,z)$ is the equilibrium temperature.  We 
 define the dimensionless cooling time 
@@
 where $r_0$ is a fiducial radius and $p$ and $q$ are the standard powerlaw indices
-for midplane density and temperature, respectively.  The polytropic index $\Gamma$ 
+for midplane density and temperature, respectively. We consider $q<0$
+without loss of generality.  The polytropic index $\Gamma$ 
 parametrizes the vertical stratification with $\Gamma = 1$ ($\Gamma = \gamma$)
@@
 The vertically isothermal case, $\Gamma = 1$, can be calculated either 
-as a special case or as the $\Gamma \rightarrow 1$ limit of Eq. \ref{eqm_dens},    
+as a special case or by taking the limit of Eq. \ref{eqm_dens},    
 \begin{subequations}\begin{align}
@@
 
-A more general expression for vertical shear holds for vertical polytopes 
+A more general expression for vertical shear holds for vertical polytropes 
 (which satisfy Eq.\ \ref{eqm_press}), with no assumed radial structure:
@@
 \subsection{Entropy gradients and vertical buoyancy}
-The gradients of entropy, $S\equiv
+The gradients of specific entropy, $S\equiv
 C_P\ln{P^{1/\gamma}/\rho}$, in our disk models are
@@
 \end{align} 
-implies stability. For typical model parameters the right hand
-side (RHS) is $O( h^2)$. Hence, instability in the adiabatic limit  
-requires $\gamma < \Gamma + O( h^2)$. Since we consider
-$ h\ll1$, this implies the equilibrium disk must have a small
-entropy gradient in order to be unstable to axisymmetric adiabatic
-perturbations. This is generally not true for passively irradiated, stably
+implies stability. For typical model parameters, 
+Eq. \ref{solberg2} indicates that axisymmetric instability in the adiabatic limit  
+requires $\gamma < \Gamma + O( h^2)$, corresponding to a small entropy 
+gradient.  
+%This implies the equilibrium
+%disk must have a small 
+%entropy gradient in order to be unstable to axisymmetric adiabatic
+%perturbations. 
+This is generally not true for passively irradiated, stably
 stratified disks \citep{chiang97}, in which the VSI requires rapid
@@
  standard two step process to obtain linearized equations in the radially local 
- approximation (e.g., \citealp{goldreich67}, who also consider vertically local perturbations).
+ approximation (e.g., \citealp{goldreich67}, who also consider vertically localized perturbations).
 

### 2015-05-08 summary.tex
mkl going through entire manucript, merge with youdin's updates

@@
 \section{Summary and discussion}\label{summary}
-In this paper we study the vertical shear instability (VSI) with a focus the 
+In this paper, we have studied the vertical shear instability (VSI) with a focus the 
 role of radiative cooling.  In turn we assess the viability of the VSI 
@@
 optically thick radiative diffusion to optically thin Newtonian cooling.  The largest obstacle to VSI occurs in 
-high density inner disk, $\lesssim 1$ -- 5 AU, where radiative diffusion times are slow.  Shorter wavelength disturbances
+high density inner disk, $\lesssim 1$ -- 5AU, where radiative diffusion times are slow.  Shorter wavelength disturbances
  cool faster, so long as they remain optically thick.  In the inner disk, however, our VSI cooling criterion requires 
@@
  is high enough for optically thin cooling to be sufficiently fast.  Our standard opacity assumes a Solar abundances of small dust.
- In this case, the VSI can operate from $\sim 5$ -- 100 AU.  A factor of 10 reduction in the opacity, for instance by locking
+ In this case, the VSI can operate from $\sim 5$ -- 100AU.  
+A factor of 10 reduction in the opacity, for instance by locking
  small dust into planetesimals and planetary cores \citep{youdin13}, makes cooling times too slow for VSI growth.  
@@
  in the outer disk at the expense of the inner disk, pushing the inner limit of VSI growth further out.
- The standard choice -- corresponding to Solar abundances in a standard MMSN disk -- allows the VSI 
+ The standard choice --- corresponding to Solar abundances in a standard MMSN disk --- allows the VSI 
  to grow over the widest range of relevant disk radii.

### 2015-05-08 linear_adia.tex

@@
 
-\subsection{An upper limit to the cooling time}\label{iso_vsi_beta_crit}
+\subsection{The critical cooling time derived}\label{iso_vsi_beta_crit}
 Here we estimate the maximum thermal relaxation timescale 
@@
 for any $M$. In Appendix \ref{max_cool}, we find that if
-$|hq\khat|\leq 1$, then $\beta_c$ decreases with increasing $M$.   
+$|hq\khat|\leq 1$, so that marginal stability exists, then $\beta_c$
+decreases with increasing $M$.    
 %\begin{align}

### 2015-05-08 paper.tex

--- 2015-05-08 paper.tex
@@
 %results to PPDs.     
+>>>>>>> 3c00bb4e2b076179e280f2c07728086a110764ef
 

### 2015-05-08 setup.tex
minor to 3.4

--- 2015-05-07 setup.tex
@@
 if both Solberg-Hoiland criteria  are satisfied:
-\begin{align}
+\begin{subequations}\begin{align}
   \kappa^2 - \frac{1}{\rho C_P}\nabla P \cdot \nabla S &> 0,\\
@@
     \frac{\p j^2}{\p z}\frac{\p S}{\p r} \right)&>0. \label{second_SH} 
-\end{align}
+\end{align}\end{subequations}
 \citep{tassoul78}. %, where $j=r^2\Omega$ is the specific angular
@@
 % compared to rotation, so 
-the first criterion is generally satisfied. 
-Thus, to assess stability, we consider the second
+the first criterion is easily satisfied. 
+Thus, we consider the second
 criterion. 
-Without loss of generality, we evaluate Eq. \ref{second_SH} in $z>0$
-where $\p_zP<0$. Here we approximate $\p_r j^2\simeq r^3\OmK^2$ and
-use Eq. \ref{vertical_shear} for $\p_zj^2$. Then 
+Without loss of generality (since Eq. \ref{second_SH} is even in $z$), we consider $z>0$
+so that $\p_zP<0$. With $\p_r j^2\simeq r^3\OmK^2$ and
+using Eq. \ref{vertical_shear} for $\p_zj^2$  we find that
 \begin{align}\label{solberg2}
@@
 \end{align} 
-implies stability. For typical model parameters the right hand
-side (RHS) is $O( h^2)$. Hence, instability in the adiabatic limit  
-requires $\gamma < \Gamma + O( h^2)$. Since we consider
-$ h\ll1$, this implies the equilibrium disk must have a small
-entropy gradient in order to be unstable to axisymmetric adiabatic
-perturbations. This is generally not true for passively irradiated, stably
-stratified disks \citep{chiang97}, in which the VSI requires rapid
-cooling, as shown below.
+implies stability. For typical model parameters, the right hand
+side (RHS) is $O( h^2) \ll 1$.
+The left hand side is positive and order unity ($O(0.1)$ at least) if the disk is convectively stable, and not very close to convective instability.
+This stable stratification is expected for irradiated disks  \citep{chiang97}.
+Thus adiabatic disturbances are stable to a disk's vertical shear, explaining why the VSI requires cooling.
 

### 2015-05-08 appendix.tex
re-shuffled 6.4, updated fig 2, poster files update

@@
 obtained by explicitly solving 
-Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} with $\theta$
-treated as a small parameter. For fixed $\khat$ the 
-condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
-All all such modes are stabilized if
+Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} with $\theta$ as a small parameter. 
+For fixed $\khat$ the condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
+All such modes are stabilized if
 $\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$.
@@
 for $\beta_\mathrm{crit}$ from the dispersion relations 
-Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} without  
+Eq. \ref{relax_cond1}---\ref{relax_cond2} without  
 assuming it is much less than unity from the onset or place

### 2015-05-08 summary.tex
minor edits 2,3,4
minor to 3.4

@@
-\section{Caveats and Neglected Effects} 
+\section{Caveats and Neglected Effects}\label{caveats} 
 %We highlight below some outstanding issues that should be addressed in
@@
 \section{Summary and discussion}\label{summary}
-In this paper, we have studied the vertical shear instability (VSI) with a focus the 
+In this paper we study the vertical shear instability (VSI) with a focus the 
 role of radiative cooling.  In turn we assess the viability of the VSI 

### 2015-05-08 paper.tex
minor

@@
 where $H=hr$ is the characteristic disk scale height.
-%for several disk scale heights, $H = hr$. 
-%As noted in
-%\S\ref{intro} this vertical variation is a source of free energy and
-%thus vertical shear ($\p_z\Omega \neq0$) is destabilizing. 
-%
-%For $\gamma>1$ the disk is stably stratified with a real 
+
 This destabilizing shear competes with the stabilizing vertical
@@
 in a fiducial PPD model, see \S\ref{toy_relax} for details.
-%, whose parameters are in Appendix \ref{mmsn} . %Here $\rho_0$
-%is the midplane density and vertical shear rate and buoyancy
-%frequencies are normalized by the local Keplerian frequency. 
+
 Vertical shear is 
 generally weak compared to buoyancy, suggesting stability. In fact, 
-without any cooling, the Solberg-Hoiland criteria \citep{tassoul78}  
+without any cooling, the Solberg-Hoiland criteria %\citep{tassoul78}  Tassoul cited in S-H section is enough
 confirms that vertical buoyancy is strong
@@
 thermal fluctuations decay which, combined with pressure equilibration, reduces the effective buoyancy.
-
-%However, the stabilizing effect of buoyancy can be mitigated for 
-%effectively isothermal disturbances. Consider a fluid element displaced upwards, away from
-%the disk midplane. In a stably stratified adiabatic disk, it would
-%sink because of buoyancy. However, if thermal timescales are short, in
-%that temperature flucations decay rapidly, then the fluid element would attain its surrounding 
-%density and  
-%%(assuming an ideal gas and it reaches pressure 
-%%equilibrium with its surroundings). 
-%thus buoyancy is rendered ineffective. Qualitatively, a short thermal
-%timescale $t_c$ reduces the effective (stabilizing) buoyancy force.
 
@@
 \end{align}
- %This result is far from rigorous, but it contains some useful insights.
 If we interpret Eq.\ \ref{bcrit_est} as a local criterion, then we see that 
@@
   t_c <  &\underbrace{\frac{h |q| }{\gamma - 1}}_{\beta_\mathrm{crit}}\OmK^{-1}
- % &\equiv \beta_\mathrm{crit}\OmK^{-1},\notag
 \end{align}   
@@
 (but non-rigourously) obtain Eq. \ref{prelim_bcrit}
-  by evaulating  Eq. \ref{bcrit_est} at $|z|=\gamma H/2$. %i.e., at $z=O(H)$
-
-%This estimate agrees with 
-%our central result, the cooling criterion that we derive more rigorously in \S\ref{iso_vsi_beta_crit}.
-
-%
-%The left-hand side represents an effective buoyancy, which is
-%stabilizing; while the right-hand side represents vertical shear,
-%which is destabilizing. In a thin disk $ h \ll 1$, so the destabilizing effect of 
-%vertical shear is rather weak because $|r\p_z\Omega|\sim h\OmK
-%\ll \OmK$. By comparison, the buoyancy frequency $N_z\sim
-%\OmK$. It follows that $t_c\lesssim O( h\OmK^{-1})$ is
-%required for instability.  \emph{In a thin  
-%  disk the thermal timescale must be shorter than the dynamical
-%  timescale by $O( h)\ll 1$ for the VSI to operate.}   
-%
-%
-%Then even a small (positive)
-%vertical entropy gradient can significantly stabilize the VSI. 
-%Note that in astrophysical disks both $N_z^2$ and $\p_z\Omega$ varies
-%with height. Eq. \ref{bcrit_est} then suggest that with
-%increasing height, buoyancy would eventually stabilize the VSI (since
-%$N_z^2\propto z^2$ and $\p_z\Omega\propto z$).
-%% This means that 
-%% disturbances localized away from the disk midplane are easily stabilized by
-%% buoyancy. 
-%This argument assumes vertically localized disturbances. It is,
-%however, not clear how to apply Eq. \ref{bcrit_est} to vertically
-%global disturbances, as considered in this work.
-%
-%
-%
-%
-%Our central result is the requirement 
-%\begin{align}\label{prelim_bcrit}
-%  (\gamma - 1)t_c <  h |q| \OmK^{-1}
-%\end{align}   
-%for the VSI in a vertically isothermal, thin disk. This equation
-%reflects Eq. \ref{bcrit_est} but there is no $z$ dependence. 
-%% While assumptions and simplifications will be made 
-%% to derive this, 
-%Eq. \ref{prelim_bcrit} can be 
-%readily applied to evaluate where in a disk the VSI is likely to
-%operate.
+  by evaulating  Eq. \ref{bcrit_est} at $|z|=\gamma H/2$.
 
@@
 %results to PPDs.
->>>>>>> 3c00bb4e2b076179e280f2c07728086a110764ef
 

### 2015-05-08 numerical.tex
edits to 6.4.  tried to more clearly state goal at outset and remove some repetitive discussion that repeats 6.3.  added speculation at end for non-polytropes.  I guess I'm officially happy now.

@@
 
-%we plot the maximum VSI growth rates
-%found in the fiducial disk model as a function of the thermal
-%relaxation timescale $\beta$ for several values of $\khat$ and
-%$\zmax$. We also plot the critical timescale
-%$\beta_\mathrm{crit}=0.125$ for the fiducial disk parameters.   
-%
-%Note that the curves are not neccessarily smooth because the most unstable
-%mode may not always correspond to the same type of disturbance. For example,
-%as seen above, for $\khat \leq 10$ higher order body modes are
-%dominant as $\beta\to 0$, but the fundamental 
-%mode becomes dominant as $\beta$ is increased. For $\khat=1$ the VSI
-%can operate slightly beyond $\beta_\mathrm{crit}$  (recall that
-%$\beta_\mathrm{crit}$ was derived assuming $\khat^2\gg 1$). For
-%$\khat=5,\,10$ (and $30$ for $\zmax=3H$), we find
-%$\beta_\mathrm{crit}$ accurately predicts the thermal timescal beyond
-%which the VSI is suppressed.   
-%
-%Complications from vertical boundaries arise when considering large
-%$\khat$. For $\khat=30$ and $\zmax=5H,\,7H$ we find for $\beta\gtrsim
-%0.1$ the dominant disturbance correspond to the leftmost eigenvalues 
-%seen in the lower two panels of Fig. \ref{compare_modes_cool_kx30},
-%which do not exist for $\zmax=3H$. For $\khat=50,\,100$ and
-%$\beta\to0$, we find surface modes dominate so the maximum growth rate
-%depends on $\zmax$ in the limit of rapid thermal relaxation. 
-%However, the maximum growth rate is insensitive to 
-%$\zmax$ for moderately large $\beta$ ($\gtrsim 0.1$), i.e. close to
-%the stabilization of the fundamental mode.
+\subsection{Vertically non-isothermal disk model} 
+We now generalize our critical cooling time by considering disks which
+are not vertically isothermal.  We rely on physical arguments for this generalization.
+In disks with weaker vertical stratification, i.e.\ $\Gamma > 1$, we expect $\beta_\mathrm{crit}$ to be larger.
+We thus transform $1/(\gamma -1) \to 1/(\gamma - \Gamma)$.
 
-%Our result is consistent with numerical simulations performed by
-%\citetalias{nelson13} for which a thermal relaxation timescale $\beta\gtrsim
-%0.6$ stabilized the VSI; as Fig. \ref{bcrit_compare1} show that growth
-%rates for $\beta\gtrsim 0.6$ are small, and are restricted to very
-%small radial lengthscales $\sim 10^{-2}H$, for which current numerical
-%simulations may not be able to resolve.
-%
-%As pointed out by \citetalias{barker15}, in vertically isothermal disks with
-%an imposed boundary, surface modes are dependent on $\zmax$, but they
-%are often the most unstable. Despite this caveat, and that surface
-%modes are not accounted in our analysis, we see from
-%Fig. \ref{bcrit_compare1} that $\beta_\mathrm{crit}$ nevertheless
-%provides a reasonable estimate for an upper limit to the thermal
-%timescale for the VSI.  
-%
-%We further  test the dependence of $\beta_\mathrm{crit}$ on disk parameters
-%(Eq. \ref{iso_vsi_cond}).  We use the fiducial setup as a  reference
-%point and vary $q\in[-0.2,-1.2]$,  $\gamma\in[1.2,2.0]$, and
-%$ h\in[0.02,0.1]$ separately. The numerically obtained
-%$\beta_\mathrm{crit}$ are shown in Fig. \ref{bcrit_compare} in
-%comparison with Eq. \ref{iso_vsi_cond}, which generally agree. This
-%result, together with Fig. \ref{bcrit_compare1}, shows that  
-%$\beta \lesssim \beta_\mathrm{crit}$ is a practical
-%criteria for the VSI to operate effectively in stably stratified,
-%vertically isothermal disks.
+We also expect $\beta_\mathrm{crit}$ to scale with the vertical shear.   
+In general the vertical shear rate $\p_z\Omega\propto s$, the radial
+entropy gradient, see Eq.\ \ref{vertical_shear_ex}.  We thus 
+ transform $q \to s = q +p(1-\Gamma)$.
 
-\subsection{Vertically non-isothermal disk model} 
-We briefly examine a vertically non-isothermal disk model with 
-a physical surface. Recall from Eq. \ref{vertical_shear_ex} that in
-general the vertical shear rate $\p_z\Omega\propto s$, the radial
-entropy gradient. From the expression for the critical cooling timescale,
-Eq. \ref{iso_vsi_cond}, it is reasonable to expect a similar criteria
-for vertically non-isothermal disks in the form 
+Our estimate for the generalized cooling 
+ time criterion -- valid for both vertically isothermal and non-isothermal disks (cf.\ Eq.\ \ref{iso_vsi_cond}) -- is thus
 \begin{align}\label{bcrit_noniso}
@@
 
-Here, we consider a disk model with $(p,q, h)=(-0.5,-1,0.05)$ and $(\gamma,
-\Gamma)=(1.4,1.3)$. We set the vertical domain size close to the
-physical disk surface, $\zmax = 0.99H_s\simeq 2.6H$, which is smaller
-than our standard domain size. For this setup, $s=-0.85$ and
+We test $\beta_\mathrm{crit, noniso}$ using a disk model with 
+$(p,q, h)=(-0.5,-1,0.05)$ and $(\gamma,
+\Gamma)=(1.4,1.3)$. For this setup, $s=-0.85$ and
 $\beta_\mathrm{crit,noniso}=0.425$, which is $\sim 3$ times larger than
@@
 
-% A disk surface at finite height
-% means that higher order body modes and surface modes are better
-% defined (cf. in vertically isothermal disk models where the disk
-% boundary $\zmax$ is a free parameter, see \citetalias{barker15}).
+We set the vertical domain size just inside the
+physical, zero-density disk surface, $\zmax = 0.99H_s\simeq 2.6H$, see Eq.\ \ref{eqm_dens}. 
+At this surface, the vertical buoyancy frequency, $N_z$,  diverges as $|z|\to H_s$.  While we might expect
+pathological behavior for this model, we fortunately do not find it.
 
 In Fig. \ref{compare_modes_vnoniso_kx10} we plot the mode diagram for
-$\khat=30$ for several values of $\beta$. This plot is 
-qualitatively similar to Fig. \ref{compare_modes_cool_kx10}. As before
-we find as $\beta$ is increased, higher order modes are rapidly
-stabilized, eventually only the fundamental mode can operate before
-the VSI is entirely suppressed. (Note the significant reduction in
-growth rates for $\beta=0.4$, which is close to
-$\beta_\mathrm{crit,noniso}$.) However, this behavior is not seen for 
-large wavenumbers (see below).
+$\khat=30$ and several values of $\beta$. This plot is 
+qualitatively similar to the vertically isothermal case, Fig. \ref{compare_modes_cool_kx10}. As before, 
+larger $\beta$ values rapidly
+stabilize higher order modes.  For sufficiently large $\beta$ only the fundamental mode is unstable, with a growth rate
+that decreases as $\beta$ approaches 
+$\beta_\mathrm{crit,noniso}$.
 
@@
 
-In Fig. \ref{bcrit_compare2} we plot the maximum VSI growth
-rates in the vertically non-isothermal disk model, which is
-qualitatively similar to Fig. \ref{bcrit_compare1} for the vertically
-isothermal disk. For $\khat=O(10)$, there is a   
-precise cut-off thermal timescale close to that given by
-Eq. \ref{bcrit_noniso}. We found that very large wavenumbers, 
-$\khat > O(10^2)$, were needed to enable surface modes, and modes with
-such values of $\khat$ do not have a maximum cooling time. However,
-these small lengthscales may not be important in practice
-(see \S\ref{caveats_visc}). Thus, $\beta<\beta_\mathrm{crit,noniso}$ 
-provides an excellent criteria for instability.
+Fig.\ \ref{bcrit_compare2} shows how maximum VSI growth
+rates vary with $\beta$ in the vertically non-isothermal model. The qualitative
+behavior similar to Fig.\ \ref{bcrit_compare1} for the vertically
+isothermal disk. For $\khat=O(10)$, growth rates drop to zero at the
+predicted $\beta_\mathrm{crit,noniso}$.  As in the vertically isothermal case,
+modes with both lower and higher wavenumbers  exhibit slow growth rates 
+beyond the critical cooling time. As argued in \S\ref{bcrit_num_test},
+the critical cooling time is useful despite not being an absolute stability limit.
 
@@
 
-The above calculation demonstrates that the effect of introducing
-finite cooling on the VSI operates very similarly in vertically
-isothermal disks with an imposed surface, and in vertically
-non-isothermal disks with a physical surface. This is despite the fact
-that the vertical buoyancy frequency, $N_z$, formally diverges as $|z|\to H_s$.
+The generalized cooling time criterion in Eq.\ \ref{bcrit_noniso}
+thus appears to be valid.  If a disk's vertical structure is
+not characterized by a single polytropic index, $\Gamma$, we 
+expect that a density weighted average of  $\beta_\mathrm{crit,noniso}$
+should be a good approximation.  This expectation remains to be tested.
 

### 2015-05-08 paper.tex
mkl going through paper

@@
 %to the Solar Nebula}  
-\author{Min-Kai Lin\altaffilmark{$\dagger$} \& Andrew N Youdin}
+\author{Min-Kai Lin\altaffilmark{1} \& Andrew N Youdin}
 \affil{Department of Astronomy and Steward Observatory, \\
@@
   Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA}
-\altaffiltext{$\dagger$}{Steward Theory Fellow}
+\altaffiltext{1}{Steward Theory Fellow}
 \email{minkailin@email.arizona.edu, youdin@email.arizona.edu}
@@
   The VSI is driven by a weak vertical gradient in the disk's orbital motion.
-  In order to grow, the VSI must overcome vertical buoyancy, a strong
+  In order to grow, the VSI must overcome vertical buoyancy, a strongly
   stabilizing influence in cold disks, where  
@@
   aspect ratio, $ h \ll 1$; the radial  
-  powerlaw temperature gradient, $q$; and the adiabatic index,
+  power-law temperature gradient, $q$; and the adiabatic index,
   $\gamma$.  For longer cooling times, the VSI  
@@
 We discuss caveats and extensions in \S\ref{caveats}.
- We conlude in \S\ref{summary} with a summary.  Some technical and background 
+ We conclude in \S\ref{summary} with a summary.  Some technical and background 
  developments are explained in the appendices.
@@
 
-We are more interested, however, the the global instability criterion for disturbances at all heights.
-Our desired result 
+We are more interested, however, in the global instability criterion for disturbances at all heights.
+Our key result 
 is the cooling requirement 
 \begin{align}\label{prelim_bcrit}
-  t_c <  &\underbrace{\frac{h |q| }{\gamma - 1}}_{\beta_\mathrm{crit}}\OmK^{-1}
+  t_c <  &\frac{h |q| }{\gamma - 1}\OmK^{-1}
 \end{align}   
 for VSI growth that we derive in \S\ref{iso_vsi_beta_crit}.  We can simply 
-(but non-rigourously) obtain Eq. \ref{prelim_bcrit}
-  by evaulating  Eq. \ref{bcrit_est} at $|z|=\gamma H/2$.
+(but non-rigorously) obtain Eq. \ref{prelim_bcrit}
+by evaluating  Eq. \ref{bcrit_est} at $|z|=\gamma H/2$.
 
@@
   0.73cm]{figures/bcrit_mmsn2d_kx50}
-  \caption{Estimate of the thermal timescale  in a fiducial 
-    PPD model, the Minimum Mass Solar Nebula. The red
-    lines corresponds to a maximum thermal timescale, derived in
-    this work, below which the VSI is expected to operate effectively
-    (the region bounded by the red lines).     
+  \caption{Estimate of the thermal timescale $t_c$  in a fiducial 
+    PPD model, the Minimum Mass Solar Nebula. Here, $\beta_\mathrm{crit}=h|q|/(\gamma-1)$  
+    is a maximum dimensionless cooling timescale below which the VSI
+    is expected to operate effectively. This corresponds to   
+    the region bounded by the red lines. 
     \label{bcrit_mmsn2d} 
@@
  The regions between the red lines satisfy the cooling requirement of Eq.\ \ref{prelim_bcrit}.
- Even without a detailed analysis we can expect VSI growth between $\sim 10$ -- 150 AU.
- The inner disk is complicated by the fact that the optically thick midplane has longer cooling times. 
- Thus the cooling criterion fails in the midplane but is satisfied above.  This case requires the 
+ Even without a detailed analysis we can expect VSI growth between $\sim 10$ -- 150AU.
+ The inner disk is complicated by the fact that the optically thick
+ midplane has longer cooling times.  
+ Thus the cooling criterion fails in the midplane but is satisfied
+ above.  This case requires the  
  more detailed analysis of \S\ref{application}.
@@
   Fig. \ref{bcrit_mmsn2d} also compares two different radial wavenumbers, 
-  $k_x = 10/H$ and $50/H$.  In the inner disk, the higher wavenumber mode cools faster, due to radiative diffusion.
-  Thus in the inner disk, shorter wavelength VSI modes are more like to grow. In the outer disk, the cooling time is the 
-  same for different $k_x$ and $z$, because optically thin cooling is independent of both lengthscale and density.
-  Thus the outer barrier to growth, beyond $\sim 150$ AU in this model, is independent of  wavelength. 
-This example highlights the fact that optically thin cooling sets a lower limit to the cooling time; a fact that is sometimes obscured when 
+  $k_x = 10/H$ and $50/H$.  In the inner disk, the higher wavenumber
+  mode cools faster, due to radiative diffusion. 
+  Thus in the inner disk, shorter wavelength VSI modes are more likely
+  to grow. In the outer disk, the cooling time is the  
+  same for different $k_x$ and $z$, because optically thin cooling is
+  independent of both lengthscale and density. 
+  Thus the outer barrier to growth, beyond $\sim 150$ AU in this
+  model, is independent of  wavelength.  
+This example highlights the fact that optically thin cooling sets a
+lower limit to the cooling time; a fact that is sometimes obscured
+when  
 the radiative diffusion approximation is made at the outset.

### 2015-05-09 appendix.tex
updated to apj submitted version

@@
 in Eq. \ref{lin_mass}, \ref{lin_vx} and \ref{lin_energy}, i.e. setting 
-$\zeta=0$. (Nomially $\zeta=1$, which is used in our numerical study.)
+$\zeta=0$. (Nominally $\zeta=1$, which is used in our numerical study.)
 
@@
 
-We comment that this approximtaion is equivalent to 
+We comment that this approximation is equivalent to 
 that adopted in the vertically global shearing box formalism
@@
 \end{align}
-as expected on physical grounds. %A similar conclusion can be reached in the
+as expected on physical grounds. Note that if the thin-disk
+approximation is used in an infinite domain, then the maximum growth
+rate is unbounded (since in that case $\mathrm{max}|f|\to\infty$).
+
+%A similar conclusion can be reached in the
 %equivalent linear problem in global cylindrical disk geometry.
@@
 In practice, one might consider a vertical domain of a few scale 
-heights in a thin disk with $|q|=O(1)$. In this case, $|h\zhat|\ll1$ and  
+heights in a thin disk with $|q|=O(1)$. Then 
 $f\simeq \zhat$, so that $\mathrm{max}|f| = O(1)$, implying a
@@
 which is equivalent to Eq. 34 of \citetalias{barker15} in the limit
-$\khat^2\gg L$. This signifies instability for
+$\khat^2\gg L,\,1$. This signifies instability for
 $L\geq1$ since we can choose the sign 
-of the square root to make $\hat{\sigma} = \imag\hat{\upsilon}>0$.  These
+of the square root to make $\imag\hat{\upsilon}>0$.  These
 are the low frequency unstable modes seen in
@@
 Eq. \ref{relax_cond1}---\ref{relax_cond2} without  
-assuming it is much less than unity from the onset or place
+assuming it is much less than unity at the outset or place
 restrictions on $\theta$.

### 2015-05-09 application.tex
updated to apj submitted version

@@
 In the inner disk, e.g. $r_\mathrm{AU}\sim 1$, only
-extremely small-scale perturabtions are in the optically-thin
+extremely small-scale perturbations are in the optically-thin
 regime near the midplane. However, in the outer disk, e.g. $r_\mathrm{AU}\sim 10$, 
@@
 \end{figure}
+
+\begin{figure*}
+  \includegraphics[scale=.47,clip=true,trim=0cm 1.8cm 0cm
+  0cm]{figures/bcrit_mmsn_kap0d1_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm 1.8cm 0cm
+  0cm]{figures/bcrit_mmsn_kap0d1_z3}\\ 
+  \includegraphics[scale=.47,clip=true,trim=0cm 1.8cm 0cm
+  1cm]{figures/bcrit_mmsn_kap1_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm
+  1.8cm 0cm 1cm]{figures/bcrit_mmsn_kap1_z3}\\
+  \includegraphics[scale=.47,clip=true,trim=0cm 0cm 0cm
+  1cm]{figures/bcrit_mmsn_kap10_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm 0cm 0cm
+  1cm]{figures/bcrit_mmsn_kap10_z3} 
+  \caption{Dimensionless thermal relaxation timescales $\beta$,
+    evaluated at the midplane (left) and at $z=3H$ (right) in the
+    fiducial PPD. Eq. \ref{beta_mmsn_simp} is plotted  
+    for three values of the 
+    perturbation radial wavenumber: $\khat=1$ (dotted), $\khat=10$
+    (dashed) and $\khat=100$ (dashed-dot), for three values of the
+    opacity relative to the MMSN: $\hat{\kappa}_d=0.1$ (top),
+    $\hat{\kappa}_d=1$ (middle) and $\hat{\kappa}_d=10$ (bottom).  
+    The solid line is the 
+    critical thermal relaxation timescale $\beta_\mathrm{crit}$.  
+    \label{mmsn_bcrit_bcool}}   
+\end{figure*}
 
@@
 we require the analysis of \S\ref{vsi_mmsn_grow} to determine if the
-VSI can grow.  (That calculation will show that growth is stongly
+VSI can grow.  (That calculation will show that growth is strongly
 suppressed for this case.)  We proceed with the awareness that
@@
 
-\begin{figure*}
-  \includegraphics[scale=.47,clip=true,trim=0cm 1.8cm 0cm
-  0cm]{figures/bcrit_mmsn_kap0d1_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm 1.8cm 0cm
-  0cm]{figures/bcrit_mmsn_kap0d1_z3}\\ 
-  \includegraphics[scale=.47,clip=true,trim=0cm 1.8cm 0cm
-  1cm]{figures/bcrit_mmsn_kap1_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm
-  1.8cm 0cm 1cm]{figures/bcrit_mmsn_kap1_z3}\\
-  \includegraphics[scale=.47,clip=true,trim=0cm 0cm 0cm
-  1cm]{figures/bcrit_mmsn_kap10_z0}\includegraphics[scale=.47,clip=true,trim=2.5cm 0cm 0cm
-  1cm]{figures/bcrit_mmsn_kap10_z3} 
-  \caption{Dimensionless thermal relaxation timescales $\beta$,
-    evaluated at the midplane (left) and at $z=3H$ (right) in the
-    fiducial PPD. Eq. \ref{beta_mmsn_simp} is plotted  
-    for three values of the 
-    perturbation radial wavenumber: $\khat=1$ (dotted), $\khat=10$
-    (dashed) and $\khat=100$ (dashed-dot), for three values of the
-    opacity relative to the MMSN: $\hat{\kappa}_d=0.1$ (top),
-    $\hat{\kappa}_d=1$ (middle) and $\hat{\kappa}_d=10$ (bottom).  
-    The solid line is the 
-    critical thermal relaxation timescale $\beta_\mathrm{crit}$.  
-    \label{mmsn_bcrit_bcool}}   
-\end{figure*}
-
 In Fig. \ref{mmsn_bcrit_bcool}, we compare $\beta$ to
@@
 wavenumbers overlap, as expected from Eq.\ \ref{beta_mmsn_simp}. 
-Since the (optically thin) slope of $\beta$ is slightly steeper than
+Since the (optically thin) slope of $\beta$ is steeper than
 for $\beta_\mathrm{crit}$, VSI growth can be suppressed at large radii
@@
 shorter and even more problematic, as shown by the red curves in
-Fig. \ref{mmsn_bcrit_bcool_mink}.  At lower opacties, even optically
+Fig. \ref{mmsn_bcrit_bcool_mink}.  At lower opacities, even optically
 thin cooling (the fastest possible) is too slow, as discussed above.
@@
 \subsection{VSI growth in the MMSN}\label{vsi_mmsn_grow}
-We now consider the growth of the VSI in a MMSN disk model.  The
-linear analysis is similar to  
+We now consider the growth of the VSI in a MMSN disk model.  This numerical calculation is similar to   
 \S\ref{numerical} but with different disk parameters and with the
@@
 because it is the fastest growing mode except for some surface modes at high 
-wavenumbers. We neglect these surface modes for reasons disucssed in
+wavenumbers. We neglect these surface modes for reasons discussed in
 \S\ref{numerical} and \S\ref{caveats_visc}.
@@
 from Fig.\ \ref{mmsn_bcrit_bcool_mink}.  Growth at smaller disk radii
-is possible for yet larger wavenumbers, yet we remain skeptical about
-the non-linear viability of such short modes.
+is possible for yet larger wavenumbers, but we remain skeptical about
+the non-linear significance of such short lengthscales.
 

### 2015-05-09 linear_adia.tex
updated to apj submitted version

@@
 dominate in numerical calculations with increasing $\beta$
-(\S\ref{therm_relax_eff}) for moderately large    
+(\S\ref{therm_relax_eff}) for moderate    
 values of $\khat$, as well as in PPDs 

### 2015-05-09 numerical.tex
updated to apj submitted version

@@
 $\zeta=1$, i.e.\ retaining  radial gradients of the background disk
-structure. The equilbrium disk structure given is in \S\ref{eqm}. This
+structure. The equilibrium disk structure given is in \S\ref{eqm}. This
 numerical study does not make the simplifying approximations  
@@
 
-Unless otherwise stated, we adopt fiducial disk parameters are 
+Unless otherwise stated, we adopt fiducial disk parameters 
 $(\gamma, \Gamma) = (1.4, 1.011)$, $(p,q, h)=(-1.5,-1,0.05)$, and 
@@
   \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10_analytic.ps}
-  \caption{Growth rate, $\sigma$, vs.\ osciallation frequency, $\omega$, for
+  \caption{Growth rate, $\sigma$, vs.\ oscillation frequency, $\omega$, for
   VSI modes with wavenumber $k_x = \khat/H=10/H$ in the fiducial disk model
@@
 show below that with slower cooling, the growth of higher order modes
-is preferentially damped, consistent with the analytic analysis in
+is preferentially damped, consistent with the analysis in
 \S\ref{iso_vsi_beta_crit}.  Second, the fastest growing modes may not
@@
 wavenumbers. Fig. \ref{compare_modes_iso_kx30} maps the $\khat = 30$
-eigenvalues, with the surface modes labelled.  The surface modes
+eigenvalues, with the surface modes labeled.  The surface modes
 strongly depend on the location of the imposed vertical boundary, and
@@
 that surface modes arise whether the surface is artificial (as for our
-imposed $\zmax$) or natural (for disk models with a zero density surface).
+imposed $\zmax$) or natural (for disk models with a zero-density surface).
 Physical surface modes are in principle possible; BL15 mention
@@
 modes, specifically $\khat=30$.   We only show the larger domains with $\zmax = 5H, 7H$. 
-These cases again show good convergence for $\beta \geq 0.03$, which the
+These cases again show good convergence for $\beta \geq 0.03$, which 
 the smaller domain with  $\zmax = 3H$ lacks.
@@
 as required in the analytic derivation of \S\ref{iso_vsi_beta_crit}.  Though  inconvenient,  given that the 
-fundemental mode is easy to identify and the most numerically converged, this complication is not ultimately significant
+fundamental mode is easy to identify and the most numerically converged, this complication is not ultimately significant
 for the operability of the VSI.
@@
 For longer wavelength modes, with $\khat = 1$, growth persists to slightly longer cooling times.  
-This difference is not surprising since our analytic derivation assumed $\khat^2 \gg1$, and 
+This difference is not surprising since our analytic derivation assumed $\khat^2 \gg1$, but  
 the change is quantitatively minor.
@@
 
-Our results agree with the vertically isothermal analysis of \citetalias{nelson13} 
+Our results agree with the vertically isothermal simulations of \citetalias{nelson13} 
 which used the same disk parameters as our fiducial model.  The expected $\beta_\mathrm{crit} = 0.125$ 
@@
 
-\subsection{Vertically non-isothermal disk model} 
+\subsection{Vertically non-isothermal disk model}\label{nonvertiso}
 We now generalize our critical cooling time by considering disks which
 are not vertically isothermal.  We rely on physical arguments for this generalization.
-In disks with weaker vertical stratification, i.e.\ $\Gamma > 1$, we expect $\beta_\mathrm{crit}$ to be larger.
+In disks with weaker vertical stratification, i.e.\ $\Gamma > 1$ for
+fixed $\gamma$, we expect $\beta_\mathrm{crit}$ to be larger.
 We thus transform $1/(\gamma -1) \to 1/(\gamma - \Gamma)$.
@@
 Our estimate for the generalized cooling 
- time criterion -- valid for both vertically isothermal and non-isothermal disks (cf.\ Eq.\ \ref{iso_vsi_cond}) -- is thus
+ time criterion --- valid for both vertically isothermal and non-isothermal disks (cf.\ Eq.\ \ref{iso_vsi_cond}) --- is thus
 \begin{align}\label{bcrit_noniso}
- \beta_\mathrm{crit}\to\beta_\mathrm{crit, noniso} =
+ \beta_\mathrm{crit}\to\beta_\mathrm{crit, gen} =
  \frac{h|s|}{\gamma - \Gamma}. 
@@
 
-We test $\beta_\mathrm{crit, noniso}$ using a disk model with 
+We test $\beta_\mathrm{crit, gen}$ using a disk model with 
 $(p,q, h)=(-0.5,-1,0.05)$ and $(\gamma,
 \Gamma)=(1.4,1.3)$. For this setup, $s=-0.85$ and
-$\beta_\mathrm{crit,noniso}=0.425$, which is $\sim 3$ times larger than
+$\beta_\mathrm{crit,gen}=0.425$, which is $\sim 3$ times larger than
 our fiducial, vertically isothermal disk with $\beta_\mathrm{crit}=0.125$.
@@
 that decreases as $\beta$ approaches 
-$\beta_\mathrm{crit,noniso}$.
+$\beta_\mathrm{crit,gen}$.
 
@@
 rates vary with $\beta$ in the vertically non-isothermal model. The qualitative
-behavior similar to Fig.\ \ref{bcrit_compare1} for the vertically
+behavior is similar to Fig.\ \ref{bcrit_compare1} for the vertically
 isothermal disk. For $\khat=O(10)$, growth rates drop to zero at the
-predicted $\beta_\mathrm{crit,noniso}$.  As in the vertically isothermal case,
+predicted $\beta_\mathrm{crit,gen}$.  As in the vertically isothermal case,
 modes with both lower and higher wavenumbers  exhibit slow growth rates 
@@
     timescale accounting for a non-isothermal background, given by 
-     $\beta_\mathrm{crit,noniso}= 0.425$ (see Eq. \ref{bcrit_noniso}),   
+     $\beta_\mathrm{crit,gen}= 0.425$ (see Eq. \ref{bcrit_noniso}),   
     is also shown as the vertical dotted line. 
@@
 not characterized by a single polytropic index, $\Gamma$, we 
-expect that a density weighted average of  $\beta_\mathrm{crit,noniso}$
+expect that a density weighted average of  $\beta_\mathrm{crit,gen}$
 should be a good approximation.  This expectation remains to be tested.

### 2015-05-09 setup.tex
minor
mkl going through paper
updated to apj submitted version

@@
 \end{align} \end{subequations}
-where $r_0$ is a fiducial radius and $p$ and $q$ are the standard powerlaw indices
-for midplane density and temperature, respectively.  The polytropic index $\Gamma$ 
+where $r_0$ is a fiducial radius and $p$ and $q$ are the standard power-law indices
+for midplane density and temperature, respectively. Without loss of
+generality we take $q<0$, as is typical in PPDs.   
+The polytropic index $\Gamma$ 
 parametrizes the vertical stratification with $\Gamma = 1$ ($\Gamma = \gamma$)
@@
 The vertically isothermal case, $\Gamma = 1$, can be calculated either 
-as a special case or by taking the limit of Eq. \ref{eqm_dens},    
+as a special case or by taking the limit of Eq. \ref{eqm_dens} as $\Gamma\to 1$,    
 \begin{subequations}\begin{align}
-\lim_{\Gamma\to1}  \frac{\rho(r,z)}{\rho_0(r)} &=
+ \frac{\rho(r,z)}{\rho_0(r)} &=
   \exp{\left[\frac{1}{ h^2}\left(\frac{1}{\sqrt{1+z^2/r^2}}-1\right)\right]},   \\
@@
   -\frac{\p P}{\p z}\left(\frac{\p j^2}{\p r}\frac{\p S}{ \p z} -
-    \frac{\p j^2}{\p z}\frac{\p S}{\p r} \right)&>0. \label{second_SH} 
+    \frac{\p j^2}{\p z}\frac{\p S}{\p r} \right)&>0 \label{second_SH} 
 \end{align}\end{subequations}
@@
 implies stability. For typical model parameters, the right hand
-side (RHS) is $O( h^2) \ll 1$.
-The left hand side is positive and order unity ($O(0.1)$ at least) if the disk is convectively stable, and not very close to convective instability.
-This stable stratification is expected for irradiated disks  \citep{chiang97}.
-Thus adiabatic disturbances are stable to a disk's vertical shear, explaining why the VSI requires cooling.
+side is $O(h^2) \lesssim 10^{-2}$. The left hand side is positive and order
+unity ($O(0.1)$ at least) if the disk is convectively stable, and not
+very close to convective instability.  This stable stratification is
+expected for irradiated disks  \citep{chiang97}. 
+Thus adiabatic disturbances are stable to a disk's vertical shear,
+explaining why the VSI requires cooling.
 
@@
 The first set, presented in \S\ref{sec:radlocal}, is more general and used 
-for precise numerical calculations.  The second set, in \S\ref{sec:simplified}, makes
+for numerical calculations.  The second set, in \S\ref{sec:simplified}, makes
 additional approximations  about disk structure and wave frequency.  This simplified
@@
 are presented in \S\ref{numerical} and \S\ref{application}.   The coefficient $ \zeta = 1$ is 
-introduced simply to label terms containing radial gradients of the equilibrium state, 
-which again are evaluated at $(r_0, z)$. For clarity, hereafter we drop the subcript $0$.    
+introduced simply to label terms with explicit radial gradients of the equilibrium state, 
+which again are evaluated at $(r_0, z)$. For clarity, hereafter we drop the subscript $0$.    
 We will consider the effects of ignoring these radial gradients below.

### 2015-05-09 summary.tex
tweak to summary, brief mention of non-vertically iso case.
updated to apj submitted version

@@
 Our neglect of viscosity is valid in a laminar disk, because molecular viscosity is so small.
-However turbulence could act as an effective viscosity which preferentially damps 
+However, turbulence could act as an effective viscosity which preferentially damps 
 small scale modes.   Since the goal of VSI is to drive turbulent transport,
@@
 the standard
-precription for the kinematic viscosity $\nu = \alpha c_s H$
-\citep{shakura73}, in terms of the dimsionless $\alpha$ 
+prescription for the kinematic viscosity $\nu = \alpha c_s H$
+\citep{shakura73}, in terms of the dimensionless $\alpha$ 
 parameter. The viscous timescale for a perturbation lengthscale  
@@
 For significant growth, $t_\mathrm{visc}$ should be longer than the
-characterstic VSI growth timescale $1/(h|q|\Omega)$, i.e.\ 
+characteristic VSI growth timescale $1/(h|q|\Omega)$, i.e.\ 
 $\khat^2\lesssim |q|h/\alpha$. With $h \sim 0.05$, $q = -1$  and
@@
 \end{align}
-Since vertical buoyancy is an important stabilitizing influnce for the VSI, it should
+Since vertical buoyancy is an important stabilizing influence for the VSI, it should
 be considered in future studies of CONO.
@@
 The critical cooling time is thus short, shorter than the orbital time by a factor of the disk
-aspect ratio.  This finding is consistent with, and helps explain, the results of 3D numerical simulations \citepalias{nelson13}.
+aspect ratio.  This finding is consistent with, and helps explain, the results of recent numerical simulations \citepalias{nelson13}.
+We briefly consider  vertically non-isothermal disks in \S\ref{nonvertiso}.
 
@@
  cool faster, so long as they remain optically thick.  In the inner disk, however, our VSI cooling criterion requires 
- wavelengths that are too short to drive significant transport.  The best hope for the VSI in inner disk appears to 
- be a low surface density, which speeds radiative diffusion by lengthening the photon mean free path.  This option naturally begs 
+ wavelengths that are too short to drive significant transport.
+The best hope for the VSI in inner disk is a low surface density, which speeds radiative diffusion by lengthening the photon mean free path.
+This option naturally begs
  the question of what accretion mechanism lowered the surface density in the first place.

### 2015-06-05 iaus/proceed.tex
added iaus proceeding (in progress)

% iaus2esa.tex -- sample pages for Proceedings IAU Symposium document class
% (based on v1.0 cca2esam.tex)
% v1.04 released 17 May 2004 by TechBooks
%% small changes and additions made by KAvdH/IAU 4 June 2004
% Copyright (2004) International Astronomical Union

\NeedsTeXFormat{LaTeX2e}

\documentclass{iau}
\usepackage{graphicx}
\usepackage{amsmath}
\newcommand{\apj}{ApJ}
\newcommand{\mnras}{MNRAS}
\newcommand{\aap}{A\&A}
%\usepackage{bibtex}

\title[Young Stars Near the Sun] %% give here short title %%
{Vertical Shear Instability in the Solar Nebula}

\author[M.-K. Lin \& A. N. Youdin]   %% give here short author list %%
{Min-Kai Lin \and Andrew N. Youdin}

\affiliation{Department of Astronomy and Steward Observatory,\\ 
  University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721, USA 
  \\ email: {\tt minkailin@email.arizona.edu}, {\tt youdin@email.arizona.edu }}

\pubyear{2015}
\volume{314}  %% insert here IAU Symposium No.
\pagerange{119--126}
% \date{?? and in revised form ??}
\setcounter{page}{1}
\jname{Young Stars \& Planets Near the Sun}
\editors{J. H. Kastner, B. Stelzer, \& S. A. Metchev, eds.}
\begin{document}

\maketitle

\begin{abstract}
  We quantify the thermodynamic requirement for the Vertical Shear
  Instability, and evaluate its relevance to realistic protoplanetary
  disks as a potential route to hydrodynamic turbulence.

\keywords{accretion, accretion disks, hydrodynamics, instabilities, methods: analytical}
\end{abstract}

\firstsection % if your document starts with a section,
              % remove some space above using this command.
\section{Introduction}
A purely hydrodynamic route to turbulence has important implications
for the evolution of cold protoplanetary disks and solids within
them. One such candidate is the Vertical Shear Instability (VSI) 
operating in disks where the orbital frequency $\Omega$ depends on the
height $z$ away from the disk midplane. While astrophysical disks generally
possess such vertical shear, VSI also requires short thermal 
timescales. We determine a quantitative thermodynamic requirement for
the VSI, and apply our results to realistic disk conditions. We find 
the VSI can operate at $5$---$50$AU in a typical protoplanetary disk,
with characteristic growth times of $30$ orbits.

\section{The need for rapid cooling for the VSI}
Astrophysical disks generally have vertical shear, $\p_z\Omega\neq
0$. This is a source of free energy, and may thus lead to
instability (\cite[Goldreic \& Schubert 1967]{goldreich67};
\cite[Urpin 2003]{urpin03}; \cite[Barker \& Latter 2015]{barker15}).  
However, if the disk is stably stratified, as is typical
in irradiated protoplanetary disks, the VSI also requires rapid
cooling to overcome the stabilizing influence of vertical buoyancy
(\cite[Nelson et al. 2013]{nelson13}).

More specifically, for a vertically isothermal disk with radial
temperature dependence $T\propto r^{q}$, the cooling or thermal
relaxation timescale $t_c$ must be sufficiently small for the VSI to
operate effectively:
\begin{align}\label{crit}
  t_c\Omega_\mathrm{K} < \frac{h|q|}{\gamma - 1}\equiv
  \beta_\mathrm{crit},  
\end{align}
where $\Omega_\mathrm{K}$ is the Keplerian frequency, $h$  is the disk
aspect ratio, and $\gamma$ is the adiabatic index (\cite[Lin \& Youdin
2015]{lin15}).

Since protoplanetary disks are typically thin, $h\ll 1$,
Eq. \ref{crit} implies that $t_c \ll \Omega_\mathrm{K}$ is required
for the VSI. That is, the thermal timescale must be significantly
shorter than the dynamical timescale, which reflects the fact that
vertical shear is weakly destablizing, while vertical buoyancy is
strongly stabilizing.

%gcorr_compare_iso_maxrate_z5.ps

% \begin{figure}
%   \begin{center}
%     \includegraphics[width=0.5\linewidth]{gcorr_compare_iso_maxrate_z5.ps} 
%     \caption{Maximum VSI growth rates in a vertically isothermal disk
%       with $(h,q,\gamma)=(0.05,-1,1.4)$ as a function of the
%       dimensionless cooling timescale
%       $\beta=t_c\Omega_\mathrm{K}$. The vertical line denotes the
%       characteristic maximum timescale according to Eq. \ref{crit}. 
%       \label{linear}}
%   \end{center}
% \end{figure}

\section{Applicability of the VSI in realistic protoplanetary disks}

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{eigen_compare_grow.ps} 
    \caption{stuff
      \label{mmsn}}
  \end{center}
\end{figure}

\section{Conclusions}

\begin{thebibliography}{}

\bibitem[{{Barker} \& {Latter}(2015)}]{barker15}
  {Barker}, A.~J., \& {Latter}, H.~N. 2015, \mnras, 450, 21

\bibitem[{{Chiang} \& {Youdin}(2010)}]{chiang10}
  {Chiang}, E., \& {Youdin}, A.~N. 2010, Annual Review of Earth and Planetary
  Sciences, 38, 493

\bibitem[{{Goldreich} \& {Schubert}(1967)}]{goldreich67}
  {Goldreich}, P., \& {Schubert}, G. 1967, \apj, 150, 571

\bibitem[{{Lin} \& {Youdin}(2015)}]{lin15}
  {Lin}, M.-L., \& {Youdin}, A.~N. 2015, \mnras, submitted, ArXiv: 1505.02163

\bibitem[{{Nelson} {et~al.}(2013){Nelson}, {Gressel}, \& {Umurhan}}]{nelson13}
  {Nelson}, R.~P., {Gressel}, O., \& {Umurhan}, O.~M. 2013, \mnras, 435, 2610

\bibitem[{{Urpin}(2003)}]{urpin03}
  {Urpin}, V. 2003, \aap, 404, 397
\end{thebibliography}

% \bibliographystyle{apj}
% \bibliography{ref}

\end{document}

### 2015-06-05 iaus/makepdf.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+latex proceed.tex
+dvips -o proceed.ps proceed.dvi
+ps2pdf -sPAPERSIZE=a4 proceed.ps
+echo 'done'

### 2015-06-08 iaus/proceed.tex
rough draft of proceeding

@@
   We quantify the thermodynamic requirement for the Vertical Shear
-  Instability, and evaluate its relevance to realistic protoplanetary
+  Instability and evaluate its relevance to realistic protoplanetary 
   disks as a potential route to hydrodynamic turbulence.
@@
 height $z$ away from the disk midplane. While astrophysical disks generally
-possess such vertical shear, VSI also requires short thermal 
+possess vertical shear, VSI also requires short thermal 
 timescales. We determine a quantitative thermodynamic requirement for
 the VSI, and apply our results to realistic disk conditions. We find 
-the VSI can operate at $5$---$50$AU in a typical protoplanetary disk,
+the VSI can operate effectively at $5$---$50$AU in a typical protoplanetary disk,
 with characteristic growth times of $30$ orbits.
@@
 instability (\cite[Goldreic \& Schubert 1967]{goldreich67};
-\cite[Urpin 2003]{urpin03}; \cite[Barker \& Latter 2015]{barker15}).  
-However, if the disk is stably stratified, as is typical
-in irradiated protoplanetary disks, the VSI also requires rapid
+\cite[Urpin 2003]{urpin03}; \cite[Barker \& Latter 2015]{barker15}),
+provided the disturbance vertical lengthscale is much larger than its
+radial lengthscale. However, if the disk is stably stratified, as is typical
+for irradiated protoplanetary disks, the VSI also requires rapid
 cooling to overcome the stabilizing influence of vertical buoyancy
@@
 2015]{lin15}). 
-
 Since protoplanetary disks are typically thin, $h\ll 1$,
@@
 %   \begin{center}
-%     \includegraphics[width=0.5\linewidth]{gcorr_compare_iso_maxrate_z5.ps} 
+%     \includegraphics[width=\linewidth]{gcorr_compare_iso_maxrate_z5.ps} 
 %     \caption{Maximum VSI growth rates in a vertically isothermal disk
@@
 \section{Applicability of the VSI in realistic protoplanetary disks} 
+We solve the linear stability problem for axisymmetric
+disturbances in a model of the Solar Nebula described in \cite[Chiang
+\& Youdin (2010)]{chiang10}. We include in the energy equation a
+realistic cooling function, based on dust opacity, which depends on
+$z$ as well as the disturbance lengthscale.
+
+In Fig. \ref{mmsn} we plot characteristic VSI growth times for a range
+of radial perturbation wavenumbers. Growth times increase rapidly at $\lesssim5$AU
+and $\gtrsim 100$AU as the cooling times violate Eq. \ref{crit} in
+these regions. Thus, the VSI is most effective at intermediate
+radii. Notice also the growth times are strongly dependent on
+lengthscale toward the optically-thick inner disk, with smaller-scale
+disturbances being able to operate further inwards. In the
+optically-thin outer disk, cooling times are independent of scale and
+all scales are stabilized beyond approximate the same radius.
 
@@
     \includegraphics[width=\linewidth]{eigen_compare_grow.ps} 
-    \caption{stuff
+    \caption{Characteristic growth times of the Vertical Shear
+      Instability in a typical protoplanetary disk model. Here, $k_x$
+      is the wavenumber of the disturbance and $H$ is the disk
+      scale-height. 
       \label{mmsn}}
@@
 \section{Conclusions}
+Vertical shear is ubiquitous in astrophysical disks. Such
+configurations are unstable if the cooling timescales are
+significantly shorter than the dynamical timescale (Eq. \ref{crit}). 
+This can be satisified for small-scale disturbances at few 10s
+of AU in typical protoplanetary disk 
+models with realistic cooling. The VSI is thus dynamically important
+in the outer disk, with potentially significantly implications for
+transport and dust evolution.
 
@@
 \bibitem[{{Lin} \& {Youdin}(2015)}]{lin15}
-  {Lin}, M.-L., \& {Youdin}, A.~N. 2015, \mnras, submitted, ArXiv: 1505.02163
+  {Lin}, M.-K., \& {Youdin}, A.~N. 2015, \apj, submitted, ArXiv: 1505.02163
 

### 2015-06-12 iaus/proceed.tex
minor update to iaus314

@@
 0$. This is a source of free energy, and may thus lead to
-instability (\cite[Goldreic \& Schubert 1967]{goldreich67};
+instability (\cite[Goldreich \& Schubert 1967]{goldreich67};
 \cite[Urpin 2003]{urpin03}; \cite[Barker \& Latter 2015]{barker15}),
@@
 shorter than the dynamical timescale, which reflects the fact that
-vertical shear is weakly destablizing, while vertical buoyancy is
+vertical shear is weakly destabilizing, while vertical buoyancy is
 strongly stabilizing.
@@
 significantly shorter than the dynamical timescale (Eq. \ref{crit}). 
-This can be satisified for small-scale disturbances at few 10s
+This can be satisfied for small-scale disturbances at few 10s
 of AU in typical protoplanetary disk 

### 2015-07-06 appendix.tex
first revised draft, only reviewer comments addressed, still want to go through entire paper once more before re-submit

@@
 In the presence of vertical shear $q\neq0$, Eq. \ref{adia_integral}
-suggests $\hat{\upsilon}^2$ is complex for real $\khat$, which implies the
-possibility of instability, regardless of the value of
-$|\gamma-1|$. This is contradicts the second Solberg-Hoiland criterion 
-(Eq. \ref{solberg2}), which imply instability for adiabatic
-perturbations in a thin, vertically isothermal disk would require 
-$|\gamma-1|\ll1$. 
+{\bf shows that} $\hat{\upsilon}^2$ is complex for real $\khat$, which
+indicates instability {\bf for any value of  $\gamma>1$. This is inconsistent
+  with the second Solberg-Hoiland criterion  
+(Eq. \ref{solberg2}), which implies that instability requires the disk
+to be close to neutral stratification (i.e. $\gamma-1\ll 1$ for a
+vertically isothermal disk).} 
+%This is contradicts the second Solberg-Hoiland criterion 
+%(Eq. \ref{solberg2}), which imply instability for adiabatic
+%perturbations in a thin, vertically isothermal disk would require 
+%$|\gamma-1|\ll1$. 
 %also need real freq^2 in adia limit 
@@
 rate is unbounded (since in that case $\mathrm{max}|f|\to\infty$).
+{\bf However, large growth rates would invalidate the low-frequency approximation.}
 
@@
   &c_3 = \beta\left\{h q \khat + \gamma \left[\ii + h q
-      k \left(1+2\khat^2\right)\right] - 3\ii - 2\ii
+      \khat \left(1+2\khat^2\right)\right] - 3\ii - 2\ii
     \khat^2\right\},\\
@@
 for all $M\geq 0$ if $\theta\leq 1$, which imply 
-$\mathrm{max}(\beta_c)$ occurs at $M=0$. This result may also be
-obtained by explicitly solving 
+$\mathrm{max}(\beta_c)$ occurs at $M=0$.  {\bf This conclusion may also be reached} by explicitly solving 
 Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} with $\theta$ as a small parameter. 

### 2015-07-06 linear_iso.tex
first revised draft, only reviewer comments addressed, still want to go through entire paper once more before re-submit

@@
 
- In Appendix \ref{global_corr}  we discuss the limits of these approximations
- and show that they are valid in the parameter space of interest for the VSI.
+ In Appendix \ref{global_corr}  we discuss the limits of these approximations.
+% and show that they are valid in the parameter space of interest for the VSI.
 In particular we point out that the fully-radially-local approximation is only valid for
-short cooling times, $\beta \lesssim O(1)$.
+short cooling times, $\beta \lesssim O(1)${\bf, which is the 
+  appropriate regime for the VSI to operate, as demonstrated below. 
+  On the other hand, for
+  $\beta\gtrsim O(1)$, this approximation can introduce artificial
+  instability due to the non-self-consistent neglect of global radial gradients.}
 

### 2015-07-06 numerical.tex
first revised draft, only reviewer comments addressed, still want to go through entire paper once more before re-submit

@@
 boundaries is a free surface,  
-\begin{align}
+\begin{align}\label{bc}
   \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0 \quad \text{at } z=\pm\zmax,
@@
   \caption{Visualization of the fundamental VSI mode in
-    Fig. \ref{lowfreq_eigenfunc}. The color scale corresponds to the
-    pressure perturbation $\delta P=\rho W$ scaled by its maximum value.
-    Arrows show the flow, scaled as $\sqrt{\rho}(\delta
-    v_x,\delta v_z)$. The horizontal axis has been stretched 
+    Fig. \ref{lowfreq_eigenfunc}. {\bf For this plot the mode has been
+    transformed back to real space. Thus, the color scale corresponds to the
+    \emph{real} 
+    pressure perturbation scaled by its maximum value.
+    Arrows show the \emph{real} meridional flow  multiplied by
+    $\sqrt{\rho}$.}
+  The horizontal axis has been stretched 
     for clarity.  
@@
 Fig. \ref{lowfreq_eigenfunc} shows 
-eigenfunctions $W = \delta P/\rho$ and $\delta   v_z$ for the lowest order fundamental mode.  
+eigenfunctions $W = \delta P/\rho$ and $\delta   v_z$ for the lowest
+order fundamental mode.  {\bf (In this and similar plots below, we
+  normalize eigenfunctions such that its maximum amplitude is unity
+  and vanishing imaginary part at the lower boundary.)} 
 Fig. \ref{lowfreq_eigenfunc_2d} maps the pressure perturbation and meridional flow,
@@
 wavenumbers. Fig. \ref{compare_modes_iso_kx30} maps the $\khat = 30$
-eigenvalues, with the surface modes labeled.  The surface modes
+eigenvalues, with the surface modes labeled.  Surface modes
 strongly depend on the location of the imposed vertical boundary, and
@@
 domains \citepalias{nelson13, mcnally14}. \citetalias{barker15} show
-that surface modes arise whether the surface is artificial (as for our
-imposed $\zmax$) or natural (for disk models with a zero-density surface).
-Physical surface modes are in principle possible; BL15 mention
-surfaces between a disk's interior and corona.  In models where the
-surface is imposed, like ours, surface modes are a numerical
-artifact.
+that surface modes arise whether the surface is {\bf imposed (as in
+  our models)} or natural (for disk models with a zero-density surface).
+{\bf BL15 mention the interface between a disk's interior and corona
+  as a physical site for surface modes develop. While surface modes in
+  our current disk model are a genuine mathematical feature of the 
+  problem defined by Eq. \ref{lin_all} and Eq. \ref{bc}, they are
+  artificial in the sense that there is no natural choice for $\zmax$.  
+}
+
+%Physical surface modes are in principle possible; BL15 mention
+%surfaces between a disk's interior and corona. In models where the
+%surface is imposed, like ours, surface modes are {\bf artificial.}
+%{\bf They are, however, a genuine mathematical feature of the linear
+ % problem defined by Eq. \ref{lin_all} and Eq. \ref{bc}.}
 
@@
 The motion in surface modes is indeed concentrated near the surface,
-as shown in Fig. \ref{lowfreq_eigenfunc_surf}.   
-Thus even if surface modes were physical, their contribution to
+as shown in Fig. \ref{lowfreq_eigenfunc_surf}, {\bf where the density
+  is low.} 
+Thus, their contribution to
 transport might (by themselves) be weak,  following the arguments in 
@@
 boundary has little effect on the critical cooling time.  This agreement occurs despite the fact that peak growth rates
-differ as $\beta \to 0$.  We are thus confident that artificial boundary effects, including surface modes, do not
+differ as $\beta \to 0$.  We are thus confident that boundary effects, including surface modes, do not
 affect our analysis of the critical cooling time.

### 2015-07-06 setup.tex
first revised draft, only reviewer comments addressed, still want to go through entire paper once more before re-submit

@@
 side is $O(h^2) \lesssim 10^{-2}$. The left hand side is positive and order
-unity ($O(0.1)$ at least) if the disk is convectively stable, and not
-very close to convective instability.  This stable stratification is
+unity if the disk is convectively stable, and not
+very close to convective instability  {\bf( $\gamma-\Gamma\geq0.1$ in
+  the disk models considered)}. This stable stratification is
 expected for irradiated disks  \citep{chiang97}. 
-Thus adiabatic disturbances are stable to a disk's vertical shear,
-explaining why the VSI requires cooling.
+Thus, adiabatic disturbances are stable to a disk's vertical shear
+{\bf for typical values of  $\gamma-\Gamma$},  
+explaining why the VSI requires cooling in realistic disks.
 

### 2015-07-08 appendix.tex
revised manuscript, with referee response.txt

@@
 (Eq. \ref{solberg2}), which implies that instability requires the disk
-to be close to neutral stratification (i.e. $\gamma-1\ll 1$ for a
+to be close to neutral stratification (i.e. $|\gamma-1|\ll 1$ for a
 vertically isothermal disk).} 
@@
 rate is unbounded (since in that case $\mathrm{max}|f|\to\infty$).
-{\bf However, large growth rates would invalidate the low-frequency approximation.}
+{\bf However, large growth rates invalidate the low-frequency
+  approximation and the above analysis breaks down.}
 
@@
 \subsection{Maximum critical cooling time}\label{max_cool}   
-Here we show that, under appropriate conditions, the longest critical 
+Here we show that {\bf for sufficiently thin disks} the longest critical 
 cooling time is that for the $M=0$ or fundamental mode. This allows
@@
 Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} with $\theta$ as a small parameter. 
-For fixed $\khat$ the condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
-All such modes are stabilized if
+For fixed $\khat$ the condition $\theta\leq 1$ can be satisfied for 
+sufficiently {\bf small $h$}. All such modes are stabilized if
 $\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$.

### 2015-07-08 linear_iso.tex
revised manuscript, with referee response.txt

@@
 short cooling times, $\beta \lesssim O(1)${\bf, which is the 
-  appropriate regime for the VSI to operate, as demonstrated below. 
+  regime in which the VSI operates, as demonstrated below. 
   On the other hand, for 

### 2015-07-08 numerical.tex
revised manuscript, with referee response.txt

@@
   normalize eigenfunctions such that its maximum amplitude is unity
-  and vanishing imaginary part at the lower boundary.)} 
+  with vanishing imaginary part at the lower boundary.)} 
 Fig. \ref{lowfreq_eigenfunc_2d} maps the pressure perturbation and meridional flow,
@@
   as a physical site for surface modes develop. While surface modes in
-  our current disk model are a genuine mathematical feature of the 
+  a vertically isothermal disk model are a genuine mathematical feature of the 
   problem defined by Eq. \ref{lin_all} and Eq. \ref{bc}, they are

### 2015-07-08 paper.tex
revised manuscript, with referee response.txt

@@
 \begin{abstract}
-  It is difficult to understand how cold circumstellar disks accrete onto their central stars.  Due to 
-  lower levels of ionization, magnetic transport mechanisms that
-  operate in warmer disks are less effective.  
-  A hydrodynamic mechanism, the vertical shear instability (VSI), offers a means to drive angular momentum 
-  transport in cold accretion disks such as protoplanetary disks (PPDs).  
-  The VSI is driven by a weak vertical gradient in the disk's orbital motion.
-  In order to grow, the VSI must overcome vertical buoyancy, a strongly
+{\bf
+The vertical shear instability (VSI) offers a potential hydrodynamic mechanism to drive angular momentum transport in cold protoplanetary disks (PPDs).
+The VSI is driven by a weak vertical gradient in the disk's orbital motion, but must overcome vertical buoyancy, a strongly
   stabilizing influence in cold disks, where
-  heating is dominated by external irradiation.  Rapid cooling, via
-  radiative losses, reduces the effective buoyancy 
-  and allows the VSI to operate.  In this paper, we quantify the
-  cooling timescale, $t_c$, needed for growth of the VSI.
-   We perform a linear analysis of the VSI with cooling in
-  vertically global and radially local disk models. 
-  For irradiated disks, we find that the VSI is most vigorous for
+  heating is dominated by external irradiation.  Rapid radiative cooling      reduces the effective buoyancy
+  and allows the VSI to operate.  We quantify the
+  cooling timescale, $t_c$, needed for growth of the VSI. We perform a linear analysis of the VSI with cooling in
+  vertically global, radially local disk models. We find the VSI is most vigorous for
   rapid cooling with $t_c < \OmK^{-1}  h |q| / (\gamma -1)$
@@
   power-law temperature gradient, $q$; and the adiabatic index,
-  $\gamma$.  For longer cooling times, the VSI  
+  $\gamma$.  For longer $t_c$, the VSI
   is much less effective because growth slows and shifts to smaller length scales, which are more prone to
@@
   solid protoplanets, a sink for dust grains, can impede the VSI.
+}
+
+%  It is difficult to understand how cold circumstellar disks accrete onto their central stars.  Due to 
+%  lower levels of ionization, magnetic transport mechanisms that
+%  operate in warmer disks are less effective.  
+%  A hydrodynamic mechanism, the vertical shear instability (VSI), offers a means to drive angular momentum 
+ % transport in cold accretion disks such as protoplanetary disks (PPDs).  
+%  The VSI is driven by a weak vertical gradient in the disk's orbital motion.
+%  In order to grow, the VSI must overcome vertical buoyancy, a strongly
+%  stabilizing influence in cold disks, where  
+%  heating is dominated by external irradiation.  Rapid cooling, via
+%  radiative losses, reduces the effective buoyancy 
+%  and allows the VSI to operate.  In this paper, we quantify the
+%  cooling timescale, $t_c$, needed for growth of the VSI.
+%   We perform a linear analysis of the VSI with cooling in
+%  vertically global and radially local disk models. 
+%  For irradiated disks, we find that the VSI is most vigorous for
+%  rapid cooling with $t_c < \OmK^{-1}  h |q| / (\gamma -1)$ 
+%  in terms of the Keplerian orbital frequency, $\OmK$; the disk's
+%  aspect ratio, $ h \ll 1$; the radial  
+%  power-law temperature gradient, $q$; and the adiabatic index,
+%  $\gamma$.  For longer cooling times, the VSI  
+%  is much less effective because growth slows and shifts to smaller length scales, which are more prone to 
+%  viscous or turbulent decay.  We apply our results to PPD models 
+%  where $t_c$ is determined by the opacity of dust grains.  We
+%  find that the VSI is most effective at intermediate radii, from $\sim 5$AU to $\sim 50$AU with a
+%  characteristic growth time of $\sim 30$ local orbital periods.
+%  Growth is suppressed by long cooling times both in the opaque  
+%  inner disk and the optically thin outer disk.  A reduction in the
+%  dust opacity by a factor of 10 increases cooling times  
+%  enough to quench the VSI at all disk radii.  Thus the formation of
+%  solid protoplanets, a sink for dust grains, can impede the VSI. 
 \end{abstract}

### 2015-07-08 setup.tex
revised manuscript, with referee response.txt

@@
 unity if the disk is convectively stable, and not
-very close to convective instability  {\bf( $\gamma-\Gamma\geq0.1$ in
+very close to convective instability  {\bf($\gamma-\Gamma\geq0.1$ in
   the disk models considered)}. This stable stratification is

### 2015-07-08 appendix.tex
edits to reply to referee's point 1

@@
 {\bf shows that} $\hat{\upsilon}^2$ is complex for real $\khat$, which
-indicates instability {\bf for any value of  $\gamma>1$. This is inconsistent
+indicates instability {\bf for any value of  $\gamma>1$. This stability condition is inconsistent
   with the second Solberg-Hoiland criterion  
-(Eq. \ref{solberg2}), which implies that instability requires the disk
+(Eq. \ref{solberg2}), which states that instability requires the disk
 to be close to neutral stratification (i.e. $|\gamma-1|\ll 1$ for a
@@
 %also need real freq^2 in adia limit 
-This is because we have retained the global radial temperature 
+This {\bf spurious growth in the $\zeta = 0$ model arises}
+ because we have retained the global radial temperature 
 gradient to obtain vertical shear, 
@@
 that this inconsistency is unimportant for the VSI, which occurs 
-for $\beta\ll1$.   %and small scale
+for $\beta\ll1$, {\bf not for adiabatic perturbations}.   %and small scale
 

### 2015-07-08 setup.tex
edits to reply to referee's point 1

--- 2015-05-09 setup.tex
@@
 side is $O(h^2) \lesssim 10^{-2}$. The left hand side is positive and order
-unity ($O(0.1)$ at least) if the disk is convectively stable, and not
-very close to convective instability.  This stable stratification is
+unity %if the disk is convectively stable, and not
+%very close to convective instability  {\bf($\gamma-\Gamma\geq0.1$ in
+  %the disk models considered)}. 
+  {\bf in our disk models (with $\gamma-\Gamma\geq0.1$)
+  implying strong stability to convection, see Eq.\ \ref{nzsq_def}.}
+  This stable stratification is
 expected for irradiated disks  \citep{chiang97}. 
-Thus adiabatic disturbances are stable to a disk's vertical shear,
+Thus, adiabatic disturbances {\bf in a standard, irradiated disk}
+ are stable to a disk's vertical shear,
+%{\bf for typical values of  $\gamma-\Gamma$},  
 explaining why the VSI requires cooling.

### 2015-07-08 linear_iso.tex
more reply edits

@@
   $\beta\gtrsim O(1)$, this approximation can introduce artificial
-  instability due to the non-self-consistent neglect of global radial gradients.}
+  instability due to the non-self-consistent neglect of global radial gradients (see \S\ref{analytic_adia}).}
 

### 2015-07-08 numerical.tex
more reply edits

@@
 {\bf BL15 mention the interface between a disk's interior and corona
-  as a physical site for surface modes develop. While surface modes in
-  a vertically isothermal disk model are a genuine mathematical feature of the 
-  problem defined by Eq. \ref{lin_all} and Eq. \ref{bc}, they are
-  artificial in the sense that there is no natural choice for $\zmax$.
-}
+  as a physical boundary that can trigger  surface modes. 
+  Since our vertically isothermal models lack a surface, modes which
+  depend on the adopted value of $\zmax$ are not physically meaningful.}
+  %While surface modes in
+  %a vertically isothermal disk model are a genuine mathematical feature of the 
+  %problem defined by Eq. \ref{lin_all} and Eq. \ref{bc}, they are
+  %artificial in the sense that there is no natural choice for $\zmax$.}
 

### 2015-07-09 paper.tex
updated to re-submitted version

@@
 {\bf
-The vertical shear instability (VSI) offers a potential hydrodynamic mechanism to drive angular momentum transport in cold protoplanetary disks (PPDs).
-The VSI is driven by a weak vertical gradient in the disk's orbital motion, but must overcome vertical buoyancy, a strongly
-  stabilizing influence in cold disks, where
-  heating is dominated by external irradiation.  Rapid radiative cooling      reduces the effective buoyancy
-  and allows the VSI to operate.  We quantify the
-  cooling timescale, $t_c$, needed for growth of the VSI. We perform a linear analysis of the VSI with cooling in
-  vertically global, radially local disk models. We find the VSI is most vigorous for
-  rapid cooling with $t_c < \OmK^{-1}  h |q| / (\gamma -1)$
-  in terms of the Keplerian orbital frequency, $\OmK$; the disk's
-  aspect ratio, $ h \ll 1$; the radial
-  power-law temperature gradient, $q$; and the adiabatic index,
-  $\gamma$.  For longer $t_c$, the VSI
-  is much less effective because growth slows and shifts to smaller length scales, which are more prone to
-  viscous or turbulent decay.  We apply our results to PPD models
-  where $t_c$ is determined by the opacity of dust grains.  We
-  find that the VSI is most effective at intermediate radii, from $\sim 5$AU to $\sim 50$AU with a
-  characteristic growth time of $\sim 30$ local orbital periods.
-  Growth is suppressed by long cooling times both in the opaque
-  inner disk and the optically thin outer disk.  A reduction in the
-  dust opacity by a factor of 10 increases cooling times
-  enough to quench the VSI at all disk radii.  Thus the formation of
-  solid protoplanets, a sink for dust grains, can impede the VSI.
+The vertical shear instability (VSI) offers a potential hydrodynamic mechanism for angular momentum transport in protoplanetary disks (PPDs).
+The VSI is driven by a weak vertical gradient in the disk's orbital motion, but must overcome vertical buoyancy, a strongly stabilizing influence in cold disks, where heating is dominated by external irradiation.
+Rapid radiative cooling reduces the effective buoyancy and allows the VSI to operate.
+We quantify the cooling timescale $t_c$ needed for efficient VSI growth, through a linear analysis of the VSI with cooling in vertically global, radially local disk models.
+We find the VSI is most vigorous for rapid cooling with $t_c<\OmK^{-1}h|q|/(\gamma -1)$ in terms of the Keplerian orbital frequency, $\OmK$; the disk's aspect-ratio, $h\ll1$; the radial power-law temperature gradient, $q$; and the adiabatic index, $\gamma$.
+For longer $t_c$, the VSI  is much less effective because growth slows and shifts to smaller length scales, which are more prone to viscous or turbulent decay.
+We apply our results to PPD models where $t_c$ is determined by the opacity of dust grains.
+We find that the VSI is most effective at intermediate radii, from $\sim5$AU to $\sim50$AU with a characteristic growth time of $\sim30$ local orbital periods.
+Growth is suppressed by long cooling times both in the opaque inner disk and the optically thin outer disk.
+Reducing the dust opacity by a factor of 10 increases cooling times enough to quench the VSI at all disk radii.
+Thus the formation of solid protoplanets, a sink for dust grains, can impede the VSI.
 }
